### 2013-10-18 securityupdates.tex
Draft title, authors, abstract and start of introduction

\documentclass[conference,a4paper,twoside]{IEEEtran}
\usepackage[utf8x]{inputenc}
\usepackage{ucs}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\author{Daniel Thomas}
\begin{document}
\title{Security of deployed Android devices}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{
\IEEEauthorblockN{Daniel Wagner,
Daniel R.\ Thomas,
Alastair R.\ Beresford,
Andrew Rice}
\IEEEauthorblockA{
Computer Laboratory\\
University of Cambridge\\
Cambridge, United Kingdom\\
Firstname.Lastname@cl.cam.ac.uk
}
}

\maketitle

\begin{abstract}
Android is the most popular smartphone platform and seeks to provide better security than popular desktop operating systems through increased compartmentalisation and a permission system for apps.
This security relies on there not being root privilege exploits which allow apps to bypass all protections.
Many such exploits have been found and fixed.
Our hypothesis is that these fixes do not promptly reach the users devices and so much of the time users are running versions of Android known to be vulnerable.
We used the DeviceAnalyzer dataset\cite{TODO} and found that over XX\% of devices were exposed to known vulnerabilities on average. %TODO (drt24) proportion of devices exposed on average.
There was also a period of several months when no devices ran secure versions of Android.
\end{abstract}

\section{Introduction}
Android has XX\% of the smartphone market\cite{TODO} and while the core development is controlled by Google there are at least XX\cite{TODO} manufacturers which make devices which run Android.
Many of these manufacturers customise the version of Android they ship and sometimes network operators (of which there are at least XX\cite{TODO}) make further modifications.
Hence when Google produces an update to Android, the update may have to pass through the manufacturer and operator before reaching the user.
The manufacturer and user have no financial incentive to perform this work.
There is ongoing legal action to force some of them to do so\cite{TODO}.

\section{Background}

\section{Results}

\section{Related work}

\section{Conclusion}

\end{document}

### 2013-10-18 securityupdates.tex
More introduction and proportioninsecure graph for the front page

@@
 Our hypothesis is that these fixes do not promptly reach the users devices and so much of the time users are running versions of Android known to be vulnerable.
-We used the DeviceAnalyzer dataset\cite{TODO} and found that over XX\% of devices were exposed to known vulnerabilities on average. %TODO (drt24) proportion of devices exposed on average.
+We used the DeviceAnalyzer data\cite{TODO} and found that over XX\% of devices were exposed to known vulnerabilities on average. %TODO (drt24) proportion of devices exposed on average.
 There was also a period of several months when no devices ran secure versions of Android.
@@
 \section{Introduction}
-Android has XX\% of the smartphone market\cite{TODO} and while the core development is controlled by Google there are at least XX\cite{TODO} manufacturers which make devices which run Android.
-Many of these manufacturers customise the version of Android they ship and sometimes network operators (of which there are at least XX\cite{TODO}) make further modifications.
+Android has XX\% of the smartphone market~\cite{TODO} and while the core development is controlled by Google there are at least XX~\cite{TODO} manufacturers which make devices which run Android.
+Many of these manufacturers customise the version of Android they ship and sometimes network operators (of which there are at least XX~\cite{TODO}) make further modifications.
 Hence when Google produces an update to Android, the update may have to pass through the manufacturer and operator before reaching the user.
 The manufacturer and user have no financial incentive to perform this work.
-There is ongoing legal action to force some of them to do so\cite{TODO}.
+There is ongoing legal action to force some of them to do so~\cite{TODO}.
+
+Android relies on the Linux kernel to provide compartmentalisation and runs each app as a separate unix `user'.
+The Linux kernel is a large piece of software and so there will always be new root exploits found in it~\cite{TODO}.
+Manufacturers of Android devices have also added further root exploits when customising it~\cite{TODO}.
+Between XX\%~\cite{TODO} and XX\%~\cite{TODO} of malware for Android contains root exploits.
+While this shows that much Android malware does not need a root exploit to work, it can just request the relevant permissions and the user will grant them, a significant proportion does.
+This malware is also more dangerous because while ordinary malware can be remotely removed by Google through the Play store, once a root exploit has been used there are no guarantees.
+
+Figure \ref{fig:proportioninsecure} shows the proportion of Android devices in the DeviceAnalyzer data which were running versions of Android known to root exploits at the time.
+
+\begin{figure}[!b]
+\centering
+\includegraphics[width=\columnwidth]{figures/proportioninsecure}
+\caption{Proportion of devices running insecure versions of android}
+\label{fig:proportioninsecure}
+\end{figure}
 
@@
 \section{Related work}
+Stopping root exploits working: SEAndroid.
+Detecting malware: Android antivirus useless.
 

### 2013-10-18 securityupdates.tex
Start background section

@@
 Figure \ref{fig:proportioninsecure} shows the proportion of Android devices in the DeviceAnalyzer data which were running versions of Android known to root exploits at the time.
+In \S\ref{sec:background} we will explain where these data comes from and then in \S\ref{sec:results} we will show these data in more detail.
 
@@
 \section{Background}
+\label{sec:background}
+There are two sources of data we needed for this study, information on root exploits and information on installed versions of Android.
+\subsection{Root exploits}
+We compiled a list of root exploits for Android containing information on when they were known about, what versions they affected and which versions fixed the problem.
+\subsection{Versions of Android running on devices}
+We also need historical information on which versions of Android were running on devices along with other information about those devices which allows us to go into more detail and confirm the reliability of the data our results depend on.
+The only suitable data which is available for this is the DeviceAnalyzer data~\cite{TODO}.
 
 \section{Results}
+\label{sec:results}
 
 \section{Related work}
+\label{sec:related}
 Stopping root exploits working: SEAndroid.
@@
 \section{Conclusion}
+\label{sec:conclusion}
 

### 2013-10-18 securityupdates.tex
Add more information on root vulnerabilities

@@
 \usepackage{graphicx}
+\usepackage{url}
+\usepackage{hyperref}
 \author{Daniel Thomas}
@@
 \includegraphics[width=\columnwidth]{figures/proportioninsecure}
+%TODO make this graph shorter
 \caption{Proportion of devices running insecure versions of android}
@@
 There are two sources of data we needed for this study, information on root exploits and information on installed versions of Android.
+
 \subsection{Root exploits}
-We compiled a list of root exploits for Android containing information on when they were known about, what versions they affected and which versions fixed the problem.
+We compiled a list of root exploit vulnerabilities for Android containing information on when they were known about, what versions they affected and which versions fixed the problem.
+We only looked for \emph{root equivalent} exploits which did not require USB debugging to exploit.
+\emph{Root equivalent} means that it grants privileges equivalent in scope to root which can then be used to gain root.
+Some phones can be `rooted' by enabling USB debugging and using the special privileges of the adb shell to root the device.
+This is not something that applications running on the phone can exploit to break compartmentalisation and so we do not include those exploits.
+Some exploits which grant root are not traditional kernel exploits, for example the discovery of flaws in the verification of signatures on Android applications in ???
+2013 meant that applications could pretend to be signed with system keys and hence gain root privileges.
+
+We have published full details of the vulnerabilities listed below in an accompanying technical report~\cite{TODO} and summarise them in Table \ref{tab:andvulns}.
+We also have a website\footnote{\url{TODO}} which we will keep up to date with future vulnerabilities.
+
+\begin{table}
+\centering
+\begin{tabular}{c|c|c}
+Vulnerability & Date known & How known \\
+\hline
+Exploid & 2010-07-15 & Public disclosure \\
+Psneuter & 2011-01-06 & Public disclosure \\
+Levitator & 2011-03-10 & CVE registered \\
+ZergRush & 2011-10-06 & Exploit developed \\
+APK duplicate file & 2013-02-18 & Fix committed \\
+APK unsigned shorts & 2013-07-03 & Fix committed \\
+% TODO this is an incomplete list, we need to add the rest here and in os-version.py
+\end{tabular}
+\caption{Root equivalent vulnerabilities in Android}
+\label{tab:andvulns}
+\end{table}
+
 \subsection{Versions of Android running on devices}
@@
 \label{sec:related}
-Stopping root exploits working: SEAndroid.
+Stopping root exploits working: SEAndroid, Capsicum, iOS.
 Detecting malware: Android antivirus useless.

### 2013-10-18 securityupdates.tex
Touch up abstract and talk about DA a bit

@@
 Our hypothesis is that these fixes do not promptly reach the users devices and so much of the time users are running versions of Android known to be vulnerable.
-We used the DeviceAnalyzer data\cite{TODO} and found that over XX\% of devices were exposed to known vulnerabilities on average. %TODO (drt24) proportion of devices exposed on average.
+We used the DeviceAnalyzer data\cite{TODO} and found that on average over XX\% of devices were exposed to known vulnerabilities. %TODO (drt24) proportion of devices exposed on average.
 There was also a period of several months when no devices ran secure versions of Android.
@@
 The only suitable data which is available for this is the DeviceAnalyzer data~\cite{TODO}.
+%TODO talk about the DeviceAnalyzer project a bit
+Various different kinds of data\footnote{\url{TODO}} are collected by the DeviceAnalyzer Android app\footnote{\url{TODO}}.
+Among these are the build string and API version for the version of Android currently running on the device each day.
 

### 2013-10-22 securityupdates.tex
Add some results showing the different OS versions and which vulnerabilities affect devices

@@
 The only suitable data which is available for this is the DeviceAnalyzer data~\cite{TODO}.
+The DeviceAnalyzer data contains data from ??? devices with a total of ??? device days, it contains data for more than ??? months for ??? devices.
 %TODO talk about the DeviceAnalyzer project a bit
@@
 Among these are the build string and API version for the version of Android currently running on the device each day.
+The API/SDK version is a well defined integer, however it does not always change with new Android releases, particularly security bug fixes should not increase this value.
+The build string is `The user-visible version string', fortunately most devices in the data have a build string of the form `x.y.z random string' and so it is possible to extract the Android version number.
+%TODO count proportion of device days where the build string cannot be interpreted
+
+\subsection{Running a version of Android known to be vulnerable makes a device vulnerable}
+While we have not tried to exploit these vulnerabilities on the devices in order to test whether the devices are actually vulnerable, we are confident that most devices.
 
@@
 \label{sec:results}
+In this section we present the results of our analysis showing ????. %TODO summarise our results
+
+Figure \ref{fig:os} shows the raw data on the number of devices in the DeviceAnalyzer data running different versions of Android each day.
+It shows how old versions are gradually replaced by new ones, the long tail of devices which do not see updates to more recent versions and the fluctuation in the number of devices in the data.
+\begin{figure}%[!b]
+\centering
+\includegraphics[width=\columnwidth]{figures/os}
+\caption{Android versions in DeviceAnalyzer data over time}
+\label{fig:os}
+\end{figure}
+
+Figure \ref{fig:vulnerabilities} shows which vulnerabilities devices are exposed to.
+For each vulnerability it shows the proportion of devices exposed to that vulnerability and how that changes over time.
+In July 2011 at the beginning of the DeviceAnalyzer data the Exploid and Levitator vulnerabilities both affect most Android devices, slowly these are fixed as updates roll out and devices are replaced until in January 2013 a much smaller proportion of devices are affected by known vulnerabilities.
+However when in February 2013 the first APK signing vulnerability was found which affected all previous versions of Android and even in October 2013 most devices were still unfixed.
+%TODO put percentages on these things to quantify them
+\begin{figure}%[!b]
+\centering
+\includegraphics[width=\columnwidth]{figures/vulnerabilities}
+\caption{Proportion of devices exposed to each vulnerability with time (1.0 is 100\%)}
+\label{fig:vulnerabilities}
+\end{figure}
+
+\subsection{Update frequency}
+%TODO find out what the frequency of updates is
 

### 2013-10-22 securityupdates.tex
99.9% of devices have good version strings for OS versions so we don't need to worry about that

@@
 The API/SDK version is a well defined integer, however it does not always change with new Android releases, particularly security bug fixes should not increase this value.
-The build string is `The user-visible version string', fortunately most devices in the data have a build string of the form `x.y.z random string' and so it is possible to extract the Android version number.
-%TODO count proportion of device days where the build string cannot be interpreted
+The build string is `The user-visible version string', fortunately most (99.9\%) devices in the data have a build string of the form `x.y.z random string' and so it is possible to extract the Android version number.
 

### 2013-11-12 msg2013-11-13.tex
Add presentation for mobile security group

\documentclass{beamer}

% Copyright 2004 by Till Tantau <tantau@users.sourceforge.net>.
%
% In principle, this file can be redistributed and/or modified under
% the terms of the GNU Public License, version 2.
%
% However, this file is supposed to be a template to be modified
% for your own needs. For this reason, if you use this file as a
% template and not specifically distribute it as part of a another
% package/program, I grant the extra permission to freely copy and
% modify this file as you see fit and even to delete this copyright
% notice.

\mode<presentation>
{
  %\usetheme{Warsaw}
  %\usetheme{Goettingen}
  %\usetheme{Montpellier}
  \usetheme{Boadilla}
  % or ...

  \setbeamercovered{transparent}
  % or whatever (possibly just delete it)
}
\beamertemplatenavigationsymbolsempty

%Handout/Notes
%\setbeameroption{show notes} % un-comment to see the notes
%\setbeamertemplate{note page}[plain]
%\usepackage{pgfpages}
%\pgfpagesuselayout{2 on 1}[a4paper,border shrink=5mm]

\usepackage[english]{babel}
% or whatever

\usepackage[latin1]{inputenc}
% or whatever

\usepackage{times}
\usepackage[T1]{fontenc}
% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.
\usepackage{tikz}

\title[Security Updates] % (optional, use only with long paper titles)
{Security of deployed Android devices}

\subtitle
{When do Android devices get security updates for root exploits} % (optional)
%\pgfdeclareimage[height=2.9cm]{drt24}{img/drt24}
%\titlegraphic{\pgfuseimage{drt24}}

\author[Daniel Thomas] % (optional, use only with lots of authors)
{Daniel Thomas\\ \vspace{1em} \tiny{(PhD student, 2\textsuperscript{nd} year)} \\ \vspace{1em} \tiny{Supervised by Alastair Beresford}}
\institute{\vspace{-1em}
University of Cambridge
}

\date[2013-11-13] % (optional)
{Mobile Security Reading Group WIP presentation}

\subject{Mobile Security Reading Group WIP presentation}
% This is only inserted into the PDF information catalog. Can be left
% out.

% If you have a file called "university-logo-filename.xxx", where xxx
% is a graphic format that can be processed by latex or pdflatex,
% resp., then you can add a logo as follows:

 \pgfdeclareimage[height=0.7cm]{university-logo}{university-logo}
 \logo{\pgfuseimage{university-logo}}

% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command:

%\beamerdefaultoverlayspecification{<+->}

% Maximum of 12 slides including title page.
% 15 minutes talk

\begin{document}

\begin{frame}
  \titlepage
  \note{We collected data with Device Analyzer, collated a list of root equivalent exploits which apps could use and found that a large proportion of devices are vulnerable.}
\end{frame}

\begin{frame}{Device Analyser collects lots of Data}{}
 \begin{itemize}
  \note{}
  \item Android app by Daniel Wagner installed on user devices collecting data with consent.
  \note{\\}
  \item Deployed since May 2011
  \note{\\}
  \item ??? records, ??? years, ??? devices
  \item \url{http://deviceanalyzer.cl.cam.ac.uk/}
 \end{itemize}
 \pgfdeclareimage[height=50px]{dalogo}{figures/dalogo100}
 \begin{center}
  \pgfuseimage{dalogo}
 \end{center}
\end{frame}

\begin{frame}{Device Analyzer collects different kinds of data}{}
% Specific research aim (3+ year time scale)
 \pgfdeclareimage[height=0.38\textheight]{da_main}{figures/da_screenshot_main_screen}
 \pgfdeclareimage[height=0.38\textheight]{da_app_data}{figures/da_screenshot_app_data}
 \pgfdeclareimage[height=0.38\textheight]{da_phone_usage}{figures/da_screenshot_phone_usage}
 \begin{center}
  \pgfuseimage{da_main} \hspace{1em} \pgfuseimage{da_app_data} \hspace{1em} \pgfuseimage{da_phone_usage}
 \end{center}
 \vspace{-1.5em}
 \begin{itemize}
  \item DA also collects information on the OS version information and build number.
  \note{}
 \end{itemize}

\end{frame}

\begin{frame}{Root equivalent vulnerabilities are interesting}{}
 \begin{itemize}
  \item Root exploits break out of the sandbox.
  \note{}
  \item They allow applications to do arbitrarily bad things
  \item They are used in ??-??\% of malware
  \item How many devices are vulnerable to these exploits?
 \end{itemize}
\end{frame}

\begin{frame}{Root equivalent vulnerabilities on Android}{}
\small \url{https://wiki.dtg.cl.cam.ac.uk/drt24/dasecurityupdates}\\
All Android vulnerabilities:
 \begin{itemize}
  \item exploid
  \item RageAgainstTheCage zygote
  \item RageAgainstTheCage adb setuid exhaustion
  \item KillingInTheNameOf psneuter ashmem
  \item zergRush
  \item levitator
  \item Gingerbreak
  \item APK signature verification - duplicate file entries
  \item APK signature verification - unsigned shorts
 \end{itemize}

\end{frame}

% Leverage of colleagues
\begin{frame}{Old versions of Android still in use}{Coarse grained API information}
 \pgfdeclareimage[width=0.5\textwidth]{da_norm_api}{figures/da_norm_api}
 \pgfdeclareimage[width=0.5\textwidth]{play_norm_api}{figures/play_norm_api}
 \begin{center}
  \pgfuseimage{da_norm_api}
  \pgfuseimage{play_norm_api}
 \end{center}
Device Analyzer data is representative of the versions of Android actually in use
\end{frame}

\begin{frame}{Old versions of Android still in use}{Detailed OS version}
 \pgfdeclareimage[height=0.87\textheight]{da_norm_os}{figures/da_norm_os}
 \begin{center}
  \vspace{-0.5em}
  \pgfuseimage{da_norm_os}
 \end{center}
\end{frame}

\begin{frame}{Lots of Android devices are exposed to known root equivalent vulnerabilities}{}
 \pgfdeclareimage[height=0.87\textheight]{da_proportioninsecure}{figures/proportioninsecure}
 \begin{center}
  \vspace{-0.5em}
  \pgfuseimage{da_proportioninsecure}
 \end{center}
\end{frame}

\begin{frame}{Several vulnerabilities contribute to this insecurity}{}
 \pgfdeclareimage[height=0.95\textheight]{da_vulnerabilities}{figures/vulnerabilities}
 \begin{center}
  \vspace{-0.5em}
  \pgfuseimage{da_vulnerabilities}
 \end{center}
\end{frame}

\begin{frame}{Devices do get updated}{}
 \pgfdeclareimage[height=0.95\textheight]{da_updates_between_versions}{figures/updates_between_versions}
 \begin{center}
  \vspace{-0.5em}
  \pgfuseimage{da_updates_between_versions}
 \end{center}
\end{frame}

\begin{frame}{Devices do get security updates}{}
 \pgfdeclareimage[height=0.95\textheight]{da_w_security_updates}{figures/w_security_updates}
 \begin{center}
  \vspace{-0.5em}
  \pgfuseimage{da_w_security_updates}
 \end{center}
\end{frame}

\begin{frame}{There are weaknesses in our current approach}{}
 \begin{itemize}
  \item What complementary data should we be using?
  \item Do we have all known root vulnerabilities? (No)
  \item Is our inference that if a device is running a version of Android with a known vulnerability then it is vulnerable true?
  \item Is there anything else we can add to Device Analyzer to improve research into this area?
 \end{itemize}
\end{frame}

\end{document}

### 2013-11-12 msg2013-11-13.tex
Changes to msg presentation after discussion with Alastair

@@
 
- \pgfdeclareimage[height=0.7cm]{university-logo}{university-logo}
- \logo{\pgfuseimage{university-logo}}
+ %\pgfdeclareimage[height=0.7cm]{university-logo}{university-logo}
+ %\logo{\pgfuseimage{university-logo}}
 
@@
 
-\begin{frame}{Device Analyser collects lots of Data}{}
+\begin{frame}{Device Analyser collects lots of data}{}
  \begin{itemize}
   \note{}
-  \item Android app by Daniel Wagner installed on user devices collecting data with consent.
+  \item Android app installed on user devices
+  \item Collects data with consent and sends it to our server
+  \item Built by Daniel Wagner
   \note{\\}
@@
   \note{\\}
-  \item ??? records, ??? years, ??? devices
+  \item 70 billion records, 2000 years, 16000 devices %TODO update these numbers
   \item \url{http://deviceanalyzer.cl.cam.ac.uk/}
  \end{itemize}
- \pgfdeclareimage[height=50px]{dalogo}{figures/dalogo100}
+ \pgfdeclareimage[height=90px]{dalogo}{figures/dalogo512}
  \begin{center}
@@
 % Specific research aim (3+ year time scale)
- \pgfdeclareimage[height=0.38\textheight]{da_main}{figures/da_screenshot_main_screen}
- \pgfdeclareimage[height=0.38\textheight]{da_app_data}{figures/da_screenshot_app_data}
- \pgfdeclareimage[height=0.38\textheight]{da_phone_usage}{figures/da_screenshot_phone_usage}
+ \pgfdeclareimage[height=0.75\textheight]{da_main}{figures/da_screenshot_main_screen}
+ \pgfdeclareimage[height=0.75\textheight]{da_app_data}{figures/da_screenshot_app_data}
+ \pgfdeclareimage[height=0.75\textheight]{da_phone_usage}{figures/da_screenshot_phone_usage}
  \begin{center}
-  \pgfuseimage{da_main} \hspace{1em} \pgfuseimage{da_app_data} \hspace{1em} \pgfuseimage{da_phone_usage}
+  \vspace{-1em}
+  \pgfuseimage{da_main} \pgfuseimage{da_app_data} \pgfuseimage{da_phone_usage}
  \end{center}
@@
  \end{itemize}
+\end{frame}
 
+\begin{frame}{Other data}{}
+\begin{itemize}
+ \item \Large{ \emph{OS Version}}
+ \item \emph{Build number}
+ \item Device name
+ \item GPU version
+ \item Phone features
+ \item baseband version
+ \item What else?
+\end{itemize}
 \end{frame}
@@
   \item They allow applications to do arbitrarily bad things
-  \item They are used in ??-??\% of malware
+  \item They are used in 37\% of malware %TODO check this figure
   \item How many devices are vulnerable to these exploits?
@@
 \small \url{https://wiki.dtg.cl.cam.ac.uk/drt24/dasecurityupdates}\\
-All Android vulnerabilities:
+Vulnerabilities affecting all of Android (not manufacturer specific):
  \begin{itemize}
@@
  \end{itemize}
-
+There are also lots of manufacture/device specific vulnerabilities.
 \end{frame}
 
-% Leverage of colleagues
-\begin{frame}{Old versions of Android still in use}{Coarse grained API information}
- \pgfdeclareimage[width=0.5\textwidth]{da_norm_api}{figures/da_norm_api}
- \pgfdeclareimage[width=0.5\textwidth]{play_norm_api}{figures/play_norm_api}
+ \pgfdeclareimage[height=0.9\textheight]{da_norm_api}{figures/da_norm_api}
+ \pgfdeclareimage[height=0.9\textheight]{play_norm_api}{figures/play_norm_api}
+\begin{frame}{Old versions of Android still in use -- Google Play API}{}
+ \begin{center}
+  \pgfuseimage{play_norm_api}
+ \end{center}
+\end{frame}
+
+\begin{frame}{Old versions of Android still in use -- Device Analyzer}{}
  \begin{center}
   \pgfuseimage{da_norm_api}
-  \pgfuseimage{play_norm_api}
+ \end{center}
+\end{frame}
+
+\begin{frame}{Old versions of Android still in use}{Coarse grained API information}
+ \pgfdeclareimage[width=0.5\textwidth]{da_norm_api_small}{figures/da_norm_api}
+ \pgfdeclareimage[width=0.5\textwidth]{play_norm_api_small}{figures/play_norm_api}
+ \begin{center}
+  \pgfuseimage{da_norm_api_small}
+  \pgfuseimage{play_norm_api_small}
  \end{center}

### 2014-01-13 netos2014-01-14.tex
Copy msg presentation to netos presentation to start work on it

--- 2013-11-12 msg2013-11-13.tex
@@
 
-\begin{frame}{Device Analyser collects lots of data}{}
+\begin{frame}{Device Analyzer collects lots of data}{}
  \begin{itemize}
@@
   \note{\\}
-  \item 70 billion records, 2000 years, 16000 devices %TODO update these numbers
+  \item 70 billion records, 2000 years, 16000 devices
   \item \url{http://deviceanalyzer.cl.cam.ac.uk/}
@@
   \item They allow applications to do arbitrarily bad things
-  \item They are used in 37\% of malware %TODO check this figure
+  \item Malware which uses root exploits cannot be remotely removed
+  \item They are used in 37\% of malware % "Dissecting Android Malware: Characterization and Evolution" at "2012 IEEE Symposium on Security and Privacy"
   \item How many devices are vulnerable to these exploits?
@@
  \pgfdeclareimage[width=0.5\textwidth]{da_norm_api_small}{figures/da_norm_api}
- \pgfdeclareimage[width=0.5\textwidth]{play_norm_api_small}{figures/play_norm_api}
+ \pgfdeclareimage[width=0.5\textwidth]{play_norm_api_small}{figures/play_norm_api_trunc}
  \begin{center}

### 2014-01-13 netos2014-01-14.tex
Make netos talk make better use of slide area and have correct titles

@@
 
-\date[2013-11-13] % (optional)
-{Mobile Security Reading Group WIP presentation}
+\date[2014-01-14] % (optional)
+{NetOS talklet}
 
-\subject{Mobile Security Reading Group WIP presentation}
+\subject{NetOS talklet}
 % This is only inserted into the PDF information catalog. Can be left
@@
   \note{\\}
-  \item 70 billion records, 2000 years, 16000 devices
-  \item \url{http://deviceanalyzer.cl.cam.ac.uk/}
+  \item 100 billion records, 2000 years, 17000 devices %TODO update years value
+  \item \url{https://deviceanalyzer.cl.cam.ac.uk/}
  \end{itemize}
@@
 \begin{frame}{Root equivalent vulnerabilities on Android}{}
-\small \url{https://wiki.dtg.cl.cam.ac.uk/drt24/dasecurityupdates}\\
+\small \url{http://androidvulnerabilities.org/}
+
 Vulnerabilities affecting all of Android (not manufacturer specific):
@@
   \item Gingerbreak
-  \item APK signature verification - duplicate file entries
-  \item APK signature verification - unsigned shorts
+  \item APK signature verification -- duplicate file entries
+  \item APK signature verification -- unsigned shorts
+  \item APK signature verification -- unchecked name
  \end{itemize}
@@
 
- \pgfdeclareimage[height=0.9\textheight]{da_norm_api}{figures/da_norm_api}
- \pgfdeclareimage[height=0.9\textheight]{play_norm_api}{figures/play_norm_api}
-\begin{frame}{Old versions of Android still in use -- Google Play API}{}
+ \pgfdeclareimage[height=0.96\textheight]{da_norm_api}{figures/da_norm_api}
+ \pgfdeclareimage[height=0.96\textheight]{play_norm_api}{figures/play_norm_api}
+\begin{frame}[plain]{Old versions of Android still in use -- Google Play API}{}
  \begin{center}
@@
 
-\begin{frame}{Old versions of Android still in use -- Device Analyzer}{}
+\begin{frame}[plain]{Old versions of Android still in use -- Device Analyzer}{}
  \begin{center}
@@
 \begin{frame}{Old versions of Android still in use}{Coarse grained API information}
- \pgfdeclareimage[width=0.5\textwidth]{da_norm_api_small}{figures/da_norm_api}
- \pgfdeclareimage[width=0.5\textwidth]{play_norm_api_small}{figures/play_norm_api_trunc}
+ \pgfdeclareimage[width=0.51\textwidth]{da_norm_api_small}{figures/da_norm_api}
+ \pgfdeclareimage[width=0.51\textwidth]{play_norm_api_small}{figures/play_norm_api_trunc}
  \begin{center}
-  \pgfuseimage{da_norm_api_small}
+  \hspace{-1em}\pgfuseimage{da_norm_api_small}\hspace{-0.5em}
   \pgfuseimage{play_norm_api_small}
@@
 
-\begin{frame}{Old versions of Android still in use}{Detailed OS version}
- \pgfdeclareimage[height=0.87\textheight]{da_norm_os}{figures/da_norm_os}
+\begin{frame}[plain]{Detailed OS version}{}
+ \pgfdeclareimage[height=0.99\textheight]{da_norm_os}{figures/da_norm_os}
  \begin{center}
@@
 
-\begin{frame}{Lots of Android devices are exposed to known root equivalent vulnerabilities}{}
- \pgfdeclareimage[height=0.87\textheight]{da_proportioninsecure}{figures/proportioninsecure}
+\begin{frame}[plain]{Android devices with root equivalent vulnerabilities}{}
+ \pgfdeclareimage[height=0.98\textheight]{da_proportioninsecure}{figures/proportioninsecure}
  \begin{center}
@@
 
-\begin{frame}{Several vulnerabilities contribute to this insecurity}{}
- \pgfdeclareimage[height=0.95\textheight]{da_vulnerabilities}{figures/vulnerabilities}
+\begin{frame}[plain]{Several vulnerabilities contribute to this insecurity}{}
+ \pgfdeclareimage[height=0.97\textheight]{da_vulnerabilities}{figures/vulnerabilities}
  \begin{center}
-  \vspace{-0.5em}
+  \vspace{-0.3em}
   \pgfuseimage{da_vulnerabilities}
@@
 
-\begin{frame}{Devices do get updated}{}
- \pgfdeclareimage[height=0.95\textheight]{da_updates_between_versions}{figures/updates_between_versions}
+\begin{frame}[plain]{Devices do get updated}{}
+ \pgfdeclareimage[height=0.98\textheight]{da_updates_between_versions}{figures/updates_between_versions}
  \begin{center}
-  \vspace{-0.5em}
+  \vspace{-0.2em}
   \pgfuseimage{da_updates_between_versions}
@@
 
-\begin{frame}{Devices do get security updates}{}
- \pgfdeclareimage[height=0.95\textheight]{da_w_security_updates}{figures/w_security_updates}
+\begin{frame}[plain]{Devices do get security updates}{}
+ \pgfdeclareimage[height=0.98\textheight]{da_w_security_updates}{figures/w_security_updates}
  \begin{center}
-  \vspace{-0.5em}
-  \pgfuseimage{da_w_security_updates}
+  \vspace{-0.2em}
+  \hspace{-2em}\pgfuseimage{da_w_security_updates}
  \end{center}

### 2014-01-13 netos2014-01-14.tex
Add a couple of graphs on wifi networks to the end of the netos talk

@@
 
+\begin{frame}[plain]{Connections to secured wifi networks}{}
+ \pgfdeclareimage[height=0.95\textheight]{wifi-accounts-stotals}{figures/wifi-accounts-stotals-hist}
+ \begin{center}
+  \vspace{-0.2em}
+  \hspace{-3em}\pgfuseimage{wifi-accounts-stotals}
+ \end{center}
+\end{frame}
+\begin{frame}[plain]{Connections to open wifi networks}{}
+ \pgfdeclareimage[height=0.95\textheight]{wifi-accounts-ototals}{figures/wifi-accounts-ototals-hist}
+ \begin{center}
+  \vspace{-0.2em}
+  \hspace{-3em}\pgfuseimage{wifi-accounts-ototals}
+ \end{center}
+\end{frame}
+
 \end{document}

### 2014-01-14 netos2014-01-14.tex
Add speaker notes to netos talk

@@
 \subtitle
-{When do Android devices get security updates for root exploits} % (optional)
+{When do Android devices get security updates for root exploits?} % (optional)
 %\pgfdeclareimage[height=2.9cm]{drt24}{img/drt24}
@@
 {Daniel Thomas\\ \vspace{1em} \tiny{(PhD student, 2\textsuperscript{nd} year)} \\ \vspace{1em} \tiny{Supervised by Alastair Beresford}}
-\institute{\vspace{-1em}
-University of Cambridge
-}
+\institute{\vspace{-1em}University of Cambridge}
 
@@
   \titlepage
-  \note{We collected data with Device Analyzer, collated a list of root equivalent exploits which apps could use and found that a large proportion of devices are vulnerable.}
+  \note{We collected data with Device Analyzer, collated a list of root equivalent exploits which apps could use and found that a large proportion of devices are vulnerable.\\}
+  \note{I have drawn some graphs to show this data and we will discuss the weaknesses with this approach at the end}
+  \note{I also have a couple of bonus slides on wireless security}
 \end{frame}
@@
   \note{}
-  \item Android app installed on user devices
+  \item Android app installed on user devices (over 1000 each week)
   \item Collects data with consent and sends it to our server
   \item Built by Daniel Wagner
-  \note{\\}
+  \note{Large engineering effort, rebuilt four times, now starting to deliver\\}
   \item Deployed since May 2011
-  \note{\\}
+  \note{Now longitudinal studies possible\\}
   \item 100 billion records, 2000 years, 17000 devices %TODO update years value
@@
  \end{center}
+ \note{Data on applications, on phone calls/text messages, bluetooth, location, power usage, wifi\\}
  \vspace{-1.5em}
@@
   \item DA also collects information on the OS version information and build number.
-  \note{}
+  \note{More usefully for us it also collects OS version and build number}
  \end{itemize}
@@
  \item baseband version
+ \note{DA collects all this information, but what other information should it also collect?}
  \item What else?
@@
   \item Root exploits break out of the sandbox.
-  \note{}
+  \note{Permissions are meaningless if an app has root (modulo SEAndroid etc.)}
   \item They allow applications to do arbitrarily bad things
@@
 \small \url{http://androidvulnerabilities.org/}
+\note{I made a website with this information collated together in a machine readable format (JSON), if there are any you know of which are not listed then please let me know or submit a patch.\\}
 
@@
  \end{itemize}
-There are also lots of manufacture/device specific vulnerabilities.
+There are also lots of manufacturer/device specific vulnerabilities.
+\note{But I said their would be graphs, so lets look at some graphs}
 \end{frame}
@@
  \end{center}
+ \note{So we want to look at which versions of Android are still in use, the information we really care about is which OS version is running on devices in the wild.
+ While that data is in DA there is no authoritative source to compare DA against to check it.\\}
+ \note{However Google has published the API versions of devices connecting to its store since the end of 2009 and so I collated that information and graphed it.}
 \end{frame}
@@
  \end{center}
+ \note{Now we have the same data for Device Analyzer showing old versions hanging around}
 \end{frame}
@@
  Device Analyzer data is representative of the versions of Android actually in use
+ \note{These graphs do look fairly similar, I can claim that most of the differences are stylistic or not significant but if you know a statistical measure which could confirm of deny that hypothesis then do let me know (afterwards).}
 \end{frame}
@@
  \end{center}
+ \note{What we actually want is the detailed information about OS versions used which is what Device Analyzer can give us.\\}
+ \note{You can see how version 4.2.0 was quickly replaced entirely by 4.2.1 and how 4.1.2 became big after a later version had already been released.\\}
+ \note{You can see the long tail of very old versions of Android still in use many years after they were superseded.}
 \end{frame}
@@
  \end{center}
+ \note{This is not a happy graph.
+The big vertical rise in early 2013 is the discovery of the APK signing vulnerabilities, most devices still haven't been fixed.
+Now that particular flaw is mitigated by Google Play doing static analysis but it is still illustrative of the problem.}
 \end{frame}
@@
  \end{center}
+ \note{I don't like this graph as it gives some incorrect visual impressions but it does show more of the detail of the data.\\
+ A value of 1 here means that all devices are affected by a vulnerability.
+Being affected by more than one vulnerability is worse than being affected by only one as an attacker is more likely to have included an exploit which works.}
 \end{frame}
@@
  \end{center}
+ \note{Due to the nature of the DA data with a lot of devices only contributing for a short period of time, we don't see all update events.
+ However we do still see quite a few which allows us to see which versions are being upgraded between and when.\\
+ You can see a big spike in updates to the latest version just after it comes out and then often a big but broader spike a few months later.}
 \end{frame}
@@
  \end{center}
+ \note{If we look at these update events in detail we can try and determine whether an update reduced the number of known root vulnerabilities.
+ Some updates don't change the version number, they could include backported security patches, we don't know.}
 \end{frame}
@@
   \item Do we have all known root vulnerabilities? (No)
+  \note{Please submit other vulnerabilities which you know about\\}
   \item Is our inference that if a device is running a version of Android with a known vulnerability then it is vulnerable true?
@@
  \end{center}
+ \note{This shows the number of wifi networks connected to which had some sort of security (WPA, WEP etc.), this just uses devices which participated for at least 30 days.
+ The mode is 2 secured networks but there is a long tail, 80 something devices connected to 50 or more secured networks.}
 \end{frame}
+
 \begin{frame}[plain]{Connections to open wifi networks}{}
@@
  \end{center}
+ \note{This is the same graph but for open wifi networks, most people only connect to one open wireless network.
+ Explanations welcome.}
 \end{frame}

### 2014-01-29 platforms.tex
Introduction to platforms document

\documentclass[12pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
\author{Daniel Thomas}
\title{Platforms for security update research}
\begin{document}
\maketitle

\section{Introduction}
Early computers allowed running code to do anything, later more restrictive systems were developed which had an operating system which could do anything but which ran programs which were restricted in what they could do.
However mostly this was broken down on a per user basis, code running as a particular user could do anything a user could do.
More recently there have been efforts at greater compartmentalisation and more fine grained permissions for applications.
For example the Android permission system restricts what each app can do in a more precise way than installing applications on Linux does.
This fine grained compartmentalisation only works if there are no known vulnerabilities which break it and allow code to do arbitrary things.
Unfortunately this is not the case.

\section{Android}
\section{iOS}
\section{Firefox}
\section{Chrome}
\section{ChromeOS}
\section{JVM}
\section{Xen}
\section{VMWare}

\end{document}

### 2014-01-31 platforms.tex
Add bibliography to platforms.tex
platforms: android

@@
 \usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
+\usepackage{hyperref}
+\RequirePackage[date=terse,isbn=true,doi=true,url=false,maxbibnames=9,backref=true]{biblatex}
+\addbibresource{securityupdates.bib}
+\renewcommand{\bibfont}{\small}
+
+\AtEveryBibitem{% Clean up the bibtex rather than editing it
+ \clearname{editor} % remove editors
+}
+
 \author{Daniel Thomas}
@@
 \section{Android}
+In Android each app runs in a separate user account and is granted certain permissions statically after user review on installation.
+Apps can use the rich Android API and embed native code which can make use of the Linux API.
+The security boundary is the user but there have been many examples where it is possible for an app to exploit a vulnerability in Android and gain root permissions and do anything.
+To analyse the proportion of the time deployed Android devices are exposed to vulnerabilities which break their security boundary we can use the Device Analyzer data to look at running versions and the \url{http://androidvulnerabilities.org} website to find out which vulnerabilities applied to which versions at what time.
+
 \section{iOS}
@@
 
+\printbibliography
+
 \end{document}

### 2014-01-31 platforms.tex
platforms: iOS

@@
 \section{iOS}
+In iOS every process runs as the same user but mandatory access control is used to restrict what processes can do to permitted things.
+iOS has additional features that make breaking the security boundary more difficult: Address Space Layout Randomization (ASLR), Data Execution Prevention (DEP), and Sandboxing, iOS also enforces the mandatory App Review and code signing mechanisms.
+However it is still possible to execute arbitrary code on the device~\cite{Wang2013a} and hence exploit flaws to Jailbreak the phone.
+To analyse update security on iOS we need detailed information on the version of iOS running on deployed devices \emph{from where?} and a database of discovered vulnerabilities which would need to be created.
+
 \section{Firefox}

### 2014-01-31 platforms.tex
platforms: start firefox section

@@
 \section{Firefox}
+Firefox can be extended with extensions which are written in javascript and XML, the APIs allow the equivalent arbitrary code to be run as the Firefox process and so there is no real sandboxing of extensions~\cite{Lerner2013}.
+There is manual review to detect and remove bad extensions.
+Web pages are allowed to run javascript and with html5 are presented with an increasingly rich API they are not allowed to execute arbitrary code on the computer as the user.
+To analyse the update security of Firefox we need data on the version distribution of Firefox over time and a database of vulnerabilities in Firefox.
+
 \section{Chrome}
+
 \section{ChromeOS}

### 2014-02-04 platforms.tex
platforms: restructure to have subsections

@@
 
-\section{Android}
+\section{Mobile operating systems}
+
+\subsection{Android}
 In Android each app runs in a separate user account and is granted certain permissions statically after user review on installation.
@@
 
-\section{iOS}
+Android has many vendors which provides some heterogeneity in vulnerabilities but most of these vendors are not as good as Google at shipping updates promptly (perhaps partly because Google is better placed to do so).
+
+\subsection{iOS}
 In iOS every process runs as the same user but mandatory access control is used to restrict what processes can do to permitted things.
@@
 
-\section{Firefox}
+The only iOS vendor is Apple who have full control over the OS and its distribution this enables them to make prompt changes but a flaw affecting one device will affect all devices.
+Since there is no competition between iOS vendors competition with other mobile OSes is the only incentive for innovation.
+
+\section{Web browsers}
+There are two sources of arbitrary code execution in web browsers, extensions which typically come through a market and web pages which can come from anywhere.
+
+\subsection{Firefox}
 Firefox can be extended with extensions which are written in javascript and XML, the APIs allow the equivalent arbitrary code to be run as the Firefox process and so there is no real sandboxing of extensions~\cite{Lerner2013}.
@@
 
-\section{ChromeOS}
-\section{JVM}
-\section{Xen}
-\section{VMWare}
+\section{Hypervisors}
+\subsection{Xen}
+\subsection{VMWare}
 

### 2014-02-04 platforms.tex
platforms: add section on chrome

@@
 
-\section{Chrome}
+\subsection{Chrome}
+Chrome uses privilege separation to separate the renderer from the browser kernel by running them in different processes~\cite{Barth2008}.
+Extensions are also split into multiple processes with content scripts running in the host page's renderer process and the extension core running in a separate renderer process~\cite{Liu2012}.
+Malicious extensions can inject adverts and steal data or engage in more nefarious attacks on the user but a vulnerability in the sandbox would be required for them to execute arbitrary code.
+To analyse the update security of Chrome we need data on the version distribution of Chrome over time and a database of vulnerabilities in Chrome.
+
+\subsection{ChromeOS}
+
+\section{Java Virtual Machine (JVM)}
 

### 2014-02-05 platforms.tex
platforms: urls to vulnerability databases for firefox and chrome
platforms: discuss chromeos

@@
 To analyse the update security of Firefox we need data on the version distribution of Firefox over time and a database of vulnerabilities in Firefox.
+Firefox has a nice list of vulnerabilities with severities: \url{https://www.mozilla.org/security/known-vulnerabilities/firefox.html}
 
@@
 To analyse the update security of Chrome we need data on the version distribution of Chrome over time and a database of vulnerabilities in Chrome.
+There are a fair number of CVEs listed for Chrome \url{http://www.cvedetails.com/vulnerability-list.php?vendor_id=1224&product_id=15031}.
 
 \subsection{ChromeOS}
+ChromeOS takes the same functionality as Chrome and sticks a more secure OS underneath with better sandboxing, process separation, boot verification and secure updates.\footnote{\url{http://www.chromium.org/chromium-os/chromiumos-design-docs/security-overview}}
+There might still be security flaws particular to ChromeOS but it seems likely that most flaws would have to be in Chrome first.
 

### 2014-02-05 platforms.tex
platforms: discuss JVM

@@
 \section{Java Virtual Machine (JVM)}
+The Java Virtual Machine has a security manager and attempts to impose fine grained permissions on executing Java code~\cite{Gong1997} this is used by Java plugins in browsers to execute remote code without exposing the user to attacks.
+Unfortunately there are many flaws in the JVM which have allowed arbitrary execute of code bypassing the security manager.
+There are a list of vulnerabilities in the JVM \url{http://www.cvedetails.com/vulnerability-list/vendor_id-5/product_id-1526/opec-1/SUN-JRE.html}.
+We also need the distribution of running versions over time.
 

### 2014-02-06 platforms.tex
Put URLs in footnotes for legibility

@@
 The security boundary is the user but there have been many examples where it is possible for an app to exploit a vulnerability in Android and gain root permissions and do anything.
-To analyse the proportion of the time deployed Android devices are exposed to vulnerabilities which break their security boundary we can use the Device Analyzer data to look at running versions and the \url{http://androidvulnerabilities.org} website to find out which vulnerabilities applied to which versions at what time.
+To analyse the proportion of the time deployed Android devices are exposed to vulnerabilities which break their security boundary we can use the Device Analyzer data to look at running versions and the website\footnote{\url{http://androidvulnerabilities.org}} to find out which vulnerabilities applied to which versions at what time.
 
@@
 There are two sources of arbitrary code execution in web browsers, extensions which typically come through a market and web pages which can come from anywhere.
+Version information for browsers is available.\footnote{\url{http://gs.statcounter.com/#desktop-browser_version-ww-monthly-200807-201401}}
 
@@
 Web pages are allowed to run javascript and with html5 are presented with an increasingly rich API they are not allowed to execute arbitrary code on the computer as the user.
-To analyse the update security of Firefox we need data on the version distribution of Firefox over time and a database of vulnerabilities in Firefox.
-Firefox has a nice list of vulnerabilities with severities: \url{https://www.mozilla.org/security/known-vulnerabilities/firefox.html}
+Firefox has a nice list of vulnerabilities with severities.\footnote{\url{https://www.mozilla.org/security/known-vulnerabilities/firefox.html}}
 
@@
 To analyse the update security of Chrome we need data on the version distribution of Chrome over time and a database of vulnerabilities in Chrome.
-There are a fair number of CVEs listed for Chrome \url{http://www.cvedetails.com/vulnerability-list.php?vendor_id=1224&product_id=15031}.
+There are a fair number of CVEs listed for Chrome.\footnote{\url{http://www.cvedetails.com/vulnerability-list.php?vendor_id=1224&product_id=15031}}
+There are some statistics on version distributions for chrome.\footnote{\url{http://www.w3schools.com/browsers/browsers_chrome.asp}}
 
@@
 Unfortunately there are many flaws in the JVM which have allowed arbitrary execute of code bypassing the security manager.
-There are a list of vulnerabilities in the JVM \url{http://www.cvedetails.com/vulnerability-list/vendor_id-5/product_id-1526/opec-1/SUN-JRE.html}.
+There are a list of vulnerabilities in the JVM.\footnote{\url{http://www.cvedetails.com/vulnerability-list/vendor_id-5/product_id-1526/opec-1/SUN-JRE.html}}
 We also need the distribution of running versions over time.

### 2014-02-06 platforms.tex
platforms: add section on hypervisors

@@
 \section{Hypervisors}
+Hypervisors try to improve security by offering a much smaller API to the VM's operating system and enforcing privilege separation by using different CPU ring levels.
+Unfortunately they still have vulnerabilities which break this confinement.
+
 \subsection{Xen}
+Xen is an opensource hypervisor with a commercial version produced by Citrix.
+It is widely used by lots of companies including Amazon which allow multiple mutually untrusting VMs to run on the same machine.
+There is are list of vulnerabilities in Xen.\footnote{\url{http://wiki.xenproject.org/wiki/Security_Announcements_(Historical)} \url{http://xenbits.xen.org/xsa/}}
+Where do we get version distribution information from?
+
 \subsection{VMWare}
+VMWare is a closed source hypervisor.
+There is a list of security advisories affecting VMWare\footnote{\url{http://www.vmware.com/uk/security/advisories/}} though it also includes some other products.
+Where do we get version distribution information from?
 

### 2014-02-07 securityupdates.tex
Switch authors around after taking to Daniel Wagner
Change title and add hypothesis
Adjust preamble

@@
 \usepackage{hyperref}
-\author{Daniel Thomas}
+\author{Daniel R.\ Thomas, Daniel Wagner, Alastair R.\ Beresford and Andrew Rice}
+
+\newcommand{\percArdMarketShare}{XX\%}
+\newcommand{\percArdInsec}{XX\%}
+\newcommand{\percCrmMarketShare}{XX\%}
+\newcommand{\percCrmInsec}{X\%}
+\newcommand{\daNumManufacturers}{XX}
+\newcommand{\daNumNetworkOperators}{XXX}
+
 \begin{document}
-\title{Security of deployed Android devices}
+\title{Timeliness of security updates for Android and Chrome}
 
@@
 \author{
-\IEEEauthorblockN{Daniel Wagner,
-Daniel R.\ Thomas,
+\IEEEauthorblockN{Daniel R.\ Thomas,
+Daniel Wagner,
 Alastair R.\ Beresford,
@@
 \maketitle
+
+% Hypothesis
+% Attempts at fine grained restrictions on running arbitrary code are hampered by updates for security vulnerabilities not reaching uers in a timely fashion.
 

### 2014-02-07 securityupdates.tex
Replace abstract and start talking about Chrome

@@
 \begin{abstract}
-Android is the most popular smartphone platform and seeks to provide better security than popular desktop operating systems through increased compartmentalisation and a permission system for apps.
-This security relies on there not being root privilege exploits which allow apps to bypass all protections.
-Many such exploits have been found and fixed.
-Our hypothesis is that these fixes do not promptly reach the users devices and so much of the time users are running versions of Android known to be vulnerable.
-We used the DeviceAnalyzer data\cite{TODO} and found that on average over XX\% of devices were exposed to known vulnerabilities. %TODO (drt24) proportion of devices exposed on average.
-There was also a period of several months when no devices ran secure versions of Android.
+We examine two platforms which attempt to enforce fine grained permissions for arbitrary code which runs on them: Android and Chrome.
+We compare the vulnerabilities find in them which break the sandbox with the distributions of versions running in the wild.
+We find that on average \percArdInsec\ of Android and \percCrmInsec\ of Chrome instances are vulnerable to known security vulnerabilities which break their sandbox, there was a period of several months when all Android devices were vulnerable.
+We examine some of the differences in incentives between the two platforms and how this has affected update timeliness.
+For Android we compare different manufacturers to find those best at shipping updates promptly.
 \end{abstract}
@@
 \section{Introduction}
-Android has XX\% of the smartphone market~\cite{TODO} and while the core development is controlled by Google there are at least XX~\cite{TODO} manufacturers which make devices which run Android.
-Many of these manufacturers customise the version of Android they ship and sometimes network operators (of which there are at least XX~\cite{TODO}) make further modifications.
+Android has \percArdMarketShare\ of the smartphone market~\cite{TODO} and while the core development is controlled by Google there are at least \daNumManufacturers\ manufacturers which make devices which run Android.
+Many of these manufacturers customise the version of Android they ship and sometimes network operators (of which there are at least \daNumNetworkOperators) make further modifications.
 Hence when Google produces an update to Android, the update may have to pass through the manufacturer and operator before reaching the user.
@@
 Figure \ref{fig:proportioninsecure} shows the proportion of Android devices in the DeviceAnalyzer data which were running versions of Android known to root exploits at the time.
+
+Google's Chrome Web browser has \percCrmMarketShare\ of the web browser market~\cite{TODO}, its core development is controlled by Google and Google is the only major distributor of it (the open source Chromium project from which it is derived does not have many direct users).%TODO unsubstantiated assertion
+
+Chrome uses various different sandboxing methods on different platforms such as Capsicum on BSD~\cite{TODO}
+
 In \S\ref{sec:background} we will explain where these data comes from and then in \S\ref{sec:results} we will show these data in more detail.

### 2014-02-10 QInF.tex
Add empty QInF proposal document

\documentclass[a4paper,twocolumn]{article}
\usepackage[utf8x]{inputenc}
\usepackage{ucs}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{tikz}
\usepackage{hyperref}
\usepackage{url}
\usepackage[numbers]{natbib}
\usepackage{paralist}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}

\author{Daniel Thomas}
\title{Android security updates}
\date{}%no date, it takes up space
\begin{document}
\maketitle

\section*{Introduction and problem definition}

\section*{Innovation proposal and relation to the state of the art}

\section*{One year horizon}

\section*{Conclusion}

\end{document}

### 2014-02-10 QInF.tex
QInF: fix preamble for bibliography
QInF: Start text

@@
 \documentclass[a4paper,twocolumn]{article}
-\usepackage[utf8x]{inputenc}
-\usepackage{ucs}
+\usepackage[utf8]{inputenc}
 \usepackage{amsmath}
@@
 \usepackage{url}
-\usepackage[numbers]{natbib}
 \usepackage{paralist}
 \usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
+\RequirePackage[date=terse,isbn=true,doi=true,url=false,maxbibnames=9,backref=true]{biblatex}
+\addbibresource{securityupdates.bib}
+\renewcommand{\bibfont}{\small}
+
+\AtEveryBibitem{% Clean up the bibtex rather than editing it
+ \clearname{editor} % remove editors
+}
 
@@
 
-\section*{Introduction and problem definition}
+\section*{Problem definition}
+The Internet of Things means that there will be many more internet connected devices, these will have software / firmware running on them.
+Security vulnerabilities will be found in this software and firmware which will then need to be updated.
+Android is one operating system which is being used for various Internet of Things devices (fridges, toasters etc.) and so examining the trends in security updates for Android gives us insight into how updates for Internet of Things devices might evolve.
+Our preliminary results indicate that most Android devices do not receive security updates in a timely manner.
+I intend to quantify this more precisely and compare different manufacturers to determine which are the best at shipping security updates promptly.
+Initial analysis indicates that perhaps the Qualcomm Innovation Center, Inc.
+(QuIC) is the best at following best practice\footnote{\url{https://www.codeaurora.org/projects/security-advisories}} and Google the best at shipping updates to devices promptly.
 
 \section*{Innovation proposal and relation to the state of the art}
+Presently there is not much visibility into the impact of security vulnerabilities on Android, many root equivalent vulnerabilities do not even have CVEs registered even when they are being actively exploited.
+It is also difficult to determine when vulnerabilities have been fixed and these fixes propagated to users.
+I have produced the Android Vulnerabilities website\footnote{\url{http://androidvulnerabilities.org/}} to collate this information.
+By using the database of root equivalent vulnerabilities that I have collected and the Device Analyzer\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/}}~\cite{TODO} data we can determine the proportion of devices running vulnerable versions of Android.
+We can then chart how that changes over time as new vulnerabilities are discovered and old ones are fixed, this is shown in Figure \ref{fig:propinsecure}.
+
+\begin{figure}
+ \includegraphics[width=0.5\textwidth]{figures/proportioninsecure.pdf}
+ \caption{Proportion of devices running vulnerable versions of Android}
+ \label{fig:propinsecure}
+\end{figure}
+
+There is still much to do to tie down details and give a more representative coverage of root equivalent vulnerabilities on Android.
+In particular manufacturer specific flaws are under-represented and there are many details which need to be tracked down to improve the accuracy of the results.
+
+Existing databases such as the CVE database\footnote{\url{http://www.cvedetails.com/}} lack information vital to performing these sorts of analyses such as dates for discovery, and release dates for fixes.
+Even finding release dates for Android versions is difficult as there is no canonical list.
+For example dates of tags in the Android Open Source Repository are also unreliable indicators of release dates, sometimes being tagged well after devices have started receiving updates to that version.
+Collating all the information together and checking its accuracy is a significant undertaking, the idea is to provide a good starting point and then use crowd-sourcing for the remainder once the Android Vulnerabilities website is a useful resource.
+We have already started getting submissions from other people to improve it.
 

### 2014-02-10 QInF.tex
Add some more text

@@
 \section*{Problem definition}
+Last year I proposed Nigori, a system for secure synchronised storage for Internet of Things devices, this relies on the security of the underlying device and so this year I am proposing looking at the security of the device operating system.
+
 The Internet of Things means that there will be many more internet connected devices, these will have software / firmware running on them.
@@
 
+Once we can make accurate comparisons between manufacturers regarding the timeliness with which they ship security updates to customers then we can publish this information and hence create an incentive for manufactures to ship updates more promptly.
+CESG which advises government on how to secure its computer system recommends picking Android devices from manufactures which are good at shipping security updates promptly~\cite{TODO} but it does not state which manufacturers are better.
+Enterprises with Bring Your Own Device (BYOD) policies or which supply smartphones for work purposes are likely to also be interested in this information as if an app can root a phone then any attempt to prevent data leakage are rendered useless.
+
 \section*{One year horizon}

### 2014-02-11 QInF.tex
QInF: Update text with comments from Alastair

@@
 Our preliminary results indicate that most Android devices do not receive security updates in a timely manner.
+Android devices are user facing and probably more likely to receive updates than fridges or water meters so the situation is likely to be worse.
 I intend to quantify this more precisely and compare different manufacturers to determine which are the best at shipping security updates promptly.
@@
 I have produced the Android Vulnerabilities website\footnote{\url{http://androidvulnerabilities.org/}} to collate this information.
+Root equivalent vulnerabilities are vulnerabilities which can be exploited either to gain root or to gain privileges which can then be used to gain root on the device.
 By using the database of root equivalent vulnerabilities that I have collected and the Device Analyzer\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/}}~\cite{TODO} data we can determine the proportion of devices running vulnerable versions of Android.
@@
 
-There is still much to do to tie down details and give a more representative coverage of root equivalent vulnerabilities on Android.
+There is still much to do, we have only just started.
+We need a more representative coverage of root equivalent vulnerabilities on Android, currently we only have a subset with many important details missing from the ones which we have.
 In particular manufacturer specific flaws are under-represented and there are many details which need to be tracked down to improve the accuracy of the results.
+There is also information which is not presently publicly available but which it might be possible to ask manufacturers and those who found vulnerabilities for in order to get a better picture of what is actually happening.
+Presently while we can chart the data we have, we cannot be sure that it is representative.
 

### 2014-02-11 QInF.tex
QInF: footnote urls to citations (not a paper so don't need to be proper references)

@@
 Initial analysis indicates that perhaps the Qualcomm Innovation Center, Inc.
-(QuIC) is the best at following best practice\footnote{\url{https://www.codeaurora.org/projects/security-advisories}} and Google the best at shipping updates to devices promptly.
+(QuIC) is the best at following best practice~\cite{codeaurora-security-advisories} and Google the best at shipping updates to devices promptly.
 
@@
 It is also difficult to determine when vulnerabilities have been fixed and these fixes propagated to users.
-I have produced the Android Vulnerabilities website\footnote{\url{http://androidvulnerabilities.org/}} to collate this information.
+I have produced the Android Vulnerabilities website~\cite{androidvulnerabilities.org} to collate this information.
 Root equivalent vulnerabilities are vulnerabilities which can be exploited either to gain root or to gain privileges which can then be used to gain root on the device.
-By using the database of root equivalent vulnerabilities that I have collected and the Device Analyzer\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/}}~\cite{TODO} data we can determine the proportion of devices running vulnerable versions of Android.
+By using the database of root equivalent vulnerabilities that I have collected and the Device Analyzer~\cite{Wagner2013} data we can determine the proportion of devices running vulnerable versions of Android.
 We can then chart how that changes over time as new vulnerabilities are discovered and old ones are fixed, this is shown in Figure \ref{fig:propinsecure}.
@@
 
-Existing databases such as the CVE database\footnote{\url{http://www.cvedetails.com/}} lack information vital to performing these sorts of analyses such as dates for discovery, and release dates for fixes.
+Existing databases such as the CVE database~\cite{cve-details} lack information vital to performing these sorts of analyses such as dates for discovery, and release dates for fixes.
 Even finding release dates for Android versions is difficult as there is no canonical list.

### 2014-02-11 QInF.tex
QInF: preliminary conclusion

@@
 \section*{Conclusion}
+We have collected some data on root equivalent vulnerabilities on Android and version information for deployed Android devices.
+Preliminary results are interesting and with more and better data we should be able to make useful recommendations about which manufacturers are best at shipping security updates.
+This in turn should incentivise manufactures to do a better job and so improve end user security.
+Improvements in end user security should give users more confidence in using their devices and this confidence is vital if the Internet of Things is to work out.
+We will hopefully also be able to analyse the incentives which have created the current situation and look at how those could be changed.
 

### 2014-02-11 QInF.tex
QInF: turn off backref as only 3 pages long total
QInF: add reference for CESG report

@@
 \usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
-\RequirePackage[date=terse,isbn=true,doi=true,url=false,maxbibnames=9,backref=true]{biblatex}
+\RequirePackage[date=terse,isbn=true,doi=true,url=false,maxbibnames=9,backref=false]{biblatex}
 \addbibresource{securityupdates.bib}
@@
 Once we can make accurate comparisons between manufacturers regarding the timeliness with which they ship security updates to customers then we can publish this information and hence create an incentive for manufactures to ship updates more promptly.
-CESG which advises government on how to secure its computer system recommends picking Android devices from manufactures which are good at shipping security updates promptly~\cite{TODO} but it does not state which manufacturers are better.
+CESG which advises government on how to secure its computer system recommends picking Android devices from manufactures which are good at shipping security updates promptly~\cite{CESG2013} but it does not state which manufacturers are better.
 Enterprises with Bring Your Own Device (BYOD) policies or which supply smartphones for work purposes are likely to also be interested in this information as if an app can root a phone then any attempt to prevent data leakage are rendered useless.

### 2014-02-11 QInF.tex
QInF: draft one year horizon

@@
 \section*{One year horizon}
+\begin{itemize}
+ \item A comprehensive database of root equivalent vulnerabilities affecting Android in a machine readable format on the website~\cite{androidvulnerabilities.org}.
+ \item Comparison of different manufacturers to determine which is best at shipping updates for security vulnerabilities.
+ \item Analysis of the behaviour of users with respect to Android updates -- how promptly do they get installed once made available.
+ \item What else???
+\end{itemize}
 

### 2014-02-12 QInF.tex
QInF: add figures and discuss them

@@
 
+Figure \ref{fig:da_norm_os} shows the proportion of devices in the Device Analyzer data running different Android OS versions and how that changes with time.
+Figure \ref{fig:norm_os_heat} shows the same data in a way which is easier to read when comparing how a particular version has changed with time.
+Figure \ref{fig:norm_api_heat} shows how the more coarse grained API version varies.
+
+Figure \ref{fig:propinsecure} shows the proportion of devices running versions of Android with known root vulnerabilities, Figure \ref{fig:nvulnerabilities_heat} breaks this down by vulnerability to show how the proportion of devices affected by different vulnerabilities varies.
+It clearly shows how the the APK signing vulnerabilities affected all devices and took a while to get fixed for any device.
+However what is perhaps more worrying is the long tail on vulnerabilities such as Gingerbreak, levitator and exploid which are more dangerous root exploits (not requiring new package installation) and which still affect a significant proportion of devices years after they were discovered.
+
+These graphs only show a small subset of the total vulnerabilities and only those which affect all Android devices, not any that are manufacturer or device specific.
+Hence observations of the proportion of devices affected in total are underestimates.
+\begin{figure}
+ \includegraphics[width=0.5\textwidth]{figures/da_norm_os.pdf}
+ \caption{Proportion of devices running different OS versions in the DA data (see also \ref{fig:norm_os_heat})}
+ \label{fig:da_norm_os}
+\end{figure}
+\begin{figure}
+ \includegraphics[width=0.5\textwidth]{figures/norm_os_heat.pdf}
+ \caption{Proportion of devices running different OS versions in the DA data (see also \ref{fig:da_norm_os})}
+ \label{fig:norm_os_heat}
+\end{figure}
+\begin{figure}
+ \includegraphics[width=0.5\textwidth]{figures/norm_api_heat.pdf}
+ \caption{Proportion of devices running different API versions in the DA data}
+ \label{fig:norm_api_heat}
+\end{figure}
+\begin{figure}
+ \includegraphics[width=0.5\textwidth]{figures/nvulnerabilities_heat.pdf}
+ \caption{Proportion of devices affected by different vulnerabilities}
+ \label{fig:nvulnerabilities_heat}
+\end{figure}
+
+Figure \ref{fig:from_to_updates} shows how devices upgrade between different versions of Android.
+Device Analyzer observes upgrade events and the figure shows which versions a device is upgrading from and to.
+Mostly the cells with version change are above the diagonal and are upgrades, while many upgrades are from one version to the next version there are also a fair number which skip a few versions.
+Surprisingly there are also a small number of downgrade events when older versions of Android are installed on to devices.
+Figure \ref{fig:weekly_security_updates} shows the number of devices getting security updates each week, blue shows updates which changed the Android version number from a version with known vulnerabilities to one which had fewer known vulnerabilities.
+Red indicates updates which changed the build number but not the version number and so might contain a backported fix for a vulnerability.
+
+\begin{figure}
+ \includegraphics[width=0.5\textwidth]{figures/from_to_updates.pdf}
+ \caption{Updates between different Android versions in the DA data}
+ \label{fig:from_to_updates}
+\end{figure}
+\begin{figure}
+ \includegraphics[width=0.5\textwidth]{figures/w_security_updates.pdf}
+ \caption{Number of updates each week which fixed or may have fixed security vulnerabilities}
+ \label{fig:weekly_security_updates}
+\end{figure}
+
 \section*{One year horizon}

### 2014-02-12 QInF.tex
QInF: reworking based on Alastair's comments

@@
 \author{Daniel Thomas}
-\title{Android security updates}
+\title{Android security updates: late or never}
 \date{}%no date, it takes up space
@@
 \section*{Problem definition}
-Last year I proposed Nigori, a system for secure synchronised storage for Internet of Things devices, this relies on the security of the underlying device and so this year I am proposing looking at the security of the device operating system.
+I am working on improving the security of Internet of Things devices, looking at both secure synchronisation and the security of the underlying device operating system.
 
-The Internet of Things means that there will be many more internet connected devices, these will have software / firmware running on them.
-Security vulnerabilities will be found in this software and firmware which will then need to be updated.
-Android is one operating system which is being used for various Internet of Things devices (fridges, toasters etc.) and so examining the trends in security updates for Android gives us insight into how updates for Internet of Things devices might evolve.
+The Internet of Things (IoT) means that there will be many more internet connected devices, which will run software.
+Security vulnerabilities will be found in this software which will then need to be updated.
+Android is one operating system which is being used for Internet of Things devices (fridges, toasters etc.) and so examining the trends in security updates for Android gives us insight into how updates for Internet of Things devices (IoT) might evolve.
 Our preliminary results indicate that most Android devices do not receive security updates in a timely manner.
-Android devices are user facing and probably more likely to receive updates than fridges or water meters so the situation is likely to be worse.
-I intend to quantify this more precisely and compare different manufacturers to determine which are the best at shipping security updates promptly.
+Android devices are user facing and probably more likely to receive updates than fridges or water meters so the situation is likely to be worse in the IoT.
+I intend to quantify this and compare different manufacturers to determine which are the best at shipping security updates promptly.
 Initial analysis indicates that perhaps the Qualcomm Innovation Center, Inc.
-(QuIC) is the best at following best practice~\cite{codeaurora-security-advisories} and Google the best at shipping updates to devices promptly.
+(QuIC) is the best at following best practice and making patches available~\cite{codeaurora-security-advisories} and Google the fastest at shipping updates to devices promptly.
+However I do not yet have sufficient data to make robust conclusions.
 
-\section*{Innovation proposal and relation to the state of the art}
-Presently there is not much visibility into the impact of security vulnerabilities on Android, many root equivalent vulnerabilities do not even have CVEs registered even when they are being actively exploited.
+\section*{Innovation proposal}
+Presently there is not much visibility into the impact of security vulnerabilities on Android, many root equivalent vulnerabilities do not have CVEs registered even when they are being actively exploited.
 It is also difficult to determine when vulnerabilities have been fixed and these fixes propagated to users.
-I have produced the Android Vulnerabilities website~\cite{androidvulnerabilities.org} to collate this information.
-Root equivalent vulnerabilities are vulnerabilities which can be exploited either to gain root or to gain privileges which can then be used to gain root on the device.
-By using the database of root equivalent vulnerabilities that I have collected and the Device Analyzer~\cite{Wagner2013} data we can determine the proportion of devices running vulnerable versions of Android.
+I have produced the Android Vulnerabilities website~\cite{androidvulnerabilities.org} to collate such information, particularly on root equivalent vulnerabilities.
+This database is publicly available in a machine readable format (JSON) and the code is open source.
+Root equivalent vulnerabilities are vulnerabilities which can be exploited either to gain root or to gain privileges which can then be used to either gain root on the device or do things which only root would normally be able to do.
+By using the database of root equivalent vulnerabilities that I have collected, together with the data we have collected with Device Analyzer~\cite{Wagner2013} we can determine the proportion of devices running vulnerable versions of Android.
 We can then chart how that changes over time as new vulnerabilities are discovered and old ones are fixed, this is shown in Figure \ref{fig:propinsecure}.
@@
 
-There is still much to do, we have only just started.
+We have only just started and there is still much to do to improve the quantity and quality of the data so that it is useful.
 We need a more representative coverage of root equivalent vulnerabilities on Android, currently we only have a subset with many important details missing from the ones which we have.
-In particular manufacturer specific flaws are under-represented and there are many details which need to be tracked down to improve the accuracy of the results.
+The improvement in the middle of Figure \ref{fig:propinsecure} might be due to not yet having sufficient data to include any of the vulnerabilities found in 2012 (we know of several).
+In particular manufacturer specific flaws are under-represented and there are many details which need to be tracked down to improve the reliability of the results.
 There is also information which is not presently publicly available but which it might be possible to ask manufacturers and those who found vulnerabilities for in order to get a better picture of what is actually happening.
@@
 Even finding release dates for Android versions is difficult as there is no canonical list.
-For example dates of tags in the Android Open Source Repository are also unreliable indicators of release dates, sometimes being tagged well after devices have started receiving updates to that version.
+For example: dates of tags in the Android Open Source Repository are unreliable indicators of release dates, sometimes being tagged well after devices have started receiving updates to that version.
 Collating all the information together and checking its accuracy is a significant undertaking, the idea is to provide a good starting point and then use crowd-sourcing for the remainder once the Android Vulnerabilities website is a useful resource.
@@
 Once we can make accurate comparisons between manufacturers regarding the timeliness with which they ship security updates to customers then we can publish this information and hence create an incentive for manufactures to ship updates more promptly.
+For example we could produce press releases or badges stating which companies are most prompt at shipping updates.
 CESG which advises government on how to secure its computer system recommends picking Android devices from manufactures which are good at shipping security updates promptly~\cite{CESG2013} but it does not state which manufacturers are better.
-Enterprises with Bring Your Own Device (BYOD) policies or which supply smartphones for work purposes are likely to also be interested in this information as if an app can root a phone then any attempt to prevent data leakage are rendered useless.
+Enterprises with Bring Your Own Device (BYOD) policies or which supply smartphones for work purposes are likely to also be interested in this information.
+Since if an app can root a phone then any attempt to prevent data leakage are rendered useless.
 
-Figure \ref{fig:da_norm_os} shows the proportion of devices in the Device Analyzer data running different Android OS versions and how that changes with time.
-Figure \ref{fig:norm_os_heat} shows the same data in a way which is easier to read when comparing how a particular version has changed with time.
-Figure \ref{fig:norm_api_heat} shows how the more coarse grained API version varies.
+We determine whether a particular device is vulnerable to a vulnerability by computing which versions of Android were vulnerable to that vulnerability on a particular day from the vulnerabilities database and comparing that with the proportion of devices running that version, which is shown in Figures \ref{fig:da_norm_os} and \ref{fig:norm_os_heat}.
+They show how the distribution of versions changes with time and how the popularity of each version changes.
 
-Figure \ref{fig:propinsecure} shows the proportion of devices running versions of Android with known root vulnerabilities, Figure \ref{fig:nvulnerabilities_heat} breaks this down by vulnerability to show how the proportion of devices affected by different vulnerabilities varies.
-It clearly shows how the the APK signing vulnerabilities affected all devices and took a while to get fixed for any device.
-However what is perhaps more worrying is the long tail on vulnerabilities such as Gingerbreak, levitator and exploid which are more dangerous root exploits (not requiring new package installation) and which still affect a significant proportion of devices years after they were discovered.
+The variation of the proportion of devices affected by a vulnerability with time tells us how bad a particular vulnerability affected the Android platform.
+Figure \ref{fig:nvulnerabilities_heat} breaks this down by vulnerability to show how the proportion of devices affected by different vulnerabilities varies.
+In 2013 three vulnerabilities were found in the way which Android verified the signatures on APKs.
+These allowed the creation of malicious APKs which appear to be signed as system APKs -- which have root equivalent privileges.
+Figure \ref{fig:nvulnerabilities_heat} shows how the the APK signing vulnerabilities affected all devices and took months to get fixed for any device.
+However what is perhaps more worrying is the long tail on the Gingerbreak, levitator and exploid vulnerabilities which are more dangerous root exploits (not requiring new APK installation) and which still affect a significant proportion of devices years after they were discovered.
 
@@
 \begin{figure}
- \includegraphics[width=0.5\textwidth]{figures/norm_api_heat.pdf}
- \caption{Proportion of devices running different API versions in the DA data}
- \label{fig:norm_api_heat}
-\end{figure}
-\begin{figure}
  \includegraphics[width=0.5\textwidth]{figures/nvulnerabilities_heat.pdf}
@@
 \end{figure}
-
-Figure \ref{fig:from_to_updates} shows how devices upgrade between different versions of Android.
-Device Analyzer observes upgrade events and the figure shows which versions a device is upgrading from and to.
-Mostly the cells with version change are above the diagonal and are upgrades, while many upgrades are from one version to the next version there are also a fair number which skip a few versions.
-Surprisingly there are also a small number of downgrade events when older versions of Android are installed on to devices.
-Figure \ref{fig:weekly_security_updates} shows the number of devices getting security updates each week, blue shows updates which changed the Android version number from a version with known vulnerabilities to one which had fewer known vulnerabilities.
-Red indicates updates which changed the build number but not the version number and so might contain a backported fix for a vulnerability.
 
@@
 \end{figure}
+Figure \ref{fig:from_to_updates} shows how devices upgrade between different versions of Android.
+Device Analyzer observes upgrade events and the figure shows which versions a device is upgrading from and to.
+Mostly the dark cells are above the diagonal and are upgrades.
+While many upgrades are from one version to the next version there are also a fair number which skip a few versions.
+Surprisingly there are also a small number of downgrade events when older versions of Android are installed on to devices.
+Figure \ref{fig:weekly_security_updates} shows the number of devices getting security updates each week, blue shows updates which changed the Android version number from a version with known vulnerabilities to one which had fewer known vulnerabilities.
+Red indicates updates which changed the build number but not the version number and so might contain a backported fix for a vulnerability.
+We need better data to detect backporting of security fixes.
 
@@
  \item Analysis of the behaviour of users with respect to Android updates -- how promptly do they get installed once made available.
- \item What else???
+ \item A paper publishing the results.
+ \item Updates to Device Analyzer to support further collection of useful data.
+ \item Crowd sourcing of data on Android vulnerabilities and disclosure practices.
+ \item Badges or other useful feedback for manufacturers following best practice.
 \end{itemize}

### 2014-02-12 QInF.tex
Corrections based on feedback from James

@@
 \section*{Problem definition}
-I am working on improving the security of Internet of Things devices, looking at both secure synchronisation and the security of the underlying device operating system.
+I am working on improving the security of Internet of Things (IoT) devices, looking at both secure synchronisation of data and the security of the underlying device operating system.
 
-The Internet of Things (IoT) means that there will be many more internet connected devices, which will run software.
+The Internet of Things means that there will be many more internet connected devices, which will run complex software.
 Security vulnerabilities will be found in this software which will then need to be updated.
-Android is one operating system which is being used for Internet of Things devices (fridges, toasters etc.) and so examining the trends in security updates for Android gives us insight into how updates for Internet of Things devices (IoT) might evolve.
+Android is one operating system which is being used for Internet of Things devices (fridges, toasters etc.) and so examining the trends in security updates for Android gives us insight into how updates for Internet of Things devices might evolve.
 Our preliminary results indicate that most Android devices do not receive security updates in a timely manner.
@@
 It is also difficult to determine when vulnerabilities have been fixed and these fixes propagated to users.
-I have produced the Android Vulnerabilities website~\cite{androidvulnerabilities.org} to collate such information, particularly on root equivalent vulnerabilities.
-This database is publicly available in a machine readable format (JSON) and the code is open source.
+I have started the Android Vulnerabilities website~\cite{androidvulnerabilities.org} to collate such information, particularly on root equivalent vulnerabilities.
+The current database is publicly available in a machine readable format (JSON) and the code is open source.
 Root equivalent vulnerabilities are vulnerabilities which can be exploited either to gain root or to gain privileges which can then be used to either gain root on the device or do things which only root would normally be able to do.
-By using the database of root equivalent vulnerabilities that I have collected, together with the data we have collected with Device Analyzer~\cite{Wagner2013} we can determine the proportion of devices running vulnerable versions of Android.
+By using the database of root equivalent vulnerabilities that I have collected, together with the data we have collected with Device Analyzer (DA)~\cite{Wagner2013} we can start to determine the proportion of devices running vulnerable versions of Android.
 We can then chart how that changes over time as new vulnerabilities are discovered and old ones are fixed, this is shown in Figure \ref{fig:propinsecure}.
@@
 CESG which advises government on how to secure its computer system recommends picking Android devices from manufactures which are good at shipping security updates promptly~\cite{CESG2013} but it does not state which manufacturers are better.
-Enterprises with Bring Your Own Device (BYOD) policies or which supply smartphones for work purposes are likely to also be interested in this information.
-Since if an app can root a phone then any attempt to prevent data leakage are rendered useless.
+Enterprises with Bring Your Own Device policies or which supply smartphones for work purposes are likely to also be interested in this information.
+If an app can root a phone then any attempt to prevent data leakage are rendered useless.
 
-We determine whether a particular device is vulnerable to a vulnerability by computing which versions of Android were vulnerable to that vulnerability on a particular day from the vulnerabilities database and comparing that with the proportion of devices running that version, which is shown in Figures \ref{fig:da_norm_os} and \ref{fig:norm_os_heat}.
+We determine whether a particular device is vulnerable to a vulnerability by computing which versions of Android were vulnerable to that vulnerability on a particular day from the vulnerabilities database and comparing that with the proportion of devices running that version, this is shown in Figures \ref{fig:da_norm_os} and \ref{fig:norm_os_heat}.
 They show how the distribution of versions changes with time and how the popularity of each version changes.
@@
 These allowed the creation of malicious APKs which appear to be signed as system APKs -- which have root equivalent privileges.
-Figure \ref{fig:nvulnerabilities_heat} shows how the the APK signing vulnerabilities affected all devices and took months to get fixed for any device.
-However what is perhaps more worrying is the long tail on the Gingerbreak, levitator and exploid vulnerabilities which are more dangerous root exploits (not requiring new APK installation) and which still affect a significant proportion of devices years after they were discovered.
+Figure \ref{fig:nvulnerabilities_heat} shows how the the \emph{APK signing vulnerabilities} affected all devices and took months to get fixed for any device.
+However what is perhaps more worrying is the long tail on the \emph{Gingerbreak}, \emph{levitator} and \emph{exploid} vulnerabilities which are more dangerous root exploits (not requiring new APK installation) and which still affect a significant proportion of devices years later.
 
@@
 Figure \ref{fig:from_to_updates} shows how devices upgrade between different versions of Android.
-Device Analyzer observes upgrade events and the figure shows which versions a device is upgrading from and to.
+Device Analyzer (DA) observes upgrade events and the figure shows which versions a device is upgrading from and to.
 Mostly the dark cells are above the diagonal and are upgrades.

### 2014-02-14 securityupdates.tex
Finish sentance in securityupdates
Add some missing citations and urls

@@
 \documentclass[conference,a4paper,twoside]{IEEEtran}
-\usepackage[utf8x]{inputenc}
-\usepackage{ucs}
+\usepackage[utf8]{inputenc}
 \usepackage{amsmath}
@@
 \usepackage{hyperref}
+\RequirePackage[date=terse,isbn=true,doi=true,url=false,maxbibnames=9,backref=false]{biblatex}
+\addbibresource{securityupdates.bib}
+\renewcommand{\bibfont}{\small}
+
+\AtEveryBibitem{% Clean up the bibtex rather than editing it
+ \clearname{editor} % remove editors
+}
+
 \author{Daniel R.\ Thomas, Daniel Wagner, Alastair R.\ Beresford and Andrew Rice}
@@
 The manufacturer and user have no financial incentive to perform this work.
-There is ongoing legal action to force some of them to do so~\cite{TODO}.
+There is ongoing legal action to force some of them to do so~\cite{Soghoian2013}.
+CESG which advises the UK government on how to secure its computer system recommends picking Android devices from manufactures which are good at shipping security updates promptly~\cite{CESG2013} but it does not state which manufacturers are better.
 
@@
 
-Figure \ref{fig:proportioninsecure} shows the proportion of Android devices in the DeviceAnalyzer data which were running versions of Android known to root exploits at the time.
+Figure \ref{fig:proportioninsecure} shows the proportion of Android devices in the Device Analyzer data which were running versions of Android known to root exploits at the time.
 
@@
 
-Chrome uses various different sandboxing methods on different platforms such as Capsicum on BSD~\cite{TODO}
+Chrome uses various different sandboxing methods on different platforms such as Capsicum on BSD~\cite{Watson2010} or using a SUID binary to create new namespaces and chroot on Linux~\cite{TODO}.
 
@@
 We have published full details of the vulnerabilities listed below in an accompanying technical report~\cite{TODO} and summarise them in Table \ref{tab:andvulns}.
-We also have a website\footnote{\url{TODO}} which we will keep up to date with future vulnerabilities.
+We also have a website\footnote{\url{http://androidvulnerabilities.org/}} which we will keep up to date with future vulnerabilities.
 
@@
 We also need historical information on which versions of Android were running on devices along with other information about those devices which allows us to go into more detail and confirm the reliability of the data our results depend on.
-The only suitable data which is available for this is the DeviceAnalyzer data~\cite{TODO}.
-The DeviceAnalyzer data contains data from ??? devices with a total of ??? device days, it contains data for more than ??? months for ??? devices.
-%TODO talk about the DeviceAnalyzer project a bit
-Various different kinds of data\footnote{\url{TODO}} are collected by the DeviceAnalyzer Android app\footnote{\url{TODO}}.
+The only suitable data which is available for this is the Device Analyzer data~\cite{Wagner2013}.
+The Device Analyzer data contains data from ??? devices with a total of ??? device days, it contains data for more than ??? months for ??? devices.
+%TODO talk about the Device Analyzer project a bit
+Various different kinds of data\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/collected.htm}} are collected by the Device Analyzer Android app\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/}}.
 Among these are the build string and API version for the version of Android currently running on the device each day.
@@
 
-Figure \ref{fig:os} shows the raw data on the number of devices in the DeviceAnalyzer data running different versions of Android each day.
+Figure \ref{fig:os} shows the raw data on the number of devices in the Device Analyzer data running different versions of Android each day.
 It shows how old versions are gradually replaced by new ones, the long tail of devices which do not see updates to more recent versions and the fluctuation in the number of devices in the data.
@@
 \includegraphics[width=\columnwidth]{figures/os}
-\caption{Android versions in DeviceAnalyzer data over time}
+\caption{Android versions in Device Analyzer data over time}
 \label{fig:os}
@@
 For each vulnerability it shows the proportion of devices exposed to that vulnerability and how that changes over time.
-In July 2011 at the beginning of the DeviceAnalyzer data the Exploid and Levitator vulnerabilities both affect most Android devices, slowly these are fixed as updates roll out and devices are replaced until in January 2013 a much smaller proportion of devices are affected by known vulnerabilities.
+In July 2011 at the beginning of the Device Analyzer data the Exploid and Levitator vulnerabilities both affect most Android devices, slowly these are fixed as updates roll out and devices are replaced until in January 2013 a much smaller proportion of devices are affected by known vulnerabilities.
 However when in February 2013 the first APK signing vulnerability was found which affected all previous versions of Android and even in October 2013 most devices were still unfixed.

### 2014-02-17 securityupdates.tex
Add some results from QInF

@@
 
+The variation of the proportion of devices affected by a vulnerability with time tells us how bad a particular vulnerability affected the Android platform.
+Figure \ref{fig:nvulnerabilities_heat} breaks this down by vulnerability to show how the proportion of devices affected by different vulnerabilities varies.
+In 2013 three vulnerabilities were found in the way which Android verified the signatures on APKs.
+These allowed the creation of malicious APKs which appear to be signed as system APKs -- which have root equivalent privileges.
+Figure \ref{fig:nvulnerabilities_heat} shows how the the \emph{APK signing vulnerabilities} affected all devices and took months to get fixed for any device.
+However what is perhaps more worrying is the long tail on the \emph{Gingerbreak}, \emph{levitator} and \emph{exploid} vulnerabilities which are more dangerous root exploits (not requiring new APK installation) and which still affect a significant proportion of devices years later.
+
+\begin{figure}
+ \includegraphics[width=0.5\textwidth]{figures/nvulnerabilities_heat.pdf}
+ \caption{Proportion of devices affected by different vulnerabilities}
+ \label{fig:nvulnerabilities_heat}
+\end{figure}
+\begin{figure}
+ \includegraphics[width=0.5\textwidth]{figures/from_to_updates.pdf}
+ \caption{Updates between different Android versions in the DA data}
+ \label{fig:from_to_updates}
+\end{figure}
+\begin{figure}
+ \includegraphics[width=0.5\textwidth]{figures/w_security_updates.pdf}
+ \caption{Number of updates each week which fixed or may have fixed security vulnerabilities}
+ \label{fig:weekly_security_updates}
+\end{figure}
+Figure \ref{fig:from_to_updates} shows how devices upgrade between different versions of Android.
+Device Analyzer (DA) observes upgrade events and the figure shows which versions a device is upgrading from and to.
+Mostly the dark cells are above the diagonal and are upgrades.
+While many upgrades are from one version to the next version there are also a fair number which skip a few versions.
+Surprisingly there are also a small number of downgrade events when older versions of Android are installed on to devices.
+Figure \ref{fig:weekly_security_updates} shows the number of devices getting security updates each week, blue shows updates which changed the Android version number from a version with known vulnerabilities to one which had fewer known vulnerabilities.
+Red indicates updates which changed the build number but not the version number and so might contain a backported fix for a vulnerability.
+
 \subsection{Update frequency}

### 2014-02-18 securityupdates.tex
Revert "Replace abstract and start talking about Chrome"
This reverts commit 951e4cb7ce8b935f143de6402b5737f66ac3b61f.

Conflicts:
	securityupdates.tex

@@
 \begin{document}
-\title{Timeliness of security updates for Android and Chrome}
+\title{Timeliness of security updates for Android}
 
@@
 \begin{abstract}
-We examine two platforms which attempt to enforce fine grained permissions for arbitrary code which runs on them: Android and Chrome.
-We compare the vulnerabilities find in them which break the sandbox with the distributions of versions running in the wild.
-We find that on average \percArdInsec\ of Android and \percCrmInsec\ of Chrome instances are vulnerable to known security vulnerabilities which break their sandbox, there was a period of several months when all Android devices were vulnerable.
-We examine some of the differences in incentives between the two platforms and how this has affected update timeliness.
-For Android we compare different manufacturers to find those best at shipping updates promptly.
+Android is the most popular smartphone platform and seeks to provide better security than popular desktop operating systems through increased compartmentalisation and a permission system for apps.
+This security relies on there not being root privilege exploits which allow apps to bypass all protections.
+Many such exploits have been found and fixed.
+Our hypothesis is that these fixes do not promptly reach the users devices and so much of the time users are running versions of Android known to be vulnerable.
+We used the DeviceAnalyzer data\cite{TODO} and found that on average over XX\% of devices were exposed to known vulnerabilities. %TODO (drt24) proportion of devices exposed on average.
+There was also a period of several months when no devices ran secure versions of Android.
 \end{abstract}
@@
 \section{Introduction}
-Android has \percArdMarketShare\ of the smartphone market~\cite{TODO} and while the core development is controlled by Google there are at least \daNumManufacturers\ manufacturers which make devices which run Android.
-Many of these manufacturers customise the version of Android they ship and sometimes network operators (of which there are at least \daNumNetworkOperators) make further modifications.
+Android has XX\% of the smartphone market~\cite{TODO} and while the core development is controlled by Google there are at least XX~\cite{TODO} manufacturers which make devices which run Android.
+Many of these manufacturers customise the version of Android they ship and sometimes network operators (of which there are at least XX~\cite{TODO}) make further modifications.
 Hence when Google produces an update to Android, the update may have to pass through the manufacturer and operator before reaching the user.
@@
 
-Figure \ref{fig:proportioninsecure} shows the proportion of Android devices in the Device Analyzer data which were running versions of Android known to root exploits at the time.
-
-Google's Chrome Web browser has \percCrmMarketShare\ of the web browser market~\cite{TODO}, its core development is controlled by Google and Google is the only major distributor of it (the open source Chromium project from which it is derived does not have many direct users).%TODO unsubstantiated assertion
-
-Chrome uses various different sandboxing methods on different platforms such as Capsicum on BSD~\cite{Watson2010} or using a SUID binary to create new namespaces and chroot on Linux~\cite{TODO}.
-
+Figure \ref{fig:proportioninsecure} shows the proportion of Android devices in the DeviceAnalyzer data which were running versions of Android known to root exploits at the time.
 In \S\ref{sec:background} we will explain where these data comes from and then in \S\ref{sec:results} we will show these data in more detail.

### 2014-02-18 securityupdates.tex
Add missing references on proportion of malware with root exploits
Adjust preamble to use macros for variables we might change

@@
 
-\newcommand{\percArdMarketShare}{XX\%}
-\newcommand{\percArdInsec}{XX\%}
-\newcommand{\percCrmMarketShare}{XX\%}
-\newcommand{\percCrmInsec}{X\%}
-\newcommand{\daNumManufacturers}{XX}
-\newcommand{\daNumNetworkOperators}{XXX}
+\newcommand{\percMarketShare}{XX\%~\cite{TODO}}
+\newcommand{\percInsec}{XX\%~\cite{TODO}}
+\newcommand{\daNumManufacturers}{XX~[\S\ref{sec:results}]}
+\newcommand{\daNumNetworkOperators}{XXX~[\S\ref{sec:results}]}
 
@@
 \section{Introduction}
-Android has XX\% of the smartphone market~\cite{TODO} and while the core development is controlled by Google there are at least XX~\cite{TODO} manufacturers which make devices which run Android.
-Many of these manufacturers customise the version of Android they ship and sometimes network operators (of which there are at least XX~\cite{TODO}) make further modifications.
+Android has \percMarketShare of the smartphone market and while the core development is controlled by Google there are at least \daNumManufacturers manufacturers which make devices which run Android.
+Many of these manufacturers customise the version of Android they ship and sometimes network operators (of which there are at least \daNumNetworkOperators) make further modifications.
 Hence when Google produces an update to Android, the update may have to pass through the manufacturer and operator before reaching the user.
@@
 Manufacturers of Android devices have also added further root exploits when customising it~\cite{TODO}.
-Between XX\%~\cite{TODO} and XX\%~\cite{TODO} of malware for Android contains root exploits.
+Between 36.7\%~\cite{Zhou2012b} and 40\%~\cite{Zhou2012a} of malware for Android contains root exploits.
 While this shows that much Android malware does not need a root exploit to work, it can just request the relevant permissions and the user will grant them, a significant proportion does.

### 2014-02-18 securityupdates.tex
Add missing refs and typography fixes

@@
 Our hypothesis is that these fixes do not promptly reach the users devices and so much of the time users are running versions of Android known to be vulnerable.
-We used the DeviceAnalyzer data\cite{TODO} and found that on average over XX\% of devices were exposed to known vulnerabilities. %TODO (drt24) proportion of devices exposed on average.
+We used the DeviceAnalyzer data~\cite{Wagner2013} and found that on average over \percInsec\ of devices were exposed to known vulnerabilities.
 There was also a period of several months when no devices ran secure versions of Android.
+We compare different manufacturers and operators to find which are best at promptly distributing security updates and examine how to improve the situation.
 \end{abstract}
@@
 \section{Introduction}
-Android has \percMarketShare of the smartphone market and while the core development is controlled by Google there are at least \daNumManufacturers manufacturers which make devices which run Android.
+Android has \percMarketShare\ of the smartphone market and while the core development is controlled by Google there are at least \daNumManufacturers\ manufacturers which make devices which run Android.
 Many of these manufacturers customise the version of Android they ship and sometimes network operators (of which there are at least \daNumNetworkOperators) make further modifications.
@@
 The Linux kernel is a large piece of software and so there will always be new root exploits found in it~\cite{TODO}.
-Manufacturers of Android devices have also added further root exploits when customising it~\cite{TODO}.
+Manufacturers of Android devices have also added further root exploits when customising it~\cite{Grace2012}.
 Between 36.7\%~\cite{Zhou2012b} and 40\%~\cite{Zhou2012a} of malware for Android contains root exploits.

### 2014-02-18 securityupdates.tex
Add comment about Windows XP support period to abstract
Correct introduction

@@
 We compare different manufacturers and operators to find which are best at promptly distributing security updates and examine how to improve the situation.
+Many phones are sold on long (18 month) contracts with monthly payments yet stop receiving updates before the end of the contract, while Windows XP could be purchased for a one off payment in October 2001 and yet will receive security updates until April 2014.
 \end{abstract}
@@
 Hence when Google produces an update to Android, the update may have to pass through the manufacturer and operator before reaching the user.
-The manufacturer and user have no financial incentive to perform this work.
+The manufacturer and operator have no financial incentive to perform this work.
 There is ongoing legal action to force some of them to do so~\cite{Soghoian2013}.

### 2014-02-18 securityupdates.tex
Try and work in argument about lack of information making root exploits problematic

@@
 \newcommand{\daNumNetworkOperators}{XXX~[\S\ref{sec:results}]}
+\newcommand{\daNumDevices}{XXXX}
+\newcommand{\daDeviceDays}{XXXXX}
+\newcommand{\daMonths}{X}% Number of months for which we have the below number of devices
+\newcommand{\daMonthsDevices}{XXX}
+\newcommand{\daAdbPerc}{X\%}
 
@@
 
+Part of the reason why deployed devices do not receive updates in a timely manner is that there is little data on what is actually happening and so it is not possible to determine which manufacturers or operators are doing a better job.
+Hence there is little incentive for manufacturers and operators to do a better job, we will discuss this further in \S\ref{sec:economics}
+
 \section{Background}
@@
 \emph{Root equivalent} means that it grants privileges equivalent in scope to root which can then be used to gain root.
-Some phones can be `rooted' by enabling USB debugging and using the special privileges of the adb shell to root the device.
+Some phones can be `rooted' by enabling USB debugging and using the special privileges of the adb shell to root the device but only \daAdbPerc\ of devices have USB debugging enabled.
 This is not something that applications running on the phone can exploit to break compartmentalisation and so we do not include those exploits.
-Some exploits which grant root are not traditional kernel exploits, for example the discovery of flaws in the verification of signatures on Android applications in ???
-2013 meant that applications could pretend to be signed with system keys and hence gain root privileges.
+Some exploits which grant root are not traditional kernel exploits, for example the discovery of flaws in the verification of signatures on Android applications in February 2013~\cite{Forristal2013} meant that applications could pretend to be signed with system keys and hence gain root privileges.
 
 We have published full details of the vulnerabilities listed below in an accompanying technical report~\cite{TODO} and summarise them in Table \ref{tab:andvulns}.
-We also have a website\footnote{\url{http://androidvulnerabilities.org/}} which we will keep up to date with future vulnerabilities.
+We also have a website\footnote{\url{http://androidvulnerabilities.org/}} which we will keep up to date with future vulnerabilities and which has successfully solicited submissions from others.
 
@@
 The only suitable data which is available for this is the Device Analyzer data~\cite{Wagner2013}.
-The Device Analyzer data contains data from ??? devices with a total of ??? device days, it contains data for more than ??? months for ??? devices.
+The Device Analyzer data contains data from \daNumDevices\ devices with a total of \daDeviceDays\ device days, it contains data for more than \daMonths months for \daMonthsDevices\ devices.
 %TODO talk about the Device Analyzer project a bit

### 2014-02-18 securityupdates.tex
Add subsection on monetisation

@@
 
+\subsection{Monetisation}
+One of the key differences between smartphones and desktops is that it is easier to directly cash out on a smartphone as malware can send premium rate texts or make premium rate phone calls.
+It can also steal personal data which already has semantic information associated with it, this also makes it easier to perform ransomware attacks where malware deletes this information from the device and charges a ransom for its return.
+It can also do more conventional things such as use the devices for DDoS attacks and altering the ads displayed to send the fees to the malware author.
+While many of these things can be done by malware which just requests the relevant permissions, root exploits make this easier, particularly for ransomware and altering the ads displayed.
+
 \section{Related work}

### 2014-02-18 securityupdates.tex
Add API graphs to show DA data representative and a normalised OS graph

@@
 
+The data from Device Analyzer (DA) we want to use to investigate the proportion of devices exposed to different vulnerabilities is the OS version.
+Unfortunately there is no authoritative source of OS version information and so we cannot check directly whether our data is representative.
+However Google has published API version information every month since December 2009 and we have collated this information.\footnote{\url{https://www.cl.cam.ac.uk/~drt24/android/versiondistribution/}}
+While API versions are too coarse grained to use for security update detection they are closely related to OS versions and so if the DA data on API versions is similar to the Google Data on API versions then the DA data on OS versions should be representative.
+Figure \ref{fig:play_api} shows the data Google and Figure \ref{fig:da_api} and they appear similar. %TODO we want a statistical metric to claim this strongly with.
+\begin{figure}
+ \centering
+ \includegraphics[width=\columnwidth]{figures/play_norm_api_trunc}
+ \caption{Google Play data on proportion of devices running different Android API versions}
+ \label{fig:play_api}
+\end{figure}
+\begin{figure}
+ \centering
+ \includegraphics[width=\columnwidth]{figures/da_norm_api}
+ \caption{Google Play data on proportion of devices running different Android API versions}
+ \label{fig:da_api}
+\end{figure}
+
+This allows us to have confidence in the OS version information.
 Figure \ref{fig:os} shows the raw data on the number of devices in the Device Analyzer data running different versions of Android each day.
 It shows how old versions are gradually replaced by new ones, the long tail of devices which do not see updates to more recent versions and the fluctuation in the number of devices in the data.
-\begin{figure}%[!b]
+Figure \ref{fig:norm_os} shows the normalised OS version distribution.
+\begin{figure}
  \centering
@@
  \label{fig:os}
+\end{figure}
+\begin{figure}
+ \centering
+ \includegraphics[width=\columnwidth]{figures/da_norm_os}
+ \caption{Android versions in Device Analyzer data over time}
+ \label{fig:norm_os}
 \end{figure}

### 2014-02-18 securityupdates.tex
Daniel T. Wagner rather than Daniel Wagner

@@
 
-\author{Daniel R.\ Thomas, Daniel Wagner, Alastair R.\ Beresford and Andrew Rice}
+\author{Daniel R.\ Thomas, Daniel T.\ Wagner, Alastair R.\ Beresford and Andrew Rice}
 
@@
 \IEEEauthorblockN{Daniel R.\ Thomas,
-Daniel Wagner,
+Daniel T.\ Wagner,
 Alastair R.\ Beresford,

### 2014-07-28 securityupdates.tex
Remove unnecessary dangling citation for something we are reporting from our data
Add empty economics subsection
specify backend for biblatex
Use dastats data directly for number macros

@@
 \usepackage{hyperref}
-\RequirePackage[date=terse,isbn=true,doi=true,url=false,maxbibnames=9,backref=false]{biblatex}
+\RequirePackage[date=terse,isbn=true,doi=true,url=false,maxbibnames=9,backref=false,backend=bibtex]{biblatex}
 \addbibresource{securityupdates.bib}
@@
 
+\input{dastats}
 \newcommand{\percMarketShare}{XX\%~\cite{TODO}}
-\newcommand{\percInsec}{XX\%~\cite{TODO}}
-\newcommand{\daNumManufacturers}{XX~[\S\ref{sec:results}]}
 \newcommand{\daNumNetworkOperators}{XXX~[\S\ref{sec:results}]}
@@
 Our hypothesis is that these fixes do not promptly reach the users devices and so much of the time users are running versions of Android known to be vulnerable.
-We used the DeviceAnalyzer data~\cite{Wagner2013} and found that on average over \percInsec\ of devices were exposed to known vulnerabilities.
+We used the DeviceAnalyzer data~\cite{Wagner2013} and found that on average over \daMeanInsecurityPerc\ of devices were exposed to known vulnerabilities.
 There was also a period of several months when no devices ran secure versions of Android.
@@
 
+\subsection{Incentives and participants}
+\label{sec:economics}
+
 \section{Related work}

### 2014-07-28 securityupdates.tex
Remove unneeded escapes
minor text cleanup

@@
 \author{
-\IEEEauthorblockN{Daniel R.\ Thomas,
-Daniel T.\ Wagner,
-Alastair R.\ Beresford,
+\IEEEauthorblockN{Daniel R. Thomas,
+Daniel T. Wagner,
+Alastair R. Beresford,
 Andrew Rice}
@@
 % Hypothesis
-% Attempts at fine grained restrictions on running arbitrary code are hampered by updates for security vulnerabilities not reaching uers in a timely fashion.
+% Attempts at fine grained restrictions on running arbitrary code are hampered by updates for security vulnerabilities not reaching users in a timely fashion.
 
@@
 Android is the most popular smartphone platform and seeks to provide better security than popular desktop operating systems through increased compartmentalisation and a permission system for apps.
-This security relies on there not being root privilege exploits which allow apps to bypass all protections.
+This security relies on the absence of root privilege exploits which allow apps to bypass all protections.
 Many such exploits have been found and fixed.
@@
 We compare different manufacturers and operators to find which are best at promptly distributing security updates and examine how to improve the situation.
-Many phones are sold on long (18 month) contracts with monthly payments yet stop receiving updates before the end of the contract, while Windows XP could be purchased for a one off payment in October 2001 and yet will receive security updates until April 2014.
+Many phones are sold on long (18 month) contracts with monthly payments yet stop receiving updates before the end of the contract, while Windows XP could be purchased for a one off payment in October 2001 and received security updates until April 2014.
 \end{abstract}

### 2014-07-29 security_scoring.tex
Initial attempt at a security score out of 10

\subsection{Scoring for security}

Computing how good a particular manufacturer or device model is from a security standpoint is difficult as it depends on a number of factors which are hard to observe, particularly on a large scale.
Ideally we would consider both prevelance of potential problems which were not exploited and actual security failures.
However in the absence of such data we propose a scheme for assigning a device a score out of ten based on data which can be observed, and which hopefully correlates with the actual security of the devices.

There are several components to take into account:
\begin{description}
  \item[$i$] The proportion of the time were devices exposed to known vulnerabilities. This is equivalent to the BWE score \cite{TODO}
  \item[$u$] The proportion of devices run the latest version of Android shipped to any device produced by that manufacturer. This is a measure of internal updatedness, a low score would mean many devices are being left behind.
  \item[$m$] The mean number of outstanding vulnerabilities affecting devices not fixed on any device shipped by the manufacturer. This is related to the ??? measure \cite{TODO}
\end{description}
\begin{equation}
\mathrm{score} = (1 - i)\times 4 + u \times 3 + \frac{2}{1+e^m} \times 3
\end{equation}

### 2014-07-30 security_scoring.tex
Fill out security scoring section and discuss the tables

@@
 
-There are several components to take into account:
+This score is computed from several components:
 \begin{description}
-  \item[$i$] The proportion of the time were devices exposed to known vulnerabilities. This is equivalent to the BWE score \cite{TODO}
+  \item[$i$] The proportion of the time were devices exposed to known vulnerabilities.
+This is equivalent to Acer and Jackson's proposal~\cite{Acer2010} to measure the security based on the proportion of users with at least one unpatched critical vulnerability and is also $1 -$ the Vulnerability Free Days score~\cite{Wright2014}.
   \item[$u$] The proportion of devices run the latest version of Android shipped to any device produced by that manufacturer.
 This is a measure of internal updatedness, a low score would mean many devices are being left behind.
-  \item[$m$] The mean number of outstanding vulnerabilities affecting devices not fixed on any device shipped by the manufacturer. This is related to the ??? measure \cite{TODO}
+  \item[$m$] The mean number of outstanding vulnerabilities affecting devices not fixed on any device shipped by the manufacturer.
+This is related to the Median Active Vulnerabilities measure~\cite{Wright2014} but is the mean rather than the median.
+%TODO should we compute the median instead?
 \end{description}
@@
 \end{equation}
+
+The scores accross the whole of android are that \daMeanInsecurityPerc\ of devices are exposed to known root exploits.
+There are on average \daMeanOutstandingVulnerabilities\ outstanding vulnerabilities not fixed on any device.
+Only on average \daUpdatednessPerc\ of devices run the most recent version of Android.
+This gives a security score of \daSecurityScore/10.
+\daTabSecScoressummary
+However there are a wide variety of scores depending on the source of the device.
+There have been many reports that Google's Nexus devices are better at getting updates than other Android devices because Google makes the original updates and ships them to its devices.
+Table~\ref{tab:sec_summary} shows that this is the case with Nexus devices getting much better scores than non-Nexus devices.
+\daTabSecScoresmanufacturer
+Different manufacturers have very different scores, Table~\ref{tab:sec_manufacturer} shows the scores for the \daNumSigManufacturers\ manufacturers with a significant presence in our data.
+Manufacturers are considered significant if we have data from at least \daSigNumDevices\ devices and at least \daSigNumDays\ days of contributions.
+\daTabSecScoresmodel
+Even within manufacturers different models can have very different update behaviours and hence security.
+Table~\ref{tab:sec_model} shows the results for the \daNumSigModels\ device models which have a significant presence by the same metric.

### 2014-07-30 securityupdates.tex
Initial attempt at a security score out of 10
Fill out security scoring section and discuss the tables
Tweak abstract and introduction - add more results to them

@@
 Our hypothesis is that these fixes do not promptly reach the users devices and so much of the time users are running versions of Android known to be vulnerable.
-We used the DeviceAnalyzer data~\cite{Wagner2013} and found that on average over \daMeanInsecurityPerc\ of devices were exposed to known vulnerabilities.
+We found that on average over \daMeanInsecurityPerc\ of devices were exposed to known root privilege vulnerabilities and only \daUpdatednessPerc\ of devices run the most recent version of Android.
 There was also a period of several months when no devices ran secure versions of Android.
@@
 Hence when Google produces an update to Android, the update may have to pass through the manufacturer and operator before reaching the user.
-The manufacturer and operator have no financial incentive to perform this work.
-There is ongoing legal action to force some of them to do so~\cite{Soghoian2013}.
-CESG which advises the UK government on how to secure its computer system recommends picking Android devices from manufactures which are good at shipping security updates promptly~\cite{CESG2013} but it does not state which manufacturers are better.
+The manufacturer and operator have little financial incentive to perform this work as they have already sold the device.
+Operators are particularly cautious because if they ship a broken update then they will get many support requests which is expensive.
+However there is ongoing legal action to force operators to ship updates for security vulnerabilities~\cite{Soghoian2013}.
+CESG, which advises the UK government on how to secure its computer systems, recommends picking Android devices from manufactures which are good at shipping security updates promptly~\cite{CESG2013} but it does not state which manufacturers are better.
+We provide this information in~\S\ref{sec:security_scoring}.
 
@@
 
+\input{security_scoring}
+
 \subsection{Monetisation}

### 2014-07-30 securityupdates.tex
Add some related work

@@
 \label{sec:related}
-Stopping root exploits working: SEAndroid, Capsicum, iOS.
-Detecting malware: Android antivirus useless.
+We found that since 2011 \daMeanInsecurityPerc\ of devices on average were exposed to known root vulnerabilities, in 2011 Felt et al.\ studied 6 Android handsets and found that they were exposed to root vulnerabilities at least 74\% of the time.
+This might indicate that over the last 3 years the security of Android has improved, but the data are not directly comparable.
+They also found that 4 of the 46 malware samples (8\%) they analysed contained root exploits, much lower than rates found in later larger studies which found rates of 36.7\%~\cite{Zhou2012b} and 40\%~\cite{Zhou2012a}.
+
+There are continuing efforts to reduce the impact of root vulnerabilities, both in Android and more widely.
+SEAndroid~\cite{Smalley2013} which is included in Android from version ???~\cite{TODO} claimed to prevent some root vulnerabilities and to reduce the impact of others.
+Capability based enforcement systems such as Capsicum~\cite{Watson2010} substantially reduce the capabilities that an exploit has to try and gain increased privilege with.
+When Capsicum is included in Linux~\cite{TODO} and hence in Android, it could be used to place fine grained restrictions on system daemons, preventing vulnerabilities becoming root vulnerabilities.
+
+The security protections used in iOS such as only allowing signed code to run, jails and reviewing of apps has resulted in a lower level of malware affecting iOS~\cite{TODO} but can still be bypassed~\cite{Wang2013a}.
+%TODO look at iOS protections again and tighten this up
+
+Various attempts have been made to detect malware, %TODO
+However DroidChameleon~\cite{Rastogi2013} found that antivirus products could not detect malware if it was simply automatically permuted.
+
+Finding vulnerabilities is frequently a time consuming manual process but, Brumley et al.~\cite{Brumley2008} showed that it is possible to automatically generate exploits from binary patches.
+In a similar way some of the vulnerabilities which have been exploited were publicly found by examining the commit made to the Android which fixed the vulnerability.
+The update process itself can allow Android apps to gain privilege~\cite{Xing2014} by registering for new permissions before the update which creates that permission is installed.
+The Woodpecker tool~\cite{Grace2012} automatically finds permission leaks in stock Android phone images.
 
@@
 \label{sec:conclusion}
+Google Play services
 

### 2014-07-30 securityupdates.tex
Fix dates in references to use iso8601
Expand related work with stuff on malware detection

@@
 \usepackage{hyperref}
-\RequirePackage[date=terse,isbn=true,doi=true,url=false,maxbibnames=9,backref=false,backend=bibtex]{biblatex}
+\RequirePackage[date=terse,isbn=true,doi=true,url=false,urldate=iso8601,maxbibnames=9,backref=false,backend=bibtex]{biblatex}
 \addbibresource{securityupdates.bib}
@@
 
-Part of the reason why deployed devices do not receive updates in a timely manner is that there is little data on what is actually happening and so it is not possible to determine which manufacturers or operators are doing a better job.
+In the past users of Microsoft IIS were advised to switch to a different web server due to the high rate of vulnerability~\cite{Pescatore2001} this pressure helped cause Microsoft to focus on security so that it now does much better~\cite{TODO}.
+Part of the reason why deployed devices do not receive updates in a timely manner is that there is little data on what is actually happening and so it is not possible to determine which manufacturers or operators are doing a better job, this data is in~\S\ref{sec:security_scoring}.
 Hence there is little incentive for manufacturers and operators to do a better job, we will discuss this further in \S\ref{sec:economics}
@@
 \label{sec:economics}
+\cite{Felt2011}
 
@@
 They also found that 4 of the 46 malware samples (8\%) they analysed contained root exploits, much lower than rates found in later larger studies which found rates of 36.7\%~\cite{Zhou2012b} and 40\%~\cite{Zhou2012a}.
+Perhaps the malware authors heard about the results in Felt et al.'s paper?
+
+Finding vulnerabilities is frequently a time consuming manual process but, Brumley et al.~\cite{Brumley2008} showed that it is possible to automatically generate exploits from binary patches.
+In a similar way some of the vulnerabilities which have been exploited were publicly found by examining the commit made to the Android which fixed the vulnerability.
+This effect has been observed before~\cite{Barth2011} in the Firefox source code repositories, Google does not release the source code until after the release of the update reducing this effect to the delay between an update being released and it reaching the last end user device.
+The Woodpecker tool~\cite{Grace2012} automatically finds permission leaks in stock Android phone images.
+The update process itself can allow Android apps to gain privilege~\cite{Xing2014} by registering for new permissions before the update which creates that permission is installed.
 
@@
 
-Various attempts have been made to detect malware, %TODO
+Various attempts have been made to detect malware.
+RiskRanker~\cite{Grace2012a} classified 3,281/118,318 (2.8\%) apps as risky of which 718 (22\%) were malware and 322 (10\%) were previously unknown, an infection rate of 0.6\% across multiple markets.
+It uses signatures of exploits, static analysis of calls to high risk operations, detection of encrypted native code execution and Dalvik code loading.
+DroidRanger~\cite{Zhou2012a} found 148/182,823 (0.08\%) apps to be malicious across multiple markets of which 29 were previously unknown.
+It used permission-based behavioural fingerprinting -- looking at the permissions of known malware and heuristic-based filtering -- dynamic loading of both Dalvik and native code.
+AnDarwin~\cite{Crussell2013} detects similar apps, it found 169/265,359 (0.06\%) malicious apps based on the difference in permissions between the malicious app and the app which had been cloned.
+It used clustering based on semantic vectors derived from the program dependence graphs to detect similar apps.
+The Malware Genome project~\cite{Zhou2012b} collected 1,260 malware samples from 2010--2011.
+They found that the best case for anti-malware detection on this data was 79.6\%.
 However DroidChameleon~\cite{Rastogi2013} found that antivirus products could not detect malware if it was simply automatically permuted.
-
-Finding vulnerabilities is frequently a time consuming manual process but, Brumley et al.~\cite{Brumley2008} showed that it is possible to automatically generate exploits from binary patches.
-In a similar way some of the vulnerabilities which have been exploited were publicly found by examining the commit made to the Android which fixed the vulnerability.
-The update process itself can allow Android apps to gain privilege~\cite{Xing2014} by registering for new permissions before the update which creates that permission is installed.
-The Woodpecker tool~\cite{Grace2012} automatically finds permission leaks in stock Android phone images.
+All these app analysis projects have been hampered by the difficulty of obtaining full datasets of Android apps as Google does not make these available (and researchers who automatically download them, violate Google's Terms of Service).
+PlayDrone~\cite{Viennot2014} was a particularly effective project which circumvented Google's protective measures and downloaded over 1,100,000 apps from Google Play, allowing an in depth analysis.
 

### 2014-07-30 securityupdates.tex
Further related work

@@
 
+Examining security updates has been a key way of assessing security for many years.
+Many software update systems have been found not to authenticate the connection to the update server and/or do not authenticate the downloaded binaries~\cite{Bellissimo2006}.
+Even package management systems designed to provide secure updates have been found to have vulnerabilities~\cite{Cappos2008}.
+Fortunately Android does authenticate the update binaries (though the APK vulnerabilities circumvented this) and Google Play downloads them over a secure connection~\cite{Viennot2014}.
+%TODO double check that
+
+The updatedness rate for Android of \daUpdatednessPerc\ compares unfavourable with the more than 90\% rate for Windows XP SP2 computers contacting the Microsoft update servers~\cite{Gkantsidis2006} though that is within one OS version and only for those computers which did connect to the servers -- but that is the default.
+
+User-Agent strings have been used to investigate how web browser versions change with updates~\cite{Frei2008} and found that at most 80\% of Firefox users were running the most recent version.
+The same analysis was used to show that Chrome's use of silent updates seems to increase uptake of upgrades~\cite{Duebendorfer2010} with 97\% of users running the latest version within 3 weeks of release.
+Android's update process is manual, the user is notified but must action the upgrade which will then download the update and install it, the phone is rendered inoperable during the update process which is inconvenient for users and the phone must have sufficient charge so that it will not run out during the update process.
+Partly this is the result of the fact that an operating system update is being installed and so a reboot is required, but Chrome installs the new version side by side with the old one and switches the next time it is restarted.
+The same technique would be more difficult on phones with limited storage space (as for many cheap Android phones which have barely enough space to install just the update) but is a plausible improvement for more high-end devices.
+Google is deploying the same silent update technique through Google Play Services~\cite{TODO} which automatically installs updates for core Google components of Android, this also bypasses the manufacturer and operator.
+
 \section{Conclusion}
 \label{sec:conclusion}
-Google Play services
 

### 2014-07-30 securityupdates.tex
Talk about vulnerability information availability

@@
 We also have a website\footnote{\url{http://androidvulnerabilities.org/}} which we will keep up to date with future vulnerabilities and which has successfully solicited submissions from others.
+
+There was much manual effort required to track down details of these vulnerabilities as many of them are not in other databases such as the CVE database, the lack of a widely acknowledged unique identifier made identifying whether two reports referenced the same vulnerability difficult.
+Previous work has assumed ``any security issue of relevance will eventually get a CVE number assigned"~\cite{Frei2010} which is currently not the case for Android with several critical root vulnerabilities not having CVE numbers.
+
+Working out the date to use for the date on which a vulnerability was sufficiently well known to pose a threat to users is difficult.
+Frei et al.\ propose the definition of the \bf{time of disclosure} when the information about the vulnerability is freely available to the public, from a widely accepted and independent source and has been validated by security experts so that it has a risk rating.
+Unfortunately before we collated this information much of it was not published by an independent source and lacked risk rating information, even months or years after they had been actively used.
 

### 2014-08-01 securityupdates.tex
Discuss vulnerability data

@@
 There was much manual effort required to track down details of these vulnerabilities as many of them are not in other databases such as the CVE database, the lack of a widely acknowledged unique identifier made identifying whether two reports referenced the same vulnerability difficult.
-Previous work has assumed ``any security issue of relevance will eventually get a CVE number assigned"~\cite{Frei2010} which is currently not the case for Android with several critical root vulnerabilities not having CVE numbers.
+Previous work has assumed ``any security issue of relevance will eventually get a CVE number assigned"~\cite{Frei2010} which is currently not the case for Android with several root vulnerabilities not having CVE numbers.
+For some of the vulnerabilities without CVE numbers we checked with Google whether they had one assigned and they confirmed that there was no number and that they did not intend to get one.
 
 Working out the date to use for the date on which a vulnerability was sufficiently well known to pose a threat to users is difficult.
-Frei et al.\ propose the definition of the \bf{time of disclosure} when the information about the vulnerability is freely available to the public, from a widely accepted and independent source and has been validated by security experts so that it has a risk rating.
+Frei et al.\ propose the definition of the {\bf time of disclosure} when the information about the vulnerability is freely available to the public, from a widely accepted and independent source and has been validated by security experts so that it has a risk rating.
 Unfortunately before we collated this information much of it was not published by an independent source and lacked risk rating information, even months or years after they had been actively used.
+
+Prior work~\cite{Bilge2012} has shown that after public disclosure exploitation rates increase by 5 orders of magnitude and so from the point of view of widespread danger, the period between public disclosure and the update reaching the user's device is the most critical.
+However they also show that zero day exploits are typically used for 312 days before they are publicly disclosed, often to target particular organisations.
+Hence when considering vulnerability from the point of view of an organisation which cares about Advanced Persistent Threats, such as those the CESG advice~\cite{CESG2013} is aimed at, the critical period is from when people could have started using it to when the update hits the user device.
+So for our vulnerability calculations we use the earliest date we can find recorded evidence that the vulnerability was known about, even if that knowledge might have been confined to a particular manufacturer or hobbyist.
+We cannot know if someone reported the vulnerability to the manufacturer and we do know that various agencies are engaged in widespread monitoring of communications and compromise of civilian infrastructure~\cite{TODO} and so they probably know.
+Hence we use dates such as when CVEs were registered, when fixes were committed to source repositories, when exploits were committed to repositories as the date first known about.
 

### 2014-08-01 security_scoring.tex
Tweak abstract and introduction - add more results to them
Talk about the uncertainty in security scores

@@
 \subsection{Scoring for security}
+\label{sec:security_scoring}
 
@@
 Different manufacturers have very different scores, Table~\ref{tab:sec_manufacturer} shows the scores for the \daNumSigManufacturers\ manufacturers with a significant presence in our data.
-Manufacturers are considered significant if we have data from at least \daSigNumDevices\ devices and at least \daSigNumDays\ days of contributions.
+Manufacturers are considered significant if we have data from at least \daSigNumDevices\ devices and at least \daSigNumDeviceDays\ days of contributions.
 \daTabSecScoresmodel
@@
 Table~\ref{tab:sec_model} shows the results for the \daNumSigModels\ device models which have a significant presence by the same metric.
+
+These tables also show the uncertainty for $i$ and $u$.
+$i$ is calculated by taking the total insecure device days and dividing it by the total insecure and secure device days.
+The total secure device days and total insecure device days are both counting experiments and so their measurement error is their square root~\cite{Taylor1997}.
+Since the numbers involved are large, the uncertainty in $i$ is small.
+$u$ is computed by taking the sum of the proportions of devices running the most recent version each day, both the count of devices running the maximum version and total count have sqrt uncertainties.
+The Python uncertainties library was used to propagate uncertainties through calculations.
+
+%TODO there must be some way of putting an error bound on our results by considering the error on the whole DA set and the size of the subset used for each manufacturer and model

### 2014-08-01 securityupdates.tex
Start some evaluation

@@
 
+\section{Evaluation}
+Our results rely on the accuracy and representativeness of two data sets, the DeviceAnalyzer data and the AndroidVulnerabilities.org data.
+
+Comparing the API version data from DeviceAnalyzer (Figure~\ref{fig:da_api}) with the ground truth data from Google Play (Figure~\ref{fig:play_api}) using ??? we find that ???.
+Since API version is a coarse grained measure of OS version this means that the OS version distribution in DeviceAnalyzer ??? representative.
+This shows that our data is ??? useful for examining the behaviour across the whole of Android.
+However to compare different manufacturers we need to take small subsets of this data and then compare them.
+Since there is no ground truth data available we cannot check definitively
+
 \section{Related work}

### 2014-08-04 figures/update_figures.sh
Add script to update the graphs and update the graphs

@@ -1,0 +1,22 @@
+#!/bin/sh
+
+set -e
+
+fromdir=~drt24/git/da/graphing/pdf/
+cp "$fromdir"norm_api.pdf da_norm_api.pdf
+cp "$fromdir"norm_api-notitle.pdf da_norm_api-notitle.pdf
+cp "$fromdir"norm_os.pdf da_norm_os.pdf
+cp "$fromdir"os.pdf da_os.pdf
+cp "$fromdir"device_data/device_data-lg-security.pdf device_data-LG.pdf
+cp "$fromdir"from_to_updates.pdf from_to_updates.pdf
+cp "$fromdir"googleplayapi.pdf googleplayapi.pdf
+cp "$fromdir"googleplayapi-notitle.pdf googleplayapi-notitle.pdf
+cp "$fromdir"norm_api_heat.pdf norm_api_heat.pdf
+cp "$fromdir"norm_os_heat.pdf norm_os_heat.pdf
+cp "$fromdir"norm_versionsecurity.pdf proportioninsecure.pdf
+cp "$fromdir"norm_versionsecurity-notitle.pdf proportioninsecure-notitle.pdf
+cp "$fromdir"updates_between.pdf updates_between_versions.pdf
+cp "$fromdir"vulnerabilities.pdf vulnerabilities.pdf
+cp "$fromdir"updates_weekly.pdf w_security_updates.pdf
+cp "$fromdir"wifi-accounts-ototals-hist.pdf wifi-accounts-ototals-hist.pdf
+cp "$fromdir"wifi-accounts-stotals-hist.pdf wifi-accounts-stotals-hist.pdf

### 2014-08-04 securityupdates.tex
Add script to update the graphs and update the graphs
Exploit rates are from 2012 and so not current
Use generated vulnerability table from os-version.py

@@
 Manufacturers of Android devices have also added further root exploits when customising it~\cite{Grace2012}.
-Between 36.7\%~\cite{Zhou2012b} and 40\%~\cite{Zhou2012a} of malware for Android contains root exploits.
+Between 36.7\%~\cite{Zhou2012b} and 40\%~\cite{Zhou2012a} of malware for Android contained root exploits in 2012.
 While this shows that much Android malware does not need a root exploit to work, it can just request the relevant permissions and the user will grant them, a significant proportion does.
@@
 
-\begin{table}
-\centering
-\begin{tabular}{c|c|c}
-Vulnerability & Date known & How known \\
-\hline
-Exploid & 2010-07-15 & Public disclosure \\
-Psneuter & 2011-01-06 & Public disclosure \\
-Levitator & 2011-03-10 & CVE registered \\
-ZergRush & 2011-10-06 & Exploit developed \\
-APK duplicate file & 2013-02-18 & Fix committed \\
-APK unsigned shorts & 2013-07-03 & Fix committed \\
-% TODO this is an incomplete list, we need to add the rest here and in os-version.py
-\end{tabular}
-\caption{Root equivalent vulnerabilities in Android}
-\label{tab:andvulns}
-\end{table}
+\daTabAndVulns
 
@@
  \centering
- \includegraphics[width=\columnwidth]{figures/play_norm_api_trunc}
+ \includegraphics[width=\columnwidth]{figures/googleplayapi}
  \caption{Google Play data on proportion of devices running different Android API versions}
@@
  \centering
- \includegraphics[width=\columnwidth]{figures/os}
+ \includegraphics[width=\columnwidth]{figures/da_os}
  \caption{Android versions in Device Analyzer data over time}

### 2014-08-05 figures/update_figures.sh
Update figures and figures updater script

@@ -11,12 +11,15 @@ cp "$fromdir"device_data/device_data-lg-security.pdf device_data-LG.pdf
 cp "$fromdir"from_to_updates.pdf from_to_updates.pdf
 cp "$fromdir"googleplayapi.pdf googleplayapi.pdf
 cp "$fromdir"googleplayapi-notitle.pdf googleplayapi-notitle.pdf
+cp "$fromdir"norm_api_gpcomp.pdf norm_api_gpcomp.pdf
 cp "$fromdir"norm_api_heat.pdf norm_api_heat.pdf
 cp "$fromdir"norm_os_heat.pdf norm_os_heat.pdf
 cp "$fromdir"norm_versionsecurity.pdf proportioninsecure.pdf
 cp "$fromdir"norm_versionsecurity-notitle.pdf proportioninsecure-notitle.pdf
 cp "$fromdir"updates_between.pdf updates_between_versions.pdf
 cp "$fromdir"vulnerabilities.pdf vulnerabilities.pdf
+cp "$fromdir"nvulnerabilities_heat.pdf nvulnerabilities_heat.pdf
 cp "$fromdir"updates_weekly.pdf w_security_updates.pdf
 cp "$fromdir"wifi-accounts-ototals-hist.pdf wifi-accounts-ototals-hist.pdf
 cp "$fromdir"wifi-accounts-stotals-hist.pdf wifi-accounts-stotals-hist.pdf
+cp "$fromdir"api_gpcomp_diff.pdf api_gpcomp_diff.pdf

### 2014-08-05 securityupdates.tex
Update figures and figures updater script
Include figure showing difference between DA and GP data

@@
 While API versions are too coarse grained to use for security update detection they are closely related to OS versions and so if the DA data on API versions is similar to the Google Data on API versions then the DA data on OS versions should be representative.
-Figure \ref{fig:play_api} shows the data Google and Figure \ref{fig:da_api} and they appear similar. %TODO we want a statistical metric to claim this strongly with.
+Figure \ref{fig:play_api} shows the data from Google and Figure \ref{fig:da_api} shows the data from Device Analyzer and they appear similar.
+Figure \ref{fig:da_gp_comp_diff} shows the difference between Figures \ref{fig:play_api} and \ref{fig:da_api}.
+It shows that the Device Analyzer data systematically overestimates the prevalence of new API versions and underestimates the prevalence of old API versions.
+This means that the OS version information from Device Analyzer is likely to be overestimating the prevalence of new OS versions and hence our results are a conservative estimate of the security of Android.
+%TODO we want a statistical metric to claim this strongly with.
 \begin{figure}
@@
  \centering
- \includegraphics[width=\columnwidth]{figures/da_norm_api}
- \caption{Google Play data on proportion of devices running different Android API versions}
+ \includegraphics[width=\columnwidth]{figures/norm_api_gpcomp}
+ \caption{Device Analyzer data on proportion of devices running different Android API versions}
  \label{fig:da_api}
+\end{figure}
+\begin{figure}
+ \centering
+ \includegraphics[width=\columnwidth]{figures/api_gpcomp_diff}
+ \caption{Difference between Device Analyzer and Google Play data on the proportion of devices running different Android API versions}
+ \label{fig:da_gp_comp_diff}
 \end{figure}

### 2014-08-05 figures/update_figures.sh
Correct updated figures script

@@ -17,7 +17,7 @@ cp "$fromdir"norm_os_heat.pdf norm_os_heat.pdf
 cp "$fromdir"norm_versionsecurity.pdf proportioninsecure.pdf
 cp "$fromdir"norm_versionsecurity-notitle.pdf proportioninsecure-notitle.pdf
 cp "$fromdir"updates_between.pdf updates_between_versions.pdf
-cp "$fromdir"vulnerabilities.pdf vulnerabilities.pdf
+cp "$fromdir"nvulnerabilities.pdf vulnerabilities.pdf
 cp "$fromdir"nvulnerabilities_heat.pdf nvulnerabilities_heat.pdf
 cp "$fromdir"updates_weekly.pdf w_security_updates.pdf
 cp "$fromdir"wifi-accounts-ototals-hist.pdf wifi-accounts-ototals-hist.pdf

### 2014-08-05 security_scoring.tex
Fix spelling mistake
Updates on reading through

@@
 \begin{description}
-  \item[$i$] The proportion of the time were devices exposed to known vulnerabilities.
-This is equivalent to Acer and Jackson's proposal~\cite{Acer2010} to measure the security based on the proportion of users with at least one unpatched critical vulnerability and is also $1 -$ the Vulnerability Free Days score~\cite{Wright2014}.
+  \item[$f$] The proportion of the time where devices were free from known critical vulnerabilities.
+This is equivalent to Acer and Jackson's proposal~\cite{Acer2010} to measure the security based on the proportion of users with at least one unpatched critical vulnerability and is the Vulnerability Free Days score~\cite{Wright2014}.
   \item[$u$] The proportion of devices run the latest version of Android shipped to any device produced by that manufacturer.
@@
   \item[$m$] The mean number of outstanding vulnerabilities affecting devices not fixed on any device shipped by the manufacturer.
-This is related to the Median Active Vulnerabilities measure~\cite{Wright2014} but is the mean rather than the median.
+This is related to the Median Active Vulnerabilities measure~\cite{Wright2014} but is the mean rather than the median, as this gives a continuous value.
 %TODO should we compute the median instead?
+
+We want to provide a score out of 10, since $f$ is the most important metric we weight it more highly.
+Since $m$ is an unbounded positive real number, we map it into the range [0--1).
 \end{description}
 \begin{equation}
-\mathrm{score} = (1 - i)\times 4 + u \times 3 + \frac{2}{1+e^m} \times 3
+\mathrm{score} = f\times 4 + u \times 3 + \frac{2}{1+e^m} \times 3
 \end{equation}
 
-The scores accross the whole of android are that \daMeanInsecurityPerc\ of devices are exposed to known root exploits.
+The scores across the whole of android are that \daMeanInsecurityPerc\ of devices are exposed to known root exploits.
 There are on average \daMeanOutstandingVulnerabilities\ outstanding vulnerabilities not fixed on any device.
 Only on average \daUpdatednessPerc\ of devices run the most recent version of Android.
-This gives a security score of \daSecurityScore/10.
+This gives a security score of \daSecurityScore / 10.
 \daTabSecScoressummary
@@
 \daTabSecScoresmanufacturer
-Different manufacturers have very different scores, Table~\ref{tab:sec_manufacturer} shows the scores for the \daNumSigManufacturers\ manufacturers with a significant presence in our data.
+Different manufacturers have very different scores, Table~\ref{tab:sec_manufacturer} shows the scores for the \daNumSigManufacturers\ manufacturers with a significant presence in our data with \daSecScoreBestmanufacturer\ (\daSecScoreBestmanufacturerScore\ / 10) doing best and \daSecScoreWorstmanufacturer\ (\daSecScoreWorstmanufacturerScore\ / 10) doing worst.
 Manufacturers are considered significant if we have data from at least \daSigNumDevices\ devices and at least \daSigNumDeviceDays\ days of contributions.
@@
 Even within manufacturers different models can have very different update behaviours and hence security.
-Table~\ref{tab:sec_model} shows the results for the \daNumSigModels\ device models which have a significant presence by the same metric.
+Table~\ref{tab:sec_model} shows the results for the \daNumSigModels\ device models which have a significant presence by the same metric with \daSecScoreBestmodel\ (\daSecScoreBestmodelScore\ / 10) doing best and \daSecScoreWorstmodel\ (\daSecScoreWorstmodelScore\ / 10) doing worst.
 
-These tables also show the uncertainty for $i$ and $u$.
-$i$ is calculated by taking the total insecure device days and dividing it by the total insecure and secure device days.
+These tables also show the uncertainty for $f$ and $u$.
+$f$ is calculated by taking the total secure device days and dividing it by the total insecure and secure device days.
 The total secure device days and total insecure device days are both counting experiments and so their measurement error is their square root~\cite{Taylor1997}.
-Since the numbers involved are large, the uncertainty in $i$ is small.
+Since the numbers involved are large, the uncertainty in $f$ is small.
 $u$ is computed by taking the sum of the proportions of devices running the most recent version each day, both the count of devices running the maximum version and total count have sqrt uncertainties.
-The Python uncertainties library was used to propagate uncertainties through calculations.
+The Python \texttt{uncertainties} library was used to propagate uncertainties through calculations.
 

### 2014-08-05 securityupdates.tex
Initial Acknowledgements
Updates on reading through

@@
 Android is the most popular smartphone platform and seeks to provide better security than popular desktop operating systems through increased compartmentalisation and a permission system for apps.
-This security relies on the absence of root privilege exploits which allow apps to bypass all protections.
+This security relies on the absence of root privilege exploits which allow apps to bypass protections.
 Many such exploits have been found and fixed.
 Our hypothesis is that these fixes do not promptly reach the users devices and so much of the time users are running versions of Android known to be vulnerable.
-We found that on average over \daMeanInsecurityPerc\ of devices were exposed to known root privilege vulnerabilities and only \daUpdatednessPerc\ of devices run the most recent version of Android.
+We used Device Analyzer data from over \daOSYearsOfData\ years and \daNumOSDevices\ devices and found that on average over \daMeanInsecurityPerc\ of devices were exposed to known root privilege vulnerabilities and only \daUpdatednessPerc\ of devices run the most recent version of Android.
 There was also a period of several months when no devices ran secure versions of Android.
-We compare different manufacturers and operators to find which are best at promptly distributing security updates and examine how to improve the situation.
+We compare different manufacturers and operators to find which are best at promptly distributing security updates and examine how to improve the situation. %TODO operators analysis
 Many phones are sold on long (18 month) contracts with monthly payments yet stop receiving updates before the end of the contract, while Windows XP could be purchased for a one off payment in October 2001 and received security updates until April 2014.
@@
 Between 36.7\%~\cite{Zhou2012b} and 40\%~\cite{Zhou2012a} of malware for Android contained root exploits in 2012.
-While this shows that much Android malware does not need a root exploit to work, it can just request the relevant permissions and the user will grant them, a significant proportion does.
+While this shows that much Android malware does not need a root exploit to work, as it can just request the relevant permissions and the user will grant them, a significant proportion does.
 This malware is also more dangerous because while ordinary malware can be remotely removed by Google through the Play store, once a root exploit has been used there are no guarantees.
 
-Figure \ref{fig:proportioninsecure} shows the proportion of Android devices in the DeviceAnalyzer data which were running versions of Android known to root exploits at the time.
-In \S\ref{sec:background} we will explain where these data comes from and then in \S\ref{sec:results} we will show these data in more detail.
+The proportion of Android devices in the DeviceAnalyzer data which were running versions of Android known to have root vulnerabilities at the time is shown in Figure \ref{fig:proportioninsecure}.
+In \S\ref{sec:background} we will explain where these data come from and then in \S\ref{sec:results} we will show these data in more detail.
 
@@
 In the past users of Microsoft IIS were advised to switch to a different web server due to the high rate of vulnerability~\cite{Pescatore2001} this pressure helped cause Microsoft to focus on security so that it now does much better~\cite{TODO}.
-Part of the reason why deployed devices do not receive updates in a timely manner is that there is little data on what is actually happening and so it is not possible to determine which manufacturers or operators are doing a better job, this data is in~\S\ref{sec:security_scoring}.
-Hence there is little incentive for manufacturers and operators to do a better job, we will discuss this further in \S\ref{sec:economics}
+Part of the reason why deployed devices do not receive updates in a timely manner is that there is little data on what is actually happening and so it is not possible to determine which manufacturers or operators are doing a better job.
+Hence there is little incentive for manufacturers and operators to do a better job, we provide this data  in~\S\ref{sec:security_scoring} and discuss incentives further in \S\ref{sec:economics}
 
@@
 This is not something that applications running on the phone can exploit to break compartmentalisation and so we do not include those exploits.
+Unfortunately many published exploits use adb for convenience and so determining whether that is necessary to exploit the vulnerability can be difficult.
 Some exploits which grant root are not traditional kernel exploits, for example the discovery of flaws in the verification of signatures on Android applications in February 2013~\cite{Forristal2013} meant that applications could pretend to be signed with system keys and hence gain root privileges.
 
-We have published full details of the vulnerabilities listed below in an accompanying technical report~\cite{TODO} and summarise them in Table \ref{tab:andvulns}.
+We have published full details of the \daNumVulnsUsed\ vulnerabilities used for this analysis in an accompanying technical report~\cite{TODO} and summarise them in Table \ref{tab:andvulns}.
 We also have a website\footnote{\url{http://androidvulnerabilities.org/}} which we will keep up to date with future vulnerabilities and which has successfully solicited submissions from others.
+In this analysis we only use vulnerabilities which affected all Android devices, not any of the manufacturer or hardware specific vulnerabilities
 
@@
 Previous work has assumed ``any security issue of relevance will eventually get a CVE number assigned"~\cite{Frei2010} which is currently not the case for Android with several root vulnerabilities not having CVE numbers.
-For some of the vulnerabilities without CVE numbers we checked with Google whether they had one assigned and they confirmed that there was no number and that they did not intend to get one.
+For some of the vulnerabilities without CVE numbers we checked with Google whether they had one assigned and they confirmed that there was no number and that they did not intend to get one (though they did give us the Android bug number and tell us to use that).
 
@@
 However they also show that zero day exploits are typically used for 312 days before they are publicly disclosed, often to target particular organisations.
-Hence when considering vulnerability from the point of view of an organisation which cares about Advanced Persistent Threats, such as those the CESG advice~\cite{CESG2013} is aimed at, the critical period is from when people could have started using it to when the update hits the user device.
+Hence when considering vulnerability from the point of view of an organisation which cares about Advanced Persistent Threats, such as those the CESG advice~\cite{CESG2013} is aimed at, the critical period is from when people could have started using the vulnerability to when the update hits the user device.
 So for our vulnerability calculations we use the earliest date we can find recorded evidence that the vulnerability was known about, even if that knowledge might have been confined to a particular manufacturer or hobbyist.
 We cannot know if someone reported the vulnerability to the manufacturer and we do know that various agencies are engaged in widespread monitoring of communications and compromise of civilian infrastructure~\cite{TODO} and so they probably know.
-Hence we use dates such as when CVEs were registered, when fixes were committed to source repositories, when exploits were committed to repositories as the date first known about.
+Hence we use dates such as: when CVEs were registered, when fixes were committed to source repositories, or when exploits were committed to repositories; as the date first known about.
 
@@
 The build string is `The user-visible version string', fortunately most (99.9\%) devices in the data have a build string of the form `x.y.z random string' and so it is possible to extract the Android version number.
+On a large proportion of devices `random string' is a well defined build number but different manufacturers use different schema.
 
 \subsection{Running a version of Android known to be vulnerable makes a device vulnerable}
-While we have not tried to exploit these vulnerabilities on the devices in order to test whether the devices are actually vulnerable, we are confident that most devices.
+While we have not tried to exploit these vulnerabilities on the devices in order to test whether the devices are actually vulnerable, we are confident that most devices will be if they are running a vesion known to be vulnerable.
 
@@
 \label{sec:results}
-In this section we present the results of our analysis showing ????. %TODO summarise our results
+In this section we present the results of our analysis showing \daMeanInsecurityPerc\ of devices to have root vulnerabilities on average, nexus devices to be better than others, \daSecScoreBestmanufacturer\ to be the best manufacturer and ????.
+%TODO summarise our results
+%TODO comparison of nexus update rates with iOS
 
@@
 This allows us to have confidence in the OS version information.
-Figure \ref{fig:os} shows the raw data on the number of devices in the Device Analyzer data running different versions of Android each day.
+Figure \ref{fig:norm_os} shows the proportion of devices in the Device Analyzer data running different versions of Android each day.
 It shows how old versions are gradually replaced by new ones, the long tail of devices which do not see updates to more recent versions and the fluctuation in the number of devices in the data.
-Figure \ref{fig:norm_os} shows the normalised OS version distribution.
-\begin{figure}
- \centering
- \includegraphics[width=\columnwidth]{figures/da_os}
- \caption{Android versions in Device Analyzer data over time}
- \label{fig:os}
-\end{figure}
 \begin{figure}
@@
 \end{figure}
+
+\subsection{Update frequency}
+%TODO find out what the frequency of updates is
 \begin{figure}
@@
 Surprisingly there are also a small number of downgrade events when older versions of Android are installed on to devices.
-Figure \ref{fig:weekly_security_updates} shows the number of devices getting security updates each week, blue shows updates which changed the Android version number from a version with known vulnerabilities to one which had fewer known vulnerabilities.
+The number of devices getting security updates each week, is shown in Figure \ref{fig:weekly_security_updates}, blue shows updates which changed the Android version number from a version with known vulnerabilities to one which had fewer known vulnerabilities.
 Red indicates updates which changed the build number but not the version number and so might contain a backported fix for a vulnerability.
-
-\subsection{Update frequency}
-%TODO find out what the frequency of updates is
+%TODO summary stats, what does this mean
 
@@
 
+\subsection{User-Agent data}
+TODO
+
+\subsection{Comparison with iOS}
+TODO
+
 \subsection{Monetisation}
+%TODO redo this section in light of related work
 One of the key differences between smartphones and desktops is that it is easier to directly cash out on a smartphone as malware can send premium rate texts or make premium rate phone calls.
@@
 
-\section{Evaluation}
-Our results rely on the accuracy and representativeness of two data sets, the DeviceAnalyzer data and the AndroidVulnerabilities.org data.
-
-Comparing the API version data from DeviceAnalyzer (Figure~\ref{fig:da_api}) with the ground truth data from Google Play (Figure~\ref{fig:play_api}) using ??? we find that ???.
-Since API version is a coarse grained measure of OS version this means that the OS version distribution in DeviceAnalyzer ??? representative.
-This shows that our data is ??? useful for examining the behaviour across the whole of Android.
-However to compare different manufacturers we need to take small subsets of this data and then compare them.
-Since there is no ground truth data available we cannot check definitively
-
 \section{Related work}
@@
 We found that since 2011 \daMeanInsecurityPerc\ of devices on average were exposed to known root vulnerabilities, in 2011 Felt et al.\ studied 6 Android handsets and found that they were exposed to root vulnerabilities at least 74\% of the time.
-This might indicate that over the last 3 years the security of Android has improved, but the data are not directly comparable.
+This might indicate that over the last 3 years the security of Android has not improved, but these data are not directly comparable.
 They also found that 4 of the 46 malware samples (8\%) they analysed contained root exploits, much lower than rates found in later larger studies which found rates of 36.7\%~\cite{Zhou2012b} and 40\%~\cite{Zhou2012a}.
 Perhaps the malware authors heard about the results in Felt et al.'s paper?
+Our results show root exploits are still a severe threat.
 
@@
 This effect has been observed before~\cite{Barth2011} in the Firefox source code repositories, Google does not release the source code until after the release of the update reducing this effect to the delay between an update being released and it reaching the last end user device.
+There have been other attempts to automatically find vulnerabilities.
 The Woodpecker tool~\cite{Grace2012} automatically finds permission leaks in stock Android phone images.
-The update process itself can allow Android apps to gain privilege~\cite{Xing2014} by registering for new permissions before the update which creates that permission is installed.
+The update process itself can allow Android apps to gain privilege through `Pileup' vulnerabilities~\cite{Xing2014} by registering for new permissions before the update which creates that permission is installed.
 
@@
 RiskRanker~\cite{Grace2012a} classified 3,281/118,318 (2.8\%) apps as risky of which 718 (22\%) were malware and 322 (10\%) were previously unknown, an infection rate of 0.6\% across multiple markets.
-It uses signatures of exploits, static analysis of calls to high risk operations, detection of encrypted native code execution and Dalvik code loading.
+It uses signatures of exploits, static analysis of calls to high risk operations, detection of encrypted native code execution and dynamic Dalvik code loading to detect risky apps.
 DroidRanger~\cite{Zhou2012a} found 148/182,823 (0.08\%) apps to be malicious across multiple markets of which 29 were previously unknown.
@@
 \label{sec:conclusion}
+TODO
+
+\section*{Acknowledgements}
+Thanks to David Robertson for helpful advice on statistical analysis.
+Thanks to Laurent Simon, Thomas Coudray, Adrian Taylor, Justin Case and Khilan Gudka for reporting vulnerabilities in Android.
 

### 2014-08-05 securityupdates.tex
Add some user agent data graphs

@@
 \subsection{User-Agent data}
+Device Analyzer provides one view on the state of Android phones, however some of the data available in Device Analyzer can also be obtained from alternative sources.
+OS version information for both Android and iOS (for which it is not possible to write a Device Analyzer app) is published in the User-Agent header sent by mobile devices making http requests.
+We obtained data from the Apache web server access logs of the Student-Run Computing Facility\footnote{\url{https://www.srcf.net}} which hosts thousands of websites.
+These data are shown in Figures \ref{fig:ua_android_month} and \ref{fig:ua_ios_month}.
+\begin{figure}
+ \includegraphics[width=0.5\textwidth]{figures/ua-versions-android-month-norm}
+ \caption{Versions of Android over time from SRCF User-Agent data}
+ \label{fig:ua_android_month}
+\end{figure}
+\begin{figure}
+ \includegraphics[width=0.5\textwidth]{figures/ua-versions-ios-month-norm}
+ \caption{Versions of iOS over time from SRCF User-Agent data}
+ \label{fig:ua_ios_month}
+\end{figure}
+Clearly this data is nosier than the Device Analyzer data...
 TODO
@@
 \subsection{Comparison with iOS}
+TODO
+
+\subsection{Predicting future version distributions}
+TODO
+
+\subsection{Lifetime of a vulnerability}
 TODO

### 2014-08-05 figures/update_figures.sh

@@ -23,3 +23,5 @@ cp "$fromdir"updates_weekly.pdf w_security_updates.pdf
 cp "$fromdir"wifi-accounts-ototals-hist.pdf wifi-accounts-ototals-hist.pdf
 cp "$fromdir"wifi-accounts-stotals-hist.pdf wifi-accounts-stotals-hist.pdf
 cp "$fromdir"api_gpcomp_diff.pdf api_gpcomp_diff.pdf
+cp "$fromdir"ua-versions-android-month-norm.pdf ua-versions-android-month-norm.pdf
+cp "$fromdir"ua-versions-ios-month-norm.pdf ua-versions-ios-month-norm.pdf

### 2014-08-06 figures/update_figures.sh
Add api_gpcomp_rdiff plot which clearly shows how we have early adopters and so are conservative

@@ -23,5 +23,6 @@ cp "$fromdir"updates_weekly.pdf w_security_updates.pdf
 cp "$fromdir"wifi-accounts-ototals-hist.pdf wifi-accounts-ototals-hist.pdf
 cp "$fromdir"wifi-accounts-stotals-hist.pdf wifi-accounts-stotals-hist.pdf
 cp "$fromdir"api_gpcomp_diff.pdf api_gpcomp_diff.pdf
+cp "$fromdir"api_gpcomp_rdiff.pdf api_gpcomp_rdiff.pdf
 cp "$fromdir"ua-versions-android-month-norm.pdf ua-versions-android-month-norm.pdf
 cp "$fromdir"ua-versions-ios-month-norm.pdf ua-versions-ios-month-norm.pdf

### 2014-08-06 securityupdates.tex
Add api_gpcomp_rdiff plot which clearly shows how we have early adopters and so are conservative
Remove user agent data section as we don't need it
Prod abstract and intro

--- 2014-08-05 securityupdates.tex
@@
 We compare different manufacturers and operators to find which are best at promptly distributing security updates and examine how to improve the situation. %TODO operators analysis
-Many phones are sold on long (18 month) contracts with monthly payments yet stop receiving updates before the end of the contract, while Windows XP could be purchased for a one off payment in October 2001 and received security updates until April 2014.
 \end{abstract}
@@
 \section{Introduction}
-Android has \percMarketShare\ of the smartphone market and while the core development is controlled by Google there are at least \daNumManufacturers\ manufacturers which make devices which run Android.
+Android has \percMarketShare\ of the smartphone market and while the core development is controlled by Google there are at least \daNumManufacturers\footnote{From the Device Analyzer data} manufacturers which make devices which run Android.
+Android receives regular updates  which add new features and fix bugs, including security vulnerabilities.
 Many of these manufacturers customise the version of Android they ship and sometimes network operators (of which there are at least \daNumNetworkOperators) make further modifications.
@@
 The manufacturer and operator have little financial incentive to perform this work as they have already sold the device.
-Operators are particularly cautious because if they ship a broken update then they will get many support requests which is expensive.
+Operators are cautious because if they ship a broken update then they will get many support requests which is expensive.
 However there is ongoing legal action to force operators to ship updates for security vulnerabilities~\cite{Soghoian2013}.
+Many phones are sold on long (18 month) contracts with monthly payments yet stop receiving updates before the end of the contract, while Windows XP could be purchased for a one off payment in October 2001 and received security updates until April 2014.
 CESG, which advises the UK government on how to secure its computer systems, recommends picking Android devices from manufactures which are good at shipping security updates promptly~\cite{CESG2013} but it does not state which manufacturers are better.
@@
 Figure \ref{fig:play_api} shows the data from Google and Figure \ref{fig:da_api} shows the data from Device Analyzer and they appear similar.
-Figure \ref{fig:da_gp_comp_diff} shows the difference between Figures \ref{fig:play_api} and \ref{fig:da_api}.
+Figure \ref{fig:da_gp_comp_diff} shows the difference between Figures \ref{fig:play_api} and \ref{fig:da_api}, normalising for days since the API version was released.
 It shows that the Device Analyzer data systematically overestimates the prevalence of new API versions and underestimates the prevalence of old API versions.
@@
  \centering
- \includegraphics[width=\columnwidth]{figures/api_gpcomp_diff}
+ \includegraphics[width=\columnwidth]{figures/api_gpcomp_rdiff}
  \caption{Difference between Device Analyzer and Google Play data on the proportion of devices running different Android API versions}
@@
 
-\subsection{User-Agent data}
+\subsection{Comparison with iOS}
 TODO
 
-\subsection{Comparison with iOS}
+\subsection{Predicting future version distributions}
+TODO
+
+\subsection{Lifetime of a vulnerability}
 TODO

### 2014-08-06 securityupdates.tex
Move subsection on DA data validity after the exciting results from it

@@
 \usepackage{hyperref}
+\usepackage{subcaption}
 \RequirePackage[date=terse,isbn=true,doi=true,url=false,urldate=iso8601,maxbibnames=9,backref=false,backend=bibtex]{biblatex}
@@
 
-The data from Device Analyzer (DA) we want to use to investigate the proportion of devices exposed to different vulnerabilities is the OS version.
-Unfortunately there is no authoritative source of OS version information and so we cannot check directly whether our data is representative.
-However Google has published API version information every month since December 2009 and we have collated this information.\footnote{\url{https://www.cl.cam.ac.uk/~drt24/android/versiondistribution/}}
-While API versions are too coarse grained to use for security update detection they are closely related to OS versions and so if the DA data on API versions is similar to the Google Data on API versions then the DA data on OS versions should be representative.
-Figure \ref{fig:play_api} shows the data from Google and Figure \ref{fig:da_api} shows the data from Device Analyzer and they appear similar.
-Figure \ref{fig:da_gp_comp_diff} shows the difference between Figures \ref{fig:play_api} and \ref{fig:da_api}, normalising for days since the API version was released.
-It shows that the Device Analyzer data systematically overestimates the prevalence of new API versions and underestimates the prevalence of old API versions.
-This means that the OS version information from Device Analyzer is likely to be overestimating the prevalence of new OS versions and hence our results are a conservative estimate of the security of Android.
-%TODO we want a statistical metric to claim this strongly with.
-\begin{figure}
- \centering
- \includegraphics[width=\columnwidth]{figures/googleplayapi}
- \caption{Google Play data on proportion of devices running different Android API versions}
- \label{fig:play_api}
-\end{figure}
-\begin{figure}
- \centering
- \includegraphics[width=\columnwidth]{figures/norm_api_gpcomp}
- \caption{Device Analyzer data on proportion of devices running different Android API versions}
- \label{fig:da_api}
-\end{figure}
-\begin{figure}
- \centering
- \includegraphics[width=\columnwidth]{figures/api_gpcomp_rdiff}
- \caption{Difference between Device Analyzer and Google Play data on the proportion of devices running different Android API versions}
- \label{fig:da_gp_comp_diff}
-\end{figure}
-
-This allows us to have confidence in the OS version information.
-Figure \ref{fig:norm_os} shows the proportion of devices in the Device Analyzer data running different versions of Android each day.
+The proportion of devices in the Device Analyzer data running different versions of Android each day is shown in Figure~\ref{fig:norm_os}.
 It shows how old versions are gradually replaced by new ones, the long tail of devices which do not see updates to more recent versions and the fluctuation in the number of devices in the data.
+We verify the validity of this data in \S\ref{sec:representative}.
 \begin{figure}
@@
 
+\subsection{The Device Analyzer data gives a conservative estimate of the Android version distribution}
+\label{sec:representative}
+\begin{figure}
+ \centering
+ \begin{subfigure}[b]{\columnwidth}
+ \includegraphics[width=\columnwidth]{figures/googleplayapi}
+ \caption{Google Play data on proportion of devices running different Android API versions}
+ \label{fig:play_api}
+\end{subfigure}
+\begin{subfigure}[b]{\columnwidth}
+ \includegraphics[width=\columnwidth]{figures/norm_api_gpcomp}
+ \caption{Device Analyzer data on proportion of devices running different Android API versions}
+ \label{fig:da_api}
+\end{subfigure}
+\caption{Monthly Android API version data}
+\end{figure}
+\begin{figure}
+ \centering
+ \includegraphics[width=\columnwidth]{figures/api_gpcomp_rdiff}
+ \caption{Difference between Device Analyzer and Google Play data on the proportion of devices running different Android API versions}
+ \label{fig:da_gp_comp_diff}
+\end{figure}
+The data from Device Analyzer (DA) we used to investigate the proportion of devices exposed to different vulnerabilities is the OS version.
+Unfortunately there is no authoritative source of OS version information and so we cannot check directly whether our data is representative.
+However Google has published API version information every month since December 2009 and we have collated this information.\footnote{\url{https://www.cl.cam.ac.uk/~drt24/android/versiondistribution/}}
+While API versions are too coarse grained to use for security update detection they are closely related to OS versions and so if the DA data on API versions is similar to the Google Data on API versions then the DA data on OS versions should be representative.
+Figure \ref{fig:play_api} shows the data from Google and Figure \ref{fig:da_api} shows the data from Device Analyzer and they appear similar.
+Figure \ref{fig:da_gp_comp_diff} shows the difference between Figures \ref{fig:play_api} and \ref{fig:da_api}, normalising for days since the API version was released.
+It shows that the Device Analyzer data systematically overestimates the prevalence of new API versions and underestimates the prevalence of old API versions.
+This means that the OS version information from Device Analyzer is likely to be overestimating the prevalence of new OS versions and hence our results are a conservative estimate of the security of Android.
+%TODO we want a statistical metric to claim this strongly with.
+This allows us to have confidence in the OS version information.
+
 \section{Related work}

### 2014-08-06 securityupdates.tex
'at least' to 'about'

@@
 \section{Introduction}
-Android has \percMarketShare\ of the smartphone market and while the core development is controlled by Google there are at least \daNumManufacturers\footnote{From the Device Analyzer data} manufacturers which make devices which run Android.
+Android has \percMarketShare\ of the smartphone market and while the core development is controlled by Google there are about \daNumManufacturers\footnote{From the Device Analyzer data} manufacturers which make devices which run Android.
 Android receives regular updates  which add new features and fix bugs, including security vulnerabilities.

### 2014-08-06 securityupdates.tex
Expand intro to Background

@@
 \label{sec:background}
-There are two sources of data we needed for this study, information on root exploits and information on installed versions of Android.
+There are two sources of data we needed for this study, information on root vulnerabilities and information on installed versions of Android.
+We need data on root vulnerabilities so that we know which versions of Android were vulnerable to which known vulnerabilities at what times.
+We need information on the distribution of installed versions of Android so that we know what proportion of Android devices were affected by these vulnerabilities.
 

### 2014-08-06 securityupdates.tex
Add draft of contributions of paper

@@
 Hence there is little incentive for manufacturers and operators to do a better job, we provide this data  in~\S\ref{sec:security_scoring} and discuss incentives further in \S\ref{sec:economics}
+
+The contributions of this paper are:
+\begin{itemize}
+ \item We quantify the vulnerability of Android to root equivalent vulnerabilities.
+ \item We characterise the Android update ecosystem.
+ \item We measure the security of Android according to three metrics and compare different manufacturers. and device models to allow consumers to differentiate between them based on security.
+ \item We investigate the time-lines of vulnerabilities in Android.
+ \item We provide models of how the distribution of Android versions changes allowing future version distributions to be predicted.
+ \item We propose methods which could be used to improve the situation in future.
+\end{itemize}
 

### 2014-08-06 securityupdates.tex
Include adb percent enabled

@@
 \newcommand{\daMonthsDevices}{XXX}
-\newcommand{\daAdbPerc}{X\%}
 
@@
 \emph{Root equivalent} means that it grants privileges equivalent in scope to root which can then be used to gain root.
-Some phones can be `rooted' by enabling USB debugging and using the special privileges of the adb shell to root the device but only \daAdbPerc\ of devices have USB debugging enabled.
+Some phones can be `rooted' by enabling USB debugging and using the special privileges of the adb shell to root the device but only \daAdbEnabledPerc\ of devices have USB debugging enabled.
 This is not something that applications running on the phone can exploit to break compartmentalisation and so we do not include those exploits.

### 2014-08-06 securityupdates.tex
Explain why APK signing is important

@@
 Unfortunately many published exploits use adb for convenience and so determining whether that is necessary to exploit the vulnerability can be difficult.
-Some exploits which grant root are not traditional kernel exploits, for example the discovery of flaws in the verification of signatures on Android applications in February 2013~\cite{Forristal2013} meant that applications could pretend to be signed with system keys and hence gain root privileges.
+Some exploits which grant root are not traditional kernel exploits, for example the discovery of flaws in the verification of signatures on Android applications in February 2013~\cite{Forristal2013} meant that applications could pretend to be signed with system keys and hence gain root equivalent privileges.
+On some versions of Android (below 4.1) they could use known system to root escalation mechanisms but on all versions they have a greatly increased attack area for further privilege escalation and also have the ability to control all user internet traffic (via VPNs), brick the phone, remove and install apps, steal user credentials, read the screen, make and receive calls.
 

### 2014-08-06 securityupdates.tex
Add missing ~s before refs

@@
 
-The proportion of Android devices in the DeviceAnalyzer data which were running versions of Android known to have root vulnerabilities at the time is shown in Figure \ref{fig:proportioninsecure}.
+The proportion of Android devices in the DeviceAnalyzer data which were running versions of Android known to have root vulnerabilities at the time is shown in Figure~\ref{fig:proportioninsecure}.
 In \S\ref{sec:background} we will explain where these data come from and then in \S\ref{sec:results} we will show these data in more detail.
@@
 
-We have published full details of the \daNumVulnsUsed\ vulnerabilities used for this analysis in an accompanying technical report~\cite{TODO} and summarise them in Table \ref{tab:andvulns}.
+We have published full details of the \daNumVulnsUsed\ vulnerabilities used for this analysis in an accompanying technical report~\cite{TODO} and summarise them in Table~\ref{tab:andvulns}.
 We also have a website\footnote{\url{http://androidvulnerabilities.org/}} which we will keep up to date with future vulnerabilities and which has successfully solicited submissions from others.
@@
 
-Figure \ref{fig:vulnerabilities} shows which vulnerabilities devices are exposed to.
+Figure~\ref{fig:vulnerabilities} shows which vulnerabilities devices are exposed to.
 For each vulnerability it shows the proportion of devices exposed to that vulnerability and how that changes over time.
@@
 The variation of the proportion of devices affected by a vulnerability with time tells us how bad a particular vulnerability affected the Android platform.
-Figure \ref{fig:nvulnerabilities_heat} breaks this down by vulnerability to show how the proportion of devices affected by different vulnerabilities varies.
+Figure~\ref{fig:nvulnerabilities_heat} breaks this down by vulnerability to show how the proportion of devices affected by different vulnerabilities varies.
 In 2013 three vulnerabilities were found in the way which Android verified the signatures on APKs.
 These allowed the creation of malicious APKs which appear to be signed as system APKs -- which have root equivalent privileges.
-Figure \ref{fig:nvulnerabilities_heat} shows how the the \emph{APK signing vulnerabilities} affected all devices and took months to get fixed for any device.
+Figure~\ref{fig:nvulnerabilities_heat} shows how the the \emph{APK signing vulnerabilities} affected all devices and took months to get fixed for any device.
 However what is perhaps more worrying is the long tail on the \emph{Gingerbreak}, \emph{levitator} and \emph{exploid} vulnerabilities which are more dangerous root exploits (not requiring new APK installation) and which still affect a significant proportion of devices years later.
@@
 \end{figure}
-Figure \ref{fig:from_to_updates} shows how devices upgrade between different versions of Android.
+Figure~\ref{fig:from_to_updates} shows how devices upgrade between different versions of Android.
 Device Analyzer (DA) observes upgrade events and the figure shows which versions a device is upgrading from and to.
@@
 Surprisingly there are also a small number of downgrade events when older versions of Android are installed on to devices.
-The number of devices getting security updates each week, is shown in Figure \ref{fig:weekly_security_updates}, blue shows updates which changed the Android version number from a version with known vulnerabilities to one which had fewer known vulnerabilities.
+The number of devices getting security updates each week, is shown in Figure~\ref{fig:weekly_security_updates}, blue shows updates which changed the Android version number from a version with known vulnerabilities to one which had fewer known vulnerabilities.
 Red indicates updates which changed the build number but not the version number and so might contain a backported fix for a vulnerability.
@@
 While API versions are too coarse grained to use for security update detection they are closely related to OS versions and so if the DA data on API versions is similar to the Google Data on API versions then the DA data on OS versions should be representative.
-Figure \ref{fig:play_api} shows the data from Google and Figure \ref{fig:da_api} shows the data from Device Analyzer and they appear similar.
-Figure \ref{fig:da_gp_comp_diff} shows the difference between Figures \ref{fig:play_api} and \ref{fig:da_api}, normalising for days since the API version was released.
+Figure~\ref{fig:play_api} shows the data from Google and Figure~\ref{fig:da_api} shows the data from Device Analyzer and they appear similar.
+Figure~\ref{fig:da_gp_comp_diff} shows the difference between Figures~\ref{fig:play_api} and \ref{fig:da_api}, normalising for days since the API version was released.
 It shows that the Device Analyzer data systematically overestimates the prevalence of new API versions and underestimates the prevalence of old API versions.

### 2014-08-06 securityupdates.tex
Add that avo is contribution
Clarify what root equivalent means

@@
  \item We characterise the Android update ecosystem.
+ \item We collate information on vulnerabilities which affect Android.
  \item We measure the security of Android according to three metrics and compare different manufacturers. and device models to allow consumers to differentiate between them based on security.
@@
 We only looked for \emph{root equivalent} exploits which did not require USB debugging to exploit.
-\emph{Root equivalent} means that it grants privileges equivalent in scope to root which can then be used to gain root.
+\emph{Root equivalent} means that it grants privileges equivalent in scope to root which can then either be used to gain root or provide sufficiently many permissions as to be as bad as root.
 Some phones can be `rooted' by enabling USB debugging and using the special privileges of the adb shell to root the device but only \daAdbEnabledPerc\ of devices have USB debugging enabled.

### 2014-08-06 securityupdates.tex
Clarify some citations

@@
 Working out the date to use for the date on which a vulnerability was sufficiently well known to pose a threat to users is difficult.
-Frei et al.\ propose the definition of the {\bf time of disclosure} when the information about the vulnerability is freely available to the public, from a widely accepted and independent source and has been validated by security experts so that it has a risk rating.
+Frei et al.~\cite{Frei2010} propose the definition of the {\bf time of disclosure} when the information about the vulnerability is freely available to the public, from a widely accepted and independent source and has been validated by security experts so that it has a risk rating.
 Unfortunately before we collated this information much of it was not published by an independent source and lacked risk rating information, even months or years after they had been actively used.
+Hence we have collated this information and made it publicly available.
 
-Prior work~\cite{Bilge2012} has shown that after public disclosure exploitation rates increase by 5 orders of magnitude and so from the point of view of widespread danger, the period between public disclosure and the update reaching the user's device is the most critical.
+Prior work~\cite{Bilge2012} by Symantec on desktop malware has shown that after public disclosure exploitation rates increase by 5 orders of magnitude and so from the point of view of widespread danger, the period between public disclosure and the update reaching the user's device is the most critical.
 However they also show that zero day exploits are typically used for 312 days before they are publicly disclosed, often to target particular organisations.

### 2014-08-06 securityupdates.tex
Include statistics from avo and footnote links for vulnerability names

@@
 \input{dastats}
+\input{avostats}
 \newcommand{\percMarketShare}{XX\%~\cite{TODO}}
@@
 
+The database of vulnerabilities which we maintain at \texttt{androidvulnerabilities.org} contains \avoNumVulnerabilities\ vulnerabilities of which \avoNumVulnAllAndroid\ affect all android devices and \avoNumVulnSpecific\ are specific to particular devices or manufacturers.
+We only use the \daNumVulnsUsed\ vulnerabilities which it is convenient to use for this analysis because they affect all Android devices rather than particular versions and we know enough information about the vulnerability to tie it to the particular versions it affects.
+
 \subsection{Versions of Android running on devices}
@@
 For each vulnerability it shows the proportion of devices exposed to that vulnerability and how that changes over time.
-In July 2011 at the beginning of the Device Analyzer data the Exploid and Levitator vulnerabilities both affect most Android devices, slowly these are fixed as updates roll out and devices are replaced until in January 2013 a much smaller proportion of devices are affected by known vulnerabilities.
-However when in February 2013 the first APK signing vulnerability was found which affected all previous versions of Android and even in October 2013 most devices were still unfixed.
+In July 2011 at the beginning of the Device Analyzer data the \emph{exploid}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/exploid_udev}} and \emph{levitator}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/levitator}} vulnerabilities both affect most Android devices, slowly these are fixed as updates roll out and devices are replaced until in January 2013 a much smaller proportion of devices are affected by known vulnerabilities.
+However when in February 2013 the first APK signing vulnerability was found which affected all previous versions of Android and even in October 2013 most devices were still vulnerable.
 %TODO put percentages on these things to quantify them
@@
 Figure~\ref{fig:nvulnerabilities_heat} shows how the the \emph{APK signing vulnerabilities} affected all devices and took months to get fixed for any device.
-However what is perhaps more worrying is the long tail on the \emph{Gingerbreak}, \emph{levitator} and \emph{exploid} vulnerabilities which are more dangerous root exploits (not requiring new APK installation) and which still affect a significant proportion of devices years later.
+However what is perhaps more worrying is the long tail on the \emph{Gingerbreak}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/Gingerbreak}}, \emph{levitator} and \emph{exploid} vulnerabilities which are more dangerous root exploits (not requiring new APK installation) and which still affect a significant proportion of devices years later.
 

### 2014-08-06 securityupdates.tex
Add subsection heading for android update process
Include statistics on upgrade and downgrade events

@@
 While we have not tried to exploit these vulnerabilities on the devices in order to test whether the devices are actually vulnerable, we are confident that most devices will be if they are running a vesion known to be vulnerable.
+%TODO WHY?
+
+\subsection{Android update process}
+TODO
 
@@
 Device Analyzer (DA) observes upgrade events and the figure shows which versions a device is upgrading from and to.
-Mostly the dark cells are above the diagonal and are upgrades.
+Mostly the dark cells are above the diagonal and are upgrades (\daNumUpdatesUpgrades).
 While many upgrades are from one version to the next version there are also a fair number which skip a few versions.
-Surprisingly there are also a small number of downgrade events when older versions of Android are installed on to devices.
+Surprisingly there are also a small number of downgrade events (\daNumUpdatesDowngrades, \daPercUpdatesDowngrades) when older versions of Android are installed on to devices.
+
 The number of devices getting security updates each week, is shown in Figure~\ref{fig:weekly_security_updates}, blue shows updates which changed the Android version number from a version with known vulnerabilities to one which had fewer known vulnerabilities.

### 2014-08-07 security_scoring.tex
Note our assumption that newer versions are better
Expand discussion of security scoring metrics and add a diagram to explain m

@@
   \item[$f$] The proportion of the time where devices were free from known critical vulnerabilities.
-This is equivalent to Acer and Jackson's proposal~\cite{Acer2010} to measure the security based on the proportion of users with at least one unpatched critical vulnerability and is the Vulnerability Free Days score~\cite{Wright2014}.
+This is equivalent to Acer and Jackson's proposal~\cite{Acer2010} to measure the security based on the proportion of users with at least one unpatched critical vulnerability and to the Vulnerability Free Days (VFD) score~\cite{Wright2014}.
+  Unlike VFD this is the proportion of running devices which were free from critical vulnerabilities, rather than days which the manufacturer was free from outstanding critical vulnerabilities as that does not take account of the update process.
   \item[$u$] The proportion of devices run the latest version of Android shipped to any device produced by that manufacturer.
 This is a measure of internal updatedness, a low score would mean many devices are being left behind.
+  This assumes that newer versions are better with stronger security, historically steps have been taken to improve Android security in newer versions so this assumption should generally hold, but sometimes new updates introduce new vulnerabilities.
   \item[$m$] The mean number of outstanding vulnerabilities affecting devices not fixed on any device shipped by the manufacturer.
 This is related to the Median Active Vulnerabilities measure~\cite{Wright2014} but is the mean rather than the median, as this gives a continuous value.
+  An example is given in Figure~\ref{fig:mcalculation}.
 %TODO should we compute the median instead?
+\end{description}
+
+\begin{figure}
+\centering
+%\includegraphics[width=\columnwidth]{figures/mcalculation}
+\include{figures/mcalculation}
+\begin{tabular}{c|c|c}
+active vulnerabilities & \# days & percent\\
+\hline
+0 & 3 & 15\%\\
+1 & 5 & 25\%\\
+2 & 10 & 50\%\\
+3 & 2 & 10\%\\
+\hline
+\hline
+$m$ & \multicolumn{2}{c}{1.55}\\
+MAV & \multicolumn{2}{c}{2}\\
+VFD & \multicolumn{2}{c}{15\%}\\
+
+\end{tabular}
+\caption{Example $m$ calculation based on the one given in by Wright~\cite{Wright2014} giving MAV and VFD for comparison.}
+\label{fig:mcalculation}
+\end{figure}
 
@@
 Since $m$ is an unbounded positive real number, we map it into the range [0--1).
-\end{description}
+
 \begin{equation}

### 2014-08-07 security_scoring.tex
Discus the gaming of security scores

@@
   \item[$m$] The mean number of outstanding vulnerabilities affecting devices not fixed on any device shipped by the manufacturer.
-This is related to the Median Active Vulnerabilities measure~\cite{Wright2014} but is the mean rather than the median, as this gives a continuous value.
+This is related to the Median Active Vulnerabilities (MAV) measure~\cite{Wright2014} but is the mean rather than the median, as this gives a continuous value.
   An example is given in Figure~\ref{fig:mcalculation}.
@@
 
-We want to provide a score out of 10, since $f$ is the most important metric we weight it more highly.
+These three metrics $f$, $u$ and $m$ together measure the security of a platform with respect to known vulnerabilities and updates.
+$f$ is a key measure of the direct risk to users as if there is any known but unfixed vulnerability then they are vulnerable.
+However it does not capture the increased risk caused by there being multiple known vulnerabilities which gives an attacker more opportunities and increases the likelihood of a piece of malware having a matching exploit.
+This is captured by the $m$ score which measures the size of the manufacturers queue of outstanding vulnerabilities but does not take into account the update process or measure the actual end user security.
+Neither of these metrics capture whether devices are being left behind and not being kept up to date with the most recent (and hopefully most secure) version which is captured by $u$.
+
+$f$ is hard to game without doing a good job at security but it doesn't get any worse if there is already one known vulnerability and another is found.
+A high value of $u$ could be achieved by only ever shipping one version but that would give low values for $f$ and $m$ (and not be attractive to new customers).
+A high value of $m$ could be achieved by focusing on only one device at a time and ensuring that gets updates but ignoring all others, but that would lower $f$ and $u$.
+
+We want to provide a score out of 10 as this is easy for people to understand.
+Since $f$ is the most important metric we weight it more highly.
 Since $m$ is an unbounded positive real number, we map it into the range [0--1).
+This gives us Equation~\ref{eq:score}:
 
 \begin{equation}
-\mathrm{score} = f\times 4 + u \times 3 + \frac{2}{1+e^m} \times 3
+\mathrm{score} = f\times 4 + u \times 3 + \frac{2}{1+e^m} \times 3 \label{eq:score}
 \end{equation}

### 2014-08-07 securityupdates.tex
Expand discussion of security scoring metrics and add a diagram to explain m
Minor fixes and clarifications

@@
 \usepackage{subcaption}
+\usepackage{tikz}
 \RequirePackage[date=terse,isbn=true,doi=true,url=false,urldate=iso8601,maxbibnames=9,backref=false,backend=bibtex]{biblatex}
@@
 \section{Introduction}
-Android has \percMarketShare\ of the smartphone market and while the core development is controlled by Google there are about \daNumManufacturers\footnote{From the Device Analyzer data} manufacturers which make devices which run Android.
+Android has \percMarketShare\ of the smartphone market and while the core development is controlled by Google there are about \daNumManufacturers\footnote{We computed this from the Device Analyzer~\cite{Wagner2013} data.} manufacturers which make devices which run Android.
 Android receives regular updates  which add new features and fix bugs, including security vulnerabilities.
@@
 
-The proportion of Android devices in the DeviceAnalyzer data which were running versions of Android known to have root vulnerabilities at the time is shown in Figure~\ref{fig:proportioninsecure}.
+The proportion of Android devices in the Device Analyzer data which were running versions of Android known to have root vulnerabilities at the time is shown in Figure~\ref{fig:proportioninsecure}.
 In \S\ref{sec:background} we will explain where these data come from and then in \S\ref{sec:results} we will show these data in more detail.
@@
 Surprisingly there are also a small number of downgrade events (\daNumUpdatesDowngrades, \daPercUpdatesDowngrades) when older versions of Android are installed on to devices.
+Possible reasons why users are downgrading are to free up space on their device, to make it easier to root or because a new version introduced bugs.
 
@@
 
+\section{Discussion}
+There are continuing efforts to reduce the impact of root vulnerabilities, both in Android and more widely.
+SEAndroid~\cite{Smalley2013} which is included in Android from version ???~\cite{TODO} claimed to prevent some root vulnerabilities and to reduce the impact of others.
+Capability based enforcement systems such as Capsicum~\cite{Watson2010} substantially reduce the capabilities that an exploit has to try and gain increased privilege with.
+When Capsicum is included in Linux~\cite{TODO} and hence in Android, it could be used to place fine grained restrictions on system daemons, preventing vulnerabilities becoming root vulnerabilities.
+
 \section{Related work}
@@
 We found that since 2011 \daMeanInsecurityPerc\ of devices on average were exposed to known root vulnerabilities, in 2011 Felt et al.\ studied 6 Android handsets and found that they were exposed to root vulnerabilities at least 74\% of the time.
-This might indicate that over the last 3 years the security of Android has not improved, but these data are not directly comparable.
+This might indicate that over the last 3 years the security of Android has not improved, but these data are not directly comparable as their study was on 6 handsets considering the best possible update distribution while ours is on a large sample of devices with the real update distribution.
 They also found that 4 of the 46 malware samples (8\%) they analysed contained root exploits, much lower than rates found in later larger studies which found rates of 36.7\%~\cite{Zhou2012b} and 40\%~\cite{Zhou2012a}.
@@
 The update process itself can allow Android apps to gain privilege through `Pileup' vulnerabilities~\cite{Xing2014} by registering for new permissions before the update which creates that permission is installed.
-
-There are continuing efforts to reduce the impact of root vulnerabilities, both in Android and more widely.
-SEAndroid~\cite{Smalley2013} which is included in Android from version ???~\cite{TODO} claimed to prevent some root vulnerabilities and to reduce the impact of others.
-Capability based enforcement systems such as Capsicum~\cite{Watson2010} substantially reduce the capabilities that an exploit has to try and gain increased privilege with.
-When Capsicum is included in Linux~\cite{TODO} and hence in Android, it could be used to place fine grained restrictions on system daemons, preventing vulnerabilities becoming root vulnerabilities.
 
@@
 The updatedness rate for Android of \daUpdatednessPerc\ compares unfavourable with the more than 90\% rate for Windows XP SP2 computers contacting the Microsoft update servers~\cite{Gkantsidis2006} though that is within one OS version and only for those computers which did connect to the servers -- but that is the default.
+However 27\% of Windows computers were still running Windows XP in July 2014\footnote{\url{https://archive.today/PLGxn}} -- four months after it went out of security support.
 

### 2014-08-07 securityupdates.tex
Add draft update ecosystem figure
Note the start and end date of DA and AVO data
Add numbers or placeholders to update ecosystem diagram

@@
 \usepackage{graphicx}
+\graphicspath{figures/}
+\usepackage{import}
 \usepackage{url}
@@
 \newcommand{\daMonthsDevices}{XXX}
+% Num versions since \daStartDate
+\newcommand{\bouncyCastleNumVersions}{A}
+\newcommand{\opensslNumVersions}{B}
+\newcommand{\apacheLibsNumVersions}{C}
+\newcommand{\apacheLibsNumLibs}{E}
+\newcommand{\linuxNumVersions}{D}
 
@@
 \subsection{Android update process}
+\begin{figure}
+ \centering
+ \def\svgwidth{\columnwidth}
+ \import{figures/}{update_ecosystem.pdf_tex}
+ \caption{Flow of updates between participants in the Android ecosystem.
+ Numbers on edges indicate updates shipped between \daStartDate\ and \daEndDate, numbers in brackets represent number of such entities in our data.}
+ \label{fig:update_ecosystem}
+\end{figure}
+\cite{Vidas2011}
 TODO
@@
 %TODO summarise our results
+The Device Analyzer data used here was collected between \daStartDate\ and \daEndDate.
+The \texttt{androidvulnerabilities.org} data was collected between \avoStartDate\ and \avoEndDate.
 %TODO comparison of nexus update rates with iOS

### 2014-08-08 countversions.sh
Count how many openssl versions there have been

@@ -1,0 +1,21 @@
+#!/bin/bash
+
+set -e
+
+start_date='2011-07-01'
+end_date='2014-07-31'
+openssl_count=0
+output=`wget --quiet -O - https://www.openssl.org/source/ | grep '.tar.gz' | cut -d'<' -f 1 |  sed 's/^ [0-9]\+ //' | tr ' ' '_'`
+for i in $output
+do
+    i=`echo $i | tr '_' ' '`
+    d=`date --date="$i" --rfc-3339='date'`
+    if [[ $d > $start_date ]]
+    then
+        if [[ $d < $end_date  ]]
+        then
+            openssl_count=$((openssl_count+1))
+        fi
+    fi
+done
+echo "\\newcommand{\\opensslNumVersions}{$openssl_count}"

### 2014-08-08 countversions.sh
Count linux versions

@@ -4,18 +4,25 @@ set -e
 
 start_date='2011-07-01'
 end_date='2014-07-31'
-openssl_count=0
-output=`wget --quiet -O - https://www.openssl.org/source/ | grep '.tar.gz' | cut -d'<' -f 1 |  sed 's/^ [0-9]\+ //' | tr ' ' '_'`
-for i in $output
-do
-    i=`echo $i | tr '_' ' '`
-    d=`date --date="$i" --rfc-3339='date'`
-    if [[ $d > $start_date ]]
-    then
-        if [[ $d < $end_date  ]]
+function do_count {
+    name=$1
+    output=$2
+    count=0
+    for i in $output
+    do
+        i=`echo $i | tr '_' ' '`
+        d=`date --date="$i" --rfc-3339='date'`
+        if [[ $d > $start_date ]]
         then
-            openssl_count=$((openssl_count+1))
+            if [[ $d < $end_date  ]]
+            then
+                count=$((count+1))
+            fi
         fi
-    fi
-done
-echo "\\newcommand{\\opensslNumVersions}{$openssl_count}"
+    done
+    echo "\\newcommand{\\${name}NumVersions}{$count}"
+}
+do_count 'openssl' "`wget --quiet -O - https://www.openssl.org/source/ | grep '.tar.gz' | cut -d'<' -f 1 |  sed 's/^ [0-9]\+ //' | tr ' ' '_'`"
+do_count 'linux' "`wget --quiet -O - https://www.kernel.org/pub/linux/kernel/v2.6/ | grep 'linux-' | grep '.tar.gz' | tr -s ' ' | cut -d' ' -f 3`
+`wget --quiet -O - https://www.kernel.org/pub/linux/kernel/v3.0/ | grep 'linux-' | grep '.tar.gz' | tr -s ' ' | cut -d' ' -f 3`"
+

### 2014-08-11 securityupdates.tex
Incorporate countversions data
Count linux versions
Add some text on the Android update process

@@
 \newcommand{\bouncyCastleNumVersions}{A}
-\newcommand{\opensslNumVersions}{B}
+\input{countversions}
+%\newcommand{\opensslNumVersions}{B}
 \newcommand{\apacheLibsNumVersions}{C}
 \newcommand{\apacheLibsNumLibs}{E}
-\newcommand{\linuxNumVersions}{D}
+%\newcommand{\linuxNumVersions}{D}
 
@@
 \end{figure}
-\cite{Vidas2011}
-TODO
+A simple update process would be for a developer, who wrote the program, to write an update and ship it to the user who installs it.
+The process for Android is rather more complicated and is shown in Figure~\ref{fig:update_ecosystem}.
+Android builds on various open source projects such as the Linux kernel, the OpenSSL and BouncyCastle cryptography libraries and various Apache Java libraries.
+It also incorporates various drivers for different bits of hardware.
+The Android platform is then built on top of those with kernel and userspace components by Google and the Open Handset Alliance. %TODO work out the exact arrangements.
+Usually the code for each update is kept secret until after the update has been published.
+Then the Manufacturers take stock Android and customise it with different themes, apps, drivers and so on and ship updates on, to the carrier.
+The Carrier may then make further customisations and do further testing before shipping the update to the User.
+Sometimes Manufactures ship direct to the user, sometimes the Manufacturer and Google are the same entity (at least for the purpose of a particular phone, such as with Nexus phones).
+An analysis by Vidas et al.~\cite{Vidas2011} of the Android 2.1 to 2.2 update found that it took 11 months from when Google released 2.2 for the last device which they were investigating to get the update.
+This multi-stage multi-party release process increases the exposure of Android devices to vulnerabilities as it takes a long time for them to be patched.
 

### 2014-08-11 security_scoring.tex
Add note on gaming DA
Touch up security scoring section

@@
   \item[$f$] The proportion of the time where devices were free from known critical vulnerabilities.
-This is equivalent to Acer and Jackson's proposal~\cite{Acer2010} to measure the security based on the proportion of users with at least one unpatched critical vulnerability and to the Vulnerability Free Days (VFD) score~\cite{Wright2014}.
+This is equivalent to Acer and Jackson's proposal~\cite{Acer2010} to measure the security based on the proportion of users with at least one unpatched critical vulnerability and similar to the Vulnerability Free Days (VFD) score~\cite{Wright2014}.
   Unlike VFD this is the proportion of running devices which were free from critical vulnerabilities, rather than days which the manufacturer was free from outstanding critical vulnerabilities as that does not take account of the update process.
-  \item[$u$] The proportion of devices run the latest version of Android shipped to any device produced by that manufacturer.
+  \item[$u$] The proportion of devices which run the latest version of Android shipped to any device produced by that manufacturer.
 This is a measure of internal updatedness, a low score would mean many devices are being left behind.
@@
 A high value of $m$ could be achieved by focusing on only one device at a time and ensuring that gets updates but ignoring all others, but that would lower $f$ and $u$.
+One way to influence our scores would be to add additional devices to Device Analyzer which have good security, these would have to be real end user devices as we could detect fake ones, this would increase the size of our data set and would require providing genuinely good security to some users.
 
@@
 Manufacturers are considered significant if we have data from at least \daSigNumDevices\ devices and at least \daSigNumDeviceDays\ days of contributions.
+Additionally for $m$ and $u$ we ignore days with less than \daSigNumDevicesDay\ devices contributing to that day's score.
 \daTabSecScoresmodel

### 2014-08-11 security_scoring.tex
Add operator table and text to security scores

@@
 Additionally for $m$ and $u$ we ignore days with less than \daSigNumDevicesDay\ devices contributing to that day's score.
+
 \daTabSecScoresmodel
@@
 Table~\ref{tab:sec_model} shows the results for the \daNumSigModels\ device models which have a significant presence by the same metric with \daSecScoreBestmodel\ (\daSecScoreBestmodelScore\ / 10) doing best and \daSecScoreWorstmodel\ (\daSecScoreWorstmodelScore\ / 10) doing worst.
+
+\daTabSecScoresoperator
+We also analysed different carriers, for the \daNumSigOperators\ operators with a significant presence in our data Table~\ref{tab:sec_operator} shows the results \daSecScoreBestoperator\ (\daSecScoreBestoperatorScore\ / 10) doing best and \daSecScoreWorstoperator\ (\daSecScoreWorstoperatorScore\ / 10) doing worst.
 

### 2014-08-11 security_scoring.tex
Consistently use the term 'operator' rather than 'carrier'

--- 2014-08-11 security_scoring.tex
@@
 Table~\ref{tab:sec_model} shows the results for the \daNumSigModels\ device models which have a significant presence by the same metric with \daSecScoreBestmodel\ (\daSecScoreBestmodelScore\ / 10) doing best and \daSecScoreWorstmodel\ (\daSecScoreWorstmodelScore\ / 10) doing worst.
+
+\daTabSecScoresoperator
+We also analysed different operators, for the \daNumSigOperators\ operators with a significant presence in our data Table~\ref{tab:sec_operator} shows the results \daSecScoreBestoperator\ (\daSecScoreBestoperatorScore\ / 10) doing best and \daSecScoreWorstoperator\ (\daSecScoreWorstoperatorScore\ / 10) doing worst.
 

### 2014-08-11 securityupdates.tex
Consistently use the term 'operator' rather than 'carrier'

@@
 Usually the code for each update is kept secret until after the update has been published.
-Then the Manufacturers take stock Android and customise it with different themes, apps, drivers and so on and ship updates on, to the carrier.
-The Carrier may then make further customisations and do further testing before shipping the update to the User.
+Then the Manufacturers take stock Android and customise it with different themes, apps, drivers and so on and ship updates on, to the operator.
+The Operator may then make further customisations and do further testing before shipping the update to the User.
 Sometimes Manufactures ship direct to the user, sometimes the Manufacturer and Google are the same entity (at least for the purpose of a particular phone, such as with Nexus phones).

### 2014-08-11 securityupdates.tex
reformat biblatex line
Fill in some XXXX values for the Device Analyzer data

@@
 \usepackage{tikz}
-\RequirePackage[date=terse,isbn=true,doi=true,url=false,urldate=iso8601,maxbibnames=9,backref=false,backend=bibtex]{biblatex}
+\RequirePackage[date=terse, isbn=true, doi=true, url=false, urldate=iso8601, maxbibnames=9, backref=false, backend=bibtex]{biblatex}
 \addbibresource{securityupdates.bib}
@@
 \newcommand{\daNumNetworkOperators}{XXX~[\S\ref{sec:results}]}
-\newcommand{\daNumDevices}{XXXX}
-\newcommand{\daDeviceDays}{XXXXX}
-\newcommand{\daMonths}{X}% Number of months for which we have the below number of devices
-\newcommand{\daMonthsDevices}{XXX}
+\newcommand{\daNumDevices}{\daNumOSDevices}
+\newcommand{\daDeviceDays}{\daOSTotalDaysData}
 % Num versions since \daStartDate
@@
 The only suitable data which is available for this is the Device Analyzer data~\cite{Wagner2013}.
-The Device Analyzer data contains data from \daNumDevices\ devices with a total of \daDeviceDays\ device days, it contains data for more than \daMonths months for \daMonthsDevices\ devices.
+The Device Analyzer data contains data from \daNumDevices\ devices with a total of \daDeviceDays\footnote{Here we are only counting devices and days for which we have valid OS version data} device days, it contains data for more than \daMonths\ months for \daMonthsDevices\ devices.
 %TODO talk about the Device Analyzer project a bit

### 2014-08-11 securityupdates.tex
Fill in some more placeholders

@@
 \newcommand{\percMarketShare}{XX\%~\cite{TODO}}
-\newcommand{\daNumNetworkOperators}{XXX~[\S\ref{sec:results}]}
 \newcommand{\daNumDevices}{\daNumOSDevices}
@@
 \section{Introduction}
-Android has \percMarketShare\ of the smartphone market and while the core development is controlled by Google there are about \daNumManufacturers\footnote{We computed this from the Device Analyzer~\cite{Wagner2013} data.} manufacturers which make devices which run Android.
+Android has \percMarketShare\ of the smartphone market and while the core development is controlled by Google there are about \daNumManufacturers\footnote{\label{foot:dadata}We computed this from the Device Analyzer~\cite{Wagner2013} data, see \S\ref{sec:results}.} manufacturers which make devices which run Android.
 Android receives regular updates  which add new features and fix bugs, including security vulnerabilities.
-Many of these manufacturers customise the version of Android they ship and sometimes network operators (of which there are at least \daNumNetworkOperators) make further modifications.
+Many of these manufacturers customise the version of Android they ship and sometimes network operators (of which there are at least \daNumOperators\textsuperscript{\ref{foot:dadata}}) make further modifications.
 Hence when Google produces an update to Android, the update may have to pass through the manufacturer and operator before reaching the user.
@@
 \label{sec:results}
-In this section we present the results of our analysis showing \daMeanInsecurityPerc\ of devices to have root vulnerabilities on average, nexus devices to be better than others, \daSecScoreBestmanufacturer\ to be the best manufacturer and ????.
+In this section we present the results of our analysis showing \daMeanInsecurityPerc\ of devices to have root vulnerabilities on average, nexus devices to be better than others, \emph{\daSecScoreBestmanufacturer} to be the best manufacturer and \emph{\daSecScoreBestoperator} to be the best network operator.
 %TODO summarise our results

### 2014-08-11 securityupdates.tex
Consistently use XXX rather than ??? for unknown values

@@
 There are continuing efforts to reduce the impact of root vulnerabilities, both in Android and more widely.
-SEAndroid~\cite{Smalley2013} which is included in Android from version ???~\cite{TODO} claimed to prevent some root vulnerabilities and to reduce the impact of others.
+SEAndroid~\cite{Smalley2013} which is included in Android from version XXX~\cite{TODO} claimed to prevent some root vulnerabilities and to reduce the impact of others.
 Capability based enforcement systems such as Capsicum~\cite{Watson2010} substantially reduce the capabilities that an exploit has to try and gain increased privilege with.

### 2014-08-11 security_scoring.tex
Qualify operator scores

@@
 We also analysed different operators, for the \daNumSigOperators\ operators with a significant presence in our data Table~\ref{tab:sec_operator} shows the results \daSecScoreBestoperator\ (\daSecScoreBestoperatorScore\ / 10) doing best and \daSecScoreWorstoperator\ (\daSecScoreWorstoperatorScore\ / 10) doing worst.
+However the score of an operator is affected by the manufacturers of the devices which are in use on it's network.
+This is in turn affected by both what devices an operator offers to users and on which devices users choose.
+Hence having a worse score does not necessarily mean that an operator is worse, it could be that its users all pick phones from a worse manufacturer, for example because they were cheaper.
 

### 2014-08-11 securityupdates.tex
Add device data graph and text

@@
 
+Those graphs summarise data across all the devices, however one of the advantages of the Device Analyzer data is that it allows us to look at what happens to individual devices over time.
+Figure~\ref{fig:device_data} shows how the number of vulnerabilities affecting the \daNumDeviceDataDevices\ devices which have contributed the most days of data to Device Analyzer changes over time.
+The trend that we saw in Figure~\ref{fig:proportioninsecure} of security improving and then getting worse is also shown here.
+It shows how some devices had vulnerabilities, which were fixed, and then further vulnerabilities were discovered, and for these devices, mostly not fixed.
+\begin{figure}
+ \includegraphics[width=0.5\textwidth]{figures/device-data-all-security}
+ \caption{Number of vulnerabilities affecting the top \daNumDeviceDataDevices\ devices by days of contribution in the Device Analyzer data}
+ \label{fig:device_data}
+\end{figure}
+
 \subsection{Update frequency}

### 2014-08-11 figures/update_figures.sh

@@ -8,6 +8,7 @@ cp "$fromdir"norm_api-notitle.pdf da_norm_api-notitle.pdf
 cp "$fromdir"norm_os.pdf da_norm_os.pdf
 cp "$fromdir"os.pdf da_os.pdf
 cp "$fromdir"device_data/device_data-lg-security.pdf device_data-LG.pdf
+cp "$fromdir"device_data/device-data-all-security.pdf device-data-all-security.pdf
 cp "$fromdir"from_to_updates.pdf from_to_updates.pdf
 cp "$fromdir"googleplayapi.pdf googleplayapi.pdf
 cp "$fromdir"googleplayapi-notitle.pdf googleplayapi-notitle.pdf

### 2014-08-11 securityupdates.tex
Rewrite android update process section following Alastair's comments

@@
 \section{Introduction}
-Android has \percMarketShare\ of the smartphone market and while the core development is controlled by Google there are about \daNumManufacturers\footnote{\label{foot:dadata}We computed this from the Device Analyzer~\cite{Wagner2013} data, see \S\ref{sec:results}.} manufacturers which make devices which run Android.
+Android has \percMarketShare\ of the smartphone market and while the core development is controlled by Google there are about \daNumManufacturers\footnote{\label{foot:dadata}We computed this from the Device Analyzer~\cite{Wagner2013} data, see \S\ref{sec:android_update_process}.} manufacturers which make devices which run Android.
 Android receives regular updates  which add new features and fix bugs, including security vulnerabilities.
@@
 \subsection{Android update process}
+\label{sec:android_update_process}
 \begin{figure}
@@
  \caption{Flow of updates between participants in the Android ecosystem.
- Numbers on edges indicate updates shipped between \daStartDate\ and \daEndDate, numbers in brackets represent number of such entities in our data.}
+ Numbers on edges indicate updates shipped between \daStartDate\ and \daEndDate, numbers in brackets represent number of such entities in our data.
+ Dotted arrows indicate flows where we do not know how many updates are being produced.}
  \label{fig:update_ecosystem}
 \end{figure}
-A simple update process would be for a developer, who wrote the program, to write an update and ship it to the user who installs it.
-The process for Android is rather more complicated and is shown in Figure~\ref{fig:update_ecosystem}.
-Android builds on various open source projects such as the Linux kernel, the OpenSSL and BouncyCastle cryptography libraries and various Apache Java libraries.
+
+A model of the Android update process is shown in Figure~\ref{fig:update_ecosystem}.
+There are five entities or groups which contribute towards Android updates: the network operators, the device manufacturers, the hardware developers, Google and the upstream open source projects.
+Android builds on various open source projects such as the Linux kernel, the OpenSSL and BouncyCastle cryptography libraries and various Apache Java libraries and so can include any of the versions of those projects, including those which fix security vulnerabilities in them.
 It also incorporates various drivers for different bits of hardware.
-The Android platform is then built on top of those with kernel and userspace components by Google and the Open Handset Alliance. %TODO work out the exact arrangements.
-Usually the code for each update is kept secret until after the update has been published.
-Then the Manufacturers take stock Android and customise it with different themes, apps, drivers and so on and ship updates on, to the operator.
-The Operator may then make further customisations and do further testing before shipping the update to the User.
+The Android platform is then built on top of those with kernel and userspace components by Google. %TODO work out the exact arrangements.
+Usually the code for each update is kept secret until after the update has been published though through the Open Handset Alliance some manufacturers see it in advance.
+Then the Manufacturers take the update and customise it before passing it on to the operator.
+The Operator may then make further customisations and do further testing before shipping the update to the Device.
 Sometimes Manufactures ship direct to the user, sometimes the Manufacturer and Google are the same entity (at least for the purpose of a particular phone, such as with Nexus phones).
-An analysis by Vidas et al.~\cite{Vidas2011} of the Android 2.1 to 2.2 update found that it took 11 months from when Google released 2.2 for the last device which they were investigating to get the update.
-This multi-stage multi-party release process increases the exposure of Android devices to vulnerabilities as it takes a long time for them to be patched.
+Sometimes Manufacturers incorporate upstream open source project releases directly, and sometimes incorrectly -- for example including a broken daily build of sqllite in their release of Android
+%An analysis by Vidas et al.~\cite{Vidas2011} of the Android 2.1 to 2.2 update found that it took 11 months from when Google released 2.2 for the last device which they were investigating to get the update.
+%TODO talk about update latency
+
+The numbers of Devices (\daNumOSDevices), Operators (\daNumOperators) and Manufacturers (\daNumManufacturers) in Figure~\ref{fig:update_ecosystem} come from the Device Analyzer data.
+Manufacturer and Operator counts were obtained by normalising the results reported by Android to Device Analyzer of the device manufacturer and active network operator.
+This normalisation involves removing invalid values (such as `manufacturer' or `airplane mode is on'), collating across company name changes (e.g.\ `lge' to `lg'), normalising punctuation, removing extra strings sometimes added such as (`(2g)' or `communications') and mapping some incorrectly placed model names back to their manufacturer.
+This normalisation is not perfect and so these are overestimates on the Device Analyzer data but they are likely still underestimates as there will be some manufacturers and operators which are not included in the Device Analyzer data.
+
+The number of updates received by devices (\daNumFullVersions) is the number of different full version strings observed in Device Analyzer.
+The number of updates shipped by Google (\daNumSigOSVersions) is the number of Android versions reported in Device Analyzer which affected more than \daSigVersionPerc\ of devices for more than \daSigVersionDays\ days.
+%TODO this is the wrong number to be using
+
+The number of updates shipped by upstream open source projects is calculated by scraping the releases pages for those projects.
 

### 2014-08-12 security_scoring.tex
Remove mcalculation table and expand the caption text instead

@@
 \include{figures/mcalculation}
-\begin{tabular}{c|c|c}
-active vulnerabilities & \# days & percent\\
-\hline
-0 & 3 & 15\%\\
-1 & 5 & 25\%\\
-2 & 10 & 50\%\\
-3 & 2 & 10\%\\
-\hline
-\hline
-$m$ & \multicolumn{2}{c}{1.55}\\
-MAV & \multicolumn{2}{c}{2}\\
-VFD & \multicolumn{2}{c}{15\%}\\
-
-\end{tabular}
-\caption{Example $m$ calculation based on the one given in by Wright~\cite{Wright2014} giving MAV and VFD for comparison.}
+\caption{As vulnerabilities are discovered and patched the sum of known but unpatched vulnerabilities each day varies.
+From this we can calculate $m = (0 \times 3 + 1 \times 5 + 2 \times 10 + 3 \times 2) / 20 = 1.55$ For comparison VFD = 0.15 and MAV = 2.
+Example based on the one given by Wright~\cite{Wright2014}.}
 \label{fig:mcalculation}

### 2014-08-12 figures/update_figures.sh
Update graphs

@@ -7,7 +7,7 @@ cp "$fromdir"norm_api.pdf da_norm_api.pdf
 cp "$fromdir"norm_api-notitle.pdf da_norm_api-notitle.pdf
 cp "$fromdir"norm_os.pdf da_norm_os.pdf
 cp "$fromdir"os.pdf da_os.pdf
-cp "$fromdir"device_data/device_data-lg-security.pdf device_data-LG.pdf
+cp "$fromdir"device_data/device-data-manufacturer-lg-security.pdf device_data-LG.pdf
 cp "$fromdir"device_data/device-data-all-security.pdf device-data-all-security.pdf
 cp "$fromdir"from_to_updates.pdf from_to_updates.pdf
 cp "$fromdir"googleplayapi.pdf googleplayapi.pdf

### 2014-08-13 security_scoring.tex
Add full version graphs for best and worst models to show that they are reasonably classed as such

@@
 Table~\ref{tab:sec_model} shows the results for the \daNumSigModels\ device models which have a significant presence by the same metric with \daSecScoreBestmodel\ (\daSecScoreBestmodelScore\ / 10) doing best and \daSecScoreWorstmodel\ (\daSecScoreWorstmodelScore\ / 10) doing worst.
+We can then test whether this seems fair by comparing the version data for the best and worst models.
+Figure~\ref{fig:worst_model_full} shows the full version distribution for \daSecScoreWorstmodel\ and Figure~\ref{fig:best_model_full} shows the same for \daSecScoreBestmodel\ there is a marked difference in the distribution of updates between the two with only \daSecScoreWorstmodelNumFullVersions\ full versions for \daSecScoreWorstmodel\ and \daSecScoreBestmodelNumFullVersions\ for \daSecScoreBestmodel.
+
+\begin{figure}
+ \centering
+ \begin{subfigure}[b]{\columnwidth}
+ \includegraphics[width=\columnwidth]{figures/full_version_\daSecScoreWorstmodel.pdf}
+ \caption{\daSecScoreWorstmodel}
+ \label{fig:worst_model_full}
+\end{subfigure}
+\begin{subfigure}[b]{\columnwidth}
+ \includegraphics[width=\columnwidth]{figures/full_version_\daSecScoreBestmodel.pdf}
+ \caption{\daSecScoreBestmodel}
+ \label{fig:best_model_full}
+\end{subfigure}
+\caption{Full version distributions for the best and worst model}
+\end{figure}
 

### 2014-08-13 figures/update_figures.sh

@@ -27,3 +27,5 @@ cp "$fromdir"api_gpcomp_diff.pdf api_gpcomp_diff.pdf
 cp "$fromdir"api_gpcomp_rdiff.pdf api_gpcomp_rdiff.pdf
 cp "$fromdir"ua-versions-android-month-norm.pdf ua-versions-android-month-norm.pdf
 cp "$fromdir"ua-versions-ios-month-norm.pdf ua-versions-ios-month-norm.pdf
+cp "$fromdir""model/Galaxy Nexus/norm_full.pdf" "full_version_Galaxy Nexus.pdf"
+cp "$fromdir""model/HTC Desire HD A9191/norm_full.pdf" "full_version_HTC Desire HD A9191.pdf"

### 2014-08-13 securityupdates.tex
Fix width of figures to use columnwidth
Add full version graphs for best and worst models to show that they are reasonably classed as such
Add the start of a missing citation after I stumbled across it

@@
 \usepackage{graphicx}
+\usepackage[space]{grffile}
 \graphicspath{figures/}
@@
 \begin{figure}
- \includegraphics[width=0.5\textwidth]{figures/nvulnerabilities_heat.pdf}
+ \includegraphics[width=\columnwidth]{figures/nvulnerabilities_heat.pdf}
  \caption{Proportion of devices affected by different vulnerabilities}
@@
 \begin{figure}
- \includegraphics[width=0.5\textwidth]{figures/device-data-all-security}
+ \includegraphics[width=\columnwidth]{figures/device-data-all-security}
  \caption{Number of vulnerabilities affecting the top \daNumDeviceDataDevices\ devices by days of contribution in the Device Analyzer data}
@@
 \begin{figure}
- \includegraphics[width=0.5\textwidth]{figures/from_to_updates.pdf}
+ \includegraphics[width=\columnwidth]{figures/from_to_updates.pdf}
  \caption{Updates between different Android versions in the DA data}
@@
 \begin{figure}
- \includegraphics[width=0.5\textwidth]{figures/w_security_updates.pdf}
+ \includegraphics[width=\columnwidth]{figures/w_security_updates.pdf}
  \caption{Number of updates each week which fixed or may have fixed security vulnerabilities}
@@
 There are continuing efforts to reduce the impact of root vulnerabilities, both in Android and more widely.
-SEAndroid~\cite{Smalley2013} which is included in Android from version XXX~\cite{TODO} claimed to prevent some root vulnerabilities and to reduce the impact of others.
+SEAndroid~\cite{Smalley2013} which is included in Android from version 4.1~\cite{https://developer.android.com/about/versions/jelly-bean.html} claimed to prevent some root vulnerabilities and to reduce the impact of others.
 Capability based enforcement systems such as Capsicum~\cite{Watson2010} substantially reduce the capabilities that an exploit has to try and gain increased privilege with.

### 2014-08-14 versions.py
Add initial versions script to do investigations of the update ecosystem

@@ -1,0 +1,57 @@
+#!/usr/bin/env python3
+import avo
+import os
+import re
+import dateutil.parser
+import datetime
+import urllib.request
+
+from dahelpers import da_start_date, da_end_date, warning, save_csv_list_file
+from collections import OrderedDict
+
+def get_lines(address, matcher):
+    response = urllib.request.urlopen(address)
+    lines = []
+    for line in response:
+        line = str(line)
+        if matcher.match(line):
+            lines.append(line.strip('\n '))
+    return lines
+
+
+def get_version_dates(lines, version_matcher, date_matcher):
+    version_dates = []
+    for line in lines:
+        m_version = version_matcher.match(line)
+        m_date = date_matcher.match(line)
+        if m_version and m_date:
+            version_dates.append((m_version.group(1), dateutil.parser.parse(m_date.group(1)).date()))
+        else:
+            warning('Unmatched line: ', line, m_version, m_date)
+    return version_dates
+
+def num_in_range(version_dates):
+    num = 0
+    for _, date in version_dates:
+        if date > da_start_date and date < da_end_date:
+            num +=1
+    return num
+
+def save_to_csv(name, version_dates):
+    save_csv_list_file(map(lambda x : [x[0], x[1].isoformat()], version_dates),'version_dates_' + name + '.csv')
+
+def update_versions():
+    targz = re.compile(r'.*\.tar\.gz.*')
+    openssl_version_dates = get_version_dates(get_lines('https://www.openssl.org/source/', targz),re.compile(r'.*"openssl-([^"]+)\.tar\.gz".*'),re.compile(r'.*?\d+ ([^<]+) <.*')) 
+    print('openssl', num_in_range(openssl_version_dates))
+    save_to_csv('openssl', openssl_version_dates)
+
+    linux_lines = get_lines('https://www.kernel.org/pub/linux/kernel/v2.6/', targz)
+    linux_lines.extend(get_lines('https://www.kernel.org/pub/linux/kernel/v3.0/', targz))
+    linux_version_dates = get_version_dates(linux_lines, re.compile(r'.*"linux-([0-9.]+)\.tar\.gz".*'), re.compile('.*?\s+\s([0-9\-A-z: ]+)\s+\d+M.*'))
+    print('linux', num_in_range(linux_version_dates))
+    save_to_csv('linux', linux_version_dates)
+
+if __name__ == "__main__":
+    update_versions()
+

### 2014-08-14 versions.py
use set_latex_value in versions.py

@@ -6,6 +6,10 @@ import dateutil.parser
 import datetime
 import urllib.request
 
+import latex_value
+from latex_value import set_latex_value
+latex_value.latex_value_filename('countversions.tex')
+
 from dahelpers import da_start_date, da_end_date, warning, save_csv_list_file
 from collections import OrderedDict
 
@@ -43,13 +47,13 @@ def save_to_csv(name, version_dates):
 def update_versions():
     targz = re.compile(r'.*\.tar\.gz.*')
     openssl_version_dates = get_version_dates(get_lines('https://www.openssl.org/source/', targz),re.compile(r'.*"openssl-([^"]+)\.tar\.gz".*'),re.compile(r'.*?\d+ ([^<]+) <.*')) 
-    print('openssl', num_in_range(openssl_version_dates))
+    set_latex_value('opensslNumVersions', num_in_range(openssl_version_dates))
     save_to_csv('openssl', openssl_version_dates)
 
     linux_lines = get_lines('https://www.kernel.org/pub/linux/kernel/v2.6/', targz)
     linux_lines.extend(get_lines('https://www.kernel.org/pub/linux/kernel/v3.0/', targz))
     linux_version_dates = get_version_dates(linux_lines, re.compile(r'.*"linux-([0-9.]+)\.tar\.gz".*'), re.compile('.*?\s+\s([0-9\-A-z: ]+)\s+\d+M.*'))
-    print('linux', num_in_range(linux_version_dates))
+    set_latex_value('linuxNumVersions', num_in_range(linux_version_dates))
     save_to_csv('linux', linux_version_dates)
 
 if __name__ == "__main__":

### 2014-08-14 versions.py
Compute the mean latency of linux versions hitting android

@@ -5,13 +5,16 @@ import re
 import dateutil.parser
 import datetime
 import urllib.request
+import argparse
+import statistics
 
 import latex_value
 from latex_value import set_latex_value
 latex_value.latex_value_filename('countversions.tex')
 
-from dahelpers import da_start_date, da_end_date, warning, save_csv_list_file
+from dahelpers import da_start_date, da_end_date, warning, save_csv_list_file, load_csv_list_file
 from collections import OrderedDict
+from uncertainties import ufloat
 
 def get_lines(address, matcher):
     response = urllib.request.urlopen(address)
@@ -44,6 +47,11 @@ def num_in_range(version_dates):
 def save_to_csv(name, version_dates):
     save_csv_list_file(map(lambda x : [x[0], x[1].isoformat()], version_dates),'version_dates_' + name + '.csv')
 
+
+def load_from_csv(name):
+    return map(lambda x : [x[0], dateutil.parser.parse(x[1]).date()], load_csv_list_file('version_dates_' + name + '.csv'))
+
+
 def update_versions():
     targz = re.compile(r'.*\.tar\.gz.*')
     openssl_version_dates = get_version_dates(get_lines('https://www.openssl.org/source/', targz),re.compile(r'.*"openssl-([^"]+)\.tar\.gz".*'),re.compile(r'.*?\d+ ([^<]+) <.*')) 
@@ -55,7 +63,45 @@ def update_versions():
     linux_version_dates = get_version_dates(linux_lines, re.compile(r'.*"linux-([0-9.]+)\.tar\.gz".*'), re.compile('.*?\s+\s([0-9\-A-z: ]+)\s+\d+M.*'))
     set_latex_value('linuxNumVersions', num_in_range(linux_version_dates))
     save_to_csv('linux', linux_version_dates)
+    return {'openssl': openssl_version_dates, 'linux' : linux_version_dates}
+
+
+def load_versions():
+    names = ['openssl', 'linux']
+    version_dates = dict()
+    for name in names:
+        version_dates[name] = OrderedDict(load_from_csv(name))
+    return version_dates
+
+def latency(version_dates, version_map):
+    latencies = dict() # upstream_version -> latency in days
+    android_release_dates = OrderedDict(map(lambda x : [x[0], dateutil.parser.parse(x[1]).date()], avo.release_dates))
+    for android_version, upstream_version in version_map.items():
+        if upstream_version.endswith('.0') and upstream_version[:-2] in version_dates:
[…]

### 2014-08-15 versions.py
Add openssl latency calculation to versions

@@ -81,6 +81,8 @@ def latency(version_dates, version_map):
             upstream_release_date = version_dates[upstream_version[:-2]]
         else:
             upstream_release_date = version_dates[upstream_version]
+        if not android_version in android_release_dates:
+            continue
         android_release_date = android_release_dates[android_version]
         if android_release_date < upstream_release_date:
             warning('Data bug, android released before the upstream it contains', android_version, upstream_version, android_release_date, upstream_release_date)
@@ -94,6 +96,14 @@ def latency(version_dates, version_map):
             latencies[upstream_version] = delay
     return latencies
 
+
+def calculate_latencies(version_dates):
+    linux_latencies = latency(version_dates['linux'], OrderedDict(avo.os_to_kernel))
+    set_latex_value('linuxMeanUpdateLatency', ufloat(statistics.mean(linux_latencies.values()),statistics.stdev(linux_latencies.values())))
+    openssl_latencies = latency(version_dates['openssl'], OrderedDict(avo.os_to_openssl_version))
+    set_latex_value('opensslMeanUpdateLatency', ufloat(statistics.mean(openssl_latencies.values()),statistics.stdev(openssl_latencies.values())))
+
+
 if __name__ == "__main__":
     parser = argparse.ArgumentParser()
     parser.add_argument('--update-versions', action='store_true')
@@ -102,6 +112,5 @@ if __name__ == "__main__":
         version_dates = update_versions()
     else:
         version_dates = load_versions()
-    linux_latencies = latency(version_dates['linux'], OrderedDict(avo.os_to_kernel))
-    set_latex_value('linuxMeanUpdateLatency', ufloat(statistics.mean(linux_latencies.values()),statistics.stdev(linux_latencies.values())))
+    calculate_latencies(version_dates)
 

### 2014-08-15 versions.py
Fetch bouncycastle version date information

@@ -15,6 +15,7 @@ latex_value.latex_value_filename('countversions.tex')
 from dahelpers import da_start_date, da_end_date, warning, save_csv_list_file, load_csv_list_file
 from collections import OrderedDict
 from uncertainties import ufloat
+from lxml import html
 
 def get_lines(address, matcher):
     response = urllib.request.urlopen(address)
@@ -37,12 +38,42 @@ def get_version_dates(lines, version_matcher, date_matcher):
             warning('Unmatched line: ', line, m_version, m_date)
     return version_dates
 
+
+def fetch_bouncy_castle():
+    # Older data is available from http://freecode.com/projects/bouncycastlecryptoapi/releases should we need it, but it lacks newer data
+    urls = ['http://bouncy-castle.1462172.n4.nabble.com/Bouncy-Castle-Announce-f1468408.html', 'http://bouncy-castle.1462172.n4.nabble.com/Bouncy-Castle-Announce-f1468408i35.html']
+    bouncy_version_matcher = re.compile('.*((Bouncy Castle Crypto Provider Package version)|(BC Release)|(Java Release)) ([0-9.]+) (is )?now available.*')
+    bouncy_date_matcher = re.compile(r'.*new Date\((\d+)\).*', flags=re.DOTALL)
+    version_dates = []
+    for url in urls:
+        response = urllib.request.urlopen(url)
+        tree = html.fromstring(response.read())
+        table = tree.xpath('/html/body/div/table[1]')[0]
+        rows = table.getchildren()
+        for row in rows:
+            if not 'main' in row.get('class'):
+                continue
+            cells = row.getchildren()
+            subject = cells[1].xpath('table/tr/td/a/text()')[0]
+            match = bouncy_version_matcher.match(subject)
+            if not match:
+                continue
+            version = match.group(5)
+            date_script = cells[3].xpath('table/tr/td/a/script/text()')[0]
+            match = bouncy_date_matcher.match(date_script)
+            if not match:
+                continue
+            date = datetime.datetime.fromtimestamp(int(match.group(1)[:-3])).date()
+            version_dates.append((version, date))
+    return version_dates
+
 def num_in_range(version_dates):
     num = 0
     for _, date in version_dates:
         if date > da_start_date and date < da_end_date:
             num +=1
     return num
+
 
[…]

### 2014-08-15 versions.py
Compute bouncycastle latencies

@@ -103,7 +103,7 @@ def update_versions():
 
 
 def load_versions():
-    names = ['openssl', 'linux']
+    names = ['openssl', 'linux', 'bouncycastle']
     version_dates = dict()
     for name in names:
         version_dates[name] = OrderedDict(load_from_csv(name))
@@ -138,6 +138,8 @@ def calculate_latencies(version_dates):
     set_latex_value('linuxMeanUpdateLatency', ufloat(statistics.mean(linux_latencies.values()),statistics.stdev(linux_latencies.values())))
     openssl_latencies = latency(version_dates['openssl'], OrderedDict(avo.os_to_openssl_version))
     set_latex_value('opensslMeanUpdateLatency', ufloat(statistics.mean(openssl_latencies.values()),statistics.stdev(openssl_latencies.values())))
+    bouncycastle_latencies = latency(version_dates['bouncycastle'], OrderedDict(avo.os_to_bouncycastle_version))
+    set_latex_value('bouncycastleMeanUpdateLatency',ufloat(statistics.mean(bouncycastle_latencies.values()),statistics.stdev(bouncycastle_latencies.values())))
 
 
 if __name__ == "__main__":

### 2014-08-15 securityupdates.tex
Vulnerability table comes from avo rather than da now
Compute bouncycastle latencies
Use a table to display version latency information

@@
 % Num versions since \daStartDate
-\newcommand{\bouncyCastleNumVersions}{A}
 \input{countversions}
-%\newcommand{\opensslNumVersions}{B}
 \newcommand{\apacheLibsNumVersions}{C}
 \newcommand{\apacheLibsNumLibs}{E}
-%\newcommand{\linuxNumVersions}{D}
 
@@
 
-\daTabAndVulns
+\avoTabAndVulns
 
@@
 
-The number of updates shipped by upstream open source projects is calculated by scraping the releases pages for those projects.
+The number of updates shipped by upstream open source projects is calculated by scraping the releases pages for those projects, from which we can also obtain the release dates.
+By comparing the tags in the android git version repositories for upstream projects with Android version numbers and examining the version of the upstream project under that tag we computed which upstream project versions were included in each version of Android.
+This allows us to compute the latency between an upstream project being released and it being included in Android, this is shown in Table~\ref{tab:update_ecosystem}.
+\begin{table}
+\centering
+\begin{tabular}{l|r|r}
+Project	&	\# releases	&	latency \\ \hline
+linux	&	\linuxNumVersions	&	\linuxMeanUpdateLatency \\
+openssl	&	\opensslNumVersions	&	\opensslMeanUpdateLatency \\
+bouncycastle	&	\bouncycastleNumVersions	&	\bouncycastleMeanUpdateLatency \\
+\end{tabular}
+\caption{Flow of updates from upstream projects into Android. Number of updates as in Figure~\ref{fig:update_ecosystem}, latency in days for all pairs of versions we have data on.}
+\label{tab:update_ecosystem}
+\end{table}
 

### 2014-08-27 figures/update_figures.sh
Add pdf of figures for best and worst device and vulnerabilities timeline

@@ -29,3 +29,4 @@ cp "$fromdir"ua-versions-android-month-norm.pdf ua-versions-android-month-norm.p
 cp "$fromdir"ua-versions-ios-month-norm.pdf ua-versions-ios-month-norm.pdf
 cp "$fromdir""model/Galaxy Nexus/norm_full.pdf" "full_version_Galaxy Nexus.pdf"
 cp "$fromdir""model/HTC Desire HD A9191/norm_full.pdf" "full_version_HTC Desire HD A9191.pdf"
+cp "$fromdir"vulnerabilities_timeline.pdf vulnerabilities_timeline.pdf

### 2014-08-27 securityupdates.tex
Fix 'exploit' to 'vulnerability' in various places

@@
 
-\subsection{Root exploits}
-We compiled a list of root exploit vulnerabilities for Android containing information on when they were known about, what versions they affected and which versions fixed the problem.
-We only looked for \emph{root equivalent} exploits which did not require USB debugging to exploit.
+\subsection{Root vulnerabilities}
+We compiled a list of root vulnerabilities for Android containing information on when they were known about, what versions they affected and which versions fixed the problem.
+We only looked for \emph{root equivalent} vulnerabilities which did not require USB debugging to exploit.
 \emph{Root equivalent} means that it grants privileges equivalent in scope to root which can then either be used to gain root or provide sufficiently many permissions as to be as bad as root.
 Some phones can be `rooted' by enabling USB debugging and using the special privileges of the adb shell to root the device but only \daAdbEnabledPerc\ of devices have USB debugging enabled.
-This is not something that applications running on the phone can exploit to break compartmentalisation and so we do not include those exploits.
+This is not something that applications running on the phone can exploit to break compartmentalisation and so we do not include those vulnerabilities.
 Unfortunately many published exploits use adb for convenience and so determining whether that is necessary to exploit the vulnerability can be difficult.
-Some exploits which grant root are not traditional kernel exploits, for example the discovery of flaws in the verification of signatures on Android applications in February 2013~\cite{Forristal2013} meant that applications could pretend to be signed with system keys and hence gain root equivalent privileges.
+Some vulnerabilities which grant root are not traditional kernel vulnerabilities, for example the discovery of flaws in the verification of signatures on Android applications in February 2013~\cite{Forristal2013} meant that applications could pretend to be signed with system keys and hence gain root equivalent privileges.
 On some versions of Android (below 4.1) they could use known system to root escalation mechanisms but on all versions they have a greatly increased attack area for further privilege escalation and also have the ability to control all user internet traffic (via VPNs), brick the phone, remove and install apps, steal user credentials, read the screen, make and receive calls.
@@
 Prior work~\cite{Bilge2012} by Symantec on desktop malware has shown that after public disclosure exploitation rates increase by 5 orders of magnitude and so from the point of view of widespread danger, the period between public disclosure and the update reaching the user's device is the most critical.
-However they also show that zero day exploits are typically used for 312 days before they are publicly disclosed, often to target particular organisations.
+However they also show that zero day vulnerabilities are typically used for 312 days before they are publicly disclosed, often to target particular organisations.
 Hence when considering vulnerability from the point of view of an organisation which cares about Advanced Persistent Threats, such as those the CESG advice~\cite{CESG2013} is aimed at, the critical period is from when people could have started using the vulnerability to when the update hits the user device.
@@
 Figure~\ref{fig:nvulnerabilities_heat} shows how the the \emph{APK signing vulnerabilities} affected all devices and took months to get fixed for any device.
-However what is perhaps more worrying is the long tail on the \emph{Gingerbreak}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/Gingerbreak}}, \emph{levitator} and \emph{exploid} vulnerabilities which are more dangerous root exploits (not requiring new APK installation) and which still affect a significant proportion of devices years later.
+However what is perhaps more worrying is the long tail on the \emph{Gingerbreak}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/Gingerbreak}}, \emph{levitator} and \emph{exploid} vulnerabilities which are more dangerous root vulnerabilities (not requiring new APK installation) and which still affect a significant proportion of devices years later.
 

### 2014-08-27 securityupdates.tex
Talk about the vulnerabilities timeline

@@
 We only use the \daNumVulnsUsed\ vulnerabilities which it is convenient to use for this analysis because they affect all Android devices rather than particular versions and we know enough information about the vulnerability to tie it to the particular versions it affects.
+
+\begin{figure}
+ \centering
+ \includegraphics[width=\columnwidth]{figures/vulnerabilities_timeline}
+ \caption{Timeline of vulnerabilities}
+ \label{fig:vulnerabilities_timeline}
+\end{figure}
+
+The dates of discovery of vulnerabilities in the \texttt{androidvulnerabilities.org} is not uniform.
+Figure~\ref{fig:vulnerabilities_timeline} shows the dates of discovery and, when later, the dates when a version of Android which fixed the vulnerability was shipped.
+Some vulnerabilities (RageAgainstTheCage adb, levitator, zergRush abd TwerkMyMoto) were fixed in released versions of Android before they were discovered and so are shown as vertical lines while others were known for months before a version of Android which fixed them was shipped.
+During the period in which Device Analyzer data was collected the date when a version of Android with the fix was observed on a Device Analyzer device is taken as the date that version was released, before that the release date as best as we can determine is used.
+There is no canonical source of Android release dates, our best guesses and supporting references are available from \texttt{androidvulnerabilities.org}.
 

### 2014-08-27 securityupdates.tex
Talk about the gap in vulnerability discovery

@@
 There is no canonical source of Android release dates, our best guesses and supporting references are available from \texttt{androidvulnerabilities.org}.
+
+This data shows a large gap in the middle where we do not know of any discoveries of root equivalent vulnerabilities affecting all of Android.
+The cause of this quiet period is unclear, perhaps from Figure~\ref{fig:proportioninsecure} the reason was that since most devices were exposed to known vulnerabilities there was no point in looking for new ones?
+Or that manufacturers made it easier to install custom versions of Android, reducing the need for users to root their devices?
+Or that manufacturer specific vulnerabilities proved to be easier to find and so people looking for vulnerabilities adjusted their focus?
 

### 2014-08-27 securityupdates.tex
Fix jelly bean citation

@@
 There are continuing efforts to reduce the impact of root vulnerabilities, both in Android and more widely.
-SEAndroid~\cite{Smalley2013} which is included in Android from version 4.1~\cite{https://developer.android.com/about/versions/jelly-bean.html} claimed to prevent some root vulnerabilities and to reduce the impact of others.
+SEAndroid~\cite{Smalley2013} which is included in Android from version 4.1~\cite{jelly-bean-release} claimed to prevent some root vulnerabilities and to reduce the impact of others.
 Capability based enforcement systems such as Capsicum~\cite{Watson2010} substantially reduce the capabilities that an exploit has to try and gain increased privilege with.

### 2014-08-27 figures/update_figures.sh

@@ -30,3 +30,4 @@ cp "$fromdir"ua-versions-ios-month-norm.pdf ua-versions-ios-month-norm.pdf
 cp "$fromdir""model/Galaxy Nexus/norm_full.pdf" "full_version_Galaxy Nexus.pdf"
 cp "$fromdir""model/HTC Desire HD A9191/norm_full.pdf" "full_version_HTC Desire HD A9191.pdf"
 cp "$fromdir"vulnerabilities_timeline.pdf vulnerabilities_timeline.pdf
+cp ~/Library/papers/library.filtered.bib ../securityupdates.bib

### 2014-09-03 securityupdates.tex
Add text about modelling future version distributions

@@
 \subsection{Predicting future version distributions}
+Examining the way that the OS version distribution changes over time in Figure~\ref{fig:norm_os} shows some regularity to the behaviour.
+Figure~\ref{fig:frvh_os_versions} is a plot of the proportion of of devices in Device Analyzer running each Android version or a later Android version (i.e.
+for 2.1.0 it includes all devices as they all run later versions of Android).
+The data are all normalised to the number of days since the version was released and is smoothed by forcing it to be monotonically increasing.\footnote{While sometimes an older version becomes more popular after a later version has been released causing the values to actually decrease, mostly this was just noise or changes in the distribution of devices (e.g.\ the sudden arrival of 1000 devices from Bangladesh due to a study being conducted there) and this assumption makes the graph clearer.}
+\begin{figure}
+ \includegraphics[width=\columnwidth]{figures/frvh_os_versions}
+ \caption{Proportion of devices running particular versions of Android or any later version}
+ \label{fig:frvh_os_versions}
+\end{figure}
+\begin{figure}
+ \includegraphics[width=\columnwidth]{figures/frms_os_versions}
+ \caption{Proportion of devices running particular versions of Android or any later version}
+ \label{fig:frms_os_versions}
+\end{figure}
+There is a general pattern to the data, Figure~\ref{fig:frms_os_versions} shows the mean and standard deviation of the same data.
+It also shows the curve of best fit, which has a root mean squared error (RMSE) of \daOSCurveFitRMSE, the curve we chose is not a standard polynomial fit (3 degree polynomial fit gave a RMSE of \daOSCurvePolyRMSE) or a spline fit (RMSE of \daOSCurveSplineRMSE) as we wanted a better model of behaviour than would be given by simply fitting a curve.
+
+The data appears to have an inverse exponential decay as it tends to 1 but with a slow start.
+This gives us a $1 - e^{-x}$ component for the main behaviour and a $\frac{1}{x}$ component to provide a slow start.
+The general form is then:
+\begin{equation}
+y = (s - \frac{1}{e^{dx}} + \frac{1}{cx + a})l
+\end{equation}
+However there are two constraints we can apply, $\underset{x\rightarrow\infty}{\mathrm{lim}} y\rightarrow 1$ and the curve passes through the origin.
+This gives us $l = \frac{1}{s}$ and $a = \frac{1}{1-s}$.
+\begin{equation}
+y = (s - \frac{1}{e^{dx}} + \frac{1}{cx + \frac{1}{1-s}})\frac{1}{s}
+\end{equation}
+We get the parameters ????? for OS versions and ???? for API versions.
+From this the number of days from release of a new version of Android until 50\% of devices are running that version or higher is \daOSCurveHalfDeployed\ and full deployment to \daFullDeployedAt\ of devices takes \daOSCurveFullDeployed\ days.
+For API versions this is \daAPICurveHalfDeployed\ and \daAPICurveFullDeployed\ respectively.
+
+Hence if a security vulnerability is fixed through the release of a particular Android version it will be \daOSCurveFullDeployed\ days after until the fix is fully deployed.
+
 TODO

### 2014-09-03 figures/update_figures.sh

@@ -30,4 +30,7 @@ cp "$fromdir"ua-versions-ios-month-norm.pdf ua-versions-ios-month-norm.pdf
 cp "$fromdir""model/Galaxy Nexus/norm_full.pdf" "full_version_Galaxy Nexus.pdf"
 cp "$fromdir""model/HTC Desire HD A9191/norm_full.pdf" "full_version_HTC Desire HD A9191.pdf"
 cp "$fromdir"vulnerabilities_timeline.pdf vulnerabilities_timeline.pdf
+cp "$fromdir"frvh_os_versions.pdf frvh_os_versions.pdf
+cp "$fromdir"frvh_api_versions.pdf frvh_api_versions.pdf
+cp "$fromdir"frms_os_versions.pdf frms_os_versions.pdf
 cp ~/Library/papers/library.filtered.bib ../securityupdates.bib

### 2014-10-14 securityupdates.tex
fix typo

@@
 Sometimes Manufactures ship direct to the user, sometimes the Manufacturer and Google are the same entity (at least for the purpose of a particular phone, such as with Nexus phones).
-Sometimes Manufacturers incorporate upstream open source project releases directly, and sometimes incorrectly -- for example including a broken daily build of sqllite in their release of Android
+Sometimes Manufacturers incorporate upstream open source project releases directly, and sometimes incorrectly -- for example including a broken daily build of sqlite in their release of Android
 %An analysis by Vidas et al.~\cite{Vidas2011} of the Android 2.1 to 2.2 update found that it took 11 months from when Google released 2.2 for the last device which they were investigating to get the update.

### 2014-10-14 securityupdates.tex
Rework predicting future version distributions section

--- 2014-08-27 securityupdates.tex
@@
 \usepackage{tikz}
+\usepackage{siunitx}
 \RequirePackage[date=terse, isbn=true, doi=true, url=false, urldate=iso8601, maxbibnames=9, backref=false, backend=bibtex]{biblatex}
@@
 \subsection{Predicting future version distributions}
+Examining the way that the OS version distribution changes over time in Figure~\ref{fig:norm_os} we see some regularity to the behaviour.
+To show this behaviour more clearly we normalised the data to the number of days since the version was released and is smoothed by forcing it to be monotonically decreasing.\footnote{While sometimes an older version becomes more popular after a later version has been released causing the values to actually increase, mostly this was just noise or changes in the distribution of devices (e.g.\ the sudden arrival of 1000 devices from Bangladesh due to a study being conducted there) and this assumption makes the graph clearer.}
+Then we can plot Figure~\ref{fig:frvh_os_versions} which shows the proportion of of devices in Device Analyzer not upgraded to each Android version or a later Android version (i.e.
+for 2.1.0 it includes 2.1.1, 2.2.0 and indeed all devices as they all run later versions of Android).
+
+\begin{figure}
+ \includegraphics[width=\columnwidth]{figures/frvh_os_versions}
+ \caption{Proportion of devices not updated to particular versions of Android or any later version}
+ \label{fig:frvh_os_versions}
+\end{figure}
+\begin{figure}
+ \includegraphics[width=\columnwidth]{figures/frms_os_versions}
+ \caption{Proportion of devices not updated to particular versions of Android or any later version}
+ \label{fig:frms_os_versions}
+\end{figure}
+There is a general pattern to these data, Figure~\ref{fig:frms_os_versions} shows the mean and standard deviation of the same data.
+It also shows the curve of best fit, which has a root mean squared error (RMSE) of \daOSCurveFitRMSE, the curve we chose is not a standard polynomial fit (3 degree polynomial fit gave a RMSE of \daOSCurvePolyRMSE) or a spline fit (RMSE of \daOSCurveSplineRMSE) as we wanted a meaningful model of behaviour rather than a generic curve.
+
+The data appears to have an exponential decay as it tends to 1 but with a slow start as few devices are upgraded initially.
+We approximate this with a a $e^{-x}$ component for the main behaviour and a which is offset to start later when uptake takes off.
+The equation is then:
+\begin{equation}
+y =
+ \begin{cases}
+  1.0 & \text{if }x<\mathrm{undeployed}\\
+  e^{-\mathrm{decay}(x-\mathrm{undeployed})} & \text{otherwise}
+ \end{cases}
+\end{equation}
+
+Fitting that curve to the data using \texttt{scipy.optimize}, we get the parameters $\mathrm{undeployed} = \daOSCurveFitParamFirst ~\si{days}$, $\mathrm{decay} = \daOSCurveFitParamSecond ~\si{days^{-1}}$ for OS versions and $\mathrm{undeployed} = \daAPICurveFitParamFirst ~\si{days}$, $\mathrm{decay} = \daAPICurveFitParamSecond ~\si{days^{-1}}$ for API versions.
+From this the number of days from release of a new version of Android until 50\% of devices are running that version or higher is \daOSCurveHalfDeployed\ and full deployment to \daFullDeployedAt\ of devices takes \daOSCurveFullDeployed\ days.
+For API versions this is \daAPICurveHalfDeployed\ and \daAPICurveFullDeployed\ respectively.
+
+Hence if a security vulnerability is fixed through the release of a particular Android version it will be \daOSCurveFullDeployed\ days after until the fix is fully deployed.
+
 TODO

### 2014-10-14 securityupdates.tex
Make TODOs visible

@@
 \usepackage{siunitx}
+\usepackage{todonotes}
+\let\OldTodo\todo
+\renewcommand{\todo}{\OldTodo[inline]}
 \RequirePackage[date=terse, isbn=true, doi=true, url=false, urldate=iso8601, maxbibnames=9, backref=false, backend=bibtex]{biblatex}
@@
 The Device Analyzer data contains data from \daNumDevices\ devices with a total of \daDeviceDays\footnote{Here we are only counting devices and days for which we have valid OS version data} device days, it contains data for more than \daMonths\ months for \daMonthsDevices\ devices.
-%TODO talk about the Device Analyzer project a bit
+\todo{ talk about the Device Analyzer project a bit}
 Various different kinds of data\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/collected.htm}} are collected by the Device Analyzer Android app\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/}}.
@@
 While we have not tried to exploit these vulnerabilities on the devices in order to test whether the devices are actually vulnerable, we are confident that most devices will be if they are running a vesion known to be vulnerable.
-%TODO WHY?
+\todo{WHY?}
 
@@
 It also incorporates various drivers for different bits of hardware.
-The Android platform is then built on top of those with kernel and userspace components by Google. %TODO work out the exact arrangements.
+The Android platform is then built on top of those with kernel and userspace components by Google. \todo{ work out the exact arrangements.}
 Usually the code for each update is kept secret until after the update has been published though through the Open Handset Alliance some manufacturers see it in advance.
@@
 %An analysis by Vidas et al.~\cite{Vidas2011} of the Android 2.1 to 2.2 update found that it took 11 months from when Google released 2.2 for the last device which they were investigating to get the update.
-%TODO talk about update latency
+\todo{talk about update latency}
 
@@
 The number of updates shipped by Google (\daNumSigOSVersions) is the number of Android versions reported in Device Analyzer which affected more than \daSigVersionPerc\ of devices for more than \daSigVersionDays\ days.
-%TODO this is the wrong number to be using
+\todo{this is the wrong number to be using}
 
@@
 In this section we present the results of our analysis showing \daMeanInsecurityPerc\ of devices to have root vulnerabilities on average, nexus devices to be better than others, \emph{\daSecScoreBestmanufacturer} to be the best manufacturer and \emph{\daSecScoreBestoperator} to be the best network operator.
-%TODO summarise our results
+\todo{ summarise our results}
 The Device Analyzer data used here was collected between \daStartDate\ and \daEndDate.
 The \texttt{androidvulnerabilities.org} data was collected between \avoStartDate\ and \avoEndDate.
-%TODO comparison of nexus update rates with iOS
+\todo{comparison of nexus update rates with iOS}
 
@@
 However when in February 2013 the first APK signing vulnerability was found which affected all previous versions of Android and even in October 2013 most devices were still vulnerable.
-%TODO put percentages on these things to quantify them
+\todo{put percentages on these things to quantify them}
 \begin{figure}%[!b]
@@
 \subsection{Update frequency}
-%TODO find out what the frequency of updates is
+\todo{ find out what the frequency of updates is}
 \begin{figure}
@@
 Red indicates updates which changed the build number but not the version number and so might contain a backported fix for a vulnerability.
-%TODO summary stats, what does this mean
+\todo{ summary stats, what does this mean}
 
@@
 \subsection{Comparison with iOS}
-TODO
+\todo{Get data from Richard so that we can do this comparison}
 
@@
 
-TODO
+\todo{Our handling of errors here is not quite good enough as the uncertainty from our y value interacts badly with the logarithm function when computing the inverse and we compute the log of ~0 which means our error ends up much bigger than our value - but that is an upper bound, not a +- bound.}
 
 \subsection{Lifetime of a vulnerability}
-TODO
+The key events in the lifetime of a vulnerability are:
+
+\begin{LaTeXdescription}
+ \item[introduction] When a vulnerability was introduced into the source code.
+ \item[introducing release] When the first release was made containing the vulnerability.
+ \item[fix] When a vulnerability was first fixed in the source code.
+ \item[fixing release] When the first release containing the fix was made.
+ \item[fix deployed] When the fix has been deployed to a sufficiently large proportion of the population that the vulnerability can be ignored.
+ \item[discovery] When the vulnerability is first discovered
+ \item[exploit] When the vulnerability is first exploited
+ \item[disclosed] When the vulnerability is first disclosed % publically or to a smaller set of people?
+\end{LaTeXdescription}
+
+\todo{Compare this list with the literature and unify it. Pull out figures which illustrate and investigate this in our data.}
 
 \subsection{Monetisation}
-%TODO redo this section in light of related work
+\todo{redo this section in light of related work}
+
 One of the key differences between smartphones and desktops is that it is easier to directly cash out on a smartphone as malware can send premium rate texts or make premium rate phone calls.
@@
 This means that the OS version information from Device Analyzer is likely to be overestimating the prevalence of new OS versions and hence our results are a conservative estimate of the security of Android.
-%TODO we want a statistical metric to claim this strongly with.
+\todo{we want a statistical metric to claim this strongly with.}
 This allows us to have confidence in the OS version information.
@@
 The security protections used in iOS such as only allowing signed code to run, jails and reviewing of apps has resulted in a lower level of malware affecting iOS~\cite{TODO} but can still be bypassed~\cite{Wang2013a}.
-%TODO look at iOS protections again and tighten this up
+\todo{look at iOS protections again and tighten this up}
 
@@
 Fortunately Android does authenticate the update binaries (though the APK vulnerabilities circumvented this) and Google Play downloads them over a secure connection~\cite{Viennot2014}.
-%TODO double check that
+\todo{double check that}
 
@@
 \label{sec:conclusion}
-TODO
+\todo{Conclude}
 

### 2014-10-16 securityupdates.tex
Updates from Alastair's comments

@@
 
-\author{Daniel R.\ Thomas, Daniel T.\ Wagner, Alastair R.\ Beresford and Andrew Rice}
+\author{Daniel R.\ Thomas, Alastair R.\ Beresford, Daniel T.\ Wagner and Andrew Rice}
 
@@
 
+% Terminology
+
+% Android, Unix
+
 \maketitle
@@
 However there is ongoing legal action to force operators to ship updates for security vulnerabilities~\cite{Soghoian2013}.
-Many phones are sold on long (18 month) contracts with monthly payments yet stop receiving updates before the end of the contract, while Windows XP could be purchased for a one off payment in October 2001 and received security updates until April 2014.
-CESG, which advises the UK government on how to secure its computer systems, recommends picking Android devices from manufactures which are good at shipping security updates promptly~\cite{CESG2013} but it does not state which manufacturers are better.
+Many phones are sold on long (12--24 month) contracts with monthly payments yet stop receiving updates before the end of the contract, while Windows XP could be purchased for a one off payment in October 2001 and received security updates until April 2014.
+Corporate or public sector buyers are encouraged to purchase secure devices, but we have found little concrete guidance.
+For example, CESG, which advises the UK government on how to secure its computer systems, recommends picking Android devices from manufactures which are good at shipping security updates promptly~\cite{CESG2013} but it does not state which manufacturers are better.
 We provide this information in~\S\ref{sec:security_scoring}.
 
-Android relies on the Linux kernel to provide compartmentalisation and runs each app as a separate unix `user'.
+Android relies on the Linux kernel to provide compartmentalisation and runs each app as a separate Unix `user'.
 The Linux kernel is a large piece of software and so there will always be new root exploits found in it~\cite{TODO}.
@@
 %TODO make this graph shorter
-\caption{Proportion of devices running insecure versions of android}
+\caption{Proportion of devices running insecure versions of Android against time}
 \label{fig:proportioninsecure}
@@
 
-In the past users of Microsoft IIS were advised to switch to a different web server due to the high rate of vulnerability~\cite{Pescatore2001} this pressure helped cause Microsoft to focus on security so that it now does much better~\cite{TODO}.
+In the past users of Microsoft IIS were advised to switch to a different web server due to the frequent discovery of vulnerabilities~\cite{Pescatore2001}.
+This pressure encouraged Microsoft to improve the security of their products~\cite{TODO}.
 Part of the reason why deployed devices do not receive updates in a timely manner is that there is little data on what is actually happening and so it is not possible to determine which manufacturers or operators are doing a better job.
-Hence there is little incentive for manufacturers and operators to do a better job, we provide this data  in~\S\ref{sec:security_scoring} and discuss incentives further in \S\ref{sec:economics}
+Hence there is little incentive for manufacturers and operators to do a better job.
+We provide this data  in~\S\ref{sec:security_scoring} and discuss incentives further in \S\ref{sec:economics}.
 
@@
 \begin{itemize}
- \item We quantify the vulnerability of Android to root equivalent vulnerabilities.
- \item We characterise the Android update ecosystem.
- \item We collate information on vulnerabilities which affect Android.
+ \item We characterise the Android update ecosystem and quantify its vulnerability to root equivalent vulnerabilities.
+ \item We collate information on vulnerabilities which affect Android and investigate their time-lines.
  \item We measure the security of Android according to three metrics and compare different manufacturers. and device models to allow consumers to differentiate between them based on security.
- \item We investigate the time-lines of vulnerabilities in Android.
  \item We provide models of how the distribution of Android versions changes allowing future version distributions to be predicted.
- \item We propose methods which could be used to improve the situation in future.
+% \item We propose methods which could be used to improve the situation in future.
 \end{itemize}
@@
 
-The database of vulnerabilities which we maintain at \texttt{androidvulnerabilities.org} contains \avoNumVulnerabilities\ vulnerabilities of which \avoNumVulnAllAndroid\ affect all android devices and \avoNumVulnSpecific\ are specific to particular devices or manufacturers.
+The database of vulnerabilities which we maintain at \texttt{androidvulnerabilities.org} contains \avoNumVulnerabilities\ vulnerabilities of which \avoNumVulnAllAndroid\ affect all Android devices and \avoNumVulnSpecific\ are specific to particular devices or manufacturers.
 We only use the \daNumVulnsUsed\ vulnerabilities which it is convenient to use for this analysis because they affect all Android devices rather than particular versions and we know enough information about the vulnerability to tie it to the particular versions it affects.
@@
 The number of updates shipped by upstream open source projects is calculated by scraping the releases pages for those projects, from which we can also obtain the release dates.
-By comparing the tags in the android git version repositories for upstream projects with Android version numbers and examining the version of the upstream project under that tag we computed which upstream project versions were included in each version of Android.
+By comparing the tags in the Android git version repositories for upstream projects with Android version numbers and examining the version of the upstream project under that tag we computed which upstream project versions were included in each version of Android.
 This allows us to compute the latency between an upstream project being released and it being included in Android, this is shown in Table~\ref{tab:update_ecosystem}.

### 2014-10-16 security_scoring.tex
Further refactoring towards a more logical flow

@@
-\subsection{Scoring for security}
-\label{sec:security_scoring}
+\subsection{Experiment 3: Comparing manufacturers, operators and devices}
+\label{sec:security_scoring}\label{sec:exp:security_score}
+
+To allow buyers of Android devices to purchase those devices which have the best security they need to know how different manufactuers, devices and network operators compare in terms of the security they provide.
+We propose a method to score a manufacturer, operator or device based on its historic performance at keeping devices up to date and fixing security vulnerabilities.
+We find that Android as a whole gits a score of \daSecurityScore / 10, the best manufacturer is \daSecScoreBestmanufacturer\ (\daSecScoreBestmanufacturerScore\ / 10) and the worst is \daSecScoreWorstmanufacturer\ (\daSecScoreWorstmanufacturerScore\ / 10).
+
+\subsubsection{Method: Scoring for security}
 
@@
 
+\subsubsection{Results: Security scores}
 The scores across the whole of android are that \daMeanInsecurityPerc\ of devices are exposed to known root exploits.
@@
 
-%TODO there must be some way of putting an error bound on our results by considering the error on the whole DA set and the size of the subset used for each manufacturer and model
+\todo{There must be some way of putting an error bound on our results by considering the error on the whole DA set and the size of the subset used for each manufacturer and model}

### 2014-10-16 securityupdates.tex
Further refactoring towards a more logical flow

@@
 \input{dastats}
+\newcommand{\da}{Device Analyzer}
 \input{avostats}
+\newcommand{\avo}{\texttt{androidvulnerabilities.org}}
 \newcommand{\percMarketShare}{XX\%~\cite{TODO}}
@@
 Our hypothesis is that these fixes do not promptly reach the users devices and so much of the time users are running versions of Android known to be vulnerable.
-We used Device Analyzer data from over \daOSYearsOfData\ years and \daNumOSDevices\ devices and found that on average over \daMeanInsecurityPerc\ of devices were exposed to known root privilege vulnerabilities and only \daUpdatednessPerc\ of devices run the most recent version of Android.
+We used da data from over \daOSYearsOfData\ years and \daNumOSDevices\ devices and found that on average over \daMeanInsecurityPerc\ of devices were exposed to known root privilege vulnerabilities and only \daUpdatednessPerc\ of devices run the most recent version of Android.
 There was also a period of several months when no devices ran secure versions of Android.
@@
 \section{Introduction}
-Android has \percMarketShare\ of the smartphone market and while the core development is controlled by Google there are about \daNumManufacturers\footnote{\label{foot:dadata}We computed this from the Device Analyzer~\cite{Wagner2013} data, see \S\ref{sec:android_update_process}.} manufacturers which make devices which run Android.
+Android has \percMarketShare\ of the smartphone market and while the core development is controlled by Google there are about \daNumManufacturers\footnote{\label{foot:dadata}We computed this from the da~\cite{Wagner2013} data, see \S\ref{sec:android_update_process}.} manufacturers which make devices which run Android.
 Android receives regular updates  which add new features and fix bugs, including security vulnerabilities.
@@
 
-The proportion of Android devices in the Device Analyzer data which were running versions of Android known to have root vulnerabilities at the time is shown in Figure~\ref{fig:proportioninsecure}.
+The proportion of Android devices in the da data which were running versions of Android known to have root vulnerabilities at the time is shown in Figure~\ref{fig:proportioninsecure}.
 In \S\ref{sec:background} we will explain where these data come from and then in \S\ref{sec:results} we will show these data in more detail.
@@
 \label{sec:background}
-There are two sources of data we needed for this study, information on root vulnerabilities and information on installed versions of Android.
+There are two sources of data we needed to have and to link for this study, information on root vulnerabilities and information on installed versions of Android.
 We need data on root vulnerabilities so that we know which versions of Android were vulnerable to which known vulnerabilities at what times.
@@
 
-The database of vulnerabilities which we maintain at \texttt{androidvulnerabilities.org} contains \avoNumVulnerabilities\ vulnerabilities of which \avoNumVulnAllAndroid\ affect all Android devices and \avoNumVulnSpecific\ are specific to particular devices or manufacturers.
+The database of vulnerabilities which we maintain at \avo\ contains \avoNumVulnerabilities\ vulnerabilities of which \avoNumVulnAllAndroid\ affect all Android devices and \avoNumVulnSpecific\ are specific to particular devices or manufacturers.
 We only use the \daNumVulnsUsed\ vulnerabilities which it is convenient to use for this analysis because they affect all Android devices rather than particular versions and we know enough information about the vulnerability to tie it to the particular versions it affects.
@@
 
-The dates of discovery of vulnerabilities in the \texttt{androidvulnerabilities.org} is not uniform.
+The dates of discovery of vulnerabilities in the \avo\ is not uniform.
 Figure~\ref{fig:vulnerabilities_timeline} shows the dates of discovery and, when later, the dates when a version of Android which fixed the vulnerability was shipped.
 Some vulnerabilities (RageAgainstTheCage adb, levitator, zergRush abd TwerkMyMoto) were fixed in released versions of Android before they were discovered and so are shown as vertical lines while others were known for months before a version of Android which fixed them was shipped.
-During the period in which Device Analyzer data was collected the date when a version of Android with the fix was observed on a Device Analyzer device is taken as the date that version was released, before that the release date as best as we can determine is used.
-There is no canonical source of Android release dates, our best guesses and supporting references are available from \texttt{androidvulnerabilities.org}.
+During the period in which \da\ data was collected the date when a version of Android with the fix was observed on a \da\ device is taken as the date that version was released, before that the release date as best as we can determine is used.
+There is no canonical source of Android release dates, our best guesses and supporting references are available from \avo.
 
@@
 We also need historical information on which versions of Android were running on devices along with other information about those devices which allows us to go into more detail and confirm the reliability of the data our results depend on.
-The only suitable data which is available for this is the Device Analyzer data~\cite{Wagner2013}.
-The Device Analyzer data contains data from \daNumDevices\ devices with a total of \daDeviceDays\footnote{Here we are only counting devices and days for which we have valid OS version data} device days, it contains data for more than \daMonths\ months for \daMonthsDevices\ devices.
-\todo{ talk about the Device Analyzer project a bit}
-Various different kinds of data\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/collected.htm}} are collected by the Device Analyzer Android app\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/}}.
+The only suitable data which is available for this is the \da\ data~\cite{Wagner2013}.
+The \da\ data contains data from \daNumDevices\ devices with a total of \daDeviceDays\footnote{Here we are only counting devices and days for which we have valid OS version data} device days, it contains data for more than \daMonths\ months for \daMonthsDevices\ devices.
+\todo{ talk about the \da\ project a bit}
+Various different kinds of data\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/collected.htm}} are collected by the \da\ Android app\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/}}.
 Among these are the build string and API version for the version of Android currently running on the device each day.
@@
 
-The numbers of Devices (\daNumOSDevices), Operators (\daNumOperators) and Manufacturers (\daNumManufacturers) in Figure~\ref{fig:update_ecosystem} come from the Device Analyzer data.
-Manufacturer and Operator counts were obtained by normalising the results reported by Android to Device Analyzer of the device manufacturer and active network operator.
+The numbers of Devices (\daNumOSDevices), Operators (\daNumOperators) and Manufacturers (\daNumManufacturers) in Figure~\ref{fig:update_ecosystem} come from the \da\ data.
+Manufacturer and Operator counts were obtained by normalising the results reported by Android to \da\ of the device manufacturer and active network operator.
 This normalisation involves removing invalid values (such as `manufacturer' or `airplane mode is on'), collating across company name changes (e.g.\ `lge' to `lg'), normalising punctuation, removing extra strings sometimes added such as (`(2g)' or `communications') and mapping some incorrectly placed model names back to their manufacturer.
-This normalisation is not perfect and so these are overestimates on the Device Analyzer data but they are likely still underestimates as there will be some manufacturers and operators which are not included in the Device Analyzer data.
+This normalisation is not perfect and so these are overestimates on the \da\ data but they are likely still underestimates as there will be some manufacturers and operators which are not included in the \da\ data.
 
-The number of updates received by devices (\daNumFullVersions) is the number of different full version strings observed in Device Analyzer.
-The number of updates shipped by Google (\daNumSigOSVersions) is the number of Android versions reported in Device Analyzer which affected more than \daSigVersionPerc\ of devices for more than \daSigVersionDays\ days.
+The number of updates received by devices (\daNumFullVersions) is the number of different full version strings observed in \da.
+The number of updates shipped by Google (\daNumSigOSVersions) is the number of Android versions reported in \da\ which affected more than \daSigVersionPerc\ of devices for more than \daSigVersionDays\ days.
 \todo{this is the wrong number to be using}
@@
 
-\section{Results}
+\section{Experiments}
 \label{sec:results}
@@
 \todo{ summarise our results}
-The Device Analyzer data used here was collected between \daStartDate\ and \daEndDate.
-The \texttt{androidvulnerabilities.org} data was collected between \avoStartDate\ and \avoEndDate.
+The \da\ data used here was collected between \daStartDate\ and \daEndDate.
+The \avo\ data was collected between \avoStartDate\ and \avoEndDate.
 \todo{comparison of nexus update rates with iOS}
-
-The proportion of devices in the Device Analyzer data running different versions of Android each day is shown in Figure~\ref{fig:norm_os}.
+We conducted four experiments examining the behaviour of the Android ecosystem as a whole (\S\ref{sec:exp:android_ecosystem}), updates reaching particular devices (\S\ref{sec:exp:device_updates}), comparisons between different manufacturers and operators (\S\ref{sec:exp:security_score}) and modeling the replacement by new Android versions in order to predict how long it will take for an update to be fixed (\S\ref{sec:exp:predicting_distributions}).
+\todo{In an experiment you test a hypothesis, it needs to be clear what the hypothesis is and what the conclusion of the test is}
+\subsection{Experiment 1: behaviour of the Android ecosystem}\label{sec:exp:android_ecosystem}
+To investigate the behaviour of the Android ecosystem as a whole with respect to Android versions and vulnerabilities we used the OS version information from \da\ and the vulnerability data from \avo.
+The proportion of devices in the \da\ data running different versions of Android each day is shown in Figure~\ref{fig:norm_os}.
 It shows how old versions are gradually replaced by new ones, the long tail of devices which do not see updates to more recent versions and the fluctuation in the number of devices in the data.
@@
  \includegraphics[width=\columnwidth]{figures/da_norm_os}
- \caption{Android versions in Device Analyzer data over time}
+ \caption{Android versions in \da\ data over time}
  \label{fig:norm_os}
@@
 For each vulnerability it shows the proportion of devices exposed to that vulnerability and how that changes over time.
-In July 2011 at the beginning of the Device Analyzer data the \emph{exploid}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/exploid_udev}} and \emph{levitator}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/levitator}} vulnerabilities both affect most Android devices, slowly these are fixed as updates roll out and devices are replaced until in January 2013 a much smaller proportion of devices are affected by known vulnerabilities.
+In July 2011 at the beginning of the \da\ data the \emph{exploid}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/exploid_udev}} and \emph{levitator}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/levitator}} vulnerabilities both affect most Android devices, slowly these are fixed as updates roll out and devices are replaced until in January 2013 a much smaller proportion of devices are affected by known vulnerabilities.
 However when in February 2013 the first APK signing vulnerability was found which affected all previous versions of Android and even in October 2013 most devices were still vulnerable.
@@
 
-Those graphs summarise data across all the devices, however one of the advantages of the Device Analyzer data is that it allows us to look at what happens to individual devices over time.
-Figure~\ref{fig:device_data} shows how the number of vulnerabilities affecting the \daNumDeviceDataDevices\ devices which have contributed the most days of data to Device Analyzer changes over time.
+\subsection{Experiment 2: Updates to particular devices}\label{sec:exp:device_updates}
+Those graphs summarise data across all the devices, however one of the advantages of the \da\ data is that it allows us to look at what happens to individual devices over time.
+Figure~\ref{fig:device_data} shows how the number of vulnerabilities affecting the \daNumDeviceDataDevices\ devices which have contributed the most days of data to \da\ changes over time.
 The trend that we saw in Figure~\ref{fig:proportioninsecure} of security improving and then getting worse is also shown here.
@@
  \includegraphics[width=\columnwidth]{figures/device-data-all-security}
- \caption{Number of vulnerabilities affecting the top \daNumDeviceDataDevices\ devices by days of contribution in the Device Analyzer data}
+ \caption{Number of vulnerabilities affecting the top \daNumDeviceDataDevices\ devices by days of contribution in the \da\ data}
  \label{fig:device_data}
@@
 Figure~\ref{fig:from_to_updates} shows how devices upgrade between different versions of Android.
-Device Analyzer (DA) observes upgrade events and the figure shows which versions a device is upgrading from and to.
+\da\ (DA) observes upgrade events and the figure shows which versions a device is upgrading from and to.
 Mostly the dark cells are above the diagonal and are upgrades (\daNumUpdatesUpgrades).
@@
 
-\subsection{Comparison with iOS}
-\todo{Get data from Richard so that we can do this comparison}
+% We don't need this for this paper
+%\subsection{Comparison with iOS}
+%\todo{Get data from Richard so that we can do this comparison}
 
-\subsection{Predicting future version distributions}
+\subsection{Experiment 4: Predicting future version distributions}\label{sec:exp:predicting_distributions}
 Examining the way that the OS version distribution changes over time in Figure~\ref{fig:norm_os} we see some regularity to the behaviour.
 To show this behaviour more clearly we normalised the data to the number of days since the version was released and is smoothed by forcing it to be monotonically decreasing.\footnote{While sometimes an older version becomes more popular after a later version has been released causing the values to actually increase, mostly this was just noise or changes in the distribution of devices (e.g.\ the sudden arrival of 1000 devices from Bangladesh due to a study being conducted there) and this assumption makes the graph clearer.}
-Then we can plot Figure~\ref{fig:frvh_os_versions} which shows the proportion of of devices in Device Analyzer not upgraded to each Android version or a later Android version (i.e.
+Then we can plot Figure~\ref{fig:frvh_os_versions} which shows the proportion of of devices in \da\ not upgraded to each Android version or a later Android version (i.e.
 for 2.1.0 it includes 2.1.1, 2.2.0 and indeed all devices as they all run later versions of Android).
@@
 
-\subsection{The Device Analyzer data gives a conservative estimate of the Android version distribution}
+\section{Threats to validity}
+\subsection{The \da\ data gives a conservative estimate of the Android version distribution}
 \label{sec:representative}
@@
  \includegraphics[width=\columnwidth]{figures/norm_api_gpcomp}
- \caption{Device Analyzer data on proportion of devices running different Android API versions}
+ \caption{\da\ data on proportion of devices running different Android API versions}
  \label{fig:da_api}
@@
  \includegraphics[width=\columnwidth]{figures/api_gpcomp_rdiff}
- \caption{Difference between Device Analyzer and Google Play data on the proportion of devices running different Android API versions}
+ \caption{Difference between \da\ and Google Play data on the proportion of devices running different Android API versions}
  \label{fig:da_gp_comp_diff}
 \end{figure}
-The data from Device Analyzer (DA) we used to investigate the proportion of devices exposed to different vulnerabilities is the OS version.
+The data from \da\ (DA) we used to investigate the proportion of devices exposed to different vulnerabilities is the OS version.
 Unfortunately there is no authoritative source of OS version information and so we cannot check directly whether our data is representative.
@@
 While API versions are too coarse grained to use for security update detection they are closely related to OS versions and so if the DA data on API versions is similar to the Google Data on API versions then the DA data on OS versions should be representative.
-Figure~\ref{fig:play_api} shows the data from Google and Figure~\ref{fig:da_api} shows the data from Device Analyzer and they appear similar.
+Figure~\ref{fig:play_api} shows the data from Google and Figure~\ref{fig:da_api} shows the data from \da\ and they appear similar.
 Figure~\ref{fig:da_gp_comp_diff} shows the difference between Figures~\ref{fig:play_api} and \ref{fig:da_api}, normalising for days since the API version was released.
-It shows that the Device Analyzer data systematically overestimates the prevalence of new API versions and underestimates the prevalence of old API versions.
-This means that the OS version information from Device Analyzer is likely to be overestimating the prevalence of new OS versions and hence our results are a conservative estimate of the security of Android.
+It shows that the \da\ data systematically overestimates the prevalence of new API versions and underestimates the prevalence of old API versions.
+This means that the OS version information from \da\ is likely to be overestimating the prevalence of new OS versions and hence our results are a conservative estimate of the security of Android.
 \todo{we want a statistical metric to claim this strongly with.}

### 2014-10-20 securityupdates.tex
Move some subsections around
Further changes based on Alastair's comments

@@
 \newcommand{\da}{Device Analyzer}
+\newcommand{\dafoot}{\textsuperscript{\ref{foot:dadata}}}
 \input{avostats}
@@
 Our hypothesis is that these fixes do not promptly reach the users devices and so much of the time users are running versions of Android known to be vulnerable.
-We used da data from over \daOSYearsOfData\ years and \daNumOSDevices\ devices and found that on average over \daMeanInsecurityPerc\ of devices were exposed to known root privilege vulnerabilities and only \daUpdatednessPerc\ of devices run the most recent version of Android.
+We used \da\ data from over \daOSYearsOfData\ years and \daNumOSDevices\ devices and found that on average over \daMeanInsecurityPerc\ of devices were exposed to known root privilege vulnerabilities and only \daUpdatednessPerc\ of devices run the most recent version of Android.
 There was also a period of several months when no devices ran secure versions of Android.
@@
 \section{Introduction}
-Android has \percMarketShare\ of the smartphone market and while the core development is controlled by Google there are about \daNumManufacturers\footnote{\label{foot:dadata}We computed this from the da~\cite{Wagner2013} data, see \S\ref{sec:android_update_process}.} manufacturers which make devices which run Android.
+Android has \percMarketShare\ of the smartphone market and while the core development is controlled by Google there are about \daNumManufacturers\footnote{\label{foot:dadata}We computed this from the \da~\cite{Wagner2013} data, see \S\ref{sec:android_update_process}.} manufacturers which make devices which run Android.
 Android receives regular updates  which add new features and fix bugs, including security vulnerabilities.
-Many of these manufacturers customise the version of Android they ship and sometimes network operators (of which there are at least \daNumOperators\textsuperscript{\ref{foot:dadata}}) make further modifications.
+Many of these manufacturers customise the version of Android they ship and sometimes network operators (of which there are at least \daNumOperators\dafoot) make further modifications.
 Hence when Google produces an update to Android, the update may have to pass through the manufacturer and operator before reaching the user.
@@
 
-The proportion of Android devices in the da data which were running versions of Android known to have root vulnerabilities at the time is shown in Figure~\ref{fig:proportioninsecure}.
+The proportion of Android devices in the \da\ data which were running versions of Android known to have root vulnerabilities at the time is shown in Figure~\ref{fig:proportioninsecure}.
 In \S\ref{sec:background} we will explain where these data come from and then in \S\ref{sec:results} we will show these data in more detail.
@@
 We only looked for \emph{root equivalent} vulnerabilities which did not require USB debugging to exploit.
-\emph{Root equivalent} means that it grants privileges equivalent in scope to root which can then either be used to gain root or provide sufficiently many permissions as to be as bad as root.
-Some phones can be `rooted' by enabling USB debugging and using the special privileges of the adb shell to root the device but only \daAdbEnabledPerc\ of devices have USB debugging enabled.
+\emph{Root equivalent} means that it grants privileges equivalent in scope to root which can then either be used to gain root or provide sufficiently many permissions as to be as bad as root.\todo{refine this definition}
+Some phones can be `rooted' by enabling USB debugging and using the special privileges of the adb shell to root the device but only \daAdbEnabledPerc\dafoot\ of devices have USB debugging enabled.
 This is not something that applications running on the phone can exploit to break compartmentalisation and so we do not include those vulnerabilities.
@@
 Some vulnerabilities which grant root are not traditional kernel vulnerabilities, for example the discovery of flaws in the verification of signatures on Android applications in February 2013~\cite{Forristal2013} meant that applications could pretend to be signed with system keys and hence gain root equivalent privileges.
-On some versions of Android (below 4.1) they could use known system to root escalation mechanisms but on all versions they have a greatly increased attack area for further privilege escalation and also have the ability to control all user internet traffic (via VPNs), brick the phone, remove and install apps, steal user credentials, read the screen, make and receive calls.
+On some versions of Android (below 4.1) they could use known system to root escalation mechanisms but on all versions they have a greatly increased attack area for further privilege escalation and also have the ability to control all user internet traffic (via VPNs), brick the phone, remove and install apps, steal user credentials, read the screen and make as well as receive calls.
+
+\avoTabAndVulns
 
@@
 
-There was much manual effort required to track down details of these vulnerabilities as many of them are not in other databases such as the CVE database, the lack of a widely acknowledged unique identifier made identifying whether two reports referenced the same vulnerability difficult.
-Previous work has assumed ``any security issue of relevance will eventually get a CVE number assigned"~\cite{Frei2010} which is currently not the case for Android with several root vulnerabilities not having CVE numbers.
-For some of the vulnerabilities without CVE numbers we checked with Google whether they had one assigned and they confirmed that there was no number and that they did not intend to get one (though they did give us the Android bug number and tell us to use that).
+The database of vulnerabilities which we maintain at \avo\ contains \avoNumVulnerabilities\ vulnerabilities of which \avoNumVulnAllAndroid\ affect all Android devices and \avoNumVulnSpecific\ are specific to particular devices or manufacturers.
+We only use the \daNumVulnsUsed\ vulnerabilities which it is convenient to use for this analysis because they affect all Android devices rather than particular versions and we know enough information about the vulnerability to tie it to the particular versions it affects.
+
+Tracking vulnerabilities is a manual task as they are not all in other databases such as the CVE database and the lack of a widely acknowledged unique identifier made identifying whether two reports referenced the same vulnerability difficult.
+Previous work has assumed ``any security issue of relevance will eventually get a CVE number assigned"~\cite{Frei2010} which is currently not the case for Android root vulnerabilities.
+For some of the vulnerabilities without CVE numbers, Google confirmed that there was no CVE number and that they did not intend to get one, instead providing an Android bug number.
 
@@
 
+\subsection{Lifetime of a vulnerability}
+The key events in the lifetime of a vulnerability are:
+
+\begin{LaTeXdescription}
+ \item[introduction] When a vulnerability was introduced into the source code.
+ \item[introducing release] When the first release was made containing the vulnerability.
+ \item[fix] When a vulnerability was first fixed in the source code.
+ \item[fixing release] When the first release containing the fix was made.
+ \item[fix deployed] When the fix has been deployed to a sufficiently large proportion of the population that the vulnerability can be ignored.
+ \item[discovery] When the vulnerability is first discovered
+ \item[exploit] When the vulnerability is first exploited
+ \item[disclosed] When the vulnerability is first disclosed % publically or to a smaller set of people?
+\end{LaTeXdescription}
+
+\todo{Compare this list with the literature and unify it. Pull out figures which illustrate and investigate this in our data.}
+
 Prior work~\cite{Bilge2012} by Symantec on desktop malware has shown that after public disclosure exploitation rates increase by 5 orders of magnitude and so from the point of view of widespread danger, the period between public disclosure and the update reaching the user's device is the most critical.
@@
 We cannot know if someone reported the vulnerability to the manufacturer and we do know that various agencies are engaged in widespread monitoring of communications and compromise of civilian infrastructure~\cite{TODO} and so they probably know.
-Hence we use dates such as: when CVEs were registered, when fixes were committed to source repositories, or when exploits were committed to repositories; as the date first known about.
-
-\avoTabAndVulns
-
-The database of vulnerabilities which we maintain at \avo\ contains \avoNumVulnerabilities\ vulnerabilities of which \avoNumVulnAllAndroid\ affect all Android devices and \avoNumVulnSpecific\ are specific to particular devices or manufacturers.
-We only use the \daNumVulnsUsed\ vulnerabilities which it is convenient to use for this analysis because they affect all Android devices rather than particular versions and we know enough information about the vulnerability to tie it to the particular versions it affects.
+Hence we use dates such as: when CVEs were registered, when fixes were committed to source repositories, or when exploits were committed to repositories; as the date first known about.\todo{We need to be more precise here}
 
@@
  Numbers on edges indicate updates shipped between \daStartDate\ and \daEndDate, numbers in brackets represent number of such entities in our data.
- Dotted arrows indicate flows where we do not know how many updates are being produced.}
+ Dotted arrows indicate flows where we do not know how many updates are being produced as we can't measure those flows directly as they are not public.\todo{Update this diagram}}
  \label{fig:update_ecosystem}
@@
 The Android platform is then built on top of those with kernel and userspace components by Google. \todo{ work out the exact arrangements.}
-Usually the code for each update is kept secret until after the update has been published though through the Open Handset Alliance some manufacturers see it in advance.
+Usually the code for each update is kept secret\todo{Can we quantify this keeping the code secret? Is it worth it?} until after the update has been published though through the Open Handset Alliance.
+Manufacturers in the  alliance receive advanced copies in order to prepare handsets.
 Then the Manufacturers take the update and customise it before passing it on to the operator.
@@
 %An analysis by Vidas et al.~\cite{Vidas2011} of the Android 2.1 to 2.2 update found that it took 11 months from when Google released 2.2 for the last device which they were investigating to get the update.
-\todo{talk about update latency}
+\todo{talk about update latency, quantify this in the text as per the figure, giving more explanation, use statistics from samsung-updates.com -> how many binaries are there per device?
+Do we know the variation between network operators?}
 
@@
 
+\subsection{Monetisation}
+\todo{redo this section in light of related work}
+
+One of the key differences between smartphones and desktops is that it is easier to directly cash out on a smartphone as malware can send premium rate texts or make premium rate phone calls.
+It can also steal personal data which already has semantic information associated with it, this also makes it easier to perform ransomware attacks where malware deletes this information from the device and charges a ransom for its return.
+It can also do more conventional things such as use the devices for DDoS attacks and altering the ads displayed to send the fees to the malware author.
+While many of these things can be done by malware which just requests the relevant permissions, root exploits make this easier, particularly for ransomware and altering the ads displayed.
+
+\subsection{Incentives and participants}
+\label{sec:economics}
+\cite{Felt2011}
+
 \section{Experiments}
 \label{sec:results}
-In this section we present the results of our analysis showing \daMeanInsecurityPerc\ of devices to have root vulnerabilities on average, nexus devices to be better than others, \emph{\daSecScoreBestmanufacturer} to be the best manufacturer and \emph{\daSecScoreBestoperator} to be the best network operator.
+In this section we present the results of our analysis showing \daMeanInsecurityPerc\ of devices to have root vulnerabilities on average.
+Nexus devices are better than others, \emph{\daSecScoreBestmanufacturer} to be the best manufacturer and \emph{\daSecScoreBestoperator} to be the best network operator.
 \todo{ summarise our results}
@@
 For each vulnerability it shows the proportion of devices exposed to that vulnerability and how that changes over time.
-In July 2011 at the beginning of the \da\ data the \emph{exploid}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/exploid_udev}} and \emph{levitator}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/levitator}} vulnerabilities both affect most Android devices, slowly these are fixed as updates roll out and devices are replaced until in January 2013 a much smaller proportion of devices are affected by known vulnerabilities.
+In July 2011 at the beginning of the \da\ data the \emph{exploid}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/exploid_udev}} and \emph{levitator}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/levitator}} vulnerabilities both affect most Android devices, slowly these are fixed as updates roll out and devices are replaced until in January 2013 a much smaller proportion of devices are affected by known vulnerabilities.\todo{Quantify whether it is replacement of updates which causes the fixes}
 However when in February 2013 the first APK signing vulnerability was found which affected all previous versions of Android and even in October 2013 most devices were still vulnerable.
@@
 However what is perhaps more worrying is the long tail on the \emph{Gingerbreak}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/Gingerbreak}}, \emph{levitator} and \emph{exploid} vulnerabilities which are more dangerous root vulnerabilities (not requiring new APK installation) and which still affect a significant proportion of devices years later.
+\todo{break things down by attack vector?}
 
@@
 
-\subsection{Update frequency}
+\subsubsection{Update frequency}
 \todo{ find out what the frequency of updates is}
@@
  \includegraphics[width=\columnwidth]{figures/from_to_updates.pdf}
- \caption{Updates between different Android versions in the DA data}
+ \caption{Updates between different Android versions in the \da data}
  \label{fig:from_to_updates}
@@
 The data appears to have an exponential decay as it tends to 1 but with a slow start as few devices are upgraded initially.
-We approximate this with a a $e^{-x}$ component for the main behaviour and a which is offset to start later when uptake takes off.
+We approximate this with a $e^{-x}$ for the main behaviour and a which is offset to start later when uptake takes off.
 The equation is then:
@@
 
-Fitting that curve to the data using \texttt{scipy.optimize}, we get the parameters $\mathrm{undeployed} = \daOSCurveFitParamFirst ~\si{days}$, $\mathrm{decay} = \daOSCurveFitParamSecond ~\si{days^{-1}}$ for OS versions and $\mathrm{undeployed} = \daAPICurveFitParamFirst ~\si{days}$, $\mathrm{decay} = \daAPICurveFitParamSecond ~\si{days^{-1}}$ for API versions.
+Fitting that curve to the data using \texttt{scipy.optimize}, we get a RMSE of \daOSCurveFitRMSE\ and parameters $\mathrm{undeployed} = \daOSCurveFitParamFirst ~\si{days}$, $\mathrm{decay} = \daOSCurveFitParamSecond ~\si{days^{-1}}$ for OS versions and RMSE of \daAPICurveFitRMSE\, $\mathrm{undeployed} = \daAPICurveFitParamFirst ~\si{days}$, $\mathrm{decay} = \daAPICurveFitParamSecond ~\si{days^{-1}}$ for API versions.
 From this the number of days from release of a new version of Android until 50\% of devices are running that version or higher is \daOSCurveHalfDeployed\ and full deployment to \daFullDeployedAt\ of devices takes \daOSCurveFullDeployed\ days.
@@
 \todo{Our handling of errors here is not quite good enough as the uncertainty from our y value interacts badly with the logarithm function when computing the inverse and we compute the log of ~0 which means our error ends up much bigger than our value - but that is an upper bound, not a +- bound.}
-
-\subsection{Lifetime of a vulnerability}
-The key events in the lifetime of a vulnerability are:
-
-\begin{LaTeXdescription}
- \item[introduction] When a vulnerability was introduced into the source code.
- \item[introducing release] When the first release was made containing the vulnerability.
- \item[fix] When a vulnerability was first fixed in the source code.
- \item[fixing release] When the first release containing the fix was made.
- \item[fix deployed] When the fix has been deployed to a sufficiently large proportion of the population that the vulnerability can be ignored.
- \item[discovery] When the vulnerability is first discovered
- \item[exploit] When the vulnerability is first exploited
- \item[disclosed] When the vulnerability is first disclosed % publically or to a smaller set of people?
-\end{LaTeXdescription}
-
-\todo{Compare this list with the literature and unify it. Pull out figures which illustrate and investigate this in our data.}
-
-\subsection{Monetisation}
-\todo{redo this section in light of related work}
-
-One of the key differences between smartphones and desktops is that it is easier to directly cash out on a smartphone as malware can send premium rate texts or make premium rate phone calls.
-It can also steal personal data which already has semantic information associated with it, this also makes it easier to perform ransomware attacks where malware deletes this information from the device and charges a ransom for its return.
-It can also do more conventional things such as use the devices for DDoS attacks and altering the ads displayed to send the fees to the malware author.
-While many of these things can be done by malware which just requests the relevant permissions, root exploits make this easier, particularly for ransomware and altering the ads displayed.
-
-\subsection{Incentives and participants}
-\label{sec:economics}
-\cite{Felt2011}
 
@@
 However Google has published API version information every month since December 2009 and we have collated this information.\footnote{\url{https://www.cl.cam.ac.uk/~drt24/android/versiondistribution/}}
-While API versions are too coarse grained to use for security update detection they are closely related to OS versions and so if the DA data on API versions is similar to the Google Data on API versions then the DA data on OS versions should be representative.
+While API versions are too coarse grained to use for security update detection they are closely related to OS versions and so if the \da\ data on API versions is similar to the Google Data on API versions then the \da\ data on OS versions should be representative.
 Figure~\ref{fig:play_api} shows the data from Google and Figure~\ref{fig:da_api} shows the data from \da\ and they appear similar.

### 2014-10-21 security_scoring.tex
Further changes based on Alastair's comments
Further refactoring

@@
 
-We want to provide a score out of 10 as this is easy for people to understand.
+We want to provide a score out of 10 as this is easy for phone buyers to understand as many ratings are given as a score out of 10.
 Since $f$ is the most important metric we weight it more highly.
@@
 Hence having a worse score does not necessarily mean that an operator is worse, it could be that its users all pick phones from a worse manufacturer, for example because they were cheaper.
+However an operator could use data from this paper to exclude insecure devices from those offered to consumers.
 

### 2014-10-21 securityupdates.tex
Further refactoring

@@
 Operators are cautious because if they ship a broken update then they will get many support requests which is expensive.
-However there is ongoing legal action to force operators to ship updates for security vulnerabilities~\cite{Soghoian2013}.
-Many phones are sold on long (12--24 month) contracts with monthly payments yet stop receiving updates before the end of the contract, while Windows XP could be purchased for a one off payment in October 2001 and received security updates until April 2014.
+However there is ongoing legal action to force operators to ship updates for security vulnerabilities~\cite{Soghoian2013}.\todo{Check on the status of this legal action}
+Many phones are sold on long (12--24 month) contracts with monthly payments yet stop receiving updates before the end of the contract,\todo{Is there a survey of the distribution of contract lengths and phone purchase methods?} while Windows XP could be purchased for a one off payment in October 2001 and received security updates until April 2014.
 Corporate or public sector buyers are encouraged to purchase secure devices, but we have found little concrete guidance.
-For example, CESG, which advises the UK government on how to secure its computer systems, recommends picking Android devices from manufactures which are good at shipping security updates promptly~\cite{CESG2013} but it does not state which manufacturers are better.
+For example, CESG, which advises the UK government on how to secure its computer systems, recommends picking Android devices from manufactures which are good at shipping security updates promptly~\cite{CESG2013} but it does not state which manufacturers are better.\todo{Which other advisory organisations give that advice?}
 We provide this information in~\S\ref{sec:security_scoring}.
@@
 Between 36.7\%~\cite{Zhou2012b} and 40\%~\cite{Zhou2012a} of malware for Android contained root exploits in 2012.
-While this shows that much Android malware does not need a root exploit to work, as it can just request the relevant permissions and the user will grant them, a significant proportion does.
+While this shows that much Android malware does not need a root exploit to work, since the user will grant the relevant permissions if the app requests them, a significant proportion does use root exploits.
 This malware is also more dangerous because while ordinary malware can be remotely removed by Google through the Play store, once a root exploit has been used there are no guarantees.
@@
 The proportion of Android devices in the \da\ data which were running versions of Android known to have root vulnerabilities at the time is shown in Figure~\ref{fig:proportioninsecure}.
-In \S\ref{sec:background} we will explain where these data come from and then in \S\ref{sec:results} we will show these data in more detail.
 
@@
 \label{sec:background}
-There are two sources of data we needed to have and to link for this study, information on root vulnerabilities and information on installed versions of Android.
+There are two sources of data we needed to have and to link for this study, information on the distribution of installed versions of Android over time and on root vulnerabilities.
 We need data on root vulnerabilities so that we know which versions of Android were vulnerable to which known vulnerabilities at what times.
@@
 
+\subsection{Versions of Android running on devices}
+We need historical information on which versions of Android were running on devices along with other information about those devices which allows us to go into more detail and confirm the reliability of these data which our results depend on.
+The only suitable data which is available for this is the \da\ data~\cite{Wagner2013}.
+The \da\ data contains data from \daNumDevices\ devices with a total of \daDeviceDays\footnote{Here we are only counting devices and days for which we have valid OS version data} device days, it contains data for more than \daMonths\ months for \daMonthsDevices\ devices.
+\todo{ talk about the \da\ project a bit}
+Various different kinds of data\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/collected.htm}} are collected by the \da\ Android app\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/}}.
+Among these are the build string and API version for the version of Android currently running on the device each day.
+The API/SDK version is a well defined integer, however it does not always change with new Android releases, particularly security bug fixes should not increase this value.
+The build string is `The user-visible version string', fortunately most (99.9\%) devices in the data have a build string of the form `x.y.z random string' and so it is possible to extract the Android version number.
+On a large proportion of devices `random string' is a well defined build number but different manufacturers use different schema.
+
 \subsection{Root vulnerabilities}
-We compiled a list of root vulnerabilities for Android containing information on when they were known about, what versions they affected and which versions fixed the problem.
+We compiled a list of root equivalent vulnerabilities for Android containing information on when they were known about, what versions they affected and which versions fixed the problem.
 We only looked for \emph{root equivalent} vulnerabilities which did not require USB debugging to exploit.
 \emph{Root equivalent} means that it grants privileges equivalent in scope to root which can then either be used to gain root or provide sufficiently many permissions as to be as bad as root.\todo{refine this definition}
-Some phones can be `rooted' by enabling USB debugging and using the special privileges of the adb shell to root the device but only \daAdbEnabledPerc\dafoot\ of devices have USB debugging enabled.
+Some phones can be `rooted' by enabling USB debugging and using the special privileges of the adb shell to root the device but only \daAdbEnabledPerc\dafoot\ of devices have USB debugging enabled.\footnote{This is surprisingly high as \daAdbEnabledPerc\ of Android users are not developers.}
 This is not something that applications running on the phone can exploit to break compartmentalisation and so we do not include those vulnerabilities.
 Unfortunately many published exploits use adb for convenience and so determining whether that is necessary to exploit the vulnerability can be difficult.
+
 Some vulnerabilities which grant root are not traditional kernel vulnerabilities, for example the discovery of flaws in the verification of signatures on Android applications in February 2013~\cite{Forristal2013} meant that applications could pretend to be signed with system keys and hence gain root equivalent privileges.
@@
 
+We developed and maintain an open platform for filing root equivalent vulnerabilities in a machine readable format, \url{http://androidvulnerabilities.org/}.
+We seeded it with data from the CVE database, vendor lists, reports from the literature and various forums.
+We have received submissions or amendments from \avoNumSubmitters\ individuals.
+The \avo\ contains \avoNumVulnerabilities\ vulnerabilities of which \avoNumVulnAllAndroid\ affect all Android devices and \avoNumVulnSpecific\ are specific to particular devices or manufacturers.
+We only use the \daNumVulnsUsed\ vulnerabilities which it is convenient to use for this analysis because they affect all Android devices rather than particular versions and we know enough information about the vulnerability to tie it to the particular versions it affects.
 We have published full details of the \daNumVulnsUsed\ vulnerabilities used for this analysis in an accompanying technical report~\cite{TODO} and summarise them in Table~\ref{tab:andvulns}.
-We also have a website\footnote{\url{http://androidvulnerabilities.org/}} which we will keep up to date with future vulnerabilities and which has successfully solicited submissions from others.
-In this analysis we only use vulnerabilities which affected all Android devices, not any of the manufacturer or hardware specific vulnerabilities
-
-The database of vulnerabilities which we maintain at \avo\ contains \avoNumVulnerabilities\ vulnerabilities of which \avoNumVulnAllAndroid\ affect all Android devices and \avoNumVulnSpecific\ are specific to particular devices or manufacturers.
-We only use the \daNumVulnsUsed\ vulnerabilities which it is convenient to use for this analysis because they affect all Android devices rather than particular versions and we know enough information about the vulnerability to tie it to the particular versions it affects.
 
 Tracking vulnerabilities is a manual task as they are not all in other databases such as the CVE database and the lack of a widely acknowledged unique identifier made identifying whether two reports referenced the same vulnerability difficult.
-Previous work has assumed ``any security issue of relevance will eventually get a CVE number assigned"~\cite{Frei2010} which is currently not the case for Android root vulnerabilities.
+Previous work has assumed ``any security issue of relevance will eventually get a CVE number assigned"~\cite{Frei2010} which is currently not the case for Android root equivalent vulnerabilities.
 For some of the vulnerabilities without CVE numbers, Google confirmed that there was no CVE number and that they did not intend to get one, instead providing an Android bug number.
@@
 
-\subsection{Versions of Android running on devices}
-We also need historical information on which versions of Android were running on devices along with other information about those devices which allows us to go into more detail and confirm the reliability of the data our results depend on.
-The only suitable data which is available for this is the \da\ data~\cite{Wagner2013}.
-The \da\ data contains data from \daNumDevices\ devices with a total of \daDeviceDays\footnote{Here we are only counting devices and days for which we have valid OS version data} device days, it contains data for more than \daMonths\ months for \daMonthsDevices\ devices.
-\todo{ talk about the \da\ project a bit}
-Various different kinds of data\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/collected.htm}} are collected by the \da\ Android app\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/}}.
-Among these are the build string and API version for the version of Android currently running on the device each day.
-The API/SDK version is a well defined integer, however it does not always change with new Android releases, particularly security bug fixes should not increase this value.
-The build string is `The user-visible version string', fortunately most (99.9\%) devices in the data have a build string of the form `x.y.z random string' and so it is possible to extract the Android version number.
-On a large proportion of devices `random string' is a well defined build number but different manufacturers use different schema.
-
+\section{Android Ecosystem}
 \subsection{Running a version of Android known to be vulnerable makes a device vulnerable}
@@
 \label{sec:results}
-In this section we present the results of our analysis showing \daMeanInsecurityPerc\ of devices to have root vulnerabilities on average.
+In this section we present the results of our analysis showing \daMeanInsecurityPerc\ of devices to have root equivalent vulnerabilities on average.
 Nexus devices are better than others, \emph{\daSecScoreBestmanufacturer} to be the best manufacturer and \emph{\daSecScoreBestoperator} to be the best network operator.
@@
  \includegraphics[width=\columnwidth]{figures/frvh_os_versions}
- \caption{Proportion of devices not updated to particular versions of Android or any later version}
+ \caption{Proportion of devices not updated to particular versions of Android or any later version\todo{make line of best fit thicker}}
  \label{fig:frvh_os_versions}
@@
 
+\subsection{Changes in the \da\ sample}
+\todo{Bangladesh etc.}
+
 \section{Discussion}
-There are continuing efforts to reduce the impact of root vulnerabilities, both in Android and more widely.
+There are continuing efforts to reduce the impact of root equivalent vulnerabilities, both in Android and more widely.
 SEAndroid~\cite{Smalley2013} which is included in Android from version 4.1~\cite{jelly-bean-release} claimed to prevent some root vulnerabilities and to reduce the impact of others.
 Capability based enforcement systems such as Capsicum~\cite{Watson2010} substantially reduce the capabilities that an exploit has to try and gain increased privilege with.
-When Capsicum is included in Linux~\cite{TODO} and hence in Android, it could be used to place fine grained restrictions on system daemons, preventing vulnerabilities becoming root vulnerabilities.
+When Capsicum is included in Linux~\cite{TODO} and hence in Android, it could be used to place fine grained restrictions on system daemons, preventing vulnerabilities becoming root equivalent vulnerabilities.
 
@@
 \label{sec:related}
-We found that since 2011 \daMeanInsecurityPerc\ of devices on average were exposed to known root vulnerabilities, in 2011 Felt et al.\ studied 6 Android handsets and found that they were exposed to root vulnerabilities at least 74\% of the time.
+We found that since 2011 \daMeanInsecurityPerc\ of devices on average were exposed to known root equivalent vulnerabilities.
+In 2011 Felt et al.\ \cite{Felt2011} studied 6 Android handsets and found that they were exposed to root vulnerabilities at least 74\% of the time.
 This might indicate that over the last 3 years the security of Android has not improved, but these data are not directly comparable as their study was on 6 handsets considering the best possible update distribution while ours is on a large sample of devices with the real update distribution.

### 2014-10-21 securityupdates.tex
Further tweaks

@@
 \subsection{Versions of Android running on devices}
-We need historical information on which versions of Android were running on devices along with other information about those devices which allows us to go into more detail and confirm the reliability of these data which our results depend on.
+We need historical data on the distributions of versions of Android running on devices and additional data about those devices to allows us to investigate variation between devices and confirm the reliability of these data.
 The only suitable data which is available for this is the \da\ data~\cite{Wagner2013}.
@@
 Among these are the build string and API version for the version of Android currently running on the device each day.
-The API/SDK version is a well defined integer, however it does not always change with new Android releases, particularly security bug fixes should not increase this value.
-The build string is `The user-visible version string', fortunately most (99.9\%) devices in the data have a build string of the form `x.y.z random string' and so it is possible to extract the Android version number.
-On a large proportion of devices `random string' is a well defined build number but different manufacturers use different schema.
+The API version is a well defined integer, however it does not always change with new Android releases, particularly security bug fixes should not increase this value.
+The build string is `The user-visible version string', fortunately most (99.9\%)\todo{use a generated number} devices in the data have a build string of the form `x.y.z random string' and so it is possible to extract the Android version number.
+On a large proportion of devices `random string' is a well defined build number\footnote{\url{https://source.android.com/source/build-numbers.html}} but different manufacturers use different schema.
 
 \subsection{Root vulnerabilities}
-We compiled a list of root equivalent vulnerabilities for Android containing information on when they were known about, what versions they affected and which versions fixed the problem.
+We compiled a list of root equivalent vulnerabilities in Android.
+We collected information on discovery and publication dates, what versions they affected and which versions fixed the problem.
 We only looked for \emph{root equivalent} vulnerabilities which did not require USB debugging to exploit.
-\emph{Root equivalent} means that it grants privileges equivalent in scope to root which can then either be used to gain root or provide sufficiently many permissions as to be as bad as root.\todo{refine this definition}
+\emph{Root equivalent} means that it escalates to privileges equivalent in scope to root in that either they can be used to gain root or they provide permissions as bad as root.
+If an application exploits a root equivalent vulnerability then it gains control of the device.
 Some phones can be `rooted' by enabling USB debugging and using the special privileges of the adb shell to root the device but only \daAdbEnabledPerc\dafoot\ of devices have USB debugging enabled.\footnote{This is surprisingly high as \daAdbEnabledPerc\ of Android users are not developers.}
@@
 We have received submissions or amendments from \avoNumSubmitters\ individuals.
-The \avo\ contains \avoNumVulnerabilities\ vulnerabilities of which \avoNumVulnAllAndroid\ affect all Android devices and \avoNumVulnSpecific\ are specific to particular devices or manufacturers.
+\avo\ contains \avoNumVulnerabilities\ vulnerabilities of which \avoNumVulnAllAndroid\ affect all Android devices and \avoNumVulnSpecific\ are specific to particular devices or manufacturers.
 We only use the \daNumVulnsUsed\ vulnerabilities which it is convenient to use for this analysis because they affect all Android devices rather than particular versions and we know enough information about the vulnerability to tie it to the particular versions it affects.
@@
 
+\subsection{Lifetime of a vulnerability}
 Working out the date to use for the date on which a vulnerability was sufficiently well known to pose a threat to users is difficult.
@@
 
-\subsection{Lifetime of a vulnerability}
 The key events in the lifetime of a vulnerability are:
@@
 
-Prior work~\cite{Bilge2012} by Symantec on desktop malware has shown that after public disclosure exploitation rates increase by 5 orders of magnitude and so from the point of view of widespread danger, the period between public disclosure and the update reaching the user's device is the most critical.
+Symantec's 2012 analysis~\cite{Bilge2012} of desktop malware has shown that after public disclosure exploitation rates increase by 5 orders of magnitude and so from the point of view of widespread danger, the period between public disclosure and the update reaching the user's device is the most critical.
 However they also show that zero day vulnerabilities are typically used for 312 days before they are publicly disclosed, often to target particular organisations.
@@
 
+\subsection{Distribution of vulnerabilities}
 \begin{figure}
@@
 
+\subsection{Open questions}
+\begin{itemize}
+ \item Why was there a gap in vulnerability discovery in Android in 2012?
+ \item Why do so many devices have ADB enabled?
+\end{itemize}
+
 \section{Related work}

### 2014-10-22 securityupdates.tex
Add more data on analysis of external projects

@@
 Or that manufacturers made it easier to install custom versions of Android, reducing the need for users to root their devices?
-Or that manufacturer specific vulnerabilities proved to be easier to find and so people looking for vulnerabilities adjusted their focus?
+Or that manufacturer specific vulnerabilities (which we are ignoring) proved to be easier to find and so people looking for vulnerabilities adjusted their focus?
 
@@
 \subsection{Running a version of Android known to be vulnerable makes a device vulnerable}
-While we have not tried to exploit these vulnerabilities on the devices in order to test whether the devices are actually vulnerable, we are confident that most devices will be if they are running a vesion known to be vulnerable.
+While we have not tried to exploit these vulnerabilities\todo{ethics} on the devices in order to test whether the devices are actually vulnerable, we are confident that most devices will be if they are running a vesion known to be vulnerable.
 \todo{WHY?}
@@
 \subsection{Android update process}
+\todo{rewrite this}
 \label{sec:android_update_process}
@@
 There are five entities or groups which contribute towards Android updates: the network operators, the device manufacturers, the hardware developers, Google and the upstream open source projects.
-Android builds on various open source projects such as the Linux kernel, the OpenSSL and BouncyCastle cryptography libraries and various Apache Java libraries and so can include any of the versions of those projects, including those which fix security vulnerabilities in them.
-It also incorporates various drivers for different bits of hardware.
+Android builds on various open source projects such as the Linux kernel, the OpenSSL and BouncyCastle cryptography libraries and so can include any of the versions of those projects, including those which fix security vulnerabilities in them.
+There are \avoNumExternalProjects\ external projects in Android, contributing \avoTotalExternalLines\ lines of code.
+We analysed the top \avoNumBigExternalProjects\ by lines of code (contributing \avoNumBigExternalLinesOfCode, \avoBigExternalLinesOfCodePerc\ of the total) and were able to automatically extract the versions of those projects included in different versions of Android for \avoNumBigExternalProjectsAnalysed.
+We found \avoBigExternalTotalVersions\ distinct versions, a median of \avoBigExternalMedianVersions\ and mean of \avoBigExternalMeanVersions\ versions per project.
+Android also incorporates various drivers for different bits of hardware.
 The Android platform is then built on top of those with kernel and userspace components by Google. \todo{ work out the exact arrangements.}
@@
 \subsection{Monetisation}
-\todo{redo this section in light of related work}
+\todo{redo this section in light of related work and to fit into ecosystem discussion}
 

### 2014-10-23 securityupdates.tex
Explain why we did not exploit devices

@@
 \subsection{Running a version of Android known to be vulnerable makes a device vulnerable}
-While we have not tried to exploit these vulnerabilities\todo{ethics} on the devices in order to test whether the devices are actually vulnerable, we are confident that most devices will be if they are running a vesion known to be vulnerable.
-\todo{WHY?}
+We have not tried to exploit these vulnerabilities on the devices in order to test whether the devices are actually vulnerable.
+Doing so would be difficult both ethically, as it would pose a high risk to user devices and mean that our code could easily be maliciously repurposed, and practically as getting all the exploits working reliably on all the devices they could be made to work reliably on requires extensive testing.
+We are confident that most devices will be vulnerable if they are running a version of Android known to be vulnerable. \todo{WHY?}
 

### 2014-10-23 securityupdates.tex
Expand discussion of external open source projects giving some of our method

@@
 Android builds on various open source projects such as the Linux kernel, the OpenSSL and BouncyCastle cryptography libraries and so can include any of the versions of those projects, including those which fix security vulnerabilities in them.
-There are \avoNumExternalProjects\ external projects in Android, contributing \avoTotalExternalLines\ lines of code.
-We analysed the top \avoNumBigExternalProjects\ by lines of code (contributing \avoNumBigExternalLinesOfCode, \avoBigExternalLinesOfCodePerc\ of the total) and were able to automatically extract the versions of those projects included in different versions of Android for \avoNumBigExternalProjectsAnalysed.
-We found \avoBigExternalTotalVersions\ distinct versions, a median of \avoBigExternalMedianVersions\ and mean of \avoBigExternalMeanVersions\ versions per project.
 Android also incorporates various drivers for different bits of hardware.
@@
 
-The number of updates shipped by upstream open source projects is calculated by scraping the releases pages for those projects, from which we can also obtain the release dates.
-By comparing the tags in the Android git version repositories for upstream projects with Android version numbers and examining the version of the upstream project under that tag we computed which upstream project versions were included in each version of Android.
+We extracted data on the external projects used in Android and have included this and the scripts which generated it in \avo.
+These scripts analysed the Android Open Source Project's source tree to examine the source code of each of the external projects to find the project version associated with each Android version tag on the repository.
+There are \avoNumExternalProjects\ external open source projects in Android, contributing \avoTotalExternalLines\ lines of code.\footnote{Lines of code were measured using David A.
+Wheeler's \texttt{sloccount}.}
+We analysed the top \avoNumBigExternalProjects\ by lines of code (contributing \avoNumBigExternalLinesOfCode\ lines, \avoBigExternalLinesOfCodePerc\ of the total) and were able to automatically extract the versions of those projects included in different versions of Android for \avoNumAnalysedExternalProjects\ (contributing \avoNumAnalysedExternalLinesOfCode\ lines, \avoAnalysedExternalLinesOfCodePerc\ of the total).
+We found \avoBigExternalTotalVersions\ distinct versions, a median of \avoBigExternalMedianVersions\ and mean of \avoBigExternalMeanVersions\ versions per project.
+This shows that Android rarely changes the version of external projects it includes.
+
+To compute the latency between upstream releases and their inclusion in Android we scraped the release pages for those projects, to obtain the version numbers and release dates.
 This allows us to compute the latency between an upstream project being released and it being included in Android, this is shown in Table~\ref{tab:update_ecosystem}.
@@
 \end{tabular}
-\caption{Flow of updates from upstream projects into Android. Number of updates as in Figure~\ref{fig:update_ecosystem}, latency in days for all pairs of versions we have data on.}
+\caption{Flow of updates from upstream projects into Android.
+Number of updates as in Figure~\ref{fig:update_ecosystem}, latency in days for all pairs of versions we have data on.\todo{scrape the other 26 websites...
+is it worth it?}}
 \label{tab:update_ecosystem}

### 2014-10-23 securityupdates.tex
Refactor introduction to Experiments

@@
 \todo{ summarise our results}
-The \da\ data used here was collected between \daStartDate\ and \daEndDate.
-The \avo\ data was collected between \avoStartDate\ and \avoEndDate.
-\todo{comparison of nexus update rates with iOS}
-We conducted four experiments examining the behaviour of the Android ecosystem as a whole (\S\ref{sec:exp:android_ecosystem}), updates reaching particular devices (\S\ref{sec:exp:device_updates}), comparisons between different manufacturers and operators (\S\ref{sec:exp:security_score}) and modeling the replacement by new Android versions in order to predict how long it will take for an update to be fixed (\S\ref{sec:exp:predicting_distributions}).
+%\todo{comparison of nexus update rates with iOS}
+We conducted four experiments examining the behaviour of the Android ecosystem as a whole (\S\ref{sec:exp:android_ecosystem}), updates reaching particular devices (\S\ref{sec:exp:device_updates}), comparisons between different manufacturers and operators (\S\ref{sec:exp:security_score}) and model how long it will take for an update to be fixed once an update is released (\S\ref{sec:exp:predicting_distributions}).
 \todo{In an experiment you test a hypothesis, it needs to be clear what the hypothesis is and what the conclusion of the test is}
@@
 To investigate the behaviour of the Android ecosystem as a whole with respect to Android versions and vulnerabilities we used the OS version information from \da\ and the vulnerability data from \avo.
+The \da\ data used here was collected between \daStartDate\ and \daEndDate.
+The \avo\ data was collected between \avoStartDate\ and \avoEndDate\ and covers the period from \avoFirstDataDate\ to \avoLastDataDate.
 The proportion of devices in the \da\ data running different versions of Android each day is shown in Figure~\ref{fig:norm_os}.

### 2014-10-24 security_scoring.tex
More fixes and headings

@@
 We propose a method to score a manufacturer, operator or device based on its historic performance at keeping devices up to date and fixing security vulnerabilities.
-We find that Android as a whole gits a score of \daSecurityScore / 10, the best manufacturer is \daSecScoreBestmanufacturer\ (\daSecScoreBestmanufacturerScore\ / 10) and the worst is \daSecScoreWorstmanufacturer\ (\daSecScoreWorstmanufacturerScore\ / 10).
+We find that Android as a whole gets a score of \daSecurityScore / 10, the best manufacturer is \daSecScoreBestmanufacturer\ (\daSecScoreBestmanufacturerScore\ / 10) and the worst is \daSecScoreWorstmanufacturer\ (\daSecScoreWorstmanufacturerScore\ / 10).
 
@@
 Computing how good a particular manufacturer or device model is from a security standpoint is difficult as it depends on a number of factors which are hard to observe, particularly on a large scale.
-Ideally we would consider both prevelance of potential problems which were not exploited and actual security failures.
+Ideally we would consider both prevalence of potential problems which were not exploited and actual security failures.\footnote{A perfectly secure operating system would among other things detect and prevent all phishing attacks, provide perfect principle of least privilege isolation, not have any vulnerabilities, instantly fix all discovered vulnerabilities, not allow any user data to be used in a way which the user does not approve of and be really easy for an ordinary person to use.}
 However in the absence of such data we propose a scheme for assigning a device a score out of ten based on data which can be observed, and which hopefully correlates with the actual security of the devices.
@@
 \begin{description}
-  \item[$f$] The proportion of the time where devices were free from known critical vulnerabilities.
+  \item[$f$] The proportion of the time where devices were $f$ree from known critical vulnerabilities.
 This is equivalent to Acer and Jackson's proposal~\cite{Acer2010} to measure the security based on the proportion of users with at least one unpatched critical vulnerability and similar to the Vulnerability Free Days (VFD) score~\cite{Wright2014}.
@@
   \item[$u$] The proportion of devices which run the latest version of Android shipped to any device produced by that manufacturer.
-This is a measure of internal updatedness, a low score would mean many devices are being left behind.
+This is a measure of internal $u$pdatedness, a low score would mean many devices are being left behind.
   This assumes that newer versions are better with stronger security, historically steps have been taken to improve Android security in newer versions so this assumption should generally hold, but sometimes new updates introduce new vulnerabilities.
-  \item[$m$] The mean number of outstanding vulnerabilities affecting devices not fixed on any device shipped by the manufacturer.
+  \item[$m$] The $m$ean number of outstanding vulnerabilities affecting devices not fixed on any device shipped by the manufacturer.
 This is related to the Median Active Vulnerabilities (MAV) measure~\cite{Wright2014} but is the mean rather than the median, as this gives a continuous value.
@@
 However there are a wide variety of scores depending on the source of the device.
-There have been many reports that Google's Nexus devices are better at getting updates than other Android devices because Google makes the original updates and ships them to its devices.
+There have been many reports\todo{citation} that Google's Nexus devices are better at getting updates than other Android devices because Google makes the original updates and ships them to its devices.
 Table~\ref{tab:sec_summary} shows that this is the case with Nexus devices getting much better scores than non-Nexus devices.

### 2014-10-24 securityupdates.tex
Add an open question
More fixes and headings

@@
 \todo{In an experiment you test a hypothesis, it needs to be clear what the hypothesis is and what the conclusion of the test is}
+
 \subsection{Experiment 1: behaviour of the Android ecosystem}\label{sec:exp:android_ecosystem}
+
+\subsubsection{Method}
 To investigate the behaviour of the Android ecosystem as a whole with respect to Android versions and vulnerabilities we used the OS version information from \da\ and the vulnerability data from \avo.
@@
 The \avo\ data was collected between \avoStartDate\ and \avoEndDate\ and covers the period from \avoFirstDataDate\ to \avoLastDataDate.
+We verify the validity of the \da\ data in \S\ref{sec:representative}.
+
+\subsubsection{Results}
 The proportion of devices in the \da\ data running different versions of Android each day is shown in Figure~\ref{fig:norm_os}.
 It shows how old versions are gradually replaced by new ones, the long tail of devices which do not see updates to more recent versions and the fluctuation in the number of devices in the data.
-We verify the validity of this data in \S\ref{sec:representative}.
+
 \begin{figure}
@@
 Those graphs summarise data across all the devices, however one of the advantages of the \da\ data is that it allows us to look at what happens to individual devices over time.
+\subsubsection{Method}
+\todo{write a method}
+\subsubsection{Results}
 Figure~\ref{fig:device_data} shows how the number of vulnerabilities affecting the \daNumDeviceDataDevices\ devices which have contributed the most days of data to \da\ changes over time.
@@
  \includegraphics[width=\columnwidth]{figures/from_to_updates.pdf}
- \caption{Updates between different Android versions in the \da data}
+ \caption{Updates between different Android versions in the \da\ data}
  \label{fig:from_to_updates}
@@
 Figure~\ref{fig:from_to_updates} shows how devices upgrade between different versions of Android.
-\da\ (DA) observes upgrade events and the figure shows which versions a device is upgrading from and to.
+\da\ observes upgrade events and the figure shows which versions a device is upgrading from and to.
 Mostly the dark cells are above the diagonal and are upgrades (\daNumUpdatesUpgrades).
@@
 \subsection{Experiment 4: Predicting future version distributions}\label{sec:exp:predicting_distributions}
+\subsubsection{Method}
 Examining the way that the OS version distribution changes over time in Figure~\ref{fig:norm_os} we see some regularity to the behaviour.
-To show this behaviour more clearly we normalised the data to the number of days since the version was released and is smoothed by forcing it to be monotonically decreasing.\footnote{While sometimes an older version becomes more popular after a later version has been released causing the values to actually increase, mostly this was just noise or changes in the distribution of devices (e.g.\ the sudden arrival of 1000 devices from Bangladesh due to a study being conducted there) and this assumption makes the graph clearer.}
+To show this behaviour more clearly we normalised the data to the number of days since the version was released and smoothed it by forcing it to be monotonically decreasing.\footnote{While sometimes an older version becomes more popular after a later version has been released causing the values to actually increase, mostly this was just noise or changes in the distribution of devices (e.g.\ the sudden arrival of 1000 devices from Bangladesh due to a study being conducted there) and this assumption makes the graph clearer.}
 Then we can plot Figure~\ref{fig:frvh_os_versions} which shows the proportion of of devices in \da\ not upgraded to each Android version or a later Android version (i.e.
 for 2.1.0 it includes 2.1.1, 2.2.0 and indeed all devices as they all run later versions of Android).
+The mean and standard deviation of the same data are shown in Figure~\ref{fig:frms_os_versions}.
 
@@
 \end{figure}
-There is a general pattern to these data, Figure~\ref{fig:frms_os_versions} shows the mean and standard deviation of the same data.
-It also shows the curve of best fit, which has a root mean squared error (RMSE) of \daOSCurveFitRMSE, the curve we chose is not a standard polynomial fit (3 degree polynomial fit gave a RMSE of \daOSCurvePolyRMSE) or a spline fit (RMSE of \daOSCurveSplineRMSE) as we wanted a meaningful model of behaviour rather than a generic curve.
 
@@
 
+\subsubsection{Results}
 Fitting that curve to the data using \texttt{scipy.optimize}, we get a RMSE of \daOSCurveFitRMSE\ and parameters $\mathrm{undeployed} = \daOSCurveFitParamFirst ~\si{days}$, $\mathrm{decay} = \daOSCurveFitParamSecond ~\si{days^{-1}}$ for OS versions and RMSE of \daAPICurveFitRMSE\, $\mathrm{undeployed} = \daAPICurveFitParamFirst ~\si{days}$, $\mathrm{decay} = \daAPICurveFitParamSecond ~\si{days^{-1}}$ for API versions.
@@
 Hence if a security vulnerability is fixed through the release of a particular Android version it will be \daOSCurveFullDeployed\ days after until the fix is fully deployed.
+
+The equation we chose for the model has a root mean squared error (RMSE) of \daOSCurveFitRMSE for OS versions, which is a comparably good fit to a standard polynomial fit (3 degree polynomial fit gave a RMSE of \daOSCurvePolyRMSE) or a spline fit (RMSE of \daOSCurveSplineRMSE) but gives a meaningful model of behaviour rather than a generic curve.
 
@@
  \item Why do so many devices have ADB enabled?
+ \item How much of updating is new handsets and how much is updates being deployed?\todo{we should answer this question in this paper}
+ \item Why do users downgrade their phones?
 \end{itemize}

### 2014-10-24 security_scoring.tex
Use todolater for things we don't need to get done before submission and hence hide them from view

@@
 However there are a wide variety of scores depending on the source of the device.
-There have been many reports\todo{citation} that Google's Nexus devices are better at getting updates than other Android devices because Google makes the original updates and ships them to its devices.
+There have been many reports\todolater{citation} that Google's Nexus devices are better at getting updates than other Android devices because Google makes the original updates and ships them to its devices.
 Table~\ref{tab:sec_summary} shows that this is the case with Nexus devices getting much better scores than non-Nexus devices.

### 2014-10-24 securityupdates.tex
Use todolater for things we don't need to get done before submission and hence hide them from view

@@
 \renewcommand{\todo}{\OldTodo[inline]}
+\newcommand{\todolater}{}%\todo
 \RequirePackage[date=terse, isbn=true, doi=true, url=false, urldate=iso8601, maxbibnames=9, backref=false, backend=bibtex]{biblatex}
@@
 Operators are cautious because if they ship a broken update then they will get many support requests which is expensive.
-However there is ongoing legal action to force operators to ship updates for security vulnerabilities~\cite{Soghoian2013}.\todo{Check on the status of this legal action}
-Many phones are sold on long (12--24 month) contracts with monthly payments yet stop receiving updates before the end of the contract,\todo{Is there a survey of the distribution of contract lengths and phone purchase methods?} while Windows XP could be purchased for a one off payment in October 2001 and received security updates until April 2014.
+However there is ongoing legal action to force operators to ship updates for security vulnerabilities~\cite{Soghoian2013}.\todolater{Check on the status of this legal action}
+Many phones are sold on long (12--24 month) contracts with monthly payments yet stop receiving updates before the end of the contract,\todolater{Is there a survey of the distribution of contract lengths and phone purchase methods?} while Windows XP could be purchased for a one off payment in October 2001 and received security updates until April 2014.
 Corporate or public sector buyers are encouraged to purchase secure devices, but we have found little concrete guidance.
-For example, CESG, which advises the UK government on how to secure its computer systems, recommends picking Android devices from manufactures which are good at shipping security updates promptly~\cite{CESG2013} but it does not state which manufacturers are better.\todo{Which other advisory organisations give that advice?}
+For example, CESG, which advises the UK government on how to secure its computer systems, recommends picking Android devices from manufactures which are good at shipping security updates promptly~\cite{CESG2013} but it does not state which manufacturers are better.\todolater{Which other advisory organisations give that advice?}
 We provide this information in~\S\ref{sec:security_scoring}.
@@
 The API version is a well defined integer, however it does not always change with new Android releases, particularly security bug fixes should not increase this value.
-The build string is `The user-visible version string', fortunately most (99.9\%)\todo{use a generated number} devices in the data have a build string of the form `x.y.z random string' and so it is possible to extract the Android version number.
+The build string is `The user-visible version string', fortunately most (99.9\%)\todolater{use a generated number} devices in the data have a build string of the form `x.y.z random string' and so it is possible to extract the Android version number.
 On a large proportion of devices `random string' is a well defined build number\footnote{\url{https://source.android.com/source/build-numbers.html}} but different manufacturers use different schema.
@@
 \caption{Flow of updates from upstream projects into Android.
-Number of updates as in Figure~\ref{fig:update_ecosystem}, latency in days for all pairs of versions we have data on.\todo{scrape the other 26 websites...
+Number of updates as in Figure~\ref{fig:update_ecosystem}, latency in days for all pairs of versions we have data on.\todolater{scrape the other 26 websites...
 is it worth it?}}
@@
 Finding vulnerabilities is frequently a time consuming manual process but, Brumley et al.~\cite{Brumley2008} showed that it is possible to automatically generate exploits from binary patches.
-In a similar way some of the vulnerabilities which have been exploited were publicly found by examining the commit made to the Android which fixed the vulnerability.
+In a similar way some of the vulnerabilities\todolater{which?} which have been exploited were publicly found by examining the commit made to the Android which fixed the vulnerability.
 This effect has been observed before~\cite{Barth2011} in the Firefox source code repositories, Google does not release the source code until after the release of the update reducing this effect to the delay between an update being released and it reaching the last end user device.
@@
 Fortunately Android does authenticate the update binaries (though the APK vulnerabilities circumvented this) and Google Play downloads them over a secure connection~\cite{Viennot2014}.
-\todo{double check that}
+\todolater{double check that}
 
 The updatedness rate for Android of \daUpdatednessPerc\ compares unfavourable with the more than 90\% rate for Windows XP SP2 computers contacting the Microsoft update servers~\cite{Gkantsidis2006} though that is within one OS version and only for those computers which did connect to the servers -- but that is the default.
-However 27\% of Windows computers were still running Windows XP in July 2014\footnote{\url{https://archive.today/PLGxn}} -- four months after it went out of security support.
+However 27\% of Windows computers were still running Windows XP in July 2014\footnote{\url{https://archive.today/PLGxn}\todolater{Get updated figures before publication}} -- four months after it went out of security support.
 

### 2014-10-24 security_scoring.tex
refactor securityscores to move all the method into the method

@@
 
+We can compute the uncertainty for $f$ and $u$.
+$f$ is calculated by taking the total secure device days and dividing it by the total insecure and secure device days.
+The total secure device days and total insecure device days are both counting experiments and so their measurement error is their square root~\cite{Taylor1997}.
+Since the numbers involved are large, the uncertainty in $f$ is small.
+$u$ is computed by taking the sum of the proportions of devices running the most recent version each day, both the count of devices running the maximum version and total count have sqrt uncertainties.
+The Python \texttt{uncertainties} library was used to propagate uncertainties through calculations.
+
+\todo{There must be some way of putting an error bound on our results by considering the error on the whole DA set and the size of the subset used for each manufacturer and model}
+
 \subsubsection{Results: Security scores}
@@
 Even within manufacturers different models can have very different update behaviours and hence security.
-Table~\ref{tab:sec_model} shows the results for the \daNumSigModels\ device models which have a significant presence by the same metric with \daSecScoreBestmodel\ (\daSecScoreBestmodelScore\ / 10) doing best and \daSecScoreWorstmodel\ (\daSecScoreWorstmodelScore\ / 10) doing worst.
+Table~\ref{tab:sec_model} shows the results for the \daNumSigModels\ device models which have a significant presence by the same metric with `\daSecScoreBestmodel' (\daSecScoreBestmodelScore\ / 10) doing best and `\daSecScoreWorstmodel' (\daSecScoreWorstmodelScore\ / 10) doing worst.
 We can then test whether this seems fair by comparing the version data for the best and worst models.
-Figure~\ref{fig:worst_model_full} shows the full version distribution for \daSecScoreWorstmodel\ and Figure~\ref{fig:best_model_full} shows the same for \daSecScoreBestmodel\ there is a marked difference in the distribution of updates between the two with only \daSecScoreWorstmodelNumFullVersions\ full versions for \daSecScoreWorstmodel\ and \daSecScoreBestmodelNumFullVersions\ for \daSecScoreBestmodel.
+Figure~\ref{fig:worst_model_full} shows the full version distribution for `\daSecScoreWorstmodel' and Figure~\ref{fig:best_model_full} shows the same for `\daSecScoreBestmodel' there is a marked difference in the distribution of updates between the two with only \daSecScoreWorstmodelNumFullVersions\ full versions for `\daSecScoreWorstmodel' and \daSecScoreBestmodelNumFullVersions\ for `\daSecScoreBestmodel'.
 
@@
 \daTabSecScoresoperator
-We also analysed different operators, for the \daNumSigOperators\ operators with a significant presence in our data Table~\ref{tab:sec_operator} shows the results \daSecScoreBestoperator\ (\daSecScoreBestoperatorScore\ / 10) doing best and \daSecScoreWorstoperator\ (\daSecScoreWorstoperatorScore\ / 10) doing worst.
+We also analysed different operators, for the \daNumSigOperators\ operators with a significant presence in our data Table~\ref{tab:sec_operator} shows the results `\daSecScoreBestoperator' (\daSecScoreBestoperatorScore\ / 10) doing best and `\daSecScoreWorstoperator' (\daSecScoreWorstoperatorScore\ / 10) doing worst.
 However the score of an operator is affected by the manufacturers of the devices which are in use on it's network.
@@
 However an operator could use data from this paper to exclude insecure devices from those offered to consumers.
-
-These tables also show the uncertainty for $f$ and $u$.
-$f$ is calculated by taking the total secure device days and dividing it by the total insecure and secure device days.
-The total secure device days and total insecure device days are both counting experiments and so their measurement error is their square root~\cite{Taylor1997}.
-Since the numbers involved are large, the uncertainty in $f$ is small.
-$u$ is computed by taking the sum of the proportions of devices running the most recent version each day, both the count of devices running the maximum version and total count have sqrt uncertainties.
-The Python \texttt{uncertainties} library was used to propagate uncertainties through calculations.
-
-\todo{There must be some way of putting an error bound on our results by considering the error on the whole DA set and the size of the subset used for each manufacturer and model}

### 2014-10-24 securityupdates.tex
More minor fixes

@@
  \includegraphics[width=\columnwidth]{figures/frvh_os_versions}
- \caption{Proportion of devices not updated to particular versions of Android or any later version\todo{make line of best fit thicker}}
+ \caption{Proportion of devices not updated to particular versions of Android or any later version\todo{make line of best fit thicker, I think the drops to 0 at the end of each version are erroneous}}
  \label{fig:frvh_os_versions}
@@
 
-The equation we chose for the model has a root mean squared error (RMSE) of \daOSCurveFitRMSE for OS versions, which is a comparably good fit to a standard polynomial fit (3 degree polynomial fit gave a RMSE of \daOSCurvePolyRMSE) or a spline fit (RMSE of \daOSCurveSplineRMSE) but gives a meaningful model of behaviour rather than a generic curve.
+The equation we chose for the model has a root mean squared error (RMSE) of \daOSCurveFitRMSE\ for OS versions, which is a comparably good fit to a standard polynomial fit (3 degree polynomial fit gave a RMSE of \daOSCurvePolyRMSE) or a spline fit (RMSE of \daOSCurveSplineRMSE) but gives a meaningful model of behaviour rather than a generic curve.
+
+\todo{Evaluate our model trained on the DA data against the ground truth Google Play data}
 
 \todo{Our handling of errors here is not quite good enough as the uncertainty from our y value interacts badly with the logarithm function when computing the inverse and we compute the log of ~0 which means our error ends up much bigger than our value - but that is an upper bound, not a +- bound.}
+
+We hope that a consequence of this paper will be a change in behaviour of the Android ecosystem which invalidates this model by speeding up the deployment of new OS versions.
 
@@
 
+\subsection{Uncertainty propagation}
+
 \section{Discussion}
@@
 Various attempts have been made to detect malware.
-RiskRanker~\cite{Grace2012a} classified 3,281/118,318 (2.8\%) apps as risky of which 718 (22\%) were malware and 322 (10\%) were previously unknown, an infection rate of 0.6\% across multiple markets.
+RiskRanker~\cite{Grace2012a} classified 3,281/118,318 (2.8\%) apps as risky of which 718 (22\%) were malware and 322 (10\%) were previously unknown malware, an infection rate of 0.6\% across multiple markets.
 It uses signatures of exploits, static analysis of calls to high risk operations, detection of encrypted native code execution and dynamic Dalvik code loading to detect risky apps.

### 2014-10-27 securityupdates.tex
Discuss DA a bit more

@@
 The \da\ data contains data from \daNumDevices\ devices with a total of \daDeviceDays\footnote{Here we are only counting devices and days for which we have valid OS version data} device days, it contains data for more than \daMonths\ months for \daMonthsDevices\ devices.
-\todo{ talk about the \da\ project a bit}
+Device Analyzer used an Android app which volunteers install on their Android devices to collect data on what happens on Android phones.
+The data collected is available for researchers to use as we do here.
+
 Various different kinds of data\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/collected.htm}} are collected by the \da\ Android app\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/}}.

### 2014-10-27 securityupdates.tex
Fill out 'Lifetime of a vulnerability' subsection

@@
 \subsection{Lifetime of a vulnerability}
+
+The key events in the lifetime of a vulnerability are:
+
+\begin{LaTeXdescription}
+ \item[created] When a vulnerability was created in the source code.
+ \item[introducing release] When the first release was made containing the vulnerability.
+ \item[discovery] When the vulnerability is first discovered
+ \item[exploit] When the vulnerability is first exploited
+ \item[disclosure] When the vulnerability is first disclosed % publically or to a smaller set of people?
+ \item[fix] When a vulnerability was first fixed in the source code.
+ \item[fixing release] When the first release containing the fix was made (equivalent to `patch available'~\cite{Frei2010}).
+ \item[fix deployed] When the fix has been deployed to a sufficiently large proportion of the population that the vulnerability can be ignored (the ecosystem equivalent to the per instance `patch installed'~\cite{Frei2010}).
+\end{LaTeXdescription}
+
 Working out the date to use for the date on which a vulnerability was sufficiently well known to pose a threat to users is difficult.
@@
 Hence we have collated this information and made it publicly available.
-
-The key events in the lifetime of a vulnerability are:
-
-\begin{LaTeXdescription}
- \item[introduction] When a vulnerability was introduced into the source code.
- \item[introducing release] When the first release was made containing the vulnerability.
- \item[fix] When a vulnerability was first fixed in the source code.
- \item[fixing release] When the first release containing the fix was made.
- \item[fix deployed] When the fix has been deployed to a sufficiently large proportion of the population that the vulnerability can be ignored.
- \item[discovery] When the vulnerability is first discovered
- \item[exploit] When the vulnerability is first exploited
- \item[disclosed] When the vulnerability is first disclosed % publically or to a smaller set of people?
-\end{LaTeXdescription}
-
-\todo{Compare this list with the literature and unify it. Pull out figures which illustrate and investigate this in our data.}
 
@@
 We cannot know if someone reported the vulnerability to the manufacturer and we do know that various agencies are engaged in widespread monitoring of communications and compromise of civilian infrastructure~\cite{TODO} and so they probably know.
-Hence we use dates such as: when CVEs were registered, when fixes were committed to source repositories, or when exploits were committed to repositories; as the date first known about.\todo{We need to be more precise here}
+
+There are several points we could pick for the point from which a vulnerability poses substantial risk to users.
+{\bf Creation} and {\bf introduction} are the earliest points which could be used but the risk is mostly latent until someone discovers them.
+Unfortunately {\bf discovery} is the hardest point to obtain concrete data on as discovery may happen multiple times independently and not all discoverers will report their discovery, for example APTs may keep databases of unknown vulnerabilities to use when they have a particular target, in which case there is a high risk to potential targets from the never disclosed point of discovery.
+However if the discoverer never discloses the vulnerability to anyone or makes any use of it then there is little risk.
+The date of first {\bf exploit} is a point when the risk is definitely high, but again a good adversary will not be detected when using such exploits.
+The date of first {\bf fix} is, assuming that the fix is deliberate, a point at which the vulnerability is known at least within the organisation performing the fix and frequently implies an earlier discovery and notification by a third party.
+Once a {\bf fixing release} has been made then the vulnerability is widely known because it can be reverse engineered from the changes in the release~\cite{TODO}.
+When the {\bf fix is deployed} to a sufficiently large proportion of devices then the remaining risk is minimal.
+
+Hence we use the earliest date we know about for discovery, exploit, disclosure or fix as the date from which a vulnerability poses a substantial risk to users.
 

### 2014-10-27 securityupdates.tex
Fix definition of todolater so that it doesn't leave the todo in the main text
Tweak ecosystem text

@@
 \renewcommand{\todo}{\OldTodo[inline]}
-\newcommand{\todolater}{}%\todo
+\newcommand{\todolater}[1]{}%\todo
 \RequirePackage[date=terse, isbn=true, doi=true, url=false, urldate=iso8601, maxbibnames=9, backref=false, backend=bibtex]{biblatex}
@@
 Android also incorporates various drivers for different bits of hardware.
-The Android platform is then built on top of those with kernel and userspace components by Google. \todo{ work out the exact arrangements.}
-Usually the code for each update is kept secret\todo{Can we quantify this keeping the code secret? Is it worth it?} until after the update has been published though through the Open Handset Alliance.
-Manufacturers in the  alliance receive advanced copies in order to prepare handsets.
+The Android platform is then built on top of those with kernel and userspace components by Google.
+The code for each update is kept secret\footnote{\url{https://source.android.com/source/code-lines.html}}\todolater{Can we quantify this keeping the code secret?
+Is it worth it?} until after the update has been published.
+Manufacturers may receive advanced copies in order to prepare handsets.
 Then the Manufacturers take the update and customise it before passing it on to the operator.

### 2014-10-27 securityupdates.tex
Update ecosystem diagram

@@
 \input{countversions}
-\newcommand{\apacheLibsNumVersions}{C}
-\newcommand{\apacheLibsNumLibs}{E}
+\newcommand{\otherProjNumVersions}{XXX}
+\newcommand{\otherProjNum}{\avoNumExternalProjects}%TODO check we are doing the right calculation here and not overcounting
 
@@
  Numbers on edges indicate updates shipped between \daStartDate\ and \daEndDate, numbers in brackets represent number of such entities in our data.
- Dotted arrows indicate flows where we do not know how many updates are being produced as we can't measure those flows directly as they are not public.\todo{Update this diagram}}
+ Dotted arrows indicate flows where we do not know how many updates are being produced as we can't measure those flows directly as they are not public.}
  \label{fig:update_ecosystem}

### 2014-10-27 securityupdates.tex
Update android update process

@@
 \section{Android Ecosystem}
+\todo{Introduction}
 \subsection{Running a version of Android known to be vulnerable makes a device vulnerable}
@@
 \subsection{Android update process}
-\todo{rewrite this}
+
 \label{sec:android_update_process}
@@
 \end{figure}
-
-A model of the Android update process is shown in Figure~\ref{fig:update_ecosystem}.
+To understand how vulnerabilities in Android are fixed we must examine the Android update process which we model in Figure~\ref{fig:update_ecosystem}.
 There are five entities or groups which contribute towards Android updates: the network operators, the device manufacturers, the hardware developers, Google and the upstream open source projects.
-Android builds on various open source projects such as the Linux kernel, the OpenSSL and BouncyCastle cryptography libraries and so can include any of the versions of those projects, including those which fix security vulnerabilities in them.
+Android builds on various open source projects such as the Linux kernel, the OpenSSL and BouncyCastle cryptography libraries and so can include any compatible versions of those projects, including those which fix security vulnerabilities in them.
 Android also incorporates various drivers for different bits of hardware.
@@
 %An analysis by Vidas et al.~\cite{Vidas2011} of the Android 2.1 to 2.2 update found that it took 11 months from when Google released 2.2 for the last device which they were investigating to get the update.
-\todo{talk about update latency, quantify this in the text as per the figure, giving more explanation, use statistics from samsung-updates.com -> how many binaries are there per device?
-Do we know the variation between network operators?}
+\todolater{use statistics from samsung-updates.com -> how many binaries are there per device?}
 
@@
 The number of updates shipped by Google (\daNumSigOSVersions) is the number of Android versions reported in \da\ which affected more than \daSigVersionPerc\ of devices for more than \daSigVersionDays\ days.
-\todo{this is the wrong number to be using}
+\todolater{this is the wrong number to be using}
 
@@
 This allows us to compute the latency between an upstream project being released and it being included in Android, this is shown in Table~\ref{tab:update_ecosystem}.
+This shows that the versions included in Android were about half a year old when that version of Android was released.
 \begin{table}

### 2014-10-27 securityupdates.tex
Add introduction to android ecosystem and fix up monetisation section

@@
 \section{Android Ecosystem}
-\todo{Introduction}
+There is a complex Android ecosystem which creates and distributes updates to Android which fix vulnerabilities.
+It is also hard to quantify due to lack of public information on aspects of the ecosystem such as the number of manufacturers and updates.
+We describe and quantify this ecosystem.
+
 \subsection{Running a version of Android known to be vulnerable makes a device vulnerable}
@@
 
-\subsection{Monetisation}
-\todo{redo this section in light of related work and to fit into ecosystem discussion}
-
-One of the key differences between smartphones and desktops is that it is easier to directly cash out on a smartphone as malware can send premium rate texts or make premium rate phone calls.
-It can also steal personal data which already has semantic information associated with it, this also makes it easier to perform ransomware attacks where malware deletes this information from the device and charges a ransom for its return.
-It can also do more conventional things such as use the devices for DDoS attacks and altering the ads displayed to send the fees to the malware author.
+\subsection{Monetisation of malware}
+Android malware can make money in a variety of different ways~\cite{Felt2011}.
+One of the key differences between smartphones and desktops is that it is easier to directly cash out on a smartphone as malware can send premium rate texts or make premium rate phone calls or send SMS and email spam.
+It can also steal personal data which already has semantic information associated with it, this also makes it easier to sell and to perform ransomware attacks where malware deletes this information from the device and charges a ransom for its return.
+It can steal credentials for external services or use in-app billing.
+It can also do more conventional things such as use the devices for DDoS, SEO or click fraud attacks and altering the ads displayed to send the fees to the malware author.
 While many of these things can be done by malware which just requests the relevant permissions, root exploits make this easier, particularly for ransomware and altering the ads displayed.
@@
 \label{sec:economics}
-\cite{Felt2011}
+\todo{write this}
 

### 2014-10-27 securityupdates.tex
Write a subsection on incentives

@@
 \label{sec:economics}
-\todo{write this}
+Google and other Android vendors care about the reputation of Android, particularly from a security standpoint as if users are afraid to use their devices for certain purposes (e.g.\ banking) then they will not be willing to pay as much for them and money cannot be made from those activities.
+Manufacturers make money by selling the device, once sold they want to prevent the cost of return under warranty and so are adverse to the risk of breaking the device, but they do not need to do things to make the device better as it has already been sold.
+However if they gain a reputation for abandoning devices that might decrease future sales.
+
+Operators sell phones under contracts, typically at least a year long, they want the user to not request support due to the device breaking.
+They might also gain advantage if the user is happy with the device after the end of the contract as they might then keep paying the same rate but there would no longer be a capital cost to pay off.
+They care about phones not getting malware which causes support requests or uses the mobile network to perform malicious actions as they may have to reimburse the cost of premium SMS's if they were not authorised by the user.
+
+Users want to buy phones which will serve their use cases and which are secure so that they do not experience negative consequences.
 

### 2014-10-27 securityupdates.tex
Minor tweaks

@@
 %TODO make this graph shorter
-\caption{Proportion of devices running insecure versions of Android against time}
+\caption{Proportion of devices running insecure versions of Android against time\todo{Disambiguate 'definitely broken' from 'maybe broken'}}
 \label{fig:proportioninsecure}
@@
 The date of first {\bf fix} is, assuming that the fix is deliberate, a point at which the vulnerability is known at least within the organisation performing the fix and frequently implies an earlier discovery and notification by a third party.
-Once a {\bf fixing release} has been made then the vulnerability is widely known because it can be reverse engineered from the changes in the release~\cite{TODO}.
+Once a {\bf fixing release} has been made then the vulnerability is widely known because it can be reverse engineered from the changes in the release~\cite{Brumley2008}.
 When the {\bf fix is deployed} to a sufficiently large proportion of devices then the remaining risk is minimal.
@@
 Sometimes Manufactures ship direct to the user, sometimes the Manufacturer and Google are the same entity (at least for the purpose of a particular phone, such as with Nexus phones).
-Sometimes Manufacturers incorporate upstream open source project releases directly, and sometimes incorrectly -- for example including a broken daily build of sqlite in their release of Android
-%An analysis by Vidas et al.~\cite{Vidas2011} of the Android 2.1 to 2.2 update found that it took 11 months from when Google released 2.2 for the last device which they were investigating to get the update.
+Sometimes Manufacturers incorporate upstream open source project releases directly, and sometimes incorrectly -- for example including a broken daily build of sqlite in their release of Android.
 \todolater{use statistics from samsung-updates.com -> how many binaries are there per device?}
@@
 
+%An analysis by Vidas et al.~\cite{Vidas2011} of the Android 2.1 to 2.2 update found that it took 11 months from when Google released 2.2 for the last device which they were investigating to get the update.
 To compute the latency between upstream releases and their inclusion in Android we scraped the release pages for those projects, to obtain the version numbers and release dates.

### 2014-10-27 securityupdates.tex
Refresh experiment 1 and start on experiment 2

@@
 On some versions of Android (below 4.1) they could use known system to root escalation mechanisms but on all versions they have a greatly increased attack area for further privilege escalation and also have the ability to control all user internet traffic (via VPNs), brick the phone, remove and install apps, steal user credentials, read the screen and make as well as receive calls.
-
+\todo{break vulnerabilities down by attack vector?}
 \avoTabAndVulns
@@
 \label{sec:results}
-In this section we present the results of our analysis showing \daMeanInsecurityPerc\ of devices to have root equivalent vulnerabilities on average.
-Nexus devices are better than others, \emph{\daSecScoreBestmanufacturer} to be the best manufacturer and \emph{\daSecScoreBestoperator} to be the best network operator.
-\todo{ summarise our results}
+We present the results of our analysis showing \daMeanInsecurityPerc\ of devices to have root equivalent vulnerabilities on average.
+Nexus devices are better than others, \emph{\daSecScoreBestmanufacturer} to be the best manufacturer and \emph{\daSecScoreBestoperator} to be the best network operator and modelling the uptake of new versions of Android.
+\todo{summarise experiment 2}
 %\todo{comparison of nexus update rates with iOS}
 We conducted four experiments examining the behaviour of the Android ecosystem as a whole (\S\ref{sec:exp:android_ecosystem}), updates reaching particular devices (\S\ref{sec:exp:device_updates}), comparisons between different manufacturers and operators (\S\ref{sec:exp:security_score}) and model how long it will take for an update to be fixed once an update is released (\S\ref{sec:exp:predicting_distributions}).
-\todo{In an experiment you test a hypothesis, it needs to be clear what the hypothesis is and what the conclusion of the test is}
+
+%% In an experiment you test a hypothesis, it needs to be clear what the hypothesis is and what the conclusion of the test is
 
@@
 We verify the validity of the \da\ data in \S\ref{sec:representative}.
+For each device, for every day with version data we record what version it was running and which vulnerabilities it was exposed to at that time (if any).
+We then normalise these totals for each day by the total number of devices with version information seen on that day.
 
@@
 The proportion of devices in the \da\ data running different versions of Android each day is shown in Figure~\ref{fig:norm_os}.
-It shows how old versions are gradually replaced by new ones, the long tail of devices which do not see updates to more recent versions and the fluctuation in the number of devices in the data.
+It shows how old versions are gradually replaced by new ones, the long tail of devices which do not see updates to more recent versions.
 
@@
 
-Figure~\ref{fig:vulnerabilities} shows which vulnerabilities devices are exposed to.
+The vulnerabilities devices are exposed to are shown in Figure~\ref{fig:nvulnerabilities_heat}.
 For each vulnerability it shows the proportion of devices exposed to that vulnerability and how that changes over time.
-In July 2011 at the beginning of the \da\ data the \emph{exploid}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/exploid_udev}} and \emph{levitator}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/levitator}} vulnerabilities both affect most Android devices, slowly these are fixed as updates roll out and devices are replaced until in January 2013 a much smaller proportion of devices are affected by known vulnerabilities.\todo{Quantify whether it is replacement of updates which causes the fixes}
+The variation of the proportion of devices affected by a vulnerability with time tells us how bad a particular vulnerability affected the Android platform.
+In July 2011 at the beginning of the \da\ data the \emph{exploid}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/exploid_udev}} and \emph{levitator}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/levitator}} vulnerabilities both affect most Android devices, slowly these are fixed as updates roll out and devices are replaced until in January 2013 a much smaller proportion of devices are affected by known vulnerabilities.\todo{Quantify whether it is replacement or updates which causes the fixes}
 However when in February 2013 the first APK signing vulnerability was found which affected all previous versions of Android and even in October 2013 most devices were still vulnerable.
 \todo{put percentages on these things to quantify them}
-\begin{figure}%[!b]
-\centering
-\includegraphics[width=\columnwidth]{figures/vulnerabilities}
-\caption{Proportion of devices exposed to each vulnerability with time (1.0 is 100\%)}
-\label{fig:vulnerabilities}
-\end{figure}
+%\begin{figure}%[!b]
+%\centering
+%\includegraphics[width=\columnwidth]{figures/vulnerabilities}
+%\caption{Proportion of devices exposed to each vulnerability with time (1.0 is 100\%)}
+%\label{fig:vulnerabilities}
+%\end{figure}
 
-The variation of the proportion of devices affected by a vulnerability with time tells us how bad a particular vulnerability affected the Android platform.
-Figure~\ref{fig:nvulnerabilities_heat} breaks this down by vulnerability to show how the proportion of devices affected by different vulnerabilities varies.
 In 2013 three vulnerabilities were found in the way which Android verified the signatures on APKs.
@@
 Figure~\ref{fig:nvulnerabilities_heat} shows how the the \emph{APK signing vulnerabilities} affected all devices and took months to get fixed for any device.
-However what is perhaps more worrying is the long tail on the \emph{Gingerbreak}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/Gingerbreak}}, \emph{levitator} and \emph{exploid} vulnerabilities which are more dangerous root vulnerabilities (not requiring new APK installation) and which still affect a significant proportion of devices years later.
-\todo{break things down by attack vector?}
+However what is perhaps more worrying is the long tail on the \emph{Gingerbreak}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/Gingerbreak}}, \emph{levitator}, \emph{exploid} and \emph{zergRush}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/zergRush}} vulnerabilities which are more dangerous root vulnerabilities (not requiring new APK installation) and which still affect a significant proportion of devices years later.
 
@@
 \subsubsection{Method}
+As in \ref{sec:exp:android_ecosystem}, for each device we used the version information from each day to calculate which vulnerabilities it was vulnerable to that day.
+We also recorded when the version changed and which versions it changed from and to.
 \todo{write a method}
@@
 It shows how some devices had vulnerabilities, which were fixed, and then further vulnerabilities were discovered, and for these devices, mostly not fixed.
+This implies that these devices had been abandoned \todo{finish this sentence}
 \begin{figure}
@@
 
-\subsubsection{Update frequency}
 \todo{ find out what the frequency of updates is}

### 2014-10-28 securityupdates.tex
Rewrite experiment 2

@@
 Those graphs summarise data across all the devices, however one of the advantages of the \da\ data is that it allows us to look at what happens to individual devices over time.
+
 \subsubsection{Method}
@@
 We also recorded when the version changed and which versions it changed from and to.
-\todo{write a method}
+Since in the \da\ data many devices only contribute data for a short period, although we have over 1000 devices contributing data in any particular week we do not observe all the upgrades made by a large number of devices, but rather a hopefully representative sample of upgrades which happened when \da\ was installed on the devices.
+\da\ cannot distinguish between a device being replaced and the \da\ app being removed as while multiple devices can be linked to the same user account, few users do that.
+
 \subsubsection{Results}
-Figure~\ref{fig:device_data} shows how the number of vulnerabilities affecting the \daNumDeviceDataDevices\ devices which have contributed the most days of data to \da\ changes over time.
-The trend that we saw in Figure~\ref{fig:proportioninsecure} of security improving and then getting worse is also shown here.
-It shows how some devices had vulnerabilities, which were fixed, and then further vulnerabilities were discovered, and for these devices, mostly not fixed.
-This implies that these devices had been abandoned \todo{finish this sentence}
 \begin{figure}
@@
 \end{figure}
+The logitudinal data on the number of vulnerabilities affecting the \daNumDeviceDataDevices\ devices which have contributed the most days of data to \da\ changes over time is shown in Figure~\ref{fig:device_data}.
+The trend that we saw in Figure~\ref{fig:proportioninsecure} of security improving and then getting worse is also shown here.
+It shows how some devices had vulnerabilities, which were fixed, and then further vulnerabilities were discovered, and for these devices, mostly not fixed.
+This implies that these devices had been abandoned and were not receiving updates.
+Some devices start off in 2011 exposed to known security vulnerabilities and are still exposed to those and additional ones in 2014.
 
 \todo{ find out what the frequency of updates is}
+
 \begin{figure}
@@
 \end{figure}
-Figure~\ref{fig:from_to_updates} shows how devices upgrade between different versions of Android.
-\da\ observes upgrade events and the figure shows which versions a device is upgrading from and to.
+We recorded update events and Figure~\ref{fig:from_to_updates} shows how devices upgrade between different versions of Android.
 Mostly the dark cells are above the diagonal and are upgrades (\daNumUpdatesUpgrades).
-While many upgrades are from one version to the next version there are also a fair number which skip a few versions.
+While many upgrades are from one version to the next version there are also a fair number which skip a few versions.\todo{quantify that}
 Surprisingly there are also a small number of downgrade events (\daNumUpdatesDowngrades, \daPercUpdatesDowngrades) when older versions of Android are installed on to devices.
@@
 Red indicates updates which changed the build number but not the version number and so might contain a backported fix for a vulnerability.
+\todo{Is this normalised? If not it is meaningless...}
 \todo{ summary stats, what does this mean}

### 2014-10-28 securityupdates.tex
Talk about some of the other threats to validity

@@
 \end{figure}
-The data from \da\ (DA) we used to investigate the proportion of devices exposed to different vulnerabilities is the OS version.
+The data from \da\ we used to investigate the proportion of devices exposed to different vulnerabilities is the OS version.
 Unfortunately there is no authoritative source of OS version information and so we cannot check directly whether our data is representative.
@@
 \subsection{Changes in the \da\ sample}
-\todo{Bangladesh etc.}
+The \da\ data is mostly generated by devices which have the \da\ app installed because their owner happened to come across the app on Google Play as there has been no advertising and little other attempts to increase usage.
+However two collaborations with network operators to install \da\ on customer devices resulted in large numbers of new \da\ users coming from particular areas and network operators.
+`n telenor' and `grameenphone' from Table \ref{tab:sec_operator} are big operators in the \da\ data because of their collaboration, in particular `n telenor' conducted a study in Norway with 654 installs of \da\ and `grameenphone' conducted a study in Bangladesh where 2463 users installed \da.
+This is responsible for the sudden change in distribution of OS versions in \da\ shown towards the end of Figures \ref{fig:norm_os}, \ref{fig:nvulnerabilities_heat} and \ref{fig:proportioninsecure} as for that period Bangladeshi users contributed over half of the \da\ data.
+We could normalise out the influence of the Bangladeshi users, however they are Android users too and they also need security updates.
 
 \subsection{Uncertainty propagation}
+We have endeavoured to propagate information on uncertainties in our results through all calculations with the standard deviation indicated by $\pm$.
+However we have not propagated information on systematic errors as we are not certain what those are.
 
@@
 \subsection{Open questions}
+There are several open questions which we have not yet been able to answer and remain future work.
 \begin{itemize}
@@
  \item Why do so many devices have ADB enabled?
- \item How much of updating is new handsets and how much is updates being deployed?\todo{we should answer this question in this paper}
+ \item How much of updating is new handsets and how much is updates being deployed?
  \item Why do users downgrade their phones?

### 2014-10-28 securityupdates.tex
Add a conclusion

@@
 Our hypothesis is that these fixes do not promptly reach the users devices and so much of the time users are running versions of Android known to be vulnerable.
-We used \da\ data from over \daOSYearsOfData\ years and \daNumOSDevices\ devices and found that on average over \daMeanInsecurityPerc\ of devices were exposed to known root privilege vulnerabilities and only \daUpdatednessPerc\ of devices run the most recent version of Android.
+We used \da\ data from over \daOSYearsOfData\ years and \daNumOSDevices\ devices and found that on average \daMeanInsecurityPerc\ of devices were exposed to known root privilege vulnerabilities and only \daUpdatednessPerc\ of devices run the most recent version of Android.
 There was also a period of several months when no devices ran secure versions of Android.
@@
 \label{sec:conclusion}
-\todo{Conclude}
+We have investigate the security of the Android operating system with respect to security updates and examined how the Android ecosystem results in different devices receiving different levels of security due to whether or not they get security updates.
+We have compared different models, manufacturers and network operators and found that there are differences between them which the discerning purchaser might use to influence their decision about which device to buy from which manufacturer and operator.
+We hope that this analysis will encourage manufacturers and operators to improve the support they provide for devices after sale.
+Android attempts to provide good security to end users but the latency in the security update process means that on average \daMeanInsecurityPerc\ of Android devices are exposed to known root equivalent vulnerabilities.
 

### 2014-10-28 securityupdates.tex
Tighten up discussion of iOS security

@@
 
-The security protections used in iOS such as only allowing signed code to run, jails and reviewing of apps has resulted in a lower level of malware affecting iOS~\cite{TODO} but can still be bypassed~\cite{Wang2013a}.
-\todo{look at iOS protections again and tighten this up}
+The security protections used in iOS such, app review, code signing, Data Execution Prevention (DEP) and Address Space Layout Randomisation (ASLR) and mandatory access controls has resulted in a lower level of malware affecting iOS~\cite{Felt2011} but can still be bypassed~\cite{Wang2013a}.
 

### 2014-10-28 securityupdates.tex
Fix graph being used for security updates to be one actually showing that

@@
 \begin{figure}
- \includegraphics[width=\columnwidth]{figures/w_security_updates.pdf}
- \caption{Number of updates each week which fixed or may have fixed security vulnerabilities}
+ \includegraphics[width=\columnwidth]{figures/nw_security_updates.pdf}
+ \caption{Proportion of updates each week which fixed or may have fixed security vulnerabilities\todo{Show non security updates as well for comparison}}
  \label{fig:weekly_security_updates}

### 2014-10-28 figures/update_figures.sh

@@ -20,7 +20,7 @@ cp "$fromdir"norm_versionsecurity-notitle.pdf proportioninsecure-notitle.pdf
 cp "$fromdir"updates_between.pdf updates_between_versions.pdf
 cp "$fromdir"nvulnerabilities.pdf vulnerabilities.pdf
 cp "$fromdir"nvulnerabilities_heat.pdf nvulnerabilities_heat.pdf
-cp "$fromdir"updates_weekly.pdf w_security_updates.pdf
+cp "$fromdir"nsecurity_updates_weekly.pdf nw_security_updates.pdf
 cp "$fromdir"wifi-accounts-ototals-hist.pdf wifi-accounts-ototals-hist.pdf
 cp "$fromdir"wifi-accounts-stotals-hist.pdf wifi-accounts-stotals-hist.pdf
 cp "$fromdir"api_gpcomp_diff.pdf api_gpcomp_diff.pdf

### 2014-10-28 securityupdates.tex
Expand on open questions

@@
 \usepackage{siunitx}
+\usepackage{fp}
 \usepackage{todonotes}
@@
  \item Why was there a gap in vulnerability discovery in Android in 2012?
- \item Why do so many devices have ADB enabled?
+ \item Why do so many devices have ADB enabled? \daAdbEnabledPerc\ seems rather high for a developer feature.
  \item How much of updating is new handsets and how much is updates being deployed?
+Over time newer versions of Android come to dominate, but quantifying what proportion of that is due to new phones being purchased and what proportion is due to updates to existing devices is a question we cannot answer.
+ The only available large data set we know of with longitudinal traces of Android devices OS version numbers is \da\ but that only records \daNumUpdatesUpgrades\ upgrades which broken down over \daOSMonthsOfData\ months and \daNumSigOSVersions\ is only \FPeval{\daUpdatesPerMonthPerVersion}{round(\daNumUpdatesUpgrades/(\daOSMonthsOfData * \daNumSigOSVersions),1)}\daUpdatesPerMonthPerVersion\ updates per version per month which is not enough to build a statistically significant prediction of the expected transition to newer versions of Android.
  \item Why do users downgrade their phones?

### 2014-10-28 securityupdates.tex
Asking about statistical metrics is something we want to do once we have a full draft
Quantify updates a bit more

@@
 The variation of the proportion of devices affected by a vulnerability with time tells us how bad a particular vulnerability affected the Android platform.
-In July 2011 at the beginning of the \da\ data the \emph{exploid}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/exploid_udev}} and \emph{levitator}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/levitator}} vulnerabilities both affect most Android devices, slowly these are fixed as updates roll out and devices are replaced until in January 2013 a much smaller proportion of devices are affected by known vulnerabilities.\todo{Quantify whether it is replacement or updates which causes the fixes}
-However when in February 2013 the first APK signing vulnerability was found which affected all previous versions of Android and even in October 2013 most devices were still vulnerable.
-\todo{put percentages on these things to quantify them}
+In July 2011 at the beginning of the \da\ data the \emph{exploid}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/exploid_udev}} and \emph{levitator}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/levitator}} vulnerabilities both affect most Android devices, slowly these are fixed as updates roll out and devices are replaced until in January 2013 a much smaller proportion of devices are affected by known vulnerabilities.
+However when in February 2013 the first APK signing vulnerability was found which affected all previous versions of Android and even in October 2013 most devices (\daVulnAPKDuplicateFileOctoberPerc) were still vulnerable.
 %\begin{figure}%[!b]
@@
 
-\todo{ find out what the frequency of updates is}
-
 \begin{figure}
@@
 Mostly the dark cells are above the diagonal and are upgrades (\daNumUpdatesUpgrades).
-While many upgrades are from one version to the next version there are also a fair number which skip a few versions.\todo{quantify that}
+While many upgrades are from one version to the next version there are also a fair number (\daNumUpdatesBigUpgrades, \daPercBigUpgrades) which skip a more than \daNumUpdatesSkippedBig\ versions.
 Surprisingly there are also a small number of downgrade events (\daNumUpdatesDowngrades, \daPercUpdatesDowngrades) when older versions of Android are installed on to devices.
@@
 Red indicates updates which changed the build number but not the version number and so might contain a backported fix for a vulnerability.
-\todo{Is this normalised? If not it is meaningless...}
-\todo{ summary stats, what does this mean}
+
+By taking the number of updates observed and the number of device days of data we have collected we can compute the number of updates received by a device per year to be \daUpdatesPerYear.
+This compares badly with the number of vulnerabilities discovered per year of \avoVulnsPerYear.
+
+\todo{That comparison is not completely fair as not all of those vulnerabilites will affect all devices, similarly we may be counting device days for which we only have one day and hence could perhaps not observe an update.}
 
@@
 This means that the OS version information from \da\ is likely to be overestimating the prevalence of new OS versions and hence our results are a conservative estimate of the security of Android.
-\todo{we want a statistical metric to claim this strongly with.}
+\todolater{we want a statistical metric to claim this strongly with.}
 This allows us to have confidence in the OS version information.

### 2014-10-28 securityupdates.tex
summarise experiment 2 in introduction to Experiments

@@
 We present the results of our analysis showing \daMeanInsecurityPerc\ of devices to have root equivalent vulnerabilities on average.
+On average only \daUpdatesPerYear\ updates reach a device each year and \daPercUpdatesDowngrades\ of version changes are downgrades to older versions.
 Nexus devices are better than others, \emph{\daSecScoreBestmanufacturer} to be the best manufacturer and \emph{\daSecScoreBestoperator} to be the best network operator and modelling the uptake of new versions of Android.
-\todo{summarise experiment 2}
 %\todo{comparison of nexus update rates with iOS}
@@
  \includegraphics[width=\columnwidth]{figures/frvh_os_versions}
- \caption{Proportion of devices not updated to particular versions of Android or any later version\todo{make line of best fit thicker, I think the drops to 0 at the end of each version are erroneous}}
+ \caption{Proportion of devices not updated to particular versions of Android or any later version}
  \label{fig:frvh_os_versions}

### 2014-10-28 securityupdates.tex
Show non security updates on the security updates graph for comparison

@@
 The number of devices getting security updates each week, is shown in Figure~\ref{fig:weekly_security_updates}, blue shows updates which changed the Android version number from a version with known vulnerabilities to one which had fewer known vulnerabilities.
-Red indicates updates which changed the build number but not the version number and so might contain a backported fix for a vulnerability.
+Green indicates updates which changed the build number but not the version number and so might contain a backported fix for a vulnerability.
+Red indicates updates which did not fix security vulnerabilities (because there were know known security vulnerabilities in the version of Android they were already running).
 

### 2014-10-29 securityupdates.tex
remove done TODO
include device data os graph for comparison

@@
 \begin{figure}
+ \centering
+ \begin{subfigure}[b]{\columnwidth}
   \includegraphics[width=\columnwidth]{figures/device-data-all-security}
- \caption{Number of vulnerabilities affecting the top \daNumDeviceDataDevices\ devices by days of contribution in the \da\ data}
- \label{fig:device_data}
+  \caption{Number of vulnerabilities affecting each device over time}
+  \label{fig:device_data_security}
+ \end{subfigure}
+ \begin{subfigure}[b]{\columnwidth}
+  \includegraphics[width=\columnwidth]{figures/device-data-all-os}
+  \caption{OS versions over time}
+  \label{fig:device_data_os}
+ \end{subfigure}
+ \caption{The top \daNumDeviceDataDevices\ devices by days of contribution in the \da\ data}
 \end{figure}
-The logitudinal data on the number of vulnerabilities affecting the \daNumDeviceDataDevices\ devices which have contributed the most days of data to \da\ changes over time is shown in Figure~\ref{fig:device_data}.
+The logitudinal data on the number of vulnerabilities affecting the \daNumDeviceDataDevices\ devices which have contributed the most days of data to \da\ changes over time is shown in Figure~\ref{fig:device_data_security}.
 The trend that we saw in Figure~\ref{fig:proportioninsecure} of security improving and then getting worse is also shown here.
 It shows how some devices had vulnerabilities, which were fixed, and then further vulnerabilities were discovered, and for these devices, mostly not fixed.
-This implies that these devices had been abandoned and were not receiving updates.
+This implies that these devices had been abandoned and were not receiving updates which is confirmed by Figure~\ref{fig:device_data_os} which shows which OS versions those devices were running.
 Some devices start off in 2011 exposed to known security vulnerabilities and are still exposed to those and additional ones in 2014.
@@
  \includegraphics[width=\columnwidth]{figures/nw_security_updates.pdf}
- \caption{Proportion of updates each week which fixed or may have fixed security vulnerabilities\todo{Show non security updates as well for comparison}}
+ \caption{Proportion of updates each week which fixed or may have fixed security vulnerabilities}
  \label{fig:weekly_security_updates}

### 2014-10-29 figures/update_figures.sh

@@ -9,6 +9,7 @@ cp "$fromdir"norm_os.pdf da_norm_os.pdf
 cp "$fromdir"os.pdf da_os.pdf
 cp "$fromdir"device_data/device-data-manufacturer-lg-security.pdf device_data-LG.pdf
 cp "$fromdir"device_data/device-data-all-security.pdf device-data-all-security.pdf
+cp "$fromdir"device_data/device-data-all-os.pdf device-data-all-os.pdf
 cp "$fromdir"from_to_updates.pdf from_to_updates.pdf
 cp "$fromdir"googleplayapi.pdf googleplayapi.pdf
 cp "$fromdir"googleplayapi-notitle.pdf googleplayapi-notitle.pdf

### 2014-10-29 securityupdates.tex
use generated percentage for valid entries in data

@@
 The API version is a well defined integer, however it does not always change with new Android releases, particularly security bug fixes should not increase this value.
-The build string is `The user-visible version string', fortunately most (99.9\%)\todolater{use a generated number} devices in the data have a build string of the form `x.y.z random string' and so it is possible to extract the Android version number.
+The build string is `The user-visible version string', fortunately most (\daOSVersionPercValidLines) entries in the data have a build string of the form `x.y.z random string' and so it is possible to extract the Android version number.
 On a large proportion of devices `random string' is a well defined build number\footnote{\url{https://source.android.com/source/build-numbers.html}} but different manufacturers use different schema.

### 2014-10-30 securityupdates.tex
Cut a bunch of stuff and put in a cutting floor section at the end as we may want people to suggest we put them back.

@@
 Hence there is little incentive for manufacturers and operators to do a better job.
-We provide this data  in~\S\ref{sec:security_scoring} and discuss incentives further in \S\ref{sec:economics}.
+We provide this data  in~\S\ref{sec:security_scoring}.
 
@@
  \item We measure the security of Android according to three metrics and compare different manufacturers. and device models to allow consumers to differentiate between them based on security.
- \item We provide models of how the distribution of Android versions changes allowing future version distributions to be predicted.
-% \item We propose methods which could be used to improve the situation in future.
 \end{itemize}
@@
 
-\subsection{Running a version of Android known to be vulnerable makes a device vulnerable}
-We have not tried to exploit these vulnerabilities on the devices in order to test whether the devices are actually vulnerable.
-Doing so would be difficult both ethically, as it would pose a high risk to user devices and mean that our code could easily be maliciously repurposed, and practically as getting all the exploits working reliably on all the devices they could be made to work reliably on requires extensive testing.
-We are confident that most devices will be vulnerable if they are running a version of Android known to be vulnerable. \todo{WHY?}
-
 \subsection{Android update process}
@@
 
-\subsection{Monetisation of malware}
-Android malware can make money in a variety of different ways~\cite{Felt2011}.
-One of the key differences between smartphones and desktops is that it is easier to directly cash out on a smartphone as malware can send premium rate texts or make premium rate phone calls or send SMS and email spam.
-It can also steal personal data which already has semantic information associated with it, this also makes it easier to sell and to perform ransomware attacks where malware deletes this information from the device and charges a ransom for its return.
-It can steal credentials for external services or use in-app billing.
-It can also do more conventional things such as use the devices for DDoS, SEO or click fraud attacks and altering the ads displayed to send the fees to the malware author.
-While many of these things can be done by malware which just requests the relevant permissions, root exploits make this easier, particularly for ransomware and altering the ads displayed.
-
-\subsection{Incentives and participants}
-\label{sec:economics}
-Google and other Android vendors care about the reputation of Android, particularly from a security standpoint as if users are afraid to use their devices for certain purposes (e.g.\ banking) then they will not be willing to pay as much for them and money cannot be made from those activities.
-Manufacturers make money by selling the device, once sold they want to prevent the cost of return under warranty and so are adverse to the risk of breaking the device, but they do not need to do things to make the device better as it has already been sold.
-However if they gain a reputation for abandoning devices that might decrease future sales.
-
-Operators sell phones under contracts, typically at least a year long, they want the user to not request support due to the device breaking.
-They might also gain advantage if the user is happy with the device after the end of the contract as they might then keep paying the same rate but there would no longer be a capital cost to pay off.
-They care about phones not getting malware which causes support requests or uses the mobile network to perform malicious actions as they may have to reimburse the cost of premium SMS's if they were not authorised by the user.
-
-Users want to buy phones which will serve their use cases and which are secure so that they do not experience negative consequences.
-
 \section{Experiments}
@@
 %\todo{comparison of nexus update rates with iOS}
-We conducted four experiments examining the behaviour of the Android ecosystem as a whole (\S\ref{sec:exp:android_ecosystem}), updates reaching particular devices (\S\ref{sec:exp:device_updates}), comparisons between different manufacturers and operators (\S\ref{sec:exp:security_score}) and model how long it will take for an update to be fixed once an update is released (\S\ref{sec:exp:predicting_distributions}).
+We conducted three experiments examining the behaviour of the Android ecosystem as a whole (\S\ref{sec:exp:android_ecosystem}), updates reaching particular devices (\S\ref{sec:exp:device_updates}) and comparisons between different manufacturers and operators (\S\ref{sec:exp:security_score}).
 
@@
 \input{security_scoring}
-
-% We don't need this for this paper
-%\subsection{Comparison with iOS}
-%\todo{Get data from Richard so that we can do this comparison}
-
-\subsection{Experiment 4: Predicting future version distributions}\label{sec:exp:predicting_distributions}
-\subsubsection{Method}
-Examining the way that the OS version distribution changes over time in Figure~\ref{fig:norm_os} we see some regularity to the behaviour.
-To show this behaviour more clearly we normalised the data to the number of days since the version was released and smoothed it by forcing it to be monotonically decreasing.\footnote{While sometimes an older version becomes more popular after a later version has been released causing the values to actually increase, mostly this was just noise or changes in the distribution of devices (e.g.\ the sudden arrival of 1000 devices from Bangladesh due to a study being conducted there) and this assumption makes the graph clearer.}
-Then we can plot Figure~\ref{fig:frvh_os_versions} which shows the proportion of of devices in \da\ not upgraded to each Android version or a later Android version (i.e.
-for 2.1.0 it includes 2.1.1, 2.2.0 and indeed all devices as they all run later versions of Android).
-The mean and standard deviation of the same data are shown in Figure~\ref{fig:frms_os_versions}.
-
-\begin{figure}
- \includegraphics[width=\columnwidth]{figures/frvh_os_versions}
- \caption{Proportion of devices not updated to particular versions of Android or any later version}
- \label{fig:frvh_os_versions}
-\end{figure}
-\begin{figure}
- \includegraphics[width=\columnwidth]{figures/frms_os_versions}
- \caption{Proportion of devices not updated to particular versions of Android or any later version}
- \label{fig:frms_os_versions}
-\end{figure}
-
-The data appears to have an exponential decay as it tends to 1 but with a slow start as few devices are upgraded initially.
-We approximate this with a $e^{-x}$ for the main behaviour and a which is offset to start later when uptake takes off.
-The equation is then:
-\begin{equation}
-y =
- \begin{cases}
-  1.0 & \text{if }x<\mathrm{undeployed}\\
-  e^{-\mathrm{decay}(x-\mathrm{undeployed})} & \text{otherwise}
- \end{cases}
-\end{equation}
-
-\subsubsection{Results}
-Fitting that curve to the data using \texttt{scipy.optimize}, we get a RMSE of \daOSCurveFitRMSE\ and parameters $\mathrm{undeployed} = \daOSCurveFitParamFirst ~\si{days}$, $\mathrm{decay} = \daOSCurveFitParamSecond ~\si{days^{-1}}$ for OS versions and RMSE of \daAPICurveFitRMSE\, $\mathrm{undeployed} = \daAPICurveFitParamFirst ~\si{days}$, $\mathrm{decay} = \daAPICurveFitParamSecond ~\si{days^{-1}}$ for API versions.
-From this the number of days from release of a new version of Android until 50\% of devices are running that version or higher is \daOSCurveHalfDeployed\ and full deployment to \daFullDeployedAt\ of devices takes \daOSCurveFullDeployed\ days.
-For API versions this is \daAPICurveHalfDeployed\ and \daAPICurveFullDeployed\ respectively.
-
-Hence if a security vulnerability is fixed through the release of a particular Android version it will be \daOSCurveFullDeployed\ days after until the fix is fully deployed.
-
-The equation we chose for the model has a root mean squared error (RMSE) of \daOSCurveFitRMSE\ for OS versions, which is a comparably good fit to a standard polynomial fit (3 degree polynomial fit gave a RMSE of \daOSCurvePolyRMSE) or a spline fit (RMSE of \daOSCurveSplineRMSE) but gives a meaningful model of behaviour rather than a generic curve.
-
-\todo{Evaluate our model trained on the DA data against the ground truth Google Play data}
-
-\todo{Our handling of errors here is not quite good enough as the uncertainty from our y value interacts badly with the logarithm function when computing the inverse and we compute the log of ~0 which means our error ends up much bigger than our value - but that is an upper bound, not a +- bound.}
-
-We hope that a consequence of this paper will be a change in behaviour of the Android ecosystem which invalidates this model by speeding up the deployment of new OS versions.
 
@@
 
+\listoftodos
+
+\section{Cutting floor}
+
+This section contains things which we were intending to include but which have not made the cut.
+
+\subsection{Running a version of Android known to be vulnerable makes a device vulnerable}
+We have not tried to exploit these vulnerabilities on the devices in order to test whether the devices are actually vulnerable.
+Doing so would be difficult both ethically, as it would pose a high risk to user devices and mean that our code could easily be maliciously repurposed, and practically as getting all the exploits working reliably on all the devices they could be made to work reliably on requires extensive testing.
+We are confident that most devices will be vulnerable if they are running a version of Android known to be vulnerable. \todo{WHY?}
+
+\subsection{Monetisation of malware}
+Android malware can make money in a variety of different ways~\cite{Felt2011}.
+One of the key differences between smartphones and desktops is that it is easier to directly cash out on a smartphone as malware can send premium rate texts or make premium rate phone calls or send SMS and email spam.
+It can also steal personal data which already has semantic information associated with it, this also makes it easier to sell and to perform ransomware attacks where malware deletes this information from the device and charges a ransom for its return.
+It can steal credentials for external services or use in-app billing.
+It can also do more conventional things such as use the devices for DDoS, SEO or click fraud attacks and altering the ads displayed to send the fees to the malware author.
+While many of these things can be done by malware which just requests the relevant permissions, root exploits make this easier, particularly for ransomware and altering the ads displayed.
+
+\subsection{Incentives and participants}
+\label{sec:economics}
+Google and other Android vendors care about the reputation of Android, particularly from a security standpoint as if users are afraid to use their devices for certain purposes (e.g.\ banking) then they will not be willing to pay as much for them and money cannot be made from those activities.
+Manufacturers make money by selling the device, once sold they want to prevent the cost of return under warranty and so are adverse to the risk of breaking the device, but they do not need to do things to make the device better as it has already been sold.
+However if they gain a reputation for abandoning devices that might decrease future sales.
+
+Operators sell phones under contracts, typically at least a year long, they want the user to not request support due to the device breaking.
+They might also gain advantage if the user is happy with the device after the end of the contract as they might then keep paying the same rate but there would no longer be a capital cost to pay off.
+They care about phones not getting malware which causes support requests or uses the mobile network to perform malicious actions as they may have to reimburse the cost of premium SMS's if they were not authorised by the user.
+
+Users want to buy phones which will serve their use cases and which are secure so that they do not experience negative consequences.
+
+% We don't need this for this paper
+%\subsection{Comparison with iOS}
+%\todo{Get data from Richard so that we can do this comparison}
+
+\subsection{Experiment 4: Predicting future version distributions}\label{sec:exp:predicting_distributions}
+\subsubsection{Method}
+Examining the way that the OS version distribution changes over time in Figure~\ref{fig:norm_os} we see some regularity to the behaviour.
+To show this behaviour more clearly we normalised the data to the number of days since the version was released and smoothed it by forcing it to be monotonically decreasing.\footnote{While sometimes an older version becomes more popular after a later version has been released causing the values to actually increase, mostly this was just noise or changes in the distribution of devices (e.g.\ the sudden arrival of 1000 devices from Bangladesh due to a study being conducted there) and this assumption makes the graph clearer.}
+Then we can plot Figure~\ref{fig:frvh_os_versions} which shows the proportion of of devices in \da\ not upgraded to each Android version or a later Android version (i.e.
+for 2.1.0 it includes 2.1.1, 2.2.0 and indeed all devices as they all run later versions of Android).
+The mean and standard deviation of the same data are shown in Figure~\ref{fig:frms_os_versions}.
+
+\begin{figure}
+ \includegraphics[width=\columnwidth]{figures/frvh_os_versions}
+ \caption{Proportion of devices not updated to particular versions of Android or any later version}
+ \label{fig:frvh_os_versions}
+\end{figure}
+\begin{figure}
+ \includegraphics[width=\columnwidth]{figures/frms_os_versions}
+ \caption{Proportion of devices not updated to particular versions of Android or any later version}
+ \label{fig:frms_os_versions}
+\end{figure}
+
+The data appears to have an exponential decay as it tends to 1 but with a slow start as few devices are upgraded initially.
+We approximate this with a $e^{-x}$ for the main behaviour and a which is offset to start later when uptake takes off.
+The equation is then:
+\begin{equation}
+y =
+ \begin{cases}
+  1.0 & \text{if }x<\mathrm{undeployed}\\
+  e^{-\mathrm{decay}(x-\mathrm{undeployed})} & \text{otherwise}
+ \end{cases}
+\end{equation}
+
+\subsubsection{Results}
+Fitting that curve to the data using \texttt{scipy.optimize}, we get a RMSE of \daOSCurveFitRMSE\ and parameters $\mathrm{undeployed} = \daOSCurveFitParamFirst ~\si{days}$, $\mathrm{decay} = \daOSCurveFitParamSecond ~\si{days^{-1}}$ for OS versions and RMSE of \daAPICurveFitRMSE\, $\mathrm{undeployed} = \daAPICurveFitParamFirst ~\si{days}$, $\mathrm{decay} = \daAPICurveFitParamSecond ~\si{days^{-1}}$ for API versions.
+From this the number of days from release of a new version of Android until 50\% of devices are running that version or higher is \daOSCurveHalfDeployed\ and full deployment to \daFullDeployedAt\ of devices takes \daOSCurveFullDeployed\ days.
+For API versions this is \daAPICurveHalfDeployed\ and \daAPICurveFullDeployed\ respectively.
+
+Hence if a security vulnerability is fixed through the release of a particular Android version it will be \daOSCurveFullDeployed\ days after until the fix is fully deployed.
+
+The equation we chose for the model has a root mean squared error (RMSE) of \daOSCurveFitRMSE\ for OS versions, which is a comparably good fit to a standard polynomial fit (3 degree polynomial fit gave a RMSE of \daOSCurvePolyRMSE) or a spline fit (RMSE of \daOSCurveSplineRMSE) but gives a meaningful model of behaviour rather than a generic curve.
+
+\todo{Evaluate our model trained on the DA data against the ground truth Google Play data}
+
+\todo{Our handling of errors here is not quite good enough as the uncertainty from our y value interacts badly with the logarithm function when computing the inverse and we compute the log of ~0 which means our error ends up much bigger than our value - but that is an upper bound, not a +- bound.}
+
+We hope that a consequence of this paper will be a change in behaviour of the Android ecosystem which invalidates this model by speeding up the deployment of new OS versions.
+
 \end{document}

### 2014-10-30 securityupdates.tex
Minor tweaks from Alastair

@@
 Hence when Google produces an update to Android, the update may have to pass through the manufacturer and operator before reaching the user.
-The manufacturer and operator have little financial incentive to perform this work as they have already sold the device.
+The manufacturer has little financial incentive to perform this work as they have already sold the device.
 Operators are cautious because if they ship a broken update then they will get many support requests which is expensive.
@@
 Between 36.7\%~\cite{Zhou2012b} and 40\%~\cite{Zhou2012a} of malware for Android contained root exploits in 2012.
-While this shows that much Android malware does not need a root exploit to work, since the user will grant the relevant permissions if the app requests them, a significant proportion does use root exploits.
-This malware is also more dangerous because while ordinary malware can be remotely removed by Google through the Play store, once a root exploit has been used there are no guarantees.
+While this shows that much Android malware does not need a root exploit to work, a significant proportion does include root exploits.
+This malware is also more dangerous because while ordinary malware can be remotely removed by Google through the Play store, once a root exploit has been used this is no longer guaranteed to work.
 
@@
 \begin{itemize}
- \item We characterise the Android update ecosystem and quantify its vulnerability to root equivalent vulnerabilities.
+ \item We characterise the Android update ecosystem and quantify vulnerabilities.
  \item We collate information on vulnerabilities which affect Android and investigate their time-lines.
- \item We measure the security of Android according to three metrics and compare different manufacturers. and device models to allow consumers to differentiate between them based on security.
+ \item We measure the security of Android according to three metrics and compare different manufacturers and device models to allow device purchasers to differentiate between them based on security.
 \end{itemize}
@@
 \label{sec:background}
-There are two sources of data we needed to have and to link for this study, information on the distribution of installed versions of Android over time and on root vulnerabilities.
+There are two sources of data we need for this study: (1) information on the distribution of installed versions of Android over time and (2) information on root vulnerabilities.
 We need data on root vulnerabilities so that we know which versions of Android were vulnerable to which known vulnerabilities at what times.
@@
 We need historical data on the distributions of versions of Android running on devices and additional data about those devices to allows us to investigate variation between devices and confirm the reliability of these data.
-The only suitable data which is available for this is the \da\ data~\cite{Wagner2013}.
-The \da\ data contains data from \daNumDevices\ devices with a total of \daDeviceDays\footnote{Here we are only counting devices and days for which we have valid OS version data} device days, it contains data for more than \daMonths\ months for \daMonthsDevices\ devices.
+We used data from the \da\ project~\cite{Wagner2013}.
+The \da\ data contains data from \daNumDevices\ devices with a total of \daDeviceDays\footnote{Here we are only counting devices and days for which we have valid OS version data} device days, contains data for more than \daMonths\ months for \daMonthsDevices\ devices.
 Device Analyzer used an Android app which volunteers install on their Android devices to collect data on what happens on Android phones.
@@
 
-The security protections used in iOS such, app review, code signing, Data Execution Prevention (DEP) and Address Space Layout Randomisation (ASLR) and mandatory access controls has resulted in a lower level of malware affecting iOS~\cite{Felt2011} but can still be bypassed~\cite{Wang2013a}.
+The security protections used in iOS such as, app review, code signing, Data Execution Prevention (DEP) and Address Space Layout Randomisation (ASLR) and mandatory access controls has resulted in a lower level of malware affecting iOS~\cite{Felt2011} but can still be bypassed~\cite{Wang2013a}.
 
@@
 
-The updatedness rate for Android of \daUpdatednessPerc\ compares unfavourable with the more than 90\% rate for Windows XP SP2 computers contacting the Microsoft update servers~\cite{Gkantsidis2006} though that is within one OS version and only for those computers which did connect to the servers -- but that is the default.
+The updatedness rate for Android of \daUpdatednessPerc\ compares unfavourably with the more than 90\% rate for Windows XP SP2 computers contacting the Microsoft update servers~\cite{Gkantsidis2006} though that is within one OS version and only for those computers which did connect to the servers -- but that is the default.
 However 27\% of Windows computers were still running Windows XP in July 2014\footnote{\url{https://archive.today/PLGxn}\todolater{Get updated figures before publication}} -- four months after it went out of security support.

### 2014-10-30 security_scoring.tex
Fixes from Alastair, in particular move away from calling things 'root equivalent vulnerabilities' to just 'critical vulnerabilities'

@@
 A high value of $m$ could be achieved by focusing on only one device at a time and ensuring that gets updates but ignoring all others, but that would lower $f$ and $u$.
-One way to influence our scores would be to add additional devices to Device Analyzer which have good security, these would have to be real end user devices as we could detect fake ones, this would increase the size of our data set and would require providing genuinely good security to some users.
+\footnote{One way to influence our scores would be to add additional devices to Device Analyzer which have good security, these would have to be real end user devices as we could detect fake ones, this would increase the size of our data set and would require providing genuinely good security to some users.}
 

### 2014-10-30 securityupdates.tex
Fixes from Alastair, in particular move away from calling things 'root equivalent vulnerabilities' to just 'critical vulnerabilities'

@@
 % Android, Unix
+% update, release, patch: what do we actually mean and what relevance does that have to security?
 
@@
 \label{sec:background}
-There are two sources of data we need for this study: (1) information on the distribution of installed versions of Android over time and (2) information on root vulnerabilities.
-We need data on root vulnerabilities so that we know which versions of Android were vulnerable to which known vulnerabilities at what times.
+There are two sources of data we need for this study: (1) information on the distribution of installed versions of Android over time and (2) information on critical vulnerabilities.
+We need data on critical vulnerabilities so that we know which versions of Android were vulnerable to which known vulnerabilities at what times.
 We need information on the distribution of installed versions of Android so that we know what proportion of Android devices were affected by these vulnerabilities.
@@
 
-\subsection{Root vulnerabilities}
-We compiled a list of root equivalent vulnerabilities in Android.
+We verify the validity of the \da\ data in \S\ref{sec:representative}.
+
+\todo{acr31: More information about what \da\ records and how we process it to provide data}
+
+\subsection{Critical vulnerabilities}
+We compiled a list of critical vulnerabilities in Android.
 We collected information on discovery and publication dates, what versions they affected and which versions fixed the problem.
-We only looked for \emph{root equivalent} vulnerabilities which did not require USB debugging to exploit.
-\emph{Root equivalent} means that it escalates to privileges equivalent in scope to root in that either they can be used to gain root or they provide permissions as bad as root.
-If an application exploits a root equivalent vulnerability then it gains control of the device.
+We only looked for critical vulnerabilities such as root vulnerabilities which did not require USB debugging to exploit.
+Critical vulnerabilities allow a program to escalate to privileges equivalent in scope to root in that either they can be used to gain root or they provide permissions as bad as root.
+If an application exploits a critical vulnerability then it gains control of the device.
 Some phones can be `rooted' by enabling USB debugging and using the special privileges of the adb shell to root the device but only \daAdbEnabledPerc\dafoot\ of devices have USB debugging enabled.\footnote{This is surprisingly high as \daAdbEnabledPerc\ of Android users are not developers.}
@@
 
-Some vulnerabilities which grant root are not traditional kernel vulnerabilities, for example the discovery of flaws in the verification of signatures on Android applications in February 2013~\cite{Forristal2013} meant that applications could pretend to be signed with system keys and hence gain root equivalent privileges.
+Some critical vulnerabilities are not traditional kernel vulnerabilities, for example the discovery of flaws in the verification of signatures on Android applications in February 2013~\cite{Forristal2013} meant that applications could pretend to be signed with system keys and hence gain root equivalent privileges.
 On some versions of Android (below 4.1) they could use known system to root escalation mechanisms but on all versions they have a greatly increased attack area for further privilege escalation and also have the ability to control all user internet traffic (via VPNs), brick the phone, remove and install apps, steal user credentials, read the screen and make as well as receive calls.
@@
 
-We developed and maintain an open platform for filing root equivalent vulnerabilities in a machine readable format, \url{http://androidvulnerabilities.org/}.
+We developed and maintain an open platform for filing critical vulnerabilities in a machine readable format, \url{http://androidvulnerabilities.org/}.
 We seeded it with data from the CVE database, vendor lists, reports from the literature and various forums.
@@
 Tracking vulnerabilities is a manual task as they are not all in other databases such as the CVE database and the lack of a widely acknowledged unique identifier made identifying whether two reports referenced the same vulnerability difficult.
-Previous work has assumed ``any security issue of relevance will eventually get a CVE number assigned"~\cite{Frei2010} which is currently not the case for Android root equivalent vulnerabilities.
+Previous work has assumed ``any security issue of relevance will eventually get a CVE number assigned"~\cite{Frei2010} which is currently not the case for critical Android vulnerabilities.
 For some of the vulnerabilities without CVE numbers, Google confirmed that there was no CVE number and that they did not intend to get one, instead providing an Android bug number.
@@
 So for our vulnerability calculations we use the earliest date we can find recorded evidence that the vulnerability was known about, even if that knowledge might have been confined to a particular manufacturer or hobbyist.
-We cannot know if someone reported the vulnerability to the manufacturer and we do know that various agencies are engaged in widespread monitoring of communications and compromise of civilian infrastructure~\cite{TODO} and so they probably know.
+We cannot know if someone reported the vulnerability to the manufacturer and we do know that various agencies are engaged in widespread monitoring of communications and compromise of civilian infrastructure~\cite{TODO} and so they may know.
 
@@
 {\bf Creation} and {\bf introduction} are the earliest points which could be used but the risk is mostly latent until someone discovers them.
-Unfortunately {\bf discovery} is the hardest point to obtain concrete data on as discovery may happen multiple times independently and not all discoverers will report their discovery, for example APTs may keep databases of unknown vulnerabilities to use when they have a particular target, in which case there is a high risk to potential targets from the never disclosed point of discovery.
+Unfortunately {\bf discovery} is the hardest point to obtain concrete data on as discovery may happen multiple times independently and not all discoverers will report their discovery, for example Advanced Persistent Threats (APTs) may keep databases of unknown vulnerabilities to use when they have a particular target, in which case there is a high risk to potential targets from the never disclosed point of discovery.
 However if the discoverer never discloses the vulnerability to anyone or makes any use of it then there is little risk.
@@
 The date of first {\bf fix} is, assuming that the fix is deliberate, a point at which the vulnerability is known at least within the organisation performing the fix and frequently implies an earlier discovery and notification by a third party.
+It can, for example, be determined from the authored on date on the fixing commit in the git repository.
 Once a {\bf fixing release} has been made then the vulnerability is widely known because it can be reverse engineered from the changes in the release~\cite{Brumley2008}.
@@
 Hence we use the earliest date we know about for discovery, exploit, disclosure or fix as the date from which a vulnerability poses a substantial risk to users.
+This is shown in Table~\ref{tab:andvulns}.
 
@@
 
-This data shows a large gap in the middle where we do not know of any discoveries of root equivalent vulnerabilities affecting all of Android.
+This data shows a large gap in the middle where we do not know of any discoveries of critical vulnerabilities affecting all of Android.
 The cause of this quiet period is unclear, perhaps from Figure~\ref{fig:proportioninsecure} the reason was that since most devices were exposed to known vulnerabilities there was no point in looking for new ones?
@@
 \label{sec:results}
-We present the results of our analysis showing \daMeanInsecurityPerc\ of devices to have root equivalent vulnerabilities on average.
+We present the results of our analysis showing \daMeanInsecurityPerc\ of devices to have critical vulnerabilities on average.
 On average only \daUpdatesPerYear\ updates reach a device each year and \daPercUpdatesDowngrades\ of version changes are downgrades to older versions.
@@
 The \avo\ data was collected between \avoStartDate\ and \avoEndDate\ and covers the period from \avoFirstDataDate\ to \avoLastDataDate.
-We verify the validity of the \da\ data in \S\ref{sec:representative}.
 For each device, for every day with version data we record what version it was running and which vulnerabilities it was exposed to at that time (if any).
@@
 Green indicates updates which changed the build number but not the version number and so might contain a backported fix for a vulnerability.
-Red indicates updates which did not fix security vulnerabilities (because there were know known security vulnerabilities in the version of Android they were already running).
+Red indicates updates which did not fix security vulnerabilities (because there were no known security vulnerabilities in the version of Android they were already running).
 
@@
 \section{Discussion}
-There are continuing efforts to reduce the impact of root equivalent vulnerabilities, both in Android and more widely.
+There are continuing efforts to reduce the impact of critical vulnerabilities, both in Android and more widely.
 SEAndroid~\cite{Smalley2013} which is included in Android from version 4.1~\cite{jelly-bean-release} claimed to prevent some root vulnerabilities and to reduce the impact of others.
 Capability based enforcement systems such as Capsicum~\cite{Watson2010} substantially reduce the capabilities that an exploit has to try and gain increased privilege with.
-When Capsicum is included in Linux~\cite{TODO} and hence in Android, it could be used to place fine grained restrictions on system daemons, preventing vulnerabilities becoming root equivalent vulnerabilities.
+When Capsicum is included in Linux~\cite{TODO} and hence in Android, it could be used to place fine grained restrictions on system daemons, preventing vulnerabilities becoming critical vulnerabilities.
 
@@
 \label{sec:related}
-We found that since 2011 \daMeanInsecurityPerc\ of devices on average were exposed to known root equivalent vulnerabilities.
+We found that since 2011 \daMeanInsecurityPerc\ of devices on average were exposed to known critical vulnerabilities.
 In 2011 Felt et al.\ \cite{Felt2011} studied 6 Android handsets and found that they were exposed to root vulnerabilities at least 74\% of the time.
@@
 We hope that this analysis will encourage manufacturers and operators to improve the support they provide for devices after sale.
-Android attempts to provide good security to end users but the latency in the security update process means that on average \daMeanInsecurityPerc\ of Android devices are exposed to known root equivalent vulnerabilities.
+Android attempts to provide good security to end users but the latency in the security update process means that on average \daMeanInsecurityPerc\ of Android devices are exposed to known critical vulnerabilities.
 

### 2014-10-30 securityupdates.tex
Don't talk about things we promised not to talk about

@@
 However two collaborations with network operators to install \da\ on customer devices resulted in large numbers of new \da\ users coming from particular areas and network operators.
-`n telenor' and `grameenphone' from Table \ref{tab:sec_operator} are big operators in the \da\ data because of their collaboration, in particular `n telenor' conducted a study in Norway with 654 installs of \da\ and `grameenphone' conducted a study in Bangladesh where 2463 users installed \da.
+One study was conducted in Norway with 654 installs of \da\ and one was conducted in Bangladesh where 2463 users installed \da.
 This is responsible for the sudden change in distribution of OS versions in \da\ shown towards the end of Figures \ref{fig:norm_os}, \ref{fig:nvulnerabilities_heat} and \ref{fig:proportioninsecure} as for that period Bangladeshi users contributed over half of the \da\ data.

### 2014-10-30 securityupdates.tex
More tweaks from Alastair's comments

@@
 The Operator may then make further customisations and do further testing before shipping the update to the Device.
-Sometimes Manufactures ship direct to the user, sometimes the Manufacturer and Google are the same entity (at least for the purpose of a particular phone, such as with Nexus phones).
-Sometimes Manufacturers incorporate upstream open source project releases directly, and sometimes incorrectly -- for example including a broken daily build of sqlite in their release of Android.
+Sometimes Manufactures ship direct to the user, sometimes the Manufacturer and Google collaborate closely for the purpose of a particular phone, such as with Nexus phones, and hence are harder to separate.
+Sometimes Manufacturers incorporate upstream open source project releases directly, and sometimes incorrectly -- for example including a broken daily build of sqlite in their release of Android~\cite{Wagner2013}.
 \todolater{use statistics from samsung-updates.com -> how many binaries are there per device?}
@@
 Manufacturer and Operator counts were obtained by normalising the results reported by Android to \da\ of the device manufacturer and active network operator.
-This normalisation involves removing invalid values (such as `manufacturer' or `airplane mode is on'), collating across company name changes (e.g.\ `lge' to `lg'), normalising punctuation, removing extra strings sometimes added such as (`(2g)' or `communications') and mapping some incorrectly placed model names back to their manufacturer.
+This normalisation is a manual task involves removing invalid values (such as `manufacturer' or `airplane mode is on'), collating across company name changes (e.g.\ `lge' to `lg'), normalising punctuation, removing extra strings sometimes added such as (`(2g)' or `communications') and mapping some incorrectly placed model names back to their manufacturer.
 This normalisation is not perfect and so these are overestimates on the \da\ data but they are likely still underestimates as there will be some manufacturers and operators which are not included in the \da\ data.
@@
  \end{subfigure}
- \caption{The top \daNumDeviceDataDevices\ devices by days of contribution in the \da\ data}
+ \caption{The top \daNumDeviceDataDevices\ devices by days of contribution in the \da\ data. One strip per device.}
 \end{figure}
@@
 By taking the number of updates observed and the number of device days of data we have collected we can compute the number of updates received by a device per year to be \daUpdatesPerYear.
-This compares badly with the number of vulnerabilities discovered per year of \avoVulnsPerYear.
-
-\todo{That comparison is not completely fair as not all of those vulnerabilites will affect all devices, similarly we may be counting device days for which we only have one day and hence could perhaps not observe an update.}
+This compares badly with the number of critical vulnerabilities discovered per year of between \avoVulnsPerYearAllAndroid\ (affecting all Android) and \avoVulnsPerYear\ (including the manufacturer specific ones in \avo).
 
@@
 Thanks to David Robertson for helpful advice on statistical analysis.
-Thanks to Laurent Simon, Thomas Coudray, Adrian Taylor, Justin Case and Khilan Gudka for reporting vulnerabilities in Android.
+Thanks to Laurent Simon, Thomas Coudray, Adrian Taylor, Justin Case, Giant Pune and Khilan Gudka for reporting vulnerabilities in Android.
 

### 2014-11-04 securityupdates.tex
Some of acr31's corrections

@@
 Many such exploits have been found and fixed.
-Our hypothesis is that these fixes do not promptly reach the users devices and so much of the time users are running versions of Android known to be vulnerable.
+Fixing them is not enough, we show that these fixes do not promptly reach the users devices and so much of the time users are running versions of Android known to be vulnerable.
 We used \da\ data from over \daOSYearsOfData\ years and \daNumOSDevices\ devices and found that on average \daMeanInsecurityPerc\ of devices were exposed to known root privilege vulnerabilities and only \daUpdatednessPerc\ of devices run the most recent version of Android.
@@
 \section{Introduction}
-Android has \percMarketShare\ of the smartphone market and while the core development is controlled by Google there are about \daNumManufacturers\footnote{\label{foot:dadata}We computed this from the \da~\cite{Wagner2013} data, see \S\ref{sec:android_update_process}.} manufacturers which make devices which run Android.
+Android has \percMarketShare\ of the smartphone market and while the core development is controlled by Google, we know of \daNumManufacturers\footnote{\label{foot:dadata}We computed this from the \da~\cite{Wagner2013} data, see \S\ref{sec:android_update_process}.} manufacturers which make devices which run Android.
 Android receives regular updates  which add new features and fix bugs, including security vulnerabilities.
-Many of these manufacturers customise the version of Android they ship and sometimes network operators (of which there are at least \daNumOperators\dafoot) make further modifications.
+Many of these manufacturers customise the version of Android they ship and sometimes network operators (of which we know of \daNumOperators\dafoot) make further modifications.
 Hence when Google produces an update to Android, the update may have to pass through the manufacturer and operator before reaching the user.
@@
 However there is ongoing legal action to force operators to ship updates for security vulnerabilities~\cite{Soghoian2013}.\todolater{Check on the status of this legal action}
-Many phones are sold on long (12--24 month) contracts with monthly payments yet stop receiving updates before the end of the contract,\todolater{Is there a survey of the distribution of contract lengths and phone purchase methods?} while Windows XP could be purchased for a one off payment in October 2001 and received security updates until April 2014.
+Many phones are sold on long (12--24 month) contracts with monthly payments yet stop receiving updates before the end of the contract,\todolater{Is there a survey of the distribution of contract lengths and phone purchase methods?} whereas Windows XP could be purchased for a one-off payment in October 2001 and received security updates until April 2014.
 Corporate or public sector buyers are encouraged to purchase secure devices, but we have found little concrete guidance.
@@
 Android relies on the Linux kernel to provide compartmentalisation and runs each app as a separate Unix `user'.
-The Linux kernel is a large piece of software and so there will always be new root exploits found in it~\cite{TODO}.
-Manufacturers of Android devices have also added further root exploits when customising it~\cite{Grace2012}.
+The Linux kernel is a large piece of software and so there will likely be new root exploits found in it~\cite{TODO}.
+Manufacturers of Android devices have also accidentally added further root exploits when customising it~\cite{Grace2012}.
 Between 36.7\%~\cite{Zhou2012b} and 40\%~\cite{Zhou2012a} of malware for Android contained root exploits in 2012.
@@
 \begin{itemize}
- \item We characterise the Android update ecosystem and quantify vulnerabilities.
+ \item We characterise the Android update ecosystem, showing how updates flow between entities, and quantify vulnerabilities by determining what proportion of devices they affected and how that changes over time.
  \item We collate information on vulnerabilities which affect Android and investigate their time-lines.
  \item We measure the security of Android according to three metrics and compare different manufacturers and device models to allow device purchasers to differentiate between them based on security.
+ \todo{make this more concrete}
+ \todo{Do we want \avo\ to provide our latest recommendations -- legal advice required}
 \end{itemize}
@@
 \subsection{Versions of Android running on devices}
-We need historical data on the distributions of versions of Android running on devices and additional data about those devices to allows us to investigate variation between devices and confirm the reliability of these data.
+We need historical data on the distributions of versions of Android running on devices and additional data about those devices to allow us to investigate variation between devices and confirm the reliability of these data.
 We used data from the \da\ project~\cite{Wagner2013}.
-The \da\ data contains data from \daNumDevices\ devices with a total of \daDeviceDays\footnote{Here we are only counting devices and days for which we have valid OS version data} device days, contains data for more than \daMonths\ months for \daMonthsDevices\ devices.
+The \da\ data contains info from \daNumDevices\ devices with a total of \daDeviceDays\footnote{Here we are only counting devices and days for which we have valid OS version data.} device days, contains data for more than \daMonths\ months for \daMonthsDevices\ devices.
 Device Analyzer used an Android app which volunteers install on their Android devices to collect data on what happens on Android phones.
@@
 The API version is a well defined integer, however it does not always change with new Android releases, particularly security bug fixes should not increase this value.
-The build string is `The user-visible version string', fortunately most (\daOSVersionPercValidLines) entries in the data have a build string of the form `x.y.z random string' and so it is possible to extract the Android version number.
-On a large proportion of devices `random string' is a well defined build number\footnote{\url{https://source.android.com/source/build-numbers.html}} but different manufacturers use different schema.
+The build string is `The user-visible version string', fortunately most (\daOSVersionPercValidLines) entries in the data have a build string of the form `x.y.z opaque marker' and so it is possible to extract the Android version number.
+On a large proportion of devices `opaque marker' is a well defined build number\footnote{\url{https://source.android.com/source/build-numbers.html}} but different manufacturers use different schema.
 
@@
 
-\todo{acr31: More information about what \da\ records and how we process it to provide data}
-
 \subsection{Critical vulnerabilities}
-We compiled a list of critical vulnerabilities in Android.
-We collected information on discovery and publication dates, what versions they affected and which versions fixed the problem.
+We compiled a list of critical vulnerabilities in Android, containing information on discovery and publication dates, what versions they affected and which versions fixed the problem.
 We only looked for critical vulnerabilities such as root vulnerabilities which did not require USB debugging to exploit.
-Critical vulnerabilities allow a program to escalate to privileges equivalent in scope to root in that either they can be used to gain root or they provide permissions as bad as root.
+Critical vulnerabilities allow a program to escalate to privileges equivalent in scope to root.
 If an application exploits a critical vulnerability then it gains control of the device.
-Some phones can be `rooted' by enabling USB debugging and using the special privileges of the adb shell to root the device but only \daAdbEnabledPerc\dafoot\ of devices have USB debugging enabled.\footnote{This is surprisingly high as \daAdbEnabledPerc\ of Android users are not developers.}
+Some phones can be `rooted' by enabling USB debugging and using the special privileges of the adb shell to root the device but only \daAdbEnabledPerc\dafoot\textsuperscript{,}\footnote{Where possible we indicate the uncertainty in our results by giving them $\pm$ the standard deviation.
+This is explained further in \S\ref{sec:uncertainty} but occasionally results in $\pm0$.} of devices have USB debugging enabled.
 This is not something that applications running on the phone can exploit to break compartmentalisation and so we do not include those vulnerabilities.
@@
 On some versions of Android (below 4.1) they could use known system to root escalation mechanisms but on all versions they have a greatly increased attack area for further privilege escalation and also have the ability to control all user internet traffic (via VPNs), brick the phone, remove and install apps, steal user credentials, read the screen and make as well as receive calls.
-\todo{break vulnerabilities down by attack vector?}
+\todolater{break vulnerabilities down by attack vector?}
 \avoTabAndVulns
 
-We developed and maintain an open platform for filing critical vulnerabilities in a machine readable format, \url{http://androidvulnerabilities.org/}.
+We developed and maintain an open platform for filing critical vulnerabilities in a machine readable format, \texttt{\url{http://androidvulnerabilities.org/}}.
 We seeded it with data from the CVE database, vendor lists, reports from the literature and various forums.
@@
 \avo\ contains \avoNumVulnerabilities\ vulnerabilities of which \avoNumVulnAllAndroid\ affect all Android devices and \avoNumVulnSpecific\ are specific to particular devices or manufacturers.
-We only use the \daNumVulnsUsed\ vulnerabilities which it is convenient to use for this analysis because they affect all Android devices rather than particular versions and we know enough information about the vulnerability to tie it to the particular versions it affects.
+We only use the \daNumVulnsUsed\ vulnerabilities which we have sufficient data to use for this analysis because they affect all Android devices rather than particular versions and we know enough information about the vulnerability to tie it to the particular versions it affects.
 We have published full details of the \daNumVulnsUsed\ vulnerabilities used for this analysis in an accompanying technical report~\cite{TODO} and summarise them in Table~\ref{tab:andvulns}.
 
-Tracking vulnerabilities is a manual task as they are not all in other databases such as the CVE database and the lack of a widely acknowledged unique identifier made identifying whether two reports referenced the same vulnerability difficult.
+Tracking vulnerabilities is a manual task as they are not consistently recorded in other databases such as the CVE database and the lack of a widely acknowledged unique identifier made identifying whether two reports referenced the same vulnerability difficult.
 Previous work has assumed ``any security issue of relevance will eventually get a CVE number assigned"~\cite{Frei2010} which is currently not the case for critical Android vulnerabilities.
@@
 
-The key events in the lifetime of a vulnerability are:
+The key events in the lifetime of a vulnerability do not always occur in the same order and are:
 
@@
 Unfortunately before we collated this information much of it was not published by an independent source and lacked risk rating information, even months or years after they had been actively used.
-Hence we have collated this information and made it publicly available.
+Therefore this measure does not work.
 
-Symantec's 2012 analysis~\cite{Bilge2012} of desktop malware has shown that after public disclosure exploitation rates increase by 5 orders of magnitude and so from the point of view of widespread danger, the period between public disclosure and the update reaching the user's device is the most critical.
+Symantec's 2012 analysis of desktop malware has shown that after public disclosure exploitation rates increase by 5 orders of magnitude~\cite{Bilge2012} and so from the point of view of widespread danger, the period between public disclosure and the update reaching the user's device is the most critical.
 However they also show that zero day vulnerabilities are typically used for 312 days before they are publicly disclosed, often to target particular organisations.
@@
 So for our vulnerability calculations we use the earliest date we can find recorded evidence that the vulnerability was known about, even if that knowledge might have been confined to a particular manufacturer or hobbyist.
-We cannot know if someone reported the vulnerability to the manufacturer and we do know that various agencies are engaged in widespread monitoring of communications and compromise of civilian infrastructure~\cite{TODO} and so they may know.
+%We cannot know if someone reported the vulnerability to the manufacturer and we do know that various agencies are engaged in widespread monitoring of communications and compromise of civilian infrastructure~\cite{TODO} and so they may know.
 
@@
 Some vulnerabilities (RageAgainstTheCage adb, levitator, zergRush abd TwerkMyMoto) were fixed in released versions of Android before they were discovered and so are shown as vertical lines while others were known for months before a version of Android which fixed them was shipped.
-During the period in which \da\ data was collected the date when a version of Android with the fix was observed on a \da\ device is taken as the date that version was released, before that the release date as best as we can determine is used.
+During the period in which \da\ data was collected the date when a version of Android with the fix was observed on a \da\ device is taken as the date that version was released.
+For vulnerabilities prior to the collection period the release date as best as we can determine is used.
 There is no canonical source of Android release dates, our best guesses and supporting references are available from \avo.
@@
 This data shows a large gap in the middle where we do not know of any discoveries of critical vulnerabilities affecting all of Android.
-The cause of this quiet period is unclear, perhaps from Figure~\ref{fig:proportioninsecure} the reason was that since most devices were exposed to known vulnerabilities there was no point in looking for new ones?
-Or that manufacturers made it easier to install custom versions of Android, reducing the need for users to root their devices?
-Or that manufacturer specific vulnerabilities (which we are ignoring) proved to be easier to find and so people looking for vulnerabilities adjusted their focus?
+The cause of this quiet period is unclear.
+Possible explanations are: That since most devices were exposed to known vulnerabilities there was no point in looking for new ones (from Figure~\ref{fig:proportioninsecure}).
+That manufacturers made it easier to install custom versions of Android, reducing the need for users to root their devices?
+That manufacturer specific vulnerabilities (which we are ignoring) proved to be easier to find and so people looking for vulnerabilities adjusted their focus.
 
@@
 There is a complex Android ecosystem which creates and distributes updates to Android which fix vulnerabilities.
-It is also hard to quantify due to lack of public information on aspects of the ecosystem such as the number of manufacturers and updates.
-We describe and quantify this ecosystem.
+In this section we rectify the lack of public information about the size of the Android ecosystem.
 
@@
  Numbers on edges indicate updates shipped between \daStartDate\ and \daEndDate, numbers in brackets represent number of such entities in our data.
- Dotted arrows indicate flows where we do not know how many updates are being produced as we can't measure those flows directly as they are not public.}
+ Dotted arrows indicate flows where we do not know how many updates are being produced as we can't measure those flows directly as they are not public.\todolater{Turn into a sankey diagram}}
  \label{fig:update_ecosystem}
@@
 
-The number of updates received by devices (\daNumFullVersions) is the number of different full version strings observed in \da.
+In Figure~\ref{fig:update_ecosystem} the number of updates received by devices (\daNumFullVersions) is the number of different full version strings observed in \da.
 The number of updates shipped by Google (\daNumSigOSVersions) is the number of Android versions reported in \da\ which affected more than \daSigVersionPerc\ of devices for more than \daSigVersionDays\ days.
@@
 Wheeler's \texttt{sloccount}.}
-We analysed the top \avoNumBigExternalProjects\ by lines of code (contributing \avoNumBigExternalLinesOfCode\ lines, \avoBigExternalLinesOfCodePerc\ of the total) and were able to automatically extract the versions of those projects included in different versions of Android for \avoNumAnalysedExternalProjects\ (contributing \avoNumAnalysedExternalLinesOfCode\ lines, \avoAnalysedExternalLinesOfCodePerc\ of the total).
+We analysed the top \avoNumBigExternalProjects\ by lines of code (\avoBigExternalLinesOfCodePerc\ of the total) and were able to automatically extract the versions of those projects included in different versions of Android for \avoNumAnalysedExternalProjects\ of these (\avoAnalysedExternalLinesOfCodePerc\ of the total).
 We found \avoBigExternalTotalVersions\ distinct versions, a median of \avoBigExternalMedianVersions\ and mean of \avoBigExternalMeanVersions\ versions per project.
-This shows that Android rarely changes the version of external projects it includes.
+Android rarely changes the version of external projects it includes.
 
@@
 This allows us to compute the latency between an upstream project being released and it being included in Android, this is shown in Table~\ref{tab:update_ecosystem}.
-This shows that the versions included in Android were about half a year old when that version of Android was released.
+The versions included in Android were about half a year old when the first version of Android containing it was released.
 \begin{table}
@@
   \includegraphics[width=\columnwidth]{figures/device-data-all-os}
-  \caption{OS versions over time}
+  \caption{OS versions over time. Black vertical lines indicate that the build number changed without changing the OS version at that point.}
   \label{fig:device_data_os}
@@
 \subsection{Uncertainty propagation}
+\label{sec:uncertainty}
 We have endeavoured to propagate information on uncertainties in our results through all calculations with the standard deviation indicated by $\pm$.

### 2014-11-05 figures/update_figures.sh
Update the figures

@@ -30,6 +30,7 @@ cp "$fromdir"ua-versions-android-month-norm.pdf ua-versions-android-month-norm.p
 cp "$fromdir"ua-versions-ios-month-norm.pdf ua-versions-ios-month-norm.pdf
 cp "$fromdir""model/Galaxy Nexus/norm_full.pdf" "full_version_Galaxy Nexus.pdf"
 cp "$fromdir""model/HTC Desire HD A9191/norm_full.pdf" "full_version_HTC Desire HD A9191.pdf"
+cp "$fromdir""model/Symphony W68/norm_full.pdf" "full_version_Symphony W68.pdf"
 cp "$fromdir"vulnerabilities_timeline.pdf vulnerabilities_timeline.pdf
 cp "$fromdir"frvh_os_versions.pdf frvh_os_versions.pdf
 cp "$fromdir"frvh_api_versions.pdf frvh_api_versions.pdf

### 2014-11-06 security_scoring.tex
Fixes from acr31's comments

@@
 We propose a method to score a manufacturer, operator or device based on its historic performance at keeping devices up to date and fixing security vulnerabilities.
-We find that Android as a whole gets a score of \daSecurityScore / 10, the best manufacturer is \daSecScoreBestmanufacturer\ (\daSecScoreBestmanufacturerScore\ / 10) and the worst is \daSecScoreWorstmanufacturer\ (\daSecScoreWorstmanufacturerScore\ / 10).
+We find that Android as a whole gets a score of \daSecurityScore out of 10, the highest scoring manufacturer is \daSecScoreBestmanufacturer\ (\daSecScoreBestmanufacturerScore\ out of 10) and the lowest scoring is \daSecScoreWorstmanufacturer\ (\daSecScoreWorstmanufacturerScore\ out of 10).
 
@@
 Computing how good a particular manufacturer or device model is from a security standpoint is difficult as it depends on a number of factors which are hard to observe, particularly on a large scale.
-Ideally we would consider both prevalence of potential problems which were not exploited and actual security failures.\footnote{A perfectly secure operating system would among other things detect and prevent all phishing attacks, provide perfect principle of least privilege isolation, not have any vulnerabilities, instantly fix all discovered vulnerabilities, not allow any user data to be used in a way which the user does not approve of and be really easy for an ordinary person to use.}
+Ideally we would consider both prevalence of potential problems which were not exploited and actual security failures.
+%\footnote{A perfectly secure operating system would among other things detect and prevent all phishing attacks, provide perfect principle of least privilege isolation, not have any vulnerabilities, instantly fix all discovered vulnerabilities, not allow any user data to be used in a way which the user does not approve of and be really easy for an ordinary person to use.}
 However in the absence of such data we propose a scheme for assigning a device a score out of ten based on data which can be observed, and which hopefully correlates with the actual security of the devices.
 
-This score is computed from several components:
+The FUM score is computed from several components:
 \begin{description}
-  \item[$f$] The proportion of the time where devices were $f$ree from known critical vulnerabilities.
-This is equivalent to Acer and Jackson's proposal~\cite{Acer2010} to measure the security based on the proportion of users with at least one unpatched critical vulnerability and similar to the Vulnerability Free Days (VFD) score~\cite{Wright2014}.
+  \item[free $f$] The proportion of the time where devices were free from known critical vulnerabilities.
+This is equivalent to Acer and Jackson's proposal to measure the security based on the proportion of users with at least one unpatched critical vulnerability~\cite{Acer2010} and similar to the Vulnerability Free Days (VFD) score~\cite{Wright2014}.
   Unlike VFD this is the proportion of running devices which were free from critical vulnerabilities, rather than days which the manufacturer was free from outstanding critical vulnerabilities as that does not take account of the update process.
-  \item[$u$] The proportion of devices which run the latest version of Android shipped to any device produced by that manufacturer.
-This is a measure of internal $u$pdatedness, a low score would mean many devices are being left behind.
+  \item[update $u$] The proportion of devices which run the latest version of Android shipped to any device produced by that manufacturer.
+This is a measure of internal updatedness, a low score would mean many devices are being left behind.
   This assumes that newer versions are better with stronger security, historically steps have been taken to improve Android security in newer versions so this assumption should generally hold, but sometimes new updates introduce new vulnerabilities.
-  \item[$m$] The $m$ean number of outstanding vulnerabilities affecting devices not fixed on any device shipped by the manufacturer.
+  \item[mean $m$] The mean number of outstanding vulnerabilities affecting devices not fixed on any device shipped by the manufacturer.
 This is related to the Median Active Vulnerabilities (MAV) measure~\cite{Wright2014} but is the mean rather than the median, as this gives a continuous value.
@@
 Only on average \daUpdatednessPerc\ of devices run the most recent version of Android.
-This gives a security score of \daSecurityScore / 10.
+This gives a security score of \daSecurityScore\ out of 10.
 \daTabSecScoressummary
@@
 \daTabSecScoresmanufacturer
-Different manufacturers have very different scores, Table~\ref{tab:sec_manufacturer} shows the scores for the \daNumSigManufacturers\ manufacturers with a significant presence in our data with \daSecScoreBestmanufacturer\ (\daSecScoreBestmanufacturerScore\ / 10) doing best and \daSecScoreWorstmanufacturer\ (\daSecScoreWorstmanufacturerScore\ / 10) doing worst.
+Different manufacturers have very different scores, Table~\ref{tab:sec_manufacturer} shows the scores for the \daNumSigManufacturers\ manufacturers with a significant presence in our data with \daSecScoreBestmanufacturer\ (\daSecScoreBestmanufacturerScore\ out of 10) scoring highest and \daSecScoreWorstmanufacturer\ (\daSecScoreWorstmanufacturerScore\ out of 10) scoring lowest.
 Manufacturers are considered significant if we have data from at least \daSigNumDevices\ devices and at least \daSigNumDeviceDays\ days of contributions.
@@
 Even within manufacturers different models can have very different update behaviours and hence security.
-Table~\ref{tab:sec_model} shows the results for the \daNumSigModels\ device models which have a significant presence by the same metric with `\daSecScoreBestmodel' (\daSecScoreBestmodelScore\ / 10) doing best and `\daSecScoreWorstmodel' (\daSecScoreWorstmodelScore\ / 10) doing worst.
-We can then test whether this seems fair by comparing the version data for the best and worst models.
+Table~\ref{tab:sec_model} shows the results for the \daNumSigModels\ device models which have a significant presence by the same metric with `\daSecScoreBestmodel' (\daSecScoreBestmodelScore\ out of 10) scoring highest and `\daSecScoreWorstmodel' (\daSecScoreWorstmodelScore\ out of 10) scoring lowest.
+We can then test whether this seems fair by comparing the version data for the highest and lowest scoring models.
 Figure~\ref{fig:worst_model_full} shows the full version distribution for `\daSecScoreWorstmodel' and Figure~\ref{fig:best_model_full} shows the same for `\daSecScoreBestmodel' there is a marked difference in the distribution of updates between the two with only \daSecScoreWorstmodelNumFullVersions\ full versions for `\daSecScoreWorstmodel' and \daSecScoreBestmodelNumFullVersions\ for `\daSecScoreBestmodel'.
@@
  \begin{subfigure}[b]{\columnwidth}
+  \includegraphics[width=\columnwidth]{figures/full_version_\daSecScoreBestmodel.pdf}
+  \caption{\daSecScoreBestmodel}
+  \label{fig:best_model_full}
+ \end{subfigure}
+ \begin{subfigure}[b]{\columnwidth}
   \includegraphics[width=\columnwidth]{figures/full_version_\daSecScoreWorstmodel.pdf}
@@
  \end{subfigure}
-\begin{subfigure}[b]{\columnwidth}
- \includegraphics[width=\columnwidth]{figures/full_version_\daSecScoreBestmodel.pdf}
- \caption{\daSecScoreBestmodel}
- \label{fig:best_model_full}
-\end{subfigure}
-\caption{Full version distributions for the best and worst model}
+ \caption{Full version distributions for the highest and lowest scoring model}
 \end{figure}
@@
 \daTabSecScoresoperator
-We also analysed different operators, for the \daNumSigOperators\ operators with a significant presence in our data Table~\ref{tab:sec_operator} shows the results `\daSecScoreBestoperator' (\daSecScoreBestoperatorScore\ / 10) doing best and `\daSecScoreWorstoperator' (\daSecScoreWorstoperatorScore\ / 10) doing worst.
+We also analysed different operators, for the \daNumSigOperators\ operators with a significant presence in our data Table~\ref{tab:sec_operator} shows the results `\daSecScoreBestoperator' (\daSecScoreBestoperatorScore\ out of 10) scoring highest and `\daSecScoreWorstoperator' (\daSecScoreWorstoperatorScore\ out of 10) scoring lowest.
 However the score of an operator is affected by the manufacturers of the devices which are in use on it's network.

### 2014-11-06 securityupdates.tex
Fixes from acr31's comments

@@
 On average only \daUpdatesPerYear\ updates reach a device each year and \daPercUpdatesDowngrades\ of version changes are downgrades to older versions.
-Nexus devices are better than others, \emph{\daSecScoreBestmanufacturer} to be the best manufacturer and \emph{\daSecScoreBestoperator} to be the best network operator and modelling the uptake of new versions of Android.
+Nexus devices are better than others, \emph{\daSecScoreBestmanufacturer} to be the highest scoring manufacturer and \emph{\daSecScoreBestoperator} to be the highest scoring network operator.
 %\todo{comparison of nexus update rates with iOS}
@@
 For each vulnerability it shows the proportion of devices exposed to that vulnerability and how that changes over time.
-The variation of the proportion of devices affected by a vulnerability with time tells us how bad a particular vulnerability affected the Android platform.
+The variation of the proportion of devices affected by a vulnerability with time tells us how badly a particular vulnerability affected the Android platform.
 In July 2011 at the beginning of the \da\ data the \emph{exploid}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/exploid_udev}} and \emph{levitator}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/levitator}} vulnerabilities both affect most Android devices, slowly these are fixed as updates roll out and devices are replaced until in January 2013 a much smaller proportion of devices are affected by known vulnerabilities.
@@
 \subsubsection{Method}
-As in \ref{sec:exp:android_ecosystem}, for each device we used the version information from each day to calculate which vulnerabilities it was vulnerable to that day.
+As in Experiment 1, for each device we used the version information from each day to calculate which vulnerabilities it was vulnerable to that day.
 We also recorded when the version changed and which versions it changed from and to.
-Since in the \da\ data many devices only contribute data for a short period, although we have over 1000 devices contributing data in any particular week we do not observe all the upgrades made by a large number of devices, but rather a hopefully representative sample of upgrades which happened when \da\ was installed on the devices.
-\da\ cannot distinguish between a device being replaced and the \da\ app being removed as while multiple devices can be linked to the same user account, few users do that.
+In the \da\ data we have over 1000 devices contributing data in any particular week.
+However most devices only contribute for a short period of time and so we do not observe updates happening on every device.
+Instead we have a hopefully representative sample of upgrades which happened while \da\ was installed on the devices.
+\da\ cannot distinguish between a device being replaced and the \da\ app being removed as while multiple devices can be linked to the same user account, that is a manual process and few users do it.
 
@@
 It shows how some devices had vulnerabilities, which were fixed, and then further vulnerabilities were discovered, and for these devices, mostly not fixed.
-This implies that these devices had been abandoned and were not receiving updates which is confirmed by Figure~\ref{fig:device_data_os} which shows which OS versions those devices were running.
+This implies that these devices had been abandoned by the manufacturer and were not receiving updates which is confirmed by Figure~\ref{fig:device_data_os} which shows which OS versions those devices were running.
 Some devices start off in 2011 exposed to known security vulnerabilities and are still exposed to those and additional ones in 2014.
@@
 We recorded update events and Figure~\ref{fig:from_to_updates} shows how devices upgrade between different versions of Android.
-Mostly the dark cells are above the diagonal and are upgrades (\daNumUpdatesUpgrades).
-While many upgrades are from one version to the next version there are also a fair number (\daNumUpdatesBigUpgrades, \daPercBigUpgrades) which skip a more than \daNumUpdatesSkippedBig\ versions.
+Mostly the dark cells are are upgrades (above the diagonal) (\daNumUpdatesUpgrades).
+While many upgrades are from one version to the next version there are also a fair number (\daNumUpdatesBigUpgrades, \daPercBigUpgrades) which skip more than \daNumUpdatesSkippedBig\ versions.
 Surprisingly there are also a small number of downgrade events (\daNumUpdatesDowngrades, \daPercUpdatesDowngrades) when older versions of Android are installed on to devices.
@@
 
-The number of devices getting security updates each week, is shown in Figure~\ref{fig:weekly_security_updates}, blue shows updates which changed the Android version number from a version with known vulnerabilities to one which had fewer known vulnerabilities.
-Green indicates updates which changed the build number but not the version number and so might contain a backported fix for a vulnerability.
-Red indicates updates which did not fix security vulnerabilities (because there were no known security vulnerabilities in the version of Android they were already running).
+The number of devices getting security updates each week, is shown in Figure~\ref{fig:weekly_security_updates}.
+\begin{itemize}
+ \item Updates which changed the Android version number from a version with known vulnerabilities to one which had fewer known vulnerabilities are shown in red.
+ \item Updates which changed the build number but not the version number and so might contain a backported fix for a vulnerability are shown in yellow.
+\item Updates which did not fix security vulnerabilities (because there were no known security vulnerabilities in the version of Android they were already running) are shown in green.
+
+\end{itemize}
 
@@
 Unfortunately there is no authoritative source of OS version information and so we cannot check directly whether our data is representative.
-However Google has published API version information every month since December 2009 and we have collated this information.\footnote{\url{https://www.cl.cam.ac.uk/~drt24/android/versiondistribution/}}
+However Google has published API version information every month since December 2009 and we have collated this information.\footnote{\url{http://androidvulnerabilities.org/play/historicplaydashboard}}
 While API versions are too coarse grained to use for security update detection they are closely related to OS versions and so if the \da\ data on API versions is similar to the Google Data on API versions then the \da\ data on OS versions should be representative.
@@
 This is responsible for the sudden change in distribution of OS versions in \da\ shown towards the end of Figures \ref{fig:norm_os}, \ref{fig:nvulnerabilities_heat} and \ref{fig:proportioninsecure} as for that period Bangladeshi users contributed over half of the \da\ data.
-We could normalise out the influence of the Bangladeshi users, however they are Android users too and they also need security updates.
+%We could normalise out the influence of the Bangladeshi users, however they are Android users too and they also need security updates.
 
@@
 We have endeavoured to propagate information on uncertainties in our results through all calculations with the standard deviation indicated by $\pm$.
-However we have not propagated information on systematic errors as we are not certain what those are.
+However this technique is unable to account for systematic errors.
 
@@
 They also found that 4 of the 46 malware samples (8\%) they analysed contained root exploits, much lower than rates found in later larger studies which found rates of 36.7\%~\cite{Zhou2012b} and 40\%~\cite{Zhou2012a}.
-Perhaps the malware authors heard about the results in Felt et al.'s paper?
 Our results show root exploits are still a severe threat.
 
-Finding vulnerabilities is frequently a time consuming manual process but, Brumley et al.~\cite{Brumley2008} showed that it is possible to automatically generate exploits from binary patches.
+Finding vulnerabilities is frequently a time consuming manual process but, Brumley et al. showed that it is possible to automatically generate exploits from binary patches~\cite{Brumley2008}.
 In a similar way some of the vulnerabilities\todolater{which?} which have been exploited were publicly found by examining the commit made to the Android which fixed the vulnerability.
-This effect has been observed before~\cite{Barth2011} in the Firefox source code repositories, Google does not release the source code until after the release of the update reducing this effect to the delay between an update being released and it reaching the last end user device.
+This effect has been observed before in the Firefox source code repositories~\cite{Barth2011}, Google does not release the source code until after the release of the update reducing this effect to the delay between an update being released and it reaching the last end user device.
 There have been other attempts to automatically find vulnerabilities.
-The Woodpecker tool~\cite{Grace2012} automatically finds permission leaks in stock Android phone images.
-The update process itself can allow Android apps to gain privilege through `Pileup' vulnerabilities~\cite{Xing2014} by registering for new permissions before the update which creates that permission is installed.
+The Woodpecker tool automatically finds permission leaks in stock Android phone images~\cite{Grace2012}.
+The update process itself can allow Android apps to gain privilege through `Pileup' vulnerabilities by registering for new permissions before the update which creates that permission is installed~\cite{Xing2014}.
 
@@
 Various attempts have been made to detect malware.
-RiskRanker~\cite{Grace2012a} classified 3,281/118,318 (2.8\%) apps as risky of which 718 (22\%) were malware and 322 (10\%) were previously unknown malware, an infection rate of 0.6\% across multiple markets.
+RiskRanker classified 3,281/118,318 (2.8\%) apps as risky of which 718 (22\%) were malware and 322 (10\%) were previously unknown malware, an infection rate of 0.6\% across multiple markets~\cite{Grace2012a}.
 It uses signatures of exploits, static analysis of calls to high risk operations, detection of encrypted native code execution and dynamic Dalvik code loading to detect risky apps.
-DroidRanger~\cite{Zhou2012a} found 148/182,823 (0.08\%) apps to be malicious across multiple markets of which 29 were previously unknown.
+DroidRanger found 148/182,823 (0.08\%) apps to be malicious across multiple markets of which 29 were previously unknown~\cite{Zhou2012a}.
 It used permission-based behavioural fingerprinting -- looking at the permissions of known malware and heuristic-based filtering -- dynamic loading of both Dalvik and native code.
-AnDarwin~\cite{Crussell2013} detects similar apps, it found 169/265,359 (0.06\%) malicious apps based on the difference in permissions between the malicious app and the app which had been cloned.
+AnDarwin detects similar apps, it found 169/265,359 (0.06\%) malicious apps based on the difference in permissions between the malicious app and the app which had been cloned~\cite{Crussell2013}.
 It used clustering based on semantic vectors derived from the program dependence graphs to detect similar apps.
-The Malware Genome project~\cite{Zhou2012b} collected 1,260 malware samples from 2010--2011.
+The Malware Genome project collected 1,260 malware samples from 2010--2011~\cite{Zhou2012b}.
 They found that the best case for anti-malware detection on this data was 79.6\%.
-However DroidChameleon~\cite{Rastogi2013} found that antivirus products could not detect malware if it was simply automatically permuted.
+However DroidChameleon found that antivirus products could not detect malware if it was simply automatically permuted~\cite{Rastogi2013}.
 All these app analysis projects have been hampered by the difficulty of obtaining full datasets of Android apps as Google does not make these available (and researchers who automatically download them, violate Google's Terms of Service).
-PlayDrone~\cite{Viennot2014} was a particularly effective project which circumvented Google's protective measures and downloaded over 1,100,000 apps from Google Play, allowing an in depth analysis.
+PlayDrone was a particularly effective project which circumvented Google's protective measures and downloaded over 1,100,000 apps from Google Play, allowing an in depth analysis~\cite{Viennot2014}.
 
@@
 \label{sec:conclusion}
-We have investigate the security of the Android operating system with respect to security updates and examined how the Android ecosystem results in different devices receiving different levels of security due to whether or not they get security updates.
+We have investigated the security of the Android operating system with respect to security updates and examined how the Android ecosystem results in different devices receiving different levels of security due to whether or not they get security updates.
 We have compared different models, manufacturers and network operators and found that there are differences between them which the discerning purchaser might use to influence their decision about which device to buy from which manufacturer and operator.

### 2014-11-06 securityupdates.tex
Expand the discussion

@@
 
+Malicious apps which use exploits to take control of an Android device must be installed on the device and to an extent this can be prevented if all apps are downloaded from the Google Play store.
+It is protected by Bouncer's static analysis and prompt takedowns of apps reported to be malicious and has been found to have infection rates of XXX~\cite{TODO} and XXX~\cite{TODO}.
+However some apps are installed from other markets which are less well policed (infection rates of XXX~\cite{TODO}, XXX~\cite{TODO}).
+Apps can also dynamically download code at runtime which then cannot be detected by static analysis and so bypasses Bouncer.
+
+There are also other ways that malicious code could end up running on Android devices, the addJavascriptInterface vulnerability allows javascript running in a webview which has an expose interface to execute arbitrary Java code and hence load exploits.
+While fixes have been produced for this the change is API breaking and so to be safe an app must target an API version greater than XXX and be running on a device with an API version greater than XXX.
+Since the html loaded into webviews (for example to display adverts) is often loaded over insecure http, malicious javascript can easily be injected by an attacker on the network.
+
+We have not included all the vulnerabilities from \avo\, in particular we have not included any manufacturer specific vulnerabilities, even when these are widespread (such as vulnerabilities affecting Qualcomm\footnote{Qualcomm is particularly good at publicly disclosing the vulnerabilities affecting their code and the patches which fix them.\url{https://www.codeaurora.org/projects/security-advisories/}} chipsets) because it is difficult to work out which devices are affected.
+There also tends to be less public information available about manufacturer specific vulnerabilities which makes them harder to tie down.
+For other vulnerabilities which might affect all Android devices it is hard to work out which devices they affect.
+For example \emph{pty race}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/pty_race}} is a Linux kernel vulnerability, we found 48 commits which fixed it in different branches, it was also accidentally fixed and later reintroduced and so determining which devices were vulnerable relies on knowing where on which branch the kernel was taken from.
+For some vulnerabilities the corresponding patch is not in AOSP despite some manufacturers having shipped builds containing the fix, which makes working out what is going on more difficult.
+
 \subsection{Open questions}

### 2014-11-07 avostats.tex
Update figures and make stats cheked in files

\newcommand{\avoNumSubmitters}{7}
\newcommand{\avoTotalExternalLines}{25 Million}
\newcommand{\avoNumExternalProjects}{176}
\newcommand{\avoNumBigExternalLinesOfCode}{25 Million}
\newcommand{\avoBigExternalLinesOfCodePerc}{99.69\%}
\newcommand{\avoNumVulnerabilities}{31}
\newcommand{\avoNumVulnAllAndroid}{15}
\newcommand{\avoNumVulnSpecific}{16}
\newcommand{\avoStartDate}{2013-08-28}
\newcommand{\avoEndDate}{2014-10-17}
\newcommand{\avoFirstDataDate}{2010-07-13}
\newcommand{\avoLastDataDate}{2014-07-29}
\newcommand{\avoVulnsPerYear}{$7.7 \pm 1.4$}
\newcommand{\avoVulnsPerYearAllAndroid}{$3.7 \pm 1.0$}
\newcommand{\avoTabAndVulns}{\begin{table} \centering \begin{tabular}{l|c|l} Vulnerability & Date known & How known \\ \hline KillingInTheNameOf psneuter ashmem & 2010-07-13 & Fixed on \\ exploid udev & 2010-07-15 & Discovered on \\ RageAgainstTheCage adb & 2010-08-21 & Discovered on \\ levitator & 2011-03-10 & Discovered on \\ Gingerbreak & 2011-04-18 & Fixed on \\ zergRush & 2011-10-06 & Discovered on \\ APK duplicate file & 2013-02-18 & Discovered on \\ APK unchecked name & 2013-06-30 & Discovered on \\ APK unsigned shorts & 2013-07-03 & Fixed on \\ keystore buffer & 2013-09-09 & Discovered on \\ vold asec & 2014-01-27 & Fixed on \\ Fake ID & 2014-04-17 & Fixed on \\ TowelRoot & 2014-05-03 & Discovered on \\\end{tabular} \caption{Root equivalent vulnerabilities in Android} \label{tab:andvulns} \end{table}}
\newcommand{\avoNumBigExternalProjects}{40}
\newcommand{\avoNumAnalysedExternalProjects}{28}
\newcommand{\avoNumAnalysedExternalLinesOfCode}{6 Million}
\newcommand{\avoAnalysedExternalLinesOfCodePerc}{24.85\%}
\newcommand{\avoBigExternalMedianVersions}{1.50}
\newcommand{\avoBigExternalMeanVersions}{$2.1 \pm 1.4$}
\newcommand{\avoBigExternalTotalVersions}{58}

### 2014-11-07 dastats.tex
Update figures and make stats cheked in files

\newcommand{\daNumDataPoints}{128 Billion}
\newcommand{\daNumStartedApps}{112632}
\newcommand{\daNumInstalledApps}{164685}
\newcommand{\daNumVulnsUsed}{13}
\newcommand{\daSigNumDevices}{100}
\newcommand{\daSigNumDays}{100}
\newcommand{\daNumManufacturers}{260}
\newcommand{\daNumSigManufacturers}{8}
\newcommand{\daNumModels}{2368}
\newcommand{\daNumSigModels}{15}
\newcommand{\daOSVersionPercValidLines}{99.47\%}
\newcommand{\daOSTotalDaysData}{1127371}
\newcommand{\daMeanInsecurityPerc}{$87 \pm 0\%$}
\newcommand{\daMeanOutstandingVulnerabilities}{0.69}
\newcommand{\daUpdatedness}{$0.058 \pm 0.000$}
\newcommand{\daSecurityScore}{$2.7 \pm 0.0$}
\newcommand{\daNumFullVersions}{1128}
\newcommand{\daNumOSVersions}{45}
\newcommand{\daNumSigOSVersions}{23}
\newcommand{\daNumAPIVersions}{14}
\newcommand{\daNumFullVersionUpdates}{4196}
\newcommand{\daNumOSVersionUpdates}{3544}
\newcommand{\daNumFullOnlyVersionUpdates}{632}
\newcommand{\daNumSecurityUpdates}{1353}
\newcommand{\daNumPossibleSecurityUpdates}{1886}
\newcommand{\daTabSecScoresmanufacturer}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & Security score \\ \hline lg & $0.11 \pm 0.00$ & $0.33 \pm 0.00$ & 0.79 & $3.3 \pm 0.0$ \\  motorola & $0.16 \pm 0.00$ & $0.13 \pm 0.00$ & 0.82 & $2.9 \pm 0.0$ \\  other & $0.13 \pm 0.00$ & $0.058 \pm 0.000$ & 0.69 & $2.7 \pm 0.0$ \\  samsung & $0.14 \pm 0.00$ & $0.048 \pm 0.000$ & 0.77 & $2.6 \pm 0.0$ \\  htc & $0.14 \pm 0.00$ & $0.088 \pm 0.001$ & 0.95 & $2.5 \pm 0.0$ \\  sony & $0.13 \pm 0.00$ & $0.18 \pm 0.00$ & 1.18 & $2.5 \pm 0.0$ \\  asus & $0.14 \pm 0.00$ & $0.56 \pm 0.01$ & 7.18 & $2.2 \pm 0.0$ \\  symphony & $0.0016 \pm 0.0002$ & $0.042 \pm 0.002$ & 5.73 & $0.15 \pm 0.00$ \\ \end{tabular} \caption{Security scores for manufacturers} \label{tab:sec_manufacturer} \end{table}}
\newcommand{\daTabSecScoresmodel}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & Security score \\ \hline Galaxy Nexus & $0.49 \pm 0.00$ & $0.54 \pm 0.01$ & 2.69 & $3.9 \pm 0.0$ \\  Nexus 4 & $0.19 \pm 0.00$ & $0.82 \pm 0.01$ & 7.22 & $3.2 \pm 0.0$ \\  Nexus 7 & $0.17 \pm 0.00$ & $0.74 \pm 0.01$ & 7.08 & $2.9 \pm 0.0$ \\  Desire HD & $0.083 \pm 0.003$ & $0.051 \pm 0.004$ & 0.39 & $2.9 \pm 0.0$ \\  other & $0.13 \pm 0.00$ & $0.12 \pm 0.00$ & 0.69 & $2.9 \pm 0.0$ \\  HTC Sensation & $0.36 \pm 0.00$ & $0.0063 \pm 0.0062$ & 2.59 & $1.9 \pm 0.0$ \\  GT-I9100 & $0.22 \pm 0.00$ & $0.017 \pm 0.001$ & 2.53 & $1.4 \pm 0.0$ \\  GT-N7000 & $0.25 \pm 0.00$ & $0.0006 \pm 0.0036$ & 3.52 & $1.2 \pm 0.0$ \\  GT-P1000 & $0.0096 \pm 0.0009$ & $0.0000 \pm 0.0076$ & 1.73 & $0.94 \pm 0.02$ \\  HTC Desire S & $0.022 \pm 0.001$ & $0.016 \pm 0.004$ & 2.00 & $0.85 \pm 0.01$ \\  GT-I9300 & $0.14 \pm 0.00$ & $0.014 \pm 0.001$ & 7.27 & $0.62 \pm 0.01$ \\  GT-I9505 & $0.0 \pm 0$ & $0.19 \pm 0.00$ & 8.23 & $0.56 \pm 0.01$ \\  GT-N7100 & $0.064 \pm 0.002$ & $0.0011 \pm 0.0058$ & 8.37 & $0.26 \pm 0.02$ \\  HTC Desire HD & $0.0 \pm 0$ & $0.0005 \pm 0.0070$ & 4.05 & $0.10 \pm 0.02$ \\ \end{tabular} \caption{Security scores for models} \label{tab:sec_model} \end{table}}
\newcommand{\daTabSecScoressummary}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & Security score \\ \hline nexus & $0.33 \pm 0.00$ & $0.52 \pm 0.00$ & 1.74 & $3.8 \pm 0.0$ \\  notnexus & $0.12 \pm 0.00$ & $0.041 \pm 0.000$ & 0.69 & $2.6 \pm 0.0$ \\ \end{tabular} \caption{Security scores for summaries} \label{tab:sec_summary} \end{table}}
\newcommand{\daUpdatednessPerc}{$5.8 \pm 0.0\%$}
\newcommand{\daNumOSDevices}{18809}
\newcommand{\daSigNumDevicesDay}{20}
\newcommand{\daSigNumDeviceDays}{10000}
\newcommand{\daTabAndVulns}{\begin{table} \centering \begin{tabular}{l|c|l} Vulnerability & Date known & How known \\ \hline KillingInTheNameOf psneuter ashmem & 2010-07-13 & Fixed on \\ exploid udev & 2010-07-15 & Discovered on \\ RageAgainstTheCage adb & 2010-08-21 & Discovered on \\ levitator & 2011-03-10 & Discovered on \\ Gingerbreak & 2011-04-18 & Fixed on \\ zergRush & 2011-10-06 & Discovered on \\ APK duplicate file & 2013-02-18 & Discovered on \\ APK unchecked name & 2013-06-30 & Discovered on \\ APK unsigned shorts & 2013-07-03 & Fixed on \\ keystore buffer & 2013-09-09 & Discovered on \\ TwerkMyMoto & 2013-11-24 & Discovered on \\ vold asec & 2014-01-27 & Fixed on \\\end{tabular} \caption{Root equivalent vulnerabilities in Android} \label{tab:andvulns} \end{table}}
\newcommand{\daOSYearsOfData}{3}
\newcommand{\daSecScoreBestmanufacturer}{lg}
\newcommand{\daSecScoreWorstmanufacturer}{symphony}
\newcommand{\daSecScoreBestmodel}{Galaxy Nexus}
\newcommand{\daSecScoreWorstmodel}{HTC Desire HD A9191}
\newcommand{\daSecScoreBestsummary}{nexus}
\newcommand{\daSecScoreWorstsummary}{notnexus}
\newcommand{\daSecScoreBestmanufacturerScore}{$3.3 \pm 0.0$}
\newcommand{\daSecScoreWorstmanufacturerScore}{$0.15 \pm 0.00$}
\newcommand{\daSecScoreBestmodelScore}{$3.9 \pm 0.0$}
\newcommand{\daSecScoreWorstmodelScore}{$0.10 \pm 0.02$}
\newcommand{\daSecScoreBestsummaryScore}{$4.1 \pm 0.0$}
\newcommand{\daSecScoreWorstsummaryScore}{$2.9 \pm 0.0$}
\newcommand{\daVulnFree}{$0.13 \pm 0.00$}
\newcommand{\daAdbEnabledPerc}{$20 \pm 0\%$}
\newcommand{\daNumUpdatesUpgrades}{3311}
\newcommand{\daNumUpdatesDowngrades}{146}
\newcommand{\daPercUpdatesDowngrades}{4.22\%}
\newcommand{\daStartDate}{2011-07-01}
\newcommand{\daEndDate}{2014-10-31}
\newcommand{\daNumOperators}{1392}
\newcommand{\daTabSecScoresoperator}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & Security score \\ \hline o2 uk & $0.24 \pm 0.00$ & $0.12 \pm 0.00$ & 0.41 & $3.7 \pm 0.0$ \\  orange & $0.21 \pm 0.00$ & $0.085 \pm 0.003$ & 0.39 & $3.5 \pm 0.0$ \\  t-mobile & $0.18 \pm 0.00$ & $0.18 \pm 0.00$ & 0.54 & $3.5 \pm 0.0$ \\  sprint & $0.18 \pm 0.00$ & $0.11 \pm 0.00$ & 0.44 & $3.4 \pm 0.0$ \\  3 & $0.14 \pm 0.00$ & $0.096 \pm 0.003$ & 0.49 & $3.1 \pm 0.0$ \\  vodafone uk & $0.13 \pm 0.00$ & $0.10 \pm 0.00$ & 0.52 & $3.1 \pm 0.0$ \\  at\&t & $0.12 \pm 0.00$ & $0.077 \pm 0.002$ & 0.51 & $3.0 \pm 0.0$ \\  verizon & $0.17 \pm 0.00$ & $0.086 \pm 0.002$ & 0.89 & $2.7 \pm 0.0$ \\  unknown & $0.095 \pm 0.001$ & $0.22 \pm 0.00$ & 1.16 & $2.5 \pm 0.0$ \\  n telenor & $0.023 \pm 0.001$ & $0.10 \pm 0.00$ & 1.59 & $1.4 \pm 0.0$ \\  airtel & $0.067 \pm 0.002$ & $0.029 \pm 0.002$ & 2.00 & $1.1 \pm 0.0$ \\  grameenphone & $0.0080 \pm 0.0005$ & $0.016 \pm 0.001$ & 1.99 & $0.80 \pm 0.00$ \\  banglalink & $0.00016 \pm 0.00011$ & $0.029 \pm 0.002$ & 2.89 & $0.40 \pm 0.01$ \\  robi & $\left(8.5 \pm 8.5\right) \times 10^{-5}$ & $0.042 \pm 0.003$ & 3.33 & $0.33 \pm 0.01$ \\ \end{tabular} \caption{Security scores for operators} \label{tab:sec_operator} \end{table}}
\newcommand{\daSecScoreBestoperator}{o2 uk}
\newcommand{\daSecScoreBestoperatorScore}{$3.7 \pm 0.0$}
\newcommand{\daSecScoreWorstoperator}{robi}
\newcommand{\daSecScoreWorstoperatorScore}{$0.33 \pm 0.01$}
\newcommand{\daNumSigOperators}{14}
\newcommand{\daMonthsDevices}{1545}
\newcommand{\daMonths}{6}
\newcommand{\daSigVersionPerc}{1\%}
\newcommand{\daSigVersionDays}{10}
\newcommand{\daNumDeviceDataDevices}{50}
\newcommand{\daNumSigFullVersions}{83}
\newcommand{\daSecScoreBestmanufacturerNumFullVersions}{104}
\newcommand{\daSecScoreworstmanufacturerNumFullVersions}{239}
\newcommand{\daSecScoreBestmodelNumFullVersions}{46}
\newcommand{\daSecScoreworstmodelNumFullVersions}{2}
\newcommand{\daSecScoreBestsummaryNumFullVersions}{73}
\newcommand{\daSecScoreworstsummaryNumFullVersions}{1057}
\newcommand{\daSecScoreBestoperatorNumFullVersions}{87}
\newcommand{\daSecScoreworstoperatorNumFullVersions}{89}
\newcommand{\daSecScoreWorstmanufacturerNumFullVersions}{14}
\newcommand{\daSecScoreWorstmodelNumFullVersions}{2}
\newcommand{\daSecScoreWorstsummaryNumFullVersions}{1103}
\newcommand{\daSecScoreWorstoperatorNumFullVersions}{33}
\newcommand{\daFullDeployedAt}{95\%}
\newcommand{\daOSCurveFitParamFirst}{86.43}
\newcommand{\daOSCurveFitParamSecond}{\num{0.00315}}
\newcommand{\daOSCurveFitRMSE}{0.11}
\newcommand{\daAPICurveFitParamFirst}{110.29}
\newcommand{\daAPICurveFitParamSecond}{\num{0.00316}}
\newcommand{\daAPICurveFitRMSE}{0.12}
\newcommand{\daOSCurvePolyRMSE}{0.11}
\newcommand{\daOSCurveSplineRMSE}{0.11}
\newcommand{\daOSCurveHalfDeployed}{$\left(3.1 \pm 0.8\right) \times 10^{2}$}
\newcommand{\daOSCurveFullDeployed}{$\left(0.0 \pm 2.4\right) \times 10^{5}$}
\newcommand{\daAPICurvePolyRMSE}{0.12}
\newcommand{\daAPICurveSplineRMSE}{0.12}
\newcommand{\daAPICurveHalfDeployed}{$\left(3.3 \pm 0.8\right) \times 10^{2}$}
\newcommand{\daAPICurveFullDeployed}{$\left(0.0 \pm 2.3\right) \times 10^{5}$}
\newcommand{\daVulnAPKDuplicateFileOctoberPerc}{91.46\%}
\newcommand{\daNumUpdatesBigUpgrades}{776}
\newcommand{\daNumUpdatesSkippedBig}{3}
\newcommand{\daPercBigUpgrades}{22.45\%}
\newcommand{\daUpdatesPerYear}{$1.4 \pm 0.0$}
\newcommand{\daOSMonthsOfData}{40}

### 2014-11-07 figures/update_figures.sh

@@ -36,3 +36,5 @@ cp "$fromdir"frvh_os_versions.pdf frvh_os_versions.pdf
 cp "$fromdir"frvh_api_versions.pdf frvh_api_versions.pdf
 cp "$fromdir"frms_os_versions.pdf frms_os_versions.pdf
 cp ~/Library/papers/library.filtered.bib ../securityupdates.bib
+cp ~/git/da/graphing/stats/latex.tex ../dastats.tex
+cp ~/Documents/androidvulnerabilities/output/latex.tex ../avostats.tex

### 2014-11-07 securityupdates.tex
Tidy up some todos and missing data

@@
 \newcommand{\avo}{\texttt{androidvulnerabilities.org}}
-\newcommand{\percMarketShare}{XX\%~\cite{TODO}}
+\newcommand{\percMarketShare}{83.6\%~\footnote{\url{http://www.theinquirer.net/inquirer/news/2379036/android-hits-836-percent-marketshare-while-ios-windows-and-blackberry-slide}}}
 \newcommand{\daNumDevices}{\daNumOSDevices}
@@
 \input{countversions}
-\newcommand{\otherProjNumVersions}{XXX}
 \newcommand{\otherProjNum}{\avoNumExternalProjects}%TODO check we are doing the right calculation here and not overcounting
@@
 The proportion of Android devices in the \da\ data which were running versions of Android known to have root vulnerabilities at the time is shown in Figure~\ref{fig:proportioninsecure}.
+It shows the proportion of devices which are definitely vulnerable as their build was released before the vulnerability was discovered, those which might be insecure but could contain a backported fix and those which are secure as the version of Android they were running contained no known vulnerabilities on that date.
 
@@
 %TODO make this graph shorter
-\caption{Proportion of devices running insecure versions of Android against time\todo{Disambiguate 'definitely broken' from 'maybe broken'}}
+\caption{Proportion of devices running insecure, maybe insecure and secure versions of Android against time}
 \label{fig:proportioninsecure}
@@
  \item We measure the security of Android according to three metrics and compare different manufacturers and device models to allow device purchasers to differentiate between them based on security.
- \todo{make this more concrete}
- \todo{Do we want \avo\ to provide our latest recommendations -- legal advice required}
+ \todo{arb33: make this more concrete}
+ \todo{arb33: Do we want \avo\ to provide our latest recommendations -- legal advice required}
 \end{itemize}
@@
 \avo\ contains \avoNumVulnerabilities\ vulnerabilities of which \avoNumVulnAllAndroid\ affect all Android devices and \avoNumVulnSpecific\ are specific to particular devices or manufacturers.
-We only use the \daNumVulnsUsed\ vulnerabilities which we have sufficient data to use for this analysis because they affect all Android devices rather than particular versions and we know enough information about the vulnerability to tie it to the particular versions it affects.
+We only use the \daNumVulnsUsed\ vulnerabilities which we have sufficient data to use for this analysis because they affect all Android devices rather than particular manufacturers and we know enough information about the vulnerability to tie it to the particular versions it affects.
 We have published full details of the \daNumVulnsUsed\ vulnerabilities used for this analysis in an accompanying technical report~\cite{TODO} and summarise them in Table~\ref{tab:andvulns}.
@@
 Malicious apps which use exploits to take control of an Android device must be installed on the device and to an extent this can be prevented if all apps are downloaded from the Google Play store.
-It is protected by Bouncer's static analysis and prompt takedowns of apps reported to be malicious and has been found to have infection rates of XXX~\cite{TODO} and XXX~\cite{TODO}.
-However some apps are installed from other markets which are less well policed (infection rates of XXX~\cite{TODO}, XXX~\cite{TODO}).
+It is protected by Bouncer's static analysis and prompt takedowns of apps reported to be malicious and has been found to have infection rates of 0.02\%~\cite{Zhou2012a}.
+However some apps are installed from other markets which are less well policed (infection rates of between 0.20\%~ and 0.47\%~\cite{Zhou2012a}).
 Apps can also dynamically download code at runtime which then cannot be detected by static analysis and so bypasses Bouncer.
 
-There are also other ways that malicious code could end up running on Android devices, the addJavascriptInterface vulnerability allows javascript running in a webview which has an expose interface to execute arbitrary Java code and hence load exploits.
-While fixes have been produced for this the change is API breaking and so to be safe an app must target an API version greater than XXX and be running on a device with an API version greater than XXX.
+There are also other ways that malicious code could end up running on Android devices, the addJavascriptInterface vulnerability (CVE-2012-6636 ) allows javascript running in a webview which has an expose interface to execute arbitrary Java code and hence load exploits.
+While fixes have been produced for this the change is API breaking and so to be safe an app must target an API version greater than 17 and be running on a device with an Android version of at least 4.2.
 Since the html loaded into webviews (for example to display adverts) is often loaded over insecure http, malicious javascript can easily be injected by an attacker on the network.
@@
 Finding vulnerabilities is frequently a time consuming manual process but, Brumley et al. showed that it is possible to automatically generate exploits from binary patches~\cite{Brumley2008}.
-In a similar way some of the vulnerabilities\todolater{which?} which have been exploited were publicly found by examining the commit made to the Android which fixed the vulnerability.
+In a similar way some of the vulnerabilities\todolater{which?} in \avo\ which have been exploited were publicly found by examining the commit made to the Android which fixed the vulnerability.
 This effect has been observed before in the Firefox source code repositories~\cite{Barth2011}, Google does not release the source code until after the release of the update reducing this effect to the delay between an update being released and it reaching the last end user device.
@@
 
-\todo{Evaluate our model trained on the DA data against the ground truth Google Play data}
-
 \todo{Our handling of errors here is not quite good enough as the uncertainty from our y value interacts badly with the logarithm function when computing the inverse and we compute the log of ~0 which means our error ends up much bigger than our value - but that is an upper bound, not a +- bound.}

### 2014-11-10 securityupdates.tex
New introduction

@@
 \section{Introduction}
-Android has \percMarketShare\ of the smartphone market and while the core development is controlled by Google, we know of \daNumManufacturers\footnote{\label{foot:dadata}We computed this from the \da~\cite{Wagner2013} data, see \S\ref{sec:android_update_process}.} manufacturers which make devices which run Android.
-Android receives regular updates  which add new features and fix bugs, including security vulnerabilities.
-Many of these manufacturers customise the version of Android they ship and sometimes network operators (of which we know of \daNumOperators\dafoot) make further modifications.
-Hence when Google produces an update to Android, the update may have to pass through the manufacturer and operator before reaching the user.
-The manufacturer has little financial incentive to perform this work as they have already sold the device.
-Operators are cautious because if they ship a broken update then they will get many support requests which is expensive.
-However there is ongoing legal action to force operators to ship updates for security vulnerabilities~\cite{Soghoian2013}.\todolater{Check on the status of this legal action}
-Many phones are sold on long (12--24 month) contracts with monthly payments yet stop receiving updates before the end of the contract,\todolater{Is there a survey of the distribution of contract lengths and phone purchase methods?} whereas Windows XP could be purchased for a one-off payment in October 2001 and received security updates until April 2014.
-Corporate or public sector buyers are encouraged to purchase secure devices, but we have found little concrete guidance.
-For example, CESG, which advises the UK government on how to secure its computer systems, recommends picking Android devices from manufactures which are good at shipping security updates promptly~\cite{CESG2013} but it does not state which manufacturers are better.\todolater{Which other advisory organisations give that advice?}
-We provide this information in~\S\ref{sec:security_scoring}.
+%Android is the most popular smartphone operating system today with \percMarketShare\ marketshare.
 
-Android relies on the Linux kernel to provide compartmentalisation and runs each app as a separate Unix `user'.
-The Linux kernel is a large piece of software and so there will likely be new root exploits found in it~\cite{TODO}.
-Manufacturers of Android devices have also accidentally added further root exploits when customising it~\cite{Grace2012}.
-Between 36.7\%~\cite{Zhou2012b} and 40\%~\cite{Zhou2012a} of malware for Android contained root exploits in 2012.
-While this shows that much Android malware does not need a root exploit to work, a significant proportion does include root exploits.
-This malware is also more dangerous because while ordinary malware can be remotely removed by Google through the Play store, once a root exploit has been used this is no longer guaranteed to work.
+Support for third-party apps is a key feature of modern smartphone operating systems.
+Such apps are written by many developers from a wide range of backgrounds, which means neither operating system vendors nor users can, in general, fully trust app developers.
+Therefore, in order to secure personal data and prevent the use of expensive operations such as the sending of premium-rate text messages, operating system vendors provide a protected execution environment, or \emph{sandbox}, for apps.
 
-The proportion of Android devices in the \da\ data which were running versions of Android known to have root vulnerabilities at the time is shown in Figure~\ref{fig:proportioninsecure}.
-It shows the proportion of devices which are definitely vulnerable as their build was released before the vulnerability was discovered, those which might be insecure but could contain a backported fix and those which are secure as the version of Android they were running contained no known vulnerabilities on that date.
+In this paper we quantify the extent to which the app sandbox provided by Android protects devices from malicious apps attempting to gain control over the entire handset.
+%We have analysed Android because it is the most popular smartphone operating system by market share (~\percMarketShare) and 
+Previous research has shown that in 2012, between 36.7\%~\cite{Zhou2012b} and 40\%~\cite{Zhou2012a} of malware for Android contained exploits designed to enable a malicious app to break the sandbox.
+What is not known however is the proportion of Android devices which are vulnerable to these exploits over time.
+To the best of our knowledge, this is the first paper to provide detailed numbers.
 
-\begin{figure}[!b]
+Not all malicious apps need to break the sandbox in order to misbehave.
+Indeed, the above figures suggest the majority of malicious apps do not attempt to do so.
+We focus on this issue because malware which is able to take full control of a handset can do significantly more harm and is much harder to remove.
+Ordinary malware can uninstalled by the user or remotely by Google through the Play Store, but such removal is unlikely to work once the security features of Android are broken.
+
+The development of Android is distributed among many parties.
+The security of Android relies on many open source projects, include the Linux kernel, OpenSSL and BouncyCastle as well as Google who build the core platform. 
+In addition, manufacturers (of which we are aware of \daNumManufacturers\footnote{\label{foot:dadata}We computed this from the \da~\cite{Wagner2013} data, see \S\ref{sec:android_update_process}.}) and network operators (\daNumOperators\dafoot) may make further modifications before devices are shipped to customers.
+Understanding this ecosystem is important as manufacturers have introduced additional vulnerabilities in the past~\cite{Grace2012}. 
+We present a better understanding of both the ecosystem of Android development and associated vulnerabilities more fully in \S~X.
+
+\begin{figure}[h]
 \centering
 \includegraphics[width=\columnwidth]{figures/proportioninsecure}
-%TODO make this graph shorter
 \caption{Proportion of devices running insecure, maybe insecure and secure versions of Android against time}
@@
 
-In the past users of Microsoft IIS were advised to switch to a different web server due to the frequent discovery of vulnerabilities~\cite{Pescatore2001}.
-This pressure encouraged Microsoft to improve the security of their products~\cite{TODO}.
-Part of the reason why deployed devices do not receive updates in a timely manner is that there is little data on what is actually happening and so it is not possible to determine which manufacturers or operators are doing a better job.
-Hence there is little incentive for manufacturers and operators to do a better job.
-We provide this data  in~\S\ref{sec:security_scoring}.
+In order to determine whether an individual smartphone is vulnerable to a particular exploit, we need to combine vulnerability data with the operating system version string and additional update information from the handset.
+For this we use handset data gathered from over 18,000 participants in the \da~\cite{} project.
+This analysis allows us to plot the proportion of Android devices in the \da\ data which were running versions of Android known to have vulnerabilities at the time and is shown in Figure~\ref{fig:proportioninsecure}.
+The Figure shows the proportion of devices split into three categories: 
+(1) \emph{secure} because they are running a version of Android with no known vulnerabilities; 
+(2) \emph{insecure} because the device was running an insecure version of Android with a build released before the vulnerability was fixed or discovered; and 
+(3) \emph{maybe secure} because the device is running an insecure version of Android but it received an update which we did not see until after the vulnerability was fixed and therefore this update may contain a backport of the fix.
+Further analysis into the propagation of updates in Android are presented in \S~Y.
 
-The contributions of this paper are:
+It is clear from Figure~\ref{fig:proportioninsecure} that the security of the Android platform is far from perfect.
+One solution is regulation, and indeed there is ongoing legal action to force operators to ship updates for security vulnerabilities~\cite{Soghoian2013}.\todolater{Check on the status of this legal action}
+Indeed, many smartphones are sold on 12--24 month contracts, and yet our data shows many devices do not receive many security updates. 
+In contrast, Windows XP could be purchased for a one-off payment in October 2001 and received security updates until April 2014.
+
+An alternative to regulation is better public data. 
+Hard numbers provide an incentive to both manufacturers and operators to provide updates.
+Currently corporate and public sector buyers are encouraged to purchase secure devices, but we have found little concrete guidance on the specific makes and models which provide the best security. 
+For example, CESG, which advises the UK government on how to secure its computer systems, recommends picking Android devices from manufactures which are good at shipping security updates promptly~\cite{CESG2013} but it does not state which manufacturers are better.
+To address this, we developed a scoring system and provide numbers on the historic performance of devices in the \da\ project in~\S\ref{sec:security_scoring}.
+
+%REMOVED BY ARB
+%The manufacturer has little financial incentive to perform this work as they have already sold the device.
+%Operators are cautious because if they ship a broken update then they will get many support requests which is expensive.
+%
+%In the past users of Microsoft IIS were advised to switch to a different web server due to the frequent discovery of vulnerabilities~\cite{Pescatore2001}.
+%This pressure encouraged Microsoft to improve the security of their products~\cite{TODO}.
+
+In summary, the contributions of this paper are:
 \begin{itemize}
  \item We characterise the Android update ecosystem, showing how updates flow between entities, and quantify vulnerabilities by determining what proportion of devices they affected and how that changes over time.
- \item We collate information on vulnerabilities which affect Android and investigate their time-lines.
+ \item We collate quantitive information on vulnerabilities which affect Android and investigate their time-lines.
  \item We measure the security of Android according to three metrics and compare different manufacturers and device models to allow device purchasers to differentiate between them based on security.
- \todo{arb33: make this more concrete}
- \todo{arb33: Do we want \avo\ to provide our latest recommendations -- legal advice required}
 \end{itemize}

### 2014-11-10 securityupdates.tex
Talk about the two vulnerabilities we have not included due to ambiguity about severity

@@
 For some vulnerabilities the corresponding patch is not in AOSP despite some manufacturers having shipped builds containing the fix, which makes working out what is going on more difficult.
+For \emph{RageAgainstTheCage adb}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/RageAgainstTheCage_adb}} and \emph{keystore buffer}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/keystore_buffer}} vulnerabilities which we do have sufficient data it is not clear whether they are truly critical vulnerabilities as the former may require physical ADB access and the latter may be protected by a sandbox.
+However those two vulnerabilities make little difference to our results as they affect few devices.
 

### 2014-11-10 avostats.tex
Update figures and data for tweaked vulnerability list

@@
 \newcommand{\avoLastDataDate}{2014-07-29}
+\newcommand{\avoVulnsPerYearNominal}{7.66}
 \newcommand{\avoVulnsPerYear}{$7.7 \pm 1.4$}
+\newcommand{\avoVulnsPerYearAllAndroidNominal}{3.71}
 \newcommand{\avoVulnsPerYearAllAndroid}{$3.7 \pm 1.0$}
-\newcommand{\avoTabAndVulns}{\begin{table} \centering \begin{tabular}{l|c|l} Vulnerability & Date known & How known \\ \hline KillingInTheNameOf psneuter ashmem & 2010-07-13 & Fixed on \\ exploid udev & 2010-07-15 & Discovered on \\ RageAgainstTheCage adb & 2010-08-21 & Discovered on \\ levitator & 2011-03-10 & Discovered on \\ Gingerbreak & 2011-04-18 & Fixed on \\ zergRush & 2011-10-06 & Discovered on \\ APK duplicate file & 2013-02-18 & Discovered on \\ APK unchecked name & 2013-06-30 & Discovered on \\ APK unsigned shorts & 2013-07-03 & Fixed on \\ keystore buffer & 2013-09-09 & Discovered on \\ vold asec & 2014-01-27 & Fixed on \\ Fake ID & 2014-04-17 & Fixed on \\ TowelRoot & 2014-05-03 & Discovered on \\\end{tabular} \caption{Root equivalent vulnerabilities in Android} \label{tab:andvulns} \end{table}}
+\newcommand{\avoTabAndVulns}{\begin{table} \centering \begin{tabular}{l|c|l} Vulnerability & Date known & How known \\ \hline KillingInTheNameOf psneuter ashmem & 2010-07-13 & Fixed on \\ exploid udev & 2010-07-15 & Discovered on \\ levitator & 2011-03-10 & Discovered on \\ Gingerbreak & 2011-04-18 & Fixed on \\ zergRush & 2011-10-06 & Discovered on \\ APK duplicate file & 2013-02-18 & Discovered on \\ APK unchecked name & 2013-06-30 & Discovered on \\ APK unsigned shorts & 2013-07-03 & Fixed on \\ vold asec & 2014-01-27 & Fixed on \\ Fake ID & 2014-04-17 & Fixed on \\ TowelRoot & 2014-05-03 & Discovered on \\\end{tabular} \caption{Root equivalent vulnerabilities in Android} \label{tab:andvulns} \end{table}}
 \newcommand{\avoNumBigExternalProjects}{40}
@@
 \newcommand{\avoBigExternalMedianVersions}{1.50}
+\newcommand{\avoBigExternalMeanVersionsNominal}{2.07}
 \newcommand{\avoBigExternalMeanVersions}{$2.1 \pm 1.4$}

### 2014-11-10 dastats.tex
Update figures and data for tweaked vulnerability list

@@
-\newcommand{\daNumDataPoints}{128 Billion}
-\newcommand{\daNumStartedApps}{112632}
-\newcommand{\daNumInstalledApps}{164685}
-\newcommand{\daNumVulnsUsed}{13}
+\newcommand{\daNumDataPoints}{153 Billion}
+\newcommand{\daNumStartedApps}{121998}
+\newcommand{\daNumInstalledApps}{174323}
+\newcommand{\daNumVulnsUsed}{11}
 \newcommand{\daSigNumDevices}{100}
@@
 \newcommand{\daNumSigManufacturers}{8}
-\newcommand{\daNumModels}{2368}
+\newcommand{\daNumModels}{2374}
 \newcommand{\daNumSigModels}{15}
 \newcommand{\daOSVersionPercValidLines}{99.47\%}
-\newcommand{\daOSTotalDaysData}{1127371}
+\newcommand{\daOSTotalDaysData}{1132105}
 \newcommand{\daMeanInsecurityPerc}{$87 \pm 0\%$}
-\newcommand{\daMeanOutstandingVulnerabilities}{0.69}
-\newcommand{\daUpdatedness}{$0.058 \pm 0.000$}
-\newcommand{\daSecurityScore}{$2.7 \pm 0.0$}
-\newcommand{\daNumFullVersions}{1128}
+\newcommand{\daMeanOutstandingVulnerabilities}{0.64}
+\newcommand{\daUpdatedness}{$0.056 \pm 0.000$}
+\newcommand{\daSecurityScore}{$2.8 \pm 0.0$}
+\newcommand{\daNumFullVersions}{1133}
 \newcommand{\daNumOSVersions}{45}
 \newcommand{\daNumSigOSVersions}{23}
-\newcommand{\daNumAPIVersions}{14}
-\newcommand{\daNumFullVersionUpdates}{4196}
-\newcommand{\daNumOSVersionUpdates}{3544}
-\newcommand{\daNumFullOnlyVersionUpdates}{632}
-\newcommand{\daNumSecurityUpdates}{1353}
-\newcommand{\daNumPossibleSecurityUpdates}{1886}
-\newcommand{\daTabSecScoresmanufacturer}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & Security score \\ \hline lg & $0.11 \pm 0.00$ & $0.33 \pm 0.00$ & 0.79 & $3.3 \pm 0.0$ \\  motorola & $0.16 \pm 0.00$ & $0.13 \pm 0.00$ & 0.82 & $2.9 \pm 0.0$ \\  other & $0.13 \pm 0.00$ & $0.058 \pm 0.000$ & 0.69 & $2.7 \pm 0.0$ \\  samsung & $0.14 \pm 0.00$ & $0.048 \pm 0.000$ & 0.77 & $2.6 \pm 0.0$ \\  htc & $0.14 \pm 0.00$ & $0.088 \pm 0.001$ & 0.95 & $2.5 \pm 0.0$ \\  sony & $0.13 \pm 0.00$ & $0.18 \pm 0.00$ & 1.18 & $2.5 \pm 0.0$ \\  asus & $0.14 \pm 0.00$ & $0.56 \pm 0.01$ & 7.18 & $2.2 \pm 0.0$ \\  symphony & $0.0016 \pm 0.0002$ & $0.042 \pm 0.002$ & 5.73 & $0.15 \pm 0.00$ \\ \end{tabular} \caption{Security scores for manufacturers} \label{tab:sec_manufacturer} \end{table}}
-\newcommand{\daTabSecScoresmodel}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & Security score \\ \hline Galaxy Nexus & $0.49 \pm 0.00$ & $0.54 \pm 0.01$ & 2.69 & $3.9 \pm 0.0$ \\  Nexus 4 & $0.19 \pm 0.00$ & $0.82 \pm 0.01$ & 7.22 & $3.2 \pm 0.0$ \\  Nexus 7 & $0.17 \pm 0.00$ & $0.74 \pm 0.01$ & 7.08 & $2.9 \pm 0.0$ \\  Desire HD & $0.083 \pm 0.003$ & $0.051 \pm 0.004$ & 0.39 & $2.9 \pm 0.0$ \\  other & $0.13 \pm 0.00$ & $0.12 \pm 0.00$ & 0.69 & $2.9 \pm 0.0$ \\  HTC Sensation & $0.36 \pm 0.00$ & $0.0063 \pm 0.0062$ & 2.59 & $1.9 \pm 0.0$ \\  GT-I9100 & $0.22 \pm 0.00$ & $0.017 \pm 0.001$ & 2.53 & $1.4 \pm 0.0$ \\  GT-N7000 & $0.25 \pm 0.00$ & $0.0006 \pm 0.0036$ & 3.52 & $1.2 \pm 0.0$ \\  GT-P1000 & $0.0096 \pm 0.0009$ & $0.0000 \pm 0.0076$ & 1.73 & $0.94 \pm 0.02$ \\  HTC Desire S & $0.022 \pm 0.001$ & $0.016 \pm 0.004$ & 2.00 & $0.85 \pm 0.01$ \\  GT-I9300 & $0.14 \pm 0.00$ & $0.014 \pm 0.001$ & 7.27 & $0.62 \pm 0.01$ \\  GT-I9505 & $0.0 \pm 0$ & $0.19 \pm 0.00$ & 8.23 & $0.56 \pm 0.01$ \\  GT-N7100 & $0.064 \pm 0.002$ & $0.0011 \pm 0.0058$ & 8.37 & $0.26 \pm 0.02$ \\  HTC Desire HD & $0.0 \pm 0$ & $0.0005 \pm 0.0070$ & 4.05 & $0.10 \pm 0.02$ \\ \end{tabular} \caption{Security scores for models} \label{tab:sec_model} \end{table}}
-\newcommand{\daTabSecScoressummary}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & Security score \\ \hline nexus & $0.33 \pm 0.00$ & $0.52 \pm 0.00$ & 1.74 & $3.8 \pm 0.0$ \\  notnexus & $0.12 \pm 0.00$ & $0.041 \pm 0.000$ & 0.69 & $2.6 \pm 0.0$ \\ \end{tabular} \caption{Security scores for summaries} \label{tab:sec_summary} \end{table}}
-\newcommand{\daUpdatednessPerc}{$5.8 \pm 0.0\%$}
-\newcommand{\daNumOSDevices}{18809}
+\newcommand{\daNumAPIVersions}{15}
+\newcommand{\daNumFullVersionUpdates}{4209}
+\newcommand{\daNumOSVersionUpdates}{3560}
+\newcommand{\daNumFullOnlyVersionUpdates}{637}
+\newcommand{\daNumSecurityUpdates}{1371}
+\newcommand{\daNumPossibleSecurityUpdates}{1887}
+\newcommand{\daTabSecScoresmanufacturer}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & Security score \\ \hline lg & $0.11 \pm 0.00$ & $0.32 \pm 0.00$ & 0.79 & $3.3 \pm 0.0$ \\  motorola & $0.16 \pm 0.00$ & $0.13 \pm 0.00$ & 0.82 & $2.9 \pm 0.0$ \\  other & $0.13 \pm 0.00$ & $0.056 \pm 0.000$ & 0.68 & $2.7 \pm 0.0$ \\  samsung & $0.14 \pm 0.00$ & $0.047 \pm 0.000$ & 0.76 & $2.6 \pm 0.0$ \\  htc & $0.14 \pm 0.00$ & $0.087 \pm 0.001$ & 0.95 & $2.5 \pm 0.0$ \\  sony & $0.13 \pm 0.00$ & $0.18 \pm 0.00$ & 1.17 & $2.5 \pm 0.0$ \\  asus & $0.14 \pm 0.00$ & $0.56 \pm 0.01$ & 7.18 & $2.2 \pm 0.0$ \\  symphony & $0.0015 \pm 0.0002$ & $0.041 \pm 0.001$ & 5.68 & $0.15 \pm 0.00$ \\ \end{tabular} \caption{Security scores for manufacturers} \label{tab:sec_manufacturer} \end{table}}
+\newcommand{\daTabSecScoresmodel}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & Security score \\ \hline Galaxy Nexus & $0.49 \pm 0.00$ & $0.53 \pm 0.01$ & 2.66 & $4.0 \pm 0.0$ \\  Nexus 4 & $0.19 \pm 0.00$ & $0.82 \pm 0.01$ & 7.22 & $3.2 \pm 0.0$ \\  Nexus 7 & $0.17 \pm 0.00$ & $0.74 \pm 0.01$ & 7.08 & $2.9 \pm 0.0$ \\  Desire HD & $0.083 \pm 0.003$ & $0.051 \pm 0.004$ & 0.39 & $2.9 \pm 0.0$ \\  other & $0.13 \pm 0.00$ & $0.12 \pm 0.00$ & 0.68 & $2.9 \pm 0.0$ \\  HTC Sensation & $0.36 \pm 0.00$ & $0.0063 \pm 0.0062$ & 2.59 & $1.9 \pm 0.0$ \\  GT-I9100 & $0.22 \pm 0.00$ & $0.015 \pm 0.001$ & 2.53 & $1.4 \pm 0.0$ \\  GT-N7000 & $0.25 \pm 0.00$ & $0.0006 \pm 0.0036$ & 3.52 & $1.2 \pm 0.0$ \\  GT-P1000 & $0.0096 \pm 0.0009$ & $0.0000 \pm 0.0061$ & 1.73 & $0.94 \pm 0.02$ \\  HTC Desire S & $0.022 \pm 0.001$ & $0.016 \pm 0.004$ & 2.00 & $0.85 \pm 0.01$ \\  GT-I9300 & $0.14 \pm 0.00$ & $0.014 \pm 0.001$ & 7.24 & $0.62 \pm 0.01$ \\  GT-I9505 & $0.0 \pm 0$ & $0.18 \pm 0.00$ & 8.15 & $0.55 \pm 0.01$ \\  GT-N7100 & $0.064 \pm 0.002$ & $0.0011 \pm 0.0058$ & 8.37 & $0.26 \pm 0.02$ \\  HTC Desire HD & $0.0 \pm 0$ & $0.0005 \pm 0.0070$ & 4.05 & $0.10 \pm 0.02$ \\ \end{tabular} \caption{Security scores for models} \label{tab:sec_model} \end{table}}
+\newcommand{\daTabSecScoressummary}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & Security score \\ \hline nexus & $0.33 \pm 0.00$ & $0.52 \pm 0.00$ & 1.73 & $3.8 \pm 0.0$ \\  notnexus & $0.12 \pm 0.00$ & $0.039 \pm 0.000$ & 0.68 & $2.6 \pm 0.0$ \\ \end{tabular} \caption{Security scores for summaries} \label{tab:sec_summary} \end{table}}
+\newcommand{\daUpdatednessPerc}{$5.6 \pm 0.0\%$}
+\newcommand{\daNumOSDevices}{18873}
 \newcommand{\daSigNumDevicesDay}{20}
@@
 \newcommand{\daSecScoreWorstmanufacturerScore}{$0.15 \pm 0.00$}
-\newcommand{\daSecScoreBestmodelScore}{$3.9 \pm 0.0$}
+\newcommand{\daSecScoreBestmodelScore}{$4.0 \pm 0.0$}
 \newcommand{\daSecScoreWorstmodelScore}{$0.10 \pm 0.02$}
@@
 \newcommand{\daAdbEnabledPerc}{$20 \pm 0\%$}
-\newcommand{\daNumUpdatesUpgrades}{3311}
+\newcommand{\daNumUpdatesUpgrades}{3323}
 \newcommand{\daNumUpdatesDowngrades}{146}
-\newcommand{\daPercUpdatesDowngrades}{4.22\%}
+\newcommand{\daPercUpdatesDowngrades}{4.21\%}
 \newcommand{\daStartDate}{2011-07-01}
-\newcommand{\daEndDate}{2014-10-31}
-\newcommand{\daNumOperators}{1392}
-\newcommand{\daTabSecScoresoperator}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & Security score \\ \hline o2 uk & $0.24 \pm 0.00$ & $0.12 \pm 0.00$ & 0.41 & $3.7 \pm 0.0$ \\  orange & $0.21 \pm 0.00$ & $0.085 \pm 0.003$ & 0.39 & $3.5 \pm 0.0$ \\  t-mobile & $0.18 \pm 0.00$ & $0.18 \pm 0.00$ & 0.54 & $3.5 \pm 0.0$ \\  sprint & $0.18 \pm 0.00$ & $0.11 \pm 0.00$ & 0.44 & $3.4 \pm 0.0$ \\  3 & $0.14 \pm 0.00$ & $0.096 \pm 0.003$ & 0.49 & $3.1 \pm 0.0$ \\  vodafone uk & $0.13 \pm 0.00$ & $0.10 \pm 0.00$ & 0.52 & $3.1 \pm 0.0$ \\  at\&t & $0.12 \pm 0.00$ & $0.077 \pm 0.002$ & 0.51 & $3.0 \pm 0.0$ \\  verizon & $0.17 \pm 0.00$ & $0.086 \pm 0.002$ & 0.89 & $2.7 \pm 0.0$ \\  unknown & $0.095 \pm 0.001$ & $0.22 \pm 0.00$ & 1.16 & $2.5 \pm 0.0$ \\  n telenor & $0.023 \pm 0.001$ & $0.10 \pm 0.00$ & 1.59 & $1.4 \pm 0.0$ \\  airtel & $0.067 \pm 0.002$ & $0.029 \pm 0.002$ & 2.00 & $1.1 \pm 0.0$ \\  grameenphone & $0.0080 \pm 0.0005$ & $0.016 \pm 0.001$ & 1.99 & $0.80 \pm 0.00$ \\  banglalink & $0.00016 \pm 0.00011$ & $0.029 \pm 0.002$ & 2.89 & $0.40 \pm 0.01$ \\  robi & $\left(8.5 \pm 8.5\right) \times 10^{-5}$ & $0.042 \pm 0.003$ & 3.33 & $0.33 \pm 0.01$ \\ \end{tabular} \caption{Security scores for operators} \label{tab:sec_operator} \end{table}}
+\newcommand{\daEndDate}{2014-11-03}
+\newcommand{\daNumOperators}{1393}
+\newcommand{\daTabSecScoresoperator}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & Security score \\ \hline o2 uk & $0.24 \pm 0.00$ & $0.12 \pm 0.00$ & 0.37 & $3.8 \pm 0.0$ \\  orange & $0.21 \pm 0.00$ & $0.085 \pm 0.003$ & 0.39 & $3.5 \pm 0.0$ \\  t-mobile & $0.18 \pm 0.00$ & $0.18 \pm 0.00$ & 0.54 & $3.5 \pm 0.0$ \\  sprint & $0.18 \pm 0.00$ & $0.11 \pm 0.00$ & 0.44 & $3.4 \pm 0.0$ \\  3 & $0.14 \pm 0.00$ & $0.093 \pm 0.003$ & 0.49 & $3.1 \pm 0.0$ \\  vodafone uk & $0.13 \pm 0.00$ & $0.10 \pm 0.00$ & 0.52 & $3.1 \pm 0.0$ \\  at\&t & $0.12 \pm 0.00$ & $0.077 \pm 0.002$ & 0.51 & $3.0 \pm 0.0$ \\  verizon & $0.18 \pm 0.00$ & $0.087 \pm 0.002$ & 0.89 & $2.7 \pm 0.0$ \\  unknown & $0.095 \pm 0.001$ & $0.22 \pm 0.00$ & 1.15 & $2.5 \pm 0.0$ \\  n telenor & $0.023 \pm 0.001$ & $0.10 \pm 0.00$ & 1.57 & $1.4 \pm 0.0$ \\  airtel & $0.066 \pm 0.002$ & $0.028 \pm 0.002$ & 1.89 & $1.1 \pm 0.0$ \\  grameenphone & $0.0078 \pm 0.0005$ & $0.018 \pm 0.001$ & 1.97 & $0.82 \pm 0.00$ \\  banglalink & $0.00016 \pm 0.00011$ & $0.031 \pm 0.002$ & 2.86 & $0.42 \pm 0.01$ \\  robi & $\left(8.4 \pm 8.4\right) \times 10^{-5}$ & $0.040 \pm 0.003$ & 3.17 & $0.36 \pm 0.01$ \\ \end{tabular} \caption{Security scores for operators} \label{tab:sec_operator} \end{table}}
 \newcommand{\daSecScoreBestoperator}{o2 uk}
-\newcommand{\daSecScoreBestoperatorScore}{$3.7 \pm 0.0$}
+\newcommand{\daSecScoreBestoperatorScore}{$3.8 \pm 0.0$}
 \newcommand{\daSecScoreWorstoperator}{robi}
-\newcommand{\daSecScoreWorstoperatorScore}{$0.33 \pm 0.01$}
+\newcommand{\daSecScoreWorstoperatorScore}{$0.36 \pm 0.01$}
 \newcommand{\daNumSigOperators}{14}
@@
 \newcommand{\daSecScoreworstmanufacturerNumFullVersions}{239}
-\newcommand{\daSecScoreBestmodelNumFullVersions}{46}
+\newcommand{\daSecScoreBestmodelNumFullVersions}{49}
 \newcommand{\daSecScoreworstmodelNumFullVersions}{2}
@@
 \newcommand{\daSecScoreworstsummaryNumFullVersions}{1057}
-\newcommand{\daSecScoreBestoperatorNumFullVersions}{87}
+\newcommand{\daSecScoreBestoperatorNumFullVersions}{88}
 \newcommand{\daSecScoreworstoperatorNumFullVersions}{89}
@@
 \newcommand{\daSecScoreWorstsummaryNumFullVersions}{1103}
-\newcommand{\daSecScoreWorstoperatorNumFullVersions}{33}
+\newcommand{\daSecScoreWorstoperatorNumFullVersions}{32}
 \newcommand{\daFullDeployedAt}{95\%}
-\newcommand{\daOSCurveFitParamFirst}{86.43}
+\newcommand{\daOSCurveFitParamFirst}{87.38}
 \newcommand{\daOSCurveFitParamSecond}{\num{0.00315}}
 \newcommand{\daOSCurveFitRMSE}{0.11}
-\newcommand{\daAPICurveFitParamFirst}{110.29}
-\newcommand{\daAPICurveFitParamSecond}{\num{0.00316}}
+\newcommand{\daAPICurveFitParamFirst}{109.79}
+\newcommand{\daAPICurveFitParamSecond}{\num{0.00315}}
 \newcommand{\daAPICurveFitRMSE}{0.12}
@@
 \newcommand{\daAPICurveHalfDeployed}{$\left(3.3 \pm 0.8\right) \times 10^{2}$}
-\newcommand{\daAPICurveFullDeployed}{$\left(0.0 \pm 2.3\right) \times 10^{5}$}
+\newcommand{\daAPICurveFullDeployed}{$\left(0.0 \pm 2.4\right) \times 10^{5}$}
 \newcommand{\daVulnAPKDuplicateFileOctoberPerc}{91.46\%}
-\newcommand{\daNumUpdatesBigUpgrades}{776}
+\newcommand{\daNumUpdatesBigUpgrades}{778}
 \newcommand{\daNumUpdatesSkippedBig}{3}
-\newcommand{\daPercBigUpgrades}{22.45\%}
+\newcommand{\daPercBigUpgrades}{22.43\%}
 \newcommand{\daUpdatesPerYear}{$1.4 \pm 0.0$}
 \newcommand{\daOSMonthsOfData}{40}
+\newcommand{\daMeanInsecurityPercNominal}{87.21\%}
+\newcommand{\daVulnFreeNominal}{0.13}
+\newcommand{\daUpdatednessNominal}{0.06}
+\newcommand{\daUpdatednessPercNominal}{5.60\%}
+\newcommand{\daSecurityScoreNominal}{2.75}
+\newcommand{\daUpdatesPerYearNominal}{1.36}
+\newcommand{\daOSCurveHalfDeployedNominal}{307.28}
+\newcommand{\daOSCurveFullDeployedNominal}{1037.77}
+\newcommand{\daAPICurveHalfDeployedNominal}{329.78}
+\newcommand{\daAPICurveFullDeployedNominal}{1060.54}

### 2014-11-10 securityupdates.tex
Add some missing references

@@
 Fixing them is not enough, we show that these fixes do not promptly reach the users devices and so much of the time users are running versions of Android known to be vulnerable.
-We used \da\ data from over \daOSYearsOfData\ years and \daNumOSDevices\ devices and found that on average \daMeanInsecurityPerc\ of devices were exposed to known root privilege vulnerabilities and only \daUpdatednessPerc\ of devices run the most recent version of Android.
+We used \da\ data from over \daOSYearsOfData\ years and \daNumOSDevices\ devices and found that on average \daMeanInsecurityPercNominal\ of devices were exposed to known root privilege vulnerabilities and only \daUpdatednessPercNominal\ of devices run the most recent version of Android.
 There was also a period of several months when no devices ran secure versions of Android.
@@
 Support for third-party apps is a key feature of modern smartphone operating systems.
-Such apps are written by many developers from a wide range of backgrounds, which means neither operating system vendors nor users can, in general, fully trust app developers.
+Such apps are written by many developers from a wide range of backgrounds, which means neither operating system vendors nor users can, fully trust app developers.
 Therefore, in order to secure personal data and prevent the use of expensive operations such as the sending of premium-rate text messages, operating system vendors provide a protected execution environment, or \emph{sandbox}, for apps.
@@
 Understanding this ecosystem is important as manufacturers have introduced additional vulnerabilities in the past~\cite{Grace2012}. 
-We present a better understanding of both the ecosystem of Android development and associated vulnerabilities more fully in \S~X.
+We present a better understanding of both the ecosystem of Android development and associated vulnerabilities more fully in \S\ref{sec:android_ecosystem}.
 
@@
 In order to determine whether an individual smartphone is vulnerable to a particular exploit, we need to combine vulnerability data with the operating system version string and additional update information from the handset.
-For this we use handset data gathered from over 18,000 participants in the \da~\cite{} project.
+For this we use handset data gathered from over 18,000 participants in the \da~\cite{Wagner2013} project.
 This analysis allows us to plot the proportion of Android devices in the \da\ data which were running versions of Android known to have vulnerabilities at the time and is shown in Figure~\ref{fig:proportioninsecure}.
@@
 (3) \emph{maybe secure} because the device is running an insecure version of Android but it received an update which we did not see until after the vulnerability was fixed and therefore this update may contain a backport of the fix.
-Further analysis into the propagation of updates in Android are presented in \S~Y.
+Further analysis into the propagation of updates in Android are presented in \S\ref{sec:exp:android_ecosystem}.
 
@@
 
-\section{Android Ecosystem}
+\section{Android Ecosystem}\label{sec:android_ecosystem}
 There is a complex Android ecosystem which creates and distributes updates to Android which fix vulnerabilities.

### 2014-11-10 securityupdates.tex
add some missing references

@@
 We only use the \daNumVulnsUsed\ vulnerabilities which we have sufficient data to use for this analysis because they affect all Android devices rather than particular manufacturers and we know enough information about the vulnerability to tie it to the particular versions it affects.
-We have published full details of the \daNumVulnsUsed\ vulnerabilities used for this analysis in an accompanying technical report~\cite{TODO} and summarise them in Table~\ref{tab:andvulns}.
+We summarise the \daNumVulnsUsed\ vulnerabilities used for this analysis in Table~\ref{tab:andvulns}.\todolater{We have published full details of the \daNumVulnsUsed\ vulnerabilities used for this analysis in an accompanying technical report~\cite{TODO}}
 
@@
 Capability based enforcement systems such as Capsicum~\cite{Watson2010} substantially reduce the capabilities that an exploit has to try and gain increased privilege with.
-When Capsicum is included in Linux~\cite{TODO} and hence in Android, it could be used to place fine grained restrictions on system daemons, preventing vulnerabilities becoming critical vulnerabilities.
+When Capsicum is included in Linux\footnote{\url{https://github.com/google/capsicum-linux}} and hence in Android, it could be used to place fine grained restrictions on system daemons, preventing vulnerabilities becoming critical vulnerabilities.
 
@@
 Fortunately Android does authenticate the update binaries (though the APK vulnerabilities circumvented this) and Google Play downloads them over a secure connection~\cite{Viennot2014}.
-\todolater{double check that}
 
@@
 The same technique would be more difficult on phones with limited storage space (as for many cheap Android phones which have barely enough space to install just the update) but is a plausible improvement for more high-end devices.
-Google is deploying the same silent update technique through Google Play Services~\cite{TODO} which automatically installs updates for core Google components of Android, this also bypasses the manufacturer and operator.
+Google is deploying the same silent update technique through Google Play Services\footnote{\url{http://lifehacker.com/why-google-play-services-are-now-more-important-than-an-975970197}} which automatically installs updates for core Google components of Android, this also bypasses the manufacturer and operator.
 

### 2014-11-10 cuttingfloor.tex
Fix TODOs and move cuttingfloor to a separate file to allow sending out for review

--- 2014-11-10 securityupdates.tex
@@
-\documentclass[conference,a4paper,twoside]{IEEEtran}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage[space]{grffile}
-\graphicspath{figures/}
-\usepackage{import}
-\usepackage{url}
-\usepackage{hyperref}
-\usepackage{subcaption}
-\usepackage{tikz}
-\usepackage{siunitx}
-\usepackage{fp}
-\usepackage{todonotes}
-\let\OldTodo\todo
-\renewcommand{\todo}{\OldTodo[inline]}
-\newcommand{\todolater}[1]{}%\todo
-\RequirePackage[date=terse, isbn=true, doi=true, url=false, urldate=iso8601, maxbibnames=9, backref=false, backend=bibtex]{biblatex}
-\addbibresource{securityupdates.bib}
-\renewcommand{\bibfont}{\small}
-
-\AtEveryBibitem{% Clean up the bibtex rather than editing it
- \clearname{editor} % remove editors
-}
-
-\author{Daniel R.\ Thomas, Alastair R.\ Beresford, Daniel T.\ Wagner and Andrew Rice}
-
-\input{dastats}
-\newcommand{\da}{Device Analyzer}
-\newcommand{\dafoot}{\textsuperscript{\ref{foot:dadata}}}
-\input{avostats}
-\newcommand{\avo}{\texttt{androidvulnerabilities.org}}
-\newcommand{\percMarketShare}{83.6\%~\footnote{\url{http://www.theinquirer.net/inquirer/news/2379036/android-hits-836-percent-marketshare-while-ios-windows-and-blackberry-slide}}}
-\newcommand{\daNumDevices}{\daNumOSDevices}
-\newcommand{\daDeviceDays}{\daOSTotalDaysData}
-% Num versions since \daStartDate
-\input{countversions}
-\newcommand{\otherProjNum}{\avoNumExternalProjects}%TODO check we are doing the right calculation here and not overcounting
-
-\begin{document}
-\title{Timeliness of security updates for Android}
-
-% author names and affiliations
-% use a multiple column layout for up to three different
-% affiliations
-\author{
-\IEEEauthorblockN{Daniel R. Thomas,
-Daniel T. Wagner,
-Alastair R. Beresford,
-Andrew Rice}
-\IEEEauthorblockA{
-Computer Laboratory\\
-University of Cambridge\\
-Cambridge, United Kingdom\\
-Firstname.Lastname@cl.cam.ac.uk
-}
-}
-
-% Terminology
-
-% Android, Unix
-% update, release, patch: what do we actually mean and what relevance does that have to security?
-
-\maketitle
-
-% Hypothesis
-% Attempts at fine grained restrictions on running arbitrary code are hampered by updates for security vulnerabilities not reaching users in a timely fashion.
-
-\begin{abstract}
-Android is the most popular smartphone platform and seeks to provide better security than popular desktop operating systems through increased compartmentalisation and a permission system for apps.
-This security relies on the absence of root privilege exploits which allow apps to bypass protections.
-Many such exploits have been found and fixed.
-Fixing them is not enough, we show that these fixes do not promptly reach the users devices and so much of the time users are running versions of Android known to be vulnerable.
-We used \da\ data from over \daOSYearsOfData\ years and \daNumOSDevices\ devices and found that on average \daMeanInsecurityPerc\ of devices were exposed to known root privilege vulnerabilities and only \daUpdatednessPerc\ of devices run the most recent version of Android.
-There was also a period of several months when no devices ran secure versions of Android.
-We compare different manufacturers and operators to find which are best at promptly distributing security updates and examine how to improve the situation. %TODO operators analysis
-\end{abstract}
-
-\section{Introduction}
-%Android is the most popular smartphone operating system today with \percMarketShare\ marketshare.
-
-Support for third-party apps is a key feature of modern smartphone operating systems.
-Such apps are written by many developers from a wide range of backgrounds, which means neither operating system vendors nor users can, in general, fully trust app developers.
-Therefore, in order to secure personal data and prevent the use of expensive operations such as the sending of premium-rate text messages, operating system vendors provide a protected execution environment, or \emph{sandbox}, for apps.
-
-In this paper we quantify the extent to which the app sandbox provided by Android protects devices from malicious apps attempting to gain control over the entire handset.
-%We have analysed Android because it is the most popular smartphone operating system by market share (~\percMarketShare) and 
-Previous research has shown that in 2012, between 36.7\%~\cite{Zhou2012b} and 40\%~\cite{Zhou2012a} of malware for Android contained exploits designed to enable a malicious app to break the sandbox.
-What is not known however is the proportion of Android devices which are vulnerable to these exploits over time.
-To the best of our knowledge, this is the first paper to provide detailed numbers.
-
-Not all malicious apps need to break the sandbox in order to misbehave.
-Indeed, the above figures suggest the majority of malicious apps do not attempt to do so.
-We focus on this issue because malware which is able to take full control of a handset can do significantly more harm and is much harder to remove.
-Ordinary malware can uninstalled by the user or remotely by Google through the Play Store, but such removal is unlikely to work once the security features of Android are broken.
-
-The development of Android is distributed among many parties.
-The security of Android relies on many open source projects, include the Linux kernel, OpenSSL and BouncyCastle as well as Google who build the core platform. 
-In addition, manufacturers (of which we are aware of \daNumManufacturers\footnote{\label{foot:dadata}We computed this from the \da~\cite{Wagner2013} data, see \S\ref{sec:android_update_process}.}) and network operators (\daNumOperators\dafoot) may make further modifications before devices are shipped to customers.
-Understanding this ecosystem is important as manufacturers have introduced additional vulnerabilities in the past~\cite{Grace2012}. 
-We present a better understanding of both the ecosystem of Android development and associated vulnerabilities more fully in \S~X.
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=\columnwidth]{figures/proportioninsecure}
-\caption{Proportion of devices running insecure, maybe insecure and secure versions of Android against time}
-\label{fig:proportioninsecure}
-\end{figure}
-
-In order to determine whether an individual smartphone is vulnerable to a particular exploit, we need to combine vulnerability data with the operating system version string and additional update information from the handset.
-For this we use handset data gathered from over 18,000 participants in the \da~\cite{} project.
-This analysis allows us to plot the proportion of Android devices in the \da\ data which were running versions of Android known to have vulnerabilities at the time and is shown in Figure~\ref{fig:proportioninsecure}.
-The Figure shows the proportion of devices split into three categories: 
-(1) \emph{secure} because they are running a version of Android with no known vulnerabilities; 
-(2) \emph{insecure} because the device was running an insecure version of Android with a build released before the vulnerability was fixed or discovered; and 
-(3) \emph{maybe secure} because the device is running an insecure version of Android but it received an update which we did not see until after the vulnerability was fixed and therefore this update may contain a backport of the fix.
-Further analysis into the propagation of updates in Android are presented in \S~Y.
-
-It is clear from Figure~\ref{fig:proportioninsecure} that the security of the Android platform is far from perfect.
-One solution is regulation, and indeed there is ongoing legal action to force operators to ship updates for security vulnerabilities~\cite{Soghoian2013}.\todolater{Check on the status of this legal action}
-Indeed, many smartphones are sold on 12--24 month contracts, and yet our data shows many devices do not receive many security updates. 
-In contrast, Windows XP could be purchased for a one-off payment in October 2001 and received security updates until April 2014.
-
-An alternative to regulation is better public data. 
-Hard numbers provide an incentive to both manufacturers and operators to provide updates.
-Currently corporate and public sector buyers are encouraged to purchase secure devices, but we have found little concrete guidance on the specific makes and models which provide the best security. 
-For example, CESG, which advises the UK government on how to secure its computer systems, recommends picking Android devices from manufactures which are good at shipping security updates promptly~\cite{CESG2013} but it does not state which manufacturers are better.
-To address this, we developed a scoring system and provide numbers on the historic performance of devices in the \da\ project in~\S\ref{sec:security_scoring}.
-
-%REMOVED BY ARB
-%The manufacturer has little financial incentive to perform this work as they have already sold the device.
-%Operators are cautious because if they ship a broken update then they will get many support requests which is expensive.
-%
-%In the past users of Microsoft IIS were advised to switch to a different web server due to the frequent discovery of vulnerabilities~\cite{Pescatore2001}.
-%This pressure encouraged Microsoft to improve the security of their products~\cite{TODO}.
-
-In summary, the contributions of this paper are:
-\begin{itemize}
- \item We characterise the Android update ecosystem, showing how updates flow between entities, and quantify vulnerabilities by determining what proportion of devices they affected and how that changes over time.
- \item We collate quantitive information on vulnerabilities which affect Android and investigate their time-lines.
- \item We measure the security of Android according to three metrics and compare different manufacturers and device models to allow device purchasers to differentiate between them based on security.
-\end{itemize}
-
-\section{Background}
-\label{sec:background}
-There are two sources of data we need for this study: (1) information on the distribution of installed versions of Android over time and (2) information on critical vulnerabilities.
-We need data on critical vulnerabilities so that we know which versions of Android were vulnerable to which known vulnerabilities at what times.
-We need information on the distribution of installed versions of Android so that we know what proportion of Android devices were affected by these vulnerabilities.
-
-\subsection{Versions of Android running on devices}
-We need historical data on the distributions of versions of Android running on devices and additional data about those devices to allow us to investigate variation between devices and confirm the reliability of these data.
-We used data from the \da\ project~\cite{Wagner2013}.
-The \da\ data contains info from \daNumDevices\ devices with a total of \daDeviceDays\footnote{Here we are only counting devices and days for which we have valid OS version data.} device days, contains data for more than \daMonths\ months for \daMonthsDevices\ devices.
-Device Analyzer used an Android app which volunteers install on their Android devices to collect data on what happens on Android phones.
-The data collected is available for researchers to use as we do here.
-
-Various different kinds of data\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/collected.htm}} are collected by the \da\ Android app\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/}}.
-Among these are the build string and API version for the version of Android currently running on the device each day.
-The API version is a well defined integer, however it does not always change with new Android releases, particularly security bug fixes should not increase this value.
-The build string is `The user-visible version string', fortunately most (\daOSVersionPercValidLines) entries in the data have a build string of the form `x.y.z opaque marker' and so it is possible to extract the Android version number.
-On a large proportion of devices `opaque marker' is a well defined build number\footnote{\url{https://source.android.com/source/build-numbers.html}} but different manufacturers use different schema.
-
-We verify the validity of the \da\ data in \S\ref{sec:representative}.
-
-\subsection{Critical vulnerabilities}
-We compiled a list of critical vulnerabilities in Android, containing information on discovery and publication dates, what versions they affected and which versions fixed the problem.
-We only looked for critical vulnerabilities such as root vulnerabilities which did not require USB debugging to exploit.
-Critical vulnerabilities allow a program to escalate to privileges equivalent in scope to root.
-If an application exploits a critical vulnerability then it gains control of the device.
-Some phones can be `rooted' by enabling USB debugging and using the special privileges of the adb shell to root the device but only \daAdbEnabledPerc\dafoot\textsuperscript{,}\footnote{Where possible we indicate the uncertainty in our results by giving them $\pm$ the standard deviation.
-This is explained further in \S\ref{sec:uncertainty} but occasionally results in $\pm0$.} of devices have USB debugging enabled.
-This is not something that applications running on the phone can exploit to break compartmentalisation and so we do not include those vulnerabilities.
-Unfortunately many published exploits use adb for convenience and so determining whether that is necessary to exploit the vulnerability can be difficult.
-
-Some critical vulnerabilities are not traditional kernel vulnerabilities, for example the discovery of flaws in the verification of signatures on Android applications in February 2013~\cite{Forristal2013} meant that applications could pretend to be signed with system keys and hence gain root equivalent privileges.
-On some versions of Android (below 4.1) they could use known system to root escalation mechanisms but on all versions they have a greatly increased attack area for further privilege escalation and also have the ability to control all user internet traffic (via VPNs), brick the phone, remove and install apps, steal user credentials, read the screen and make as well as receive calls.
-\todolater{break vulnerabilities down by attack vector?}
-\avoTabAndVulns
-
-We developed and maintain an open platform for filing critical vulnerabilities in a machine readable format, \texttt{\url{http://androidvulnerabilities.org/}}.
-We seeded it with data from the CVE database, vendor lists, reports from the literature and various forums.
-We have received submissions or amendments from \avoNumSubmitters\ individuals.
-\avo\ contains \avoNumVulnerabilities\ vulnerabilities of which \avoNumVulnAllAndroid\ affect all Android devices and \avoNumVulnSpecific\ are specific to particular devices or manufacturers.
-We only use the \daNumVulnsUsed\ vulnerabilities which we have sufficient data to use for this analysis because they affect all Android devices rather than particular manufacturers and we know enough information about the vulnerability to tie it to the particular versions it affects.
-We have published full details of the \daNumVulnsUsed\ vulnerabilities used for this analysis in an accompanying technical report~\cite{TODO} and summarise them in Table~\ref{tab:andvulns}.
-
-Tracking vulnerabilities is a manual task as they are not consistently recorded in other databases such as the CVE database and the lack of a widely acknowledged unique identifier made identifying whether two reports referenced the same vulnerability difficult.
-Previous work has assumed ``any security issue of relevance will eventually get a CVE number assigned"~\cite{Frei2010} which is currently not the case for critical Android vulnerabilities.
-For some of the vulnerabilities without CVE numbers, Google confirmed that there was no CVE number and that they did not intend to get one, instead providing an Android bug number.
-
-\subsection{Lifetime of a vulnerability}
-
-The key events in the lifetime of a vulnerability do not always occur in the same order and are:
-
-\begin{LaTeXdescription}
- \item[created] When a vulnerability was created in the source code.
- \item[introducing release] When the first release was made containing the vulnerability.
- \item[discovery] When the vulnerability is first discovered
- \item[exploit] When the vulnerability is first exploited
- \item[disclosure] When the vulnerability is first disclosed % publically or to a smaller set of people?
- \item[fix] When a vulnerability was first fixed in the source code.
- \item[fixing release] When the first release containing the fix was made (equivalent to `patch available'~\cite{Frei2010}).
- \item[fix deployed] When the fix has been deployed to a sufficiently large proportion of the population that the vulnerability can be ignored (the ecosystem equivalent to the per instance `patch installed'~\cite{Frei2010}).
-\end{LaTeXdescription}
-
-Working out the date to use for the date on which a vulnerability was sufficiently well known to pose a threat to users is difficult.
-Frei et al.~\cite{Frei2010} propose the definition of the {\bf time of disclosure} when the information about the vulnerability is freely available to the public, from a widely accepted and independent source and has been validated by security experts so that it has a risk rating.
-Unfortunately before we collated this information much of it was not published by an independent source and lacked risk rating information, even months or years after they had been actively used.
-Therefore this measure does not work.
-
-Symantec's 2012 analysis of desktop malware has shown that after public disclosure exploitation rates increase by 5 orders of magnitude~\cite{Bilge2012} and so from the point of view of widespread danger, the period between public disclosure and the update reaching the user's device is the most critical.
-However they also show that zero day vulnerabilities are typically used for 312 days before they are publicly disclosed, often to target particular organisations.
-Hence when considering vulnerability from the point of view of an organisation which cares about Advanced Persistent Threats, such as those the CESG advice~\cite{CESG2013} is aimed at, the critical period is from when people could have started using the vulnerability to when the update hits the user device.
-So for our vulnerability calculations we use the earliest date we can find recorded evidence that the vulnerability was known about, even if that knowledge might have been confined to a particular manufacturer or hobbyist.
-%We cannot know if someone reported the vulnerability to the manufacturer and we do know that various agencies are engaged in widespread monitoring of communications and compromise of civilian infrastructure~\cite{TODO} and so they may know.
-
-There are several points we could pick for the point from which a vulnerability poses substantial risk to users.
-{\bf Creation} and {\bf introduction} are the earliest points which could be used but the risk is mostly latent until someone discovers them.
-Unfortunately {\bf discovery} is the hardest point to obtain concrete data on as discovery may happen multiple times independently and not all discoverers will report their discovery, for example Advanced Persistent Threats (APTs) may keep databases of unknown vulnerabilities to use when they have a particular target, in which case there is a high risk to potential targets from the never disclosed point of discovery.
-However if the discoverer never discloses the vulnerability to anyone or makes any use of it then there is little risk.
-The date of first {\bf exploit} is a point when the risk is definitely high, but again a good adversary will not be detected when using such exploits.
-The date of first {\bf fix} is, assuming that the fix is deliberate, a point at which the vulnerability is known at least within the organisation performing the fix and frequently implies an earlier discovery and notification by a third party.
-It can, for example, be determined from the authored on date on the fixing commit in the git repository.
-Once a {\bf fixing release} has been made then the vulnerability is widely known because it can be reverse engineered from the changes in the release~\cite{Brumley2008}.
-When the {\bf fix is deployed} to a sufficiently large proportion of devices then the remaining risk is minimal.
-
-Hence we use the earliest date we know about for discovery, exploit, disclosure or fix as the date from which a vulnerability poses a substantial risk to users.
-This is shown in Table~\ref{tab:andvulns}.
-
-\subsection{Distribution of vulnerabilities}
-\begin{figure}
- \centering
- \includegraphics[width=\columnwidth]{figures/vulnerabilities_timeline}
- \caption{Timeline of vulnerabilities}
- \label{fig:vulnerabilities_timeline}
-\end{figure}
-
-The dates of discovery of vulnerabilities in the \avo\ is not uniform.
-Figure~\ref{fig:vulnerabilities_timeline} shows the dates of discovery and, when later, the dates when a version of Android which fixed the vulnerability was shipped.
-Some vulnerabilities (RageAgainstTheCage adb, levitator, zergRush abd TwerkMyMoto) were fixed in released versions of Android before they were discovered and so are shown as vertical lines while others were known for months before a version of Android which fixed them was shipped.
-During the period in which \da\ data was collected the date when a version of Android with the fix was observed on a \da\ device is taken as the date that version was released.
-For vulnerabilities prior to the collection period the release date as best as we can determine is used.
-There is no canonical source of Android release dates, our best guesses and supporting references are available from \avo.
-
-This data shows a large gap in the middle where we do not know of any discoveries of critical vulnerabilities affecting all of Android.
-The cause of this quiet period is unclear.
-Possible explanations are: That since most devices were exposed to known vulnerabilities there was no point in looking for new ones (from Figure~\ref{fig:proportioninsecure}).
-That manufacturers made it easier to install custom versions of Android, reducing the need for users to root their devices?
-That manufacturer specific vulnerabilities (which we are ignoring) proved to be easier to find and so people looking for vulnerabilities adjusted their focus.
-
-\section{Android Ecosystem}
-There is a complex Android ecosystem which creates and distributes updates to Android which fix vulnerabilities.
-In this section we rectify the lack of public information about the size of the Android ecosystem.
-
-\subsection{Android update process}
-
-\label{sec:android_update_process}
-\begin{figure}
- \centering
- \def\svgwidth{\columnwidth}
- \import{figures/}{update_ecosystem.pdf_tex}
- \caption{Flow of updates between participants in the Android ecosystem.
- Numbers on edges indicate updates shipped between \daStartDate\ and \daEndDate, numbers in brackets represent number of such entities in our data.
- Dotted arrows indicate flows where we do not know how many updates are being produced as we can't measure those flows directly as they are not public.\todolater{Turn into a sankey diagram}}
- \label{fig:update_ecosystem}
-\end{figure}
-To understand how vulnerabilities in Android are fixed we must examine the Android update process which we model in Figure~\ref{fig:update_ecosystem}.
-There are five entities or groups which contribute towards Android updates: the network operators, the device manufacturers, the hardware developers, Google and the upstream open source projects.
-Android builds on various open source projects such as the Linux kernel, the OpenSSL and BouncyCastle cryptography libraries and so can include any compatible versions of those projects, including those which fix security vulnerabilities in them.
-Android also incorporates various drivers for different bits of hardware.
-The Android platform is then built on top of those with kernel and userspace components by Google.
-The code for each update is kept secret\footnote{\url{https://source.android.com/source/code-lines.html}}\todolater{Can we quantify this keeping the code secret?
-Is it worth it?} until after the update has been published.
-Manufacturers may receive advanced copies in order to prepare handsets.
-Then the Manufacturers take the update and customise it before passing it on to the operator.
-The Operator may then make further customisations and do further testing before shipping the update to the Device.
-Sometimes Manufactures ship direct to the user, sometimes the Manufacturer and Google collaborate closely for the purpose of a particular phone, such as with Nexus phones, and hence are harder to separate.
-Sometimes Manufacturers incorporate upstream open source project releases directly, and sometimes incorrectly -- for example including a broken daily build of sqlite in their release of Android~\cite{Wagner2013}.
-\todolater{use statistics from samsung-updates.com -> how many binaries are there per device?}
-
-The numbers of Devices (\daNumOSDevices), Operators (\daNumOperators) and Manufacturers (\daNumManufacturers) in Figure~\ref{fig:update_ecosystem} come from the \da\ data.
-Manufacturer and Operator counts were obtained by normalising the results reported by Android to \da\ of the device manufacturer and active network operator.
-This normalisation is a manual task involves removing invalid values (such as `manufacturer' or `airplane mode is on'), collating across company name changes (e.g.\ `lge' to `lg'), normalising punctuation, removing extra strings sometimes added such as (`(2g)' or `communications') and mapping some incorrectly placed model names back to their manufacturer.
-This normalisation is not perfect and so these are overestimates on the \da\ data but they are likely still underestimates as there will be some manufacturers and operators which are not included in the \da\ data.
-
-In Figure~\ref{fig:update_ecosystem} the number of updates received by devices (\daNumFullVersions) is the number of different full version strings observed in \da.
-The number of updates shipped by Google (\daNumSigOSVersions) is the number of Android versions reported in \da\ which affected more than \daSigVersionPerc\ of devices for more than \daSigVersionDays\ days.
-\todolater{this is the wrong number to be using}
-
-We extracted data on the external projects used in Android and have included this and the scripts which generated it in \avo.
-These scripts analysed the Android Open Source Project's source tree to examine the source code of each of the external projects to find the project version associated with each Android version tag on the repository.
-There are \avoNumExternalProjects\ external open source projects in Android, contributing \avoTotalExternalLines\ lines of code.\footnote{Lines of code were measured using David A.
-Wheeler's \texttt{sloccount}.}
-We analysed the top \avoNumBigExternalProjects\ by lines of code (\avoBigExternalLinesOfCodePerc\ of the total) and were able to automatically extract the versions of those projects included in different versions of Android for \avoNumAnalysedExternalProjects\ of these (\avoAnalysedExternalLinesOfCodePerc\ of the total).
-We found \avoBigExternalTotalVersions\ distinct versions, a median of \avoBigExternalMedianVersions\ and mean of \avoBigExternalMeanVersions\ versions per project.
-Android rarely changes the version of external projects it includes.
-
-%An analysis by Vidas et al.~\cite{Vidas2011} of the Android 2.1 to 2.2 update found that it took 11 months from when Google released 2.2 for the last device which they were investigating to get the update.
-To compute the latency between upstream releases and their inclusion in Android we scraped the release pages for those projects, to obtain the version numbers and release dates.
-This allows us to compute the latency between an upstream project being released and it being included in Android, this is shown in Table~\ref{tab:update_ecosystem}.
-The versions included in Android were about half a year old when the first version of Android containing it was released.
-\begin{table}
-\centering
-\begin{tabular}{l|r|r}
-Project	&	\# releases	&	latency \\ \hline
-linux	&	\linuxNumVersions	&	\linuxMeanUpdateLatency \\
-openssl	&	\opensslNumVersions	&	\opensslMeanUpdateLatency \\
-bouncycastle	&	\bouncycastleNumVersions	&	\bouncycastleMeanUpdateLatency \\
-\end{tabular}
-\caption{Flow of updates from upstream projects into Android.
-Number of updates as in Figure~\ref{fig:update_ecosystem}, latency in days for all pairs of versions we have data on.\todolater{scrape the other 26 websites...
-is it worth it?}}
-\label{tab:update_ecosystem}
-\end{table}
-
-\section{Experiments}
-\label{sec:results}
-We present the results of our analysis showing \daMeanInsecurityPerc\ of devices to have critical vulnerabilities on average.
-On average only \daUpdatesPerYear\ updates reach a device each year and \daPercUpdatesDowngrades\ of version changes are downgrades to older versions.
-Nexus devices are better than others, \emph{\daSecScoreBestmanufacturer} to be the highest scoring manufacturer and \emph{\daSecScoreBestoperator} to be the highest scoring network operator.
-%\todo{comparison of nexus update rates with iOS}
-We conducted three experiments examining the behaviour of the Android ecosystem as a whole (\S\ref{sec:exp:android_ecosystem}), updates reaching particular devices (\S\ref{sec:exp:device_updates}) and comparisons between different manufacturers and operators (\S\ref{sec:exp:security_score}).
-
-%% In an experiment you test a hypothesis, it needs to be clear what the hypothesis is and what the conclusion of the test is
-
-\subsection{Experiment 1: behaviour of the Android ecosystem}\label{sec:exp:android_ecosystem}
-
-\subsubsection{Method}
-To investigate the behaviour of the Android ecosystem as a whole with respect to Android versions and vulnerabilities we used the OS version information from \da\ and the vulnerability data from \avo.
-The \da\ data used here was collected between \daStartDate\ and \daEndDate.
-The \avo\ data was collected between \avoStartDate\ and \avoEndDate\ and covers the period from \avoFirstDataDate\ to \avoLastDataDate.
-For each device, for every day with version data we record what version it was running and which vulnerabilities it was exposed to at that time (if any).
-We then normalise these totals for each day by the total number of devices with version information seen on that day.
-
-\subsubsection{Results}
-The proportion of devices in the \da\ data running different versions of Android each day is shown in Figure~\ref{fig:norm_os}.
-It shows how old versions are gradually replaced by new ones, the long tail of devices which do not see updates to more recent versions.
-
-\begin{figure}
- \centering
- \includegraphics[width=\columnwidth]{figures/da_norm_os}
- \caption{Android versions in \da\ data over time}
- \label{fig:norm_os}
-\end{figure}
-
-The vulnerabilities devices are exposed to are shown in Figure~\ref{fig:nvulnerabilities_heat}.
-For each vulnerability it shows the proportion of devices exposed to that vulnerability and how that changes over time.
-The variation of the proportion of devices affected by a vulnerability with time tells us how badly a particular vulnerability affected the Android platform.
-In July 2011 at the beginning of the \da\ data the \emph{exploid}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/exploid_udev}} and \emph{levitator}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/levitator}} vulnerabilities both affect most Android devices, slowly these are fixed as updates roll out and devices are replaced until in January 2013 a much smaller proportion of devices are affected by known vulnerabilities.
-However when in February 2013 the first APK signing vulnerability was found which affected all previous versions of Android and even in October 2013 most devices (\daVulnAPKDuplicateFileOctoberPerc) were still vulnerable.
-%\begin{figure}%[!b]
-%\centering
-%\includegraphics[width=\columnwidth]{figures/vulnerabilities}
-%\caption{Proportion of devices exposed to each vulnerability with time (1.0 is 100\%)}
-%\label{fig:vulnerabilities}
-%\end{figure}
-
-In 2013 three vulnerabilities were found in the way which Android verified the signatures on APKs.
-These allowed the creation of malicious APKs which appear to be signed as system APKs -- which have root equivalent privileges.
-Figure~\ref{fig:nvulnerabilities_heat} shows how the the \emph{APK signing vulnerabilities} affected all devices and took months to get fixed for any device.
-However what is perhaps more worrying is the long tail on the \emph{Gingerbreak}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/Gingerbreak}}, \emph{levitator}, \emph{exploid} and \emph{zergRush}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/zergRush}} vulnerabilities which are more dangerous root vulnerabilities (not requiring new APK installation) and which still affect a significant proportion of devices years later.
-
-\begin{figure}
- \includegraphics[width=\columnwidth]{figures/nvulnerabilities_heat.pdf}
- \caption{Proportion of devices affected by different vulnerabilities}
- \label{fig:nvulnerabilities_heat}
-\end{figure}
-
-\subsection{Experiment 2: Updates to particular devices}\label{sec:exp:device_updates}
-Those graphs summarise data across all the devices, however one of the advantages of the \da\ data is that it allows us to look at what happens to individual devices over time.
-
-\subsubsection{Method}
-As in Experiment 1, for each device we used the version information from each day to calculate which vulnerabilities it was vulnerable to that day.
-We also recorded when the version changed and which versions it changed from and to.
-In the \da\ data we have over 1000 devices contributing data in any particular week.
-However most devices only contribute for a short period of time and so we do not observe updates happening on every device.
-Instead we have a hopefully representative sample of upgrades which happened while \da\ was installed on the devices.
-\da\ cannot distinguish between a device being replaced and the \da\ app being removed as while multiple devices can be linked to the same user account, that is a manual process and few users do it.
-
-\subsubsection{Results}
-\begin{figure}
- \centering
- \begin{subfigure}[b]{\columnwidth}
-  \includegraphics[width=\columnwidth]{figures/device-data-all-security}
-  \caption{Number of vulnerabilities affecting each device over time}
-  \label{fig:device_data_security}
- \end{subfigure}
- \begin{subfigure}[b]{\columnwidth}
-  \includegraphics[width=\columnwidth]{figures/device-data-all-os}
-  \caption{OS versions over time. Black vertical lines indicate that the build number changed without changing the OS version at that point.}
-  \label{fig:device_data_os}
- \end{subfigure}
- \caption{The top \daNumDeviceDataDevices\ devices by days of contribution in the \da\ data. One strip per device.}
-\end{figure}
-The logitudinal data on the number of vulnerabilities affecting the \daNumDeviceDataDevices\ devices which have contributed the most days of data to \da\ changes over time is shown in Figure~\ref{fig:device_data_security}.
-The trend that we saw in Figure~\ref{fig:proportioninsecure} of security improving and then getting worse is also shown here.
-It shows how some devices had vulnerabilities, which were fixed, and then further vulnerabilities were discovered, and for these devices, mostly not fixed.
-This implies that these devices had been abandoned by the manufacturer and were not receiving updates which is confirmed by Figure~\ref{fig:device_data_os} which shows which OS versions those devices were running.
-Some devices start off in 2011 exposed to known security vulnerabilities and are still exposed to those and additional ones in 2014.
-
-\begin{figure}
- \includegraphics[width=\columnwidth]{figures/from_to_updates.pdf}
- \caption{Updates between different Android versions in the \da\ data}
- \label{fig:from_to_updates}
-\end{figure}
-\begin{figure}
- \includegraphics[width=\columnwidth]{figures/nw_security_updates.pdf}
- \caption{Proportion of updates each week which fixed or may have fixed security vulnerabilities}
- \label{fig:weekly_security_updates}
-\end{figure}
-We recorded update events and Figure~\ref{fig:from_to_updates} shows how devices upgrade between different versions of Android.
-Mostly the dark cells are are upgrades (above the diagonal) (\daNumUpdatesUpgrades).
-While many upgrades are from one version to the next version there are also a fair number (\daNumUpdatesBigUpgrades, \daPercBigUpgrades) which skip more than \daNumUpdatesSkippedBig\ versions.
-Surprisingly there are also a small number of downgrade events (\daNumUpdatesDowngrades, \daPercUpdatesDowngrades) when older versions of Android are installed on to devices.
-Possible reasons why users are downgrading are to free up space on their device, to make it easier to root or because a new version introduced bugs.
-
-The number of devices getting security updates each week, is shown in Figure~\ref{fig:weekly_security_updates}.
-\begin{itemize}
- \item Updates which changed the Android version number from a version with known vulnerabilities to one which had fewer known vulnerabilities are shown in red.
- \item Updates which changed the build number but not the version number and so might contain a backported fix for a vulnerability are shown in yellow.
-\item Updates which did not fix security vulnerabilities (because there were no known security vulnerabilities in the version of Android they were already running) are shown in green.
-
-\end{itemize}
-
-By taking the number of updates observed and the number of device days of data we have collected we can compute the number of updates received by a device per year to be \daUpdatesPerYear.
-This compares badly with the number of critical vulnerabilities discovered per year of between \avoVulnsPerYearAllAndroid\ (affecting all Android) and \avoVulnsPerYear\ (including the manufacturer specific ones in \avo).
-
-\input{security_scoring}
-
-\section{Threats to validity}
-\subsection{The \da\ data gives a conservative estimate of the Android version distribution}
-\label{sec:representative}
-\begin{figure}
- \centering
- \begin{subfigure}[b]{\columnwidth}
- \includegraphics[width=\columnwidth]{figures/googleplayapi}
- \caption{Google Play data on proportion of devices running different Android API versions}
- \label{fig:play_api}
-\end{subfigure}
-\begin{subfigure}[b]{\columnwidth}
- \includegraphics[width=\columnwidth]{figures/norm_api_gpcomp}
- \caption{\da\ data on proportion of devices running different Android API versions}
- \label{fig:da_api}
-\end{subfigure}
-\caption{Monthly Android API version data}
-\end{figure}
-\begin{figure}
- \centering
- \includegraphics[width=\columnwidth]{figures/api_gpcomp_rdiff}
- \caption{Difference between \da\ and Google Play data on the proportion of devices running different Android API versions}
- \label{fig:da_gp_comp_diff}
-\end{figure}
-The data from \da\ we used to investigate the proportion of devices exposed to different vulnerabilities is the OS version.
-Unfortunately there is no authoritative source of OS version information and so we cannot check directly whether our data is representative.
-However Google has published API version information every month since December 2009 and we have collated this information.\footnote{\url{http://androidvulnerabilities.org/play/historicplaydashboard}}
-While API versions are too coarse grained to use for security update detection they are closely related to OS versions and so if the \da\ data on API versions is similar to the Google Data on API versions then the \da\ data on OS versions should be representative.
-Figure~\ref{fig:play_api} shows the data from Google and Figure~\ref{fig:da_api} shows the data from \da\ and they appear similar.
-Figure~\ref{fig:da_gp_comp_diff} shows the difference between Figures~\ref{fig:play_api} and \ref{fig:da_api}, normalising for days since the API version was released.
-It shows that the \da\ data systematically overestimates the prevalence of new API versions and underestimates the prevalence of old API versions.
-This means that the OS version information from \da\ is likely to be overestimating the prevalence of new OS versions and hence our results are a conservative estimate of the security of Android.
-\todolater{we want a statistical metric to claim this strongly with.}
-This allows us to have confidence in the OS version information.
-
-\subsection{Changes in the \da\ sample}
-The \da\ data is mostly generated by devices which have the \da\ app installed because their owner happened to come across the app on Google Play as there has been no advertising and little other attempts to increase usage.
-However two collaborations with network operators to install \da\ on customer devices resulted in large numbers of new \da\ users coming from particular areas and network operators.
-One study was conducted in Norway with 654 installs of \da\ and one was conducted in Bangladesh where 2463 users installed \da.
-This is responsible for the sudden change in distribution of OS versions in \da\ shown towards the end of Figures \ref{fig:norm_os}, \ref{fig:nvulnerabilities_heat} and \ref{fig:proportioninsecure} as for that period Bangladeshi users contributed over half of the \da\ data.
-%We could normalise out the influence of the Bangladeshi users, however they are Android users too and they also need security updates.
-
-\subsection{Uncertainty propagation}
-\label{sec:uncertainty}
-We have endeavoured to propagate information on uncertainties in our results through all calculations with the standard deviation indicated by $\pm$.
-However this technique is unable to account for systematic errors.
-
-\section{Discussion}
-There are continuing efforts to reduce the impact of critical vulnerabilities, both in Android and more widely.
-SEAndroid~\cite{Smalley2013} which is included in Android from version 4.1~\cite{jelly-bean-release} claimed to prevent some root vulnerabilities and to reduce the impact of others.
-Capability based enforcement systems such as Capsicum~\cite{Watson2010} substantially reduce the capabilities that an exploit has to try and gain increased privilege with.
-When Capsicum is included in Linux~\cite{TODO} and hence in Android, it could be used to place fine grained restrictions on system daemons, preventing vulnerabilities becoming critical vulnerabilities.
-
-Malicious apps which use exploits to take control of an Android device must be installed on the device and to an extent this can be prevented if all apps are downloaded from the Google Play store.
-It is protected by Bouncer's static analysis and prompt takedowns of apps reported to be malicious and has been found to have infection rates of 0.02\%~\cite{Zhou2012a}.
-However some apps are installed from other markets which are less well policed (infection rates of between 0.20\%~ and 0.47\%~\cite{Zhou2012a}).
-Apps can also dynamically download code at runtime which then cannot be detected by static analysis and so bypasses Bouncer.
-
-There are also other ways that malicious code could end up running on Android devices, the addJavascriptInterface vulnerability (CVE-2012-6636 ) allows javascript running in a webview which has an expose interface to execute arbitrary Java code and hence load exploits.
-While fixes have been produced for this the change is API breaking and so to be safe an app must target an API version greater than 17 and be running on a device with an Android version of at least 4.2.
-Since the html loaded into webviews (for example to display adverts) is often loaded over insecure http, malicious javascript can easily be injected by an attacker on the network.
-
-We have not included all the vulnerabilities from \avo\, in particular we have not included any manufacturer specific vulnerabilities, even when these are widespread (such as vulnerabilities affecting Qualcomm\footnote{Qualcomm is particularly good at publicly disclosing the vulnerabilities affecting their code and the patches which fix them.\url{https://www.codeaurora.org/projects/security-advisories/}} chipsets) because it is difficult to work out which devices are affected.
-There also tends to be less public information available about manufacturer specific vulnerabilities which makes them harder to tie down.
-For other vulnerabilities which might affect all Android devices it is hard to work out which devices they affect.
-For example \emph{pty race}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/pty_race}} is a Linux kernel vulnerability, we found 48 commits which fixed it in different branches, it was also accidentally fixed and later reintroduced and so determining which devices were vulnerable relies on knowing where on which branch the kernel was taken from.
-For some vulnerabilities the corresponding patch is not in AOSP despite some manufacturers having shipped builds containing the fix, which makes working out what is going on more difficult.
-For \emph{RageAgainstTheCage adb}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/RageAgainstTheCage_adb}} and \emph{keystore buffer}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/keystore_buffer}} vulnerabilities which we do have sufficient data it is not clear whether they are truly critical vulnerabilities as the former may require physical ADB access and the latter may be protected by a sandbox.
-However those two vulnerabilities make little difference to our results as they affect few devices.
-
-\subsection{Open questions}
-There are several open questions which we have not yet been able to answer and remain future work.
-\begin{itemize}
- \item Why was there a gap in vulnerability discovery in Android in 2012?
- \item Why do so many devices have ADB enabled? \daAdbEnabledPerc\ seems rather high for a developer feature.
- \item How much of updating is new handsets and how much is updates being deployed?
-Over time newer versions of Android come to dominate, but quantifying what proportion of that is due to new phones being purchased and what proportion is due to updates to existing devices is a question we cannot answer.
- The only available large data set we know of with longitudinal traces of Android devices OS version numbers is \da\ but that only records \daNumUpdatesUpgrades\ upgrades which broken down over \daOSMonthsOfData\ months and \daNumSigOSVersions\ is only \FPeval{\daUpdatesPerMonthPerVersion}{round(\daNumUpdatesUpgrades/(\daOSMonthsOfData * \daNumSigOSVersions),1)}\daUpdatesPerMonthPerVersion\ updates per version per month which is not enough to build a statistically significant prediction of the expected transition to newer versions of Android.
- \item Why do users downgrade their phones?
-\end{itemize}
-
-\section{Related work}
-\label{sec:related}
-We found that since 2011 \daMeanInsecurityPerc\ of devices on average were exposed to known critical vulnerabilities.
-In 2011 Felt et al.\ \cite{Felt2011} studied 6 Android handsets and found that they were exposed to root vulnerabilities at least 74\% of the time.
-This might indicate that over the last 3 years the security of Android has not improved, but these data are not directly comparable as their study was on 6 handsets considering the best possible update distribution while ours is on a large sample of devices with the real update distribution.
-They also found that 4 of the 46 malware samples (8\%) they analysed contained root exploits, much lower than rates found in later larger studies which found rates of 36.7\%~\cite{Zhou2012b} and 40\%~\cite{Zhou2012a}.
-Our results show root exploits are still a severe threat.
-
-Finding vulnerabilities is frequently a time consuming manual process but, Brumley et al. showed that it is possible to automatically generate exploits from binary patches~\cite{Brumley2008}.
-In a similar way some of the vulnerabilities\todolater{which?} in \avo\ which have been exploited were publicly found by examining the commit made to the Android which fixed the vulnerability.
-This effect has been observed before in the Firefox source code repositories~\cite{Barth2011}, Google does not release the source code until after the release of the update reducing this effect to the delay between an update being released and it reaching the last end user device.
-There have been other attempts to automatically find vulnerabilities.
-The Woodpecker tool automatically finds permission leaks in stock Android phone images~\cite{Grace2012}.
-The update process itself can allow Android apps to gain privilege through `Pileup' vulnerabilities by registering for new permissions before the update which creates that permission is installed~\cite{Xing2014}.
-
-The security protections used in iOS such as, app review, code signing, Data Execution Prevention (DEP) and Address Space Layout Randomisation (ASLR) and mandatory access controls has resulted in a lower level of malware affecting iOS~\cite{Felt2011} but can still be bypassed~\cite{Wang2013a}.
-
-Various attempts have been made to detect malware.
-RiskRanker classified 3,281/118,318 (2.8\%) apps as risky of which 718 (22\%) were malware and 322 (10\%) were previously unknown malware, an infection rate of 0.6\% across multiple markets~\cite{Grace2012a}.
-It uses signatures of exploits, static analysis of calls to high risk operations, detection of encrypted native code execution and dynamic Dalvik code loading to detect risky apps.
-DroidRanger found 148/182,823 (0.08\%) apps to be malicious across multiple markets of which 29 were previously unknown~\cite{Zhou2012a}.
-It used permission-based behavioural fingerprinting -- looking at the permissions of known malware and heuristic-based filtering -- dynamic loading of both Dalvik and native code.
-AnDarwin detects similar apps, it found 169/265,359 (0.06\%) malicious apps based on the difference in permissions between the malicious app and the app which had been cloned~\cite{Crussell2013}.
-It used clustering based on semantic vectors derived from the program dependence graphs to detect similar apps.
-The Malware Genome project collected 1,260 malware samples from 2010--2011~\cite{Zhou2012b}.
-They found that the best case for anti-malware detection on this data was 79.6\%.
-However DroidChameleon found that antivirus products could not detect malware if it was simply automatically permuted~\cite{Rastogi2013}.
-All these app analysis projects have been hampered by the difficulty of obtaining full datasets of Android apps as Google does not make these available (and researchers who automatically download them, violate Google's Terms of Service).
-PlayDrone was a particularly effective project which circumvented Google's protective measures and downloaded over 1,100,000 apps from Google Play, allowing an in depth analysis~\cite{Viennot2014}.
-
-Examining security updates has been a key way of assessing security for many years.
-Many software update systems have been found not to authenticate the connection to the update server and/or do not authenticate the downloaded binaries~\cite{Bellissimo2006}.
-Even package management systems designed to provide secure updates have been found to have vulnerabilities~\cite{Cappos2008}.
-Fortunately Android does authenticate the update binaries (though the APK vulnerabilities circumvented this) and Google Play downloads them over a secure connection~\cite{Viennot2014}.
-\todolater{double check that}
-
-The updatedness rate for Android of \daUpdatednessPerc\ compares unfavourably with the more than 90\% rate for Windows XP SP2 computers contacting the Microsoft update servers~\cite{Gkantsidis2006} though that is within one OS version and only for those computers which did connect to the servers -- but that is the default.
-However 27\% of Windows computers were still running Windows XP in July 2014\footnote{\url{https://archive.today/PLGxn}\todolater{Get updated figures before publication}} -- four months after it went out of security support.
-
-User-Agent strings have been used to investigate how web browser versions change with updates~\cite{Frei2008} and found that at most 80\% of Firefox users were running the most recent version.
-The same analysis was used to show that Chrome's use of silent updates seems to increase uptake of upgrades~\cite{Duebendorfer2010} with 97\% of users running the latest version within 3 weeks of release.
-Android's update process is manual, the user is notified but must action the upgrade which will then download the update and install it, the phone is rendered inoperable during the update process which is inconvenient for users and the phone must have sufficient charge so that it will not run out during the update process.
-Partly this is the result of the fact that an operating system update is being installed and so a reboot is required, but Chrome installs the new version side by side with the old one and switches the next time it is restarted.
-The same technique would be more difficult on phones with limited storage space (as for many cheap Android phones which have barely enough space to install just the update) but is a plausible improvement for more high-end devices.
-Google is deploying the same silent update technique through Google Play Services~\cite{TODO} which automatically installs updates for core Google components of Android, this also bypasses the manufacturer and operator.
-
-\section{Conclusion}
-\label{sec:conclusion}
-We have investigated the security of the Android operating system with respect to security updates and examined how the Android ecosystem results in different devices receiving different levels of security due to whether or not they get security updates.
-We have compared different models, manufacturers and network operators and found that there are differences between them which the discerning purchaser might use to influence their decision about which device to buy from which manufacturer and operator.
-We hope that this analysis will encourage manufacturers and operators to improve the support they provide for devices after sale.
-Android attempts to provide good security to end users but the latency in the security update process means that on average \daMeanInsecurityPerc\ of Android devices are exposed to known critical vulnerabilities.
-
-\section*{Acknowledgements}
-Thanks to David Robertson for helpful advice on statistical analysis.
-Thanks to Laurent Simon, Thomas Coudray, Adrian Taylor, Justin Case, Giant Pune and Khilan Gudka for reporting vulnerabilities in Android.
-
-\printbibliography
-
-\listoftodos
-
 \section{Cutting floor}
@@
 We hope that a consequence of this paper will be a change in behaviour of the Android ecosystem which invalidates this model by speeding up the deployment of new OS versions.
-
-\end{document}

### 2014-11-10 securityupdates.tex
Fix TODOs and move cuttingfloor to a separate file to allow sending out for review

@@
 \usepackage{fp}
-\usepackage{todonotes}
+\usepackage[disable]{todonotes}
 \let\OldTodo\todo
@@
 There was also a period of several months when no devices ran secure versions of Android.
-We compare different manufacturers and operators to find which are best at promptly distributing security updates and examine how to improve the situation. %TODO operators analysis
+We compare different manufacturers and operators to find which are best at promptly distributing security updates and examine how to improve the situation.
 \end{abstract}
@@
 
-\section{Cutting floor}
-
-This section contains things which we were intending to include but which have not made the cut.
-
-\subsection{Running a version of Android known to be vulnerable makes a device vulnerable}
-We have not tried to exploit these vulnerabilities on the devices in order to test whether the devices are actually vulnerable.
-Doing so would be difficult both ethically, as it would pose a high risk to user devices and mean that our code could easily be maliciously repurposed, and practically as getting all the exploits working reliably on all the devices they could be made to work reliably on requires extensive testing.
-We are confident that most devices will be vulnerable if they are running a version of Android known to be vulnerable. \todo{WHY?}
-
-\subsection{Monetisation of malware}
-Android malware can make money in a variety of different ways~\cite{Felt2011}.
-One of the key differences between smartphones and desktops is that it is easier to directly cash out on a smartphone as malware can send premium rate texts or make premium rate phone calls or send SMS and email spam.
-It can also steal personal data which already has semantic information associated with it, this also makes it easier to sell and to perform ransomware attacks where malware deletes this information from the device and charges a ransom for its return.
-It can steal credentials for external services or use in-app billing.
-It can also do more conventional things such as use the devices for DDoS, SEO or click fraud attacks and altering the ads displayed to send the fees to the malware author.
-While many of these things can be done by malware which just requests the relevant permissions, root exploits make this easier, particularly for ransomware and altering the ads displayed.
-
-\subsection{Incentives and participants}
-\label{sec:economics}
-Google and other Android vendors care about the reputation of Android, particularly from a security standpoint as if users are afraid to use their devices for certain purposes (e.g.\ banking) then they will not be willing to pay as much for them and money cannot be made from those activities.
-Manufacturers make money by selling the device, once sold they want to prevent the cost of return under warranty and so are adverse to the risk of breaking the device, but they do not need to do things to make the device better as it has already been sold.
-However if they gain a reputation for abandoning devices that might decrease future sales.
-
-Operators sell phones under contracts, typically at least a year long, they want the user to not request support due to the device breaking.
-They might also gain advantage if the user is happy with the device after the end of the contract as they might then keep paying the same rate but there would no longer be a capital cost to pay off.
-They care about phones not getting malware which causes support requests or uses the mobile network to perform malicious actions as they may have to reimburse the cost of premium SMS's if they were not authorised by the user.
-
-Users want to buy phones which will serve their use cases and which are secure so that they do not experience negative consequences.
-
-% We don't need this for this paper
-%\subsection{Comparison with iOS}
-%\todo{Get data from Richard so that we can do this comparison}
-
-\subsection{Experiment 4: Predicting future version distributions}\label{sec:exp:predicting_distributions}
-\subsubsection{Method}
-Examining the way that the OS version distribution changes over time in Figure~\ref{fig:norm_os} we see some regularity to the behaviour.
-To show this behaviour more clearly we normalised the data to the number of days since the version was released and smoothed it by forcing it to be monotonically decreasing.\footnote{While sometimes an older version becomes more popular after a later version has been released causing the values to actually increase, mostly this was just noise or changes in the distribution of devices (e.g.\ the sudden arrival of 1000 devices from Bangladesh due to a study being conducted there) and this assumption makes the graph clearer.}
-Then we can plot Figure~\ref{fig:frvh_os_versions} which shows the proportion of of devices in \da\ not upgraded to each Android version or a later Android version (i.e.
-for 2.1.0 it includes 2.1.1, 2.2.0 and indeed all devices as they all run later versions of Android).
-The mean and standard deviation of the same data are shown in Figure~\ref{fig:frms_os_versions}.
-
-\begin{figure}
- \includegraphics[width=\columnwidth]{figures/frvh_os_versions}
- \caption{Proportion of devices not updated to particular versions of Android or any later version}
- \label{fig:frvh_os_versions}
-\end{figure}
-\begin{figure}
- \includegraphics[width=\columnwidth]{figures/frms_os_versions}
- \caption{Proportion of devices not updated to particular versions of Android or any later version}
- \label{fig:frms_os_versions}
-\end{figure}
-
-The data appears to have an exponential decay as it tends to 1 but with a slow start as few devices are upgraded initially.
-We approximate this with a $e^{-x}$ for the main behaviour and a which is offset to start later when uptake takes off.
-The equation is then:
-\begin{equation}
-y =
- \begin{cases}
-  1.0 & \text{if }x<\mathrm{undeployed}\\
-  e^{-\mathrm{decay}(x-\mathrm{undeployed})} & \text{otherwise}
- \end{cases}
-\end{equation}
-
-\subsubsection{Results}
-Fitting that curve to the data using \texttt{scipy.optimize}, we get a RMSE of \daOSCurveFitRMSE\ and parameters $\mathrm{undeployed} = \daOSCurveFitParamFirst ~\si{days}$, $\mathrm{decay} = \daOSCurveFitParamSecond ~\si{days^{-1}}$ for OS versions and RMSE of \daAPICurveFitRMSE\, $\mathrm{undeployed} = \daAPICurveFitParamFirst ~\si{days}$, $\mathrm{decay} = \daAPICurveFitParamSecond ~\si{days^{-1}}$ for API versions.
-From this the number of days from release of a new version of Android until 50\% of devices are running that version or higher is \daOSCurveHalfDeployed\ and full deployment to \daFullDeployedAt\ of devices takes \daOSCurveFullDeployed\ days.
-For API versions this is \daAPICurveHalfDeployed\ and \daAPICurveFullDeployed\ respectively.
-
-Hence if a security vulnerability is fixed through the release of a particular Android version it will be \daOSCurveFullDeployed\ days after until the fix is fully deployed.
-
-The equation we chose for the model has a root mean squared error (RMSE) of \daOSCurveFitRMSE\ for OS versions, which is a comparably good fit to a standard polynomial fit (3 degree polynomial fit gave a RMSE of \daOSCurvePolyRMSE) or a spline fit (RMSE of \daOSCurveSplineRMSE) but gives a meaningful model of behaviour rather than a generic curve.
-
-\todo{Our handling of errors here is not quite good enough as the uncertainty from our y value interacts badly with the logarithm function when computing the inverse and we compute the log of ~0 which means our error ends up much bigger than our value - but that is an upper bound, not a +- bound.}
-
-We hope that a consequence of this paper will be a change in behaviour of the Android ecosystem which invalidates this model by speeding up the deployment of new OS versions.
-
 \end{document}

### 2014-11-10 securityupdates.tex
New abstract; minor fixes to the intro.

--- 2014-11-10 securityupdates.tex
@@
 \begin{abstract}
-Android is the most popular smartphone platform and seeks to provide better security than popular desktop operating systems through increased compartmentalisation and a permission system for apps.
-This security relies on the absence of root privilege exploits which allow apps to bypass protections.
-Many such exploits have been found and fixed.
-Fixing them is not enough, we show that these fixes do not promptly reach the users devices and so much of the time users are running versions of Android known to be vulnerable.
-We used \da\ data from over \daOSYearsOfData\ years and \daNumOSDevices\ devices and found that on average \daMeanInsecurityPercNominal\ of devices were exposed to known root privilege vulnerabilities and only \daUpdatednessPercNominal\ of devices run the most recent version of Android.
-There was also a period of several months when no devices ran secure versions of Android.
-We compare different manufacturers and operators to find which are best at promptly distributing security updates and examine how to improve the situation. %TODO operators analysis
+Modern smartphone operating systems protect users from malicious applications by providing a protected execution environment or sandbox.
+This approach relies on the absence of active exploits which allow apps to break out of the sandbox.
+Many such exploits have been found, and therefore the security of a smartphone operating system relies on the prompt delivery of security updates.
+In this paper we tease apart the Android software ecosystem and quantify the extent to which the security sandbox provided by Android protects devices from malicious apps.
+We analysed \da\ data from over \daOSYearsOfData\ years and \daNumOSDevices\ devices and found that on average \daMeanInsecurityPercNominal\ of devices were exposed to known root privilege vulnerabilities and only \daUpdatednessPercNominal\ of devices run the most recent version of Android.
+We find there is significant variability in the timely delivery of security updates across different manufacturers and mobile network operators, and we define a security metric to rank the performance of manufacturers and operators in our data set.
 \end{abstract}
@@
 Support for third-party apps is a key feature of modern smartphone operating systems.
-Such apps are written by many developers from a wide range of backgrounds, which means neither operating system vendors nor users can, fully trust app developers.
-Therefore, in order to secure personal data and prevent the use of expensive operations such as the sending of premium-rate text messages, operating system vendors provide a protected execution environment, or \emph{sandbox}, for apps.
+Such apps are written by many developers from a wide range of backgrounds, which means neither operating system vendors nor users can fully trust app developers.
+Therefore, in order to secure personal data and prevent the use of expensive operations, such as the sending of premium-rate text messages, operating system vendors provide a protected execution environment, or \emph{sandbox}, for apps.
 
@@
 We focus on this issue because malware which is able to take full control of a handset can do significantly more harm and is much harder to remove.
-Ordinary malware can uninstalled by the user or remotely by Google through the Play Store, but such removal is unlikely to work once the security features of Android are broken.
+Ordinary malware can be uninstalled by the user, or remotely by Google through the Play Store, and returns the device to a secure state. 
+Removing an app which has used an exploit to take control of the handset is unlikely to return the device to secure state.
 
-The development of Android is distributed among many parties.
-The security of Android relies on many open source projects, include the Linux kernel, OpenSSL and BouncyCastle as well as Google who build the core platform. 
+The development of Android is distributed among many parties and the security of Android relies on many open source projects, include the Linux kernel, OpenSSL and BouncyCastle as well as Google who build the core platform.
 In addition, manufacturers (of which we are aware of \daNumManufacturers\footnote{\label{foot:dadata}We computed this from the \da~\cite{Wagner2013} data, see \S\ref{sec:android_update_process}.}) and network operators (\daNumOperators\dafoot) may make further modifications before devices are shipped to customers.
@@
 In order to determine whether an individual smartphone is vulnerable to a particular exploit, we need to combine vulnerability data with the operating system version string and additional update information from the handset.
-For this we use handset data gathered from over 18,000 participants in the \da~\cite{Wagner2013} project.
+For this we use handset data gathered from \daNumOSDevices\ devices in the \da~\cite{Wagner2013} project.
 This analysis allows us to plot the proportion of Android devices in the \da\ data which were running versions of Android known to have vulnerabilities at the time and is shown in Figure~\ref{fig:proportioninsecure}.
@@
 
-An alternative to regulation is better public data. 
+An alternative to regulation is access to publicly available data. 
 Hard numbers provide an incentive to both manufacturers and operators to provide updates.

### 2014-11-10 securityupdates.tex

--- 2014-11-10 securityupdates.tex
@@
 \begin{abstract}
-Android is the most popular smartphone platform and seeks to provide better security than popular desktop operating systems through increased compartmentalisation and a permission system for apps.
-This security relies on the absence of root privilege exploits which allow apps to bypass protections.
-Many such exploits have been found and fixed.
-Fixing them is not enough, we show that these fixes do not promptly reach the users devices and so much of the time users are running versions of Android known to be vulnerable.
-We used \da\ data from over \daOSYearsOfData\ years and \daNumOSDevices\ devices and found that on average \daMeanInsecurityPercNominal\ of devices were exposed to known root privilege vulnerabilities and only \daUpdatednessPercNominal\ of devices run the most recent version of Android.
-There was also a period of several months when no devices ran secure versions of Android.
-We compare different manufacturers and operators to find which are best at promptly distributing security updates and examine how to improve the situation.
+Modern smartphone operating systems protect users from malicious applications by providing a protected execution environment or sandbox.
+This approach relies on the absence of active exploits which allow apps to break out of the sandbox.
+Many such exploits have been found, and therefore the security of a smartphone operating system relies on the prompt delivery of security updates.
+In this paper we tease apart the Android software ecosystem and quantify the extent to which the security sandbox provided by Android protects devices from malicious apps.
+We analysed \da\ data from over \daOSYearsOfData\ years and \daNumOSDevices\ devices and found that on average \daMeanInsecurityPercNominal\ of devices were exposed to known root privilege vulnerabilities and only \daUpdatednessPercNominal\ of devices run the most recent version of Android.
+We find there is significant variability in the timely delivery of security updates across different manufacturers and mobile network operators, and we define a security metric to rank the performance of manufacturers and operators in our data set.
 \end{abstract}

### 2014-11-10 securityupdates.tex
Spelling and minor fixes from Daniel Wagner

@@
  \item We characterise the Android update ecosystem, showing how updates flow between entities, and quantify vulnerabilities by determining what proportion of devices they affected and how that changes over time.
- \item We collate quantitive information on vulnerabilities which affect Android and investigate their time-lines.
+ \item We created an open database to collate quantitative information on vulnerabilities which affect Android, and investigate their time-lines.
  \item We measure the security of Android according to three metrics and compare different manufacturers and device models to allow device purchasers to differentiate between them based on security.
@@
 
-Various different kinds of data\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/collected.htm}} are collected by the \da\ Android app\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/}}.
+Various different kinds of data\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/collected.html}} are collected by the \da\ Android app\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/}}.
 Among these are the build string and API version for the version of Android currently running on the device each day.
@@
 If an application exploits a critical vulnerability then it gains control of the device.
-Some phones can be `rooted' by enabling USB debugging and using the special privileges of the adb shell to root the device but only \daAdbEnabledPerc\dafoot\textsuperscript{,}\footnote{Where possible we indicate the uncertainty in our results by giving them $\pm$ the standard deviation.
+Some phones can be `rooted' by enabling USB debugging and using the special privileges of the \texttt{adb} shell to root the device but only \daAdbEnabledPerc\dafoot\textsuperscript{,}\footnote{Where possible we indicate the uncertainty in our results by giving them $\pm$ the standard deviation.
 This is explained further in \S\ref{sec:uncertainty} but occasionally results in $\pm0$.} of devices have USB debugging enabled.
 This is not something that applications running on the phone can exploit to break compartmentalisation and so we do not include those vulnerabilities.
-Unfortunately many published exploits use adb for convenience and so determining whether that is necessary to exploit the vulnerability can be difficult.
+Unfortunately many published exploits use \texttt{adb} for convenience and so determining whether that is necessary to exploit the vulnerability can be difficult.
 
@@
 
-Working out the date to use for the date on which a vulnerability was sufficiently well known to pose a threat to users is difficult.
+Establishing when a vulnerability started posing a threat to users is difficult.
 Frei et al.~\cite{Frei2010} propose the definition of the {\bf time of disclosure} when the information about the vulnerability is freely available to the public, from a widely accepted and independent source and has been validated by security experts so that it has a risk rating.
@@
 
-There are several points we could pick for the point from which a vulnerability poses substantial risk to users.
+There are several plausible points from which a vulnerability poses a threat to users.
 {\bf Creation} and {\bf introduction} are the earliest points which could be used but the risk is mostly latent until someone discovers them.
@@
 Figure~\ref{fig:vulnerabilities_timeline} shows the dates of discovery and, when later, the dates when a version of Android which fixed the vulnerability was shipped.
-Some vulnerabilities (RageAgainstTheCage adb, levitator, zergRush abd TwerkMyMoto) were fixed in released versions of Android before they were discovered and so are shown as vertical lines while others were known for months before a version of Android which fixed them was shipped.
+Some vulnerabilities (\emph{levitator}, \emph{zergRush}) were fixed in released versions of Android before they were discovered and so are shown as vertical lines while others were known for months before a version of Android which fixed them was shipped.
 During the period in which \da\ data was collected the date when a version of Android with the fix was observed on a \da\ device is taken as the date that version was released.
@@
 For some vulnerabilities the corresponding patch is not in AOSP despite some manufacturers having shipped builds containing the fix, which makes working out what is going on more difficult.
-For \emph{RageAgainstTheCage adb}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/RageAgainstTheCage_adb}} and \emph{keystore buffer}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/keystore_buffer}} vulnerabilities which we do have sufficient data it is not clear whether they are truly critical vulnerabilities as the former may require physical ADB access and the latter may be protected by a sandbox.
+For \emph{RageAgainstTheCage adb}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/RageAgainstTheCage_adb}} and \emph{keystore buffer}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/keystore_buffer}} vulnerabilities which we do have sufficient data it is not clear whether they are truly critical vulnerabilities as the former may require physical \texttt{adb} access and the latter may be protected by a sandbox.
 However those two vulnerabilities make little difference to our results as they affect few devices.
@@
  \item Why was there a gap in vulnerability discovery in Android in 2012?
- \item Why do so many devices have ADB enabled? \daAdbEnabledPerc\ seems rather high for a developer feature.
+ \item Why do so many devices have \texttt{adb} enabled? \daAdbEnabledPerc\ seems rather high for a developer feature.
  \item How much of updating is new handsets and how much is updates being deployed?

### 2014-11-10 securityupdates.tex
A couple of minor fixes

@@
 Manufacturer and Operator counts were obtained by normalising the results reported by Android to \da\ of the device manufacturer and active network operator.
-This normalisation is a manual task involves removing invalid values (such as `manufacturer' or `airplane mode is on'), collating across company name changes (e.g.\ `lge' to `lg'), normalising punctuation, removing extra strings sometimes added such as (`(2g)' or `communications') and mapping some incorrectly placed model names back to their manufacturer.
+This normalisation is a manual task involves removing invalid values (such as `manufacturer' or `airplane mode is on'), collating across company name changes (e.g.\ `lge' to `LG'), normalising punctuation, removing extra strings sometimes added such as (`(2g)' or `communications') and mapping some incorrectly placed model names back to their manufacturer.
 This normalisation is not perfect and so these are overestimates on the \da\ data but they are likely still underestimates as there will be some manufacturers and operators which are not included in the \da\ data.
@@
 We recorded update events and Figure~\ref{fig:from_to_updates} shows how devices upgrade between different versions of Android.
-Mostly the dark cells are are upgrades (above the diagonal) (\daNumUpdatesUpgrades).
+Mostly the dark cells are upgrades (above the diagonal) (\daNumUpdatesUpgrades).
 While many upgrades are from one version to the next version there are also a fair number (\daNumUpdatesBigUpgrades, \daPercBigUpgrades) which skip more than \daNumUpdatesSkippedBig\ versions.

### 2014-11-11 avostats.tex
Updates following comments from Richard Clayton

@@
 \newcommand{\avoVulnsPerYearAllAndroid}{$3.7 \pm 1.0$}
-\newcommand{\avoTabAndVulns}{\begin{table} \centering \begin{tabular}{l|c|l} Vulnerability & Date known & How known \\ \hline KillingInTheNameOf psneuter ashmem & 2010-07-13 & Fixed on \\ exploid udev & 2010-07-15 & Discovered on \\ levitator & 2011-03-10 & Discovered on \\ Gingerbreak & 2011-04-18 & Fixed on \\ zergRush & 2011-10-06 & Discovered on \\ APK duplicate file & 2013-02-18 & Discovered on \\ APK unchecked name & 2013-06-30 & Discovered on \\ APK unsigned shorts & 2013-07-03 & Fixed on \\ vold asec & 2014-01-27 & Fixed on \\ Fake ID & 2014-04-17 & Fixed on \\ TowelRoot & 2014-05-03 & Discovered on \\\end{tabular} \caption{Root equivalent vulnerabilities in Android} \label{tab:andvulns} \end{table}}
+\newcommand{\avoTabAndVulns}{\begin{table} \centering \begin{tabular}{l|l|c} Vulnerability & How known & Date \\ \hline KillingInTheNameOf psneuter ashmem & Fixed on & 2010-07-13 \\ exploid udev & Discovered on & 2010-07-15 \\ levitator & Discovered on & 2011-03-10 \\ Gingerbreak & Fixed on & 2011-04-18 \\ zergRush & Discovered on & 2011-10-06 \\ APK duplicate file & Discovered on & 2013-02-18 \\ APK unchecked name & Discovered on & 2013-06-30 \\ APK unsigned shorts & Fixed on & 2013-07-03 \\ vold asec & Fixed on & 2014-01-27 \\ Fake ID & Fixed on & 2014-04-17 \\ TowelRoot & Discovered on & 2014-05-03 \\\end{tabular} \caption{Root equivalent vulnerabilities in Android} \label{tab:andvulns} \end{table}}
 \newcommand{\avoNumBigExternalProjects}{40}

### 2014-11-11 dastats.tex
Updates following comments from Richard Clayton

@@
 \newcommand{\daNumDataPoints}{153 Billion}
-\newcommand{\daNumStartedApps}{121998}
-\newcommand{\daNumInstalledApps}{174323}
+\newcommand{\daNumStartedApps}{122032}
+\newcommand{\daNumInstalledApps}{174390}
 \newcommand{\daNumVulnsUsed}{11}
@@
 \newcommand{\daSigNumDays}{100}
-\newcommand{\daNumManufacturers}{260}
-\newcommand{\daNumSigManufacturers}{8}
-\newcommand{\daNumModels}{2374}
+\newcommand{\daNumManufacturers}{1}
+\newcommand{\daNumSigManufacturers}{-1}
+\newcommand{\daNumModels}{2379}
 \newcommand{\daNumSigModels}{15}
-\newcommand{\daOSVersionPercValidLines}{99.47\%}
-\newcommand{\daOSTotalDaysData}{1132105}
+\newcommand{\daOSVersionPercValidLines}{99.46\%}
+\newcommand{\daOSTotalDaysData}{1134626}
 \newcommand{\daMeanInsecurityPerc}{$87 \pm 0\%$}
@@
 \newcommand{\daSecurityScore}{$2.8 \pm 0.0$}
-\newcommand{\daNumFullVersions}{1133}
+\newcommand{\daNumFullVersions}{1136}
 \newcommand{\daNumOSVersions}{45}
@@
 \newcommand{\daNumAPIVersions}{15}
-\newcommand{\daNumFullVersionUpdates}{4209}
-\newcommand{\daNumOSVersionUpdates}{3560}
+\newcommand{\daNumFullVersionUpdates}{4212}
+\newcommand{\daNumOSVersionUpdates}{3563}
 \newcommand{\daNumFullOnlyVersionUpdates}{637}
-\newcommand{\daNumSecurityUpdates}{1371}
-\newcommand{\daNumPossibleSecurityUpdates}{1887}
-\newcommand{\daTabSecScoresmanufacturer}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & Security score \\ \hline lg & $0.11 \pm 0.00$ & $0.32 \pm 0.00$ & 0.79 & $3.3 \pm 0.0$ \\  motorola & $0.16 \pm 0.00$ & $0.13 \pm 0.00$ & 0.82 & $2.9 \pm 0.0$ \\  other & $0.13 \pm 0.00$ & $0.056 \pm 0.000$ & 0.68 & $2.7 \pm 0.0$ \\  samsung & $0.14 \pm 0.00$ & $0.047 \pm 0.000$ & 0.76 & $2.6 \pm 0.0$ \\  htc & $0.14 \pm 0.00$ & $0.087 \pm 0.001$ & 0.95 & $2.5 \pm 0.0$ \\  sony & $0.13 \pm 0.00$ & $0.18 \pm 0.00$ & 1.17 & $2.5 \pm 0.0$ \\  asus & $0.14 \pm 0.00$ & $0.56 \pm 0.01$ & 7.18 & $2.2 \pm 0.0$ \\  symphony & $0.0015 \pm 0.0002$ & $0.041 \pm 0.001$ & 5.68 & $0.15 \pm 0.00$ \\ \end{tabular} \caption{Security scores for manufacturers} \label{tab:sec_manufacturer} \end{table}}
-\newcommand{\daTabSecScoresmodel}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & Security score \\ \hline Galaxy Nexus & $0.49 \pm 0.00$ & $0.53 \pm 0.01$ & 2.66 & $4.0 \pm 0.0$ \\  Nexus 4 & $0.19 \pm 0.00$ & $0.82 \pm 0.01$ & 7.22 & $3.2 \pm 0.0$ \\  Nexus 7 & $0.17 \pm 0.00$ & $0.74 \pm 0.01$ & 7.08 & $2.9 \pm 0.0$ \\  Desire HD & $0.083 \pm 0.003$ & $0.051 \pm 0.004$ & 0.39 & $2.9 \pm 0.0$ \\  other & $0.13 \pm 0.00$ & $0.12 \pm 0.00$ & 0.68 & $2.9 \pm 0.0$ \\  HTC Sensation & $0.36 \pm 0.00$ & $0.0063 \pm 0.0062$ & 2.59 & $1.9 \pm 0.0$ \\  GT-I9100 & $0.22 \pm 0.00$ & $0.015 \pm 0.001$ & 2.53 & $1.4 \pm 0.0$ \\  GT-N7000 & $0.25 \pm 0.00$ & $0.0006 \pm 0.0036$ & 3.52 & $1.2 \pm 0.0$ \\  GT-P1000 & $0.0096 \pm 0.0009$ & $0.0000 \pm 0.0061$ & 1.73 & $0.94 \pm 0.02$ \\  HTC Desire S & $0.022 \pm 0.001$ & $0.016 \pm 0.004$ & 2.00 & $0.85 \pm 0.01$ \\  GT-I9300 & $0.14 \pm 0.00$ & $0.014 \pm 0.001$ & 7.24 & $0.62 \pm 0.01$ \\  GT-I9505 & $0.0 \pm 0$ & $0.18 \pm 0.00$ & 8.15 & $0.55 \pm 0.01$ \\  GT-N7100 & $0.064 \pm 0.002$ & $0.0011 \pm 0.0058$ & 8.37 & $0.26 \pm 0.02$ \\  HTC Desire HD & $0.0 \pm 0$ & $0.0005 \pm 0.0070$ & 4.05 & $0.10 \pm 0.02$ \\ \end{tabular} \caption{Security scores for models} \label{tab:sec_model} \end{table}}
-\newcommand{\daTabSecScoressummary}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & Security score \\ \hline nexus & $0.33 \pm 0.00$ & $0.52 \pm 0.00$ & 1.73 & $3.8 \pm 0.0$ \\  notnexus & $0.12 \pm 0.00$ & $0.039 \pm 0.000$ & 0.68 & $2.6 \pm 0.0$ \\ \end{tabular} \caption{Security scores for summaries} \label{tab:sec_summary} \end{table}}
+\newcommand{\daNumSecurityUpdates}{1110}
+\newcommand{\daNumPossibleSecurityUpdates}{2141}
+\newcommand{\daTabSecScoresmanufacturer}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{Security score} \\ &&&& (out of 10) \\ \hline LG & $0.13 \pm 0.00$ & $0.056 \pm 0.000$ & 0.64 & $2.8 \pm 0.0$ \\ \end{tabular} \caption{Security scores for manufacturers} \label{tab:sec_manufacturer} \end{table}}
+\newcommand{\daTabSecScoresmodel}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{Security score} \\ &&&& (out of 10) \\ \hline Galaxy Nexus & $0.54 \pm 0.00$ & $0.52 \pm 0.01$ & 1.53 & $4.8 \pm 0.0$ \\  Nexus 4 & $0.23 \pm 0.00$ & $0.76 \pm 0.01$ & 6.06 & $3.2 \pm 0.0$ \\  Nexus 7 & $0.19 \pm 0.00$ & $0.73 \pm 0.01$ & 5.92 & $3.0 \pm 0.0$ \\  Desire HD & $0.083 \pm 0.003$ & $0.025 \pm 0.007$ & 0.39 & $2.8 \pm 0.0$ \\  HTC Sensation & $0.36 \pm 0.00$ & $0.0000 \pm 0.0078$ & 1.59 & $2.4 \pm 0.0$ \\  GT-I9100 & $0.22 \pm 0.00$ & $0.012 \pm 0.002$ & 1.20 & $2.3 \pm 0.0$ \\  HTC Desire S & $0.022 \pm 0.001$ & $0.001 \pm 0.010$ & 1.00 & $1.7 \pm 0.0$ \\  GT-N7000 & $0.25 \pm 0.00$ & $0.0006 \pm 0.0035$ & 2.52 & $1.4 \pm 0.0$ \\  GT-P1000 & $0.0096 \pm 0.0009$ & $0.0000 \pm 0.0074$ & 1.73 & $0.94 \pm 0.02$ \\  GT-I9300 & $0.15 \pm 0.00$ & $0.013 \pm 0.001$ & 6.16 & $0.64 \pm 0.01$ \\  GT-I9505 & $0.0 \pm 0$ & $0.18 \pm 0.00$ & 6.83 & $0.56 \pm 0.01$ \\  GT-N7100 & $0.067 \pm 0.002$ & $0.0011 \pm 0.0058$ & 6.91 & $0.28 \pm 0.02$ \\  HTC Desire HD & $0.0 \pm 0$ & $0.0000 \pm 0.0073$ & 3.05 & $0.27 \pm 0.02$ \\ \end{tabular} \caption{Security scores for models} \label{tab:sec_model} \end{table}}
+\newcommand{\daTabSecScoressummary}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{Security score} \\ &&&& (out of 10) \\ \hline nexus & $0.36 \pm 0.00$ & $0.52 \pm 0.00$ & 0.68 & $5.0 \pm 0.0$ \\  notnexus & $0.12 \pm 0.00$ & $0.039 \pm 0.000$ & 0.64 & $2.7 \pm 0.0$ \\ \end{tabular} \caption{Security scores for summaries} \label{tab:sec_summary} \end{table}}
 \newcommand{\daUpdatednessPerc}{$5.6 \pm 0.0\%$}
-\newcommand{\daNumOSDevices}{18873}
+\newcommand{\daNumOSDevices}{18899}
 \newcommand{\daSigNumDevicesDay}{20}
@@
 \newcommand{\daSecScoreWorstmanufacturerScore}{$0.15 \pm 0.00$}
-\newcommand{\daSecScoreBestmodelScore}{$4.0 \pm 0.0$}
-\newcommand{\daSecScoreWorstmodelScore}{$0.10 \pm 0.02$}
+\newcommand{\daSecScoreBestmodelScore}{$4.8 \pm 0.0$}
+\newcommand{\daSecScoreWorstmodelScore}{$0.27 \pm 0.02$}
 \newcommand{\daSecScoreBestsummaryScore}{$4.1 \pm 0.0$}
@@
 \newcommand{\daAdbEnabledPerc}{$20 \pm 0\%$}
-\newcommand{\daNumUpdatesUpgrades}{3323}
-\newcommand{\daNumUpdatesDowngrades}{146}
-\newcommand{\daPercUpdatesDowngrades}{4.21\%}
+\newcommand{\daNumUpdatesUpgrades}{3325}
+\newcommand{\daNumUpdatesDowngrades}{147}
+\newcommand{\daPercUpdatesDowngrades}{4.23\%}
 \newcommand{\daStartDate}{2011-07-01}
-\newcommand{\daEndDate}{2014-11-03}
+\newcommand{\daEndDate}{2014-11-04}
 \newcommand{\daNumOperators}{1393}
-\newcommand{\daTabSecScoresoperator}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & Security score \\ \hline o2 uk & $0.24 \pm 0.00$ & $0.12 \pm 0.00$ & 0.37 & $3.8 \pm 0.0$ \\  orange & $0.21 \pm 0.00$ & $0.085 \pm 0.003$ & 0.39 & $3.5 \pm 0.0$ \\  t-mobile & $0.18 \pm 0.00$ & $0.18 \pm 0.00$ & 0.54 & $3.5 \pm 0.0$ \\  sprint & $0.18 \pm 0.00$ & $0.11 \pm 0.00$ & 0.44 & $3.4 \pm 0.0$ \\  3 & $0.14 \pm 0.00$ & $0.093 \pm 0.003$ & 0.49 & $3.1 \pm 0.0$ \\  vodafone uk & $0.13 \pm 0.00$ & $0.10 \pm 0.00$ & 0.52 & $3.1 \pm 0.0$ \\  at\&t & $0.12 \pm 0.00$ & $0.077 \pm 0.002$ & 0.51 & $3.0 \pm 0.0$ \\  verizon & $0.18 \pm 0.00$ & $0.087 \pm 0.002$ & 0.89 & $2.7 \pm 0.0$ \\  unknown & $0.095 \pm 0.001$ & $0.22 \pm 0.00$ & 1.15 & $2.5 \pm 0.0$ \\  n telenor & $0.023 \pm 0.001$ & $0.10 \pm 0.00$ & 1.57 & $1.4 \pm 0.0$ \\  airtel & $0.066 \pm 0.002$ & $0.028 \pm 0.002$ & 1.89 & $1.1 \pm 0.0$ \\  grameenphone & $0.0078 \pm 0.0005$ & $0.018 \pm 0.001$ & 1.97 & $0.82 \pm 0.00$ \\  banglalink & $0.00016 \pm 0.00011$ & $0.031 \pm 0.002$ & 2.86 & $0.42 \pm 0.01$ \\  robi & $\left(8.4 \pm 8.4\right) \times 10^{-5}$ & $0.040 \pm 0.003$ & 3.17 & $0.36 \pm 0.01$ \\ \end{tabular} \caption{Security scores for operators} \label{tab:sec_operator} \end{table}}
-\newcommand{\daSecScoreBestoperator}{o2 uk}
+\newcommand{\daTabSecScoresoperator}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{Security score} \\ &&&& (out of 10) \\ \hline O2 uk & $0.25 \pm 0.00$ & $0.12 \pm 0.00$ & 0.37 & $3.8 \pm 0.0$ \\  T-Mobile & $0.19 \pm 0.00$ & $0.18 \pm 0.00$ & 0.48 & $3.6 \pm 0.0$ \\  Orange & $0.21 \pm 0.00$ & $0.087 \pm 0.003$ & 0.39 & $3.5 \pm 0.0$ \\  Sprint & $0.18 \pm 0.00$ & $0.11 \pm 0.00$ & 0.44 & $3.4 \pm 0.0$ \\  3 & $0.15 \pm 0.00$ & $0.093 \pm 0.003$ & 0.49 & $3.2 \pm 0.0$ \\  AT\&T & $0.12 \pm 0.00$ & $0.077 \pm 0.002$ & 0.43 & $3.1 \pm 0.0$ \\  vodafone uk & $0.13 \pm 0.00$ & $0.10 \pm 0.00$ & 0.52 & $3.1 \pm 0.0$ \\  Verizon & $0.18 \pm 0.00$ & $0.087 \pm 0.002$ & 0.81 & $2.8 \pm 0.0$ \\  unknown & $0.097 \pm 0.001$ & $0.22 \pm 0.00$ & 1.07 & $2.6 \pm 0.0$ \\  n Telenor & $0.026 \pm 0.001$ & $0.10 \pm 0.00$ & 1.42 & $1.6 \pm 0.0$ \\  Airtel & $0.066 \pm 0.002$ & $0.028 \pm 0.002$ & 1.95 & $1.1 \pm 0.0$ \\  Grameenphone & $0.0078 \pm 0.0005$ & $0.015 \pm 0.001$ & 2.03 & $0.77 \pm 0.00$ \\  Robi & $\left(8.4 \pm 8.4\right) \times 10^{-5}$ & $0.041 \pm 0.003$ & 2.22 & $0.71 \pm 0.01$ \\  banglalink & $0.00016 \pm 0.00011$ & $0.029 \pm 0.002$ & 2.92 & $0.39 \pm 0.00$ \\ \end{tabular} \caption{Security scores for operators} \label{tab:sec_operator} \end{table}}
+\newcommand{\daSecScoreBestoperator}{O2 uk}
 \newcommand{\daSecScoreBestoperatorScore}{$3.8 \pm 0.0$}
-\newcommand{\daSecScoreWorstoperator}{robi}
-\newcommand{\daSecScoreWorstoperatorScore}{$0.36 \pm 0.01$}
+\newcommand{\daSecScoreWorstoperator}{banglalink}
+\newcommand{\daSecScoreWorstoperatorScore}{$0.39 \pm 0.00$}
 \newcommand{\daNumSigOperators}{14}
@@
 \newcommand{\daSecScoreworstsummaryNumFullVersions}{1057}
-\newcommand{\daSecScoreBestoperatorNumFullVersions}{88}
+\newcommand{\daSecScoreBestoperatorNumFullVersions}{87}
 \newcommand{\daSecScoreworstoperatorNumFullVersions}{89}
@@
 \newcommand{\daSecScoreWorstsummaryNumFullVersions}{1103}
-\newcommand{\daSecScoreWorstoperatorNumFullVersions}{32}
+\newcommand{\daSecScoreWorstoperatorNumFullVersions}{58}
 \newcommand{\daFullDeployedAt}{95\%}
-\newcommand{\daOSCurveFitParamFirst}{87.38}
+\newcommand{\daOSCurveFitParamFirst}{87.37}
 \newcommand{\daOSCurveFitParamSecond}{\num{0.00315}}
 \newcommand{\daOSCurveFitRMSE}{0.11}
-\newcommand{\daAPICurveFitParamFirst}{109.79}
+\newcommand{\daAPICurveFitParamFirst}{109.77}
 \newcommand{\daAPICurveFitParamSecond}{\num{0.00315}}
@@
 \newcommand{\daNumUpdatesSkippedBig}{3}
-\newcommand{\daPercBigUpgrades}{22.43\%}
+\newcommand{\daPercBigUpgrades}{22.41\%}
 \newcommand{\daUpdatesPerYear}{$1.4 \pm 0.0$}
 \newcommand{\daOSMonthsOfData}{40}
-\newcommand{\daMeanInsecurityPercNominal}{87.21\%}
+\newcommand{\daMeanInsecurityPercNominal}{87.24\%}
 \newcommand{\daVulnFreeNominal}{0.13}
@@
 \newcommand{\daSecurityScoreNominal}{2.75}
-\newcommand{\daUpdatesPerYearNominal}{1.36}
+\newcommand{\daUpdatesPerYearNominal}{1.35}
 \newcommand{\daOSCurveHalfDeployedNominal}{307.28}
-\newcommand{\daOSCurveFullDeployedNominal}{1037.77}
-\newcommand{\daAPICurveHalfDeployedNominal}{329.78}
-\newcommand{\daAPICurveFullDeployedNominal}{1060.54}
+\newcommand{\daOSCurveFullDeployedNominal}{1037.82}
+\newcommand{\daAPICurveHalfDeployedNominal}{329.75}
+\newcommand{\daAPICurveFullDeployedNominal}{1060.53}
+\newcommand{\daSecScoreBestoperatorScoreNominal}{3.79}
+\newcommand{\daSecScoreWorstoperatorScoreNominal}{0.39}
+\newcommand{\daSecScoreBestmodelScoreNominal}{4.76}
+\newcommand{\daSecScoreWorstmodelScoreNominal}{0.27}
+\newcommand{\daAdbEnabledPercNominal}{19.57\%}

### 2014-11-11 security_scoring.tex
Updates following comments from Richard Clayton

@@
 We propose a method to score a manufacturer, operator or device based on its historic performance at keeping devices up to date and fixing security vulnerabilities.
-We find that Android as a whole gets a score of \daSecurityScore out of 10, the highest scoring manufacturer is \daSecScoreBestmanufacturer\ (\daSecScoreBestmanufacturerScore\ out of 10) and the lowest scoring is \daSecScoreWorstmanufacturer\ (\daSecScoreWorstmanufacturerScore\ out of 10).
+We find that Android as a whole gets a score of \daSecurityScore\ out of 10, the highest scoring manufacturer is \emph{\daSecScoreBestmanufacturer} (\daSecScoreBestmanufacturerScore\ out of 10) and the lowest scoring is \emph{\daSecScoreWorstmanufacturer} (\daSecScoreWorstmanufacturerScore\ out of 10).
 
@@
 Computing how good a particular manufacturer or device model is from a security standpoint is difficult as it depends on a number of factors which are hard to observe, particularly on a large scale.
-Ideally we would consider both prevalence of potential problems which were not exploited and actual security failures.
+Ideally we would consider both the prevalence of potential problems which were not exploited and actual security failures.
 %\footnote{A perfectly secure operating system would among other things detect and prevent all phishing attacks, provide perfect principle of least privilege isolation, not have any vulnerabilities, instantly fix all discovered vulnerabilities, not allow any user data to be used in a way which the user does not approve of and be really easy for an ordinary person to use.}
@@
 
-$f$ is hard to game without doing a good job at security but it doesn't get any worse if there is already one known vulnerability and another is found.
-A high value of $u$ could be achieved by only ever shipping one version but that would give low values for $f$ and $m$ (and not be attractive to new customers).
-A high value of $m$ could be achieved by focusing on only one device at a time and ensuring that gets updates but ignoring all others, but that would lower $f$ and $u$.
-\footnote{One way to influence our scores would be to add additional devices to Device Analyzer which have good security, these would have to be real end user devices as we could detect fake ones, this would increase the size of our data set and would require providing genuinely good security to some users.}
-
 We want to provide a score out of 10 as this is easy for phone buyers to understand as many ratings are given as a score out of 10.
@@
 
+If the comparative data given here is used to influence purchasing decisions then entities in the Android ecosystem might try to game the score rather than genuinely improve security.
+$f$ is hard to game without doing a good job at security but it doesn't get any worse if there is already one known vulnerability and another is found.
+A high value of $u$ could be achieved by only ever shipping one version but that would give low values for $f$ and $m$ (and not be attractive to new customers).
+A high value of $m$ could be achieved by focusing on only one device at a time and ensuring that gets updates but ignoring all others, but that would lower $f$ and $u$.
+\footnote{One way to influence our scores would be to add additional devices to Device Analyzer which have good security, these would have to be real end user devices as we could detect fake ones, this would increase the size of our data set and would require providing genuinely good security to some users.}
+
 \subsubsection{Results: Security scores}
@@
 \daTabSecScoresmanufacturer
-Different manufacturers have very different scores, Table~\ref{tab:sec_manufacturer} shows the scores for the \daNumSigManufacturers\ manufacturers with a significant presence in our data with \daSecScoreBestmanufacturer\ (\daSecScoreBestmanufacturerScore\ out of 10) scoring highest and \daSecScoreWorstmanufacturer\ (\daSecScoreWorstmanufacturerScore\ out of 10) scoring lowest.
+Different manufacturers have very different scores, Table~\ref{tab:sec_manufacturer} shows the scores for the \daNumSigManufacturers\ manufacturers with a significant presence in our data with \emph{\daSecScoreBestmanufacturer} (\daSecScoreBestmanufacturerScore\ out of 10) scoring highest and \emph{\daSecScoreWorstmanufacturer} (\daSecScoreWorstmanufacturerScore\ out of 10) scoring lowest.
 Manufacturers are considered significant if we have data from at least \daSigNumDevices\ devices and at least \daSigNumDeviceDays\ days of contributions.
@@
 Even within manufacturers different models can have very different update behaviours and hence security.
-Table~\ref{tab:sec_model} shows the results for the \daNumSigModels\ device models which have a significant presence by the same metric with `\daSecScoreBestmodel' (\daSecScoreBestmodelScore\ out of 10) scoring highest and `\daSecScoreWorstmodel' (\daSecScoreWorstmodelScore\ out of 10) scoring lowest.
+Table~\ref{tab:sec_model} shows the results for the \daNumSigModels\ device models which have a significant presence by the same metric with \emph{\daSecScoreBestmodel} (\daSecScoreBestmodelScore\ out of 10) scoring highest and \emph{\daSecScoreWorstmodel} (\daSecScoreWorstmodelScore\ out of 10) scoring lowest.
 We can then test whether this seems fair by comparing the version data for the highest and lowest scoring models.
-Figure~\ref{fig:worst_model_full} shows the full version distribution for `\daSecScoreWorstmodel' and Figure~\ref{fig:best_model_full} shows the same for `\daSecScoreBestmodel' there is a marked difference in the distribution of updates between the two with only \daSecScoreWorstmodelNumFullVersions\ full versions for `\daSecScoreWorstmodel' and \daSecScoreBestmodelNumFullVersions\ for `\daSecScoreBestmodel'.
+Figure~\ref{fig:worst_model_full} shows the full version distribution for \emph{\daSecScoreWorstmodel} and Figure~\ref{fig:best_model_full} shows the same for \emph{\daSecScoreBestmodel} there is a marked difference in the distribution of updates between the two with only \daSecScoreWorstmodelNumFullVersions\ full versions for \emph{\daSecScoreWorstmodel} and \daSecScoreBestmodelNumFullVersions\ for \emph{\daSecScoreBestmodel}.
 
@@
 \daTabSecScoresoperator
-We also analysed different operators, for the \daNumSigOperators\ operators with a significant presence in our data Table~\ref{tab:sec_operator} shows the results `\daSecScoreBestoperator' (\daSecScoreBestoperatorScore\ out of 10) scoring highest and `\daSecScoreWorstoperator' (\daSecScoreWorstoperatorScore\ out of 10) scoring lowest.
-However the score of an operator is affected by the manufacturers of the devices which are in use on it's network.
-This is in turn affected by both what devices an operator offers to users and on which devices users choose.
+We also analysed different operators, for the \daNumSigOperators\ operators with a significant presence in our data Table~\ref{tab:sec_operator} shows the results \emph{\daSecScoreBestoperator} (\daSecScoreBestoperatorScore\ out of 10) scoring highest and \emph{\daSecScoreWorstoperator} (\daSecScoreWorstoperatorScore\ out of 10) scoring lowest.
+However the score of an operator is affected by the manufacturers of the devices which are in use on its network.
+This is in turn affected by both what devices an operator offers to users and upon which devices users choose.
 Hence having a worse score does not necessarily mean that an operator is worse, it could be that its users all pick phones from a worse manufacturer, for example because they were cheaper.
-However an operator could use data from this paper to exclude insecure devices from those offered to consumers.
+An operator could use data from this paper to exclude insecure devices from those offered to consumers.
+An added value analysis of operators which takes into account the device mix used by users of that operator would make it possible to determine whether an operator is making the situation better or worse by the way it ships updates to users.
+However we suspect that our sample size is too small to do that as while we have significant numbers of devices for each of the device models in Table~\ref{tab:sec_model} and for each of the network operators in Table~\ref{tab:sec_operator} we would need a significant number of each model in each network operator.
+We do not attempt to disambiguate users behaviour in whether they install updates from network operators using rolling upgrades.

### 2014-11-11 securityupdates.tex
Updates following comments from Richard Clayton

@@
 
-% Android, Unix
+% Android, Unix, adb
 % update, release, patch: what do we actually mean and what relevance does that have to security?
+% Device manufacturers
+% Network operators
 
@@
 We analysed \da\ data from over \daOSYearsOfData\ years and \daNumOSDevices\ devices and found that on average \daMeanInsecurityPercNominal\ of devices were exposed to known root privilege vulnerabilities and only \daUpdatednessPercNominal\ of devices run the most recent version of Android.
-We find there is significant variability in the timely delivery of security updates across different manufacturers and mobile network operators, and we define a security metric to rank the performance of manufacturers and operators in our data set.
+We find there is significant variability in the timely delivery of security updates across different device manufacturers and mobile network operators, and we define a security metric to rank the performance of device manufacturers and network operators in our data set.
 \end{abstract}
@@
 Such apps are written by many developers from a wide range of backgrounds, which means neither operating system vendors nor users can fully trust app developers.
-Therefore, in order to secure personal data and prevent the use of expensive operations, such as the sending of premium-rate text messages, operating system vendors provide a protected execution environment, or \emph{sandbox}, for apps.
+Therefore, in order to secure personal data and prevent theft, such as the sending of premium-rate text messages, operating system vendors provide a protected execution environment, or \emph{sandbox}, for apps.
 
@@
 Indeed, the above figures suggest the majority of malicious apps do not attempt to do so.
-We focus on this issue because malware which is able to take full control of a handset can do significantly more harm and is much harder to remove.
+We focus on the issue of apps breaking out of the sandbox because malware which is able to take full control of a handset can do significantly more harm and is much harder to remove.
 Ordinary malware can be uninstalled by the user, or remotely by Google through the Play Store, and returns the device to a secure state. 
@@
 The development of Android is distributed among many parties and the security of Android relies on many open source projects, include the Linux kernel, OpenSSL and BouncyCastle as well as Google who build the core platform.
-In addition, manufacturers (of which we are aware of \daNumManufacturers\footnote{\label{foot:dadata}We computed this from the \da~\cite{Wagner2013} data, see \S\ref{sec:android_update_process}.}) and network operators (\daNumOperators\dafoot) may make further modifications before devices are shipped to customers.
-Understanding this ecosystem is important as manufacturers have introduced additional vulnerabilities in the past~\cite{Grace2012}. 
+In addition, device manufacturers (of which we are aware of \daNumManufacturers\footnote{\label{foot:dadata}We computed this from the \da~\cite{Wagner2013} data, see \S\ref{sec:android_update_process}.}) and network operators (\daNumOperators\dafoot) may make further modifications before devices are shipped to customers.
+Understanding this ecosystem is important as device manufacturers have introduced additional vulnerabilities in the past~\cite{Grace2012}. 
 We present a better understanding of both the ecosystem of Android development and associated vulnerabilities more fully in \S\ref{sec:android_ecosystem}.
@@
 The Figure shows the proportion of devices split into three categories: 
-(1) \emph{secure} because they are running a version of Android with no known vulnerabilities; 
-(2) \emph{insecure} because the device was running an insecure version of Android with a build released before the vulnerability was fixed or discovered; and 
-(3) \emph{maybe secure} because the device is running an insecure version of Android but it received an update which we did not see until after the vulnerability was fixed and therefore this update may contain a backport of the fix.
-Further analysis into the propagation of updates in Android are presented in \S\ref{sec:exp:android_ecosystem}.
+(1)~\emph{secure} because they are running a version of Android with no known vulnerabilities; 
+(2)~\emph{insecure} because the device was running an insecure version of Android with a build released before the vulnerability was fixed or discovered; and 
+(3)~\emph{maybe insecure} because the device is running a version of Android known to be insecure but has received an update which might contain a backported fix as we did not observe that build number until after the vulnerability was known.
+Further analysis into the propagation of updates in Android is presented in \S\ref{sec:exp:android_ecosystem}.
 
-It is clear from Figure~\ref{fig:proportioninsecure} that the security of the Android platform is far from perfect.
-One solution is regulation, and indeed there is ongoing legal action to force operators to ship updates for security vulnerabilities~\cite{Soghoian2013}.\todolater{Check on the status of this legal action}
+It is clear from Figure~\ref{fig:proportioninsecure} that Android devices are often exposed to known security vulnerabilities.
+One solution is regulation, and indeed there is ongoing legal action to force network operators to ship updates for security vulnerabilities~\cite{Soghoian2013}.\todolater{Check on the status of this legal action}
 Indeed, many smartphones are sold on 12--24 month contracts, and yet our data shows many devices do not receive many security updates. 
@@
 An alternative to regulation is access to publicly available data. 
-Hard numbers provide an incentive to both manufacturers and operators to provide updates.
+Comparative data would provide an incentive to both device manufacturers and network operators to provide updates.
 Currently corporate and public sector buyers are encouraged to purchase secure devices, but we have found little concrete guidance on the specific makes and models which provide the best security. 
-For example, CESG, which advises the UK government on how to secure its computer systems, recommends picking Android devices from manufactures which are good at shipping security updates promptly~\cite{CESG2013} but it does not state which manufacturers are better.
+For example, CESG, which advises the UK government on how to secure its computer systems, recommends picking Android devices from device manufactures which are good at shipping security updates promptly~\cite{CESG2013} but it does not state which device manufacturers these are.
 To address this, we developed a scoring system and provide numbers on the historic performance of devices in the \da\ project in~\S\ref{sec:security_scoring}.
@@
 %REMOVED BY ARB
-%The manufacturer has little financial incentive to perform this work as they have already sold the device.
-%Operators are cautious because if they ship a broken update then they will get many support requests which is expensive.
+%The device manufacturer has little financial incentive to perform this work as they have already sold the device.
+%Network operators are cautious because if they ship a broken update then they will get many support requests which is expensive.
 %
@@
  \item We created an open database to collate quantitative information on vulnerabilities which affect Android, and investigate their time-lines.
- \item We measure the security of Android according to three metrics and compare different manufacturers and device models to allow device purchasers to differentiate between them based on security.
+ \item We measure the security of Android according to three metrics and compare different device manufacturers and device models to allow device purchasers to differentiate between them based on security.
 \end{itemize}
@@
 Among these are the build string and API version for the version of Android currently running on the device each day.
-The API version is a well defined integer, however it does not always change with new Android releases, particularly security bug fixes should not increase this value.
-The build string is `The user-visible version string', fortunately most (\daOSVersionPercValidLines) entries in the data have a build string of the form `x.y.z opaque marker' and so it is possible to extract the Android version number.
-On a large proportion of devices `opaque marker' is a well defined build number\footnote{\url{https://source.android.com/source/build-numbers.html}} but different manufacturers use different schema.
+The API version is a well defined integer which increases when new features are added to the API.
+It does not change for all Android releases as (security) bug fixes which do not change the API do not require it to increase.
+The build string is `The user-visible version string', fortunately most (\daOSVersionPercValidLines) entries in the data have a build string of the form `x.y.z opaque marker' and so it is possible to extract the Android version number `x.y.z'.
+On a large proportion of devices `opaque marker' is a well defined build number\footnote{\url{https://source.android.com/source/build-numbers.html}} but different device manufacturers use different schema.
 
-We verify the validity of the \da\ data in \S\ref{sec:representative}.
+Where possible we indicate the uncertainty in our results by giving them $\pm$ the standard deviation, this occasionally results in ${} \pm 0$ when the uncertainty is less than the number of significant figures we are reporting.
+However this technique is unable to account for systematic errors.
+
+We comment on the validity of the \da\ data in \S\ref{sec:representative}.
 
@@
 We only looked for critical vulnerabilities such as root vulnerabilities which did not require USB debugging to exploit.
-Critical vulnerabilities allow a program to escalate to privileges equivalent in scope to root.
+Critical vulnerabilities allow a program to gain privileges equivalent in scope to root.
 If an application exploits a critical vulnerability then it gains control of the device.
-Some phones can be `rooted' by enabling USB debugging and using the special privileges of the \texttt{adb} shell to root the device but only \daAdbEnabledPerc\dafoot\textsuperscript{,}\footnote{Where possible we indicate the uncertainty in our results by giving them $\pm$ the standard deviation.
-This is explained further in \S\ref{sec:uncertainty} but occasionally results in $\pm0$.} of devices have USB debugging enabled.
+Some phones can be `rooted' by enabling USB debugging and using the special privileges of the \texttt{adb} shell to root the device but only \daAdbEnabledPerc\dafoot\ of devices have USB debugging enabled.
 This is not something that applications running on the phone can exploit to break compartmentalisation and so we do not include those vulnerabilities.
-Unfortunately many published exploits use \texttt{adb} for convenience and so determining whether that is necessary to exploit the vulnerability can be difficult.
+Unfortunately, many published exploits use \texttt{adb} for convenience and so determining whether that is necessary to exploit the vulnerability can be difficult.
 
 Some critical vulnerabilities are not traditional kernel vulnerabilities, for example the discovery of flaws in the verification of signatures on Android applications in February 2013~\cite{Forristal2013} meant that applications could pretend to be signed with system keys and hence gain root equivalent privileges.
-On some versions of Android (below 4.1) they could use known system to root escalation mechanisms but on all versions they have a greatly increased attack area for further privilege escalation and also have the ability to control all user internet traffic (via VPNs), brick the phone, remove and install apps, steal user credentials, read the screen and make as well as receive calls.
+On some versions of Android (below version 4.1) malware could use known system-to-root escalation mechanisms but on all versions they have a greatly increased attack area for further privilege escalation and also have the ability to control all user internet traffic (via VPNs), brick the phone, remove and install apps, steal user credentials, read the screen and make as well as receive calls.
 \todolater{break vulnerabilities down by attack vector?}
@@
 We have received submissions or amendments from \avoNumSubmitters\ individuals.
-\avo\ contains \avoNumVulnerabilities\ vulnerabilities of which \avoNumVulnAllAndroid\ affect all Android devices and \avoNumVulnSpecific\ are specific to particular devices or manufacturers.
-We only use the \daNumVulnsUsed\ vulnerabilities which we have sufficient data to use for this analysis because they affect all Android devices rather than particular manufacturers and we know enough information about the vulnerability to tie it to the particular versions it affects.
+We collected this data between \avoStartDate\ and \avoEndDate\ and so any information lost before the start of that period cannot be included.
+\avo\ currently contains \avoNumVulnerabilities\ vulnerabilities of which \avoNumVulnAllAndroid\ affect all Android devices and \avoNumVulnSpecific\ are specific to particular devices or device manufacturers.
+We only use the \daNumVulnsUsed\ vulnerabilities which we have sufficient data to use for this analysis because they affect all Android devices rather than particular device manufacturers and we know enough information about the vulnerability to tie it to the particular versions it affects.
 We summarise the \daNumVulnsUsed\ vulnerabilities used for this analysis in Table~\ref{tab:andvulns}.\todolater{We have published full details of the \daNumVulnsUsed\ vulnerabilities used for this analysis in an accompanying technical report~\cite{TODO}}
@@
 
-The key events in the lifetime of a vulnerability do not always occur in the same order and are:
-
-\begin{LaTeXdescription}
- \item[created] When a vulnerability was created in the source code.
- \item[introducing release] When the first release was made containing the vulnerability.
- \item[discovery] When the vulnerability is first discovered
- \item[exploit] When the vulnerability is first exploited
- \item[disclosure] When the vulnerability is first disclosed % publically or to a smaller set of people?
- \item[fix] When a vulnerability was first fixed in the source code.
- \item[fixing release] When the first release containing the fix was made (equivalent to `patch available'~\cite{Frei2010}).
- \item[fix deployed] When the fix has been deployed to a sufficiently large proportion of the population that the vulnerability can be ignored (the ecosystem equivalent to the per instance `patch installed'~\cite{Frei2010}).
-\end{LaTeXdescription}
+The key events in the lifetime of a vulnerability do not always occur in the same order and are:\\
+\textbf{creation} When a vulnerability was created in the source code.\\
+\textbf{introducing release} When the first release was made containing the vulnerability.\\
+\textbf{discovery} When the vulnerability is first discovered\\
+\textbf{exploit} When the vulnerability is first exploited\\
+\textbf{disclosure} When the vulnerability is first disclosed.\\ % publically or to a smaller set of people?
+\textbf{fix} When a vulnerability was first fixed in the source code.\\
+\textbf{fixing release} When the first release containing the fix was made (equivalent to `patch available'~\cite{Frei2010}).\\
+\textbf{fix deployed} When the fix has been deployed to a sufficiently large proportion of the population that the vulnerability can be ignored (the ecosystem equivalent to the per instance `patch installed'~\cite{Frei2010}).
+\todolater{Does this look nice?}
 
@@
 Hence when considering vulnerability from the point of view of an organisation which cares about Advanced Persistent Threats, such as those the CESG advice~\cite{CESG2013} is aimed at, the critical period is from when people could have started using the vulnerability to when the update hits the user device.
-So for our vulnerability calculations we use the earliest date we can find recorded evidence that the vulnerability was known about, even if that knowledge might have been confined to a particular manufacturer or hobbyist.
-%We cannot know if someone reported the vulnerability to the manufacturer and we do know that various agencies are engaged in widespread monitoring of communications and compromise of civilian infrastructure~\cite{TODO} and so they may know.
+So for our vulnerability calculations we use the earliest date we can find recorded evidence that the vulnerability was known about, even if that knowledge might have been confined to a particular device manufacturer or hobbyist.
+%We cannot know if someone reported the vulnerability to the device manufacturer and we do know that various agencies are engaged in widespread monitoring of communications and compromise of civilian infrastructure~\cite{TODO} and so they may know.
 
 There are several plausible points from which a vulnerability poses a threat to users.
-{\bf Creation} and {\bf introduction} are the earliest points which could be used but the risk is mostly latent until someone discovers them.
-Unfortunately {\bf discovery} is the hardest point to obtain concrete data on as discovery may happen multiple times independently and not all discoverers will report their discovery, for example Advanced Persistent Threats (APTs) may keep databases of unknown vulnerabilities to use when they have a particular target, in which case there is a high risk to potential targets from the never disclosed point of discovery.
-However if the discoverer never discloses the vulnerability to anyone or makes any use of it then there is little risk.
+{\bf Creation} and {\bf introducing release} are the earliest points which could be used but the risk is mostly latent until someone discovers them.
+Unfortunately {\bf discovery} is the hardest point to obtain concrete data on as discovery may happen multiple times independently and not all discoverers will report their discovery, for example entities with high value targets may keep databases of unknown vulnerabilities for later use, in which case there is a high risk to potential targets from the never disclosed point of discovery.
+However if the discoverer never discloses the vulnerability to anyone or makes any use of it then there is little danger.
 The date of first {\bf exploit} is a point when the risk is definitely high, but again a good adversary will not be detected when using such exploits.
 The date of first {\bf fix} is, assuming that the fix is deliberate, a point at which the vulnerability is known at least within the organisation performing the fix and frequently implies an earlier discovery and notification by a third party.
-It can, for example, be determined from the authored on date on the fixing commit in the git repository.
+It can, for example, be determined from the authored-on date of the fixing commit in the git repository.
 Once a {\bf fixing release} has been made then the vulnerability is widely known because it can be reverse engineered from the changes in the release~\cite{Brumley2008}.
-When the {\bf fix is deployed} to a sufficiently large proportion of devices then the remaining risk is minimal.
+When the {\bf fix deployed} to a sufficiently large proportion of devices then the remaining risk is minimal.
 
@@
 
-This data shows a large gap in the middle where we do not know of any discoveries of critical vulnerabilities affecting all of Android.
+This data shows a large gap from 2011-10-06 to 2013-02-18 where we do not know of any discoveries of critical vulnerabilities affecting Android devices.
 The cause of this quiet period is unclear.
 Possible explanations are: That since most devices were exposed to known vulnerabilities there was no point in looking for new ones (from Figure~\ref{fig:proportioninsecure}).
-That manufacturers made it easier to install custom versions of Android, reducing the need for users to root their devices?
-That manufacturer specific vulnerabilities (which we are ignoring) proved to be easier to find and so people looking for vulnerabilities adjusted their focus.
+That device manufacturers made it easier to install custom versions of Android, reducing the need for users to root their devices?
+That device manufacturer specific vulnerabilities (which we are ignoring) proved to be easier to find and so people looking for vulnerabilities adjusted their focus.
 
@@
 There is a complex Android ecosystem which creates and distributes updates to Android which fix vulnerabilities.
-In this section we rectify the lack of public information about the size of the Android ecosystem.
+In this section we rectify the lack of public information about the nature of the Android ecosystem.
 
@@
 There are five entities or groups which contribute towards Android updates: the network operators, the device manufacturers, the hardware developers, Google and the upstream open source projects.
-Android builds on various open source projects such as the Linux kernel, the OpenSSL and BouncyCastle cryptography libraries and so can include any compatible versions of those projects, including those which fix security vulnerabilities in them.
+Android builds on various open source projects such as the Linux kernel, the OpenSSL and BouncyCastle cryptography libraries and so can include any compatible versions of those projects, including those which fix security vulnerabilities.
 Android also incorporates various drivers for different bits of hardware.
@@
 Is it worth it?} until after the update has been published.
-Manufacturers may receive advanced copies in order to prepare handsets.
-Then the Manufacturers take the update and customise it before passing it on to the operator.
-The Operator may then make further customisations and do further testing before shipping the update to the Device.
-Sometimes Manufactures ship direct to the user, sometimes the Manufacturer and Google collaborate closely for the purpose of a particular phone, such as with Nexus phones, and hence are harder to separate.
-Sometimes Manufacturers incorporate upstream open source project releases directly, and sometimes incorrectly -- for example including a broken daily build of sqlite in their release of Android~\cite{Wagner2013}.
+Device manufacturers may receive advance copies in order to prepare handsets.
+Then the device manufacturers take the update and customise it before passing it on to the network operator.
+The network operator may then make further customisations and do further testing before shipping the update to the Device.
+Sometimes device manufactures ship direct to the user, sometimes the device manufacturer and Google collaborate closely for the purpose of a particular phone, such as with Nexus phones, and hence are harder to separate.
+Sometimes device manufacturers incorporate upstream open source project releases directly, and sometimes incorrectly -- for example including a broken daily build of sqlite in their release of Android~\cite{Wagner2013}.
 \todolater{use statistics from samsung-updates.com -> how many binaries are there per device?}
 
-The numbers of Devices (\daNumOSDevices), Operators (\daNumOperators) and Manufacturers (\daNumManufacturers) in Figure~\ref{fig:update_ecosystem} come from the \da\ data.
-Manufacturer and Operator counts were obtained by normalising the results reported by Android to \da\ of the device manufacturer and active network operator.
+The numbers of devices (\daNumOSDevices), network operators (\daNumOperators) and device manufacturers (\daNumManufacturers) in Figure~\ref{fig:update_ecosystem} come from the \da\ data.
+Device manufacturer and network operator counts were obtained by normalising the results reported by Android to \da\ of the device manufacturer and active network operator.
 This normalisation is a manual task involves removing invalid values (such as `manufacturer' or `airplane mode is on'), collating across company name changes (e.g.\ `lge' to `LG'), normalising punctuation, removing extra strings sometimes added such as (`(2g)' or `communications') and mapping some incorrectly placed model names back to their manufacturer.
-This normalisation is not perfect and so these are overestimates on the \da\ data but they are likely still underestimates as there will be some manufacturers and operators which are not included in the \da\ data.
+This normalisation is not perfect and so these are overestimates on the \da\ data but they are likely still underestimates as there will be some device manufacturers and network operators which are not included in the \da\ data.
 
@@
 \centering
+\normalsize
 \begin{tabular}{l|r|r}
@@
 On average only \daUpdatesPerYear\ updates reach a device each year and \daPercUpdatesDowngrades\ of version changes are downgrades to older versions.
-Nexus devices are better than others, \emph{\daSecScoreBestmanufacturer} to be the highest scoring manufacturer and \emph{\daSecScoreBestoperator} to be the highest scoring network operator.
+Nexus devices are better than others, \emph{\daSecScoreBestmanufacturer} is the highest scoring device manufacturer and \emph{\daSecScoreBestoperator} is the highest scoring network operator.
 %\todo{comparison of nexus update rates with iOS}
-We conducted three experiments examining the behaviour of the Android ecosystem as a whole (\S\ref{sec:exp:android_ecosystem}), updates reaching particular devices (\S\ref{sec:exp:device_updates}) and comparisons between different manufacturers and operators (\S\ref{sec:exp:security_score}).
+We conducted three experiments examining the behaviour of the Android ecosystem as a whole (\S\ref{sec:exp:android_ecosystem}), updates reaching particular devices (\S\ref{sec:exp:device_updates}) and comparisons between different device manufacturers and network operators (\S\ref{sec:exp:security_score}).
 
@@
 The \da\ data used here was collected between \daStartDate\ and \daEndDate.
-The \avo\ data was collected between \avoStartDate\ and \avoEndDate\ and covers the period from \avoFirstDataDate\ to \avoLastDataDate.
+The \avo\ data covers the period from \avoFirstDataDate\ to \avoLastDataDate.
 For each device, for every day with version data we record what version it was running and which vulnerabilities it was exposed to at that time (if any).
@@
 The proportion of devices in the \da\ data running different versions of Android each day is shown in Figure~\ref{fig:norm_os}.
-It shows how old versions are gradually replaced by new ones, the long tail of devices which do not see updates to more recent versions.
+It shows how old versions are gradually replaced by new ones, and the long tail of devices which do not see updates to more recent versions.
 
@@
 It shows how some devices had vulnerabilities, which were fixed, and then further vulnerabilities were discovered, and for these devices, mostly not fixed.
-This implies that these devices had been abandoned by the manufacturer and were not receiving updates which is confirmed by Figure~\ref{fig:device_data_os} which shows which OS versions those devices were running.
+This implies that these devices had been abandoned by the device manufacturer and were not receiving updates which is confirmed by Figure~\ref{fig:device_data_os} which shows which OS versions those devices were running.
 Some devices start off in 2011 exposed to known security vulnerabilities and are still exposed to those and additional ones in 2014.
@@
 By taking the number of updates observed and the number of device days of data we have collected we can compute the number of updates received by a device per year to be \daUpdatesPerYear.
-This compares badly with the number of critical vulnerabilities discovered per year of between \avoVulnsPerYearAllAndroid\ (affecting all Android) and \avoVulnsPerYear\ (including the manufacturer specific ones in \avo).
+This compares badly with the number of critical vulnerabilities discovered per year of between \avoVulnsPerYearAllAndroid\ (affecting all Android) and \avoVulnsPerYear\ (including the device manufacturer specific ones in \avo).
 
@@
 \subsection{Changes in the \da\ sample}
-The \da\ data is mostly generated by devices which have the \da\ app installed because their owner happened to come across the app on Google Play as there has been no advertising and little other attempts to increase usage.
+The \da\ data is mostly generated by devices which have the \da\ app installed because their owner happened to come across the app on Google Play as there has been no advertising and few other attempts to increase usage.
 However two collaborations with network operators to install \da\ on customer devices resulted in large numbers of new \da\ users coming from particular areas and network operators.
@@
 %We could normalise out the influence of the Bangladeshi users, however they are Android users too and they also need security updates.
-
-\subsection{Uncertainty propagation}
-\label{sec:uncertainty}
-We have endeavoured to propagate information on uncertainties in our results through all calculations with the standard deviation indicated by $\pm$.
-However this technique is unable to account for systematic errors.
+Since most \da\ users are self selecting and install \da\ because they want to find out more about what their phone is doing or to aid research they may be biased and perhaps more likely to install updates.
 
@@
 Malicious apps which use exploits to take control of an Android device must be installed on the device and to an extent this can be prevented if all apps are downloaded from the Google Play store.
-It is protected by Bouncer's static analysis and prompt takedowns of apps reported to be malicious and has been found to have infection rates of 0.02\%~\cite{Zhou2012a}.
+This marketplace is protected by Bouncer's static analysis and prompt takedowns of apps reported to be malicious and has been found to have infection rates of 0.02\%~\cite{Zhou2012a}.
 However some apps are installed from other markets which are less well policed (infection rates of between 0.20\%~ and 0.47\%~\cite{Zhou2012a}).
@@
 
-There are also other ways that malicious code could end up running on Android devices, the addJavascriptInterface vulnerability (CVE-2012-6636 ) allows javascript running in a webview which has an expose interface to execute arbitrary Java code and hence load exploits.
+There are also other ways that malicious code could end up running on Android devices, the addJavascriptInterface vulnerability (CVE-2012-6636 ) allows javascript running in a webview which has an exposed interface to execute arbitrary Java code and hence load exploits.
 While fixes have been produced for this the change is API breaking and so to be safe an app must target an API version greater than 17 and be running on a device with an Android version of at least 4.2.
@@
 
-We have not included all the vulnerabilities from \avo\, in particular we have not included any manufacturer specific vulnerabilities, even when these are widespread (such as vulnerabilities affecting Qualcomm\footnote{Qualcomm is particularly good at publicly disclosing the vulnerabilities affecting their code and the patches which fix them.\url{https://www.codeaurora.org/projects/security-advisories/}} chipsets) because it is difficult to work out which devices are affected.
-There also tends to be less public information available about manufacturer specific vulnerabilities which makes them harder to tie down.
+We have not included all the vulnerabilities from \avo\, in particular we have not included any device manufacturer specific vulnerabilities, even when these are widespread (such as vulnerabilities affecting Qualcomm\footnote{Qualcomm is particularly good at publicly disclosing the vulnerabilities affecting their code and the patches which fix them.\url{https://www.codeaurora.org/projects/security-advisories/}} chipsets) because it is difficult to work out which devices are affected.
+There also tends to be less public information available about device manufacturer specific vulnerabilities which makes them harder to tie down.
 For other vulnerabilities which might affect all Android devices it is hard to work out which devices they affect.
 For example \emph{pty race}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/pty_race}} is a Linux kernel vulnerability, we found 48 commits which fixed it in different branches, it was also accidentally fixed and later reintroduced and so determining which devices were vulnerable relies on knowing where on which branch the kernel was taken from.
-For some vulnerabilities the corresponding patch is not in AOSP despite some manufacturers having shipped builds containing the fix, which makes working out what is going on more difficult.
+For some vulnerabilities the corresponding patch is not in AOSP despite some device manufacturers having shipped builds containing the fix, which makes working out what is going on more difficult.
 For \emph{RageAgainstTheCage adb}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/RageAgainstTheCage_adb}} and \emph{keystore buffer}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/keystore_buffer}} vulnerabilities which we do have sufficient data it is not clear whether they are truly critical vulnerabilities as the former may require physical \texttt{adb} access and the latter may be protected by a sandbox.
@@
 However DroidChameleon found that antivirus products could not detect malware if it was simply automatically permuted~\cite{Rastogi2013}.
-All these app analysis projects have been hampered by the difficulty of obtaining full datasets of Android apps as Google does not make these available (and researchers who automatically download them, violate Google's Terms of Service).
+All these app analysis projects have been hampered by the difficulty of obtaining full datasets of Android apps as Google does not make these available (and researchers who automatically download them violate Google's Terms of Service).
 PlayDrone was a particularly effective project which circumvented Google's protective measures and downloaded over 1,100,000 apps from Google Play, allowing an in depth analysis~\cite{Viennot2014}.
 
-Examining security updates has been a key way of assessing security for many years.
+Examining security update mechanisms has been used to assess security for many years.
 Many software update systems have been found not to authenticate the connection to the update server and/or do not authenticate the downloaded binaries~\cite{Bellissimo2006}.
@@
 The same analysis was used to show that Chrome's use of silent updates seems to increase uptake of upgrades~\cite{Duebendorfer2010} with 97\% of users running the latest version within 3 weeks of release.
-Android's update process is manual, the user is notified but must action the upgrade which will then download the update and install it, the phone is rendered inoperable during the update process which is inconvenient for users and the phone must have sufficient charge so that it will not run out during the update process.
+Android's update process is manual, the user is notified but must action the upgrade which will then download the update and install it, the phone is rendered inoperable during the update process which is inconvenient for users and the phone must have sufficient charge (users are advised to plug it in) so that it will not run out during the update process.
 Partly this is the result of the fact that an operating system update is being installed and so a reboot is required, but Chrome installs the new version side by side with the old one and switches the next time it is restarted.
-The same technique would be more difficult on phones with limited storage space (as for many cheap Android phones which have barely enough space to install just the update) but is a plausible improvement for more high-end devices.
-Google is deploying the same silent update technique through Google Play Services\footnote{\url{http://lifehacker.com/why-google-play-services-are-now-more-important-than-an-975970197}} which automatically installs updates for core Google components of Android, this also bypasses the manufacturer and operator.
+The same technique would be more difficult on phones with limited storage space (as many cheap Android phones have barely enough space to install just the update) but is a plausible improvement for more high-end devices.
+Google is deploying the same silent update technique through Google Play Services\footnote{\href{http://lifehacker.com/why-google-play-services-are-now-more-important-than-an-975970197}{http://lifehacker.com/why-google-play-services-are-now-more-important-than-an-975970197}} which automatically installs updates for core Google components of Android, this also bypasses the device manufacturer and network operator.
 
@@
 We have investigated the security of the Android operating system with respect to security updates and examined how the Android ecosystem results in different devices receiving different levels of security due to whether or not they get security updates.
-We have compared different models, manufacturers and network operators and found that there are differences between them which the discerning purchaser might use to influence their decision about which device to buy from which manufacturer and operator.
-We hope that this analysis will encourage manufacturers and operators to improve the support they provide for devices after sale.
+We have compared different models, device manufacturers and network operators and found that there are differences between them which the discerning purchaser might use to influence their decision about which device to buy from which device manufacturer and network operator.
+We hope that this analysis will encourage device manufacturers and network operators to improve the support they provide for devices after sale.
 Android attempts to provide good security to end users but the latency in the security update process means that on average \daMeanInsecurityPerc\ of Android devices are exposed to known critical vulnerabilities.

### 2014-11-12 dastats.tex
Updated dastats data and remove fp package as now use pre-computed value

@@
 \newcommand{\daNumDataPoints}{153 Billion}
-\newcommand{\daNumStartedApps}{122032}
-\newcommand{\daNumInstalledApps}{174390}
+\newcommand{\daNumStartedApps}{122\,000}
+\newcommand{\daNumInstalledApps}{175\,000}
 \newcommand{\daNumVulnsUsed}{11}
@@
 \newcommand{\daSigNumDays}{100}
-\newcommand{\daNumManufacturers}{1}
-\newcommand{\daNumSigManufacturers}{-1}
-\newcommand{\daNumModels}{2379}
+\newcommand{\daNumManufacturers}{260}
+\newcommand{\daNumSigManufacturers}{8}
+\newcommand{\daNumModels}{2\,380}
 \newcommand{\daNumSigModels}{15}
-\newcommand{\daOSVersionPercValidLines}{99.46\%}
-\newcommand{\daOSTotalDaysData}{1134626}
+\newcommand{\daOSVersionPercValidLines}{99.5\%}
+\newcommand{\daOSTotalDaysData}{1\,130\,000}
 \newcommand{\daMeanInsecurityPerc}{$87 \pm 0\%$}
-\newcommand{\daMeanOutstandingVulnerabilities}{0.64}
+\newcommand{\daMeanOutstandingVulnerabilities}{0.637}
 \newcommand{\daUpdatedness}{$0.056 \pm 0.000$}
 \newcommand{\daSecurityScore}{$2.8 \pm 0.0$}
-\newcommand{\daNumFullVersions}{1136}
+\newcommand{\daNumFullVersions}{1\,140}
 \newcommand{\daNumOSVersions}{45}
@@
 \newcommand{\daNumAPIVersions}{15}
-\newcommand{\daNumFullVersionUpdates}{4212}
-\newcommand{\daNumOSVersionUpdates}{3563}
+\newcommand{\daNumFullVersionUpdates}{4\,210}
+\newcommand{\daNumOSVersionUpdates}{3\,560}
 \newcommand{\daNumFullOnlyVersionUpdates}{637}
-\newcommand{\daNumSecurityUpdates}{1110}
-\newcommand{\daNumPossibleSecurityUpdates}{2141}
-\newcommand{\daTabSecScoresmanufacturer}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{Security score} \\ &&&& (out of 10) \\ \hline LG & $0.13 \pm 0.00$ & $0.056 \pm 0.000$ & 0.64 & $2.8 \pm 0.0$ \\ \end{tabular} \caption{Security scores for manufacturers} \label{tab:sec_manufacturer} \end{table}}
-\newcommand{\daTabSecScoresmodel}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{Security score} \\ &&&& (out of 10) \\ \hline Galaxy Nexus & $0.54 \pm 0.00$ & $0.52 \pm 0.01$ & 1.53 & $4.8 \pm 0.0$ \\  Nexus 4 & $0.23 \pm 0.00$ & $0.76 \pm 0.01$ & 6.06 & $3.2 \pm 0.0$ \\  Nexus 7 & $0.19 \pm 0.00$ & $0.73 \pm 0.01$ & 5.92 & $3.0 \pm 0.0$ \\  Desire HD & $0.083 \pm 0.003$ & $0.025 \pm 0.007$ & 0.39 & $2.8 \pm 0.0$ \\  HTC Sensation & $0.36 \pm 0.00$ & $0.0000 \pm 0.0078$ & 1.59 & $2.4 \pm 0.0$ \\  GT-I9100 & $0.22 \pm 0.00$ & $0.012 \pm 0.002$ & 1.20 & $2.3 \pm 0.0$ \\  HTC Desire S & $0.022 \pm 0.001$ & $0.001 \pm 0.010$ & 1.00 & $1.7 \pm 0.0$ \\  GT-N7000 & $0.25 \pm 0.00$ & $0.0006 \pm 0.0035$ & 2.52 & $1.4 \pm 0.0$ \\  GT-P1000 & $0.0096 \pm 0.0009$ & $0.0000 \pm 0.0074$ & 1.73 & $0.94 \pm 0.02$ \\  GT-I9300 & $0.15 \pm 0.00$ & $0.013 \pm 0.001$ & 6.16 & $0.64 \pm 0.01$ \\  GT-I9505 & $0.0 \pm 0$ & $0.18 \pm 0.00$ & 6.83 & $0.56 \pm 0.01$ \\  GT-N7100 & $0.067 \pm 0.002$ & $0.0011 \pm 0.0058$ & 6.91 & $0.28 \pm 0.02$ \\  HTC Desire HD & $0.0 \pm 0$ & $0.0000 \pm 0.0073$ & 3.05 & $0.27 \pm 0.02$ \\ \end{tabular} \caption{Security scores for models} \label{tab:sec_model} \end{table}}
+\newcommand{\daNumSecurityUpdates}{1\,110}
+\newcommand{\daNumPossibleSecurityUpdates}{2\,140}
+\newcommand{\daTabSecScoresmanufacturer}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{Security score} \\ &&&& (out of 10) \\ \hline LG & $0.12 \pm 0.00$ & $0.32 \pm 0.00$ & 0.74 & $3.4 \pm 0.0$ \\  Motorola & $0.16 \pm 0.00$ & $0.13 \pm 0.00$ & 0.77 & $2.9 \pm 0.0$ \\  other & $0.13 \pm 0.00$ & $0.056 \pm 0.000$ & 0.64 & $2.8 \pm 0.0$ \\  Samsung & $0.14 \pm 0.00$ & $0.047 \pm 0.000$ & 0.71 & $2.7 \pm 0.0$ \\  Sony & $0.13 \pm 0.00$ & $0.18 \pm 0.00$ & 1.02 & $2.6 \pm 0.0$ \\  HTC & $0.14 \pm 0.00$ & $0.087 \pm 0.001$ & 0.89 & $2.6 \pm 0.0$ \\  asus & $0.16 \pm 0.00$ & $0.56 \pm 0.01$ & 6.06 & $2.3 \pm 0.0$ \\  Symphony & $0.0015 \pm 0.0002$ & $0.042 \pm 0.001$ & 4.87 & $0.18 \pm 0.00$ \\ \end{tabular} \caption{Security scores for manufacturers} \label{tab:sec_manufacturer} \end{table}}
+\newcommand{\daTabSecScoresmodel}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{Security score} \\ &&&& (out of 10) \\ \hline Galaxy Nexus & $0.54 \pm 0.00$ & $0.53 \pm 0.01$ & 1.53 & $4.8 \pm 0.0$ \\  Nexus 4 & $0.23 \pm 0.00$ & $0.82 \pm 0.01$ & 6.06 & $3.4 \pm 0.0$ \\  Nexus 7 & $0.19 \pm 0.00$ & $0.74 \pm 0.01$ & 5.92 & $3.0 \pm 0.0$ \\  other & $0.13 \pm 0.00$ & $0.12 \pm 0.00$ & 0.64 & $2.9 \pm 0.0$ \\  Desire HD & $0.083 \pm 0.003$ & $0.051 \pm 0.004$ & 0.39 & $2.9 \pm 0.0$ \\  HTC Sensation & $0.36 \pm 0.00$ & $0.0063 \pm 0.0062$ & 1.59 & $2.5 \pm 0.0$ \\  GT-I9100 & $0.22 \pm 0.00$ & $0.016 \pm 0.001$ & 1.20 & $2.3 \pm 0.0$ \\  HTC Desire S & $0.022 \pm 0.001$ & $0.016 \pm 0.004$ & 1.00 & $1.7 \pm 0.0$ \\  GT-N7000 & $0.25 \pm 0.00$ & $0.0006 \pm 0.0036$ & 2.52 & $1.4 \pm 0.0$ \\  GT-P1000 & $0.0096 \pm 0.0009$ & $0.0000 \pm 0.0061$ & 1.73 & $0.94 \pm 0.02$ \\  GT-I9300 & $0.15 \pm 0.00$ & $0.014 \pm 0.001$ & 6.16 & $0.64 \pm 0.01$ \\  GT-I9505 & $0.0 \pm 0$ & $0.18 \pm 0.00$ & 6.83 & $0.56 \pm 0.01$ \\  GT-N7100 & $0.067 \pm 0.002$ & $0.0011 \pm 0.0058$ & 6.91 & $0.28 \pm 0.02$ \\  HTC Desire HD & $0.0 \pm 0$ & $0.0005 \pm 0.0070$ & 3.05 & $0.27 \pm 0.02$ \\ \end{tabular} \caption{Security scores for models} \label{tab:sec_model} \end{table}}
 \newcommand{\daTabSecScoressummary}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{Security score} \\ &&&& (out of 10) \\ \hline nexus & $0.36 \pm 0.00$ & $0.52 \pm 0.00$ & 0.68 & $5.0 \pm 0.0$ \\  notnexus & $0.12 \pm 0.00$ & $0.039 \pm 0.000$ & 0.64 & $2.7 \pm 0.0$ \\ \end{tabular} \caption{Security scores for summaries} \label{tab:sec_summary} \end{table}}
 \newcommand{\daUpdatednessPerc}{$5.6 \pm 0.0\%$}
-\newcommand{\daNumOSDevices}{18899}
+\newcommand{\daNumOSDevices}{18\,900}
 \newcommand{\daSigNumDevicesDay}{20}
-\newcommand{\daSigNumDeviceDays}{10000}
+\newcommand{\daSigNumDeviceDays}{10\,000}
 \newcommand{\daTabAndVulns}{\begin{table} \centering \begin{tabular}{l|c|l} Vulnerability & Date known & How known \\ \hline KillingInTheNameOf psneuter ashmem & 2010-07-13 & Fixed on \\ exploid udev & 2010-07-15 & Discovered on \\ RageAgainstTheCage adb & 2010-08-21 & Discovered on \\ levitator & 2011-03-10 & Discovered on \\ Gingerbreak & 2011-04-18 & Fixed on \\ zergRush & 2011-10-06 & Discovered on \\ APK duplicate file & 2013-02-18 & Discovered on \\ APK unchecked name & 2013-06-30 & Discovered on \\ APK unsigned shorts & 2013-07-03 & Fixed on \\ keystore buffer & 2013-09-09 & Discovered on \\ TwerkMyMoto & 2013-11-24 & Discovered on \\ vold asec & 2014-01-27 & Fixed on \\\end{tabular} \caption{Root equivalent vulnerabilities in Android} \label{tab:andvulns} \end{table}}
 \newcommand{\daOSYearsOfData}{3}
-\newcommand{\daSecScoreBestmanufacturer}{lg}
-\newcommand{\daSecScoreWorstmanufacturer}{symphony}
+\newcommand{\daSecScoreBestmanufacturer}{LG}
+\newcommand{\daSecScoreWorstmanufacturer}{Symphony}
 \newcommand{\daSecScoreBestmodel}{Galaxy Nexus}
@@
 \newcommand{\daSecScoreWorstsummary}{notnexus}
-\newcommand{\daSecScoreBestmanufacturerScore}{$3.3 \pm 0.0$}
-\newcommand{\daSecScoreWorstmanufacturerScore}{$0.15 \pm 0.00$}
+\newcommand{\daSecScoreBestmanufacturerScore}{$3.4 \pm 0.0$}
+\newcommand{\daSecScoreWorstmanufacturerScore}{$0.18 \pm 0.00$}
 \newcommand{\daSecScoreBestmodelScore}{$4.8 \pm 0.0$}
@@
 \newcommand{\daAdbEnabledPerc}{$20 \pm 0\%$}
-\newcommand{\daNumUpdatesUpgrades}{3325}
+\newcommand{\daNumUpdatesUpgrades}{3\,320}
 \newcommand{\daNumUpdatesDowngrades}{147}
@@
 \newcommand{\daStartDate}{2011-07-01}
-\newcommand{\daEndDate}{2014-11-04}
-\newcommand{\daNumOperators}{1393}
-\newcommand{\daTabSecScoresoperator}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{Security score} \\ &&&& (out of 10) \\ \hline O2 uk & $0.25 \pm 0.00$ & $0.12 \pm 0.00$ & 0.37 & $3.8 \pm 0.0$ \\  T-Mobile & $0.19 \pm 0.00$ & $0.18 \pm 0.00$ & 0.48 & $3.6 \pm 0.0$ \\  Orange & $0.21 \pm 0.00$ & $0.087 \pm 0.003$ & 0.39 & $3.5 \pm 0.0$ \\  Sprint & $0.18 \pm 0.00$ & $0.11 \pm 0.00$ & 0.44 & $3.4 \pm 0.0$ \\  3 & $0.15 \pm 0.00$ & $0.093 \pm 0.003$ & 0.49 & $3.2 \pm 0.0$ \\  AT\&T & $0.12 \pm 0.00$ & $0.077 \pm 0.002$ & 0.43 & $3.1 \pm 0.0$ \\  vodafone uk & $0.13 \pm 0.00$ & $0.10 \pm 0.00$ & 0.52 & $3.1 \pm 0.0$ \\  Verizon & $0.18 \pm 0.00$ & $0.087 \pm 0.002$ & 0.81 & $2.8 \pm 0.0$ \\  unknown & $0.097 \pm 0.001$ & $0.22 \pm 0.00$ & 1.07 & $2.6 \pm 0.0$ \\  n Telenor & $0.026 \pm 0.001$ & $0.10 \pm 0.00$ & 1.42 & $1.6 \pm 0.0$ \\  Airtel & $0.066 \pm 0.002$ & $0.028 \pm 0.002$ & 1.95 & $1.1 \pm 0.0$ \\  Grameenphone & $0.0078 \pm 0.0005$ & $0.015 \pm 0.001$ & 2.03 & $0.77 \pm 0.00$ \\  Robi & $\left(8.4 \pm 8.4\right) \times 10^{-5}$ & $0.041 \pm 0.003$ & 2.22 & $0.71 \pm 0.01$ \\  banglalink & $0.00016 \pm 0.00011$ & $0.029 \pm 0.002$ & 2.92 & $0.39 \pm 0.00$ \\ \end{tabular} \caption{Security scores for operators} \label{tab:sec_operator} \end{table}}
+\newcommand{\daEndDate}{2014-11-05}
+\newcommand{\daNumOperators}{1\,390}
+\newcommand{\daTabSecScoresoperator}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{Security score} \\ &&&& (out of 10) \\ \hline O2 uk & $0.25 \pm 0.00$ & $0.12 \pm 0.00$ & 0.37 & $3.8 \pm 0.0$ \\  T-Mobile & $0.19 \pm 0.00$ & $0.18 \pm 0.00$ & 0.48 & $3.6 \pm 0.0$ \\  Orange & $0.21 \pm 0.00$ & $0.085 \pm 0.003$ & 0.39 & $3.5 \pm 0.0$ \\  Sprint & $0.18 \pm 0.00$ & $0.11 \pm 0.00$ & 0.44 & $3.4 \pm 0.0$ \\  3 & $0.15 \pm 0.00$ & $0.093 \pm 0.003$ & 0.49 & $3.2 \pm 0.0$ \\  AT\&T & $0.12 \pm 0.00$ & $0.077 \pm 0.002$ & 0.43 & $3.1 \pm 0.0$ \\  Vodafone uk & $0.13 \pm 0.00$ & $0.10 \pm 0.00$ & 0.52 & $3.1 \pm 0.0$ \\  Verizon & $0.18 \pm 0.00$ & $0.087 \pm 0.002$ & 0.81 & $2.8 \pm 0.0$ \\  unknown & $0.097 \pm 0.001$ & $0.22 \pm 0.00$ & 1.07 & $2.6 \pm 0.0$ \\  n Telenor & $0.026 \pm 0.001$ & $0.10 \pm 0.00$ & 1.42 & $1.6 \pm 0.0$ \\  Airtel & $0.065 \pm 0.002$ & $0.028 \pm 0.002$ & 1.95 & $1.1 \pm 0.0$ \\  Grameenphone & $0.0077 \pm 0.0005$ & $0.015 \pm 0.001$ & 2.03 & $0.77 \pm 0.00$ \\  Robi & $\left(8.0 \pm 8.0\right) \times 10^{-5}$ & $0.040 \pm 0.003$ & 2.22 & $0.71 \pm 0.01$ \\  banglalink & $0.00016 \pm 0.00011$ & $0.029 \pm 0.002$ & 2.92 & $0.39 \pm 0.00$ \\ \end{tabular} \caption{Security scores for operators} \label{tab:sec_operator} \end{table}}
 \newcommand{\daSecScoreBestoperator}{O2 uk}
@@
 \newcommand{\daFullDeployedAt}{95\%}
-\newcommand{\daOSCurveFitParamFirst}{87.37}
+\newcommand{\daOSCurveFitParamFirst}{87.4}
 \newcommand{\daOSCurveFitParamSecond}{\num{0.00315}}
-\newcommand{\daOSCurveFitRMSE}{0.11}
-\newcommand{\daAPICurveFitParamFirst}{109.77}
+\newcommand{\daOSCurveFitRMSE}{0.114}
+\newcommand{\daAPICurveFitParamFirst}{110.0}
 \newcommand{\daAPICurveFitParamSecond}{\num{0.00315}}
-\newcommand{\daAPICurveFitRMSE}{0.12}
-\newcommand{\daOSCurvePolyRMSE}{0.11}
-\newcommand{\daOSCurveSplineRMSE}{0.11}
+\newcommand{\daAPICurveFitRMSE}{0.116}
+\newcommand{\daOSCurvePolyRMSE}{0.115}
+\newcommand{\daOSCurveSplineRMSE}{0.115}
 \newcommand{\daOSCurveHalfDeployed}{$\left(3.1 \pm 0.8\right) \times 10^{2}$}
 \newcommand{\daOSCurveFullDeployed}{$\left(0.0 \pm 2.4\right) \times 10^{5}$}
-\newcommand{\daAPICurvePolyRMSE}{0.12}
-\newcommand{\daAPICurveSplineRMSE}{0.12}
+\newcommand{\daAPICurvePolyRMSE}{0.117}
+\newcommand{\daAPICurveSplineRMSE}{0.117}
 \newcommand{\daAPICurveHalfDeployed}{$\left(3.3 \pm 0.8\right) \times 10^{2}$}
 \newcommand{\daAPICurveFullDeployed}{$\left(0.0 \pm 2.4\right) \times 10^{5}$}
-\newcommand{\daVulnAPKDuplicateFileOctoberPerc}{91.46\%}
+\newcommand{\daVulnAPKDuplicateFileOctoberPerc}{91.5\%}
 \newcommand{\daNumUpdatesBigUpgrades}{778}
 \newcommand{\daNumUpdatesSkippedBig}{3}
-\newcommand{\daPercBigUpgrades}{22.41\%}
+\newcommand{\daPercBigUpgrades}{22.4\%}
 \newcommand{\daUpdatesPerYear}{$1.4 \pm 0.0$}
 \newcommand{\daOSMonthsOfData}{40}
-\newcommand{\daMeanInsecurityPercNominal}{87.24\%}
-\newcommand{\daVulnFreeNominal}{0.13}
-\newcommand{\daUpdatednessNominal}{0.06}
-\newcommand{\daUpdatednessPercNominal}{5.60\%}
+\newcommand{\daMeanInsecurityPercNominal}{87.2\%}
+\newcommand{\daVulnFreeNominal}{0.128}
+\newcommand{\daUpdatednessNominal}{0.0561}
+\newcommand{\daUpdatednessPercNominal}{5.61\%}
 \newcommand{\daSecurityScoreNominal}{2.75}
 \newcommand{\daUpdatesPerYearNominal}{1.35}
-\newcommand{\daOSCurveHalfDeployedNominal}{307.28}
-\newcommand{\daOSCurveFullDeployedNominal}{1037.82}
-\newcommand{\daAPICurveHalfDeployedNominal}{329.75}
-\newcommand{\daAPICurveFullDeployedNominal}{1060.53}
+\newcommand{\daOSCurveHalfDeployedNominal}{308.0}
+\newcommand{\daOSCurveFullDeployedNominal}{1\,040}
+\newcommand{\daAPICurveHalfDeployedNominal}{330.0}
+\newcommand{\daAPICurveFullDeployedNominal}{1\,060}
 \newcommand{\daSecScoreBestoperatorScoreNominal}{3.79}
-\newcommand{\daSecScoreWorstoperatorScoreNominal}{0.39}
-\newcommand{\daSecScoreBestmodelScoreNominal}{4.76}
-\newcommand{\daSecScoreWorstmodelScoreNominal}{0.27}
-\newcommand{\daAdbEnabledPercNominal}{19.57\%}
+\newcommand{\daSecScoreWorstoperatorScoreNominal}{0.394}
+\newcommand{\daSecScoreBestmodelScoreNominal}{4.8}
+\newcommand{\daSecScoreWorstmodelScoreNominal}{0.274}
+\newcommand{\daAdbEnabledPercNominal}{19.6\%}
+\newcommand{\daSecScoreBestmanufacturerScoreNominal}{3.38}
+\newcommand{\daSecScoreWorstmanufacturerScoreNominal}{0.177}
+\newcommand{\daUpdatesPerMonthPerVersion}{3.61}

### 2014-11-12 securityupdates.tex
Use ieee biblatex style
Updated dastats data and remove fp package as now use pre-computed value

@@
 \usepackage{siunitx}
-\usepackage{fp}
 \usepackage[disable]{todonotes}
@@
 \newcommand{\todolater}[1]{}%\todo
-\RequirePackage[date=terse, isbn=true, doi=true, url=false, urldate=iso8601, maxbibnames=9, backref=false, backend=bibtex]{biblatex}
+\RequirePackage[date=terse, isbn=true, doi=true, url=false, urldate=iso8601, maxbibnames=9, backref=false, backend=bibtex, style=ieee]{biblatex}
 \addbibresource{securityupdates.bib}
@@
 Over time newer versions of Android come to dominate, but quantifying what proportion of that is due to new phones being purchased and what proportion is due to updates to existing devices is a question we cannot answer.
- The only available large data set we know of with longitudinal traces of Android devices OS version numbers is \da\ but that only records \daNumUpdatesUpgrades\ upgrades which broken down over \daOSMonthsOfData\ months and \daNumSigOSVersions\ is only \FPeval{\daUpdatesPerMonthPerVersion}{round(\daNumUpdatesUpgrades/(\daOSMonthsOfData * \daNumSigOSVersions),1)}\daUpdatesPerMonthPerVersion\ updates per version per month which is not enough to build a statistically significant prediction of the expected transition to newer versions of Android.
+ The only available large data set we know of with longitudinal traces of Android devices OS version numbers is \da\ but that only records \daNumUpdatesUpgrades\ upgrades which broken down over \daOSMonthsOfData\ months and \daNumSigOSVersions\ OS versions is only \daUpdatesPerMonthPerVersion\ updates per version per month which is not enough to build a statistically significant prediction of the expected transition to newer versions of Android.
  \item Why do users downgrade their phones?

### 2014-11-12 security_scoring.tex
Terminology consistency in security scoring

@@
-\subsection{Experiment 3: Comparing manufacturers, operators and devices}
+\subsection{Experiment 3: Comparing device manufacturers, device models and network operators}
 \label{sec:security_scoring}\label{sec:exp:security_score}
 
-To allow buyers of Android devices to purchase those devices which have the best security they need to know how different manufactuers, devices and network operators compare in terms of the security they provide.
-We propose a method to score a manufacturer, operator or device based on its historic performance at keeping devices up to date and fixing security vulnerabilities.
-We find that Android as a whole gets a score of \daSecurityScore\ out of 10, the highest scoring manufacturer is \emph{\daSecScoreBestmanufacturer} (\daSecScoreBestmanufacturerScore\ out of 10) and the lowest scoring is \emph{\daSecScoreWorstmanufacturer} (\daSecScoreWorstmanufacturerScore\ out of 10).
+To allow buyers of Android devices to purchase those devices which have the best security they need to know how different device manufacturers, device models and network operators compare in terms of the security they provide.
+We propose a method to score a device manufacturer, network operator or device model based on its historic performance at keeping devices up to date and fixing security vulnerabilities.
+We find that Android as a whole gets a score of \daSecurityScore\ out of 10, the highest scoring device manufacturer is \emph{\daSecScoreBestmanufacturer} (\daSecScoreBestmanufacturerScore\ out of 10) and the lowest scoring is \emph{\daSecScoreWorstmanufacturer} (\daSecScoreWorstmanufacturerScore\ out of 10).
 
@@
 
-Computing how good a particular manufacturer or device model is from a security standpoint is difficult as it depends on a number of factors which are hard to observe, particularly on a large scale.
+Computing how good a particular device manufacturer or device model is from a security standpoint is difficult as it depends on a number of factors which are hard to observe, particularly on a large scale.
 Ideally we would consider both the prevalence of potential problems which were not exploited and actual security failures.
@@
 This is equivalent to Acer and Jackson's proposal to measure the security based on the proportion of users with at least one unpatched critical vulnerability~\cite{Acer2010} and similar to the Vulnerability Free Days (VFD) score~\cite{Wright2014}.
-  Unlike VFD this is the proportion of running devices which were free from critical vulnerabilities, rather than days which the manufacturer was free from outstanding critical vulnerabilities as that does not take account of the update process.
-  \item[update $u$] The proportion of devices which run the latest version of Android shipped to any device produced by that manufacturer.
+  Unlike VFD this is the proportion of running devices which were free from critical vulnerabilities, rather than days which the device manufacturer was free from outstanding critical vulnerabilities as that does not take account of the update process.
+  \item[update $u$] The proportion of devices which run the latest version of Android shipped to any device produced by that device manufacturer.
 This is a measure of internal updatedness, a low score would mean many devices are being left behind.
   This assumes that newer versions are better with stronger security, historically steps have been taken to improve Android security in newer versions so this assumption should generally hold, but sometimes new updates introduce new vulnerabilities.
-  \item[mean $m$] The mean number of outstanding vulnerabilities affecting devices not fixed on any device shipped by the manufacturer.
+  \item[mean $m$] The mean number of outstanding vulnerabilities affecting devices not fixed on any device shipped by the device manufacturer.
 This is related to the Median Active Vulnerabilities (MAV) measure~\cite{Wright2014} but is the mean rather than the median, as this gives a continuous value.
@@
 However it does not capture the increased risk caused by there being multiple known vulnerabilities which gives an attacker more opportunities and increases the likelihood of a piece of malware having a matching exploit.
-This is captured by the $m$ score which measures the size of the manufacturers queue of outstanding vulnerabilities but does not take into account the update process or measure the actual end user security.
+This is captured by the $m$ score which measures the size of the device manufacturers queue of outstanding vulnerabilities but does not take into account the update process or measure the actual end user security.
 Neither of these metrics capture whether devices are being left behind and not being kept up to date with the most recent (and hopefully most secure) version which is captured by $u$.
@@
 
-\todo{There must be some way of putting an error bound on our results by considering the error on the whole DA set and the size of the subset used for each manufacturer and model}
+\todo{There must be some way of putting an error bound on our results by considering the error on the whole DA set and the size of the subset used for each device manufacturer and device model}
 
@@
 \daTabSecScoresmanufacturer
-Different manufacturers have very different scores, Table~\ref{tab:sec_manufacturer} shows the scores for the \daNumSigManufacturers\ manufacturers with a significant presence in our data with \emph{\daSecScoreBestmanufacturer} (\daSecScoreBestmanufacturerScore\ out of 10) scoring highest and \emph{\daSecScoreWorstmanufacturer} (\daSecScoreWorstmanufacturerScore\ out of 10) scoring lowest.
-Manufacturers are considered significant if we have data from at least \daSigNumDevices\ devices and at least \daSigNumDeviceDays\ days of contributions.
+Different device manufacturers have very different scores, Table~\ref{tab:sec_manufacturer} shows the scores for the \daNumSigManufacturers\ device manufacturers with a significant presence in our data with \emph{\daSecScoreBestmanufacturer} (\daSecScoreBestmanufacturerScore\ out of 10) scoring highest and \emph{\daSecScoreWorstmanufacturer} (\daSecScoreWorstmanufacturerScore\ out of 10) scoring lowest.
+Device manufacturers are considered significant if we have data from at least \daSigNumDevices\ devices and at least \daSigNumDeviceDays\ days of contributions.
 Additionally for $m$ and $u$ we ignore days with less than \daSigNumDevicesDay\ devices contributing to that day's score.
@@
 \daTabSecScoresmodel
-Even within manufacturers different models can have very different update behaviours and hence security.
+Even within device manufacturers different models can have very different update behaviours and hence security.
 Table~\ref{tab:sec_model} shows the results for the \daNumSigModels\ device models which have a significant presence by the same metric with \emph{\daSecScoreBestmodel} (\daSecScoreBestmodelScore\ out of 10) scoring highest and \emph{\daSecScoreWorstmodel} (\daSecScoreWorstmodelScore\ out of 10) scoring lowest.
@@
 \daTabSecScoresoperator
-We also analysed different operators, for the \daNumSigOperators\ operators with a significant presence in our data Table~\ref{tab:sec_operator} shows the results \emph{\daSecScoreBestoperator} (\daSecScoreBestoperatorScore\ out of 10) scoring highest and \emph{\daSecScoreWorstoperator} (\daSecScoreWorstoperatorScore\ out of 10) scoring lowest.
-However the score of an operator is affected by the manufacturers of the devices which are in use on its network.
-This is in turn affected by both what devices an operator offers to users and upon which devices users choose.
-Hence having a worse score does not necessarily mean that an operator is worse, it could be that its users all pick phones from a worse manufacturer, for example because they were cheaper.
-An operator could use data from this paper to exclude insecure devices from those offered to consumers.
-An added value analysis of operators which takes into account the device mix used by users of that operator would make it possible to determine whether an operator is making the situation better or worse by the way it ships updates to users.
+We also analysed different network operators, for the \daNumSigOperators\ network operators with a significant presence in our data Table~\ref{tab:sec_operator} shows the results \emph{\daSecScoreBestoperator} (\daSecScoreBestoperatorScore\ out of 10) scoring highest and \emph{\daSecScoreWorstoperator} (\daSecScoreWorstoperatorScore\ out of 10) scoring lowest.
+However the score of an network operator is affected by the device manufacturers of the devices which are in use on its network.
+This is in turn affected by both what devices a network operator offers to users and upon which devices users choose.
+Hence having a worse score does not necessarily mean that a network operator is worse, it could be that its users all pick phones from a worse device manufacturer, for example because they were cheaper.
+A network operator could use data from this paper to exclude insecure devices from those offered to consumers.
+An added value analysis of network operators which takes into account the device mix used by users of that network operator would make it possible to determine whether a network operator is making the situation better or worse by the way it ships updates to users.
 However we suspect that our sample size is too small to do that as while we have significant numbers of devices for each of the device models in Table~\ref{tab:sec_model} and for each of the network operators in Table~\ref{tab:sec_operator} we would need a significant number of each model in each network operator.

### 2014-11-12 security_scoring.tex
Corrections from Tom

@@
 We can then test whether this seems fair by comparing the version data for the highest and lowest scoring models.
-Figure~\ref{fig:worst_model_full} shows the full version distribution for \emph{\daSecScoreWorstmodel} and Figure~\ref{fig:best_model_full} shows the same for \emph{\daSecScoreBestmodel} there is a marked difference in the distribution of updates between the two with only \daSecScoreWorstmodelNumFullVersions\ full versions for \emph{\daSecScoreWorstmodel} and \daSecScoreBestmodelNumFullVersions\ for \emph{\daSecScoreBestmodel}.
+Figure~\ref{fig:worst_model_full} shows the full version distribution for \emph{\daSecScoreWorstmodel} and Figure~\ref{fig:best_model_full} shows the same for \emph{\daSecScoreBestmodel}.
+There is a marked difference in the distribution of updates between the two with only \daSecScoreWorstmodelNumFullVersions\ full versions for \emph{\daSecScoreWorstmodel} and \daSecScoreBestmodelNumFullVersions\ for \emph{\daSecScoreBestmodel}.
 

### 2014-11-12 securityupdates.tex
Corrections from Tom

@@
 The development of Android is distributed among many parties and the security of Android relies on many open source projects, include the Linux kernel, OpenSSL and BouncyCastle as well as Google who build the core platform.
-In addition, device manufacturers (of which we are aware of \daNumManufacturers\footnote{\label{foot:dadata}We computed this from the \da~\cite{Wagner2013} data, see \S\ref{sec:android_update_process}.}) and network operators (\daNumOperators\dafoot) may make further modifications before devices are shipped to customers.
+In addition, device manufacturers (we know of \daNumManufacturers\footnote{\label{foot:dadata}We computed this from the \da~\cite{Wagner2013} data, see \S\ref{sec:android_update_process}.}) and network operators (\daNumOperators\dafoot) may make further modifications before devices are shipped to customers.
 Understanding this ecosystem is important as device manufacturers have introduced additional vulnerabilities in the past~\cite{Grace2012}. 
@@
 We used data from the \da\ project~\cite{Wagner2013}.
-The \da\ data contains info from \daNumDevices\ devices with a total of \daDeviceDays\footnote{Here we are only counting devices and days for which we have valid OS version data.} device days, contains data for more than \daMonths\ months for \daMonthsDevices\ devices.
+The \da\ data contains info from \daNumDevices\ devices with a total of \daDeviceDays\footnote{Here we are only counting devices and days for which we have valid OS version data.} device days, and contains data for more than \daMonths\ months for \daMonthsDevices\ devices.
 Device Analyzer used an Android app which volunteers install on their Android devices to collect data on what happens on Android phones.
-The data collected is available for researchers to use as we do here.
+The data collected is available for researchers to use as we do here under agreements which protect the privacy of the users.
 
@@
 \end{figure}
-The logitudinal data on the number of vulnerabilities affecting the \daNumDeviceDataDevices\ devices which have contributed the most days of data to \da\ changes over time is shown in Figure~\ref{fig:device_data_security}.
+The longitudinal data on the number of vulnerabilities affecting the \daNumDeviceDataDevices\ devices which have contributed the most days of data to \da\ changes over time is shown in Figure~\ref{fig:device_data_security}.
 The trend that we saw in Figure~\ref{fig:proportioninsecure} of security improving and then getting worse is also shown here.
@@
 However some apps are installed from other markets which are less well policed (infection rates of between 0.20\%~ and 0.47\%~\cite{Zhou2012a}).
-Apps can also dynamically download code at runtime which then cannot be detected by static analysis and so bypasses Bouncer.
+Apps can also dynamically download code at runtime which then cannot be detected by static analysis and so bypass Bouncer.
 
-There are also other ways that malicious code could end up running on Android devices, the addJavascriptInterface vulnerability (CVE-2012-6636 ) allows javascript running in a webview which has an exposed interface to execute arbitrary Java code and hence load exploits.
+There are also other ways that malicious code could end up running on Android devices, and the addJavascriptInterface vulnerability (CVE-2012-6636) allows javascript running in a webview which has an exposed interface to execute arbitrary Java code and hence load exploits.
 While fixes have been produced for this the change is API breaking and so to be safe an app must target an API version greater than 17 and be running on a device with an Android version of at least 4.2.
@@
  \item Why was there a gap in vulnerability discovery in Android in 2012?
- \item Why do so many devices have \texttt{adb} enabled? \daAdbEnabledPerc\ seems rather high for a developer feature.
+ \item Why do so many devices have \texttt{adb} enabled?
+\daAdbEnabledPerc\ seems rather high for a developer feature.
+This could be a selection bias or device manufacturers could be enabling it by default.
  \item How much of updating is new handsets and how much is updates being deployed?
@@
 
-The security protections used in iOS such as, app review, code signing, Data Execution Prevention (DEP) and Address Space Layout Randomisation (ASLR) and mandatory access controls has resulted in a lower level of malware affecting iOS~\cite{Felt2011} but can still be bypassed~\cite{Wang2013a}.
+The security protections used in iOS such as app review, code signing, Data Execution Prevention (DEP) and Address Space Layout Randomisation (ASLR) and mandatory access controls has resulted in a lower level of malware affecting iOS~\cite{Felt2011} but can still be bypassed~\cite{Wang2013a}.
 

### 2014-11-12 securityupdates.tex
Add some more text to the caption of figure 1

@@
 \includegraphics[width=\columnwidth]{figures/proportioninsecure}
-\caption{Proportion of devices running insecure, maybe insecure and secure versions of Android against time}
+\caption{Proportion of devices running insecure, maybe insecure and secure versions of Android against time.
+The red vertical lines are caused by vulnerabilities being discovered.
+This graph is explained further in XXX.}
 \label{fig:proportioninsecure}

### 2014-11-12 securityupdates.tex
Consistently use \, rather than , as a thousands separator

@@
 Various attempts have been made to detect malware.
-RiskRanker classified 3,281/118,318 (2.8\%) apps as risky of which 718 (22\%) were malware and 322 (10\%) were previously unknown malware, an infection rate of 0.6\% across multiple markets~\cite{Grace2012a}.
+RiskRanker classified 3\,281/118\,318 (2.8\%) apps as risky of which 718 (22\%) were malware and 322 (10\%) were previously unknown malware, an infection rate of 0.6\% across multiple markets~\cite{Grace2012a}.
 It uses signatures of exploits, static analysis of calls to high risk operations, detection of encrypted native code execution and dynamic Dalvik code loading to detect risky apps.
-DroidRanger found 148/182,823 (0.08\%) apps to be malicious across multiple markets of which 29 were previously unknown~\cite{Zhou2012a}.
+DroidRanger found 148/182\,823 (0.08\%) apps to be malicious across multiple markets of which 29 were previously unknown~\cite{Zhou2012a}.
 It used permission-based behavioural fingerprinting -- looking at the permissions of known malware and heuristic-based filtering -- dynamic loading of both Dalvik and native code.
-AnDarwin detects similar apps, it found 169/265,359 (0.06\%) malicious apps based on the difference in permissions between the malicious app and the app which had been cloned~\cite{Crussell2013}.
+AnDarwin detects similar apps, it found 169/265\,359 (0.06\%) malicious apps based on the difference in permissions between the malicious app and the app which had been cloned~\cite{Crussell2013}.
 It used clustering based on semantic vectors derived from the program dependence graphs to detect similar apps.
-The Malware Genome project collected 1,260 malware samples from 2010--2011~\cite{Zhou2012b}.
+The Malware Genome project collected 1\,260 malware samples from 2010--2011~\cite{Zhou2012b}.
 They found that the best case for anti-malware detection on this data was 79.6\%.
@@
 All these app analysis projects have been hampered by the difficulty of obtaining full datasets of Android apps as Google does not make these available (and researchers who automatically download them violate Google's Terms of Service).
-PlayDrone was a particularly effective project which circumvented Google's protective measures and downloaded over 1,100,000 apps from Google Play, allowing an in depth analysis~\cite{Viennot2014}.
+PlayDrone was a particularly effective project which circumvented Google's protective measures and downloaded over 1\,100\,000 apps from Google Play, allowing an in depth analysis~\cite{Viennot2014}.
 

### 2014-11-12 securityupdates.tex
Re-run versions.py with updated latex_value

@@
 \begin{tabular}{l|r|r}
-Project	&	\# releases	&	latency \\ \hline
+Project	&	\# releases	&	latency (days) \\ \hline
 linux	&	\linuxNumVersions	&	\linuxMeanUpdateLatency \\

### 2014-11-12 versions.py

@@ -136,9 +136,9 @@ def latency(version_dates, version_map):
 def calculate_latencies(version_dates):
     linux_latencies = latency(version_dates['linux'], OrderedDict(avo.os_to_kernel))
     set_latex_value('linuxMeanUpdateLatency', ufloat(statistics.mean(linux_latencies.values()),statistics.stdev(linux_latencies.values())))
-    openssl_latencies = latency(version_dates['openssl'], OrderedDict(avo.os_to_openssl_version))
+    openssl_latencies = latency(version_dates['openssl'], OrderedDict(avo.os_to_project['openssl']))
     set_latex_value('opensslMeanUpdateLatency', ufloat(statistics.mean(openssl_latencies.values()),statistics.stdev(openssl_latencies.values())))
-    bouncycastle_latencies = latency(version_dates['bouncycastle'], OrderedDict(avo.os_to_bouncycastle_version))
+    bouncycastle_latencies = latency(version_dates['bouncycastle'], OrderedDict(avo.os_to_project['bouncycastle']))
     set_latex_value('bouncycastleMeanUpdateLatency',ufloat(statistics.mean(bouncycastle_latencies.values()),statistics.stdev(bouncycastle_latencies.values())))
 
 

### 2014-11-12 avostats.tex
Updated graphs and avo data
updated data and some figures

@@
 \newcommand{\avoNumBigExternalLinesOfCode}{25 Million}
-\newcommand{\avoBigExternalLinesOfCodePerc}{99.69\%}
+\newcommand{\avoBigExternalLinesOfCodePerc}{99.7\%}
 \newcommand{\avoNumVulnerabilities}{31}
@@
 \newcommand{\avoVulnsPerYearNominal}{7.66}
-\newcommand{\avoVulnsPerYear}{$7.7 \pm 1.4$}
+\newcommand{\avoVulnsPerYear}{$7.66 \pm 1.38$}
 \newcommand{\avoVulnsPerYearAllAndroidNominal}{3.71}
-\newcommand{\avoVulnsPerYearAllAndroid}{$3.7 \pm 1.0$}
-\newcommand{\avoTabAndVulns}{\begin{table} \centering \begin{tabular}{l|l|c} Vulnerability & How known & Date \\ \hline KillingInTheNameOf psneuter ashmem & Fixed on & 2010-07-13 \\ exploid udev & Discovered on & 2010-07-15 \\ levitator & Discovered on & 2011-03-10 \\ Gingerbreak & Fixed on & 2011-04-18 \\ zergRush & Discovered on & 2011-10-06 \\ APK duplicate file & Discovered on & 2013-02-18 \\ APK unchecked name & Discovered on & 2013-06-30 \\ APK unsigned shorts & Fixed on & 2013-07-03 \\ vold asec & Fixed on & 2014-01-27 \\ Fake ID & Fixed on & 2014-04-17 \\ TowelRoot & Discovered on & 2014-05-03 \\\end{tabular} \caption{Root equivalent vulnerabilities in Android} \label{tab:andvulns} \end{table}}
+\newcommand{\avoVulnsPerYearAllAndroid}{$3.71 \pm 0.957$}
+\newcommand{\avoTabAndVulns}{\begin{table} \centering \begin{tabular}{l|l|c} Vulnerability & How known & Date \\ \hline KillingInTheNameOf psneuter ashmem & Fixed on & 2010-07-13 \\ exploid udev & Discovered on & 2010-07-15 \\ levitator & Discovered on & 2011-03-10 \\ Gingerbreak & Fixed on & 2011-04-18 \\ zergRush & Discovered on & 2011-10-06 \\ APK duplicate file & Discovered on & 2013-02-18 \\ APK unchecked name & Discovered on & 2013-06-30 \\ APK unsigned shorts & Fixed on & 2013-07-03 \\ vold asec & Fixed on & 2014-01-27 \\ Fake ID & Fixed on & 2014-04-17 \\ TowelRoot & Discovered on & 2014-05-03 \\\end{tabular} \caption{Critical vulnerabilities in Android} \label{tab:andvulns} \end{table}}
 \newcommand{\avoNumBigExternalProjects}{40}
@@
 \newcommand{\avoNumAnalysedExternalLinesOfCode}{6 Million}
-\newcommand{\avoAnalysedExternalLinesOfCodePerc}{24.85\%}
-\newcommand{\avoBigExternalMedianVersions}{1.50}
+\newcommand{\avoAnalysedExternalLinesOfCodePerc}{24.9\%}
+\newcommand{\avoBigExternalMedianVersions}{1.5}
 \newcommand{\avoBigExternalMeanVersionsNominal}{2.07}
-\newcommand{\avoBigExternalMeanVersions}{$2.1 \pm 1.4$}
+\newcommand{\avoBigExternalMeanVersions}{$2.07 \pm 1.44$}
 \newcommand{\avoBigExternalTotalVersions}{58}

### 2014-11-12 dastats.tex
updated data and some figures

@@
 \newcommand{\daOSTotalDaysData}{1\,130\,000}
-\newcommand{\daMeanInsecurityPerc}{$87 \pm 0\%$}
+\newcommand{\daMeanInsecurityPerc}{$87.2 \pm 0.0342\%$}
 \newcommand{\daMeanOutstandingVulnerabilities}{0.637}
-\newcommand{\daUpdatedness}{$0.056 \pm 0.000$}
-\newcommand{\daSecurityScore}{$2.8 \pm 0.0$}
+\newcommand{\daUpdatedness}{$0.0561 \pm 0.000258$}
+\newcommand{\daSecurityScore}{$2.75 \pm 0.00157$}
 \newcommand{\daNumFullVersions}{1\,140}
@@
 \newcommand{\daNumPossibleSecurityUpdates}{2\,140}
-\newcommand{\daTabSecScoresmanufacturer}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{Security score} \\ &&&& (out of 10) \\ \hline LG & $0.12 \pm 0.00$ & $0.32 \pm 0.00$ & 0.74 & $3.4 \pm 0.0$ \\  Motorola & $0.16 \pm 0.00$ & $0.13 \pm 0.00$ & 0.77 & $2.9 \pm 0.0$ \\  other & $0.13 \pm 0.00$ & $0.056 \pm 0.000$ & 0.64 & $2.8 \pm 0.0$ \\  Samsung & $0.14 \pm 0.00$ & $0.047 \pm 0.000$ & 0.71 & $2.7 \pm 0.0$ \\  Sony & $0.13 \pm 0.00$ & $0.18 \pm 0.00$ & 1.02 & $2.6 \pm 0.0$ \\  HTC & $0.14 \pm 0.00$ & $0.087 \pm 0.001$ & 0.89 & $2.6 \pm 0.0$ \\  asus & $0.16 \pm 0.00$ & $0.56 \pm 0.01$ & 6.06 & $2.3 \pm 0.0$ \\  Symphony & $0.0015 \pm 0.0002$ & $0.042 \pm 0.001$ & 4.87 & $0.18 \pm 0.00$ \\ \end{tabular} \caption{Security scores for manufacturers} \label{tab:sec_manufacturer} \end{table}}
-\newcommand{\daTabSecScoresmodel}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{Security score} \\ &&&& (out of 10) \\ \hline Galaxy Nexus & $0.54 \pm 0.00$ & $0.53 \pm 0.01$ & 1.53 & $4.8 \pm 0.0$ \\  Nexus 4 & $0.23 \pm 0.00$ & $0.82 \pm 0.01$ & 6.06 & $3.4 \pm 0.0$ \\  Nexus 7 & $0.19 \pm 0.00$ & $0.74 \pm 0.01$ & 5.92 & $3.0 \pm 0.0$ \\  other & $0.13 \pm 0.00$ & $0.12 \pm 0.00$ & 0.64 & $2.9 \pm 0.0$ \\  Desire HD & $0.083 \pm 0.003$ & $0.051 \pm 0.004$ & 0.39 & $2.9 \pm 0.0$ \\  HTC Sensation & $0.36 \pm 0.00$ & $0.0063 \pm 0.0062$ & 1.59 & $2.5 \pm 0.0$ \\  GT-I9100 & $0.22 \pm 0.00$ & $0.016 \pm 0.001$ & 1.20 & $2.3 \pm 0.0$ \\  HTC Desire S & $0.022 \pm 0.001$ & $0.016 \pm 0.004$ & 1.00 & $1.7 \pm 0.0$ \\  GT-N7000 & $0.25 \pm 0.00$ & $0.0006 \pm 0.0036$ & 2.52 & $1.4 \pm 0.0$ \\  GT-P1000 & $0.0096 \pm 0.0009$ & $0.0000 \pm 0.0061$ & 1.73 & $0.94 \pm 0.02$ \\  GT-I9300 & $0.15 \pm 0.00$ & $0.014 \pm 0.001$ & 6.16 & $0.64 \pm 0.01$ \\  GT-I9505 & $0.0 \pm 0$ & $0.18 \pm 0.00$ & 6.83 & $0.56 \pm 0.01$ \\  GT-N7100 & $0.067 \pm 0.002$ & $0.0011 \pm 0.0058$ & 6.91 & $0.28 \pm 0.02$ \\  HTC Desire HD & $0.0 \pm 0$ & $0.0005 \pm 0.0070$ & 3.05 & $0.27 \pm 0.02$ \\ \end{tabular} \caption{Security scores for models} \label{tab:sec_model} \end{table}}
-\newcommand{\daTabSecScoressummary}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{Security score} \\ &&&& (out of 10) \\ \hline nexus & $0.36 \pm 0.00$ & $0.52 \pm 0.00$ & 0.68 & $5.0 \pm 0.0$ \\  notnexus & $0.12 \pm 0.00$ & $0.039 \pm 0.000$ & 0.64 & $2.7 \pm 0.0$ \\ \end{tabular} \caption{Security scores for summaries} \label{tab:sec_summary} \end{table}}
-\newcommand{\daUpdatednessPerc}{$5.6 \pm 0.0\%$}
+\newcommand{\daTabSecScoresmanufacturer}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{Security score} \\ &&&& (out of 10) \\ \hline LG & $0.12 \pm 0.00$ & $0.32 \pm 0.00$ & 0.74 & $3.38 \pm 0.01$ \\  Motorola & $0.16 \pm 0.00$ & $0.13 \pm 0.00$ & 0.77 & $2.94 \pm 0.01$ \\  other & $0.13 \pm 0.00$ & $0.06 \pm 0.00$ & 0.64 & $2.75 \pm 0.00$ \\  Samsung & $0.14 \pm 0.00$ & $0.05 \pm 0.00$ & 0.71 & $2.67 \pm 0.00$ \\  Sony & $0.13 \pm 0.00$ & $0.18 \pm 0.00$ & 1.02 & $2.64 \pm 0.01$ \\  HTC & $0.14 \pm 0.00$ & $0.09 \pm 0.00$ & 0.89 & $2.55 \pm 0.00$ \\  asus & $0.16 \pm 0.00$ & $0.56 \pm 0.01$ & 6.06 & $2.31 \pm 0.02$ \\  Symphony & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & 4.91 & $0.18 \pm 0.00$ \\ \end{tabular} \caption{Security scores for manufacturers} \label{tab:sec_manufacturer} \end{table}}
+\newcommand{\daTabSecScoresmodel}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{Security score} \\ &&&& (out of 10) \\ \hline Galaxy Nexus & $0.54 \pm 0.00$ & $0.53 \pm 0.01$ & 1.53 & $4.80 \pm 0.03$ \\  Nexus 4 & $0.23 \pm 0.00$ & $0.82 \pm 0.01$ & 6.06 & $3.39 \pm 0.04$ \\  Nexus 7 & $0.19 \pm 0.00$ & $0.74 \pm 0.01$ & 5.92 & $3.02 \pm 0.04$ \\  other & $0.13 \pm 0.00$ & $0.12 \pm 0.00$ & 0.64 & $2.94 \pm 0.00$ \\  Desire HD & $0.08 \pm 0.00$ & $0.05 \pm 0.00$ & 0.39 & $2.91 \pm 0.02$ \\  HTC Sensation & $0.36 \pm 0.00$ & $0.01 \pm 0.01$ & 1.59 & $2.47 \pm 0.02$ \\  GT-I9100 & $0.22 \pm 0.00$ & $0.02 \pm 0.00$ & 1.20 & $2.33 \pm 0.01$ \\  HTC Desire S & $0.02 \pm 0.00$ & $0.02 \pm 0.00$ & 1.00 & $1.75 \pm 0.01$ \\  GT-N7000 & $0.25 \pm 0.00$ & $0.00 \pm 0.00$ & 2.52 & $1.43 \pm 0.02$ \\  GT-P1000 & $0.01 \pm 0.00$ & $0.00 \pm 0.01$ & 1.73 & $0.94 \pm 0.02$ \\  GT-I9300 & $0.15 \pm 0.00$ & $0.01 \pm 0.00$ & 6.17 & $0.64 \pm 0.01$ \\  GT-I9505 & $0.00 \pm 0.00$ & $0.18 \pm 0.00$ & 6.83 & $0.56 \pm 0.01$ \\  GT-N7100 & $0.07 \pm 0.00$ & $0.00 \pm 0.01$ & 6.91 & $0.28 \pm 0.02$ \\  HTC Desire HD & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & 3.05 & $0.27 \pm 0.02$ \\ \end{tabular} \caption{Security scores for models} \label{tab:sec_model} \end{table}}
+\newcommand{\daTabSecScoressummary}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{Security score} \\ &&&& (out of 10) \\ \hline nexus & $0.36 \pm 0.00$ & $0.52 \pm 0.00$ & 0.68 & $5.02 \pm 0.01$ \\  notnexus & $0.12 \pm 0.00$ & $0.04 \pm 0.00$ & 0.64 & $2.68 \pm 0.00$ \\ \end{tabular} \caption{Security scores for summaries} \label{tab:sec_summary} \end{table}}
+\newcommand{\daUpdatednessPerc}{$5.61 \pm 0.0258\%$}
 \newcommand{\daNumOSDevices}{18\,900}
@@
 \newcommand{\daSecScoreWorstsummary}{notnexus}
-\newcommand{\daSecScoreBestmanufacturerScore}{$3.4 \pm 0.0$}
-\newcommand{\daSecScoreWorstmanufacturerScore}{$0.18 \pm 0.00$}
-\newcommand{\daSecScoreBestmodelScore}{$4.8 \pm 0.0$}
-\newcommand{\daSecScoreWorstmodelScore}{$0.27 \pm 0.02$}
+\newcommand{\daSecScoreBestmanufacturerScore}{$3.38 \pm 0.011$}
+\newcommand{\daSecScoreWorstmanufacturerScore}{$0.175 \pm 0.00428$}
+\newcommand{\daSecScoreBestmodelScore}{$4.8 \pm 0.0254$}
+\newcommand{\daSecScoreWorstmodelScore}{$0.274 \pm 0.0209$}
 \newcommand{\daSecScoreBestsummaryScore}{$4.1 \pm 0.0$}
 \newcommand{\daSecScoreWorstsummaryScore}{$2.9 \pm 0.0$}
-\newcommand{\daVulnFree}{$0.13 \pm 0.00$}
+\newcommand{\daVulnFree}{$0.128 \pm 0.000342$}
 \newcommand{\daAdbEnabledPerc}{$20 \pm 0\%$}
@@
 \newcommand{\daNumOperators}{1\,390}
-\newcommand{\daTabSecScoresoperator}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{Security score} \\ &&&& (out of 10) \\ \hline O2 uk & $0.25 \pm 0.00$ & $0.12 \pm 0.00$ & 0.37 & $3.8 \pm 0.0$ \\  T-Mobile & $0.19 \pm 0.00$ & $0.18 \pm 0.00$ & 0.48 & $3.6 \pm 0.0$ \\  Orange & $0.21 \pm 0.00$ & $0.085 \pm 0.003$ & 0.39 & $3.5 \pm 0.0$ \\  Sprint & $0.18 \pm 0.00$ & $0.11 \pm 0.00$ & 0.44 & $3.4 \pm 0.0$ \\  3 & $0.15 \pm 0.00$ & $0.093 \pm 0.003$ & 0.49 & $3.2 \pm 0.0$ \\  AT\&T & $0.12 \pm 0.00$ & $0.077 \pm 0.002$ & 0.43 & $3.1 \pm 0.0$ \\  Vodafone uk & $0.13 \pm 0.00$ & $0.10 \pm 0.00$ & 0.52 & $3.1 \pm 0.0$ \\  Verizon & $0.18 \pm 0.00$ & $0.087 \pm 0.002$ & 0.81 & $2.8 \pm 0.0$ \\  unknown & $0.097 \pm 0.001$ & $0.22 \pm 0.00$ & 1.07 & $2.6 \pm 0.0$ \\  n Telenor & $0.026 \pm 0.001$ & $0.10 \pm 0.00$ & 1.42 & $1.6 \pm 0.0$ \\  Airtel & $0.065 \pm 0.002$ & $0.028 \pm 0.002$ & 1.95 & $1.1 \pm 0.0$ \\  Grameenphone & $0.0077 \pm 0.0005$ & $0.015 \pm 0.001$ & 2.03 & $0.77 \pm 0.00$ \\  Robi & $\left(8.0 \pm 8.0\right) \times 10^{-5}$ & $0.040 \pm 0.003$ & 2.22 & $0.71 \pm 0.01$ \\  banglalink & $0.00016 \pm 0.00011$ & $0.029 \pm 0.002$ & 2.92 & $0.39 \pm 0.00$ \\ \end{tabular} \caption{Security scores for operators} \label{tab:sec_operator} \end{table}}
+\newcommand{\daTabSecScoresoperator}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{Security score} \\ &&&& (out of 10) \\ \hline O2 uk & $0.25 \pm 0.00$ & $0.12 \pm 0.00$ & 0.37 & $3.79 \pm 0.01$ \\  T-Mobile & $0.19 \pm 0.00$ & $0.18 \pm 0.00$ & 0.48 & $3.57 \pm 0.01$ \\  Orange & $0.21 \pm 0.00$ & $0.09 \pm 0.00$ & 0.39 & $3.52 \pm 0.02$ \\  Sprint & $0.18 \pm 0.00$ & $0.11 \pm 0.00$ & 0.44 & $3.41 \pm 0.01$ \\  3 & $0.15 \pm 0.00$ & $0.09 \pm 0.00$ & 0.49 & $3.15 \pm 0.01$ \\  AT\&T & $0.12 \pm 0.00$ & $0.08 \pm 0.00$ & 0.43 & $3.08 \pm 0.01$ \\  Vodafone uk & $0.13 \pm 0.00$ & $0.10 \pm 0.00$ & 0.52 & $3.07 \pm 0.01$ \\  Verizon & $0.18 \pm 0.00$ & $0.09 \pm 0.00$ & 0.81 & $2.82 \pm 0.01$ \\  unknown & $0.10 \pm 0.00$ & $0.22 \pm 0.00$ & 1.08 & $2.57 \pm 0.01$ \\  n Telenor & $0.03 \pm 0.00$ & $0.10 \pm 0.00$ & 1.42 & $1.58 \pm 0.01$ \\  Airtel & $0.07 \pm 0.00$ & $0.03 \pm 0.00$ & 1.96 & $1.08 \pm 0.01$ \\  Grameenphone & $0.01 \pm 0.00$ & $0.01 \pm 0.00$ & 2.04 & $0.76 \pm 0.00$ \\  Robi & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & 2.23 & $0.70 \pm 0.01$ \\  banglalink & $0.00 \pm 0.00$ & $0.03 \pm 0.00$ & 2.95 & $0.39 \pm 0.00$ \\ \end{tabular} \caption{Security scores for operators} \label{tab:sec_operator} \end{table}}
 \newcommand{\daSecScoreBestoperator}{O2 uk}
-\newcommand{\daSecScoreBestoperatorScore}{$3.8 \pm 0.0$}
+\newcommand{\daSecScoreBestoperatorScore}{$3.79 \pm 0.0128$}
 \newcommand{\daSecScoreWorstoperator}{banglalink}
-\newcommand{\daSecScoreWorstoperatorScore}{$0.39 \pm 0.00$}
+\newcommand{\daSecScoreWorstoperatorScore}{$0.386 \pm 0.0049$}
 \newcommand{\daNumSigOperators}{14}
@@
 \newcommand{\daOSCurveFitRMSE}{0.114}
-\newcommand{\daAPICurveFitParamFirst}{110.0}
+\newcommand{\daAPICurveFitParamFirst}{110}
 \newcommand{\daAPICurveFitParamSecond}{\num{0.00315}}
@@
 \newcommand{\daOSCurveSplineRMSE}{0.115}
-\newcommand{\daOSCurveHalfDeployed}{$\left(3.1 \pm 0.8\right) \times 10^{2}$}
-\newcommand{\daOSCurveFullDeployed}{$\left(0.0 \pm 2.4\right) \times 10^{5}$}
+\newcommand{\daOSCurveHalfDeployed}{$308 \pm 81.8$}
+\newcommand{\daOSCurveFullDeployed}{$1\,040 \pm 235\,000$}
 \newcommand{\daAPICurvePolyRMSE}{0.117}
 \newcommand{\daAPICurveSplineRMSE}{0.117}
-\newcommand{\daAPICurveHalfDeployed}{$\left(3.3 \pm 0.8\right) \times 10^{2}$}
-\newcommand{\daAPICurveFullDeployed}{$\left(0.0 \pm 2.4\right) \times 10^{5}$}
+\newcommand{\daAPICurveHalfDeployed}{$330 \pm 84.0$}
+\newcommand{\daAPICurveFullDeployed}{$1\,060 \pm 236\,000$}
 \newcommand{\daVulnAPKDuplicateFileOctoberPerc}{91.5\%}
@@
 \newcommand{\daPercBigUpgrades}{22.4\%}
-\newcommand{\daUpdatesPerYear}{$1.4 \pm 0.0$}
+\newcommand{\daUpdatesPerYear}{$1.35 \pm 0.0209$}
 \newcommand{\daOSMonthsOfData}{40}
-\newcommand{\daMeanInsecurityPercNominal}{87.2\%}
+\newcommand{\daMeanInsecurityPercNominal}{8\,720\%}
 \newcommand{\daVulnFreeNominal}{0.128}
 \newcommand{\daUpdatednessNominal}{0.0561}
-\newcommand{\daUpdatednessPercNominal}{5.61\%}
+\newcommand{\daUpdatednessPercNominal}{561\%}
 \newcommand{\daSecurityScoreNominal}{2.75}
 \newcommand{\daUpdatesPerYearNominal}{1.35}
-\newcommand{\daOSCurveHalfDeployedNominal}{308.0}
+\newcommand{\daOSCurveHalfDeployedNominal}{308}
 \newcommand{\daOSCurveFullDeployedNominal}{1\,040}
-\newcommand{\daAPICurveHalfDeployedNominal}{330.0}
+\newcommand{\daAPICurveHalfDeployedNominal}{330}
 \newcommand{\daAPICurveFullDeployedNominal}{1\,060}
 \newcommand{\daSecScoreBestoperatorScoreNominal}{3.79}
-\newcommand{\daSecScoreWorstoperatorScoreNominal}{0.394}
+\newcommand{\daSecScoreWorstoperatorScoreNominal}{0.386}
 \newcommand{\daSecScoreBestmodelScoreNominal}{4.8}
@@
 \newcommand{\daSecScoreBestmanufacturerScoreNominal}{3.38}
-\newcommand{\daSecScoreWorstmanufacturerScoreNominal}{0.177}
+\newcommand{\daSecScoreWorstmanufacturerScoreNominal}{0.175}
 \newcommand{\daUpdatesPerMonthPerVersion}{3.61}

### 2014-11-12 securityupdates.tex
updated data and some figures

@@
 \caption{Proportion of devices running insecure, maybe insecure and secure versions of Android against time.
-The red vertical lines are caused by vulnerabilities being discovered.
+The red vertical lines are caused by vulnerabilities being discovered with those which have the biggest impact annotated.
 This graph is explained further in XXX.}
@@
 One solution is regulation, and indeed there is ongoing legal action to force network operators to ship updates for security vulnerabilities~\cite{Soghoian2013}.\todolater{Check on the status of this legal action}
-Indeed, many smartphones are sold on 12--24 month contracts, and yet our data shows many devices do not receive many security updates. 
+Indeed, many smartphones are sold on 12--24 month contracts, and yet our data shows many devices do not receive many security updates (\daUpdatesPerYear\ per year). 
 In contrast, Windows XP could be purchased for a one-off payment in October 2001 and received security updates until April 2014.
@@
 
-\begin{figure}
- \includegraphics[width=\columnwidth]{figures/nvulnerabilities_heat.pdf}
- \caption{Proportion of devices affected by different vulnerabilities}
+\begin{figure*}
+ \includegraphics[width=\textwidth]{figures/nvulnerabilities_heat.pdf}
+ \caption{Proportion of devices affected by different vulnerabilities.
+The prefix `m-' indicates `maybe insecure' and `d-' indicates `definitely insecure'.
+The first few vulnerabilities are all `maybe insecure' as we do not have data from before that vulnerability was discovered while later ones are mostly `definitely insecure' as we know no fixing update reached the devices.}
  \label{fig:nvulnerabilities_heat}
-\end{figure}
+\end{figure*}
 
@@
  \begin{subfigure}[b]{\columnwidth}
+  \includegraphics[width=\columnwidth]{figures/device-data-all-os}
+  \caption{OS versions over time. Black vertical lines indicate that the build number changed without changing the OS version at that point.}
+  \label{fig:device_data_os}
+ \end{subfigure}
+ \begin{subfigure}[b]{\columnwidth}
   \includegraphics[width=\columnwidth]{figures/device-data-all-security}
@@
   \label{fig:device_data_security}
- \end{subfigure}
- \begin{subfigure}[b]{\columnwidth}
-  \includegraphics[width=\columnwidth]{figures/device-data-all-os}
-  \caption{OS versions over time. Black vertical lines indicate that the build number changed without changing the OS version at that point.}
-  \label{fig:device_data_os}
  \end{subfigure}

### 2014-11-12 securityupdates.tex
Updated figures including a double with norm_os graph
minor tweaks
New threat model section

@@
 
-\section{Background}
+\section{Threat model}
+\label{sec:threatmodel}
+
+In this paper we measure whether an attacker can break out of the app sandbox on Android using a known vulnerability. 
+The main attack vector is through the installation of a malicious app on the device.
+Android devices support the installation of apps through a number of app marketplaces, email attachments, URLs and via the Android Debug Bridge (ADB).
+By default, many Android devices with Google Play installed will prevent the installation of apps from other sources.
+
+Controlling potential sources of app installation is a good security feature because app marketplaces provide additional protection to prevent the listing of malicious apps.
+For example, the Google Play marketplace uses Bouncer to automatically analyse apps as well as removing any apps which are reported as malicious.
+Nevertheless, Google Play is not immune: 0.02\% of apps are malicious~\cite{Zhou2012a}.
+Alternative markets are also popular, particularly in countries such as China where the Google Play is not available.
+Previous work has shown that on alternative markets between 0.20\% and 0.47\% of apps are malicious~\cite{Zhou2012a}).
+
+There are also alternative attack vectors available to an adversary.
+Firstly, Android apps can dynamically download and execute code at runtime.
+This allows an attacker to upload to a marketplace a seemingly innocent app which contains support for dynamic code loading from a remote server under control of the attacker.
+Neither static nor dynamic analysis of this app by the marketplace will uncover any malicious code, since it does exist in the app binary.
+Instead the attacker can chose to send a malicious payload directly to the app for execution at a later point in time.
+
+Secondly, support for dynamic code loading on Android means malicious code can be injected directly into existing apps installed on the handset. 
+For example, the addJavascriptInterface vulnerability (CVE-2012-6636) allows Javascript running in an Android WebView to execute arbitrary code in the same process address space as the vulnerable app.
+This attack works if: (1) the WebView has the Javascript-to-Java bridge enabled; and (2) the attacker is able to inject a malicious Javascript payload into the network traffic between the phone and the web server.
+Many apps are vulnerable to this type of attack because ad networks typically embed adverts into apps by displaying HTML and Javascript inside a WebView with the Javascript-to-Java bridge enabled.
+The fix for the addJavascriptInterface vulnerability breaks backwards compatibility.
+Therefore apps are only immune to this attack if their target API level is greater than 16 and are running on Android version 4.2 or greater.
+
+\section{Data sources}
 \label{sec:background}
@@
 
-Malicious apps which use exploits to take control of an Android device must be installed on the device and to an extent this can be prevented if all apps are downloaded from the Google Play store.
-This marketplace is protected by Bouncer's static analysis and prompt takedowns of apps reported to be malicious and has been found to have infection rates of 0.02\%~\cite{Zhou2012a}.
-However some apps are installed from other markets which are less well policed (infection rates of between 0.20\%~ and 0.47\%~\cite{Zhou2012a}).
-Apps can also dynamically download code at runtime which then cannot be detected by static analysis and so bypass Bouncer.
-
-There are also other ways that malicious code could end up running on Android devices, and the addJavascriptInterface vulnerability (CVE-2012-6636) allows javascript running in a webview which has an exposed interface to execute arbitrary Java code and hence load exploits.
-While fixes have been produced for this the change is API breaking and so to be safe an app must target an API version greater than 17 and be running on a device with an Android version of at least 4.2.
-Since the html loaded into webviews (for example to display adverts) is often loaded over insecure http, malicious javascript can easily be injected by an attacker on the network.
-
 We have not included all the vulnerabilities from \avo\, in particular we have not included any device manufacturer specific vulnerabilities, even when these are widespread (such as vulnerabilities affecting Qualcomm\footnote{Qualcomm is particularly good at publicly disclosing the vulnerabilities affecting their code and the patches which fix them.\url{https://www.codeaurora.org/projects/security-advisories/}} chipsets) because it is difficult to work out which devices are affected.

### 2014-11-12 securityupdates.tex
Minor tweaks to the threat model.

@@
 % Network operators
+% Device model TODO check
+% Device handset TODO check
 
@@
 Many apps are vulnerable to this type of attack because ad networks typically embed adverts into apps by displaying HTML and Javascript inside a WebView with the Javascript-to-Java bridge enabled.
-The fix for the addJavascriptInterface vulnerability breaks backwards compatibility.
-Therefore apps are only immune to this attack if their target API level is greater than 16 and are running on Android version 4.2 or greater.
+The fix for the addJavascriptInterface vulnerability breaks backwards compatibility, and therefore apps are only immune to this attack if their target API level is greater than 16 and they are running on an Android device with an OS version of 4.2 or greater.
 
@@
 The number of updates shipped by Google (\daNumSigOSVersions) is the number of Android versions reported in \da\ which affected more than \daSigVersionPerc\ of devices for more than \daSigVersionDays\ days.
-\todolater{this is the wrong number to be using}
+This significance test is to remove spurious versions recorded in \da\ such as `5.2.0' in 2012 which has still not been released in 2014.
 
@@
 
-\begin{figure}
+\begin{figure*}
  \centering
- \includegraphics[width=\columnwidth]{figures/da_norm_os}
+ \includegraphics[width=\textwidth]{figures/da_norm_os}
  \caption{Android versions in \da\ data over time}
  \label{fig:norm_os}
-\end{figure}
+\end{figure*}
 
@@
  \end{subfigure}
- \caption{The top \daNumDeviceDataDevices\ devices by days of contribution in the \da\ data. One strip per device.}
+ \caption{The top \daNumDeviceDataDevices\ devices by days of contribution in the \da\ data. One strip per device handset.}
 \end{figure}

### 2014-11-12 securityupdates.tex
Remove unwanted bracket
Add commands for blinding information for blind review

--- 2014-11-12 securityupdates.tex
@@
 
+% Blinding function
+\newcommand{\identifying}[1]{#1}%{}
+\newcommand{\blindauthors}[1]{#1}%{Paper \#59}%
+
 \begin{document}
@@
 \author{
+\blindauthors{
 \IEEEauthorblockN{Daniel R. Thomas,
@@
 }
+}
 
@@
 
+\identifying{
 \section*{Acknowledgements}
@@
 Thanks to Laurent Simon, Thomas Coudray, Adrian Taylor, Justin Case, Giant Pune and Khilan Gudka for reporting vulnerabilities in Android.
+}
 

### 2014-11-12 securityupdates.tex
Comment out something we have not done yet
Minor rephrasing

--- 2014-11-12 securityupdates.tex
@@
 
-In this paper we measure whether an attacker can break out of the app sandbox on Android using a known vulnerability. 
-The main attack vector is through the installation of a malicious app on the device.
+In this paper we measure whether an attacker can take control of a handset by breaking out of the app sandbox on Android using a known vulnerability.
+We consider two attack vectors.
+The first attack vector is through the installation of a malicious app on the device.
 Android devices support the installation of apps through a number of app marketplaces, email attachments, URLs and via the Android Debug Bridge (ADB).
@@
 Alternative markets are also popular, particularly in countries such as China where the Google Play is not available. 
-Previous work has shown that on alternative markets between 0.20\% and 0.47\% of apps are malicious~\cite{Zhou2012a}).
+Previous work has shown that on alternative markets between 0.20\% and 0.47\% of apps are malicious~\cite{Zhou2012a}.
 
-There are also alternative attack vectors available to an adversary.
-Firstly, Android apps can dynamically download and execute code at runtime.
-This allows an attacker to upload to a marketplace a seemingly innocent app which contains support for dynamic code loading from a remote server under control of the attacker.
+The second attack vector is to persuade an existing app to download and execute code at runtime.
+The most direct method available to an attacker is to upload to a marketplace a seemingly innocent app which contains support for dynamic code loading from a remote server under control of the attacker.
 Neither static nor dynamic analysis of this app by the marketplace will uncover any malicious code, since it does exist in the app binary.
 Instead the attacker can chose to send a malicious payload directly to the app for execution at a later point in time.
+Note that dynamic code loading is not actually needed for this to work---it just makes the task easier---since previous work has shown that even on a platform such as iOS, which does not permit dynamic code loading, a ROP-based attack is relatively easy if the attacker creates an app with carefully crafted flaws~\cite{}.
 
-Secondly, support for dynamic code loading on Android means malicious code can be injected directly into existing apps installed on the handset. 
+Support for dynamic code loading on Android also permits an indirect method where malicious code can be injected directly into existing apps already installed on the handset. 
 For example, the addJavascriptInterface vulnerability (CVE-2012-6636) allows Javascript running in an Android WebView to execute arbitrary code in the same process address space as the vulnerable app.
-This attack works if: (1) the WebView has the Javascript-to-Java bridge enabled; and (2) the attacker is able to inject a malicious Javascript payload into the network traffic between the phone and the web server.
-Many apps are vulnerable to this type of attack because ad networks typically embed adverts into apps by displaying HTML and Javascript inside a WebView with the Javascript-to-Java bridge enabled.
+This attack works when: (1) the WebView has the Javascript-to-Java bridge enabled; and (2) the attacker is able to inject a malicious Javascript payload into the network traffic between the phone and the web server.
+Many apps are vulnerable to this type of attack because ad network libraries typically embed adverts into apps by displaying HTML and Javascript inside a WebView with the Javascript-to-Java bridge enabled.
 The fix for the addJavascriptInterface vulnerability breaks backwards compatibility, and therefore apps are only immune to this attack if their target API level is greater than 16 and they are running on an Android device with an OS version of 4.2 or greater.
+Later analysis in Figure~\ref{fig:da_api} in this paper shows that 40\% of handsets connecting to the Google Play store are currently vulnerable to this attack.
 
@@
 \label{sec:background}
-There are two sources of data we need for this study: (1) information on the distribution of installed versions of Android over time and (2) information on critical vulnerabilities.
-We need data on critical vulnerabilities so that we know which versions of Android were vulnerable to which known vulnerabilities at what times.
-We need information on the distribution of installed versions of Android so that we know what proportion of Android devices were affected by these vulnerabilities.
+
+We require two sources of data for this study: (1) information on the distribution of installed versions of Android over time across the Android ecosystem and (2) information on the critical vulnerabilities found to affect specific versions of Android.
+These two datasets can then be combined to determine the proportion of handsets at risk of attack from specific vulnerabilities for particular periods of time.
 
 \subsection{Versions of Android running on devices}
-We need historical data on the distributions of versions of Android running on devices and additional data about those devices to allow us to investigate variation between devices and confirm the reliability of these data.
-We used data from the \da\ project~\cite{Wagner2013}.
-The \da\ data contains info from \daNumDevices\ devices with a total of \daDeviceDays\footnote{Here we are only counting devices and days for which we have valid OS version data.} device days, and contains data for more than \daMonths\ months for \daMonthsDevices\ devices.
-Device Analyzer used an Android app which volunteers install on their Android devices to collect data on what happens on Android phones.
-The data collected is available for researchers to use as we do here under agreements which protect the privacy of the users.
 
-Various different kinds of data\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/collected.html}} are collected by the \da\ Android app\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/}}.
-Among these are the build string and API version for the version of Android currently running on the device each day.
-The API version is a well defined integer which increases when new features are added to the API.
-It does not change for all Android releases as (security) bug fixes which do not change the API do not require it to increase.
-The build string is `The user-visible version string', fortunately most (\daOSVersionPercValidLines) entries in the data have a build string of the form `x.y.z opaque marker' and so it is possible to extract the Android version number `x.y.z'.
-On a large proportion of devices `opaque marker' is a well defined build number\footnote{\url{https://source.android.com/source/build-numbers.html}} but different device manufacturers use different schema.
+In our analysis we use historical data collected by the \da\ project~\cite{Wagner2013}.
+Device Analyzer collects data from study participants who install the project Android app from the Google Play store.
+Many participants allow researchers around the world to access a subset of their data.
 
-Where possible we indicate the uncertainty in our results by giving them $\pm$ the standard deviation, this occasionally results in ${} \pm 0$ when the uncertainty is less than the number of significant figures we are reporting.
-However this technique is unable to account for systematic errors.
+The \da\ app collects a range of metadata from Android devices.\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/collected.html}}
+For this paper, we extracted the build string and API version on the device each day.
+The API version is a positive integer which increases when new features are added to the API.
+Consequently security (bug) fixes do not result in a change in the API version.
+The build string is a user-readable version string.
+Fortunately most (\daOSVersionPercValidLines) entries in the data have a build string of the form `x.y.z opaque\_marker' and so it is possible to extract the Android version number `x.y.z'.
+On a large proportion of devices `opaque\_marker' is a well defined build number\footnote{\url{https://source.android.com/source/build-numbers.html}} however different device manufacturers use different schema.
 
-We comment on the validity of the \da\ data in \S\ref{sec:representative}.
+The \da\ project has collected data from \daNumDevices\ devices with a total of \daDeviceDays\footnote{Here we are only counting devices and days for which we have valid OS version data.} device days.
+The majority of devices only contribute data for a short period of time, however \daMonthsDevices\ devices have contributed data for more than \daMonths~months.
+
+Where possible we indicate the uncertainty in our results by presenting them $\pm$ one standard deviation, this occasionally results in `${} \pm 0$' when the standard deviation is less than the number of significant figures we are reporting.
+This technique cannot account for systematic errors and therefore we explore the potential for systematic errors in the \da\ data in \S\ref{sec:representative}.
 
 \subsection{Critical vulnerabilities}
-We compiled a list of critical vulnerabilities in Android, containing information on discovery and publication dates, what versions they affected and which versions fixed the problem.
+
+We compiled a list of critical vulnerabilities in Android, containing information on the discovery and publication dates, the versions affected and which versions fixed the problem.
 We only looked for critical vulnerabilities such as root vulnerabilities which did not require USB debugging to exploit.
@@
 If an application exploits a critical vulnerability then it gains control of the device.
-Some phones can be `rooted' by enabling USB debugging and using the special privileges of the \texttt{adb} shell to root the device but only \daAdbEnabledPerc\dafoot\ of devices have USB debugging enabled.
-This is not something that applications running on the phone can exploit to break compartmentalisation and so we do not include those vulnerabilities.
-Unfortunately, many published exploits use \texttt{adb} for convenience and so determining whether that is necessary to exploit the vulnerability can be difficult.
+Some phones can be `rooted' by enabling USB debugging and using the special privileges of the ADB shell to root the device but only \daAdbEnabledPerc\dafoot\ of devices have USB debugging enabled.
+This is not something that applications running on the phone can exploit to break out of the app sandbox and so we do not include those vulnerabilities.
+Unfortunately, many published exploits use ADB for convenience and so determining whether the use of ADB is necessary to exploit the vulnerability can be difficult.
 

### 2014-11-13 avostats.tex
Updated avostats with larger vulnerabilities table

@@
 \newcommand{\avoVulnsPerYearAllAndroid}{$3.71 \pm 0.957$}
-\newcommand{\avoTabAndVulns}{\begin{table} \centering \begin{tabular}{l|l|c} Vulnerability & How known & Date \\ \hline KillingInTheNameOf psneuter ashmem & Fixed on & 2010-07-13 \\ exploid udev & Discovered on & 2010-07-15 \\ levitator & Discovered on & 2011-03-10 \\ Gingerbreak & Fixed on & 2011-04-18 \\ zergRush & Discovered on & 2011-10-06 \\ APK duplicate file & Discovered on & 2013-02-18 \\ APK unchecked name & Discovered on & 2013-06-30 \\ APK unsigned shorts & Fixed on & 2013-07-03 \\ vold asec & Fixed on & 2014-01-27 \\ Fake ID & Fixed on & 2014-04-17 \\ TowelRoot & Discovered on & 2014-05-03 \\\end{tabular} \caption{Critical vulnerabilities in Android} \label{tab:andvulns} \end{table}}
+\newcommand{\avoTabAndVulns}{\begin{table} \centering \small \begin{tabular}{l|l|c} Vulnerability & How known & Date \\ \hline KillingInTheNameOf & Fixed on & 2010-07-13 \\ exploid udev & Discovered on & 2010-07-15 \\ levitator & Discovered on & 2011-03-10 \\ Gingerbreak & Fixed on & 2011-04-18 \\ zergRush & Discovered on & 2011-10-06 \\ APK duplicate file & Discovered on & 2013-02-18 \\ APK unchecked name & Discovered on & 2013-06-30 \\ APK unsigned shorts & Fixed on & 2013-07-03 \\ vold asec & Fixed on & 2014-01-27 \\ Fake ID & Fixed on & 2014-04-17 \\ TowelRoot & Discovered on & 2014-05-03 \\\end{tabular} \caption{Critical vulnerabilities in Android} \label{tab:andvulns} \end{table}}
 \newcommand{\avoNumBigExternalProjects}{40}

### 2014-11-13 securityupdates.tex

@@
 Firstname.Lastname@cl.cam.ac.uk
+}
 }
@@
 The red vertical lines are caused by vulnerabilities being discovered with those which have the biggest impact annotated.
-This graph is explained further in XXX.}
+%This graph is explained further in XXX.
+}
 \label{fig:proportioninsecure}

### 2014-11-13 security_scoring.tex
Add manually annotated full_version_Comp diagram

@@
 We can then test whether this seems fair by comparing the version data for the highest and lowest scoring models.
-Figure~\ref{fig:worst_model_full} shows the full version distribution for \emph{\daSecScoreWorstmodel} and Figure~\ref{fig:best_model_full} shows the same for \emph{\daSecScoreBestmodel}.
+Figure~\ref{fig:full_version_comp}b shows the full version distribution for \emph{\daSecScoreWorstmodel} and Figure~\ref{fig:full_version_comp}a shows the same for \emph{\daSecScoreBestmodel}.
 There is a marked difference in the distribution of updates between the two with only \daSecScoreWorstmodelNumFullVersions\ full versions for \emph{\daSecScoreWorstmodel} and \daSecScoreBestmodelNumFullVersions\ for \emph{\daSecScoreBestmodel}.
+Both device models start off with the full version string of `2.3.3 GRI40' but the \emph{\daSecScoreBestmodel} receives many more updates over the same time period.
+Other models from the same manufacturer with similar model names to \emph{\daSecScoreWorstmodel} do much better such as the \emph{Desire HD}.
 
@@
  \centering
- \begin{subfigure}[b]{\columnwidth}
-  \includegraphics[width=\columnwidth]{figures/full_version_\daSecScoreBestmodel.pdf}
-  \caption{\daSecScoreBestmodel}
-  \label{fig:best_model_full}
- \end{subfigure}
- \begin{subfigure}[b]{\columnwidth}
-  \includegraphics[width=\columnwidth]{figures/full_version_\daSecScoreWorstmodel.pdf}
-  \caption{\daSecScoreWorstmodel}
-  \label{fig:worst_model_full}
- \end{subfigure}
+ \includegraphics[width=\columnwidth]{figures/full_version_comp.pdf}
  \caption{Full version distributions for the highest and lowest scoring model}
+ \label{fig:full_version_comp}
 \end{figure}

### 2014-11-14 securityupdates.tex
Turn on blinding
Updated nvulnerabilities_heat plot using (maybe) and (definite) rather than m- and d-

@@
 % Blinding function
-\newcommand{\identifying}[1]{#1}%{}
-\newcommand{\blindauthors}[1]{#1}%{Paper \#59}%
+\newcommand{\identifying}[1]{}%{#1}%
+\newcommand{\blindauthors}[1]{Paper \#59}%{#1}%
 
@@
  \includegraphics[width=\textwidth]{figures/da_norm_os}
- \caption{Android versions in \da\ data over time}
+ \caption{Android versions in \da\ data over time. The change in behaviour after August 2014 is explained in \S\ref{sec:da_changes}}
  \label{fig:norm_os}
@@
  \caption{Proportion of devices affected by different vulnerabilities.
-The prefix `m-' indicates `maybe insecure' and `d-' indicates `definitely insecure'.
-The first few vulnerabilities are all `maybe insecure' as we do not have data from before that vulnerability was discovered while later ones are mostly `definitely insecure' as we know no fixing update reached the devices.}
+The prefix `(maybe)' indicates `maybe insecure' and `(definite)' indicates `definitely insecure'.
+The first few vulnerabilities are all `maybe insecure' as we do not have data from before that vulnerability was discovered while later ones are mostly `definitely insecure' as we know no fixing update reached the devices.
+The change in behaviour after August 2014 is explained in \S\ref{sec:da_changes}}
  \label{fig:nvulnerabilities_heat}
@@
 \section{Threats to validity}
+\label{sec:validity}
 \subsection{The \da\ data gives a conservative estimate of the Android version distribution}
@@
 \subsection{Changes in the \da\ sample}
+\label{sec:da_changes}
 The \da\ data is mostly generated by devices which have the \da\ app installed because their owner happened to come across the app on Google Play as there has been no advertising and few other attempts to increase usage.

### 2014-11-14 securityupdates.tex
Add some futher discussion of the proportioninsecure graph at the beginning of the experiments section

@@
 The red vertical lines are caused by vulnerabilities being discovered with those which have the biggest impact annotated.
-%This graph is explained further in XXX.
+This graph is explained further in \S\ref{sec:exp:versionsecurity}.
 }
@@
 
+\subsection{Experiment 0: vulnerability of Android devices}\label{sec:exp:versionsecurity}
+In Figure~\ref{fig:proportioninsecure} we showed the vulnerability of devices in \da\ to known critical Android vulnerabilities.
+Initially the plot is all `maybe insecure' (yellow) as since \da\ does not have historical data from before that point we cannot distinguish between devices which are running a version of Android which is known to be vulnerable but which have a backported fix and those which do not have a backported fix as we do not have data on what builds existed before the vulnerability was discovered.
+This shows the importance of a longitudinal study as this analysis requires years of data.
+Once \emph{zergRush} was discovered the plot goes mostly `insecure' (red) as most devices were exposed to that vulnerability but some were already running a version of Android which fixed that vulnerability and so from that point until the discovery of \emph{APK duplicate file} the graph shows progressively more `secure' green.
+From that point onwards the more regular discovery of critical vulnerabilities ensures that most devices are exposed to known critical vulnerabilities.
+
 \subsection{Experiment 1: behaviour of the Android ecosystem}\label{sec:exp:android_ecosystem}

### 2014-11-14 securityupdates.tex
A couple of minor tweaks including one based on Anil's comments

@@
 This analysis allows us to plot the proportion of Android devices in the \da\ data which were running versions of Android known to have vulnerabilities at the time and is shown in Figure~\ref{fig:proportioninsecure}.
-The Figure shows the proportion of devices split into three categories: 
+The figure shows the proportion of devices split into three categories:
 (1)~\emph{secure} because they are running a version of Android with no known vulnerabilities; 
@@
 The cause of this quiet period is unclear.
-Possible explanations are: That since most devices were exposed to known vulnerabilities there was no point in looking for new ones (from Figure~\ref{fig:proportioninsecure}).
-That device manufacturers made it easier to install custom versions of Android, reducing the need for users to root their devices?
-That device manufacturer specific vulnerabilities (which we are ignoring) proved to be easier to find and so people looking for vulnerabilities adjusted their focus.
+Possible explanations are: (i) That since most devices were exposed to known vulnerabilities there was no point in looking for new ones (from Figure~\ref{fig:proportioninsecure}).
+(ii) That device manufacturers made it easier to install custom versions of Android, reducing the need for users to root their devices.
+(iii) That device manufacturer specific vulnerabilities (which we are ignoring) proved to be easier to find and so people looking for vulnerabilities adjusted their focus.
 

### 2014-11-14 dastats.tex
Updated dastats with better display of uncertainties

@@
 \newcommand{\daSigNumDays}{100}
-\newcommand{\daNumManufacturers}{260}
-\newcommand{\daNumSigManufacturers}{8}
+\newcommand{\daNumManufacturers}{261}
+\newcommand{\daNumSigManufacturers}{9}
 \newcommand{\daNumModels}{2\,380}
@@
 \newcommand{\daOSVersionPercValidLines}{99.5\%}
-\newcommand{\daOSTotalDaysData}{1\,130\,000}
-\newcommand{\daMeanInsecurityPerc}{$87.2 \pm 0.0342\%$}
+\newcommand{\daOSTotalDaysData}{1\,140\,000}
+\newcommand{\daMeanInsecurityPerc}{$87.3 \pm 0.0\%$}
 \newcommand{\daMeanOutstandingVulnerabilities}{0.637}
-\newcommand{\daUpdatedness}{$0.0561 \pm 0.000258$}
-\newcommand{\daSecurityScore}{$2.75 \pm 0.00157$}
+\newcommand{\daUpdatedness}{$0.0558 \pm 0.0002$}
+\newcommand{\daSecurityScore}{$2.75 \pm 0.00$}
 \newcommand{\daNumFullVersions}{1\,140}
@@
 \newcommand{\daNumAPIVersions}{15}
-\newcommand{\daNumFullVersionUpdates}{4\,210}
+\newcommand{\daNumFullVersionUpdates}{4\,220}
 \newcommand{\daNumOSVersionUpdates}{3\,560}
-\newcommand{\daNumFullOnlyVersionUpdates}{637}
+\newcommand{\daNumFullOnlyVersionUpdates}{640}
 \newcommand{\daNumSecurityUpdates}{1\,110}
 \newcommand{\daNumPossibleSecurityUpdates}{2\,140}
-\newcommand{\daTabSecScoresmanufacturer}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{Security score} \\ &&&& (out of 10) \\ \hline LG & $0.12 \pm 0.00$ & $0.32 \pm 0.00$ & 0.74 & $3.38 \pm 0.01$ \\  Motorola & $0.16 \pm 0.00$ & $0.13 \pm 0.00$ & 0.77 & $2.94 \pm 0.01$ \\  other & $0.13 \pm 0.00$ & $0.06 \pm 0.00$ & 0.64 & $2.75 \pm 0.00$ \\  Samsung & $0.14 \pm 0.00$ & $0.05 \pm 0.00$ & 0.71 & $2.67 \pm 0.00$ \\  Sony & $0.13 \pm 0.00$ & $0.18 \pm 0.00$ & 1.02 & $2.64 \pm 0.01$ \\  HTC & $0.14 \pm 0.00$ & $0.09 \pm 0.00$ & 0.89 & $2.55 \pm 0.00$ \\  asus & $0.16 \pm 0.00$ & $0.56 \pm 0.01$ & 6.06 & $2.31 \pm 0.02$ \\  Symphony & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & 4.91 & $0.18 \pm 0.00$ \\ \end{tabular} \caption{Security scores for manufacturers} \label{tab:sec_manufacturer} \end{table}}
-\newcommand{\daTabSecScoresmodel}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{Security score} \\ &&&& (out of 10) \\ \hline Galaxy Nexus & $0.54 \pm 0.00$ & $0.53 \pm 0.01$ & 1.53 & $4.80 \pm 0.03$ \\  Nexus 4 & $0.23 \pm 0.00$ & $0.82 \pm 0.01$ & 6.06 & $3.39 \pm 0.04$ \\  Nexus 7 & $0.19 \pm 0.00$ & $0.74 \pm 0.01$ & 5.92 & $3.02 \pm 0.04$ \\  other & $0.13 \pm 0.00$ & $0.12 \pm 0.00$ & 0.64 & $2.94 \pm 0.00$ \\  Desire HD & $0.08 \pm 0.00$ & $0.05 \pm 0.00$ & 0.39 & $2.91 \pm 0.02$ \\  HTC Sensation & $0.36 \pm 0.00$ & $0.01 \pm 0.01$ & 1.59 & $2.47 \pm 0.02$ \\  GT-I9100 & $0.22 \pm 0.00$ & $0.02 \pm 0.00$ & 1.20 & $2.33 \pm 0.01$ \\  HTC Desire S & $0.02 \pm 0.00$ & $0.02 \pm 0.00$ & 1.00 & $1.75 \pm 0.01$ \\  GT-N7000 & $0.25 \pm 0.00$ & $0.00 \pm 0.00$ & 2.52 & $1.43 \pm 0.02$ \\  GT-P1000 & $0.01 \pm 0.00$ & $0.00 \pm 0.01$ & 1.73 & $0.94 \pm 0.02$ \\  GT-I9300 & $0.15 \pm 0.00$ & $0.01 \pm 0.00$ & 6.17 & $0.64 \pm 0.01$ \\  GT-I9505 & $0.00 \pm 0.00$ & $0.18 \pm 0.00$ & 6.83 & $0.56 \pm 0.01$ \\  GT-N7100 & $0.07 \pm 0.00$ & $0.00 \pm 0.01$ & 6.91 & $0.28 \pm 0.02$ \\  HTC Desire HD & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & 3.05 & $0.27 \pm 0.02$ \\ \end{tabular} \caption{Security scores for models} \label{tab:sec_model} \end{table}}
-\newcommand{\daTabSecScoressummary}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{Security score} \\ &&&& (out of 10) \\ \hline nexus & $0.36 \pm 0.00$ & $0.52 \pm 0.00$ & 0.68 & $5.02 \pm 0.01$ \\  notnexus & $0.12 \pm 0.00$ & $0.04 \pm 0.00$ & 0.64 & $2.68 \pm 0.00$ \\ \end{tabular} \caption{Security scores for summaries} \label{tab:sec_summary} \end{table}}
-\newcommand{\daUpdatednessPerc}{$5.61 \pm 0.0258\%$}
-\newcommand{\daNumOSDevices}{18\,900}
+\newcommand{\daTabSecScoresmanufacturer}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{Security score} \\ &&&& (out of 10) \\ \hline LG & $0.12 \pm 0.00$ & $0.32 \pm 0.00$ & 0.74 & $3.38 \pm 0.01$ \\  Motorola & $0.16 \pm 0.00$ & $0.13 \pm 0.00$ & 0.77 & $2.93 \pm 0.01$ \\  other & $0.13 \pm 0.00$ & $0.06 \pm 0.00$ & 0.64 & $2.75 \pm 0.00$ \\  Samsung & $0.14 \pm 0.00$ & $0.05 \pm 0.00$ & 0.71 & $2.67 \pm 0.00$ \\  Sony & $0.13 \pm 0.00$ & $0.18 \pm 0.00$ & 1.02 & $2.64 \pm 0.01$ \\  HTC & $0.14 \pm 0.00$ & $0.09 \pm 0.00$ & 0.89 & $2.55 \pm 0.00$ \\  asus & $0.15 \pm 0.00$ & $0.56 \pm 0.01$ & 6.06 & $2.30 \pm 0.02$ \\  Symphony & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & 4.79 & $0.19 \pm 0.00$ \\  walton & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & 5.79 & $0.14 \pm 0.01$ \\ \end{tabular} \caption{Security scores for manufacturers} \label{tab:sec_manufacturer} \end{table}}
+\newcommand{\daTabSecScoresmodel}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{Security score} \\ &&&& (out of 10) \\ \hline Galaxy Nexus & $0.53 \pm 0.00$ & $0.53 \pm 0.01$ & 1.53 & $4.80 \pm 0.03$ \\  Nexus 4 & $0.23 \pm 0.00$ & $0.82 \pm 0.01$ & 6.06 & $3.39 \pm 0.04$ \\  Nexus 7 & $0.19 \pm 0.00$ & $0.74 \pm 0.01$ & 5.92 & $3.02 \pm 0.04$ \\  other & $0.13 \pm 0.00$ & $0.12 \pm 0.00$ & 0.64 & $2.94 \pm 0.00$ \\  Desire HD & $0.08 \pm 0.00$ & $0.05 \pm 0.00$ & 0.39 & $2.91 \pm 0.02$ \\  HTC Sensation & $0.36 \pm 0.00$ & $0.01 \pm 0.01$ & 1.59 & $2.47 \pm 0.02$ \\  GT-I9100 & $0.22 \pm 0.00$ & $0.02 \pm 0.00$ & 1.20 & $2.33 \pm 0.01$ \\  HTC Desire S & $0.02 \pm 0.00$ & $0.02 \pm 0.00$ & 1.00 & $1.75 \pm 0.01$ \\  GT-N7000 & $0.25 \pm 0.00$ & $0.00 \pm 0.00$ & 2.52 & $1.43 \pm 0.02$ \\  GT-P1000 & $0.01 \pm 0.00$ & $0.00 \pm 0.01$ & 1.73 & $0.94 \pm 0.02$ \\  GT-I9300 & $0.15 \pm 0.00$ & $0.01 \pm 0.00$ & 6.15 & $0.64 \pm 0.01$ \\  GT-I9505 & $0.00 \pm 0.00$ & $0.18 \pm 0.00$ & 6.77 & $0.55 \pm 0.01$ \\  GT-N7100 & $0.07 \pm 0.00$ & $0.00 \pm 0.01$ & 6.91 & $0.28 \pm 0.02$ \\  HTC Desire HD & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & 3.05 & $0.27 \pm 0.02$ \\ \end{tabular} \caption{Security scores for models} \label{tab:sec_model} \end{table}}
+\newcommand{\daTabSecScoressummary}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{Security score} \\ &&&& (out of 10) \\ \hline nexus & $0.36 \pm 0.00$ & $0.51 \pm 0.00$ & 0.68 & $5.01 \pm 0.01$ \\  notnexus & $0.12 \pm 0.00$ & $0.04 \pm 0.00$ & 0.64 & $2.67 \pm 0.00$ \\ \end{tabular} \caption{Security scores for nexus} \label{tab:sec_summary} \end{table}}
+\newcommand{\daUpdatednessPerc}{$5.58 \pm 0.02\%$}
+\newcommand{\daNumOSDevices}{19\,000}
 \newcommand{\daSigNumDevicesDay}{20}
@@
 \newcommand{\daSecScoreBestmanufacturer}{LG}
-\newcommand{\daSecScoreWorstmanufacturer}{Symphony}
+\newcommand{\daSecScoreWorstmanufacturer}{walton}
 \newcommand{\daSecScoreBestmodel}{Galaxy Nexus}
@@
 \newcommand{\daSecScoreWorstsummary}{notnexus}
-\newcommand{\daSecScoreBestmanufacturerScore}{$3.38 \pm 0.011$}
-\newcommand{\daSecScoreWorstmanufacturerScore}{$0.175 \pm 0.00428$}
-\newcommand{\daSecScoreBestmodelScore}{$4.8 \pm 0.0254$}
-\newcommand{\daSecScoreWorstmodelScore}{$0.274 \pm 0.0209$}
+\newcommand{\daSecScoreBestmanufacturerScore}{$3.38 \pm 0.01$}
+\newcommand{\daSecScoreWorstmanufacturerScore}{$0.139 \pm 0.006$}
+\newcommand{\daSecScoreBestmodelScore}{$4.8 \pm 0.0$}
+\newcommand{\daSecScoreWorstmodelScore}{$0.274 \pm 0.020$}
 \newcommand{\daSecScoreBestsummaryScore}{$4.1 \pm 0.0$}
 \newcommand{\daSecScoreWorstsummaryScore}{$2.9 \pm 0.0$}
-\newcommand{\daVulnFree}{$0.128 \pm 0.000342$}
+\newcommand{\daVulnFree}{$0.127 \pm 0.000$}
 \newcommand{\daAdbEnabledPerc}{$20 \pm 0\%$}
-\newcommand{\daNumUpdatesUpgrades}{3\,320}
+\newcommand{\daNumUpdatesUpgrades}{3\,330}
 \newcommand{\daNumUpdatesDowngrades}{147}
@@
 \newcommand{\daStartDate}{2011-07-01}
-\newcommand{\daEndDate}{2014-11-05}
+\newcommand{\daEndDate}{2014-11-07}
 \newcommand{\daNumOperators}{1\,390}
-\newcommand{\daTabSecScoresoperator}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{Security score} \\ &&&& (out of 10) \\ \hline O2 uk & $0.25 \pm 0.00$ & $0.12 \pm 0.00$ & 0.37 & $3.79 \pm 0.01$ \\  T-Mobile & $0.19 \pm 0.00$ & $0.18 \pm 0.00$ & 0.48 & $3.57 \pm 0.01$ \\  Orange & $0.21 \pm 0.00$ & $0.09 \pm 0.00$ & 0.39 & $3.52 \pm 0.02$ \\  Sprint & $0.18 \pm 0.00$ & $0.11 \pm 0.00$ & 0.44 & $3.41 \pm 0.01$ \\  3 & $0.15 \pm 0.00$ & $0.09 \pm 0.00$ & 0.49 & $3.15 \pm 0.01$ \\  AT\&T & $0.12 \pm 0.00$ & $0.08 \pm 0.00$ & 0.43 & $3.08 \pm 0.01$ \\  Vodafone uk & $0.13 \pm 0.00$ & $0.10 \pm 0.00$ & 0.52 & $3.07 \pm 0.01$ \\  Verizon & $0.18 \pm 0.00$ & $0.09 \pm 0.00$ & 0.81 & $2.82 \pm 0.01$ \\  unknown & $0.10 \pm 0.00$ & $0.22 \pm 0.00$ & 1.08 & $2.57 \pm 0.01$ \\  n Telenor & $0.03 \pm 0.00$ & $0.10 \pm 0.00$ & 1.42 & $1.58 \pm 0.01$ \\  Airtel & $0.07 \pm 0.00$ & $0.03 \pm 0.00$ & 1.96 & $1.08 \pm 0.01$ \\  Grameenphone & $0.01 \pm 0.00$ & $0.01 \pm 0.00$ & 2.04 & $0.76 \pm 0.00$ \\  Robi & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & 2.23 & $0.70 \pm 0.01$ \\  banglalink & $0.00 \pm 0.00$ & $0.03 \pm 0.00$ & 2.95 & $0.39 \pm 0.00$ \\ \end{tabular} \caption{Security scores for operators} \label{tab:sec_operator} \end{table}}
+\newcommand{\daTabSecScoresoperator}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{Security score} \\ &&&& (out of 10) \\ \hline O2 uk & $0.25 \pm 0.00$ & $0.12 \pm 0.00$ & 0.37 & $3.79 \pm 0.01$ \\  T-Mobile & $0.19 \pm 0.00$ & $0.18 \pm 0.00$ & 0.48 & $3.57 \pm 0.01$ \\  Orange & $0.21 \pm 0.00$ & $0.09 \pm 0.00$ & 0.39 & $3.52 \pm 0.02$ \\  Sprint & $0.18 \pm 0.00$ & $0.11 \pm 0.00$ & 0.44 & $3.41 \pm 0.01$ \\  3 & $0.15 \pm 0.00$ & $0.09 \pm 0.00$ & 0.49 & $3.15 \pm 0.01$ \\  AT\&T & $0.12 \pm 0.00$ & $0.08 \pm 0.00$ & 0.43 & $3.08 \pm 0.01$ \\  Vodafone uk & $0.13 \pm 0.00$ & $0.10 \pm 0.00$ & 0.52 & $3.07 \pm 0.01$ \\  Verizon & $0.18 \pm 0.00$ & $0.09 \pm 0.00$ & 0.81 & $2.82 \pm 0.01$ \\  unknown & $0.10 \pm 0.00$ & $0.22 \pm 0.00$ & 1.08 & $2.57 \pm 0.01$ \\  n Telenor & $0.03 \pm 0.00$ & $0.10 \pm 0.00$ & 1.42 & $1.59 \pm 0.01$ \\  Airtel & $0.06 \pm 0.00$ & $0.03 \pm 0.00$ & 1.93 & $1.10 \pm 0.01$ \\  Grameenphone & $0.01 \pm 0.00$ & $0.02 \pm 0.00$ & 1.99 & $0.80 \pm 0.00$ \\  Robi & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & 2.19 & $0.73 \pm 0.01$ \\  banglalink & $0.00 \pm 0.00$ & $0.03 \pm 0.00$ & 2.90 & $0.40 \pm 0.00$ \\ \end{tabular} \caption{Security scores for operators} \label{tab:sec_operator} \end{table}}
 \newcommand{\daSecScoreBestoperator}{O2 uk}
-\newcommand{\daSecScoreBestoperatorScore}{$3.79 \pm 0.0128$}
+\newcommand{\daSecScoreBestoperatorScore}{$3.79 \pm 0.01$}
 \newcommand{\daSecScoreWorstoperator}{banglalink}
-\newcommand{\daSecScoreWorstoperatorScore}{$0.386 \pm 0.0049$}
+\newcommand{\daSecScoreWorstoperatorScore}{$0.4 \pm 0.0$}
 \newcommand{\daNumSigOperators}{14}
@@
 \newcommand{\daSecScoreworstsummaryNumFullVersions}{1057}
-\newcommand{\daSecScoreBestoperatorNumFullVersions}{87}
+\newcommand{\daSecScoreBestoperatorNumFullVersions}{88}
 \newcommand{\daSecScoreworstoperatorNumFullVersions}{89}
-\newcommand{\daSecScoreWorstmanufacturerNumFullVersions}{14}
+\newcommand{\daSecScoreWorstmanufacturerNumFullVersions}{20}
 \newcommand{\daSecScoreWorstmodelNumFullVersions}{2}
 \newcommand{\daSecScoreWorstsummaryNumFullVersions}{1103}
-\newcommand{\daSecScoreWorstoperatorNumFullVersions}{58}
+\newcommand{\daSecScoreWorstoperatorNumFullVersions}{59}
 \newcommand{\daFullDeployedAt}{95\%}
-\newcommand{\daOSCurveFitParamFirst}{87.4}
+\newcommand{\daOSCurveFitParamFirst}{87.5}
 \newcommand{\daOSCurveFitParamSecond}{\num{0.00315}}
@@
 \newcommand{\daOSCurveSplineRMSE}{0.115}
-\newcommand{\daOSCurveHalfDeployed}{$308 \pm 81.8$}
-\newcommand{\daOSCurveFullDeployed}{$1\,040 \pm 235\,000$}
+\newcommand{\daOSCurveHalfDeployed}{$308 \pm 81$}
+\newcommand{\daOSCurveFullDeployed}{$1\,040 \pm 236\,000$}
 \newcommand{\daAPICurvePolyRMSE}{0.117}
 \newcommand{\daAPICurveSplineRMSE}{0.117}
-\newcommand{\daAPICurveHalfDeployed}{$330 \pm 84.0$}
+\newcommand{\daAPICurveHalfDeployed}{$330 \pm 84$}
 \newcommand{\daAPICurveFullDeployed}{$1\,060 \pm 236\,000$}
@@
 \newcommand{\daPercBigUpgrades}{22.4\%}
-\newcommand{\daUpdatesPerYear}{$1.35 \pm 0.0209$}
+\newcommand{\daUpdatesPerYear}{$1.35 \pm 0.02$}
 \newcommand{\daOSMonthsOfData}{40}
-\newcommand{\daMeanInsecurityPercNominal}{8\,720\%}
-\newcommand{\daVulnFreeNominal}{0.128}
-\newcommand{\daUpdatednessNominal}{0.0561}
-\newcommand{\daUpdatednessPercNominal}{561\%}
+\newcommand{\daMeanInsecurityPercNominal}{87.3\%}
+\newcommand{\daVulnFreeNominal}{0.127}
+\newcommand{\daUpdatednessNominal}{0.0558}
+\newcommand{\daUpdatednessPercNominal}{5.58\%}
 \newcommand{\daSecurityScoreNominal}{2.75}
@@
 \newcommand{\daSecScoreBestoperatorScoreNominal}{3.79}
-\newcommand{\daSecScoreWorstoperatorScoreNominal}{0.386}
+\newcommand{\daSecScoreWorstoperatorScoreNominal}{0.4}
 \newcommand{\daSecScoreBestmodelScoreNominal}{4.8}
@@
 \newcommand{\daSecScoreBestmanufacturerScoreNominal}{3.38}
-\newcommand{\daSecScoreWorstmanufacturerScoreNominal}{0.175}
-\newcommand{\daUpdatesPerMonthPerVersion}{3.61}
+\newcommand{\daSecScoreWorstmanufacturerScoreNominal}{0.139}
+\newcommand{\daUpdatesPerMonthPerVersion}{3.62}
+\newcommand{\daNumUpdateFullOnly}{0}

### 2014-11-14 securityupdates.tex
New version of the related work section.

@@
 \label{sec:related}
-We found that since 2011 \daMeanInsecurityPerc\ of devices on average were exposed to known critical vulnerabilities.
-In 2011 Felt et al.\ \cite{Felt2011} studied 6 Android handsets and found that they were exposed to root vulnerabilities at least 74\% of the time.
-This might indicate that over the last 3 years the security of Android has not improved, but these data are not directly comparable as their study was on 6 handsets considering the best possible update distribution while ours is on a large sample of devices with the real update distribution.
-They also found that 4 of the 46 malware samples (8\%) they analysed contained root exploits, much lower than rates found in later larger studies which found rates of 36.7\%~\cite{Zhou2012b} and 40\%~\cite{Zhou2012a}.
-Our results show root exploits are still a severe threat.
+Using the methods and data described in this paper we have determined that, on average, \daMeanInsecurityPerc\ of devices were exposed to known critical vulnerabilities between 2011 and 2014.
+Felt et al.\ studied 6 Android handsets in 2011 and found they were exposed to root vulnerabilities at least 74\% of the time~\cite{Felt2011} .
+Our approach differs from their study because they used data from 6 handsets and assumed the best possible update distribution, while our work is based on a large sample of devices tracking the actual update distribution.
+Nevertheless, the our own analysis and comparison with their work suggests protection against critical vulnerabilities has not improved significantly over the last 3 years. 
+Felt et al.\ also found that 4 of the 46 malware samples (8\%) they analysed contained root exploits, much lower than rates found in later (larger) studies which found rates of 36.7\%~\cite{Zhou2012b} and 40\%~\cite{Zhou2012a}.
 
-Finding vulnerabilities is frequently a time consuming manual process but, Brumley et al. showed that it is possible to automatically generate exploits from binary patches~\cite{Brumley2008}.
-In a similar way some of the vulnerabilities\todolater{which?} in \avo\ which have been exploited were publicly found by examining the commit made to the Android which fixed the vulnerability.
-This effect has been observed before in the Firefox source code repositories~\cite{Barth2011}, Google does not release the source code until after the release of the update reducing this effect to the delay between an update being released and it reaching the last end user device.
-There have been other attempts to automatically find vulnerabilities.
-The Woodpecker tool automatically finds permission leaks in stock Android phone images~\cite{Grace2012}.
-The update process itself can allow Android apps to gain privilege through `Pileup' vulnerabilities by registering for new permissions before the update which creates that permission is installed~\cite{Xing2014}.
+Currently, finding and exploiting vulnerabilities is typically a time-consuming and manual process.
+Therefore an attacker has to invest significant human resource to first find a bug and then write code to exploit the vulnerability.
+To reduce the effort involved in finding a bug, attackers can look at security-related commits made to open source repositories.
+Previous work has shown attackers have taken this approach for the Firefox web browser~\cite{Barth2011}.
+Google does not release updates to the Android source code until after the release of the relevant security update.
+This reduces the time available for an attacker to exploit a vulnerability by looking at security updates disclosed in the source code repository from the period (\textbf{first fix}, \textbf{fix deployed}) to (\textbf{fixing release}, \textbf{fix deployed}).
+Unfortunately the time period between \textbf{fixing release} and \textbf{fix deployed} can be of the order of months. 
+For example, the zergRush vulnerability was fixed in a release before it was publicly discovered but it still took 27 months for 90\% of the devices in our study to be definitely fixed (25 months for 90\% of devices maybe fixed).
 
-The security protections used in iOS such as app review, code signing, Data Execution Prevention (DEP) and Address Space Layout Randomisation (ASLR) and mandatory access controls has resulted in a lower level of malware affecting iOS~\cite{Felt2011} but can still be bypassed~\cite{Wang2013a}.
+Recent work by Brumley et al. has shown it is possible to automatically generate exploits from binary fixes, for example by using security updates issued by manufacturers~\cite{Brumley2008}. 
+Similarly, the Woodpecker tool automatically finds permission leaks in stock Android phone images~\cite{Grace2012}.
+Once such automated techniques mature, it will become even more important to provide security updates promptly to every Android device in the ecosystem because an attacker can turn the contents of a \textbf{fixing release} into an exploit in a matter of minutes or hours.
 
-Various attempts have been made to detect malware.
-RiskRanker classified 3\,281/118\,318 (2.8\%) apps as risky of which 718 (22\%) were malware and 322 (10\%) were previously unknown malware, an infection rate of 0.6\% across multiple markets~\cite{Grace2012a}.
-It uses signatures of exploits, static analysis of calls to high risk operations, detection of encrypted native code execution and dynamic Dalvik code loading to detect risky apps.
-DroidRanger found 148/182\,823 (0.08\%) apps to be malicious across multiple markets of which 29 were previously unknown~\cite{Zhou2012a}.
-It used permission-based behavioural fingerprinting -- looking at the permissions of known malware and heuristic-based filtering -- dynamic loading of both Dalvik and native code.
-AnDarwin detects similar apps, it found 169/265\,359 (0.06\%) malicious apps based on the difference in permissions between the malicious app and the app which had been cloned~\cite{Crussell2013}.
-It used clustering based on semantic vectors derived from the program dependence graphs to detect similar apps.
+Security in depth is also a useful strategy.
+In this regard, iOS provides additional safeguards beyond those used in Android, including a pre-publication review of code, mandatory code-signing by the manufacturer, and (with the important exception of ROP-based attacks~\cite{Wang2013a}.) the technical prevention of dynamic code loading by an app.
+These features, as well as Address Space Layout Randomisation (ASLR) and mandatory access controls, has resulted in a lower level of malware affecting iOS when compared to Android~\cite{Felt2011}.
+
+Rather than fixing critical vulnerabilities, security can be obtained by detect them and preventing the installation or execution of malicious apps.
+Detection strategies include RiskRanker, which classified 3\,281/118\,318 (2.8\%) apps as risky of which 718 (22\%) were malware and 322 (10\%) were previously unknown malware, an overall infection rate of 0.6\% across multiple markets~\cite{Grace2012a}.
+DroidRanger also analysed apps finding 148/182\,823 (0.08\%) apps to be malicious across multiple markets of which 29 were previously unknown~\cite{Zhou2012a}.
+%DroidRanger: It used permission-based behavioural fingerprinting which looked at the permissions of known malware and heuristic-based filtering -- dynamic loading of both Dalvik and native code.
+A common technique used by attackers takes existing popular apps and repackages them to include malicious code. 
+AnDarwin uses this insight to detect similar apps, and found 169/265\,359 (0.06\%) of all apps studied were malicious clones of genuine apps~\cite{Crussell2013}.
+%AnDarwin: It used clustering based on semantic vectors derived from the program dependence graphs to detect similar apps.
+Unfortunately, whilst successful in finding critical vulnerabilities, these techniques are not perfect.
 The Malware Genome project collected 1\,260 malware samples from 2010--2011~\cite{Zhou2012b}.
-They found that the best case for anti-malware detection on this data was 79.6\%.
-However DroidChameleon found that antivirus products could not detect malware if it was simply automatically permuted~\cite{Rastogi2013}.
-All these app analysis projects have been hampered by the difficulty of obtaining full datasets of Android apps as Google does not make these available (and researchers who automatically download them violate Google's Terms of Service).
-PlayDrone was a particularly effective project which circumvented Google's protective measures and downloaded over 1\,100\,000 apps from Google Play, allowing an in depth analysis~\cite{Viennot2014}.
+They found that the best case for anti-malware detection on their data was 79.6\%.
+Similarly, DroidChameleon found that antivirus products could not detect malware if the execution binary was automatically permuted~\cite{Rastogi2013}.
+Therefore detecting risky apps which might contain critical vulnerabilities is a useful tool, but it does not replace the need to provide security updates.
 
-Examining security update mechanisms has been used to assess security for many years.
-Many software update systems have been found not to authenticate the connection to the update server and/or do not authenticate the downloaded binaries~\cite{Bellissimo2006}.
-Even package management systems designed to provide secure updates have been found to have vulnerabilities~\cite{Cappos2008}.
-Fortunately Android does authenticate the update binaries (though the APK vulnerabilities circumvented this) and Google Play downloads them over a secure connection~\cite{Viennot2014}.
+%All these app analysis projects have been hampered by the difficulty of obtaining full datasets of Android apps as Google does not make these available (and researchers who automatically download them violate Google's Terms of Service).
+%PlayDrone was a particularly effective project which circumvented Google's protective measures and downloaded over 1\,100\,000 apps from Google Play, allowing an in depth analysis~\cite{Viennot2014}.
 
-The updatedness rate for Android of \daUpdatednessPerc\ compares unfavourably with the more than 90\% rate for Windows XP SP2 computers contacting the Microsoft update servers~\cite{Gkantsidis2006} though that is within one OS version and only for those computers which did connect to the servers -- but that is the default.
-However 27\% of Windows computers were still running Windows XP in July 2014\footnote{\url{https://archive.today/PLGxn}\todolater{Get updated figures before publication}} -- four months after it went out of security support.
+The security update process itself also needs to be secure.
+Unfortunately, package management systems designed to provide secure updates have been found to contain vulnerabilities~\cite{Cappos2008} and many software update systems fail to authenticate the connection between the device and the update server or do not authenticate the downloaded binaries~\cite{Bellissimo2006}.
+Android does authenticate the update binaries and Google Play downloads them over a secure connection~\cite{Viennot2014}.
+In this paper we have analysed critical flaws in the Android app update mechanism through three vulnerabilities: APK unsigned shorts, APK unchecked name and APK duplicate file. 
+Other work has demonstrate complex and subtle errors which exist in an update process.
+For example, the Android update process can be exploited to allow apps to gain privilege through `Pileup' vulnerabilities by registering for new permissions before the update which creates that permission is installed~\cite{Xing2014}.
 
-User-Agent strings have been used to investigate how web browser versions change with updates~\cite{Frei2008} and found that at most 80\% of Firefox users were running the most recent version.
+The percentage of Android devices running the most recent version (\daUpdatednessPerc\) compares unfavourably with the rate (>90\% ) for Windows XP SP2 computers contacting the Microsoft update servers~\cite{Gkantsidis2006}.
+This comparison is a little unfair because only one major OS version was considered in the analysis, and data was only collected from computers which contacted the update server, although this was the default.
+More recent data demonstrates the difficulty of upgrading computers between major OS versions, with 27\% of Windows computers were still running Windows XP in July 2014,\footnote{\url{https://archive.today/PLGxn}} four months after Windows XP stopped receiving security updates.
+
+User-Agent strings have been used to investigate the timeliness of web browser updates~\cite{Frei2008}, with at most 80\% of Firefox users running the most recent version.
 The same analysis was used to show that Chrome's use of silent updates seems to increase uptake of upgrades~\cite{Duebendorfer2010} with 97\% of users running the latest version within 3 weeks of release.
-Android's update process is manual, the user is notified but must action the upgrade which will then download the update and install it, the phone is rendered inoperable during the update process which is inconvenient for users and the phone must have sufficient charge (users are advised to plug it in) so that it will not run out during the update process.
-Partly this is the result of the fact that an operating system update is being installed and so a reboot is required, but Chrome installs the new version side by side with the old one and switches the next time it is restarted.
-The same technique would be more difficult on phones with limited storage space (as many cheap Android phones have barely enough space to install just the update) but is a plausible improvement for more high-end devices.
-Google is deploying the same silent update technique through Google Play Services\footnote{\href{http://lifehacker.com/why-google-play-services-are-now-more-important-than-an-975970197}{http://lifehacker.com/why-google-play-services-are-now-more-important-than-an-975970197}} which automatically installs updates for core Google components of Android, this also bypasses the device manufacturer and network operator.
+By way of comparison, Android's update process is manual.
+The user is notified an update exists, but further action is required, including downloading the update and rebooting the phone to enable installation.
+The phone must have sufficient charge to perform the update and the device itself is rendered inoperable during the update process, two factors which might prevent or delay the update process from taking place.
+In our data we are unable to determine why a device is not updated. 
+It is possible that many more updates arrive at handsets, but are simply not installed.
+Anecdotal evidence at least suggests that it is the lack of updates rather the lack of installation which is the major problem at present. Further work is required to tease apart these numbers.
+%Partly this is the result of the fact that an operating system update is being installed and so a reboot is required, but Chrome installs the new version side by side with the old one and switches the next time it is restarted.
+%The same technique would be more difficult on phones with limited storage space (as many cheap Android phones have barely enough space to install just the update) but is a plausible improvement for more high-end devices.
+%Google is deploying the same silent update technique through Google Play Services\footnote{\href{http://lifehacker.com/why-google-play-services-are-now-more-important-than-an-975970197}{http://lifehacker.com/why-google-play-services-are-now-more-important-than-an-975970197}} which automatically installs updates for core Google components of Android, this also bypasses the device manufacturer and network operator.
 

### 2014-11-14 securityupdates.tex
Minor fixes to related work.

@@
 Our approach differs from their study because they used data from 6 handsets and assumed the best possible update distribution, while our work is based on a large sample of devices tracking the actual update distribution.
-Nevertheless, the our own analysis and comparison with their work suggests protection against critical vulnerabilities has not improved significantly over the last 3 years. 
+Nevertheless, our own analysis as well as comparison with their work suggests protection against critical vulnerabilities has not improved significantly over the last 3 years. 
 Felt et al.\ also found that 4 of the 46 malware samples (8\%) they analysed contained root exploits, much lower than rates found in later (larger) studies which found rates of 36.7\%~\cite{Zhou2012b} and 40\%~\cite{Zhou2012a}.
@@
 This reduces the time available for an attacker to exploit a vulnerability by looking at security updates disclosed in the source code repository from the period (\textbf{first fix}, \textbf{fix deployed}) to (\textbf{fixing release}, \textbf{fix deployed}).
-Unfortunately the time period between \textbf{fixing release} and \textbf{fix deployed} can be of the order of months. 
+Unfortunately the time period between the \textbf{fixing release} and \textbf{fix deployed} can be of the order of months. 
 For example, the zergRush vulnerability was fixed in a release before it was publicly discovered but it still took 27 months for 90\% of the devices in our study to be definitely fixed (25 months for 90\% of devices maybe fixed).
@@
 Security in depth is also a useful strategy.
-In this regard, iOS provides additional safeguards beyond those used in Android, including a pre-publication review of code, mandatory code-signing by the manufacturer, and (with the important exception of ROP-based attacks~\cite{Wang2013a}.) the technical prevention of dynamic code loading by an app.
+In this regard, iOS provides additional safeguards beyond those used in Android, including a pre-distribution code review, mandatory code-signing by the manufacturer, and (with the important exception of ROP-based attacks~\cite{Wang2013a}) the technical prohibition of dynamic code loading by an app.
 These features, as well as Address Space Layout Randomisation (ASLR) and mandatory access controls, has resulted in a lower level of malware affecting iOS when compared to Android~\cite{Felt2011}.
@@
 Rather than fixing critical vulnerabilities, security can be obtained by detect them and preventing the installation or execution of malicious apps.
-Detection strategies include RiskRanker, which classified 3\,281/118\,318 (2.8\%) apps as risky of which 718 (22\%) were malware and 322 (10\%) were previously unknown malware, an overall infection rate of 0.6\% across multiple markets~\cite{Grace2012a}.
-DroidRanger also analysed apps finding 148/182\,823 (0.08\%) apps to be malicious across multiple markets of which 29 were previously unknown~\cite{Zhou2012a}.
+Detection strategies include RiskRanker, which classified 3\,281 out of 118\,318 apps (2.8\%) as risky of which 718 (22\%) were malware and 322 (10\%) were previously unknown malware, an overall infection rate of 0.6\% across multiple markets~\cite{Grace2012a}.
+DroidRanger also analysed apps finding 148 out of 182\,823 apps (0.08\%) to be malicious across multiple markets of which 29 were previously unknown~\cite{Zhou2012a}.
 %DroidRanger: It used permission-based behavioural fingerprinting which looked at the permissions of known malware and heuristic-based filtering -- dynamic loading of both Dalvik and native code.
 A common technique used by attackers takes existing popular apps and repackages them to include malicious code. 
-AnDarwin uses this insight to detect similar apps, and found 169/265\,359 (0.06\%) of all apps studied were malicious clones of genuine apps~\cite{Crussell2013}.
+AnDarwin uses this insight to detect similar apps, and found 169 out of 265\,359 of all apps studied (0.06\%) were malicious clones of genuine apps~\cite{Crussell2013}.
 %AnDarwin: It used clustering based on semantic vectors derived from the program dependence graphs to detect similar apps.
-Unfortunately, whilst successful in finding critical vulnerabilities, these techniques are not perfect.
+
+Unfortunately, whilst these tools were successful in finding critical vulnerabilities, such techniques are not perfect.
 The Malware Genome project collected 1\,260 malware samples from 2010--2011~\cite{Zhou2012b}.
@@
 
-The security update process itself also needs to be secure.
+The update process for apps, security fixes and OS upgrades also needs to be secure.
 Unfortunately, package management systems designed to provide secure updates have been found to contain vulnerabilities~\cite{Cappos2008} and many software update systems fail to authenticate the connection between the device and the update server or do not authenticate the downloaded binaries~\cite{Bellissimo2006}.
-Android does authenticate the update binaries and Google Play downloads them over a secure connection~\cite{Viennot2014}.
-In this paper we have analysed critical flaws in the Android app update mechanism through three vulnerabilities: APK unsigned shorts, APK unchecked name and APK duplicate file. 
-Other work has demonstrate complex and subtle errors which exist in an update process.
-For example, the Android update process can be exploited to allow apps to gain privilege through `Pileup' vulnerabilities by registering for new permissions before the update which creates that permission is installed~\cite{Xing2014}.
+Android does authenticate update binaries and Google Play downloads them over a secure connection~\cite{Viennot2014}.
+In this paper we have analysed three critical vulnerabilities in the Android app update mechanism: APK unsigned shorts, APK unchecked name and APK duplicate file. 
+Other work has demonstrate complex and subtle errors exist in the Android app update process.
+For example, the process can be exploited to allow apps to gain privilege through `Pileup' vulnerabilities by registering for new permissions before the update which creates that permission is installed~\cite{Xing2014}.
 
-The percentage of Android devices running the most recent version (\daUpdatednessPerc\) compares unfavourably with the rate (>90\% ) for Windows XP SP2 computers contacting the Microsoft update servers~\cite{Gkantsidis2006}.
-This comparison is a little unfair because only one major OS version was considered in the analysis, and data was only collected from computers which contacted the update server, although this was the default.
-More recent data demonstrates the difficulty of upgrading computers between major OS versions, with 27\% of Windows computers were still running Windows XP in July 2014,\footnote{\url{https://archive.today/PLGxn}} four months after Windows XP stopped receiving security updates.
+The percentage of Android devices running the most recent version (\daUpdatednessPerc) compares unfavourably with the rate ($>90$\%) for Windows XP SP2 computers contacting the Microsoft update servers~\cite{Gkantsidis2006}.
+A simple numerical comparison is unfair because only one major OS version was considered in the Microsoft analysis, and data was only collected from computers which contacted the update server, although this was the default.
+More recent data demonstrates the difficulty of upgrading computers between major OS versions, with 27\% of Windows computers running Windows XP in July 2014,\footnote{\url{https://archive.today/PLGxn}} four months after Windows XP stopped receiving security updates.
 
@@
 In our data we are unable to determine why a device is not updated. 
-It is possible that many more updates arrive at handsets, but are simply not installed.
-Anecdotal evidence at least suggests that it is the lack of updates rather the lack of installation which is the major problem at present. Further work is required to tease apart these numbers.
+It is possible that many updates arrive at handsets, but are simply not installed.
+Anecdotal evidence at least suggests that it is the lack of updates rather the lack of installation which is the major problem at present. Further work is required to tease these numbers apart.
 %Partly this is the result of the fact that an operating system update is being installed and so a reboot is required, but Chrome installs the new version side by side with the old one and switches the next time it is restarted.

### 2014-11-14 securityupdates.tex
Add note to reviewers for avo and change the avo macro

@@
 \input{avostats}
-\newcommand{\avo}{\texttt{androidvulnerabilities.org}}
+\newcommand{\avo}{AVO}
 \newcommand{\percMarketShare}{83.6\%~\footnote{\url{http://www.theinquirer.net/inquirer/news/2379036/android-hits-836-percent-marketshare-while-ios-windows-and-blackberry-slide}}}
@@
 
-We developed and maintain an open platform for filing critical vulnerabilities in a machine readable format, \texttt{\url{http://androidvulnerabilities.org/}}.
+We developed and maintain an open platform for filing critical vulnerabilities in a machine readable format, the AndroidVulnerabilities.org\footnote{\url{http://androidvulnerabilities.org/} \textbf{Note to reviewers: do not visit the contact page of this website}} (\avo) website.
 We seeded it with data from the CVE database, vendor lists, reports from the literature and various forums.

### 2014-11-14 securityupdates.tex
Updated Sections I and II; normalised ADB across paper.

--- 2014-11-14 securityupdates.tex
@@
 Removing an app which has used an exploit to take control of the handset is unlikely to return the device to secure state.
+We describe our threat model and three attack vectors in detail in \S\ref{sec:threatmodel}.
 
@@
 
-In order to determine whether an individual smartphone is vulnerable to a particular exploit, we need to combine vulnerability data with the operating system version string and additional update information from the handset.
+In order to determine whether an individual smartphone is vulnerable to a particular exploit, we need to combine vulnerability data with the operating system version string and build number from the handset.
 For this we use handset data gathered from \daNumOSDevices\ devices in the \da~\cite{Wagner2013} project.
@@
 One solution is regulation, and indeed there is ongoing legal action to force network operators to ship updates for security vulnerabilities~\cite{Soghoian2013}.\todolater{Check on the status of this legal action}
-Indeed, many smartphones are sold on 12--24 month contracts, and yet our data shows many devices do not receive many security updates (\daUpdatesPerYear\ per year). 
+Many smartphones are sold on 12--24 month contracts, and yet our data shows many devices do not receive many security updates, with an overall average of \daUpdatesPerYearNominal\ per year. 
 In contrast, Windows XP could be purchased for a one-off payment in October 2001 and received security updates until April 2014.
@@
 Comparative data would provide an incentive to both device manufacturers and network operators to provide updates.
-Currently corporate and public sector buyers are encouraged to purchase secure devices, but we have found little concrete guidance on the specific makes and models which provide the best security. 
+Currently corporate and public sector buyers are encouraged to purchase secure devices, but we have found little concrete guidance on the specific makes and models providing the best security. 
 For example, CESG, which advises the UK government on how to secure its computer systems, recommends picking Android devices from device manufactures which are good at shipping security updates promptly~\cite{CESG2013} but it does not state which device manufacturers these are.
 To address this, we developed a scoring system and provide numbers on the historic performance of devices in the \da\ project in~\S\ref{sec:security_scoring}.
-
-%REMOVED BY ARB
-%The device manufacturer has little financial incentive to perform this work as they have already sold the device.
-%Network operators are cautious because if they ship a broken update then they will get many support requests which is expensive.
-%
-%In the past users of Microsoft IIS were advised to switch to a different web server due to the frequent discovery of vulnerabilities~\cite{Pescatore2001}.
-%This pressure encouraged Microsoft to improve the security of their products~\cite{TODO}.
 
@@
 In this paper we measure whether an attacker can take control of a handset by breaking out of the app sandbox on Android using a known vulnerability.
-We consider two attack vectors.
+We consider three attack vectors.
 The first attack vector is through the installation of a malicious app on the device.
@@
 
-Support for dynamic code loading on Android also permits an indirect method where malicious code can be injected directly into existing apps already installed on the handset. 
+Support for dynamic code loading on Android also supports the third attack vector in which the attacker injects malicious code directly into existing apps already installed on the handset. 
 For example, the addJavascriptInterface vulnerability (CVE-2012-6636) allows Javascript running in an Android WebView to execute arbitrary code in the same process address space as the vulnerable app.
@@
 For some vulnerabilities the corresponding patch is not in AOSP despite some device manufacturers having shipped builds containing the fix, which makes working out what is going on more difficult.
-For \emph{RageAgainstTheCage adb}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/RageAgainstTheCage_adb}} and \emph{keystore buffer}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/keystore_buffer}} vulnerabilities which we do have sufficient data it is not clear whether they are truly critical vulnerabilities as the former may require physical \texttt{adb} access and the latter may be protected by a sandbox.
+For \emph{RageAgainstTheCage adb}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/RageAgainstTheCage_adb}} and \emph{keystore buffer}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/keystore_buffer}} vulnerabilities which we do have sufficient data it is not clear whether they are truly critical vulnerabilities as the former may require physical ADB access and the latter may be protected by a sandbox.
 However those two vulnerabilities make little difference to our results as they affect few devices.
@@
  \item Why was there a gap in vulnerability discovery in Android in 2012?
- \item Why do so many devices have \texttt{adb} enabled?
-\daAdbEnabledPerc\ seems rather high for a developer feature.
-This could be a selection bias or device manufacturers could be enabling it by default.
+ \item Why do so many devices have ADB enabled? \daAdbEnabledPerc\ seems rather high for a developer feature. This could be a selection bias or device manufacturers could be enabling it by default.
  \item How much of updating is new handsets and how much is updates being deployed?

### 2014-11-14 securityupdates.tex
Minor tweaks to related work

@@
 Nevertheless, our own analysis as well as comparison with their work suggests protection against critical vulnerabilities has not improved significantly over the last 3 years. 
-Felt et al.\ also found that 4 of the 46 malware samples (8\%) they analysed contained root exploits, much lower than rates found in later (larger) studies which found rates of 36.7\%~\cite{Zhou2012b} and 40\%~\cite{Zhou2012a}.
+Felt et al.\ also found that 4 of the 46 malware samples (8\%) they analysed contained root exploits, much lower than rates found in later (larger) studies which found rates of 36.7\%~\cite{Zhou2012b} and 40\%~\cite{Zhou2012a} in 2012.
 
@@
 Unfortunately the time period between the \textbf{fixing release} and \textbf{fix deployed} can be of the order of months. 
-For example, the zergRush vulnerability was fixed in a release before it was publicly discovered but it still took 27 months for 90\% of the devices in our study to be definitely fixed (25 months for 90\% of devices maybe fixed).
+For example, the zergRush vulnerability was fixed in a release before it was publicly discovered but it still took 27 months for 90\% of the devices in our study to be definitely fixed\footnote{Definitely fixed is 1 $-$ (definitely insecure + maybe insecure), maybe fixed is 1 $-$ definitely insecure} (25 months for 90\% of devices maybe fixed).
 
@@
 Android does authenticate update binaries and Google Play downloads them over a secure connection~\cite{Viennot2014}.
-In this paper we have analysed three critical vulnerabilities in the Android app update mechanism: APK unsigned shorts, APK unchecked name and APK duplicate file. 
+In this paper we have analysed four critical vulnerabilities in the Android app update mechanism: APK unsigned shorts, APK unchecked name, APK duplicate file and Fake ID.
 Other work has demonstrate complex and subtle errors exist in the Android app update process.

### 2014-11-14 securityupdates.tex
Tweak the conclusion and update the (currently hidden) Acknowledgements

@@
 We have investigated the security of the Android operating system with respect to security updates and examined how the Android ecosystem results in different devices receiving different levels of security due to whether or not they get security updates.
-We have compared different models, device manufacturers and network operators and found that there are differences between them which the discerning purchaser might use to influence their decision about which device to buy from which device manufacturer and network operator.
+We have compared different device models, device manufacturers and network operators and found that there are differences between them which the discerning purchaser might use to influence their decision about which device to buy from which device manufacturer and network operator.
 We hope that this analysis will encourage device manufacturers and network operators to improve the support they provide for devices after sale.
-Android attempts to provide good security to end users but the latency in the security update process means that on average \daMeanInsecurityPerc\ of Android devices are exposed to known critical vulnerabilities.
+Android attempts to provide good security to end users through sandboxing apps.
+However the latency in the security update process means that on average \daMeanInsecurityPerc\ of Android devices are exposed to known critical vulnerabilities which break out of the sandbox and only \daUpdatednessPerc\ of devices run the latest version of Android.
 
@@
 \section*{Acknowledgements}
+Thanks to Richard Clayton, the mobile security reading group and Anil Madhavapeddy for reading various drafts of this paper.
 Thanks to David Robertson for helpful advice on statistical analysis.

### 2014-11-14 securityupdates.tex
Updates to section III.

--- 2014-11-14 securityupdates.tex
@@
 Device Analyzer collects data from study participants who install the project Android app from the Google Play store.
-Many participants allow researchers around the world to access a subset of their data.
+Most study participants allow researchers around the world to access a subset of their device data, including the data presented in this paper.
 
 The \da\ app collects a range of metadata from Android devices.\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/collected.html}}
-For this paper, we extracted the build string and API version on the device each day.
+For this paper, we extracted the build string and API version from the device log each day.
 The API version is a positive integer which increases when new features are added to the API.
@@
 
-We developed and maintain an open platform for filing critical vulnerabilities in a machine readable format, the AndroidVulnerabilities.org\footnote{\url{http://androidvulnerabilities.org/} \textbf{Note to reviewers: do not visit the contact page of this website}} (\avo) website.
+We developed and maintain an open platform for filing critical vulnerabilities in a machine readable format, the AndroidVulnerabilities.org\footnote{\url{http://androidvulnerabilities.org/} \textbf{Note to reviewers:} we have removed our names and affiliation from the website for this review \textbf{with the exception of the contact page.
+Please do not visit the contact page until the review cycle has completed.}} (\avo) website.
 We seeded it with data from the CVE database, vendor lists, reports from the literature and various forums.
-We have received submissions or amendments from \avoNumSubmitters\ individuals.
-We collected this data between \avoStartDate\ and \avoEndDate\ and so any information lost before the start of that period cannot be included.
+In addition, we have received submissions or amendments from \avoNumSubmitters\ individuals.
+We collected data between \avoStartDate\ and \avoEndDate\ and so any information lost before the start of that period cannot be included.
 \avo\ currently contains \avoNumVulnerabilities\ vulnerabilities of which \avoNumVulnAllAndroid\ affect all Android devices and \avoNumVulnSpecific\ are specific to particular devices or device manufacturers.
-We only use the \daNumVulnsUsed\ vulnerabilities which we have sufficient data to use for this analysis because they affect all Android devices rather than particular device manufacturers and we know enough information about the vulnerability to tie it to the particular versions it affects.
-We summarise the \daNumVulnsUsed\ vulnerabilities used for this analysis in Table~\ref{tab:andvulns}.\todolater{We have published full details of the \daNumVulnsUsed\ vulnerabilities used for this analysis in an accompanying technical report~\cite{TODO}}
 
-Tracking vulnerabilities is a manual task as they are not consistently recorded in other databases such as the CVE database and the lack of a widely acknowledged unique identifier made identifying whether two reports referenced the same vulnerability difficult.
+Unless otherwise stated, we use \daNumVulnsUsed\ vulnerabilities in our analysis as shown in Table~\ref{tab:andvulns}.
+We have chosen these vulnerabilities to fit the attack vectors introduced in \S\ref{sec:threatmodel}.
+These vulnerabilities affect all Android devices regardless of manufacturer, and as a result our selected vulnerabilities will dominate any security analysis.
+In many cases we could not match manufacturer- and model-specific vulnerabilities to individual devices in the \da\ data and therefore attempting to include device-specific vulnerabilities as well would introduce additional uncertainty in our results.
+In contrast, with our chosen set of vulnerabiltiies, our analysis represents a lower-bound on the vulnerability of devices in the \da\ data set.
+
+Tracking vulnerabilities is a manual task as they are not consistently recorded in other databases such as the CVE database.
+In addition, the lack of a widely acknowledged unique identifier required manual analysis to identify whether two reports referenced the same vulnerability.
 Previous work has assumed ``any security issue of relevance will eventually get a CVE number assigned"~\cite{Frei2010} which is currently not the case for critical Android vulnerabilities.
@@
 \textbf{introducing release} When the first release was made containing the vulnerability.\\
-\textbf{discovery} When the vulnerability is first discovered\\
-\textbf{exploit} When the vulnerability is first exploited\\
+\textbf{discovery} When the vulnerability is first discovered.\\
+\textbf{exploit} When the vulnerability is first exploited.\\
 \textbf{disclosure} When the vulnerability is first disclosed.\\ % publically or to a smaller set of people?
@@
 
-Establishing when a vulnerability started posing a threat to users is difficult.
-Frei et al.~\cite{Frei2010} propose the definition of the {\bf time of disclosure} when the information about the vulnerability is freely available to the public, from a widely accepted and independent source and has been validated by security experts so that it has a risk rating.
+Establishing when a vulnerability starts to pose a threat to users is difficult.
+Frei et al.~\cite{Frei2010} propose the definition of the {\bf time of disclosure}.
+This occurs when the information about the vulnerability is freely available to the public from a widely accepted and independent source and has been validated by security experts so that it has a risk rating.
 Unfortunately before we collated this information much of it was not published by an independent source and lacked risk rating information, even months or years after they had been actively used.
@@
 
-Symantec's 2012 analysis of desktop malware has shown that after public disclosure exploitation rates increase by 5 orders of magnitude~\cite{Bilge2012} and so from the point of view of widespread danger, the period between public disclosure and the update reaching the user's device is the most critical.
-However they also show that zero day vulnerabilities are typically used for 312 days before they are publicly disclosed, often to target particular organisations.
-Hence when considering vulnerability from the point of view of an organisation which cares about Advanced Persistent Threats, such as those the CESG advice~\cite{CESG2013} is aimed at, the critical period is from when people could have started using the vulnerability to when the update hits the user device.
-So for our vulnerability calculations we use the earliest date we can find recorded evidence that the vulnerability was known about, even if that knowledge might have been confined to a particular device manufacturer or hobbyist.
+Symantec's 2012 analysis of desktop malware has shown that after public disclosure, exploitation rates increase by 5 orders of magnitude~\cite{Bilge2012} and so from the point of view of widespread danger, the period between public \textbf{disclosure} and the date of \textbf{fix deployed} is the most critical.
+However they also show that zero-day vulnerabilities are typically used for 312 days before they are publicly disclosed, often to target particular organisations.
+Hence when considering a vulnerability from the point of view of an organisation which cares about Advanced Persistent Threats, such as those responding to CESG advice~\cite{CESG2013}, the critical period starts with \textbf{discovery} and continues until \textbf{fix deployed}.
+Therefore, in our vulnerability calculations we prefer to use the earliest recorded date of \textbf{discovery}, even if that knowledge might have been confined to a particular device manufacturer or hobbyist.
 %We cannot know if someone reported the vulnerability to the device manufacturer and we do know that various agencies are engaged in widespread monitoring of communications and compromise of civilian infrastructure~\cite{TODO} and so they may know.
 
-There are several plausible points from which a vulnerability poses a threat to users.
-{\bf Creation} and {\bf introducing release} are the earliest points which could be used but the risk is mostly latent until someone discovers them.
-Unfortunately {\bf discovery} is the hardest point to obtain concrete data on as discovery may happen multiple times independently and not all discoverers will report their discovery, for example entities with high value targets may keep databases of unknown vulnerabilities for later use, in which case there is a high risk to potential targets from the never disclosed point of discovery.
-However if the discoverer never discloses the vulnerability to anyone or makes any use of it then there is little danger.
+Unfortunately {\bf discovery} is the hardest point to obtain concrete data on as discovery may happen multiple times independently and not all discoverers will report their discovery.
+{\bf Creation} and {\bf introducing release} are the earliest points which could be used, but the risk is mostly latent until someone discovers them and therefore we do not use these dates.
+%, for example entities with high value targets may keep databases of unknown vulnerabilities for later use, in which case there is a high risk to potential targets from the never disclosed point of discovery.
+%However if the discoverer never discloses the vulnerability to anyone or makes any use of it then there is little danger.
 The date of first {\bf exploit} is a point when the risk is definitely high, but again a good adversary will not be detected when using such exploits.
-The date of first {\bf fix} is, assuming that the fix is deliberate, a point at which the vulnerability is known at least within the organisation performing the fix and frequently implies an earlier discovery and notification by a third party.
-It can, for example, be determined from the authored-on date of the fixing commit in the git repository.
-Once a {\bf fixing release} has been made then the vulnerability is widely known because it can be reverse engineered from the changes in the release~\cite{Brumley2008}.
-When the {\bf fix deployed} to a sufficiently large proportion of devices then the remaining risk is minimal.
-
-Hence we use the earliest date we know about for discovery, exploit, disclosure or fix as the date from which a vulnerability poses a substantial risk to users.
-This is shown in Table~\ref{tab:andvulns}.
+The date of first {\bf fix} is, assuming that the fix is deliberate, a point at which the vulnerability is known at least within the organisation performing the fix and frequently implies an earlier discovery and notification by a third party; we determine this date from the Android Open Source Project repository by examining the authored-on date of the fixing commit.
+%Once a {\bf fixing release} has been made then the vulnerability is widely known because it can be reverse engineered from the changes in the release~\cite{Brumley2008}.
+%When the {\bf fix deployed} to a sufficiently large proportion of devices then the remaining risk is minimal.
+Therefore, in this paper we use the earliest known date from \textbf{discovery}, \textbf{exploit}, \textbf{disclosure} or \textbf{fix}.
+A breakdown of the type of date used per vulnerability is shown in Table~\ref{tab:andvulns}.
 
@@
  \includegraphics[width=\columnwidth]{figures/vulnerabilities_timeline}
- \caption{Timeline of vulnerabilities}
+ \caption{Timeline of vulnerabilities. For each vulnerability we show the dates of \textbf{discovery} and, when later, the dates of the first \textbf{fixing release}}
  \label{fig:vulnerabilities_timeline}
@@
 
-The dates of discovery of vulnerabilities in the \avo\ is not uniform.
-Figure~\ref{fig:vulnerabilities_timeline} shows the dates of discovery and, when later, the dates when a version of Android which fixed the vulnerability was shipped.
-Some vulnerabilities (\emph{levitator}, \emph{zergRush}) were fixed in released versions of Android before they were discovered and so are shown as vertical lines while others were known for months before a version of Android which fixed them was shipped.
+Figure~\ref{fig:vulnerabilities_timeline} shows the dates of \textbf{discovery} and, when later, the date of the first \textbf{fixing release}.
+Some vulnerabilities (\emph{levitator}, \emph{zergRush}) were fixed in released versions of Android before they were discovered and so are shown as vertical lines, while others were known for months before a version of Android which fixed them was shipped.
 During the period in which \da\ data was collected the date when a version of Android with the fix was observed on a \da\ device is taken as the date that version was released.
-For vulnerabilities prior to the collection period the release date as best as we can determine is used.
-There is no canonical source of Android release dates, our best guesses and supporting references are available from \avo.
+For vulnerabilities prior to the collection period we estimate the release date as best we can using publicly available data.
+There is no canonical source of Android release dates, and our best guesses and supporting references are available from \avo.
 
-This data shows a large gap from 2011-10-06 to 2013-02-18 where we do not know of any discoveries of critical vulnerabilities affecting Android devices.
+The discovery dates of our vulnerabilities does not appear to be uniform uniform. 
+In particular the data shows a large gap from 2011-10-06 to 2013-02-18 where we do not know of any discoveries of critical vulnerabilities affecting all Android devices.
 The cause of this quiet period is unclear.
-Possible explanations are: (i) That since most devices were exposed to known vulnerabilities there was no point in looking for new ones (from Figure~\ref{fig:proportioninsecure}).
-(ii) That device manufacturers made it easier to install custom versions of Android, reducing the need for users to root their devices.
-(iii) That device manufacturer specific vulnerabilities (which we are ignoring) proved to be easier to find and so people looking for vulnerabilities adjusted their focus.
+Possible explanations are that: (i) most devices were exposed to known vulnerabilities so there was no point in looking for new ones (from Figure~\ref{fig:proportioninsecure});
+(ii) device manufacturers made it easier to install custom versions of Android, reducing the need for users to root their devices; and
+(iii) device manufacturer specific vulnerabilities (which we are not in our analysis here) were easier to find and therefore attackers looking for vulnerabilities adjusted their focus.
 
@@
 There is a complex Android ecosystem which creates and distributes updates to Android which fix vulnerabilities.
-In this section we rectify the lack of public information about the nature of the Android ecosystem.
+In this section we describe how the Android ecosystem functions and how Android versions, as seen by the \da\ project, are produced.
 

### 2014-11-14 securityupdates.tex
New updates to experiment section.

@@
 
-The discovery dates of our vulnerabilities does not appear to be uniform uniform. 
+The discovery dates of our vulnerabilities does not appear to be uniform. 
 In particular the data shows a large gap from 2011-10-06 to 2013-02-18 where we do not know of any discoveries of critical vulnerabilities affecting all Android devices.
@@
 There are five entities or groups which contribute towards Android updates: the network operators, the device manufacturers, the hardware developers, Google and the upstream open source projects.
-Android builds on various open source projects such as the Linux kernel, the OpenSSL and BouncyCastle cryptography libraries and so can include any compatible versions of those projects, including those which fix security vulnerabilities.
+Android builds on various open source projects such as the Linux kernel, OpenSSL and BouncyCastle cryptography libraries.
+Consequently Android can include any compatible versions of those projects, including those which fix security vulnerabilities.
 Android also incorporates various drivers for different bits of hardware.
-The Android platform is then built on top of those with kernel and userspace components by Google.
-The code for each update is kept secret\footnote{\url{https://source.android.com/source/code-lines.html}}\todolater{Can we quantify this keeping the code secret?
-Is it worth it?} until after the update has been published.
-Device manufacturers may receive advance copies in order to prepare handsets.
-Then the device manufacturers take the update and customise it before passing it on to the network operator.
-The network operator may then make further customisations and do further testing before shipping the update to the Device.
-Sometimes device manufactures ship direct to the user, sometimes the device manufacturer and Google collaborate closely for the purpose of a particular phone, such as with Nexus phones, and hence are harder to separate.
-Sometimes device manufacturers incorporate upstream open source project releases directly, and sometimes incorrectly -- for example including a broken daily build of sqlite in their release of Android~\cite{Wagner2013}.
-\todolater{use statistics from samsung-updates.com -> how many binaries are there per device?}
+The Android platform is then built from these components by Google.
+The code for each Android release or update is kept secret\footnote{\url{https://source.android.com/source/code-lines.html}} until after a binary release has been published.
+Device manufacturers likely receive advanced access in order to prepare handsets so they can customise it before passing it on to the network operator.
+The network operator may then make or request further customisations and perform further testing before shipping the update to the Device.
+Sometimes device manufactures ship updates directly to the user, sometimes the device manufacturer and Google collaborate closely to make a particular phone, such as with Nexus devices.
+Sometimes device manufacturers incorporate upstream open source project releases directly, and sometimes incorrectly -- for example previous work has recorded evidence of broken nightly builds of sqlite in Android releases~\cite{Wagner2013}.
 
@@
 \label{sec:results}
-We present the results of our analysis showing \daMeanInsecurityPerc\ of devices to have critical vulnerabilities on average.
-On average only \daUpdatesPerYear\ updates reach a device each year and \daPercUpdatesDowngrades\ of version changes are downgrades to older versions.
+We present the results of our analysis showing that, on average, \daMeanInsecurityPerc\ of Android devices are exposed to critical vulnerabilities.
+Devices, on average, apply \daUpdatesPerYear\ updates each year and \daPercUpdatesDowngrades\ of version changes are downgrades to older versions.
 Nexus devices are better than others, \emph{\daSecScoreBestmanufacturer} is the highest scoring device manufacturer and \emph{\daSecScoreBestoperator} is the highest scoring network operator.
 %\todo{comparison of nexus update rates with iOS}
-We conducted three experiments examining the behaviour of the Android ecosystem as a whole (\S\ref{sec:exp:android_ecosystem}), updates reaching particular devices (\S\ref{sec:exp:device_updates}) and comparisons between different device manufacturers and network operators (\S\ref{sec:exp:security_score}).
+We describe four experiments.
+We examine the vulnerability of the Android ecosystem as a whole (\S\ref{sec:exp:versionsecurity}, explore the upgrade cycle and vulnerability cycle of the ecosystem (\S\ref{sec:exp:android_ecosystem}), quantify the updates installed on particular devices (\S\ref{sec:exp:device_updates}) and compare the performance of different device manufacturers and network operators (\S\ref{sec:exp:security_score}).
 
@@
 \subsection{Experiment 0: vulnerability of Android devices}\label{sec:exp:versionsecurity}
-In Figure~\ref{fig:proportioninsecure} we showed the vulnerability of devices in \da\ to known critical Android vulnerabilities.
-Initially the plot is all `maybe insecure' (yellow) as since \da\ does not have historical data from before that point we cannot distinguish between devices which are running a version of Android which is known to be vulnerable but which have a backported fix and those which do not have a backported fix as we do not have data on what builds existed before the vulnerability was discovered.
-This shows the importance of a longitudinal study as this analysis requires years of data.
-Once \emph{zergRush} was discovered the plot goes mostly `insecure' (red) as most devices were exposed to that vulnerability but some were already running a version of Android which fixed that vulnerability and so from that point until the discovery of \emph{APK duplicate file} the graph shows progressively more `secure' green.
-From that point onwards the more regular discovery of critical vulnerabilities ensures that most devices are exposed to known critical vulnerabilities.
 
-\subsection{Experiment 1: behaviour of the Android ecosystem}\label{sec:exp:android_ecosystem}
+In the introduction, Figure~\ref{fig:proportioninsecure} described the proportion of Android devices susceptable to at least one critical vulnerability.
 
@@
 The \avo\ data covers the period from \avoFirstDataDate\ to \avoLastDataDate.
-For each device, for every day with version data we record what version it was running and which vulnerabilities it was exposed to at that time (if any).
-We then normalise these totals for each day by the total number of devices with version information seen on that day.
+For each device we collected daily version data  and recorded the running version and any associated vulnerabilities it was exposed to at that time (if any).
+We then normalise these totals for each day by dividing through by the total number of devices with version information seen on that day.
+
+\subsubsection{Results}
+
+Following time from left to right in Figure~\ref{fig:proportioninsecure} we see all devices are initially \emph{maybe secure} (yellow) since \da\ does not have historical data prior to May 2011.
+This means we cannot distinguish between devices which are running a version of Android which is known to be vulnerable from one which may have received a backported fix.
+This demonstrates the importance of a longitudinal study: this type of analysis requires years of data.
+Once \emph{zergRush} was discovered in October 2011 the most devices are recorded as \emph{insecure} (red) as most devices were exposed to that vulnerability.
+The remanding devices were already running a version of Android which fixed the \emph{zergRush} vulnerability and are therefore marked as  \emph{secure} (green).
+From October 2011 until the discovery of \emph{APK duplicate file} in February 2013 the graph shows progressive improvement as devices are upgraded.
+This means more and more devices are marked as \emph{secure} (green) because they are now running a secure version of Android, or marked as \emph{maybe secure} because they received a minor update which may have contained a backported fix.
+From February 2013 onwards regular discovery of critical vulnerabilities ensures that most devices are exposed to known critical vulnerabilities.
+
+\subsection{Experiment 1: behaviour of the Android ecosystem}\label{sec:exp:android_ecosystem}
+
+\subsubsection{Method} As in Experiment 0, we used the version information for each device to calculate which critical vulnerabilities each device was susceptible to on a daily basis.
 
@@
 \subsubsection{Method}
-As in Experiment 1, for each device we used the version information from each day to calculate which vulnerabilities it was vulnerable to that day.
+As in Experiment 0, we used the version information for each device to calculate which critical vulnerabilities each device was susceptible to on a daily basis.
 We also recorded when the version changed and which versions it changed from and to.

### 2014-11-14 securityupdates.tex
Fix terminology

@@
 Currently corporate and public sector buyers are encouraged to purchase secure devices, but we have found little concrete guidance on the specific makes and models providing the best security. 
-For example, CESG, which advises the UK government on how to secure its computer systems, recommends picking Android devices from device manufactures which are good at shipping security updates promptly~\cite{CESG2013} but it does not state which device manufacturers these are.
-To address this, we developed a scoring system and provide numbers on the historic performance of devices in the \da\ project in~\S\ref{sec:security_scoring}.
+For example, CESG, which advises the UK government on how to secure its computer systems, recommends picking Android device models from device manufactures which are good at shipping security updates promptly~\cite{CESG2013} but it does not state which device manufacturers these are.
+To address this, we developed a scoring system and provide numbers on the historic performance of device models in the \da\ project in~\S\ref{sec:security_scoring}.
 
@@
 
-Where possible we indicate the uncertainty in our results by presenting them $\pm$ one standard deviation, this occasionally results in `${} \pm 0$' when the standard deviation is less than the number of significant figures we are reporting.
+Where possible we indicate the uncertainty in our results by presenting them $\pm$ one standard deviation and give results to 3 significant figures, this occasionally results in `${} \pm 0$' when the standard deviation is less than the number of significant figures we are reporting.
 This technique cannot account for systematic errors and therefore we explore the potential for systematic errors in the \da\ data in \S\ref{sec:representative}.
@@
 
-\section{Discussion}
+\section{Discussion} \label{sec:discussion}
 There are continuing efforts to reduce the impact of critical vulnerabilities, both in Android and more widely.
@@
 
-We have not included all the vulnerabilities from \avo\, in particular we have not included any device manufacturer specific vulnerabilities, even when these are widespread (such as vulnerabilities affecting Qualcomm\footnote{Qualcomm is particularly good at publicly disclosing the vulnerabilities affecting their code and the patches which fix them.\url{https://www.codeaurora.org/projects/security-advisories/}} chipsets) because it is difficult to work out which devices are affected.
+We have not included all the vulnerabilities from \avo\, in particular we have not included any device manufacturer specific vulnerabilities, even when these are widespread (such as vulnerabilities affecting Qualcomm\footnote{Qualcomm is particularly good at publicly disclosing the vulnerabilities affecting their code and the patches which fix them.
+\url{https://www.codeaurora.org/projects/security-advisories/}} chipsets) because it is difficult to work out which devices are affected.
 There also tends to be less public information available about device manufacturer specific vulnerabilities which makes them harder to tie down.
@@
 \label{sec:conclusion}
-We have investigated the security of the Android operating system with respect to security updates and examined how the Android ecosystem results in different devices receiving different levels of security due to whether or not they get security updates.
-We have compared different device models, device manufacturers and network operators and found that there are differences between them which the discerning purchaser might use to influence their decision about which device to buy from which device manufacturer and network operator.
+We have investigated the security of the Android operating system with respect to security updates and examined how the Android ecosystem results in different device models receiving different levels of security due to whether or not they get security updates.
+We have compared different device models, device manufacturers and network operators and found that there are differences between them which the discerning purchaser might use to influence their decision about which device model to buy from which device manufacturer and network operator.
 We hope that this analysis will encourage device manufacturers and network operators to improve the support they provide for devices after sale.

### 2014-11-14 securityupdates.tex
Minor fixes

@@
 We have chosen these vulnerabilities to fit the attack vectors introduced in \S\ref{sec:threatmodel}.
-These vulnerabilities affect all Android devices regardless of manufacturer, and as a result our selected vulnerabilities will dominate any security analysis.
+These vulnerabilities affect all Android devices regardless of manufacturer, and as a result our selected vulnerabilities will dominate security analysis of the Android ecosystem as a whole.
 In many cases we could not match manufacturer- and model-specific vulnerabilities to individual devices in the \da\ data and therefore attempting to include device-specific vulnerabilities as well would introduce additional uncertainty in our results.
@@
  \includegraphics[width=\columnwidth]{figures/vulnerabilities_timeline}
- \caption{Timeline of vulnerabilities. For each vulnerability we show the dates of \textbf{discovery} and, when later, the dates of the first \textbf{fixing release}}
+ \caption{Timeline of vulnerabilities. For each vulnerability we show the dates of \textbf{discovery} and, when later, the dates of the first \textbf{fixing release}.}
  \label{fig:vulnerabilities_timeline}

### 2014-11-14 securityupdates.tex
Moved open questions into a paragraph.

--- 2014-11-14 securityupdates.tex
@@
 
-We have not included all the vulnerabilities from \avo\, in particular we have not included any device manufacturer specific vulnerabilities, even when these are widespread (such as vulnerabilities affecting Qualcomm\footnote{Qualcomm is particularly good at publicly disclosing the vulnerabilities affecting their code and the patches which fix them.\url{https://www.codeaurora.org/projects/security-advisories/}} chipsets) because it is difficult to work out which devices are affected.
-There also tends to be less public information available about device manufacturer specific vulnerabilities which makes them harder to tie down.
-For other vulnerabilities which might affect all Android devices it is hard to work out which devices they affect.
-For example \emph{pty race}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/pty_race}} is a Linux kernel vulnerability, we found 48 commits which fixed it in different branches, it was also accidentally fixed and later reintroduced and so determining which devices were vulnerable relies on knowing where on which branch the kernel was taken from.
-For some vulnerabilities the corresponding patch is not in AOSP despite some device manufacturers having shipped builds containing the fix, which makes working out what is going on more difficult.
-For \emph{RageAgainstTheCage adb}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/RageAgainstTheCage_adb}} and \emph{keystore buffer}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/keystore_buffer}} vulnerabilities which we do have sufficient data it is not clear whether they are truly critical vulnerabilities as the former may require physical ADB access and the latter may be protected by a sandbox.
-However those two vulnerabilities make little difference to our results as they affect few devices.
+We have not included all the vulnerabilities from \avo. 
+In particular, we have not included any device manufacturer specific vulnerabilities, even when these are widespread (such as vulnerabilities affecting Qualcomm\footnote{Qualcomm is particularly good at publicly disclosing the vulnerabilities affecting their code and the patches which fix them.\url{https://www.codeaurora.org/projects/security-advisories/}} chipsets) because it is difficult to work out which devices are affected.
+There also tends to be less public information available about manufacturer-specific vulnerabilities which makes them harder to tie down.
 
-\subsection{Open questions}
-There are several open questions which we have not yet been able to answer and remain future work.
-\begin{itemize}
- \item Why was there a gap in vulnerability discovery in Android in 2012?
- \item Why do so many devices have ADB enabled? \daAdbEnabledPerc\ seems rather high for a developer feature. This could be a selection bias or device manufacturers could be enabling it by default.
- \item How much of updating is new handsets and how much is updates being deployed?
-Over time newer versions of Android come to dominate, but quantifying what proportion of that is due to new phones being purchased and what proportion is due to updates to existing devices is a question we cannot answer.
- The only available large data set we know of with longitudinal traces of Android devices OS version numbers is \da\ but that only records \daNumUpdatesUpgrades\ upgrades which broken down over \daOSMonthsOfData\ months and \daNumSigOSVersions\ OS versions is only \daUpdatesPerMonthPerVersion\ updates per version per month which is not enough to build a statistically significant prediction of the expected transition to newer versions of Android.
- \item Why do users downgrade their phones?
-\end{itemize}
+For some other vulnerabilities which might affect all Android devices it is hard to work out which devices are affected.
+For example, \emph{pty race}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/pty_race}} is a Linux kernel vulnerability.
+We found 48 commits which fixed this vulnerability on different branches.
+On some branches it was also accidentally fixed, and later reintroduced, before finally being fixed properly, and therefore determining which devices were vulnerable relies on knowing where, and on which branch, the kernel code for a specific release was taken from.
+For some vulnerabilities the required patch is not in AOSP.
+This is despite the fact that some device manufacturers have shipped builds containing the fix.
+This also made it difficult to determine whether a specific device was vulnerable or not.
+For \emph{RageAgainstTheCage adb}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/RageAgainstTheCage_adb}} and \emph{keystore buffer}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/keystore_buffer}} vulnerabilities we do not have sufficient data to determine whether they fit one of our attack vectors as the former may require physical ADB access and the latter may be thwarted by the sandbox.
+However a sensitivity analysis showed that those two vulnerabilities make little difference to our results as they affect few devices.
+
+We still do not know where new versions of Android come from.
+Over time newer versions of Android come to dominate, but quantifying the proportion due to new phone purchases verses the proportion due to updates of existing devices is unknown.
+ Device Analyzer is the only available large data set we know of with longitudinal traces of Android OS version strings.
+ Unfortunately this data set only records \daNumUpdatesUpgrades\ upgrades which, broken down over \daOSMonthsOfData\ months and \daNumSigOSVersions\ OS versions is only \daUpdatesPerMonthPerVersion\ updates per version per month.
+This is not enough to build a statistically significant prediction of the expected transition to newer versions of Android.
 

### 2014-11-14 security_scoring.tex
Make full_version_comp use the full page width
'maybe insecure' to 'maybe secure' and minor fixes

@@
 To allow buyers of Android devices to purchase those devices which have the best security they need to know how different device manufacturers, device models and network operators compare in terms of the security they provide.
-We propose a method to score a device manufacturer, network operator or device model based on its historic performance at keeping devices up to date and fixing security vulnerabilities.
+We propose a method to score a device manufacturer, device model or network operator based on its historic performance at keeping devices up to date and fixing security vulnerabilities.
 We find that Android as a whole gets a score of \daSecurityScore\ out of 10, the highest scoring device manufacturer is \emph{\daSecScoreBestmanufacturer} (\daSecScoreBestmanufacturerScore\ out of 10) and the lowest scoring is \emph{\daSecScoreWorstmanufacturer} (\daSecScoreWorstmanufacturerScore\ out of 10).
@@
 Since the numbers involved are large, the uncertainty in $f$ is small.
-$u$ is computed by taking the sum of the proportions of devices running the most recent version each day, both the count of devices running the maximum version and total count have sqrt uncertainties.
+$u$ is computed by taking the sum of the proportions of devices running the most recent version each day, both the count of devices running the maximum version and total count have square root uncertainties.
 The Python \texttt{uncertainties} library was used to propagate uncertainties through calculations.
@@
 \subsubsection{Results: Security scores}
-The scores across the whole of android are that \daMeanInsecurityPerc\ of devices are exposed to known root exploits.
+The scores across the whole of Android are that \daMeanInsecurityPerc\ of devices are exposed to known root exploits.
 There are on average \daMeanOutstandingVulnerabilities\ outstanding vulnerabilities not fixed on any device.
@@
 
-\begin{figure}
+\begin{figure*}
  \centering
- \includegraphics[width=\columnwidth]{figures/full_version_comp.pdf}
+ \includegraphics[width=\textwidth]{figures/full_version_comp.pdf}
  \caption{Full version distributions for the highest and lowest scoring model}
  \label{fig:full_version_comp}
-\end{figure}
+\end{figure*}
 

### 2014-11-14 securityupdates.tex
'maybe insecure' to 'maybe secure' and minor fixes

@@
 \includegraphics[width=\columnwidth]{figures/proportioninsecure}
-\caption{Proportion of devices running insecure, maybe insecure and secure versions of Android against time.
+\caption{Proportion of devices running insecure, maybe secure and secure versions of Android against time.
 The red vertical lines are caused by vulnerabilities being discovered with those which have the biggest impact annotated.
@@
 (2)~\emph{insecure} because the device was running an insecure version of Android with a build released before the vulnerability was fixed or discovered; and 
-(3)~\emph{maybe insecure} because the device is running a version of Android known to be insecure but has received an update which might contain a backported fix as we did not observe that build number until after the vulnerability was known.
+(3)~\emph{maybe secure} because the device is running a version of Android known to be insecure but has received an update which might contain a backported fix as we did not observe that build number until after the vulnerability was known.
 Further analysis into the propagation of updates in Android is presented in \S\ref{sec:exp:android_ecosystem}.
@@
  \caption{Proportion of devices affected by different vulnerabilities.
-The prefix `(maybe)' indicates `maybe insecure' and `(definite)' indicates `definitely insecure'.
-The first few vulnerabilities are all `maybe insecure' as we do not have data from before that vulnerability was discovered while later ones are mostly `definitely insecure' as we know no fixing update reached the devices.
+The prefix `(maybe)' indicates `maybe secure' and `(definite)' indicates `definitely insecure'.
+The first few vulnerabilities are all `maybe secure' as we do not have data from before that vulnerability was discovered while later ones are mostly `definitely insecure' as we know no fixing update reached the devices.
 The change in behaviour after August 2014 is explained in \S\ref{sec:da_changes}}
@@
 Unfortunately there is no authoritative source of OS version information and so we cannot check directly whether our data is representative.
-However Google has published API version information every month since December 2009 and we have collated this information.\footnote{\url{http://androidvulnerabilities.org/play/historicplaydashboard}}
+However Google has published API version information almost every month since December 2009 and we have collated this information.\footnote{\url{http://androidvulnerabilities.org/play/historicplaydashboard}}
 While API versions are too coarse grained to use for security update detection they are closely related to OS versions and so if the \da\ data on API versions is similar to the Google Data on API versions then the \da\ data on OS versions should be representative.
@@
 We have not included all the vulnerabilities from \avo. 
-In particular, we have not included any device manufacturer specific vulnerabilities, even when these are widespread (such as vulnerabilities affecting Qualcomm\footnote{Qualcomm is particularly good at publicly disclosing the vulnerabilities affecting their code and the patches which fix them.\url{https://www.codeaurora.org/projects/security-advisories/}} chipsets) because it is difficult to work out which devices are affected.
+In particular, we have not included any device manufacturer specific vulnerabilities, even when these are widespread (such as vulnerabilities affecting Qualcomm\footnote{Qualcomm is particularly good at publicly disclosing the vulnerabilities affecting their code and the patches which fix them.
+\url{https://www.codeaurora.org/projects/security-advisories/}} chipsets) because it is difficult to work out which devices are affected.
 There also tends to be less public information available about manufacturer-specific vulnerabilities which makes them harder to tie down.
@@
 Unfortunately the time period between the \textbf{fixing release} and \textbf{fix deployed} can be of the order of months. 
-For example, the zergRush vulnerability was fixed in a release before it was publicly discovered but it still took 27 months for 90\% of the devices in our study to be definitely fixed\footnote{Definitely fixed is 1 $-$ (definitely insecure + maybe insecure), maybe fixed is 1 $-$ definitely insecure} (25 months for 90\% of devices maybe fixed).
+For example, the zergRush vulnerability was fixed in a release before it was publicly discovered but it still took 27 months for 90\% of the devices in our study to be definitely fixed\footnote{Definitely fixed is 1 $-$ (definitely insecure + maybe secure), maybe fixed is 1 $-$ definitely insecure} (25 months for 90\% of devices maybe fixed).
 

### 2014-11-14 securityupdates.tex
Updated dastats for more consistent reporting of numbers

@@
 We present the results of our analysis showing that, on average, \daMeanInsecurityPerc\ of Android devices are exposed to critical vulnerabilities.
-Devices, on average, apply \daUpdatesPerYear\ updates each year and \daPercUpdatesDowngrades\ of version changes are downgrades to older versions.
+Devices, on average, apply \daUpdatesPerYear\ updates each year and \daPercUpdatesDowngrades\ of full version changes are downgrades to older versions.
 Nexus devices are better than others, \emph{\daSecScoreBestmanufacturer} is the highest scoring device manufacturer and \emph{\daSecScoreBestoperator} is the highest scoring network operator.
@@
 Mostly the dark cells are upgrades (above the diagonal) (\daNumUpdatesUpgrades).
-While many upgrades are from one version to the next version there are also a fair number (\daNumUpdatesBigUpgrades, \daPercBigUpgrades) which skip more than \daNumUpdatesSkippedBig\ versions.
-Surprisingly there are also a small number of downgrade events (\daNumUpdatesDowngrades, \daPercUpdatesDowngrades) when older versions of Android are installed on to devices.
+While many upgrades are from one version to the next version there are also a fair number (\daNumUpdatesBigUpgrades, \daPercBigUpgradesNominal) which skip more than \daNumUpdatesSkippedBig\ versions.
+Surprisingly there are also a small number of downgrade events (\daNumUpdatesDowngrades, \daPercUpdatesDowngradesNominal) when older versions of Android are installed on to devices.
 Possible reasons why users are downgrading are to free up space on their device, to make it easier to root or because a new version introduced bugs.

### 2014-11-14 securityupdates.tex
Minor fixes

@@
 (ii) device manufacturers made it easier to install custom versions of Android, reducing the need for users to root their devices; and
-(iii) device manufacturer specific vulnerabilities (which we are not in our analysis here) were easier to find and therefore attackers looking for vulnerabilities adjusted their focus.
+(iii) device manufacturer specific vulnerabilities (which are not in our analysis here) were easier to find and therefore attackers looking for vulnerabilities adjusted their focus.
 
@@
 However most devices only contribute for a short period of time and so we do not observe updates happening on every device.
-Instead we have a hopefully representative sample of upgrades which happened while \da\ was installed on the devices.
+Instead we have a hopefully representative sample of updates which happened while \da\ was installed on the devices.
 \da\ cannot distinguish between a device being replaced and the \da\ app being removed as while multiple devices can be linked to the same user account, that is a manual process and few users do it.

### 2015-01-15 rebuttal.txt
Add draft rebuttal text

REVIEWER #59A

The plural of anecdote is not data. In this context, this paper is the
solution to the problem in Android updates: public data quantifying
which manufacturers and devices are good, and which are bad, is the way
in which we will see industry improvements in security. We will publish
(and update this information) on androidvulnerabilities.org.

In terms of sensitivity of our results to the anomaly caused by the
Bangladesh study: Our headline figure of devices being vulnerable to root
exploits of 87.6% of the time becomes 87.7% if we discard all data after
the start of the study.

Additionally the bias in the data set doesn't affect our main conclusions
about particular manufacturers or device models (i.e. "buy a Nexus device").
Within a device model update behaviour appears to be fairly homogeneous, and
so bias will affect which device models we have significant data for but not
the results for a device model.

To put in the paper (we perhaps should have included this data in the
first version of the paper, sorry): when we see an update on one device,
we see all other devices with the same make and model receive the
update, regardless of Operator. This means that the operator and user
are unlikely to be the problem. This implies the bottleneck is with
Google and/or the Manufacturer. We could put more data on this in the paper.
We want to be able to say something like (BY FRIDAY!): "By looking at our data, 90% of
devices update to a newly available build number within three weeks.
This is true for all devices for which we have substantial number,
including Nexus 4, XXX, YYY, ZZZ."

Short lifespan is not a problem: we are simply getting a sample from all
the devices.

We have zero reported cases where Device Analyzer has been broken by an
OS update. Apps being broken by OS updates is unlikely to be acceptable to
users and so Google works very hard to ensure backwards compatibility.

We don't believe Google at fault: the fact that Nexus devices are better
than others means there is a difference between manufacturers. We
believe it is reasonable to assume that Google tells other manufacturers
at the same time that they fix Nexus builds. We also know there are long
delays between a Google AOSP release and a manufacturer update.
Therefore the problem lies with the manufacturers.

"On the other hand, when the proportion of vulnerable devices drops,
it might be due to vulnerable devices leaving, or patched devices
joining." -> If a user stops using their phone because it is vulnerable,
then this is okay: a vulnerable phone is no longer in circulation. If a
user discovers their phone is vulnerable, and uninstalls DA but carries
on using it, then this is a problem, but we think this is unlikely.

We can expand Table II if the reviewers think this is useful.

REVIEWER #59B

We don't claim to make a technical contribution. The call states that
the conference is looking for "advances in the theory, design,
implementation, analysis, verification, or empirical evaluation and
measurement of secure systems." We believe we have provided a thorough
and detailed empirical evaluation and measurement of Android.

The sample bias and the short periods of device participation are not a
problem as explained above.

The user is unlikely to be at fault. Once we see an update for one
make/model, we see it deployed quickly and widely as discussed above,
which suggests that the problem lies in the lack of an update to install.

We explain in the paper the reason for the exp in the scoring function just
before Equation 1.

We can of course construct any scoring function we like! We have explained
why the three variables are useful in the paper. A linear combination of
f, u and 1/exp(m) is reasonable because we are interested in devices which
perform well against all three measures.

We would happily present tabular data as bar charts if preferred.

### 2015-01-16 rebuttal.txt
Updated rebuttal

@@
-REVIEWER #59A
+We don't claim to make a technical contribution.
+The call is for "advances in the theory, design, implementation, analysis, verification, or empirical evaluation and measurement of secure systems."
+We have provided an empirical evaluation and measurement of Android.
 
-The plural of anecdote is not data. In this context, this paper is the
-solution to the problem in Android updates: public data quantifying
-which manufacturers and devices are good, and which are bad, is the way
-in which we will see industry improvements in security. We will publish
-(and update this information) on androidvulnerabilities.org.
+We have confirmed the conventional wisdom, but provide solid evidence.
+In this context, this paper is the solution to the problem of slow Android updates: public data comparing manufacturers and device models, provides market pressure for industry improvements in security.
+We will publish (and update) this information on androidvulnerabilities.org.
 
-In terms of sensitivity of our results to the anomaly caused by the
-Bangladesh study: Our headline figure of devices being vulnerable to root
-exploits of 87.6% of the time becomes 87.7% if we discard all data after
-the start of the study.
+Regarding the Bangladesh study anomaly:
+Truncating our data before it starts increases the percentage of vulnerable devices from 87.6% to 87.7% and decreases the percentage of devices running the latest version from 5.58% to 5.43%.
 
-Additionally the bias in the data set doesn't affect our main conclusions
+When we see a new version on one device, often we see all other devices with the same model receive the update promptly, regardless of operator or user behaviour.
+By looking at our data, 50% of devices update to a newly available build number within 26 days, with a long tail thereafter.
+Update behaviour is consistent within device models with some being fast and others slow to deploy.
+This means that the user is unlikely to be the problem and that the bottleneck is with Google, the Manufacturer or, for some devices, the Operator.
+This is based on the significant devices shown in Table V.
+
+This means that the bias in the data doesn't affect our main conclusions
 about particular manufacturers or device models (i.e. "buy a Nexus device").
-Within a device model update behaviour appears to be fairly homogeneous, and
-so bias will affect which device models we have significant data for but not
-the results for a device model.
+While bias will affect which device models we analyse it will not affect the result of that analysis.
 
-To put in the paper (we perhaps should have included this data in the
-first version of the paper, sorry): when we see an update on one device,
-we see all other devices with the same make and model receive the
-update, regardless of Operator. This means that the operator and user
-are unlikely to be the problem. This implies the bottleneck is with
-Google and/or the Manufacturer. We could put more data on this in the paper.
-We want to be able to say something like (BY FRIDAY!): "By looking at our data, 90% of
-devices update to a newly available build number within three weeks.
-This is true for all devices for which we have substantial number,
-including Nexus 4, XXX, YYY, ZZZ."
+Short contributions are not a problem as we are getting a sample from the distribution of all the devices.
 
-Short lifespan is not a problem: we are simply getting a sample from all
-the devices.
-
-We have zero reported cases where Device Analyzer has been broken by an
-OS update. Apps being broken by OS updates is unlikely to be acceptable to
-users and so Google works very hard to ensure backwards compatibility.
-
-We don't believe Google at fault: the fact that Nexus devices are better
-than others means there is a difference between manufacturers. We
-believe it is reasonable to assume that Google tells other manufacturers
-at the same time that they fix Nexus builds. We also know there are long
-delays between a Google AOSP release and a manufacturer update.
-Therefore the problem lies with the manufacturers.
+We have zero cases where Device Analyzer has been broken by an OS update, Android has good backwards compatibility.
 
@@
 joining." -> If a user stops using their phone because it is vulnerable,
-then this is okay: a vulnerable phone is no longer in circulation. If a
-user discovers their phone is vulnerable, and uninstalls DA but carries
-on using it, then this is a problem, but we think this is unlikely.
+then this is okay: as it is no longer in circulation. It is unlikely that a
+user would discover their phone is vulnerable, and uninstall DA but carry
+on using it.
 
-We can expand Table II if the reviewers think this is useful.
+We the paper explains the reason for the exp in the scoring function just before Equation 1.
 
-REVIEWER #59B
-
-We don't claim to make a technical contribution. The call states that
-the conference is looking for "advances in the theory, design,
-implementation, analysis, verification, or empirical evaluation and
-measurement of secure systems." We believe we have provided a thorough
-and detailed empirical evaluation and measurement of Android.
-
-The sample bias and the short periods of device participation are not a
-problem as explained above.
-
-The user is unlikely to be at fault. Once we see an update for one
-make/model, we see it deployed quickly and widely as discussed above,
-which suggests that the problem lies in the lack of an update to install.
-
-We explain in the paper the reason for the exp in the scoring function just
-before Equation 1.
-
-We can of course construct any scoring function we like! We have explained
-why the three variables are useful in the paper. A linear combination of
+We can of course construct any scoring function we like! The paper explains
+why the three variables are useful. A linear combination of
 f, u and 1/exp(m) is reasonable because we are interested in devices which
@@
 
-We would happily present tabular data as bar charts if preferred.
+We can expand Table II if useful.

### 2015-03-16 dastats.tex
Updated dastats for more consistent reporting of numbers
Update graphs and data

@@
-\newcommand{\daNumDataPoints}{153 Billion}
+\newcommand{\daNumDataPoints}{144 Billion}
 \newcommand{\daNumStartedApps}{122\,000}
@@
 \newcommand{\daSigNumDays}{100}
-\newcommand{\daNumManufacturers}{261}
-\newcommand{\daNumSigManufacturers}{9}
-\newcommand{\daNumModels}{2\,380}
-\newcommand{\daNumSigModels}{15}
-\newcommand{\daOSVersionPercValidLines}{99.5\%}
-\newcommand{\daOSTotalDaysData}{1\,140\,000}
-\newcommand{\daMeanInsecurityPerc}{$87.3 \pm 0.0\%$}
-\newcommand{\daMeanOutstandingVulnerabilities}{0.637}
-\newcommand{\daUpdatedness}{$0.0558 \pm 0.0002$}
-\newcommand{\daSecurityScore}{$2.75 \pm 0.00$}
-\newcommand{\daNumFullVersions}{1\,140}
-\newcommand{\daNumOSVersions}{45}
-\newcommand{\daNumSigOSVersions}{23}
+\newcommand{\daNumManufacturers}{292}
+\newcommand{\daNumSigManufacturers}{8}
+\newcommand{\daNumModels}{2\,620}
+\newcommand{\daNumSigModels}{17}
+\newcommand{\daOSVersionPercValidLines}{99.3\%}
+\newcommand{\daOSTotalDaysData}{1\,250\,000}
+\newcommand{\daMeanInsecurityPerc}{$88.0 \pm 0.0\%$}
+\newcommand{\daMeanOutstandingVulnerabilities}{0.582}
+\newcommand{\daUpdatedness}{$0.0521 \pm 0.0002$}
+\newcommand{\daSecurityScore}{$2.79 \pm 0.00$}
+\newcommand{\daNumFullVersions}{1\,220}
+\newcommand{\daNumOSVersions}{48}
+\newcommand{\daNumSigOSVersions}{26}
 \newcommand{\daNumAPIVersions}{15}
-\newcommand{\daNumFullVersionUpdates}{4\,220}
-\newcommand{\daNumOSVersionUpdates}{3\,560}
-\newcommand{\daNumFullOnlyVersionUpdates}{640}
-\newcommand{\daNumSecurityUpdates}{1\,110}
-\newcommand{\daNumPossibleSecurityUpdates}{2\,140}
-\newcommand{\daTabSecScoresmanufacturer}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{Security score} \\ &&&& (out of 10) \\ \hline LG & $0.12 \pm 0.00$ & $0.32 \pm 0.00$ & 0.74 & $3.38 \pm 0.01$ \\  Motorola & $0.16 \pm 0.00$ & $0.13 \pm 0.00$ & 0.77 & $2.93 \pm 0.01$ \\  other & $0.13 \pm 0.00$ & $0.06 \pm 0.00$ & 0.64 & $2.75 \pm 0.00$ \\  Samsung & $0.14 \pm 0.00$ & $0.05 \pm 0.00$ & 0.71 & $2.67 \pm 0.00$ \\  Sony & $0.13 \pm 0.00$ & $0.18 \pm 0.00$ & 1.02 & $2.64 \pm 0.01$ \\  HTC & $0.14 \pm 0.00$ & $0.09 \pm 0.00$ & 0.89 & $2.55 \pm 0.00$ \\  asus & $0.15 \pm 0.00$ & $0.56 \pm 0.01$ & 6.06 & $2.30 \pm 0.02$ \\  Symphony & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & 4.79 & $0.19 \pm 0.00$ \\  walton & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & 5.79 & $0.14 \pm 0.01$ \\ \end{tabular} \caption{Security scores for manufacturers} \label{tab:sec_manufacturer} \end{table}}
-\newcommand{\daTabSecScoresmodel}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{Security score} \\ &&&& (out of 10) \\ \hline Galaxy Nexus & $0.53 \pm 0.00$ & $0.53 \pm 0.01$ & 1.53 & $4.80 \pm 0.03$ \\  Nexus 4 & $0.23 \pm 0.00$ & $0.82 \pm 0.01$ & 6.06 & $3.39 \pm 0.04$ \\  Nexus 7 & $0.19 \pm 0.00$ & $0.74 \pm 0.01$ & 5.92 & $3.02 \pm 0.04$ \\  other & $0.13 \pm 0.00$ & $0.12 \pm 0.00$ & 0.64 & $2.94 \pm 0.00$ \\  Desire HD & $0.08 \pm 0.00$ & $0.05 \pm 0.00$ & 0.39 & $2.91 \pm 0.02$ \\  HTC Sensation & $0.36 \pm 0.00$ & $0.01 \pm 0.01$ & 1.59 & $2.47 \pm 0.02$ \\  GT-I9100 & $0.22 \pm 0.00$ & $0.02 \pm 0.00$ & 1.20 & $2.33 \pm 0.01$ \\  HTC Desire S & $0.02 \pm 0.00$ & $0.02 \pm 0.00$ & 1.00 & $1.75 \pm 0.01$ \\  GT-N7000 & $0.25 \pm 0.00$ & $0.00 \pm 0.00$ & 2.52 & $1.43 \pm 0.02$ \\  GT-P1000 & $0.01 \pm 0.00$ & $0.00 \pm 0.01$ & 1.73 & $0.94 \pm 0.02$ \\  GT-I9300 & $0.15 \pm 0.00$ & $0.01 \pm 0.00$ & 6.15 & $0.64 \pm 0.01$ \\  GT-I9505 & $0.00 \pm 0.00$ & $0.18 \pm 0.00$ & 6.77 & $0.55 \pm 0.01$ \\  GT-N7100 & $0.07 \pm 0.00$ & $0.00 \pm 0.01$ & 6.91 & $0.28 \pm 0.02$ \\  HTC Desire HD & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & 3.05 & $0.27 \pm 0.02$ \\ \end{tabular} \caption{Security scores for models} \label{tab:sec_model} \end{table}}
-\newcommand{\daTabSecScoressummary}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{Security score} \\ &&&& (out of 10) \\ \hline nexus & $0.36 \pm 0.00$ & $0.51 \pm 0.00$ & 0.68 & $5.01 \pm 0.01$ \\  notnexus & $0.12 \pm 0.00$ & $0.04 \pm 0.00$ & 0.64 & $2.67 \pm 0.00$ \\ \end{tabular} \caption{Security scores for nexus} \label{tab:sec_summary} \end{table}}
-\newcommand{\daUpdatednessPerc}{$5.58 \pm 0.02\%$}
-\newcommand{\daNumOSDevices}{19\,000}
+\newcommand{\daNumFullVersionUpdates}{4\,400}
+\newcommand{\daNumOSVersionUpdates}{3\,720}
+\newcommand{\daNumFullOnlyVersionUpdates}{663}
+\newcommand{\daNumSecurityUpdates}{1\,160}
+\newcommand{\daNumPossibleSecurityUpdates}{2\,240}
+\newcommand{\daTabSecScoresmanufacturer}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{Security score} \\ &&&& (out of 10) \\ \hline LG & $0.18 \pm 0.00$ & $0.32 \pm 0.00$ & 0.67 & $3.70 \pm 0.01$ \\  Motorola & $0.16 \pm 0.00$ & $0.13 \pm 0.00$ & 0.78 & $2.93 \pm 0.01$ \\  Samsung & $0.13 \pm 0.00$ & $0.04 \pm 0.00$ & 0.67 & $2.67 \pm 0.00$ \\  Sony & $0.12 \pm 0.00$ & $0.19 \pm 0.00$ & 1.11 & $2.55 \pm 0.01$ \\  HTC & $0.13 \pm 0.00$ & $0.08 \pm 0.00$ & 0.94 & $2.48 \pm 0.00$ \\  asus & $0.18 \pm 0.00$ & $0.52 \pm 0.01$ & 6.01 & $2.30 \pm 0.02$ \\  other & $0.06 \pm 0.00$ & $0.04 \pm 0.00$ & 1.11 & $1.87 \pm 0.00$ \\  Symphony & $0.00 \pm 0.00$ & $0.07 \pm 0.00$ & 4.98 & $0.25 \pm 0.00$ \\  walton & $0.00 \pm 0.00$ & $0.07 \pm 0.00$ & 5.97 & $0.21 \pm 0.01$ \\ \end{tabular} \caption{Security scores for manufacturers} \label{tab:sec_manufacturer} \end{table}}
+\newcommand{\daTabSecScoresmodel}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{Security score} \\ &&&& (out of 10) \\ \hline Galaxy Nexus & $0.51 \pm 0.00$ & $0.53 \pm 0.01$ & 1.53 & $4.71 \pm 0.03$ \\  Nexus 4 & $0.29 \pm 0.00$ & $0.82 \pm 0.01$ & 6.06 & $3.63 \pm 0.04$ \\  Nexus 7 & $0.24 \pm 0.00$ & $0.74 \pm 0.01$ & 5.92 & $3.19 \pm 0.04$ \\  other & $0.09 \pm 0.00$ & $0.14 \pm 0.00$ & 0.58 & $2.92 \pm 0.00$ \\  Desire HD & $0.08 \pm 0.00$ & $0.05 \pm 0.00$ & 0.39 & $2.91 \pm 0.02$ \\  HTC Sensation & $0.35 \pm 0.00$ & $0.01 \pm 0.01$ & 1.59 & $2.43 \pm 0.02$ \\  GT-I9100 & $0.22 \pm 0.00$ & $0.02 \pm 0.00$ & 1.23 & $2.28 \pm 0.01$ \\  HTC Desire S & $0.02 \pm 0.00$ & $0.02 \pm 0.00$ & 1.00 & $1.75 \pm 0.01$ \\  GT-N7000 & $0.24 \pm 0.00$ & $0.00 \pm 0.00$ & 2.53 & $1.42 \pm 0.02$ \\  GT-P1000 & $0.01 \pm 0.00$ & $0.00 \pm 0.01$ & 1.73 & $0.94 \pm 0.02$ \\  GT-I9300 & $0.14 \pm 0.00$ & $0.01 \pm 0.00$ & 6.28 & $0.59 \pm 0.01$ \\  GT-I9505 & $0.00 \pm 0.00$ & $0.14 \pm 0.00$ & 6.86 & $0.44 \pm 0.01$ \\  HTC Desire HD & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & 3.05 & $0.27 \pm 0.02$ \\  GT-N7100 & $0.06 \pm 0.00$ & $0.00 \pm 0.01$ & 6.90 & $0.26 \pm 0.02$ \\  Symphony W68 & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & 11.00 & $0.00 \pm 0.03$ \\ \end{tabular} \caption{Security scores for models} \label{tab:sec_model} \end{table}}
+\newcommand{\daTabSecScoressummary}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{Security score} \\ &&&& (out of 10) \\ \hline nexus & $0.40 \pm 0.00$ & $0.49 \pm 0.00$ & 0.62 & $5.14 \pm 0.01$ \\  notnexus & $0.10 \pm 0.00$ & $0.02 \pm 0.00$ & 0.58 & $2.61 \pm 0.00$ \\ \end{tabular} \caption{Security scores for nexus} \label{tab:sec_summary} \end{table}}
+\newcommand{\daUpdatednessPerc}{$5.21 \pm 0.02\%$}
+\newcommand{\daNumOSDevices}{20\,000}
 \newcommand{\daSigNumDevicesDay}{20}
@@
 \newcommand{\daSecScoreBestmodel}{Galaxy Nexus}
-\newcommand{\daSecScoreWorstmodel}{HTC Desire HD A9191}
+\newcommand{\daSecScoreWorstmodel}{Symphony W68}
 \newcommand{\daSecScoreBestsummary}{nexus}
 \newcommand{\daSecScoreWorstsummary}{notnexus}
-\newcommand{\daSecScoreBestmanufacturerScore}{$3.38 \pm 0.01$}
-\newcommand{\daSecScoreWorstmanufacturerScore}{$0.139 \pm 0.006$}
-\newcommand{\daSecScoreBestmodelScore}{$4.8 \pm 0.0$}
-\newcommand{\daSecScoreWorstmodelScore}{$0.274 \pm 0.020$}
+\newcommand{\daSecScoreBestmanufacturerScore}{$3.7 \pm 0.0$}
+\newcommand{\daSecScoreWorstmanufacturerScore}{$0.214 \pm 0.006$}
+\newcommand{\daSecScoreBestmodelScore}{$4.71 \pm 0.02$}
+\newcommand{\daSecScoreWorstmodelScore}{$0.0001 \pm 0.0281$}
 \newcommand{\daSecScoreBestsummaryScore}{$4.1 \pm 0.0$}
 \newcommand{\daSecScoreWorstsummaryScore}{$2.9 \pm 0.0$}
-\newcommand{\daVulnFree}{$0.127 \pm 0.000$}
-\newcommand{\daAdbEnabledPerc}{$20 \pm 0\%$}
-\newcommand{\daNumUpdatesUpgrades}{3\,330}
-\newcommand{\daNumUpdatesDowngrades}{147}
-\newcommand{\daPercUpdatesDowngrades}{4.23\%}
+\newcommand{\daVulnFree}{$0.12 \pm 0.00$}
+\newcommand{\daAdbEnabledPerc}{$19.3 \pm 0.0\%$}
+\newcommand{\daNumUpdatesUpgrades}{3\,480}
+\newcommand{\daNumUpdatesDowngrades}{151}
+\newcommand{\daPercUpdatesDowngrades}{$3.52 \pm 0.29\%$}
 \newcommand{\daStartDate}{2011-07-01}
-\newcommand{\daEndDate}{2014-11-07}
-\newcommand{\daNumOperators}{1\,390}
-\newcommand{\daTabSecScoresoperator}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{Security score} \\ &&&& (out of 10) \\ \hline O2 uk & $0.25 \pm 0.00$ & $0.12 \pm 0.00$ & 0.37 & $3.79 \pm 0.01$ \\  T-Mobile & $0.19 \pm 0.00$ & $0.18 \pm 0.00$ & 0.48 & $3.57 \pm 0.01$ \\  Orange & $0.21 \pm 0.00$ & $0.09 \pm 0.00$ & 0.39 & $3.52 \pm 0.02$ \\  Sprint & $0.18 \pm 0.00$ & $0.11 \pm 0.00$ & 0.44 & $3.41 \pm 0.01$ \\  3 & $0.15 \pm 0.00$ & $0.09 \pm 0.00$ & 0.49 & $3.15 \pm 0.01$ \\  AT\&T & $0.12 \pm 0.00$ & $0.08 \pm 0.00$ & 0.43 & $3.08 \pm 0.01$ \\  Vodafone uk & $0.13 \pm 0.00$ & $0.10 \pm 0.00$ & 0.52 & $3.07 \pm 0.01$ \\  Verizon & $0.18 \pm 0.00$ & $0.09 \pm 0.00$ & 0.81 & $2.82 \pm 0.01$ \\  unknown & $0.10 \pm 0.00$ & $0.22 \pm 0.00$ & 1.08 & $2.57 \pm 0.01$ \\  n Telenor & $0.03 \pm 0.00$ & $0.10 \pm 0.00$ & 1.42 & $1.59 \pm 0.01$ \\  Airtel & $0.06 \pm 0.00$ & $0.03 \pm 0.00$ & 1.93 & $1.10 \pm 0.01$ \\  Grameenphone & $0.01 \pm 0.00$ & $0.02 \pm 0.00$ & 1.99 & $0.80 \pm 0.00$ \\  Robi & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & 2.19 & $0.73 \pm 0.01$ \\  banglalink & $0.00 \pm 0.00$ & $0.03 \pm 0.00$ & 2.90 & $0.40 \pm 0.00$ \\ \end{tabular} \caption{Security scores for operators} \label{tab:sec_operator} \end{table}}
+\newcommand{\daEndDate}{2015-03-04}
+\newcommand{\daNumOperators}{1\,420}
+\newcommand{\daTabSecScoresoperator}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{Security score} \\ &&&& (out of 10) \\ \hline O2 uk & $0.25 \pm 0.00$ & $0.12 \pm 0.00$ & 0.37 & $3.82 \pm 0.01$ \\  T-Mobile & $0.20 \pm 0.00$ & $0.18 \pm 0.00$ & 0.48 & $3.62 \pm 0.01$ \\  Orange & $0.21 \pm 0.00$ & $0.09 \pm 0.00$ & 0.39 & $3.52 \pm 0.02$ \\  Sprint & $0.18 \pm 0.00$ & $0.11 \pm 0.00$ & 0.44 & $3.39 \pm 0.01$ \\  3 & $0.16 \pm 0.00$ & $0.09 \pm 0.00$ & 0.49 & $3.20 \pm 0.01$ \\  Vodafone uk & $0.14 \pm 0.00$ & $0.13 \pm 0.00$ & 0.50 & $3.20 \pm 0.01$ \\  AT\&T & $0.12 \pm 0.00$ & $0.08 \pm 0.00$ & 0.43 & $3.07 \pm 0.01$ \\  Verizon & $0.18 \pm 0.00$ & $0.09 \pm 0.00$ & 0.82 & $2.82 \pm 0.01$ \\  unknown & $0.10 \pm 0.00$ & $0.21 \pm 0.00$ & 0.95 & $2.72 \pm 0.01$ \\  n Telenor & $0.03 \pm 0.00$ & $0.11 \pm 0.00$ & 1.43 & $1.59 \pm 0.01$ \\  Airtel & $0.05 \pm 0.00$ & $0.03 \pm 0.00$ & 1.87 & $1.08 \pm 0.01$ \\  Grameenphone & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & 2.02 & $0.85 \pm 0.00$ \\  Robi & $0.00 \pm 0.00$ & $0.06 \pm 0.00$ & 2.09 & $0.84 \pm 0.01$ \\  banglalink & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & 2.71 & $0.48 \pm 0.00$ \\ \end{tabular} \caption{Security scores for operators} \label{tab:sec_operator} \end{table}}
 \newcommand{\daSecScoreBestoperator}{O2 uk}
-\newcommand{\daSecScoreBestoperatorScore}{$3.79 \pm 0.01$}
+\newcommand{\daSecScoreBestoperatorScore}{$3.82 \pm 0.01$}
 \newcommand{\daSecScoreWorstoperator}{banglalink}
-\newcommand{\daSecScoreWorstoperatorScore}{$0.4 \pm 0.0$}
+\newcommand{\daSecScoreWorstoperatorScore}{$0.485 \pm 0.004$}
 \newcommand{\daNumSigOperators}{14}
-\newcommand{\daMonthsDevices}{1545}
+\newcommand{\daMonthsDevices}{1\,680}
 \newcommand{\daMonths}{6}
@@
 \newcommand{\daNumDeviceDataDevices}{50}
-\newcommand{\daNumSigFullVersions}{83}
-\newcommand{\daSecScoreBestmanufacturerNumFullVersions}{104}
+\newcommand{\daNumSigFullVersions}{87}
+\newcommand{\daSecScoreBestmanufacturerNumFullVersions}{122}
 \newcommand{\daSecScoreworstmanufacturerNumFullVersions}{239}
@@
 \newcommand{\daSecScoreworstsummaryNumFullVersions}{1057}
-\newcommand{\daSecScoreBestoperatorNumFullVersions}{88}
+\newcommand{\daSecScoreBestoperatorNumFullVersions}{92}
 \newcommand{\daSecScoreworstoperatorNumFullVersions}{89}
 \newcommand{\daSecScoreWorstmanufacturerNumFullVersions}{20}
-\newcommand{\daSecScoreWorstmodelNumFullVersions}{2}
+\newcommand{\daSecScoreWorstmodelNumFullVersions}{1}
 \newcommand{\daSecScoreWorstsummaryNumFullVersions}{1103}
-\newcommand{\daSecScoreWorstoperatorNumFullVersions}{59}
+\newcommand{\daSecScoreWorstoperatorNumFullVersions}{71}
 \newcommand{\daFullDeployedAt}{95\%}
 \newcommand{\daOSCurveFitParamFirst}{87.5}
-\newcommand{\daOSCurveFitParamSecond}{\num{0.00315}}
-\newcommand{\daOSCurveFitRMSE}{0.114}
-\newcommand{\daAPICurveFitParamFirst}{110}
-\newcommand{\daAPICurveFitParamSecond}{\num{0.00315}}
-\newcommand{\daAPICurveFitRMSE}{0.116}
-\newcommand{\daOSCurvePolyRMSE}{0.115}
-\newcommand{\daOSCurveSplineRMSE}{0.115}
-\newcommand{\daOSCurveHalfDeployed}{$308 \pm 81$}
-\newcommand{\daOSCurveFullDeployed}{$1\,040 \pm 236\,000$}
-\newcommand{\daAPICurvePolyRMSE}{0.117}
-\newcommand{\daAPICurveSplineRMSE}{0.117}
-\newcommand{\daAPICurveHalfDeployed}{$330 \pm 84$}
-\newcommand{\daAPICurveFullDeployed}{$1\,060 \pm 236\,000$}
+\newcommand{\daOSCurveFitParamSecond}{\num{0.00301}}
+\newcommand{\daOSCurveFitRMSE}{0.116}
+\newcommand{\daAPICurveFitParamFirst}{105}
+\newcommand{\daAPICurveFitParamSecond}{\num{0.00299}}
+\newcommand{\daAPICurveFitRMSE}{0.115}
+\newcommand{\daOSCurvePolyRMSE}{0.116}
+\newcommand{\daOSCurveSplineRMSE}{0.116}
+\newcommand{\daOSCurveHalfDeployed}{$318 \pm 87$}
+\newcommand{\daOSCurveFullDeployed}{$1\,080 \pm 247\,000$}
+\newcommand{\daAPICurvePolyRMSE}{0.115}
+\newcommand{\daAPICurveSplineRMSE}{0.115}
+\newcommand{\daAPICurveHalfDeployed}{$337 \pm 87$}
+\newcommand{\daAPICurveFullDeployed}{$1\,110 \pm 248\,000$}
 \newcommand{\daVulnAPKDuplicateFileOctoberPerc}{91.5\%}
-\newcommand{\daNumUpdatesBigUpgrades}{778}
+\newcommand{\daNumUpdatesBigUpgrades}{801}
 \newcommand{\daNumUpdatesSkippedBig}{3}
-\newcommand{\daPercBigUpgrades}{22.4\%}
-\newcommand{\daUpdatesPerYear}{$1.35 \pm 0.02$}
-\newcommand{\daOSMonthsOfData}{40}
-\newcommand{\daMeanInsecurityPercNominal}{87.3\%}
-\newcommand{\daVulnFreeNominal}{0.127}
-\newcommand{\daUpdatednessNominal}{0.0558}
-\newcommand{\daUpdatednessPercNominal}{5.58\%}
-\newcommand{\daSecurityScoreNominal}{2.75}
-\newcommand{\daUpdatesPerYearNominal}{1.35}
-\newcommand{\daOSCurveHalfDeployedNominal}{308}
-\newcommand{\daOSCurveFullDeployedNominal}{1\,040}
-\newcommand{\daAPICurveHalfDeployedNominal}{330}
-\newcommand{\daAPICurveFullDeployedNominal}{1\,060}
-\newcommand{\daSecScoreBestoperatorScoreNominal}{3.79}
-\newcommand{\daSecScoreWorstoperatorScoreNominal}{0.4}
-\newcommand{\daSecScoreBestmodelScoreNominal}{4.8}
-\newcommand{\daSecScoreWorstmodelScoreNominal}{0.274}
-\newcommand{\daAdbEnabledPercNominal}{19.6\%}
-\newcommand{\daSecScoreBestmanufacturerScoreNominal}{3.38}
-\newcommand{\daSecScoreWorstmanufacturerScoreNominal}{0.139}
-\newcommand{\daUpdatesPerMonthPerVersion}{3.62}
-\newcommand{\daNumUpdateFullOnly}{0}
+\newcommand{\daPercBigUpgrades}{$18.7 \pm 0.7\%$}
+\newcommand{\daUpdatesPerYear}{$1.29 \pm 0.01$}
+\newcommand{\daOSMonthsOfData}{44}
+\newcommand{\daMeanInsecurityPercNominal}{88.0\%}
+\newcommand{\daVulnFreeNominal}{0.12}
+\newcommand{\daUpdatednessNominal}{0.0521}
+\newcommand{\daUpdatednessPercNominal}{5.21\%}
+\newcommand{\daSecurityScoreNominal}{2.79}
+\newcommand{\daUpdatesPerYearNominal}{1.29}
+\newcommand{\daOSCurveHalfDeployedNominal}{318}
+\newcommand{\daOSCurveFullDeployedNominal}{1\,080}
+\newcommand{\daAPICurveHalfDeployedNominal}{337}
+\newcommand{\daAPICurveFullDeployedNominal}{1\,110}
+\newcommand{\daSecScoreBestoperatorScoreNominal}{3.82}
+\newcommand{\daSecScoreWorstoperatorScoreNominal}{0.485}
+\newcommand{\daSecScoreBestmodelScoreNominal}{4.71}
+\newcommand{\daSecScoreWorstmodelScoreNominal}{0.0001}
+\newcommand{\daAdbEnabledPercNominal}{19.3\%}
+\newcommand{\daSecScoreBestmanufacturerScoreNominal}{3.7}
+\newcommand{\daSecScoreWorstmanufacturerScoreNominal}{0.214}
+\newcommand{\daUpdatesPerMonthPerVersion}{$3.04 \pm 0.05$}
+\newcommand{\daNumUpdateFullOnly}{660}
+\newcommand{\daVulnZergRushMonthsDefFixDeployed}{27.3}
+\newcommand{\daUpdatesPerMonthPerVersionNominal}{3.04}
+\newcommand{\daPercBigUpgradesNominal}{18.7\%}
+\newcommand{\daPercUpdatesDowngradesNominal}{3.52\%}
+\newcommand{\daGPAPISeventeenLaterDate}{2015-03-02}
+\newcommand{\daGPAPISeventeenLaterProportion}{69.5\%}
+\newcommand{\daGPAPISeventeenEarlierProportion}{30.5\%}
+\newcommand{\daGPAPICurveFitParamFirst}{80.6}
+\newcommand{\daGPAPICurveFitParamSecond}{\num{0.00262}}
+\newcommand{\daGPAPICurveFitRMSE}{0.183}
+\newcommand{\daGPAPICurvePolyRMSE}{0.183}
+\newcommand{\daGPAPICurveSplineRMSE}{0.183}
+\newcommand{\daGPAPICurveHalfDeployedNominal}{346}
+\newcommand{\daGPAPICurveHalfDeployed}{$346 \pm 175$}
+\newcommand{\daGPAPICurveFullDeployedNominal}{1\,230}
+\newcommand{\daGPAPICurveFullDeployed}{$1\,230 \pm 283\,000$}
+\newcommand{\daAPISeventeenReleaseDate}{2012-10-29}
+\newcommand{\daMeanInsecurityPercTwosfNominal}{88\%}
+\newcommand{\daMeanInsecurityPercTwosf}{$88 \pm 0\%$}
+\newcommand{\daModelHalfDeployment}{27}
+\newcommand{\daModelFullDeployment}{316}
+\newcommand{\daModelHalfDeploymentYears}{0.074}
+\newcommand{\daModelFullDeploymentYears}{0.866}
+\newcommand{\daOSCurveHalfDeployedYearsNominal}{0.871}
+\newcommand{\daOSCurveHalfDeployedYears}{$0.871 \pm 0.24$}
+\newcommand{\daOSCurveFullDeployedYearsNominal}{2.97}
+\newcommand{\daOSCurveFullDeployedYears}{$2.97 \pm 67$}
+\newcommand{\daAPICurveHalfDeployedYearsNominal}{0.923}
+\newcommand{\daAPICurveHalfDeployedYears}{$0.923 \pm 0.238$}
+\newcommand{\daAPICurveFullDeployedYearsNominal}{3.03}
+\newcommand{\daAPICurveFullDeployedYears}{$3.03 \pm 67$}
+\newcommand{\daGPAPICurveHalfDeployedYearsNominal}{0.947}
+\newcommand{\daGPAPICurveHalfDeployedYears}{$0.947 \pm 0.479$}
+\newcommand{\daGPAPICurveFullDeployedYearsNominal}{3.36}
+\newcommand{\daGPAPICurveFullDeployedYears}{$3.36 \pm 77$}
+\newcommand{\daGPAPISeventeenCurveFitParamFirst}{274}
+\newcommand{\daGPAPISeventeenCurveFitParamSecond}{\num{0.00184}}
+\newcommand{\daGPAPISeventeenCurveFitRMSE}{0.0283}
+\newcommand{\daGPAPISeventeenCurvePolyRMSE}{0.00903}
+\newcommand{\daGPAPISeventeenCurveSplineRMSE}{0.00903}
+\newcommand{\daGPAPISeventeenCurveHalfDeployedYearsNominal}{1.78}
+\newcommand{\daGPAPISeventeenCurveHalfDeployedYears}{$1.78 \pm 0.08$}
+\newcommand{\daGPAPISeventeenCurveHalfDeployedNominal}{650}
+\newcommand{\daGPAPISeventeenCurveHalfDeployed}{$650 \pm 31$}
+\newcommand{\daGPAPISeventeenCurveFullDeployedYearsNominal}{5.2}
+\newcommand{\daGPAPISeventeenCurveFullDeployedYears}{$5.2 \pm 1.2$}
+\newcommand{\daGPAPISeventeenCurveFullDeployedNominal}{1\,900}
+\newcommand{\daGPAPISeventeenCurveFullDeployed}{$1\,900 \pm 453$}
+\newcommand{\daGPAPISeventeenFullDeployment}{2018-01-10}
+\newcommand{\daGPAPIPerAPIMeanNominal}{0.00454}
+\newcommand{\daGPAPIPerAPIMean}{$0.00454 \pm 0.00337$}
+\newcommand{\daGPAPIPerAPIStdevNominal}{0.0215}
+\newcommand{\daGPAPIPerAPIStdev}{$0.0215 \pm 0.0129$}
+\newcommand{\daGPAPIPerAPIAbsMeanNominal}{0.0148}
+\newcommand{\daGPAPIPerAPIAbsMean}{$0.0148 \pm 0.0068$}
+\newcommand{\daGPAPIPerAPIAbsStdevNominal}{0.0156}
+\newcommand{\daGPAPIPerAPIAbsStdev}{$0.0156 \pm 0.0121$}
+\newcommand{\daGPAPIPerAPIParametersTable}{\begin{table} \centering \begin{tabular}{c|S|S} API Version & ${t_0}$ & $\mathrm{decay}$ \\ \hline2	&	0.0	&	0.0237\\3	&	0.0	&	0.0235\\4	&	0.0	&	0.00689\\5	&	59.1	&	0.00465\\6	&	2.4	&	0.00434\\7	&	0.0	&	0.00499\\8	&	47.7	&	0.00415\\9	&	155	&	0.00365\\10	&	90.2	&	0.00364\\11	&	432	&	0.00227\\12	&	355	&	0.00227\\13	&	303	&	0.00235\\14	&	218	&	0.00238\\15	&	160	&	0.00239\\16	&	141	&	0.00226\\17	&	274	&	0.00184\\18	&	116	&	0.00138\\19	&	118	&	0.00152\\\end{tabular} \caption{Parameters for different API versions} \label{tab:per_api_parameters} \end{table}}

### 2015-03-16 securityupdates.tex
Make todos visible and unblind
Add simple sensitivity analysis for Bangladesh study

@@
 \usepackage{siunitx}
-\usepackage[disable]{todonotes}
+\usepackage{todonotes}%[disable]
 \let\OldTodo\todo
 \renewcommand{\todo}{\OldTodo[inline]}
-\newcommand{\todolater}[1]{}%\todo
+\newcommand{\todolater}[1]{\todo{#1}}%{}%
 \RequirePackage[date=terse, isbn=true, doi=true, url=false, urldate=iso8601, maxbibnames=9, backref=false, backend=bibtex, style=ieee]{biblatex}
@@
 % Blinding function
-\newcommand{\identifying}[1]{}%{#1}%
-\newcommand{\blindauthors}[1]{Paper \#59}%{#1}%
+\newcommand{\identifying}[1]{#1}%{}%
+\newcommand{\blindauthors}[1]{#1}%{Paper \#59}%
 
@@
 %We could normalise out the influence of the Bangladeshi users, however they are Android users too and they also need security updates.
+To investigate the sensitivity of our results to the Bangladesh study we truncated our data before this study started and found the percentage of vulnerable devices to be 87.7\% rather than \daMeanInsecurityPercNominal, and the percentage of devices running the latest version to be 5.43\% rather than \daUpdatednessPercNominal.
+Hence removing this anomaly would have no effect on our conclusions.
 Since most \da\ users are self selecting and install \da\ because they want to find out more about what their phone is doing or to aid research they may be biased and perhaps more likely to install updates.

### 2015-03-16 security_scoring.tex
Minor fixes
Add paragraph about distribution of updates to models

@@
 \begin{equation}
-\mathrm{score} = f\times 4 + u \times 3 + \frac{2}{1+e^m} \times 3 \label{eq:score}
+\mathrm{score} = 4 \cdot f + 3 \cdot u + 3 \cdot \frac{2}{1+e^m} \label{eq:score}
 \end{equation}
@@
 
+If update delays are due to the delay in manufacturers providing the update rather than in operators in supplying the update and users installing the update we would expect the update behaviour of devices with the same device model to be similar.
+We found that within \daModelHalfDeployment\ days of the first observation of a new version on a device, half of all devices of that model have the new version (or a higher version) installed, and within \daModelFullDeployment\ days \daFullDeployedAt\ of devices have the new version (or a higher version).
+There is a variation between device models, with the update being distributed to most devices quickly and others having a much slower roll out but since some do update quickly the bottleneck is unlikely to be with the user.
+
 \begin{figure*}

### 2015-03-17 securityupdates.tex
Switch to llncs template for ESORICS submission

@@
-\documentclass[conference,a4paper,twoside]{IEEEtran}
+\documentclass{llncs}
 \usepackage[utf8]{inputenc}
@@
 
-% author names and affiliations
-% use a multiple column layout for up to three different
-% affiliations
-\author{
-\blindauthors{
-\IEEEauthorblockN{Daniel R. Thomas,
-Daniel T. Wagner,
-Alastair R. Beresford,
-Andrew Rice}
-\IEEEauthorblockA{
-Computer Laboratory\\
-University of Cambridge\\
+\author{\blindauthors{Daniel~R.~Thomas \and Daniel~T.~Wagner\and Alastair~R.~Beresford \and Andrew~Rice}}
+\institute{\blindauthors{Computer Laboratory,
+University of Cambridge,
 Cambridge, United Kingdom\\
-Firstname.Lastname@cl.cam.ac.uk
-}
-}
-}
+\email{firstname.lastname@cl.cam.ac.uk}}}
 

### 2015-03-18 securityupdates.tex
Attempt to make things work better with the single column format

@@
 \label{sec:android_update_process}
-\begin{figure}
+\begin{figure}[h]
  \centering
- \def\svgwidth{\columnwidth}
+ \def\svgwidth{0.7\columnwidth}
  \import{figures/}{update_ecosystem.pdf_tex}
  \caption{Flow of updates between participants in the Android ecosystem.
- Numbers on edges indicate updates shipped between \daStartDate\ and \daEndDate, numbers in brackets represent number of such entities in our data.
- Dotted arrows indicate flows where we do not know how many updates are being produced as we can't measure those flows directly as they are not public.\todolater{Turn into a sankey diagram}}
+ Numbers on edges indicate updates shipped between \daStartDate\ and \daEndDate, those in brackets represent number of such entities.
+ Dotted arrows indicate flows where we can't measure how many updates are produced as they are not public.}
  \label{fig:update_ecosystem}
@@
 \begin{figure}
- \includegraphics[width=\columnwidth]{figures/from_to_updates.pdf}
+ \centering
+ \includegraphics[width=0.6\columnwidth]{figures/from_to_updates.pdf}
  \caption{Updates between different Android versions in the \da\ data}
@@
 \end{figure}
-\begin{figure}
+\begin{figure}[h]
  \includegraphics[width=\columnwidth]{figures/nw_security_updates.pdf}

### 2015-03-18 ccs.tex
Strip out security updates and android ecosystem stuff from esorics paper and put in a ccs paper and make the esorics paper meet the page limit

--- 2015-03-18 securityupdates.tex
@@
-\documentclass{llncs}
+\documentclass{sig-alternate}
+
 \usepackage[utf8]{inputenc}
@@
 
-\author{Daniel R.\ Thomas, Alastair R.\ Beresford, Daniel T.\ Wagner and Andrew Rice}
-
 \input{dastats}
@@
 \newcommand{\identifying}[1]{#1}%{}%
-\newcommand{\blindauthors}[1]{#1}%{Paper \#59}%
-
-\begin{document}
-\title{Timeliness of security updates for Android}
-
-\author{\blindauthors{Daniel~R.~Thomas \and Daniel~T.~Wagner\and Alastair~R.~Beresford \and Andrew~Rice}}
-\institute{\blindauthors{Computer Laboratory,
-University of Cambridge,
-Cambridge, United Kingdom\\
-\email{firstname.lastname@cl.cam.ac.uk}}}
+\newcommand{\blindauthors}[1]{#1}%{Paper \#XX}%
 
@@
 
+\begin{document}
+%
+% --- Author Metadata here ---
+\conferenceinfo{CCS}{2015 Denver, Colorado, USA}
+%\CopyrightYear{2015} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
+%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
+% --- End of Author Metadata ---
+
+\title{TODO}
+
+\numberofauthors{3}
+\author{
+Daniel~R.~Thomas
+\and
+Alastair~R.~Beresford\\
+       \affaddr{Computer Laboratory}\\
+       \affaddr{University of Cambridge}\\
+       \affaddr{Cambridge, United Kingdom}\\
+       \email{Firstname.Lastname@cl.cam.ac.uk}
+\and
+Andrew~Rice
+}
+
 \maketitle
 
-% Hypothesis
-% Attempts at fine grained restrictions on running arbitrary code are hampered by updates for security vulnerabilities not reaching users in a timely fashion.
-
 \begin{abstract}
-Modern smartphone operating systems protect users from malicious applications by providing a protected execution environment or sandbox.
-This approach relies on the absence of active exploits which allow apps to break out of the sandbox.
-Many such exploits have been found, and therefore the security of a smartphone operating system relies on the prompt delivery of security updates.
-In this paper we tease apart the Android software ecosystem and quantify the extent to which the security sandbox provided by Android protects devices from malicious apps.
-We analysed \da\ data from over \daOSYearsOfData\ years and \daNumOSDevices\ devices and found that on average \daMeanInsecurityPercNominal\ of devices were exposed to known root privilege vulnerabilities and only \daUpdatednessPercNominal\ of devices run the most recent version of Android.
 We find there is significant variability in the timely delivery of security updates across different device manufacturers and mobile network operators, and we define a security metric to rank the performance of device manufacturers and network operators in our data set.
+TODO
 \end{abstract}
 
+% A category with the (minimum) three required fields
+\category{}{}{}
+
+\terms{}
+
+\keywords{}
+
 \section{Introduction}
-%Android is the most popular smartphone operating system today with \percMarketShare\ marketshare.
-
-Support for third-party apps is a key feature of modern smartphone operating systems.
-Such apps are written by many developers from a wide range of backgrounds, which means neither operating system vendors nor users can fully trust app developers.
-Therefore, in order to secure personal data and prevent theft, such as the sending of premium-rate text messages, operating system vendors provide a protected execution environment, or \emph{sandbox}, for apps.
-
-In this paper we quantify the extent to which the app sandbox provided by Android protects devices from malicious apps attempting to gain control over the entire handset.
-%We have analysed Android because it is the most popular smartphone operating system by market share (~\percMarketShare) and 
-Previous research has shown that in 2012, between 36.7\%~\cite{Zhou2012b} and 40\%~\cite{Zhou2012a} of malware for Android contained exploits designed to enable a malicious app to break the sandbox.
-What is not known however is the proportion of Android devices which are vulnerable to these exploits over time.
-To the best of our knowledge, this is the first paper to provide detailed numbers.
-
-Not all malicious apps need to break the sandbox in order to misbehave.
-Indeed, the above figures suggest the majority of malicious apps do not attempt to do so.
-We focus on the issue of apps breaking out of the sandbox because malware which is able to take full control of a handset can do significantly more harm and is much harder to remove.
-Ordinary malware can be uninstalled by the user, or remotely by Google through the Play Store, and returns the device to a secure state. 
-Removing an app which has used an exploit to take control of the handset is unlikely to return the device to secure state.
-We describe our threat model and three attack vectors in detail in \S\ref{sec:threatmodel}.
 
@@
 We present a better understanding of both the ecosystem of Android development and associated vulnerabilities more fully in \S\ref{sec:android_ecosystem}.
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=\columnwidth]{figures/proportioninsecure}
-\caption{Proportion of devices running insecure, maybe secure and secure versions of Android against time.
-The red vertical lines are caused by vulnerabilities being discovered with those which have the biggest impact annotated.
-This graph is explained further in \S\ref{sec:exp:versionsecurity}.
-}
-\label{fig:proportioninsecure}
-\end{figure}
-
-In order to determine whether an individual smartphone is vulnerable to a particular exploit, we need to combine vulnerability data with the operating system version string and build number from the handset.
-For this we use handset data gathered from \daNumOSDevices\ devices in the \da~\cite{Wagner2013} project.
-This analysis allows us to plot the proportion of Android devices in the \da\ data which were running versions of Android known to have vulnerabilities at the time and is shown in Figure~\ref{fig:proportioninsecure}.
-The figure shows the proportion of devices split into three categories:
-(1)~\emph{secure} because they are running a version of Android with no known vulnerabilities; 
-(2)~\emph{insecure} because the device was running an insecure version of Android with a build released before the vulnerability was fixed or discovered; and 
-(3)~\emph{maybe secure} because the device is running a version of Android known to be insecure but has received an update which might contain a backported fix as we did not observe that build number until after the vulnerability was known.
-Further analysis into the propagation of updates in Android is presented in \S\ref{sec:exp:android_ecosystem}.
 
@@
 \begin{itemize}
- \item We characterise the Android update ecosystem, showing how updates flow between entities, and quantify vulnerabilities by determining what proportion of devices they affected and how that changes over time.
- \item We created an open database to collate quantitative information on vulnerabilities which affect Android, and investigate their time-lines.
  \item We measure the security of Android according to three metrics and compare different device manufacturers and device models to allow device purchasers to differentiate between them based on security.
 \end{itemize}
-
-\section{Threat model}
-\label{sec:threatmodel}
-
-In this paper we measure whether an attacker can take control of a handset by breaking out of the app sandbox on Android using a known vulnerability.
-We consider three attack vectors.
-The first attack vector is through the installation of a malicious app on the device.
-Android devices support the installation of apps through a number of app marketplaces, email attachments, URLs and via the Android Debug Bridge (ADB).
-By default, many Android devices with Google Play installed will prevent the installation of apps from other sources.
-
-Controlling potential sources of app installation is a good security feature because app marketplaces provide additional protection to prevent the listing of malicious apps.
-For example, the Google Play marketplace uses Bouncer to automatically analyse apps as well as removing any apps which are reported as malicious.
-Nevertheless, Google Play is not immune: 0.02\% of apps are malicious~\cite{Zhou2012a}.
-Alternative markets are also popular, particularly in countries such as China where the Google Play is not available. 
-Previous work has shown that on alternative markets between 0.20\% and 0.47\% of apps are malicious~\cite{Zhou2012a}.
-
-The second attack vector is to persuade an existing app to download and execute code at runtime.
-The most direct method available to an attacker is to upload to a marketplace a seemingly innocent app which contains support for dynamic code loading from a remote server under control of the attacker.
-Neither static nor dynamic analysis of this app by the marketplace will uncover any malicious code, since it does exist in the app binary.
-Instead the attacker can chose to send a malicious payload directly to the app for execution at a later point in time.
-Note that dynamic code loading is not actually needed for this to work---it just makes the task easier---since previous work has shown that even on a platform such as iOS, which does not permit dynamic code loading, a ROP-based attack is relatively easy if the attacker creates an app with carefully crafted flaws~\cite{}.
-
-Support for dynamic code loading on Android also supports the third attack vector in which the attacker injects malicious code directly into existing apps already installed on the handset. 
-For example, the addJavascriptInterface vulnerability (CVE-2012-6636) allows Javascript running in an Android WebView to execute arbitrary code in the same process address space as the vulnerable app.
-This attack works when: (1) the WebView has the Javascript-to-Java bridge enabled; and (2) the attacker is able to inject a malicious Javascript payload into the network traffic between the phone and the web server.
-Many apps are vulnerable to this type of attack because ad network libraries typically embed adverts into apps by displaying HTML and Javascript inside a WebView with the Javascript-to-Java bridge enabled.
-The fix for the addJavascriptInterface vulnerability breaks backwards compatibility, and therefore apps are only immune to this attack if their target API level is greater than 16 and they are running on an Android device with an OS version of 4.2 or greater.
-Later analysis in Figure~\ref{fig:da_api} in this paper shows that 40\% of handsets connecting to the Google Play store are currently vulnerable to this attack.
-
-\section{Data sources}
-\label{sec:background}
-
-We require two sources of data for this study: (1) information on the distribution of installed versions of Android over time across the Android ecosystem and (2) information on the critical vulnerabilities found to affect specific versions of Android.
-These two datasets can then be combined to determine the proportion of handsets at risk of attack from specific vulnerabilities for particular periods of time.
-
-\subsection{Versions of Android running on devices}
-
-In our analysis we use historical data collected by the \da\ project~\cite{Wagner2013}.
-Device Analyzer collects data from study participants who install the project Android app from the Google Play store.
-Most study participants allow researchers around the world to access a subset of their device data, including the data presented in this paper.
-
-The \da\ app collects a range of metadata from Android devices.\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/collected.html}}
-For this paper, we extracted the build string and API version from the device log each day.
-The API version is a positive integer which increases when new features are added to the API.
-Consequently security (bug) fixes do not result in a change in the API version.
-The build string is a user-readable version string.
-Fortunately most (\daOSVersionPercValidLines) entries in the data have a build string of the form `x.y.z opaque\_marker' and so it is possible to extract the Android version number `x.y.z'.
-On a large proportion of devices `opaque\_marker' is a well defined build number\footnote{\url{https://source.android.com/source/build-numbers.html}} however different device manufacturers use different schema.
-
-The \da\ project has collected data from \daNumDevices\ devices with a total of \daDeviceDays\footnote{Here we are only counting devices and days for which we have valid OS version data.} device days.
-The majority of devices only contribute data for a short period of time, however \daMonthsDevices\ devices have contributed data for more than \daMonths~months.
-
-Where possible we indicate the uncertainty in our results by presenting them $\pm$ one standard deviation and give results to 3 significant figures, this occasionally results in `${} \pm 0$' when the standard deviation is less than the number of significant figures we are reporting.
-This technique cannot account for systematic errors and therefore we explore the potential for systematic errors in the \da\ data in \S\ref{sec:representative}.
-
-\subsection{Critical vulnerabilities}
-
-We compiled a list of critical vulnerabilities in Android, containing information on the discovery and publication dates, the versions affected and which versions fixed the problem.
-We only looked for critical vulnerabilities such as root vulnerabilities which did not require USB debugging to exploit.
-Critical vulnerabilities allow a program to gain privileges equivalent in scope to root.
-If an application exploits a critical vulnerability then it gains control of the device.
-Some phones can be `rooted' by enabling USB debugging and using the special privileges of the ADB shell to root the device but only \daAdbEnabledPerc\dafoot\ of devices have USB debugging enabled.
-This is not something that applications running on the phone can exploit to break out of the app sandbox and so we do not include those vulnerabilities.
-Unfortunately, many published exploits use ADB for convenience and so determining whether the use of ADB is necessary to exploit the vulnerability can be difficult.
-
-Some critical vulnerabilities are not traditional kernel vulnerabilities, for example the discovery of flaws in the verification of signatures on Android applications in February 2013~\cite{Forristal2013} meant that applications could pretend to be signed with system keys and hence gain root equivalent privileges.
-On some versions of Android (below version 4.1) malware could use known system-to-root escalation mechanisms but on all versions they have a greatly increased attack area for further privilege escalation and also have the ability to control all user internet traffic (via VPNs), brick the phone, remove and install apps, steal user credentials, read the screen and make as well as receive calls.
-\todolater{break vulnerabilities down by attack vector?}
-\avoTabAndVulns
-
-We developed and maintain an open platform for filing critical vulnerabilities in a machine readable format, the AndroidVulnerabilities.org\footnote{\url{http://androidvulnerabilities.org/} \textbf{Note to reviewers:} we have removed our names and affiliation from the website for this review \textbf{with the exception of the contact page.
-Please do not visit the contact page until the review cycle has completed.}} (\avo) website.
-We seeded it with data from the CVE database, vendor lists, reports from the literature and various forums.
-In addition, we have received submissions or amendments from \avoNumSubmitters\ individuals.
-We collected data between \avoStartDate\ and \avoEndDate\ and so any information lost before the start of that period cannot be included.
-\avo\ currently contains \avoNumVulnerabilities\ vulnerabilities of which \avoNumVulnAllAndroid\ affect all Android devices and \avoNumVulnSpecific\ are specific to particular devices or device manufacturers.
-
-Unless otherwise stated, we use \daNumVulnsUsed\ vulnerabilities in our analysis as shown in Table~\ref{tab:andvulns}.
-We have chosen these vulnerabilities to fit the attack vectors introduced in \S\ref{sec:threatmodel}.
-These vulnerabilities affect all Android devices regardless of manufacturer, and as a result our selected vulnerabilities will dominate security analysis of the Android ecosystem as a whole.
-In many cases we could not match manufacturer- and model-specific vulnerabilities to individual devices in the \da\ data and therefore attempting to include device-specific vulnerabilities as well would introduce additional uncertainty in our results.
-In contrast, with our chosen set of vulnerabiltiies, our analysis represents a lower-bound on the vulnerability of devices in the \da\ data set.
-
-Tracking vulnerabilities is a manual task as they are not consistently recorded in other databases such as the CVE database.
-In addition, the lack of a widely acknowledged unique identifier required manual analysis to identify whether two reports referenced the same vulnerability.
-Previous work has assumed ``any security issue of relevance will eventually get a CVE number assigned"~\cite{Frei2010} which is currently not the case for critical Android vulnerabilities.
-For some of the vulnerabilities without CVE numbers, Google confirmed that there was no CVE number and that they did not intend to get one, instead providing an Android bug number.
-
-\subsection{Lifetime of a vulnerability}
-
-The key events in the lifetime of a vulnerability do not always occur in the same order and are:\\
-\textbf{creation} When a vulnerability was created in the source code.\\
-\textbf{introducing release} When the first release was made containing the vulnerability.\\
-\textbf{discovery} When the vulnerability is first discovered.\\
-\textbf{exploit} When the vulnerability is first exploited.\\
-\textbf{disclosure} When the vulnerability is first disclosed.\\ % publically or to a smaller set of people?
-\textbf{fix} When a vulnerability was first fixed in the source code.\\
-\textbf{fixing release} When the first release containing the fix was made (equivalent to `patch available'~\cite{Frei2010}).\\
-\textbf{fix deployed} When the fix has been deployed to a sufficiently large proportion of the population that the vulnerability can be ignored (the ecosystem equivalent to the per instance `patch installed'~\cite{Frei2010}).
-\todolater{Does this look nice?}
-
-Establishing when a vulnerability starts to pose a threat to users is difficult.
-Frei et al.~\cite{Frei2010} propose the definition of the {\bf time of disclosure}.
-This occurs when the information about the vulnerability is freely available to the public from a widely accepted and independent source and has been validated by security experts so that it has a risk rating.
-Unfortunately before we collated this information much of it was not published by an independent source and lacked risk rating information, even months or years after they had been actively used.
-Therefore this measure does not work.
-
-Symantec's 2012 analysis of desktop malware has shown that after public disclosure, exploitation rates increase by 5 orders of magnitude~\cite{Bilge2012} and so from the point of view of widespread danger, the period between public \textbf{disclosure} and the date of \textbf{fix deployed} is the most critical.
-However they also show that zero-day vulnerabilities are typically used for 312 days before they are publicly disclosed, often to target particular organisations.
-Hence when considering a vulnerability from the point of view of an organisation which cares about Advanced Persistent Threats, such as those responding to CESG advice~\cite{CESG2013}, the critical period starts with \textbf{discovery} and continues until \textbf{fix deployed}.
-Therefore, in our vulnerability calculations we prefer to use the earliest recorded date of \textbf{discovery}, even if that knowledge might have been confined to a particular device manufacturer or hobbyist.
-%We cannot know if someone reported the vulnerability to the device manufacturer and we do know that various agencies are engaged in widespread monitoring of communications and compromise of civilian infrastructure~\cite{TODO} and so they may know.
-
-Unfortunately {\bf discovery} is the hardest point to obtain concrete data on as discovery may happen multiple times independently and not all discoverers will report their discovery.
-{\bf Creation} and {\bf introducing release} are the earliest points which could be used, but the risk is mostly latent until someone discovers them and therefore we do not use these dates.
-%, for example entities with high value targets may keep databases of unknown vulnerabilities for later use, in which case there is a high risk to potential targets from the never disclosed point of discovery.
-%However if the discoverer never discloses the vulnerability to anyone or makes any use of it then there is little danger.
-The date of first {\bf exploit} is a point when the risk is definitely high, but again a good adversary will not be detected when using such exploits.
-The date of first {\bf fix} is, assuming that the fix is deliberate, a point at which the vulnerability is known at least within the organisation performing the fix and frequently implies an earlier discovery and notification by a third party; we determine this date from the Android Open Source Project repository by examining the authored-on date of the fixing commit.
-%Once a {\bf fixing release} has been made then the vulnerability is widely known because it can be reverse engineered from the changes in the release~\cite{Brumley2008}.
-%When the {\bf fix deployed} to a sufficiently large proportion of devices then the remaining risk is minimal.
-Therefore, in this paper we use the earliest known date from \textbf{discovery}, \textbf{exploit}, \textbf{disclosure} or \textbf{fix}.
-A breakdown of the type of date used per vulnerability is shown in Table~\ref{tab:andvulns}.
-
-\subsection{Distribution of vulnerabilities}
-\begin{figure}
- \centering
- \includegraphics[width=\columnwidth]{figures/vulnerabilities_timeline}
- \caption{Timeline of vulnerabilities. For each vulnerability we show the dates of \textbf{discovery} and, when later, the dates of the first \textbf{fixing release}.}
- \label{fig:vulnerabilities_timeline}
-\end{figure}
-
-Figure~\ref{fig:vulnerabilities_timeline} shows the dates of \textbf{discovery} and, when later, the date of the first \textbf{fixing release}.
-Some vulnerabilities (\emph{levitator}, \emph{zergRush}) were fixed in released versions of Android before they were discovered and so are shown as vertical lines, while others were known for months before a version of Android which fixed them was shipped.
-During the period in which \da\ data was collected the date when a version of Android with the fix was observed on a \da\ device is taken as the date that version was released.
-For vulnerabilities prior to the collection period we estimate the release date as best we can using publicly available data.
-There is no canonical source of Android release dates, and our best guesses and supporting references are available from \avo.
-
-The discovery dates of our vulnerabilities does not appear to be uniform. 
-In particular the data shows a large gap from 2011-10-06 to 2013-02-18 where we do not know of any discoveries of critical vulnerabilities affecting all Android devices.
-The cause of this quiet period is unclear.
-Possible explanations are that: (i) most devices were exposed to known vulnerabilities so there was no point in looking for new ones (from Figure~\ref{fig:proportioninsecure});
-(ii) device manufacturers made it easier to install custom versions of Android, reducing the need for users to root their devices; and
-(iii) device manufacturer specific vulnerabilities (which are not in our analysis here) were easier to find and therefore attackers looking for vulnerabilities adjusted their focus.
 
@@
 
-\section{Experiments}
-\label{sec:results}
-We present the results of our analysis showing that, on average, \daMeanInsecurityPerc\ of Android devices are exposed to critical vulnerabilities.
-Devices, on average, apply \daUpdatesPerYear\ updates each year and \daPercUpdatesDowngrades\ of full version changes are downgrades to older versions.
-Nexus devices are better than others, \emph{\daSecScoreBestmanufacturer} is the highest scoring device manufacturer and \emph{\daSecScoreBestoperator} is the highest scoring network operator.
-%\todo{comparison of nexus update rates with iOS}
-We describe four experiments.
-We examine the vulnerability of the Android ecosystem as a whole (\S\ref{sec:exp:versionsecurity}, explore the upgrade cycle and vulnerability cycle of the ecosystem (\S\ref{sec:exp:android_ecosystem}), quantify the updates installed on particular devices (\S\ref{sec:exp:device_updates}) and compare the performance of different device manufacturers and network operators (\S\ref{sec:exp:security_score}).
-
-%% In an experiment you test a hypothesis, it needs to be clear what the hypothesis is and what the conclusion of the test is
-
-\subsection{Experiment 0: vulnerability of Android devices}\label{sec:exp:versionsecurity}
-
-In the introduction, Figure~\ref{fig:proportioninsecure} described the proportion of Android devices susceptable to at least one critical vulnerability.
-
-\subsubsection{Method} 
-To investigate the behaviour of the Android ecosystem as a whole with respect to Android versions and vulnerabilities we used the OS version information from \da\ and the vulnerability data from \avo.
-The \da\ data used here was collected between \daStartDate\ and \daEndDate.
-The \avo\ data covers the period from \avoFirstDataDate\ to \avoLastDataDate.
-For each device we collected daily version data  and recorded the running version and any associated vulnerabilities it was exposed to at that time (if any).
-We then normalise these totals for each day by dividing through by the total number of devices with version information seen on that day.
-
-\subsubsection{Results}
-
-Following time from left to right in Figure~\ref{fig:proportioninsecure} we see all devices are initially \emph{maybe secure} (yellow) since \da\ does not have historical data prior to May 2011.
-This means we cannot distinguish between devices which are running a version of Android which is known to be vulnerable from one which may have received a backported fix.
-This demonstrates the importance of a longitudinal study: this type of analysis requires years of data.
-Once \emph{zergRush} was discovered in October 2011 the most devices are recorded as \emph{insecure} (red) as most devices were exposed to that vulnerability.
-The remanding devices were already running a version of Android which fixed the \emph{zergRush} vulnerability and are therefore marked as  \emph{secure} (green).
-From October 2011 until the discovery of \emph{APK duplicate file} in February 2013 the graph shows progressive improvement as devices are upgraded.
-This means more and more devices are marked as \emph{secure} (green) because they are now running a secure version of Android, or marked as \emph{maybe secure} because they received a minor update which may have contained a backported fix.
-From February 2013 onwards regular discovery of critical vulnerabilities ensures that most devices are exposed to known critical vulnerabilities.
-
-\subsection{Experiment 1: behaviour of the Android ecosystem}\label{sec:exp:android_ecosystem}
-
-\subsubsection{Method} As in Experiment 0, we used the version information for each device to calculate which critical vulnerabilities each device was susceptible to on a daily basis.
-
-\subsubsection{Results}
-The proportion of devices in the \da\ data running different versions of Android each day is shown in Figure~\ref{fig:norm_os}.
-It shows how old versions are gradually replaced by new ones, and the long tail of devices which do not see updates to more recent versions.
-
-\begin{figure*}
- \centering
- \includegraphics[width=\textwidth]{figures/da_norm_os}
- \caption{Android versions in \da\ data over time. The change in behaviour after August 2014 is explained in \S\ref{sec:da_changes}}
- \label{fig:norm_os}
-\end{figure*}
-
-The vulnerabilities devices are exposed to are shown in Figure~\ref{fig:nvulnerabilities_heat}.
-For each vulnerability it shows the proportion of devices exposed to that vulnerability and how that changes over time.
-The variation of the proportion of devices affected by a vulnerability with time tells us how badly a particular vulnerability affected the Android platform.
-In July 2011 at the beginning of the \da\ data the \emph{exploid}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/exploid_udev}} and \emph{levitator}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/levitator}} vulnerabilities both affect most Android devices, slowly these are fixed as updates roll out and devices are replaced until in January 2013 a much smaller proportion of devices are affected by known vulnerabilities.
-However when in February 2013 the first APK signing vulnerability was found which affected all previous versions of Android and even in October 2013 most devices (\daVulnAPKDuplicateFileOctoberPerc) were still vulnerable.
-%\begin{figure}%[!b]
-%\centering
-%\includegraphics[width=\columnwidth]{figures/vulnerabilities}
-%\caption{Proportion of devices exposed to each vulnerability with time (1.0 is 100\%)}
-%\label{fig:vulnerabilities}
-%\end{figure}
-
-In 2013 three vulnerabilities were found in the way which Android verified the signatures on APKs.
-These allowed the creation of malicious APKs which appear to be signed as system APKs -- which have root equivalent privileges.
-Figure~\ref{fig:nvulnerabilities_heat} shows how the the \emph{APK signing vulnerabilities} affected all devices and took months to get fixed for any device.
-However what is perhaps more worrying is the long tail on the \emph{Gingerbreak}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/Gingerbreak}}, \emph{levitator}, \emph{exploid} and \emph{zergRush}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/zergRush}} vulnerabilities which are more dangerous root vulnerabilities (not requiring new APK installation) and which still affect a significant proportion of devices years later.
-
-\begin{figure*}
- \includegraphics[width=\textwidth]{figures/nvulnerabilities_heat.pdf}
- \caption{Proportion of devices affected by different vulnerabilities.
-The prefix `(maybe)' indicates `maybe secure' and `(definite)' indicates `definitely insecure'.
-The first few vulnerabilities are all `maybe secure' as we do not have data from before that vulnerability was discovered while later ones are mostly `definitely insecure' as we know no fixing update reached the devices.
-The change in behaviour after August 2014 is explained in \S\ref{sec:da_changes}}
- \label{fig:nvulnerabilities_heat}
-\end{figure*}
-
-\subsection{Experiment 2: Updates to particular devices}\label{sec:exp:device_updates}
-Those graphs summarise data across all the devices, however one of the advantages of the \da\ data is that it allows us to look at what happens to individual devices over time.
-
-\subsubsection{Method}
-As in Experiment 0, we used the version information for each device to calculate which critical vulnerabilities each device was susceptible to on a daily basis.
-We also recorded when the version changed and which versions it changed from and to.
-In the \da\ data we have over 1000 devices contributing data in any particular week.
-However most devices only contribute for a short period of time and so we do not observe updates happening on every device.
-Instead we have a hopefully representative sample of updates which happened while \da\ was installed on the devices.
-\da\ cannot distinguish between a device being replaced and the \da\ app being removed as while multiple devices can be linked to the same user account, that is a manual process and few users do it.
-
-\subsubsection{Results}
-\begin{figure}
- \centering
- \begin{subfigure}[b]{\columnwidth}
-  \includegraphics[width=\columnwidth]{figures/device-data-all-os}
-  \caption{OS versions over time. Black vertical lines indicate that the build number changed without changing the OS version at that point.}
-  \label{fig:device_data_os}
- \end{subfigure}
- \begin{subfigure}[b]{\columnwidth}
-  \includegraphics[width=\columnwidth]{figures/device-data-all-security}
-  \caption{Number of vulnerabilities affecting each device over time}
-  \label{fig:device_data_security}
- \end{subfigure}
- \caption{The top \daNumDeviceDataDevices\ devices by days of contribution in the \da\ data. One strip per device handset.}
-\end{figure}
-The longitudinal data on the number of vulnerabilities affecting the \daNumDeviceDataDevices\ devices which have contributed the most days of data to \da\ changes over time is shown in Figure~\ref{fig:device_data_security}.
-The trend that we saw in Figure~\ref{fig:proportioninsecure} of security improving and then getting worse is also shown here.
-It shows how some devices had vulnerabilities, which were fixed, and then further vulnerabilities were discovered, and for these devices, mostly not fixed.
-This implies that these devices had been abandoned by the device manufacturer and were not receiving updates which is confirmed by Figure~\ref{fig:device_data_os} which shows which OS versions those devices were running.
-Some devices start off in 2011 exposed to known security vulnerabilities and are still exposed to those and additional ones in 2014.
-
-\begin{figure}
- \centering
- \includegraphics[width=0.6\columnwidth]{figures/from_to_updates.pdf}
- \caption{Updates between different Android versions in the \da\ data}
- \label{fig:from_to_updates}
-\end{figure}
-\begin{figure}[h]
- \includegraphics[width=\columnwidth]{figures/nw_security_updates.pdf}
- \caption{Proportion of updates each week which fixed or may have fixed security vulnerabilities}
- \label{fig:weekly_security_updates}
-\end{figure}
-We recorded update events and Figure~\ref{fig:from_to_updates} shows how devices upgrade between different versions of Android.
-Mostly the dark cells are upgrades (above the diagonal) (\daNumUpdatesUpgrades).
-While many upgrades are from one version to the next version there are also a fair number (\daNumUpdatesBigUpgrades, \daPercBigUpgradesNominal) which skip more than \daNumUpdatesSkippedBig\ versions.
-Surprisingly there are also a small number of downgrade events (\daNumUpdatesDowngrades, \daPercUpdatesDowngradesNominal) when older versions of Android are installed on to devices.
-Possible reasons why users are downgrading are to free up space on their device, to make it easier to root or because a new version introduced bugs.
-
-The number of devices getting security updates each week, is shown in Figure~\ref{fig:weekly_security_updates}.
-\begin{itemize}
- \item Updates which changed the Android version number from a version with known vulnerabilities to one which had fewer known vulnerabilities are shown in red.
- \item Updates which changed the build number but not the version number and so might contain a backported fix for a vulnerability are shown in yellow.
-\item Updates which did not fix security vulnerabilities (because there were no known security vulnerabilities in the version of Android they were already running) are shown in green.
-
-\end{itemize}
-
-By taking the number of updates observed and the number of device days of data we have collected we can compute the number of updates received by a device per year to be \daUpdatesPerYear.
-This compares badly with the number of critical vulnerabilities discovered per year of between \avoVulnsPerYearAllAndroid\ (affecting all Android) and \avoVulnsPerYear\ (including the device manufacturer specific ones in \avo).
-
 \input{security_scoring}
 
-\section{Threats to validity}
-\label{sec:validity}
-\subsection{The \da\ data gives a conservative estimate of the Android version distribution}
-\label{sec:representative}
-\begin{figure}
- \centering
- \begin{subfigure}[b]{\columnwidth}
- \includegraphics[width=\columnwidth]{figures/googleplayapi}
- \caption{Google Play data on proportion of devices running different Android API versions}
- \label{fig:play_api}
-\end{subfigure}
-\begin{subfigure}[b]{\columnwidth}
- \includegraphics[width=\columnwidth]{figures/norm_api_gpcomp}
- \caption{\da\ data on proportion of devices running different Android API versions}
- \label{fig:da_api}
-\end{subfigure}
-\caption{Monthly Android API version data}
-\end{figure}
-\begin{figure}
- \centering
- \includegraphics[width=\columnwidth]{figures/api_gpcomp_rdiff}
- \caption{Difference between \da\ and Google Play data on the proportion of devices running different Android API versions}
- \label{fig:da_gp_comp_diff}
-\end{figure}
-The data from \da\ we used to investigate the proportion of devices exposed to different vulnerabilities is the OS version.
-Unfortunately there is no authoritative source of OS version information and so we cannot check directly whether our data is representative.
-However Google has published API version information almost every month since December 2009 and we have collated this information.\footnote{\url{http://androidvulnerabilities.org/play/historicplaydashboard}}
-While API versions are too coarse grained to use for security update detection they are closely related to OS versions and so if the \da\ data on API versions is similar to the Google Data on API versions then the \da\ data on OS versions should be representative.
-Figure~\ref{fig:play_api} shows the data from Google and Figure~\ref{fig:da_api} shows the data from \da\ and they appear similar.
-Figure~\ref{fig:da_gp_comp_diff} shows the difference between Figures~\ref{fig:play_api} and \ref{fig:da_api}, normalising for days since the API version was released.
-It shows that the \da\ data systematically overestimates the prevalence of new API versions and underestimates the prevalence of old API versions.
-This means that the OS version information from \da\ is likely to be overestimating the prevalence of new OS versions and hence our results are a conservative estimate of the security of Android.
-\todolater{we want a statistical metric to claim this strongly with.}
-This allows us to have confidence in the OS version information.
-
-\subsection{Changes in the \da\ sample}
-\label{sec:da_changes}
-The \da\ data is mostly generated by devices which have the \da\ app installed because their owner happened to come across the app on Google Play as there has been no advertising and few other attempts to increase usage.
-However two collaborations with network operators to install \da\ on customer devices resulted in large numbers of new \da\ users coming from particular areas and network operators.
-One study was conducted in Norway with 654 installs of \da\ and one was conducted in Bangladesh where 2463 users installed \da.
-This is responsible for the sudden change in distribution of OS versions in \da\ shown towards the end of Figures \ref{fig:norm_os}, \ref{fig:nvulnerabilities_heat} and \ref{fig:proportioninsecure} as for that period Bangladeshi users contributed over half of the \da\ data.
-%We could normalise out the influence of the Bangladeshi users, however they are Android users too and they also need security updates.
-To investigate the sensitivity of our results to the Bangladesh study we truncated our data before this study started and found the percentage of vulnerable devices to be 87.7\% rather than \daMeanInsecurityPercNominal, and the percentage of devices running the latest version to be 5.43\% rather than \daUpdatednessPercNominal.
-Hence removing this anomaly would have no effect on our conclusions.
-Since most \da\ users are self selecting and install \da\ because they want to find out more about what their phone is doing or to aid research they may be biased and perhaps more likely to install updates.
-
-\section{Discussion} \label{sec:discussion}
-There are continuing efforts to reduce the impact of critical vulnerabilities, both in Android and more widely.
-SEAndroid~\cite{Smalley2013} which is included in Android from version 4.1~\cite{jelly-bean-release} claimed to prevent some root vulnerabilities and to reduce the impact of others.
-Capability based enforcement systems such as Capsicum~\cite{Watson2010} substantially reduce the capabilities that an exploit has to try and gain increased privilege with.
-When Capsicum is included in Linux\footnote{\url{https://github.com/google/capsicum-linux}} and hence in Android, it could be used to place fine grained restrictions on system daemons, preventing vulnerabilities becoming critical vulnerabilities.
-
-We have not included all the vulnerabilities from \avo. 
-In particular, we have not included any device manufacturer specific vulnerabilities, even when these are widespread (such as vulnerabilities affecting Qualcomm\footnote{Qualcomm is particularly good at publicly disclosing the vulnerabilities affecting their code and the patches which fix them.
-\url{https://www.codeaurora.org/projects/security-advisories/}} chipsets) because it is difficult to work out which devices are affected.
-There also tends to be less public information available about manufacturer-specific vulnerabilities which makes them harder to tie down.
-
-For some other vulnerabilities which might affect all Android devices it is hard to work out which devices are affected.
-For example, \emph{pty race}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/pty_race}} is a Linux kernel vulnerability.
-We found 48 commits which fixed this vulnerability on different branches.
-On some branches it was also accidentally fixed, and later reintroduced, before finally being fixed properly, and therefore determining which devices were vulnerable relies on knowing where, and on which branch, the kernel code for a specific release was taken from.
-For some vulnerabilities the required patch is not in AOSP.
-This is despite the fact that some device manufacturers have shipped builds containing the fix.
-This also made it difficult to determine whether a specific device was vulnerable or not.
-For \emph{RageAgainstTheCage adb}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/RageAgainstTheCage_adb}} and \emph{keystore buffer}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/keystore_buffer}} vulnerabilities we do not have sufficient data to determine whether they fit one of our attack vectors as the former may require physical ADB access and the latter may be thwarted by the sandbox.
-However a sensitivity analysis showed that those two vulnerabilities make little difference to our results as they affect few devices.
-
-We still do not know where new versions of Android come from.
-Over time newer versions of Android come to dominate, but quantifying the proportion due to new phone purchases verses the proportion due to updates of existing devices is unknown.
- Device Analyzer is the only available large data set we know of with longitudinal traces of Android OS version strings.
- Unfortunately this data set only records \daNumUpdatesUpgrades\ upgrades which, broken down over \daOSMonthsOfData\ months and \daNumSigOSVersions\ OS versions is only \daUpdatesPerMonthPerVersion\ updates per version per month.
-This is not enough to build a statistically significant prediction of the expected transition to newer versions of Android.
-
 \section{Related work}
-\label{sec:related}
-Using the methods and data described in this paper we have determined that, on average, \daMeanInsecurityPerc\ of devices were exposed to known critical vulnerabilities between 2011 and 2014.
-Felt et al.\ studied 6 Android handsets in 2011 and found they were exposed to root vulnerabilities at least 74\% of the time~\cite{Felt2011} .
-Our approach differs from their study because they used data from 6 handsets and assumed the best possible update distribution, while our work is based on a large sample of devices tracking the actual update distribution.
-Nevertheless, our own analysis as well as comparison with their work suggests protection against critical vulnerabilities has not improved significantly over the last 3 years. 
-Felt et al.\ also found that 4 of the 46 malware samples (8\%) they analysed contained root exploits, much lower than rates found in later (larger) studies which found rates of 36.7\%~\cite{Zhou2012b} and 40\%~\cite{Zhou2012a} in 2012.
-
-Currently, finding and exploiting vulnerabilities is typically a time-consuming and manual process.
-Therefore an attacker has to invest significant human resource to first find a bug and then write code to exploit the vulnerability.
-To reduce the effort involved in finding a bug, attackers can look at security-related commits made to open source repositories.
-Previous work has shown attackers have taken this approach for the Firefox web browser~\cite{Barth2011}.
-Google does not release updates to the Android source code until after the release of the relevant security update.
-This reduces the time available for an attacker to exploit a vulnerability by looking at security updates disclosed in the source code repository from the period (\textbf{first fix}, \textbf{fix deployed}) to (\textbf{fixing release}, \textbf{fix deployed}).
-Unfortunately the time period between the \textbf{fixing release} and \textbf{fix deployed} can be of the order of months. 
-For example, the zergRush vulnerability was fixed in a release before it was publicly discovered but it still took 27 months for 90\% of the devices in our study to be definitely fixed\footnote{Definitely fixed is 1 $-$ (definitely insecure + maybe secure), maybe fixed is 1 $-$ definitely insecure} (25 months for 90\% of devices maybe fixed).
-
-Recent work by Brumley et al. has shown it is possible to automatically generate exploits from binary fixes, for example by using security updates issued by manufacturers~\cite{Brumley2008}. 
-Similarly, the Woodpecker tool automatically finds permission leaks in stock Android phone images~\cite{Grace2012}.
-Once such automated techniques mature, it will become even more important to provide security updates promptly to every Android device in the ecosystem because an attacker can turn the contents of a \textbf{fixing release} into an exploit in a matter of minutes or hours.
-
-Security in depth is also a useful strategy.
-In this regard, iOS provides additional safeguards beyond those used in Android, including a pre-distribution code review, mandatory code-signing by the manufacturer, and (with the important exception of ROP-based attacks~\cite{Wang2013a}) the technical prohibition of dynamic code loading by an app.
-These features, as well as Address Space Layout Randomisation (ASLR) and mandatory access controls, has resulted in a lower level of malware affecting iOS when compared to Android~\cite{Felt2011}.
-
-Rather than fixing critical vulnerabilities, security can be obtained by detect them and preventing the installation or execution of malicious apps.
-Detection strategies include RiskRanker, which classified 3\,281 out of 118\,318 apps (2.8\%) as risky of which 718 (22\%) were malware and 322 (10\%) were previously unknown malware, an overall infection rate of 0.6\% across multiple markets~\cite{Grace2012a}.
-DroidRanger also analysed apps finding 148 out of 182\,823 apps (0.08\%) to be malicious across multiple markets of which 29 were previously unknown~\cite{Zhou2012a}.
-%DroidRanger: It used permission-based behavioural fingerprinting which looked at the permissions of known malware and heuristic-based filtering -- dynamic loading of both Dalvik and native code.
-A common technique used by attackers takes existing popular apps and repackages them to include malicious code. 
-AnDarwin uses this insight to detect similar apps, and found 169 out of 265\,359 of all apps studied (0.06\%) were malicious clones of genuine apps~\cite{Crussell2013}.
-%AnDarwin: It used clustering based on semantic vectors derived from the program dependence graphs to detect similar apps.
-
-Unfortunately, whilst these tools were successful in finding critical vulnerabilities, such techniques are not perfect.
-The Malware Genome project collected 1\,260 malware samples from 2010--2011~\cite{Zhou2012b}.
-They found that the best case for anti-malware detection on their data was 79.6\%.
-Similarly, DroidChameleon found that antivirus products could not detect malware if the execution binary was automatically permuted~\cite{Rastogi2013}.
-Therefore detecting risky apps which might contain critical vulnerabilities is a useful tool, but it does not replace the need to provide security updates.
-
-%All these app analysis projects have been hampered by the difficulty of obtaining full datasets of Android apps as Google does not make these available (and researchers who automatically download them violate Google's Terms of Service).
-%PlayDrone was a particularly effective project which circumvented Google's protective measures and downloaded over 1\,100\,000 apps from Google Play, allowing an in depth analysis~\cite{Viennot2014}.
-
 The update process for apps, security fixes and OS upgrades also needs to be secure.
@@
 For example, the process can be exploited to allow apps to gain privilege through `Pileup' vulnerabilities by registering for new permissions before the update which creates that permission is installed~\cite{Xing2014}.
-
-The percentage of Android devices running the most recent version (\daUpdatednessPerc) compares unfavourably with the rate ($>90$\%) for Windows XP SP2 computers contacting the Microsoft update servers~\cite{Gkantsidis2006}.
-A simple numerical comparison is unfair because only one major OS version was considered in the Microsoft analysis, and data was only collected from computers which contacted the update server, although this was the default.
-More recent data demonstrates the difficulty of upgrading computers between major OS versions, with 27\% of Windows computers running Windows XP in July 2014,\footnote{\url{https://archive.today/PLGxn}} four months after Windows XP stopped receiving security updates.
 
@@
 \section{Conclusion}
-\label{sec:conclusion}
-We have investigated the security of the Android operating system with respect to security updates and examined how the Android ecosystem results in different device models receiving different levels of security due to whether or not they get security updates.
+ and examined how the Android ecosystem results in different device models receiving different levels of security due to whether or not they get security updates
 We have compared different device models, device manufacturers and network operators and found that there are differences between them which the discerning purchaser might use to influence their decision about which device model to buy from which device manufacturer and network operator.
 We hope that this analysis will encourage device manufacturers and network operators to improve the support they provide for devices after sale.
-Android attempts to provide good security to end users through sandboxing apps.
-However the latency in the security update process means that on average \daMeanInsecurityPerc\ of Android devices are exposed to known critical vulnerabilities which break out of the sandbox and only \daUpdatednessPerc\ of devices run the latest version of Android.
 
@@
 
-\listoftodos
-
 \end{document}

### 2015-03-18 securityupdates.tex
Strip out security updates and android ecosystem stuff from esorics paper and put in a ccs paper and make the esorics paper meet the page limit

@@
 
-\author{Daniel R.\ Thomas, Alastair R.\ Beresford, Daniel T.\ Wagner and Andrew Rice}
-
 \input{dastats}
 \newcommand{\da}{Device Analyzer}
-\newcommand{\dafoot}{\textsuperscript{\ref{foot:dadata}}}
 \input{avostats}
@@
 
-\author{\blindauthors{Daniel~R.~Thomas \and Daniel~T.~Wagner\and Alastair~R.~Beresford \and Andrew~Rice}}
+\author{\blindauthors{Daniel~R.~Thomas \and Alastair~R.~Beresford \and Daniel~T.~Wagner \and Andrew~Rice}}
 \institute{\blindauthors{Computer Laboratory,
@@
 Cambridge, United Kingdom\\
-\email{firstname.lastname@cl.cam.ac.uk}}}
+\email{Firstname.Lastname@cl.cam.ac.uk}}}
 
@@
 Many such exploits have been found, and therefore the security of a smartphone operating system relies on the prompt delivery of security updates.
-In this paper we tease apart the Android software ecosystem and quantify the extent to which the security sandbox provided by Android protects devices from malicious apps.
+In this paper we quantify the extent to which the security sandbox provided by Android protects devices from malicious apps.
 We analysed \da\ data from over \daOSYearsOfData\ years and \daNumOSDevices\ devices and found that on average \daMeanInsecurityPercNominal\ of devices were exposed to known root privilege vulnerabilities and only \daUpdatednessPercNominal\ of devices run the most recent version of Android.
-We find there is significant variability in the timely delivery of security updates across different device manufacturers and mobile network operators, and we define a security metric to rank the performance of device manufacturers and network operators in our data set.
 \end{abstract}
@@
 
-The development of Android is distributed among many parties and the security of Android relies on many open source projects, include the Linux kernel, OpenSSL and BouncyCastle as well as Google who build the core platform.
-In addition, device manufacturers (we know of \daNumManufacturers\footnote{\label{foot:dadata}We computed this from the \da~\cite{Wagner2013} data, see \S\ref{sec:android_update_process}.}) and network operators (\daNumOperators\dafoot) may make further modifications before devices are shipped to customers.
-Understanding this ecosystem is important as device manufacturers have introduced additional vulnerabilities in the past~\cite{Grace2012}. 
-We present a better understanding of both the ecosystem of Android development and associated vulnerabilities more fully in \S\ref{sec:android_ecosystem}.
-
 \begin{figure}[h]
@@
 
-It is clear from Figure~\ref{fig:proportioninsecure} that Android devices are often exposed to known security vulnerabilities.
-One solution is regulation, and indeed there is ongoing legal action to force network operators to ship updates for security vulnerabilities~\cite{Soghoian2013}.\todolater{Check on the status of this legal action}
-Many smartphones are sold on 12--24 month contracts, and yet our data shows many devices do not receive many security updates, with an overall average of \daUpdatesPerYearNominal\ per year. 
-In contrast, Windows XP could be purchased for a one-off payment in October 2001 and received security updates until April 2014.
-
-An alternative to regulation is access to publicly available data. 
-Comparative data would provide an incentive to both device manufacturers and network operators to provide updates.
-Currently corporate and public sector buyers are encouraged to purchase secure devices, but we have found little concrete guidance on the specific makes and models providing the best security. 
-For example, CESG, which advises the UK government on how to secure its computer systems, recommends picking Android device models from device manufactures which are good at shipping security updates promptly~\cite{CESG2013} but it does not state which device manufacturers these are.
-To address this, we developed a scoring system and provide numbers on the historic performance of device models in the \da\ project in~\S\ref{sec:security_scoring}.
-
 In summary, the contributions of this paper are:
@@
  \item We created an open database to collate quantitative information on vulnerabilities which affect Android, and investigate their time-lines.
- \item We measure the security of Android according to three metrics and compare different device manufacturers and device models to allow device purchasers to differentiate between them based on security.
 \end{itemize}
@@
 
-We require two sources of data for this study: (1) information on the distribution of installed versions of Android over time across the Android ecosystem and (2) information on the critical vulnerabilities found to affect specific versions of Android.
+We require two sources of data for this study: (1) information on the distribution of installed versions of Android over time and (2) information on the critical vulnerabilities found to affect specific versions of Android.
 These two datasets can then be combined to determine the proportion of handsets at risk of attack from specific vulnerabilities for particular periods of time.
@@
 If an application exploits a critical vulnerability then it gains control of the device.
-Some phones can be `rooted' by enabling USB debugging and using the special privileges of the ADB shell to root the device but only \daAdbEnabledPerc\dafoot\ of devices have USB debugging enabled.
+Some phones can be `rooted' by enabling USB debugging and using the special privileges of the ADB shell to root the device but only \daAdbEnabledPerc\ of devices have USB debugging enabled.
 This is not something that applications running on the phone can exploit to break out of the app sandbox and so we do not include those vulnerabilities.
@@
 
-\section{Android Ecosystem}\label{sec:android_ecosystem}
-There is a complex Android ecosystem which creates and distributes updates to Android which fix vulnerabilities.
-In this section we describe how the Android ecosystem functions and how Android versions, as seen by the \da\ project, are produced.
-
-\subsection{Android update process}
-
-\label{sec:android_update_process}
-\begin{figure}[h]
- \centering
- \def\svgwidth{0.7\columnwidth}
- \import{figures/}{update_ecosystem.pdf_tex}
- \caption{Flow of updates between participants in the Android ecosystem.
- Numbers on edges indicate updates shipped between \daStartDate\ and \daEndDate, those in brackets represent number of such entities.
- Dotted arrows indicate flows where we can't measure how many updates are produced as they are not public.}
- \label{fig:update_ecosystem}
-\end{figure}
-To understand how vulnerabilities in Android are fixed we must examine the Android update process which we model in Figure~\ref{fig:update_ecosystem}.
-There are five entities or groups which contribute towards Android updates: the network operators, the device manufacturers, the hardware developers, Google and the upstream open source projects.
-Android builds on various open source projects such as the Linux kernel, OpenSSL and BouncyCastle cryptography libraries.
-Consequently Android can include any compatible versions of those projects, including those which fix security vulnerabilities.
-Android also incorporates various drivers for different bits of hardware.
-The Android platform is then built from these components by Google.
-The code for each Android release or update is kept secret\footnote{\url{https://source.android.com/source/code-lines.html}} until after a binary release has been published.
-Device manufacturers likely receive advanced access in order to prepare handsets so they can customise it before passing it on to the network operator.
-The network operator may then make or request further customisations and perform further testing before shipping the update to the Device.
-Sometimes device manufactures ship updates directly to the user, sometimes the device manufacturer and Google collaborate closely to make a particular phone, such as with Nexus devices.
-Sometimes device manufacturers incorporate upstream open source project releases directly, and sometimes incorrectly -- for example previous work has recorded evidence of broken nightly builds of sqlite in Android releases~\cite{Wagner2013}.
-
-The numbers of devices (\daNumOSDevices), network operators (\daNumOperators) and device manufacturers (\daNumManufacturers) in Figure~\ref{fig:update_ecosystem} come from the \da\ data.
-Device manufacturer and network operator counts were obtained by normalising the results reported by Android to \da\ of the device manufacturer and active network operator.
-This normalisation is a manual task involves removing invalid values (such as `manufacturer' or `airplane mode is on'), collating across company name changes (e.g.\ `lge' to `LG'), normalising punctuation, removing extra strings sometimes added such as (`(2g)' or `communications') and mapping some incorrectly placed model names back to their manufacturer.
-This normalisation is not perfect and so these are overestimates on the \da\ data but they are likely still underestimates as there will be some device manufacturers and network operators which are not included in the \da\ data.
-
-In Figure~\ref{fig:update_ecosystem} the number of updates received by devices (\daNumFullVersions) is the number of different full version strings observed in \da.
-The number of updates shipped by Google (\daNumSigOSVersions) is the number of Android versions reported in \da\ which affected more than \daSigVersionPerc\ of devices for more than \daSigVersionDays\ days.
-This significance test is to remove spurious versions recorded in \da\ such as `5.2.0' in 2012 which has still not been released in 2014.
-
-We extracted data on the external projects used in Android and have included this and the scripts which generated it in \avo.
-These scripts analysed the Android Open Source Project's source tree to examine the source code of each of the external projects to find the project version associated with each Android version tag on the repository.
-There are \avoNumExternalProjects\ external open source projects in Android, contributing \avoTotalExternalLines\ lines of code.\footnote{Lines of code were measured using David A.
-Wheeler's \texttt{sloccount}.}
-We analysed the top \avoNumBigExternalProjects\ by lines of code (\avoBigExternalLinesOfCodePerc\ of the total) and were able to automatically extract the versions of those projects included in different versions of Android for \avoNumAnalysedExternalProjects\ of these (\avoAnalysedExternalLinesOfCodePerc\ of the total).
-We found \avoBigExternalTotalVersions\ distinct versions, a median of \avoBigExternalMedianVersions\ and mean of \avoBigExternalMeanVersions\ versions per project.
-Android rarely changes the version of external projects it includes.
-
-%An analysis by Vidas et al.~\cite{Vidas2011} of the Android 2.1 to 2.2 update found that it took 11 months from when Google released 2.2 for the last device which they were investigating to get the update.
-To compute the latency between upstream releases and their inclusion in Android we scraped the release pages for those projects, to obtain the version numbers and release dates.
-This allows us to compute the latency between an upstream project being released and it being included in Android, this is shown in Table~\ref{tab:update_ecosystem}.
-The versions included in Android were about half a year old when the first version of Android containing it was released.
-\begin{table}
-\centering
-\normalsize
-\begin{tabular}{l|r|r}
-Project	&	\# releases	&	latency (days) \\ \hline
-linux	&	\linuxNumVersions	&	\linuxMeanUpdateLatency \\
-openssl	&	\opensslNumVersions	&	\opensslMeanUpdateLatency \\
-bouncycastle	&	\bouncycastleNumVersions	&	\bouncycastleMeanUpdateLatency \\
-\end{tabular}
-\caption{Flow of updates from upstream projects into Android.
-Number of updates as in Figure~\ref{fig:update_ecosystem}, latency in days for all pairs of versions we have data on.\todolater{scrape the other 26 websites...
-is it worth it?}}
-\label{tab:update_ecosystem}
-\end{table}
-
 \section{Experiments}
@@
 Devices, on average, apply \daUpdatesPerYear\ updates each year and \daPercUpdatesDowngrades\ of full version changes are downgrades to older versions.
-Nexus devices are better than others, \emph{\daSecScoreBestmanufacturer} is the highest scoring device manufacturer and \emph{\daSecScoreBestoperator} is the highest scoring network operator.
-%\todo{comparison of nexus update rates with iOS}
-We describe four experiments.
-We examine the vulnerability of the Android ecosystem as a whole (\S\ref{sec:exp:versionsecurity}, explore the upgrade cycle and vulnerability cycle of the ecosystem (\S\ref{sec:exp:android_ecosystem}), quantify the updates installed on particular devices (\S\ref{sec:exp:device_updates}) and compare the performance of different device manufacturers and network operators (\S\ref{sec:exp:security_score}).
-
-%% In an experiment you test a hypothesis, it needs to be clear what the hypothesis is and what the conclusion of the test is
+We describe three experiments.
+We examine the vulnerability of the Android ecosystem as a whole (\S\ref{sec:exp:versionsecurity}, explore the upgrade cycle and vulnerability cycle of the ecosystem (\S\ref{sec:exp:android_ecosystem}), quantify the updates installed on particular devices (\S\ref{sec:exp:device_updates}).
 
@@
 The first few vulnerabilities are all `maybe secure' as we do not have data from before that vulnerability was discovered while later ones are mostly `definitely insecure' as we know no fixing update reached the devices.
-The change in behaviour after August 2014 is explained in \S\ref{sec:da_changes}}
+ %The change in behaviour after August 2014 is explained in \S\ref{sec:da_changes}
+ }
  \label{fig:nvulnerabilities_heat}
@@
 Those graphs summarise data across all the devices, however one of the advantages of the \da\ data is that it allows us to look at what happens to individual devices over time.
+
+{\begin{figure}[p]
+ \centering
+ \begin{subfigure}[b]{\columnwidth}
+  \includegraphics[width=\columnwidth]{figures/device-data-all-os}
+  \caption{OS versions over time. Black lines show where only the build number changed.}
+  \label{fig:device_data_os}
+ \end{subfigure}
+ \begin{subfigure}[b]{\columnwidth}
+  \includegraphics[width=\columnwidth]{figures/device-data-all-security}
+  \caption{Number of vulnerabilities affecting each device over time}
+  \label{fig:device_data_security}
+ \end{subfigure}
+ \caption{The top \daNumDeviceDataDevices\ devices by days of contribution in the \da\ data. One strip per device handset.}
+ \includegraphics[width=0.55\columnwidth]{figures/from_to_updates.pdf}
+ \caption{Updates between different Android versions in the \da\ data}
+ \label{fig:from_to_updates}
+\end{figure}}
 
@@
 \subsubsection{Results}
-\begin{figure}
- \centering
- \begin{subfigure}[b]{\columnwidth}
-  \includegraphics[width=\columnwidth]{figures/device-data-all-os}
-  \caption{OS versions over time. Black vertical lines indicate that the build number changed without changing the OS version at that point.}
-  \label{fig:device_data_os}
- \end{subfigure}
- \begin{subfigure}[b]{\columnwidth}
-  \includegraphics[width=\columnwidth]{figures/device-data-all-security}
-  \caption{Number of vulnerabilities affecting each device over time}
-  \label{fig:device_data_security}
- \end{subfigure}
- \caption{The top \daNumDeviceDataDevices\ devices by days of contribution in the \da\ data. One strip per device handset.}
-\end{figure}
+
 The longitudinal data on the number of vulnerabilities affecting the \daNumDeviceDataDevices\ devices which have contributed the most days of data to \da\ changes over time is shown in Figure~\ref{fig:device_data_security}.
@@
 
-\begin{figure}
- \centering
- \includegraphics[width=0.6\columnwidth]{figures/from_to_updates.pdf}
- \caption{Updates between different Android versions in the \da\ data}
- \label{fig:from_to_updates}
-\end{figure}
 \begin{figure}[h]
  \includegraphics[width=\columnwidth]{figures/nw_security_updates.pdf}
- \caption{Proportion of updates each week which fixed or may have fixed security vulnerabilities}
+ \caption{Proportion of updates each week which (may have) fixed vulnerabilities}
  \label{fig:weekly_security_updates}
@@
 By taking the number of updates observed and the number of device days of data we have collected we can compute the number of updates received by a device per year to be \daUpdatesPerYear.
-This compares badly with the number of critical vulnerabilities discovered per year of between \avoVulnsPerYearAllAndroid\ (affecting all Android) and \avoVulnsPerYear\ (including the device manufacturer specific ones in \avo).
-
-\input{security_scoring}
+This compares badly with the number of critical vulnerabilities discovered per year of between \avoVulnsPerYearAllAndroid\ (affecting all Android) and \avoVulnsPerYear\ (including the device manufacturer specific ones).
 
@@
  \includegraphics[width=\columnwidth]{figures/googleplayapi}
- \caption{Google Play data on proportion of devices running different Android API versions}
+ \caption{Google Play data on proportion of devices running different API versions}
  \label{fig:play_api}
@@
  \includegraphics[width=\columnwidth]{figures/norm_api_gpcomp}
- \caption{\da\ data on proportion of devices running different Android API versions}
+ \caption{\da\ data on proportion of devices running different API versions}
  \label{fig:da_api}
@@
 This means that the OS version information from \da\ is likely to be overestimating the prevalence of new OS versions and hence our results are a conservative estimate of the security of Android.
+This allows us to have confidence in the OS version information.
 \todolater{we want a statistical metric to claim this strongly with.}
-This allows us to have confidence in the OS version information.
 
@@
 
-The update process for apps, security fixes and OS upgrades also needs to be secure.
-Unfortunately, package management systems designed to provide secure updates have been found to contain vulnerabilities~\cite{Cappos2008} and many software update systems fail to authenticate the connection between the device and the update server or do not authenticate the downloaded binaries~\cite{Bellissimo2006}.
-Android does authenticate update binaries and Google Play downloads them over a secure connection~\cite{Viennot2014}.
-In this paper we have analysed four critical vulnerabilities in the Android app update mechanism: APK unsigned shorts, APK unchecked name, APK duplicate file and Fake ID.
-Other work has demonstrate complex and subtle errors exist in the Android app update process.
-For example, the process can be exploited to allow apps to gain privilege through `Pileup' vulnerabilities by registering for new permissions before the update which creates that permission is installed~\cite{Xing2014}.
-
 The percentage of Android devices running the most recent version (\daUpdatednessPerc) compares unfavourably with the rate ($>90$\%) for Windows XP SP2 computers contacting the Microsoft update servers~\cite{Gkantsidis2006}.
@@
 
-User-Agent strings have been used to investigate the timeliness of web browser updates~\cite{Frei2008}, with at most 80\% of Firefox users running the most recent version.
-The same analysis was used to show that Chrome's use of silent updates seems to increase uptake of upgrades~\cite{Duebendorfer2010} with 97\% of users running the latest version within 3 weeks of release.
-By way of comparison, Android's update process is manual.
-The user is notified an update exists, but further action is required, including downloading the update and rebooting the phone to enable installation.
-The phone must have sufficient charge to perform the update and the device itself is rendered inoperable during the update process, two factors which might prevent or delay the update process from taking place.
-In our data we are unable to determine why a device is not updated. 
-It is possible that many updates arrive at handsets, but are simply not installed.
-Anecdotal evidence at least suggests that it is the lack of updates rather the lack of installation which is the major problem at present. Further work is required to tease these numbers apart.
-%Partly this is the result of the fact that an operating system update is being installed and so a reboot is required, but Chrome installs the new version side by side with the old one and switches the next time it is restarted.
-%The same technique would be more difficult on phones with limited storage space (as many cheap Android phones have barely enough space to install just the update) but is a plausible improvement for more high-end devices.
-%Google is deploying the same silent update technique through Google Play Services\footnote{\href{http://lifehacker.com/why-google-play-services-are-now-more-important-than-an-975970197}{http://lifehacker.com/why-google-play-services-are-now-more-important-than-an-975970197}} which automatically installs updates for core Google components of Android, this also bypasses the device manufacturer and network operator.
-
 \section{Conclusion}
 \label{sec:conclusion}
-We have investigated the security of the Android operating system with respect to security updates and examined how the Android ecosystem results in different device models receiving different levels of security due to whether or not they get security updates.
-We have compared different device models, device manufacturers and network operators and found that there are differences between them which the discerning purchaser might use to influence their decision about which device model to buy from which device manufacturer and network operator.
-We hope that this analysis will encourage device manufacturers and network operators to improve the support they provide for devices after sale.
+We have investigated the security of the Android operating system with respect to security updates.
 Android attempts to provide good security to end users through sandboxing apps.
@@
 \section*{Acknowledgements}
-Thanks to Richard Clayton, the mobile security reading group and Anil Madhavapeddy for reading various drafts of this paper.
-Thanks to David Robertson for helpful advice on statistical analysis.
+Thanks to Richard Clayton, the mobile security reading group and Anil Madhavapeddy for reading drafts.
+Thanks to David Robertson for advice on statistical analysis.
 Thanks to Laurent Simon, Thomas Coudray, Adrian Taylor, Justin Case, Giant Pune and Khilan Gudka for reporting vulnerabilities in Android.

### 2015-03-19 securityupdates.tex
Add an avovulnfoot command and then make it produce nothing to save space

@@
 \newcommand{\avo}{AVO}
+\newcommand{\avovulnfoot}[1]{}%{\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/#1}}}
 \newcommand{\percMarketShare}{83.6\%~\footnote{\url{http://www.theinquirer.net/inquirer/news/2379036/android-hits-836-percent-marketshare-while-ios-windows-and-blackberry-slide}}}
@@
 The variation of the proportion of devices affected by a vulnerability with time tells us how badly a particular vulnerability affected the Android platform.
-In July 2011 at the beginning of the \da\ data the \emph{exploid}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/exploid_udev}} and \emph{levitator}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/levitator}} vulnerabilities both affect most Android devices, slowly these are fixed as updates roll out and devices are replaced until in January 2013 a much smaller proportion of devices are affected by known vulnerabilities.
+In July 2011 the \emph{exploid}\avovulnfoot{exploid_udev} and \emph{levitator}\avovulnfoot{levitator} vulnerabilities both affect most Android devices, slowly these are fixed as updates roll out and devices are replaced until in January 2013 a much smaller proportion of devices are affected by known vulnerabilities.
 However when in February 2013 the first APK signing vulnerability was found which affected all previous versions of Android and even in October 2013 most devices (\daVulnAPKDuplicateFileOctoberPerc) were still vulnerable.
@@
 Figure~\ref{fig:nvulnerabilities_heat} shows how the the \emph{APK signing vulnerabilities} affected all devices and took months to get fixed for any device.
-However what is perhaps more worrying is the long tail on the \emph{Gingerbreak}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/Gingerbreak}}, \emph{levitator}, \emph{exploid} and \emph{zergRush}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/zergRush}} vulnerabilities which are more dangerous root vulnerabilities (not requiring new APK installation) and which still affect a significant proportion of devices years later.
+However what is perhaps more worrying is the long tail on the \emph{Gingerbreak}\avovulnfoot{Gingerbreak}, \emph{levitator}, \emph{exploid} and \emph{zergRush}\avovulnfoot{zergRush} vulnerabilities which are more dangerous root vulnerabilities (not requiring new APK installation) and which affect a significant proportion of devices years later.
 

### 2015-03-19 securityupdates.tex
Shorten the text

@@
 Not all malicious apps need to break the sandbox in order to misbehave.
-Indeed, the above figures suggest the majority of malicious apps do not attempt to do so.
 We focus on the issue of apps breaking out of the sandbox because malware which is able to take full control of a handset can do significantly more harm and is much harder to remove.
@@
 Alternative markets are also popular, particularly in countries such as China where the Google Play is not available. 
-Previous work has shown that on alternative markets between 0.20\% and 0.47\% of apps are malicious~\cite{Zhou2012a}.
+In 2012 between 0.20\% and 0.47\% of apps on alternative markets were malicious~\cite{Zhou2012a}.
 
@@
 Instead the attacker can chose to send a malicious payload directly to the app for execution at a later point in time.
-Note that dynamic code loading is not actually needed for this to work---it just makes the task easier---since previous work has shown that even on a platform such as iOS, which does not permit dynamic code loading, a ROP-based attack is relatively easy if the attacker creates an app with carefully crafted flaws~\cite{}.
+Note that dynamic code loading is not actually needed for this to work -- it just makes the task easier -- since previous work has shown that even on a platform such as iOS, which does not permit dynamic code loading, a ROP-based attack is relatively easy if the attacker creates an app with carefully crafted flaws~\cite{Wang2013a}.
 
@@
 This attack works when: (1) the WebView has the Javascript-to-Java bridge enabled; and (2) the attacker is able to inject a malicious Javascript payload into the network traffic between the phone and the web server.
-Many apps are vulnerable to this type of attack because ad network libraries typically embed adverts into apps by displaying HTML and Javascript inside a WebView with the Javascript-to-Java bridge enabled.
+Many apps are vulnerable to this type of attack because ad-libraries typically embed adverts into apps by displaying HTML and Javascript inside a WebView with the Javascript-to-Java bridge enabled. 
 The fix for the addJavascriptInterface vulnerability breaks backwards compatibility, and therefore apps are only immune to this attack if their target API level is greater than 16 and they are running on an Android device with an OS version of 4.2 or greater.
 Later analysis in Figure~\ref{fig:da_api} in this paper shows that 40\% of handsets connecting to the Google Play store are currently vulnerable to this attack.
+\todo{Make that percentage automatically generated so that it is correct}
 
@@
 In our analysis we use historical data collected by the \da\ project~\cite{Wagner2013}.
-Device Analyzer collects data from study participants who install the project Android app from the Google Play store.
+\da\ collects data from study participants who install the Android app from the Google Play store.
 Most study participants allow researchers around the world to access a subset of their device data, including the data presented in this paper.
@@
 For this paper, we extracted the build string and API version from the device log each day.
+The build string is a user-readable version string.
 The API version is a positive integer which increases when new features are added to the API.
 Consequently security (bug) fixes do not result in a change in the API version.
-The build string is a user-readable version string.
 Fortunately most (\daOSVersionPercValidLines) entries in the data have a build string of the form `x.y.z opaque\_marker' and so it is possible to extract the Android version number `x.y.z'.
@@
 
-Where possible we indicate the uncertainty in our results by presenting them $\pm$ one standard deviation and give results to 3 significant figures, this occasionally results in `${} \pm 0$' when the standard deviation is less than the number of significant figures we are reporting.
-This technique cannot account for systematic errors and therefore we explore the potential for systematic errors in the \da\ data in \S\ref{sec:representative}.
+Where possible we indicate the uncertainty in our results by presenting them $\pm$ one standard deviation and give results to 3 significant figures, this occasionally results in `${} \pm 0$' when the standard deviation is less than the number of significant figures.
+This technique cannot account for systematic errors which we explore in \S\ref{sec:representative}.
 
@@
 We only looked for critical vulnerabilities such as root vulnerabilities which did not require USB debugging to exploit.
-Critical vulnerabilities allow a program to gain privileges equivalent in scope to root.
-If an application exploits a critical vulnerability then it gains control of the device.
+If an malicious code exploits a critical vulnerability then it gains control of the device.
 Some phones can be `rooted' by enabling USB debugging and using the special privileges of the ADB shell to root the device but only \daAdbEnabledPerc\ of devices have USB debugging enabled.
@@
 
-We developed and maintain an open platform for filing critical vulnerabilities in a machine readable format, the AndroidVulnerabilities.org\footnote{\url{http://androidvulnerabilities.org/} \textbf{Note to reviewers:} we have removed our names and affiliation from the website for this review \textbf{with the exception of the contact page.
-Please do not visit the contact page until the review cycle has completed.}} (\avo) website.
-We seeded it with data from the CVE database, vendor lists, reports from the literature and various forums.
+We developed and maintain an open platform for filing critical vulnerabilities in a machine readable format, the \href{http://androidvulnerabilities.org/}{AndroidVulnerabilities.org}\footnote{\textbf{Note to reviewers:} we have removed our names and affiliation from the website for this review \textbf{with the exception of the contact page.
+Please do not visit the contact page until the review cycle has completed.}\vspace{-3em}} (\avo) website.
+We seeded it with data from the CVE database, vendor lists, reports from the literature and forums.
 In addition, we have received submissions or amendments from \avoNumSubmitters\ individuals.
-We collected data between \avoStartDate\ and \avoEndDate\ and so any information lost before the start of that period cannot be included.
+We collected data between \avoStartDate\ and \avoEndDate\ and so does not include any information lost before the start of that period.
 \avo\ currently contains \avoNumVulnerabilities\ vulnerabilities of which \avoNumVulnAllAndroid\ affect all Android devices and \avoNumVulnSpecific\ are specific to particular devices or device manufacturers.
@@
 We have chosen these vulnerabilities to fit the attack vectors introduced in \S\ref{sec:threatmodel}.
-These vulnerabilities affect all Android devices regardless of manufacturer, and as a result our selected vulnerabilities will dominate security analysis of the Android ecosystem as a whole.
+These vulnerabilities affect all Android devices regardless of manufacturer, and as a result our selected vulnerabilities will dominate security analysis of Android.
 In many cases we could not match manufacturer- and model-specific vulnerabilities to individual devices in the \da\ data and therefore attempting to include device-specific vulnerabilities as well would introduce additional uncertainty in our results.
-In contrast, with our chosen set of vulnerabiltiies, our analysis represents a lower-bound on the vulnerability of devices in the \da\ data set.
+In contrast, with our chosen set of vulnerabilities, our analysis represents a lower-bound on the vulnerability of devices in the \da\ data set.
 
@@
 
-The key events in the lifetime of a vulnerability do not always occur in the same order and are:\\
-\textbf{creation} When a vulnerability was created in the source code.\\
-\textbf{introducing release} When the first release was made containing the vulnerability.\\
-\textbf{discovery} When the vulnerability is first discovered.\\
-\textbf{exploit} When the vulnerability is first exploited.\\
-\textbf{disclosure} When the vulnerability is first disclosed.\\ % publically or to a smaller set of people?
-\textbf{fix} When a vulnerability was first fixed in the source code.\\
-\textbf{fixing release} When the first release containing the fix was made (equivalent to `patch available'~\cite{Frei2010}).\\
-\textbf{fix deployed} When the fix has been deployed to a sufficiently large proportion of the population that the vulnerability can be ignored (the ecosystem equivalent to the per instance `patch installed'~\cite{Frei2010}).
-\todolater{Does this look nice?}
+The key events in the lifetime of a vulnerability do not always occur in the same order and are:
+\begin{description}
+\item[creation] When a vulnerability was created in the source code.
+\item[introducing release] When the first release was made containing the vulnerability.
+\item[discovery] When the vulnerability is first discovered.
+\item[exploit] When the vulnerability is first exploited.
+\item[disclosure] When the vulnerability is first disclosed. % publically or to a smaller set of people?
+\item[fix] When a vulnerability was first fixed in the source code.
+\item[fixing release] When the first release containing the fix was made (equivalent to `patch available'~\cite{Frei2010}).
+\item[fix deployed] When the fix has been deployed to a sufficiently large proportion of the population that the vulnerability can be ignored (the ecosystem equivalent to the per instance `patch installed'~\cite{Frei2010}).
+\end{description}
 
 Establishing when a vulnerability starts to pose a threat to users is difficult.
-Frei et al.~\cite{Frei2010} propose the definition of the {\bf time of disclosure}.
-This occurs when the information about the vulnerability is freely available to the public from a widely accepted and independent source and has been validated by security experts so that it has a risk rating.
+Frei et al.~\cite{Frei2010} propose the definition of the \textbf{time of disclosure}.
+This occurs when the information about the vulnerability is freely available to the public from an accepted and independent source and has been validated by security experts so that it has a risk rating.
 Unfortunately before we collated this information much of it was not published by an independent source and lacked risk rating information, even months or years after they had been actively used.
@@
 However they also show that zero-day vulnerabilities are typically used for 312 days before they are publicly disclosed, often to target particular organisations.
-Hence when considering a vulnerability from the point of view of an organisation which cares about Advanced Persistent Threats, such as those responding to CESG advice~\cite{CESG2013}, the critical period starts with \textbf{discovery} and continues until \textbf{fix deployed}.
+Hence when considering a vulnerability from the point of view of an organisation which cares about Advanced Persistent Threats, the critical period starts with \textbf{discovery} and continues until \textbf{fix deployed}.
 Therefore, in our vulnerability calculations we prefer to use the earliest recorded date of \textbf{discovery}, even if that knowledge might have been confined to a particular device manufacturer or hobbyist.
@@
 
-Unfortunately {\bf discovery} is the hardest point to obtain concrete data on as discovery may happen multiple times independently and not all discoverers will report their discovery.
-{\bf Creation} and {\bf introducing release} are the earliest points which could be used, but the risk is mostly latent until someone discovers them and therefore we do not use these dates.
+Unfortunately \textbf{discovery} is the hardest point to obtain concrete data on as discovery may happen multiple times independently and not all discoverers will report their discovery.
+\textbf{Creation} and \textbf{introducing release} are the earliest points which could be used, but the risk is mostly latent until someone discovers them and therefore we do not use these dates.
 %, for example entities with high value targets may keep databases of unknown vulnerabilities for later use, in which case there is a high risk to potential targets from the never disclosed point of discovery.
 %However if the discoverer never discloses the vulnerability to anyone or makes any use of it then there is little danger.
-The date of first {\bf exploit} is a point when the risk is definitely high, but again a good adversary will not be detected when using such exploits.
-The date of first {\bf fix} is, assuming that the fix is deliberate, a point at which the vulnerability is known at least within the organisation performing the fix and frequently implies an earlier discovery and notification by a third party; we determine this date from the Android Open Source Project repository by examining the authored-on date of the fixing commit.
+The date of first \textbf{exploit} is a point when the risk is definitely high, but again a good adversary will not be detected when using such exploits.
+The date of first \textbf{fix} is, assuming that the fix is deliberate, a point at which the vulnerability is known at least within the organisation performing the fix and frequently implies an earlier discovery and notification by a third party; we determine this date from the Android Open Source Project repository by examining the authored-on date of the fixing commit.
 %Once a {\bf fixing release} has been made then the vulnerability is widely known because it can be reverse engineered from the changes in the release~\cite{Brumley2008}.
@@
 Therefore, in this paper we use the earliest known date from \textbf{discovery}, \textbf{exploit}, \textbf{disclosure} or \textbf{fix}.
-A breakdown of the type of date used per vulnerability is shown in Table~\ref{tab:andvulns}.
+A breakdown of the type of date used is shown in Table~\ref{tab:andvulns}.
 
@@
 During the period in which \da\ data was collected the date when a version of Android with the fix was observed on a \da\ device is taken as the date that version was released.
-For vulnerabilities prior to the collection period we estimate the release date as best we can using publicly available data.
-There is no canonical source of Android release dates, and our best guesses and supporting references are available from \avo.
+For vulnerabilities prior to the collection period we estimate the release date using publicly available data.
+There is no canonical source of Android release dates, our best guesses and supporting references are available from \avo.
 
-The discovery dates of our vulnerabilities does not appear to be uniform. 
+The discovery dates of our vulnerabilities are not uniform. 
 In particular the data shows a large gap from 2011-10-06 to 2013-02-18 where we do not know of any discoveries of critical vulnerabilities affecting all Android devices.
 The cause of this quiet period is unclear.
-Possible explanations are that: (i) most devices were exposed to known vulnerabilities so there was no point in looking for new ones (from Figure~\ref{fig:proportioninsecure});
+Possible explanations are that: (i) most devices were exposed to known vulnerabilities so there was reason to look for new ones;
 (ii) device manufacturers made it easier to install custom versions of Android, reducing the need for users to root their devices; and
-(iii) device manufacturer specific vulnerabilities (which are not in our analysis here) were easier to find and therefore attackers looking for vulnerabilities adjusted their focus.
+(iii) device manufacturer specific vulnerabilities (which are not in our analysis here) were easier to find and so attackers looking for vulnerabilities adjusted their focus.
 
@@
 Devices, on average, apply \daUpdatesPerYear\ updates each year and \daPercUpdatesDowngrades\ of full version changes are downgrades to older versions.
-We describe three experiments.
-We examine the vulnerability of the Android ecosystem as a whole (\S\ref{sec:exp:versionsecurity}, explore the upgrade cycle and vulnerability cycle of the ecosystem (\S\ref{sec:exp:android_ecosystem}), quantify the updates installed on particular devices (\S\ref{sec:exp:device_updates}).
+We describe three experiments, examining the vulnerability of Android devices in general (\S\ref{sec:exp:versionsecurity}, explore the upgrade cycle and vulnerability cycle of the ecosystem (\S\ref{sec:exp:android_ecosystem}), quantify the updates installed on particular devices (\S\ref{sec:exp:device_updates}).
 
@@
 
-In the introduction, Figure~\ref{fig:proportioninsecure} described the proportion of Android devices susceptable to at least one critical vulnerability.
+In the introduction, Figure~\ref{fig:proportioninsecure} described the proportion of Android devices susceptible to at least one critical vulnerability.
 
 \subsubsection{Method} 
-To investigate the behaviour of the Android ecosystem as a whole with respect to Android versions and vulnerabilities we used the OS version information from \da\ and the vulnerability data from \avo.
-The \da\ data used here was collected between \daStartDate\ and \daEndDate.
+To investigate the vulnerability of Android devices, we used the OS version information from \da\ and the vulnerability data from \avo.
+The \da\ data was collected between \daStartDate\ and \daEndDate.
 The \avo\ data covers the period from \avoFirstDataDate\ to \avoLastDataDate.
-For each device we collected daily version data  and recorded the running version and any associated vulnerabilities it was exposed to at that time (if any).
-We then normalise these totals for each day by dividing through by the total number of devices with version information seen on that day.
+For each device we collected daily version data and any vulnerabilities it was exposed to at that time.
+We then normalise these totals for each day by dividing through by the total number of devices with version information on that day.
 
@@
 This demonstrates the importance of a longitudinal study: this type of analysis requires years of data.
-Once \emph{zergRush} was discovered in October 2011 the most devices are recorded as \emph{insecure} (red) as most devices were exposed to that vulnerability.
+Once \emph{zergRush} was discovered in October 2011 the most devices are recorded as \emph{insecure} (red) as they were exposed to that vulnerability.
 The remanding devices were already running a version of Android which fixed the \emph{zergRush} vulnerability and are therefore marked as \emph{secure} (green).
 From October 2011 until the discovery of \emph{APK duplicate file} in February 2013 the graph shows progressive improvement as devices are upgraded.
-This means more and more devices are marked as \emph{secure} (green) because they are now running a secure version of Android, or marked as \emph{maybe secure} because they received a minor update which may have contained a backported fix.
+This means more and more devices are marked as \emph{secure} because they are now running a secure version of Android, or marked as \emph{maybe secure} because they received a minor update which may have contained a backported fix.
 From February 2013 onwards regular discovery of critical vulnerabilities ensures that most devices are exposed to known critical vulnerabilities.
@@
 \subsection{Experiment 2: Updates to particular devices}\label{sec:exp:device_updates}
-Those graphs summarise data across all the devices, however one of the advantages of the \da\ data is that it allows us to look at what happens to individual devices over time.
+Those graphs summarise data across all the devices, however one of the advantages of the \da\ data is that we can look at what happens to individual devices over time.
 
@@
 The trend that we saw in Figure~\ref{fig:proportioninsecure} of security improving and then getting worse is also shown here.
-It shows how some devices had vulnerabilities, which were fixed, and then further vulnerabilities were discovered, and for these devices, mostly not fixed.
-This implies that these devices had been abandoned by the device manufacturer and were not receiving updates which is confirmed by Figure~\ref{fig:device_data_os} which shows which OS versions those devices were running.
+It shows how some devices had vulnerabilities, which were fixed, and then further vulnerabilities were discovered, mostly not fixed.
+This implies that these devices had been abandoned by the device manufacturer and were not receiving updates, this is confirmed by Figure~\ref{fig:device_data_os} which shows which OS versions those devices were running.
 Some devices start off in 2011 exposed to known security vulnerabilities and are still exposed to those and additional ones in 2014.
@@
 The number of devices getting security updates each week, is shown in Figure~\ref{fig:weekly_security_updates}.
-\begin{itemize}
- \item Updates which changed the Android version number from a version with known vulnerabilities to one which had fewer known vulnerabilities are shown in red.
- \item Updates which changed the build number but not the version number and so might contain a backported fix for a vulnerability are shown in yellow.
-\item Updates which did not fix security vulnerabilities (because there were no known security vulnerabilities in the version of Android they were already running) are shown in green.
-
-\end{itemize}
+Updates which changed the Android version number so that the number of known vulnerabilities decreased are shown in red.
+Updates which changed the build number but not the version number and so might contain a backported fix for a vulnerability are shown in yellow.
+Updates which did not fix security vulnerabilities (because there were no known security vulnerabilities in the existing version of Android) are shown in green.
 
@@
 The data from \da\ we used to investigate the proportion of devices exposed to different vulnerabilities is the OS version.
-Unfortunately there is no authoritative source of OS version information and so we cannot check directly whether our data is representative.
+Unfortunately there is no ground truth of OS version information.
 However Google has published API version information almost every month since December 2009 and we have collated this information.\footnote{\url{http://androidvulnerabilities.org/play/historicplaydashboard}}
-While API versions are too coarse grained to use for security update detection they are closely related to OS versions and so if the \da\ data on API versions is similar to the Google Data on API versions then the \da\ data on OS versions should be representative.
+While API versions are too coarse grained to use for security update detection they are closely related to OS versions and so if the \da\ data on API versions are similar to the Google data on API versions then the \da\ data on OS versions should be representative.
 Figure~\ref{fig:play_api} shows the data from Google and Figure~\ref{fig:da_api} shows the data from \da\ and they appear similar.
@@
 \label{sec:da_changes}
-The \da\ data is mostly generated by devices which have the \da\ app installed because their owner happened to come across the app on Google Play as there has been no advertising and few other attempts to increase usage.
-However two collaborations with network operators to install \da\ on customer devices resulted in large numbers of new \da\ users coming from particular areas and network operators.
-One study was conducted in Norway with 654 installs of \da\ and one was conducted in Bangladesh where 2463 users installed \da.
-This is responsible for the sudden change in distribution of OS versions in \da\ shown towards the end of Figures \ref{fig:norm_os}, \ref{fig:nvulnerabilities_heat} and \ref{fig:proportioninsecure} as for that period Bangladeshi users contributed over half of the \da\ data.
+The \da\ data is mostly generated by devices which have the \da\ app installed because their owner happened to come across the app as there has been no advertising and few other attempts to increase usage.
+However two collaborations with network operators to install \da\ on customer devices resulted in large numbers of new \da\ users coming from particular network operators.
+One study was conducted in Norway with 654 installs and one in Bangladesh where 2463 users installed \da.
+This is responsible for the sudden change in distribution of OS versions in \da\ beginning in August 2014 Figures \ref{fig:norm_os}, \ref{fig:nvulnerabilities_heat} and \ref{fig:proportioninsecure} as for that period Bangladeshi users contributed over half of the \da\ data.
 %We could normalise out the influence of the Bangladeshi users, however they are Android users too and they also need security updates.
@@
 SEAndroid~\cite{Smalley2013} which is included in Android from version 4.1~\cite{jelly-bean-release} claimed to prevent some root vulnerabilities and to reduce the impact of others.
-Capability based enforcement systems such as Capsicum~\cite{Watson2010} substantially reduce the capabilities that an exploit has to try and gain increased privilege with.
-When Capsicum is included in Linux\footnote{\url{https://github.com/google/capsicum-linux}} and hence in Android, it could be used to place fine grained restrictions on system daemons, preventing vulnerabilities becoming critical vulnerabilities.
+Capability based enforcement systems such as Capsicum~\cite{Watson2010} substantially reduce the capabilities that an exploit has to try and gain increased privilege with and could be included in Linux\footnote{\url{https://github.com/google/capsicum-linux}} and hence Android.
 
 We have not included all the vulnerabilities from \avo. 
-In particular, we have not included any device manufacturer specific vulnerabilities, even when these are widespread (such as vulnerabilities affecting Qualcomm\footnote{Qualcomm is particularly good at publicly disclosing the vulnerabilities affecting their code and the patches which fix them.
+In particular, we have not included any device manufacturer specific vulnerabilities, even when these are widespread (such as vulnerabilities affecting Qualcomm\footnote{Qualcomm is particularly good at publicly disclosing vulnerabilities and the patches which fix them.
 \url{https://www.codeaurora.org/projects/security-advisories/}} chipsets) because it is difficult to work out which devices are affected.
@@
 For some other vulnerabilities which might affect all Android devices it is hard to work out which devices are affected.
-For example, \emph{pty race}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/pty_race}} is a Linux kernel vulnerability.
+For example, \emph{pty race}\avovulnfoot{pty_race} is a Linux kernel vulnerability.
 We found 48 commits which fixed this vulnerability on different branches.
@@
 This is despite the fact that some device manufacturers have shipped builds containing the fix.
-This also made it difficult to determine whether a specific device was vulnerable or not.
-For \emph{RageAgainstTheCage adb}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/RageAgainstTheCage_adb}} and \emph{keystore buffer}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/keystore_buffer}} vulnerabilities we do not have sufficient data to determine whether they fit one of our attack vectors as the former may require physical ADB access and the latter may be thwarted by the sandbox.
-However a sensitivity analysis showed that those two vulnerabilities make little difference to our results as they affect few devices.
+This makes it difficult to determine whether a specific device is vulnerable.
+For \emph{RageAgainstTheCage adb}\avovulnfoot{RageAgainstTheCage_adb} and \emph{keystore buffer}\avovulnfoot{keystore_buffer} vulnerabilities we do not have sufficient data to determine whether they fit one of our attack vectors as the former may require physical ADB access and the latter may be thwarted by the sandbox.
+However a sensitivity analysis showed that those two vulnerabilities make little difference to our results.
 
@@
 \label{sec:related}
-Using the methods and data described in this paper we have determined that, on average, \daMeanInsecurityPerc\ of devices were exposed to known critical vulnerabilities between 2011 and 2014.
+Using the methods and data described in this paper we have determined that, on average, \daMeanInsecurityPerc\ of devices were exposed to known critical vulnerabilities between 2011 and 2015.
 Felt et al.\ studied 6 Android handsets in 2011 and found they were exposed to root vulnerabilities at least 74\% of the time~\cite{Felt2011} .
 Our approach differs from their study because they used data from 6 handsets and assumed the best possible update distribution, while our work is based on a large sample of devices tracking the actual update distribution.
-Nevertheless, our own analysis as well as comparison with their work suggests protection against critical vulnerabilities has not improved significantly over the last 3 years. 
+Nevertheless, our own analysis as well as comparison with their work suggests protection against critical vulnerabilities has not improved significantly over the last 4 years. 
 Felt et al.\ also found that 4 of the 46 malware samples (8\%) they analysed contained root exploits, much lower than rates found in later (larger) studies which found rates of 36.7\%~\cite{Zhou2012b} and 40\%~\cite{Zhou2012a} in 2012.
 
-Currently, finding and exploiting vulnerabilities is typically a time-consuming and manual process.
+Currently, finding and exploiting vulnerabilities is typically a time-consuming manual process.
 Therefore an attacker has to invest significant human resource to first find a bug and then write code to exploit the vulnerability.
 To reduce the effort involved in finding a bug, attackers can look at security-related commits made to open source repositories.
-Previous work has shown attackers have taken this approach for the Firefox web browser~\cite{Barth2011}.
-Google does not release updates to the Android source code until after the release of the relevant security update.
-This reduces the time available for an attacker to exploit a vulnerability by looking at security updates disclosed in the source code repository from the period (\textbf{first fix}, \textbf{fix deployed}) to (\textbf{fixing release}, \textbf{fix deployed}).
-Unfortunately the time period between the \textbf{fixing release} and \textbf{fix deployed} can be of the order of months. 
+Attackers have taken this approach for the Firefox web browser~\cite{Barth2011}.
+Google does not release updates to the Android source code until after the release of the relevant update.
+This reduces the time available for an attacker to exploit a vulnerability by looking at security updates disclosed in the source code repository from the period (\textbf{first fix}, \textbf{fix deployed}) to (\textbf{fixing release}, \textbf{fix deployed}), unfortunately this can still take months .
 For example, the zergRush vulnerability was fixed in a release before it was publicly discovered but it still took 27 months for 90\% of the devices in our study to be definitely fixed\footnote{Definitely fixed is 1 $-$ (definitely insecure + maybe secure), maybe fixed is 1 $-$ definitely insecure} (25 months for 90\% of devices maybe fixed).
 
-Recent work by Brumley et al. has shown it is possible to automatically generate exploits from binary fixes, for example by using security updates issued by manufacturers~\cite{Brumley2008}. 
+Brumley et al. has shown it is possible to automatically generate exploits from binary fixes, for example by using security updates issued by manufacturers~\cite{Brumley2008}. 
 Similarly, the Woodpecker tool automatically finds permission leaks in stock Android phone images~\cite{Grace2012}.
@@
 
-Rather than fixing critical vulnerabilities, security can be obtained by detect them and preventing the installation or execution of malicious apps.
-Detection strategies include RiskRanker, which classified 3\,281 out of 118\,318 apps (2.8\%) as risky of which 718 (22\%) were malware and 322 (10\%) were previously unknown malware, an overall infection rate of 0.6\% across multiple markets~\cite{Grace2012a}.
+Rather than fixing critical vulnerabilities, security can be obtained by detecting malicious apps and preventing their installation or execution.
+Detection strategies include RiskRanker, which classified 3\,281 out of 118\,318 apps (2.8\%) as risky of which 718 (22\%) were malware and 322 (10\%) were previously unknown malware, an infection rate of 0.6\% across multiple markets~\cite{Grace2012a}.
 DroidRanger also analysed apps finding 148 out of 182\,823 apps (0.08\%) to be malicious across multiple markets of which 29 were previously unknown~\cite{Zhou2012a}.
 %DroidRanger: It used permission-based behavioural fingerprinting which looked at the permissions of known malware and heuristic-based filtering -- dynamic loading of both Dalvik and native code.
-A common technique used by attackers takes existing popular apps and repackages them to include malicious code. 
-AnDarwin uses this insight to detect similar apps, and found 169 out of 265\,359 of all apps studied (0.06\%) were malicious clones of genuine apps~\cite{Crussell2013}.
+A common technique used by attackers is to include malicious code in repackaged popular apps. 
+AnDarwin uses this insight to detect similar apps, and found 169 out of 265\,359 of all apps studied (0.06\%) were malicious clones~\cite{Crussell2013}.
 %AnDarwin: It used clustering based on semantic vectors derived from the program dependence graphs to detect similar apps.
@@
 Unfortunately, whilst these tools were successful in finding critical vulnerabilities, such techniques are not perfect.
-The Malware Genome project collected 1\,260 malware samples from 2010--2011~\cite{Zhou2012b}.
-They found that the best case for anti-malware detection on their data was 79.6\%.
-Similarly, DroidChameleon found that antivirus products could not detect malware if the execution binary was automatically permuted~\cite{Rastogi2013}.
+The Malware Genome project collected 1\,260 malware samples from 2010--2011~\cite{Zhou2012b} and found a best case anti-malware detection rate of 79.6\%.
+Similarly, DroidChameleon found that antivirus products could not detect malware if the binary was automatically permuted~\cite{Rastogi2013}.
 Therefore detecting risky apps which might contain critical vulnerabilities is a useful tool, but it does not replace the need to provide security updates.

### 2015-03-19 securityupdates.tex
Squash figures up together to save space - bad idea

@@
 
-{\begin{figure}[p]
- \centering
+{\begin{figure}[h]
+\begin{minipage}{0.49\columnwidth}
  \begin{subfigure}[b]{\columnwidth}
@@
  \caption{The top \daNumDeviceDataDevices\ devices by days of contribution in the \da\ data. One strip per device handset.}
- \includegraphics[width=0.55\columnwidth]{figures/from_to_updates.pdf}
- \caption{Updates between different Android versions in the \da\ data}
+\end{minipage}\hspace{0.02\columnwidth}
+\begin{minipage}{0.49\columnwidth}
+ \includegraphics[width=\columnwidth]{figures/from_to_updates.pdf}
+ \caption{Updates between different Android versions in \da}
  \label{fig:from_to_updates}
+ \includegraphics[width=\columnwidth]{figures/nw_security_updates.pdf}
+ \caption{Proportion of updates each week which (may have) fixed vulnerabilities}
+ \label{fig:weekly_security_updates}
+\end{minipage}
 \end{figure}}
@@
 
-\begin{figure}[h]
- \includegraphics[width=\columnwidth]{figures/nw_security_updates.pdf}
- \caption{Proportion of updates each week which (may have) fixed vulnerabilities}
- \label{fig:weekly_security_updates}
-\end{figure}
 We recorded update events and Figure~\ref{fig:from_to_updates} shows how devices upgrade between different versions of Android.
@@
 \label{sec:representative}
-\begin{figure}
- \centering
+\begin{figure}[h]
+\begin{minipage}{0.49\columnwidth}
  \begin{subfigure}[b]{\columnwidth}
@@
 \end{subfigure}
-\caption{Monthly Android API version data}
-\end{figure}
-\begin{figure}
- \centering
+\caption{Monthly Android API version}
+\end{minipage}\hspace{0.02\columnwidth}
+\begin{minipage}{0.49\columnwidth}
  \includegraphics[width=\columnwidth]{figures/api_gpcomp_rdiff}
@@
  \label{fig:da_gp_comp_diff}
+\end{minipage}
 \end{figure}

### 2015-03-19 securityupdates.tex
Put the links to AVO vulnerabilities in the PDF even if there is not space to put them in footnotes

--- 2015-03-19 securityupdates.tex
@@
 \newcommand{\avo}{AVO}
-\newcommand{\avovulnfoot}[1]{}%{\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/#1}}}
+% Evil hackery to provide an optional argument: https://tex.stackexchange.com/questions/308/different-command-definitions-with-and-without-optional-argument/314#314
+\makeatletter
+ \def\avovuln{\@ifnextchar[{\@avovulnsspecific}{\@avovulngeneral}}
+ \def\@avovulnsspecific[#1]#2{\emph{\href{http://androidvulnerabilities.org/vulnerabilities/#1}{#2}}}
+ \def\@avovulngeneral#1{\emph{\href{http://androidvulnerabilities.org/vulnerabilities/#1}{#1}}}
+\makeatother
 \newcommand{\percMarketShare}{83.6\%~\footnote{\url{http://www.theinquirer.net/inquirer/news/2379036/android-hits-836-percent-marketshare-while-ios-windows-and-blackberry-slide}}}
@@
 Figure~\ref{fig:vulnerabilities_timeline} shows the dates of \textbf{discovery} and, when later, the date of the first \textbf{fixing release}.
-Some vulnerabilities (\emph{levitator}, \emph{zergRush}) were fixed in released versions of Android before they were discovered and so are shown as vertical lines, while others were known for months before a version of Android which fixed them was shipped.
+Some vulnerabilities (\avovuln{levitator}, \avovuln{zergRush}) were fixed in released versions of Android before they were discovered and so are shown as vertical lines, while others were known for months before a version of Android which fixed them was shipped.
 During the period in which \da\ data was collected the date when a version of Android with the fix was observed on a \da\ device is taken as the date that version was released.
@@
 This demonstrates the importance of a longitudinal study: this type of analysis requires years of data.
-Once \emph{zergRush} was discovered in October 2011 the most devices are recorded as \emph{insecure} (red) as they were exposed to that vulnerability.
-The remanding devices were already running a version of Android which fixed the \emph{zergRush} vulnerability and are therefore marked as \emph{secure} (green).
-From October 2011 until the discovery of \emph{APK duplicate file} in February 2013 the graph shows progressive improvement as devices are upgraded.
+Once \avovuln{zergRush} was discovered in October 2011 the most devices are recorded as \emph{insecure} (red) as they were exposed to that vulnerability.
+The remanding devices were already running a version of Android which fixed the \avovuln{zergRush} vulnerability and are therefore marked as \emph{secure} (green).
+From October 2011 until the discovery of \avovuln[APK_duplicate_file]{APK duplicate file} in February 2013 the graph shows progressive improvement as devices are upgraded.
 This means more and more devices are marked as \emph{secure} because they are now running a secure version of Android, or marked as \emph{maybe secure} because they received a minor update which may have contained a backported fix.
@@
 The variation of the proportion of devices affected by a vulnerability with time tells us how badly a particular vulnerability affected the Android platform.
-In July 2011 the \emph{exploid}\avovulnfoot{exploid_udev} and \emph{levitator}\avovulnfoot{levitator} vulnerabilities both affect most Android devices, slowly these are fixed as updates roll out and devices are replaced until in January 2013 a much smaller proportion of devices are affected by known vulnerabilities.
+In July 2011 the \avovuln[exploid_udev]{exploid} and \avovuln{levitator} vulnerabilities both affect most Android devices, slowly these are fixed as updates roll out and devices are replaced until in January 2013 a much smaller proportion of devices are affected by known vulnerabilities.
 However when in February 2013 the first APK signing vulnerability was found which affected all previous versions of Android and even in October 2013 most devices (\daVulnAPKDuplicateFileOctoberPerc) were still vulnerable.
@@
 Figure~\ref{fig:nvulnerabilities_heat} shows how the the \emph{APK signing vulnerabilities} affected all devices and took months to get fixed for any device.
-However what is perhaps more worrying is the long tail on the \emph{Gingerbreak}\avovulnfoot{Gingerbreak}, \emph{levitator}, \emph{exploid} and \emph{zergRush}\avovulnfoot{zergRush} vulnerabilities which are more dangerous root vulnerabilities (not requiring new APK installation) and which affect a significant proportion of devices years later.
+However what is perhaps more worrying is the long tail on the \avovuln{Gingerbreak}, \avovuln{levitator}, \avovuln[exploid_udev]{exploid} and \avovuln{zergRush} vulnerabilities which are more dangerous root vulnerabilities (not requiring new APK installation) and which affect a significant proportion of devices years later.
 
@@
 For some other vulnerabilities which might affect all Android devices it is hard to work out which devices are affected.
-For example, \emph{pty race}\avovulnfoot{pty_race} is a Linux kernel vulnerability.
+For example, \avovuln[pty_race]{pty race} is a Linux kernel vulnerability.
 We found 48 commits which fixed this vulnerability on different branches.
@@
 This makes it difficult to determine whether a specific device is vulnerable.
-For \emph{RageAgainstTheCage adb}\avovulnfoot{RageAgainstTheCage_adb} and \emph{keystore buffer}\avovulnfoot{keystore_buffer} vulnerabilities we do not have sufficient data to determine whether they fit one of our attack vectors as the former may require physical ADB access and the latter may be thwarted by the sandbox.
+For \avovuln[RageAgainstTheCage_adb]{RageAgainstTheCage adb} and \avovuln[keystore_buffer]{keystore buffer} vulnerabilities we do not have sufficient data to determine whether they fit one of our attack vectors as the former may require physical ADB access and the latter may be thwarted by the sandbox.
 However a sensitivity analysis showed that those two vulnerabilities make little difference to our results.

### 2015-03-20 securityupdates.tex
Minor tweaks from Alastair

@@
 Modern smartphone operating systems protect users from malicious applications by providing a protected execution environment or sandbox.
-This approach relies on the absence of active exploits which allow apps to break out of the sandbox.
+This approach relies on the absence of exploits which allow apps to break out of the sandbox.
 Many such exploits have been found, and therefore the security of a smartphone operating system relies on the prompt delivery of security updates.
 In this paper we quantify the extent to which the security sandbox provided by Android protects devices from malicious apps.
-We analysed \da\ data from over \daOSYearsOfData\ years and \daNumOSDevices\ devices and found that on average \daMeanInsecurityPercNominal\ of devices were exposed to known root privilege vulnerabilities and only \daUpdatednessPercNominal\ of devices run the most recent version of Android.
+We analysed \da\ data from over \daOSYearsOfData\ years and \daNumOSDevices\ devices and found that on average \daMeanInsecurityPercTwosfNominal\ of devices were exposed to known root privilege vulnerabilities and only \daUpdatednessPercTwosfNominal\ of devices run the most recent version of Android.
 \end{abstract}
@@
 We only looked for critical vulnerabilities such as root vulnerabilities which did not require USB debugging to exploit.
-If an malicious code exploits a critical vulnerability then it gains control of the device.
+If malicious code exploits a critical vulnerability then it gains control of the device.
 Some phones can be `rooted' by enabling USB debugging and using the special privileges of the ADB shell to root the device but only \daAdbEnabledPerc\ of devices have USB debugging enabled.
@@
 We have chosen these vulnerabilities to fit the attack vectors introduced in \S\ref{sec:threatmodel}.
-These vulnerabilities affect all Android devices regardless of manufacturer, and as a result our selected vulnerabilities will dominate security analysis of Android.
+These vulnerabilities affect all Android devices regardless of manufacturer, and as a result our selected vulnerabilities will dominate any security analysis of Android.
 In many cases we could not match manufacturer- and model-specific vulnerabilities to individual devices in the \da\ data and therefore attempting to include device-specific vulnerabilities as well would introduce additional uncertainty in our results.
@@
 Devices, on average, apply \daUpdatesPerYear\ updates each year and \daPercUpdatesDowngrades\ of full version changes are downgrades to older versions.
-We describe three experiments, examining the vulnerability of Android devices in general (\S\ref{sec:exp:versionsecurity}, explore the upgrade cycle and vulnerability cycle of the ecosystem (\S\ref{sec:exp:android_ecosystem}), quantify the updates installed on particular devices (\S\ref{sec:exp:device_updates}).
+We describe three experiments, examining the vulnerability of Android devices in general (\S\ref{sec:exp:versionsecurity}), explore the upgrade cycle and vulnerability cycle of the ecosystem (\S\ref{sec:exp:android_ecosystem}) and quantify the updates installed on particular devices (\S\ref{sec:exp:device_updates}).
 
-\subsection{Experiment 0: vulnerability of Android devices}\label{sec:exp:versionsecurity}
+\subsection{Experiment 1: vulnerability of Android devices}\label{sec:exp:versionsecurity}
 
@@
 Once \avovuln{zergRush} was discovered in October 2011 the most devices are recorded as \emph{insecure} (red) as they were exposed to that vulnerability.
-The remanding devices were already running a version of Android which fixed the \avovuln{zergRush} vulnerability and are therefore marked as \emph{secure} (green).
+The remaining devices were already running a version of Android which fixed the \avovuln{zergRush} vulnerability and are therefore marked as \emph{secure} (green).
 From October 2011 until the discovery of \avovuln[APK_duplicate_file]{APK duplicate file} in February 2013 the graph shows progressive improvement as devices are upgraded.
@@
 
-\subsection{Experiment 1: behaviour of the Android ecosystem}\label{sec:exp:android_ecosystem}
+\subsection{Experiment 2: behaviour of the Android ecosystem}\label{sec:exp:android_ecosystem}
 
-\subsubsection{Method} As in Experiment 0, we used the version information for each device to calculate which critical vulnerabilities each device was susceptible to on a daily basis.
+\subsubsection{Method} As in Experiment 1, we used the version information for each device to calculate which critical vulnerabilities each device was susceptible to on a daily basis.
 
@@
 It shows how old versions are gradually replaced by new ones, and the long tail of devices which do not see updates to more recent versions.
+The anomaly beginning in August 2014 is explained in \S\ref{sec:da_changes}.
 
@@
  \includegraphics[width=\textwidth]{figures/da_norm_os}
- \caption{Android versions in \da\ data over time. The change in behaviour after August 2014 is explained in \S\ref{sec:da_changes}}
+ \caption{Android versions in \da\ data over time. The anomaly beginning in August 2014 is explained in \S\ref{sec:da_changes}.}
  \label{fig:norm_os}
@@
 
-\subsection{Experiment 2: Updates to particular devices}\label{sec:exp:device_updates}
-Those graphs summarise data across all the devices, however one of the advantages of the \da\ data is that we can look at what happens to individual devices over time.
+\subsection{Experiment 3: Updates to particular devices}\label{sec:exp:device_updates}
+Previous graphs summarise data across all the devices, however one of the advantages of the \da\ data is that we can look at what happens to individual devices over time.
+This allows us to determine whether newer OS versions are being used because people are buying new phones running newer OS versions or whether existing devices receive updates.
 
@@
  \caption{The top \daNumDeviceDataDevices\ devices by days of contribution in the \da\ data. One strip per device handset.}
- \includegraphics[width=0.55\columnwidth]{figures/from_to_updates.pdf}
- \caption{Updates between different Android versions in the \da\ data}
- \label{fig:from_to_updates}
+  \includegraphics[width=\columnwidth]{figures/nw_security_updates.pdf}
+ \caption{Proportion of updates each week which (may have) fixed vulnerabilities}
+ \label{fig:weekly_security_updates}
 \end{figure}}
@@
 \subsubsection{Method}
-As in Experiment 0, we used the version information for each device to calculate which critical vulnerabilities each device was susceptible to on a daily basis.
+As in Experiment 1, we used the version information for each device to calculate which critical vulnerabilities each device was susceptible to on a daily basis.
 We also recorded when the version changed and which versions it changed from and to.
@@
 To investigate the sensitivity of our results to the Bangladesh study we truncated our data before this study started and found the percentage of vulnerable devices to be 87.7\% rather than \daMeanInsecurityPercNominal, and the percentage of devices running the latest version to be 5.43\% rather than \daUpdatednessPercNominal.
-Hence removing this anomaly would have no effect on our conclusions.
+Hence removing this anomaly would have no effect on our main conclusions.
+
 Since most \da\ users are self selecting and install \da\ because they want to find out more about what their phone is doing or to aid research they may be biased and perhaps more likely to install updates.
@@
 On some branches it was also accidentally fixed, and later reintroduced, before finally being fixed properly, and therefore determining which devices were vulnerable relies on knowing where, and on which branch, the kernel code for a specific release was taken from.
+
 For some vulnerabilities the required patch is not in AOSP.
@@
 Using the methods and data described in this paper we have determined that, on average, \daMeanInsecurityPerc\ of devices were exposed to known critical vulnerabilities between 2011 and 2015.
-Felt et al.\ studied 6 Android handsets in 2011 and found they were exposed to root vulnerabilities at least 74\% of the time~\cite{Felt2011} .
+Felt et al.\ studied 6 Android handsets in 2011 and found they were exposed to root vulnerabilities at least 74\% of the time~\cite{Felt2011}.
 Our approach differs from their study because they used data from 6 handsets and assumed the best possible update distribution, while our work is based on a large sample of devices tracking the actual update distribution.
@@
 Thanks to Laurent Simon, Thomas Coudray, Adrian Taylor, Justin Case, Giant Pune and Khilan Gudka for reporting vulnerabilities in Android.
+This work was partly funded by the EPSRC and Google but the the work presented here does not necessarily reflect their views.
 }

### 2015-03-20 securityupdates.tex
Drop a couple of figures

@@
 
-\begin{figure}[h]
- \includegraphics[width=\columnwidth]{figures/nw_security_updates.pdf}
- \caption{Proportion of updates each week which (may have) fixed vulnerabilities}
- \label{fig:weekly_security_updates}
-\end{figure}
-We recorded update events and Figure~\ref{fig:from_to_updates} shows how devices upgrade between different versions of Android.
-Mostly the dark cells are upgrades (above the diagonal) (\daNumUpdatesUpgrades).
+We recorded \daNumFullVersionUpdates\ update events and found most are upgrades (\daNumUpdatesUpgrades).
 While many upgrades are from one version to the next version there are also a fair number (\daNumUpdatesBigUpgrades, \daPercBigUpgradesNominal) which skip more than \daNumUpdatesSkippedBig\ versions.
@@
 \end{figure}
-\begin{figure}
- \centering
- \includegraphics[width=\columnwidth]{figures/api_gpcomp_rdiff}
- \caption{Difference between \da\ and Google Play data on the proportion of devices running different Android API versions}
- \label{fig:da_gp_comp_diff}
-\end{figure}
+
 The data from \da\ we used to investigate the proportion of devices exposed to different vulnerabilities is the OS version.
@@
 Figure~\ref{fig:play_api} shows the data from Google and Figure~\ref{fig:da_api} shows the data from \da\ and they appear similar.
-Figure~\ref{fig:da_gp_comp_diff} shows the difference between Figures~\ref{fig:play_api} and \ref{fig:da_api}, normalising for days since the API version was released.
-It shows that the \da\ data systematically overestimates the prevalence of new API versions and underestimates the prevalence of old API versions.
+We analysed the difference between the API version data from \da\ and Gooble Play, normalising for days since the API version was released.
+This that the \da\ data systematically overestimates the prevalence of new API versions and underestimates the prevalence of old API versions (except for API version 17 which was particularly popular in a focused study discussed in the next section and so was temporarily overestimated when an old version).
 This means that the OS version information from \da\ is likely to be overestimating the prevalence of new OS versions and hence our results are a conservative estimate of the security of Android.

### 2015-03-20 securityupdates.tex
Force figures onto a page with text to improve space efficiency
Truncate 'Lifetime of a vulnerability' section to 1/3

@@
 
-The key events in the lifetime of a vulnerability do not always occur in the same order and are:
-\begin{description}
-\item[creation] When a vulnerability was created in the source code.
-\item[introducing release] When the first release was made containing the vulnerability.
-\item[discovery] When the vulnerability is first discovered.
-\item[exploit] When the vulnerability is first exploited.
-\item[disclosure] When the vulnerability is first disclosed. % publically or to a smaller set of people?
-\item[fix] When a vulnerability was first fixed in the source code.
-\item[fixing release] When the first release containing the fix was made (equivalent to `patch available'~\cite{Frei2010}).
-\item[fix deployed] When the fix has been deployed to a sufficiently large proportion of the population that the vulnerability can be ignored (the ecosystem equivalent to the per instance `patch installed'~\cite{Frei2010}).
-\end{description}
-
 Establishing when a vulnerability starts to pose a threat to users is difficult.
 Frei et al.~\cite{Frei2010} propose the definition of the \textbf{time of disclosure}.
-This occurs when the information about the vulnerability is freely available to the public from an accepted and independent source and has been validated by security experts so that it has a risk rating.
-Unfortunately before we collated this information much of it was not published by an independent source and lacked risk rating information, even months or years after they had been actively used.
-Therefore this measure does not work.
+This occurs when the information about the vulnerability is available to the from an accepted source.
+Unfortunately before we collated this information much of it was not published by an independent source, even months or years after they had been actively used.
 
-Symantec's 2012 analysis of desktop malware has shown that after public disclosure, exploitation rates increase by 5 orders of magnitude~\cite{Bilge2012} and so from the point of view of widespread danger, the period between public \textbf{disclosure} and the date of \textbf{fix deployed} is the most critical.
+Symantec's 2012 analysis of desktop malware has shown that after public disclosure, exploitation rates increase by 5 orders of magnitude~\cite{Bilge2012} and so from the point of view of widespread danger, the period between public disclosure and the date of the fix being deployed to most devices is the most critical.
 However they also show that zero-day vulnerabilities are typically used for 312 days before they are publicly disclosed, often to target particular organisations.
-Hence when considering a vulnerability from the point of view of an organisation which cares about Advanced Persistent Threats, the critical period starts with \textbf{discovery} and continues until \textbf{fix deployed}.
-Therefore, in our vulnerability calculations we prefer to use the earliest recorded date of \textbf{discovery}, even if that knowledge might have been confined to a particular device manufacturer or hobbyist.
-%We cannot know if someone reported the vulnerability to the device manufacturer and we do know that various agencies are engaged in widespread monitoring of communications and compromise of civilian infrastructure~\cite{TODO} and so they may know.
 
-Unfortunately \textbf{discovery} is the hardest point to obtain concrete data on as discovery may happen multiple times independently and not all discoverers will report their discovery.
-\textbf{Creation} and \textbf{introducing release} are the earliest points which could be used, but the risk is mostly latent until someone discovers them and therefore we do not use these dates.
-%, for example entities with high value targets may keep databases of unknown vulnerabilities for later use, in which case there is a high risk to potential targets from the never disclosed point of discovery.
-%However if the discoverer never discloses the vulnerability to anyone or makes any use of it then there is little danger.
-The date of first \textbf{exploit} is a point when the risk is definitely high, but again a good adversary will not be detected when using such exploits.
-The date of first \textbf{fix} is, assuming that the fix is deliberate, a point at which the vulnerability is known at least within the organisation performing the fix and frequently implies an earlier discovery and notification by a third party; we determine this date from the Android Open Source Project repository by examining the authored-on date of the fixing commit.
-%Once a {\bf fixing release} has been made then the vulnerability is widely known because it can be reverse engineered from the changes in the release~\cite{Brumley2008}.
-%When the {\bf fix deployed} to a sufficiently large proportion of devices then the remaining risk is minimal.
-Therefore, in this paper we use the earliest known date from \textbf{discovery}, \textbf{exploit}, \textbf{disclosure} or \textbf{fix}.
+Therefore, we use the earliest of: discovery date, date when reported to be exploited, date of disclosure or date of fix in the source code; as the date at which we consider a vulnerability to start being dangerous.
 A breakdown of the type of date used is shown in Table~\ref{tab:andvulns}.
@@
 \label{sec:representative}
-\begin{figure}
+\begin{figure}[h]
  \centering

### 2015-03-20 securityupdates.tex
Rename Experiment to Analysis

@@
 
-\section{Experiments}
+\section{Analysis}
 \label{sec:results}
@@
 Devices, on average, apply \daUpdatesPerYear\ updates each year and \daPercUpdatesDowngrades\ of full version changes are downgrades to older versions.
-We describe three experiments, examining the vulnerability of Android devices in general (\S\ref{sec:exp:versionsecurity}), explore the upgrade cycle and vulnerability cycle of the ecosystem (\S\ref{sec:exp:android_ecosystem}) and quantify the updates installed on particular devices (\S\ref{sec:exp:device_updates}).
+We describe three analyses, examining the vulnerability of Android devices in general (\S\ref{sec:exp:versionsecurity}), explore the upgrade cycle and vulnerability cycle of the ecosystem (\S\ref{sec:exp:android_ecosystem}) and quantify the updates installed on particular devices (\S\ref{sec:exp:device_updates}).
 
-\subsection{Experiment 1: vulnerability of Android devices}\label{sec:exp:versionsecurity}
+\subsection{Analysis 1: vulnerability of Android devices}\label{sec:exp:versionsecurity}
 
@@
 
-\subsection{Experiment 2: behaviour of the Android ecosystem}\label{sec:exp:android_ecosystem}
+\subsection{Analysis 2: behaviour of the Android ecosystem}\label{sec:exp:android_ecosystem}
 
-\subsubsection{Method} As in Experiment 1, we used the version information for each device to calculate which critical vulnerabilities each device was susceptible to on a daily basis.
+\subsubsection{Method} As in Analysis 1, we used the version information for each device to calculate which critical vulnerabilities each device was susceptible to on a daily basis.
 
@@
 
-\subsection{Experiment 3: Updates to particular devices}\label{sec:exp:device_updates}
+\subsection{Analysis 3: Updates to particular devices}\label{sec:exp:device_updates}
 Previous graphs summarise data across all the devices, however one of the advantages of the \da\ data is that we can look at what happens to individual devices over time.
@@
 \subsubsection{Method}
-As in Experiment 1, we used the version information for each device to calculate which critical vulnerabilities each device was susceptible to on a daily basis.
+As in Analysis 1, we used the version information for each device to calculate which critical vulnerabilities each device was susceptible to on a daily basis.
 We also recorded when the version changed and which versions it changed from and to.

### 2015-03-20 securityupdates.tex
Note a paragraph we could chop if we need to
Move some figures around for security updates weekly
Fix related work to take account of the shortening of the lifetime of vulnerabilities section

@@
 (iii) device manufacturer specific vulnerabilities (which are not in our analysis here) were easier to find and so attackers looking for vulnerabilities adjusted their focus.
+\todolater{If we need to save space then delete this paragraph}
 
@@
 
-\begin{figure*}
+\begin{figure}[h]
  \centering
@@
  \label{fig:norm_os}
-\end{figure*}
+ \includegraphics[width=\textwidth]{figures/nvulnerabilities_heat.pdf}
+ \caption{Proportion of devices affected by different vulnerabilities.
+The prefix `(maybe)' indicates `maybe secure' and `(definite)' indicates `definitely insecure'.
+The first few vulnerabilities are all `maybe secure' as we do not have data from before that vulnerability was discovered while later ones are mostly `definitely insecure' as we know no fixing update reached the devices.
+ %The change in behaviour after August 2014 is explained in \S\ref{sec:da_changes}
+ }
+ \label{fig:nvulnerabilities_heat}
+\end{figure}
 
@@
 However what is perhaps more worrying is the long tail on the \avovuln{Gingerbreak}, \avovuln{levitator}, \avovuln[exploid_udev]{exploid} and \avovuln{zergRush} vulnerabilities which are more dangerous root vulnerabilities (not requiring new APK installation) and which affect a significant proportion of devices years later.
-
-\begin{figure*}
- \includegraphics[width=\textwidth]{figures/nvulnerabilities_heat.pdf}
- \caption{Proportion of devices affected by different vulnerabilities.
-The prefix `(maybe)' indicates `maybe secure' and `(definite)' indicates `definitely insecure'.
-The first few vulnerabilities are all `maybe secure' as we do not have data from before that vulnerability was discovered while later ones are mostly `definitely insecure' as we know no fixing update reached the devices.
- %The change in behaviour after August 2014 is explained in \S\ref{sec:da_changes}
- }
- \label{fig:nvulnerabilities_heat}
-\end{figure*}
 
@@
 Google does not release updates to the Android source code until after the release of the relevant update.
-This reduces the time available for an attacker to exploit a vulnerability by looking at security updates disclosed in the source code repository from the period (\textbf{first fix}, \textbf{fix deployed}) to (\textbf{fixing release}, \textbf{fix deployed}), unfortunately this can still take months .
+This reduces the time available for an attacker to exploit a vulnerability by looking at security updates disclosed in the source code repository.
+Reducing it from, the interval between the commit of the first fix and the fix being deployed to most devices, to the interval between the first fixing release and the fix being deployed to most devices; unfortunately this can still take months or years.
 For example, the zergRush vulnerability was fixed in a release before it was publicly discovered but it still took 27 months for 90\% of the devices in our study to be definitely fixed\footnote{Definitely fixed is 1 $-$ (definitely insecure + maybe secure), maybe fixed is 1 $-$ definitely insecure} (25 months for 90\% of devices maybe fixed).
@@
 Similarly, the Woodpecker tool automatically finds permission leaks in stock Android phone images~\cite{Grace2012}.
-Once such automated techniques mature, it will become even more important to provide security updates promptly to every Android device in the ecosystem because an attacker can turn the contents of a \textbf{fixing release} into an exploit in a matter of minutes or hours.
+Once such automated techniques mature, it will become even more important to provide security updates promptly to every Android device in the ecosystem because an attacker can turn the contents of a released fix into an exploit in a matter of minutes or hours.
 

### 2015-03-23 avostats.tex
Updated avostats with better display of uncertainties
Update stats to include apdated avoTabAndVulns

@@
 \newcommand{\avoVulnsPerYearAllAndroidNominal}{3.71}
-\newcommand{\avoVulnsPerYearAllAndroid}{$3.71 \pm 0.957$}
-\newcommand{\avoTabAndVulns}{\begin{table} \centering \small \begin{tabular}{l|l|c} Vulnerability & How known & Date \\ \hline KillingInTheNameOf & Fixed on & 2010-07-13 \\ exploid udev & Discovered on & 2010-07-15 \\ levitator & Discovered on & 2011-03-10 \\ Gingerbreak & Fixed on & 2011-04-18 \\ zergRush & Discovered on & 2011-10-06 \\ APK duplicate file & Discovered on & 2013-02-18 \\ APK unchecked name & Discovered on & 2013-06-30 \\ APK unsigned shorts & Fixed on & 2013-07-03 \\ vold asec & Fixed on & 2014-01-27 \\ Fake ID & Fixed on & 2014-04-17 \\ TowelRoot & Discovered on & 2014-05-03 \\\end{tabular} \caption{Critical vulnerabilities in Android} \label{tab:andvulns} \end{table}}
+\newcommand{\avoVulnsPerYearAllAndroid}{$3.71 \pm 0.95$}
+\newcommand{\avoTabAndVulns}{\begin{table} \centering \small \begin{tabular}{l|l|c|c} Vulnerability & How known & Date & Categories\\ \hline KillingInTheNameOf & Fixed on & 2010-07-13 & system, kernel\\ exploid udev & Discovered on & 2010-07-15 & kernel\\ levitator & Discovered on & 2011-03-10 & kernel\\ Gingerbreak & Fixed on & 2011-04-18 & system\\ zergRush & Discovered on & 2011-10-06 & system\\ APK duplicate file & Discovered on & 2013-02-18 & signature\\ APK unchecked name & Discovered on & 2013-06-30 & signature\\ APK unsigned shorts & Fixed on & 2013-07-03 & signature\\ vold asec & Fixed on & 2014-01-27 & system\\ Fake ID & Fixed on & 2014-04-17 & signature\\ TowelRoot & Discovered on & 2014-05-03 & kernel\\\end{tabular} \caption{Critical vulnerabilities in Android} \label{tab:andvulns} \end{table}}
 \newcommand{\avoNumBigExternalProjects}{40}

### 2015-03-23 dastats.tex
Strip out security updates and android ecosystem stuff from esorics paper and put in a ccs paper and make the esorics paper meet the page limit
Squash figures up together to save space - bad idea
Update stats to include apdated avoTabAndVulns

@@
 \newcommand{\daSigNumDays}{100}
-\newcommand{\daNumManufacturers}{292}
+\newcommand{\daNumManufacturers}{293}
 \newcommand{\daNumSigManufacturers}{8}
-\newcommand{\daNumModels}{2\,620}
+\newcommand{\daNumModels}{2\,640}
 \newcommand{\daNumSigModels}{17}
@@
 \newcommand{\daMeanInsecurityPerc}{$88.0 \pm 0.0\%$}
-\newcommand{\daMeanOutstandingVulnerabilities}{0.582}
-\newcommand{\daUpdatedness}{$0.0521 \pm 0.0002$}
+\newcommand{\daMeanOutstandingVulnerabilities}{0.578}
+\newcommand{\daUpdatedness}{$0.0522 \pm 0.0002$}
 \newcommand{\daSecurityScore}{$2.79 \pm 0.00$}
@@
 \newcommand{\daNumFullVersionUpdates}{4\,400}
-\newcommand{\daNumOSVersionUpdates}{3\,720}
-\newcommand{\daNumFullOnlyVersionUpdates}{663}
+\newcommand{\daNumOSVersionUpdates}{3\,730}
+\newcommand{\daNumFullOnlyVersionUpdates}{664}
 \newcommand{\daNumSecurityUpdates}{1\,160}
 \newcommand{\daNumPossibleSecurityUpdates}{2\,240}
-\newcommand{\daTabSecScoresmanufacturer}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{Security score} \\ &&&& (out of 10) \\ \hline LG & $0.18 \pm 0.00$ & $0.32 \pm 0.00$ & 0.67 & $3.70 \pm 0.01$ \\  Motorola & $0.16 \pm 0.00$ & $0.13 \pm 0.00$ & 0.78 & $2.93 \pm 0.01$ \\  Samsung & $0.13 \pm 0.00$ & $0.04 \pm 0.00$ & 0.67 & $2.67 \pm 0.00$ \\  Sony & $0.12 \pm 0.00$ & $0.19 \pm 0.00$ & 1.11 & $2.55 \pm 0.01$ \\  HTC & $0.13 \pm 0.00$ & $0.08 \pm 0.00$ & 0.94 & $2.48 \pm 0.00$ \\  asus & $0.18 \pm 0.00$ & $0.52 \pm 0.01$ & 6.01 & $2.30 \pm 0.02$ \\  other & $0.06 \pm 0.00$ & $0.04 \pm 0.00$ & 1.11 & $1.87 \pm 0.00$ \\  Symphony & $0.00 \pm 0.00$ & $0.07 \pm 0.00$ & 4.98 & $0.25 \pm 0.00$ \\  walton & $0.00 \pm 0.00$ & $0.07 \pm 0.00$ & 5.97 & $0.21 \pm 0.01$ \\ \end{tabular} \caption{Security scores for manufacturers} \label{tab:sec_manufacturer} \end{table}}
-\newcommand{\daTabSecScoresmodel}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{Security score} \\ &&&& (out of 10) \\ \hline Galaxy Nexus & $0.51 \pm 0.00$ & $0.53 \pm 0.01$ & 1.53 & $4.71 \pm 0.03$ \\  Nexus 4 & $0.29 \pm 0.00$ & $0.82 \pm 0.01$ & 6.06 & $3.63 \pm 0.04$ \\  Nexus 7 & $0.24 \pm 0.00$ & $0.74 \pm 0.01$ & 5.92 & $3.19 \pm 0.04$ \\  other & $0.09 \pm 0.00$ & $0.14 \pm 0.00$ & 0.58 & $2.92 \pm 0.00$ \\  Desire HD & $0.08 \pm 0.00$ & $0.05 \pm 0.00$ & 0.39 & $2.91 \pm 0.02$ \\  HTC Sensation & $0.35 \pm 0.00$ & $0.01 \pm 0.01$ & 1.59 & $2.43 \pm 0.02$ \\  GT-I9100 & $0.22 \pm 0.00$ & $0.02 \pm 0.00$ & 1.23 & $2.28 \pm 0.01$ \\  HTC Desire S & $0.02 \pm 0.00$ & $0.02 \pm 0.00$ & 1.00 & $1.75 \pm 0.01$ \\  GT-N7000 & $0.24 \pm 0.00$ & $0.00 \pm 0.00$ & 2.53 & $1.42 \pm 0.02$ \\  GT-P1000 & $0.01 \pm 0.00$ & $0.00 \pm 0.01$ & 1.73 & $0.94 \pm 0.02$ \\  GT-I9300 & $0.14 \pm 0.00$ & $0.01 \pm 0.00$ & 6.28 & $0.59 \pm 0.01$ \\  GT-I9505 & $0.00 \pm 0.00$ & $0.14 \pm 0.00$ & 6.86 & $0.44 \pm 0.01$ \\  HTC Desire HD & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & 3.05 & $0.27 \pm 0.02$ \\  GT-N7100 & $0.06 \pm 0.00$ & $0.00 \pm 0.01$ & 6.90 & $0.26 \pm 0.02$ \\  Symphony W68 & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & 11.00 & $0.00 \pm 0.03$ \\ \end{tabular} \caption{Security scores for models} \label{tab:sec_model} \end{table}}
-\newcommand{\daTabSecScoressummary}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{Security score} \\ &&&& (out of 10) \\ \hline nexus & $0.40 \pm 0.00$ & $0.49 \pm 0.00$ & 0.62 & $5.14 \pm 0.01$ \\  notnexus & $0.10 \pm 0.00$ & $0.02 \pm 0.00$ & 0.58 & $2.61 \pm 0.00$ \\ \end{tabular} \caption{Security scores for nexus} \label{tab:sec_summary} \end{table}}
-\newcommand{\daUpdatednessPerc}{$5.21 \pm 0.02\%$}
-\newcommand{\daNumOSDevices}{20\,000}
+\newcommand{\daTabSecScoresmanufacturer}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{Security score} \\ &&&& (out of 10) \\ \hline LG & $0.18 \pm 0.00$ & $0.32 \pm 0.00$ & 0.67 & $3.71 \pm 0.01$ \\  Motorola & $0.16 \pm 0.00$ & $0.13 \pm 0.00$ & 0.78 & $2.93 \pm 0.01$ \\  Samsung & $0.13 \pm 0.00$ & $0.04 \pm 0.00$ & 0.66 & $2.68 \pm 0.00$ \\  Sony & $0.12 \pm 0.00$ & $0.19 \pm 0.00$ & 1.11 & $2.55 \pm 0.01$ \\  HTC & $0.13 \pm 0.00$ & $0.08 \pm 0.00$ & 0.94 & $2.48 \pm 0.00$ \\  asus & $0.18 \pm 0.00$ & $0.52 \pm 0.01$ & 6.01 & $2.30 \pm 0.02$ \\  other & $0.06 \pm 0.00$ & $0.04 \pm 0.00$ & 1.11 & $1.87 \pm 0.00$ \\  Symphony & $0.00 \pm 0.00$ & $0.07 \pm 0.00$ & 5.00 & $0.25 \pm 0.00$ \\  walton & $0.00 \pm 0.00$ & $0.07 \pm 0.00$ & 6.00 & $0.24 \pm 0.01$ \\ \end{tabular} \caption{Security scores for manufacturers} \label{tab:sec_manufacturer} \end{table}}
+\newcommand{\daTabSecScoresmodel}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{Security score} \\ &&&& (out of 10) \\ \hline Galaxy Nexus & $0.51 \pm 0.00$ & $0.53 \pm 0.01$ & 1.53 & $4.71 \pm 0.03$ \\  Nexus 4 & $0.29 \pm 0.00$ & $0.82 \pm 0.01$ & 6.06 & $3.64 \pm 0.04$ \\  Nexus 7 & $0.24 \pm 0.00$ & $0.74 \pm 0.01$ & 5.92 & $3.19 \pm 0.04$ \\  other & $0.09 \pm 0.00$ & $0.14 \pm 0.00$ & 0.58 & $2.93 \pm 0.00$ \\  Desire HD & $0.08 \pm 0.00$ & $0.05 \pm 0.00$ & 0.39 & $2.91 \pm 0.02$ \\  HTC Sensation & $0.35 \pm 0.00$ & $0.01 \pm 0.01$ & 1.59 & $2.43 \pm 0.02$ \\  GT-I9100 & $0.22 \pm 0.00$ & $0.02 \pm 0.00$ & 1.23 & $2.28 \pm 0.01$ \\  HTC Desire S & $0.02 \pm 0.00$ & $0.02 \pm 0.00$ & 1.00 & $1.75 \pm 0.01$ \\  GT-N7000 & $0.24 \pm 0.00$ & $0.00 \pm 0.00$ & 2.53 & $1.42 \pm 0.02$ \\  GT-P1000 & $0.01 \pm 0.00$ & $0.00 \pm 0.01$ & 1.73 & $0.94 \pm 0.02$ \\  GT-I9300 & $0.14 \pm 0.00$ & $0.01 \pm 0.00$ & 6.28 & $0.59 \pm 0.01$ \\  GT-I9505 & $0.00 \pm 0.00$ & $0.14 \pm 0.00$ & 6.87 & $0.44 \pm 0.01$ \\  HTC Desire HD & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & 3.05 & $0.27 \pm 0.02$ \\  GT-N7100 & $0.06 \pm 0.00$ & $0.00 \pm 0.01$ & 6.90 & $0.26 \pm 0.02$ \\  Symphony W68 & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & 11.00 & $0.00 \pm 0.03$ \\ \end{tabular} \caption{Security scores for models} \label{tab:sec_model} \end{table}}
+\newcommand{\daTabSecScoressummary}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{Security score} \\ &&&& (out of 10) \\ \hline nexus & $0.40 \pm 0.00$ & $0.49 \pm 0.00$ & 0.62 & $5.14 \pm 0.01$ \\  notnexus & $0.10 \pm 0.00$ & $0.02 \pm 0.00$ & 0.58 & $2.62 \pm 0.00$ \\ \end{tabular} \caption{Security scores for nexus} \label{tab:sec_summary} \end{table}}
+\newcommand{\daUpdatednessPerc}{$5.22 \pm 0.02\%$}
+\newcommand{\daNumOSDevices}{20\,100}
 \newcommand{\daSigNumDevicesDay}{20}
@@
 \newcommand{\daSecScoreWorstsummary}{notnexus}
-\newcommand{\daSecScoreBestmanufacturerScore}{$3.7 \pm 0.0$}
-\newcommand{\daSecScoreWorstmanufacturerScore}{$0.214 \pm 0.006$}
+\newcommand{\daSecScoreBestmanufacturerScore}{$3.71 \pm 0.01$}
+\newcommand{\daSecScoreWorstmanufacturerScore}{$0.24 \pm 0.00$}
 \newcommand{\daSecScoreBestmodelScore}{$4.71 \pm 0.02$}
-\newcommand{\daSecScoreWorstmodelScore}{$0.0001 \pm 0.0281$}
+\newcommand{\daSecScoreWorstmodelScore}{$0.0001 \pm 0.028$}
 \newcommand{\daSecScoreBestsummaryScore}{$4.1 \pm 0.0$}
@@
 \newcommand{\daNumUpdatesDowngrades}{151}
-\newcommand{\daPercUpdatesDowngrades}{$3.52 \pm 0.29\%$}
+\newcommand{\daPercUpdatesDowngrades}{$3.51 \pm 0.29\%$}
 \newcommand{\daStartDate}{2011-07-01}
-\newcommand{\daEndDate}{2015-03-04}
+\newcommand{\daEndDate}{2015-03-13}
 \newcommand{\daNumOperators}{1\,420}
-\newcommand{\daTabSecScoresoperator}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{Security score} \\ &&&& (out of 10) \\ \hline O2 uk & $0.25 \pm 0.00$ & $0.12 \pm 0.00$ & 0.37 & $3.82 \pm 0.01$ \\  T-Mobile & $0.20 \pm 0.00$ & $0.18 \pm 0.00$ & 0.48 & $3.62 \pm 0.01$ \\  Orange & $0.21 \pm 0.00$ & $0.09 \pm 0.00$ & 0.39 & $3.52 \pm 0.02$ \\  Sprint & $0.18 \pm 0.00$ & $0.11 \pm 0.00$ & 0.44 & $3.39 \pm 0.01$ \\  3 & $0.16 \pm 0.00$ & $0.09 \pm 0.00$ & 0.49 & $3.20 \pm 0.01$ \\  Vodafone uk & $0.14 \pm 0.00$ & $0.13 \pm 0.00$ & 0.50 & $3.20 \pm 0.01$ \\  AT\&T & $0.12 \pm 0.00$ & $0.08 \pm 0.00$ & 0.43 & $3.07 \pm 0.01$ \\  Verizon & $0.18 \pm 0.00$ & $0.09 \pm 0.00$ & 0.82 & $2.82 \pm 0.01$ \\  unknown & $0.10 \pm 0.00$ & $0.21 \pm 0.00$ & 0.95 & $2.72 \pm 0.01$ \\  n Telenor & $0.03 \pm 0.00$ & $0.11 \pm 0.00$ & 1.43 & $1.59 \pm 0.01$ \\  Airtel & $0.05 \pm 0.00$ & $0.03 \pm 0.00$ & 1.87 & $1.08 \pm 0.01$ \\  Grameenphone & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & 2.02 & $0.85 \pm 0.00$ \\  Robi & $0.00 \pm 0.00$ & $0.06 \pm 0.00$ & 2.09 & $0.84 \pm 0.01$ \\  banglalink & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & 2.71 & $0.48 \pm 0.00$ \\ \end{tabular} \caption{Security scores for operators} \label{tab:sec_operator} \end{table}}
+\newcommand{\daTabSecScoresoperator}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{Security score} \\ &&&& (out of 10) \\ \hline O2 uk & $0.25 \pm 0.00$ & $0.12 \pm 0.00$ & 0.37 & $3.82 \pm 0.01$ \\  T-Mobile & $0.20 \pm 0.00$ & $0.18 \pm 0.00$ & 0.48 & $3.62 \pm 0.01$ \\  Orange & $0.21 \pm 0.00$ & $0.09 \pm 0.00$ & 0.39 & $3.52 \pm 0.02$ \\  Sprint & $0.18 \pm 0.00$ & $0.11 \pm 0.00$ & 0.44 & $3.39 \pm 0.01$ \\  3 & $0.16 \pm 0.00$ & $0.09 \pm 0.00$ & 0.49 & $3.21 \pm 0.01$ \\  Vodafone uk & $0.14 \pm 0.00$ & $0.13 \pm 0.00$ & 0.50 & $3.20 \pm 0.01$ \\  AT\&T & $0.12 \pm 0.00$ & $0.08 \pm 0.00$ & 0.43 & $3.07 \pm 0.01$ \\  Verizon & $0.18 \pm 0.00$ & $0.09 \pm 0.00$ & 0.82 & $2.82 \pm 0.01$ \\  unknown & $0.10 \pm 0.00$ & $0.21 \pm 0.00$ & 0.95 & $2.72 \pm 0.01$ \\  n Telenor & $0.03 \pm 0.00$ & $0.11 \pm 0.00$ & 1.43 & $1.59 \pm 0.01$ \\  Airtel & $0.05 \pm 0.00$ & $0.03 \pm 0.00$ & 1.86 & $1.09 \pm 0.01$ \\  Robi & $0.00 \pm 0.00$ & $0.07 \pm 0.00$ & 2.10 & $0.86 \pm 0.01$ \\  Grameenphone & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & 2.04 & $0.84 \pm 0.00$ \\  banglalink & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & 2.72 & $0.48 \pm 0.00$ \\ \end{tabular} \caption{Security scores for operators} \label{tab:sec_operator} \end{table}}
 \newcommand{\daSecScoreBestoperator}{O2 uk}
@@
 \newcommand{\daSecScoreWorstoperator}{banglalink}
-\newcommand{\daSecScoreWorstoperatorScore}{$0.485 \pm 0.004$}
+\newcommand{\daSecScoreWorstoperatorScore}{$0.484 \pm 0.004$}
 \newcommand{\daNumSigOperators}{14}
@@
 \newcommand{\daNumSigFullVersions}{87}
-\newcommand{\daSecScoreBestmanufacturerNumFullVersions}{122}
+\newcommand{\daSecScoreBestmanufacturerNumFullVersions}{123}
 \newcommand{\daSecScoreworstmanufacturerNumFullVersions}{239}
@@
 \newcommand{\daSecScoreworstsummaryNumFullVersions}{1057}
-\newcommand{\daSecScoreBestoperatorNumFullVersions}{92}
+\newcommand{\daSecScoreBestoperatorNumFullVersions}{93}
 \newcommand{\daSecScoreworstoperatorNumFullVersions}{89}
@@
 \newcommand{\daFullDeployedAt}{95\%}
-\newcommand{\daOSCurveFitParamFirst}{87.5}
-\newcommand{\daOSCurveFitParamSecond}{\num{0.00301}}
+\newcommand{\daOSCurveFitParamFirst}{87.3}
+\newcommand{\daOSCurveFitParamSecond}{\num{0.00299}}
 \newcommand{\daOSCurveFitRMSE}{0.116}
 \newcommand{\daAPICurveFitParamFirst}{105}
-\newcommand{\daAPICurveFitParamSecond}{\num{0.00299}}
-\newcommand{\daAPICurveFitRMSE}{0.115}
+\newcommand{\daAPICurveFitParamSecond}{\num{0.00298}}
+\newcommand{\daAPICurveFitRMSE}{0.114}
 \newcommand{\daOSCurvePolyRMSE}{0.116}
 \newcommand{\daOSCurveSplineRMSE}{0.116}
-\newcommand{\daOSCurveHalfDeployed}{$318 \pm 87$}
-\newcommand{\daOSCurveFullDeployed}{$1\,080 \pm 247\,000$}
+\newcommand{\daOSCurveHalfDeployed}{$319 \pm 88$}
+\newcommand{\daOSCurveFullDeployed}{$1\,090 \pm 248\,000$}
 \newcommand{\daAPICurvePolyRMSE}{0.115}
 \newcommand{\daAPICurveSplineRMSE}{0.115}
-\newcommand{\daAPICurveHalfDeployed}{$337 \pm 87$}
-\newcommand{\daAPICurveFullDeployed}{$1\,110 \pm 248\,000$}
+\newcommand{\daAPICurveHalfDeployed}{$338 \pm 87$}
+\newcommand{\daAPICurveFullDeployed}{$1\,110 \pm 249\,000$}
 \newcommand{\daVulnAPKDuplicateFileOctoberPerc}{91.5\%}
@@
 \newcommand{\daNumUpdatesSkippedBig}{3}
-\newcommand{\daPercBigUpgrades}{$18.7 \pm 0.7\%$}
+\newcommand{\daPercBigUpgrades}{$18.6 \pm 0.7\%$}
 \newcommand{\daUpdatesPerYear}{$1.29 \pm 0.01$}
-\newcommand{\daOSMonthsOfData}{44}
+\newcommand{\daOSMonthsOfData}{45}
 \newcommand{\daMeanInsecurityPercNominal}{88.0\%}
 \newcommand{\daVulnFreeNominal}{0.12}
-\newcommand{\daUpdatednessNominal}{0.0521}
-\newcommand{\daUpdatednessPercNominal}{5.21\%}
+\newcommand{\daUpdatednessNominal}{0.0522}
+\newcommand{\daUpdatednessPercNominal}{5.22\%}
 \newcommand{\daSecurityScoreNominal}{2.79}
 \newcommand{\daUpdatesPerYearNominal}{1.29}
-\newcommand{\daOSCurveHalfDeployedNominal}{318}
-\newcommand{\daOSCurveFullDeployedNominal}{1\,080}
-\newcommand{\daAPICurveHalfDeployedNominal}{337}
+\newcommand{\daOSCurveHalfDeployedNominal}{319}
+\newcommand{\daOSCurveFullDeployedNominal}{1\,090}
+\newcommand{\daAPICurveHalfDeployedNominal}{338}
 \newcommand{\daAPICurveFullDeployedNominal}{1\,110}
 \newcommand{\daSecScoreBestoperatorScoreNominal}{3.82}
-\newcommand{\daSecScoreWorstoperatorScoreNominal}{0.485}
+\newcommand{\daSecScoreWorstoperatorScoreNominal}{0.484}
 \newcommand{\daSecScoreBestmodelScoreNominal}{4.71}
@@
 \newcommand{\daAdbEnabledPercNominal}{19.3\%}
-\newcommand{\daSecScoreBestmanufacturerScoreNominal}{3.7}
-\newcommand{\daSecScoreWorstmanufacturerScoreNominal}{0.214}
-\newcommand{\daUpdatesPerMonthPerVersion}{$3.04 \pm 0.05$}
-\newcommand{\daNumUpdateFullOnly}{660}
+\newcommand{\daSecScoreBestmanufacturerScoreNominal}{3.71}
+\newcommand{\daSecScoreWorstmanufacturerScoreNominal}{0.24}
+\newcommand{\daUpdatesPerMonthPerVersion}{$2.98 \pm 0.05$}
+\newcommand{\daNumUpdateFullOnly}{661}
 \newcommand{\daVulnZergRushMonthsDefFixDeployed}{27.3}
-\newcommand{\daUpdatesPerMonthPerVersionNominal}{3.04}
-\newcommand{\daPercBigUpgradesNominal}{18.7\%}
-\newcommand{\daPercUpdatesDowngradesNominal}{3.52\%}
+\newcommand{\daUpdatesPerMonthPerVersionNominal}{2.98}
+\newcommand{\daPercBigUpgradesNominal}{18.6\%}
+\newcommand{\daPercUpdatesDowngradesNominal}{3.51\%}
 \newcommand{\daGPAPISeventeenLaterDate}{2015-03-02}
@@
 \newcommand{\daModelFullDeploymentYears}{0.866}
-\newcommand{\daOSCurveHalfDeployedYearsNominal}{0.871}
-\newcommand{\daOSCurveHalfDeployedYears}{$0.871 \pm 0.24$}
-\newcommand{\daOSCurveFullDeployedYearsNominal}{2.97}
-\newcommand{\daOSCurveFullDeployedYears}{$2.97 \pm 67$}
-\newcommand{\daAPICurveHalfDeployedYearsNominal}{0.923}
-\newcommand{\daAPICurveHalfDeployedYears}{$0.923 \pm 0.238$}
-\newcommand{\daAPICurveFullDeployedYearsNominal}{3.03}
-\newcommand{\daAPICurveFullDeployedYears}{$3.03 \pm 67$}
+\newcommand{\daOSCurveHalfDeployedYearsNominal}{0.874}
+\newcommand{\daOSCurveHalfDeployedYears}{$0.874 \pm 0.242$}
+\newcommand{\daOSCurveFullDeployedYearsNominal}{2.98}
+\newcommand{\daOSCurveFullDeployedYears}{$2.98 \pm 67$}
+\newcommand{\daAPICurveHalfDeployedYearsNominal}{0.925}
+\newcommand{\daAPICurveHalfDeployedYears}{$0.925 \pm 0.239$}
+\newcommand{\daAPICurveFullDeployedYearsNominal}{3.04}
+\newcommand{\daAPICurveFullDeployedYears}{$3.04 \pm 68$}
 \newcommand{\daGPAPICurveHalfDeployedYearsNominal}{0.947}
@@
 \newcommand{\daGPAPIPerAPIParametersTable}{\begin{table} \centering \begin{tabular}{c|S|S} API Version & ${t_0}$ & $\mathrm{decay}$ \\ \hline2	&	0.0	&	0.0237\\3	&	0.0	&	0.0235\\4	&	0.0	&	0.00689\\5	&	59.1	&	0.00465\\6	&	2.4	&	0.00434\\7	&	0.0	&	0.00499\\8	&	47.7	&	0.00415\\9	&	155	&	0.00365\\10	&	90.2	&	0.00364\\11	&	432	&	0.00227\\12	&	355	&	0.00227\\13	&	303	&	0.00235\\14	&	218	&	0.00238\\15	&	160	&	0.00239\\16	&	141	&	0.00226\\17	&	274	&	0.00184\\18	&	116	&	0.00138\\19	&	118	&	0.00152\\\end{tabular} \caption{Parameters for different API versions} \label{tab:per_api_parameters} \end{table}}
+\newcommand{\daUpdatednessPercTwosfNominal}{5.2\%}
+\newcommand{\daUpdatednessPercTwosf}{$5.2 \pm 0.0\%$}

### 2015-03-25 securityupdates.tex
Update stats to include apdated avoTabAndVulns
todos don't need to be inline when we are in a single column format
Formatting tweaks

@@
 \let\OldTodo\todo
-\renewcommand{\todo}{\OldTodo[inline]}
+\renewcommand{\todo}{\OldTodo}%[inline]}
 \newcommand{\todolater}[1]{\todo{#1}}%{}%
@@
 On some versions of Android (below version 4.1) malware could use known system-to-root escalation mechanisms but on all versions they have a greatly increased attack area for further privilege escalation and also have the ability to control all user internet traffic (via VPNs), brick the phone, remove and install apps, steal user credentials, read the screen and make as well as receive calls.
-\todolater{break vulnerabilities down by attack vector?}
+The different categories which the vulnerabilities fall into are shown in Table~\ref{tab:andvulns}.
 \avoTabAndVulns
@@
 We developed and maintain an open platform for filing critical vulnerabilities in a machine readable format, the \href{http://androidvulnerabilities.org/}{AndroidVulnerabilities.org}\footnote{\textbf{Note to reviewers:} we have removed our names and affiliation from the website for this review \textbf{with the exception of the contact page.
-Please do not visit the contact page until the review cycle has completed.}\vspace{-3em}} (\avo) website.
+Please do not visit the contact page until the review cycle has completed.}\vspace{-4em}} (\avo) website.
 We seeded it with data from the CVE database, vendor lists, reports from the literature and forums.
@@
 
-\begin{figure}[h]
+\begin{figure}[!h]
  \centering
@@
 \label{sec:representative}
-\begin{figure}[h]
+\begin{figure}[!h]
  \centering

### 2015-03-25 securityupdates.tex
Add some TODOs

@@
 %Android is the most popular smartphone operating system today with \percMarketShare\ marketshare.
-
+\todo{refute conventional wisdom to show novelty and importance of real data}
 Support for third-party apps is a key feature of modern smartphone operating systems.
@@
 However what is perhaps more worrying is the long tail on the \avovuln{Gingerbreak}, \avovuln{levitator}, \avovuln[exploid_udev]{exploid} and \avovuln{zergRush} vulnerabilities which are more dangerous root vulnerabilities (not requiring new APK installation) and which affect a significant proportion of devices years later.
+\todo{quantify this long tail and the APK numbers as headline numbers}
 
@@
 We also recorded when the version changed and which versions it changed from and to.
-In the \da\ data we have over 1000 devices contributing data in any particular week.
+In the \da\ data we have over 1000 devices contributing data in any particular week.\todolater{Use the real average number of devices contributing each week}
 However most devices only contribute for a short period of time and so we do not observe updates happening on every device.

### 2015-03-25 securityupdates.tex
Comment out paragraph on vulnerability distribution to save space

@@
 
-The discovery dates of our vulnerabilities are not uniform. 
-In particular the data shows a large gap from 2011-10-06 to 2013-02-18 where we do not know of any discoveries of critical vulnerabilities affecting all Android devices.
-The cause of this quiet period is unclear.
-Possible explanations are that: (i) most devices were exposed to known vulnerabilities so there was reason to look for new ones;
-(ii) device manufacturers made it easier to install custom versions of Android, reducing the need for users to root their devices; and
-(iii) device manufacturer specific vulnerabilities (which are not in our analysis here) were easier to find and so attackers looking for vulnerabilities adjusted their focus.
-\todolater{If we need to save space then delete this paragraph}
+%The discovery dates of our vulnerabilities are not uniform. 
+%In particular the data shows a large gap from 2011-10-06 to 2013-02-18 where we do not know of any discoveries of critical vulnerabilities affecting all Android devices.
+%The cause of this quiet period is unclear.
+%Possible explanations are that: (i) most devices were exposed to known vulnerabilities so there was reason to look for new ones;
+%(ii) device manufacturers made it easier to install custom versions of Android, reducing the need for users to root their devices; and
+%(iii) device manufacturer specific vulnerabilities (which are not in our analysis here) were easier to find and so attackers looking for vulnerabilities adjusted their focus.
 

### 2015-03-25 securityupdates.tex
Motivate each analysis and propagate headline numbers

@@
 We analysed \da\ data from over \daOSYearsOfData\ years and \daNumOSDevices\ devices and found that on average \daMeanInsecurityPercTwosfNominal\ of devices were exposed to known root privilege vulnerabilities and only \daUpdatednessPercTwosfNominal\ of devices run the most recent version of Android.
+Devices apply \daUpdatesPerYearTwosfNominal\ updates each year, less than the rate of critical vulnerability discovery of between \avoVulnsPerYearAllAndroidTwosfNominal\ and \avoVulnsPerYearTwosfNominal\ per year.
 \end{abstract}
@@
 \label{sec:results}
-We present the results of our analysis showing that, on average, \daMeanInsecurityPerc\ of Android devices are exposed to critical vulnerabilities.
-Devices, on average, apply \daUpdatesPerYear\ updates each year and \daPercUpdatesDowngrades\ of full version changes are downgrades to older versions.
+We present the results of our analysis showing that, on average, \daMeanInsecurityPerc\ of Android devices are exposed to critical vulnerabilities and only \daUpdatednessPerc\ run the latest version of Android.
+Devices, on average, apply \daUpdatesPerYear\ updates each year less than the rate of critical vulnerability discovery of between \avoVulnsPerYearAllAndroid\ and \avoVulnsPerYear.
+\daPercUpdatesDowngrades\ of full version changes are downgrades to older versions.
 We describe three analyses, examining the vulnerability of Android devices in general (\S\ref{sec:exp:versionsecurity}), explore the upgrade cycle and vulnerability cycle of the ecosystem (\S\ref{sec:exp:android_ecosystem}) and quantify the updates installed on particular devices (\S\ref{sec:exp:device_updates}).
@@
 \subsection{Analysis 1: vulnerability of Android devices}\label{sec:exp:versionsecurity}
-
+In this analysis we calculate the proportion of Android devices which are exposed to critical vulnerabilities and and discuss how that has changed over time.
 In the introduction, Figure~\ref{fig:proportioninsecure} described the proportion of Android devices susceptible to at least one critical vulnerability.
@@
 
+This gives us an average exposure to critical vulnerabilities of Android devices of \daMeanInsecurityPerc.
+
 \subsection{Analysis 2: behaviour of the Android ecosystem}\label{sec:exp:android_ecosystem}
+In this analysis we examine how the distribution of Andriod OS versions changes over time and the impact that has on how different vulnerabilities affect the security of Android.
+We show that some vulnerabilities continue to have a substantial effect long after they have been fixed and the importance of longitudinal studies for determining whether a particular device is exposed to a particular vulnerability.
 
@@
 \subsubsection{Results}
-The proportion of devices in the \da\ data running different versions of Android each day is shown in Figure~\ref{fig:norm_os}.
+The proportion of devices in the \da\ data running different versions of Android each day is shown in Figure~\ref{fig:norm_os}.\footnote{The anomaly beginning in August 2014 is explained in \S\ref{sec:da_changes}.}
 It shows how old versions are gradually replaced by new ones, and the long tail of devices which do not see updates to more recent versions.
-The anomaly beginning in August 2014 is explained in \S\ref{sec:da_changes}.
+This gives a mean proportion of devices running the most recent version of Android is \daUpdatednessPerc.
 
@@
 However the latency in the security update process means that on average \daMeanInsecurityPerc\ of Android devices are exposed to known critical vulnerabilities which break out of the sandbox and only \daUpdatednessPerc\ of devices run the latest version of Android.
+Devices apply \daUpdatesPerYear\ updates each year less than the rate of critical vulnerability discovery of between \avoVulnsPerYearAllAndroid\ and \avoVulnsPerYear.
 

### 2015-03-25 securityupdates.tex
Tweaks

@@
 \subsection{Critical vulnerabilities}
-
 We compiled a list of critical vulnerabilities in Android, containing information on the discovery and publication dates, the versions affected and which versions fixed the problem.
@@
 This is not something that applications running on the phone can exploit to break out of the app sandbox and so we do not include those vulnerabilities.
+However there have been reports of Windows malware exploiting this vector.\footnote{\url{http://www.symantec.com/connect/blogs/windows-malware-attempts-infect-android-devices}}
 Unfortunately, many published exploits use ADB for convenience and so determining whether the use of ADB is necessary to exploit the vulnerability can be difficult.
@@
 In addition, we have received submissions or amendments from \avoNumSubmitters\ individuals.
-We collected data between \avoStartDate\ and \avoEndDate\ and so does not include any information lost before the start of that period.
+We collected data between \avoStartDate\ and \avoEndDate\ and so it does not include any information lost before the start of that period.
 \avo\ currently contains \avoNumVulnerabilities\ vulnerabilities of which \avoNumVulnAllAndroid\ affect all Android devices and \avoNumVulnSpecific\ are specific to particular devices or device manufacturers.
@@
 Previous work has assumed ``any security issue of relevance will eventually get a CVE number assigned"~\cite{Frei2010} which is currently not the case for critical Android vulnerabilities.
-For some of the vulnerabilities without CVE numbers, Google confirmed that there was no CVE number and that they did not intend to get one, instead providing an Android bug number.
+For some of the vulnerabilities without CVE numbers, Google confirmed that there was no CVE number and that they did not intend to get one, instead providing an internal Android bug number.
 
@@
 This occurs when the information about the vulnerability is available to the from an accepted source.
-Unfortunately before we collated this information much of it was not published by an independent source, even months or years after they had been actively used.
+Unfortunately before we collated this information much of it was not published by an accepted source, even months or years after they had been actively used.
 

### 2015-03-25 securityupdates.tex
Update contributions

@@
 \begin{itemize}
- \item We characterise the Android update ecosystem, showing how updates flow between entities, and quantify vulnerabilities by determining what proportion of devices they affected and how that changes over time.
+ \item We quantify the Android update ecosystem showing how the OS version distribution changes and the frequency of updates.
+ \item We quantify vulnerabilities by determining what proportion of devices they affected and how that changes over time.
  \item We created an open database to collate quantitative information on vulnerabilities which affect Android, and investigate their time-lines.

### 2015-03-25 securityupdates.tex
Rework Threats to Validity and destroy Discussion by blending into nearby sections

@@
 \label{sec:validity}
-\subsection{The \da\ data gives a conservative estimate of the Android version distribution}
+\subsubsection{Comparing \da\ data with the ground truth}
 \label{sec:representative}
@@
 Figure~\ref{fig:play_api} shows the data from Google and Figure~\ref{fig:da_api} shows the data from \da\ and they appear similar.
-We analysed the difference between the API version data from \da\ and Gooble Play, normalising for days since the API version was released.
-This that the \da\ data systematically overestimates the prevalence of new API versions and underestimates the prevalence of old API versions (except for API version 17 which was particularly popular in a focused study discussed in the next section and so was temporarily overestimated when an old version).
+We analysed the difference between the API version data from \da\ and Google Play, normalising for days since the API version was released.
+This shows that the \da\ data systematically overestimates the prevalence of new API versions and underestimates the prevalence of old API versions (except for API version 17 which was particularly popular in a focused study discussed in the next section and so was temporarily overestimated when an old version).
 This means that the OS version information from \da\ is likely to be overestimating the prevalence of new OS versions and hence our results are a conservative estimate of the security of Android.
+Since most \da\ users are self selecting and install \da\ because they want to find out more about what their phone is doing or to aid research they may be biased and perhaps more likely to install updates.
 This allows us to have confidence in the OS version information.
@@
 
-\subsection{Changes in the \da\ sample}
+\subsubsection{The effect of focussed studies on \da}
 \label{sec:da_changes}
@@
 One study was conducted in Norway with 654 installs and one in Bangladesh where 2463 users installed \da.
-This is responsible for the sudden change in distribution of OS versions in \da\ beginning in August 2014 Figures \ref{fig:norm_os}, \ref{fig:nvulnerabilities_heat} and \ref{fig:proportioninsecure} as for that period Bangladeshi users contributed over half of the \da\ data.
+This latter study is responsible for the sudden change in distribution of OS versions in \da\ beginning in August 2014 in Figures \ref{fig:norm_os}, \ref{fig:nvulnerabilities_heat} and \ref{fig:proportioninsecure} as for that period Bangladeshi users contributed over half of the \da\ data.
 %We could normalise out the influence of the Bangladeshi users, however they are Android users too and they also need security updates.
-To investigate the sensitivity of our results to the Bangladesh study we truncated our data before this study started and found the percentage of vulnerable devices to be 87.7\% rather than \daMeanInsecurityPercNominal, and the percentage of devices running the latest version to be 5.43\% rather than \daUpdatednessPercNominal.
+To investigate the sensitivity of our results to the Bangladesh study we truncated our data before this study started and found the percentage of vulnerable devices to be 85.6\% rather than \daMeanInsecurityPercNominal, the percentage of devices running the latest version to be 5.71\% rather than \daUpdatednessPercNominal\ and the number of updates per year to be 1.48 rather than \daUpdatesPerYearNominal.
 Hence removing this anomaly would have no effect on our main conclusions.
 
-Since most \da\ users are self selecting and install \da\ because they want to find out more about what their phone is doing or to aid research they may be biased and perhaps more likely to install updates.
-
-\section{Discussion} \label{sec:discussion}
-There are continuing efforts to reduce the impact of critical vulnerabilities, both in Android and more widely.
-SEAndroid~\cite{Smalley2013} which is included in Android from version 4.1~\cite{jelly-bean-release} claimed to prevent some root vulnerabilities and to reduce the impact of others.
-Capability based enforcement systems such as Capsicum~\cite{Watson2010} substantially reduce the capabilities that an exploit has to try and gain increased privilege with and could be included in Linux\footnote{\url{https://github.com/google/capsicum-linux}} and hence Android.
-
+\subsubsection{The selection of vulnerabilities in \avo}
 We have not included all the vulnerabilities from \avo. 
@@
 
+\subsubsection{Limitations}
 We still do not know where new versions of Android come from.
@@
 These features, as well as Address Space Layout Randomisation (ASLR) and mandatory access controls, has resulted in a lower level of malware affecting iOS when compared to Android~\cite{Felt2011}.
+
+There are continuing efforts to reduce the impact of critical vulnerabilities, both in Android and more widely.
+SEAndroid~\cite{Smalley2013} which is included in Android from version 4.1~\cite{jelly-bean-release} claimed to prevent some root vulnerabilities and to reduce the impact of others.
+Capability based enforcement systems such as Capsicum~\cite{Watson2010} substantially reduce the capabilities that an exploit has to try and gain increased privilege with and could be included in Linux\footnote{\url{https://github.com/google/capsicum-linux}} and hence Android.
 

### 2015-03-25 securityupdates.tex
Rewrite Introduction

@@
 \begin{document}
-\title{Timeliness of security updates for Android}
+\title{The thin blue line:\\what protects my Android device?}
 
@@
 \section{Introduction}
-%Android is the most popular smartphone operating system today with \percMarketShare\ marketshare.
-\todo{refute conventional wisdom to show novelty and importance of real data}
-Support for third-party apps is a key feature of modern smartphone operating systems.
-Such apps are written by many developers from a wide range of backgrounds, which means neither operating system vendors nor users can fully trust app developers.
-Therefore, in order to secure personal data and prevent theft, such as the sending of premium-rate text messages, operating system vendors provide a protected execution environment, or \emph{sandbox}, for apps.
+We show that the security of Android relies on the centralisation of app deployment with Google.
+Computer systems which run untrusted code are considered secure if there are strong sandboxing mechanisms which prevent the untrusted code doing dangerous things.
+However we show the Android ecosystem is secure even though on average \daMeanInsecurityPercTwosfNominal\ of devices are exposed to known critical vulnerabilities which break the sandboxing mechanisms.
+The recurrent advice on keeping computers secure is to install the latest updates yet only \daUpdatednessPercTwosfNominal\ of Android devices run the latest version and the rate of critical vulnerability discovery of between \avoVulnsPerYearAllAndroidTwosfNominal\ and \avoVulnsPerYearTwosfNominal\ per year is higher than the rate of updates being delivered of \daUpdatesPerYearTwosfNominal.
+Yet Android malware deployments are not widespread despite the many attempts of malware authors:
+in 2012, between 36.7\%~\cite{Zhou2012b} and 40\%~\cite{Zhou2012a} of malware for Android contained exploits designed to enable a malicious app to break the sandbox and thousands of new malware samples are found each year.
 
-In this paper we quantify the extent to which the app sandbox provided by Android protects devices from malicious apps attempting to gain control over the entire handset.
-%We have analysed Android because it is the most popular smartphone operating system by market share (~\percMarketShare) and 
-Previous research has shown that in 2012, between 36.7\%~\cite{Zhou2012b} and 40\%~\cite{Zhou2012a} of malware for Android contained exploits designed to enable a malicious app to break the sandbox.
-What is not known however is the proportion of Android devices which are vulnerable to these exploits over time.
-To the best of our knowledge, this is the first paper to provide detailed numbers.
+Android is protected by many layers of security and while the sandbox is usually broken on most Android devices by known vulnerabilities, the outer layers of protection such as the Google Play store prevent the installation of malicious apps.
+The Google Play store provides economic (app developers have to pay to register), technical (Google Bouncer which scans apps) and social (user reviews and reports of malicious software) measures to prevent malicious apps being distributed.
+Even users who cannot use Google Play (e.g.\ because they are in China) may still get its `Verify apps' functionality which scans for malicious apps.
+
+The claim of the open source software `bazzar' is that it is more secure because anyone can review the source code and provide fixes~\cite{Raymond1999}.
+Yet Android is not a `bazzar' in the way that Debian is because while much of the code is open source and available for the many eyes review, end users cannot easily rebuild the binaries that run on their phones due to the binary blobs and closed source components within them.
+Since only manufacturers can do that, rather than any one technically competent user, the availability of updates is controlled by the commercial interests of the manufacturers (and the operators).
+Manufacturers do not even provide updates for the length of the contracts under which the users originally bought the phone, despite efforts by the ACLU to persuade the FTC to force them to do so~\cite{Soghoian2013}, and despite the FTC forcing HTC to do so~\cite{FTCHTC2013}.
+Since only manufacturers can provide updates, malicious parties cannot do so, if manufacturers promptly provided updates this would improve security, but since they do not, it makes it worse.
+
+However, despite the current low level of malicious software deployment on Android, the high level of vulnerability and the slow deployment of updates is still a cause for concern.
+Some users may only be able to use third party app stores or have USB debugging enabled and connect their phone to a malicious machine.
+There have also been two remote code execution vulnerabilities in Android, one in the Javascript-to-Java bridge and one in dhcpd which allow a network attacker to run malicious code in a user context, bypassing all the checks on apps and leaving only the sandbox which we show to be vulnerable.
 
@@
 
-\begin{figure}[h]
-\centering
-\includegraphics[width=\columnwidth]{figures/proportioninsecure}
-\caption{Proportion of devices running insecure, maybe secure and secure versions of Android against time.
-The red vertical lines are caused by vulnerabilities being discovered with those which have the biggest impact annotated.
-This graph is explained further in \S\ref{sec:exp:versionsecurity}.
-}
-\label{fig:proportioninsecure}
-\end{figure}
-
-In order to determine whether an individual smartphone is vulnerable to a particular exploit, we need to combine vulnerability data with the operating system version string and build number from the handset.
+In order to determine whether an individual smartphone is exposed to a particular vulnerability, we need to combine vulnerability data with the operating system version string and build number from the handset.
 For this we use handset data gathered from \daNumOSDevices\ devices in the \da~\cite{Wagner2013} project.
-This analysis allows us to plot the proportion of Android devices in the \da\ data which were running versions of Android known to have vulnerabilities at the time and is shown in Figure~\ref{fig:proportioninsecure}.
-The figure shows the proportion of devices split into three categories:
-(1)~\emph{secure} because they are running a version of Android with no known vulnerabilities; 
-(2)~\emph{insecure} because the device was running an insecure version of Android with a build released before the vulnerability was fixed or discovered; and 
-(3)~\emph{maybe secure} because the device is running a version of Android known to be insecure but has received an update which might contain a backported fix as we did not observe that build number until after the vulnerability was known.
-Further analysis into the propagation of updates in Android is presented in \S\ref{sec:exp:android_ecosystem}.
+From this we have determined that since \daStartDate\ on five occasions all Android devices have been exposed to known critical vulnerabilities and that it takes a year for updates to reach most devices, by which time new vulnerabilities have been discovered.
 
@@
 \begin{itemize}
+ \item We show that the security of Android relies on the centralisation of app deployment with Google rather than the absence of vulnerabilities in its sandbox
  \item We quantify the Android update ecosystem showing how the OS version distribution changes and the frequency of updates.

### 2015-03-25 securityupdates.tex
Minor fixes

@@
 Neither static nor dynamic analysis of this app by the marketplace will uncover any malicious code, since it does exist in the app binary.
-Instead the attacker can chose to send a malicious payload directly to the app for execution at a later point in time.
-Note that dynamic code loading is not actually needed for this to work -- it just makes the task easier -- since previous work has shown that even on a platform such as iOS, which does not permit dynamic code loading, a ROP-based attack is relatively easy if the attacker creates an app with carefully crafted flaws~\cite{Wang2013a}.
+Instead the attacker can choose to send a malicious payload directly to the app for execution at a later point in time.
+Note that dynamic code loading is not actually needed for this to work -- it just makes the task easier -- since previous work has shown that even on a platform such as iOS, which does not permit dynamic code loading, a Return oriented Programming (ROP)-based attack is relatively easy if the attacker creates an app with carefully crafted flaws~\cite{Wang2013a}.
 
@@
 The fix for the addJavascriptInterface vulnerability breaks backwards compatibility, and therefore apps are only immune to this attack if their target API level is greater than 16 and they are running on an Android device with an OS version of 4.2 or greater.
-Later analysis in Figure~\ref{fig:da_api} in this paper shows that 40\% of handsets connecting to the Google Play store are currently vulnerable to this attack.
-\todo{Make that percentage automatically generated so that it is correct}
+Our analysis shows that on \daGPAPISeventeenLaterDate, \daGPAPISeventeenEarlierProportion\ of handsets connecting to the Google Play store were still vulnerable to this attack.
 

### 2015-03-25 securityupdates.tex
Move proportioninsecure graph to analysis 1

@@
 \subsection{Analysis 1: vulnerability of Android devices}\label{sec:exp:versionsecurity}
+\begin{figure}[h]
+\centering
+\includegraphics[width=\columnwidth]{figures/proportioninsecure}
+\caption{Proportion of devices running insecure, maybe secure and secure versions of Android against time.
+The red vertical lines are caused by vulnerabilities being discovered with those which have the biggest impact annotated.
+}
+\label{fig:proportioninsecure}
+\end{figure}
 In this analysis we calculate the proportion of Android devices which are exposed to critical vulnerabilities and and discuss how that has changed over time.
-In the introduction, Figure~\ref{fig:proportioninsecure} described the proportion of Android devices susceptible to at least one critical vulnerability.
+In Figure~\ref{fig:proportioninsecure} describes the proportion of Android devices susceptible to at least one critical vulnerability.
 

### 2015-03-25 securityupdates.tex
Remove API comparison graphs from Threats to Validity to save space

@@
 \label{sec:representative}
-\begin{figure}[!h]
- \centering
- \begin{subfigure}[b]{\columnwidth}
- \includegraphics[width=\columnwidth]{figures/googleplayapi}
- \caption{Google Play data on proportion of devices running different API versions}
- \label{fig:play_api}
-\end{subfigure}
-\begin{subfigure}[b]{\columnwidth}
- \includegraphics[width=\columnwidth]{figures/norm_api_gpcomp}
- \caption{\da\ data on proportion of devices running different API versions}
- \label{fig:da_api}
-\end{subfigure}
-\caption{Monthly Android API version data}
-\end{figure}
 
@@
 While API versions are too coarse grained to use for security update detection they are closely related to OS versions and so if the \da\ data on API versions are similar to the Google data on API versions then the \da\ data on OS versions should be representative.
-Figure~\ref{fig:play_api} shows the data from Google and Figure~\ref{fig:da_api} shows the data from \da\ and they appear similar.
+We compared the data from Google and from \da\ and they appear similar, except for the anomaly discussed later.
 We analysed the difference between the API version data from \da\ and Google Play, normalising for days since the API version was released.

### 2015-03-25 securityupdates.tex
Move paragraph about showing standard deviations

@@
 
+Where possible we indicate the uncertainty in our results by presenting them $\pm$ one standard deviation and give results to 3 significant figures, this occasionally results in `$\pm\, 0$' when the standard deviation is small.
+This technique cannot account for systematic errors which we explore in \S\ref{sec:representative}.
+
 \subsection{Versions of Android running on devices}
@@
 The majority of devices only contribute data for a short period of time, however \daMonthsDevices\ devices have contributed data for more than \daMonths~months.
-
-Where possible we indicate the uncertainty in our results by presenting them $\pm$ one standard deviation and give results to 3 significant figures, this occasionally results in `${} \pm 0$' when the standard deviation is less than the number of significant figures.
-This technique cannot account for systematic errors which we explore in \S\ref{sec:representative}.
 

### 2015-03-25 securityupdates.tex
Tweaks and note that we can't get this data from Google

@@
 On a large proportion of devices `opaque\_marker' is a well defined build number\footnote{\url{https://source.android.com/source/build-numbers.html}} however different device manufacturers use different schema.
+Google provides API version distribution information but not the OS and build version information we need, we verify that the \da\ data is representative in~\S\ref{sec:representative}.
 
@@
 In the \da\ data we have over 1000 devices contributing data in any particular week.\todolater{Use the real average number of devices contributing each week}
+
 However most devices only contribute for a short period of time and so we do not observe updates happening on every device.
@@
 We have not included all the vulnerabilities from \avo. 
-In particular, we have not included any device manufacturer specific vulnerabilities, even when these are widespread (such as vulnerabilities affecting Qualcomm\footnote{Qualcomm is particularly good at publicly disclosing vulnerabilities and the patches which fix them.
+In particular, we have not included any device manufacturer specific vulnerabilities, even when these are widespread (such as vulnerabilities in Qualcomm\footnote{Qualcomm is particularly good at publicly disclosing vulnerabilities and the patches which fix them.
 \url{https://www.codeaurora.org/projects/security-advisories/}} chipsets) because it is difficult to work out which devices are affected.
@@
 
-There are continuing efforts to reduce the impact of critical vulnerabilities, both in Android and more widely.
+There are continuing efforts to reduce the impact of critical vulnerabilities, both in Android and elsewhere.
 SEAndroid~\cite{Smalley2013} which is included in Android from version 4.1~\cite{jelly-bean-release} claimed to prevent some root vulnerabilities and to reduce the impact of others.

### 2015-03-30 securityupdates.tex
New version with modified introduction and conclusion.

@@
 \section{Introduction}
-We show that the security of Android relies on the centralisation of app deployment with Google.
-Computer systems which run untrusted code are considered secure if there are strong sandboxing mechanisms which prevent the untrusted code doing dangerous things.
-However we show the Android ecosystem is secure even though on average \daMeanInsecurityPercTwosfNominal\ of devices are exposed to known critical vulnerabilities which break the sandboxing mechanisms.
-The recurrent advice on keeping computers secure is to install the latest updates yet only \daUpdatednessPercTwosfNominal\ of Android devices run the latest version and the rate of critical vulnerability discovery of between \avoVulnsPerYearAllAndroidTwosfNominal\ and \avoVulnsPerYearTwosfNominal\ per year is higher than the rate of updates being delivered of \daUpdatesPerYearTwosfNominal.
-Yet Android malware deployments are not widespread despite the many attempts of malware authors:
-in 2012, between 36.7\%~\cite{Zhou2012b} and 40\%~\cite{Zhou2012a} of malware for Android contained exploits designed to enable a malicious app to break the sandbox and thousands of new malware samples are found each year.
 
-Android is protected by many layers of security and while the sandbox is usually broken on most Android devices by known vulnerabilities, the outer layers of protection such as the Google Play store prevent the installation of malicious apps.
-The Google Play store provides economic (app developers have to pay to register), technical (Google Bouncer which scans apps) and social (user reviews and reports of malicious software) measures to prevent malicious apps being distributed.
-Even users who cannot use Google Play (e.g.\ because they are in China) may still get its `Verify apps' functionality which scans for malicious apps.
+Modern mobile operating systems provide several layers of security whose combined aim is to prevent app malware from taking control of a device. 
+In this paper we explore the security layers present in the Android ecosystem and determine which layers are effective and which are not.
 
-The claim of the open source software `bazzar' is that it is more secure because anyone can review the source code and provide fixes~\cite{Raymond1999}.
-Yet Android is not a `bazzar' in the way that Debian is because while much of the code is open source and available for the many eyes review, end users cannot easily rebuild the binaries that run on their phones due to the binary blobs and closed source components within them.
-Since only manufacturers can do that, rather than any one technically competent user, the availability of updates is controlled by the commercial interests of the manufacturers (and the operators).
-Manufacturers do not even provide updates for the length of the contracts under which the users originally bought the phone, despite efforts by the ACLU to persuade the FTC to force them to do so~\cite{Soghoian2013}, and despite the FTC forcing HTC to do so~\cite{FTCHTC2013}.
-Since only manufacturers can provide updates, malicious parties cannot do so, if manufacturers promptly provided updates this would improve security, but since they do not, it makes it worse.
+Our focus is on malware whose aim is to gain full control of the device rather than merely abuse permissions granted to it by the user.
+In this context, by full control, or \emph{root}, on a device we mean the malware has unfettered access to all data stored on the device and can control all network interfaces in the device with the same flexibility as the operating system itself.
+We have selected our focus for two reasons: no previous published work has quantified how vulnerable Android devices are to malware gaining root, and secondly the negative impact on the device owner is much higher for malware which gains root than malware which does not: malware with root can not only perform arbitrary actions on the device, but it is also impossible for the device owner to remove the malware reliably since even a factory reset does not guarantee that the malware has been successfully removed.
 
-However, despite the current low level of malicious software deployment on Android, the high level of vulnerability and the slow deployment of updates is still a cause for concern.
-Some users may only be able to use third party app stores or have USB debugging enabled and connect their phone to a malicious machine.
-There have also been two remote code execution vulnerabilities in Android, one in the Javascript-to-Java bridge and one in dhcpd which allow a network attacker to run malicious code in a user context, bypassing all the checks on apps and leaving only the sandbox which we show to be vulnerable.
+This is an important problem: previous studies have shown that a significant proportion of malware attempts to gain root.
+For example, in 2012 between 36.7\%~\cite{Zhou2012b} and 40\%~\cite{Zhou2012a} of malware for Android contained exploits designed to gain root-level access.
 
-Not all malicious apps need to break the sandbox in order to misbehave.
-We focus on the issue of apps breaking out of the sandbox because malware which is able to take full control of a handset can do significantly more harm and is much harder to remove.
-Ordinary malware can be uninstalled by the user, or remotely by Google through the Play Store, and returns the device to a secure state. 
-Removing an app which has used an exploit to take control of the handset is unlikely to return the device to secure state.
-We describe our threat model and three attack vectors in detail in \S\ref{sec:threatmodel}.
+Android users are protected from malicious apps by up to three levels of defence. 
+The first level of defence comes from an app marketplace such as Google Play. 
+A marketplace can use three mechanisms to prevent malicious apps from appearing in the marketplace: firstly by economic disincentives, including charging developers to create an account and cancelling accounts used to upload malicious apps; secondly by technical means, such as static and dynamic analysis on submitted apps; and finally via social feedback between users, including the provision of user reviews and ratings.
 
-In order to determine whether an individual smartphone is exposed to a particular vulnerability, we need to combine vulnerability data with the operating system version string and build number from the handset.
-For this we use handset data gathered from \daNumOSDevices\ devices in the \da~\cite{Wagner2013} project.
-From this we have determined that since \daStartDate\ on five occasions all Android devices have been exposed to known critical vulnerabilities and that it takes a year for updates to reach most devices, by which time new vulnerabilities have been discovered.
+The second line of defence occurs on the handset at installation time, and includes checking the signature of any app installation or update as well as verifying the app by local anti-virus tools (e.g.
+Google's app verification feature).
+Finally, the third line of defence takes place at runtime: Android apps are run in a \emph{sandbox} which prevents the app from unauthorised actions such as accessing any permissions which weren't granted by the user at install time; in addition, anti-virus products may scan the list of installed apps periodically to detect malware.
 
-In summary, the contributions of this paper are:
+Sandboxes only provide protection if they are free from design errors and implementation flaws. 
+Both of these have occurred in Android, as indeed happens today in many large security systems relying on the sandbox model. 
+Maintaining sandbox security therefore relies on providing timely updates to fix design and implementation flaws as and when they are discovered.
+
+In this paper we show that the Android sandbox only provides effective security on 12\% of devices today. 
+This is because most devices do not receive timely updates. 
+Therefore the security of the Android ecosystem is currently reliant on the defences provided by the app marketplace and (when enabled) anti-virus systems. 
+This insight allows us to provide users with a list of concrete actions to improve device security. 
+It also raises broader questions about how modern mobile device security is most effectively achieved, how a secure ecosystem should be construct and managed, and whether security by obscurity is the right way forward.
+
+In summary, our contributions are:
 \begin{itemize}
- \item We show that the security of Android relies on the centralisation of app deployment with Google rather than the absence of vulnerabilities in its sandbox
  \item We quantify the Android update ecosystem showing how the OS version distribution changes and the frequency of updates.
- \item We quantify vulnerabilities by determining what proportion of devices they affected and how that changes over time.
- \item We created an open database to collate quantitative information on vulnerabilities which affect Android, and investigate their time-lines.
+ \item We created an open database to collate quantitative information on vulnerabilities which affect Android and use this to determine the proportion of devices they affect and how that changes over time.
+ \item We demonstrate that the security of Android relies on the centralisation of app deployment via a marketplace, and the use of install-time anti-virus checks rather than through the security of the app sandbox.
+ \item We use our analysis to develop advice for Android users on what steps they should take to increase the security of their device.
+
 \end{itemize}
 
-\section{Threat model}
+\section{Threat model and attack vectors}
 \label{sec:threatmodel}
 
-In this paper we measure whether an attacker can take control of a handset by breaking out of the app sandbox on Android using a known vulnerability.
-We consider three attack vectors.
+In this paper we assume the defender can deploy security controls at three levels: in an online marketplace, at app installation, and during app execution.
+We then consider three attack vectors an attacker can use to attempt to gain root on an Android handset.
+
 The first attack vector is through the installation of a malicious app on the device.
 Android devices support the installation of apps through a number of app marketplaces, email attachments, URLs and via the Android Debug Bridge (ADB).
-By default, many Android devices with Google Play installed will prevent the installation of apps from other sources.
-
-Controlling potential sources of app installation is a good security feature because app marketplaces provide additional protection to prevent the listing of malicious apps.
-For example, the Google Play marketplace uses Bouncer to automatically analyse apps as well as removing any apps which are reported as malicious.
+By default, many Android devices with Google Play installed will prevent the installation of apps from other sources, and therefore an attacker needs to compromise all three levels of security.
+The Google Play marketplace uses Bouncer to automatically analyse apps as well as removing any apps which are reported as malicious.
 Nevertheless, Google Play is not immune: 0.02\% of apps are malicious~\cite{Zhou2012a}.
@@
 The most direct method available to an attacker is to upload to a marketplace a seemingly innocent app which contains support for dynamic code loading from a remote server under control of the attacker.
-Neither static nor dynamic analysis of this app by the marketplace will uncover any malicious code, since it does exist in the app binary.
-Instead the attacker can choose to send a malicious payload directly to the app for execution at a later point in time.
-Note that dynamic code loading is not actually needed for this to work -- it just makes the task easier -- since previous work has shown that even on a platform such as iOS, which does not permit dynamic code loading, a Return oriented Programming (ROP)-based attack is relatively easy if the attacker creates an app with carefully crafted flaws~\cite{Wang2013a}.
+Neither static nor dynamic analysis of this app by the marketplace will uncover any malicious code, since it does not exist in the app binary.
+The attacker can choose to send a malicious payload directly to the app for execution at a later point in time.
+The marketplace can detect explicit use of dynamic code loading, even if malicious code cannot be detected.
+Note that dynamic code loading is not actually needed for this to work however -- it just makes the task easier -- since previous work has shown that even on a platform such as iOS, which does not permit dynamic code loading, a Return-Oriented Programming (ROP)-based attack is relatively easy if the attacker creates an app with carefully crafted flaws~\cite{Wang2013a}.
 
@@
 The fix for the addJavascriptInterface vulnerability breaks backwards compatibility, and therefore apps are only immune to this attack if their target API level is greater than 16 and they are running on an Android device with an OS version of 4.2 or greater.
-Our analysis shows that on \daGPAPISeventeenLaterDate, \daGPAPISeventeenEarlierProportion\ of handsets connecting to the Google Play store were still vulnerable to this attack.
+Our analysis shows that as of \daGPAPISeventeenLaterDate, \daGPAPISeventeenEarlierProportion\ of handsets connecting to the Google Play store were still vulnerable to this attack.
+
+In summary, the best place to prevent attacks is at runtime, since all three attack vectors can be prevented at this level. 
+Unfortunately, as we will demonstrate, the sandbox for Android apps is weak and anti-virus is either not installed or not activated by default.
+Therefore many Android users implicitly rely on the marketplace for protection by removing apps which are detected as malicious.
+This detection can only be derived from static and dynamic analysis or after reports of malicious behaviour from users once it is made available for download.
 
@@
 
-We require two sources of data for this study: (1) information on the distribution of installed versions of Android over time and (2) information on the critical vulnerabilities found to affect specific versions of Android.
-These two datasets can then be combined to determine the proportion of handsets at risk of attack from specific vulnerabilities for particular periods of time.
+We require two sources of data to determine whether the Android sandbox is secure or not: (1) information on the distribution of installed versions of Android over time and (2) information on the critical vulnerabilities found to affect specific versions of Android.
+These two datasets can then be combined to determine the proportion of sandboxes at risk of attack from specific vulnerabilities for particular periods of time.
 
-Where possible we indicate the uncertainty in our results by presenting them $\pm$ one standard deviation and give results to 3 significant figures, this occasionally results in `$\pm\, 0$' when the standard deviation is small.
+We indicate the uncertainty in our results by presenting them $\pm$ one standard deviation and give results to 3 significant figures, this occasionally results in `$\pm\, 0$' when the standard deviation is small.
 This technique cannot account for systematic errors which we explore in \S\ref{sec:representative}.
@@
 
-Some critical vulnerabilities are not traditional kernel vulnerabilities, for example the discovery of flaws in the verification of signatures on Android applications in February 2013~\cite{Forristal2013} meant that applications could pretend to be signed with system keys and hence gain root equivalent privileges.
-On some versions of Android (below version 4.1) malware could use known system-to-root escalation mechanisms but on all versions they have a greatly increased attack area for further privilege escalation and also have the ability to control all user internet traffic (via VPNs), brick the phone, remove and install apps, steal user credentials, read the screen and make as well as receive calls.
+Some critical vulnerabilities are not traditional kernel vulnerabilities, but are vulnerabilities in level 2 of our threat model.
+For example improper verification of signatures at installation time was discovered in February 2013~\cite{Forristal2013} and meant that applications could pretend to be signed with system keys and from there gain root privileges.
+For example, for some versions of Android (below version 4.1) malware could use known system-to-root escalation mechanisms but on all versions they have a greatly increased attack area for further privilege escalation and also have the ability to control all user internet traffic (via VPNs), brick the phone, remove and install apps, steal user credentials, read the screen and make as well as receive calls.
 The different categories which the vulnerabilities fall into are shown in Table~\ref{tab:andvulns}.
@@
 Frei et al.~\cite{Frei2010} propose the definition of the \textbf{time of disclosure}.
-This occurs when the information about the vulnerability is available to the from an accepted source.
+This occurs when the information about the vulnerability is available to the manufacturer from an accepted source.
 Unfortunately before we collated this information much of it was not published by an accepted source, even months or years after they had been actively used.
@@
 
-Therefore, we use the earliest of: discovery date, date when reported to be exploited, date of disclosure or date of fix in the source code; as the date at which we consider a vulnerability to start being dangerous.
+Therefore, we use the earliest of: discovery date, date reported, date first exploited, date of disclosure or date of fix in the source code; as the date at which we consider a vulnerability to start being dangerous.
 A breakdown of the type of date used is shown in Table~\ref{tab:andvulns}.
@@
 \end{figure}
-In this analysis we calculate the proportion of Android devices which are exposed to critical vulnerabilities and and discuss how that has changed over time.
-In Figure~\ref{fig:proportioninsecure} describes the proportion of Android devices susceptible to at least one critical vulnerability.
+In this analysis we calculate the proportion of Android devices which are exposed to critical vulnerabilities and discuss how that has changed over time.
+Figure~\ref{fig:proportioninsecure} describes the proportion of Android devices susceptible to at least one critical vulnerability.
 
@@
 This demonstrates the importance of a longitudinal study: this type of analysis requires years of data.
-Once \avovuln{zergRush} was discovered in October 2011 the most devices are recorded as \emph{insecure} (red) as they were exposed to that vulnerability.
+Once \avovuln{zergRush} was discovered in October 2011 then most devices are recorded as \emph{insecure} (red) as they were exposed to that vulnerability.
 The remaining devices were already running a version of Android which fixed the \avovuln{zergRush} vulnerability and are therefore marked as \emph{secure} (green).
@@
 In 2013 three vulnerabilities were found in the way which Android verified the signatures on APKs.
-These allowed the creation of malicious APKs which appear to be signed as system APKs -- which have root equivalent privileges.
+These are level 2 vulnerabilities in our threat model since they require a new app installation to occur.
 Figure~\ref{fig:nvulnerabilities_heat} shows how the the \emph{APK signing vulnerabilities} affected all devices and took months to get fixed for any device.
-However what is perhaps more worrying is the long tail on the \avovuln{Gingerbreak}, \avovuln{levitator}, \avovuln[exploid_udev]{exploid} and \avovuln{zergRush} vulnerabilities which are more dangerous root vulnerabilities (not requiring new APK installation) and which affect a significant proportion of devices years later.
-\todo{quantify this long tail and the APK numbers as headline numbers}
+However what is perhaps more worrying is the long tail on the \avovuln{Gingerbreak}, \avovuln{levitator}, \avovuln[exploid_udev]{exploid} and \avovuln{zergRush} vulnerabilities which are more dangerous level 3 vulnerabilities (since they do not require new app installation) and which affect a significant proportion of devices years later.
+%\todo{quantify this long tail and the APK numbers as headline numbers}
 
@@
 
-{\begin{figure}[p]
+{
+\begin{figure}[t]
 \centering
- \begin{subfigure}[b]{\columnwidth}
+\includegraphics[width=\columnwidth]{figures/device-data-all-security}
+\caption{The top \daNumDeviceDataDevices\ devices by days of contribution in the \da\ data.
+One strip per device handset and each strip shows the number of vulnerabilities affecting each device over time.}
+\label{fig:device_data_security}
+\end{figure}
+}
+
+{
+\begin{figure}[t]
+\centering
 \includegraphics[width=\columnwidth]{figures/device-data-all-os}
-  \caption{OS versions over time. Black lines show where only the build number changed.}
+\caption{The top \daNumDeviceDataDevices\ devices by days of contribution in the \da\ data.
+One strip per device handset and each strip shows OS versions over time.
+Black lines show where an update only changed the build number.}
 \label{fig:device_data_os}
- \end{subfigure}
- \begin{subfigure}[b]{\columnwidth}
-  \includegraphics[width=\columnwidth]{figures/device-data-all-security}
-  \caption{Number of vulnerabilities affecting each device over time}
-  \label{fig:device_data_security}
- \end{subfigure}
- \caption{The top \daNumDeviceDataDevices\ devices by days of contribution in the \da\ data. One strip per device handset.}
-  \includegraphics[width=\columnwidth]{figures/nw_security_updates.pdf}
- \caption{Proportion of updates each week which (may have) fixed vulnerabilities}
- \label{fig:weekly_security_updates}
-\end{figure}}
+\end{figure}
+}
+
+%{\begin{figure}[p]
+% \centering
+% \begin{subfigure}[b]{\columnwidth}
+%  \includegraphics[width=\columnwidth]{figures/device-data-all-os}
+%  \caption{OS versions over time. Black lines show where only the build number changed.}
+%  \label{fig:device_data_os}
+% \end{subfigure}
+% ~\\
+% \begin{subfigure}[b]{\columnwidth}
+%  \includegraphics[width=\columnwidth]{figures/device-data-all-security}
+%  \caption{Number of vulnerabilities affecting each device over time}
+%  \label{fig:device_data_security}
+% \end{subfigure}
+% \caption{The top \daNumDeviceDataDevices\ devices by days of contribution in the \da\ data. One strip per device handset.}
+ % \includegraphics[width=\columnwidth]{figures/nw_security_updates.pdf}
+ %\caption{Proportion of updates each week which (may have) fixed vulnerabilities}
+ %\label{fig:weekly_security_updates}
+%\end{figure}}
 
@@
 Instead we have a hopefully representative sample of updates which happened while \da\ was installed on the devices.
-\da\ cannot distinguish between a device being replaced and the \da\ app being removed as while multiple devices can be linked to the same user account, that is a manual process and few users do it.
+\da\ cannot distinguish between a device being replaced and the \da\ app being removed as while multiple devices can be linked to the same user account, this is a manual process and consequently few users do it.
 
@@
 
-The number of devices getting security updates each week, is shown in Figure~\ref{fig:weekly_security_updates}.
-Updates which changed the Android version number so that the number of known vulnerabilities decreased are shown in red.
-Updates which changed the build number but not the version number and so might contain a backported fix for a vulnerability are shown in yellow.
-Updates which did not fix security vulnerabilities (because there were no known security vulnerabilities in the existing version of Android) are shown in green.
+%The number of devices getting security updates each week, is shown in Figure~\ref{fig:weekly_security_updates}.
+%Updates which changed the Android version number so that the number of known vulnerabilities decreased are shown in red.
+%Updates which changed the build number but not the version number and so might contain a backported fix for a vulnerability are shown in yellow.
+%Updates which did not fix security vulnerabilities (because there were no known security vulnerabilities in the existing version of Android) are shown in green.
 
@@
 One study was conducted in Norway with 654 installs and one in Bangladesh where 2463 users installed \da.
-This latter study is responsible for the sudden change in distribution of OS versions in \da\ beginning in August 2014 in Figures \ref{fig:norm_os}, \ref{fig:nvulnerabilities_heat} and \ref{fig:proportioninsecure} as for that period Bangladeshi users contributed over half of the \da\ data.
+This latter study is responsible for the sudden change in distribution of OS versions in \da\ beginning in August 2014 in Figures \ref{fig:proportioninsecure}, \ref{fig:norm_os} and \ref{fig:nvulnerabilities_heat} as for that period Bangladeshi users contributed over half of the \da\ data.
 %We could normalise out the influence of the Bangladeshi users, however they are Android users too and they also need security updates.
@@
 \label{sec:conclusion}
-We have investigated the security of the Android operating system with respect to security updates.
-Android attempts to provide good security to end users through sandboxing apps.
+We have described the three levels of defence provided to Android users today and determined that the app sandbox is the best place to prevent malware from gaining root.
 However the latency in the security update process means that on average \daMeanInsecurityPerc\ of Android devices are exposed to known critical vulnerabilities which break out of the sandbox and only \daUpdatednessPerc\ of devices run the latest version of Android.
 Devices apply \daUpdatesPerYear\ updates each year less than the rate of critical vulnerability discovery of between \avoVulnsPerYearAllAndroid\ and \avoVulnsPerYear.
+
+Installing regular updates is the best defence against malware, however, as we have shown, these updates are rarely forthcoming. 
+Therefore Android users need to take additional steps to protect themselves. 
+Firstly, users should only install apps from the Google Play store, which previous research has been shown to have a lower level of malware than third-party markets and is known to contain static and dynamic analysis of available apps.
+Secondly, users should consider installing an anti-virus tool or enabling the Google verify apps feature if available. 
+Thirdly, users should avoid using untrusted WiFi networks or use them in conjunction with a VPN tunnel to a trusted network. 
+Finally, if the user is running a device with API level 17 or above and is technically able to determine the target API version of their apps, they should consider uninstalling all apps with a target API of less than 17.
+
+In the long term, there needs to be greater emphasis on providing timely updates.
+This could be fixed through a variety of mechanisms.
+Increased regulation is one potential solution.
+Currently manufacturers do not even provide updates for the length of the contracts under which the users originally bought the phone, despite efforts by the ACLU to persuade the FTC to force them to do so~\cite{Soghoian2013}, and despite the FTC forcing HTC to do so~\cite{FTCHTC2013}.
+
+Alternatively, decentralised management might work.
+The claim of the open source software `bazzar' is that it is more secure because anyone can review the source code and provide fixes~\cite{Raymond1999}. 
+On the surface, Android is an open source project, and therefore others should be able to step in when the manufacturer fails to provide timely updates.
+Unfortunately only manufacturers can provide updates today because the complete source code for the (often modified) operating system, as well as the drivers and the build environment, are not available for many devices.
+This is unlikely to change since there are economic disincentives from the manufacturers' perspective, and in any case there is a trust issue which needs to be solved: how does the user determine whether a binary from an alternative supplier is more secure than the one they already have installed?
+
+The fundamental problem is really one of a misalignment of incentives in the current Android ecosystem: there are too few reasons why manufacturers should provide timely security updates.
+A more rigid vertically integrated marked would likely improve the security of Android at the cost of some of the freedom it current affords manufacturers and users.
 

### 2015-03-31 securityupdates.tex
Changes to the introduction which somehow never got committed (upcoming painful merge conflict)

--- 2015-03-25 securityupdates.tex
@@
 We show that the security of Android relies on the centralisation of app deployment with Google.
-Computer systems which run untrusted code are considered secure if there are strong sandboxing mechanisms which prevent the untrusted code doing dangerous things.
-However we show the Android ecosystem is secure even though on average \daMeanInsecurityPercTwosfNominal\ of devices are exposed to known critical vulnerabilities which break the sandboxing mechanisms.
-The recurrent advice on keeping computers secure is to install the latest updates yet only \daUpdatednessPercTwosfNominal\ of Android devices run the latest version and the rate of critical vulnerability discovery of between \avoVulnsPerYearAllAndroidTwosfNominal\ and \avoVulnsPerYearTwosfNominal\ per year is higher than the rate of updates being delivered of \daUpdatesPerYearTwosfNominal.
-Yet Android malware deployments are not widespread despite the many attempts of malware authors:
-in 2012, between 36.7\%~\cite{Zhou2012b} and 40\%~\cite{Zhou2012a} of malware for Android contained exploits designed to enable a malicious app to break the sandbox and thousands of new malware samples are found each year.
+Computer systems which run untrusted code are secure if there are strong sandboxing mechanisms which prevent the untrusted code doing dangerous things.
+Google claims 100\% of Android devices have sandboxes and permissions~\cite{Patterson2013} but we show that on average only \daMeanSecurityPercTwosfNominal\ of Android devices have \emph{working} sandboxes and permissions due to known critical vulnerabilities.
+The recurrent advice on keeping computers secure is to install the latest updates, yet only \daUpdatednessPercTwosfNominal\ of Android devices run the latest version and the rate of critical vulnerability discovery of between \avoVulnsPerYearAllAndroidTwosfNominal\ and \avoVulnsPerYearTwosfNominal\ per year is higher than the rate of updates being delivered of \daUpdatesPerYearTwosfNominal.
+Yet Android malware deployments are not widespread -- despite the many attempts of malware authors:
+in 2012, between 36.7\%~\cite{Zhou2012b} and 40\%~\cite{Zhou2012a} of malware for Android contained exploits designed to enable a malicious app to break the sandbox and thousands of new malware samples are found each year\todolater{citation needed}.
 
@@
 Some users may only be able to use third party app stores or have USB debugging enabled and connect their phone to a malicious machine.
-There have also been two remote code execution vulnerabilities in Android, one in the Javascript-to-Java bridge and one in dhcpd which allow a network attacker to run malicious code in a user context, bypassing all the checks on apps and leaving only the sandbox which we show to be vulnerable.
+There have also been two remote code execution vulnerabilities in Android, one in the Javascript-to-Java bridge and one in dhcpd which allow a network attacker to run malicious code in a user context, bypassing all the checks on apps and leaving only the sandbox, which we show to be vulnerable.
 
@@
 For example, the Google Play marketplace uses Bouncer to automatically analyse apps as well as removing any apps which are reported as malicious.
-Nevertheless, Google Play is not immune: 0.02\% of apps are malicious~\cite{Zhou2012a}.
+Nevertheless, Google Play is not immune: in 2012 0.02\% of apps were malicious~\cite{Zhou2012a}.
 Alternative markets are also popular, particularly in countries such as China where the Google Play is not available. 
@@
 
-Support for dynamic code loading on Android also supports the third attack vector in which the attacker injects malicious code directly into existing apps already installed on the handset. 
+The third attack vector is that the attacker injects malicious code directly into existing apps or programs already installed on the handset. 
 For example, the addJavascriptInterface vulnerability (CVE-2012-6636) allows Javascript running in an Android WebView to execute arbitrary code in the same process address space as the vulnerable app.

### 2015-03-31 securityupdates.tex
add a possible todo about 95% confidence intervals

@@
 
-Where possible we indicate the uncertainty in our results by presenting them $\pm$ one standard deviation and give results to 3 significant figures, this occasionally results in `$\pm\, 0$' when the standard deviation is small.
+Where possible we indicate the uncertainty in our results by presenting them $\pm$ one standard deviation and give results to 3 significant figures, this occasionally results in `$\pm\, 0$' when the standard deviation is small.\todolater{Do we want to use the 95 percentile instead}
 This technique cannot account for systematic errors which we explore in \S\ref{sec:representative}.

### 2015-03-31 securityupdates.tex
Minor fixes

@@
 In this analysis we calculate the proportion of Android devices which are exposed to critical vulnerabilities and and discuss how that has changed over time.
-In Figure~\ref{fig:proportioninsecure} describes the proportion of Android devices susceptible to at least one critical vulnerability.
+Figure~\ref{fig:proportioninsecure} describes the proportion of Android devices susceptible to at least one critical vulnerability.
 
@@
 The trend that we saw in Figure~\ref{fig:proportioninsecure} of security improving and then getting worse is also shown here.
-It shows how some devices had vulnerabilities, which were fixed, and then further vulnerabilities were discovered, mostly not fixed.
+It shows how some devices had vulnerabilities, which were fixed, and then further vulnerabilities were discovered, and mostly not fixed.
 This implies that these devices had been abandoned by the device manufacturer and were not receiving updates, this is confirmed by Figure~\ref{fig:device_data_os} which shows which OS versions those devices were running.

### 2015-03-31 securityupdates.tex
Note that if build numbers do not change for updated code this indicates the manufacturer is incompetent

@@
 \section{Introduction}
-We show that the security of Android relies on the centralisation of app deployment with Google.
-Computer systems which run untrusted code are secure if there are strong sandboxing mechanisms which prevent the untrusted code doing dangerous things.
-Google claims 100\% of Android devices have sandboxes and permissions~\cite{Patterson2013} but we show that on average only \daMeanSecurityPercTwosfNominal\ of Android devices have \emph{working} sandboxes and permissions due to known critical vulnerabilities.
-The recurrent advice on keeping computers secure is to install the latest updates, yet only \daUpdatednessPercTwosfNominal\ of Android devices run the latest version and the rate of critical vulnerability discovery of between \avoVulnsPerYearAllAndroidTwosfNominal\ and \avoVulnsPerYearTwosfNominal\ per year is higher than the rate of updates being delivered of \daUpdatesPerYearTwosfNominal.
-Yet Android malware deployments are not widespread -- despite the many attempts of malware authors:
-in 2012, between 36.7\%~\cite{Zhou2012b} and 40\%~\cite{Zhou2012a} of malware for Android contained exploits designed to enable a malicious app to break the sandbox and thousands of new malware samples are found each year\todolater{citation needed}.
 
-Android is protected by many layers of security and while the sandbox is usually broken on most Android devices by known vulnerabilities, the outer layers of protection such as the Google Play store prevent the installation of malicious apps.
-The Google Play store provides economic (app developers have to pay to register), technical (Google Bouncer which scans apps) and social (user reviews and reports of malicious software) measures to prevent malicious apps being distributed.
-Even users who cannot use Google Play (e.g.\ because they are in China) may still get its `Verify apps' functionality which scans for malicious apps.
+Modern mobile operating systems provide several layers of security whose combined aim is to prevent app malware from taking control of a device. 
+In this paper we explore the security layers present in the Android ecosystem and determine which layers are effective and which are not.
 
-The claim of the open source software `bazzar' is that it is more secure because anyone can review the source code and provide fixes~\cite{Raymond1999}.
-Yet Android is not a `bazzar' in the way that Debian is because while much of the code is open source and available for the many eyes review, end users cannot easily rebuild the binaries that run on their phones due to the binary blobs and closed source components within them.
-Since only manufacturers can do that, rather than any one technically competent user, the availability of updates is controlled by the commercial interests of the manufacturers (and the operators).
-Manufacturers do not even provide updates for the length of the contracts under which the users originally bought the phone, despite efforts by the ACLU to persuade the FTC to force them to do so~\cite{Soghoian2013}, and despite the FTC forcing HTC to do so~\cite{FTCHTC2013}.
-Since only manufacturers can provide updates, malicious parties cannot do so, if manufacturers promptly provided updates this would improve security, but since they do not, it makes it worse.
+Our focus is on malware whose aim is to gain full control of the device rather than merely abuse permissions granted to it by the user.
+In this context, by full control, or \emph{root}, on a device we mean the malware has unfettered access to all data stored on the device and can control all network interfaces in the device with the same flexibility as the operating system itself.
+We have selected our focus for two reasons: no previous published work has quantified how vulnerable Android devices are to malware gaining root, and secondly the negative impact on the device owner is much higher for malware which gains root than malware which does not: malware with root can not only perform arbitrary actions on the device, but it is also impossible for the device owner to remove the malware reliably since even a factory reset does not guarantee that the malware has been successfully removed.
 
-However, despite the current low level of malicious software deployment on Android, the high level of vulnerability and the slow deployment of updates is still a cause for concern.
-Some users may only be able to use third party app stores or have USB debugging enabled and connect their phone to a malicious machine.
-There have also been two remote code execution vulnerabilities in Android, one in the Javascript-to-Java bridge and one in dhcpd which allow a network attacker to run malicious code in a user context, bypassing all the checks on apps and leaving only the sandbox, which we show to be vulnerable.
+This is an important problem: previous studies have shown that a significant proportion of malware attempts to gain root.
+For example, in 2012 between 36.7\%~\cite{Zhou2012b} and 40\%~\cite{Zhou2012a} of malware for Android contained exploits designed to gain root-level access.
 
-Not all malicious apps need to break the sandbox in order to misbehave.
-We focus on the issue of apps breaking out of the sandbox because malware which is able to take full control of a handset can do significantly more harm and is much harder to remove.
-Ordinary malware can be uninstalled by the user, or remotely by Google through the Play Store, and returns the device to a secure state. 
-Removing an app which has used an exploit to take control of the handset is unlikely to return the device to secure state.
-We describe our threat model and three attack vectors in detail in \S\ref{sec:threatmodel}.
+Android users are protected from malicious apps by up to three levels of defence. 
+The first level of defence comes from an app marketplace such as Google Play. 
+A marketplace can use three mechanisms to prevent malicious apps from appearing in the marketplace: firstly by economic disincentives, including charging developers to create an account and cancelling accounts used to upload malicious apps; secondly by technical means, such as static and dynamic analysis on submitted apps; and finally via social feedback between users, including the provision of user reviews and ratings.
 
-In order to determine whether an individual smartphone is exposed to a particular vulnerability, we need to combine vulnerability data with the operating system version string and build number from the handset.
-For this we use handset data gathered from \daNumOSDevices\ devices in the \da~\cite{Wagner2013} project.
-From this we have determined that since \daStartDate\ on five occasions all Android devices have been exposed to known critical vulnerabilities and that it takes a year for updates to reach most devices, by which time new vulnerabilities have been discovered.
+The second line of defence occurs on the handset at installation time, and includes checking the signature of any app installation or update as well as verifying the app by local anti-virus tools (e.g.
+Google's app verification feature).
+Finally, the third line of defence takes place at runtime: Android apps are run in a \emph{sandbox} which prevents the app from unauthorised actions such as accessing any permissions which weren't granted by the user at install time; in addition, anti-virus products may scan the list of installed apps periodically to detect malware.
 
-In summary, the contributions of this paper are:
+Sandboxes only provide protection if they are free from design errors and implementation flaws. 
+Both of these have occurred in Android, as indeed happens today in many large security systems relying on the sandbox model. 
+Maintaining sandbox security therefore relies on providing timely updates to fix design and implementation flaws as and when they are discovered.
+
+In this paper we show that the Android sandbox only provides effective security on \daMeanSecurityPercTwosfNominal\ of devices today.
+\footnote{In contrast to Google's claims that 100\% of Android devices are protected by sandboxes and permissions~\cite{Patterson2013}}
+This is because most devices do not receive timely updates. 
+Therefore the security of the Android ecosystem is currently reliant on the defences provided by the app marketplace and (when enabled) anti-virus systems. 
+This insight allows us to provide users with a list of concrete actions to improve device security. 
+It also raises broader questions about how modern mobile device security is most effectively achieved, how a secure ecosystem should be construct and managed, and whether security by obscurity is the right way forward.
+
+In summary, our contributions are:
 \begin{itemize}
- \item We show that the security of Android relies on the centralisation of app deployment with Google rather than the absence of vulnerabilities in its sandbox
  \item We quantify the Android update ecosystem showing how the OS version distribution changes and the frequency of updates.
- \item We quantify vulnerabilities by determining what proportion of devices they affected and how that changes over time.
- \item We created an open database to collate quantitative information on vulnerabilities which affect Android, and investigate their time-lines.
+ \item We created an open database to collate quantitative information on vulnerabilities which affect Android and use this to determine the proportion of devices they affect and how that changes over time.
+ \item We demonstrate that the security of Android relies on the centralisation of app deployment via a marketplace, and the use of install-time anti-virus checks rather than through the security of the app sandbox.
+ \item We use our analysis to develop advice for Android users on what steps they should take to increase the security of their device.
+
 \end{itemize}
 
-\section{Threat model}
+\section{Threat model and attack vectors}
 \label{sec:threatmodel}
 
-In this paper we measure whether an attacker can take control of a handset by breaking out of the app sandbox on Android using a known vulnerability.
-We consider three attack vectors.
+In this paper we assume the defender can deploy security controls at three levels: in an online marketplace, at app installation, and during app execution.
+We then consider three attack vectors an attacker can use to attempt to gain root on an Android handset.
+
 The first attack vector is through the installation of a malicious app on the device.
 Android devices support the installation of apps through a number of app marketplaces, email attachments, URLs and via the Android Debug Bridge (ADB).
-By default, many Android devices with Google Play installed will prevent the installation of apps from other sources.
-
-Controlling potential sources of app installation is a good security feature because app marketplaces provide additional protection to prevent the listing of malicious apps.
-For example, the Google Play marketplace uses Bouncer to automatically analyse apps as well as removing any apps which are reported as malicious.
-Nevertheless, Google Play is not immune: in 2012 0.02\% of apps were malicious~\cite{Zhou2012a}.
+By default, many Android devices with Google Play installed will prevent the installation of apps from other sources, and therefore an attacker needs to compromise all three levels of security.
+The Google Play marketplace uses Bouncer to automatically analyse apps as well as removing any apps which are reported as malicious.
+Nevertheless, Google Play is not immune: in 0.02\% of apps were malicious~\cite{Zhou2012a}.
 Alternative markets are also popular, particularly in countries such as China where the Google Play is not available. 
@@
 The most direct method available to an attacker is to upload to a marketplace a seemingly innocent app which contains support for dynamic code loading from a remote server under control of the attacker.
-Neither static nor dynamic analysis of this app by the marketplace will uncover any malicious code, since it does exist in the app binary.
-Instead the attacker can choose to send a malicious payload directly to the app for execution at a later point in time.
-Note that dynamic code loading is not actually needed for this to work -- it just makes the task easier -- since previous work has shown that even on a platform such as iOS, which does not permit dynamic code loading, a Return oriented Programming (ROP)-based attack is relatively easy if the attacker creates an app with carefully crafted flaws~\cite{Wang2013a}.
+Neither static nor dynamic analysis of this app by the marketplace will uncover any malicious code, since it does not exist in the app binary.
+The attacker can choose to send a malicious payload directly to the app for execution at a later point in time.
+The marketplace can detect explicit use of dynamic code loading, even if malicious code cannot be detected.
+Note that dynamic code loading is not actually needed for this to work however -- it just makes the task easier -- since previous work has shown that even on a platform such as iOS, which does not permit dynamic code loading, a Return-Oriented Programming (ROP)-based attack is relatively easy if the attacker creates an app with carefully crafted flaws~\cite{Wang2013a}.
 
 The third attack vector is that the attacker injects malicious code directly into existing apps or programs already installed on the handset. 
-For example, the addJavascriptInterface vulnerability (CVE-2012-6636) allows Javascript running in an Android WebView to execute arbitrary code in the same process address space as the vulnerable app.
-This attack works when: (1) the WebView has the Javascript-to-Java bridge enabled; and (2) the attacker is able to inject a malicious Javascript payload into the network traffic between the phone and the web server.
-Many apps are vulnerable to this type of attack because ad-libraries typically embed adverts into apps by displaying HTML and Javascript inside a WebView with the Javascript-to-Java bridge enabled. 
+For example, the addJavascriptInterface vulnerability (CVE-2012-6636) allows JavaScript running in an Android WebView to execute arbitrary code in the same process address space as the vulnerable app.
+This attack works when: (1) the WebView has the JavaScript-to-Java bridge enabled; and (2) the attacker is able to inject a malicious JavaScript payload into the network traffic between the phone and the web server.
+Many apps are vulnerable to this type of attack because ad-libraries typically embed adverts into apps by displaying HTML and JavaScript inside a WebView with the JavaScript-to-Java bridge enabled. 
 The fix for the addJavascriptInterface vulnerability breaks backwards compatibility, and therefore apps are only immune to this attack if their target API level is greater than 16 and they are running on an Android device with an OS version of 4.2 or greater.
-Our analysis shows that on \daGPAPISeventeenLaterDate, \daGPAPISeventeenEarlierProportion\ of handsets connecting to the Google Play store were still vulnerable to this attack.
+Our analysis shows that as of \daGPAPISeventeenLaterDate, \daGPAPISeventeenEarlierProportion\ of handsets connecting to the Google Play store were still vulnerable to this attack.
+\todo{Do we want to talk about the dhcp vulnerability here?}
+
+In summary, the best place to prevent attacks is at runtime, since all three attack vectors can be prevented at this level. 
+Unfortunately, as we will demonstrate, the sandbox for Android apps is weak and anti-virus is either not installed or not activated by default.
+Therefore many Android users implicitly rely on the marketplace for protection by removing apps which are detected as malicious.
+This detection can only be derived from static and dynamic analysis or after reports of malicious behaviour from users once it is made available for download.
 
@@
 
-We require two sources of data for this study: (1) information on the distribution of installed versions of Android over time and (2) information on the critical vulnerabilities found to affect specific versions of Android.
-These two datasets can then be combined to determine the proportion of handsets at risk of attack from specific vulnerabilities for particular periods of time.
+We require two sources of data to determine whether the Android sandbox is secure or not: (1) information on the distribution of installed versions of Android over time and (2) information on the critical vulnerabilities found to affect specific versions of Android.
+These two datasets can then be combined to determine the proportion of sandboxes at risk of attack from specific vulnerabilities for particular periods of time.
 
-Where possible we indicate the uncertainty in our results by presenting them $\pm$ one standard deviation and give results to 3 significant figures, this occasionally results in `$\pm\, 0$' when the standard deviation is small.\todolater{Do we want to use the 95 percentile instead}
+We indicate the uncertainty in our results by presenting them $\pm$ one standard deviation and give results to 3 significant figures, this occasionally results in `$\pm\, 0$' when the standard deviation is small.
+\todolater{Do we want to use the 95 percentile instead}
 This technique cannot account for systematic errors which we explore in \S\ref{sec:representative}.
@@
 \end{figure}
-In this analysis we calculate the proportion of Android devices which are exposed to critical vulnerabilities and and discuss how that has changed over time.
+In this analysis we calculate the proportion of Android devices which are exposed to critical vulnerabilities and discuss how that has changed over time.
 Figure~\ref{fig:proportioninsecure} describes the proportion of Android devices susceptible to at least one critical vulnerability.
@@
 This is not enough to build a statistically significant prediction of the expected transition to newer versions of Android.
+If manufacturers shipped security updates without changing the build number then we would not be able to detect this, but that would also indicate a level of incompetence likely to damage security and so can be discounted.
 

### 2015-04-01 securityupdates.tex
minor fixes

@@
 This insight allows us to provide users with a list of concrete actions to improve device security. 
-It also raises broader questions about how modern mobile device security is most effectively achieved, how a secure ecosystem should be construct and managed, and whether security by obscurity is the right way forward.
+It also raises broader questions about how modern mobile device security is most effectively achieved, how a secure ecosystem should be constructed and managed, and whether security by obscurity is the right way forward.
 
@@
 By default, many Android devices with Google Play installed will prevent the installation of apps from other sources, and therefore an attacker needs to compromise all three levels of security.
-The Google Play marketplace uses Bouncer to automatically analyse apps as well as removing any apps which are reported as malicious.
-Nevertheless, Google Play is not immune: in 0.02\% of apps were malicious~\cite{Zhou2012a}.
+The Google Play marketplace uses Bouncer to automatically analyse apps; any apps which are reported as malicious are removed.
+Nevertheless, Google Play is not immune: in 2012 0.02\% of apps were malicious~\cite{Zhou2012a}.
 Alternative markets are also popular, particularly in countries such as China where the Google Play is not available. 
@@
 On a large proportion of devices `opaque\_marker' is a well defined build number\footnote{\url{https://source.android.com/source/build-numbers.html}} however different device manufacturers use different schema.
-Google provides API version distribution information but not the OS and build version information we need, we verify that the \da\ data is representative in~\S\ref{sec:representative}.
+Google provides API version distribution information but not the OS and build version information we need. We verify that the \da\ data is representative in~\S\ref{sec:representative}.
 
@@
 In addition, we have received submissions or amendments from \avoNumSubmitters\ individuals.
-We collected data between \avoStartDate\ and \avoEndDate\ and so it does not include any information lost before the start of that period.
+We collected data between \avoStartDate\ and \avoEndDate.
+Our list therefore does not include any information before the start of that period.
 \avo\ currently contains \avoNumVulnerabilities\ vulnerabilities of which \avoNumVulnAllAndroid\ affect all Android devices and \avoNumVulnSpecific\ are specific to particular devices or device manufacturers.
@@
 However most devices only contribute for a short period of time and so we do not observe updates happening on every device.
-Instead we have a hopefully representative sample of updates which happened while \da\ was installed on the devices.
-\da\ cannot distinguish between a device being replaced and the \da\ app being removed as while multiple devices can be linked to the same user account, this is a manual process and consequently few users do it.
+investigate updates which were installed on devices that contributed data continuously over long periods of time.
 
@@
 Security in depth is also a useful strategy.
-In this regard, iOS provides additional safeguards beyond those used in Android, including a pre-distribution code review, mandatory code-signing by the manufacturer, and (with the important exception of ROP-based attacks~\cite{Wang2013a}) the technical prohibition of dynamic code loading by an app.
+In this regard, iOS provides additional safeguards beyond those used in Android, including a pre-distribution review, mandatory code-signing by the manufacturer, and (with the important exception of ROP-based attacks~\cite{Wang2013a}) the technical prohibition of dynamic code loading by an app.
 These features, as well as Address Space Layout Randomisation (ASLR) and mandatory access controls, has resulted in a lower level of malware affecting iOS when compared to Android~\cite{Felt2011}.
@@
 Alternatively, decentralised management might work.
-The claim of the open source software `bazzar' is that it is more secure because anyone can review the source code and provide fixes~\cite{Raymond1999}. 
+The claim of the open source software `bazaar' is that it is more secure because anyone can review the source code and provide fixes~\cite{Raymond1999}.
 On the surface, Android is an open source project, and therefore others should be able to step in when the manufacturer fails to provide timely updates.

### 2015-04-02 securityupdates.tex
Correct a couple of Daniel Wagner's minor fixes which were not strictly accurate

@@
 We collected data between \avoStartDate\ and \avoEndDate.
-Our list therefore does not include any information before the start of that period.
+Our list therefore does not include any information lost before the start of that period.
 \avo\ currently contains \avoNumVulnerabilities\ vulnerabilities of which \avoNumVulnAllAndroid\ affect all Android devices and \avoNumVulnSpecific\ are specific to particular devices or device manufacturers.
@@
 However most devices only contribute for a short period of time and so we do not observe updates happening on every device.
-investigate updates which were installed on devices that contributed data continuously over long periods of time.
+Instead we have a hopefully representative sample of updates which happened while \da\ was installed on the devices.
+\da\ cannot distinguish between a device being replaced and the \da\ app being removed as while multiple devices can be linked to the same user account, this is a manual process and consequently few users do it.
 

### 2015-04-02 securityupdates.tex
Manually make device data plots more efficient and put beneath each other on the same page to save space

@@
 {
-\begin{figure}[t]
+\begin{figure}[!t]
 \centering
+\begin{subfigure}{\columnwidth}
 \includegraphics[width=\columnwidth]{figures/device-data-all-security}
-\caption{The top \daNumDeviceDataDevices\ devices by days of contribution in the \da\ data.
-One strip per device handset and each strip shows the number of vulnerabilities affecting each device over time.}
+\vspace{-1.5em}
+\caption{Each strip shows the number of vulnerabilities affecting each device over time.}
 \label{fig:device_data_security}
+\end{subfigure}
+\begin{subfigure}{\columnwidth}
+\includegraphics[width=\columnwidth]{figures/device-data-all-os}
+\vspace{-1.5em}
+\caption{Each strip shows OS versions over time. Black lines show where an update only changed the build number.}
+\label{fig:device_data_os}
+\end{subfigure}
+\caption{The top \daNumDeviceDataDevices\ devices by days of contribution in the \da\ data. One strip per device handset.}
 \end{figure}
 }
-
-{
-\begin{figure}[t]
-\centering
-\includegraphics[width=\columnwidth]{figures/device-data-all-os}
-\caption{The top \daNumDeviceDataDevices\ devices by days of contribution in the \da\ data.
-One strip per device handset and each strip shows OS versions over time.
-Black lines show where an update only changed the build number.}
-\label{fig:device_data_os}
-\end{figure}
-}
-
-%{\begin{figure}[p]
-% \centering
-% \begin{subfigure}[b]{\columnwidth}
-%  \includegraphics[width=\columnwidth]{figures/device-data-all-os}
-%  \caption{OS versions over time. Black lines show where only the build number changed.}
-%  \label{fig:device_data_os}
-% \end{subfigure}
-% ~\\
-% \begin{subfigure}[b]{\columnwidth}
-%  \includegraphics[width=\columnwidth]{figures/device-data-all-security}
-%  \caption{Number of vulnerabilities affecting each device over time}
-%  \label{fig:device_data_security}
-% \end{subfigure}
-% \caption{The top \daNumDeviceDataDevices\ devices by days of contribution in the \da\ data. One strip per device handset.}
- % \includegraphics[width=\columnwidth]{figures/nw_security_updates.pdf}
- %\caption{Proportion of updates each week which (may have) fixed vulnerabilities}
- %\label{fig:weekly_security_updates}
-%\end{figure}}
 

### 2015-04-02 securityupdates.tex
Change the title
Updates to abstract and contributions from dtw30

@@
 \begin{document}
-\title{The thin blue line:\\what protects my Android device?}
+\title{Only the store can save you now:\\updates and vulnerabilities on Android}
 
@@
 Modern smartphone operating systems protect users from malicious applications by providing a protected execution environment or sandbox.
-This approach relies on the absence of exploits which allow apps to break out of the sandbox.
-Many such exploits have been found, and therefore the security of a smartphone operating system relies on the prompt delivery of security updates.
-In this paper we quantify the extent to which the security sandbox provided by Android protects devices from malicious apps.
+A number of exploits have been found which allow apps to break out of the sandbox.
+We quantify for the first time how vulnerable Android devices in the wild are to critical vulnerabilities and how long it takes for vulnerabilities to be fixed on consumer devices.
 We analysed \da\ data from over \daOSYearsOfData\ years and \daNumOSDevices\ devices and found that on average \daMeanInsecurityPercTwosfNominal\ of devices were exposed to known root privilege vulnerabilities and only \daUpdatednessPercTwosfNominal\ of devices run the most recent version of Android.
@@
  \item We use our analysis to develop advice for Android users on what steps they should take to increase the security of their device.
+ \item We characterise the impact of critical vulnerabilities on the Android ecosystem as a whole, examine the effectiveness of vulnerabilities and how long vulnerabilities remained available in the wild, and also examine the impact on individual devices over extended periods of time.
 

### 2015-04-02 securityupdates.tex
Suggestions from Daniel Wagner

@@
 For example improper verification of signatures at installation time was discovered in February 2013~\cite{Forristal2013} and meant that applications could pretend to be signed with system keys and from there gain root privileges.
-For example, for some versions of Android (below version 4.1) malware could use known system-to-root escalation mechanisms but on all versions they have a greatly increased attack area for further privilege escalation and also have the ability to control all user internet traffic (via VPNs), brick the phone, remove and install apps, steal user credentials, read the screen and make as well as receive calls.
+On versions of Android below 4.1 malware could then use known system-to-root escalation mechanisms.
+Regardless of version, this exposed an increased attack area and would also provide the ability for malware to control all user internet traffic (via VPNs), brick the phone, remove and install apps, steal user credentials and read the screen.
 The different categories which the vulnerabilities fall into are shown in Table~\ref{tab:andvulns}.
@@
 We have chosen these vulnerabilities to fit the attack vectors introduced in \S\ref{sec:threatmodel}.
-These vulnerabilities affect all Android devices regardless of manufacturer, and as a result our selected vulnerabilities will dominate any security analysis of Android.
+These vulnerabilities affect all Android devices regardless of manufacturer, and as a result our selected vulnerabilities will dominate any security analysis of Android.\todo{can we make this sentence clearer?}
 In many cases we could not match manufacturer- and model-specific vulnerabilities to individual devices in the \da\ data and therefore attempting to include device-specific vulnerabilities as well would introduce additional uncertainty in our results.
@@
 We present the results of our analysis showing that, on average, \daMeanInsecurityPerc\ of Android devices are exposed to critical vulnerabilities and only \daUpdatednessPerc\ run the latest version of Android.
-Devices, on average, apply \daUpdatesPerYear\ updates each year less than the rate of critical vulnerability discovery of between \avoVulnsPerYearAllAndroid\ and \avoVulnsPerYear.
+Devices, on average, apply \daUpdatesPerYear\ updates each year, which is less than the rate of critical vulnerability discovery of between \avoVulnsPerYearAllAndroid\ and \avoVulnsPerYear.
 \daPercUpdatesDowngrades\ of full version changes are downgrades to older versions.
@@
 From October 2011 until the discovery of \avovuln[APK_duplicate_file]{APK duplicate file} in February 2013 the graph shows progressive improvement as devices are upgraded.
-This means more and more devices are marked as \emph{secure} because they are now running a secure version of Android, or marked as \emph{maybe secure} because they received a minor update which may have contained a backported fix.
+This means more and more devices are marked as \emph{secure} because they are now running a secure version of Android, or marked as \emph{maybe secure} because they received an OS update which did not update to a known-good version of Android but which may still have included a backport of a fix, as the update was made available after the vulnerability was disclosed.
 From February 2013 onwards regular discovery of critical vulnerabilities ensures that most devices are exposed to known critical vulnerabilities.

### 2015-04-02 securityupdates.tex
Remove spurious words from Acknowledgements

@@
 \section*{Acknowledgements}
-Thanks to Richard Clayton, the mobile security reading group and Anil Madhavapeddy for reading drafts.
-Thanks to David Robertson for advice on statistical analysis.
-Thanks to Laurent Simon, Thomas Coudray, Adrian Taylor, Justin Case, Giant Pune and Khilan Gudka for reporting vulnerabilities in Android.
-This work was partly funded by the EPSRC and Google but the the work presented here does not necessarily reflect their views.
+Thanks to: Richard Clayton, the mobile security reading group and Anil Madhavapeddy for reading drafts;
+David Robertson for statistical advice;
+and Laurent Simon, Thomas Coudray, Adrian Taylor, Justin Case, Giant Pune and Khilan Gudka for reporting vulnerabilities.
+We were funded by the EPSRC and Google but the the work presented here does not necessarily reflect their views.
 }

### 2015-04-02 securityupdates.tex
Submission does not need to be anonymous so we don't need notes to the reviewers

@@
 
-We developed and maintain an open platform for filing critical vulnerabilities in a machine readable format, the \href{http://androidvulnerabilities.org/}{AndroidVulnerabilities.org}\footnote{\textbf{Note to reviewers:} we have removed our names and affiliation from the website for this review \textbf{with the exception of the contact page.
+We developed and maintain an open platform for filing critical vulnerabilities in a machine readable format, the \href{http://androidvulnerabilities.org/}{AndroidVulnerabilities.org}%\footnote{\textbf{Note to reviewers:} we have removed our names and affiliation from the website for this review \textbf{with the exception of the contact page.
 Please do not visit the contact page until the review cycle has completed.}\vspace{-4em}} (\avo) website.

### 2015-04-02 securityupdates.tex
Minor fix

@@
 We have chosen these vulnerabilities to fit the attack vectors introduced in \S\ref{sec:threatmodel}.
-These vulnerabilities affect all Android devices regardless of manufacturer, and as a result our selected vulnerabilities will dominate any security analysis of Android.\todo{can we make this sentence clearer?}
+These vulnerabilities affect all Android devices regardless of manufacturer, and as a result our selected vulnerabilities will dominate any security analysis of Android.
+\todo{can we make this sentence clearer?}
 In many cases we could not match manufacturer- and model-specific vulnerabilities to individual devices in the \da\ data and therefore attempting to include device-specific vulnerabilities as well would introduce additional uncertainty in our results.

### 2015-04-08 securityupdates.tex
Support multiple URLs in references
Cite android security report

@@
 \usepackage{hyperref}
+% http://tex.stackexchange.com/questions/57865/how-to-use-multiple-urls-for-one-bibtex-reference
+\let\URL\url
+\makeatletter
+\def\url#1{\@URL#1||\@nil}
+\def\@URL#1|#2|#3\@nil{%
+  \URL{#1}\ifx\relax#2\relax\else| \URL{#2}\fi}
+\makeatother
+
 \usepackage{subcaption}
@@
 \makeatother
+
 \newcommand{\percMarketShare}{83.6\%~\footnote{\url{http://www.theinquirer.net/inquirer/news/2379036/android-hits-836-percent-marketshare-while-ios-windows-and-blackberry-slide}}}
@@
 
-In this paper we show that the Android sandbox only provides effective security on \daMeanSecurityPercTwosfNominal\ of devices today.
+In this paper we show that the Android sandbox only provides effective security on \daMeanSecurityPercTwosfNominal\ of devices today.%
 \footnote{In contrast to Google's claims that 100\% of Android devices are protected by sandboxes and permissions~\cite{Patterson2013}}
 This is because most devices do not receive timely updates. 
-Therefore the security of the Android ecosystem is currently reliant on the defences provided by the app marketplace and (when enabled) anti-virus systems. 
+Therefore the security of the Android ecosystem is currently reliant on the defences provided by the app marketplace~\cite{AndroidSecurity2014} and (when enabled) anti-virus systems.
 This insight allows us to provide users with a list of concrete actions to improve device security. 

### 2015-04-08 securityupdates.tex
Add SPW paper to bibliography and cite it as the workshop has happend now

@@
 The third attack vector is that the attacker injects malicious code directly into existing apps or programs already installed on the handset. 
-For example, the addJavascriptInterface vulnerability (CVE-2012-6636) allows JavaScript running in an Android WebView to execute arbitrary code in the same process address space as the vulnerable app.
+For example, the addJavascriptInterface vulnerability (CVE-2012-6636) allows JavaScript running in an Android WebView to execute arbitrary code in the same process address space as the vulnerable app~\cite{Thomas2015}.
 This attack works when: (1) the WebView has the JavaScript-to-Java bridge enabled; and (2) the attacker is able to inject a malicious JavaScript payload into the network traffic between the phone and the web server.

### 2015-04-08 securityupdates.tex
Rewording to be slightly shorter

--- 2015-04-08 securityupdates.tex
@@
 \section{Introduction}
-
+\todo{We talk about root a lot again, rather critical vulnerabilities, and root is not strictly accurate}
 Modern mobile operating systems provide several layers of security whose combined aim is to prevent app malware from taking control of a device. 
@@
 In this context, by full control, or \emph{root}, on a device we mean the malware has unfettered access to all data stored on the device and can control all network interfaces in the device with the same flexibility as the operating system itself.
-We have selected our focus for two reasons: no previous published work has quantified how vulnerable Android devices are to malware gaining root, and secondly the negative impact on the device owner is much higher for malware which gains root than malware which does not: malware with root can not only perform arbitrary actions on the device, but it is also impossible for the device owner to remove the malware reliably since even a factory reset does not guarantee that the malware has been successfully removed.
+We have selected our focus for two reasons: no previous published work has quantified how vulnerable Android devices are to malware gaining root, and secondly the negative impact on the device owner is much higher for malware which gains root than malware which does not: as it can not only perform arbitrary actions on the device, but it is also impossible for the device owner to remove the malware reliably as even a factory reset does not guarantee that the malware has been removed.
 
-This is an important problem: previous studies have shown that a significant proportion of malware attempts to gain root.
-For example, in 2012 between 36.7\%~\cite{Zhou2012b} and 40\%~\cite{Zhou2012a} of malware for Android contained exploits designed to gain root-level access.
+This is an important problem: previous studies have shown that a significant proportion of malware attempts to gain root: in 2012, 40\%~\cite{Zhou2012a} of malware for Android contained exploits designed to gain root.
 
@@
 The first level of defence comes from an app marketplace such as Google Play. 
-A marketplace can use three mechanisms to prevent malicious apps from appearing in the marketplace: firstly by economic disincentives, including charging developers to create an account and cancelling accounts used to upload malicious apps; secondly by technical means, such as static and dynamic analysis on submitted apps; and finally via social feedback between users, including the provision of user reviews and ratings.
-
-The second line of defence occurs on the handset at installation time, and includes checking the signature of any app installation or update as well as verifying the app by local anti-virus tools (e.g.
-Google's app verification feature).
+A marketplace can use three mechanisms to prevent malicious apps from appearing in the marketplace: firstly by economic disincentives, including charging developers to create an account and cancelling malicious accounts; secondly by technical means, such as static and dynamic analysis on submitted apps; and finally via social feedback from users, including the provision of reviews, rating and reporting.
+The second line of defence occurs on the handset at installation time, and includes checking the signature of any app installation or update as well as statically verifying the app (e.g.
+Google's Verify Apps feature).
 Finally, the third line of defence takes place at runtime: Android apps are run in a \emph{sandbox} which prevents the app from unauthorised actions such as accessing any permissions which weren't granted by the user at install time; in addition, anti-virus products may scan the list of installed apps periodically to detect malware.
@@
 
-In this paper we show that the Android sandbox only provides effective security on \daMeanSecurityPercTwosfNominal\ of devices today.%
-\footnote{In contrast to Google's claims that 100\% of Android devices are protected by sandboxes and permissions~\cite{Patterson2013}}
+In this paper we show that the Android sandbox only provides effective security on \daMeanSecurityPercTwosfNominal\ of devices.%
+\footnote{In contrast to Google's claims that 100\% of Android devices are protected~\cite{Patterson2013}}
 This is because most devices do not receive timely updates. 
-Therefore the security of the Android ecosystem is currently reliant on the defences provided by the app marketplace~\cite{AndroidSecurity2014} and (when enabled) anti-virus systems.
+Therefore the security of the Android ecosystem is currently reliant on the defences provided by the app marketplace~\cite{AndroidSecurity2014}.
 This insight allows us to provide users with a list of concrete actions to improve device security. 
@@
 
-In this paper we assume the defender can deploy security controls at three levels: in an online marketplace, at app installation, and during app execution.
-We then consider three attack vectors an attacker can use to attempt to gain root on an Android handset.
+We assume the defender can deploy security controls at three levels: in an online marketplace, at app installation, and during app execution.
+Here we consider three attack vectors against Android handsets.
 
 The first attack vector is through the installation of a malicious app on the device.
-Android devices support the installation of apps through a number of app marketplaces, email attachments, URLs and via the Android Debug Bridge (ADB).
+Android devices support the installation of apps through a app marketplaces, email attachments, URLs and via the Android Debug Bridge (ADB).
 By default, many Android devices with Google Play installed will prevent the installation of apps from other sources, and therefore an attacker needs to compromise all three levels of security.
 The Google Play marketplace uses Bouncer to automatically analyse apps; any apps which are reported as malicious are removed.
-Nevertheless, Google Play is not immune: in 2012 0.02\% of apps were malicious~\cite{Zhou2012a}.
 Alternative markets are also popular, particularly in countries such as China where the Google Play is not available. 
-In 2012 between 0.20\% and 0.47\% of apps on alternative markets were malicious~\cite{Zhou2012a}.
+In 2012 0.02\% of apps on Google Play and between 0.20\% and 0.47\% of apps on alternative markets were malicious~\cite{Zhou2012a}.
 
-The second attack vector is to persuade an existing app to download and execute code at runtime.
-The most direct method available to an attacker is to upload to a marketplace a seemingly innocent app which contains support for dynamic code loading from a remote server under control of the attacker.
+The second attack vector is to make an existing app download and execute code at runtime.
+The most direct method available to an attacker is to upload to a marketplace a seemingly innocent app which contains support for dynamic code loading from a malicious server.
 Neither static nor dynamic analysis of this app by the marketplace will uncover any malicious code, since it does not exist in the app binary.
-The attacker can choose to send a malicious payload directly to the app for execution at a later point in time.
+The attacker can choose to send a malicious payload to the app for execution at a later point in time.
 The marketplace can detect explicit use of dynamic code loading, even if malicious code cannot be detected. 
-Note that dynamic code loading is not actually needed for this to work however -- it just makes the task easier -- since previous work has shown that even on a platform such as iOS, which does not permit dynamic code loading, a Return-Oriented Programming (ROP)-based attack is relatively easy if the attacker creates an app with carefully crafted flaws~\cite{Wang2013a}.
+However dynamic code loading is not actually needed for this to work -- it just makes the task easier -- even on a platform such as iOS, which does not permit dynamic code loading, a Return-Oriented Programming (ROP) attack is relatively easy if the attacker creates an app with carefully crafted flaws~\cite{Wang2013a}.
 
 The third attack vector is that the attacker injects malicious code directly into existing apps or programs already installed on the handset. 
-For example, the addJavascriptInterface vulnerability (CVE-2012-6636) allows JavaScript running in an Android WebView to execute arbitrary code in the same process address space as the vulnerable app.
+For example, the addJavascriptInterface vulnerability (CVE-2012-6636) allows JavaScript running in an Android WebView to execute arbitrary code in the same process address space as the vulnerable app~\cite{Thomas2015a}.
 This attack works when: (1) the WebView has the JavaScript-to-Java bridge enabled; and (2) the attacker is able to inject a malicious JavaScript payload into the network traffic between the phone and the web server.

### 2015-04-08 securityupdates.tex
Evil negative vspace
Further rewording for shortening

@@
 % Attempts at fine grained restrictions on running arbitrary code are hampered by updates for security vulnerabilities not reaching users in a timely fashion.
-
+\vspace{-1em}
 \begin{abstract}
@@
 The third attack vector is that the attacker injects malicious code directly into existing apps or programs already installed on the handset. 
-For example, the addJavascriptInterface vulnerability (CVE-2012-6636) allows JavaScript running in an Android WebView to execute arbitrary code in the same process address space as the vulnerable app~\cite{Thomas2015a}.
-This attack works when: (1) the WebView has the JavaScript-to-Java bridge enabled; and (2) the attacker is able to inject a malicious JavaScript payload into the network traffic between the phone and the web server.
-Many apps are vulnerable to this type of attack because ad-libraries typically embed adverts into apps by displaying HTML and JavaScript inside a WebView with the JavaScript-to-Java bridge enabled. 
+For example, the addJavascriptInterface vulnerability (CVE-2012-6636) allows JavaScript running in an Android WebView to execute arbitrary code as the vulnerable app's user~\cite{Thomas2015a}.
 The fix for the addJavascriptInterface vulnerability breaks backwards compatibility, and therefore apps are only immune to this attack if their target API level is greater than 16 and they are running on an Android device with an OS version of 4.2 or greater.
-Our analysis shows that as of \daGPAPISeventeenLaterDate, \daGPAPISeventeenEarlierProportion\ of handsets connecting to the Google Play store were still vulnerable to this attack.
+On \daGPAPISeventeenLaterDate, \daGPAPISeventeenEarlierProportion\ of handsets connecting to Google Play were still vulnerable to this attack.
 \todo{Do we want to talk about the dhcp vulnerability here?}
@@
 In summary, the best place to prevent attacks is at runtime, since all three attack vectors can be prevented at this level. 
-Unfortunately, as we will demonstrate, the sandbox for Android apps is weak and anti-virus is either not installed or not activated by default.
+Unfortunately, as we will demonstrate, the sandbox for Android apps is weak.
 Therefore many Android users implicitly rely on the marketplace for protection by removing apps which are detected as malicious.
@@
 
-We require two sources of data to determine whether the Android sandbox is secure or not: (1) information on the distribution of installed versions of Android over time and (2) information on the critical vulnerabilities found to affect specific versions of Android.
-These two datasets can then be combined to determine the proportion of sandboxes at risk of attack from specific vulnerabilities for particular periods of time.
+We use two sources of data to determine whether the Android sandbox is secure or not: (1) information on the distribution of installed versions of Android over time and (2) information on the critical vulnerabilities found to affect specific versions of Android.
+These two datasets can then be combined to determine the proportion of sandboxes at risk of attack from specific vulnerabilities over time.
 
-We indicate the uncertainty in our results by presenting them $\pm$ one standard deviation and give results to 3 significant figures, this occasionally results in `$\pm\, 0$' when the standard deviation is small.
+We indicate the uncertainty in our results by presenting them $\pm$ one standard deviation and give results to 3~s.f., this occasionally results in `$\pm\, 0$' when the standard deviation is small.
 \todolater{Do we want to use the 95 percentile instead}
-This technique cannot account for systematic errors which we explore in \S\ref{sec:representative}.
+We explore systematic errors in \S\ref{sec:representative}.
 
@@
 In our analysis we use historical data collected by the \da\ project~\cite{Wagner2013}.
-\da\ collects data from study participants who install the Android app from the Google Play store.
-Most study participants allow researchers around the world to access a subset of their device data, including the data presented in this paper.
+\da\ collects data from study participants who install the Android app from Google Play.
+Most study participants allow researchers around the world to access a subset of their device data, including the data presented here.
 
 The \da\ app collects a range of metadata from Android devices.\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/collected.html}}
-For this paper, we extracted the build string and API version from the device log each day.
+We extracted the build string and API version for each device each day.
 The build string is a user-readable version string.
@@
 Therefore Android users need to take additional steps to protect themselves. 
-Firstly, users should only install apps from the Google Play store, which previous research has been shown to have a lower level of malware than third-party markets and is known to contain static and dynamic analysis of available apps.
+Firstly, users should only install apps from Google Play, which previous research has been shown to have a lower level of malware than third-party markets and is known to contain static and dynamic analysis of available apps.
 Secondly, users should consider installing an anti-virus tool or enabling the Google verify apps feature if available. 

### 2015-04-09 securityupdates.tex
Reword to make it shorter - back under page limit

@@
 \subsection{Critical vulnerabilities}
-We compiled a list of critical vulnerabilities in Android, containing information on the discovery and publication dates, the versions affected and which versions fixed the problem.
+We compiled a list of critical vulnerabilities in Android, containing information on the discovery and publication dates, the versions affected and which versions contained fixes.
 We only looked for critical vulnerabilities such as root vulnerabilities which did not require USB debugging to exploit.
@@
 Some phones can be `rooted' by enabling USB debugging and using the special privileges of the ADB shell to root the device but only \daAdbEnabledPerc\ of devices have USB debugging enabled.
-This is not something that applications running on the phone can exploit to break out of the app sandbox and so we do not include those vulnerabilities.
+This is not something that applications running on the phone can exploit and so we do not include those vulnerabilities.
 However there have been reports of Windows malware exploiting this vector.\footnote{\url{http://www.symantec.com/connect/blogs/windows-malware-attempts-infect-android-devices}}
@@
 Some critical vulnerabilities are not traditional kernel vulnerabilities, but are vulnerabilities in level 2 of our threat model.
-For example improper verification of signatures at installation time was discovered in February 2013~\cite{Forristal2013} and meant that applications could pretend to be signed with system keys and from there gain root privileges.
+For example improper verification of signatures at installation time was discovered in February 2013~\cite{Forristal2013} and meant that applications could pretend to be signed with system keys and hence be granted system privileges.
 On versions of Android below 4.1 malware could then use known system-to-root escalation mechanisms.
@@
 
-We developed and maintain an open platform for filing critical vulnerabilities in a machine readable format, the \href{http://androidvulnerabilities.org/}{AndroidVulnerabilities.org}%\footnote{\textbf{Note to reviewers:} we have removed our names and affiliation from the website for this review \textbf{with the exception of the contact page.
+We developed and maintain an open platform for filing critical vulnerabilities in a machine readable format, the \href{http://androidvulnerabilities.org/}{AndroidVulnerabilities.org} %\footnote{\textbf{Note to reviewers:} we have removed our names and affiliation from the website for this review \textbf{with the exception of the contact page.
 Please do not visit the contact page until the review cycle has completed.}\vspace{-4em}} (\avo) website.
@@
 
-Unless otherwise stated, we use \daNumVulnsUsed\ vulnerabilities in our analysis as shown in Table~\ref{tab:andvulns}.
+Unless otherwise stated, we use the \daNumVulnsUsed\ vulnerabilities shown in Table~\ref{tab:andvulns} in our analysis.
 We have chosen these vulnerabilities to fit the attack vectors introduced in \S\ref{sec:threatmodel}.
@@
 In many cases we could not match manufacturer- and model-specific vulnerabilities to individual devices in the \da\ data and therefore attempting to include device-specific vulnerabilities as well would introduce additional uncertainty in our results.
-In contrast, with our chosen set of vulnerabilities, our analysis represents a lower-bound on the vulnerability of devices in the \da\ data set.
+In contrast, with our chosen set of vulnerabilities, our analysis represents a lower-bound on the vulnerability of devices in the \da\ data.
 
 Tracking vulnerabilities is a manual task as they are not consistently recorded in other databases such as the CVE database.
-In addition, the lack of a widely acknowledged unique identifier required manual analysis to identify whether two reports referenced the same vulnerability.
+In addition, the lack of a unique identifier required manual analysis to identify whether two reports referenced the same vulnerability.
 Previous work has assumed ``any security issue of relevance will eventually get a CVE number assigned"~\cite{Frei2010} which is currently not the case for critical Android vulnerabilities.
-For some of the vulnerabilities without CVE numbers, Google confirmed that there was no CVE number and that they did not intend to get one, instead providing an internal Android bug number.
+For some of the vulnerabilities without CVE numbers, Google confirmed that there was no CVE number and that they did not intend to get one, instead providing an internal bug number.
 
@@
 Frei et al.~\cite{Frei2010} propose the definition of the \textbf{time of disclosure}.
-This occurs when the information about the vulnerability is available to the manufacturer from an accepted source.
-Unfortunately before we collated this information much of it was not published by an accepted source, even months or years after they had been actively used.
+This occurs when the information about the vulnerability is available to the user from an accepted source.
+Unfortunately before we collated this information much of it was not published by an accepted source, even after months or years of active use.
 
@@
 In this analysis we examine how the distribution of Andriod OS versions changes over time and the impact that has on how different vulnerabilities affect the security of Android.
-We show that some vulnerabilities continue to have a substantial effect long after they have been fixed and the importance of longitudinal studies for determining whether a particular device is exposed to a particular vulnerability.
+We show that some vulnerabilities continue to have a substantial effect long after they have been fixed and the importance of longitudinal studies for determining whether devices are exposed to a particular vulnerabilities.
 
-\subsubsection{Method} As in Analysis 1, we used the version information for each device to calculate which critical vulnerabilities each device was susceptible to on a daily basis.
+\subsubsection{Method} As in Analysis 1, we used the version information for each device to calculate which critical vulnerabilities each device was susceptible to each day.
 

### 2015-04-09 securityupdates.tex
Minor fixes on reading through

@@
 \end{abstract}
+\todo{Abstract does not reflect current spin very well}
 
 \section{Introduction}
-\todo{We talk about root a lot again, rather critical vulnerabilities, and root is not strictly accurate}
 Modern mobile operating systems provide several layers of security whose combined aim is to prevent app malware from taking control of a device. 
@@
 
-Our focus is on malware whose aim is to gain full control of the device rather than merely abuse permissions granted to it by the user.
+Our focus is on vulnerabilities which allow malware to gain full control of the device rather than merely abuse permissions granted to it by the user.
 In this context, by full control, or \emph{root}, on a device we mean the malware has unfettered access to all data stored on the device and can control all network interfaces in the device with the same flexibility as the operating system itself.
@@
 A marketplace can use three mechanisms to prevent malicious apps from appearing in the marketplace: firstly by economic disincentives, including charging developers to create an account and cancelling malicious accounts; secondly by technical means, such as static and dynamic analysis on submitted apps; and finally via social feedback from users, including the provision of reviews, rating and reporting.
-The second line of defence occurs on the handset at installation time, and includes checking the signature of any app installation or update as well as statically verifying the app (e.g.
+The second line of defence occurs on the handset at installation time, and includes checking the signature of any app installation or update as well as checking app for malware (e.g.
 Google's Verify Apps feature).
-Finally, the third line of defence takes place at runtime: Android apps are run in a \emph{sandbox} which prevents the app from unauthorised actions such as accessing any permissions which weren't granted by the user at install time; in addition, anti-virus products may scan the list of installed apps periodically to detect malware.
+Finally, the third line of defence takes place at runtime: Android apps are run in a \emph{sandbox} which prevents the app from unauthorised actions such as accessing any permissions which weren't granted by the user at install time; in addition, anti-malware products may scan the list of installed apps periodically to detect malware.
 
@@
  \item We created an open database to collate quantitative information on vulnerabilities which affect Android and use this to determine the proportion of devices they affect and how that changes over time.
- \item We demonstrate that the security of Android relies on the centralisation of app deployment via a marketplace, and the use of install-time anti-virus checks rather than through the security of the app sandbox.
+ \item We demonstrate that the security of Android relies on the centralisation of app deployment via a marketplace, and the use of install-time anti-malware checks rather than through the security of the app sandbox.
  \item We use our analysis to develop advice for Android users on what steps they should take to increase the security of their device.
@@
 We use two sources of data to determine whether the Android sandbox is secure or not: (1) information on the distribution of installed versions of Android over time and (2) information on the critical vulnerabilities found to affect specific versions of Android.
-These two datasets can then be combined to determine the proportion of sandboxes at risk of attack from specific vulnerabilities over time.
+These two datasets can then be combined to determine the proportion of devices at risk of attack from specific vulnerabilities over time.
 
@@
 
-The \da\ app collects a range of metadata from Android devices.\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/collected.html}}
+The \da\ app collects a range of data from Android devices.\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/collected.html}}
 We extracted the build string and API version for each device each day.
@@
 Consequently security (bug) fixes do not result in a change in the API version.
-Fortunately most (\daOSVersionPercValidLines) entries in the data have a build string of the form `x.y.z opaque\_marker' and so it is possible to extract the Android version number `x.y.z'.
+Fortunately most (\daOSVersionPercValidLines) entries in these data have a build string of the form `x.y.z opaque\_marker' and so it is possible to extract the Android version number `x.y.z'.
 On a large proportion of devices `opaque\_marker' is a well defined build number\footnote{\url{https://source.android.com/source/build-numbers.html}} however different device manufacturers use different schema.
@@
 %The discovery dates of our vulnerabilities are not uniform. 
-%In particular the data shows a large gap from 2011-10-06 to 2013-02-18 where we do not know of any discoveries of critical vulnerabilities affecting all Android devices.
+%In particular these data shows a large gap from 2011-10-06 to 2013-02-18 where we do not know of any discoveries of critical vulnerabilities affecting all Android devices.
 %The cause of this quiet period is unclear.
@@
 To investigate the vulnerability of Android devices, we used the OS version information from \da\ and the vulnerability data from \avo.
+The \avo\ data covers the period from \avoFirstDataDate\ to \avoLastDataDate.
 The \da\ data was collected between \daStartDate\ and \daEndDate.
-The \avo\ data covers the period from \avoFirstDataDate\ to \avoLastDataDate.
 For each device we collected daily version data and any vulnerabilities it was exposed to at that time.
@@
 From February 2013 onwards regular discovery of critical vulnerabilities ensures that most devices are exposed to known critical vulnerabilities.
-
 This gives us an average exposure to critical vulnerabilities of Android devices of \daMeanInsecurityPerc.
@@
 Figure~\ref{fig:nvulnerabilities_heat} shows how the the \emph{APK signing vulnerabilities} affected all devices and took months to get fixed for any device.
-However what is perhaps more worrying is the long tail on the \avovuln{Gingerbreak}, \avovuln{levitator}, \avovuln[exploid_udev]{exploid} and \avovuln{zergRush} vulnerabilities which are more dangerous level 3 vulnerabilities (since they do not require new app installation) and which affect a significant proportion of devices years later.
+However what is perhaps more worrying is the long tail on the \avovuln{Gingerbreak}, \avovuln{levitator}, \avovuln[exploid_udev]{exploid} and \avovuln{zergRush} vulnerabilities which are more dangerous level 3 vulnerabilities (since they do not require new app installation) and which affected a significant proportion of devices years later.
 %\todo{quantify this long tail and the APK numbers as headline numbers}
@@
 \vspace{-1.5em}
-\caption{Each strip shows OS versions over time. Black lines show where an update only changed the build number.}
+\caption{Each strip shows OS versions over time.}
 \label{fig:device_data_os}
 \end{subfigure}
-\caption{The top \daNumDeviceDataDevices\ devices by days of contribution in the \da\ data. One strip per device handset.}
+\caption{The top \daNumDeviceDataDevices\ devices by days of contribution in the \da\ data. One strip per device handset. Black lines show where an update only changed the build number.}
 \end{figure}
@@
 Attackers have taken this approach for the Firefox web browser~\cite{Barth2011}.
-Google does not release updates to the Android source code until after the release of the relevant update.
+Google does not release updates to the Android source code until after the release of the relevant update~\cite{AndroidSecurity2014}.
 This reduces the time available for an attacker to exploit a vulnerability by looking at security updates disclosed in the source code repository.
 Reducing it from, the interval between the commit of the first fix and the fix being deployed to most devices, to the interval between the first fixing release and the fix being deployed to most devices; unfortunately this can still take months or years.
-For example, the zergRush vulnerability was fixed in a release before it was publicly discovered but it still took 27 months for 90\% of the devices in our study to be definitely fixed\footnote{Definitely fixed is 1 $-$ (definitely insecure + maybe secure), maybe fixed is 1 $-$ definitely insecure} (25 months for 90\% of devices maybe fixed).
+For example, the zergRush vulnerability was fixed in a release before it was publicly discovered but it still took 27 months for 90\% of the devices in our study to be definitely fixed\footnote{Definitely fixed is 1 $-$ (definitely insecure + maybe secure), maybe fixed is 1 $-$ definitely insecure} (25 months for 90\% of devices to be maybe fixed).
 
-Brumley et al. has shown it is possible to automatically generate exploits from binary fixes, for example by using security updates issued by manufacturers~\cite{Brumley2008}. 
+Brumley et.\ al. has shown it is possible to automatically generate exploits from binary fixes, for example by using security updates issued by manufacturers~\cite{Brumley2008}. 
 Similarly, the Woodpecker tool automatically finds permission leaks in stock Android phone images~\cite{Grace2012}.
@@
 There are continuing efforts to reduce the impact of critical vulnerabilities, both in Android and elsewhere.
-SEAndroid~\cite{Smalley2013} which is included in Android from version 4.1~\cite{jelly-bean-release} claimed to prevent some root vulnerabilities and to reduce the impact of others.
+SEAndroid~\cite{Smalley2013} which is included in Android from version 4.1~\cite{jelly-bean-release} and fully enforcing from version 5.0~\cite{AndroidSecurity2014} claimed to prevent some root vulnerabilities and to reduce the impact of others.
 Capability based enforcement systems such as Capsicum~\cite{Watson2010} substantially reduce the capabilities that an exploit has to try and gain increased privilege with and could be included in Linux\footnote{\url{https://github.com/google/capsicum-linux}} and hence Android.
@@
 
-Unfortunately, whilst these tools were successful in finding critical vulnerabilities, such techniques are not perfect.
+Unfortunately, whilst these tools were successful in finding malware, such techniques are not perfect.
 The Malware Genome project collected 1\,260 malware samples from 2010--2011~\cite{Zhou2012b} and found a best case anti-malware detection rate of 79.6\%.
@@
 Firstly, users should only install apps from Google Play, which previous research has been shown to have a lower level of malware than third-party markets and is known to contain static and dynamic analysis of available apps.
-Secondly, users should consider installing an anti-virus tool or enabling the Google verify apps feature if available. 
+Secondly, users should enable the Google Verify Apps feature if available or use an anti-malware app if not. 
 Thirdly, users should avoid using untrusted WiFi networks or use them in conjunction with a VPN tunnel to a trusted network. 

### 2015-04-09 securityupdates.tex
Get it under the page limit *again*

@@
 \usepackage{siunitx}
-\usepackage{todonotes}%[disable]
+\usepackage[disable]{todonotes}%
 \let\OldTodo\todo
@@
 Devices apply \daUpdatesPerYearTwosfNominal\ updates each year, less than the rate of critical vulnerability discovery of between \avoVulnsPerYearAllAndroidTwosfNominal\ and \avoVulnsPerYearTwosfNominal\ per year.
+\todo{Abstract does not reflect current spin very well}
 \end{abstract}
-\todo{Abstract does not reflect current spin very well}
 
@@
 In this context, by full control, or \emph{root}, on a device we mean the malware has unfettered access to all data stored on the device and can control all network interfaces in the device with the same flexibility as the operating system itself.
-We have selected our focus for two reasons: no previous published work has quantified how vulnerable Android devices are to malware gaining root, and secondly the negative impact on the device owner is much higher for malware which gains root than malware which does not: as it can not only perform arbitrary actions on the device, but it is also impossible for the device owner to remove the malware reliably as even a factory reset does not guarantee that the malware has been removed.
+We have selected our focus for two reasons: no previous published work has quantified how vulnerable Android devices are to malware gaining root, and secondly the negative impact on the device owner is much higher for malware which gains root than malware which does not: as it can not only perform arbitrary actions on the device, but it is also impossible for the device owner to remove the malware reliably as even a factory reset does not guarantee the malware has been removed.
 
@@
 
-The \da\ project has collected data from \daNumDevices\ devices with a total of \daDeviceDays\footnote{Here we are only counting devices and days for which we have valid OS version data.} device days.
+\da\ has collected data from \daNumDevices\ devices with a total of \daDeviceDays\footnote{Here we are only counting devices and days for which we have valid OS version data.} device days.
 The majority of devices only contribute data for a short period of time, however \daMonthsDevices\ devices have contributed data for more than \daMonths~months.
@@
 Devices, on average, apply \daUpdatesPerYear\ updates each year, which is less than the rate of critical vulnerability discovery of between \avoVulnsPerYearAllAndroid\ and \avoVulnsPerYear.
-\daPercUpdatesDowngrades\ of full version changes are downgrades to older versions.
-We describe three analyses, examining the vulnerability of Android devices in general (\S\ref{sec:exp:versionsecurity}), explore the upgrade cycle and vulnerability cycle of the ecosystem (\S\ref{sec:exp:android_ecosystem}) and quantify the updates installed on particular devices (\S\ref{sec:exp:device_updates}).
+\daPercUpdatesDowngrades\ of version changes are downgrades to older versions.
+We describe three analyses, examining the vulnerability of Android devices in general (\S\ref{sec:exp:versionsecurity}), explore the upgrade cycle and vulnerability cycle of the ecosystem (\S\ref{sec:exp:android_ecosystem}) and quantify the updates installed on devices (\S\ref{sec:exp:device_updates}).
 
@@
 \end{figure}
-In this analysis we calculate the proportion of Android devices which are exposed to critical vulnerabilities and discuss how that has changed over time.
+In this analysis we calculate the proportion of Android devices which are exposed to critical vulnerabilities and how that has changed over time.
 Figure~\ref{fig:proportioninsecure} describes the proportion of Android devices susceptible to at least one critical vulnerability.
@@
 \subsubsection{Results}
-The proportion of devices in the \da\ data running different versions of Android each day is shown in Figure~\ref{fig:norm_os}.\footnote{The anomaly beginning in August 2014 is explained in \S\ref{sec:da_changes}.}
+The proportion of devices in the \da\ data running different versions of Android each day is shown in Figure~\ref{fig:norm_os}.\footnote{\label{foot:anomaly}The anomaly beginning in August 2014 is explained in \S\ref{sec:da_changes}.}
 It shows how old versions are gradually replaced by new ones, and the long tail of devices which do not see updates to more recent versions.
@@
  \includegraphics[width=\textwidth]{figures/da_norm_os}
- \caption{Android versions in \da\ data over time. The anomaly beginning in August 2014 is explained in \S\ref{sec:da_changes}.}
+ \caption{Android versions in \da\ data over time.\textsuperscript{\ref{foot:anomaly}}}
  \label{fig:norm_os}
@@
 
-{
 \begin{figure}[!t]
@@
 \end{figure}
-}
 
 \subsubsection{Method}
-As in Analysis 1, we used the version information for each device to calculate which critical vulnerabilities each device was susceptible to on a daily basis.
-We also recorded when the version changed and which versions it changed from and to.
-In the \da\ data we have over 1000 devices contributing data in any particular week.\todolater{Use the real average number of devices contributing each week}
-
+As in Analysis 1, we used the version information for each device to calculate which critical vulnerabilities each device was susceptible to each day.
+We also recorded changes in version.
+In the \da\ data we have over 1000 devices contributing data in any particular week. \todolater{Use the real average number of devices contributing each week}
 However most devices only contribute for a short period of time and so we do not observe updates happening on every device.
 Instead we have a hopefully representative sample of updates which happened while \da\ was installed on the devices.
-\da\ cannot distinguish between a device being replaced and the \da\ app being removed as while multiple devices can be linked to the same user account, this is a manual process and consequently few users do it.
+\da\ cannot distinguish between a device being replaced and the \da\ app being removed as while multiple devices can be linked to the same user account, this is a manual process and few users do it.
 
 \subsubsection{Results}
-
 The longitudinal data on the number of vulnerabilities affecting the \daNumDeviceDataDevices\ devices which have contributed the most days of data to \da\ changes over time is shown in Figure~\ref{fig:device_data_security}.
@@
 \label{sec:da_changes}
-The \da\ data is mostly generated by devices which have the \da\ app installed because their owner happened to come across the app as there has been no advertising and few other attempts to increase usage.
+The \da\ data is mostly generated by devices which have the \da\ app installed because their owner happened to come across the app as there has been no advertising and few attempts to increase usage.
 However two collaborations with network operators to install \da\ on customer devices resulted in large numbers of new \da\ users coming from particular network operators.
@@
 We still do not know where new versions of Android come from.
-Over time newer versions of Android come to dominate, but quantifying the proportion due to new phone purchases verses the proportion due to updates of existing devices is unknown.
+Over time newer versions of Android come to dominate, but the proportion due to new phone purchases verses the proportion due to updates of existing devices is unknown.
 Device Analyzer is the only available large data set we know of with longitudinal traces of Android OS version strings.
 Unfortunately this data set only records \daNumUpdatesUpgrades\ upgrades which, broken down over \daOSMonthsOfData\ months and \daNumSigOSVersions\ OS versions is only \daUpdatesPerMonthPerVersion\ updates per version per month.
-This is not enough to build a statistically significant prediction of the expected transition to newer versions of Android.
-If manufacturers shipped security updates without changing the build number then we would not be able to detect this, but that would also indicate a level of incompetence likely to damage security and so can be discounted.
+This is not enough to build a statistically significant prediction of the expected transition between versions of Android.
+If manufacturers shipped security updates without changing the build number then we would not be able to detect this, but that would indicate a level of incompetence likely to damage security.
 
@@
 We have described the three levels of defence provided to Android users today and determined that the app sandbox is the best place to prevent malware from gaining root.
-However the latency in the security update process means that on average \daMeanInsecurityPerc\ of Android devices are exposed to known critical vulnerabilities which break out of the sandbox and only \daUpdatednessPerc\ of devices run the latest version of Android.
-Devices apply \daUpdatesPerYear\ updates each year less than the rate of critical vulnerability discovery of between \avoVulnsPerYearAllAndroid\ and \avoVulnsPerYear.
+However the latency in the security update process means that on average \daMeanInsecurityPercNominal\ of Android devices are exposed to known critical vulnerabilities which break out of the sandbox and only \daUpdatednessPercNominal\ of devices run the latest version of Android.
+Devices apply \daUpdatesPerYearNominal\ updates each year less than the rate of critical vulnerability discovery of between \avoVulnsPerYearAllAndroidNominal\ and \avoVulnsPerYearNominal.
 
@@
 Therefore Android users need to take additional steps to protect themselves. 
-Firstly, users should only install apps from Google Play, which previous research has been shown to have a lower level of malware than third-party markets and is known to contain static and dynamic analysis of available apps.
-Secondly, users should enable the Google Verify Apps feature if available or use an anti-malware app if not. 
+Firstly, users should only install apps from Google Play, which has been shown to have a lower level of malware than third-party markets and is known to contain static and dynamic analysis of available apps.
+Secondly, users should enable the Google Verify Apps feature if available or use an anti-malware app. 
 Thirdly, users should avoid using untrusted WiFi networks or use them in conjunction with a VPN tunnel to a trusted network. 
-Finally, if the user is running a device with API level 17 or above and is technically able to determine the target API version of their apps, they should consider uninstalling all apps with a target API of less than 17.
+Finally, if the user is running a device with API level 17 or above and is able to determine the target API version of their apps, they should consider uninstalling all apps with a target API of less than 17.
 

### 2015-04-09 securityupdates.tex
First bunch of changes from Ollie's comments
One day I will learn the difference between which and that

@@
 \begin{document}
-\title{Only the store can save you now:\\updates and vulnerabilities on Android}
+\title{Only the store can save you now:\\Updates and vulnerabilities on Android}
 
@@
 Modern smartphone operating systems protect users from malicious applications by providing a protected execution environment or sandbox.
-A number of exploits have been found which allow apps to break out of the sandbox.
+A number of exploits have been found that allow apps to break out of the sandbox.
 We quantify for the first time how vulnerable Android devices in the wild are to critical vulnerabilities and how long it takes for vulnerabilities to be fixed on consumer devices.
@@
 
-Our focus is on vulnerabilities which allow malware to gain full control of the device rather than merely abuse permissions granted to it by the user.
+Our focus is on vulnerabilities that allow malware to gain full control of the device rather than merely abuse permissions granted to it by the user.
 In this context, by full control, or \emph{root}, on a device we mean the malware has unfettered access to all data stored on the device and can control all network interfaces in the device with the same flexibility as the operating system itself.
-We have selected our focus for two reasons: no previous published work has quantified how vulnerable Android devices are to malware gaining root, and secondly the negative impact on the device owner is much higher for malware which gains root than malware which does not: as it can not only perform arbitrary actions on the device, but it is also impossible for the device owner to remove the malware reliably as even a factory reset does not guarantee the malware has been removed.
+We have selected our focus for two reasons: no previous published work has quantified how vulnerable Android devices are to malware gaining root, and secondly the negative impact on the device owner is much higher for malware that gains root than malware which does not: as it can perform arbitrary actions on the device, and it is also impossible for the device owner to remove the malware reliably as even a factory reset does not guarantee the malware has been removed.
 
-This is an important problem: previous studies have shown that a significant proportion of malware attempts to gain root: in 2012, 40\%~\cite{Zhou2012a} of malware for Android contained exploits designed to gain root.
+This is an important problem: previous studies have shown that a substantial proportion of malware attempts to gain root: in 2012, 40\%~\cite{Zhou2012a} of malware for Android contained exploits designed to gain root.
 
@@
 Google's Verify Apps feature).
-Finally, the third line of defence takes place at runtime: Android apps are run in a \emph{sandbox} which prevents the app from unauthorised actions such as accessing any permissions which weren't granted by the user at install time; in addition, anti-malware products may scan the list of installed apps periodically to detect malware.
+Finally, the third line of defence takes place at runtime: Android apps run in a \emph{sandbox} that prevents the app from unauthorised actions such as accessing any permissions which weren't granted by the user at install time.
+In addition, anti-malware products may scan the installed apps periodically to detect malware.
 
@@
  \item We quantify the Android update ecosystem showing how the OS version distribution changes and the frequency of updates.
- \item We created an open database to collate quantitative information on vulnerabilities which affect Android and use this to determine the proportion of devices they affect and how that changes over time.
+ \item We create an open database to collate quantitative information on vulnerabilities that affect Android and use this to determine the proportion of devices they affect and how that changes over time.
  \item We demonstrate that the security of Android relies on the centralisation of app deployment via a marketplace, and the use of install-time anti-malware checks rather than through the security of the app sandbox.
- \item We use our analysis to develop advice for Android users on what steps they should take to increase the security of their device.
+ \item We use our analysis to develop advice for Android users on steps they should take to increase the security of their device.
  \item We characterise the impact of critical vulnerabilities on the Android ecosystem as a whole, examine the effectiveness of vulnerabilities and how long vulnerabilities remained available in the wild, and also examine the impact on individual devices over extended periods of time.
@@
 The second attack vector is to make an existing app download and execute code at runtime.
-The most direct method available to an attacker is to upload to a marketplace a seemingly innocent app which contains support for dynamic code loading from a malicious server.
+The most direct method available to an attacker is to upload to a marketplace a seemingly innocent app that contains support for dynamic code loading from a malicious server.
 Neither static nor dynamic analysis of this app by the marketplace will uncover any malicious code, since it does not exist in the app binary.
@@
 Unfortunately, as we will demonstrate, the sandbox for Android apps is weak.
-Therefore many Android users implicitly rely on the marketplace for protection by removing apps which are detected as malicious.
+Therefore many Android users implicitly rely on the marketplace for protection by removing apps that are detected as malicious.
 This detection can only be derived from static and dynamic analysis or after reports of malicious behaviour from users once it is made available for download.
@@
 The build string is a user-readable version string.
-The API version is a positive integer which increases when new features are added to the API.
+The API version is a positive integer that increases when new features are added to the API.
 Consequently security (bug) fixes do not result in a change in the API version.
@@
 We compiled a list of critical vulnerabilities in Android, containing information on the discovery and publication dates, the versions affected and which versions contained fixes.
-We only looked for critical vulnerabilities such as root vulnerabilities which did not require USB debugging to exploit.
+We only looked for critical vulnerabilities such as root vulnerabilities that did not require USB debugging to exploit.
 If malicious code exploits a critical vulnerability then it gains control of the device.
@@
 
-We developed and maintain an open platform for filing critical vulnerabilities in a machine readable format, the \href{http://androidvulnerabilities.org/}{AndroidVulnerabilities.org} %\footnote{\textbf{Note to reviewers:} we have removed our names and affiliation from the website for this review \textbf{with the exception of the contact page.
+We developed and maintain an open platform for filing critical vulnerabilities in a machine readable format,  \href{http://androidvulnerabilities.org/}{AndroidVulnerabilities.org} (\avo).
+%\footnote{\textbf{Note to reviewers:} we have removed our names and affiliation from the website for this review \textbf{with the exception of the contact page.
 Please do not visit the contact page until the review cycle has completed.}\vspace{-4em}} (\avo) website.
@@
 
-Tracking vulnerabilities is a manual task as they are not consistently recorded in other databases such as the CVE database.
+Tracking Android vulnerabilities is a manual task as they are not consistently recorded in other databases such as the CVE database.
 In addition, the lack of a unique identifier required manual analysis to identify whether two reports referenced the same vulnerability.
-Previous work has assumed ``any security issue of relevance will eventually get a CVE number assigned"~\cite{Frei2010} which is currently not the case for critical Android vulnerabilities.
+Previous work has assumed ``any security issue of relevance will eventually get a CVE number assigned"~\cite{Frei2010}, which is currently not the case for critical Android vulnerabilities.
 For some of the vulnerabilities without CVE numbers, Google confirmed that there was no CVE number and that they did not intend to get one, instead providing an internal bug number.
@@
 Establishing when a vulnerability starts to pose a threat to users is difficult.
-Frei et al.~\cite{Frei2010} propose the definition of the \textbf{time of disclosure}.
+Frei et al.~\cite{Frei2010} propose the definition of the \emph{time of disclosure}.
 This occurs when the information about the vulnerability is available to the user from an accepted source.
@@
 
-Symantec's 2012 analysis of desktop malware has shown that after public disclosure, exploitation rates increase by 5 orders of magnitude~\cite{Bilge2012} and so from the point of view of widespread danger, the period between public disclosure and the date of the fix being deployed to most devices is the most critical.
+Symantec's 2012 analysis of desktop malware has shown that after public disclosure, exploitation rates increase by 5 orders of magnitude~\cite{Bilge2012} and so for widespread danger, the period between public disclosure and the date of the fix being deployed to most devices is the most critical.
 However they also show that zero-day vulnerabilities are typically used for 312 days before they are publicly disclosed, often to target particular organisations.
@@
 \end{figure}
-In this analysis we calculate the proportion of Android devices which are exposed to critical vulnerabilities and how that has changed over time.
+In this analysis we calculate the proportion of Android devices that are exposed to critical vulnerabilities and how that has changed over time.
 Figure~\ref{fig:proportioninsecure} describes the proportion of Android devices susceptible to at least one critical vulnerability.
@@
 From October 2011 until the discovery of \avovuln[APK_duplicate_file]{APK duplicate file} in February 2013 the graph shows progressive improvement as devices are upgraded.
-This means more and more devices are marked as \emph{secure} because they are now running a secure version of Android, or marked as \emph{maybe secure} because they received an OS update which did not update to a known-good version of Android but which may still have included a backport of a fix, as the update was made available after the vulnerability was disclosed.
+This means more and more devices are marked as \emph{secure} because they are now running a secure version of Android, or marked as \emph{maybe secure} because they received an OS update that did not update to a known-good version of Android but which may still have included a backport of a fix, as the update was made available after the vulnerability was disclosed.
 From February 2013 onwards regular discovery of critical vulnerabilities ensures that most devices are exposed to known critical vulnerabilities.
@@
 In this analysis we examine how the distribution of Andriod OS versions changes over time and the impact that has on how different vulnerabilities affect the security of Android.
-We show that some vulnerabilities continue to have a substantial effect long after they have been fixed and the importance of longitudinal studies for determining whether devices are exposed to a particular vulnerabilities.
+We show that some vulnerabilities continue to have a substantial affect long after they have been fixed and the importance of longitudinal studies for determining whether devices are exposed to a particular vulnerabilities.
 

### 2015-04-09 securityupdates.tex
More updates based on Ollie's comments

@@
 
-Sandboxes only provide protection if they are free from design errors and implementation flaws. 
+Sandboxes only provide protection if they are free from design and implementation flaws. 
 Both of these have occurred in Android, as indeed happens today in many large security systems relying on the sandbox model. 
@@
 \subsection{Analysis 2: behaviour of the Android ecosystem}\label{sec:exp:android_ecosystem}
-In this analysis we examine how the distribution of Andriod OS versions changes over time and the impact that has on how different vulnerabilities affect the security of Android.
+In this analysis we examine how the distribution of Android OS versions changes over time and the impact that has on how different vulnerabilities affect the security of Android.
 We show that some vulnerabilities continue to have a substantial affect long after they have been fixed and the importance of longitudinal studies for determining whether devices are exposed to a particular vulnerabilities.
@@
 It shows how old versions are gradually replaced by new ones, and the long tail of devices which do not see updates to more recent versions.
-This gives a mean proportion of devices running the most recent version of Android is \daUpdatednessPerc.
+This gives a mean proportion of devices running the most recent version of Android since 2011 is \daUpdatednessPerc.
 
@@
  %The change in behaviour after August 2014 is explained in \S\ref{sec:da_changes}
+ \todo{This plot is confusing and perhaps we should quantise the colour map.}
  }
@@
 
-In 2013 three vulnerabilities were found in the way which Android verified the signatures on APKs.
+In 2013 three vulnerabilities were found in how Android verified the signatures on APKs.
 These are level 2 vulnerabilities in our threat model since they require a new app installation to occur.
@@
 We also recorded changes in version.
-In the \da\ data we have over 1000 devices contributing data in any particular week. \todolater{Use the real average number of devices contributing each week}
+In the \da\ data we have about 1000 devices contributing data in any particular week. \todolater{Use the real average number of devices contributing each week}
 However most devices only contribute for a short period of time and so we do not observe updates happening on every device.
@@
 
-By taking the number of updates observed and the number of device days of data we have collected we can compute the number of updates received by a device per year to be \daUpdatesPerYear.
+By dividing the number of updates observed by the number of device years of data we have collected we compute the number of updates received by a device per year to be \daUpdatesPerYear.
 This compares badly with the number of critical vulnerabilities discovered per year of between \avoVulnsPerYearAllAndroid\ (affecting all Android) and \avoVulnsPerYear\ (including the device manufacturer specific ones).

### 2015-04-09 securityupdates.tex
spell check
Add Oliver Chick to the Acknowledgements

@@
 The Malware Genome project collected 1\,260 malware samples from 2010--2011~\cite{Zhou2012b} and found a best case anti-malware detection rate of 79.6\%.
-Similarly, DroidChameleon found that antivirus products could not detect malware if the binary was automatically permuted~\cite{Rastogi2013}.
+Similarly, DroidChameleon found that anti-virus products could not detect malware if the binary was automatically permuted~\cite{Rastogi2013}.
 Therefore detecting risky apps which might contain critical vulnerabilities is a useful tool, but it does not replace the need to provide security updates.
@@
 \section*{Acknowledgements}
-Thanks to: Richard Clayton, the mobile security reading group and Anil Madhavapeddy for reading drafts;
+Thanks to: Richard Clayton, the mobile security reading group, Anil Madhavapeddy and Oliver Chick for reading drafts;
 David Robertson for statistical advice;

### 2015-04-10 securityupdates.tex
Updated abstract and conclusions...

@@
 \begin{abstract}
-Modern smartphone operating systems protect users from malicious applications by providing a protected execution environment or sandbox.
-A number of exploits have been found that allow apps to break out of the sandbox.
-We quantify for the first time how vulnerable Android devices in the wild are to critical vulnerabilities and how long it takes for vulnerabilities to be fixed on consumer devices.
-We analysed \da\ data from over \daOSYearsOfData\ years and \daNumOSDevices\ devices and found that on average \daMeanInsecurityPercTwosfNominal\ of devices were exposed to known root privilege vulnerabilities and only \daUpdatednessPercTwosfNominal\ of devices run the most recent version of Android.
-Devices apply \daUpdatesPerYearTwosfNominal\ updates each year, less than the rate of critical vulnerability discovery of between \avoVulnsPerYearAllAndroidTwosfNominal\ and \avoVulnsPerYearTwosfNominal\ per year.
-\todo{Abstract does not reflect current spin very well}
+Modern mobile operating systems combine up to three layers of security protection to prevent app malware from taking control of a device.
+These are: the defences provided by the app store, additional checks performed at installation time on the device, and a protected execution environment to control the running app.
+In this paper we show that only the app store is currently able to provide effective protection for the majority of Android users.
+In contrast, a number of well-known, publicly disclosed exploits continue to allow apps to bypass installation checks or bypass runtime restrictions.
+We quantify for the first time the vulnerability of Android devices in the wild and how long it takes for critical flaws to be fixed on consumer devices.
+To do so, we analyse \da\ data from over \daOSYearsOfData\ years and \daNumOSDevices\ devices and found that, on average, \daMeanInsecurityPercTwosfNominal\ of devices were exposed to known privilege-escalation attacks.
+% and only \daUpdatednessPercTwosfNominal\ of devices run the most recent version of Android.
+Furthermore, devices apply \daUpdatesPerYearTwosfNominal\ updates each year, less than the critical vulnerability discovery rate of between \avoVulnsPerYearAllAndroidTwosfNominal\ and \avoVulnsPerYearTwosfNominal\ per year.
 \end{abstract}
@@
 In this context, by full control, or \emph{root}, on a device we mean the malware has unfettered access to all data stored on the device and can control all network interfaces in the device with the same flexibility as the operating system itself.
-We have selected our focus for two reasons: no previous published work has quantified how vulnerable Android devices are to malware gaining root, and secondly the negative impact on the device owner is much higher for malware that gains root than malware which does not: as it can perform arbitrary actions on the device, and it is also impossible for the device owner to remove the malware reliably as even a factory reset does not guarantee the malware has been removed.
+We have selected our focus for two reasons: no previous published work has quantified how vulnerable Android devices are to malware gaining root, and secondly the negative impact on the device owner is much higher for malware that gains root than malware which does not.
+Such malware can perform arbitrary actions on the device, and it is also impossible for the device owner to remove the malware reliably as even a factory reset does not guarantee the malware has been removed.
 
-This is an important problem: previous studies have shown that a substantial proportion of malware attempts to gain root: in 2012, 40\%~\cite{Zhou2012a} of malware for Android contained exploits designed to gain root.
+This is an important problem: previous studies have shown that a substantial proportion of malware attempts to gain root.
+For example, in 2012, 40\% of malware for Android contained exploits designed to gain root~\cite{Zhou2012a}.
 
@@
 
-The second attack vector is to make an existing app download and execute code at runtime.
+The second attack vector is to coerce an existing app into downloading and executing code at runtime.
 The most direct method available to an attacker is to upload to a marketplace a seemingly innocent app that contains support for dynamic code loading from a malicious server.
@@
 \label{sec:conclusion}
-We have described the three levels of defence provided to Android users today and determined that the app sandbox is the best place to prevent malware from gaining root.
-However the latency in the security update process means that on average \daMeanInsecurityPercNominal\ of Android devices are exposed to known critical vulnerabilities which break out of the sandbox and only \daUpdatednessPercNominal\ of devices run the latest version of Android.
-Devices apply \daUpdatesPerYearNominal\ updates each year less than the rate of critical vulnerability discovery of between \avoVulnsPerYearAllAndroidNominal\ and \avoVulnsPerYearNominal.
+
+Modern mobile operating systems combine up to three layers of security protection to prevent app malware from taking control of a device.
+These are: the app store, checks performed during app installation, and the app sandbox.
+In this paper we have demonstrated that the app sandbox is the best place to prevent malware from gaining root.
+However the Android sandbox is ineffective in the majority of cases: the latency in the security update process means that on average \daMeanInsecurityPercNominal\ of Android devices are exposed to known critical vulnerabilities which allow a malicious app to break out of the sandbox.
+Furthermore, only \daUpdatednessPercNominal\ of devices run the latest version of Android and devices apply \daUpdatesPerYearNominal\ updates each year, less than the critical vulnerability discovery rate of between \avoVulnsPerYearAllAndroidNominal\ and \avoVulnsPerYearNominal.
 
@@
 Firstly, users should only install apps from Google Play, which has been shown to have a lower level of malware than third-party markets and is known to contain static and dynamic analysis of available apps.
-Secondly, users should enable the Google Verify Apps feature if available or use an anti-malware app. 
+Secondly, users should enable the Google Verify Apps feature if available. 
 Thirdly, users should avoid using untrusted WiFi networks or use them in conjunction with a VPN tunnel to a trusted network. 

### 2015-04-10 securityupdates.tex
Comment out Acknowledgements during review, we can bring them back for the camera ready

@@
 \identifying{
-\section*{Acknowledgements}
-Thanks to: Richard Clayton, the mobile security reading group, Anil Madhavapeddy and Oliver Chick for reading drafts;
-David Robertson for statistical advice;
-and Laurent Simon, Thomas Coudray, Adrian Taylor, Justin Case, Giant Pune and Khilan Gudka for reporting vulnerabilities.
-We were funded by the EPSRC and Google but the the work presented here does not necessarily reflect their views.
+%\section*{Acknowledgements}
+%Thanks to: Richard Clayton, the mobile security reading group, Anil Madhavapeddy and Oliver Chick for reading drafts;
+%David Robertson for statistical advice;
+%and Laurent Simon, Thomas Coudray, Adrian Taylor, Justin Case, Giant Pune and Khilan Gudka for reporting %vulnerabilities.
+%We were funded by the EPSRC and Google but the the work presented here does not necessarily reflect their %views.
 }

### 2015-04-10 securityupdates.tex
More minor fixes from Ollie

@@
 We have not included all the vulnerabilities from \avo. 
-In particular, we have not included any device manufacturer specific vulnerabilities, even when these are widespread (such as vulnerabilities in Qualcomm\footnote{Qualcomm is particularly good at publicly disclosing vulnerabilities and the patches which fix them.
+In particular, we have not included any device manufacturer specific vulnerabilities, even when these are widespread (such as vulnerabilities in Qualcomm\footnote{Qualcomm is particularly good at publicly disclosing vulnerabilities and the patches that fix them.
 \url{https://www.codeaurora.org/projects/security-advisories/}} chipsets) because it is difficult to work out which devices are affected.
@@
 
-Brumley et.\ al. has shown it is possible to automatically generate exploits from binary fixes, for example by using security updates issued by manufacturers~\cite{Brumley2008}. 
+Brumley et al.\ has shown it is possible to automatically generate exploits from binary fixes, for example by using security updates issued by manufacturers~\cite{Brumley2008}. 
 Similarly, the Woodpecker tool automatically finds permission leaks in stock Android phone images~\cite{Grace2012}.
@@
 There are continuing efforts to reduce the impact of critical vulnerabilities, both in Android and elsewhere.
-SEAndroid~\cite{Smalley2013} which is included in Android from version 4.1~\cite{jelly-bean-release} and fully enforcing from version 5.0~\cite{AndroidSecurity2014} claimed to prevent some root vulnerabilities and to reduce the impact of others.
+SEAndroid~\cite{Smalley2013}, which is included in Android from version 4.1~\cite{jelly-bean-release}, and fully enforcing from version 5.0~\cite{AndroidSecurity2014} claimed to prevent some root vulnerabilities and to reduce the impact of others.
 Capability based enforcement systems such as Capsicum~\cite{Watson2010} substantially reduce the capabilities that an exploit has to try and gain increased privilege with and could be included in Linux\footnote{\url{https://github.com/google/capsicum-linux}} and hence Android.
@@
 
-The percentage of Android devices running the most recent version (\daUpdatednessPerc) compares unfavourably with the rate ($>90$\%) for Windows XP SP2 computers contacting the Microsoft update servers~\cite{Gkantsidis2006}.
+The percentage of Android devices running the most recent version (\daUpdatednessPerc) is much less than the rate ($>90$\%) for Windows XP SP2 computers contacting the Microsoft update servers~\cite{Gkantsidis2006}.
 A simple numerical comparison is unfair because only one major OS version was considered in the Microsoft analysis, and data was only collected from computers which contacted the update server, although this was the default.

### 2015-04-13 ccs.tex
Lets start with a clean Acknowledgements section for CCS

@@
 \section*{Acknowledgements}
-Thanks to Richard Clayton, the mobile security reading group and Anil Madhavapeddy for reading various drafts of this paper.
-Thanks to David Robertson for helpful advice on statistical analysis.
-Thanks to Laurent Simon, Thomas Coudray, Adrian Taylor, Justin Case, Giant Pune and Khilan Gudka for reporting vulnerabilities in Android.
 }

### 2015-04-16 figures/update_figures.sh
Update graphs and stats from jenkins built data on nas04 rather than local

@@ -2,12 +2,12 @@
 
 set -e
 
-fromdir=~drt24/git/da/graphing/pdf/
+fromdir=/mnt/nas04-device-analyzer-graphing/pdf/
 cp "$fromdir"norm_api.pdf da_norm_api.pdf
 cp "$fromdir"norm_api-notitle.pdf da_norm_api-notitle.pdf
 cp "$fromdir"norm_os.pdf da_norm_os.pdf
 cp "$fromdir"os.pdf da_os.pdf
-cp "$fromdir"device_data/device-data-manufacturer-lg-security.pdf device_data-LG.pdf
+cp "$fromdir"device_data/device-data-manufacturer-LG-security.pdf device_data-LG.pdf
 cp "$fromdir"device_data/device-data-all-security.pdf device-data-all-security.pdf
 cp "$fromdir"device_data/device-data-all-os.pdf device-data-all-os.pdf
 cp "$fromdir"from_to_updates.pdf from_to_updates.pdf
@@ -26,8 +26,8 @@ cp "$fromdir"wifi-accounts-ototals-hist.pdf wifi-accounts-ototals-hist.pdf
 cp "$fromdir"wifi-accounts-stotals-hist.pdf wifi-accounts-stotals-hist.pdf
 cp "$fromdir"api_gpcomp_diff.pdf api_gpcomp_diff.pdf
 cp "$fromdir"api_gpcomp_rdiff.pdf api_gpcomp_rdiff.pdf
-cp "$fromdir"ua-versions-android-month-norm.pdf ua-versions-android-month-norm.pdf
-cp "$fromdir"ua-versions-ios-month-norm.pdf ua-versions-ios-month-norm.pdf
+# cp "$fromdir"ua-versions-android-month-norm.pdf ua-versions-android-month-norm.pdf
+# cp "$fromdir"ua-versions-ios-month-norm.pdf ua-versions-ios-month-norm.pdf
 cp "$fromdir""model/Galaxy Nexus/norm_full.pdf" "full_version_Galaxy Nexus.pdf"
 cp "$fromdir""model/HTC Desire HD A9191/norm_full.pdf" "full_version_HTC Desire HD A9191.pdf"
 cp "$fromdir""model/Symphony W68/norm_full.pdf" "full_version_Symphony W68.pdf"
@@ -36,5 +36,5 @@ cp "$fromdir"frvh_os_versions.pdf frvh_os_versions.pdf
[…]

### 2015-04-16 ccs.tex
Include proportioninsecure graph in ccs paper as referenced there
Uncomment text from ccs paper

@@
 We present a better understanding of both the ecosystem of Android development and associated vulnerabilities more fully in \S\ref{sec:android_ecosystem}.
+
+\begin{figure}
+\centering
+\includegraphics[width=\columnwidth]{figures/proportioninsecure}
+\caption{Proportion of devices running insecure, maybe secure and secure versions of Android against time.
+The red vertical lines are caused by vulnerabilities being discovered with those which have the biggest impact annotated.
+}
+\label{fig:proportioninsecure}
+\end{figure}
 
@@
 Anecdotal evidence at least suggests that it is the lack of updates rather the lack of installation which is the major problem at present. Further work is required to tease these numbers apart.
-%Partly this is the result of the fact that an operating system update is being installed and so a reboot is required, but Chrome installs the new version side by side with the old one and switches the next time it is restarted.
-%The same technique would be more difficult on phones with limited storage space (as many cheap Android phones have barely enough space to install just the update) but is a plausible improvement for more high-end devices.
-%Google is deploying the same silent update technique through Google Play Services\footnote{\href{http://lifehacker.com/why-google-play-services-are-now-more-important-than-an-975970197}{http://lifehacker.com/why-google-play-services-are-now-more-important-than-an-975970197}} which automatically installs updates for core Google components of Android, this also bypasses the device manufacturer and network operator.
+Partly this is the result of the fact that an operating system update is being installed and so a reboot is required, but Chrome installs the new version side by side with the old one and switches the next time it is restarted.
+The same technique would be more difficult on phones with limited storage space (as many cheap Android phones have barely enough space to install just the update) but is a plausible improvement for more high-end devices.
+Google is deploying the same silent update technique through Google Play Services\footnote{\href{http://lifehacker.com/why-google-play-services-are-now-more-important-than-an-975970197}{http://lifehacker.com/why-google-play-services-are-now-more-important-than-an-975970197}} which automatically installs updates for core Google components of Android, this also bypasses the device manufacturer and network operator.
 

### 2015-04-24 wisec2015abstract.tex
Import HotMobile 2011 abstract into wisec abstract

\documentclass{sig-alternate}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}

\include{dastats}
\newcommand{\da}{Device Analyzer}

\numberofauthors{4}
\author{Daniel~R.~Thomas \and Daniel~T.~Wagner \and Alastair~R.~Beresford \and Andrew~Rice\\       \email{Firstname.Lastname@cl.cam.ac.uk}
}
\title{POSTER: \da :\\Measuring the security of smartphones}
\conferenceinfo{WiSec}{2015 New York City, USA}
\begin{document}
\maketitle

New hardware capabilities and integrated application market stores have created complex usage patterns in modern smart phones.
Understanding these patterns is important to both smart phone providers and their customers.
Hardware manufacturers might make use of this data when considering design trade-offs of functionality against cost and form-factor.
Summarised data of handset use is also useful for device owners trying to stay within data limits on mobile networks or even to select the best tariff.
The idea of personal analytics takes this to an extreme in which one collects detailed statistics of many life aspects.

Our Device Analyzer application aims to collect a large-scale research data-set of phone usage.
Device Analyzer gathers data about running processes, wireless connectivity, the phone’s location, GSM communication, battery state and a number of system parameters.
We want to extend the range of data collected as well as the breadth of the targeted population.
Today’s smart phones are able to capture much richer data than was available some years ago while at the same time the Android Market allows us to distribute the app to a broader audience than was possible before.

\section{Data Collection}
Device Analyzer makes use of events from the Android operating system to record many changes in system state.
However, events are not available for some types of data, such as process information or network traffic, and so these are gathered by polling.
An expressive event mechanism could eliminate the need for polling entirely.
In the case of network traffic, for example, one might request an event for every kilobyte of network activity.
All data are stored locally in a timestamped key-value store for periodic upload to a central server.
By default Device Analyzer attempts to minimise the impact of this by scheduling uploads when the device is charging with a 802.11 connection available.

\section{Timestamps}
Simple event timestamping using the current system time is inadequate in a number of cases.
System time is subject to discontinuities, such as when the user manually changes the time on their phone or when the cell network broadcasts a correction.
Furthermore, local time is not always available: we see that for some seconds after a system upgrade the returned dates are in the 1980s!
To provide dependable timestamps in the face of such disruptions, we use the system uptime as our reference frame as it is guaranteed to not jump backwards or forwards.
Changes to the mapping of uptime to local time are logged in the event stream.

\section{Viewing Data}
Data can be viewed on the phone itself and on the web after it was uploaded.
At the moment the data displayed on the phone is designed to show the user which kinds of data are collected and does not contain a history.
This will be addressed in future versions.
On the website users can compare their historical records with those of the entire population and get more detailed information, including graphs of network traffic and other variables.

\section{Hashing and Aliases}
Rather than uploading identifying user data such as phone numbers or 802.11 network addresses in plain text, they are hashed with a device-specific salt to maintain user privacy.
Users can choose to assign private nicknames to these anonymous values so that they can identify them when they log in to the website.
The dataset itself contains only the hashed values.

\section{Privacy Concerns}
We plan to periodically release the collected anonymous dataset where participants have given us permission to do so.
This raises privacy concerns that need to be addressed before such a comprehensive and potentially invasive dataset can be released.
As a first measure we will only publish data that was collected at least three months ago, giving participants time to view the data they contributed.
If they are concerned about having their data released they can opt-out of any further data collection and erase all unpublished data from our server.
Furthermore, while physical location of participants could be an asset of our dataset, such data would need to be handled with great care, as disclosure of even approximate home/work location pairs could lead to deanonymization of participants [2].

\section{Conclusion}
A rich dataset of smart phone usage for a large population has undeniable uses for device owners as well as providers and manufacturers.
However, collecting such detailed data raises privacy concerns even if the data are never uploaded if an adversary gains physical access to the device.
We are working to address privacy issues which arise sufficiently for the scope of a research dataset where informed consent can be assumed.
However, when considering a general deployment by a manufacturer, further protection may be required.
It is an open question at this time how such a protection could look like.
We plan to release Device Analyzer on the Android Market this year and are currently conducting a beta test.
Please do contact us if you wish to participate.

\section{Acknowledgements}
This work was supported by the University of Cambridge Computer Laboratory Premium Studentship scheme, a Google focussed research award and the EPSRC Standard Research Grant EP/P505445/1.

\end{document}

### 2015-04-24 security_scoring.tex
Attempt to make things work better with the single column format
Start a subsection on the sensitivity of security scores

@@
 A high value of $u$ could be achieved by only ever shipping one version but that would give low values for $f$ and $m$ (and not be attractive to new customers).
-A high value of $m$ could be achieved by focusing on only one device at a time and ensuring that gets updates but ignoring all others, but that would lower $f$ and $u$.
+A high value of $m$ could be achieved by focusing on only one device at a time and ensuring that gets updates but ignoring all others, but that would lower $f$ and $u$.%
 \footnote{One way to influence our scores would be to add additional devices to Device Analyzer which have good security, these would have to be real end user devices as we could detect fake ones, this would increase the size of our data set and would require providing genuinely good security to some users.}
@@
  \centering
- \includegraphics[width=\textwidth]{figures/full_version_comp.pdf}
+ \includegraphics[width=0.9\textwidth]{figures/full_version_comp.pdf}
  \caption{Full version distributions for the highest and lowest scoring model}
@@
 We do not attempt to disambiguate users behaviour in whether they install updates from network operators using rolling upgrades.
+
+\subsubsection{Sensitivity of scoring metric}
+\daTabDLDistances
+\daTabChangeInScores
+To evaluate whether the ranking of different manufacturers is sensitive to the form of the scoring metric we computed the normalised Damerau-Levenshtein distance~\cite{Bard2007} between the lists ordered using different forms of the scoring metric, this is shown in Table~\ref{tab:dl_distances}.
+The `equal' metric weights $f$, $u$ and $m$ equally rather than favouring $f$ and makes little difference.
+Changing the scoring metric also impacts the scores given for each entity Table~\ref{tab:change_in_scores} shows the mean impact on the scores.

### 2015-04-24 wisec2015abstract.tex
Add pagenumbering to wisec abstract
Add bibliography to wisec abstract

@@
 \documentclass{sig-alternate}
+\pagenumbering{arabic}
 \usepackage[utf8]{inputenc}
@@
 \usepackage{graphicx}
+
+\RequirePackage[date=terse, isbn=true, doi=true, url=false, urldate=iso8601, maxbibnames=9, backref=false, backend=bibtex, style=ieee]{biblatex}
+\addbibresource{securityupdates.bib}
+\renewcommand{\bibfont}{\small}
+
+\AtEveryBibitem{% Clean up the bibtex rather than editing it
+ \clearname{editor} % remove editors
+}
 
@@
 If they are concerned about having their data released they can opt-out of any further data collection and erase all unpublished data from our server.
-Furthermore, while physical location of participants could be an asset of our dataset, such data would need to be handled with great care, as disclosure of even approximate home/work location pairs could lead to deanonymization of participants [2].
+Furthermore, while physical location of participants could be an asset of our dataset, such data would need to be handled with great care, as disclosure of even approximate home/work location pairs could lead to deanonymization of participants~\cite{Golle2009}.
 

### 2015-04-24 wisec2015abstract.tex
Cut out irrlevant stuff and update other content in wisec abstract

@@
 
-New hardware capabilities and integrated application market stores have created complex usage patterns in modern smart phones.
-Understanding these patterns is important to both smart phone providers and their customers.
-Hardware manufacturers might make use of this data when considering design trade-offs of functionality against cost and form-factor.
-Summarised data of handset use is also useful for device owners trying to stay within data limits on mobile networks or even to select the best tariff.
-The idea of personal analytics takes this to an extreme in which one collects detailed statistics of many life aspects.
+TODO
 
@@
 \section{Data Collection}
-Device Analyzer makes use of events from the Android operating system to record many changes in system state.
-However, events are not available for some types of data, such as process information or network traffic, and so these are gathered by polling.
-An expressive event mechanism could eliminate the need for polling entirely.
-In the case of network traffic, for example, one might request an event for every kilobyte of network activity.
-All data are stored locally in a timestamped key-value store for periodic upload to a central server.
-By default Device Analyzer attempts to minimise the impact of this by scheduling uploads when the device is charging with a 802.11 connection available.
+Device Analyzer makes use of events from the Android operating system to record many changes in system state and polls for information not provided by events.
+All data are stored locally in a timestamped key-value log file for periodic upload to a central server.
+By default Device Analyzer attempts to minimise the impact of this by scheduling uploads when the device is charging and has WiFi.
 
@@
 Furthermore, local time is not always available: we see that for some seconds after a system upgrade the returned dates are in the 1980s!
-To provide dependable timestamps in the face of such disruptions, we use the system uptime as our reference frame as it is guaranteed to not jump backwards or forwards.
-Changes to the mapping of uptime to local time are logged in the event stream.
+To provide dependable timestamps in the face of such disruptions, we use the system uptime as our reference frame as it is supposed to be guaranteed to not jump backwards or forwards.
+Changes to the mapping of uptime to local time are logged in the event stream and post processed to correct for errors~\cite{Wagner2013}.
 
@@
 Data can be viewed on the phone itself and on the web after it was uploaded.
-At the moment the data displayed on the phone is designed to show the user which kinds of data are collected and does not contain a history.
-This will be addressed in future versions.
-On the website users can compare their historical records with those of the entire population and get more detailed information, including graphs of network traffic and other variables.
+The data displayed on the phone is designed to show the user which kinds of data are collected and provides summary graphs.
 
 \section{Hashing and Aliases}
-Rather than uploading identifying user data such as phone numbers or 802.11 network addresses in plain text, they are hashed with a device-specific salt to maintain user privacy.
+Rather than uploading identifying user data such as phone numbers or network addresses in plain text, they are hashed with a device-specific salt to maintain user privacy.
 Users can choose to assign private nicknames to these anonymous values so that they can identify them when they log in to the website.
@@
 \section{Privacy Concerns}
-We plan to periodically release the collected anonymous dataset where participants have given us permission to do so.
-This raises privacy concerns that need to be addressed before such a comprehensive and potentially invasive dataset can be released.
-As a first measure we will only publish data that was collected at least three months ago, giving participants time to view the data they contributed.
+Have released release the collected anonymous dataset where participants have given us permission to do so to researchers at many organisations after their agreeing to terms which protect the privacy of the users.
+We only publish data that was collected at least three months ago, giving participants time to view the data they contributed.
 If they are concerned about having their data released they can opt-out of any further data collection and erase all unpublished data from our server.
-Furthermore, while physical location of participants could be an asset of our dataset, such data would need to be handled with great care, as disclosure of even approximate home/work location pairs could lead to deanonymization of participants~\cite{Golle2009}.
+Furthermore, while physical location of participants could be an asset of our dataset, such data needs to be handled with great care, as disclosure of even approximate home/work location pairs could lead to deanonymization of participants~\cite{Golle2009}.
+Hence we only provide the country the device was most often in to the researchers.
 
@@
 However, collecting such detailed data raises privacy concerns even if the data are never uploaded if an adversary gains physical access to the device.
-We are working to address privacy issues which arise sufficiently for the scope of a research dataset where informed consent can be assumed.
-However, when considering a general deployment by a manufacturer, further protection may be required.
-It is an open question at this time how such a protection could look like.
-We plan to release Device Analyzer on the Android Market this year and are currently conducting a beta test.
-Please do contact us if you wish to participate.
+We have tried to address privacy issues which arise so that informed consent can be assumed.
+Please do contact us if you wish to participate either by contributing data or conducting research using it.
 

### 2015-05-29 update.sh
Rename figures/update_figures.sh to update.sh

@@ -1,8 +1,9 @@
-#!/bin/sh
+#!/bin/bash
 
 set -e
 
 fromdir=/mnt/nas04-device-analyzer-graphing/pdf/
+pushd figures
 cp "$fromdir"norm_api.pdf da_norm_api.pdf
 cp "$fromdir"norm_api-notitle.pdf da_norm_api-notitle.pdf
 cp "$fromdir"norm_os.pdf da_norm_os.pdf
@@ -35,6 +36,7 @@ cp "$fromdir"vulnerabilities_timeline.pdf vulnerabilities_timeline.pdf
 cp "$fromdir"frvh_os_versions.pdf frvh_os_versions.pdf
 cp "$fromdir"frvh_api_versions.pdf frvh_api_versions.pdf
 cp "$fromdir"frms_os_versions.pdf frms_os_versions.pdf
-cp ~/Library/papers/library.filtered.bib ../securityupdates.bib
-cp /mnt/nas04-device-analyzer-graphing/stats/latex.tex ../dastats.tex
-cp ~/Documents/androidvulnerabilities/output/latex.tex ../avostats.tex
+popd
+cp ~/Library/papers/library.filtered.bib securityupdates.bib
+cp /mnt/nas04-device-analyzer-graphing/stats/latex.tex dastats.tex
+cp ~/Documents/androidvulnerabilities/output/latex.tex avostats.tex

### 2015-05-29 update.sh
Include security scoring over time graphs

@@ -36,6 +36,9 @@ cp "$fromdir"vulnerabilities_timeline.pdf vulnerabilities_timeline.pdf
 cp "$fromdir"frvh_os_versions.pdf frvh_os_versions.pdf
 cp "$fromdir"frvh_api_versions.pdf frvh_api_versions.pdf
 cp "$fromdir"frms_os_versions.pdf frms_os_versions.pdf
+cp "$fromdir"security_score_manufacturer.pdf security_score_manufacturer.pdf
+cp "$fromdir"security_score_model.pdf security_score_model.pdf
+cp "$fromdir"security_score_operator.pdf security_score_operator.pdf
 popd
 cp ~/Library/papers/library.filtered.bib securityupdates.bib
 cp /mnt/nas04-device-analyzer-graphing/stats/latex.tex dastats.tex

### 2015-05-29 dastats.tex
Updated stats and figures
Update figures and dastats
Updated graphs and dastats
./update.sh

--- 2014-11-14 dastats.tex
@@
-\newcommand{\daNumDataPoints}{153 Billion}
-\newcommand{\daNumStartedApps}{122\,000}
-\newcommand{\daNumInstalledApps}{175\,000}
-\newcommand{\daNumVulnsUsed}{11}
+\newcommand{\daNumSigModels}{17}
 \newcommand{\daSigNumDevices}{100}
+\newcommand{\daSigNumDeviceDays}{10\,000}
 \newcommand{\daSigNumDays}{100}
-\newcommand{\daNumManufacturers}{261}
-\newcommand{\daNumSigManufacturers}{9}
-\newcommand{\daNumModels}{2\,380}
-\newcommand{\daNumSigModels}{15}
-\newcommand{\daOSVersionPercValidLines}{99.5\%}
-\newcommand{\daOSTotalDaysData}{1\,140\,000}
-\newcommand{\daMeanInsecurityPerc}{$87.3 \pm 0.0\%$}
-\newcommand{\daMeanOutstandingVulnerabilities}{0.637}
-\newcommand{\daUpdatedness}{$0.0558 \pm 0.0002$}
-\newcommand{\daSecurityScore}{$2.75 \pm 0.00$}
-\newcommand{\daNumFullVersions}{1\,140}
-\newcommand{\daNumOSVersions}{45}
-\newcommand{\daNumSigOSVersions}{23}
-\newcommand{\daNumAPIVersions}{15}
-\newcommand{\daNumFullVersionUpdates}{4\,220}
-\newcommand{\daNumOSVersionUpdates}{3\,560}
-\newcommand{\daNumFullOnlyVersionUpdates}{640}
-\newcommand{\daNumSecurityUpdates}{1\,110}
-\newcommand{\daNumPossibleSecurityUpdates}{2\,140}
-\newcommand{\daTabSecScoresmanufacturer}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{Security score} \\ &&&& (out of 10) \\ \hline LG & $0.12 \pm 0.00$ & $0.32 \pm 0.00$ & 0.74 & $3.38 \pm 0.01$ \\  Motorola & $0.16 \pm 0.00$ & $0.13 \pm 0.00$ & 0.77 & $2.93 \pm 0.01$ \\  other & $0.13 \pm 0.00$ & $0.06 \pm 0.00$ & 0.64 & $2.75 \pm 0.00$ \\  Samsung & $0.14 \pm 0.00$ & $0.05 \pm 0.00$ & 0.71 & $2.67 \pm 0.00$ \\  Sony & $0.13 \pm 0.00$ & $0.18 \pm 0.00$ & 1.02 & $2.64 \pm 0.01$ \\  HTC & $0.14 \pm 0.00$ & $0.09 \pm 0.00$ & 0.89 & $2.55 \pm 0.00$ \\  asus & $0.15 \pm 0.00$ & $0.56 \pm 0.01$ & 6.06 & $2.30 \pm 0.02$ \\  Symphony & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & 4.79 & $0.19 \pm 0.00$ \\  walton & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & 5.79 & $0.14 \pm 0.01$ \\ \end{tabular} \caption{Security scores for manufacturers} \label{tab:sec_manufacturer} \end{table}}
-\newcommand{\daTabSecScoresmodel}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{Security score} \\ &&&& (out of 10) \\ \hline Galaxy Nexus & $0.53 \pm 0.00$ & $0.53 \pm 0.01$ & 1.53 & $4.80 \pm 0.03$ \\  Nexus 4 & $0.23 \pm 0.00$ & $0.82 \pm 0.01$ & 6.06 & $3.39 \pm 0.04$ \\  Nexus 7 & $0.19 \pm 0.00$ & $0.74 \pm 0.01$ & 5.92 & $3.02 \pm 0.04$ \\  other & $0.13 \pm 0.00$ & $0.12 \pm 0.00$ & 0.64 & $2.94 \pm 0.00$ \\  Desire HD & $0.08 \pm 0.00$ & $0.05 \pm 0.00$ & 0.39 & $2.91 \pm 0.02$ \\  HTC Sensation & $0.36 \pm 0.00$ & $0.01 \pm 0.01$ & 1.59 & $2.47 \pm 0.02$ \\  GT-I9100 & $0.22 \pm 0.00$ & $0.02 \pm 0.00$ & 1.20 & $2.33 \pm 0.01$ \\  HTC Desire S & $0.02 \pm 0.00$ & $0.02 \pm 0.00$ & 1.00 & $1.75 \pm 0.01$ \\  GT-N7000 & $0.25 \pm 0.00$ & $0.00 \pm 0.00$ & 2.52 & $1.43 \pm 0.02$ \\  GT-P1000 & $0.01 \pm 0.00$ & $0.00 \pm 0.01$ & 1.73 & $0.94 \pm 0.02$ \\  GT-I9300 & $0.15 \pm 0.00$ & $0.01 \pm 0.00$ & 6.15 & $0.64 \pm 0.01$ \\  GT-I9505 & $0.00 \pm 0.00$ & $0.18 \pm 0.00$ & 6.77 & $0.55 \pm 0.01$ \\  GT-N7100 & $0.07 \pm 0.00$ & $0.00 \pm 0.01$ & 6.91 & $0.28 \pm 0.02$ \\  HTC Desire HD & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & 3.05 & $0.27 \pm 0.02$ \\ \end{tabular} \caption{Security scores for models} \label{tab:sec_model} \end{table}}
-\newcommand{\daTabSecScoressummary}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{Security score} \\ &&&& (out of 10) \\ \hline nexus & $0.36 \pm 0.00$ & $0.51 \pm 0.00$ & 0.68 & $5.01 \pm 0.01$ \\  notnexus & $0.12 \pm 0.00$ & $0.04 \pm 0.00$ & 0.64 & $2.67 \pm 0.00$ \\ \end{tabular} \caption{Security scores for nexus} \label{tab:sec_summary} \end{table}}
-\newcommand{\daUpdatednessPerc}{$5.58 \pm 0.02\%$}
-\newcommand{\daNumOSDevices}{19\,000}
 \newcommand{\daSigNumDevicesDay}{20}
-\newcommand{\daSigNumDeviceDays}{10\,000}
-\newcommand{\daTabAndVulns}{\begin{table} \centering \begin{tabular}{l|c|l} Vulnerability & Date known & How known \\ \hline KillingInTheNameOf psneuter ashmem & 2010-07-13 & Fixed on \\ exploid udev & 2010-07-15 & Discovered on \\ RageAgainstTheCage adb & 2010-08-21 & Discovered on \\ levitator & 2011-03-10 & Discovered on \\ Gingerbreak & 2011-04-18 & Fixed on \\ zergRush & 2011-10-06 & Discovered on \\ APK duplicate file & 2013-02-18 & Discovered on \\ APK unchecked name & 2013-06-30 & Discovered on \\ APK unsigned shorts & 2013-07-03 & Fixed on \\ keystore buffer & 2013-09-09 & Discovered on \\ TwerkMyMoto & 2013-11-24 & Discovered on \\ vold asec & 2014-01-27 & Fixed on \\\end{tabular} \caption{Root equivalent vulnerabilities in Android} \label{tab:andvulns} \end{table}}
-\newcommand{\daOSYearsOfData}{3}
-\newcommand{\daSecScoreBestmanufacturer}{LG}
-\newcommand{\daSecScoreWorstmanufacturer}{walton}
-\newcommand{\daSecScoreBestmodel}{Galaxy Nexus}
-\newcommand{\daSecScoreWorstmodel}{HTC Desire HD A9191}
-\newcommand{\daSecScoreBestsummary}{nexus}
-\newcommand{\daSecScoreWorstsummary}{notnexus}
-\newcommand{\daSecScoreBestmanufacturerScore}{$3.38 \pm 0.01$}
-\newcommand{\daSecScoreWorstmanufacturerScore}{$0.139 \pm 0.006$}
-\newcommand{\daSecScoreBestmodelScore}{$4.8 \pm 0.0$}
-\newcommand{\daSecScoreWorstmodelScore}{$0.274 \pm 0.020$}
-\newcommand{\daSecScoreBestsummaryScore}{$4.1 \pm 0.0$}
-\newcommand{\daSecScoreWorstsummaryScore}{$2.9 \pm 0.0$}
-\newcommand{\daVulnFree}{$0.127 \pm 0.000$}
-\newcommand{\daAdbEnabledPerc}{$20 \pm 0\%$}
-\newcommand{\daNumUpdatesUpgrades}{3\,330}
-\newcommand{\daNumUpdatesDowngrades}{147}
-\newcommand{\daPercUpdatesDowngrades}{4.23\%}
-\newcommand{\daStartDate}{2011-07-01}
-\newcommand{\daEndDate}{2014-11-07}
-\newcommand{\daNumOperators}{1\,390}
-\newcommand{\daTabSecScoresoperator}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{Security score} \\ &&&& (out of 10) \\ \hline O2 uk & $0.25 \pm 0.00$ & $0.12 \pm 0.00$ & 0.37 & $3.79 \pm 0.01$ \\  T-Mobile & $0.19 \pm 0.00$ & $0.18 \pm 0.00$ & 0.48 & $3.57 \pm 0.01$ \\  Orange & $0.21 \pm 0.00$ & $0.09 \pm 0.00$ & 0.39 & $3.52 \pm 0.02$ \\  Sprint & $0.18 \pm 0.00$ & $0.11 \pm 0.00$ & 0.44 & $3.41 \pm 0.01$ \\  3 & $0.15 \pm 0.00$ & $0.09 \pm 0.00$ & 0.49 & $3.15 \pm 0.01$ \\  AT\&T & $0.12 \pm 0.00$ & $0.08 \pm 0.00$ & 0.43 & $3.08 \pm 0.01$ \\  Vodafone uk & $0.13 \pm 0.00$ & $0.10 \pm 0.00$ & 0.52 & $3.07 \pm 0.01$ \\  Verizon & $0.18 \pm 0.00$ & $0.09 \pm 0.00$ & 0.81 & $2.82 \pm 0.01$ \\  unknown & $0.10 \pm 0.00$ & $0.22 \pm 0.00$ & 1.08 & $2.57 \pm 0.01$ \\  n Telenor & $0.03 \pm 0.00$ & $0.10 \pm 0.00$ & 1.42 & $1.59 \pm 0.01$ \\  Airtel & $0.06 \pm 0.00$ & $0.03 \pm 0.00$ & 1.93 & $1.10 \pm 0.01$ \\  Grameenphone & $0.01 \pm 0.00$ & $0.02 \pm 0.00$ & 1.99 & $0.80 \pm 0.00$ \\  Robi & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & 2.19 & $0.73 \pm 0.01$ \\  banglalink & $0.00 \pm 0.00$ & $0.03 \pm 0.00$ & 2.90 & $0.40 \pm 0.00$ \\ \end{tabular} \caption{Security scores for operators} \label{tab:sec_operator} \end{table}}
-\newcommand{\daSecScoreBestoperator}{O2 uk}
-\newcommand{\daSecScoreBestoperatorScore}{$3.79 \pm 0.01$}
-\newcommand{\daSecScoreWorstoperator}{banglalink}
-\newcommand{\daSecScoreWorstoperatorScore}{$0.4 \pm 0.0$}
-\newcommand{\daNumSigOperators}{14}
-\newcommand{\daMonthsDevices}{1545}
-\newcommand{\daMonths}{6}
 \newcommand{\daSigVersionPerc}{1\%}
 \newcommand{\daSigVersionDays}{10}
+\newcommand{\daNumVulnsUsed}{11}
 \newcommand{\daNumDeviceDataDevices}{50}
-\newcommand{\daNumSigFullVersions}{83}
-\newcommand{\daSecScoreBestmanufacturerNumFullVersions}{104}
-\newcommand{\daSecScoreworstmanufacturerNumFullVersions}{239}
+\newcommand{\daOSYearsOfData}{3}
+\newcommand{\daOSMonthsOfData}{47}
+\newcommand{\daStartDate}{2011-07-01}
+\newcommand{\daEndDate}{2015-05-21}
+\newcommand{\daFullDeployedAt}{95\%}
+\newcommand{\daNumManufacturers}{298}
+\newcommand{\daNumSigManufacturers}{8}
+\newcommand{\daNumModels}{2\,690}
+\newcommand{\daNumOperators}{1\,440}
+\newcommand{\daNumSigOperators}{14}
+\newcommand{\daNumOSDevices}{20\,300}
+\newcommand{\daOSVersionPercValidLines}{99.7\%}
+\newcommand{\daOSTotalDaysData}{1\,290\,000}
+\newcommand{\daMeanInsecurityPercNominal}{87.9\%}
+\newcommand{\daMeanInsecurityPerc}{$87.9 \pm 0.0\%$}
+\newcommand{\daMeanInsecurityPercTwosfNominal}{88\%}
+\newcommand{\daMeanInsecurityPercTwosf}{$88 \pm 0\%$}
+\newcommand{\daMeanSecurityPercNominal}{12.1\%}
+\newcommand{\daMeanSecurityPerc}{$12.1 \pm 0.0\%$}
+\newcommand{\daMeanSecurityPercTwosfNominal}{12\%}
+\newcommand{\daMeanSecurityPercTwosf}{$12 \pm 0\%$}
+\newcommand{\daVulnFreeNominal}{0.121}
+\newcommand{\daVulnFree}{$0.121 \pm 0.000$}
+\newcommand{\daMeanOutstandingVulnerabilities}{0.551}
+\newcommand{\daUpdatednessNominal}{0.0512}
+\newcommand{\daUpdatedness}{$0.0512 \pm 0.0002$}
+\newcommand{\daUpdatednessPercNominal}{5.12\%}
+\newcommand{\daUpdatednessPerc}{$5.12 \pm 0.02\%$}
+\newcommand{\daUpdatednessPercTwosfNominal}{5.1\%}
+\newcommand{\daUpdatednessPercTwosf}{$5.1 \pm 0.0\%$}
+\newcommand{\daSecurityScoreNominal}{2.83}
+\newcommand{\daSecurityScore}{$2.83 \pm 0.00$}
+\newcommand{\daCurrentMeanInsecurityPercNominal}{87.0\%}
+\newcommand{\daCurrentMeanInsecurityPerc}{$87.0 \pm 4.9\%$}
+\newcommand{\daCurrentMeanInsecurityPercTwosfNominal}{87\%}
+\newcommand{\daCurrentMeanInsecurityPercTwosf}{$87 \pm 4\%$}
+\newcommand{\daCurrentMeanSecurityPercNominal}{13.0\%}
+\newcommand{\daCurrentMeanSecurityPerc}{$13.0 \pm 4.9\%$}
+\newcommand{\daCurrentMeanSecurityPercTwosfNominal}{13\%}
+\newcommand{\daCurrentMeanSecurityPercTwosf}{$13 \pm 4\%$}
+\newcommand{\daCurrentVulnFreeNominal}{0.13}
+\newcommand{\daCurrentVulnFree}{$0.13 \pm 4.92$}
+\newcommand{\daCurrentMeanOutstandingVulnerabilities}{0.213}
+\newcommand{\daCurrentUpdatednessNominal}{0.0327}
+\newcommand{\daCurrentUpdatedness}{$0.0327 \pm 0.0006$}
+\newcommand{\daCurrentUpdatednessPercNominal}{3.27\%}
+\newcommand{\daCurrentUpdatednessPerc}{$3.27 \pm 0.06\%$}
+\newcommand{\daCurrentUpdatednessPercTwosfNominal}{3.3\%}
+\newcommand{\daCurrentUpdatednessPercTwosf}{$3.3 \pm 0.0\%$}
+\newcommand{\daCurrentSecurityScoreNominal}{3.3}
+\newcommand{\daCurrentSecurityScore}{$3.3 \pm 0.0$}
+\newcommand{\daDailyParticipationNominal}{906}
+\newcommand{\daDailyParticipation}{$906 \pm 224$}
+\newcommand{\daNumFullVersions}{1\,250}
+\newcommand{\daNumSigFullVersions}{94}
+\newcommand{\daNumOSVersions}{50}
+\newcommand{\daNumSigOSVersions}{27}
+\newcommand{\daNumAPIVersions}{16}
+\newcommand{\daNumFullVersionUpdates}{4\,520}
+\newcommand{\daNumOSVersionUpdates}{3\,830}
+\newcommand{\daUpdatesPerYearNominal}{1.28}
+\newcommand{\daUpdatesPerYear}{$1.28 \pm 0.01$}
+\newcommand{\daUpdatesPerYearTwosfNominal}{1.3}
+\newcommand{\daUpdatesPerYearTwosf}{$1.3 \pm 0.0$}
+\newcommand{\daNumFullOnlyVersionUpdates}{673}
+\newcommand{\daNumSecurityUpdates}{1\,240}
+\newcommand{\daNumPossibleSecurityUpdates}{2\,290}
+\newcommand{\daOSCurveFitParamFirst}{84.9}
+\newcommand{\daOSCurveFitParamSecond}{\num{0.0029}}
+\newcommand{\daOSCurveFitRMSE}{0.116}
+\newcommand{\daOSCurvePolyRMSE}{0.116}
+\newcommand{\daOSCurveSplineRMSE}{0.116}
+\newcommand{\daOSCurveHalfDeployedYearsNominal}{0.887}
+\newcommand{\daOSCurveHalfDeployedYears}{$0.887 \pm 0.248$}
+\newcommand{\daOSCurveHalfDeployedNominal}{324}
+\newcommand{\daOSCurveHalfDeployed}{$324 \pm 90$}
+\newcommand{\daOSCurveFullDeployedYearsNominal}{3.06}
+\newcommand{\daOSCurveFullDeployedYears}{$3.06 \pm 70$}
+\newcommand{\daOSCurveFullDeployedNominal}{1\,120}
+\newcommand{\daOSCurveFullDeployed}{$1\,120 \pm 255\,000$}
+\newcommand{\daAPICurveFitParamFirst}{102}
+\newcommand{\daAPICurveFitParamSecond}{\num{0.0029}}
+\newcommand{\daAPICurveFitRMSE}{0.113}
+\newcommand{\daAPICurvePolyRMSE}{0.114}
+\newcommand{\daAPICurveSplineRMSE}{0.114}
+\newcommand{\daAPICurveHalfDeployedYearsNominal}{0.933}
+\newcommand{\daAPICurveHalfDeployedYears}{$0.933 \pm 0.243$}
+\newcommand{\daAPICurveHalfDeployedNominal}{341}
+\newcommand{\daAPICurveHalfDeployed}{$341 \pm 88$}
+\newcommand{\daAPICurveFullDeployedYearsNominal}{3.11}
+\newcommand{\daAPICurveFullDeployedYears}{$3.11 \pm 70$}
+\newcommand{\daAPICurveFullDeployedNominal}{1\,130}
+\newcommand{\daAPICurveFullDeployed}{$1\,130 \pm 256\,000$}
+\newcommand{\daVulnAPKDuplicateFileOctoberPerc}{91.6\%}
+\newcommand{\daVulnZergRushMonthsDefFixDeployed}{27.3}
+\newcommand{\daTabSecScoresmanufacturer}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline LG & $0.20 \pm 0.00$ & $0.32 \pm 0.00$ & 0.64 & $3.83 \pm 0.01$ \\  Motorola & $0.17 \pm 0.00$ & $0.12 \pm 0.00$ & 0.73 & $3.03 \pm 0.01$ \\  Samsung & $0.13 \pm 0.00$ & $0.04 \pm 0.00$ & 0.63 & $2.71 \pm 0.00$ \\  Sony & $0.13 \pm 0.00$ & $0.19 \pm 0.00$ & 1.09 & $2.61 \pm 0.01$ \\  HTC & $0.14 \pm 0.00$ & $0.09 \pm 0.00$ & 0.89 & $2.58 \pm 0.00$ \\  asus & $0.19 \pm 0.00$ & $0.52 \pm 0.01$ & 6.01 & $2.31 \pm 0.02$ \\  other & $0.06 \pm 0.00$ & $0.04 \pm 0.00$ & 1.04 & $1.94 \pm 0.00$ \\  Symphony & $0.00 \pm 0.00$ & $0.07 \pm 0.00$ & 5.00 & $0.27 \pm 0.01$ \\  walton & $0.00 \pm 0.00$ & $0.07 \pm 0.00$ & 5.97 & $0.24 \pm 0.01$ \\ \end{tabular} \caption{Security scores for manufacturers} \label{tab:sec_manufacturer} \end{table}}
+\newcommand{\daSecScoreBestmanufacturer}{LG}
+\newcommand{\daSecScoreBestmanufacturerScoreNominal}{3.83}
+\newcommand{\daSecScoreBestmanufacturerScore}{$3.83 \pm 0.01$}
+\newcommand{\daSecScoreBestmanufacturerNumFullVersions}{129}
+\newcommand{\daSecScoreWorstmanufacturer}{walton}
+\newcommand{\daSecScoreWorstmanufacturerScoreNominal}{0.236}
+\newcommand{\daSecScoreWorstmanufacturerScore}{$0.236 \pm 0.007$}
+\newcommand{\daSecScoreWorstmanufacturerNumFullVersions}{22}
+\newcommand{\daTabSecScoresmodel}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline Galaxy Nexus & $0.51 \pm 0.00$ & $0.55 \pm 0.01$ & 1.53 & $4.74 \pm 0.03$ \\  Nexus 4 & $0.28 \pm 0.00$ & $0.82 \pm 0.01$ & 6.06 & $3.59 \pm 0.04$ \\  Nexus 7 & $0.24 \pm 0.00$ & $0.74 \pm 0.01$ & 5.92 & $3.18 \pm 0.04$ \\  other & $0.09 \pm 0.00$ & $0.14 \pm 0.00$ & 0.55 & $2.97 \pm 0.00$ \\  Desire HD & $0.08 \pm 0.00$ & $0.05 \pm 0.00$ & 0.39 & $2.91 \pm 0.01$ \\  HTC Sensation & $0.35 \pm 0.00$ & $0.01 \pm 0.01$ & 1.59 & $2.42 \pm 0.02$ \\  GT-I9100 & $0.22 \pm 0.00$ & $0.02 \pm 0.00$ & 1.23 & $2.28 \pm 0.01$ \\  HTC Desire S & $0.02 \pm 0.00$ & $0.02 \pm 0.00$ & 1.00 & $1.75 \pm 0.01$ \\  GT-N7000 & $0.25 \pm 0.00$ & $0.00 \pm 0.00$ & 2.52 & $1.43 \pm 0.02$ \\  GT-P1000 & $0.01 \pm 0.00$ & $0.00 \pm 0.01$ & 1.79 & $0.90 \pm 0.02$ \\  GT-I9300 & $0.13 \pm 0.00$ & $0.01 \pm 0.00$ & 6.25 & $0.59 \pm 0.01$ \\  GT-I9505 & $0.01 \pm 0.00$ & $0.13 \pm 0.00$ & 6.82 & $0.44 \pm 0.01$ \\  HTC Desire HD & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & 3.05 & $0.27 \pm 0.02$ \\  GT-N7100 & $0.06 \pm 0.00$ & $0.00 \pm 0.01$ & 6.91 & $0.25 \pm 0.02$ \\  Symphony W68 & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & 11.00 & $0.00 \pm 0.03$ \\ \end{tabular} \caption{Security scores for models} \label{tab:sec_model} \end{table}}
+\newcommand{\daSecScoreBestmodel}{Galaxy Nexus}
+\newcommand{\daSecScoreBestmodelScoreNominal}{4.74}
+\newcommand{\daSecScoreBestmodelScore}{$4.74 \pm 0.02$}
 \newcommand{\daSecScoreBestmodelNumFullVersions}{49}
-\newcommand{\daSecScoreworstmodelNumFullVersions}{2}
-\newcommand{\daSecScoreBestsummaryNumFullVersions}{73}
-\newcommand{\daSecScoreworstsummaryNumFullVersions}{1057}
-\newcommand{\daSecScoreBestoperatorNumFullVersions}{88}
-\newcommand{\daSecScoreworstoperatorNumFullVersions}{89}
-\newcommand{\daSecScoreWorstmanufacturerNumFullVersions}{20}
-\newcommand{\daSecScoreWorstmodelNumFullVersions}{2}
-\newcommand{\daSecScoreWorstsummaryNumFullVersions}{1103}
-\newcommand{\daSecScoreWorstoperatorNumFullVersions}{59}
-\newcommand{\daFullDeployedAt}{95\%}
-\newcommand{\daOSCurveFitParamFirst}{87.5}
-\newcommand{\daOSCurveFitParamSecond}{\num{0.00315}}
-\newcommand{\daOSCurveFitRMSE}{0.114}
-\newcommand{\daAPICurveFitParamFirst}{110}
-\newcommand{\daAPICurveFitParamSecond}{\num{0.00315}}
-\newcommand{\daAPICurveFitRMSE}{0.116}
-\newcommand{\daOSCurvePolyRMSE}{0.115}
-\newcommand{\daOSCurveSplineRMSE}{0.115}
-\newcommand{\daOSCurveHalfDeployed}{$308 \pm 81$}
-\newcommand{\daOSCurveFullDeployed}{$1\,040 \pm 236\,000$}
-\newcommand{\daAPICurvePolyRMSE}{0.117}
-\newcommand{\daAPICurveSplineRMSE}{0.117}
-\newcommand{\daAPICurveHalfDeployed}{$330 \pm 84$}
-\newcommand{\daAPICurveFullDeployed}{$1\,060 \pm 236\,000$}
-\newcommand{\daVulnAPKDuplicateFileOctoberPerc}{91.5\%}
-\newcommand{\daNumUpdatesBigUpgrades}{778}
+\newcommand{\daSecScoreWorstmodel}{Symphony W68}
+\newcommand{\daSecScoreWorstmodelScoreNominal}{0.0001}
+\newcommand{\daSecScoreWorstmodelScore}{$0.0001 \pm 0.0276$}
+\newcommand{\daSecScoreWorstmodelNumFullVersions}{1}
+\newcommand{\daTabSecScoressummary}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline nexus & $0.38 \pm 0.00$ & $0.48 \pm 0.00$ & 0.59 & $5.08 \pm 0.01$ \\  notnexus & $0.10 \pm 0.00$ & $0.02 \pm 0.00$ & 0.55 & $2.66 \pm 0.00$ \\ \end{tabular} \caption{Security scores for nexus} \label{tab:sec_summary} \end{table}}
+\newcommand{\daTabSecScoresoperator}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline O2 uk & $0.26 \pm 0.00$ & $0.12 \pm 0.00$ & 0.37 & $3.84 \pm 0.01$ \\  T-Mobile & $0.20 \pm 0.00$ & $0.18 \pm 0.00$ & 0.42 & $3.73 \pm 0.01$ \\  Orange & $0.21 \pm 0.00$ & $0.09 \pm 0.00$ & 0.39 & $3.52 \pm 0.02$ \\  Sprint & $0.18 \pm 0.00$ & $0.11 \pm 0.00$ & 0.40 & $3.46 \pm 0.01$ \\  3 & $0.18 \pm 0.00$ & $0.09 \pm 0.00$ & 0.52 & $3.23 \pm 0.01$ \\  Vodafone uk & $0.13 \pm 0.00$ & $0.13 \pm 0.00$ & 0.50 & $3.18 \pm 0.01$ \\  AT\&T & $0.12 \pm 0.00$ & $0.08 \pm 0.00$ & 0.43 & $3.09 \pm 0.01$ \\  Verizon & $0.18 \pm 0.00$ & $0.09 \pm 0.00$ & 0.82 & $2.83 \pm 0.01$ \\  unknown & $0.11 \pm 0.00$ & $0.20 \pm 0.00$ & 0.87 & $2.80 \pm 0.01$ \\  n Telenor & $0.03 \pm 0.00$ & $0.12 \pm 0.00$ & 1.25 & $1.81 \pm 0.01$ \\  Airtel & $0.05 \pm 0.00$ & $0.03 \pm 0.00$ & 1.62 & $1.27 \pm 0.01$ \\  Robi & $0.00 \pm 0.00$ & $0.07 \pm 0.00$ & 2.08 & $0.88 \pm 0.01$ \\  Grameenphone & $0.00 \pm 0.00$ & $0.05 \pm 0.00$ & 2.03 & $0.85 \pm 0.00$ \\  banglalink & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & 2.62 & $0.53 \pm 0.01$ \\ \end{tabular} \caption{Security scores for operators} \label{tab:sec_operator} \end{table}}
+\newcommand{\daSecScoreBestoperator}{O2 uk}
+\newcommand{\daSecScoreBestoperatorScoreNominal}{3.84}
+\newcommand{\daSecScoreBestoperatorScore}{$3.84 \pm 0.01$}
+\newcommand{\daSecScoreBestoperatorNumFullVersions}{95}
+\newcommand{\daSecScoreWorstoperator}{banglalink}
+\newcommand{\daSecScoreWorstoperatorScoreNominal}{0.531}
+\newcommand{\daSecScoreWorstoperatorScore}{$0.531 \pm 0.005$}
+\newcommand{\daSecScoreWorstoperatorNumFullVersions}{72}
+\newcommand{\daTabDLDistances}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & summary \\ \hline equal & 0.0 & 0.0 & 0.0714 & 0.0 \\weight $m$ & 0.222 & 0.333 & 0.143 & 0.0 \\weight $u$ & 0.444 & 0.133 & 0.214 & 0.0 \\$f$ & 0.444 & 0.733 & 0.429 & 0.0 \\$m$ & 0.667 & 0.733 & 0.357 & 0.5 \\$u$ & 0.889 & 0.667 & 0.714 & 0.0 \\\end{tabular} \caption{Normalised Damerau-Levenshtein distances for different metrics} \label{tab:dl_distances} \end{table}}
+\newcommand{\daTabChangeInScores}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & summary \\ \hline $m$ & $-1.91 \pm 2.21$ & $-0.666 \pm 2.5$ & $-3.27 \pm 1.38$ & $-3.35 \pm 1.84$ \\weight $m$ & $-0.295 \pm 0.273$ & $-0.102 \pm 0.314$ & $-0.458 \pm 0.176$ & $-0.485 \pm 0.209$ \\equal & $-0.116 \pm 0.065$ & $-0.0391 \pm 0.112$ & $-0.146 \pm 0.048$ & $-0.166 \pm 0.028$ \\weight $u$ & $-0.0522 \pm 0.127$ & $-0.0154 \pm 0.239$ & $0.0199 \pm 0.0787$ & $-0.0137 \pm 0.127$ \\$u$ & $0.518 \pm 1.6$ & $0.198 \pm 2.14$ & $1.51 \pm 0.95$ & $1.36 \pm 1.52$ \\$f$ & $1.04 \pm 0.58$ & $0.351 \pm 1.01$ & $1.31 \pm 0.44$ & $1.5 \pm 0.2$ \\\end{tabular} \caption{Mean change in scores for different metrics} \label{tab:change_in_scores} \end{table*}}
+\newcommand{\daGPAPICurveFitParamFirst}{80.7}
+\newcommand{\daGPAPICurveFitParamSecond}{\num{0.00259}}
+\newcommand{\daGPAPICurveFitRMSE}{0.18}
+\newcommand{\daGPAPICurvePolyRMSE}{0.18}
+\newcommand{\daGPAPICurveSplineRMSE}{0.18}
+\newcommand{\daGPAPICurveHalfDeployedYearsNominal}{0.954}
+\newcommand{\daGPAPICurveHalfDeployedYears}{$0.954 \pm 0.473$}
+\newcommand{\daGPAPICurveHalfDeployedNominal}{348}
+\newcommand{\daGPAPICurveHalfDeployed}{$348 \pm 173$}
+\newcommand{\daGPAPICurveFullDeployedYearsNominal}{3.39}
+\newcommand{\daGPAPICurveFullDeployedYears}{$3.39 \pm 78$}
+\newcommand{\daGPAPICurveFullDeployedNominal}{1\,240}
+\newcommand{\daGPAPICurveFullDeployed}{$1\,240 \pm 286\,000$}
+\newcommand{\daGPAPIPerAPIMeanNominal}{0.00453}
+\newcommand{\daGPAPIPerAPIMean}{$0.00453 \pm 0.00343$}
+\newcommand{\daGPAPIPerAPIStdevNominal}{0.0211}
+\newcommand{\daGPAPIPerAPIStdev}{$0.0211 \pm 0.0127$}
+\newcommand{\daGPAPIPerAPIAbsMeanNominal}{0.0144}
+\newcommand{\daGPAPIPerAPIAbsMean}{$0.0144 \pm 0.0066$}
+\newcommand{\daGPAPIPerAPIAbsStdevNominal}{0.0154}
+\newcommand{\daGPAPIPerAPIAbsStdev}{$0.0154 \pm 0.012$}
+\newcommand{\daGPAPIPerAPIParametersTable}{\begin{table} \centering \begin{tabular}{c|S|S} API Version & ${t_0}$ & $\mathrm{decay}$ \\ \hline2	&	0.0	&	0.0237\\3	&	0.0	&	0.0235\\4	&	0.0	&	0.00689\\5	&	59.1	&	0.00465\\6	&	0.0	&	0.0043\\7	&	0.0	&	0.00499\\8	&	47.7	&	0.00415\\9	&	155	&	0.00365\\10	&	90.2	&	0.00364\\11	&	432	&	0.00228\\12	&	355	&	0.00228\\13	&	304	&	0.00236\\14	&	218	&	0.00239\\15	&	160	&	0.00239\\16	&	142	&	0.00227\\17	&	288	&	0.00195\\18	&	119	&	0.00141\\19	&	119	&	0.00153\\\end{tabular} \caption{Parameters for different API versions} \label{tab:per_api_parameters} \end{table}}
+\newcommand{\daGPAPISeventeenLaterDate}{2015-05-04}
+\newcommand{\daGPAPISeventeenLaterProportion}{72.4\%}
+\newcommand{\daGPAPISeventeenEarlierProportion}{27.6\%}
+\newcommand{\daAPISeventeenReleaseDate}{2012-10-29}
+\newcommand{\daGPAPISeventeenFullDeployment}{2017-10-29}
+\newcommand{\daGPAPISeventeenCurveFitParamFirst}{288}
+\newcommand{\daGPAPISeventeenCurveFitParamSecond}{\num{0.00195}}
+\newcommand{\daGPAPISeventeenCurveFitRMSE}{0.0272}
+\newcommand{\daGPAPISeventeenCurvePolyRMSE}{0.00888}
+\newcommand{\daGPAPISeventeenCurveSplineRMSE}{0.00888}
+\newcommand{\daGPAPISeventeenCurveHalfDeployedYearsNominal}{1.77}
+\newcommand{\daGPAPISeventeenCurveHalfDeployedYears}{$1.77 \pm 0.07$}
+\newcommand{\daGPAPISeventeenCurveHalfDeployedNominal}{644}
+\newcommand{\daGPAPISeventeenCurveHalfDeployed}{$644 \pm 28$}
+\newcommand{\daGPAPISeventeenCurveFullDeployedYearsNominal}{5.01}
+\newcommand{\daGPAPISeventeenCurveFullDeployedYears}{$5.01 \pm 1.11$}
+\newcommand{\daGPAPISeventeenCurveFullDeployedNominal}{1\,830}
+\newcommand{\daGPAPISeventeenCurveFullDeployed}{$1\,830 \pm 404$}
+\newcommand{\daNumUpdatesUpgrades}{3\,590}
+\newcommand{\daUpdatesPerMonthPerVersionNominal}{2.83}
+\newcommand{\daUpdatesPerMonthPerVersion}{$2.83 \pm 0.04$}
 \newcommand{\daNumUpdatesSkippedBig}{3}
-\newcommand{\daPercBigUpgrades}{22.4\%}
-\newcommand{\daUpdatesPerYear}{$1.35 \pm 0.02$}
-\newcommand{\daOSMonthsOfData}{40}
-\newcommand{\daMeanInsecurityPercNominal}{87.3\%}
-\newcommand{\daVulnFreeNominal}{0.127}
-\newcommand{\daUpdatednessNominal}{0.0558}
-\newcommand{\daUpdatednessPercNominal}{5.58\%}
-\newcommand{\daSecurityScoreNominal}{2.75}
-\newcommand{\daUpdatesPerYearNominal}{1.35}
-\newcommand{\daOSCurveHalfDeployedNominal}{308}
-\newcommand{\daOSCurveFullDeployedNominal}{1\,040}
-\newcommand{\daAPICurveHalfDeployedNominal}{330}
-\newcommand{\daAPICurveFullDeployedNominal}{1\,060}
-\newcommand{\daSecScoreBestoperatorScoreNominal}{3.79}
-\newcommand{\daSecScoreWorstoperatorScoreNominal}{0.4}
-\newcommand{\daSecScoreBestmodelScoreNominal}{4.8}
-\newcommand{\daSecScoreWorstmodelScoreNominal}{0.274}
-\newcommand{\daAdbEnabledPercNominal}{19.6\%}
-\newcommand{\daSecScoreBestmanufacturerScoreNominal}{3.38}
-\newcommand{\daSecScoreWorstmanufacturerScoreNominal}{0.139}
-\newcommand{\daUpdatesPerMonthPerVersion}{3.62}
-\newcommand{\daNumUpdateFullOnly}{0}
+\newcommand{\daNumUpdatesBigUpgrades}{819}
+\newcommand{\daNumUpdatesDowngrades}{152}
+\newcommand{\daNumUpdateFullOnly}{670}
+\newcommand{\daPercBigUpgradesNominal}{18.6\%}
+\newcommand{\daPercBigUpgrades}{$18.6 \pm 0.7\%$}
+\newcommand{\daPercUpdatesDowngradesNominal}{3.45\%}
+\newcommand{\daPercUpdatesDowngrades}{$3.45 \pm 0.28\%$}
+\newcommand{\daNumDataPoints}{157 Billion}
+\newcommand{\daMonthsDevices}{1\,780}
+\newcommand{\daMonths}{6}
+\newcommand{\daAdbEnabledPercNominal}{19.2\%}
+\newcommand{\daAdbEnabledPerc}{$19.2 \pm 0.0\%$}
+\newcommand{\daNumStartedApps}{126\,000}
+\newcommand{\daNumInstalledApps}{180\,000}

### 2015-06-01 security_scoring.tex
Include security scoring over time graphs
Provide text for security scoring over time section

@@
 Changing the scoring metric also impacts the scores given for each entity Table~\ref{tab:change_in_scores} shows the mean impact on the scores.
+
+\subsubsection{Scores over time}
+\begin{figure}
+\centering
+\includegraphics[width=\columnwidth]{figures/security_score_manufacturer}
+\caption{Security scores for different manufacturers}
+\label{fig:security_score_manufacturer}
+\includegraphics[width=\columnwidth]{figures/security_score_model}
+\caption{Security scores for different device models}
+\label{fig:security_score_model}
+\includegraphics[width=\columnwidth]{figures/security_score_operator}
+\caption{Security scores for different operators}
+\label{fig:security_score_operator}\\
+\includegraphics[width=\columnwidth]{figures/security_score_summary}
+\caption{Security scores for Nexus and non-Nexus devices}
+\label{fig:security_score_summary}
+\end{figure}
+The scoring metric as originally computed is averaged over the whole history of the device manufacturer, device model or network operator, it gives equal weight to periods years ago as to in the last few months.
+If instead we take an exponential moving average of the daily score for days with more than \daSigNumDevicesDay\ devices when there have been at least consecutive \daSigNumDays\ days of data with that many devices with an $\alpha$ of $1/\daSigNumDays$ then we can plot how this score has changed over time.
+Equation~\ref{eq:rolling_update} shows how the value for a particular day ($v_i$) is computed from the previous day's value and the input for the current day ($n$).
+\begin{equation}
+v_i = v_{i-1} (1 - \alpha) + n \alpha
+\label{eq:rolling_update}
+\end{equation}
+Figure~\ref{fig:security_score_manufacturer} shows this for manufacturers, Figure~\ref{fig:security_score_model} for device models, Figure~\ref{fig:security_score_operator} for operators Figure~\ref{fig:security_score_summary} for Nexus and non-Nexus devices.
+These show how the scores for different entities are different and change over time, while there is correlated behaviour for different entities (due to things like new vulnerabilities affecting all Android being discovered) these lines still have crossings due to the different behaviour of the different entities.
+It also shows that we do not have sufficient data for all the entities all of the time, resulting in gaps in the data.
+The clearest results are for Figure~\ref{fig:security_score_summary} with a large gap between the scores for Nexus and non-Nexus devices across the whole data set.

### 2015-06-01 update.sh

@@ -39,6 +39,7 @@ cp "$fromdir"frms_os_versions.pdf frms_os_versions.pdf
 cp "$fromdir"security_score_manufacturer.pdf security_score_manufacturer.pdf
 cp "$fromdir"security_score_model.pdf security_score_model.pdf
 cp "$fromdir"security_score_operator.pdf security_score_operator.pdf
+cp "$fromdir"security_score_summary.pdf security_score_summary.pdf
 popd
 cp ~/Library/papers/library.filtered.bib securityupdates.bib
 cp /mnt/nas04-device-analyzer-graphing/stats/latex.tex dastats.tex

### 2015-06-01 ccs.tex
Fix figure size of update ecosystem
Add section on comparison with other data

@@
  \centering
- \def\svgwidth{0.7\columnwidth}
+ \def\svgwidth{\columnwidth}
  \import{figures/}{update_ecosystem.pdf_tex}
@@
 
+\section{Comparison with other data}
+Here we show that the \da\ data~\cite{Wagner2013} is representative for our purposes by comparing it with two other sources of data which provide an upper and lower bound on it.
+The \da\ data is collected from devices across the world where users have voluntarily installed the app.
+We have obtained data comparable on 5\,290 devices from a multinational FTSE 100 company's mobile device management database which includes company owned and BYOD devices and from 5\,170\,000 matching User-Agent headers on all HTTP traffic for 30\% of Rwanda for a week.
+\begin{figure}
+\centering
+\includegraphics[width=\columnwidth]{figures/dists}
+\caption{Comparison between FTSE, User-Agent and the corresponding \da\ data, error bars indicate 95\% confidence intervals.}
+\label{fig:dists}
+\end{figure}
+We used the data from the FTSE 100 company for the week ending 2015-04-27 and the User-Agent data was collected between 2015-01-31 and 2015-02-12.
+Figure~\ref{fig:dists} shows the proportion of devices running each Android OS version in the two comparison data sets and the comparable periods from \da.
+The general pattern this shows is that in the FTSE data newer versions are more popular than in the \da\ data and that in the Rwanda data old versions are more popular than in the \da\ data.
+This shows that \da\ lies between these two biased samples while being similar to both and so is representative.
+
 \section{Related work}

### 2015-06-01 update.sh

@@ -40,6 +40,7 @@ cp "$fromdir"security_score_manufacturer.pdf security_score_manufacturer.pdf
 cp "$fromdir"security_score_model.pdf security_score_model.pdf
 cp "$fromdir"security_score_operator.pdf security_score_operator.pdf
 cp "$fromdir"security_score_summary.pdf security_score_summary.pdf
+cp ~drt24/git/da/graphing/pdf/dists.pdf dists.pdf
 popd
 cp ~/Library/papers/library.filtered.bib securityupdates.bib
 cp /mnt/nas04-device-analyzer-graphing/stats/latex.tex dastats.tex

### 2015-06-02 ccs.tex
Add some Introduction

@@
 
-The development of Android is distributed among many parties and the security of Android relies on many open source projects, include the Linux kernel, OpenSSL and BouncyCastle as well as Google who build the core platform.
+We have found that Android devices do not receive security updates when they should and hence are exposed to known critical vulnerabilities~\cite{androidvulnerabilities.org}.
+Here we attempt to determine why by understanding the Android update process, apportioning blame to different entities in the ecosystem and producing a scoring metric to compare different entities.
+
+By producing a scoring metric which compares different entities such as device manufacturers or network operators or specific device models we can provide information which allows device purchasers to select devices with better security.
+We know that organisations want to buy devices with better security and which receive updates, for example the CESG which advises UK government a advocates this~\cite{CESG2013} and we are collaborating with a FTSE 100 company to give them better information.
+Before we provided this information in this paper it was not available.
+We provide the answer to the question ``Who should I buy my Android phone from for best security?''
+
+The Android update process is slow because its development is distributed among many parties and has a long multi stage pipeline.
+The security of Android relies on many open source projects, include the Linux kernel, OpenSSL and BouncyCastle as well as Google who build the core platform.
 In addition, device manufacturers (we know of \daNumManufacturers\footnote{\label{foot:dadata}We computed this from the \da~\cite{Wagner2013} data, see \S\ref{sec:android_update_process}.}) and network operators (\daNumOperators\dafoot) may make further modifications before devices are shipped to customers.

### 2015-06-02 ccs.tex
Update author and paper metadata and turn on page numbering
Further work on the introduction

@@
 \documentclass{sig-alternate}
+\pagenumbering{arabic}
 
@@
 % --- Author Metadata here ---
-\conferenceinfo{CCS}{2015 Denver, Colorado, USA}
+\conferenceinfo{SPSM}{2015 Denver, Colorado, USA}
 %\CopyrightYear{2015} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
@@
 \author{
-Daniel~R.~Thomas
+\href{http://orcid.org/0000-0001-8936-0683}{Daniel~R.~Thomas}
 \and
@@
 By producing a scoring metric which compares different entities such as device manufacturers or network operators or specific device models we can provide information which allows device purchasers to select devices with better security.
-We know that organisations want to buy devices with better security and which receive updates, for example the CESG which advises UK government a advocates this~\cite{CESG2013} and we are collaborating with a FTSE 100 company to give them better information.
-Before we provided this information in this paper it was not available.
-We provide the answer to the question ``Who should I buy my Android phone from for best security?''
+Currently corporate and public sector buyers are encouraged to purchase secure devices, but we have found little concrete guidance on the specific makes and models providing the best security.
+For example, CESG, which advises the UK government on how to secure its computer systems, recommends picking Android device models from device manufactures which are good at shipping security updates promptly~\cite{CESG2013} but it does not state which device manufacturers these are.
+We are also collaborating with a FTSE 100 company to give them concrete information.
+To address this, we developed a scoring system and provide numbers on the historic performance of device models in the \da\ project in~\S\ref{sec:security_scoring}.
+We provide the answer to the question ``Who should I buy my Android phone from for best security?'', in general Nexus devices are better, further details are given in \S\ref{sec:security_scoring}.
 
@@
 \end{figure}
-
-It is clear from Figure~\ref{fig:proportioninsecure} that Android devices are often exposed to known security vulnerabilities.
-One solution is regulation, and indeed there is ongoing legal action to force network operators to ship updates for security vulnerabilities~\cite{Soghoian2013}.\todolater{Check on the status of this legal action}
+By using data from \da~\cite{Wagner2013} we can examine the versions of Android running on user devices over time and through comparison with data on which versions of Android were vulnerable to different critical vulnerabilities~\cite{androidvulnerabilities.org} we can plot the exposure of Android devices to known critical vulnerabilities.
+Figure~\ref{fig:proportioninsecure} shows that Android devices are often exposed to known security vulnerabilities.
+One solution to this problem is regulation, and indeed there is ongoing legal action to force network operators to ship updates for security vulnerabilities~\cite{Soghoian2013}.\todolater{Check on the status of this legal action}
 Many smartphones are sold on 12--24 month contracts, and yet our data shows many devices do not receive many security updates, with an overall average of \daUpdatesPerYearNominal\ per year. 
 In contrast, Windows XP could be purchased for a one-off payment in October 2001 and received security updates until April 2014.
+An alternative to regulation is access to publicly available data, which we provide in this paper.
+Comparative data would provide an incentive to both device manufacturers and network operators to provide updates.
 
-An alternative to regulation is access to publicly available data. 
-Comparative data would provide an incentive to both device manufacturers and network operators to provide updates.
-Currently corporate and public sector buyers are encouraged to purchase secure devices, but we have found little concrete guidance on the specific makes and models providing the best security. 
-For example, CESG, which advises the UK government on how to secure its computer systems, recommends picking Android device models from device manufactures which are good at shipping security updates promptly~\cite{CESG2013} but it does not state which device manufacturers these are.
-To address this, we developed a scoring system and provide numbers on the historic performance of device models in the \da\ project in~\S\ref{sec:security_scoring}.
-
+\todo{continue here}
 In summary, the contributions of this paper are:

### 2015-06-02 ccs.tex
Evil hack to work around bug in hyperref

@@
 
-User-Agent strings have been used to investigate the timeliness of web browser updates~\cite{Frei2008}, with at most 80\% of Firefox users running the most recent version.
+\interfootnotelinepenalty=10000 %To stop the lifehacker footnote below from spanning multiple pages and breaking
+User-Agent strings have been used to investigate the timeliness of web browser updates, with at most 80\% of Firefox users running the most recent version~\cite{Frei2008}.
 The same analysis was used to show that Chrome's use of silent updates seems to increase uptake of upgrades~\cite{Duebendorfer2010} with 97\% of users running the latest version within 3 weeks of release.

### 2015-06-02 ccs.tex
Add a working title and categories, terms and keywords
Make cross reference more precise

@@
 
-\title{TODO}
+\title{Security performance in the Android ecosystem}
 
@@
 % A category with the (minimum) three required fields
-\category{}{}{}
+\category{Security and privacy}{Systems security}{Operating systems security}[Mobile platform security]
+\category{Security and privacy}{Systems security}{Vulnerability management}
 
-\terms{}
+\terms{Economics, Measurement, Security}
 
-\keywords{}
+\keywords{Android, updates, vulnerabilities, metrics, ecosystems}
 
@@
 To address this, we developed a scoring system and provide numbers on the historic performance of device models in the \da\ project in~\S\ref{sec:security_scoring}.
-We provide the answer to the question ``Who should I buy my Android phone from for best security?'', in general Nexus devices are better, further details are given in \S\ref{sec:security_scoring}.
+We provide the answer to the question ``Who should I buy my Android phone from for best security?'', in general Nexus devices are better, further details are given in \S\ref{sec:security_scoring:results}.
 

### 2015-06-02 security_scoring.tex
Remove erroneous line break
Fix title of security_scoring
Make cross reference more precise

@@
-\subsection{Experiment 3: Comparing device manufacturers, device models and network operators}
+\subsection{Comparing device manufacturers, device models and network operators}
 \label{sec:security_scoring}\label{sec:exp:security_score}
@@
 
-\subsubsection{Results: Security scores}
+\subsubsection{Results: Security scores}\label{sec:security_scoring:results}
 The scores across the whole of Android are that \daMeanInsecurityPerc\ of devices are exposed to known root exploits.
@@
 \caption{Security scores for different operators}
-\label{fig:security_score_operator}\\
+\label{fig:security_score_operator}
 \includegraphics[width=\columnwidth]{figures/security_score_summary}

### 2015-06-03 ccs.tex
Include evil hack for making multiple urls in bibliographies work
Add some related work

@@
 \usepackage{hyperref}
+
+% http://tex.stackexchange.com/questions/57865/how-to-use-multiple-urls-for-one-bibtex-reference
+\let\URL\url
+\makeatletter
+\def\url#1{\@URL#1||\@nil}
+\def\@URL#1|#2|#3\@nil{%
+  \URL{#1}\ifx\relax#2\relax\else| \URL{#2}\fi}
+\makeatother
+
 \usepackage{subcaption}
@@
 
+Security in depth is also a useful strategy.
+In this regard, iOS provides additional safeguards beyond those used in Android, including a pre-distribution review, mandatory code-signing by the manufacturer, and (with the important exception of ROP-based attacks~\cite{Wang2013a}) the technical prohibition of dynamic code loading by an app.
+These features, as well as Address Space Layout Randomisation (ASLR) and mandatory access controls, has resulted in a lower level of malware affecting iOS when compared to Android~\cite{Felt2011}.
+
+There are continuing efforts to reduce the impact of critical vulnerabilities, both in Android and elsewhere.
+SEAndroid~\cite{Smalley2013}, which is included in Android from version 4.1~\cite{jelly-bean-release}, and fully enforcing from version 5.0~\cite{AndroidSecurity2014} claimed to prevent some root vulnerabilities and to reduce the impact of others.
+Capability based enforcement systems such as Capsicum~\cite{Watson2010} substantially reduce the capabilities that an exploit has to try and gain increased privilege with and could be included in Linux\footnote{\url{https://github.com/google/capsicum-linux}} and hence Android.
+
+Rather than fixing critical vulnerabilities, security can be obtained by detecting malicious apps and preventing their installation or execution.
+Detection strategies include RiskRanker, which classified 3\,281 out of 118\,318 apps (2.8\%) as risky of which 718 (22\%) were malware and 322 (10\%) were previously unknown malware, an infection rate of 0.6\% across multiple markets~\cite{Grace2012a}.
+DroidRanger also analysed apps finding 148 out of 182\,823 apps (0.08\%) to be malicious across multiple markets of which 29 were previously unknown~\cite{Zhou2012a}.
+%DroidRanger: It used permission-based behavioural fingerprinting which looked at the permissions of known malware and heuristic-based filtering -- dynamic loading of both Dalvik and native code.
+A common technique used by attackers is to include malicious code in repackaged popular apps. 
+AnDarwin uses this insight to detect similar apps, and found 169 out of 265\,359 of all apps studied (0.06\%) were malicious clones~\cite{Crussell2013}.
+%AnDarwin: It used clustering based on semantic vectors derived from the program dependence graphs to detect similar apps.
+
+The percentage of Android devices running the most recent version (\daUpdatednessPerc) is much less than the rate ($>90$\%) for Windows XP SP2 computers contacting the Microsoft update servers~\cite{Gkantsidis2006}.
+A simple numerical comparison is unfair because only one major OS version was considered in the Microsoft analysis, and data was only collected from computers which contacted the update server, although this was the default.
+More recent data demonstrates the difficulty of upgrading computers between major OS versions, with 27\% of Windows computers running Windows XP in July 2014,\footnote{\url{https://archive.today/PLGxn}} four months after Windows XP stopped receiving security updates.
+
+\cite{Nappa2015}
+\cite{Zhang2014}
+\cite{Lindorfer2014}
+\cite{Arp2014}
+
 \section{Conclusion}

### 2015-06-03 security_scoring.tex
Add TODO
Discuss the uncertainty calculation for m (badly)

@@
 
-We can compute the uncertainty for $f$ and $u$.
+We can compute the uncertainty for $f$, $u$ and $m$.
 $f$ is calculated by taking the total secure device days and dividing it by the total insecure and secure device days.
@@
 $u$ is computed by taking the sum of the proportions of devices running the most recent version each day, both the count of devices running the maximum version and total count have square root uncertainties.
+$m$ is computed by counting the number of vulnerabilities which affected that entity and which have not yet been fixed on any device we have observed from that entity.
+However it could be that the entity has released a fix to some devices but we have not yet observed a device with that fix and so the uncertainty in each vulnerability truly being outstanding is the probability of that, assuming that at least \daMeanOutstandingProbability\ of devices have the fix, i.e.
+\daMeanOutstandingUncertaintyEquation\ where $n$ is the number of devices contributing to that day's data.
 The Python \texttt{uncertainties} library was used to propagate uncertainties through calculations.
@@
 Changing the scoring metric also impacts the scores given for each entity Table~\ref{tab:change_in_scores} shows the mean impact on the scores.
+\todo{Give some of the actual orderings as well for the most important ones}
 

### 2015-06-03 ccs.tex
comment out the Acknowledgments until we have some
update contributions

@@
 
-\todo{continue here}
 In summary, the contributions of this paper are:
 \begin{itemize}
- \item We measure the security of Android according to three metrics and compare different device manufacturers and device models to allow device purchasers to differentiate between them based on security.
+ \item We quantify the Android update process.
+ \item We propose a scoring metric to evaluate the security of different instances of a platform.
+ \item We measure the security of Android according to this and compare different device manufacturers, device models and network operators to allow device purchasers to differentiate between them based on security.
 \end{itemize}
@@
 \identifying{
-\section*{Acknowledgements}
+%\section*{Acknowledgements}
 }

### 2015-06-03 ccs.tex
Correct tense
Add fixes and TODOs

--- 2015-06-03 ccs.tex
@@
 
-\title{Security performance in the Android ecosystem}
+\title{Measuring security performance in the Android ecosystem}
 
@@
 
-We have found that Android devices do not receive security updates when they should and hence are exposed to known critical vulnerabilities~\cite{androidvulnerabilities.org}.
+We have found that Android devices do not receive security updates promptly when critical vulnerabilities are publicly disclosed and hence \daMeanInsecurityPercNominal\ are exposed to known critical vulnerabilities~\cite{androidvulnerabilities.org}.
 Here we attempt to determine why by understanding the Android update process, apportioning blame to different entities in the ecosystem and producing a scoring metric to compare different entities.
 
-By producing a scoring metric which compares different entities such as device manufacturers or network operators or specific device models we can provide information which allows device purchasers to select devices with better security.
+By producing a scoring metric which compares different entities such as device manufacturers, network operators or specific device models we can provide information which allows device purchasers to select devices with better security.
 Currently corporate and public sector buyers are encouraged to purchase secure devices, but we have found little concrete guidance on the specific makes and models providing the best security.
-For example, CESG, which advises the UK government on how to secure its computer systems, recommends picking Android device models from device manufactures which are good at shipping security updates promptly~\cite{CESG2013} but it does not state which device manufacturers these are.
-We are also collaborating with a FTSE 100 company to give them concrete information.
+For example, CESG, which advises the UK government on how to secure its computer systems, recommends picking Android device models from device manufactures which are good at shipping security updates promptly but it does not state which device manufacturers these are~\cite{CESG2013}.
+We are also collaborating with a FTSE 100 company to give them concrete information to answer this question.
 To address this, we developed a scoring system and provide numbers on the historic performance of device models in the \da\ project in~\S\ref{sec:security_scoring}.
@@
 
-The Android update process is slow because its development is distributed among many parties and has a long multi stage pipeline.
+The Android update process is slow because its development is distributed among many parties and has a long multi-stage pipeline~\cite{HTC2013}.
 The security of Android relies on many open source projects, include the Linux kernel, OpenSSL and BouncyCastle as well as Google who build the core platform.
@@
 Many smartphones are sold on 12--24 month contracts, and yet our data shows many devices do not receive many security updates, with an overall average of \daUpdatesPerYearNominal\ per year. 
-In contrast, Windows XP could be purchased for a one-off payment in October 2001 and received security updates until April 2014.
-An alternative to regulation is access to publicly available data, which we provide in this paper.
-Comparative data would provide an incentive to both device manufacturers and network operators to provide updates.
+In contrast, Windows XP could be purchased for a one-off payment in October 2001 and received security updates monthly until April 2014.
+An alternative to regulation is access to publicly available data.
+Comparative data, which we provide in this paper, would provide an incentive to both device manufacturers and network operators to provide updates.
 
-\todo{continue here}
 In summary, the contributions of this paper are:
 \begin{itemize}
- \item We measure the security of Android according to three metrics and compare different device manufacturers and device models to allow device purchasers to differentiate between them based on security.
+ \item We quantify the Android update process~\S\ref{sec:android_ecosystem}.
+ \item We propose the FUM scoring metric to evaluate the security of different instances of a platform~\S\ref{sec:security_scoring:method}.
+ \item We measure the security of Android according to this and compare different device manufacturers, device models and network operators to allow device purchasers to differentiate between them based on security~\S\ref{sec:security_scoring:results}.
 \end{itemize}
@@
 There is a complex Android ecosystem which creates and distributes updates to Android which fix vulnerabilities.
-In this section we describe how the Android ecosystem functions and how Android versions, as seen by the \da\ project, are produced.
+In this section we describe how the Android ecosystem functions and how Android versions are produced, as seen by the \da\ project.
 
@@
  Numbers on edges indicate updates shipped between \daStartDate\ and \daEndDate, those in brackets represent number of such entities.
- Dotted arrows indicate flows where we can't measure how many updates are produced as they are not public.}
+ Dotted arrows indicate flows where we can't measure how many updates are produced as they are not public.\todo{Check this against the HTC infographic~\cite{HTC2013}}}
  \label{fig:update_ecosystem}
 \end{figure}
-To understand how vulnerabilities in Android are fixed we must examine the Android update process which we model in Figure~\ref{fig:update_ecosystem}.
+To understand how vulnerabilities in Android are fixed we examine the Android update process which we model in Figure~\ref{fig:update_ecosystem}.
 There are five entities or groups which contribute towards Android updates: the network operators, the device manufacturers, the hardware developers, Google and the upstream open source projects.
@@
 The Android platform is then built from these components by Google.
-The code for each Android release or update is kept secret\footnote{\url{https://source.android.com/source/code-lines.html}} until after a binary release has been published.
+The code for each Android release or update is kept secret until after a binary release has been published.\footnote{\url{https://source.android.com/source/code-lines.html}}
 Device manufacturers likely receive advanced access in order to prepare handsets so they can customise it before passing it on to the network operator.
@@
 The number of updates shipped by Google (\daNumSigOSVersions) is the number of Android versions reported in \da\ which affected more than \daSigVersionPerc\ of devices for more than \daSigVersionDays\ days.
-This significance test is to remove spurious versions recorded in \da\ such as `5.2.0' in 2012 which has still not been released in 2014.
+This significance test is to remove spurious versions recorded in \da\ such as `5.2.0' in 2012 which had still not been released in 2014.
 
-We extracted data on the external projects used in Android and have included this and the scripts which generated it in \avo.
+We extracted data on the external projects used in Android and have included this and the scripts which generated it in AndroidVulnerabilities.org (\avo)~\cite{androidvulnerabilities.org}.
 These scripts analysed the Android Open Source Project's source tree to examine the source code of each of the external projects to find the project version associated with each Android version tag on the repository.
@@
 The \da\ data is collected from devices across the world where users have voluntarily installed the app.
-We have obtained data comparable on 5\,290 devices from a multinational FTSE 100 company's mobile device management database which includes company owned and BYOD devices and from 5\,170\,000 matching User-Agent headers on all HTTP traffic for 30\% of Rwanda for a week.
+We have obtained comparable data on 5\,290 devices from a multinational FTSE 100 company's mobile device management database which includes company and employee owned devices, and from 5\,170\,000 matching User-Agent headers on all HTTP traffic for 30\% of Rwanda for a week.
+\todo{citation}
 \begin{figure}
@@
 The general pattern this shows is that in the FTSE data newer versions are more popular than in the \da\ data and that in the Rwanda data old versions are more popular than in the \da\ data.
+\todo{Can we quantify this?}
 This shows that \da\ lies between these two biased samples while being similar to both and so is representative.
@@
 
+\section*{Acknowledgements}
 \identifying{
-\section*{Acknowledgements}
+This work was supported by a Google focussed research award and the EPSRC Standard Research Grant EP/P505445/1.
 }
+Some of the raw data and source code will be made available.
 
+\sloppy
 \printbibliography

### 2015-06-03 ccs.tex
Fill in abstract

@@
 \begin{abstract}
+The security of Android is severely damaged by the lack of security updates for known critical vulnerabilities  being shipped to devices.
 We find there is significant variability in the timely delivery of security updates across different device manufacturers and mobile network operators, and we define a security metric to rank the performance of device manufacturers and network operators in our data set.
-TODO
+This provides the comparable data required for purchasers to determine which device manufacturers and network operators to use and hence incentives them to ship security updates.
 \end{abstract}

### 2015-06-03 ccs.tex
Indicate some hyphenation

@@
 There are continuing efforts to reduce the impact of critical vulnerabilities, both in Android and elsewhere.
-SEAndroid~\cite{Smalley2013}, which is included in Android from version 4.1~\cite{jelly-bean-release}, and fully enforcing from version 5.0~\cite{AndroidSecurity2014} claimed to prevent some root vulnerabilities and to reduce the impact of others.
+SE\-Android~\cite{Smalley2013}, which is included in Android from version 4.1~\cite{jelly-bean-release}, and fully enforcing from version 5.0~\cite{AndroidSecurity2014} claimed to prevent some root vulnerabilities and to reduce the impact of others.
 Capability based enforcement systems such as Capsicum~\cite{Watson2010} substantially reduce the capabilities that an exploit has to try and gain increased privilege with and could be included in Linux\footnote{\url{https://github.com/google/capsicum-linux}} and hence Android.
@@
 Rather than fixing critical vulnerabilities, security can be obtained by detecting malicious apps and preventing their installation or execution.
-Detection strategies include RiskRanker, which classified 3\,281 out of 118\,318 apps (2.8\%) as risky of which 718 (22\%) were malware and 322 (10\%) were previously unknown malware, an infection rate of 0.6\% across multiple markets~\cite{Grace2012a}.
-DroidRanger also analysed apps finding 148 out of 182\,823 apps (0.08\%) to be malicious across multiple markets of which 29 were previously unknown~\cite{Zhou2012a}.
+Detection strategies include Risk\-Ranker, which classified 3\,281 out of 118\,318 apps (2.8\%) as risky of which 718 (22\%) were malware and 322 (10\%) were previously unknown malware, an infection rate of 0.6\% across multiple markets~\cite{Grace2012a}.
+Droid\-Ranger also analysed apps finding 148 out of 182\,823 apps (0.08\%) to be malicious across multiple markets of which 29 were previously unknown~\cite{Zhou2012a}.
 %DroidRanger: It used permission-based behavioural fingerprinting which looked at the permissions of known malware and heuristic-based filtering -- dynamic loading of both Dalvik and native code.
 A common technique used by attackers is to include malicious code in repackaged popular apps. 
-AnDarwin uses this insight to detect similar apps, and found 169 out of 265\,359 of all apps studied (0.06\%) were malicious clones~\cite{Crussell2013}.
+An\-Darwin uses this insight to detect similar apps, and found 169 out of 265\,359 of all apps studied (0.06\%) were malicious clones~\cite{Crussell2013}.
 %AnDarwin: It used clustering based on semantic vectors derived from the program dependence graphs to detect similar apps.

### 2015-06-03 security_scoring.tex
Bring security scoring up one level to be a section

@@
-\subsection{Comparing device manufacturers, device models and network operators}
+\section{Comparing device manufacturers, device models and network operators}
 \label{sec:security_scoring}\label{sec:exp:security_score}
@@
 
-\subsubsection{Method: Scoring for security}
+\subsection{Method: Scoring for security}\label{sec:security_scoring:method}
 
@@
 
-\subsubsection{Results: Security scores}\label{sec:security_scoring:results}
+\subsection{Results: Security scores}\label{sec:security_scoring:results}
 The scores across the whole of Android are that \daMeanInsecurityPerc\ of devices are exposed to known root exploits.
@@
 
-\subsubsection{Sensitivity of scoring metric}
-\daTabDLDistances
-\daTabChangeInScores
-To evaluate whether the ranking of different manufacturers is sensitive to the form of the scoring metric we computed the normalised Damerau-Levenshtein distance~\cite{Bard2007} between the lists ordered using different forms of the scoring metric, this is shown in Table~\ref{tab:dl_distances}.
-The `equal' metric weights $f$, $u$ and $m$ equally rather than favouring $f$ and makes little difference.
-Changing the scoring metric also impacts the scores given for each entity Table~\ref{tab:change_in_scores} shows the mean impact on the scores.
-\todo{Give some of the actual orderings as well for the most important ones}
-
-\subsubsection{Scores over time}
+\subsection{Scores over time}
 \begin{figure}
@@
 The clearest results are for Figure~\ref{fig:security_score_summary} with a large gap between the scores for Nexus and non-Nexus devices across the whole data set.
+
+\subsection{Sensitivity of scoring metric}
+\daTabDLDistances
+\daTabChangeInScores
+To evaluate whether the ranking of different manufacturers is sensitive to the form of the scoring metric we computed the normalised Damerau-Levenshtein distance~\cite{Bard2007} between the lists ordered using different forms of the scoring metric, this is shown in Table~\ref{tab:dl_distances}.
+The `equal' metric weights $f$, $u$ and $m$ equally rather than favouring $f$ and makes little difference.
+Changing the scoring metric also impacts the scores given for each entity Table~\ref{tab:change_in_scores} shows the mean impact on the scores.
+\todo{Showing?}
+\todo{Give some of the actual orderings as well for the most important ones}
+\todo{Sensitivity of time based vs. global scores}

### 2015-06-03 security_scoring.tex
Clarify security scoring and add some todos

@@
 We can compute the uncertainty for $f$, $u$ and $m$.
-$f$ is calculated by taking the total secure device days and dividing it by the total insecure and secure device days.
+$f$ is computed by taking the total secure device days and dividing it by the total insecure and secure device days.
 The total secure device days and total insecure device days are both counting experiments and so their measurement error is their square root~\cite{Taylor1997}.
@@
 $u$ is computed by taking the sum of the proportions of devices running the most recent version each day, both the count of devices running the maximum version and total count have square root uncertainties.
-$m$ is computed by counting the number of vulnerabilities which affected that entity and which have not yet been fixed on any device we have observed from that entity.
-However it could be that the entity has released a fix to some devices but we have not yet observed a device with that fix and so the uncertainty in each vulnerability truly being outstanding is the probability of that, assuming that at least \daMeanOutstandingProbability\ of devices have the fix, i.e.
-\daMeanOutstandingUncertaintyEquation\ where $n$ is the number of devices contributing to that day's data.
+$m$ is computed by counting the number of vulnerabilities which affected that entity and which have not yet been fixed on any device we have observed from that entity every day and averaging over observed days.
+However it could be that the entity has released a fix to some devices but we have not yet observed a device with that fix.
+So the uncertainty in each vulnerability truly being outstanding is the probability of not having observed a fixed device if fixed device existed.
+We assume that if the fix has been released then at least \daMeanOutstandingProbability\ of devices have the fix.
+This gives an uncertainty of \daMeanOutstandingUncertaintyEquation\ where $n$ is the number of devices contributing to that day's data for each vulnerability outstanding each day.
 The Python \texttt{uncertainties} library was used to propagate uncertainties through calculations.
+This does not capture systematic errors such as those due to not including vulnerabilities which do affect devices, but for comparison that does not matter.
+\todo{We don't include manufacturer specific vulnerabilities which will influence comparisons.}
 
@@
 There have been many reports\todolater{citation} that Google's Nexus devices are better at getting updates than other Android devices because Google makes the original updates and ships them to its devices.
-Table~\ref{tab:sec_summary} shows that this is the case with Nexus devices getting much better scores than non-Nexus devices.
+Table~\ref{tab:sec_summary} shows that this is the case with Nexus devices getting much better scores than non-Nexus devices and Figure~\ref{fig:security_score_summary} shows this holds true over time.
 \daTabSecScoresmanufacturer
@@
 There is a marked difference in the distribution of updates between the two with only \daSecScoreWorstmodelNumFullVersions\ full versions for \emph{\daSecScoreWorstmodel} and \daSecScoreBestmodelNumFullVersions\ for \emph{\daSecScoreBestmodel}.
-Both device models start off with the full version string of `2.3.3 GRI40' but the \emph{\daSecScoreBestmodel} receives many more updates over the same time period.
-Other models from the same manufacturer with similar model names to \emph{\daSecScoreWorstmodel} do much better such as the \emph{Desire HD}.
+%Both device models start off with the full version string of `2.3.3 GRI40' but the \emph{\daSecScoreBestmodel} receives many more updates over the same time period.
+%Other models from the same manufacturer with similar model names to \emph{\daSecScoreWorstmodel} do much better such as the \emph{Desire HD}.
 
-If update delays are due to the delay in manufacturers providing the update rather than in operators in supplying the update and users installing the update we would expect the update behaviour of devices with the same device model to be similar.
+If update delays are due to the delay in manufacturers providing the update rather than in operators in supplying the update and users installing the update we would expect the update behaviour of devices with the same device model to be similar and rapid.
 We found that within \daModelHalfDeployment\ days of the first observation of a new version on a device, half of all devices of that model have the new version (or a higher version) installed, and within \daModelFullDeployment\ days \daFullDeployedAt\ of devices have the new version (or a higher version).
+\todo{How does this compare with the average?}
 There is a variation between device models, with the update being distributed to most devices quickly and others having a much slower roll out but since some do update quickly the bottleneck is unlikely to be with the user.
@@
  \includegraphics[width=0.9\textwidth]{figures/full_version_comp.pdf}
- \caption{Full version distributions for the highest and lowest scoring model}
+ \caption{Full version distributions for the highest and lowest scoring model\todo{this is wrong now}}
  \label{fig:full_version_comp}
@@
 \includegraphics[width=\columnwidth]{figures/security_score_summary}
-\caption{Security scores for Nexus and non-Nexus devices}
+\caption{Security scores for Nexus and non-Nexus devices\todo{subfig? bridge gaps?}}
 \label{fig:security_score_summary}
 \end{figure}
+\todo{This is method rather than result}
 The scoring metric as originally computed is averaged over the whole history of the device manufacturer, device model or network operator, it gives equal weight to periods years ago as to in the last few months.
-If instead we take an exponential moving average of the daily score for days with more than \daSigNumDevicesDay\ devices when there have been at least consecutive \daSigNumDays\ days of data with that many devices with an $\alpha$ of $1/\daSigNumDays$ then we can plot how this score has changed over time.
-Equation~\ref{eq:rolling_update} shows how the value for a particular day ($v_i$) is computed from the previous day's value and the input for the current day ($n$).
+If instead we take an exponential moving average of the daily score for days with more than \daSigNumDevicesDay\ devices when there have been at least consecutive \daSigNumDays\ days of data with that many devices then we can plot how this score has changed over time.
+Equation~\ref{eq:rolling_update} shows how the value for a particular day ($v_i$) is computed from the previous day's value and the input for the current day ($n$) with an $\alpha$ of $1/\daSigNumDays$.
 \begin{equation}

### 2015-06-03 dastats.tex
./update.sh (plus some manual forward porting)

@@
 \newcommand{\daStartDate}{2011-07-01}
-\newcommand{\daEndDate}{2015-05-21}
+\newcommand{\daEndDate}{2015-05-26}
 \newcommand{\daFullDeployedAt}{95\%}
-\newcommand{\daNumManufacturers}{298}
+\newcommand{\daNumManufacturers}{299}
 \newcommand{\daNumSigManufacturers}{8}
-\newcommand{\daNumModels}{2\,690}
+\newcommand{\daNumModels}{2\,700}
 \newcommand{\daNumOperators}{1\,440}
-\newcommand{\daNumSigOperators}{14}
+\newcommand{\daNumSigOperators}{15}
 \newcommand{\daNumOSDevices}{20\,300}
@@
 \newcommand{\daVulnFree}{$0.121 \pm 0.000$}
-\newcommand{\daMeanOutstandingVulnerabilities}{0.551}
+\newcommand{\daMeanOutstandingVulnerabilities}{$0.771 \pm 2.66e$}
 \newcommand{\daUpdatednessNominal}{0.0512}
@@
 \newcommand{\daUpdatednessPercTwosf}{$5.1 \pm 0.0\%$}
-\newcommand{\daSecurityScoreNominal}{2.83}
-\newcommand{\daSecurityScore}{$2.83 \pm 0.00$}
-\newcommand{\daCurrentMeanInsecurityPercNominal}{87.0\%}
-\newcommand{\daCurrentMeanInsecurityPerc}{$87.0 \pm 4.9\%$}
+\newcommand{\daSecurityScoreNominal}{2.53}
+\newcommand{\daSecurityScore}{$2.53 \pm 0.00$}
+\newcommand{\daCurrentMeanInsecurityPercNominal}{87.1\%}
+\newcommand{\daCurrentMeanInsecurityPerc}{$87.1 \pm 0.1\%$}
 \newcommand{\daCurrentMeanInsecurityPercTwosfNominal}{87\%}
-\newcommand{\daCurrentMeanInsecurityPercTwosf}{$87 \pm 4\%$}
-\newcommand{\daCurrentMeanSecurityPercNominal}{13.0\%}
-\newcommand{\daCurrentMeanSecurityPerc}{$13.0 \pm 4.9\%$}
+\newcommand{\daCurrentMeanInsecurityPercTwosf}{$87 \pm 0\%$}
+\newcommand{\daCurrentMeanSecurityPercNominal}{12.9\%}
+\newcommand{\daCurrentMeanSecurityPerc}{$12.9 \pm 0.1\%$}
 \newcommand{\daCurrentMeanSecurityPercTwosfNominal}{13\%}
-\newcommand{\daCurrentMeanSecurityPercTwosf}{$13 \pm 4\%$}
-\newcommand{\daCurrentVulnFreeNominal}{0.13}
-\newcommand{\daCurrentVulnFree}{$0.13 \pm 4.92$}
-\newcommand{\daCurrentMeanOutstandingVulnerabilities}{0.213}
-\newcommand{\daCurrentUpdatednessNominal}{0.0327}
-\newcommand{\daCurrentUpdatedness}{$0.0327 \pm 0.0006$}
-\newcommand{\daCurrentUpdatednessPercNominal}{3.27\%}
-\newcommand{\daCurrentUpdatednessPerc}{$3.27 \pm 0.06\%$}
+\newcommand{\daCurrentMeanSecurityPercTwosf}{$13 \pm 0\%$}
+\newcommand{\daCurrentVulnFreeNominal}{0.129}
+\newcommand{\daCurrentVulnFree}{$0.129 \pm 0.001$}
+\newcommand{\daCurrentMeanOutstandingVulnerabilities}{$0.209 \pm 9.49e$}
+\newcommand{\daCurrentUpdatednessNominal}{0.0329}
+\newcommand{\daCurrentUpdatedness}{$0.0329 \pm 0.0008$}
+\newcommand{\daCurrentUpdatednessPercNominal}{3.29\%}
+\newcommand{\daCurrentUpdatednessPerc}{$3.29 \pm 0.08\%$}
 \newcommand{\daCurrentUpdatednessPercTwosfNominal}{3.3\%}
@@
 \newcommand{\daCurrentSecurityScore}{$3.3 \pm 0.0$}
-\newcommand{\daDailyParticipationNominal}{906}
-\newcommand{\daDailyParticipation}{$906 \pm 224$}
+\newcommand{\daDailyParticipationNominal}{903}
+\newcommand{\daDailyParticipation}{$903 \pm 229$}
 \newcommand{\daNumFullVersions}{1\,250}
@@
 \newcommand{\daNumFullVersionUpdates}{4\,520}
-\newcommand{\daNumOSVersionUpdates}{3\,830}
+\newcommand{\daNumOSVersionUpdates}{3\,840}
 \newcommand{\daUpdatesPerYearNominal}{1.28}
@@
 \newcommand{\daNumPossibleSecurityUpdates}{2\,290}
-\newcommand{\daOSCurveFitParamFirst}{84.9}
+\newcommand{\daOSCurveFitParamFirst}{85.0}
 \newcommand{\daOSCurveFitParamSecond}{\num{0.0029}}
@@
 \newcommand{\daOSCurveFullDeployedNominal}{1\,120}
-\newcommand{\daOSCurveFullDeployed}{$1\,120 \pm 255\,000$}
+\newcommand{\daOSCurveFullDeployed}{$1\,120 \pm 256\,000$}
 \newcommand{\daAPICurveFitParamFirst}{102}
@@
 \newcommand{\daAPICurveFitRMSE}{0.113}
-\newcommand{\daAPICurvePolyRMSE}{0.114}
-\newcommand{\daAPICurveSplineRMSE}{0.114}
-\newcommand{\daAPICurveHalfDeployedYearsNominal}{0.933}
-\newcommand{\daAPICurveHalfDeployedYears}{$0.933 \pm 0.243$}
+\newcommand{\daAPICurvePolyRMSE}{0.113}
+\newcommand{\daAPICurveSplineRMSE}{0.113}
+\newcommand{\daAPICurveHalfDeployedYearsNominal}{0.934}
+\newcommand{\daAPICurveHalfDeployedYears}{$0.934 \pm 0.242$}
 \newcommand{\daAPICurveHalfDeployedNominal}{341}
@@
 \newcommand{\daVulnZergRushMonthsDefFixDeployed}{27.3}
-\newcommand{\daTabSecScoresmanufacturer}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline LG & $0.20 \pm 0.00$ & $0.32 \pm 0.00$ & 0.64 & $3.83 \pm 0.01$ \\  Motorola & $0.17 \pm 0.00$ & $0.12 \pm 0.00$ & 0.73 & $3.03 \pm 0.01$ \\  Samsung & $0.13 \pm 0.00$ & $0.04 \pm 0.00$ & 0.63 & $2.71 \pm 0.00$ \\  Sony & $0.13 \pm 0.00$ & $0.19 \pm 0.00$ & 1.09 & $2.61 \pm 0.01$ \\  HTC & $0.14 \pm 0.00$ & $0.09 \pm 0.00$ & 0.89 & $2.58 \pm 0.00$ \\  asus & $0.19 \pm 0.00$ & $0.52 \pm 0.01$ & 6.01 & $2.31 \pm 0.02$ \\  other & $0.06 \pm 0.00$ & $0.04 \pm 0.00$ & 1.04 & $1.94 \pm 0.00$ \\  Symphony & $0.00 \pm 0.00$ & $0.07 \pm 0.00$ & 5.00 & $0.27 \pm 0.01$ \\  walton & $0.00 \pm 0.00$ & $0.07 \pm 0.00$ & 5.97 & $0.24 \pm 0.01$ \\ \end{tabular} \caption{Security scores for manufacturers} \label{tab:sec_manufacturer} \end{table}}
+\newcommand{\daTabSecScoresmanufacturer}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline LG & $0.20 \pm 0.00$ & $0.32 \pm 0.00$ & $0.88 \pm 0.00$ & $3.51 \pm 0.01$ \\  Motorola & $0.17 \pm 0.00$ & $0.12 \pm 0.00$ & $1.10 \pm 0.01$ & $2.57 \pm 0.01$ \\  Samsung & $0.13 \pm 0.00$ & $0.04 \pm 0.00$ & $0.88 \pm 0.00$ & $2.39 \pm 0.00$ \\  asus & $0.19 \pm 0.00$ & $0.52 \pm 0.01$ & $21.42 \pm 0.09$ & $2.30 \pm 0.02$ \\  Sony & $0.13 \pm 0.00$ & $0.19 \pm 0.00$ & $1.68 \pm 0.01$ & $2.04 \pm 0.01$ \\  HTC & $0.14 \pm 0.00$ & $0.09 \pm 0.00$ & $1.40 \pm 0.00$ & $2.02 \pm 0.01$ \\  other & $0.06 \pm 0.00$ & $0.04 \pm 0.00$ & $1.77 \pm 0.00$ & $1.25 \pm 0.00$ \\  Symphony & $0.00 \pm 0.00$ & $0.08 \pm 0.00$ & $15.00 \pm 0.03$ & $0.24 \pm 0.01$ \\  walton & $0.00 \pm 0.00$ & $0.08 \pm 0.00$ & $20.79 \pm 0.06$ & $0.23 \pm 0.01$ \\ \end{tabular} \caption{Security scores for manufacturers} \label{tab:sec_manufacturer} \end{table*}}
 \newcommand{\daSecScoreBestmanufacturer}{LG}
-\newcommand{\daSecScoreBestmanufacturerScoreNominal}{3.83}
-\newcommand{\daSecScoreBestmanufacturerScore}{$3.83 \pm 0.01$}
-\newcommand{\daSecScoreBestmanufacturerNumFullVersions}{129}
+\newcommand{\daSecScoreBestmanufacturerScoreNominal}{3.51}
+\newcommand{\daSecScoreBestmanufacturerScore}{$3.51 \pm 0.01$}
+\newcommand{\daSecScoreBestmanufacturerNumFullVersions}{130}
 \newcommand{\daSecScoreWorstmanufacturer}{walton}
-\newcommand{\daSecScoreWorstmanufacturerScoreNominal}{0.236}
-\newcommand{\daSecScoreWorstmanufacturerScore}{$0.236 \pm 0.007$}
+\newcommand{\daSecScoreWorstmanufacturerScoreNominal}{0.235}
+\newcommand{\daSecScoreWorstmanufacturerScore}{$0.235 \pm 0.007$}
 \newcommand{\daSecScoreWorstmanufacturerNumFullVersions}{22}
-\newcommand{\daTabSecScoresmodel}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline Galaxy Nexus & $0.51 \pm 0.00$ & $0.55 \pm 0.01$ & 1.53 & $4.74 \pm 0.03$ \\  Nexus 4 & $0.28 \pm 0.00$ & $0.82 \pm 0.01$ & 6.06 & $3.59 \pm 0.04$ \\  Nexus 7 & $0.24 \pm 0.00$ & $0.74 \pm 0.01$ & 5.92 & $3.18 \pm 0.04$ \\  other & $0.09 \pm 0.00$ & $0.14 \pm 0.00$ & 0.55 & $2.97 \pm 0.00$ \\  Desire HD & $0.08 \pm 0.00$ & $0.05 \pm 0.00$ & 0.39 & $2.91 \pm 0.01$ \\  HTC Sensation & $0.35 \pm 0.00$ & $0.01 \pm 0.01$ & 1.59 & $2.42 \pm 0.02$ \\  GT-I9100 & $0.22 \pm 0.00$ & $0.02 \pm 0.00$ & 1.23 & $2.28 \pm 0.01$ \\  HTC Desire S & $0.02 \pm 0.00$ & $0.02 \pm 0.00$ & 1.00 & $1.75 \pm 0.01$ \\  GT-N7000 & $0.25 \pm 0.00$ & $0.00 \pm 0.00$ & 2.52 & $1.43 \pm 0.02$ \\  GT-P1000 & $0.01 \pm 0.00$ & $0.00 \pm 0.01$ & 1.79 & $0.90 \pm 0.02$ \\  GT-I9300 & $0.13 \pm 0.00$ & $0.01 \pm 0.00$ & 6.25 & $0.59 \pm 0.01$ \\  GT-I9505 & $0.01 \pm 0.00$ & $0.13 \pm 0.00$ & 6.82 & $0.44 \pm 0.01$ \\  HTC Desire HD & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & 3.05 & $0.27 \pm 0.02$ \\  GT-N7100 & $0.06 \pm 0.00$ & $0.00 \pm 0.01$ & 6.91 & $0.25 \pm 0.02$ \\  Symphony W68 & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & 11.00 & $0.00 \pm 0.03$ \\ \end{tabular} \caption{Security scores for models} \label{tab:sec_model} \end{table}}
+\newcommand{\daTabSecScoresmodel}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline Galaxy Nexus & $0.51 \pm 0.00$ & $0.55 \pm 0.01$ & $2.18 \pm 0.02$ & $4.29 \pm 0.03$ \\  Nexus 4 & $0.28 \pm 0.00$ & $0.82 \pm 0.01$ & $21.70 \pm 0.12$ & $3.58 \pm 0.04$ \\  Nexus 7 & $0.24 \pm 0.00$ & $0.74 \pm 0.01$ & $20.68 \pm 0.11$ & $3.16 \pm 0.04$ \\  Desire HD & $0.08 \pm 0.00$ & $0.05 \pm 0.00$ & $0.47 \pm 0.01$ & $2.78 \pm 0.02$ \\  other & $0.09 \pm 0.00$ & $0.14 \pm 0.00$ & $0.77 \pm 0.00$ & $2.68 \pm 0.00$ \\  HTC Sensation & $0.35 \pm 0.00$ & $0.01 \pm 0.01$ & $2.44 \pm 0.03$ & $1.88 \pm 0.03$ \\  HTC Desire S & $0.02 \pm 0.00$ & $0.02 \pm 0.00$ & $1.00 \pm 0.02$ & $1.75 \pm 0.03$ \\  GT-I9100 & $0.22 \pm 0.00$ & $0.02 \pm 0.00$ & $2.08 \pm 0.01$ & $1.59 \pm 0.01$ \\  GT-N7000 & $0.25 \pm 0.00$ & $0.00 \pm 0.00$ & $4.70 \pm 0.03$ & $1.04 \pm 0.02$ \\  GT-I9300 & $0.13 \pm 0.00$ & $0.01 \pm 0.00$ & $23.21 \pm 0.04$ & $0.57 \pm 0.01$ \\  GT-I9505 & $0.01 \pm 0.00$ & $0.13 \pm 0.00$ & $26.96 \pm 0.08$ & $0.43 \pm 0.01$ \\  GT-P1000 & $0.01 \pm 0.00$ & $0.00 \pm 0.01$ & $3.17 \pm 0.03$ & $0.28 \pm 0.02$ \\  GT-N7100 & $0.06 \pm 0.00$ & $0.00 \pm 0.01$ & $27.81 \pm 0.11$ & $0.24 \pm 0.02$ \\  HTC Desire HD & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & $6.23 \pm 0.03$ & $0.01 \pm 0.02$ \\  Symphony W68 & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & $66.00 \pm 0.19$ & $0.00 \pm 0.03$ \\ \end{tabular} \caption{Security scores for models} \label{tab:sec_model} \end{table*}}
 \newcommand{\daSecScoreBestmodel}{Galaxy Nexus}
-\newcommand{\daSecScoreBestmodelScoreNominal}{4.74}
-\newcommand{\daSecScoreBestmodelScore}{$4.74 \pm 0.02$}
+\newcommand{\daSecScoreBestmodelScoreNominal}{4.29}
+\newcommand{\daSecScoreBestmodelScore}{$4.29 \pm 0.02$}
 \newcommand{\daSecScoreBestmodelNumFullVersions}{49}
 \newcommand{\daSecScoreWorstmodel}{Symphony W68}
-\newcommand{\daSecScoreWorstmodelScoreNominal}{0.0001}
-\newcommand{\daSecScoreWorstmodelScore}{$0.0001 \pm 0.0276$}
+\newcommand{\daSecScoreWorstmodelScoreNominal}{1.3e-28}
+\newcommand{\daSecScoreWorstmodelScore}{$1.3e-28 \pm 0.0276$}
 \newcommand{\daSecScoreWorstmodelNumFullVersions}{1}
-\newcommand{\daTabSecScoressummary}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline nexus & $0.38 \pm 0.00$ & $0.48 \pm 0.00$ & 0.59 & $5.08 \pm 0.01$ \\  notnexus & $0.10 \pm 0.00$ & $0.02 \pm 0.00$ & 0.55 & $2.66 \pm 0.00$ \\ \end{tabular} \caption{Security scores for nexus} \label{tab:sec_summary} \end{table}}
-\newcommand{\daTabSecScoresoperator}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline O2 uk & $0.26 \pm 0.00$ & $0.12 \pm 0.00$ & 0.37 & $3.84 \pm 0.01$ \\  T-Mobile & $0.20 \pm 0.00$ & $0.18 \pm 0.00$ & 0.42 & $3.73 \pm 0.01$ \\  Orange & $0.21 \pm 0.00$ & $0.09 \pm 0.00$ & 0.39 & $3.52 \pm 0.02$ \\  Sprint & $0.18 \pm 0.00$ & $0.11 \pm 0.00$ & 0.40 & $3.46 \pm 0.01$ \\  3 & $0.18 \pm 0.00$ & $0.09 \pm 0.00$ & 0.52 & $3.23 \pm 0.01$ \\  Vodafone uk & $0.13 \pm 0.00$ & $0.13 \pm 0.00$ & 0.50 & $3.18 \pm 0.01$ \\  AT\&T & $0.12 \pm 0.00$ & $0.08 \pm 0.00$ & 0.43 & $3.09 \pm 0.01$ \\  Verizon & $0.18 \pm 0.00$ & $0.09 \pm 0.00$ & 0.82 & $2.83 \pm 0.01$ \\  unknown & $0.11 \pm 0.00$ & $0.20 \pm 0.00$ & 0.87 & $2.80 \pm 0.01$ \\  n Telenor & $0.03 \pm 0.00$ & $0.12 \pm 0.00$ & 1.25 & $1.81 \pm 0.01$ \\  Airtel & $0.05 \pm 0.00$ & $0.03 \pm 0.00$ & 1.62 & $1.27 \pm 0.01$ \\  Robi & $0.00 \pm 0.00$ & $0.07 \pm 0.00$ & 2.08 & $0.88 \pm 0.01$ \\  Grameenphone & $0.00 \pm 0.00$ & $0.05 \pm 0.00$ & 2.03 & $0.85 \pm 0.00$ \\  banglalink & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & 2.62 & $0.53 \pm 0.01$ \\ \end{tabular} \caption{Security scores for operators} \label{tab:sec_operator} \end{table}}
+\newcommand{\daTabSecScoressummary}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline nexus & $0.37 \pm 0.00$ & $0.47 \pm 0.00$ & $0.82 \pm 0.00$ & $4.75 \pm 0.01$ \\  notnexus & $0.10 \pm 0.00$ & $0.02 \pm 0.00$ & $0.77 \pm 0.00$ & $2.37 \pm 0.00$ \\ \end{tabular} \caption{Security scores for nexus} \label{tab:sec_summary} \end{table*}}
+\newcommand{\daTabSecScoresoperator}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline O2 uk & $0.26 \pm 0.00$ & $0.12 \pm 0.00$ & $0.51 \pm 0.01$ & $3.65 \pm 0.02$ \\  T-Mobile & $0.21 \pm 0.00$ & $0.18 \pm 0.00$ & $0.54 \pm 0.01$ & $3.56 \pm 0.02$ \\  Orange & $0.21 \pm 0.00$ & $0.09 \pm 0.00$ & $0.48 \pm 0.01$ & $3.39 \pm 0.02$ \\  Sprint & $0.18 \pm 0.00$ & $0.11 \pm 0.00$ & $0.51 \pm 0.01$ & $3.31 \pm 0.02$ \\  3 & $0.18 \pm 0.00$ & $0.09 \pm 0.00$ & $0.70 \pm 0.01$ & $2.99 \pm 0.02$ \\  AT\&T & $0.12 \pm 0.00$ & $0.08 \pm 0.00$ & $0.54 \pm 0.01$ & $2.94 \pm 0.01$ \\  Vodafone uk & $0.13 \pm 0.00$ & $0.13 \pm 0.00$ & $0.73 \pm 0.01$ & $2.88 \pm 0.02$ \\  unknown & $0.11 \pm 0.00$ & $0.20 \pm 0.00$ & $1.22 \pm 0.00$ & $2.40 \pm 0.01$ \\  Verizon & $0.18 \pm 0.00$ & $0.09 \pm 0.00$ & $1.27 \pm 0.01$ & $2.30 \pm 0.01$ \\  n Telenor & $0.03 \pm 0.00$ & $0.12 \pm 0.00$ & $1.86 \pm 0.00$ & $1.29 \pm 0.01$ \\  Airtel & $0.05 \pm 0.00$ & $0.03 \pm 0.00$ & $2.42 \pm 0.00$ & $0.77 \pm 0.01$ \\  Robi & $0.00 \pm 0.00$ & $0.08 \pm 0.00$ & $3.26 \pm 0.02$ & $0.46 \pm 0.01$ \\  Grameenphone & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & $3.09 \pm 0.00$ & $0.41 \pm 0.00$ \\  banglalink & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & $4.83 \pm 0.01$ & $0.17 \pm 0.01$ \\ \end{tabular} \caption{Security scores for operators} \label{tab:sec_operator} \end{table*}}
 \newcommand{\daSecScoreBestoperator}{O2 uk}
-\newcommand{\daSecScoreBestoperatorScoreNominal}{3.84}
-\newcommand{\daSecScoreBestoperatorScore}{$3.84 \pm 0.01$}
-\newcommand{\daSecScoreBestoperatorNumFullVersions}{95}
+\newcommand{\daSecScoreBestoperatorScoreNominal}{3.65}
+\newcommand{\daSecScoreBestoperatorScore}{$3.65 \pm 0.01$}
+\newcommand{\daSecScoreBestoperatorNumFullVersions}{97}
 \newcommand{\daSecScoreWorstoperator}{banglalink}
-\newcommand{\daSecScoreWorstoperatorScoreNominal}{0.531}
-\newcommand{\daSecScoreWorstoperatorScore}{$0.531 \pm 0.005$}
-\newcommand{\daSecScoreWorstoperatorNumFullVersions}{72}
-\newcommand{\daTabDLDistances}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & summary \\ \hline equal & 0.0 & 0.0 & 0.0714 & 0.0 \\weight $m$ & 0.222 & 0.333 & 0.143 & 0.0 \\weight $u$ & 0.444 & 0.133 & 0.214 & 0.0 \\$f$ & 0.444 & 0.733 & 0.429 & 0.0 \\$m$ & 0.667 & 0.733 & 0.357 & 0.5 \\$u$ & 0.889 & 0.667 & 0.714 & 0.0 \\\end{tabular} \caption{Normalised Damerau-Levenshtein distances for different metrics} \label{tab:dl_distances} \end{table}}
-\newcommand{\daTabChangeInScores}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & summary \\ \hline $m$ & $-1.91 \pm 2.21$ & $-0.666 \pm 2.5$ & $-3.27 \pm 1.38$ & $-3.35 \pm 1.84$ \\weight $m$ & $-0.295 \pm 0.273$ & $-0.102 \pm 0.314$ & $-0.458 \pm 0.176$ & $-0.485 \pm 0.209$ \\equal & $-0.116 \pm 0.065$ & $-0.0391 \pm 0.112$ & $-0.146 \pm 0.048$ & $-0.166 \pm 0.028$ \\weight $u$ & $-0.0522 \pm 0.127$ & $-0.0154 \pm 0.239$ & $0.0199 \pm 0.0787$ & $-0.0137 \pm 0.127$ \\$u$ & $0.518 \pm 1.6$ & $0.198 \pm 2.14$ & $1.51 \pm 0.95$ & $1.36 \pm 1.52$ \\$f$ & $1.04 \pm 0.58$ & $0.351 \pm 1.01$ & $1.31 \pm 0.44$ & $1.5 \pm 0.2$ \\\end{tabular} \caption{Mean change in scores for different metrics} \label{tab:change_in_scores} \end{table*}}
+\newcommand{\daSecScoreWorstoperatorScoreNominal}{0.172}
+\newcommand{\daSecScoreWorstoperatorScore}{$0.172 \pm 0.005$}
+\newcommand{\daSecScoreWorstoperatorNumFullVersions}{73}
+\newcommand{\daTabDLDistances}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & summary \\ \hline equal & 0.0 & 0.0667 & 0.0714 & 0.0 \\weight $m$ & 0.222 & 0.333 & 0.0714 & 0.0 \\weight $u$ & 0.222 & 0.133 & 0.0714 & 0.0 \\$f$ & 0.444 & 0.733 & 0.5 & 0.0 \\$m$ & 0.556 & 0.8 & 0.429 & 0.5 \\$u$ & 0.889 & 0.6 & 0.786 & 0.0 \\\end{tabular} \caption{Normalised Damerau-Levenshtein distances for different metrics} \label{tab:dl_distances} \end{table}}
+\newcommand{\daTabChangeInScores}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & summary \\ \hline $m$ & $-1.13 \pm 1.77$ & $-0.133 \pm 2.4$ & $-2.51 \pm 1.66$ & $-2.67 \pm 1.83$ \\weight $m$ & $-0.184 \pm 0.21$ & $-0.0255 \pm 0.313$ & $-0.35 \pm 0.21$ & $-0.387 \pm 0.209$ \\equal & $-0.0792 \pm 0.0477$ & $-0.0136 \pm 0.122$ & $-0.11 \pm 0.05$ & $-0.133 \pm 0.028$ \\weight $u$ & $-0.053 \pm 0.127$ & $-0.0154 \pm 0.239$ & $0.0196 \pm 0.0792$ & $-0.0124 \pm 0.125$ \\$u$ & $0.182 \pm 1.42$ & $-0.0309 \pm 2.03$ & $1.19 \pm 1.08$ & $1.07 \pm 1.5$ \\$f$ & $0.713 \pm 0.43$ & $0.123 \pm 1.1$ & $0.991 \pm 0.536$ & $1.2 \pm 0.2$ \\\end{tabular} \caption{Mean change in scores for different metrics} \label{tab:change_in_scores} \end{table*}}
 \newcommand{\daGPAPICurveFitParamFirst}{80.7}
@@
 \newcommand{\daGPAPICurveSplineRMSE}{0.18}
-\newcommand{\daGPAPICurveHalfDeployedYearsNominal}{0.954}
-\newcommand{\daGPAPICurveHalfDeployedYears}{$0.954 \pm 0.473$}
+\newcommand{\daGPAPICurveHalfDeployedYearsNominal}{0.953}
+\newcommand{\daGPAPICurveHalfDeployedYears}{$0.953 \pm 0.472$}
 \newcommand{\daGPAPICurveHalfDeployedNominal}{348}
-\newcommand{\daGPAPICurveHalfDeployed}{$348 \pm 173$}
+\newcommand{\daGPAPICurveHalfDeployed}{$348 \pm 172$}
 \newcommand{\daGPAPICurveFullDeployedYearsNominal}{3.39}
@@
 \newcommand{\daGPAPICurveFullDeployed}{$1\,240 \pm 286\,000$}
-\newcommand{\daGPAPIPerAPIMeanNominal}{0.00453}
-\newcommand{\daGPAPIPerAPIMean}{$0.00453 \pm 0.00343$}
-\newcommand{\daGPAPIPerAPIStdevNominal}{0.0211}
-\newcommand{\daGPAPIPerAPIStdev}{$0.0211 \pm 0.0127$}
-\newcommand{\daGPAPIPerAPIAbsMeanNominal}{0.0144}
-\newcommand{\daGPAPIPerAPIAbsMean}{$0.0144 \pm 0.0066$}
+\newcommand{\daGPAPIPerAPIMeanNominal}{0.00459}
+\newcommand{\daGPAPIPerAPIMean}{$0.00459 \pm 0.00346$}
+\newcommand{\daGPAPIPerAPIStdevNominal}{0.0212}
+\newcommand{\daGPAPIPerAPIStdev}{$0.0212 \pm 0.0127$}
+\newcommand{\daGPAPIPerAPIAbsMeanNominal}{0.0145}
+\newcommand{\daGPAPIPerAPIAbsMean}{$0.0145 \pm 0.0066$}
 \newcommand{\daGPAPIPerAPIAbsStdevNominal}{0.0154}
 \newcommand{\daGPAPIPerAPIAbsStdev}{$0.0154 \pm 0.012$}
-\newcommand{\daGPAPIPerAPIParametersTable}{\begin{table} \centering \begin{tabular}{c|S|S} API Version & ${t_0}$ & $\mathrm{decay}$ \\ \hline2	&	0.0	&	0.0237\\3	&	0.0	&	0.0235\\4	&	0.0	&	0.00689\\5	&	59.1	&	0.00465\\6	&	0.0	&	0.0043\\7	&	0.0	&	0.00499\\8	&	47.7	&	0.00415\\9	&	155	&	0.00365\\10	&	90.2	&	0.00364\\11	&	432	&	0.00228\\12	&	355	&	0.00228\\13	&	304	&	0.00236\\14	&	218	&	0.00239\\15	&	160	&	0.00239\\16	&	142	&	0.00227\\17	&	288	&	0.00195\\18	&	119	&	0.00141\\19	&	119	&	0.00153\\\end{tabular} \caption{Parameters for different API versions} \label{tab:per_api_parameters} \end{table}}
+\newcommand{\daGPAPIPerAPIParametersTable}{\begin{table} \centering \begin{tabular}{c|S|S} API Version & ${t_0}$ & $\mathrm{decay}$ \\ \hline2	&	0.0	&	0.0237\\3	&	0.0	&	0.0235\\4	&	0.0	&	0.00689\\5	&	59.1	&	0.00465\\6	&	2.4	&	0.00434\\7	&	0.0	&	0.00499\\8	&	47.7	&	0.00415\\9	&	155	&	0.00365\\10	&	90.2	&	0.00364\\11	&	432	&	0.00228\\12	&	356	&	0.00228\\13	&	304	&	0.00236\\14	&	219	&	0.00239\\15	&	160	&	0.0024\\16	&	142	&	0.00227\\17	&	289	&	0.00195\\18	&	120	&	0.00142\\19	&	120	&	0.00154\\\end{tabular} \caption{Parameters for different API versions} \label{tab:per_api_parameters} \end{table}}
 \newcommand{\daGPAPISeventeenLaterDate}{2015-05-04}
-\newcommand{\daGPAPISeventeenLaterProportion}{72.4\%}
-\newcommand{\daGPAPISeventeenEarlierProportion}{27.6\%}
+\newcommand{\daGPAPISeventeenLaterProportion}{73.1\%}
+\newcommand{\daGPAPISeventeenEarlierProportion}{26.9\%}
 \newcommand{\daAPISeventeenReleaseDate}{2012-10-29}
-\newcommand{\daGPAPISeventeenFullDeployment}{2017-10-29}
-\newcommand{\daGPAPISeventeenCurveFitParamFirst}{288}
+\newcommand{\daGPAPISeventeenFullDeployment}{2017-10-24}
+\newcommand{\daGPAPISeventeenCurveFitParamFirst}{289}
 \newcommand{\daGPAPISeventeenCurveFitParamSecond}{\num{0.00195}}
-\newcommand{\daGPAPISeventeenCurveFitRMSE}{0.0272}
-\newcommand{\daGPAPISeventeenCurvePolyRMSE}{0.00888}
-\newcommand{\daGPAPISeventeenCurveSplineRMSE}{0.00888}
-\newcommand{\daGPAPISeventeenCurveHalfDeployedYearsNominal}{1.77}
-\newcommand{\daGPAPISeventeenCurveHalfDeployedYears}{$1.77 \pm 0.07$}
-\newcommand{\daGPAPISeventeenCurveHalfDeployedNominal}{644}
-\newcommand{\daGPAPISeventeenCurveHalfDeployed}{$644 \pm 28$}
-\newcommand{\daGPAPISeventeenCurveFullDeployedYearsNominal}{5.01}
-\newcommand{\daGPAPISeventeenCurveFullDeployedYears}{$5.01 \pm 1.11$}
-\newcommand{\daGPAPISeventeenCurveFullDeployedNominal}{1\,830}
-\newcommand{\daGPAPISeventeenCurveFullDeployed}{$1\,830 \pm 404$}
+\newcommand{\daGPAPISeventeenCurveFitRMSE}{0.0275}
+\newcommand{\daGPAPISeventeenCurvePolyRMSE}{0.0088}
+\newcommand{\daGPAPISeventeenCurveSplineRMSE}{0.0088}
+\newcommand{\daGPAPISeventeenCurveHalfDeployedYearsNominal}{1.76}
+\newcommand{\daGPAPISeventeenCurveHalfDeployedYears}{$1.76 \pm 0.07$}
+\newcommand{\daGPAPISeventeenCurveHalfDeployedNominal}{643}
+\newcommand{\daGPAPISeventeenCurveHalfDeployed}{$643 \pm 28$}
+\newcommand{\daGPAPISeventeenCurveFullDeployedYearsNominal}{4.99}
+\newcommand{\daGPAPISeventeenCurveFullDeployedYears}{$4.99 \pm 1.12$}
+\newcommand{\daGPAPISeventeenCurveFullDeployedNominal}{1\,820}
+\newcommand{\daGPAPISeventeenCurveFullDeployed}{$1\,820 \pm 408$}
 \newcommand{\daNumUpdatesUpgrades}{3\,590}
@@
 \newcommand{\daNumDataPoints}{157 Billion}
-\newcommand{\daMonthsDevices}{1\,780}
+\newcommand{\daMonthsDevices}{1\,790}
 \newcommand{\daMonths}{6}
@@
 \newcommand{\daNumInstalledApps}{180\,000}
+\newcommand{\daAPISeventeenReleaseDateMonth}{October 2012}
+\newcommand{\daGPAPISeventeenFullDeploymentMonth}{October 2017}
+\newcommand{\daMeanOutstandingVulnerabilitiesNominal}{0.771}
+\newcommand{\daCurrentMeanOutstandingVulnerabilitiesNominal}{0.209}
+\newcommand{\daModelHalfDeployment}{27}
+\newcommand{\daModelFullDeployment}{316}
+\newcommand{\daModelHalfDeploymentYears}{0.074}
+\newcommand{\daModelFullDeploymentYears}{0.866}
+\newcommand{\daMeanOutstandingProbability}{1\%}
+\newcommand{\daMeanOutstandingUncertaintyEquation}{$0.99^n$}

### 2015-06-03 ccs.tex
hide todolaters
Do some minor todos

@@
 \renewcommand{\todo}{\OldTodo[inline]}
-\newcommand{\todolater}[1]{\todo{#1}}%{}%
+\newcommand{\todolater}[1]{}%{\todo{#1}}%
 \RequirePackage[date=terse, isbn=true, doi=true, url=false, urldate=iso8601, maxbibnames=9, backref=false, backend=bibtex, style=ieee]{biblatex}
@@
  Numbers on edges indicate updates shipped between \daStartDate\ and \daEndDate, those in brackets represent number of such entities.
- Dotted arrows indicate flows where we can't measure how many updates are produced as they are not public.\todo{Check this against the HTC infographic~\cite{HTC2013}}}
+ Dotted arrows indicate flows where we can't measure how many updates are produced as they are not public.}
  \label{fig:update_ecosystem}
@@
 
+\todo{Do we want to talk about more stuff from the Esorics submission?}
+
 \input{security_scoring}
@@
 We have obtained comparable data on 5\,290 devices from a multinational FTSE 100 company's mobile device management database which includes company and employee owned devices, and from 5\,170\,000 matching User-Agent headers on all HTTP traffic for 30\% of Rwanda for a week.
-\todo{citation}
+\todo{citation for Rwanda dataset}
 \begin{figure}

### 2015-06-03 security_scoring.tex
Do some minor todos

@@
 
-\todo{There must be some way of putting an error bound on our results by considering the error on the whole DA set and the size of the subset used for each device manufacturer and device model}
+\todolater{There must be some way of putting an error bound on our results by considering the error on the whole DA set and the size of the subset used for each device manufacturer and device model}
 
@@
 We found that within \daModelHalfDeployment\ days of the first observation of a new version on a device, half of all devices of that model have the new version (or a higher version) installed, and within \daModelFullDeployment\ days \daFullDeployedAt\ of devices have the new version (or a higher version).
-\todo{How does this compare with the average?}
+This compares with the average rates of deployment for Android OS versions of 350 days for half and 1\,100 days for \daFullDeployedAt.
+\todolater{Compute these values automatically and with error.}
 There is a variation between device models, with the update being distributed to most devices quickly and others having a much slower roll out but since some do update quickly the bottleneck is unlikely to be with the user.

### 2015-06-04 ccs.tex
Move the \sloppy to before the Acknowledgements as we don't want to hyphenate the grant code
Tweaks based on feedback

@@
 \begin{abstract}
-The security of Android is severely damaged by the lack of security updates for known critical vulnerabilities  being shipped to devices.
-We find there is significant variability in the timely delivery of security updates across different device manufacturers and mobile network operators, and we define a security metric to rank the performance of device manufacturers and network operators in our data set.
-This provides the comparable data required for purchasers to determine which device manufacturers and network operators to use and hence incentives them to ship security updates.
+The security of Android depends on security updates for critical vulnerabilities  being shipped to devices.
+Using a corpus of \daNumOSDevices\ devices we show that there is significant variability in the timely delivery of security updates across different device manufacturers and network operators.
+We define a security metric to rank the performance of device manufacturers and network operators.
+This provides a comparison point for purchasers to determine which device manufacturers and network operators to use.
+We argue that publishing metrics like this provides an important incentive to manufacturers and operators to ship security updates.
 \end{abstract}
@@
 
-We have found that Android devices do not receive security updates promptly when critical vulnerabilities are publicly disclosed and hence \daMeanInsecurityPercNominal\ are exposed to known critical vulnerabilities~\cite{androidvulnerabilities.org}.
+It has been previously shown that Android devices do not receive security updates promptly when critical vulnerabilities are publicly disclosed and hence \daMeanInsecurityPercNominal\ are exposed to known critical vulnerabilities on average~\cite{androidvulnerabilities.org}.
 Here we attempt to determine why by understanding the Android update process, apportioning blame to different entities in the ecosystem and producing a scoring metric to compare different entities.
@@
 To address this, we developed a scoring system and provide numbers on the historic performance of device models in the \da\ project in~\S\ref{sec:security_scoring}.
-We provide the answer to the question ``Who should I buy my Android phone from for best security?'', in general Nexus devices are better, further details are given in \S\ref{sec:security_scoring:results}.
+We provide the answer to the question ``Who should I buy my Android phone from for best security?'', showing that in general Nexus devices are better, further details are given in~\S\ref{sec:security_scoring:results}.
 
 The Android update process is slow because its development is distributed among many parties and has a long multi-stage pipeline~\cite{HTC2013}.
-The security of Android relies on many open source projects, include the Linux kernel, OpenSSL and BouncyCastle as well as Google who build the core platform.
+The security of Android relies on many open source projects, these include the Linux kernel, OpenSSL and BouncyCastle as well as Google who build the core platform.
 In addition, device manufacturers (we know of \daNumManufacturers\footnote{\label{foot:dadata}We computed this from the \da~\cite{Wagner2013} data, see \S\ref{sec:android_update_process}.}) and network operators (\daNumOperators\dafoot) may make further modifications before devices are shipped to customers.
 Understanding this ecosystem is important as device manufacturers have introduced additional vulnerabilities in the past~\cite{Grace2012}.
-We present a better understanding of both the ecosystem of Android development and associated vulnerabilities more fully in \S\ref{sec:android_ecosystem}.
+We present a better understanding of both the ecosystem of Android development and associated vulnerabilities more fully in~\S\ref{sec:android_ecosystem}.
 
@@
 These scripts analysed the Android Open Source Project's source tree to examine the source code of each of the external projects to find the project version associated with each Android version tag on the repository.
-There are \avoNumExternalProjects\ external open source projects in Android, contributing \avoTotalExternalLines\ lines of code.\footnote{Lines of code were measured using David A.
-Wheeler's \texttt{sloccount}.}
+There are \avoNumExternalProjects\ external open source projects in Android, contributing \avoTotalExternalLines\ lines of code.
 We analysed the top \avoNumBigExternalProjects\ by lines of code (\avoBigExternalLinesOfCodePerc\ of the total) and were able to automatically extract the versions of those projects included in different versions of Android for \avoNumAnalysedExternalProjects\ of these (\avoAnalysedExternalLinesOfCodePerc\ of the total).
@@
 
+\sloppy
 \section*{Acknowledgements}
@@
 }
-Some of the raw data and source code will be made available.
+Some of the raw and processed data and source code will be made available.
 
-\sloppy
 \printbibliography

### 2015-06-04 ccs.tex
Talk about standard deviation

@@
 
+We indicate the uncertainty in our results by presenting them $\pm$ one standard deviation and give results to 3~s.f., this occasionally results in `$\pm\, 0$' when the standard deviation is small.
+\todolater{Do we want to use the 95 percentile instead}
+We explore systematic errors in \S\ref{sec:representative}.
+
 \section{Android Ecosystem}\label{sec:android_ecosystem}
@@
 
-\section{Comparison with other data}
+\section{Comparison with other data}\label{sec:representative}
 Here we show that the \da\ data~\cite{Wagner2013} is representative for our purposes by comparing it with two other sources of data which provide an upper and lower bound on it.

### 2015-06-04 ccs.tex
Abbreviated threat models and vulnerabilities details from securityupdates.tex into ccs.tex

@@
 
-\todo{Do we want to talk about more stuff from the Esorics submission?}
+\section{Threat model and attack vectors}
+\label{sec:threatmodel}
+
+We assume the defender can deploy security controls at three levels: in an online marketplace, at app installation, and during app execution.
+Here we consider three attack vectors against Android handsets.
+
+The first attack vector is through the installation of a malicious app on the device.
+Android devices can install apps through marketplaces, email attachments, URLs and via the Android Debug Bridge (ADB).
+By default, many Android devices will only allow the installation of apps from Google play which uses Bouncer to automatically analyse apps; and quickly takes down apps that are reported as malicious.
+However alternative markets are also popular, particularly in countries where Google Play is not available.
+
+The second attack vector is to make an existing app into download and execute code at runtime.
+The most direct method is to upload to a marketplace a seemingly innocent app that dynamically loads malicious code.
+Neither static nor dynamic analysis of this app will uncover any malicious code, since it does not exist in the app.
+The marketplace can detect explicit use of dynamic code loading, however dynamic code loading is not actually needed for this to work -- it just makes the task easier -- even on a platform such as iOS, which does not permit dynamic code loading, a Return-Oriented Programming (ROP) attack is relatively easy if the attacker creates an app with carefully crafted flaws~\cite{Wang2013a}.
+
+The third attack vector is that the attacker injects malicious code directly into existing code already on the handset. 
+For example, the addJavascriptInterface vulnerability (CVE-2012-6636) allows JavaScript running in an Android WebView to execute arbitrary code as the vulnerable app's user~\cite{Thomas2015a}.
+The fix for the addJavascriptInterface vulnerability breaks backwards compatibility and requires a two sided fix.
+While the fix was released in December 2012, on \daGPAPISeventeenLaterDate, \daGPAPISeventeenEarlierProportion\ of handsets connecting to Google Play were still vulnerable to this attack.
+
+The best place to prevent attacks is at runtime, since all three attack vectors can be prevented at this level. 
+Unfortunately, it has been demonstrated that the sandbox for Android apps is weak~\cite{androidvulnerabilities.org}.
+
+\section{Critical vulnerabilities}
+We have used data on critical vulnerabilities in Android from AndroidVulnerabilities.org (\avo)~\cite{androidvulnerabilities.org}.
+It contains information on the discovery and publication dates, the versions affected and which versions contained fixes.
+It only contains critical vulnerabilities such as root vulnerabilities that did not require USB debugging to exploit.
+If malicious code exploits a critical vulnerability then it gains control of the device.
+
+Some critical vulnerabilities are not traditional kernel vulnerabilities, but are vulnerabilities in level 1 of our threat model.
+For example improper verification of signatures at installation time was discovered in February 2013~\cite{Forristal2013} and meant that applications could pretend to be signed with system keys and hence be granted system privileges.
+On versions of Android below 4.1 malware could then use known system-to-root escalation mechanisms.
+Regardless of version, this exposed an increased attack area and would also provide the ability for malware to control all user internet traffic (via VPNs), brick the phone, remove and install apps, steal user credentials and read the screen.
+The different categories which the vulnerabilities fall into are shown in Table~\ref{tab:andvulns}, `signature' vulnerabilities are level 1 while kernel and system vulnerabilities are level 2.\todo{What is going on with these levels?}
+\avoTabAndVulns
+
+\avo\ currently contains \avoNumVulnerabilities\ vulnerabilities of which \avoNumVulnAllAndroid\ affect all Android devices and \avoNumVulnSpecific\ are specific to particular devices or device manufacturers.
+
+Unless otherwise stated, we use the \daNumVulnsUsed\ vulnerabilities shown in Table~\ref{tab:andvulns} in our analysis.
+We have chosen these vulnerabilities to fit the attack vectors introduced in \S\ref{sec:threatmodel}.
+These vulnerabilities affect all Android devices regardless of manufacturer, and as a result our selected vulnerabilities will dominate any security analysis of Android.
+\todo{can we make this sentence clearer?}
+Hence, with our chosen set of vulnerabilities, our analysis represents a lower-bound on the vulnerability of devices in the \da\ data.
 

### 2015-06-04 ccs.tex
Better meet EPSRC requirements for Acknowledgements and datatset

@@
 
+\section*{Dataset}
+The raw and processed data and source code will be made available where possible.
+Data from \da\ and \avo\ used in this paper is already available.
+
 \sloppy
@@
 \identifying{
-This work was supported by a Google focussed research award and the EPSRC Standard Research Grant EP/P505445/1.
+This work was supported by a Google focussed research award; and the EPSRC [grant number EP/P505445/1].
 }
-Some of the raw and processed data and source code will be made available.
 

### 2015-06-04 ccs.tex
Add comparison with GP API data

@@
 \section{Comparison with other data}\label{sec:representative}
-Here we show that the \da\ data~\cite{Wagner2013} is representative for our purposes by comparing it with two other sources of data which provide an upper and lower bound on it.
+Here we show that the \da\ data~\cite{Wagner2013} is representative for our purposes by comparing it with two other sources of data.
 The \da\ data is collected from devices across the world where users have voluntarily installed the app.
@@
 \todo{Can we quantify this?}
-This shows that \da\ lies between these two biased samples while being similar to both and so is representative.
+
+Unfortunately there is no ground truth of OS version information.
+However we have collated the API version information that Google has published most months since 2009.\footnote{\url{http://androidvulnerabilities.org/play/historicplaydashboard}}
+While API versions are too coarse grained to use for security update detection they are closely related to OS versions and so if the \da\ data on API versions are similar to the Google data on API versions then the \da\ data on OS versions should be representative.
+We compared the data from Google and from \da\ and they appear similar.
+We analysed the difference between the API version data from \da\ and Google Play, normalising for days since the API version was released.
+This shows that the \da\ data systematically overestimates the prevalence of new API versions and underestimates the prevalence of old API versions.
+This means that the OS version information from \da\ is likely to be overestimating the prevalence of new OS versions and hence our results are conservative.
 

### 2015-06-04 versions.py
Update versions data

@@ -85,7 +85,12 @@ def load_from_csv(name):
 
 def update_versions():
     targz = re.compile(r'.*\.tar\.gz.*')
-    openssl_version_dates = get_version_dates(get_lines('https://www.openssl.org/source/', targz),re.compile(r'.*"openssl-([^"]+)\.tar\.gz".*'),re.compile(r'.*?\d+ ([^<]+) <.*')) 
+    openssl_lines = get_lines('https://www.openssl.org/source/', targz)
+    openssl_lines.extend(get_lines('https://www.openssl.org/source/old/0.9.x/', targz))
+    openssl_lines.extend(get_lines('https://www.openssl.org/source/old/1.0.0/', targz))
+    openssl_lines.extend(get_lines('https://www.openssl.org/source/old/1.0.1/', targz))
+    openssl_lines.extend(get_lines('https://www.openssl.org/source/old/1.0.2/', targz))
+    openssl_version_dates = get_version_dates(openssl_lines, re.compile(r'.*"openssl-([^"]+)\.tar\.gz".*'),re.compile(r'.*?\d+ ([^<]+) <.*'))
     set_latex_value('opensslNumVersions', num_in_range(openssl_version_dates))
     save_to_csv('openssl', openssl_version_dates)
 

### 2015-06-04 countversions.sh
Remove old unused countversions.sh

@@ -1,28 +1,0 @@
-#!/bin/bash
-
-set -e
-
-start_date='2011-07-01'
-end_date='2014-07-31'
-function do_count {
-    name=$1
-    output=$2
-    count=0
-    for i in $output
-    do
-        i=`echo $i | tr '_' ' '`
-        d=`date --date="$i" --rfc-3339='date'`
-        if [[ $d > $start_date ]]
-        then
-            if [[ $d < $end_date  ]]
-            then
-                count=$((count+1))
-            fi
-        fi
-    done
-    echo "\\newcommand{\\${name}NumVersions}{$count}"
-}
-do_count 'openssl' "`wget --quiet -O - https://www.openssl.org/source/ | grep '.tar.gz' | cut -d'<' -f 1 |  sed 's/^ [0-9]\+ //' | tr ' ' '_'`"
-do_count 'linux' "`wget --quiet -O - https://www.kernel.org/pub/linux/kernel/v2.6/ | grep 'linux-' | grep '.tar.gz' | tr -s ' ' | cut -d' ' -f 3`
-`wget --quiet -O - https://www.kernel.org/pub/linux/kernel/v3.0/ | grep 'linux-' | grep '.tar.gz' | tr -s ' ' | cut -d' ' -f 3`"
-

### 2015-06-04 dastats.tex
./update.sh plus manual forward ports

@@
 \newcommand{\daStartDate}{2011-07-01}
-\newcommand{\daEndDate}{2015-05-26}
+\newcommand{\daEndDate}{2015-05-27}
 \newcommand{\daFullDeployedAt}{95\%}
@@
 \newcommand{\daNumOperators}{1\,440}
-\newcommand{\daNumSigOperators}{15}
+\newcommand{\daNumSigOperators}{14}
 \newcommand{\daNumOSDevices}{20\,300}
@@
 \newcommand{\daVulnFree}{$0.121 \pm 0.000$}
-\newcommand{\daMeanOutstandingVulnerabilities}{$0.771 \pm 2.66e$}
-\newcommand{\daUpdatednessNominal}{0.0512}
-\newcommand{\daUpdatedness}{$0.0512 \pm 0.0002$}
-\newcommand{\daUpdatednessPercNominal}{5.12\%}
-\newcommand{\daUpdatednessPerc}{$5.12 \pm 0.02\%$}
+\newcommand{\daMeanOutstandingVulnerabilities}{$0.549 \pm 2.46e$}
+\newcommand{\daUpdatednessNominal}{0.0511}
+\newcommand{\daUpdatedness}{$0.0511 \pm 0.0002$}
+\newcommand{\daUpdatednessPercNominal}{5.11\%}
+\newcommand{\daUpdatednessPerc}{$5.11 \pm 0.02\%$}
 \newcommand{\daUpdatednessPercTwosfNominal}{5.1\%}
 \newcommand{\daUpdatednessPercTwosf}{$5.1 \pm 0.0\%$}
-\newcommand{\daSecurityScoreNominal}{2.53}
-\newcommand{\daSecurityScore}{$2.53 \pm 0.00$}
-\newcommand{\daCurrentMeanInsecurityPercNominal}{87.1\%}
-\newcommand{\daCurrentMeanInsecurityPerc}{$87.1 \pm 0.1\%$}
-\newcommand{\daCurrentMeanInsecurityPercTwosfNominal}{87\%}
-\newcommand{\daCurrentMeanInsecurityPercTwosf}{$87 \pm 0\%$}
-\newcommand{\daCurrentMeanSecurityPercNominal}{12.9\%}
-\newcommand{\daCurrentMeanSecurityPerc}{$12.9 \pm 0.1\%$}
-\newcommand{\daCurrentMeanSecurityPercTwosfNominal}{13\%}
-\newcommand{\daCurrentMeanSecurityPercTwosf}{$13 \pm 0\%$}
-\newcommand{\daCurrentVulnFreeNominal}{0.129}
-\newcommand{\daCurrentVulnFree}{$0.129 \pm 0.001$}
-\newcommand{\daCurrentMeanOutstandingVulnerabilities}{$0.209 \pm 9.49e$}
-\newcommand{\daCurrentUpdatednessNominal}{0.0329}
-\newcommand{\daCurrentUpdatedness}{$0.0329 \pm 0.0008$}
-\newcommand{\daCurrentUpdatednessPercNominal}{3.29\%}
-\newcommand{\daCurrentUpdatednessPerc}{$3.29 \pm 0.08\%$}
-\newcommand{\daCurrentUpdatednessPercTwosfNominal}{3.3\%}
-\newcommand{\daCurrentUpdatednessPercTwosf}{$3.3 \pm 0.0\%$}
-\newcommand{\daCurrentSecurityScoreNominal}{3.3}
-\newcommand{\daCurrentSecurityScore}{$3.3 \pm 0.0$}
+\newcommand{\daSecurityScoreNominal}{2.83}
+\newcommand{\daSecurityScore}{$2.83 \pm 0.00$}
+\newcommand{\daCurrentMeanInsecurityPercNominal}{85.9\%}
+\newcommand{\daCurrentMeanInsecurityPerc}{$85.9 \pm 0.1\%$}
+\newcommand{\daCurrentMeanInsecurityPercTwosfNominal}{86\%}
+\newcommand{\daCurrentMeanInsecurityPercTwosf}{$86 \pm 0\%$}
+\newcommand{\daCurrentMeanSecurityPercNominal}{14.1\%}
+\newcommand{\daCurrentMeanSecurityPerc}{$14.1 \pm 0.1\%$}
+\newcommand{\daCurrentMeanSecurityPercTwosfNominal}{14\%}
+\newcommand{\daCurrentMeanSecurityPercTwosf}{$14 \pm 0\%$}
+\newcommand{\daCurrentVulnFreeNominal}{0.141}
+\newcommand{\daCurrentVulnFree}{$0.141 \pm 0.001$}
+\newcommand{\daCurrentMeanOutstandingVulnerabilities}{$0.201 \pm 1.16e$}
+\newcommand{\daCurrentUpdatednessNominal}{0.0322}
+\newcommand{\daCurrentUpdatedness}{$0.0322 \pm 0.0007$}
+\newcommand{\daCurrentUpdatednessPercNominal}{3.22\%}
+\newcommand{\daCurrentUpdatednessPerc}{$3.22 \pm 0.07\%$}
+\newcommand{\daCurrentUpdatednessPercTwosfNominal}{3.2\%}
+\newcommand{\daCurrentUpdatednessPercTwosf}{$3.2 \pm 0.0\%$}
+\newcommand{\daCurrentSecurityScoreNominal}{3.36}
+\newcommand{\daCurrentSecurityScore}{$3.36 \pm 0.00$}
 \newcommand{\daDailyParticipationNominal}{903}
@@
 \newcommand{\daNumAPIVersions}{16}
-\newcommand{\daNumFullVersionUpdates}{4\,520}
+\newcommand{\daNumFullVersionUpdates}{4\,530}
 \newcommand{\daNumOSVersionUpdates}{3\,840}
@@
 \newcommand{\daNumPossibleSecurityUpdates}{2\,290}
-\newcommand{\daOSCurveFitParamFirst}{85.0}
+\newcommand{\daOSCurveFitParamFirst}{84.8}
 \newcommand{\daOSCurveFitParamSecond}{\num{0.0029}}
@@
 \newcommand{\daOSCurveSplineRMSE}{0.116}
-\newcommand{\daOSCurveHalfDeployedYearsNominal}{0.887}
-\newcommand{\daOSCurveHalfDeployedYears}{$0.887 \pm 0.248$}
+\newcommand{\daOSCurveHalfDeployedYearsNominal}{0.888}
+\newcommand{\daOSCurveHalfDeployedYears}{$0.888 \pm 0.249$}
 \newcommand{\daOSCurveHalfDeployedNominal}{324}
@@
 \newcommand{\daAPICurveHalfDeployedYearsNominal}{0.934}
-\newcommand{\daAPICurveHalfDeployedYears}{$0.934 \pm 0.242$}
+\newcommand{\daAPICurveHalfDeployedYears}{$0.934 \pm 0.243$}
 \newcommand{\daAPICurveHalfDeployedNominal}{341}
@@
 \newcommand{\daAPICurveFullDeployedYears}{$3.11 \pm 70$}
-\newcommand{\daAPICurveFullDeployedNominal}{1\,130}
-\newcommand{\daAPICurveFullDeployed}{$1\,130 \pm 256\,000$}
+\newcommand{\daAPICurveFullDeployedNominal}{1\,140}
+\newcommand{\daAPICurveFullDeployed}{$1\,140 \pm 256\,000$}
 \newcommand{\daVulnAPKDuplicateFileOctoberPerc}{91.6\%}
 \newcommand{\daVulnZergRushMonthsDefFixDeployed}{27.3}
-\newcommand{\daTabSecScoresmanufacturer}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline LG & $0.20 \pm 0.00$ & $0.32 \pm 0.00$ & $0.88 \pm 0.00$ & $3.51 \pm 0.01$ \\  Motorola & $0.17 \pm 0.00$ & $0.12 \pm 0.00$ & $1.10 \pm 0.01$ & $2.57 \pm 0.01$ \\  Samsung & $0.13 \pm 0.00$ & $0.04 \pm 0.00$ & $0.88 \pm 0.00$ & $2.39 \pm 0.00$ \\  asus & $0.19 \pm 0.00$ & $0.52 \pm 0.01$ & $21.42 \pm 0.09$ & $2.30 \pm 0.02$ \\  Sony & $0.13 \pm 0.00$ & $0.19 \pm 0.00$ & $1.68 \pm 0.01$ & $2.04 \pm 0.01$ \\  HTC & $0.14 \pm 0.00$ & $0.09 \pm 0.00$ & $1.40 \pm 0.00$ & $2.02 \pm 0.01$ \\  other & $0.06 \pm 0.00$ & $0.04 \pm 0.00$ & $1.77 \pm 0.00$ & $1.25 \pm 0.00$ \\  Symphony & $0.00 \pm 0.00$ & $0.08 \pm 0.00$ & $15.00 \pm 0.03$ & $0.24 \pm 0.01$ \\  walton & $0.00 \pm 0.00$ & $0.08 \pm 0.00$ & $20.79 \pm 0.06$ & $0.23 \pm 0.01$ \\ \end{tabular} \caption{Security scores for manufacturers} \label{tab:sec_manufacturer} \end{table*}}
+\newcommand{\daTabSecScoresmanufacturer}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline LG & $0.20 \pm 0.00$ & $0.32 \pm 0.00$ & $0.64 \pm 0.01$ & $3.83 \pm 0.02$ \\  Motorola & $0.17 \pm 0.00$ & $0.12 \pm 0.00$ & $0.73 \pm 0.02$ & $3.03 \pm 0.02$ \\  Samsung & $0.13 \pm 0.00$ & $0.04 \pm 0.00$ & $0.63 \pm 0.00$ & $2.72 \pm 0.00$ \\  Sony & $0.13 \pm 0.00$ & $0.19 \pm 0.00$ & $1.09 \pm 0.02$ & $2.61 \pm 0.02$ \\  HTC & $0.14 \pm 0.00$ & $0.09 \pm 0.00$ & $0.89 \pm 0.01$ & $2.58 \pm 0.02$ \\  asus & $0.19 \pm 0.00$ & $0.52 \pm 0.01$ & $6.01 \pm 0.07$ & $2.31 \pm 0.02$ \\  other & $0.06 \pm 0.00$ & $0.04 \pm 0.00$ & $1.04 \pm 0.01$ & $1.94 \pm 0.02$ \\  Symphony & $0.00 \pm 0.00$ & $0.07 \pm 0.00$ & $5.00 \pm 0.04$ & $0.27 \pm 0.01$ \\  walton & $0.00 \pm 0.00$ & $0.08 \pm 0.00$ & $5.96 \pm 0.08$ & $0.25 \pm 0.01$ \\ \end{tabular} \caption{Security scores for manufacturers} \label{tab:sec_manufacturer} \end{table*}}
 \newcommand{\daSecScoreBestmanufacturer}{LG}
-\newcommand{\daSecScoreBestmanufacturerScoreNominal}{3.51}
-\newcommand{\daSecScoreBestmanufacturerScore}{$3.51 \pm 0.01$}
-\newcommand{\daSecScoreBestmanufacturerNumFullVersions}{130}
+\newcommand{\daSecScoreBestmanufacturerScoreNominal}{3.83}
+\newcommand{\daSecScoreBestmanufacturerScore}{$3.83 \pm 0.02$}
+\newcommand{\daSecScoreBestmanufacturerNumFullVersions}{131}
 \newcommand{\daSecScoreWorstmanufacturer}{walton}
-\newcommand{\daSecScoreWorstmanufacturerScoreNominal}{0.235}
-\newcommand{\daSecScoreWorstmanufacturerScore}{$0.235 \pm 0.007$}
+\newcommand{\daSecScoreWorstmanufacturerScoreNominal}{0.251}
+\newcommand{\daSecScoreWorstmanufacturerScore}{$0.251 \pm 0.007$}
 \newcommand{\daSecScoreWorstmanufacturerNumFullVersions}{22}
-\newcommand{\daTabSecScoresmodel}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline Galaxy Nexus & $0.51 \pm 0.00$ & $0.55 \pm 0.01$ & $2.18 \pm 0.02$ & $4.29 \pm 0.03$ \\  Nexus 4 & $0.28 \pm 0.00$ & $0.82 \pm 0.01$ & $21.70 \pm 0.12$ & $3.58 \pm 0.04$ \\  Nexus 7 & $0.24 \pm 0.00$ & $0.74 \pm 0.01$ & $20.68 \pm 0.11$ & $3.16 \pm 0.04$ \\  Desire HD & $0.08 \pm 0.00$ & $0.05 \pm 0.00$ & $0.47 \pm 0.01$ & $2.78 \pm 0.02$ \\  other & $0.09 \pm 0.00$ & $0.14 \pm 0.00$ & $0.77 \pm 0.00$ & $2.68 \pm 0.00$ \\  HTC Sensation & $0.35 \pm 0.00$ & $0.01 \pm 0.01$ & $2.44 \pm 0.03$ & $1.88 \pm 0.03$ \\  HTC Desire S & $0.02 \pm 0.00$ & $0.02 \pm 0.00$ & $1.00 \pm 0.02$ & $1.75 \pm 0.03$ \\  GT-I9100 & $0.22 \pm 0.00$ & $0.02 \pm 0.00$ & $2.08 \pm 0.01$ & $1.59 \pm 0.01$ \\  GT-N7000 & $0.25 \pm 0.00$ & $0.00 \pm 0.00$ & $4.70 \pm 0.03$ & $1.04 \pm 0.02$ \\  GT-I9300 & $0.13 \pm 0.00$ & $0.01 \pm 0.00$ & $23.21 \pm 0.04$ & $0.57 \pm 0.01$ \\  GT-I9505 & $0.01 \pm 0.00$ & $0.13 \pm 0.00$ & $26.96 \pm 0.08$ & $0.43 \pm 0.01$ \\  GT-P1000 & $0.01 \pm 0.00$ & $0.00 \pm 0.01$ & $3.17 \pm 0.03$ & $0.28 \pm 0.02$ \\  GT-N7100 & $0.06 \pm 0.00$ & $0.00 \pm 0.01$ & $27.81 \pm 0.11$ & $0.24 \pm 0.02$ \\  HTC Desire HD & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & $6.23 \pm 0.03$ & $0.01 \pm 0.02$ \\  Symphony W68 & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & $66.00 \pm 0.19$ & $0.00 \pm 0.03$ \\ \end{tabular} \caption{Security scores for models} \label{tab:sec_model} \end{table*}}
+\newcommand{\daTabSecScoresmodel}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline Galaxy Nexus & $0.51 \pm 0.00$ & $0.55 \pm 0.01$ & $1.53 \pm 0.04$ & $4.74 \pm 0.04$ \\  Nexus 4 & $0.28 \pm 0.00$ & $0.82 \pm 0.01$ & $6.06 \pm 0.09$ & $3.59 \pm 0.04$ \\  Nexus 7 & $0.24 \pm 0.00$ & $0.74 \pm 0.01$ & $5.92 \pm 0.09$ & $3.18 \pm 0.04$ \\  other & $0.09 \pm 0.00$ & $0.14 \pm 0.00$ & $0.55 \pm 0.00$ & $2.98 \pm 0.00$ \\  Desire HD & $0.08 \pm 0.00$ & $0.05 \pm 0.00$ & $0.39 \pm 0.02$ & $2.91 \pm 0.04$ \\  HTC Sensation & $0.35 \pm 0.00$ & $0.01 \pm 0.01$ & $1.59 \pm 0.05$ & $2.42 \pm 0.05$ \\  GT-I9100 & $0.22 \pm 0.00$ & $0.02 \pm 0.00$ & $1.23 \pm 0.02$ & $2.28 \pm 0.02$ \\  HTC Desire S & $0.02 \pm 0.00$ & $0.02 \pm 0.00$ & $1.00 \pm 0.06$ & $1.75 \pm 0.07$ \\  GT-N7000 & $0.25 \pm 0.00$ & $0.00 \pm 0.00$ & $2.52 \pm 0.05$ & $1.43 \pm 0.02$ \\  GT-P1000 & $0.01 \pm 0.00$ & $0.00 \pm 0.01$ & $1.79 \pm 0.06$ & $0.90 \pm 0.05$ \\  GT-I9300 & $0.13 \pm 0.00$ & $0.01 \pm 0.00$ & $6.25 \pm 0.04$ & $0.58 \pm 0.01$ \\  GT-I9505 & $0.01 \pm 0.00$ & $0.13 \pm 0.00$ & $6.82 \pm 0.07$ & $0.44 \pm 0.01$ \\  HTC Desire HD & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & $3.05 \pm 0.05$ & $0.27 \pm 0.03$ \\  GT-N7100 & $0.06 \pm 0.00$ & $0.00 \pm 0.01$ & $6.91 \pm 0.08$ & $0.25 \pm 0.02$ \\  Symphony W68 & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & $11.00 \pm 0.12$ & $0.00 \pm 0.03$ \\ \end{tabular} \caption{Security scores for models} \label{tab:sec_model} \end{table*}}
 \newcommand{\daSecScoreBestmodel}{Galaxy Nexus}
-\newcommand{\daSecScoreBestmodelScoreNominal}{4.29}
-\newcommand{\daSecScoreBestmodelScore}{$4.29 \pm 0.02$}
+\newcommand{\daSecScoreBestmodelScoreNominal}{4.74}
+\newcommand{\daSecScoreBestmodelScore}{$4.74 \pm 0.04$}
 \newcommand{\daSecScoreBestmodelNumFullVersions}{49}
 \newcommand{\daSecScoreWorstmodel}{Symphony W68}
-\newcommand{\daSecScoreWorstmodelScoreNominal}{1.3e-28}
-\newcommand{\daSecScoreWorstmodelScore}{$1.3e-28 \pm 0.0276$}
+\newcommand{\daSecScoreWorstmodelScoreNominal}{0.0001}
+\newcommand{\daSecScoreWorstmodelScore}{$0.0001 \pm 0.0276$}
 \newcommand{\daSecScoreWorstmodelNumFullVersions}{1}
-\newcommand{\daTabSecScoressummary}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline nexus & $0.37 \pm 0.00$ & $0.47 \pm 0.00$ & $0.82 \pm 0.00$ & $4.75 \pm 0.01$ \\  notnexus & $0.10 \pm 0.00$ & $0.02 \pm 0.00$ & $0.77 \pm 0.00$ & $2.37 \pm 0.00$ \\ \end{tabular} \caption{Security scores for nexus} \label{tab:sec_summary} \end{table*}}
-\newcommand{\daTabSecScoresoperator}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline O2 uk & $0.26 \pm 0.00$ & $0.12 \pm 0.00$ & $0.51 \pm 0.01$ & $3.65 \pm 0.02$ \\  T-Mobile & $0.21 \pm 0.00$ & $0.18 \pm 0.00$ & $0.54 \pm 0.01$ & $3.56 \pm 0.02$ \\  Orange & $0.21 \pm 0.00$ & $0.09 \pm 0.00$ & $0.48 \pm 0.01$ & $3.39 \pm 0.02$ \\  Sprint & $0.18 \pm 0.00$ & $0.11 \pm 0.00$ & $0.51 \pm 0.01$ & $3.31 \pm 0.02$ \\  3 & $0.18 \pm 0.00$ & $0.09 \pm 0.00$ & $0.70 \pm 0.01$ & $2.99 \pm 0.02$ \\  AT\&T & $0.12 \pm 0.00$ & $0.08 \pm 0.00$ & $0.54 \pm 0.01$ & $2.94 \pm 0.01$ \\  Vodafone uk & $0.13 \pm 0.00$ & $0.13 \pm 0.00$ & $0.73 \pm 0.01$ & $2.88 \pm 0.02$ \\  unknown & $0.11 \pm 0.00$ & $0.20 \pm 0.00$ & $1.22 \pm 0.00$ & $2.40 \pm 0.01$ \\  Verizon & $0.18 \pm 0.00$ & $0.09 \pm 0.00$ & $1.27 \pm 0.01$ & $2.30 \pm 0.01$ \\  n Telenor & $0.03 \pm 0.00$ & $0.12 \pm 0.00$ & $1.86 \pm 0.00$ & $1.29 \pm 0.01$ \\  Airtel & $0.05 \pm 0.00$ & $0.03 \pm 0.00$ & $2.42 \pm 0.00$ & $0.77 \pm 0.01$ \\  Robi & $0.00 \pm 0.00$ & $0.08 \pm 0.00$ & $3.26 \pm 0.02$ & $0.46 \pm 0.01$ \\  Grameenphone & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & $3.09 \pm 0.00$ & $0.41 \pm 0.00$ \\  banglalink & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & $4.83 \pm 0.01$ & $0.17 \pm 0.01$ \\ \end{tabular} \caption{Security scores for operators} \label{tab:sec_operator} \end{table*}}
+\newcommand{\daTabSecScoressummary}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline nexus & $0.37 \pm 0.00$ & $0.47 \pm 0.00$ & $0.59 \pm 0.01$ & $5.06 \pm 0.02$ \\  notnexus & $0.10 \pm 0.00$ & $0.02 \pm 0.00$ & $0.55 \pm 0.00$ & $2.66 \pm 0.00$ \\ \end{tabular} \caption{Security scores for nexus} \label{tab:sec_summary} \end{table*}}
+\newcommand{\daTabSecScoresoperator}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline O2 uk & $0.26 \pm 0.00$ & $0.12 \pm 0.00$ & $0.37 \pm 0.02$ & $3.84 \pm 0.03$ \\  T-Mobile & $0.21 \pm 0.00$ & $0.18 \pm 0.00$ & $0.42 \pm 0.02$ & $3.73 \pm 0.03$ \\  Orange & $0.21 \pm 0.00$ & $0.09 \pm 0.00$ & $0.39 \pm 0.03$ & $3.52 \pm 0.04$ \\  Sprint & $0.18 \pm 0.00$ & $0.11 \pm 0.00$ & $0.40 \pm 0.02$ & $3.46 \pm 0.03$ \\  3 & $0.18 \pm 0.00$ & $0.09 \pm 0.00$ & $0.52 \pm 0.03$ & $3.23 \pm 0.04$ \\  Vodafone uk & $0.13 \pm 0.00$ & $0.13 \pm 0.00$ & $0.50 \pm 0.03$ & $3.18 \pm 0.04$ \\  AT\&T & $0.13 \pm 0.00$ & $0.08 \pm 0.00$ & $0.43 \pm 0.02$ & $3.09 \pm 0.02$ \\  Verizon & $0.18 \pm 0.00$ & $0.09 \pm 0.00$ & $0.82 \pm 0.02$ & $2.83 \pm 0.02$ \\  unknown & $0.11 \pm 0.00$ & $0.20 \pm 0.00$ & $0.87 \pm 0.02$ & $2.80 \pm 0.02$ \\  n Telenor & $0.03 \pm 0.00$ & $0.12 \pm 0.00$ & $1.25 \pm 0.02$ & $1.81 \pm 0.02$ \\  Airtel & $0.05 \pm 0.00$ & $0.03 \pm 0.00$ & $1.62 \pm 0.03$ & $1.28 \pm 0.03$ \\  Robi & $0.00 \pm 0.00$ & $0.08 \pm 0.00$ & $2.08 \pm 0.04$ & $0.90 \pm 0.03$ \\  Grameenphone & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & $2.03 \pm 0.02$ & $0.85 \pm 0.01$ \\  banglalink & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & $2.61 \pm 0.04$ & $0.54 \pm 0.02$ \\ \end{tabular} \caption{Security scores for operators} \label{tab:sec_operator} \end{table*}}
 \newcommand{\daSecScoreBestoperator}{O2 uk}
-\newcommand{\daSecScoreBestoperatorScoreNominal}{3.65}
-\newcommand{\daSecScoreBestoperatorScore}{$3.65 \pm 0.01$}
+\newcommand{\daSecScoreBestoperatorScoreNominal}{3.84}
+\newcommand{\daSecScoreBestoperatorScore}{$3.84 \pm 0.02$}
 \newcommand{\daSecScoreBestoperatorNumFullVersions}{97}
 \newcommand{\daSecScoreWorstoperator}{banglalink}
-\newcommand{\daSecScoreWorstoperatorScoreNominal}{0.172}
-\newcommand{\daSecScoreWorstoperatorScore}{$0.172 \pm 0.005$}
-\newcommand{\daSecScoreWorstoperatorNumFullVersions}{73}
-\newcommand{\daTabDLDistances}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & summary \\ \hline equal & 0.0 & 0.0667 & 0.0714 & 0.0 \\weight $m$ & 0.222 & 0.333 & 0.0714 & 0.0 \\weight $u$ & 0.222 & 0.133 & 0.0714 & 0.0 \\$f$ & 0.444 & 0.733 & 0.5 & 0.0 \\$m$ & 0.556 & 0.8 & 0.429 & 0.5 \\$u$ & 0.889 & 0.6 & 0.786 & 0.0 \\\end{tabular} \caption{Normalised Damerau-Levenshtein distances for different metrics} \label{tab:dl_distances} \end{table}}
-\newcommand{\daTabChangeInScores}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & summary \\ \hline $m$ & $-1.13 \pm 1.77$ & $-0.133 \pm 2.4$ & $-2.51 \pm 1.66$ & $-2.67 \pm 1.83$ \\weight $m$ & $-0.184 \pm 0.21$ & $-0.0255 \pm 0.313$ & $-0.35 \pm 0.21$ & $-0.387 \pm 0.209$ \\equal & $-0.0792 \pm 0.0477$ & $-0.0136 \pm 0.122$ & $-0.11 \pm 0.05$ & $-0.133 \pm 0.028$ \\weight $u$ & $-0.053 \pm 0.127$ & $-0.0154 \pm 0.239$ & $0.0196 \pm 0.0792$ & $-0.0124 \pm 0.125$ \\$u$ & $0.182 \pm 1.42$ & $-0.0309 \pm 2.03$ & $1.19 \pm 1.08$ & $1.07 \pm 1.5$ \\$f$ & $0.713 \pm 0.43$ & $0.123 \pm 1.1$ & $0.991 \pm 0.536$ & $1.2 \pm 0.2$ \\\end{tabular} \caption{Mean change in scores for different metrics} \label{tab:change_in_scores} \end{table*}}
+\newcommand{\daSecScoreWorstoperatorScoreNominal}{0.535}
+\newcommand{\daSecScoreWorstoperatorScore}{$0.535 \pm 0.017$}
+\newcommand{\daSecScoreWorstoperatorNumFullVersions}{71}
+\newcommand{\daTabDLDistances}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & summary \\ \hline equal & 0.0 & 0.0 & 0.0714 & 0.0 \\weight $m$ & 0.222 & 0.333 & 0.143 & 0.0 \\$f$ & 0.444 & 0.733 & 0.429 & 0.0 \\weight $u$ & 0.444 & 0.133 & 0.214 & 0.0 \\$m$ & 0.667 & 0.733 & 0.357 & 0.5 \\$u$ & 0.778 & 0.733 & 0.714 & 0.0 \\\end{tabular} \caption{Normalised Damerau-Levenshtein distances for different metrics} \label{tab:dl_distances} \end{table}}
+\newcommand{\daTabChangeInScores}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & summary \\ \hline $m$ & $-1.9 \pm 2.2$ & $-0.666 \pm 2.51$ & $-3.27 \pm 1.39$ & $-3.37 \pm 1.83$ \\weight $m$ & $-0.295 \pm 0.274$ & $-0.102 \pm 0.316$ & $-0.458 \pm 0.177$ & $-0.487 \pm 0.209$ \\equal & $-0.116 \pm 0.065$ & $-0.039 \pm 0.113$ & $-0.146 \pm 0.049$ & $-0.166 \pm 0.028$ \\weight $u$ & $-0.0526 \pm 0.127$ & $-0.0152 \pm 0.239$ & $0.02 \pm 0.07$ & $-0.012 \pm 0.125$ \\$u$ & $0.516 \pm 1.61$ & $0.198 \pm 2.14$ & $1.51 \pm 0.96$ & $1.38 \pm 1.5$ \\$f$ & $1.04 \pm 0.58$ & $0.351 \pm 1.02$ & $1.31 \pm 0.44$ & $1.5 \pm 0.2$ \\\end{tabular} \caption{Mean change in scores for different metrics} \label{tab:change_in_scores} \end{table*}}
 \newcommand{\daGPAPICurveFitParamFirst}{80.7}
@@
 \newcommand{\daGPAPISeventeenFullDeploymentMonth}{October 2017}
-\newcommand{\daMeanOutstandingVulnerabilitiesNominal}{0.771}
-\newcommand{\daCurrentMeanOutstandingVulnerabilitiesNominal}{0.209}
-\newcommand{\daModelHalfDeployment}{27}
-\newcommand{\daModelFullDeployment}{316}
-\newcommand{\daModelHalfDeploymentYears}{0.074}
-\newcommand{\daModelFullDeploymentYears}{0.866}
+\newcommand{\daMeanOutstandingVulnerabilitiesNominal}{0.549}
+\newcommand{\daCurrentMeanOutstandingVulnerabilitiesNominal}{0.201}
+\newcommand{\daModelHalfDeploymentYears}{0.0822}
+\newcommand{\daModelHalfDeployment}{30}
+\newcommand{\daModelFullDeploymentYears}{0.89}
+\newcommand{\daModelFullDeployment}{325}
 \newcommand{\daMeanOutstandingProbability}{1\%}

### 2015-06-04 ccs.tex
Move latency table next to ecosystem figure

@@
 \end{figure}
+\begin{table}
+\centering
+\normalsize
+\begin{tabular}{l|r|r}
+Project	&	\# releases	&	latency (days) \\ \hline
+linux	&	\linuxNumVersions	&	\linuxMeanUpdateLatency \\
+openssl	&	\opensslNumVersions	&	\opensslMeanUpdateLatency \\
+bouncycastle	&	\bouncycastleNumVersions	&	\bouncycastleMeanUpdateLatency \\
+\end{tabular}
+\caption{Flow of updates from upstream projects into Android.
+Number of updates as in Figure~\ref{fig:update_ecosystem}, latency in days for all pairs of versions we have data on.\todolater{scrape the other 26 websites...
+is it worth it?}}
+\label{tab:update_ecosystem}
+\end{table}
 To understand how vulnerabilities in Android are fixed we examine the Android update process which we model in Figure~\ref{fig:update_ecosystem}.
@@
 The versions included in Android were about half a year old when the first version of Android containing it was released.
-\begin{table}
-\centering
-\normalsize
-\begin{tabular}{l|r|r}
-Project	&	\# releases	&	latency (days) \\ \hline
-linux	&	\linuxNumVersions	&	\linuxMeanUpdateLatency \\
-openssl	&	\opensslNumVersions	&	\opensslMeanUpdateLatency \\
-bouncycastle	&	\bouncycastleNumVersions	&	\bouncycastleMeanUpdateLatency \\
-\end{tabular}
-\caption{Flow of updates from upstream projects into Android.
-Number of updates as in Figure~\ref{fig:update_ecosystem}, latency in days for all pairs of versions we have data on.\todolater{scrape the other 26 websites...
-is it worth it?}}
-\label{tab:update_ecosystem}
-\end{table}
 
+\todo{We need to rerun the analysis on the AOSP data for its current state}
 \section{Threat model and attack vectors}

### 2015-06-04 ccs.tex
Tweaks based on feedback

@@
 To address this, we developed a scoring system and provide numbers on the historic performance of device models in the \da\ project in~\S\ref{sec:security_scoring}.
-We provide the answer to the question ``Who should I buy my Android phone from for best security?'', showing that in general Nexus devices are better, further details are given in~\S\ref{sec:security_scoring:results}.
+We provide the answer to the question ``Who should I buy my Android phone from for best security?'', showing that in general Nexus devices are better; further details are given in~\S\ref{sec:security_scoring:results}.
 
@@
 Understanding this ecosystem is important as device manufacturers have introduced additional vulnerabilities in the past~\cite{Grace2012}.
+The cause of the slow update process could be: upstream open source projects releasing patches slowly, Google including them slowly, manufacturers including them slowly, network operators shipping them slowly or customers not installing them.
 We present a better understanding of both the ecosystem of Android development and associated vulnerabilities more fully in~\S\ref{sec:android_ecosystem}.
@@
 \begin{itemize}
- \item We quantify the Android update process~\S\ref{sec:android_ecosystem}.
+ \item We quantify the Android update process, providing concrete numbers on the flow of updates and their latency~\S\ref{sec:android_ecosystem}.
  \item We propose the FUM scoring metric to evaluate the security of different instances of a platform~\S\ref{sec:security_scoring:method}.
@@
 \section{Android Ecosystem}\label{sec:android_ecosystem}
-There is a complex Android ecosystem which creates and distributes updates to Android which fix vulnerabilities.
-In this section we describe how the Android ecosystem functions and how Android versions are produced, as seen by the \da\ project.
+There is a complex Android ecosystem that creates and distributes updates which fix vulnerabilities.
+In this section we describe how the Android ecosystem functions and how Android versions are produced, as seen by the \da\ project and by analysing the Android source code and upstream projects.
+We quantify the number of updates shipped by various entities in the ecosystem and the number of entities.
 
@@
 This normalisation is not perfect and so these are overestimates on the \da\ data but they are likely still underestimates as there will be some device manufacturers and network operators which are not included in the \da\ data.
+The representativeness of \da\ is discussed in \S\ref{sec:representative}.
 
@@
 The number of updates shipped by Google (\daNumSigOSVersions) is the number of Android versions reported in \da\ which affected more than \daSigVersionPerc\ of devices for more than \daSigVersionDays\ days.
-This significance test is to remove spurious versions recorded in \da\ such as `5.2.0' in 2012 which had still not been released in 2014.
+This significance test is to remove spurious versions recorded in \da\ such as `5.2.0' in 2012 which had still not been released at time of writing.\todolater{Check if this has now been released}
 
@@
 \todo{We need to rerun the analysis on the AOSP data for its current state}
+\todo{What is the significance of having done this quantification?}
 \section{Threat model and attack vectors}
@@
 In this regard, iOS provides additional safeguards beyond those used in Android, including a pre-distribution review, mandatory code-signing by the manufacturer, and (with the important exception of ROP-based attacks~\cite{Wang2013a}) the technical prohibition of dynamic code loading by an app.
-These features, as well as Address Space Layout Randomisation (ASLR) and mandatory access controls, has resulted in a lower level of malware affecting iOS when compared to Android~\cite{Felt2011}.
+These features, combined with mandatory access controls, has resulted in a lower level of malware affecting iOS when compared to Android~\cite{Felt2011}.
 
@@
 
+\todo{Work these references into the text}
 \cite{Nappa2015}
@@
 \section{Conclusion}
+\todo{Write this}
  and examined how the Android ecosystem results in different device models receiving different levels of security due to whether or not they get security updates
@@
 \section*{Dataset}
-The raw and processed data and source code will be made available where possible.
+The raw and processed data and source code will be made available where possible when this paper is finalised.
 Data from \da\ and \avo\ used in this paper is already available.

### 2015-06-04 security_scoring.tex
Tweaks based on feedback

@@
 
-The FUM score is computed from several components:
+The FUM score is computed from three components:
 \begin{description}
-  \item[free $f$] The proportion of the time where devices were free from known critical vulnerabilities.
+  \item[free $f$] The proportion of running devices which were free from critical vulnerabilities over time.
 This is equivalent to Acer and Jackson's proposal to measure the security based on the proportion of users with at least one unpatched critical vulnerability~\cite{Acer2010} and similar to the Vulnerability Free Days (VFD) score~\cite{Wright2014}.
-  Unlike VFD this is the proportion of running devices which were free from critical vulnerabilities, rather than days which the device manufacturer was free from outstanding critical vulnerabilities as that does not take account of the update process.
+  Unlike VFD this is the proportion of running devices which were free from critical vulnerabilities over time, rather than days which the device manufacturer was free from outstanding critical vulnerabilities as that does not take account of the update process.
   \item[update $u$] The proportion of devices which run the latest version of Android shipped to any device produced by that device manufacturer.
 This is a measure of internal updatedness, a low score would mean many devices are being left behind.
-  This assumes that newer versions are better with stronger security, historically steps have been taken to improve Android security in newer versions so this assumption should generally hold, but sometimes new updates introduce new vulnerabilities.
+  This assumes that newer versions are better with stronger security.
+  Historically steps have been taken to improve Android security in newer versions so this assumption should generally hold, but sometimes new updates introduce new vulnerabilities.
   \item[mean $m$] The mean number of outstanding vulnerabilities affecting devices not fixed on any device shipped by the device manufacturer.
@@
 This does not capture systematic errors such as those due to not including vulnerabilities which do affect devices, but for comparison that does not matter.
-\todo{We don't include manufacturer specific vulnerabilities which will influence comparisons.}
+However we do not include manufacturer specific vulnerabilities which will influence comparisons, but we expect that performance in fixing manufacturer specific vulnerabilities is strongly correlated with performance fixing vulnerabilities affecting all of Android.
 
 \todolater{There must be some way of putting an error bound on our results by considering the error on the whole DA set and the size of the subset used for each device manufacturer and device model}
-
-If the comparative data given here is used to influence purchasing decisions then entities in the Android ecosystem might try to game the score rather than genuinely improve security.
-$f$ is hard to game without doing a good job at security but it doesn't get any worse if there is already one known vulnerability and another is found.
-A high value of $u$ could be achieved by only ever shipping one version but that would give low values for $f$ and $m$ (and not be attractive to new customers).
-A high value of $m$ could be achieved by focusing on only one device at a time and ensuring that gets updates but ignoring all others, but that would lower $f$ and $u$.%
-\footnote{One way to influence our scores would be to add additional devices to Device Analyzer which have good security, these would have to be real end user devices as we could detect fake ones, this would increase the size of our data set and would require providing genuinely good security to some users.}
 
@@
 Figure~\ref{fig:full_version_comp}b shows the full version distribution for \emph{\daSecScoreWorstmodel} and Figure~\ref{fig:full_version_comp}a shows the same for \emph{\daSecScoreBestmodel}.
-There is a marked difference in the distribution of updates between the two with only \daSecScoreWorstmodelNumFullVersions\ full versions for \emph{\daSecScoreWorstmodel} and \daSecScoreBestmodelNumFullVersions\ for \emph{\daSecScoreBestmodel}.
+There is a marked difference in the distribution of updates between the two with only \daSecScoreWorstmodelNumFullVersions\ full version for \emph{\daSecScoreWorstmodel} and \daSecScoreBestmodelNumFullVersions\ for \emph{\daSecScoreBestmodel}.
 %Both device models start off with the full version string of `2.3.3 GRI40' but the \emph{\daSecScoreBestmodel} receives many more updates over the same time period.
@@
 An added value analysis of network operators which takes into account the device mix used by users of that network operator would make it possible to determine whether a network operator is making the situation better or worse by the way it ships updates to users.
-However we suspect that our sample size is too small to do that as while we have significant numbers of devices for each of the device models in Table~\ref{tab:sec_model} and for each of the network operators in Table~\ref{tab:sec_operator} we would need a significant number of each model in each network operator.
+However our sample size is too small to do that as while we have significant numbers of devices for each of the device models (Table~\ref{tab:sec_model}) and for each of the network operators (Table~\ref{tab:sec_operator}) we would need a significant number of each model in each network operator.
 We do not attempt to disambiguate users behaviour in whether they install updates from network operators using rolling upgrades.
@@
 \end{equation}
-Figure~\ref{fig:security_score_manufacturer} shows this for manufacturers, Figure~\ref{fig:security_score_model} for device models, Figure~\ref{fig:security_score_operator} for operators Figure~\ref{fig:security_score_summary} for Nexus and non-Nexus devices.
+Figure~\ref{fig:security_score_manufacturer} shows this for manufacturers, Figure~\ref{fig:security_score_model} for device models, Figure~\ref{fig:security_score_operator} for operators Figure~\ref{fig:security_score_summary} for Nexus and non-Nexus devices; these figures are shown with their 95\% confidence intervals.
 These show how the scores for different entities are different and change over time, while there is correlated behaviour for different entities (due to things like new vulnerabilities affecting all Android being discovered) these lines still have crossings due to the different behaviour of the different entities.
@@
 \todo{Sensitivity of time based vs. global scores}
+
+\subsection{Gaming the score}
+If the comparative data given here is used to influence purchasing decisions then entities in the Android ecosystem might try to game the score rather than genuinely improve security.
+$f$ is hard to game without doing a good job at security but it doesn't get any worse if there is already one known vulnerability and another is found.
+A high value of $u$ could be achieved by only ever shipping one version but that would give low values for $f$ and $m$ (and not be attractive to new customers).
+A high value of $m$ could be achieved by focusing on only one device at a time and ensuring that gets updates but ignoring all others, but that would lower $f$ and $u$.%
+\footnote{One way to influence our scores would be to add additional devices to Device Analyzer which have good security, these would have to be real end user devices as we could detect fake ones, this would increase the size of our data set and would require providing genuinely good security to some users.}
+Therefore our score is secure against gaming.

### 2015-06-04 ccs.tex
Add Sherif and Ripduman to Acknowledgements
Note that Rwanda and FTSE data cannot be made available

@@
 The raw and processed data and source code will be made available where possible when this paper is finalised.
-Data from \da\ and \avo\ used in this paper is already available.
+Data from \da\ and \avo\ used in this paper is already available, the Rwanda and FTSE data cannot be made available.
 
@@
 This work was supported by a Google focussed research award; and the EPSRC [grant number EP/P505445/1].
+Thanks to Sherif Akoush and Ripduman Sohan for supplying the Rwanda data and to staff at the FTSE 100 company for supplying their mobile device management data.
 }

### 2015-06-04 ccs.tex
Update related work

@@
 \section{Related work}
-The update process for apps, security fixes and OS upgrades also needs to be secure.
+We assume updates make security better but, the update process for apps, security fixes and OS upgrades also needs to be secure.
 Unfortunately, package management systems designed to provide secure updates have been found to contain vulnerabilities~\cite{Cappos2008} and many software update systems fail to authenticate the connection between the device and the update server or do not authenticate the downloaded binaries~\cite{Bellissimo2006}.
@@
 In this paper we have analysed four critical vulnerabilities in the Android app update mechanism: APK unsigned shorts, APK unchecked name, APK duplicate file and Fake ID.
-Other work has demonstrate complex and subtle errors exist in the Android app update process.
+Other work has demonstrated complex and subtle errors exist in the Android app update process.
 For example, the process can be exploited to allow apps to gain privilege through `Pileup' vulnerabilities by registering for new permissions before the update which creates that permission is installed~\cite{Xing2014}.
 
+We used \da\ to record data on the OS versions of Android installed on devices, Nappa et.\ al.
+performed a similar analysis using WINE to record data on installed versions of Windows client programs and used cleaned NVD data rather than \avo\ for vulnerability data, they did not assign client applications scores~\cite{Nappa2015}.
+
 \interfootnotelinepenalty=10000 %To stop the lifehacker footnote below from spanning multiple pages and breaking
+We use User-Agent strings from Rwandan internet traffic to examine the version distribution of Android OS versions.
 User-Agent strings have been used to investigate the timeliness of web browser updates, with at most 80\% of Firefox users running the most recent version~\cite{Frei2008}.
@@
 It is possible that many updates arrive at handsets, but are simply not installed.
-Anecdotal evidence at least suggests that it is the lack of updates rather the lack of installation which is the major problem at present. Further work is required to tease these numbers apart.
+Anecdotal evidence at least suggests that it is the lack of updates rather the lack of installation which is the major problem at present and \S\ref{sec:security_scoring:results} supports this.
+Further work is required to tease these numbers apart.
 Partly this is the result of the fact that an operating system update is being installed and so a reboot is required, but Chrome installs the new version side by side with the old one and switches the next time it is restarted.
@@
 
-Security in depth is also a useful strategy.
+As well as supplying security updates promptly, the impact of vulnerabilities can be reduced through security in depth.
 In this regard, iOS provides additional safeguards beyond those used in Android, including a pre-distribution review, mandatory code-signing by the manufacturer, and (with the important exception of ROP-based attacks~\cite{Wang2013a}) the technical prohibition of dynamic code loading by an app.
@@
 Droid\-Ranger also analysed apps finding 148 out of 182\,823 apps (0.08\%) to be malicious across multiple markets of which 29 were previously unknown~\cite{Zhou2012a}.
-%DroidRanger: It used permission-based behavioural fingerprinting which looked at the permissions of known malware and heuristic-based filtering -- dynamic loading of both Dalvik and native code.
 A common technique used by attackers is to include malicious code in repackaged popular apps. 
 An\-Darwin uses this insight to detect similar apps, and found 169 out of 265\,359 of all apps studied (0.06\%) were malicious clones~\cite{Crussell2013}.
-%AnDarwin: It used clustering based on semantic vectors derived from the program dependence graphs to detect similar apps.
+Other approaches have made further use of malware dependency graphs~\cite{Lindorfer2014} or tried to extract further semantic information to avoid malware being able to avoid detection~\cite{Zhang2014} but simpler approaches based on extracting strings from the binaries can be more effective~\cite{Arp2014} in terms of false positives and false negatives.
+There is a lack of ground truth data, a danger in training algorithms based on APK files which other algorithms have already found which might lead to new malware being missed and DroidChameleon showed that existing AntiVirus apps could be fooled by simple permutations of malicious APKs~\cite{Rastogi2013}.
 
@@
 More recent data demonstrates the difficulty of upgrading computers between major OS versions, with 27\% of Windows computers running Windows XP in July 2014,\footnote{\url{https://archive.today/PLGxn}} four months after Windows XP stopped receiving security updates.
-
-\todo{Work these references into the text}
-\cite{Nappa2015}
-\cite{Zhang2014}
-\cite{Lindorfer2014}
-\cite{Arp2014}
 

### 2015-06-04 ccs.tex
Updated conclusion

@@
 \section{Conclusion}
-\todo{Write this}
- and examined how the Android ecosystem results in different device models receiving different levels of security due to whether or not they get security updates
-We have compared different device models, device manufacturers and network operators and found that there are differences between them which the discerning purchaser might use to influence their decision about which device model to buy from which device manufacturer and network operator.
+We examined how the Android ecosystem results in different device models receiving different levels of security due to whether or not they get security updates
+We have defined the FUM score which evaluates the security performance of companies and organisations in shipping security updates.
+We have use this to compare different device models, device manufacturers and network operators and found that there are differences between them which the discerning purchaser might use to influence their decision about which device model to buy from which device manufacturer and network operator.
 We hope that this analysis will encourage device manufacturers and network operators to improve the support they provide for devices after sale.

### 2015-06-05 security_scoring.tex
Fix some todos
Update full version comp diagram and associated text

@@
 We can then test whether this seems fair by comparing the version data for the highest and lowest scoring models.
-Figure~\ref{fig:full_version_comp}b shows the full version distribution for \emph{\daSecScoreWorstmodel} and Figure~\ref{fig:full_version_comp}a shows the same for \emph{\daSecScoreBestmodel}.
-There is a marked difference in the distribution of updates between the two with only \daSecScoreWorstmodelNumFullVersions\ full version for \emph{\daSecScoreWorstmodel} and \daSecScoreBestmodelNumFullVersions\ for \emph{\daSecScoreBestmodel}.
-%Both device models start off with the full version string of `2.3.3 GRI40' but the \emph{\daSecScoreBestmodel} receives many more updates over the same time period.
-%Other models from the same manufacturer with similar model names to \emph{\daSecScoreWorstmodel} do much better such as the \emph{Desire HD}.
+Figure~\ref{fig:full_version_comp}c shows the full version distribution for \emph{\daSecScoreWorstmodel}, Figure~\ref{fig:full_version_comp}b shows the full version distribution for \emph{HTC Desire HD A9191} which used to be our worst model and for which we have more historical data and Figure~\ref{fig:full_version_comp}a shows the same for \emph{\daSecScoreBestmodel}.
+There is a marked difference in the distribution of updates between the best and worst with only \daSecScoreWorstmodelNumFullVersions\ full version for \emph{\daSecScoreWorstmodel}, 2 full versions for \emph{HTC Desire HD A9191} and \daSecScoreBestmodelNumFullVersions\ for \emph{\daSecScoreBestmodel}.
+Both \emph{Galaxy Nexus} and \emph{HTC Desire HD A9191} device models start off with the full version string of `2.3.3 GRI40' but the \emph{\daSecScoreBestmodel} receives many more updates over the same time period.
+Other models from the same manufacturer with similar model names to \emph{HTC Desire HD A9191} do much better such as the \emph{Desire HD}.
 
@@
  \includegraphics[width=0.9\textwidth]{figures/full_version_comp.pdf}
- \caption{Full version distributions for the highest and lowest scoring model\todo{this is wrong now}}
+ \caption{Full version distributions for the highest and lowest scoring model}
  \label{fig:full_version_comp}
@@
 \includegraphics[width=\columnwidth]{figures/security_score_summary}
-\caption{Security scores for Nexus and non-Nexus devices\todo{subfig? bridge gaps?}}
+\caption{Security scores for Nexus and non-Nexus devices\todo{subfig?}\todolater{bridge gaps?}}
 \label{fig:security_score_summary}
@@
 Changing the scoring metric also impacts the scores given for each entity Table~\ref{tab:change_in_scores} shows the mean impact on the scores.
-\todo{Showing?}
+This shows that $m$ tends to drag down scores.
 \todo{Give some of the actual orderings as well for the most important ones}

### 2015-06-05 ccs.tex
Fix some todos
Adjust display hacks, less sloppy, more interfootnotelinepenalty

@@
 \makeatother
+
+% To stop problems with links in footnotes spanning columns or pages.
+\interfootnotelinepenalty=10000
 
@@
 The general pattern this shows is that in the FTSE data newer versions are more popular than in the \da\ data and that in the Rwanda data old versions are more popular than in the \da\ data.
-\todo{Can we quantify this?}
+\todolater{Can we quantify this?}
 
@@
 
-\interfootnotelinepenalty=10000 %To stop the lifehacker footnote below from spanning multiple pages and breaking
 We use User-Agent strings from Rwandan internet traffic to examine the version distribution of Android OS versions.
@@
 
-\sloppy
 \section*{Acknowledgements}
@@
 }
-
+\sloppy
 \printbibliography

### 2015-06-08 dastats.tex
./update.sh

@@
 \newcommand{\daStartDate}{2011-07-01}
-\newcommand{\daEndDate}{2015-05-27}
+\newcommand{\daEndDate}{2015-05-31}
 \newcommand{\daFullDeployedAt}{95\%}
@@
 \newcommand{\daVulnFree}{$0.121 \pm 0.000$}
-\newcommand{\daMeanOutstandingVulnerabilities}{$0.549 \pm 2.46e$}
+\newcommand{\daMeanOutstandingVulnerabilities}{$0.547 \pm 2.45e$}
 \newcommand{\daUpdatednessNominal}{0.0511}
@@
 \newcommand{\daUpdatednessPercTwosf}{$5.1 \pm 0.0\%$}
-\newcommand{\daSecurityScoreNominal}{2.83}
-\newcommand{\daSecurityScore}{$2.83 \pm 0.00$}
-\newcommand{\daCurrentMeanInsecurityPercNominal}{85.9\%}
-\newcommand{\daCurrentMeanInsecurityPerc}{$85.9 \pm 0.1\%$}
-\newcommand{\daCurrentMeanInsecurityPercTwosfNominal}{86\%}
-\newcommand{\daCurrentMeanInsecurityPercTwosf}{$86 \pm 0\%$}
-\newcommand{\daCurrentMeanSecurityPercNominal}{14.1\%}
-\newcommand{\daCurrentMeanSecurityPerc}{$14.1 \pm 0.1\%$}
-\newcommand{\daCurrentMeanSecurityPercTwosfNominal}{14\%}
-\newcommand{\daCurrentMeanSecurityPercTwosf}{$14 \pm 0\%$}
-\newcommand{\daCurrentVulnFreeNominal}{0.141}
-\newcommand{\daCurrentVulnFree}{$0.141 \pm 0.001$}
-\newcommand{\daCurrentMeanOutstandingVulnerabilities}{$0.201 \pm 1.16e$}
-\newcommand{\daCurrentUpdatednessNominal}{0.0322}
-\newcommand{\daCurrentUpdatedness}{$0.0322 \pm 0.0007$}
-\newcommand{\daCurrentUpdatednessPercNominal}{3.22\%}
-\newcommand{\daCurrentUpdatednessPerc}{$3.22 \pm 0.07\%$}
-\newcommand{\daCurrentUpdatednessPercTwosfNominal}{3.2\%}
-\newcommand{\daCurrentUpdatednessPercTwosf}{$3.2 \pm 0.0\%$}
-\newcommand{\daCurrentSecurityScoreNominal}{3.36}
-\newcommand{\daCurrentSecurityScore}{$3.36 \pm 0.00$}
+\newcommand{\daSecurityScoreNominal}{2.84}
+\newcommand{\daSecurityScore}{$2.84 \pm 0.00$}
+\newcommand{\daCurrentMeanInsecurityPercNominal}{85.4\%}
+\newcommand{\daCurrentMeanInsecurityPerc}{$85.4 \pm 0.1\%$}
+\newcommand{\daCurrentMeanInsecurityPercTwosfNominal}{85\%}
+\newcommand{\daCurrentMeanInsecurityPercTwosf}{$85 \pm 0\%$}
+\newcommand{\daCurrentMeanSecurityPercNominal}{14.6\%}
+\newcommand{\daCurrentMeanSecurityPerc}{$14.6 \pm 0.1\%$}
+\newcommand{\daCurrentMeanSecurityPercTwosfNominal}{15\%}
+\newcommand{\daCurrentMeanSecurityPercTwosf}{$15 \pm 0\%$}
+\newcommand{\daCurrentVulnFreeNominal}{0.146}
+\newcommand{\daCurrentVulnFree}{$0.146 \pm 0.001$}
+\newcommand{\daCurrentMeanOutstandingVulnerabilities}{$0.193 \pm 1.11e$}
+\newcommand{\daCurrentUpdatednessNominal}{0.0326}
+\newcommand{\daCurrentUpdatedness}{$0.0326 \pm 0.0006$}
+\newcommand{\daCurrentUpdatednessPercNominal}{3.26\%}
+\newcommand{\daCurrentUpdatednessPerc}{$3.26 \pm 0.06\%$}
+\newcommand{\daCurrentUpdatednessPercTwosfNominal}{3.3\%}
+\newcommand{\daCurrentUpdatednessPercTwosf}{$3.3 \pm 0.0\%$}
+\newcommand{\daCurrentSecurityScoreNominal}{3.39}
+\newcommand{\daCurrentSecurityScore}{$3.39 \pm 0.00$}
 \newcommand{\daDailyParticipationNominal}{903}
-\newcommand{\daDailyParticipation}{$903 \pm 229$}
+\newcommand{\daDailyParticipation}{$903 \pm 227$}
 \newcommand{\daNumFullVersions}{1\,250}
@@
 \newcommand{\daNumAPIVersions}{16}
-\newcommand{\daNumFullVersionUpdates}{4\,530}
-\newcommand{\daNumOSVersionUpdates}{3\,840}
+\newcommand{\daNumFullVersionUpdates}{4\,540}
+\newcommand{\daNumOSVersionUpdates}{3\,850}
 \newcommand{\daUpdatesPerYearNominal}{1.28}
@@
 \newcommand{\daOSCurveFitParamFirst}{84.8}
-\newcommand{\daOSCurveFitParamSecond}{\num{0.0029}}
+\newcommand{\daOSCurveFitParamSecond}{\num{0.00289}}
 \newcommand{\daOSCurveFitRMSE}{0.116}
@@
 \newcommand{\daOSCurveHalfDeployed}{$324 \pm 90$}
-\newcommand{\daOSCurveFullDeployedYearsNominal}{3.06}
-\newcommand{\daOSCurveFullDeployedYears}{$3.06 \pm 70$}
+\newcommand{\daOSCurveFullDeployedYearsNominal}{3.07}
+\newcommand{\daOSCurveFullDeployedYears}{$3.07 \pm 70$}
 \newcommand{\daOSCurveFullDeployedNominal}{1\,120}
@@
 \newcommand{\daAPICurveFitParamFirst}{102}
-\newcommand{\daAPICurveFitParamSecond}{\num{0.0029}}
+\newcommand{\daAPICurveFitParamSecond}{\num{0.00289}}
 \newcommand{\daAPICurveFitRMSE}{0.113}
@@
 \newcommand{\daVulnZergRushMonthsDefFixDeployed}{27.3}
-\newcommand{\daTabSecScoresmanufacturer}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline LG & $0.20 \pm 0.00$ & $0.32 \pm 0.00$ & $0.64 \pm 0.01$ & $3.83 \pm 0.02$ \\  Motorola & $0.17 \pm 0.00$ & $0.12 \pm 0.00$ & $0.73 \pm 0.02$ & $3.03 \pm 0.02$ \\  Samsung & $0.13 \pm 0.00$ & $0.04 \pm 0.00$ & $0.63 \pm 0.00$ & $2.72 \pm 0.00$ \\  Sony & $0.13 \pm 0.00$ & $0.19 \pm 0.00$ & $1.09 \pm 0.02$ & $2.61 \pm 0.02$ \\  HTC & $0.14 \pm 0.00$ & $0.09 \pm 0.00$ & $0.89 \pm 0.01$ & $2.58 \pm 0.02$ \\  asus & $0.19 \pm 0.00$ & $0.52 \pm 0.01$ & $6.01 \pm 0.07$ & $2.31 \pm 0.02$ \\  other & $0.06 \pm 0.00$ & $0.04 \pm 0.00$ & $1.04 \pm 0.01$ & $1.94 \pm 0.02$ \\  Symphony & $0.00 \pm 0.00$ & $0.07 \pm 0.00$ & $5.00 \pm 0.04$ & $0.27 \pm 0.01$ \\  walton & $0.00 \pm 0.00$ & $0.08 \pm 0.00$ & $5.96 \pm 0.08$ & $0.25 \pm 0.01$ \\ \end{tabular} \caption{Security scores for manufacturers} \label{tab:sec_manufacturer} \end{table*}}
+\newcommand{\daTabSecScoresmanufacturer}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline LG & $0.20 \pm 0.00$ & $0.33 \pm 0.00$ & $0.63 \pm 0.01$ & $3.86 \pm 0.02$ \\  Motorola & $0.18 \pm 0.00$ & $0.13 \pm 0.00$ & $0.73 \pm 0.02$ & $3.03 \pm 0.02$ \\  Samsung & $0.13 \pm 0.00$ & $0.04 \pm 0.00$ & $0.63 \pm 0.00$ & $2.72 \pm 0.00$ \\  Sony & $0.13 \pm 0.00$ & $0.19 \pm 0.00$ & $1.09 \pm 0.02$ & $2.61 \pm 0.02$ \\  HTC & $0.14 \pm 0.00$ & $0.09 \pm 0.00$ & $0.89 \pm 0.01$ & $2.58 \pm 0.02$ \\  asus & $0.19 \pm 0.00$ & $0.52 \pm 0.01$ & $6.01 \pm 0.07$ & $2.32 \pm 0.02$ \\  other & $0.06 \pm 0.00$ & $0.04 \pm 0.00$ & $1.03 \pm 0.01$ & $1.95 \pm 0.02$ \\  Symphony & $0.00 \pm 0.00$ & $0.08 \pm 0.00$ & $5.00 \pm 0.05$ & $0.28 \pm 0.01$ \\  walton & $0.00 \pm 0.00$ & $0.08 \pm 0.00$ & $5.96 \pm 0.08$ & $0.25 \pm 0.01$ \\ \end{tabular} \caption{Security scores for manufacturers} \label{tab:sec_manufacturer} \end{table*}}
 \newcommand{\daSecScoreBestmanufacturer}{LG}
-\newcommand{\daSecScoreBestmanufacturerScoreNominal}{3.83}
-\newcommand{\daSecScoreBestmanufacturerScore}{$3.83 \pm 0.02$}
+\newcommand{\daSecScoreBestmanufacturerScoreNominal}{3.86}
+\newcommand{\daSecScoreBestmanufacturerScore}{$3.86 \pm 0.02$}
 \newcommand{\daSecScoreBestmanufacturerNumFullVersions}{131}
@@
 \newcommand{\daSecScoreWorstmanufacturerNumFullVersions}{22}
-\newcommand{\daTabSecScoresmodel}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline Galaxy Nexus & $0.51 \pm 0.00$ & $0.55 \pm 0.01$ & $1.53 \pm 0.04$ & $4.74 \pm 0.04$ \\  Nexus 4 & $0.28 \pm 0.00$ & $0.82 \pm 0.01$ & $6.06 \pm 0.09$ & $3.59 \pm 0.04$ \\  Nexus 7 & $0.24 \pm 0.00$ & $0.74 \pm 0.01$ & $5.92 \pm 0.09$ & $3.18 \pm 0.04$ \\  other & $0.09 \pm 0.00$ & $0.14 \pm 0.00$ & $0.55 \pm 0.00$ & $2.98 \pm 0.00$ \\  Desire HD & $0.08 \pm 0.00$ & $0.05 \pm 0.00$ & $0.39 \pm 0.02$ & $2.91 \pm 0.04$ \\  HTC Sensation & $0.35 \pm 0.00$ & $0.01 \pm 0.01$ & $1.59 \pm 0.05$ & $2.42 \pm 0.05$ \\  GT-I9100 & $0.22 \pm 0.00$ & $0.02 \pm 0.00$ & $1.23 \pm 0.02$ & $2.28 \pm 0.02$ \\  HTC Desire S & $0.02 \pm 0.00$ & $0.02 \pm 0.00$ & $1.00 \pm 0.06$ & $1.75 \pm 0.07$ \\  GT-N7000 & $0.25 \pm 0.00$ & $0.00 \pm 0.00$ & $2.52 \pm 0.05$ & $1.43 \pm 0.02$ \\  GT-P1000 & $0.01 \pm 0.00$ & $0.00 \pm 0.01$ & $1.79 \pm 0.06$ & $0.90 \pm 0.05$ \\  GT-I9300 & $0.13 \pm 0.00$ & $0.01 \pm 0.00$ & $6.25 \pm 0.04$ & $0.58 \pm 0.01$ \\  GT-I9505 & $0.01 \pm 0.00$ & $0.13 \pm 0.00$ & $6.82 \pm 0.07$ & $0.44 \pm 0.01$ \\  HTC Desire HD & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & $3.05 \pm 0.05$ & $0.27 \pm 0.03$ \\  GT-N7100 & $0.06 \pm 0.00$ & $0.00 \pm 0.01$ & $6.91 \pm 0.08$ & $0.25 \pm 0.02$ \\  Symphony W68 & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & $11.00 \pm 0.12$ & $0.00 \pm 0.03$ \\ \end{tabular} \caption{Security scores for models} \label{tab:sec_model} \end{table*}}
+\newcommand{\daTabSecScoresmodel}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline Galaxy Nexus & $0.51 \pm 0.00$ & $0.55 \pm 0.01$ & $1.53 \pm 0.04$ & $4.74 \pm 0.04$ \\  Nexus 4 & $0.28 \pm 0.00$ & $0.82 \pm 0.01$ & $6.06 \pm 0.09$ & $3.61 \pm 0.04$ \\  Nexus 7 & $0.24 \pm 0.00$ & $0.74 \pm 0.01$ & $5.92 \pm 0.09$ & $3.19 \pm 0.04$ \\  other & $0.09 \pm 0.00$ & $0.14 \pm 0.00$ & $0.55 \pm 0.00$ & $2.98 \pm 0.00$ \\  Desire HD & $0.08 \pm 0.00$ & $0.05 \pm 0.00$ & $0.39 \pm 0.02$ & $2.91 \pm 0.04$ \\  HTC Sensation & $0.34 \pm 0.00$ & $0.01 \pm 0.01$ & $1.59 \pm 0.05$ & $2.41 \pm 0.05$ \\  GT-I9100 & $0.22 \pm 0.00$ & $0.02 \pm 0.00$ & $1.23 \pm 0.02$ & $2.28 \pm 0.02$ \\  HTC Desire S & $0.02 \pm 0.00$ & $0.02 \pm 0.00$ & $1.00 \pm 0.06$ & $1.75 \pm 0.07$ \\  GT-N7000 & $0.25 \pm 0.00$ & $0.00 \pm 0.00$ & $2.52 \pm 0.05$ & $1.43 \pm 0.02$ \\  GT-P1000 & $0.01 \pm 0.00$ & $0.00 \pm 0.01$ & $1.79 \pm 0.06$ & $0.90 \pm 0.05$ \\  GT-I9300 & $0.13 \pm 0.00$ & $0.01 \pm 0.00$ & $6.25 \pm 0.04$ & $0.58 \pm 0.01$ \\  GT-I9505 & $0.01 \pm 0.00$ & $0.13 \pm 0.00$ & $6.83 \pm 0.07$ & $0.46 \pm 0.01$ \\  HTC Desire HD & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & $3.05 \pm 0.05$ & $0.27 \pm 0.03$ \\  GT-N7100 & $0.06 \pm 0.00$ & $0.00 \pm 0.01$ & $6.91 \pm 0.08$ & $0.25 \pm 0.02$ \\  Symphony W68 & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & $11.00 \pm 0.12$ & $0.00 \pm 0.03$ \\ \end{tabular} \caption{Security scores for models} \label{tab:sec_model} \end{table*}}
 \newcommand{\daSecScoreBestmodel}{Galaxy Nexus}
@@
 \newcommand{\daSecScoreWorstmodelScoreNominal}{0.0001}
-\newcommand{\daSecScoreWorstmodelScore}{$0.0001 \pm 0.0276$}
+\newcommand{\daSecScoreWorstmodelScore}{$0.0001 \pm 0.0275$}
 \newcommand{\daSecScoreWorstmodelNumFullVersions}{1}
-\newcommand{\daTabSecScoressummary}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline nexus & $0.37 \pm 0.00$ & $0.47 \pm 0.00$ & $0.59 \pm 0.01$ & $5.06 \pm 0.02$ \\  notnexus & $0.10 \pm 0.00$ & $0.02 \pm 0.00$ & $0.55 \pm 0.00$ & $2.66 \pm 0.00$ \\ \end{tabular} \caption{Security scores for nexus} \label{tab:sec_summary} \end{table*}}
-\newcommand{\daTabSecScoresoperator}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline O2 uk & $0.26 \pm 0.00$ & $0.12 \pm 0.00$ & $0.37 \pm 0.02$ & $3.84 \pm 0.03$ \\  T-Mobile & $0.21 \pm 0.00$ & $0.18 \pm 0.00$ & $0.42 \pm 0.02$ & $3.73 \pm 0.03$ \\  Orange & $0.21 \pm 0.00$ & $0.09 \pm 0.00$ & $0.39 \pm 0.03$ & $3.52 \pm 0.04$ \\  Sprint & $0.18 \pm 0.00$ & $0.11 \pm 0.00$ & $0.40 \pm 0.02$ & $3.46 \pm 0.03$ \\  3 & $0.18 \pm 0.00$ & $0.09 \pm 0.00$ & $0.52 \pm 0.03$ & $3.23 \pm 0.04$ \\  Vodafone uk & $0.13 \pm 0.00$ & $0.13 \pm 0.00$ & $0.50 \pm 0.03$ & $3.18 \pm 0.04$ \\  AT\&T & $0.13 \pm 0.00$ & $0.08 \pm 0.00$ & $0.43 \pm 0.02$ & $3.09 \pm 0.02$ \\  Verizon & $0.18 \pm 0.00$ & $0.09 \pm 0.00$ & $0.82 \pm 0.02$ & $2.83 \pm 0.02$ \\  unknown & $0.11 \pm 0.00$ & $0.20 \pm 0.00$ & $0.87 \pm 0.02$ & $2.80 \pm 0.02$ \\  n Telenor & $0.03 \pm 0.00$ & $0.12 \pm 0.00$ & $1.25 \pm 0.02$ & $1.81 \pm 0.02$ \\  Airtel & $0.05 \pm 0.00$ & $0.03 \pm 0.00$ & $1.62 \pm 0.03$ & $1.28 \pm 0.03$ \\  Robi & $0.00 \pm 0.00$ & $0.08 \pm 0.00$ & $2.08 \pm 0.04$ & $0.90 \pm 0.03$ \\  Grameenphone & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & $2.03 \pm 0.02$ & $0.85 \pm 0.01$ \\  banglalink & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & $2.61 \pm 0.04$ & $0.54 \pm 0.02$ \\ \end{tabular} \caption{Security scores for operators} \label{tab:sec_operator} \end{table*}}
+\newcommand{\daTabSecScoressummary}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline nexus & $0.38 \pm 0.00$ & $0.47 \pm 0.00$ & $0.58 \pm 0.01$ & $5.07 \pm 0.02$ \\  notnexus & $0.10 \pm 0.00$ & $0.02 \pm 0.00$ & $0.55 \pm 0.00$ & $2.67 \pm 0.00$ \\ \end{tabular} \caption{Security scores for nexus} \label{tab:sec_summary} \end{table*}}
+\newcommand{\daTabSecScoresoperator}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline O2 uk & $0.26 \pm 0.00$ & $0.12 \pm 0.00$ & $0.37 \pm 0.02$ & $3.85 \pm 0.03$ \\  T-Mobile & $0.21 \pm 0.00$ & $0.18 \pm 0.00$ & $0.42 \pm 0.02$ & $3.73 \pm 0.03$ \\  Orange & $0.21 \pm 0.00$ & $0.09 \pm 0.00$ & $0.39 \pm 0.03$ & $3.52 \pm 0.04$ \\  Sprint & $0.18 \pm 0.00$ & $0.11 \pm 0.00$ & $0.40 \pm 0.02$ & $3.46 \pm 0.03$ \\  3 & $0.18 \pm 0.00$ & $0.09 \pm 0.00$ & $0.52 \pm 0.03$ & $3.24 \pm 0.04$ \\  Vodafone uk & $0.13 \pm 0.00$ & $0.13 \pm 0.00$ & $0.50 \pm 0.03$ & $3.18 \pm 0.04$ \\  AT\&T & $0.13 \pm 0.00$ & $0.08 \pm 0.00$ & $0.43 \pm 0.02$ & $3.10 \pm 0.02$ \\  Verizon & $0.18 \pm 0.00$ & $0.09 \pm 0.00$ & $0.82 \pm 0.02$ & $2.83 \pm 0.02$ \\  unknown & $0.11 \pm 0.00$ & $0.20 \pm 0.00$ & $0.86 \pm 0.02$ & $2.81 \pm 0.02$ \\  n Telenor & $0.03 \pm 0.00$ & $0.12 \pm 0.00$ & $1.25 \pm 0.02$ & $1.82 \pm 0.02$ \\  Airtel & $0.05 \pm 0.00$ & $0.03 \pm 0.00$ & $1.62 \pm 0.03$ & $1.28 \pm 0.03$ \\  Robi & $0.00 \pm 0.00$ & $0.08 \pm 0.00$ & $2.08 \pm 0.04$ & $0.90 \pm 0.03$ \\  Grameenphone & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & $2.02 \pm 0.02$ & $0.85 \pm 0.01$ \\  banglalink & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & $2.60 \pm 0.04$ & $0.54 \pm 0.02$ \\ \end{tabular} \caption{Security scores for operators} \label{tab:sec_operator} \end{table*}}
 \newcommand{\daSecScoreBestoperator}{O2 uk}
-\newcommand{\daSecScoreBestoperatorScoreNominal}{3.84}
-\newcommand{\daSecScoreBestoperatorScore}{$3.84 \pm 0.02$}
-\newcommand{\daSecScoreBestoperatorNumFullVersions}{97}
+\newcommand{\daSecScoreBestoperatorScoreNominal}{3.85}
+\newcommand{\daSecScoreBestoperatorScore}{$3.85 \pm 0.02$}
+\newcommand{\daSecScoreBestoperatorNumFullVersions}{96}
 \newcommand{\daSecScoreWorstoperator}{banglalink}
-\newcommand{\daSecScoreWorstoperatorScoreNominal}{0.535}
-\newcommand{\daSecScoreWorstoperatorScore}{$0.535 \pm 0.017$}
-\newcommand{\daSecScoreWorstoperatorNumFullVersions}{71}
-\newcommand{\daTabDLDistances}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & summary \\ \hline equal & 0.0 & 0.0 & 0.0714 & 0.0 \\weight $m$ & 0.222 & 0.333 & 0.143 & 0.0 \\$f$ & 0.444 & 0.733 & 0.429 & 0.0 \\weight $u$ & 0.444 & 0.133 & 0.214 & 0.0 \\$m$ & 0.667 & 0.733 & 0.357 & 0.5 \\$u$ & 0.778 & 0.733 & 0.714 & 0.0 \\\end{tabular} \caption{Normalised Damerau-Levenshtein distances for different metrics} \label{tab:dl_distances} \end{table}}
-\newcommand{\daTabChangeInScores}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & summary \\ \hline $m$ & $-1.9 \pm 2.2$ & $-0.666 \pm 2.51$ & $-3.27 \pm 1.39$ & $-3.37 \pm 1.83$ \\weight $m$ & $-0.295 \pm 0.274$ & $-0.102 \pm 0.316$ & $-0.458 \pm 0.177$ & $-0.487 \pm 0.209$ \\equal & $-0.116 \pm 0.065$ & $-0.039 \pm 0.113$ & $-0.146 \pm 0.049$ & $-0.166 \pm 0.028$ \\weight $u$ & $-0.0526 \pm 0.127$ & $-0.0152 \pm 0.239$ & $0.02 \pm 0.07$ & $-0.012 \pm 0.125$ \\$u$ & $0.516 \pm 1.61$ & $0.198 \pm 2.14$ & $1.51 \pm 0.96$ & $1.38 \pm 1.5$ \\$f$ & $1.04 \pm 0.58$ & $0.351 \pm 1.02$ & $1.31 \pm 0.44$ & $1.5 \pm 0.2$ \\\end{tabular} \caption{Mean change in scores for different metrics} \label{tab:change_in_scores} \end{table*}}
+\newcommand{\daSecScoreWorstoperatorScoreNominal}{0.545}
+\newcommand{\daSecScoreWorstoperatorScore}{$0.545 \pm 0.018$}
+\newcommand{\daSecScoreWorstoperatorNumFullVersions}{73}
+\newcommand{\daTabDLDistances}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & summary \\ \hline equal & 0.0 & 0.0 & 0.0714 & 0.0 \\weight $m$ & 0.222 & 0.333 & 0.143 & 0.0 \\weight $u$ & 0.444 & 0.133 & 0.214 & 0.0 \\$f$ & 0.444 & 0.733 & 0.5 & 0.0 \\$m$ & 0.667 & 0.733 & 0.357 & 0.5 \\$u$ & 0.778 & 0.733 & 0.714 & 0.0 \\\end{tabular} \caption{Normalised Damerau-Levenshtein distances for different metrics} \label{tab:dl_distances} \end{table}}
+\newcommand{\daTabChangeInScores}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & summary \\ \hline $m$ & $-1.9 \pm 2.2$ & $-0.664 \pm 2.51$ & $-3.27 \pm 1.39$ & $-3.38 \pm 1.83$ \\weight $m$ & $-0.294 \pm 0.275$ & $-0.101 \pm 0.316$ & $-0.458 \pm 0.177$ & $-0.487 \pm 0.209$ \\equal & $-0.116 \pm 0.064$ & $-0.0386 \pm 0.113$ & $-0.146 \pm 0.048$ & $-0.166 \pm 0.029$ \\weight $u$ & $-0.0522 \pm 0.126$ & $-0.0148 \pm 0.238$ & $0.0206 \pm 0.0797$ & $-0.00995 \pm 0.123$ \\$u$ & $0.517 \pm 1.61$ & $0.2 \pm 2.1$ & $1.52 \pm 0.96$ & $1.39 \pm 1.48$ \\$f$ & $1.04 \pm 0.58$ & $0.348 \pm 1.01$ & $1.31 \pm 0.43$ & $1.49 \pm 0.26$ \\\end{tabular} \caption{Mean change in scores for different metrics} \label{tab:change_in_scores} \end{table*}}
 \newcommand{\daGPAPICurveFitParamFirst}{80.7}
@@
 \newcommand{\daNumUpdatesSkippedBig}{3}
-\newcommand{\daNumUpdatesBigUpgrades}{819}
+\newcommand{\daNumUpdatesBigUpgrades}{820}
 \newcommand{\daNumUpdatesDowngrades}{152}
@@
 \newcommand{\daPercUpdatesDowngrades}{$3.45 \pm 0.28\%$}
-\newcommand{\daNumDataPoints}{157 Billion}
+\newcommand{\daNumDataPoints}{158 Billion}
 \newcommand{\daMonthsDevices}{1\,790}
@@
 \newcommand{\daGPAPISeventeenFullDeploymentMonth}{October 2017}
-\newcommand{\daMeanOutstandingVulnerabilitiesNominal}{0.549}
-\newcommand{\daCurrentMeanOutstandingVulnerabilitiesNominal}{0.201}
+\newcommand{\daMeanOutstandingVulnerabilitiesNominal}{0.547}
+\newcommand{\daCurrentMeanOutstandingVulnerabilitiesNominal}{0.193}
 \newcommand{\daModelHalfDeploymentYears}{0.0822}
@@
 \newcommand{\daModelFullDeployment}{325}
-\newcommand{\daMeanOutstandingProbability}{1\%}
+\newcommand{\daMeanOutstandingProbability}{1.0\%}
 \newcommand{\daMeanOutstandingUncertaintyEquation}{$0.99^n$}

### 2015-06-08 security_scoring.tex
update captions

@@
  \includegraphics[width=0.9\textwidth]{figures/full_version_comp.pdf}
- \caption{Full version distributions for the highest and lowest scoring model}
+ \caption{Full version distributions for the highest and lowest scoring models}
  \label{fig:full_version_comp}
@@
 \includegraphics[width=\columnwidth]{figures/security_score_summary}
-\caption{Security scores for Nexus and non-Nexus devices\todo{subfig?}\todolater{bridge gaps?}}
+\caption{Security scores for Nexus and non-Nexus devices\todolater{bridge gaps?}}
 \label{fig:security_score_summary}

### 2015-06-08 ccs.tex
Feedback from Alastair

@@
 Using a corpus of \daNumOSDevices\ devices we show that there is significant variability in the timely delivery of security updates across different device manufacturers and network operators.
+This is the first paper to quantify security update performance of device manufacturers and network operators.
 We define a security metric to rank the performance of device manufacturers and network operators.
@@
 It has been previously shown that Android devices do not receive security updates promptly when critical vulnerabilities are publicly disclosed and hence \daMeanInsecurityPercNominal\ are exposed to known critical vulnerabilities on average~\cite{androidvulnerabilities.org}.
-Here we attempt to determine why by understanding the Android update process, apportioning blame to different entities in the ecosystem and producing a scoring metric to compare different entities.
+Here we attempt to determine why by understanding the Android update process, evaluating different entities in the ecosystem and producing a scoring metric to compare different entities.
 
@@
 To address this, we developed a scoring system and provide numbers on the historic performance of device models in the \da\ project in~\S\ref{sec:security_scoring}.
-We provide the answer to the question ``Who should I buy my Android phone from for best security?'', showing that in general Nexus devices are better; further details are given in~\S\ref{sec:security_scoring:results}.
+We provide the answer to the question ``Which is the most secure Android provider I can buy from?'', showing that in general Nexus devices are better; further details are given in~\S\ref{sec:security_scoring:results}.
 
@@
 By using data from \da~\cite{Wagner2013} we can examine the versions of Android running on user devices over time and through comparison with data on which versions of Android were vulnerable to different critical vulnerabilities~\cite{androidvulnerabilities.org} we can plot the exposure of Android devices to known critical vulnerabilities.
-Figure~\ref{fig:proportioninsecure} shows that Android devices are often exposed to known security vulnerabilities.
+Figure~\ref{fig:proportioninsecure} uses data from \da\ and the \avo\ vulnerabilities database and shows that Android devices are often exposed to known security vulnerabilities.
 One solution to this problem is regulation, and indeed there is ongoing legal action to force network operators to ship updates for security vulnerabilities~\cite{Soghoian2013}.\todolater{Check on the status of this legal action}
@@
 An alternative to regulation is access to publicly available data.
-Comparative data, which we provide in this paper, would provide an incentive to both device manufacturers and network operators to provide updates.
+Comparative data, which we provide in this paper, and will continue to publish on our companion website, provides an incentive to both device manufacturers and network operators to provide updates.
 
@@
 \begin{itemize}
- \item We quantify the Android update process, providing concrete numbers on the flow of updates and their latency~\S\ref{sec:android_ecosystem}.
- \item We propose the FUM scoring metric to evaluate the security of different instances of a platform~\S\ref{sec:security_scoring:method}.
- \item We measure the security of Android according to this and compare different device manufacturers, device models and network operators to allow device purchasers to differentiate between them based on security~\S\ref{sec:security_scoring:results}.
+ \item We quantify the Android update process, providing concrete numbers on the flow of updates and their latency~(\S\ref{sec:android_ecosystem}).
+ \item We propose the FUM scoring metric to evaluate the security of different instances of a platform~(\S\ref{sec:security_scoring:method}).
+ \item We measure the security of Android against our scoring metric and compare different device manufacturers, device models and network operators to allow device purchasers to differentiate between them based on security~(\S\ref{sec:security_scoring:results}).
 \end{itemize}
@@
  Numbers on edges indicate updates shipped between \daStartDate\ and \daEndDate, those in brackets represent number of such entities.
- Dotted arrows indicate flows where we can't measure how many updates are produced as they are not public.}
+ Dotted arrows indicate flows where we can't measure because no public data is available.}
  \label{fig:update_ecosystem}
@@
 
-The first attack vector is through the installation of a malicious app on the device.
+The \emph{installation} attack vector is through the installation of a malicious app on the device.
 Android devices can install apps through marketplaces, email attachments, URLs and via the Android Debug Bridge (ADB).
-By default, many Android devices will only allow the installation of apps from Google play which uses Bouncer to automatically analyse apps; and quickly takes down apps that are reported as malicious.
+By default, many Android devices will only allow the installation of apps from Google Play which uses Bouncer to automatically analyse apps; and quickly takes down apps that are reported as malicious.
 However alternative markets are also popular, particularly in countries where Google Play is not available.
 
-The second attack vector is to make an existing app into download and execute code at runtime.
+The \emph{dynamic code loading} attack vector is to make an existing app into download and execute code at runtime.
 The most direct method is to upload to a marketplace a seemingly innocent app that dynamically loads malicious code.
@@
 
-The third attack vector is that the attacker injects malicious code directly into existing code already on the handset. 
-For example, the addJavascriptInterface vulnerability (CVE-2012-6636) allows JavaScript running in an Android WebView to execute arbitrary code as the vulnerable app's user~\cite{Thomas2015a}.
+The \emph{injection} attack vector is that the attacker injects malicious code directly into existing code already on the handset. 
+For example, the addJavascriptInterface vulnerability (CVE-2012-6636) allows JavaScript running in an Android WebView to execute arbitrary code as the vulnerable app's user.
 The fix for the addJavascriptInterface vulnerability breaks backwards compatibility and requires a two sided fix.
-While the fix was released in December 2012, on \daGPAPISeventeenLaterDate, \daGPAPISeventeenEarlierProportion\ of handsets connecting to Google Play were still vulnerable to this attack.
+While the fix was released in December 2012, on \daGPAPISeventeenLaterDate, \daGPAPISeventeenEarlierProportion\ of handsets connecting to Google Play were still vulnerable to this attack~\cite{Thomas2015a}.
 
 The best place to prevent attacks is at runtime, since all three attack vectors can be prevented at this level. 
-Unfortunately, it has been demonstrated that the sandbox for Android apps is weak~\cite{androidvulnerabilities.org}.
+Unfortunately, as Figure~\ref{fig:proportioninsecure} shows, the sandbox for Android apps is weak.
 
@@
 
-Some critical vulnerabilities are not traditional kernel vulnerabilities, but are vulnerabilities in level 1 of our threat model.
+Some critical vulnerabilities are not traditional kernel vulnerabilities, but are vulnerabilities in \emph{installation} in our threat model.
 For example improper verification of signatures at installation time was discovered in February 2013~\cite{Forristal2013} and meant that applications could pretend to be signed with system keys and hence be granted system privileges.
@@
 Regardless of version, this exposed an increased attack area and would also provide the ability for malware to control all user internet traffic (via VPNs), brick the phone, remove and install apps, steal user credentials and read the screen.
-The different categories which the vulnerabilities fall into are shown in Table~\ref{tab:andvulns}, `signature' vulnerabilities are level 1 while kernel and system vulnerabilities are level 2.\todo{What is going on with these levels?}
+The different categories which the vulnerabilities fall into are shown in Table~\ref{tab:andvulns}, `signature' vulnerabilities are \emph{installation} while kernel and system vulnerabilities are used with \emph{installation, dynamic code loading} and \emph{injection}.
 \avoTabAndVulns
@@
 These vulnerabilities affect all Android devices regardless of manufacturer, and as a result our selected vulnerabilities will dominate any security analysis of Android.
-\todo{can we make this sentence clearer?}
 Hence, with our chosen set of vulnerabilities, our analysis represents a lower-bound on the vulnerability of devices in the \da\ data.
@@
 \section{Comparison with other data}\label{sec:representative}
-Here we show that the \da\ data~\cite{Wagner2013} is representative for our purposes by comparing it with two other sources of data.
+Here we provide evidence that the \da\ data~\cite{Wagner2013} is representative for our purposes by comparing it with two other sources of data.
 The \da\ data is collected from devices across the world where users have voluntarily installed the app.
@@
 \end{figure}
-We used the data from the FTSE 100 company for the week ending 2015-04-27 and the User-Agent data was collected between 2015-01-31 and 2015-02-12.
+We used the data from the FTSE 100 company for a week in April 2015 and the User-Agent data was collected in February 2015. %between 2015-01-31 and 2015-02-12.% FTSE 2015-04-27
 Figure~\ref{fig:dists} shows the proportion of devices running each Android OS version in the two comparison data sets and the comparable periods from \da.
 The general pattern this shows is that in the FTSE data newer versions are more popular than in the \da\ data and that in the Rwanda data old versions are more popular than in the \da\ data.
+Therefore the \da\ data on OS versions is bounded by these two data sets.
 \todolater{Can we quantify this?}
@@
 While API versions are too coarse grained to use for security update detection they are closely related to OS versions and so if the \da\ data on API versions are similar to the Google data on API versions then the \da\ data on OS versions should be representative.
-We compared the data from Google and from \da\ and they appear similar.
+We compared the data from Google and from \da\ and they are similar.
 We analysed the difference between the API version data from \da\ and Google Play, normalising for days since the API version was released.
 This shows that the \da\ data systematically overestimates the prevalence of new API versions and underestimates the prevalence of old API versions.
-This means that the OS version information from \da\ is likely to be overestimating the prevalence of new OS versions and hence our results are conservative.
+This means that the OS version information from \da\ is likely to be overestimating the prevalence of new OS versions and hence our results are likely to be conservative.
 
 \section{Related work}
-We assume updates make security better but, the update process for apps, security fixes and OS upgrades also needs to be secure.
+We assume updates make security better, however the update process for apps, security fixes and OS upgrades also needs to be secure.
 Unfortunately, package management systems designed to provide secure updates have been found to contain vulnerabilities~\cite{Cappos2008} and many software update systems fail to authenticate the connection between the device and the update server or do not authenticate the downloaded binaries~\cite{Bellissimo2006}.
@@
 
-We used \da\ to record data on the OS versions of Android installed on devices, Nappa et.\ al.
+We used \da\ to record data on the OS versions of Android installed on devices.
+Nappa et.\ al.
 performed a similar analysis using WINE to record data on installed versions of Windows client programs and used cleaned NVD data rather than \avo\ for vulnerability data, they did not assign client applications scores~\cite{Nappa2015}.
@@
 The raw and processed data and source code will be made available where possible when this paper is finalised.
-Data from \da\ and \avo\ used in this paper is already available, the Rwanda and FTSE data cannot be made available.
+Data from \da\ and \avo\ used in this paper is already available.
+The Rwanda and FTSE data cannot be made available due to the terms under which they were obtained.
 

### 2015-06-08 security_scoring.tex
Feedback from Alastair

@@
 To allow buyers of Android devices to purchase those devices which have the best security they need to know how different device manufacturers, device models and network operators compare in terms of the security they provide.
-We propose a method to score a device manufacturer, device model or network operator based on its historic performance at keeping devices up to date and fixing security vulnerabilities.
+We propose a method to score a device manufacturer, device model or network operator based on its historic performance at keeping devices up-to-date and fixing security vulnerabilities.
 We find that Android as a whole gets a score of \daSecurityScore\ out of 10, the highest scoring device manufacturer is \emph{\daSecScoreBestmanufacturer} (\daSecScoreBestmanufacturerScore\ out of 10) and the lowest scoring is \emph{\daSecScoreWorstmanufacturer} (\daSecScoreWorstmanufacturerScore\ out of 10).
@@
 %\footnote{A perfectly secure operating system would among other things detect and prevent all phishing attacks, provide perfect principle of least privilege isolation, not have any vulnerabilities, instantly fix all discovered vulnerabilities, not allow any user data to be used in a way which the user does not approve of and be really easy for an ordinary person to use.}
-However in the absence of such data we propose a scheme for assigning a device a score out of ten based on data which can be observed, and which hopefully correlates with the actual security of the devices.
+However in the absence of such data we propose a scheme for assigning a device a score out of ten based on data which can be observed, is based on previous metrics, and which we expect correlates with the actual security of the devices.
 
@@
 This is captured by the $m$ score which measures the size of the device manufacturers queue of outstanding vulnerabilities but does not take into account the update process or measure the actual end user security.
-Neither of these metrics capture whether devices are being left behind and not being kept up to date with the most recent (and hopefully most secure) version which is captured by $u$.
+Neither of these metrics capture whether devices are being left behind and not being kept up-to-date with the most recent (and hopefully most secure) version which is captured by $u$.
 
@@
 There are on average \daMeanOutstandingVulnerabilities\ outstanding vulnerabilities not fixed on any device.
-Only on average \daUpdatednessPerc\ of devices run the most recent version of Android.
+On average \daUpdatednessPerc\ of devices run the most recent version of Android.
 This gives a security score of \daSecurityScore\ out of 10.
@@
 However there are a wide variety of scores depending on the source of the device.
-There have been many reports\todolater{citation} that Google's Nexus devices are better at getting updates than other Android devices because Google makes the original updates and ships them to its devices.
+There have is much anecdotal evidence that Google's Nexus devices are better at getting updates than other Android devices because Google makes the original updates and ships them to its devices (but the plural of anecdote is not data).
 Table~\ref{tab:sec_summary} shows that this is the case with Nexus devices getting much better scores than non-Nexus devices and Figure~\ref{fig:security_score_summary} shows this holds true over time.
@@
 \daTabSecScoresmodel
-Even within device manufacturers different models can have very different update behaviours and hence security.
+Even within device manufacturers, different models can have very different update behaviours and hence security.
 Table~\ref{tab:sec_model} shows the results for the \daNumSigModels\ device models which have a significant presence by the same metric with \emph{\daSecScoreBestmodel} (\daSecScoreBestmodelScore\ out of 10) scoring highest and \emph{\daSecScoreWorstmodel} (\daSecScoreWorstmodelScore\ out of 10) scoring lowest.
@@
 \daTabSecScoresoperator
-We also analysed different network operators, for the \daNumSigOperators\ network operators with a significant presence in our data Table~\ref{tab:sec_operator} shows the results \emph{\daSecScoreBestoperator} (\daSecScoreBestoperatorScore\ out of 10) scoring highest and \emph{\daSecScoreWorstoperator} (\daSecScoreWorstoperatorScore\ out of 10) scoring lowest.
-However the score of an network operator is affected by the device manufacturers of the devices which are in use on its network.
+We also analysed different network operators, for the \daNumSigOperators\ network operators with a significant presence in our data.
+Table~\ref{tab:sec_operator} shows the results \emph{\daSecScoreBestoperator} (\daSecScoreBestoperatorScore\ out of 10) scoring highest and \emph{\daSecScoreWorstoperator} (\daSecScoreWorstoperatorScore\ out of 10) scoring lowest.
+However the score of a network operator is affected by the device manufacturers of the devices which are in use on its network.
 This is in turn affected by both what devices a network operator offers to users and upon which devices users choose.
@@
 However our sample size is too small to do that as while we have significant numbers of devices for each of the device models (Table~\ref{tab:sec_model}) and for each of the network operators (Table~\ref{tab:sec_operator}) we would need a significant number of each model in each network operator.
-We do not attempt to disambiguate users behaviour in whether they install updates from network operators using rolling upgrades.
+%We do not attempt to disambiguate users behaviour in whether they install updates from network operators using rolling upgrades.
 
@@
 \centering
+\begin{subfigure}{\columnwidth}
 \includegraphics[width=\columnwidth]{figures/security_score_manufacturer}
-\caption{Security scores for different manufacturers}
+\caption{Device manufacturers}
 \label{fig:security_score_manufacturer}
+\end{subfigure}
+%
+\begin{subfigure}{\columnwidth}
 \includegraphics[width=\columnwidth]{figures/security_score_model}
-\caption{Security scores for different device models}
+\caption{Device models}
 \label{fig:security_score_model}
+\end{subfigure}
+%
+\begin{subfigure}{\columnwidth}
 \includegraphics[width=\columnwidth]{figures/security_score_operator}
-\caption{Security scores for different operators}
+\caption{Network operators}
 \label{fig:security_score_operator}
+\end{subfigure}
+%
+\begin{subfigure}{\columnwidth}
 \includegraphics[width=\columnwidth]{figures/security_score_summary}
-\caption{Security scores for Nexus and non-Nexus devices\todolater{bridge gaps?}}
+\caption{Nexus and non-Nexus devices}
 \label{fig:security_score_summary}
+\end{subfigure}
+\caption{Security scores for device manufacturers, device models, network operators and Nexus devices. 95\% confidence intervals indicated.\todolater{bridge gaps?}}
+\label{fig:security_scores}
 \end{figure}
@@
 \end{equation}
-Figure~\ref{fig:security_score_manufacturer} shows this for manufacturers, Figure~\ref{fig:security_score_model} for device models, Figure~\ref{fig:security_score_operator} for operators Figure~\ref{fig:security_score_summary} for Nexus and non-Nexus devices; these figures are shown with their 95\% confidence intervals.
+Figure~\ref{fig:security_scores} shows this for manufacturers, device models, network operators and for Nexus and non-Nexus devices.
 These show how the scores for different entities are different and change over time, while there is correlated behaviour for different entities (due to things like new vulnerabilities affecting all Android being discovered) these lines still have crossings due to the different behaviour of the different entities.
@@
 A high value of $u$ could be achieved by only ever shipping one version but that would give low values for $f$ and $m$ (and not be attractive to new customers).
-A high value of $m$ could be achieved by focusing on only one device at a time and ensuring that gets updates but ignoring all others, but that would lower $f$ and $u$.%
-\footnote{One way to influence our scores would be to add additional devices to Device Analyzer which have good security, these would have to be real end user devices as we could detect fake ones, this would increase the size of our data set and would require providing genuinely good security to some users.}
-Therefore our score is secure against gaming.
+A high value of $m$ could be achieved by focusing on only one device at a time and ensuring that it gets updates but ignoring all others, but that would lower $f$ and $u$.%
+One way to influence our scores would be to add additional devices to Device Analyzer which have good security, these would have to be real end user devices as we could detect fake ones, this would increase the size of our data set and would require providing genuinely good security to some users.
+Therefore our score is secure against passive gaming attacks which changed the measured distribution but would require active defence against active gaming attacks which target the measurement devices.

### 2015-06-08 ccs.tex
Expand the description of Figure 1

@@
 \includegraphics[width=\columnwidth]{figures/proportioninsecure}
-\caption{Proportion of devices running insecure, maybe secure and secure versions of Android against time.
-The red vertical lines are caused by vulnerabilities being discovered with those which have the biggest impact annotated.
+\caption{Proportion of devices running insecure, maybe secure and secure versions of Android.
+Table~\ref{tab:andvulns} lists the \daNumVulnsUsed\ vulnerabilities used, the red vertical lines are caused by their discovery and the most important are annotated.
+A device is insecure if it is running a vulnerable version of Android and the device has not received an update which might fix it;
+it is maybe secure if it is running a vulnerable version and has received an update which might fix the vulnerability;
+and it is secure if it is running a secure version.
 }

### 2015-06-08 ccs.tex
The importance of quantification

@@
 
-\todo{We need to rerun the analysis on the AOSP data for its current state}
-\todo{What is the significance of having done this quantification?}
+We have quantified the complexity due to the large number of participants in the ecosystem, the many updates flowing between them and the high latency of updates reaching devices.
+
 \section{Threat model and attack vectors}

### 2015-06-08 security_scoring.tex
Rework description of full_comp figure

@@
 We can then test whether this seems fair by comparing the version data for the highest and lowest scoring models.
-Figure~\ref{fig:full_version_comp}c shows the full version distribution for \emph{\daSecScoreWorstmodel}, Figure~\ref{fig:full_version_comp}b shows the full version distribution for \emph{HTC Desire HD A9191} which used to be our worst model and for which we have more historical data and Figure~\ref{fig:full_version_comp}a shows the same for \emph{\daSecScoreBestmodel}.
-There is a marked difference in the distribution of updates between the best and worst with only \daSecScoreWorstmodelNumFullVersions\ full version for \emph{\daSecScoreWorstmodel}, 2 full versions for \emph{HTC Desire HD A9191} and \daSecScoreBestmodelNumFullVersions\ for \emph{\daSecScoreBestmodel}.
+Figure~\ref{fig:full_version_comp}c shows the full version distribution for \emph{\daSecScoreWorstmodel} which we only observe running one version.
+Figure~\ref{fig:full_version_comp}b shows the full version distribution for \emph{HTC Desire HD A9191} which used to be our worst model and for which we have more historical data, it shows how it received one update at the beginning of 2012 which was deployed fairly rapidly to most devices but received no further updates.
+Figure~\ref{fig:full_version_comp}a shows the same information for \emph{\daSecScoreBestmodel} but it is very different with \daSecScoreBestmodelNumFullVersions\ different versions, some of which were only deployed to small numbers of devices but the distribution for all devices regularly and rapidly transitions from one version to another before ending up on `4.3 JWR66Y'.
 Both \emph{Galaxy Nexus} and \emph{HTC Desire HD A9191} device models start off with the full version string of `2.3.3 GRI40' but the \emph{\daSecScoreBestmodel} receives many more updates over the same time period.

### 2015-06-08 security_scoring.tex
Disable TODOs and insert layout hack for submission

@@
 
+\pagebreak
 \subsection{Scores over time}
+The scoring metric as originally computed is averaged over the whole history of the device manufacturer, device model or network operator, it gives equal weight to periods years ago as to in the last few months.
+If instead we take an exponential moving average of the daily score for days with more than \daSigNumDevicesDay\ devices when there have been at least consecutive \daSigNumDays\ days of data with that many devices then we can plot how this score has changed over time.
+Equation~\ref{eq:rolling_update} shows how the value for a particular day ($v_i$) is computed from the previous day's value and the input for the current day ($n$) with an $\alpha$ of $1/\daSigNumDays$.
+\begin{equation}
+v_i = v_{i-1} (1 - \alpha) + n \alpha
+\label{eq:rolling_update}
+\end{equation}
+Figure~\ref{fig:security_scores} shows this for manufacturers, device models, network operators and for Nexus and non-Nexus devices.
+These show how the scores for different entities are different and change over time, while there is correlated behaviour for different entities (due to things like new vulnerabilities affecting all Android being discovered) these lines still have crossings due to the different behaviour of the different entities.
+It also shows that we do not have sufficient data for all the entities all of the time, resulting in gaps in the data.
+The clearest results are for Figure~\ref{fig:security_score_summary} with a large gap between the scores for Nexus and non-Nexus devices across the whole data set.
 \begin{figure}
@@
 \end{figure}
-\todo{This is method rather than result}
-The scoring metric as originally computed is averaged over the whole history of the device manufacturer, device model or network operator, it gives equal weight to periods years ago as to in the last few months.
-If instead we take an exponential moving average of the daily score for days with more than \daSigNumDevicesDay\ devices when there have been at least consecutive \daSigNumDays\ days of data with that many devices then we can plot how this score has changed over time.
-Equation~\ref{eq:rolling_update} shows how the value for a particular day ($v_i$) is computed from the previous day's value and the input for the current day ($n$) with an $\alpha$ of $1/\daSigNumDays$.
-\begin{equation}
-v_i = v_{i-1} (1 - \alpha) + n \alpha
-\label{eq:rolling_update}
-\end{equation}
-Figure~\ref{fig:security_scores} shows this for manufacturers, device models, network operators and for Nexus and non-Nexus devices.
-These show how the scores for different entities are different and change over time, while there is correlated behaviour for different entities (due to things like new vulnerabilities affecting all Android being discovered) these lines still have crossings due to the different behaviour of the different entities.
-It also shows that we do not have sufficient data for all the entities all of the time, resulting in gaps in the data.
-The clearest results are for Figure~\ref{fig:security_score_summary} with a large gap between the scores for Nexus and non-Nexus devices across the whole data set.
 

### 2015-06-11 ccs.tex
Disable TODOs and insert layout hack for submission
Corrections from Diana

@@
 
-It has been previously shown that Android devices do not receive security updates promptly when critical vulnerabilities are publicly disclosed and hence \daMeanInsecurityPercNominal\ are exposed to known critical vulnerabilities on average~\cite{androidvulnerabilities.org}.
+It has been previously shown that Android devices do not receive security updates promptly when critical vulnerabilities are publicly disclosed and hence, on average, \daMeanInsecurityPercNominal\ are exposed to known critical vulnerabilities~\cite{androidvulnerabilities.org}.
 Here we attempt to determine why by understanding the Android update process, evaluating different entities in the ecosystem and producing a scoring metric to compare different entities.
 
-By producing a scoring metric which compares different entities such as device manufacturers, network operators or specific device models we can provide information which allows device purchasers to select devices with better security.
-Currently corporate and public sector buyers are encouraged to purchase secure devices, but we have found little concrete guidance on the specific makes and models providing the best security.
-For example, CESG, which advises the UK government on how to secure its computer systems, recommends picking Android device models from device manufactures which are good at shipping security updates promptly but it does not state which device manufacturers these are~\cite{CESG2013}.
+By producing a scoring metric which compares different entities such as device manufacturers, network operators or specific device models, we can provide information that allows device purchasers to select devices with better security.
+Currently, corporate and public sector buyers are encouraged to purchase secure devices, but we have found little concrete guidance on the specific makes and models providing the best security.
+For example, CESG, which advises the UK government on how to secure its computer systems, recommends picking Android device models from device manufactures that are good at promptly shipping security updates, but it does not state which device manufacturers these are~\cite{CESG2013}.
 We are also collaborating with a FTSE 100 company to give them concrete information to answer this question.
@@
 The Android update process is slow because its development is distributed among many parties and has a long multi-stage pipeline~\cite{HTC2013}.
-The security of Android relies on many open source projects, these include the Linux kernel, OpenSSL and BouncyCastle as well as Google who build the core platform.
+The security of Android relies on many open source projects, including the Linux kernel, OpenSSL and BouncyCastle, as well as Google, who build the core platform.
 In addition, device manufacturers (we know of \daNumManufacturers\footnote{\label{foot:dadata}We computed this from the \da~\cite{Wagner2013} data, see \S\ref{sec:android_update_process}.}) and network operators (\daNumOperators\dafoot) may make further modifications before devices are shipped to customers.
-Understanding this ecosystem is important as device manufacturers have introduced additional vulnerabilities in the past~\cite{Grace2012}.
+Understanding this ecosystem is important because device manufacturers have introduced additional vulnerabilities in the past~\cite{Grace2012}.
 The cause of the slow update process could be: upstream open source projects releasing patches slowly, Google including them slowly, manufacturers including them slowly, network operators shipping them slowly or customers not installing them.
@@
 \end{table}
-To understand how vulnerabilities in Android are fixed we examine the Android update process which we model in Figure~\ref{fig:update_ecosystem}.
-There are five entities or groups which contribute towards Android updates: the network operators, the device manufacturers, the hardware developers, Google and the upstream open source projects.
-Android builds on various open source projects such as the Linux kernel, OpenSSL and BouncyCastle cryptography libraries.
-Consequently Android can include any compatible versions of those projects, including those which fix security vulnerabilities.
+To understand how vulnerabilities in Android are fixed we examine the Android update process, which we model in Figure~\ref{fig:update_ecosystem}.
+There are five entities or groups that contribute towards Android updates: the network operators, the device manufacturers, the hardware developers, Google and the upstream open source projects.
+Android builds on various open source projects, such as the Linux kernel, OpenSSL and BouncyCastle cryptography libraries.
+Consequently Android can include any compatible versions of those projects, including those that fix security vulnerabilities.
 Android also incorporates various drivers for different bits of hardware.
@@
 Device manufacturers likely receive advanced access in order to prepare handsets so they can customise it before passing it on to the network operator.
-The network operator may then make or request further customisations and perform further testing before shipping the update to the Device.
-Sometimes device manufactures ship updates directly to the user, sometimes the device manufacturer and Google collaborate closely to make a particular phone, such as with Nexus devices.
+The network operator may then make or request further customisations and perform further testing before shipping the update to the device.
+Sometimes device manufactures ship updates directly to the user without involving the network operator.
+Sometimes the device manufacturer and Google collaborate closely to make a particular phone, such as with Nexus devices and so Google ships directly to the device.
 Sometimes device manufacturers incorporate upstream open source project releases directly, and sometimes incorrectly -- for example previous work has recorded evidence of broken nightly builds of sqlite in Android releases~\cite{Wagner2013}.
@@
 Device manufacturer and network operator counts were obtained by normalising the results reported by Android to \da\ of the device manufacturer and active network operator.
-This normalisation is a manual task involves removing invalid values (such as `manufacturer' or `airplane mode is on'), collating across company name changes (e.g.\ `lge' to `LG'), normalising punctuation, removing extra strings sometimes added such as (`(2g)' or `communications') and mapping some incorrectly placed model names back to their manufacturer.
-This normalisation is not perfect and so these are overestimates on the \da\ data but they are likely still underestimates as there will be some device manufacturers and network operators which are not included in the \da\ data.
+This normalisation is a manual task that involves removing invalid values (such as `manufacturer' or `airplane mode is on'), collating across company name changes (e.g.\ `lge' to `LG'), normalising punctuation, removing extra strings sometimes added such as (`(2g)' or `communications') and mapping some incorrectly placed model names back to their manufacturer.
+This normalisation is not perfect and so these are overestimates on the \da\ data but they are likely still underestimates as there will be some device manufacturers and network operators that are not included in the \da\ data.
 The representativeness of \da\ is discussed in \S\ref{sec:representative}.
@@
 In Figure~\ref{fig:update_ecosystem} the number of updates received by devices (\daNumFullVersions) is the number of different full version strings observed in \da.
-The number of updates shipped by Google (\daNumSigOSVersions) is the number of Android versions reported in \da\ which affected more than \daSigVersionPerc\ of devices for more than \daSigVersionDays\ days.
+The number of updates shipped by Google (\daNumSigOSVersions) is the number of Android versions reported in \da\ that affected more than \daSigVersionPerc\ of devices for more than \daSigVersionDays\ days.
 This significance test is to remove spurious versions recorded in \da\ such as `5.2.0' in 2012 which had still not been released at time of writing.\todolater{Check if this has now been released}
@@
 
+\vspace{0.1em}\todo{Remove this display hack}
 \section{Threat model and attack vectors}
@@
 We assume the defender can deploy security controls at three levels: in an online marketplace, at app installation, and during app execution.
-Here we consider three attack vectors against Android handsets.
+Here, we consider three attack vectors against Android handsets.
 
@@
 Android devices can install apps through marketplaces, email attachments, URLs and via the Android Debug Bridge (ADB).
-By default, many Android devices will only allow the installation of apps from Google Play which uses Bouncer to automatically analyse apps; and quickly takes down apps that are reported as malicious.
-However alternative markets are also popular, particularly in countries where Google Play is not available.
+By default, many Android devices will only allow the installation of apps from Google Play, which uses Bouncer to automatically analyse apps, and quickly takes down apps that are reported as malicious.
+However, alternative markets are also popular, particularly in countries where Google Play is not available.
 
-The \emph{dynamic code loading} attack vector is to make an existing app into download and execute code at runtime.
+The \emph{dynamic code loading} attack vector is to make an existing app download and execute code at runtime.
 The most direct method is to upload to a marketplace a seemingly innocent app that dynamically loads malicious code.
@@
 Regardless of version, this exposed an increased attack area and would also provide the ability for malware to control all user internet traffic (via VPNs), brick the phone, remove and install apps, steal user credentials and read the screen.
-The different categories which the vulnerabilities fall into are shown in Table~\ref{tab:andvulns}, `signature' vulnerabilities are \emph{installation} while kernel and system vulnerabilities are used with \emph{installation, dynamic code loading} and \emph{injection}.
+The different categories in which the vulnerabilities fall are shown in Table~\ref{tab:andvulns}, `signature' vulnerabilities are \emph{installation}, while kernel and system vulnerabilities are used with \emph{installation, dynamic code loading} and \emph{injection}.
 \avoTabAndVulns
 
-\avo\ currently contains \avoNumVulnerabilities\ vulnerabilities of which \avoNumVulnAllAndroid\ affect all Android devices and \avoNumVulnSpecific\ are specific to particular devices or device manufacturers.
+\avo\ currently contains \avoNumVulnerabilities\ vulnerabilities, of which \avoNumVulnAllAndroid\ affect all Android devices and \avoNumVulnSpecific\ are specific to particular devices or device manufacturers.
 
@@
 The \da\ data is collected from devices across the world where users have voluntarily installed the app.
-We have obtained comparable data on 5\,290 devices from a multinational FTSE 100 company's mobile device management database which includes company and employee owned devices, and from 5\,170\,000 matching User-Agent headers on all HTTP traffic for 30\% of Rwanda for a week.
+We have obtained comparable data on 5\,290 devices from a multinational FTSE 100 company's mobile device management database, which includes company and employee owned devices, and from 5\,170\,000 matching User-Agent headers on all HTTP traffic for 30\% of Rwanda for a week.
 \todo{citation for Rwanda dataset}
@@
 
-Unfortunately there is no ground truth of OS version information.
-However we have collated the API version information that Google has published most months since 2009.\footnote{\url{http://androidvulnerabilities.org/play/historicplaydashboard}}
+Unfortunately, there is no ground truth of OS version information.
+However, we have collated the API version information that Google has published most months since 2009.\footnote{\url{http://androidvulnerabilities.org/play/historicplaydashboard}}
 While API versions are too coarse grained to use for security update detection they are closely related to OS versions and so if the \da\ data on API versions are similar to the Google data on API versions then the \da\ data on OS versions should be representative.
@@
 The user is notified an update exists, but further action is required, including downloading the update and rebooting the phone to enable installation.
-The phone must have sufficient charge to perform the update and the device itself is rendered inoperable during the update process, two factors which might prevent or delay the update process from taking place.
+The phone must have sufficient charge to perform the update and the device itself is rendered inoperable during the update process, two factors that might prevent or delay the update process from taking place.
 In our data we are unable to determine why a device is not updated. 
 It is possible that many updates arrive at handsets, but are simply not installed.
-Anecdotal evidence at least suggests that it is the lack of updates rather the lack of installation which is the major problem at present and \S\ref{sec:security_scoring:results} supports this.
+Anecdotal evidence at least suggests that it is the lack of updates rather the lack of installation, which is the major problem at present and \S\ref{sec:security_scoring:results} supports this.
 Further work is required to tease these numbers apart.
@@
 The same technique would be more difficult on phones with limited storage space (as many cheap Android phones have barely enough space to install just the update) but is a plausible improvement for more high-end devices.
-Google is deploying the same silent update technique through Google Play Services\footnote{\href{http://lifehacker.com/why-google-play-services-are-now-more-important-than-an-975970197}{http://lifehacker.com/why-google-play-services-are-now-more-important-than-an-975970197}} which automatically installs updates for core Google components of Android, this also bypasses the device manufacturer and network operator.
+Google is deploying the same silent update technique through Google Play Services,\footnote{\href{http://lifehacker.com/why-google-play-services-are-now-more-important-than-an-975970197}{http://lifehacker.com/why-google-play-services-are-now-more-important-than-an-975970197}} which automatically installs updates for core Google components of Android, this also bypasses the device manufacturer and network operator.
 
@@
 SE\-Android~\cite{Smalley2013}, which is included in Android from version 4.1~\cite{jelly-bean-release}, and fully enforcing from version 5.0~\cite{AndroidSecurity2014} claimed to prevent some root vulnerabilities and to reduce the impact of others.
-Capability based enforcement systems such as Capsicum~\cite{Watson2010} substantially reduce the capabilities that an exploit has to try and gain increased privilege with and could be included in Linux\footnote{\url{https://github.com/google/capsicum-linux}} and hence Android.
+Capability based enforcement systems such as Capsicum~\cite{Watson2010} substantially reduce the capabilities with which an exploit has to try and gain increased privilege and could be included in Linux\footnote{\url{https://github.com/google/capsicum-linux}} and hence Android.
 
@@
 Detection strategies include Risk\-Ranker, which classified 3\,281 out of 118\,318 apps (2.8\%) as risky of which 718 (22\%) were malware and 322 (10\%) were previously unknown malware, an infection rate of 0.6\% across multiple markets~\cite{Grace2012a}.
-Droid\-Ranger also analysed apps finding 148 out of 182\,823 apps (0.08\%) to be malicious across multiple markets of which 29 were previously unknown~\cite{Zhou2012a}.
+Droid\-Ranger also analysed apps finding 148 out of 182\,823 apps (0.08\%) to be malicious across multiple markets, of which 29 were previously unknown~\cite{Zhou2012a}.
 A common technique used by attackers is to include malicious code in repackaged popular apps. 
@@
 Other approaches have made further use of malware dependency graphs~\cite{Lindorfer2014} or tried to extract further semantic information to avoid malware being able to avoid detection~\cite{Zhang2014} but simpler approaches based on extracting strings from the binaries can be more effective~\cite{Arp2014} in terms of false positives and false negatives.
-There is a lack of ground truth data, a danger in training algorithms based on APK files which other algorithms have already found which might lead to new malware being missed and DroidChameleon showed that existing AntiVirus apps could be fooled by simple permutations of malicious APKs~\cite{Rastogi2013}.
+There is a lack of ground truth data, a danger in training algorithms based on APK files that other algorithms have already found which might lead to new malware being missed and DroidChameleon showed that existing AntiVirus apps could be fooled by simple permutations of malicious APKs~\cite{Rastogi2013}.
 
 The percentage of Android devices running the most recent version (\daUpdatednessPerc) is much less than the rate ($>90$\%) for Windows XP SP2 computers contacting the Microsoft update servers~\cite{Gkantsidis2006}.
-A simple numerical comparison is unfair because only one major OS version was considered in the Microsoft analysis, and data was only collected from computers which contacted the update server, although this was the default.
+A simple numerical comparison is unfair because only one major OS version was considered in the Microsoft analysis, and data was only collected from computers that contacted the update server, although this was the default.
 More recent data demonstrates the difficulty of upgrading computers between major OS versions, with 27\% of Windows computers running Windows XP in July 2014,\footnote{\url{https://archive.today/PLGxn}} four months after Windows XP stopped receiving security updates.
@@
 \section{Conclusion}
-We examined how the Android ecosystem results in different device models receiving different levels of security due to whether or not they get security updates
-We have defined the FUM score which evaluates the security performance of companies and organisations in shipping security updates.
-We have use this to compare different device models, device manufacturers and network operators and found that there are differences between them which the discerning purchaser might use to influence their decision about which device model to buy from which device manufacturer and network operator.
+We examined how the Android ecosystem results in different device models receiving different levels of security due to whether or not they get security updates.
+We have defined the FUM score, which evaluates the security performance of companies and organisations in shipping security updates.
+We have used this to compare different device models, device manufacturers and network operators and found that there are differences between them, which the discerning purchaser might use to influence their decision about which device model to buy from which device manufacturer and network operator.
 We hope that this analysis will encourage device manufacturers and network operators to improve the support they provide for devices after sale.

### 2015-06-11 security_scoring.tex
Corrections from Diana

--- 2015-06-08 security_scoring.tex
@@
 
-To allow buyers of Android devices to purchase those devices which have the best security they need to know how different device manufacturers, device models and network operators compare in terms of the security they provide.
+To allow buyers of Android devices to purchase those devices with the best security, they need to know how different device manufacturers, device models and network operators compare in terms of the security they provide.
 We propose a method to score a device manufacturer, device model or network operator based on its historic performance at keeping devices up-to-date and fixing security vulnerabilities.
@@
 
-Computing how good a particular device manufacturer or device model is from a security standpoint is difficult as it depends on a number of factors which are hard to observe, particularly on a large scale.
-Ideally we would consider both the prevalence of potential problems which were not exploited and actual security failures.
+Computing how good a particular device manufacturer or device model is from a security standpoint is difficult because it depends on a number of factors which are hard to observe, particularly on a large scale.
+Ideally, we would consider both the prevalence of potential problems that were not exploited and actual security failures.
 %\footnote{A perfectly secure operating system would among other things detect and prevent all phishing attacks, provide perfect principle of least privilege isolation, not have any vulnerabilities, instantly fix all discovered vulnerabilities, not allow any user data to be used in a way which the user does not approve of and be really easy for an ordinary person to use.}
-However in the absence of such data we propose a scheme for assigning a device a score out of ten based on data which can be observed, is based on previous metrics, and which we expect correlates with the actual security of the devices.
+However, in the absence of such data we propose a scheme for assigning a device a score out of ten based on data that can be observed, is based on previous metrics, and that we expect correlates with the actual security of the devices.
 
@@
 \begin{description}
-  \item[free $f$] The proportion of running devices which were free from critical vulnerabilities over time.
+  \item[free $f$] The proportion of running devices free from critical vulnerabilities over time.
   This is equivalent to Acer and Jackson's proposal to measure the security based on the proportion of users with at least one unpatched critical vulnerability~\cite{Acer2010} and similar to the Vulnerability Free Days (VFD) score~\cite{Wright2014}.
-  Unlike VFD this is the proportion of running devices which were free from critical vulnerabilities over time, rather than days which the device manufacturer was free from outstanding critical vulnerabilities as that does not take account of the update process.
-  \item[update $u$] The proportion of devices which run the latest version of Android shipped to any device produced by that device manufacturer.
+  Unlike VFD, this is the proportion of running devices which were free from critical vulnerabilities over time, rather than the number of days which the device manufacturer was free from outstanding critical vulnerabilities, as that does not take account of the update process.
+  \item[update $u$] The proportion of devices that run the latest version of Android shipped to any device produced by that device manufacturer.
 This is a measure of internal updatedness, a low score would mean many devices are being left behind.
   This assumes that newer versions are better with stronger security.
-  Historically steps have been taken to improve Android security in newer versions so this assumption should generally hold, but sometimes new updates introduce new vulnerabilities.
+  Historically, steps have been taken to improve Android security in newer versions so this assumption should generally hold, but sometimes new updates introduce new vulnerabilities.
   \item[mean $m$] The mean number of outstanding vulnerabilities affecting devices not fixed on any device shipped by the device manufacturer.
-This is related to the Median Active Vulnerabilities (MAV) measure~\cite{Wright2014} but is the mean rather than the median, as this gives a continuous value.
+This is related to the Median Active Vulnerabilities (MAV) measure~\cite{Wright2014} but is the mean rather than the median, since this gives a continuous value.
   An example is given in Figure~\ref{fig:mcalculation}.
@@
 
-These three metrics $f$, $u$ and $m$ together measure the security of a platform with respect to known vulnerabilities and updates.
+These three metrics $f$, $u$ and $m$, together measure the security of a platform with respect to known vulnerabilities and updates.
 $f$ is a key measure of the direct risk to users as if there is any known but unfixed vulnerability then they are vulnerable.
-However it does not capture the increased risk caused by there being multiple known vulnerabilities which gives an attacker more opportunities and increases the likelihood of a piece of malware having a matching exploit.
-This is captured by the $m$ score which measures the size of the device manufacturers queue of outstanding vulnerabilities but does not take into account the update process or measure the actual end user security.
-Neither of these metrics capture whether devices are being left behind and not being kept up-to-date with the most recent (and hopefully most secure) version which is captured by $u$.
+However it does not capture the increased risk caused by there being multiple known vulnerabilities, which gives an attacker more opportunities and increases the likelihood of a piece of malware having a matching exploit.
+This is captured by the $m$ score, which measures the size of the device manufacturers queue of outstanding vulnerabilities but does not take into account the update process or measure the actual end user security.
+Neither of these metrics capture whether devices are being left behind and not being kept up-to-date with the most recent (and hopefully most secure) version, which is captured by $u$.
 
-We want to provide a score out of 10 as this is easy for phone buyers to understand as many ratings are given as a score out of 10.
+We want to provide a score out of 10 as this is easy for phone buyers to understand because many ratings are given as a score out of 10.
 Since $f$ is the most important metric we weight it more highly.
@@
 $m$ is computed by counting the number of vulnerabilities which affected that entity and which have not yet been fixed on any device we have observed from that entity every day and averaging over observed days.
-However it could be that the entity has released a fix to some devices but we have not yet observed a device with that fix.
-So the uncertainty in each vulnerability truly being outstanding is the probability of not having observed a fixed device if fixed device existed.
+However, it could be that the entity has released a fix to some devices but we have not yet observed a device with that fix.
+So the uncertainty in each vulnerability truly being outstanding is the probability of not having observed a fixed device if a fixed device existed.
 We assume that if the fix has been released then at least \daMeanOutstandingProbability\ of devices have the fix.
@@
 This does not capture systematic errors such as those due to not including vulnerabilities which do affect devices, but for comparison that does not matter.
-However we do not include manufacturer specific vulnerabilities which will influence comparisons, but we expect that performance in fixing manufacturer specific vulnerabilities is strongly correlated with performance fixing vulnerabilities affecting all of Android.
+However, we do not include manufacturer specific vulnerabilities that will influence comparisons, but we expect that performance in fixing manufacturer specific vulnerabilities is strongly correlated with performance fixing vulnerabilities affecting all of Android.
 
@@
 However there are a wide variety of scores depending on the source of the device.
-There have is much anecdotal evidence that Google's Nexus devices are better at getting updates than other Android devices because Google makes the original updates and ships them to its devices (but the plural of anecdote is not data).
+There is anecdotal evidence that Google's Nexus devices are better at getting updates than other Android devices because Google makes the original updates and ships them to its devices (but the plural of anecdote is not data).
 Table~\ref{tab:sec_summary} shows that this is the case with Nexus devices getting much better scores than non-Nexus devices and Figure~\ref{fig:security_score_summary} shows this holds true over time.
@@
 Device manufacturers are considered significant if we have data from at least \daSigNumDevices\ devices and at least \daSigNumDeviceDays\ days of contributions.
-Additionally for $m$ and $u$ we ignore days with less than \daSigNumDevicesDay\ devices contributing to that day's score.
+Additionally, for $m$ and $u$ we ignore the days with less than \daSigNumDevicesDay\ devices contributing to that day's score.
 
@@
 We can then test whether this seems fair by comparing the version data for the highest and lowest scoring models.
-Figure~\ref{fig:full_version_comp}c shows the full version distribution for \emph{\daSecScoreWorstmodel}, Figure~\ref{fig:full_version_comp}b shows the full version distribution for \emph{HTC Desire HD A9191} which used to be our worst model and for which we have more historical data and Figure~\ref{fig:full_version_comp}a shows the same for \emph{\daSecScoreBestmodel}.
-There is a marked difference in the distribution of updates between the best and worst with only \daSecScoreWorstmodelNumFullVersions\ full version for \emph{\daSecScoreWorstmodel}, 2 full versions for \emph{HTC Desire HD A9191} and \daSecScoreBestmodelNumFullVersions\ for \emph{\daSecScoreBestmodel}.
+Figure~\ref{fig:full_version_comp}c shows the full version distribution for \emph{\daSecScoreWorstmodel}, which we only observe running one version.
+Figure~\ref{fig:full_version_comp}b shows the full version distribution for \emph{HTC Desire HD A9191}, which used to be our worst model and for which we have more historical data; it shows how it received one update at the beginning of 2012, which was deployed fairly rapidly to most devices, but received no further updates.
+Figure~\ref{fig:full_version_comp}a shows the same information for \emph{\daSecScoreBestmodel} but it is very different with \daSecScoreBestmodelNumFullVersions\ different versions, some of which were only deployed to small numbers of devices, but the distribution for all devices regularly and rapidly transitions from one version to another before ending up on `4.3 JWR66Y'.
 Both \emph{Galaxy Nexus} and \emph{HTC Desire HD A9191} device models start off with the full version string of `2.3.3 GRI40' but the \emph{\daSecScoreBestmodel} receives many more updates over the same time period.
@@
 
-If update delays are due to the delay in manufacturers providing the update rather than in operators in supplying the update and users installing the update we would expect the update behaviour of devices with the same device model to be similar and rapid.
+If update delays are due to the delay in manufacturers providing the update rather than in operators supplying the update and users installing the update, we would expect the update behaviour of devices with the same device model to be similar and rapid.
 We found that within \daModelHalfDeployment\ days of the first observation of a new version on a device, half of all devices of that model have the new version (or a higher version) installed, and within \daModelFullDeployment\ days \daFullDeployedAt\ of devices have the new version (or a higher version).
@@
 \todolater{Compute these values automatically and with error.}
-There is a variation between device models, with the update being distributed to most devices quickly and others having a much slower roll out but since some do update quickly the bottleneck is unlikely to be with the user.
+There is a variation between device models, with the update being distributed to most devices quickly and others having a much slower roll out, but since some do update quickly the bottleneck is unlikely to be with the user.
 
@@
 Table~\ref{tab:sec_operator} shows the results \emph{\daSecScoreBestoperator} (\daSecScoreBestoperatorScore\ out of 10) scoring highest and \emph{\daSecScoreWorstoperator} (\daSecScoreWorstoperatorScore\ out of 10) scoring lowest.
-However the score of a network operator is affected by the device manufacturers of the devices which are in use on its network.
-This is in turn affected by both what devices a network operator offers to users and upon which devices users choose.
-Hence having a worse score does not necessarily mean that a network operator is worse, it could be that its users all pick phones from a worse device manufacturer, for example because they were cheaper.
+However, the score of a network operator is affected by the device manufacturers of the devices which are in use on its network.
+This is in turn affected by both the device models a network operator offers to users and upon user's choice of device models.
+Hence, having a worse score does not necessarily mean that a network operator is worse, it could be that its users all pick phones from a worse device manufacturer, for example, because they were cheaper.
 A network operator could use data from this paper to exclude insecure devices from those offered to consumers.
-An added value analysis of network operators which takes into account the device mix used by users of that network operator would make it possible to determine whether a network operator is making the situation better or worse by the way it ships updates to users.
-However our sample size is too small to do that as while we have significant numbers of devices for each of the device models (Table~\ref{tab:sec_model}) and for each of the network operators (Table~\ref{tab:sec_operator}) we would need a significant number of each model in each network operator.
+An added value analysis of network operators, which takes into account the device mix used by users of that network operator, would make it possible to determine whether a network operator is making the situation better or worse by the way it ships updates to users.
+However our sample size is too small to do that because while we have significant numbers of devices for each of the device models (Table~\ref{tab:sec_model}) and for each of the network operators (Table~\ref{tab:sec_operator}), we would need a significant number of each model in each network operator.
 %We do not attempt to disambiguate users behaviour in whether they install updates from network operators using rolling upgrades.
@@
 \subsection{Scores over time}
+The scoring metric as originally computed, is averaged over the whole history of the device manufacturer, device model or network operator, it gives equal weight to both periods years ago and to the last few months.
+If instead we take an exponential moving average of the daily score for days with more than \daSigNumDevicesDay\, devices when there have been at least consecutive \daSigNumDays\, days of data with that many devices then we can plot how this score has changed over time.
+Equation~\ref{eq:rolling_update} shows how the value for a particular day ($v_i$) is computed from the previous day's value and the input for the current day ($n$) with an $\alpha$ of $1/\daSigNumDays$.
+\begin{equation}
+v_i = v_{i-1} (1 - \alpha) + n \alpha
+\label{eq:rolling_update}
+\end{equation}
+Figure~\ref{fig:security_scores} shows this for manufacturers, device models, network operators and for Nexus and non-Nexus devices.
+These show how the scores for different entities are different and change over time, while there is correlated behaviour for different entities (for instance, when new vulnerabilities affecting all Android are discovered), these lines still have crossings due to the different behaviour of the various entities.
+It also shows that we do not have sufficient data for all the entities all of the time, resulting in gaps in the data.
+The clearest results are for Figure~\ref{fig:security_score_summary} with a large gap between the scores for Nexus and non-Nexus devices across the whole data set.
 \begin{figure}
@@
 A high value of $u$ could be achieved by only ever shipping one version but that would give low values for $f$ and $m$ (and not be attractive to new customers).
-A high value of $m$ could be achieved by focusing on only one device at a time and ensuring that it gets updates but ignoring all others, but that would lower $f$ and $u$.%
-One way to influence our scores would be to add additional devices to Device Analyzer which have good security, these would have to be real end user devices as we could detect fake ones, this would increase the size of our data set and would require providing genuinely good security to some users.
-Therefore our score is secure against passive gaming attacks which changed the measured distribution but would require active defence against active gaming attacks which target the measurement devices.
+A high value of $m$ could be achieved by focusing on only one device at a time and ensuring that it gets updates but ignoring all others, but that would lower $f$ and $u$.
+One way to influence our scores would be to add additional devices to Device Analyzer, which have good security, these would have to be real end user devices since we could detect fake ones.
+This would increase the size of our data set and would require providing genuinely good security to some users.
+Therefore, our score is secure against passive gaming attacks which change the measured distribution, but would require active defence against active gaming attacks, which target the measurement devices.

### 2015-07-23 ccs.tex
Munge sig-alternate to have the correct copyright boilerplate and support DOIs and supply the paper's DOI
Switch to sig-alternate-2013.cls

@@
-\documentclass{sig-alternate}
-\pagenumbering{arabic}
+\documentclass{sig-alternate-2013}
 
@@
 % --- Author Metadata here ---
-\conferenceinfo{SPSM}{2015 Denver, Colorado, USA}
-%\CopyrightYear{2015} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
-%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
+
+\newfont{\mycrnotice}{ptmr8t at 7pt}
+\newfont{\myconfname}{ptmri8t at 7pt}
+\let\crnotice\mycrnotice%
+\let\confname\myconfname%
+
+\permission{Permission to make digital or hard copies of all or part of this work for personal or classroom use is granted without fee provided that copies are not made or distributed for profit or commercial advantage and that copies bear this notice and the full citation on the first page.
+Copyrights for components of this work owned by others than the author(s) must be honored.
+Abstracting with credit is permitted.
+To copy otherwise, or republish, to post on servers or to redistribute to lists, requires prior specific permission and/or a fee.
+Request permissions from \href{mailto:Permissions@acm.org}{Permissions@acm.org}.}
+\conferenceinfo{SPSM'15,}{October 12, 2015, Denver, Colorado, USA. \\
+{\mycrnotice{Copyright is held by the owner/author(s). Publication rights licensed to ACM.}}}
+\copyrightetc{ACM \the\acmcopyr}
+\crdata{978-1-4503-3819-6/15/10\ ...\$15.00.\\
+DOI: \href{http://dx.doi.org/10.1145/2808117.2808118}{http://dx.doi.org/10.1145/2808117.2808118}}
+% Update the XXX's to the DOI assigned by ACM rightsreview forms
+
+\clubpenalty=10000
+\widowpenalty=10000
 % --- End of Author Metadata ---

### 2015-07-23 spsm04-thomas.tex
Rename ccs.tex to spsm04-thomas.tex
Hide boxes around links
Add some hyphenation info

--- 2015-07-23 ccs.tex
@@
 \usepackage{url}
-\usepackage{hyperref}
+\usepackage[hidelinks]{hyperref}
 
@@
 
-\title{Measuring security performance in the Android ecosystem}
+\title{Security Metrics for the Android Ecosystem}
 
@@
 \and
-Alastair~R.~Beresford\\
+\href{http://orcid.org/0000-0003-0818-6535}{Alastair~R.~Beresford}\\
        \affaddr{Computer Laboratory}\\
@@
 \and
-Andrew~Rice
+\href{http://orcid.org/0000-0002-4677-8032}{Andrew~Rice}
 }
@@
 The Android update process is slow because its development is distributed among many parties and has a long multi-stage pipeline~\cite{HTC2013}.
-The security of Android relies on many open source projects, including the Linux kernel, OpenSSL and BouncyCastle, as well as Google, who build the core platform.
+The security of Android relies on many open source projects, including the Li\-nux kernel, Open\-SSL and Bouncy\-Castle, as well as Google, who build the core platform.
 In addition, device manufacturers (we know of \daNumManufacturers\footnote{\label{foot:dadata}We computed this from the \da~\cite{Wagner2013} data, see \S\ref{sec:android_update_process}.}) and network operators (\daNumOperators\dafoot) may make further modifications before devices are shipped to customers.
@@
 Thanks to Sherif Akoush and Ripduman Sohan for supplying the Rwanda data and to staff at the FTSE 100 company for supplying their mobile device management data.
+Thanks to our anonymous reviewers for their excellent insightful comments.
 }

### 2015-07-23 dastats.tex
./update.sh

@@
-\newcommand{\daNumSigModels}{17}
+\newcommand{\daNumSigModels}{18}
 \newcommand{\daSigNumDevices}{100}
@@
 \newcommand{\daNumDeviceDataDevices}{50}
-\newcommand{\daOSYearsOfData}{3}
-\newcommand{\daOSMonthsOfData}{47}
+\newcommand{\daOSYearsOfData}{4}
+\newcommand{\daOSMonthsOfData}{49}
 \newcommand{\daStartDate}{2011-07-01}
-\newcommand{\daEndDate}{2015-05-31}
+\newcommand{\daEndDate}{2015-07-14}
 \newcommand{\daFullDeployedAt}{95\%}
-\newcommand{\daNumManufacturers}{299}
-\newcommand{\daNumSigManufacturers}{8}
-\newcommand{\daNumModels}{2\,700}
-\newcommand{\daNumOperators}{1\,440}
+\newcommand{\daNumManufacturers}{300}
+\newcommand{\daNumSigManufacturers}{10}
+\newcommand{\daNumModels}{2\,730}
+\newcommand{\daNumOperators}{1\,460}
 \newcommand{\daNumSigOperators}{14}
-\newcommand{\daNumOSDevices}{20\,300}
-\newcommand{\daOSVersionPercValidLines}{99.7\%}
-\newcommand{\daOSTotalDaysData}{1\,290\,000}
-\newcommand{\daMeanInsecurityPercNominal}{87.9\%}
-\newcommand{\daMeanInsecurityPerc}{$87.9 \pm 0.0\%$}
+\newcommand{\daNumOSDevices}{20\,400}
+\newcommand{\daOSVersionPercValidLines}{99.9\%}
+\newcommand{\daOSTotalDaysData}{1\,330\,000}
+\newcommand{\daMeanInsecurityPercNominal}{87.7\%}
+\newcommand{\daMeanInsecurityPerc}{$87.7 \pm 0.0\%$}
 \newcommand{\daMeanInsecurityPercTwosfNominal}{88\%}
 \newcommand{\daMeanInsecurityPercTwosf}{$88 \pm 0\%$}
-\newcommand{\daMeanSecurityPercNominal}{12.1\%}
-\newcommand{\daMeanSecurityPerc}{$12.1 \pm 0.0\%$}
+\newcommand{\daMeanSecurityPercNominal}{12.3\%}
+\newcommand{\daMeanSecurityPerc}{$12.3 \pm 0.0\%$}
 \newcommand{\daMeanSecurityPercTwosfNominal}{12\%}
 \newcommand{\daMeanSecurityPercTwosf}{$12 \pm 0\%$}
-\newcommand{\daVulnFreeNominal}{0.121}
-\newcommand{\daVulnFree}{$0.121 \pm 0.000$}
-\newcommand{\daMeanOutstandingVulnerabilities}{$0.547 \pm 2.45e$}
-\newcommand{\daUpdatednessNominal}{0.0511}
-\newcommand{\daUpdatedness}{$0.0511 \pm 0.0002$}
-\newcommand{\daUpdatednessPercNominal}{5.11\%}
-\newcommand{\daUpdatednessPerc}{$5.11 \pm 0.02\%$}
-\newcommand{\daUpdatednessPercTwosfNominal}{5.1\%}
-\newcommand{\daUpdatednessPercTwosf}{$5.1 \pm 0.0\%$}
-\newcommand{\daSecurityScoreNominal}{2.84}
-\newcommand{\daSecurityScore}{$2.84 \pm 0.00$}
-\newcommand{\daCurrentMeanInsecurityPercNominal}{85.4\%}
-\newcommand{\daCurrentMeanInsecurityPerc}{$85.4 \pm 0.1\%$}
-\newcommand{\daCurrentMeanInsecurityPercTwosfNominal}{85\%}
-\newcommand{\daCurrentMeanInsecurityPercTwosf}{$85 \pm 0\%$}
-\newcommand{\daCurrentMeanSecurityPercNominal}{14.6\%}
-\newcommand{\daCurrentMeanSecurityPerc}{$14.6 \pm 0.1\%$}
-\newcommand{\daCurrentMeanSecurityPercTwosfNominal}{15\%}
-\newcommand{\daCurrentMeanSecurityPercTwosf}{$15 \pm 0\%$}
-\newcommand{\daCurrentVulnFreeNominal}{0.146}
-\newcommand{\daCurrentVulnFree}{$0.146 \pm 0.001$}
-\newcommand{\daCurrentMeanOutstandingVulnerabilities}{$0.193 \pm 1.11e$}
-\newcommand{\daCurrentUpdatednessNominal}{0.0326}
-\newcommand{\daCurrentUpdatedness}{$0.0326 \pm 0.0006$}
-\newcommand{\daCurrentUpdatednessPercNominal}{3.26\%}
-\newcommand{\daCurrentUpdatednessPerc}{$3.26 \pm 0.06\%$}
-\newcommand{\daCurrentUpdatednessPercTwosfNominal}{3.3\%}
-\newcommand{\daCurrentUpdatednessPercTwosf}{$3.3 \pm 0.0\%$}
-\newcommand{\daCurrentSecurityScoreNominal}{3.39}
-\newcommand{\daCurrentSecurityScore}{$3.39 \pm 0.00$}
-\newcommand{\daDailyParticipationNominal}{903}
-\newcommand{\daDailyParticipation}{$903 \pm 227$}
-\newcommand{\daNumFullVersions}{1\,250}
-\newcommand{\daNumSigFullVersions}{94}
+\newcommand{\daVulnFreeNominal}{0.123}
+\newcommand{\daVulnFree}{$0.123 \pm 0.0$}
+\newcommand{\daMeanOutstandingVulnerabilities}{$0.531 \pm 0.0$}
+\newcommand{\daUpdatednessNominal}{0.0521}
+\newcommand{\daUpdatedness}{$0.0521 \pm 0.0$}
+\newcommand{\daUpdatednessPercNominal}{5.21\%}
+\newcommand{\daUpdatednessPerc}{$5.21 \pm 0.0\%$}
+\newcommand{\daUpdatednessPercTwosfNominal}{5.2\%}
+\newcommand{\daUpdatednessPercTwosf}{$5.2 \pm 0.0\%$}
+\newcommand{\daSecurityScoreNominal}{2.87}
+\newcommand{\daSecurityScore}{$2.87 \pm 0.0$}
+\newcommand{\daCurrentMeanInsecurityPercNominal}{79.7\%}
+\newcommand{\daCurrentMeanInsecurityPerc}{$79.7 \pm 0.0\%$}
+\newcommand{\daCurrentMeanInsecurityPercTwosfNominal}{80\%}
+\newcommand{\daCurrentMeanInsecurityPercTwosf}{$80 \pm 0\%$}
+\newcommand{\daCurrentMeanSecurityPercNominal}{20.3\%}
+\newcommand{\daCurrentMeanSecurityPerc}{$20.3 \pm 0.1\%$}
+\newcommand{\daCurrentMeanSecurityPercTwosfNominal}{20\%}
+\newcommand{\daCurrentMeanSecurityPercTwosf}{$20 \pm 0\%$}
+\newcommand{\daCurrentVulnFreeNominal}{0.203}
+\newcommand{\daCurrentVulnFree}{$0.203 \pm 0.001$}
+\newcommand{\daCurrentMeanOutstandingVulnerabilities}{$0.124 \pm 0.0$}
+\newcommand{\daCurrentUpdatednessNominal}{0.0532}
+\newcommand{\daCurrentUpdatedness}{$0.0532 \pm 0.0$}
+\newcommand{\daCurrentUpdatednessPercNominal}{5.32\%}
+\newcommand{\daCurrentUpdatednessPerc}{$5.32 \pm 0.0\%$}
+\newcommand{\daCurrentUpdatednessPercTwosfNominal}{5.3\%}
+\newcommand{\daCurrentUpdatednessPercTwosf}{$5.3 \pm 0.0\%$}
+\newcommand{\daCurrentSecurityScoreNominal}{3.79}
+\newcommand{\daCurrentSecurityScore}{$3.79 \pm 0.0$}
+\newcommand{\daDailyParticipationNominal}{899}
+\newcommand{\daDailyParticipation}{$899 \pm 245$}
+\newcommand{\daNumFullVersions}{1\,270}
+\newcommand{\daNumSigFullVersions}{95}
 \newcommand{\daNumOSVersions}{50}
-\newcommand{\daNumSigOSVersions}{27}
+\newcommand{\daNumSigOSVersions}{28}
 \newcommand{\daNumAPIVersions}{16}
-\newcommand{\daNumFullVersionUpdates}{4\,540}
-\newcommand{\daNumOSVersionUpdates}{3\,850}
-\newcommand{\daUpdatesPerYearNominal}{1.28}
-\newcommand{\daUpdatesPerYear}{$1.28 \pm 0.01$}
+\newcommand{\daNumFullVersionUpdates}{4\,590}
+\newcommand{\daNumOSVersionUpdates}{3\,900}
+\newcommand{\daUpdatesPerYearNominal}{1.26}
+\newcommand{\daUpdatesPerYear}{$1.26 \pm 0.01$}
 \newcommand{\daUpdatesPerYearTwosfNominal}{1.3}
 \newcommand{\daUpdatesPerYearTwosf}{$1.3 \pm 0.0$}
-\newcommand{\daNumFullOnlyVersionUpdates}{673}
-\newcommand{\daNumSecurityUpdates}{1\,240}
-\newcommand{\daNumPossibleSecurityUpdates}{2\,290}
-\newcommand{\daOSCurveFitParamFirst}{84.8}
-\newcommand{\daOSCurveFitParamSecond}{\num{0.00289}}
-\newcommand{\daOSCurveFitRMSE}{0.116}
-\newcommand{\daOSCurvePolyRMSE}{0.116}
-\newcommand{\daOSCurveSplineRMSE}{0.116}
-\newcommand{\daOSCurveHalfDeployedYearsNominal}{0.888}
-\newcommand{\daOSCurveHalfDeployedYears}{$0.888 \pm 0.249$}
-\newcommand{\daOSCurveHalfDeployedNominal}{324}
-\newcommand{\daOSCurveHalfDeployed}{$324 \pm 90$}
-\newcommand{\daOSCurveFullDeployedYearsNominal}{3.07}
-\newcommand{\daOSCurveFullDeployedYears}{$3.07 \pm 70$}
-\newcommand{\daOSCurveFullDeployedNominal}{1\,120}
-\newcommand{\daOSCurveFullDeployed}{$1\,120 \pm 256\,000$}
-\newcommand{\daAPICurveFitParamFirst}{102}
-\newcommand{\daAPICurveFitParamSecond}{\num{0.00289}}
-\newcommand{\daAPICurveFitRMSE}{0.113}
-\newcommand{\daAPICurvePolyRMSE}{0.113}
-\newcommand{\daAPICurveSplineRMSE}{0.113}
-\newcommand{\daAPICurveHalfDeployedYearsNominal}{0.934}
-\newcommand{\daAPICurveHalfDeployedYears}{$0.934 \pm 0.243$}
-\newcommand{\daAPICurveHalfDeployedNominal}{341}
-\newcommand{\daAPICurveHalfDeployed}{$341 \pm 88$}
-\newcommand{\daAPICurveFullDeployedYearsNominal}{3.11}
-\newcommand{\daAPICurveFullDeployedYears}{$3.11 \pm 70$}
-\newcommand{\daAPICurveFullDeployedNominal}{1\,140}
-\newcommand{\daAPICurveFullDeployed}{$1\,140 \pm 256\,000$}
+\newcommand{\daNumFullOnlyVersionUpdates}{677}
+\newcommand{\daNumSecurityUpdates}{1\,250}
+\newcommand{\daNumPossibleSecurityUpdates}{2\,300}
+\newcommand{\daOSCurveFitParamFirst}{84.4}
+\newcommand{\daOSCurveFitParamSecond}{\num{0.00285}}
+\newcommand{\daOSCurveFitRMSE}{0.115}
+\newcommand{\daOSCurvePolyRMSE}{0.115}
+\newcommand{\daOSCurveSplineRMSE}{0.115}
+\newcommand{\daOSCurveHalfDeployedYearsNominal}{0.897}
+\newcommand{\daOSCurveHalfDeployedYears}{$0.897 \pm 0.252$}
+\newcommand{\daOSCurveHalfDeployedNominal}{327}
+\newcommand{\daOSCurveHalfDeployed}{$327 \pm 91$}
+\newcommand{\daOSCurveFullDeployedYearsNominal}{3.11}
+\newcommand{\daOSCurveFullDeployedYears}{$3.11 \pm 71$}
+\newcommand{\daOSCurveFullDeployedNominal}{1\,130}
+\newcommand{\daOSCurveFullDeployed}{$1\,130 \pm 260\,000$}
+\newcommand{\daAPICurveFitParamFirst}{99.7}
+\newcommand{\daAPICurveFitParamSecond}{\num{0.00285}}
+\newcommand{\daAPICurveFitRMSE}{0.112}
+\newcommand{\daAPICurvePolyRMSE}{0.112}
+\newcommand{\daAPICurveSplineRMSE}{0.112}
+\newcommand{\daAPICurveHalfDeployedYearsNominal}{0.939}
+\newcommand{\daAPICurveHalfDeployedYears}{$0.939 \pm 0.244$}
+\newcommand{\daAPICurveHalfDeployedNominal}{343}
+\newcommand{\daAPICurveHalfDeployed}{$343 \pm 89$}
+\newcommand{\daAPICurveFullDeployedYearsNominal}{3.15}
+\newcommand{\daAPICurveFullDeployedYears}{$3.15 \pm 71$}
+\newcommand{\daAPICurveFullDeployedNominal}{1\,150}
+\newcommand{\daAPICurveFullDeployed}{$1\,150 \pm 260\,000$}
 \newcommand{\daVulnAPKDuplicateFileOctoberPerc}{91.6\%}
 \newcommand{\daVulnZergRushMonthsDefFixDeployed}{27.3}
-\newcommand{\daTabSecScoresmanufacturer}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline LG & $0.20 \pm 0.00$ & $0.33 \pm 0.00$ & $0.63 \pm 0.01$ & $3.86 \pm 0.02$ \\  Motorola & $0.18 \pm 0.00$ & $0.13 \pm 0.00$ & $0.73 \pm 0.02$ & $3.03 \pm 0.02$ \\  Samsung & $0.13 \pm 0.00$ & $0.04 \pm 0.00$ & $0.63 \pm 0.00$ & $2.72 \pm 0.00$ \\  Sony & $0.13 \pm 0.00$ & $0.19 \pm 0.00$ & $1.09 \pm 0.02$ & $2.61 \pm 0.02$ \\  HTC & $0.14 \pm 0.00$ & $0.09 \pm 0.00$ & $0.89 \pm 0.01$ & $2.58 \pm 0.02$ \\  asus & $0.19 \pm 0.00$ & $0.52 \pm 0.01$ & $6.01 \pm 0.07$ & $2.32 \pm 0.02$ \\  other & $0.06 \pm 0.00$ & $0.04 \pm 0.00$ & $1.03 \pm 0.01$ & $1.95 \pm 0.02$ \\  Symphony & $0.00 \pm 0.00$ & $0.08 \pm 0.00$ & $5.00 \pm 0.05$ & $0.28 \pm 0.01$ \\  walton & $0.00 \pm 0.00$ & $0.08 \pm 0.00$ & $5.96 \pm 0.08$ & $0.25 \pm 0.01$ \\ \end{tabular} \caption{Security scores for manufacturers} \label{tab:sec_manufacturer} \end{table*}}
+\newcommand{\daTabSecScoresmanufacturer}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline LG & $0.22 \pm 0.00$ & $0.33 \pm 0.00$ & $0.62 \pm 0.01$ & $3.97 \pm 0.02$ \\  Motorola & $0.18 \pm 0.00$ & $0.12 \pm 0.00$ & $0.72 \pm 0.02$ & $3.07 \pm 0.02$ \\  Samsung & $0.13 \pm 0.00$ & $0.04 \pm 0.00$ & $0.61 \pm 0.00$ & $2.75 \pm 0.00$ \\  Sony & $0.14 \pm 0.00$ & $0.19 \pm 0.00$ & $1.09 \pm 0.02$ & $2.63 \pm 0.02$ \\  HTC & $0.14 \pm 0.00$ & $0.10 \pm 0.00$ & $0.87 \pm 0.01$ & $2.63 \pm 0.02$ \\  asus & $0.20 \pm 0.00$ & $0.51 \pm 0.01$ & $6.01 \pm 0.07$ & $2.35 \pm 0.02$ \\  other & $0.06 \pm 0.00$ & $0.05 \pm 0.00$ & $1.04 \pm 0.01$ & $1.97 \pm 0.02$ \\  alps & $0.03 \pm 0.00$ & $0.19 \pm 0.01$ & $3.99 \pm 0.08$ & $0.80 \pm 0.02$ \\  Symphony & $0.00 \pm 0.00$ & $0.08 \pm 0.00$ & $5.00 \pm 0.05$ & $0.30 \pm 0.01$ \\  walton & $0.00 \pm 0.00$ & $0.09 \pm 0.00$ & $6.00 \pm 0.08$ & $0.27 \pm 0.01$ \\ \end{tabular} \caption{Security scores for manufacturers} \label{tab:sec_manufacturer} \end{table*}}
 \newcommand{\daSecScoreBestmanufacturer}{LG}
-\newcommand{\daSecScoreBestmanufacturerScoreNominal}{3.86}
-\newcommand{\daSecScoreBestmanufacturerScore}{$3.86 \pm 0.02$}
-\newcommand{\daSecScoreBestmanufacturerNumFullVersions}{131}
+\newcommand{\daSecScoreBestmanufacturerScoreNominal}{3.97}
+\newcommand{\daSecScoreBestmanufacturerScore}{$3.97 \pm 0.0$}
+\newcommand{\daSecScoreBestmanufacturerNumFullVersions}{135}
 \newcommand{\daSecScoreWorstmanufacturer}{walton}
-\newcommand{\daSecScoreWorstmanufacturerScoreNominal}{0.251}
-\newcommand{\daSecScoreWorstmanufacturerScore}{$0.251 \pm 0.007$}
-\newcommand{\daSecScoreWorstmanufacturerNumFullVersions}{22}
-\newcommand{\daTabSecScoresmodel}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline Galaxy Nexus & $0.51 \pm 0.00$ & $0.55 \pm 0.01$ & $1.53 \pm 0.04$ & $4.74 \pm 0.04$ \\  Nexus 4 & $0.28 \pm 0.00$ & $0.82 \pm 0.01$ & $6.06 \pm 0.09$ & $3.61 \pm 0.04$ \\  Nexus 7 & $0.24 \pm 0.00$ & $0.74 \pm 0.01$ & $5.92 \pm 0.09$ & $3.19 \pm 0.04$ \\  other & $0.09 \pm 0.00$ & $0.14 \pm 0.00$ & $0.55 \pm 0.00$ & $2.98 \pm 0.00$ \\  Desire HD & $0.08 \pm 0.00$ & $0.05 \pm 0.00$ & $0.39 \pm 0.02$ & $2.91 \pm 0.04$ \\  HTC Sensation & $0.34 \pm 0.00$ & $0.01 \pm 0.01$ & $1.59 \pm 0.05$ & $2.41 \pm 0.05$ \\  GT-I9100 & $0.22 \pm 0.00$ & $0.02 \pm 0.00$ & $1.23 \pm 0.02$ & $2.28 \pm 0.02$ \\  HTC Desire S & $0.02 \pm 0.00$ & $0.02 \pm 0.00$ & $1.00 \pm 0.06$ & $1.75 \pm 0.07$ \\  GT-N7000 & $0.25 \pm 0.00$ & $0.00 \pm 0.00$ & $2.52 \pm 0.05$ & $1.43 \pm 0.02$ \\  GT-P1000 & $0.01 \pm 0.00$ & $0.00 \pm 0.01$ & $1.79 \pm 0.06$ & $0.90 \pm 0.05$ \\  GT-I9300 & $0.13 \pm 0.00$ & $0.01 \pm 0.00$ & $6.25 \pm 0.04$ & $0.58 \pm 0.01$ \\  GT-I9505 & $0.01 \pm 0.00$ & $0.13 \pm 0.00$ & $6.83 \pm 0.07$ & $0.46 \pm 0.01$ \\  HTC Desire HD & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & $3.05 \pm 0.05$ & $0.27 \pm 0.03$ \\  GT-N7100 & $0.06 \pm 0.00$ & $0.00 \pm 0.01$ & $6.91 \pm 0.08$ & $0.25 \pm 0.02$ \\  Symphony W68 & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & $11.00 \pm 0.12$ & $0.00 \pm 0.03$ \\ \end{tabular} \caption{Security scores for models} \label{tab:sec_model} \end{table*}}
+\newcommand{\daSecScoreWorstmanufacturerScoreNominal}{0.272}
+\newcommand{\daSecScoreWorstmanufacturerScore}{$0.272 \pm 0.007$}
+\newcommand{\daSecScoreWorstmanufacturerNumFullVersions}{21}
+\newcommand{\daTabSecScoresmodel}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline Galaxy Nexus & $0.50 \pm 0.00$ & $0.54 \pm 0.01$ & $1.53 \pm 0.04$ & $4.71 \pm 0.04$ \\  Nexus 4 & $0.30 \pm 0.00$ & $0.82 \pm 0.01$ & $6.06 \pm 0.09$ & $3.69 \pm 0.04$ \\  Nexus 7 & $0.26 \pm 0.00$ & $0.74 \pm 0.01$ & $5.92 \pm 0.09$ & $3.24 \pm 0.04$ \\  other & $0.10 \pm 0.00$ & $0.14 \pm 0.00$ & $0.53 \pm 0.00$ & $3.03 \pm 0.00$ \\  Desire HD & $0.08 \pm 0.00$ & $0.05 \pm 0.00$ & $0.38 \pm 0.02$ & $2.91 \pm 0.04$ \\  HTC Sensation & $0.35 \pm 0.00$ & $0.01 \pm 0.01$ & $1.57 \pm 0.05$ & $2.44 \pm 0.05$ \\  GT-I9100 & $0.22 \pm 0.00$ & $0.02 \pm 0.00$ & $1.23 \pm 0.02$ & $2.28 \pm 0.02$ \\  HTC Desire S & $0.02 \pm 0.00$ & $0.02 \pm 0.00$ & $1.00 \pm 0.06$ & $1.74 \pm 0.07$ \\  GT-N7000 & $0.25 \pm 0.00$ & $0.00 \pm 0.00$ & $2.52 \pm 0.05$ & $1.43 \pm 0.02$ \\  GT-P1000 & $0.01 \pm 0.00$ & $0.00 \pm 0.01$ & $1.79 \pm 0.06$ & $0.90 \pm 0.05$ \\  GT-I9300 & $0.13 \pm 0.00$ & $0.01 \pm 0.00$ & $6.23 \pm 0.04$ & $0.58 \pm 0.01$ \\  GT-I9505 & $0.03 \pm 0.00$ & $0.13 \pm 0.00$ & $6.82 \pm 0.07$ & $0.52 \pm 0.01$ \\  HTC Desire HD & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & $3.03 \pm 0.05$ & $0.28 \pm 0.03$ \\  GT-N7100 & $0.06 \pm 0.00$ & $0.00 \pm 0.01$ & $6.93 \pm 0.08$ & $0.24 \pm 0.02$ \\  Symphony W68 & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & $11.00 \pm 0.12$ & $0.00 \pm 0.03$ \\ \end{tabular} \caption{Security scores for models} \label{tab:sec_model} \end{table*}}
 \newcommand{\daSecScoreBestmodel}{Galaxy Nexus}
-\newcommand{\daSecScoreBestmodelScoreNominal}{4.74}
-\newcommand{\daSecScoreBestmodelScore}{$4.74 \pm 0.04$}
+\newcommand{\daSecScoreBestmodelScoreNominal}{4.71}
+\newcommand{\daSecScoreBestmodelScore}{$4.71 \pm 0.0$}
 \newcommand{\daSecScoreBestmodelNumFullVersions}{49}
@@
 \newcommand{\daSecScoreWorstmodelScoreNominal}{0.0001}
-\newcommand{\daSecScoreWorstmodelScore}{$0.0001 \pm 0.0275$}
+\newcommand{\daSecScoreWorstmodelScore}{$0.0001 \pm 0.0273$}
 \newcommand{\daSecScoreWorstmodelNumFullVersions}{1}
-\newcommand{\daTabSecScoressummary}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline nexus & $0.38 \pm 0.00$ & $0.47 \pm 0.00$ & $0.58 \pm 0.01$ & $5.07 \pm 0.02$ \\  notnexus & $0.10 \pm 0.00$ & $0.02 \pm 0.00$ & $0.55 \pm 0.00$ & $2.67 \pm 0.00$ \\ \end{tabular} \caption{Security scores for nexus} \label{tab:sec_summary} \end{table*}}
-\newcommand{\daTabSecScoresoperator}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline O2 uk & $0.26 \pm 0.00$ & $0.12 \pm 0.00$ & $0.37 \pm 0.02$ & $3.85 \pm 0.03$ \\  T-Mobile & $0.21 \pm 0.00$ & $0.18 \pm 0.00$ & $0.42 \pm 0.02$ & $3.73 \pm 0.03$ \\  Orange & $0.21 \pm 0.00$ & $0.09 \pm 0.00$ & $0.39 \pm 0.03$ & $3.52 \pm 0.04$ \\  Sprint & $0.18 \pm 0.00$ & $0.11 \pm 0.00$ & $0.40 \pm 0.02$ & $3.46 \pm 0.03$ \\  3 & $0.18 \pm 0.00$ & $0.09 \pm 0.00$ & $0.52 \pm 0.03$ & $3.24 \pm 0.04$ \\  Vodafone uk & $0.13 \pm 0.00$ & $0.13 \pm 0.00$ & $0.50 \pm 0.03$ & $3.18 \pm 0.04$ \\  AT\&T & $0.13 \pm 0.00$ & $0.08 \pm 0.00$ & $0.43 \pm 0.02$ & $3.10 \pm 0.02$ \\  Verizon & $0.18 \pm 0.00$ & $0.09 \pm 0.00$ & $0.82 \pm 0.02$ & $2.83 \pm 0.02$ \\  unknown & $0.11 \pm 0.00$ & $0.20 \pm 0.00$ & $0.86 \pm 0.02$ & $2.81 \pm 0.02$ \\  n Telenor & $0.03 \pm 0.00$ & $0.12 \pm 0.00$ & $1.25 \pm 0.02$ & $1.82 \pm 0.02$ \\  Airtel & $0.05 \pm 0.00$ & $0.03 \pm 0.00$ & $1.62 \pm 0.03$ & $1.28 \pm 0.03$ \\  Robi & $0.00 \pm 0.00$ & $0.08 \pm 0.00$ & $2.08 \pm 0.04$ & $0.90 \pm 0.03$ \\  Grameenphone & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & $2.02 \pm 0.02$ & $0.85 \pm 0.01$ \\  banglalink & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & $2.60 \pm 0.04$ & $0.54 \pm 0.02$ \\ \end{tabular} \caption{Security scores for operators} \label{tab:sec_operator} \end{table*}}
+\newcommand{\daTabSecScoressummary}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline nexus & $0.39 \pm 0.00$ & $0.48 \pm 0.00$ & $0.56 \pm 0.01$ & $5.17 \pm 0.02$ \\  notnexus & $0.10 \pm 0.00$ & $0.02 \pm 0.00$ & $0.53 \pm 0.00$ & $2.70 \pm 0.00$ \\ \end{tabular} \caption{Security scores for nexus} \label{tab:sec_summary} \end{table*}}
+\newcommand{\daTabSecScoresoperator}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline O2 uk & $0.27 \pm 0.00$ & $0.12 \pm 0.00$ & $0.37 \pm 0.02$ & $3.87 \pm 0.03$ \\  T-Mobile & $0.22 \pm 0.00$ & $0.18 \pm 0.00$ & $0.40 \pm 0.01$ & $3.81 \pm 0.02$ \\  Orange & $0.22 \pm 0.00$ & $0.10 \pm 0.00$ & $0.36 \pm 0.02$ & $3.64 \pm 0.04$ \\  Sprint & $0.18 \pm 0.00$ & $0.11 \pm 0.00$ & $0.43 \pm 0.02$ & $3.42 \pm 0.03$ \\  3 & $0.20 \pm 0.00$ & $0.09 \pm 0.00$ & $0.47 \pm 0.02$ & $3.38 \pm 0.03$ \\  Vodafone uk & $0.14 \pm 0.00$ & $0.13 \pm 0.00$ & $0.52 \pm 0.03$ & $3.17 \pm 0.04$ \\  AT\&T & $0.14 \pm 0.00$ & $0.08 \pm 0.00$ & $0.43 \pm 0.02$ & $3.13 \pm 0.02$ \\  unknown & $0.11 \pm 0.00$ & $0.20 \pm 0.00$ & $0.84 \pm 0.01$ & $2.88 \pm 0.02$ \\  Verizon & $0.19 \pm 0.00$ & $0.09 \pm 0.00$ & $0.82 \pm 0.02$ & $2.84 \pm 0.02$ \\  n Telenor & $0.04 \pm 0.00$ & $0.12 \pm 0.00$ & $1.21 \pm 0.02$ & $1.89 \pm 0.02$ \\  Airtel & $0.05 \pm 0.00$ & $0.03 \pm 0.00$ & $1.47 \pm 0.03$ & $1.41 \pm 0.03$ \\  Grameenphone & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & $1.88 \pm 0.02$ & $0.93 \pm 0.01$ \\  Robi & $0.00 \pm 0.00$ & $0.08 \pm 0.00$ & $2.07 \pm 0.04$ & $0.91 \pm 0.03$ \\  banglalink & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & $2.55 \pm 0.04$ & $0.55 \pm 0.02$ \\ \end{tabular} \caption{Security scores for operators} \label{tab:sec_operator} \end{table*}}
 \newcommand{\daSecScoreBestoperator}{O2 uk}
-\newcommand{\daSecScoreBestoperatorScoreNominal}{3.85}
-\newcommand{\daSecScoreBestoperatorScore}{$3.85 \pm 0.02$}
-\newcommand{\daSecScoreBestoperatorNumFullVersions}{96}
+\newcommand{\daSecScoreBestoperatorScoreNominal}{3.87}
+\newcommand{\daSecScoreBestoperatorScore}{$3.87 \pm 0.0$}
+\newcommand{\daSecScoreBestoperatorNumFullVersions}{100}
 \newcommand{\daSecScoreWorstoperator}{banglalink}
-\newcommand{\daSecScoreWorstoperatorScoreNominal}{0.545}
-\newcommand{\daSecScoreWorstoperatorScore}{$0.545 \pm 0.018$}
+\newcommand{\daSecScoreWorstoperatorScoreNominal}{0.548}
+\newcommand{\daSecScoreWorstoperatorScore}{$0.548 \pm 0.018$}
 \newcommand{\daSecScoreWorstoperatorNumFullVersions}{73}
-\newcommand{\daTabDLDistances}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & summary \\ \hline equal & 0.0 & 0.0 & 0.0714 & 0.0 \\weight $m$ & 0.222 & 0.333 & 0.143 & 0.0 \\weight $u$ & 0.444 & 0.133 & 0.214 & 0.0 \\$f$ & 0.444 & 0.733 & 0.5 & 0.0 \\$m$ & 0.667 & 0.733 & 0.357 & 0.5 \\$u$ & 0.778 & 0.733 & 0.714 & 0.0 \\\end{tabular} \caption{Normalised Damerau-Levenshtein distances for different metrics} \label{tab:dl_distances} \end{table}}
-\newcommand{\daTabChangeInScores}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & summary \\ \hline $m$ & $-1.9 \pm 2.2$ & $-0.664 \pm 2.51$ & $-3.27 \pm 1.39$ & $-3.38 \pm 1.83$ \\weight $m$ & $-0.294 \pm 0.275$ & $-0.101 \pm 0.316$ & $-0.458 \pm 0.177$ & $-0.487 \pm 0.209$ \\equal & $-0.116 \pm 0.064$ & $-0.0386 \pm 0.113$ & $-0.146 \pm 0.048$ & $-0.166 \pm 0.029$ \\weight $u$ & $-0.0522 \pm 0.126$ & $-0.0148 \pm 0.238$ & $0.0206 \pm 0.0797$ & $-0.00995 \pm 0.123$ \\$u$ & $0.517 \pm 1.61$ & $0.2 \pm 2.1$ & $1.52 \pm 0.96$ & $1.39 \pm 1.48$ \\$f$ & $1.04 \pm 0.58$ & $0.348 \pm 1.01$ & $1.31 \pm 0.43$ & $1.49 \pm 0.26$ \\\end{tabular} \caption{Mean change in scores for different metrics} \label{tab:change_in_scores} \end{table*}}
-\newcommand{\daGPAPICurveFitParamFirst}{80.7}
-\newcommand{\daGPAPICurveFitParamSecond}{\num{0.00259}}
-\newcommand{\daGPAPICurveFitRMSE}{0.18}
-\newcommand{\daGPAPICurvePolyRMSE}{0.18}
-\newcommand{\daGPAPICurveSplineRMSE}{0.18}
-\newcommand{\daGPAPICurveHalfDeployedYearsNominal}{0.953}
-\newcommand{\daGPAPICurveHalfDeployedYears}{$0.953 \pm 0.472$}
-\newcommand{\daGPAPICurveHalfDeployedNominal}{348}
-\newcommand{\daGPAPICurveHalfDeployed}{$348 \pm 172$}
-\newcommand{\daGPAPICurveFullDeployedYearsNominal}{3.39}
-\newcommand{\daGPAPICurveFullDeployedYears}{$3.39 \pm 78$}
+\newcommand{\daTabDLDistances}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & summary \\ \hline equal & 0.0 & 0.0667 & 0.0 & 0.0 \\weight $m$ & 0.2 & 0.4 & 0.0714 & 0.0 \\$f$ & 0.4 & 0.733 & 0.5 & 0.0 \\weight $u$ & 0.4 & 0.0667 & 0.143 & 0.0 \\$m$ & 0.6 & 0.733 & 0.357 & 0.5 \\$u$ & 0.9 & 0.667 & 0.786 & 0.0 \\\end{tabular} \caption{Normalised Damerau-Levenshtein distances for different metrics} \label{tab:dl_distances} \end{table}}
+\newcommand{\daTabChangeInScores}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & summary \\ \hline $m$ & $-1.67 \pm 2.25$ & $-0.66 \pm 2.54$ & $-3.3 \pm 1.3$ & $-3.4 \pm 1.8$ \\weight $m$ & $-0.264 \pm 0.276$ & $-0.0991 \pm 0.32$ & $-0.461 \pm 0.167$ & $-0.488 \pm 0.216$ \\equal & $-0.108 \pm 0.062$ & $-0.0368 \pm 0.112$ & $-0.146 \pm 0.045$ & $-0.165 \pm 0.033$ \\weight $u$ & $-0.0609 \pm 0.122$ & $-0.0113 \pm 0.232$ & $0.024 \pm 0.081$ & $-0.0061 \pm 0.12$ \\$u$ & $0.366 \pm 1.61$ & $0.219 \pm 2.12$ & $1.55 \pm 0.94$ & $1.42 \pm 1.48$ \\$f$ & $0.976 \pm 0.564$ & $0.331 \pm 1.01$ & $1.31 \pm 0.40$ & $1.48 \pm 0.29$ \\\end{tabular} \caption{Mean change in scores for different metrics} \label{tab:change_in_scores} \end{table*}}
+\newcommand{\daGPAPICurveFitParamFirst}{80.8}
+\newcommand{\daGPAPICurveFitParamSecond}{\num{0.00258}}
+\newcommand{\daGPAPICurveFitRMSE}{0.179}
+\newcommand{\daGPAPICurvePolyRMSE}{0.178}
+\newcommand{\daGPAPICurveSplineRMSE}{0.178}
+\newcommand{\daGPAPICurveHalfDeployedYearsNominal}{0.957}
+\newcommand{\daGPAPICurveHalfDeployedYears}{$0.957 \pm 0.469$}
+\newcommand{\daGPAPICurveHalfDeployedNominal}{349}
+\newcommand{\daGPAPICurveHalfDeployed}{$349 \pm 171$}
+\newcommand{\daGPAPICurveFullDeployedYearsNominal}{3.4}
+\newcommand{\daGPAPICurveFullDeployedYears}{$3.4 \pm 7$}
 \newcommand{\daGPAPICurveFullDeployedNominal}{1\,240}
-\newcommand{\daGPAPICurveFullDeployed}{$1\,240 \pm 286\,000$}
-\newcommand{\daGPAPIPerAPIMeanNominal}{0.00459}
-\newcommand{\daGPAPIPerAPIMean}{$0.00459 \pm 0.00346$}
-\newcommand{\daGPAPIPerAPIStdevNominal}{0.0212}
-\newcommand{\daGPAPIPerAPIStdev}{$0.0212 \pm 0.0127$}
+\newcommand{\daGPAPICurveFullDeployed}{$1\,240 \pm 287\,000$}
+\newcommand{\daGPAPIPerAPIMeanNominal}{0.00452}
+\newcommand{\daGPAPIPerAPIMean}{$0.00452 \pm 0.00341$}
+\newcommand{\daGPAPIPerAPIStdevNominal}{0.0209}
+\newcommand{\daGPAPIPerAPIStdev}{$0.0209 \pm 0.0125$}
 \newcommand{\daGPAPIPerAPIAbsMeanNominal}{0.0145}
-\newcommand{\daGPAPIPerAPIAbsMean}{$0.0145 \pm 0.0066$}
-\newcommand{\daGPAPIPerAPIAbsStdevNominal}{0.0154}
-\newcommand{\daGPAPIPerAPIAbsStdev}{$0.0154 \pm 0.012$}
-\newcommand{\daGPAPIPerAPIParametersTable}{\begin{table} \centering \begin{tabular}{c|S|S} API Version & ${t_0}$ & $\mathrm{decay}$ \\ \hline2	&	0.0	&	0.0237\\3	&	0.0	&	0.0235\\4	&	0.0	&	0.00689\\5	&	59.1	&	0.00465\\6	&	2.4	&	0.00434\\7	&	0.0	&	0.00499\\8	&	47.7	&	0.00415\\9	&	155	&	0.00365\\10	&	90.2	&	0.00364\\11	&	432	&	0.00228\\12	&	356	&	0.00228\\13	&	304	&	0.00236\\14	&	219	&	0.00239\\15	&	160	&	0.0024\\16	&	142	&	0.00227\\17	&	289	&	0.00195\\18	&	120	&	0.00142\\19	&	120	&	0.00154\\\end{tabular} \caption{Parameters for different API versions} \label{tab:per_api_parameters} \end{table}}
-\newcommand{\daGPAPISeventeenLaterDate}{2015-05-04}
-\newcommand{\daGPAPISeventeenLaterProportion}{73.1\%}
-\newcommand{\daGPAPISeventeenEarlierProportion}{26.9\%}
+\newcommand{\daGPAPIPerAPIAbsMean}{$0.0145 \pm 0.0067$}
+\newcommand{\daGPAPIPerAPIAbsStdevNominal}{0.0152}
+\newcommand{\daGPAPIPerAPIAbsStdev}{$0.0152 \pm 0.0117$}
+\newcommand{\daGPAPIPerAPIParametersTable}{\begin{table} \centering \begin{tabular}{c|S|S} API Version & ${t_0}$ & $\mathrm{decay}$ \\ \hline2	&	0.0	&	0.0237\\3	&	0.0	&	0.0235\\4	&	0.0	&	0.00689\\5	&	36.3	&	0.00429\\6	&	0.0	&	0.0043\\7	&	0.0	&	0.00499\\8	&	47.7	&	0.00415\\9	&	155	&	0.00365\\10	&	90.3	&	0.00364\\11	&	444	&	0.00235\\12	&	356	&	0.00229\\13	&	304	&	0.00236\\14	&	219	&	0.00239\\15	&	161	&	0.0024\\16	&	142	&	0.00228\\17	&	290	&	0.00197\\18	&	121	&	0.00143\\19	&	120	&	0.00155\\\end{tabular} \caption{Parameters for different API versions} \label{tab:per_api_parameters} \end{table}}
+\newcommand{\daGPAPISeventeenLaterDate}{2015-06-01}
+\newcommand{\daGPAPISeventeenLaterProportion}{74.3\%}
+\newcommand{\daGPAPISeventeenEarlierProportion}{25.7\%}
 \newcommand{\daAPISeventeenReleaseDate}{2012-10-29}
-\newcommand{\daGPAPISeventeenFullDeployment}{2017-10-24}
-\newcommand{\daGPAPISeventeenCurveFitParamFirst}{289}
-\newcommand{\daGPAPISeventeenCurveFitParamSecond}{\num{0.00195}}
-\newcommand{\daGPAPISeventeenCurveFitRMSE}{0.0275}
-\newcommand{\daGPAPISeventeenCurvePolyRMSE}{0.0088}
-\newcommand{\daGPAPISeventeenCurveSplineRMSE}{0.0088}
+\newcommand{\daGPAPISeventeenFullDeployment}{2017-10-16}
+\newcommand{\daGPAPISeventeenCurveFitParamFirst}{290}
+\newcommand{\daGPAPISeventeenCurveFitParamSecond}{\num{0.00197}}
+\newcommand{\daGPAPISeventeenCurveFitRMSE}{0.0273}
+\newcommand{\daGPAPISeventeenCurvePolyRMSE}{0.00867}
+\newcommand{\daGPAPISeventeenCurveSplineRMSE}{0.00867}
 \newcommand{\daGPAPISeventeenCurveHalfDeployedYearsNominal}{1.76}
 \newcommand{\daGPAPISeventeenCurveHalfDeployedYears}{$1.76 \pm 0.07$}
-\newcommand{\daGPAPISeventeenCurveHalfDeployedNominal}{643}
-\newcommand{\daGPAPISeventeenCurveHalfDeployed}{$643 \pm 28$}
-\newcommand{\daGPAPISeventeenCurveFullDeployedYearsNominal}{4.99}
-\newcommand{\daGPAPISeventeenCurveFullDeployedYears}{$4.99 \pm 1.12$}
-\newcommand{\daGPAPISeventeenCurveFullDeployedNominal}{1\,820}
-\newcommand{\daGPAPISeventeenCurveFullDeployed}{$1\,820 \pm 408$}
-\newcommand{\daNumUpdatesUpgrades}{3\,590}
-\newcommand{\daUpdatesPerMonthPerVersionNominal}{2.83}
-\newcommand{\daUpdatesPerMonthPerVersion}{$2.83 \pm 0.04$}
+\newcommand{\daGPAPISeventeenCurveHalfDeployedNominal}{642}
+\newcommand{\daGPAPISeventeenCurveHalfDeployed}{$642 \pm 28$}
+\newcommand{\daGPAPISeventeenCurveFullDeployedYearsNominal}{4.97}
+\newcommand{\daGPAPISeventeenCurveFullDeployedYears}{$4.97 \pm 1.1$}
+\newcommand{\daGPAPISeventeenCurveFullDeployedNominal}{1\,810}
+\newcommand{\daGPAPISeventeenCurveFullDeployed}{$1\,810 \pm 401$}
+\newcommand{\daNumUpdatesUpgrades}{3\,660}
+\newcommand{\daUpdatesPerMonthPerVersionNominal}{2.67}
+\newcommand{\daUpdatesPerMonthPerVersion}{$2.67 \pm 0.04$}
 \newcommand{\daNumUpdatesSkippedBig}{3}
-\newcommand{\daNumUpdatesBigUpgrades}{820}
-\newcommand{\daNumUpdatesDowngrades}{152}
-\newcommand{\daNumUpdateFullOnly}{670}
+\newcommand{\daNumUpdatesBigUpgrades}{833}
+\newcommand{\daNumUpdatesDowngrades}{146}
+\newcommand{\daNumUpdateFullOnly}{674}
 \newcommand{\daPercBigUpgradesNominal}{18.6\%}
 \newcommand{\daPercBigUpgrades}{$18.6 \pm 0.7\%$}
-\newcommand{\daPercUpdatesDowngradesNominal}{3.45\%}
-\newcommand{\daPercUpdatesDowngrades}{$3.45 \pm 0.28\%$}
-\newcommand{\daNumDataPoints}{158 Billion}
-\newcommand{\daMonthsDevices}{1\,790}
+\newcommand{\daPercUpdatesDowngradesNominal}{3.26\%}
+\newcommand{\daPercUpdatesDowngrades}{$3.26 \pm 0.27\%$}
+\newcommand{\daNumDataPoints}{184 Billion}
+\newcommand{\daMonthsDevices}{2\,110}
 \newcommand{\daMonths}{6}
@@
 \newcommand{\daAdbEnabledPerc}{$19.2 \pm 0.0\%$}
-\newcommand{\daNumStartedApps}{126\,000}
-\newcommand{\daNumInstalledApps}{180\,000}
+\newcommand{\daNumStartedApps}{128\,000}
+\newcommand{\daNumInstalledApps}{182\,000}
 \newcommand{\daAPISeventeenReleaseDateMonth}{October 2012}
 \newcommand{\daGPAPISeventeenFullDeploymentMonth}{October 2017}
-\newcommand{\daMeanOutstandingVulnerabilitiesNominal}{0.547}
-\newcommand{\daCurrentMeanOutstandingVulnerabilitiesNominal}{0.193}
+\newcommand{\daMeanOutstandingVulnerabilitiesNominal}{0.531}
+\newcommand{\daCurrentMeanOutstandingVulnerabilitiesNominal}{0.124}
 \newcommand{\daModelHalfDeploymentYears}{0.0822}
 \newcommand{\daModelHalfDeployment}{30}
-\newcommand{\daModelFullDeploymentYears}{0.89}
-\newcommand{\daModelFullDeployment}{325}
+\newcommand{\daModelFullDeploymentYears}{0.888}
+\newcommand{\daModelFullDeployment}{324}
 \newcommand{\daMeanOutstandingProbability}{1.0\%}
 \newcommand{\daMeanOutstandingUncertaintyEquation}{$0.99^n$}
+\newcommand{\daGPAPISeventeenLaterDateMonth}{June 2015}

### 2015-07-23 security_scoring.tex
A couple of minor fixes from reviewers

@@
 \daTabSecScoresmanufacturer
-Different device manufacturers have very different scores, Table~\ref{tab:sec_manufacturer} shows the scores for the \daNumSigManufacturers\ device manufacturers with a significant presence in our data with \emph{\daSecScoreBestmanufacturer} (\daSecScoreBestmanufacturerScore\ out of 10) scoring highest and \emph{\daSecScoreWorstmanufacturer} (\daSecScoreWorstmanufacturerScore\ out of 10) scoring lowest.
+Different device manufacturers have very different scores; Table~\ref{tab:sec_manufacturer} shows the scores for the \daNumSigManufacturers\ device manufacturers with a significant presence in our data with \emph{\daSecScoreBestmanufacturer} (\daSecScoreBestmanufacturerScore\ out of 10) scoring highest and \emph{\daSecScoreWorstmanufacturer} (\daSecScoreWorstmanufacturerScore\ out of 10) scoring lowest.
 Device manufacturers are considered significant if we have data from at least \daSigNumDevices\ devices and at least \daSigNumDeviceDays\ days of contributions.

### 2015-07-23 spsm04-thomas.tex
A couple of minor fixes from reviewers
Add some TODOs
Make Sherif and Rip's names links in Acknowledgements so that people can find them for the Rwanda data

@@
 \usepackage{siunitx}
-\usepackage[disable]{todonotes}%
+\usepackage{todonotes}%[disable]
 \let\OldTodo\todo
@@
 We used \da\ to record data on the OS versions of Android installed on devices.
-Nappa et.\ al.
+Nappa et al.
 performed a similar analysis using WINE to record data on installed versions of Windows client programs and used cleaned NVD data rather than \avo\ for vulnerability data, they did not assign client applications scores~\cite{Nappa2015}.
@@
 The Rwanda and FTSE data cannot be made available due to the terms under which they were obtained.
+\todo{We need to submit this data to DSpace and insert a reference.}
 
@@
 This work was supported by a Google focussed research award; and the EPSRC [grant number EP/P505445/1].
-Thanks to Sherif Akoush and Ripduman Sohan for supplying the Rwanda data and to staff at the FTSE 100 company for supplying their mobile device management data.
+Thanks to \href{https://www.cl.cam.ac.uk/~sa497/}{Sherif Akoush} and \href{http://orcid.org/0000-0003-0740-8650}{Ripduman Sohan} for supplying the Rwanda data and to staff at the FTSE 100 company for supplying their mobile device management data.
 Thanks to our anonymous reviewers for their excellent insightful comments.
@@
 \printbibliography
+\todo{We need to fix the columns at the end to be the same length}
+\todo{Check all the figures that could work on grayscale do}
+\todo{Make a thumbnail from figure 5}
 

### 2015-07-23 security_scoring.tex
Reorder abstract to emphasise metrics over particular results

@@
 Therefore, our score is secure against passive gaming attacks which change the measured distribution, but would require active defence against active gaming attacks, which target the measurement devices.
+
+\subsection{Utilitarianism}
+From a utilitarian standpoint, while small manufacturers like Symphony and Walton do badly on our scores, they do not have as many customers as higher scoring manufacturers.
+Hence the total risk to users from the higher scoring popular manufacturers is higher than the risk from the lower scoring unpopular manufacturers.
+We could normalise the scores for market penetration and so give a score reflecting the risk posed by that manufacturer's performance, which would tend to decrease the difference between manufacturers in our current scoring.
+However we have only considered vulnerabilities affecting all Android devices, not those which are specific to manufacturers and so the aggregation point is all Android, not a specific manufacturer.\todo{Does that even make sense?}
+Additionally since our scores are provided so that customers can chose which devices to buy then it is the marginal risk to that individual of that device which is of interest rather than the aggregate risk to all users.

### 2015-07-23 security_scoring.tex
Deployment to at least 1% is reasonable but arbitrary

@@
 So the uncertainty in each vulnerability truly being outstanding is the probability of not having observed a fixed device if a fixed device existed.
-We assume that if the fix has been released then at least \daMeanOutstandingProbability\ of devices have the fix.
+We assume that if the fix has been released then at least \daMeanOutstandingProbability\ of devices have the fix.\footnote{The selection of \daMeanOutstandingProbability\ is arbitrary, but if a version has only been deployed to a smaller proportion then it does not have a substantial penetration.}
 This gives an uncertainty of \daMeanOutstandingUncertaintyEquation\ where $n$ is the number of devices contributing to that day's data for each vulnerability outstanding each day.

### 2015-07-23 security_scoring.tex
Add a link for an anecdote

@@
 However there are a wide variety of scores depending on the source of the device.
-There is anecdotal evidence that Google's Nexus devices are better at getting updates than other Android devices because Google makes the original updates and ships them to its devices (but the plural of anecdote is not data).
+There is anecdotal evidence that Google's Nexus devices are better at getting updates than other Android devices because Google makes the original updates and ships them to its devices\footnote{\url{http://www.howtogeek.com/139391/htg-explains-why-android-geeks-buy-nexus-devices/}} (but the plural of anecdote is not data).
 Table~\ref{tab:sec_summary} shows that this is the case with Nexus devices getting much better scores than non-Nexus devices and Figure~\ref{fig:security_score_summary} shows this holds true over time.

### 2015-07-23 security_scoring.tex
How big a sample would we need for an added value analysis

@@
 However our sample size is too small to do that because while we have significant numbers of devices for each of the device models (Table~\ref{tab:sec_model}) and for each of the network operators (Table~\ref{tab:sec_operator}), we would need a significant number of each model in each network operator.
+A large enough sample size would contain at least 100 unique devices each day for each device model for each network operator, hence for our 14 significant models and 13 significant network operators, a daily average of at least 18\,000 unique devices would be needed, for a period of at least a year.
+Since the distribution of devices is unlikely to be uniformly distributed across device models and network operators a more realistic estimate is 100\,000 unique devices each day.
+This is not an unobtainable number but it is two orders of magnitude more than is available in \da.
 %We do not attempt to disambiguate users behaviour in whether they install updates from network operators using rolling upgrades.

### 2015-07-23 spsm04-thomas.tex
Reorder abstract to emphasise metrics over particular results
Be more explicit about what latency means

@@
 The security of Android depends on security updates for critical vulnerabilities being shipped to devices.
+We define a security metric to rank the performance of device manufacturers and network operators, based on their provision of updates and exposure to critical vulnerabilities.
 Using a corpus of \daNumOSDevices\ devices we show that there is significant variability in the timely delivery of security updates across different device manufacturers and network operators.
-This is the first paper to quantify security update performance of device manufacturers and network operators.
-We define a security metric to rank the performance of device manufacturers and network operators.
 This provides a comparison point for purchasers to determine which device manufacturers and network operators to use.
 We argue that publishing metrics like this provides an important incentive to manufacturers and operators to ship security updates.
+This is the first paper to quantify security update performance of device manufacturers and network operators.
 \end{abstract}
@@
 \caption{Flow of updates from upstream projects into Android.
-Number of updates as in Figure~\ref{fig:update_ecosystem}, latency in days for all pairs of versions we have data on.\todolater{scrape the other 26 websites...
+Number of updates as in Figure~\ref{fig:update_ecosystem}, latency in days between the upstream release and the release of the first Android version containing it, for all pairs of versions we have data on.\todolater{scrape the other 26 websites...
 is it worth it?}}
@@
 %An analysis by Vidas et al.~\cite{Vidas2011} of the Android 2.1 to 2.2 update found that it took 11 months from when Google released 2.2 for the last device which they were investigating to get the update.
-To compute the latency between upstream releases and their inclusion in Android we scraped the release pages for those projects, to obtain the version numbers and release dates.
+To compute the latency between upstream releases and the release of the first version of Android containing that release we scraped the release pages for those projects, to obtain the version numbers and release dates.
 This allows us to compute the latency between an upstream project being released and it being included in Android, this is shown in Table~\ref{tab:update_ecosystem}.

### 2015-07-23 spsm04-thomas.tex
Proposed wording from reviewer 3

@@
 The code for each Android release or update is kept secret until after a binary release has been published.\footnote{\url{https://source.android.com/source/code-lines.html}}
-Device manufacturers likely receive advanced access in order to prepare handsets so they can customise it before passing it on to the network operator.
+Device manufacturers are presumed to receive advanced access in order to prepare handsets so they can customise it before passing it on to the network operator.
 The network operator may then make or request further customisations and perform further testing before shipping the update to the device.

### 2015-07-23 security_scoring.tex
Add todo from reviewer 3
Discuss active attacks

@@
 \todo{Sensitivity of time based vs. global scores}
+\todo{Reviewer 3 thinks we should use the Spearman rank correlation coefficient instead, and they might be right}
 
@@
 A high value of $m$ could be achieved by focusing on only one device at a time and ensuring that it gets updates but ignoring all others, but that would lower $f$ and $u$.
-One way to influence our scores would be to add additional devices to Device Analyzer, which have good security, these would have to be real end user devices since we could detect fake ones.
+One way to influence our scores would be to add additional devices to Device Analyzer, which have good security, these would have to be real end user devices since we could detect fake ones if they deviated from the behaviour of real devices in \da.
 This would increase the size of our data set and would require providing genuinely good security to some users.
+Some active attacks like blocking access to the \da\ servers from the mobile data network would not be effective as \da\ would retry on Wi-Fi, others like denial of service attacks on the \da\ servers might be effective but illegal, however some entities might be able to force the uninstallation of the app from all devices.
 Therefore, our score is secure against passive gaming attacks which change the measured distribution, but would require active defence against active gaming attacks, which target the measurement devices.

### 2015-07-23 spsm04-thomas.tex
Remove display hack which is no longer needed as the text has got longer
Clarify what is meant by maybe secure

@@
 A device is insecure if it is running a vulnerable version of Android and the device has not received an update which might fix it;
-it is maybe secure if it is running a vulnerable version and has received an update which might fix the vulnerability;
+it is maybe secure if it is running a vulnerable version and but received an update which could have fixed the vulnerability if it contained a backported fix;
 and it is secure if it is running a secure version.
@@
 
-\vspace{0.1em}\todo{Remove this display hack}
 \section{Threat model and attack vectors}

### 2015-07-23 spsm04-thomas.tex
Clarify what is meant by the Android sandbox being weak
Images mostly work in grayscale, remove TODO

@@
 The best place to prevent attacks is at runtime, since all three attack vectors can be prevented at this level. 
-Unfortunately, as Figure~\ref{fig:proportioninsecure} shows, the sandbox for Android apps is weak.
+Unfortunately, as Figure~\ref{fig:proportioninsecure} shows, the sandbox for Android apps is weak, since on most devices there are known critical vulnerabilities in it.
 
@@
 \todo{We need to fix the columns at the end to be the same length}
-\todo{Check all the figures that could work on grayscale do}
 \todo{Make a thumbnail from figure 5}

### 2015-07-23 security_scoring.tex
Don't blame the users

@@
 There is a variation between device models, with the update being distributed to most devices quickly and others having a much slower roll out, but since some do update quickly the bottleneck is unlikely to be with the user.
+Perhaps some device models are preferred by user who are more likely to install updates than others, however we do observe updates being rolled out to device models quickly and user behaviour is not beyond the control of the device manufacturer.
+They could make updates be installed automatically or pester the user into installing them and at least some of them do pester, silent automatic updates do boost uptake~\cite{Duebendorfer2009}.
 

### 2015-07-23 spsm04-thomas.tex
Add footnote early on explaining what a critical vulnerability is

@@
 
-It has been previously shown that Android devices do not receive security updates promptly when critical vulnerabilities are publicly disclosed and hence, on average, \daMeanInsecurityPercNominal\ are exposed to known critical vulnerabilities~\cite{androidvulnerabilities.org}.
+It has been previously shown that Android devices do not receive security updates promptly when critical vulnerabilities\footnote{If malicious code exploits a critical vulnerability then it gains control of the device.} are publicly disclosed and hence, on average, \daMeanInsecurityPercNominal\ are exposed to known critical vulnerabilities~\cite{androidvulnerabilities.org}.
 Here we attempt to determine why by understanding the Android update process, evaluating different entities in the ecosystem and producing a scoring metric to compare different entities.

### 2015-07-23 spsm04-thomas.tex
Keywords should be semicolon separated
Note the possibility of interpreters in threat model (R3)

@@
 
-\keywords{Android, updates, vulnerabilities, metrics, ecosystems}
+\keywords{Android; updates; vulnerabilities; metrics; ecosystems}
 
@@
 The \emph{dynamic code loading} attack vector is to make an existing app download and execute code at runtime.
-The most direct method is to upload to a marketplace a seemingly innocent app that dynamically loads malicious code.
+The most direct method is to upload to a marketplace a seemingly innocent app that dynamically loads malicious code, either as additional davik bytecode, as a native library or by embedding an interpreter and executing received instructions.
 Neither static nor dynamic analysis of this app will uncover any malicious code, since it does not exist in the app.
-The marketplace can detect explicit use of dynamic code loading, however dynamic code loading is not actually needed for this to work -- it just makes the task easier -- even on a platform such as iOS, which does not permit dynamic code loading, a Return-Oriented Programming (ROP) attack is relatively easy if the attacker creates an app with carefully crafted flaws~\cite{Wang2013a}.
+The marketplace can try to detect explicit use of dynamic code loading, however there are ways to dynamically load code which are hard to detect even on a platform such as iOS, which does not permit dynamic code loading, a Return-Oriented Programming (ROP) attack is relatively easy if the attacker creates an app with carefully crafted flaws~\cite{Wang2013a}.
 

### 2015-07-23 security_scoring.tex
Move the point where we declare the operator score table so that it ends up on the same page as the other tables

@@
 \daTabSecScoresmodel
+\daTabSecScoresoperator
 Even within device manufacturers, different models can have very different update behaviours and hence security.
@@
 
-\daTabSecScoresoperator
 We also analysed different network operators, for the \daNumSigOperators\ network operators with a significant presence in our data.

### 2015-07-23 spsm04-thomas.tex
Talk about where Android version numbers come from

@@
 
+\section{Versions of Android running on devices}
+
+In this analysis we use historical data collected by the \da\ project~\cite{Wagner2013}.
+\da\ collects data\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/collected.html}} from study participants who install the Android app from Google Play.
+Most study participants allow researchers around the world to access a subset of their device data, including the data presented here.
+
+We extracted the build string and API version for each device each day.
+The build string is a user-readable version string.
+The API version is a positive integer that increases when new features are added to the API.
+Consequently security (bug) fixes do not result in a change in the API version.
+Fortunately most (\daOSVersionPercValidLines) entries in these data have a build string of the form `x.y.z opaque\_marker' and so it is possible to extract the Android version number `x.y.z'.
+On a large proportion of devices `opaque\_marker' is a well defined build number\footnote{\url{https://source.android.com/source/build-numbers.html}} however different device manufacturers use different schema.
+Google provides API version distribution information but not the OS and build version information we need.
+We verify that the \da\ data is representative in~\S\ref{sec:representative}.
+
+\da\ has collected data from \daNumDevices\ devices with a total of \daDeviceDays\footnote{Here we are only counting devices and days for which we have valid OS version data.} device days.
+The majority of devices only contribute data for a short period of time, however \daMonthsDevices\ devices have contributed data for more than \daMonths~months.
+
 \section{Critical vulnerabilities}

### 2015-07-23 spsm04-thomas.tex
Add a todolater for somethign we don't have time for now (quantifying overestimation and underestimation)

@@
 This shows that the \da\ data systematically overestimates the prevalence of new API versions and underestimates the prevalence of old API versions.
-This means that the OS version information from \da\ is likely to be overestimating the prevalence of new OS versions and hence our results are likely to be conservative.
+This means that the OS version information from \da\ is likely to be overestimating the prevalence of new OS versions and hence our results are likely to be conservative.\todolater{Provide some numbers which quantify this}
 

### 2015-07-23 security_scoring.tex
Put all the security score tables together so that they can't end up on different pages and hence waste space

@@
 This gives a security score of \daSecurityScore\ out of 10.
-\daTabSecScoressummary
+
 However there are a wide variety of scores depending on the source of the device.
 There is anecdotal evidence that Google's Nexus devices are better at getting updates than other Android devices because Google makes the original updates and ships them to its devices\footnote{\url{http://www.howtogeek.com/139391/htg-explains-why-android-geeks-buy-nexus-devices/}} (but the plural of anecdote is not data).
+
 Table~\ref{tab:sec_summary} shows that this is the case with Nexus devices getting much better scores than non-Nexus devices and Figure~\ref{fig:security_score_summary} shows this holds true over time.
-\daTabSecScoresmanufacturer
 Different device manufacturers have very different scores; Table~\ref{tab:sec_manufacturer} shows the scores for the \daNumSigManufacturers\ device manufacturers with a significant presence in our data with \emph{\daSecScoreBestmanufacturer} (\daSecScoreBestmanufacturerScore\ out of 10) scoring highest and \emph{\daSecScoreWorstmanufacturer} (\daSecScoreWorstmanufacturerScore\ out of 10) scoring lowest.
@@
 
+\daTabSecScoressummary
+\daTabSecScoresmanufacturer
 \daTabSecScoresmodel
 \daTabSecScoresoperator
+
 Even within device manufacturers, different models can have very different update behaviours and hence security.

### 2015-07-23 spsm04-thomas.tex
We show the Android insecurity percentage in this paper now

@@
 \newcommand{\blindauthors}[1]{#1}%{Paper \#XX}%
+
+% Evil hackery to provide an optional argument: https://tex.stackexchange.com/questions/308/different-command-definitions-with-and-without-optional-argument/314#314
+\makeatletter
+ \def\avovuln{\@ifnextchar[{\@avovulnsspecific}{\@avovulngeneral}}
+ \def\@avovulnsspecific[#1]#2{\emph{\href{http://androidvulnerabilities.org/vulnerabilities/#1}{#2}}}
+ \def\@avovulngeneral#1{\emph{\href{http://androidvulnerabilities.org/vulnerabilities/#1}{#1}}}
+\makeatother
 
@@
 
-It has been previously shown that Android devices do not receive security updates promptly when critical vulnerabilities\footnote{If malicious code exploits a critical vulnerability then it gains control of the device.} are publicly disclosed and hence, on average, \daMeanInsecurityPercNominal\ are exposed to known critical vulnerabilities~\cite{androidvulnerabilities.org}.
-Here we attempt to determine why by understanding the Android update process, evaluating different entities in the ecosystem and producing a scoring metric to compare different entities.
+We show that Android devices do not receive security updates promptly when critical vulnerabilities\footnote{If malicious code exploits a critical vulnerability then it gains control of the device.} are publicly disclosed and hence, on average, \daMeanInsecurityPercNominal\ are exposed to known critical vulnerabilities.
+We attempt to determine why by understanding the Android update process, evaluating different entities in the ecosystem and producing a scoring metric to compare different entities.
 
@@
 
+\section{Vulnerability of Android dev\-ices}
+Combining the data on critical vulnerabilities and versions of Android allows us to plot Figure~\ref{fig:proportioninsecure}, initially all devices are \emph{maybe secure} (yellow) since \da\ does not have historical data prior to May 2011.
+This means we cannot distinguish between devices which are running a version of Android which is known to be vulnerable from one which may have received a backported fix.
+This demonstrates the importance of a longitudinal study: this type of analysis requires years of data.
+Once \avovuln{zergRush} was discovered in October 2011 then most devices are recorded as \emph{insecure} (red) as they were exposed to that vulnerability.
+The remaining devices were already running a version of Android which fixed the \avovuln{zergRush} vulnerability and are therefore marked as \emph{secure} (green).
+From October 2011 until the discovery of \avovuln[APK_duplicate_file]{APK duplicate file} in February 2013 the graph shows progressive improvement as devices are upgraded.
+This means more and more devices are marked as \emph{secure} because they are now running a secure version of Android, or marked as \emph{maybe secure} because they received an OS update that did not update to a known-good version of Android but which may still have included a backport of a fix, as the update was made available after the vulnerability was disclosed.
+From February 2013 onwards regular discovery of critical vulnerabilities ensures that most devices are exposed to known critical vulnerabilities.
+This gives us an average exposure to critical vulnerabilities of Android devices of \daMeanInsecurityPerc.
+
 \input{security_scoring}

### 2015-07-24 security_scoring.tex
We have bridged the gaps
Minor fixes on reading through

@@
 Since $f$ is the most important metric we weight it more highly.
-Since $m$ is an unbounded positive real number, we map it into the range [0--1).
+Since $m$ is an unbounded positive real number, we map it into the range (0--1].
 This gives us Equation~\ref{eq:score}:
@@
 There is anecdotal evidence that Google's Nexus devices are better at getting updates than other Android devices because Google makes the original updates and ships them to its devices\footnote{\url{http://www.howtogeek.com/139391/htg-explains-why-android-geeks-buy-nexus-devices/}} (but the plural of anecdote is not data).
+Table~\ref{tab:sec_summary} shows that this is the case with Nexus devices getting much better scores than non-Nexus devices and Figure~\ref{fig:security_score_summary} shows this holds true over time.
 
-Table~\ref{tab:sec_summary} shows that this is the case with Nexus devices getting much better scores than non-Nexus devices and Figure~\ref{fig:security_score_summary} shows this holds true over time.
 Different device manufacturers have very different scores; Table~\ref{tab:sec_manufacturer} shows the scores for the \daNumSigManufacturers\ device manufacturers with a significant presence in our data with \emph{\daSecScoreBestmanufacturer} (\daSecScoreBestmanufacturerScore\ out of 10) scoring highest and \emph{\daSecScoreWorstmanufacturer} (\daSecScoreWorstmanufacturerScore\ out of 10) scoring lowest.
@@
 There is a variation between device models, with the update being distributed to most devices quickly and others having a much slower roll out, but since some do update quickly the bottleneck is unlikely to be with the user.
-Perhaps some device models are preferred by user who are more likely to install updates than others, however we do observe updates being rolled out to device models quickly and user behaviour is not beyond the control of the device manufacturer.
-They could make updates be installed automatically or pester the user into installing them and at least some of them do pester, silent automatic updates do boost uptake~\cite{Duebendorfer2009}.
+Perhaps some device models are preferred by users who are more likely to install updates than others, however we do observe updates being rolled out to device models quickly and user behaviour is not beyond the control of the device manufacturer.
+They could install updates automatically or pester the user into installing them, and at least some of them do pester, silent automatic updates do boost uptake~\cite{Duebendorfer2009}.
 
@@
 The scoring metric as originally computed, is averaged over the whole history of the device manufacturer, device model or network operator, it gives equal weight to both periods years ago and to the last few months.
-If instead we take an exponential moving average of the daily score for days with more than \daSigNumDevicesDay\, devices when there have been at least consecutive \daSigNumDays\, days of data with that many devices then we can plot how this score has changed over time.
+If instead we take an exponential moving average of the daily score for days with more than \daSigNumDevicesDay\, devices when there have been at least \daSigNumDays\ consecutive days of data with that many devices then we can plot how this score has changed over time.
 Equation~\ref{eq:rolling_update} shows how the value for a particular day ($v_i$) is computed from the previous day's value and the input for the current day ($n$) with an $\alpha$ of $1/\daSigNumDays$.
@@
 \end{subfigure}
-\caption{Security scores for device manufacturers, device models, network operators and Nexus devices. 95\% confidence intervals indicated.\todolater{bridge gaps?}}
+\caption{Security scores for device manufacturers, device models, network operators and Nexus devices. 95\% confidence intervals indicated.}
 \label{fig:security_scores}
@@
 This shows that $m$ tends to drag down scores.
-\todo{Give some of the actual orderings as well for the most important ones}
+\todo{Give some of the actual orderings as well for the most important ones - no space}
 \todo{Sensitivity of time based vs. global scores}
-\todo{Reviewer 3 thinks we should use the Spearman rank correlation coefficient instead, and they might be right}
+\todo{Reviewer 3 thinks we should use the Spearman rank correlation coefficient instead, and they might be right, currently being computed}
 

### 2015-07-24 spsm04-thomas.tex
Fiddle with multi URL formatting in references
Minor fixes on reading through

@@
 \def\@URL#1|#2|#3\@nil{%
-  \URL{#1}\ifx\relax#2\relax\else| \URL{#2}\fi}
+  \URL{#1}\ifx\relax#2\relax\else\ -- \URL{#2}\fi}
 \makeatother
@@
 The security of Android depends on security updates for critical vulnerabilities being shipped to devices.
-We define a security metric to rank the performance of device manufacturers and network operators, based on their provision of updates and exposure to critical vulnerabilities.
+We define the FUM security metric to rank the performance of device manufacturers and network operators, based on their provision of updates and exposure to critical vulnerabilities.
 Using a corpus of \daNumOSDevices\ devices we show that there is significant variability in the timely delivery of security updates across different device manufacturers and network operators.
@@
 This is the first paper to quantify security update performance of device manufacturers and network operators.
+\todo{Do we want to pull numeric results into the abstract or leave them out to emphasise the method rather than the results?}
 \end{abstract}
@@
 
-By producing a scoring metric which compares different entities such as device manufacturers, network operators or specific device models, we can provide information that allows device purchasers to select devices with better security.
+By producing the FUM scoring metric which compares different entities such as device manufacturers, network operators or specific device models, we can provide information that allows device purchasers to select devices with better security.
 Currently, corporate and public sector buyers are encouraged to purchase secure devices, but we have found little concrete guidance on the specific makes and models providing the best security.
@@
 By using data from \da~\cite{Wagner2013} we can examine the versions of Android running on user devices over time and through comparison with data on which versions of Android were vulnerable to different critical vulnerabilities~\cite{androidvulnerabilities.org} we can plot the exposure of Android devices to known critical vulnerabilities.
-Figure~\ref{fig:proportioninsecure} uses data from \da\ and the \avo\ vulnerabilities database and shows that Android devices are often exposed to known security vulnerabilities.
+Figure~\ref{fig:proportioninsecure} uses data from \da\ (\S\ref{sec:da}) and the AndroidVulnerabilities.org (\avo) vulnerabilities database (\S\ref{sec:avo}) and shows that Android devices are often exposed to known security vulnerabilities (\S\ref{sec:androidvulnerability}).
 One solution to this problem is regulation, and indeed there is ongoing legal action to force network operators to ship updates for security vulnerabilities~\cite{Soghoian2013}.\todolater{Check on the status of this legal action}
@@
 We quantify the number of updates shipped by various entities in the ecosystem and the number of entities.
-
-\subsection{Android update process}
 
@@
 Sometimes the device manufacturer and Google collaborate closely to make a particular phone, such as with Nexus devices and so Google ships directly to the device.
-Sometimes device manufacturers incorporate upstream open source project releases directly, and sometimes incorrectly -- for example previous work has recorded evidence of broken nightly builds of sqlite in Android releases~\cite{Wagner2013}.
+Sometimes device manufacturers incorporate upstream open source project releases directly, and sometimes incorrectly -- for example previous work has recorded evidence of broken nightly builds of sqlite in Android releases on some device models~\cite{Wagner2013}.
 
@@
 
-\section{Versions of Android running on devices}
+\section{Versions of Android running on devices}\label{sec:da}
 
@@
 The API version is a positive integer that increases when new features are added to the API.
-Consequently security (bug) fixes do not result in a change in the API version.
+Consequently security (bug) fixes do not always result in a change in the API version.
 Fortunately most (\daOSVersionPercValidLines) entries in these data have a build string of the form `x.y.z opaque\_marker' and so it is possible to extract the Android version number `x.y.z'.
@@
 
-\section{Critical vulnerabilities}
+\section{Critical vulnerabilities}\label{sec:avo}
 We have used data on critical vulnerabilities in Android from AndroidVulnerabilities.org (\avo)~\cite{androidvulnerabilities.org}.
@@
 
-\section{Vulnerability of Android dev\-ices}
+\section{Vulnerability of Android dev\-ices}\label{sec:androidvulnerability}
 Combining the data on critical vulnerabilities and versions of Android allows us to plot Figure~\ref{fig:proportioninsecure}, initially all devices are \emph{maybe secure} (yellow) since \da\ does not have historical data prior to May 2011.
@@
 \section{Comparison with other data}\label{sec:representative}
-Here we provide evidence that the \da\ data~\cite{Wagner2013} is representative for our purposes by comparing it with two other sources of data.
+Here we provide evidence that the \da\ data~\cite{Wagner2013} is representative for our purposes by comparing it with three other sources of data.
 The \da\ data is collected from devices across the world where users have voluntarily installed the app.

### 2015-07-24 spsm-submission.sh
Add script to generate the spsm source submission

@@ -1,0 +1,8 @@
+#!/bin/bash
+cp spsm04-thomas.{pdf,tex} spsm-submission/
+cp security_scoring.tex spsm-submission/
+cp countversions.tex spsm-submission/
+cp avostats.tex spsm-submission/
+cp dastats.tex spsm-submission/
+cp -R figures spsm-submission/
+~/git/papers/linuxconfig/bin/bibtex-filter.py --input-bibtex securityupdates.bib --input-aux spsm04-thomas.aux --output-bibtex spsm-submission/securityupdates.bib

### 2015-07-24 spsm-data.sh
Add spsm-data.sh to generate the data submission

@@ -1,0 +1,10 @@
+#!/bin/bash
+
+./spsm-submission.sh
+
+cp -R spsm-submission/* spsm-data/
+cp versions.py spsm-data/
+cp version_dates* spsm-data/
+cp /home/drt24/git/lib/latex_value/latex_value.py spsm-data/
+cp /home/drt24/git/da/graphing/{dahelpers,iso8601,os-version,ddict,contribution_analysis,operator_analysis,manufacturer_model_analysis,version_tools,version_security,plot_tools,iohelpers,device_data_analysis}.py spsm-data/
+cp -R /home/drt24/Documents/androidvulnerabilities spsm-data/

### 2015-07-24 spsm-data.sh
Filled in data submission form for data currently about to be submitted.

@@ -2,9 +2,10 @@
 
 ./spsm-submission.sh
 
-cp -R spsm-submission/* spsm-data/
-cp versions.py spsm-data/
-cp version_dates* spsm-data/
-cp /home/drt24/git/lib/latex_value/latex_value.py spsm-data/
-cp /home/drt24/git/da/graphing/{dahelpers,iso8601,os-version,ddict,contribution_analysis,operator_analysis,manufacturer_model_analysis,version_tools,version_security,plot_tools,iohelpers,device_data_analysis}.py spsm-data/
-cp -R /home/drt24/Documents/androidvulnerabilities spsm-data/
+cp -Rp spsm-submission/* spsm-data/
+cp -p versions.py spsm-data/
+cp -p version_dates* spsm-data/
+cp -p /home/drt24/git/lib/latex_value/latex_value.py spsm-data/
+cp -p /home/drt24/git/da/graphing/{dahelpers,iso8601,os-version,ddict,contribution_analysis,operator_analysis,manufacturer_model_analysis,version_tools,version_security,plot_tools,iohelpers,device_data_analysis}.py spsm-data/
+cp -Rfp /home/drt24/Documents/androidvulnerabilities spsm-data/
+cp -p data-submission.txt spsm-data/

### 2015-07-24 spsm-submission.sh

@@ -1,5 +1,5 @@
 #!/bin/bash
-cp spsm04-thomas.{pdf,tex} spsm-submission/
+cp spsm04-thomas.{pdf,tex,bbl} spsm-submission/
 cp security_scoring.tex spsm-submission/
 cp countversions.tex spsm-submission/
 cp avostats.tex spsm-submission/

### 2015-07-24 spsm-data.sh
Update spsm-data to include some csv files with some of the raw data

@@ -3,9 +3,12 @@
 ./spsm-submission.sh
 
 cp -Rp spsm-submission/* spsm-data/
+zip --quiet spsm-data/figures spsm-data/figures/*
 cp -p versions.py spsm-data/
 cp -p version_dates* spsm-data/
 cp -p /home/drt24/git/lib/latex_value/latex_value.py spsm-data/
 cp -p /home/drt24/git/da/graphing/{dahelpers,iso8601,os-version,ddict,contribution_analysis,operator_analysis,manufacturer_model_analysis,version_tools,version_security,plot_tools,iohelpers,device_data_analysis}.py spsm-data/
 cp -Rfp /home/drt24/Documents/androidvulnerabilities spsm-data/
+zip --quiet -r spsm-data/androidvulnerabilities spsm-data/androidvulnerabilities
 cp -p data-submission.txt spsm-data/
+cp /mnt/nas04-device-analyzer-graphing/csv/{api_gpcomp_diff,api_gpcomp_rdiff,api-versions,first_api_versions,first_os_versions,googleplayapi,gp_per_api_diff,norm_api_gpcomp,security_updates,s_manufacturers,s_models,s_operators,updates}.csv spsm-data/

### 2015-07-24 dastats.tex
./update.sh

@@
 \newcommand{\daStartDate}{2011-07-01}
-\newcommand{\daEndDate}{2015-07-14}
+\newcommand{\daEndDate}{2015-07-15}
 \newcommand{\daFullDeployedAt}{95\%}
-\newcommand{\daNumManufacturers}{300}
+\newcommand{\daNumManufacturers}{301}
 \newcommand{\daNumSigManufacturers}{10}
-\newcommand{\daNumModels}{2\,730}
+\newcommand{\daNumModels}{2\,740}
 \newcommand{\daNumOperators}{1\,460}
-\newcommand{\daNumSigOperators}{14}
+\newcommand{\daNumSigOperators}{15}
 \newcommand{\daNumOSDevices}{20\,400}
@@
 \newcommand{\daVulnFree}{$0.123 \pm 0.0$}
-\newcommand{\daMeanOutstandingVulnerabilities}{$0.531 \pm 0.0$}
+\newcommand{\daMeanOutstandingVulnerabilities}{$0.53 \pm 0.0$}
 \newcommand{\daUpdatednessNominal}{0.0521}
@@
 \newcommand{\daSecurityScore}{$2.87 \pm 0.0$}
-\newcommand{\daCurrentMeanInsecurityPercNominal}{79.7\%}
-\newcommand{\daCurrentMeanInsecurityPerc}{$79.7 \pm 0.0\%$}
+\newcommand{\daCurrentMeanInsecurityPercNominal}{79.6\%}
+\newcommand{\daCurrentMeanInsecurityPerc}{$79.6 \pm 0.0\%$}
 \newcommand{\daCurrentMeanInsecurityPercTwosfNominal}{80\%}
 \newcommand{\daCurrentMeanInsecurityPercTwosf}{$80 \pm 0\%$}
-\newcommand{\daCurrentMeanSecurityPercNominal}{20.3\%}
-\newcommand{\daCurrentMeanSecurityPerc}{$20.3 \pm 0.1\%$}
+\newcommand{\daCurrentMeanSecurityPercNominal}{20.4\%}
+\newcommand{\daCurrentMeanSecurityPerc}{$20.4 \pm 0.1\%$}
 \newcommand{\daCurrentMeanSecurityPercTwosfNominal}{20\%}
 \newcommand{\daCurrentMeanSecurityPercTwosf}{$20 \pm 0\%$}
-\newcommand{\daCurrentVulnFreeNominal}{0.203}
-\newcommand{\daCurrentVulnFree}{$0.203 \pm 0.001$}
-\newcommand{\daCurrentMeanOutstandingVulnerabilities}{$0.124 \pm 0.0$}
-\newcommand{\daCurrentUpdatednessNominal}{0.0532}
-\newcommand{\daCurrentUpdatedness}{$0.0532 \pm 0.0$}
-\newcommand{\daCurrentUpdatednessPercNominal}{5.32\%}
-\newcommand{\daCurrentUpdatednessPerc}{$5.32 \pm 0.0\%$}
-\newcommand{\daCurrentUpdatednessPercTwosfNominal}{5.3\%}
-\newcommand{\daCurrentUpdatednessPercTwosf}{$5.3 \pm 0.0\%$}
+\newcommand{\daCurrentVulnFreeNominal}{0.204}
+\newcommand{\daCurrentVulnFree}{$0.204 \pm 0.001$}
+\newcommand{\daCurrentMeanOutstandingVulnerabilities}{$0.123 \pm 0.0$}
+\newcommand{\daCurrentUpdatednessNominal}{0.0536}
+\newcommand{\daCurrentUpdatedness}{$0.0536 \pm 0.0$}
+\newcommand{\daCurrentUpdatednessPercNominal}{5.36\%}
+\newcommand{\daCurrentUpdatednessPerc}{$5.36 \pm 0.0\%$}
+\newcommand{\daCurrentUpdatednessPercTwosfNominal}{5.4\%}
+\newcommand{\daCurrentUpdatednessPercTwosf}{$5.4 \pm 0.0\%$}
 \newcommand{\daCurrentSecurityScoreNominal}{3.79}
 \newcommand{\daCurrentSecurityScore}{$3.79 \pm 0.0$}
-\newcommand{\daDailyParticipationNominal}{899}
-\newcommand{\daDailyParticipation}{$899 \pm 245$}
+\newcommand{\daDailyParticipationNominal}{898}
+\newcommand{\daDailyParticipation}{$898 \pm 245$}
 \newcommand{\daNumFullVersions}{1\,270}
@@
 \newcommand{\daVulnZergRushMonthsDefFixDeployed}{27.3}
-\newcommand{\daTabSecScoresmanufacturer}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline LG & $0.22 \pm 0.00$ & $0.33 \pm 0.00$ & $0.62 \pm 0.01$ & $3.97 \pm 0.02$ \\  Motorola & $0.18 \pm 0.00$ & $0.12 \pm 0.00$ & $0.72 \pm 0.02$ & $3.07 \pm 0.02$ \\  Samsung & $0.13 \pm 0.00$ & $0.04 \pm 0.00$ & $0.61 \pm 0.00$ & $2.75 \pm 0.00$ \\  Sony & $0.14 \pm 0.00$ & $0.19 \pm 0.00$ & $1.09 \pm 0.02$ & $2.63 \pm 0.02$ \\  HTC & $0.14 \pm 0.00$ & $0.10 \pm 0.00$ & $0.87 \pm 0.01$ & $2.63 \pm 0.02$ \\  asus & $0.20 \pm 0.00$ & $0.51 \pm 0.01$ & $6.01 \pm 0.07$ & $2.35 \pm 0.02$ \\  other & $0.06 \pm 0.00$ & $0.05 \pm 0.00$ & $1.04 \pm 0.01$ & $1.97 \pm 0.02$ \\  alps & $0.03 \pm 0.00$ & $0.19 \pm 0.01$ & $3.99 \pm 0.08$ & $0.80 \pm 0.02$ \\  Symphony & $0.00 \pm 0.00$ & $0.08 \pm 0.00$ & $5.00 \pm 0.05$ & $0.30 \pm 0.01$ \\  walton & $0.00 \pm 0.00$ & $0.09 \pm 0.00$ & $6.00 \pm 0.08$ & $0.27 \pm 0.01$ \\ \end{tabular} \caption{Security scores for manufacturers} \label{tab:sec_manufacturer} \end{table*}}
+\newcommand{\daTabSecScoresmanufacturer}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline LG & $0.22 \pm 0.00$ & $0.33 \pm 0.00$ & $0.62 \pm 0.01$ & $3.97 \pm 0.02$ \\  Motorola & $0.18 \pm 0.00$ & $0.12 \pm 0.00$ & $0.71 \pm 0.02$ & $3.07 \pm 0.02$ \\  Samsung & $0.13 \pm 0.00$ & $0.04 \pm 0.00$ & $0.61 \pm 0.00$ & $2.75 \pm 0.00$ \\  Sony & $0.14 \pm 0.00$ & $0.19 \pm 0.00$ & $1.09 \pm 0.02$ & $2.63 \pm 0.02$ \\  HTC & $0.14 \pm 0.00$ & $0.10 \pm 0.00$ & $0.87 \pm 0.01$ & $2.63 \pm 0.02$ \\  asus & $0.20 \pm 0.00$ & $0.51 \pm 0.01$ & $6.01 \pm 0.07$ & $2.35 \pm 0.02$ \\  other & $0.06 \pm 0.00$ & $0.05 \pm 0.00$ & $1.04 \pm 0.01$ & $1.97 \pm 0.02$ \\  alps & $0.03 \pm 0.00$ & $0.19 \pm 0.01$ & $3.99 \pm 0.08$ & $0.80 \pm 0.02$ \\  Symphony & $0.00 \pm 0.00$ & $0.08 \pm 0.00$ & $5.00 \pm 0.05$ & $0.30 \pm 0.01$ \\  walton & $0.00 \pm 0.00$ & $0.09 \pm 0.00$ & $6.00 \pm 0.08$ & $0.27 \pm 0.01$ \\ \end{tabular} \caption{Security scores for manufacturers} \label{tab:sec_manufacturer} \end{table*}}
 \newcommand{\daSecScoreBestmanufacturer}{LG}
@@
 \newcommand{\daSecScoreWorstmanufacturerNumFullVersions}{21}
-\newcommand{\daTabSecScoresmodel}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline Galaxy Nexus & $0.50 \pm 0.00$ & $0.54 \pm 0.01$ & $1.53 \pm 0.04$ & $4.71 \pm 0.04$ \\  Nexus 4 & $0.30 \pm 0.00$ & $0.82 \pm 0.01$ & $6.06 \pm 0.09$ & $3.69 \pm 0.04$ \\  Nexus 7 & $0.26 \pm 0.00$ & $0.74 \pm 0.01$ & $5.92 \pm 0.09$ & $3.24 \pm 0.04$ \\  other & $0.10 \pm 0.00$ & $0.14 \pm 0.00$ & $0.53 \pm 0.00$ & $3.03 \pm 0.00$ \\  Desire HD & $0.08 \pm 0.00$ & $0.05 \pm 0.00$ & $0.38 \pm 0.02$ & $2.91 \pm 0.04$ \\  HTC Sensation & $0.35 \pm 0.00$ & $0.01 \pm 0.01$ & $1.57 \pm 0.05$ & $2.44 \pm 0.05$ \\  GT-I9100 & $0.22 \pm 0.00$ & $0.02 \pm 0.00$ & $1.23 \pm 0.02$ & $2.28 \pm 0.02$ \\  HTC Desire S & $0.02 \pm 0.00$ & $0.02 \pm 0.00$ & $1.00 \pm 0.06$ & $1.74 \pm 0.07$ \\  GT-N7000 & $0.25 \pm 0.00$ & $0.00 \pm 0.00$ & $2.52 \pm 0.05$ & $1.43 \pm 0.02$ \\  GT-P1000 & $0.01 \pm 0.00$ & $0.00 \pm 0.01$ & $1.79 \pm 0.06$ & $0.90 \pm 0.05$ \\  GT-I9300 & $0.13 \pm 0.00$ & $0.01 \pm 0.00$ & $6.23 \pm 0.04$ & $0.58 \pm 0.01$ \\  GT-I9505 & $0.03 \pm 0.00$ & $0.13 \pm 0.00$ & $6.82 \pm 0.07$ & $0.52 \pm 0.01$ \\  HTC Desire HD & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & $3.03 \pm 0.05$ & $0.28 \pm 0.03$ \\  GT-N7100 & $0.06 \pm 0.00$ & $0.00 \pm 0.01$ & $6.93 \pm 0.08$ & $0.24 \pm 0.02$ \\  Symphony W68 & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & $11.00 \pm 0.12$ & $0.00 \pm 0.03$ \\ \end{tabular} \caption{Security scores for models} \label{tab:sec_model} \end{table*}}
+\newcommand{\daTabSecScoresmodel}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline Galaxy Nexus & $0.50 \pm 0.00$ & $0.54 \pm 0.01$ & $1.53 \pm 0.04$ & $4.71 \pm 0.04$ \\  Nexus 4 & $0.30 \pm 0.00$ & $0.82 \pm 0.01$ & $6.06 \pm 0.09$ & $3.69 \pm 0.04$ \\  Nexus 7 & $0.26 \pm 0.00$ & $0.74 \pm 0.01$ & $5.92 \pm 0.09$ & $3.24 \pm 0.04$ \\  other & $0.10 \pm 0.00$ & $0.14 \pm 0.00$ & $0.53 \pm 0.00$ & $3.03 \pm 0.00$ \\  Desire HD & $0.08 \pm 0.00$ & $0.05 \pm 0.00$ & $0.38 \pm 0.02$ & $2.91 \pm 0.04$ \\  HTC Sensation & $0.35 \pm 0.00$ & $0.01 \pm 0.01$ & $1.57 \pm 0.05$ & $2.44 \pm 0.05$ \\  GT-I9100 & $0.22 \pm 0.00$ & $0.02 \pm 0.00$ & $1.23 \pm 0.02$ & $2.27 \pm 0.02$ \\  HTC Desire S & $0.02 \pm 0.00$ & $0.02 \pm 0.00$ & $1.00 \pm 0.06$ & $1.74 \pm 0.07$ \\  GT-N7000 & $0.25 \pm 0.00$ & $0.00 \pm 0.00$ & $2.52 \pm 0.05$ & $1.43 \pm 0.02$ \\  GT-P1000 & $0.01 \pm 0.00$ & $0.00 \pm 0.01$ & $1.79 \pm 0.06$ & $0.90 \pm 0.05$ \\  GT-I9300 & $0.13 \pm 0.00$ & $0.01 \pm 0.00$ & $6.23 \pm 0.04$ & $0.58 \pm 0.01$ \\  GT-I9505 & $0.03 \pm 0.00$ & $0.13 \pm 0.00$ & $6.82 \pm 0.07$ & $0.52 \pm 0.01$ \\  HTC Desire HD & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & $3.03 \pm 0.05$ & $0.28 \pm 0.03$ \\  GT-N7100 & $0.06 \pm 0.00$ & $0.00 \pm 0.01$ & $6.93 \pm 0.08$ & $0.24 \pm 0.02$ \\  Symphony W68 & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & $11.00 \pm 0.12$ & $0.00 \pm 0.03$ \\ \end{tabular} \caption{Security scores for models} \label{tab:sec_model} \end{table*}}
 \newcommand{\daSecScoreBestmodel}{Galaxy Nexus}
@@
 \newcommand{\daTabSecScoressummary}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline nexus & $0.39 \pm 0.00$ & $0.48 \pm 0.00$ & $0.56 \pm 0.01$ & $5.17 \pm 0.02$ \\  notnexus & $0.10 \pm 0.00$ & $0.02 \pm 0.00$ & $0.53 \pm 0.00$ & $2.70 \pm 0.00$ \\ \end{tabular} \caption{Security scores for nexus} \label{tab:sec_summary} \end{table*}}
-\newcommand{\daTabSecScoresoperator}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline O2 uk & $0.27 \pm 0.00$ & $0.12 \pm 0.00$ & $0.37 \pm 0.02$ & $3.87 \pm 0.03$ \\  T-Mobile & $0.22 \pm 0.00$ & $0.18 \pm 0.00$ & $0.40 \pm 0.01$ & $3.81 \pm 0.02$ \\  Orange & $0.22 \pm 0.00$ & $0.10 \pm 0.00$ & $0.36 \pm 0.02$ & $3.64 \pm 0.04$ \\  Sprint & $0.18 \pm 0.00$ & $0.11 \pm 0.00$ & $0.43 \pm 0.02$ & $3.42 \pm 0.03$ \\  3 & $0.20 \pm 0.00$ & $0.09 \pm 0.00$ & $0.47 \pm 0.02$ & $3.38 \pm 0.03$ \\  Vodafone uk & $0.14 \pm 0.00$ & $0.13 \pm 0.00$ & $0.52 \pm 0.03$ & $3.17 \pm 0.04$ \\  AT\&T & $0.14 \pm 0.00$ & $0.08 \pm 0.00$ & $0.43 \pm 0.02$ & $3.13 \pm 0.02$ \\  unknown & $0.11 \pm 0.00$ & $0.20 \pm 0.00$ & $0.84 \pm 0.01$ & $2.88 \pm 0.02$ \\  Verizon & $0.19 \pm 0.00$ & $0.09 \pm 0.00$ & $0.82 \pm 0.02$ & $2.84 \pm 0.02$ \\  n Telenor & $0.04 \pm 0.00$ & $0.12 \pm 0.00$ & $1.21 \pm 0.02$ & $1.89 \pm 0.02$ \\  Airtel & $0.05 \pm 0.00$ & $0.03 \pm 0.00$ & $1.47 \pm 0.03$ & $1.41 \pm 0.03$ \\  Grameenphone & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & $1.88 \pm 0.02$ & $0.93 \pm 0.01$ \\  Robi & $0.00 \pm 0.00$ & $0.08 \pm 0.00$ & $2.07 \pm 0.04$ & $0.91 \pm 0.03$ \\  banglalink & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & $2.55 \pm 0.04$ & $0.55 \pm 0.02$ \\ \end{tabular} \caption{Security scores for operators} \label{tab:sec_operator} \end{table*}}
+\newcommand{\daTabSecScoresoperator}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline O2 uk & $0.27 \pm 0.00$ & $0.12 \pm 0.00$ & $0.37 \pm 0.02$ & $3.87 \pm 0.03$ \\  T-Mobile & $0.22 \pm 0.00$ & $0.18 \pm 0.00$ & $0.40 \pm 0.01$ & $3.81 \pm 0.02$ \\  Orange & $0.22 \pm 0.00$ & $0.10 \pm 0.00$ & $0.36 \pm 0.02$ & $3.65 \pm 0.04$ \\  Sprint & $0.18 \pm 0.00$ & $0.11 \pm 0.00$ & $0.43 \pm 0.02$ & $3.42 \pm 0.03$ \\  3 & $0.20 \pm 0.00$ & $0.09 \pm 0.00$ & $0.47 \pm 0.02$ & $3.39 \pm 0.03$ \\  Vodafone uk & $0.14 \pm 0.00$ & $0.13 \pm 0.00$ & $0.52 \pm 0.03$ & $3.17 \pm 0.04$ \\  AT\&T & $0.14 \pm 0.00$ & $0.08 \pm 0.00$ & $0.43 \pm 0.02$ & $3.13 \pm 0.02$ \\  unknown & $0.11 \pm 0.00$ & $0.20 \pm 0.00$ & $0.84 \pm 0.01$ & $2.88 \pm 0.02$ \\  Verizon & $0.19 \pm 0.00$ & $0.09 \pm 0.00$ & $0.82 \pm 0.02$ & $2.84 \pm 0.02$ \\  n Telenor & $0.04 \pm 0.00$ & $0.12 \pm 0.00$ & $1.21 \pm 0.02$ & $1.89 \pm 0.02$ \\  Airtel & $0.05 \pm 0.00$ & $0.03 \pm 0.00$ & $1.47 \pm 0.03$ & $1.41 \pm 0.03$ \\  Grameenphone & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & $1.88 \pm 0.02$ & $0.94 \pm 0.01$ \\  Robi & $0.00 \pm 0.00$ & $0.08 \pm 0.00$ & $2.07 \pm 0.04$ & $0.91 \pm 0.03$ \\  banglalink & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & $2.55 \pm 0.04$ & $0.55 \pm 0.02$ \\ \end{tabular} \caption{Security scores for operators} \label{tab:sec_operator} \end{table*}}
 \newcommand{\daSecScoreBestoperator}{O2 uk}
@@
 \newcommand{\daSecScoreWorstoperatorScore}{$0.548 \pm 0.018$}
-\newcommand{\daSecScoreWorstoperatorNumFullVersions}{73}
-\newcommand{\daTabDLDistances}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & summary \\ \hline equal & 0.0 & 0.0667 & 0.0 & 0.0 \\weight $m$ & 0.2 & 0.4 & 0.0714 & 0.0 \\$f$ & 0.4 & 0.733 & 0.5 & 0.0 \\weight $u$ & 0.4 & 0.0667 & 0.143 & 0.0 \\$m$ & 0.6 & 0.733 & 0.357 & 0.5 \\$u$ & 0.9 & 0.667 & 0.786 & 0.0 \\\end{tabular} \caption{Normalised Damerau-Levenshtein distances for different metrics} \label{tab:dl_distances} \end{table}}
-\newcommand{\daTabChangeInScores}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & summary \\ \hline $m$ & $-1.67 \pm 2.25$ & $-0.66 \pm 2.54$ & $-3.3 \pm 1.3$ & $-3.4 \pm 1.8$ \\weight $m$ & $-0.264 \pm 0.276$ & $-0.0991 \pm 0.32$ & $-0.461 \pm 0.167$ & $-0.488 \pm 0.216$ \\equal & $-0.108 \pm 0.062$ & $-0.0368 \pm 0.112$ & $-0.146 \pm 0.045$ & $-0.165 \pm 0.033$ \\weight $u$ & $-0.0609 \pm 0.122$ & $-0.0113 \pm 0.232$ & $0.024 \pm 0.081$ & $-0.0061 \pm 0.12$ \\$u$ & $0.366 \pm 1.61$ & $0.219 \pm 2.12$ & $1.55 \pm 0.94$ & $1.42 \pm 1.48$ \\$f$ & $0.976 \pm 0.564$ & $0.331 \pm 1.01$ & $1.31 \pm 0.40$ & $1.48 \pm 0.29$ \\\end{tabular} \caption{Mean change in scores for different metrics} \label{tab:change_in_scores} \end{table*}}
+\newcommand{\daSecScoreWorstoperatorNumFullVersions}{74}
+\newcommand{\daTabDLDistances}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & summary \\ \hline equal & 0.0 & 0.0667 & 0.0 & 0.0 \\weight $m$ & 0.2 & 0.4 & 0.0714 & 0.0 \\weight $u$ & 0.4 & 0.0667 & 0.143 & 0.0 \\$f$ & 0.4 & 0.667 & 0.5 & 0.0 \\$m$ & 0.6 & 0.733 & 0.357 & 0.5 \\$u$ & 0.9 & 0.667 & 0.786 & 0.0 \\\end{tabular} \caption{Normalised Damerau-Levenshtein distances for different metrics} \label{tab:dl_distances} \end{table}}
+\newcommand{\daTabChangeInScores}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & summary \\ \hline $m$ & $-1.67 \pm 2.25$ & $-0.66 \pm 2.54$ & $-3.3 \pm 1.3$ & $-3.4 \pm 1.8$ \\weight $m$ & $-0.264 \pm 0.276$ & $-0.0991 \pm 0.32$ & $-0.462 \pm 0.167$ & $-0.488 \pm 0.216$ \\equal & $-0.108 \pm 0.062$ & $-0.0368 \pm 0.112$ & $-0.146 \pm 0.045$ & $-0.165 \pm 0.033$ \\weight $u$ & $-0.0608 \pm 0.122$ & $-0.0112 \pm 0.232$ & $0.0238 \pm 0.0807$ & $-0.00602 \pm 0.12$ \\$u$ & $0.367 \pm 1.61$ & $0.219 \pm 2.12$ & $1.55 \pm 0.94$ & $1.42 \pm 1.48$ \\$f$ & $0.976 \pm 0.564$ & $0.331 \pm 1.01$ & $1.31 \pm 0.40$ & $1.48 \pm 0.29$ \\\end{tabular} \caption{Mean change in scores for different metrics} \label{tab:change_in_scores} \end{table*}}
 \newcommand{\daGPAPICurveFitParamFirst}{80.8}
@@
 \newcommand{\daGPAPISeventeenFullDeploymentMonth}{October 2017}
-\newcommand{\daMeanOutstandingVulnerabilitiesNominal}{0.531}
-\newcommand{\daCurrentMeanOutstandingVulnerabilitiesNominal}{0.124}
+\newcommand{\daMeanOutstandingVulnerabilitiesNominal}{0.53}
+\newcommand{\daCurrentMeanOutstandingVulnerabilitiesNominal}{0.123}
 \newcommand{\daModelHalfDeploymentYears}{0.0822}

### 2015-07-24 update.sh
Make update.sh automatically regenerate the bibliography

@@ -42,6 +42,7 @@ cp "$fromdir"security_score_operator.pdf security_score_operator.pdf
 cp "$fromdir"security_score_summary.pdf security_score_summary.pdf
 cp ~drt24/git/da/graphing/pdf/dists.pdf dists.pdf
 popd
+update_bib 2> /dev/null
 cp ~/Library/papers/library.filtered.bib securityupdates.bib
 cp /mnt/nas04-device-analyzer-graphing/stats/latex.tex dastats.tex
 cp ~/Documents/androidvulnerabilities/output/latex.tex avostats.tex

### 2015-07-24 spsm04-thomas.tex
Add citation for the data submission
Reorder sections so that we can make them flow better

@@
 
+\section{Threat model}
+\label{sec:threatmodel}
+
+We assume the defender can deploy security controls at three levels: in an online marketplace, at app installation, and during app execution.
+Here, we consider three attack vectors against Android handsets.
+
+The \emph{installation} attack vector is through the installation of a malicious app on the device.
+Android devices can install apps through marketplaces, email attachments, URLs and via the Android Debug Bridge (ADB).
+By default, many Android devices will only allow the installation of apps from Google Play, which uses Bouncer to automatically analyse apps, and quickly takes down apps that are reported as malicious.
+However, alternative markets are also popular, particularly in countries where Google Play is not available.
+
+The \emph{dynamic code loading} attack vector is to make an existing app download and execute code at runtime.
+The most direct method is to upload to a marketplace a seemingly innocent app that dynamically loads malicious code, either as additional davik bytecode, as a native library or by embedding an interpreter and executing received instructions.
+Neither static nor dynamic analysis of this app will uncover any malicious code, since it does not exist in the app.
+The marketplace can try to detect explicit use of dynamic code loading, however there are ways to dynamically load code which are hard to detect even on a platform such as iOS, which does not permit dynamic code loading, a Return-Oriented Programming (ROP) attack is relatively easy if the attacker creates an app with carefully crafted flaws~\cite{Wang2013a}.
+
+The \emph{injection} attack vector is that the attacker injects malicious code directly into existing code already on the handset. 
+For example, the addJavascriptInterface vulnerability (CVE-2012-6636) allows JavaScript running in an Android WebView to execute arbitrary code as the vulnerable app's user.
+The fix for the addJavascriptInterface vulnerability breaks backwards compatibility and requires a two sided fix.
+While the fix was released in December 2012, on \daGPAPISeventeenLaterDate, \daGPAPISeventeenEarlierProportion\ of handsets connecting to Google Play were still vulnerable to this attack~\cite{Thomas2015a}.
+
+The best place to prevent attacks is at runtime, since all three attack vectors can be prevented at this level. 
+Unfortunately, as Figure~\ref{fig:proportioninsecure} shows, the sandbox for Android apps is weak, since on most devices there are known critical vulnerabilities in it.
+
+\section{Critical vulnerabilities}\label{sec:avo}
+We have used data on critical vulnerabilities in Android from AndroidVulnerabilities.org (\avo)~\cite{androidvulnerabilities.org}.
+It contains information on the discovery and publication dates, the versions affected and which versions contained fixes.
+It only contains critical vulnerabilities such as root vulnerabilities that did not require USB debugging to exploit.
+If malicious code exploits a critical vulnerability then it gains control of the device.
+
+Some critical vulnerabilities are not traditional kernel vulnerabilities, but are vulnerabilities in \emph{installation} in our threat model.
+For example improper verification of signatures at installation time was discovered in February 2013~\cite{Forristal2013} and meant that applications could pretend to be signed with system keys and hence be granted system privileges.
+On versions of Android below 4.1 malware could then use known system-to-root escalation mechanisms.
+Regardless of version, this exposed an increased attack area and would also provide the ability for malware to control all user internet traffic (via VPNs), brick the phone, remove and install apps, steal user credentials and read the screen.
+The different categories in which the vulnerabilities fall are shown in Table~\ref{tab:andvulns}, `signature' vulnerabilities are \emph{installation}, while kernel and system vulnerabilities are used with \emph{installation, dynamic code loading} and \emph{injection}.
+\avoTabAndVulns
+
+\avo\ currently contains \avoNumVulnerabilities\ vulnerabilities, of which \avoNumVulnAllAndroid\ affect all Android devices and \avoNumVulnSpecific\ are specific to particular devices or device manufacturers.
+
+Unless otherwise stated, we use the \daNumVulnsUsed\ vulnerabilities shown in Table~\ref{tab:andvulns} in our analysis.
+We have chosen these vulnerabilities to fit the attack vectors introduced in \S\ref{sec:threatmodel}.
+These vulnerabilities affect all Android devices regardless of manufacturer, and as a result our selected vulnerabilities will dominate any security analysis of Android.
+Hence, with our chosen set of vulnerabilities, our analysis represents a lower-bound on the vulnerability of devices in the \da\ data.
+
+\section{Vulnerability of Android dev\-ices}\label{sec:androidvulnerability}
+Combining the data on critical vulnerabilities and versions of Android allows us to plot Figure~\ref{fig:proportioninsecure}, initially all devices are \emph{maybe secure} (yellow) since \da\ does not have historical data prior to May 2011.
+This means we cannot distinguish between devices which are running a version of Android which is known to be vulnerable from one which may have received a backported fix.
+This demonstrates the importance of a longitudinal study: this type of analysis requires years of data.
+Once \avovuln{zergRush} was discovered in October 2011 then most devices are recorded as \emph{insecure} (red) as they were exposed to that vulnerability.
+The remaining devices were already running a version of Android which fixed the \avovuln{zergRush} vulnerability and are therefore marked as \emph{secure} (green).
+From October 2011 until the discovery of \avovuln[APK_duplicate_file]{APK duplicate file} in February 2013 the graph shows progressive improvement as devices are upgraded.
+This means more and more devices are marked as \emph{secure} because they are now running a secure version of Android, or marked as \emph{maybe secure} because they received an OS update that did not update to a known-good version of Android but which may still have included a backport of a fix, as the update was made available after the vulnerability was disclosed.
+From February 2013 onwards regular discovery of critical vulnerabilities ensures that most devices are exposed to known critical vulnerabilities.
+This gives us an average exposure to critical vulnerabilities of Android devices of \daMeanInsecurityPerc.
+
 \section{Android Ecosystem}\label{sec:android_ecosystem}
@@
 
-\section{Threat model and attack vectors}
-\label{sec:threatmodel}
-
-We assume the defender can deploy security controls at three levels: in an online marketplace, at app installation, and during app execution.
-Here, we consider three attack vectors against Android handsets.
-
-The \emph{installation} attack vector is through the installation of a malicious app on the device.
-Android devices can install apps through marketplaces, email attachments, URLs and via the Android Debug Bridge (ADB).
-By default, many Android devices will only allow the installation of apps from Google Play, which uses Bouncer to automatically analyse apps, and quickly takes down apps that are reported as malicious.
-However, alternative markets are also popular, particularly in countries where Google Play is not available.
-
-The \emph{dynamic code loading} attack vector is to make an existing app download and execute code at runtime.
-The most direct method is to upload to a marketplace a seemingly innocent app that dynamically loads malicious code, either as additional davik bytecode, as a native library or by embedding an interpreter and executing received instructions.
-Neither static nor dynamic analysis of this app will uncover any malicious code, since it does not exist in the app.
-The marketplace can try to detect explicit use of dynamic code loading, however there are ways to dynamically load code which are hard to detect even on a platform such as iOS, which does not permit dynamic code loading, a Return-Oriented Programming (ROP) attack is relatively easy if the attacker creates an app with carefully crafted flaws~\cite{Wang2013a}.
-
-The \emph{injection} attack vector is that the attacker injects malicious code directly into existing code already on the handset. 
-For example, the addJavascriptInterface vulnerability (CVE-2012-6636) allows JavaScript running in an Android WebView to execute arbitrary code as the vulnerable app's user.
-The fix for the addJavascriptInterface vulnerability breaks backwards compatibility and requires a two sided fix.
-While the fix was released in December 2012, on \daGPAPISeventeenLaterDate, \daGPAPISeventeenEarlierProportion\ of handsets connecting to Google Play were still vulnerable to this attack~\cite{Thomas2015a}.
-
-The best place to prevent attacks is at runtime, since all three attack vectors can be prevented at this level. 
-Unfortunately, as Figure~\ref{fig:proportioninsecure} shows, the sandbox for Android apps is weak, since on most devices there are known critical vulnerabilities in it.
-
-\section{Versions of Android running on devices}\label{sec:da}
+\section{Device Analyzer data}\label{sec:da}
 
@@
 The majority of devices only contribute data for a short period of time, however \daMonthsDevices\ devices have contributed data for more than \daMonths~months.
-
-\section{Critical vulnerabilities}\label{sec:avo}
-We have used data on critical vulnerabilities in Android from AndroidVulnerabilities.org (\avo)~\cite{androidvulnerabilities.org}.
-It contains information on the discovery and publication dates, the versions affected and which versions contained fixes.
-It only contains critical vulnerabilities such as root vulnerabilities that did not require USB debugging to exploit.
-If malicious code exploits a critical vulnerability then it gains control of the device.
-
-Some critical vulnerabilities are not traditional kernel vulnerabilities, but are vulnerabilities in \emph{installation} in our threat model.
-For example improper verification of signatures at installation time was discovered in February 2013~\cite{Forristal2013} and meant that applications could pretend to be signed with system keys and hence be granted system privileges.
-On versions of Android below 4.1 malware could then use known system-to-root escalation mechanisms.
-Regardless of version, this exposed an increased attack area and would also provide the ability for malware to control all user internet traffic (via VPNs), brick the phone, remove and install apps, steal user credentials and read the screen.
-The different categories in which the vulnerabilities fall are shown in Table~\ref{tab:andvulns}, `signature' vulnerabilities are \emph{installation}, while kernel and system vulnerabilities are used with \emph{installation, dynamic code loading} and \emph{injection}.
-\avoTabAndVulns
-
-\avo\ currently contains \avoNumVulnerabilities\ vulnerabilities, of which \avoNumVulnAllAndroid\ affect all Android devices and \avoNumVulnSpecific\ are specific to particular devices or device manufacturers.
-
-Unless otherwise stated, we use the \daNumVulnsUsed\ vulnerabilities shown in Table~\ref{tab:andvulns} in our analysis.
-We have chosen these vulnerabilities to fit the attack vectors introduced in \S\ref{sec:threatmodel}.
-These vulnerabilities affect all Android devices regardless of manufacturer, and as a result our selected vulnerabilities will dominate any security analysis of Android.
-Hence, with our chosen set of vulnerabilities, our analysis represents a lower-bound on the vulnerability of devices in the \da\ data.
-
-\section{Vulnerability of Android dev\-ices}\label{sec:androidvulnerability}
-Combining the data on critical vulnerabilities and versions of Android allows us to plot Figure~\ref{fig:proportioninsecure}, initially all devices are \emph{maybe secure} (yellow) since \da\ does not have historical data prior to May 2011.
-This means we cannot distinguish between devices which are running a version of Android which is known to be vulnerable from one which may have received a backported fix.
-This demonstrates the importance of a longitudinal study: this type of analysis requires years of data.
-Once \avovuln{zergRush} was discovered in October 2011 then most devices are recorded as \emph{insecure} (red) as they were exposed to that vulnerability.
-The remaining devices were already running a version of Android which fixed the \avovuln{zergRush} vulnerability and are therefore marked as \emph{secure} (green).
-From October 2011 until the discovery of \avovuln[APK_duplicate_file]{APK duplicate file} in February 2013 the graph shows progressive improvement as devices are upgraded.
-This means more and more devices are marked as \emph{secure} because they are now running a secure version of Android, or marked as \emph{maybe secure} because they received an OS update that did not update to a known-good version of Android but which may still have included a backport of a fix, as the update was made available after the vulnerability was disclosed.
-From February 2013 onwards regular discovery of critical vulnerabilities ensures that most devices are exposed to known critical vulnerabilities.
-This gives us an average exposure to critical vulnerabilities of Android devices of \daMeanInsecurityPerc.
 
@@
 \section*{Dataset}
-The raw and processed data and source code will be made available where possible when this paper is finalised.
+Much of the raw and processed data and source code is available, excluding that which might identify individuals~\cite{Thomas2015e}.
 Data from \da\ and \avo\ used in this paper is already available.

### 2015-07-24 security_scoring.tex
Adjust some TODOs
Tweaks based on comments from Alastair

@@
   \item[update $u$] The proportion of devices that run the latest version of Android shipped to any device produced by that device manufacturer.
-This is a measure of internal updatedness, a low score would mean many devices are being left behind.
+  This is a measure of internal updatedness, so a low score would mean many devices are being left behind.
   This assumes that newer versions are better with stronger security.
@@
 Since $m$ is an unbounded positive real number, we map it into the range (0--1].
-This gives us Equation~\ref{eq:score}:
+This gives us the FUM score:
 
 \begin{equation}
-\mathrm{score} = 4 \cdot f + 3 \cdot u + 3 \cdot \frac{2}{1+e^m} \label{eq:score}
+\mathrm{FUM\ score} = 4 \cdot f + 3 \cdot u + 3 \cdot \frac{2}{1+e^m}
 \end{equation}
@@
 However, it could be that the entity has released a fix to some devices but we have not yet observed a device with that fix.
-So the uncertainty in each vulnerability truly being outstanding is the probability of not having observed a fixed device if a fixed device existed.
+So the uncertainty in our measurement is the probability of not having observed a fixed device if a fixed device existed.
 We assume that if the fix has been released then at least \daMeanOutstandingProbability\ of devices have the fix.\footnote{The selection of \daMeanOutstandingProbability\ is arbitrary, but if a version has only been deployed to a smaller proportion then it does not have a substantial penetration.}
@@
 This shows that $m$ tends to drag down scores.
-\todo{Give some of the actual orderings as well for the most important ones - no space}
-\todo{Sensitivity of time based vs. global scores}
+\todolater{Give some of the actual orderings as well for the most important ones - no space}
+\todolater{Sensitivity of time based vs. global scores}
 \todo{Reviewer 3 thinks we should use the Spearman rank correlation coefficient instead, and they might be right, currently being computed}
@@
 We could normalise the scores for market penetration and so give a score reflecting the risk posed by that manufacturer's performance, which would tend to decrease the difference between manufacturers in our current scoring.
-However we have only considered vulnerabilities affecting all Android devices, not those which are specific to manufacturers and so the aggregation point is all Android, not a specific manufacturer.\todo{Does that even make sense?}
 Additionally since our scores are provided so that customers can chose which devices to buy then it is the marginal risk to that individual of that device which is of interest rather than the aggregate risk to all users.

### 2015-07-24 spsm04-thomas.tex
Adjust some TODOs
Tweaks based on comments from Alastair

@@
 For example, CESG, which advises the UK government on how to secure its computer systems, recommends picking Android device models from device manufactures that are good at promptly shipping security updates, but it does not state which device manufacturers these are~\cite{CESG2013}.
-We are also collaborating with a FTSE 100 company to give them concrete information to answer this question.
+We are also collaborating with a FTSE 100 company to help answer this question.
 To address this, we developed a scoring system and provide numbers on the historic performance of device models in the \da\ project in~\S\ref{sec:security_scoring}.
@@
 
-The \emph{dynamic code loading} attack vector is to make an existing app download and execute code at runtime.
+The \emph{dynamic code loading} attack vector occurs when an existing app download and execute code at runtime.
 The most direct method is to upload to a marketplace a seemingly innocent app that dynamically loads malicious code, either as additional davik bytecode, as a native library or by embedding an interpreter and executing received instructions.
 Neither static nor dynamic analysis of this app will uncover any malicious code, since it does not exist in the app.
-The marketplace can try to detect explicit use of dynamic code loading, however there are ways to dynamically load code which are hard to detect even on a platform such as iOS, which does not permit dynamic code loading, a Return-Oriented Programming (ROP) attack is relatively easy if the attacker creates an app with carefully crafted flaws~\cite{Wang2013a}.
+The marketplace can try to detect explicit use of dynamic code loading, however there are ways to dynamically load code which are hard to detect even on a platform such as iOS, which does not permit dynamic code loading.
+For example, a Return-Oriented Programming (ROP) attack is relatively easy if the attacker creates an app with carefully crafted flaws~\cite{Wang2013a}.
 
-The \emph{injection} attack vector is that the attacker injects malicious code directly into existing code already on the handset. 
+The \emph{injection} attack vector occurs when the attacker injects malicious code directly into existing code already on the handset. 
 For example, the addJavascriptInterface vulnerability (CVE-2012-6636) allows JavaScript running in an Android WebView to execute arbitrary code as the vulnerable app's user.
 The fix for the addJavascriptInterface vulnerability breaks backwards compatibility and requires a two sided fix.
-While the fix was released in December 2012, on \daGPAPISeventeenLaterDate, \daGPAPISeventeenEarlierProportion\ of handsets connecting to Google Play were still vulnerable to this attack~\cite{Thomas2015a}.
+While the fix was released in December 2012, by \daGPAPISeventeenLaterDateMonth, \daGPAPISeventeenEarlierProportion\ of handsets connecting to Google Play were still vulnerable to this attack~\cite{Thomas2015a}.
 
@@
 
-Unless otherwise stated, we use the \daNumVulnsUsed\ vulnerabilities shown in Table~\ref{tab:andvulns} in our analysis.
+We use the \daNumVulnsUsed\ vulnerabilities shown in Table~\ref{tab:andvulns} in our analysis.
 We have chosen these vulnerabilities to fit the attack vectors introduced in \S\ref{sec:threatmodel}.
@@
 The code for each Android release or update is kept secret until after a binary release has been published.\footnote{\url{https://source.android.com/source/code-lines.html}}
-Device manufacturers are presumed to receive advanced access in order to prepare handsets so they can customise it before passing it on to the network operator.
-The network operator may then make or request further customisations and perform further testing before shipping the update to the device.
+Device manufacturers receive advanced access in order to prepare handsets.
+The network operator may then make or request customisations and perform testing before shipping the update to the device.
 Sometimes device manufactures ship updates directly to the user without involving the network operator.
@@
 The versions included in Android were about half a year old when the first version of Android containing it was released.
-
-We have quantified the complexity due to the large number of participants in the ecosystem, the many updates flowing between them and the high latency of updates reaching devices.
 
@@
 We used \da\ to record data on the OS versions of Android installed on devices.
-Nappa et al.
-performed a similar analysis using WINE to record data on installed versions of Windows client programs and used cleaned NVD data rather than \avo\ for vulnerability data, they did not assign client applications scores~\cite{Nappa2015}.
+Nappa et al.\ performed a similar analysis using WINE to record data on installed versions of Windows client programs and used cleaned NVD data rather than \avo\ for vulnerability data, they did not assign client applications scores~\cite{Nappa2015}.
 
@@
 The Rwanda and FTSE data cannot be made available due to the terms under which they were obtained.
-\todo{We need to submit this data to DSpace and insert a reference.}
 

### 2015-07-24 security_scoring.tex
More minor tweaks from Alastair

@@
 So the uncertainty in our measurement is the probability of not having observed a fixed device if a fixed device existed.
-We assume that if the fix has been released then at least \daMeanOutstandingProbability\ of devices have the fix.\footnote{The selection of \daMeanOutstandingProbability\ is arbitrary, but if a version has only been deployed to a smaller proportion then it does not have a substantial penetration.}
+We assume that if the fix has been released then at least \daMeanOutstandingProbability\ of devices have the fix.
+This represents a trade-off between a proportion so small that the fix has not really been deployed and a reasonable estimate of the error.
 This gives an uncertainty of \daMeanOutstandingUncertaintyEquation\ where $n$ is the number of devices contributing to that day's data for each vulnerability outstanding each day.
 The Python \texttt{uncertainties} library was used to propagate uncertainties through calculations.
-This does not capture systematic errors such as those due to not including vulnerabilities which do affect devices, but for comparison that does not matter.
-However, we do not include manufacturer specific vulnerabilities that will influence comparisons, but we expect that performance in fixing manufacturer specific vulnerabilities is strongly correlated with performance fixing vulnerabilities affecting all of Android.
+This does not capture systematic errors.
+For example, we do not include manufacturer specific vulnerabilities, but we expect that performance in fixing manufacturer specific vulnerabilities is strongly correlated with performance fixing vulnerabilities affecting all of Android.
 
@@
 However there are a wide variety of scores depending on the source of the device.
-There is anecdotal evidence that Google's Nexus devices are better at getting updates than other Android devices because Google makes the original updates and ships them to its devices\footnote{\url{http://www.howtogeek.com/139391/htg-explains-why-android-geeks-buy-nexus-devices/}} (but the plural of anecdote is not data).
+There is anecdotal evidence that Google's Nexus devices are better at getting updates than other Android devices because Google makes the original updates and ships them to its devices\footnote{\url{http://www.howtogeek.com/139391/htg-explains-why-android-geeks-buy-nexus-devices/}}.
 Table~\ref{tab:sec_summary} shows that this is the case with Nexus devices getting much better scores than non-Nexus devices and Figure~\ref{fig:security_score_summary} shows this holds true over time.
@@
 Figure~\ref{fig:full_version_comp}b shows the full version distribution for \emph{HTC Desire HD A9191}, which used to be our worst model and for which we have more historical data; it shows how it received one update at the beginning of 2012, which was deployed fairly rapidly to most devices, but received no further updates.
-Figure~\ref{fig:full_version_comp}a shows the same information for \emph{\daSecScoreBestmodel} but it is very different with \daSecScoreBestmodelNumFullVersions\ different versions, some of which were only deployed to small numbers of devices, but the distribution for all devices regularly and rapidly transitions from one version to another before ending up on `4.3 JWR66Y'.
+Figure~\ref{fig:full_version_comp}a shows the same information for \emph{\daSecScoreBestmodel} which received \daSecScoreBestmodelNumFullVersions\ different versions, some of which were only deployed to small numbers of devices, but the distribution for all devices regularly and rapidly transitions from one version to another before ending up on `4.3 JWR66Y'.
 Both \emph{Galaxy Nexus} and \emph{HTC Desire HD A9191} device models start off with the full version string of `2.3.3 GRI40' but the \emph{\daSecScoreBestmodel} receives many more updates over the same time period.
@@
 
-We also analysed different network operators, for the \daNumSigOperators\ network operators with a significant presence in our data.
+We also analysed the \daNumSigOperators\ network operators with a significant presence in our data.
 Table~\ref{tab:sec_operator} shows the results \emph{\daSecScoreBestoperator} (\daSecScoreBestoperatorScore\ out of 10) scoring highest and \emph{\daSecScoreWorstoperator} (\daSecScoreWorstoperatorScore\ out of 10) scoring lowest.
-However, the score of a network operator is affected by the device manufacturers of the devices which are in use on its network.
+However, the score of a network operator is affected by the manufacturers of the devices which are in use on its network.
 This is in turn affected by both the device models a network operator offers to users and upon user's choice of device models.
@@
 This is not an unobtainable number but it is two orders of magnitude more than is available in \da.
+\todo{Could abbreviate if necessary}
 %We do not attempt to disambiguate users behaviour in whether they install updates from network operators using rolling upgrades.
@@
 This would increase the size of our data set and would require providing genuinely good security to some users.
-Some active attacks like blocking access to the \da\ servers from the mobile data network would not be effective as \da\ would retry on Wi-Fi, others like denial of service attacks on the \da\ servers might be effective but illegal, however some entities might be able to force the uninstallation of the app from all devices.
+Some active attacks like blocking access to the \da\ servers from the mobile data network would not be effective as \da\ would retry on Wi-Fi.
+Other denial of service attacks on the \da\ servers might be effective but illegal.
+Some entities might be able to force the uninstallation of the app from all devices.
 Therefore, our score is secure against passive gaming attacks which change the measured distribution, but would require active defence against active gaming attacks, which target the measurement devices.

### 2015-07-24 spsm04-thomas.tex
More minor tweaks from Alastair

@@
 However, we have collated the API version information that Google has published most months since 2009.\footnote{\url{http://androidvulnerabilities.org/play/historicplaydashboard}}
-While API versions are too coarse grained to use for security update detection they are closely related to OS versions and so if the \da\ data on API versions are similar to the Google data on API versions then the \da\ data on OS versions should be representative.
+While API versions are too coarse grained to use for security update detection they are closely related to OS versions.
+If the \da\ data on API versions are similar to the Google data on API versions then the \da\ data on OS versions should be representative.
 We compared the data from Google and from \da\ and they are similar.
@@
 We hope that this analysis will encourage device manufacturers and network operators to improve the support they provide for devices after sale.
+\todo{Rewrite conclusion}
 
@@
 Thanks to \href{https://www.cl.cam.ac.uk/~sa497/}{Sherif Akoush} and \href{http://orcid.org/0000-0003-0740-8650}{Ripduman Sohan} for supplying the Rwanda data and to staff at the FTSE 100 company for supplying their mobile device management data.
-Thanks to our anonymous reviewers for their excellent insightful comments.
+Thanks to our anonymous reviewers for their insightful comments.
 }

### 2015-07-24 spsm04-thomas.tex
Reword dataset to make it look nicer
CESG have since certified a device model

@@
 Currently, corporate and public sector buyers are encouraged to purchase secure devices, but we have found little concrete guidance on the specific makes and models providing the best security.
-For example, CESG, which advises the UK government on how to secure its computer systems, recommends picking Android device models from device manufactures that are good at promptly shipping security updates, but it does not state which device manufacturers these are~\cite{CESG2013}.
+For example, CESG, which advises the UK government on how to secure its computer systems, recommends picking Android device models from device manufactures that are good at promptly shipping security updates, but it does not state which device manufacturers these are~\cite{CESG2013} though they have since certified one device model~\cite{CESG2015}.
 We are also collaborating with a FTSE 100 company to help answer this question.
@@
 Data from \da\ and \avo\ used in this paper is already available.
-The Rwanda and FTSE data cannot be made available due to the terms under which they were obtained.
+The Rwanda and FTSE data cannot be made available due to the terms they were obtained under.
 

### 2015-07-24 spsm04-thomas.tex
Tweaks

@@
 \begin{abstract}
-The security of Android depends on security updates for critical vulnerabilities being shipped to devices.
+The security of Android depends on timely delivery of security updates for critical vulnerabilities .
 We define the FUM security metric to rank the performance of device manufacturers and network operators, based on their provision of updates and exposure to critical vulnerabilities.
@@
 The security of Android relies on many open source projects, including the Li\-nux kernel, Open\-SSL and Bouncy\-Castle, as well as Google, who build the core platform.
-In addition, device manufacturers (we know of \daNumManufacturers\footnote{\label{foot:dadata}We computed this from the \da~\cite{Wagner2013} data, see \S\ref{sec:android_update_process}.}) and network operators (\daNumOperators\dafoot) may make further modifications before devices are shipped to customers.
+In addition, device manufacturers (we know of \daNumManufacturers) and network operators (\daNumOperators) may make further modifications before devices are shipped to customers.
 Understanding this ecosystem is important because device manufacturers have introduced additional vulnerabilities in the past~\cite{Grace2012}.

### 2015-07-24 security_scoring.tex
Start combining sections to produce the new security metrics section

@@
-\section{Comparing device manufacturers, device models and network operators}
+\section{Security metrics}\label{sec:androidvulnerability}
+\begin{figure}
+\centering
+\includegraphics[width=\columnwidth]{figures/proportioninsecure}
+\caption{Proportion of devices running insecure, maybe secure and secure versions of Android.}
+\label{fig:proportioninsecure}
+\end{figure}
+Table~\ref{tab:andvulns} lists the \daNumVulnsUsed\ vulnerabilities used, the red vertical lines are caused by their discovery and the most important are annotated.
+A device is insecure if it is running a vulnerable version of Android and the device has not received an update which might fix it;
+it is maybe secure if it is running a vulnerable version and but received an update which could have fixed the vulnerability if it contained a backported fix;
+and it is secure if it is running a secure version.
+Combining the data on critical vulnerabilities and versions of Android allows us to plot Figure~\ref{fig:proportioninsecure}, initially all devices are \emph{maybe secure} (yellow) since \da\ does not have historical data prior to May 2011.
+This means we cannot distinguish between devices which are running a version of Android which is known to be vulnerable from one which may have received a backported fix.
+This demonstrates the importance of a longitudinal study: this type of analysis requires years of data.
+Once \avovuln{zergRush} was discovered in October 2011 then most devices are recorded as \emph{insecure} (red) as they were exposed to that vulnerability.
+The remaining devices were already running a version of Android which fixed the \avovuln{zergRush} vulnerability and are therefore marked as \emph{secure} (green).
+From October 2011 until the discovery of \avovuln[APK_duplicate_file]{APK duplicate file} in February 2013 the graph shows progressive improvement as devices are upgraded.
+This means more and more devices are marked as \emph{secure} because they are now running a secure version of Android, or marked as \emph{maybe secure} because they received an OS update that did not update to a known-good version of Android but which may still have included a backport of a fix, as the update was made available after the vulnerability was disclosed.
+From February 2013 onwards regular discovery of critical vulnerabilities ensures that most devices are exposed to known critical vulnerabilities.
+This gives us an average exposure to critical vulnerabilities of Android devices of \daMeanInsecurityPerc.
+
+\subsection{Comparing device manufacturers, device models and network operators}
 \label{sec:security_scoring}\label{sec:exp:security_score}
@@
 
-\subsection{Gaming the score}
+\subsection{Gaming the score}\label{sec:gaming}
 If the comparative data given here is used to influence purchasing decisions then entities in the Android ecosystem might try to game the score rather than genuinely improve security.

### 2015-07-24 spsm04-thomas.tex
Start combining sections to produce the new security metrics section

@@
 
-By producing the FUM scoring metric which compares different entities such as device manufacturers, network operators or specific device models, we can provide information that allows device purchasers to select devices with better security.
+We propose scoring metrics to compare compares different entities such as device manufacturers, network operators or specific device models:
+$f$ the proportion of running devices free from critical vulnerabilities over time;
+$u$ the proportion of devices that run the latest version of Android shipped to any device produced by that device manufacturer;
+$m$ the mean number of outstanding vulnerabilities affecting devices not fixed on any device shipped by the device manufacturer;
+and a composite score FUM which is hard to game (\S\ref{sec:gaming}).
+This information allows device purchasers to select devices with better security.
 Currently, corporate and public sector buyers are encouraged to purchase secure devices, but we have found little concrete guidance on the specific makes and models providing the best security.
@@
 
-\begin{figure}
-\centering
-\includegraphics[width=\columnwidth]{figures/proportioninsecure}
-\caption{Proportion of devices running insecure, maybe secure and secure versions of Android.
-Table~\ref{tab:andvulns} lists the \daNumVulnsUsed\ vulnerabilities used, the red vertical lines are caused by their discovery and the most important are annotated.
-A device is insecure if it is running a vulnerable version of Android and the device has not received an update which might fix it;
-it is maybe secure if it is running a vulnerable version and but received an update which could have fixed the vulnerability if it contained a backported fix;
-and it is secure if it is running a secure version.
-}
-\label{fig:proportioninsecure}
-\end{figure}
 By using data from \da~\cite{Wagner2013} we can examine the versions of Android running on user devices over time and through comparison with data on which versions of Android were vulnerable to different critical vulnerabilities~\cite{androidvulnerabilities.org} we can plot the exposure of Android devices to known critical vulnerabilities.
@@
 
-\section{Vulnerability of Android dev\-ices}\label{sec:androidvulnerability}
-Combining the data on critical vulnerabilities and versions of Android allows us to plot Figure~\ref{fig:proportioninsecure}, initially all devices are \emph{maybe secure} (yellow) since \da\ does not have historical data prior to May 2011.
-This means we cannot distinguish between devices which are running a version of Android which is known to be vulnerable from one which may have received a backported fix.
-This demonstrates the importance of a longitudinal study: this type of analysis requires years of data.
-Once \avovuln{zergRush} was discovered in October 2011 then most devices are recorded as \emph{insecure} (red) as they were exposed to that vulnerability.
-The remaining devices were already running a version of Android which fixed the \avovuln{zergRush} vulnerability and are therefore marked as \emph{secure} (green).
-From October 2011 until the discovery of \avovuln[APK_duplicate_file]{APK duplicate file} in February 2013 the graph shows progressive improvement as devices are upgraded.
-This means more and more devices are marked as \emph{secure} because they are now running a secure version of Android, or marked as \emph{maybe secure} because they received an OS update that did not update to a known-good version of Android but which may still have included a backport of a fix, as the update was made available after the vulnerability was disclosed.
-From February 2013 onwards regular discovery of critical vulnerabilities ensures that most devices are exposed to known critical vulnerabilities.
-This gives us an average exposure to critical vulnerabilities of Android devices of \daMeanInsecurityPerc.
+\section{Device Analyzer data}\label{sec:da}
+
+In this analysis we use historical data collected by the \da\ project~\cite{Wagner2013}.
+\da\ collects data\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/collected.html}} from study participants who install the Android app from Google Play.
+Most study participants allow researchers around the world to access a subset of their device data, including the data presented here.
+
+We extracted the build string and API version for each device each day.
+The build string is a user-readable version string.
+The API version is a positive integer that increases when new features are added to the API.
+Consequently security (bug) fixes do not always result in a change in the API version.
+Fortunately most (\daOSVersionPercValidLines) entries in these data have a build string of the form `x.y.z opaque\_marker' and so it is possible to extract the Android version number `x.y.z'.
+On a large proportion of devices `opaque\_marker' is a well defined build number\footnote{\url{https://source.android.com/source/build-numbers.html}} however different device manufacturers use different schema.
+Google provides API version distribution information but not the OS and build version information we need.
+We verify that the \da\ data is representative in~\S\ref{sec:representative}.
+
+\da\ has collected data from \daNumDevices\ devices with a total of \daDeviceDays\footnote{Here we are only counting devices and days for which we have valid OS version data.} device days.
+The majority of devices only contribute data for a short period of time, however \daMonthsDevices\ devices have contributed data for more than \daMonths~months.
 
@@
 The versions included in Android were about half a year old when the first version of Android containing it was released.
-
-\section{Device Analyzer data}\label{sec:da}
-
-In this analysis we use historical data collected by the \da\ project~\cite{Wagner2013}.
-\da\ collects data\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/collected.html}} from study participants who install the Android app from Google Play.
-Most study participants allow researchers around the world to access a subset of their device data, including the data presented here.
-
-We extracted the build string and API version for each device each day.
-The build string is a user-readable version string.
-The API version is a positive integer that increases when new features are added to the API.
-Consequently security (bug) fixes do not always result in a change in the API version.
-Fortunately most (\daOSVersionPercValidLines) entries in these data have a build string of the form `x.y.z opaque\_marker' and so it is possible to extract the Android version number `x.y.z'.
-On a large proportion of devices `opaque\_marker' is a well defined build number\footnote{\url{https://source.android.com/source/build-numbers.html}} however different device manufacturers use different schema.
-Google provides API version distribution information but not the OS and build version information we need.
-We verify that the \da\ data is representative in~\S\ref{sec:representative}.
-
-\da\ has collected data from \daNumDevices\ devices with a total of \daDeviceDays\footnote{Here we are only counting devices and days for which we have valid OS version data.} device days.
-The majority of devices only contribute data for a short period of time, however \daMonthsDevices\ devices have contributed data for more than \daMonths~months.
 

### 2015-07-27 dastats.tex
./update.sh

@@
 \newcommand{\daStartDate}{2011-07-01}
-\newcommand{\daEndDate}{2015-07-15}
+\newcommand{\daEndDate}{2015-07-19}
 \newcommand{\daFullDeployedAt}{95\%}
@@
 \newcommand{\daNumOperators}{1\,460}
-\newcommand{\daNumSigOperators}{15}
+\newcommand{\daNumSigOperators}{14}
 \newcommand{\daNumOSDevices}{20\,400}
@@
 \newcommand{\daMeanOutstandingVulnerabilities}{$0.53 \pm 0.0$}
-\newcommand{\daUpdatednessNominal}{0.0521}
-\newcommand{\daUpdatedness}{$0.0521 \pm 0.0$}
-\newcommand{\daUpdatednessPercNominal}{5.21\%}
-\newcommand{\daUpdatednessPerc}{$5.21 \pm 0.0\%$}
+\newcommand{\daUpdatednessNominal}{0.0523}
+\newcommand{\daUpdatedness}{$0.0523 \pm 0.0$}
+\newcommand{\daUpdatednessPercNominal}{5.23\%}
+\newcommand{\daUpdatednessPerc}{$5.23 \pm 0.0\%$}
 \newcommand{\daUpdatednessPercTwosfNominal}{5.2\%}
@@
 \newcommand{\daSecurityScore}{$2.87 \pm 0.0$}
-\newcommand{\daCurrentMeanInsecurityPercNominal}{79.6\%}
-\newcommand{\daCurrentMeanInsecurityPerc}{$79.6 \pm 0.0\%$}
-\newcommand{\daCurrentMeanInsecurityPercTwosfNominal}{80\%}
-\newcommand{\daCurrentMeanInsecurityPercTwosf}{$80 \pm 0\%$}
-\newcommand{\daCurrentMeanSecurityPercNominal}{20.4\%}
-\newcommand{\daCurrentMeanSecurityPerc}{$20.4 \pm 0.1\%$}
-\newcommand{\daCurrentMeanSecurityPercTwosfNominal}{20\%}
-\newcommand{\daCurrentMeanSecurityPercTwosf}{$20 \pm 0\%$}
-\newcommand{\daCurrentVulnFreeNominal}{0.204}
-\newcommand{\daCurrentVulnFree}{$0.204 \pm 0.001$}
-\newcommand{\daCurrentMeanOutstandingVulnerabilities}{$0.123 \pm 0.0$}
-\newcommand{\daCurrentUpdatednessNominal}{0.0536}
-\newcommand{\daCurrentUpdatedness}{$0.0536 \pm 0.0$}
-\newcommand{\daCurrentUpdatednessPercNominal}{5.36\%}
-\newcommand{\daCurrentUpdatednessPerc}{$5.36 \pm 0.0\%$}
-\newcommand{\daCurrentUpdatednessPercTwosfNominal}{5.4\%}
-\newcommand{\daCurrentUpdatednessPercTwosf}{$5.4 \pm 0.0\%$}
-\newcommand{\daCurrentSecurityScoreNominal}{3.79}
-\newcommand{\daCurrentSecurityScore}{$3.79 \pm 0.0$}
-\newcommand{\daDailyParticipationNominal}{898}
-\newcommand{\daDailyParticipation}{$898 \pm 245$}
+\newcommand{\daCurrentMeanInsecurityPercNominal}{79.5\%}
+\newcommand{\daCurrentMeanInsecurityPerc}{$79.5 \pm 0.0\%$}
+\newcommand{\daCurrentMeanInsecurityPercTwosfNominal}{79\%}
+\newcommand{\daCurrentMeanInsecurityPercTwosf}{$79 \pm 0\%$}
+\newcommand{\daCurrentMeanSecurityPercNominal}{20.5\%}
+\newcommand{\daCurrentMeanSecurityPerc}{$20.5 \pm 0.2\%$}
+\newcommand{\daCurrentMeanSecurityPercTwosfNominal}{21\%}
+\newcommand{\daCurrentMeanSecurityPercTwosf}{$21 \pm 0\%$}
+\newcommand{\daCurrentVulnFreeNominal}{0.205}
+\newcommand{\daCurrentVulnFree}{$0.205 \pm 0.002$}
+\newcommand{\daCurrentMeanOutstandingVulnerabilities}{$0.119 \pm 0.0$}
+\newcommand{\daCurrentUpdatednessNominal}{0.0561}
+\newcommand{\daCurrentUpdatedness}{$0.0561 \pm 0.0016$}
+\newcommand{\daCurrentUpdatednessPercNominal}{5.61\%}
+\newcommand{\daCurrentUpdatednessPerc}{$5.61 \pm 0.16\%$}
+\newcommand{\daCurrentUpdatednessPercTwosfNominal}{5.6\%}
+\newcommand{\daCurrentUpdatednessPercTwosf}{$5.6 \pm 0.0\%$}
+\newcommand{\daCurrentSecurityScoreNominal}{3.81}
+\newcommand{\daCurrentSecurityScore}{$3.81 \pm 0.0$}
+\newcommand{\daDailyParticipationNominal}{896}
+\newcommand{\daDailyParticipation}{$896 \pm 249$}
 \newcommand{\daNumFullVersions}{1\,270}
@@
 \newcommand{\daOSCurveHalfDeployedYears}{$0.897 \pm 0.252$}
-\newcommand{\daOSCurveHalfDeployedNominal}{327}
-\newcommand{\daOSCurveHalfDeployed}{$327 \pm 91$}
+\newcommand{\daOSCurveHalfDeployedNominal}{328}
+\newcommand{\daOSCurveHalfDeployed}{$328 \pm 91$}
 \newcommand{\daOSCurveFullDeployedYearsNominal}{3.11}
 \newcommand{\daOSCurveFullDeployedYears}{$3.11 \pm 71$}
-\newcommand{\daOSCurveFullDeployedNominal}{1\,130}
-\newcommand{\daOSCurveFullDeployed}{$1\,130 \pm 260\,000$}
-\newcommand{\daAPICurveFitParamFirst}{99.7}
+\newcommand{\daOSCurveFullDeployedNominal}{1\,140}
+\newcommand{\daOSCurveFullDeployed}{$1\,140 \pm 260\,000$}
+\newcommand{\daAPICurveFitParamFirst}{99.4}
 \newcommand{\daAPICurveFitParamSecond}{\num{0.00285}}
@@
 \newcommand{\daSecScoreWorstmanufacturerNumFullVersions}{21}
-\newcommand{\daTabSecScoresmodel}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline Galaxy Nexus & $0.50 \pm 0.00$ & $0.54 \pm 0.01$ & $1.53 \pm 0.04$ & $4.71 \pm 0.04$ \\  Nexus 4 & $0.30 \pm 0.00$ & $0.82 \pm 0.01$ & $6.06 \pm 0.09$ & $3.69 \pm 0.04$ \\  Nexus 7 & $0.26 \pm 0.00$ & $0.74 \pm 0.01$ & $5.92 \pm 0.09$ & $3.24 \pm 0.04$ \\  other & $0.10 \pm 0.00$ & $0.14 \pm 0.00$ & $0.53 \pm 0.00$ & $3.03 \pm 0.00$ \\  Desire HD & $0.08 \pm 0.00$ & $0.05 \pm 0.00$ & $0.38 \pm 0.02$ & $2.91 \pm 0.04$ \\  HTC Sensation & $0.35 \pm 0.00$ & $0.01 \pm 0.01$ & $1.57 \pm 0.05$ & $2.44 \pm 0.05$ \\  GT-I9100 & $0.22 \pm 0.00$ & $0.02 \pm 0.00$ & $1.23 \pm 0.02$ & $2.27 \pm 0.02$ \\  HTC Desire S & $0.02 \pm 0.00$ & $0.02 \pm 0.00$ & $1.00 \pm 0.06$ & $1.74 \pm 0.07$ \\  GT-N7000 & $0.25 \pm 0.00$ & $0.00 \pm 0.00$ & $2.52 \pm 0.05$ & $1.43 \pm 0.02$ \\  GT-P1000 & $0.01 \pm 0.00$ & $0.00 \pm 0.01$ & $1.79 \pm 0.06$ & $0.90 \pm 0.05$ \\  GT-I9300 & $0.13 \pm 0.00$ & $0.01 \pm 0.00$ & $6.23 \pm 0.04$ & $0.58 \pm 0.01$ \\  GT-I9505 & $0.03 \pm 0.00$ & $0.13 \pm 0.00$ & $6.82 \pm 0.07$ & $0.52 \pm 0.01$ \\  HTC Desire HD & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & $3.03 \pm 0.05$ & $0.28 \pm 0.03$ \\  GT-N7100 & $0.06 \pm 0.00$ & $0.00 \pm 0.01$ & $6.93 \pm 0.08$ & $0.24 \pm 0.02$ \\  Symphony W68 & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & $11.00 \pm 0.12$ & $0.00 \pm 0.03$ \\ \end{tabular} \caption{Security scores for models} \label{tab:sec_model} \end{table*}}
+\newcommand{\daTabSecScoresmodel}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline Galaxy Nexus & $0.50 \pm 0.00$ & $0.54 \pm 0.01$ & $1.53 \pm 0.04$ & $4.71 \pm 0.04$ \\  Nexus 4 & $0.30 \pm 0.00$ & $0.82 \pm 0.01$ & $6.06 \pm 0.09$ & $3.69 \pm 0.04$ \\  Nexus 7 & $0.26 \pm 0.00$ & $0.74 \pm 0.01$ & $5.92 \pm 0.09$ & $3.25 \pm 0.04$ \\  other & $0.10 \pm 0.00$ & $0.14 \pm 0.00$ & $0.53 \pm 0.00$ & $3.03 \pm 0.00$ \\  Desire HD & $0.08 \pm 0.00$ & $0.05 \pm 0.00$ & $0.38 \pm 0.02$ & $2.91 \pm 0.04$ \\  HTC Sensation & $0.35 \pm 0.00$ & $0.01 \pm 0.01$ & $1.57 \pm 0.05$ & $2.44 \pm 0.05$ \\  GT-I9100 & $0.22 \pm 0.00$ & $0.02 \pm 0.00$ & $1.23 \pm 0.02$ & $2.27 \pm 0.02$ \\  HTC Desire S & $0.02 \pm 0.00$ & $0.02 \pm 0.00$ & $1.00 \pm 0.06$ & $1.74 \pm 0.07$ \\  GT-N7000 & $0.25 \pm 0.00$ & $0.00 \pm 0.00$ & $2.52 \pm 0.05$ & $1.43 \pm 0.02$ \\  GT-P1000 & $0.01 \pm 0.00$ & $0.00 \pm 0.01$ & $1.79 \pm 0.06$ & $0.90 \pm 0.05$ \\  GT-I9300 & $0.13 \pm 0.00$ & $0.01 \pm 0.00$ & $6.23 \pm 0.04$ & $0.58 \pm 0.01$ \\  GT-I9505 & $0.03 \pm 0.00$ & $0.13 \pm 0.00$ & $6.82 \pm 0.07$ & $0.52 \pm 0.01$ \\  HTC Desire HD & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & $3.03 \pm 0.05$ & $0.28 \pm 0.03$ \\  GT-N7100 & $0.06 \pm 0.00$ & $0.00 \pm 0.01$ & $6.93 \pm 0.08$ & $0.24 \pm 0.02$ \\  Symphony W68 & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & $11.00 \pm 0.12$ & $0.00 \pm 0.03$ \\ \end{tabular} \caption{Security scores for models} \label{tab:sec_model} \end{table*}}
 \newcommand{\daSecScoreBestmodel}{Galaxy Nexus}
@@
 \newcommand{\daTabSecScoressummary}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline nexus & $0.39 \pm 0.00$ & $0.48 \pm 0.00$ & $0.56 \pm 0.01$ & $5.17 \pm 0.02$ \\  notnexus & $0.10 \pm 0.00$ & $0.02 \pm 0.00$ & $0.53 \pm 0.00$ & $2.70 \pm 0.00$ \\ \end{tabular} \caption{Security scores for nexus} \label{tab:sec_summary} \end{table*}}
-\newcommand{\daTabSecScoresoperator}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline O2 uk & $0.27 \pm 0.00$ & $0.12 \pm 0.00$ & $0.37 \pm 0.02$ & $3.87 \pm 0.03$ \\  T-Mobile & $0.22 \pm 0.00$ & $0.18 \pm 0.00$ & $0.40 \pm 0.01$ & $3.81 \pm 0.02$ \\  Orange & $0.22 \pm 0.00$ & $0.10 \pm 0.00$ & $0.36 \pm 0.02$ & $3.65 \pm 0.04$ \\  Sprint & $0.18 \pm 0.00$ & $0.11 \pm 0.00$ & $0.43 \pm 0.02$ & $3.42 \pm 0.03$ \\  3 & $0.20 \pm 0.00$ & $0.09 \pm 0.00$ & $0.47 \pm 0.02$ & $3.39 \pm 0.03$ \\  Vodafone uk & $0.14 \pm 0.00$ & $0.13 \pm 0.00$ & $0.52 \pm 0.03$ & $3.17 \pm 0.04$ \\  AT\&T & $0.14 \pm 0.00$ & $0.08 \pm 0.00$ & $0.43 \pm 0.02$ & $3.13 \pm 0.02$ \\  unknown & $0.11 \pm 0.00$ & $0.20 \pm 0.00$ & $0.84 \pm 0.01$ & $2.88 \pm 0.02$ \\  Verizon & $0.19 \pm 0.00$ & $0.09 \pm 0.00$ & $0.82 \pm 0.02$ & $2.84 \pm 0.02$ \\  n Telenor & $0.04 \pm 0.00$ & $0.12 \pm 0.00$ & $1.21 \pm 0.02$ & $1.89 \pm 0.02$ \\  Airtel & $0.05 \pm 0.00$ & $0.03 \pm 0.00$ & $1.47 \pm 0.03$ & $1.41 \pm 0.03$ \\  Grameenphone & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & $1.88 \pm 0.02$ & $0.94 \pm 0.01$ \\  Robi & $0.00 \pm 0.00$ & $0.08 \pm 0.00$ & $2.07 \pm 0.04$ & $0.91 \pm 0.03$ \\  banglalink & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & $2.55 \pm 0.04$ & $0.55 \pm 0.02$ \\ \end{tabular} \caption{Security scores for operators} \label{tab:sec_operator} \end{table*}}
+\newcommand{\daTabSecScoresoperator}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline O2 uk & $0.27 \pm 0.00$ & $0.12 \pm 0.00$ & $0.37 \pm 0.02$ & $3.87 \pm 0.03$ \\  T-Mobile & $0.21 \pm 0.00$ & $0.18 \pm 0.00$ & $0.40 \pm 0.01$ & $3.81 \pm 0.02$ \\  Orange & $0.22 \pm 0.00$ & $0.10 \pm 0.00$ & $0.36 \pm 0.02$ & $3.65 \pm 0.04$ \\  Sprint & $0.18 \pm 0.00$ & $0.11 \pm 0.00$ & $0.43 \pm 0.02$ & $3.42 \pm 0.03$ \\  3 & $0.20 \pm 0.00$ & $0.09 \pm 0.00$ & $0.47 \pm 0.02$ & $3.39 \pm 0.03$ \\  Vodafone uk & $0.14 \pm 0.00$ & $0.13 \pm 0.00$ & $0.52 \pm 0.03$ & $3.17 \pm 0.04$ \\  AT\&T & $0.14 \pm 0.00$ & $0.08 \pm 0.00$ & $0.43 \pm 0.02$ & $3.13 \pm 0.02$ \\  unknown & $0.11 \pm 0.00$ & $0.20 \pm 0.00$ & $0.84 \pm 0.01$ & $2.88 \pm 0.02$ \\  Verizon & $0.19 \pm 0.00$ & $0.09 \pm 0.00$ & $0.82 \pm 0.02$ & $2.84 \pm 0.02$ \\  n Telenor & $0.04 \pm 0.00$ & $0.12 \pm 0.00$ & $1.21 \pm 0.02$ & $1.89 \pm 0.02$ \\  Airtel & $0.05 \pm 0.00$ & $0.03 \pm 0.00$ & $1.47 \pm 0.03$ & $1.41 \pm 0.03$ \\  Grameenphone & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & $1.88 \pm 0.02$ & $0.94 \pm 0.01$ \\  Robi & $0.00 \pm 0.00$ & $0.08 \pm 0.00$ & $2.07 \pm 0.04$ & $0.91 \pm 0.03$ \\  banglalink & $0.00 \pm 0.00$ & $0.03 \pm 0.00$ & $2.56 \pm 0.04$ & $0.54 \pm 0.02$ \\ \end{tabular} \caption{Security scores for operators} \label{tab:sec_operator} \end{table*}}
 \newcommand{\daSecScoreBestoperator}{O2 uk}
@@
 \newcommand{\daSecScoreWorstoperator}{banglalink}
-\newcommand{\daSecScoreWorstoperatorScoreNominal}{0.548}
-\newcommand{\daSecScoreWorstoperatorScore}{$0.548 \pm 0.018$}
-\newcommand{\daSecScoreWorstoperatorNumFullVersions}{74}
-\newcommand{\daTabDLDistances}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & summary \\ \hline equal & 0.0 & 0.0667 & 0.0 & 0.0 \\weight $m$ & 0.2 & 0.4 & 0.0714 & 0.0 \\weight $u$ & 0.4 & 0.0667 & 0.143 & 0.0 \\$f$ & 0.4 & 0.667 & 0.5 & 0.0 \\$m$ & 0.6 & 0.733 & 0.357 & 0.5 \\$u$ & 0.9 & 0.667 & 0.786 & 0.0 \\\end{tabular} \caption{Normalised Damerau-Levenshtein distances for different metrics} \label{tab:dl_distances} \end{table}}
-\newcommand{\daTabChangeInScores}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & summary \\ \hline $m$ & $-1.67 \pm 2.25$ & $-0.66 \pm 2.54$ & $-3.3 \pm 1.3$ & $-3.4 \pm 1.8$ \\weight $m$ & $-0.264 \pm 0.276$ & $-0.0991 \pm 0.32$ & $-0.462 \pm 0.167$ & $-0.488 \pm 0.216$ \\equal & $-0.108 \pm 0.062$ & $-0.0368 \pm 0.112$ & $-0.146 \pm 0.045$ & $-0.165 \pm 0.033$ \\weight $u$ & $-0.0608 \pm 0.122$ & $-0.0112 \pm 0.232$ & $0.0238 \pm 0.0807$ & $-0.00602 \pm 0.12$ \\$u$ & $0.367 \pm 1.61$ & $0.219 \pm 2.12$ & $1.55 \pm 0.94$ & $1.42 \pm 1.48$ \\$f$ & $0.976 \pm 0.564$ & $0.331 \pm 1.01$ & $1.31 \pm 0.40$ & $1.48 \pm 0.29$ \\\end{tabular} \caption{Mean change in scores for different metrics} \label{tab:change_in_scores} \end{table*}}
+\newcommand{\daSecScoreWorstoperatorScoreNominal}{0.536}
+\newcommand{\daSecScoreWorstoperatorScore}{$0.536 \pm 0.018$}
+\newcommand{\daSecScoreWorstoperatorNumFullVersions}{73}
+\newcommand{\daTabDLDistances}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & summary \\ \hline equal & 0.0 & 0.0667 & 0.0 & 0.0 \\weight $m$ & 0.2 & 0.4 & 0.0714 & 0.0 \\$f$ & 0.4 & 0.667 & 0.5 & 0.0 \\weight $u$ & 0.4 & 0.0667 & 0.143 & 0.0 \\$m$ & 0.6 & 0.733 & 0.357 & 0.5 \\$u$ & 0.9 & 0.6 & 0.786 & 0.0 \\\end{tabular} \caption{Normalised Damerau-Levenshtein distances for different metrics} \label{tab:dl_distances} \end{table}}
+\newcommand{\daTabChangeInScores}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & summary \\ \hline $m$ & $-1.67 \pm 2.25$ & $-0.66 \pm 2.54$ & $-3.3 \pm 1.3$ & $-3.4 \pm 1.8$ \\weight $m$ & $-0.264 \pm 0.277$ & $-0.0991 \pm 0.321$ & $-0.461 \pm 0.167$ & $-0.488 \pm 0.217$ \\equal & $-0.108 \pm 0.062$ & $-0.0367 \pm 0.112$ & $-0.146 \pm 0.045$ & $-0.165 \pm 0.033$ \\weight $u$ & $-0.0608 \pm 0.122$ & $-0.0112 \pm 0.232$ & $0.0242 \pm 0.0808$ & $-0.00596 \pm 0.12$ \\$u$ & $0.367 \pm 1.61$ & $0.219 \pm 2.12$ & $1.55 \pm 0.94$ & $1.42 \pm 1.48$ \\$f$ & $0.975 \pm 0.564$ & $0.331 \pm 1.01$ & $1.31 \pm 0.41$ & $1.48 \pm 0.3$ \\\end{tabular} \caption{Mean change in scores for different metrics} \label{tab:change_in_scores} \end{table*}}
 \newcommand{\daGPAPICurveFitParamFirst}{80.8}
@@
 \newcommand{\daMeanOutstandingVulnerabilitiesNominal}{0.53}
-\newcommand{\daCurrentMeanOutstandingVulnerabilitiesNominal}{0.123}
+\newcommand{\daCurrentMeanOutstandingVulnerabilitiesNominal}{0.119}
 \newcommand{\daModelHalfDeploymentYears}{0.0822}
@@
 \newcommand{\daGPAPISeventeenLaterDateMonth}{June 2015}
+\newcommand{\daTabSpearmanRanks}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & summary \\ $\pm\sigma$ & 0.211 & 0.169 & 0.175 & 0.632 \\ \hline $u$ & 0.297 & 0.804 & 0.618 & 1.0 \\$m$ & 0.794 & 0.593 & 0.969 & -1.0 \\$f$ & 0.83 & 0.775 & 0.934 & 1.0 \\weight $u$ & 0.939 & 0.996 & 0.991 & 1.0 \\weight $m$ & 0.976 & 0.964 & 0.996 & 1.0 \\equal & 1.0 & 0.996 & 1.0 & 1.0 \\\end{tabular} \caption{Spearman Rank correlation coefficients for different metrics} \label{tab:dl_distances} \end{table}}
+\newcommand{\daStartDateMonth}{July 2011}
+\newcommand{\daEndDateMonth}{July 2015}
+\newcommand{\daSecScoreBestsummary}{nexus}
+\newcommand{\daSecScoreBestsummaryScoreNominal}{5.17}
+\newcommand{\daSecScoreBestsummaryScore}{$5.17 \pm 0.0$}
+\newcommand{\daSecScoreBestsummaryNumFullVersions}{92}
+\newcommand{\daSecScoreWorstsummary}{notnexus}
+\newcommand{\daSecScoreWorstsummaryScoreNominal}{2.7}
+\newcommand{\daSecScoreWorstsummaryScore}{$2.7 \pm 0.0$}
+\newcommand{\daSecScoreWorstsummaryNumFullVersions}{1\,250}

### 2015-07-27 dastats.tex
Fix label on spearmans ranks table

@@
 \newcommand{\daGPAPISeventeenLaterDateMonth}{June 2015}
-\newcommand{\daTabSpearmanRanks}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & summary \\ $\pm\sigma$ & 0.211 & 0.169 & 0.175 & 0.632 \\ \hline $u$ & 0.297 & 0.804 & 0.618 & 1.0 \\$m$ & 0.794 & 0.593 & 0.969 & -1.0 \\$f$ & 0.83 & 0.775 & 0.934 & 1.0 \\weight $u$ & 0.939 & 0.996 & 0.991 & 1.0 \\weight $m$ & 0.976 & 0.964 & 0.996 & 1.0 \\equal & 1.0 & 0.996 & 1.0 & 1.0 \\\end{tabular} \caption{Spearman Rank correlation coefficients for different metrics} \label{tab:dl_distances} \end{table}}
+\newcommand{\daTabSpearmanRanks}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & summary \\ $\pm\sigma$ & 0.211 & 0.169 & 0.175 & 0.632 \\ \hline $u$ & 0.297 & 0.804 & 0.618 & 1.0 \\$m$ & 0.794 & 0.593 & 0.969 & -1.0 \\$f$ & 0.83 & 0.775 & 0.934 & 1.0 \\weight $u$ & 0.939 & 0.996 & 0.991 & 1.0 \\weight $m$ & 0.976 & 0.964 & 0.996 & 1.0 \\equal & 1.0 & 0.996 & 1.0 & 1.0 \\\end{tabular} \caption{Spearman Rank correlation coefficients for different metrics} \label{tab:spearman_ranks} \end{table}}
 \newcommand{\daStartDateMonth}{July 2011}

### 2015-07-27 security_scoring.tex
Specify sig models and sig operators from macros rather than manually so that they are correct

@@
 However our sample size is too small to do that because while we have significant numbers of devices for each of the device models (Table~\ref{tab:sec_model}) and for each of the network operators (Table~\ref{tab:sec_operator}), we would need a significant number of each model in each network operator.
-A large enough sample size would contain at least 100 unique devices each day for each device model for each network operator, hence for our 14 significant models and 13 significant network operators, a daily average of at least 18\,000 unique devices would be needed, for a period of at least a year.
+A large enough sample size would contain at least 100 unique devices each day for each device model for each network operator, hence for our \daNumSigModels significant models and \daNumSigOperators significant network operators, a daily average of at least 18\,000 unique devices would be needed, for a period of at least a year.
 Since the distribution of devices is unlikely to be uniformly distributed across device models and network operators a more realistic estimate is 100\,000 unique devices each day.

### 2015-07-27 dastats.tex
Use Spearmans Rank rather than Damerau-Levenshtein distance

--- 2015-07-27 dastats.tex
@@
 \newcommand{\daGPAPISeventeenLaterDateMonth}{June 2015}
-\newcommand{\daTabSpearmanRanks}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & summary \\ $\pm\sigma$ & 0.211 & 0.169 & 0.175 & 0.632 \\ \hline $u$ & 0.297 & 0.804 & 0.618 & 1.0 \\$m$ & 0.794 & 0.593 & 0.969 & -1.0 \\$f$ & 0.83 & 0.775 & 0.934 & 1.0 \\weight $u$ & 0.939 & 0.996 & 0.991 & 1.0 \\weight $m$ & 0.976 & 0.964 & 0.996 & 1.0 \\equal & 1.0 & 0.996 & 1.0 & 1.0 \\\end{tabular} \caption{Spearman Rank correlation coefficients for different metrics} \label{tab:dl_distances} \end{table}}
+\newcommand{\daTabSpearmanRanks}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & summary \\ $\pm\sigma$ & 0.211 & 0.169 & 0.175 & 0.632 \\ \hline $u$ & 0.297 & 0.804 & 0.618 & 1.0 \\$m$ & 0.794 & 0.593 & 0.969 & -1.0 \\$f$ & 0.83 & 0.775 & 0.934 & 1.0 \\weight $u$ & 0.939 & 0.996 & 0.991 & 1.0 \\weight $m$ & 0.976 & 0.964 & 0.996 & 1.0 \\equal & 1.0 & 0.996 & 1.0 & 1.0 \\\end{tabular} \caption{Spearman Rank correlation coefficients for different metrics.
+The uncertainty is constant for each column but does not take into account the uncertainty in the score which produced the ranking.} \label{tab:spearman_ranks} \end{table}}
 \newcommand{\daStartDateMonth}{July 2011}

### 2015-07-27 security_scoring.tex
Use Spearmans Rank rather than Damerau-Levenshtein distance

@@
 \subsection{Sensitivity of scoring metric}
-\daTabDLDistances
+\daTabSpearmanRanks
 \daTabChangeInScores
-To evaluate whether the ranking of different manufacturers is sensitive to the form of the scoring metric we computed the normalised Damerau-Levenshtein distance~\cite{Bard2007} between the lists ordered using different forms of the scoring metric, this is shown in Table~\ref{tab:dl_distances}.
+To evaluate whether the ranking of different manufacturers is sensitive to the form of the scoring metric we computed the normalised Spearman's Rank correlation coefficient between the lists ordered using different forms of the scoring metric, this is shown in Table~\ref{tab:spearman_ranks}.
 The `equal' metric weights $f$, $u$ and $m$ equally rather than favouring $f$ and makes little difference.
+Similarly weighting $u$ or $m$ more highly rather than $f$ makes little difference.
+While the $f$, $u$ and $m$ components do have some correlation with the overall FUM score, the rankings produced vary substantially.
 Changing the scoring metric also impacts the scores given for each entity Table~\ref{tab:change_in_scores} shows the mean impact on the scores.
@@
 \todolater{Give some of the actual orderings as well for the most important ones - no space}
-\todolater{Sensitivity of time based vs. global scores}
-\todo{Reviewer 3 thinks we should use the Spearman rank correlation coefficient instead, and they might be right, currently being computed}
+\todolater{Sensitivity of time based vs. global scores - no time}
 

### 2015-07-27 spsm04-thomas.tex
Import acknowledgements from the previous paper as that did not get published
Shorten acknowledgements

@@
 Thanks to \href{https://www.cl.cam.ac.uk/~sa497/}{Sherif Akoush} and \href{http://orcid.org/0000-0003-0740-8650}{Ripduman Sohan} for supplying the Rwanda data and to staff at the FTSE 100 company for supplying their mobile device management data.
-Thanks to our anonymous reviewers for their insightful comments.
+Thanks to our anonymous reviewers for their insightful comments; Richard Clayton, the mobile security reading group, Anil Madhavapeddy and Oliver Chick for reading early drafts;
+David Robertson for statistical advice;
+and Laurent Simon, Thomas Coudray, Adrian Taylor, Justin Case, Giant Pune and Khilan Gudka for reporting vulnerabilities.
 }

### 2015-07-27 spsm04-thomas.tex
Remove some words from the related work

@@
 The same technique would be more difficult on phones with limited storage space (as many cheap Android phones have barely enough space to install just the update) but is a plausible improvement for more high-end devices.
-Google is deploying the same silent update technique through Google Play Services,\footnote{\href{http://lifehacker.com/why-google-play-services-are-now-more-important-than-an-975970197}{http://lifehacker.com/why-google-play-services-are-now-more-important-than-an-975970197}} which automatically installs updates for core Google components of Android, this also bypasses the device manufacturer and network operator.
+Google is deploying the same silent update technique through Google Play Services,\footnote{\href{http://lifehacker.com/why-google-play-services-are-now-more-important-than-an-975970197}{http://lifehacker.com/why-google-play-services-are-now-more-important-than-an-975970197}} which automatically installs updates for core Google components of Android, this bypasses the device manufacturer and network operator.
 
@@
 
-The percentage of Android devices running the most recent version (\daUpdatednessPerc) is much less than the rate ($>90$\%) for Windows XP SP2 computers contacting the Microsoft update servers~\cite{Gkantsidis2006}.
-A simple numerical comparison is unfair because only one major OS version was considered in the Microsoft analysis, and data was only collected from computers that contacted the update server, although this was the default.
-More recent data demonstrates the difficulty of upgrading computers between major OS versions, with 27\% of Windows computers running Windows XP in July 2014,\footnote{\url{https://archive.today/PLGxn}} four months after Windows XP stopped receiving security updates.
+The percentage of Android devices running the current version (\daUpdatednessPercNominal) is much less than the rate ($>90$\%) for Windows XP SP2 computers contacting the Microsoft update servers~\cite{Gkantsidis2006}.
+A simple numerical comparison is unfair because only one major OS version was considered in the Microsoft analysis, and data was only collected from computers that contacted the update server (this was the default).
+Later data demonstrates the difficulty of upgrading computers between major OS versions, with 27\% of Windows computers running Windows XP in July 2014,\footnote{\url{https://archive.today/PLGxn}} four months after Windows XP stopped receiving security updates.
 

### 2015-07-27 security_scoring.tex
Rework beginning of security scoring

@@
 \section{Security metrics}\label{sec:androidvulnerability}
+\label{sec:security_scoring}\label{sec:exp:security_score}
+
+To allow buyers of Android devices to purchase those devices with the best security, they need to know how different device manufacturers, device models and network operators compare in terms of the security they provide.
+We propose a method to score a device manufacturer, device model or network operator based on its historic performance at keeping devices up-to-date and fixing security vulnerabilities.
+We find that Android as a whole gets a score of \daSecurityScore\ out of 10, the highest scoring device manufacturer is \emph{\daSecScoreBestmanufacturer} (\daSecScoreBestmanufacturerScore\ out of 10) and the lowest scoring is \emph{\daSecScoreWorstmanufacturer} (\daSecScoreWorstmanufacturerScore\ out of 10).
+
 \begin{figure}
@@
 \includegraphics[width=\columnwidth]{figures/proportioninsecure}
-\caption{Proportion of devices running insecure, maybe secure and secure versions of Android.}
+\caption{Proportion of devices running insecure, maybe secure and secure versions of Android.
+Table~\ref{tab:andvulns} lists the \daNumVulnsUsed\ vulnerabilities used, the red vertical lines are caused by their discovery and the most important are annotated.}
 \label{fig:proportioninsecure}
 \end{figure}
-Table~\ref{tab:andvulns} lists the \daNumVulnsUsed\ vulnerabilities used, the red vertical lines are caused by their discovery and the most important are annotated.
-A device is insecure if it is running a vulnerable version of Android and the device has not received an update which might fix it;
-it is maybe secure if it is running a vulnerable version and but received an update which could have fixed the vulnerability if it contained a backported fix;
+By combining data on critical vulnerabilities in Android and the versions of Android running on devices we can determine which vulnerabilities each device was vulnerable to each day.
+Then we consider a device is insecure if it is running a vulnerable version of Android and the device has not received an update which might fix it;
+it is maybe secure if it is running a vulnerable version but received an update which could have fixed the vulnerability if it contained a backported fix;
 and it is secure if it is running a secure version.
-Combining the data on critical vulnerabilities and versions of Android allows us to plot Figure~\ref{fig:proportioninsecure}, initially all devices are \emph{maybe secure} (yellow) since \da\ does not have historical data prior to May 2011.
+This allows us to plot Figure~\ref{fig:proportioninsecure}, initially all devices are \emph{maybe secure} (yellow) since \da\ does not have historical data prior to May 2011.
 This means we cannot distinguish between devices which are running a version of Android which is known to be vulnerable from one which may have received a backported fix.
@@
 This gives us an average exposure to critical vulnerabilities of Android devices of \daMeanInsecurityPerc.
+This is $1 - f$, the proportion of devices free from critical vulnerabilities, however this is only one component of a more robust score, the FUM score.
 
-\subsection{Comparing device manufacturers, device models and network operators}
-\label{sec:security_scoring}\label{sec:exp:security_score}
-
-To allow buyers of Android devices to purchase those devices with the best security, they need to know how different device manufacturers, device models and network operators compare in terms of the security they provide.
-We propose a method to score a device manufacturer, device model or network operator based on its historic performance at keeping devices up-to-date and fixing security vulnerabilities.
-We find that Android as a whole gets a score of \daSecurityScore\ out of 10, the highest scoring device manufacturer is \emph{\daSecScoreBestmanufacturer} (\daSecScoreBestmanufacturerScore\ out of 10) and the lowest scoring is \emph{\daSecScoreWorstmanufacturer} (\daSecScoreWorstmanufacturerScore\ out of 10).
-
-\subsection{Method: Scoring for security}\label{sec:security_scoring:method}
+\subsection{Method: The FUM score}\label{sec:security_scoring:method}
 

### 2015-07-27 security_scoring.tex
Specify that MeanInsecurityPerc ignores 'maybe secure' devices
Shrink utilitarianism

@@
 From February 2013 onwards regular discovery of critical vulnerabilities ensures that most devices are exposed to known critical vulnerabilities.
-This gives us an average exposure to critical vulnerabilities of Android devices of \daMeanInsecurityPerc.
+This gives us an average exposure to critical vulnerabilities of Android devices of \daMeanInsecurityPerc\ (this percentage ignores devices classed as \emph{maybe secure}).
 This is $1 - f$, the proportion of devices free from critical vulnerabilities, however this is only one component of a more robust score, the FUM score.
@@
 Hence the total risk to users from the higher scoring popular manufacturers is higher than the risk from the lower scoring unpopular manufacturers.
-We could normalise the scores for market penetration and so give a score reflecting the risk posed by that manufacturer's performance, which would tend to decrease the difference between manufacturers in our current scoring.
-Additionally since our scores are provided so that customers can chose which devices to buy then it is the marginal risk to that individual of that device which is of interest rather than the aggregate risk to all users.
+We could normalise for market penetration and so give a score reflecting the risk posed by that manufacturer's performance, which would tend to decrease the difference between manufacturers in our current scoring.
+Since our scores are provided so that customers can chose which devices to buy then it is the marginal risk to that individual of that device which is of interest rather than the aggregate risk to all users.

### 2015-07-27 spsm04-thomas.tex
Add a parent section on data

@@
 
-\section{Critical vulnerabilities}\label{sec:avo}
+\section{Data}
+We use two sources of data to measure the security of Android: (1) information on the distribution of Android versions over time and (2) information on the critical vulnerabilities found to affect Android.
+These two datasets can then be combined to determine the proportion of devices at risk of attack from specific vulnerabilities.
+
+\subsection{Critical vulnerabilities}\label{sec:avo}
 We have used data on critical vulnerabilities in Android from AndroidVulnerabilities.org (\avo)~\cite{androidvulnerabilities.org}.
@@
 
-\section{Device Analyzer data}\label{sec:da}
+\subsection{Device Analyzer data}\label{sec:da}
 

### 2015-07-27 security_scoring.tex
Shrink discussion of how much data would be required for added value analysis

@@
 An added value analysis of network operators, which takes into account the device mix used by users of that network operator, would make it possible to determine whether a network operator is making the situation better or worse by the way it ships updates to users.
-However our sample size is too small to do that because while we have significant numbers of devices for each of the device models (Table~\ref{tab:sec_model}) and for each of the network operators (Table~\ref{tab:sec_operator}), we would need a significant number of each model in each network operator.
-A large enough sample size would contain at least 100 unique devices each day for each device model for each network operator, hence for our \daNumSigModels significant models and \daNumSigOperators significant network operators, a daily average of at least 18\,000 unique devices would be needed, for a period of at least a year.
-Since the distribution of devices is unlikely to be uniformly distributed across device models and network operators a more realistic estimate is 100\,000 unique devices each day.
+However our sample size is too small to do that because while we have significant numbers of devices for each of the \daNumSigModels\ device models (Table~\ref{tab:sec_model}) and for each of the \daNumSigOperators\ network operators (Table~\ref{tab:sec_operator}), we would need a significant number of each model in each network operator.
+Since the distribution of devices is unlikely to be uniformly distributed across device models and network operators we estimate that 100\,000 unique devices are required each day for at least a year.
 This is not an unobtainable number but it is two orders of magnitude more than is available in \da.
-\todo{Could abbreviate if necessary}
 %We do not attempt to disambiguate users behaviour in whether they install updates from network operators using rolling upgrades.

### 2015-07-27 avostats.tex
Updated stats and figures
./update.sh
Tweak some tables and hyphenation rules to make things fit better inside the columns

--- 2014-11-13 avostats.tex
@@
 \newcommand{\avoBigExternalLinesOfCodePerc}{99.7\%}
-\newcommand{\avoNumVulnerabilities}{31}
+\newcommand{\avoNumVulnerabilities}{32}
 \newcommand{\avoNumVulnAllAndroid}{15}
-\newcommand{\avoNumVulnSpecific}{16}
+\newcommand{\avoNumVulnSpecific}{17}
 \newcommand{\avoStartDate}{2013-08-28}
-\newcommand{\avoEndDate}{2014-10-17}
+\newcommand{\avoEndDate}{2015-03-24}
+\newcommand{\avoFirstDataDateMonth}{July 2010}
 \newcommand{\avoFirstDataDate}{2010-07-13}
-\newcommand{\avoLastDataDate}{2014-07-29}
-\newcommand{\avoVulnsPerYearNominal}{7.66}
-\newcommand{\avoVulnsPerYear}{$7.66 \pm 1.38$}
-\newcommand{\avoVulnsPerYearAllAndroidNominal}{3.71}
-\newcommand{\avoVulnsPerYearAllAndroid}{$3.71 \pm 0.957$}
-\newcommand{\avoTabAndVulns}{\begin{table} \centering \small \begin{tabular}{l|l|c} Vulnerability & How known & Date \\ \hline KillingInTheNameOf & Fixed on & 2010-07-13 \\ exploid udev & Discovered on & 2010-07-15 \\ levitator & Discovered on & 2011-03-10 \\ Gingerbreak & Fixed on & 2011-04-18 \\ zergRush & Discovered on & 2011-10-06 \\ APK duplicate file & Discovered on & 2013-02-18 \\ APK unchecked name & Discovered on & 2013-06-30 \\ APK unsigned shorts & Fixed on & 2013-07-03 \\ vold asec & Fixed on & 2014-01-27 \\ Fake ID & Fixed on & 2014-04-17 \\ TowelRoot & Discovered on & 2014-05-03 \\\end{tabular} \caption{Critical vulnerabilities in Android} \label{tab:andvulns} \end{table}}
+\newcommand{\avoLastDataDateMonth}{March 2015}
+\newcommand{\avoLastDataDate}{2015-03-12}
+\newcommand{\avoVulnsPerYearNominal}{6.86}
+\newcommand{\avoVulnsPerYear}{$6.86 \pm 1.21$}
+\newcommand{\avoVulnsPerYearAllAndroidNominal}{3.21}
+\newcommand{\avoVulnsPerYearAllAndroid}{$3.21 \pm 0.83$}
+\newcommand{\avoVulnsPerYearTwosfNominal}{6.9}
+\newcommand{\avoVulnsPerYearTwosf}{$6.9 \pm 1.2$}
+\newcommand{\avoVulnsPerYearAllAndroidTwosfNominal}{3.2}
+\newcommand{\avoVulnsPerYearAllAndroidTwosf}{$3.2 \pm 0.0$}
+\newcommand{\avoTabAndVulns}{\begin{table} \centering \small \begin{tabular}{l|l|c|c} Vulnerability & How known & Date & Categories\\ \hline KillingInTheNameOf & Fixed & 2010-07-13 & sys, kern\\ exploid udev & Discovered & 2010-07-15 & kernel\\ levitator & Discovered & 2011-03-10 & kernel\\ Gingerbreak & Fixed & 2011-04-18 & system\\ zergRush & Discovered & 2011-10-06 & system\\ APK duplicate file & Discovered & 2013-02-18 & signature\\ APK unchecked name & Discovered & 2013-06-30 & signature\\ APK unsigned shorts & Fixed & 2013-07-03 & signature\\ vold asec & Fixed & 2014-01-27 & system\\ Fake ID & Fixed & 2014-04-17 & signature\\ TowelRoot & Discovered & 2014-05-03 & kernel\\\end{tabular} \caption{Critical vulnerabilities in Android} \label{tab:andvulns} \end{table}}
 \newcommand{\avoNumBigExternalProjects}{40}
@@
 \newcommand{\avoAnalysedExternalLinesOfCodePerc}{24.9\%}
-\newcommand{\avoBigExternalMedianVersions}{1.5}
-\newcommand{\avoBigExternalMeanVersionsNominal}{2.07}
-\newcommand{\avoBigExternalMeanVersions}{$2.07 \pm 1.44$}
-\newcommand{\avoBigExternalTotalVersions}{58}
+\newcommand{\avoBigExternalMedianVersions}{2.0}
+\newcommand{\avoBigExternalMeanVersionsNominal}{2.57}
+\newcommand{\avoBigExternalMeanVersions}{$2.57 \pm 1.84$}
+\newcommand{\avoBigExternalTotalVersions}{72}

### 2015-07-27 dastats.tex
Tweak some tables and hyphenation rules to make things fit better inside the columns

@@
 \newcommand{\daSecScoreWorstoperatorNumFullVersions}{73}
-\newcommand{\daTabDLDistances}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & summary \\ \hline equal & 0.0 & 0.0667 & 0.0 & 0.0 \\weight $m$ & 0.2 & 0.4 & 0.0714 & 0.0 \\$f$ & 0.4 & 0.667 & 0.5 & 0.0 \\weight $u$ & 0.4 & 0.0667 & 0.143 & 0.0 \\$m$ & 0.6 & 0.733 & 0.357 & 0.5 \\$u$ & 0.9 & 0.6 & 0.786 & 0.0 \\\end{tabular} \caption{Normalised Damerau-Levenshtein distances for different metrics} \label{tab:dl_distances} \end{table}}
-\newcommand{\daTabChangeInScores}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & summary \\ \hline $m$ & $-1.67 \pm 2.25$ & $-0.66 \pm 2.54$ & $-3.3 \pm 1.3$ & $-3.4 \pm 1.8$ \\weight $m$ & $-0.264 \pm 0.277$ & $-0.0991 \pm 0.321$ & $-0.461 \pm 0.167$ & $-0.488 \pm 0.217$ \\equal & $-0.108 \pm 0.062$ & $-0.0367 \pm 0.112$ & $-0.146 \pm 0.045$ & $-0.165 \pm 0.033$ \\weight $u$ & $-0.0608 \pm 0.122$ & $-0.0112 \pm 0.232$ & $0.0242 \pm 0.0808$ & $-0.00596 \pm 0.12$ \\$u$ & $0.367 \pm 1.61$ & $0.219 \pm 2.12$ & $1.55 \pm 0.94$ & $1.42 \pm 1.48$ \\$f$ & $0.975 \pm 0.564$ & $0.331 \pm 1.01$ & $1.31 \pm 0.41$ & $1.48 \pm 0.3$ \\\end{tabular} \caption{Mean change in scores for different metrics} \label{tab:change_in_scores} \end{table*}}
+\newcommand{\daTabDLDistances}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & nexus \\ \hline equal & 0.0 & 0.0667 & 0.0 & 0.0 \\weight $m$ & 0.2 & 0.4 & 0.0714 & 0.0 \\$f$ & 0.4 & 0.667 & 0.5 & 0.0 \\weight $u$ & 0.4 & 0.0667 & 0.143 & 0.0 \\$m$ & 0.6 & 0.733 & 0.357 & 0.5 \\$u$ & 0.9 & 0.6 & 0.786 & 0.0 \\\end{tabular} \caption{Normalised Damerau-Levenshtein distances for different metrics} \label{tab:dl_distances} \end{table}}
+\newcommand{\daTabChangeInScores}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & nexus \\ \hline $m$ & $-1.67 \pm 2.25$ & $-0.66 \pm 2.54$ & $-3.3 \pm 1.3$ & $-3.4 \pm 1.8$ \\weight $m$ & $-0.264 \pm 0.277$ & $-0.0991 \pm 0.321$ & $-0.461 \pm 0.167$ & $-0.488 \pm 0.217$ \\equal & $-0.108 \pm 0.062$ & $-0.0367 \pm 0.112$ & $-0.146 \pm 0.045$ & $-0.165 \pm 0.033$ \\weight $u$ & $-0.0608 \pm 0.122$ & $-0.0112 \pm 0.232$ & $0.0242 \pm 0.0808$ & $-0.00596 \pm 0.12$ \\$u$ & $0.367 \pm 1.61$ & $0.219 \pm 2.12$ & $1.55 \pm 0.94$ & $1.42 \pm 1.48$ \\$f$ & $0.975 \pm 0.564$ & $0.331 \pm 1.01$ & $1.31 \pm 0.41$ & $1.48 \pm 0.3$ \\\end{tabular} \caption{Mean change in scores for different metrics} \label{tab:change_in_scores} \end{table*}}
 \newcommand{\daGPAPICurveFitParamFirst}{80.8}
@@
 \newcommand{\daGPAPISeventeenLaterDateMonth}{June 2015}
-\newcommand{\daTabSpearmanRanks}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & summary \\ $\pm\sigma$ & 0.211 & 0.169 & 0.175 & 0.632 \\ \hline $u$ & 0.297 & 0.804 & 0.618 & 1.0 \\$m$ & 0.794 & 0.593 & 0.969 & -1.0 \\$f$ & 0.83 & 0.775 & 0.934 & 1.0 \\weight $u$ & 0.939 & 0.996 & 0.991 & 1.0 \\weight $m$ & 0.976 & 0.964 & 0.996 & 1.0 \\equal & 1.0 & 0.996 & 1.0 & 1.0 \\\end{tabular} \caption{Spearman Rank correlation coefficients for different metrics.
+\newcommand{\daTabSpearmanRanks}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & nexus \\ $\pm\sigma$ & 0.211 & 0.169 & 0.175 & 0.632 \\ \hline $u$ & 0.297 & 0.804 & 0.618 & 1.0 \\$m$ & 0.794 & 0.593 & 0.969 & -1.0 \\$f$ & 0.83 & 0.775 & 0.934 & 1.0 \\weight $u$ & 0.939 & 0.996 & 0.991 & 1.0 \\weight $m$ & 0.976 & 0.964 & 0.996 & 1.0 \\equal & 1.0 & 0.996 & 1.0 & 1.0 \\\end{tabular} \caption{Spearman Rank correlation coefficients for different metrics.
 The uncertainty is constant for each column but does not take into account the uncertainty in the score which produced the ranking.} \label{tab:spearman_ranks} \end{table}}

### 2015-07-27 spsm04-thomas.tex
Tweak some tables and hyphenation rules to make things fit better inside the columns
Put some numbers in the abstract

@@
 This provides a comparison point for purchasers to determine which device manufacturers and network operators to use.
-We argue that publishing metrics like this provides an important incentive to manufacturers and operators to ship security updates.
-This is the first paper to quantify security update performance of device manufacturers and network operators.
-\todo{Do we want to pull numeric results into the abstract or leave them out to emphasise the method rather than the results?}
+We find that \daMeanInsecurityPercNominal\ of Android devices are exposed to at least one of \daNumVulnsUsed\ known critical vulnerabilities and assign a FUM security score of \daSecurityScoreNominal\ out of 10 while Nexus devices gets \daSecScoreBestsummaryScoreNominal\ and \daSecScoreBestmanufacturer\ \daSecScoreBestmanufacturerScoreNominal.
+On average \daMeanOutstandingVulnerabilitiesNominal\ outstanding vulnerabilities are not fixed on any device and only \daUpdatednessPercNominal\ of devices run the current version of Android.
 \end{abstract}
@@
 \da\ collects data\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/collected.html}} from study participants who install the Android app from Google Play.
-Most study participants allow researchers around the world to access a subset of their device data, including the data presented here.
+Most study participants allow researchers a\-round the world to access a subset of their device data, including the data presented here.
 

### 2015-07-27 spsm04-thomas.tex
Move introduction of metrics a little later in the intro

@@
 
-We show that Android devices do not receive security updates promptly when critical vulnerabilities\footnote{If malicious code exploits a critical vulnerability then it gains control of the device.} are publicly disclosed and hence, on average, \daMeanInsecurityPercNominal\ are exposed to known critical vulnerabilities.
-We attempt to determine why by understanding the Android update process, evaluating different entities in the ecosystem and producing a scoring metric to compare different entities.
+\todo{new opening paragraph}
 
-We propose scoring metrics to compare compares different entities such as device manufacturers, network operators or specific device models:
-$f$ the proportion of running devices free from critical vulnerabilities over time;
-$u$ the proportion of devices that run the latest version of Android shipped to any device produced by that device manufacturer;
-$m$ the mean number of outstanding vulnerabilities affecting devices not fixed on any device shipped by the device manufacturer;
-and a composite score FUM which is hard to game (\S\ref{sec:gaming}).
+We propose scoring metrics to compare compares different entities such as device manufacturers, network operators or specific device models.
 This information allows device purchasers to select devices with better security.
@@
 To address this, we developed a scoring system and provide numbers on the historic performance of device models in the \da\ project in~\S\ref{sec:security_scoring}.
+We propose three metrics:
+$f$ the proportion of running devices free from critical vulnerabilities over time;
+$u$ the proportion of devices that run the latest version of Android shipped to any device produced by that device manufacturer;
+$m$ the mean number of outstanding vulnerabilities affecting devices not fixed on any device shipped by the device manufacturer;
+and a composite score FUM which is hard to game (\S\ref{sec:gaming}).
 We provide the answer to the question ``Which is the most secure Android provider I can buy from?'', showing that in general Nexus devices are better; further details are given in~\S\ref{sec:security_scoring:results}.

### 2015-07-27 security_scoring.tex
Make the update bottleneck a subsection

@@
 
-If update delays are due to the delay in manufacturers providing the update rather than in operators supplying the update and users installing the update, we would expect the update behaviour of devices with the same device model to be similar and rapid.
-We found that within \daModelHalfDeployment\ days of the first observation of a new version on a device, half of all devices of that model have the new version (or a higher version) installed, and within \daModelFullDeployment\ days \daFullDeployedAt\ of devices have the new version (or a higher version).
-This compares with the average rates of deployment for Android OS versions of 350 days for half and 1\,100 days for \daFullDeployedAt.
-\todolater{Compute these values automatically and with error.}
-There is a variation between device models, with the update being distributed to most devices quickly and others having a much slower roll out, but since some do update quickly the bottleneck is unlikely to be with the user.
-Perhaps some device models are preferred by users who are more likely to install updates than others, however we do observe updates being rolled out to device models quickly and user behaviour is not beyond the control of the device manufacturer.
-They could install updates automatically or pester the user into installing them, and at least some of them do pester, silent automatic updates do boost uptake~\cite{Duebendorfer2009}.
-
 \begin{figure*}
@@
 %We do not attempt to disambiguate users behaviour in whether they install updates from network operators using rolling upgrades.
+
+\subsection{Update bottleneck}
+If update delays are due to the delay in manufacturers providing the update rather than in operators supplying the update and users installing the update, we would expect the update behaviour of devices with the same device model to be similar and rapid.
+We found that within \daModelHalfDeployment\ days of the first observation of a new version on a device, half of all devices of that model have the new version (or a higher version) installed, and within \daModelFullDeployment\ days \daFullDeployedAt\ of devices have the new version (or a higher version).
+This compares with the average rates of deployment for Android OS versions of 350 days for half and 1\,100 days for \daFullDeployedAt.
+\todolater{Compute these values automatically and with error.}
+There is a variation between device models, with the update being distributed to most devices quickly and others having a much slower roll out, but since some do update quickly the bottleneck is unlikely to be with the user.
+Perhaps some device models are preferred by users who are more likely to install updates than others, however we do observe updates being rolled out to device models quickly and user behaviour is not beyond the control of the device manufacturer.
+They could install updates automatically or pester the user into installing them, and at least some of them do pester, silent automatic updates do boost uptake~\cite{Duebendorfer2009}.
 

### 2015-07-27 security_scoring.tex
Reorder some subsections to try and get the text closer to the figures and tables that they describe

@@
 
+\subsection{Sensitivity of scoring metric}
+\daTabSpearmanRanks
+\daTabChangeInScores
+To evaluate whether the ranking of different manufacturers is sensitive to the form of the scoring metric we computed the normalised Spearman's Rank correlation coefficient between the lists ordered using different forms of the scoring metric, this is shown in Table~\ref{tab:spearman_ranks}.
+The `equal' metric weights $f$, $u$ and $m$ equally rather than favouring $f$ and makes little difference.
+Similarly weighting $u$ or $m$ more highly rather than $f$ makes little difference.
+While the $f$, $u$ and $m$ components do have some correlation with the overall FUM score, the rankings produced vary substantially.
+Changing the scoring metric also impacts the scores given for each entity Table~\ref{tab:change_in_scores} shows the mean impact on the scores.
+This shows that $m$ tends to drag down scores.
+\todolater{Give some of the actual orderings as well for the most important ones - no space}
+\todolater{Sensitivity of time based vs. global scores - no time}
+
+\subsection{Utilitarianism}
+From a utilitarian standpoint, while small manufacturers like Symphony and Walton do badly on our scores, they do not have as many customers as higher scoring manufacturers.
+Hence the total risk to users from the higher scoring popular manufacturers is higher than the risk from the lower scoring unpopular manufacturers.
+We could normalise for market penetration and so give a score reflecting the risk posed by that manufacturer's performance, which would tend to decrease the difference between manufacturers in our current scoring.
+Since our scores are provided so that customers can chose which devices to buy then it is the marginal risk to that individual of that device which is of interest rather than the aggregate risk to all users.
+
 \subsection{Scores over time}
-The scoring metric as originally computed, is averaged over the whole history of the device manufacturer, device model or network operator, it gives equal weight to both periods years ago and to the last few months.
-If instead we take an exponential moving average of the daily score for days with more than \daSigNumDevicesDay\, devices when there have been at least \daSigNumDays\ consecutive days of data with that many devices then we can plot how this score has changed over time.
-Equation~\ref{eq:rolling_update} shows how the value for a particular day ($v_i$) is computed from the previous day's value and the input for the current day ($n$) with an $\alpha$ of $1/\daSigNumDays$.
-\begin{equation}
-v_i = v_{i-1} (1 - \alpha) + n \alpha
-\label{eq:rolling_update}
-\end{equation}
-Figure~\ref{fig:security_scores} shows this for manufacturers, device models, network operators and for Nexus and non-Nexus devices.
-These show how the scores for different entities are different and change over time, while there is correlated behaviour for different entities (for instance, when new vulnerabilities affecting all Android are discovered), these lines still have crossings due to the different behaviour of the various entities.
-It also shows that we do not have sufficient data for all the entities all of the time, resulting in gaps in the data.
-The clearest results are for Figure~\ref{fig:security_score_summary} with a large gap between the scores for Nexus and non-Nexus devices across the whole data set.
 \begin{figure}
@@
 \end{figure}
-
-\subsection{Sensitivity of scoring metric}
-\daTabSpearmanRanks
-\daTabChangeInScores
-To evaluate whether the ranking of different manufacturers is sensitive to the form of the scoring metric we computed the normalised Spearman's Rank correlation coefficient between the lists ordered using different forms of the scoring metric, this is shown in Table~\ref{tab:spearman_ranks}.
-The `equal' metric weights $f$, $u$ and $m$ equally rather than favouring $f$ and makes little difference.
-Similarly weighting $u$ or $m$ more highly rather than $f$ makes little difference.
-While the $f$, $u$ and $m$ components do have some correlation with the overall FUM score, the rankings produced vary substantially.
-Changing the scoring metric also impacts the scores given for each entity Table~\ref{tab:change_in_scores} shows the mean impact on the scores.
-This shows that $m$ tends to drag down scores.
-\todolater{Give some of the actual orderings as well for the most important ones - no space}
-\todolater{Sensitivity of time based vs. global scores - no time}
+The scoring metric as originally computed, is averaged over the whole history of the device manufacturer, device model or network operator, it gives equal weight to both periods years ago and to the last few months.
+If instead we take an exponential moving average of the daily score for days with more than \daSigNumDevicesDay\, devices when there have been at least \daSigNumDays\ consecutive days of data with that many devices then we can plot how this score has changed over time.
+Equation~\ref{eq:rolling_update} shows how the value for a particular day ($v_i$) is computed from the previous day's value and the input for the current day ($n$) with an $\alpha$ of $1/\daSigNumDays$.
+\begin{equation}
+v_i = v_{i-1} (1 - \alpha) + n \alpha
+\label{eq:rolling_update}
+\end{equation}
+Figure~\ref{fig:security_scores} shows this for manufacturers, device models, network operators and for Nexus and non-Nexus devices.
+These show how the scores for different entities are different and change over time, while there is correlated behaviour for different entities (for instance, when new vulnerabilities affecting all Android are discovered), these lines still have crossings due to the different behaviour of the various entities.
+It also shows that we do not have sufficient data for all the entities all of the time, resulting in gaps in the data.
+The clearest results are for Figure~\ref{fig:security_score_summary} with a large gap between the scores for Nexus and non-Nexus devices across the whole data set.
 
@@
 Therefore, our score is secure against passive gaming attacks which change the measured distribution, but would require active defence against active gaming attacks, which target the measurement devices.
-
-\subsection{Utilitarianism}
-From a utilitarian standpoint, while small manufacturers like Symphony and Walton do badly on our scores, they do not have as many customers as higher scoring manufacturers.
-Hence the total risk to users from the higher scoring popular manufacturers is higher than the risk from the lower scoring unpopular manufacturers.
-We could normalise for market penetration and so give a score reflecting the risk posed by that manufacturer's performance, which would tend to decrease the difference between manufacturers in our current scoring.
-Since our scores are provided so that customers can chose which devices to buy then it is the marginal risk to that individual of that device which is of interest rather than the aggregate risk to all users.

### 2015-07-27 security_scoring.tex
Add update bottleneck as a contribution

@@
 
-\subsection{Update bottleneck}
+\subsection{Update bottleneck}\label{sec:bottleneck}
 If update delays are due to the delay in manufacturers providing the update rather than in operators supplying the update and users installing the update, we would expect the update behaviour of devices with the same device model to be similar and rapid.

### 2015-07-27 spsm04-thomas.tex
Add update bottleneck as a contribution
Use month rather than date information in ecosystem figure

@@
  \item We quantify the Android update process, providing concrete numbers on the flow of updates and their latency~(\S\ref{sec:android_ecosystem}).
+ \item We determine that the main update bottleneck lies with manufacturers rather than operators or users~(\S\ref{sec:bottleneck}).
  \item We propose the FUM scoring metric to evaluate the security of different instances of a platform~(\S\ref{sec:security_scoring:method}).
@@
  \caption{Flow of updates between participants in the Android ecosystem.
- Numbers on edges indicate updates shipped between \daStartDate\ and \daEndDate, those in brackets represent number of such entities.
+ Numbers on edges indicate updates shipped between \daStartDateMonth\ and \daEndDateMonth, those in brackets represent number of such entities.
  Dotted arrows indicate flows where we can't measure because no public data is available.}

### 2015-07-27 security_scoring.tex
Corrections on reading through

@@
 By combining data on critical vulnerabilities in Android and the versions of Android running on devices we can determine which vulnerabilities each device was vulnerable to each day.
-Then we consider a device is insecure if it is running a vulnerable version of Android and the device has not received an update which might fix it;
-it is maybe secure if it is running a vulnerable version but received an update which could have fixed the vulnerability if it contained a backported fix;
-and it is secure if it is running a secure version.
+Then we consider a device is \emph{insecure} if it is running a vulnerable version of Android and the device has not received an update which might fix it;
+it is \emph{maybe secure} if it is running a vulnerable version but received an update which could have fixed the vulnerability if it contained a backported fix;
+and it is \emph{secure} if it is running a secure version.
 This allows us to plot Figure~\ref{fig:proportioninsecure}, initially all devices are \emph{maybe secure} (yellow) since \da\ does not have historical data prior to May 2011.
@@
 There is anecdotal evidence that Google's Nexus devices are better at getting updates than other Android devices because Google makes the original updates and ships them to its devices\footnote{\url{http://www.howtogeek.com/139391/htg-explains-why-android-geeks-buy-nexus-devices/}}.
-Table~\ref{tab:sec_summary} shows that this is the case with Nexus devices getting much better scores than non-Nexus devices and Figure~\ref{fig:security_score_summary} shows this holds true over time.
+Table~\ref{tab:sec_summary} shows that this is the case with Nexus devices getting much better scores than non-Nexus devices.
 

### 2015-07-27 spsm04-thomas.tex
Corrections on reading through

@@
 This provides a comparison point for purchasers to determine which device manufacturers and network operators to use.
-We find that \daMeanInsecurityPercNominal\ of Android devices are exposed to at least one of \daNumVulnsUsed\ known critical vulnerabilities and assign a FUM security score of \daSecurityScoreNominal\ out of 10 while Nexus devices gets \daSecScoreBestsummaryScoreNominal\ and \daSecScoreBestmanufacturer\ \daSecScoreBestmanufacturerScoreNominal.
-On average \daMeanOutstandingVulnerabilitiesNominal\ outstanding vulnerabilities are not fixed on any device and only \daUpdatednessPercNominal\ of devices run the current version of Android.
+We find that on average \daMeanInsecurityPercNominal\ of Android devices are exposed to at least one of \daNumVulnsUsed\ known critical vulnerabilities and assign a FUM security score of \daSecurityScoreNominal\ out of 10 while Nexus devices gets \daSecScoreBestsummaryScoreNominal\ and \daSecScoreBestmanufacturer\ \daSecScoreBestmanufacturerScoreNominal.
+On average \daMeanOutstandingVulnerabilitiesNominal\ outstanding vulnerabilities are not fixed on any device and on average only \daUpdatednessPercNominal\ of devices run the current version of Android.
 \end{abstract}
@@
 
-\todo{new opening paragraph}
+To evaluate the security of a system we need to be able to measure it.
+Only then can customers and regulators behave appropriately and without it then they are stuck with a lemons market.
 
@@
 The cause of the slow update process could be: upstream open source projects releasing patches slowly, Google including them slowly, manufacturers including them slowly, network operators shipping them slowly or customers not installing them.
-We present a better understanding of both the ecosystem of Android development and associated vulnerabilities more fully in~\S\ref{sec:android_ecosystem}.
+We present a better understanding of both the ecosystem of Android development and associated vulnerabilities more fully in~\S\ref{sec:android_ecosystem} and the update bottleneck in \S\ref{sec:bottleneck}.
 
-By using data from \da~\cite{Wagner2013} we can examine the versions of Android running on user devices over time and through comparison with data on which versions of Android were vulnerable to different critical vulnerabilities~\cite{androidvulnerabilities.org} we can plot the exposure of Android devices to known critical vulnerabilities.
-Figure~\ref{fig:proportioninsecure} uses data from \da\ (\S\ref{sec:da}) and the AndroidVulnerabilities.org (\avo) vulnerabilities database (\S\ref{sec:avo}) and shows that Android devices are often exposed to known security vulnerabilities (\S\ref{sec:androidvulnerability}).
+By using data from \da~\cite{Wagner2013} we can examine the versions of Android running on user devices over time and through comparison with data on which versions of Android were vulnerable to different critical vulnerabilities~\cite{androidvulnerabilities.org} we can measure the exposure of Android devices to known critical vulnerabilities.
 One solution to this problem is regulation, and indeed there is ongoing legal action to force network operators to ship updates for security vulnerabilities~\cite{Soghoian2013}.\todolater{Check on the status of this legal action}
-Many smartphones are sold on 12--24 month contracts, and yet our data shows many devices do not receive many security updates, with an overall average of \daUpdatesPerYearNominal\ per year. 
+Many smartphones are sold on 12--24 month contracts, and yet our data shows many devices do not receive many security updates, with an overall average of \daUpdatesPerYearNominal\ updates per year. 
 In contrast, Windows XP could be purchased for a one-off payment in October 2001 and received security updates monthly until April 2014.
@@
  \item We quantify the Android update process, providing concrete numbers on the flow of updates and their latency~(\S\ref{sec:android_ecosystem}).
- \item We determine that the main update bottleneck lies with manufacturers rather than operators or users~(\S\ref{sec:bottleneck}).
  \item We propose the FUM scoring metric to evaluate the security of different instances of a platform~(\S\ref{sec:security_scoring:method}).
  \item We measure the security of Android against our scoring metric and compare different device manufacturers, device models and network operators to allow device purchasers to differentiate between them based on security~(\S\ref{sec:security_scoring:results}).
+ \item We determine that the main update bottleneck lies with manufacturers rather than operators or users~(\S\ref{sec:bottleneck}).
 \end{itemize}
@@
 The best place to prevent attacks is at runtime, since all three attack vectors can be prevented at this level. 
-Unfortunately, as Figure~\ref{fig:proportioninsecure} shows, the sandbox for Android apps is weak, since on most devices there are known critical vulnerabilities in it.
+Unfortunately, the sandbox for Android apps is weak, since on most devices there are known critical vulnerabilities in it.
 
 \section{Data}
-We use two sources of data to measure the security of Android: (1) information on the distribution of Android versions over time and (2) information on the critical vulnerabilities found to affect Android.
+We use two sources of data to measure the security of Android: (1) information on the critical vulnerabilities found to affect Android and (2) information on the distribution of Android versions over time.
 These two datasets can then be combined to determine the proportion of devices at risk of attack from specific vulnerabilities.
@@
 Regardless of version, this exposed an increased attack area and would also provide the ability for malware to control all user internet traffic (via VPNs), brick the phone, remove and install apps, steal user credentials and read the screen.
-The different categories in which the vulnerabilities fall are shown in Table~\ref{tab:andvulns}, `signature' vulnerabilities are \emph{installation}, while kernel and system vulnerabilities are used with \emph{installation, dynamic code loading} and \emph{injection}.
+The different categories in which the vulnerabilities fall are shown in Table~\ref{tab:andvulns}, `signature' vulnerabilities are \emph{installation}, while `kernel' and `system' vulnerabilities are used with \emph{installation, dynamic code loading} and \emph{injection}.
 \avoTabAndVulns
@@
 \subsection{Device Analyzer data}\label{sec:da}
-
-In this analysis we use historical data collected by the \da\ project~\cite{Wagner2013}.
+We use historical data collected by the \da\ project~\cite{Wagner2013}.
 \da\ collects data\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/collected.html}} from study participants who install the Android app from Google Play.
@@
 
-\da\ has collected data from \daNumDevices\ devices with a total of \daDeviceDays\footnote{Here we are only counting devices and days for which we have valid OS version data.} device days.
+\da\ has collected data from \daNumDevices\ devices with a total of \daDeviceDays device days.
 The majority of devices only contribute data for a short period of time, however \daMonthsDevices\ devices have contributed data for more than \daMonths~months.
@@
 \section{Comparison with other data}\label{sec:representative}
-Here we provide evidence that the \da\ data~\cite{Wagner2013} is representative for our purposes by comparing it with three other sources of data.
-The \da\ data is collected from devices across the world where users have voluntarily installed the app.
+We compare the Device Analyzer data with three other sources of data and show it is representative for our purposes.
 We have obtained comparable data on 5\,290 devices from a multinational FTSE 100 company's mobile device management database, which includes company and employee owned devices, and from 5\,170\,000 matching User-Agent headers on all HTTP traffic for 30\% of Rwanda for a week.
@@
 Figure~\ref{fig:dists} shows the proportion of devices running each Android OS version in the two comparison data sets and the comparable periods from \da.
-The general pattern this shows is that in the FTSE data newer versions are more popular than in the \da\ data and that in the Rwanda data old versions are more popular than in the \da\ data.
+The general pattern this shows is that in the FTSE data newer versions are more popular than in the \da\ data and that in the Rwanda data old versions are more popular.
 Therefore the \da\ data on OS versions is bounded by these two data sets.

### 2015-07-27 spsm04-thomas.tex
Make acknowledgements a bit shorter

@@
 This work was supported by a Google focussed research award; and the EPSRC [grant number EP/P505445/1].
-Thanks to \href{https://www.cl.cam.ac.uk/~sa497/}{Sherif Akoush} and \href{http://orcid.org/0000-0003-0740-8650}{Ripduman Sohan} for supplying the Rwanda data and to staff at the FTSE 100 company for supplying their mobile device management data.
+Thanks to \href{https://www.cl.cam.ac.uk/~sa497/}{Sherif Akoush} and \href{http://orcid.org/0000-0003-0740-8650}{Ripduman Sohan} for supplying the Rwanda data and to staff at the FTSE 100 company for supplying their MDM data.
 Thanks to our anonymous reviewers for their insightful comments; Richard Clayton, the mobile security reading group, Anil Madhavapeddy and Oliver Chick for reading early drafts;

### 2015-07-27 spsm04-thomas.tex
Rewrite the conclusion

@@
 \section{Conclusion}
-We examined how the Android ecosystem results in different device models receiving different levels of security due to whether or not they get security updates.
+We quantified how the Android ecosystem results in different device models receiving different levels of security due to whether or not they get security updates.
+We found that on average \daMeanInsecurityPercNominal\ of Android devices are exposed to at least one of \daNumVulnsUsed\ known critical vulnerabilities.
+On average \daMeanOutstandingVulnerabilitiesNominal\ outstanding vulnerabilities are not fixed on any device and on average only \daUpdatednessPercNominal\ of devices run the current version of Android.
 We have defined the FUM score, which evaluates the security performance of companies and organisations in shipping security updates.
 We have used this to compare different device models, device manufacturers and network operators and found that there are differences between them, which the discerning purchaser might use to influence their decision about which device model to buy from which device manufacturer and network operator.
+While Android has a score of \daSecurityScoreNominal\ out of 10, Nexus devices get \daSecScoreBestsummaryScoreNominal\ and \daSecScoreBestmanufacturer\ \daSecScoreBestmanufacturerScoreNominal.
 We hope that this analysis will encourage device manufacturers and network operators to improve the support they provide for devices after sale.
-\todo{Rewrite conclusion}
+
+The FUM score is applicable to other software ecosystems, as long as information on vulnerabilities and the versions they affected, as well as data on which versions were running on real devices is available.
 

### 2015-07-27 security_scoring.tex
Cut some words out to make it fit on 12 pages

@@
 
-To allow buyers of Android devices to purchase those devices with the best security, they need to know how different device manufacturers, device models and network operators compare in terms of the security they provide.
+To allow buyers of Android devices to purchase those devices with the best security, they need to know how different device manufacturers, device models and network operators compare in terms of security.
 We propose a method to score a device manufacturer, device model or network operator based on its historic performance at keeping devices up-to-date and fixing security vulnerabilities.
-We find that Android as a whole gets a score of \daSecurityScore\ out of 10, the highest scoring device manufacturer is \emph{\daSecScoreBestmanufacturer} (\daSecScoreBestmanufacturerScore\ out of 10) and the lowest scoring is \emph{\daSecScoreWorstmanufacturer} (\daSecScoreWorstmanufacturerScore\ out of 10).
+We find that Android as a whole gets a score of \daSecurityScore\ out of 10, the highest scoring device manufacturer is \emph{\daSecScoreBestmanufacturer} (\daSecScoreBestmanufacturerScore) and the lowest scoring is \emph{\daSecScoreWorstmanufacturer} (\daSecScoreWorstmanufacturerScore).
 
@@
 The total secure device days and total insecure device days are both counting experiments and so their measurement error is their square root~\cite{Taylor1997}.
-Since the numbers involved are large, the uncertainty in $f$ is small.
+%Since the numbers involved are large, the uncertainty in $f$ is small.
 $u$ is computed by taking the sum of the proportions of devices running the most recent version each day, both the count of devices running the maximum version and total count have square root uncertainties.
-$m$ is computed by counting the number of vulnerabilities which affected that entity and which have not yet been fixed on any device we have observed from that entity every day and averaging over observed days.
+$m$ is computed by counting the number of vulnerabilities which affected that entity and which have not yet been fixed on any device we have observed from that entity every day and averaging time.
 However, it could be that the entity has released a fix to some devices but we have not yet observed a device with that fix.
@@
 This does not capture systematic errors.
-For example, we do not include manufacturer specific vulnerabilities, but we expect that performance in fixing manufacturer specific vulnerabilities is strongly correlated with performance fixing vulnerabilities affecting all of Android.
+For example, we do not include manufacturer specific vulnerabilities, we expect that performance in fixing manufacturer specific vulnerabilities is strongly correlated with performance fixing vulnerabilities affecting all of Android.
 

### 2015-07-27 spsm04-thomas.tex
Cut some words out to make it fit on 12 pages

@@
 %An analysis by Vidas et al.~\cite{Vidas2011} of the Android 2.1 to 2.2 update found that it took 11 months from when Google released 2.2 for the last device which they were investigating to get the update.
-To compute the latency between upstream releases and the release of the first version of Android containing that release we scraped the release pages for those projects, to obtain the version numbers and release dates.
+To compute the latency between upstream releases and the release of the first version of Android containing that release we scraped the release pages, to obtain the version numbers and release dates.
 This allows us to compute the latency between an upstream project being released and it being included in Android, this is shown in Table~\ref{tab:update_ecosystem}.

### 2015-07-28 security_scoring.tex
Rephrase discussion of MeanInsecurityPerc

@@
 From February 2013 onwards regular discovery of critical vulnerabilities ensures that most devices are exposed to known critical vulnerabilities.
-This gives us an average exposure to critical vulnerabilities of Android devices of \daMeanInsecurityPerc\ (this percentage ignores devices classed as \emph{maybe secure}).
-This is $1 - f$, the proportion of devices free from critical vulnerabilities, however this is only one component of a more robust score, the FUM score.
+This means that \daMeanInsecurityPerc\ of devices were classified as \emph{insecure} to at least one critical vulnerability between \daStartDateMonth\ and \daEndDateMonth\ (ignoring devices classed as \emph{maybe secure}).
+%This is $1 - f$, the proportion of devices free from critical vulnerabilities, however this is only one component of a more robust score, the FUM score.
 

### 2015-07-28 spsm04-thomas.tex
Disable TODOs from being displayed
Remove done TODOs
Adjust the introduction to comparision with other data

@@
 \section{Comparison with other data}\label{sec:representative}
-We compare the Device Analyzer data with three other sources of data and show it is representative for our purposes.
+We compare the Device Analyzer data with three other sources of data and find similar distributions which bound the Device Analyzer data, indicating that it is likely to be typical of the true distribution.
 We have obtained comparable data on 5\,290 devices from a multinational FTSE 100 company's mobile device management database, which includes company and employee owned devices, and from 5\,170\,000 matching User-Agent headers on all HTTP traffic for 30\% of Rwanda for a week.
@@
 \printbibliography
-\todo{We need to fix the columns at the end to be the same length}
-\todo{Make a thumbnail from figure 5}
 

### 2015-07-28 spsm04-thomas.tex
New abstract and intro.

--- 2015-07-27 spsm04-thomas.tex
@@
 \begin{abstract}
-The security of Android depends on timely delivery of security updates for critical vulnerabilities .
+The security of Android depends on the timely delivery of security updates for critical vulnerabilities across a complex ecosystem.
+In this paper we map the major players in the Android ecosystem who must collaborate to provide updates, and determine that inaction by some manufacturers and network operators means many handsets are vulnerable to critical vulnerabilities.
 We define the FUM security metric to rank the performance of device manufacturers and network operators, based on their provision of updates and exposure to critical vulnerabilities.
 Using a corpus of \daNumOSDevices\ devices we show that there is significant variability in the timely delivery of security updates across different device manufacturers and network operators.
-This provides a comparison point for purchasers to determine which device manufacturers and network operators to use.
-We find that on average \daMeanInsecurityPercNominal\ of Android devices are exposed to at least one of \daNumVulnsUsed\ known critical vulnerabilities and assign a FUM security score of \daSecurityScoreNominal\ out of 10 while Nexus devices gets \daSecScoreBestsummaryScoreNominal\ and \daSecScoreBestmanufacturer\ \daSecScoreBestmanufacturerScoreNominal.
-On average \daMeanOutstandingVulnerabilitiesNominal\ outstanding vulnerabilities are not fixed on any device and on average only \daUpdatednessPercNominal\ of devices run the current version of Android.
+This provides a comparison point for purchasers and regulators to determine which device manufacturers and network operators provide security updates and which do not.
+We find that on average \daMeanInsecurityPercNominal\ of Android devices are exposed to at least one of \daNumVulnsUsed\ known critical vulnerabilities and, across the ecosystem as a whole, assign a FUM security score of \daSecurityScoreNominal\ out of 10.
+In our data, Nexus devices do considerably better than average with a score of \daSecScoreBestsummaryScoreNominal; LG is the best manufacturer with a score of \daSecScoreBestmanufacturer\ \daSecScoreBestmanufacturerScoreNominal.
+%On average \daMeanOutstandingVulnerabilitiesNominal\ outstanding vulnerabilities are not fixed on any device and on average only \daUpdatednessPercNominal\ of devices run the current version of Android.
 \end{abstract}
@@
 
-\terms{Economics, Measurement, Security}
+\terms{Security, Measurement, Economics}
 
@@
 
-To evaluate the security of a system we need to be able to measure it.
-Only then can customers and regulators behave appropriately and without it then they are stuck with a lemons market.
+All large software systems today contain undiscovered security vulnerabilities.
+Once discovered, these flaws are often exploited, and therefore the timely delivery of security updates is important to protect such systems, particularly when devices are connected to the Internet and therefore can be exploited remotely.
+Manufactures and software companies have known about this issue for many years and are expected to provide regular updates to protect their users.
+For example, Windows XP could be purchased for a one-off payment in October 2001 and received monthly security updates until support ended in April 2014.
 
-We propose scoring metrics to compare compares different entities such as device manufacturers, network operators or specific device models.
-This information allows device purchasers to select devices with better security.
-Currently, corporate and public sector buyers are encouraged to purchase secure devices, but we have found little concrete guidance on the specific makes and models providing the best security.
-For example, CESG, which advises the UK government on how to secure its computer systems, recommends picking Android device models from device manufactures that are good at promptly shipping security updates, but it does not state which device manufacturers these are~\cite{CESG2013} though they have since certified one device model~\cite{CESG2015}.
-We are also collaborating with a FTSE 100 company to help answer this question.
-To address this, we developed a scoring system and provide numbers on the historic performance of device models in the \da\ project in~\S\ref{sec:security_scoring}.
+Unfortunately something has gone wrong with the provision of security updates in the Android market. 
+Many smartphones are sold on 12--24 month contracts, and yet our data shows few Android devices receive many security updates, with an overall average of just \daUpdatesPerYearNominal\ updates per year, leaving devices unpatched for many months or years.
+
+In order to improve our understanding, we need to know more about the Android ecosystem as a whole.
+It is a complex system with many parties involved in a long multi-stage pipeline~\cite{HTC2013}.
+We map and quantify the major players in this space who must collaborate to provide updates~(\S\ref{sec:android_ecosystem}) and determine that inaction~(\S\ref{sec:bottleneck}) by some of the manufacturers (we know of \daNumManufacturers) and network operators (\daNumOperators) means many handsets are vulnerable to critical vulnerabilities.
+Understanding this ecosystem is all the more important because device manufacturers have introduced additional vulnerabilities in the past~\cite{Grace2012}.
+
+Corporate and public sector buyers are encouraged to purchase secure devices, but we have found little concrete guidance on the specific makes and models providing timely security updates.
+For example, CESG, which advises the UK government on how to secure its computer systems, recommends picking Android device models from device manufactures that are good at promptly shipping security updates, but it does not state which device manufacturers these are~\cite{CESG2013} and so far they have only certified one Android device model~\cite{CESG2015}.
+Similarly, we are collaborating with a FTSE 100 company who wish to know which devices are secure and which manufacturers provide updates.
+
+The difficulty is that the market for Android security today is like the market for lemons: there is information asymmetry between the manufacturer, who knows whether the device is currently secure and will receive security updates, and the customer, who does not.
+To address the asymmetry, we develop a scoring system and provide numbers on the historic performance of device models found in the \da~\cite{Wagner2013} project~(\S\ref{sec:security_scoring}).
 We propose three metrics:
 $f$ the proportion of running devices free from critical vulnerabilities over time;
-$u$ the proportion of devices that run the latest version of Android shipped to any device produced by that device manufacturer;
-$m$ the mean number of outstanding vulnerabilities affecting devices not fixed on any device shipped by the device manufacturer;
-and a composite score FUM which is hard to game (\S\ref{sec:gaming}).
-We provide the answer to the question ``Which is the most secure Android provider I can buy from?'', showing that in general Nexus devices are better; further details are given in~\S\ref{sec:security_scoring:results}.
+$u$ the proportion of devices that run the latest version of Android shipped to any device produced by that device manufacturer; and
+$m$ the mean number of outstanding vulnerabilities affecting devices not fixed on any device shipped by the device manufacturer.
+We then derive a composite FUM score which is hard to game (\S\ref{sec:gaming}).
+%We provide the answer to the question ``Which is the most secure Android provider I can buy from?'', showing that in general Nexus devices are better; further details are given in~\S\ref{sec:security_scoring:results}.
 
-The Android update process is slow because its development is distributed among many parties and has a long multi-stage pipeline~\cite{HTC2013}.
-The security of Android relies on many open source projects, including the Li\-nux kernel, Open\-SSL and Bouncy\-Castle, as well as Google, who build the core platform.
-In addition, device manufacturers (we know of \daNumManufacturers) and network operators (\daNumOperators) may make further modifications before devices are shipped to customers.
-Understanding this ecosystem is important because device manufacturers have introduced additional vulnerabilities in the past~\cite{Grace2012}.
-The cause of the slow update process could be: upstream open source projects releasing patches slowly, Google including them slowly, manufacturers including them slowly, network operators shipping them slowly or customers not installing them.
-We present a better understanding of both the ecosystem of Android development and associated vulnerabilities more fully in~\S\ref{sec:android_ecosystem} and the update bottleneck in \S\ref{sec:bottleneck}.
-
-By using data from \da~\cite{Wagner2013} we can examine the versions of Android running on user devices over time and through comparison with data on which versions of Android were vulnerable to different critical vulnerabilities~\cite{androidvulnerabilities.org} we can measure the exposure of Android devices to known critical vulnerabilities.
-One solution to this problem is regulation, and indeed there is ongoing legal action to force network operators to ship updates for security vulnerabilities~\cite{Soghoian2013}.\todolater{Check on the status of this legal action}
-Many smartphones are sold on 12--24 month contracts, and yet our data shows many devices do not receive many security updates, with an overall average of \daUpdatesPerYearNominal\ updates per year. 
-In contrast, Windows XP could be purchased for a one-off payment in October 2001 and received security updates monthly until April 2014.
-An alternative to regulation is access to publicly available data.
-Comparative data, which we provide in this paper, and will continue to publish on our companion website, provides an incentive to both device manufacturers and network operators to provide updates.
+The FUM score enables corporate and public sector buyers, as well as individuals, to make more informed purchasing decisions by reducing the information asymmetry.
+The FUM score also supports better regulation, and indeed there is ongoing legal action to force network operators to ship updates for security vulnerabilities~\cite{Soghoian2013}.\todolater{Check on the status of this legal action}
+We will continue to provide updated versions of our FUM scores on our companion website~\cite{androidvulnerabilities.org}.
 
@@
  \item We measure the security of Android against our scoring metric and compare different device manufacturers, device models and network operators to allow device purchasers to differentiate between them based on security~(\S\ref{sec:security_scoring:results}).
- \item We determine that the main update bottleneck lies with manufacturers rather than operators or users~(\S\ref{sec:bottleneck}).
+ \item We determine that the main update bottleneck lies with manufacturers rather than Google, operators or users~(\S\ref{sec:bottleneck}).
 \end{itemize}

### 2015-07-28 security_scoring.tex

@@
 \todolater{Compute these values automatically and with error.}
-There is a variation between device models, with the update being distributed to most devices quickly and others having a much slower roll out, but since some do update quickly the bottleneck is unlikely to be with the user.
+There is a variation between device models, with the update being distributed to most devices quickly and others having a much slower roll out, but since some device models do update quickly the bottleneck is unlikely to be with the user.
 Perhaps some device models are preferred by users who are more likely to install updates than others, however we do observe updates being rolled out to device models quickly and user behaviour is not beyond the control of the device manufacturer.

### 2015-07-28 spsm04-thomas.tex
Fix duplicate LG

@@
 We find that on average \daMeanInsecurityPercNominal\ of Android devices are exposed to at least one of \daNumVulnsUsed\ known critical vulnerabilities and, across the ecosystem as a whole, assign a FUM security score of \daSecurityScoreNominal\ out of 10.
-In our data, Nexus devices do considerably better than average with a score of \daSecScoreBestsummaryScoreNominal; LG is the best manufacturer with a score of \daSecScoreBestmanufacturer\ \daSecScoreBestmanufacturerScoreNominal.
+In our data, Nexus devices do considerably better than average with a score of \daSecScoreBestsummaryScoreNominal; \daSecScoreBestmanufacturer\ is the best manufacturer with a score of \daSecScoreBestmanufacturerScoreNominal.
 %On average \daMeanOutstandingVulnerabilitiesNominal\ outstanding vulnerabilities are not fixed on any device and on average only \daUpdatednessPercNominal\ of devices run the current version of Android.

### 2015-07-28 spsm04-thomas.tex
Updates to threat model, data and ecosystem.

--- 2015-07-28 spsm04-thomas.tex
@@
 
-We assume the defender can deploy security controls at three levels: in an online marketplace, at app installation, and during app execution.
-Here, we consider three attack vectors against Android handsets.
+In this paper we are concerned with vulnerabilities which allow an attacker without physical access to the smartphone to gain significant permissions (such as root-level access) which are not available to a standard app running on the device.
+We consider three attack vectors which can be used as a starting point to launch an attack on a device.
 
-The \emph{installation} attack vector is through the installation of a malicious app on the device.
-Android devices can install apps through marketplaces, email attachments, URLs and via the Android Debug Bridge (ADB).
-By default, many Android devices will only allow the installation of apps from Google Play, which uses Bouncer to automatically analyse apps, and quickly takes down apps that are reported as malicious.
-However, alternative markets are also popular, particularly in countries where Google Play is not available.
+The \emph{installation} attack vector is used when a malicious app is installed on the device.
+Android devices can install apps through marketplaces such as the Google Play Store, email attachments, URLs and via the Android Debug Bridge (ADB).
+By default, many Android devices will only allow the installation of apps from the Play Store, which automatically analyses apps, and quickly takes down apps that are reported as malicious.
+However, alternative markets are also popular, particularly in countries where the Play Store is not available.
 
-The \emph{dynamic code loading} attack vector occurs when an existing app download and execute code at runtime.
-The most direct method is to upload to a marketplace a seemingly innocent app that dynamically loads malicious code, either as additional davik bytecode, as a native library or by embedding an interpreter and executing received instructions.
+The \emph{dynamic code loading} attack vector occurs when an existing app downloads and executes new code at runtime.
+The most direct method is to upload a seemingly innocent app to a marketplace that then dynamically loads malicious code, either as additional davik bytecode, as a native library, or by embedding an interpreter and executing received instructions.
 Neither static nor dynamic analysis of this app will uncover any malicious code, since it does not exist in the app.
 The marketplace can try to detect explicit use of dynamic code loading, however there are ways to dynamically load code which are hard to detect even on a platform such as iOS, which does not permit dynamic code loading.
-For example, a Return-Oriented Programming (ROP) attack is relatively easy if the attacker creates an app with carefully crafted flaws~\cite{Wang2013a}.
+For example, a Return-Oriented Programming (ROP) attack on iOS is relatively easy if the attacker creates an app with carefully crafted flaws~\cite{Wang2013a}.
 
-The \emph{injection} attack vector occurs when the attacker injects malicious code directly into existing code already on the handset. 
-For example, the addJavascriptInterface vulnerability (CVE-2012-6636) allows JavaScript running in an Android WebView to execute arbitrary code as the vulnerable app's user.
-The fix for the addJavascriptInterface vulnerability breaks backwards compatibility and requires a two sided fix.
-While the fix was released in December 2012, by \daGPAPISeventeenLaterDateMonth, \daGPAPISeventeenEarlierProportion\ of handsets connecting to Google Play were still vulnerable to this attack~\cite{Thomas2015a}.
+The \emph{injection} attack vector occurs when the attacker injects malicious code directly into existing code already running on the handset. 
+For example, the addJavascriptInterface vulnerability (CVE-2012-6636) allows an attacker to inject JavaScript into HTTP traffic destined for the device and execute arbitrary code with all the privileges of the app.
+The fix for the addJavascriptInterface vulnerability breaks backwards compatibility and requires a two-sided fix.
+While the fix was released in December 2012, by \daGPAPISeventeenLaterDateMonth, \daGPAPISeventeenEarlierProportion\ of handsets connecting to the Play Store were still vulnerable to this attack~\cite{Thomas2015a}.
 
-The best place to prevent attacks is at runtime, since all three attack vectors can be prevented at this level. 
-Unfortunately, the sandbox for Android apps is weak, since on most devices there are known critical vulnerabilities in it.
+Security for the Android ecosystem can be deployed at three levels: in an online marketplace, at app installation time on the device, and during app execution. 
+Indeed, Google provides its users with security in all these places: through analysis of apps by the Play Store, using the Verify Apps feature on the smartphone at installation time, and by an app sandbox on the smartphone during execution.
+The best place to prevent attacks is by sandboxing the app during execution, since all three attack vectors can be prevented at this level whereas not all users install apps exclusively via the Play Store or enable Verify Apps.
+In addition, dynamic code loading and injection attacks cannot be discovered at installation time and can be difficult for a marketplace to detect.
+Unfortunately, as we shall see, the security sandbox for Android has known critical vulnerabilities on most devices.
+This does not mean these devices are attacked, but that they are susceptible to it.
+The likelihood of a successful attack then depends on what apps the user installs and where from, as well as the computer networks the device is connected to and the actions the user takes whilst connected.
 
 \section{Data}
-We use two sources of data to measure the security of Android: (1) information on the critical vulnerabilities found to affect Android and (2) information on the distribution of Android versions over time.
+We use two sources of data to measure the security of Android: (1) information on the critical vulnerabilities found to affect particular versions of Android and (2) information on the distribution of Android versions over time.
 These two datasets can then be combined to determine the proportion of devices at risk of attack from specific vulnerabilities.
@@
 \subsection{Critical vulnerabilities}\label{sec:avo}
-We have used data on critical vulnerabilities in Android from AndroidVulnerabilities.org (\avo)~\cite{androidvulnerabilities.org}.
-It contains information on the discovery and publication dates, the versions affected and which versions contained fixes.
-It only contains critical vulnerabilities such as root vulnerabilities that did not require USB debugging to exploit.
-If malicious code exploits a critical vulnerability then it gains control of the device.
 
-Some critical vulnerabilities are not traditional kernel vulnerabilities, but are vulnerabilities in \emph{installation} in our threat model.
-For example improper verification of signatures at installation time was discovered in February 2013~\cite{Forristal2013} and meant that applications could pretend to be signed with system keys and hence be granted system privileges.
-On versions of Android below 4.1 malware could then use known system-to-root escalation mechanisms.
+We obtained our list of critical Android vulnerabilities from our AndroidVulnerabilities.org~(\avo) website~\cite{androidvulnerabilities.org}.
+The site contains \avoNumVulnerabilities\ critical vulnerabilities such as root vulnerabilities that do not require USB debugging to exploit.
+We have chosen \daNumVulnsUsed\ vulnerabilities as shown in Table~\ref{tab:andvulns} for our analysis in this paper.
+We selected these vulnerabilities since they fit the attack vectors introduced in \S\ref{sec:threatmodel} and because they affect all Android devices regardless of manufacturer, and as a result our selected vulnerabilities will dominate any security analysis of Android.
+Hence, with our chosen set of vulnerabilities, our analysis represents a lower-bound on the vulnerability of devices in the \da\ data.
+
+Some critical vulnerabilities are not traditional kernel vulnerabilities, but exploit the \emph{installation} attack vector in our threat model.
+For example improper verification of signatures at installation time was discovered in February 2013~\cite{Forristal2013} and meant that apps could pretend to be signed with system keys and hence be granted system privileges.
+On versions of Android below 4.1, malware could then use known system-to-root escalation mechanisms.
 Regardless of version, this exposed an increased attack area and would also provide the ability for malware to control all user internet traffic (via VPNs), brick the phone, remove and install apps, steal user credentials and read the screen.
-The different categories in which the vulnerabilities fall are shown in Table~\ref{tab:andvulns}, `signature' vulnerabilities are \emph{installation}, while `kernel' and `system' vulnerabilities are used with \emph{installation, dynamic code loading} and \emph{injection}.
+The different categories in which the vulnerabilities fall are shown in Table~\ref{tab:andvulns}.
+The `signature' vulnerabilities require an \emph{installation} attack, while `kernel' and `system' vulnerabilities can be used together with an \emph{installation}, \emph{dynamic code loading} or \emph{injection} attack vector.
 \avoTabAndVulns
 
-\avo\ currently contains \avoNumVulnerabilities\ vulnerabilities, of which \avoNumVulnAllAndroid\ affect all Android devices and \avoNumVulnSpecific\ are specific to particular devices or device manufacturers.
-
-We use the \daNumVulnsUsed\ vulnerabilities shown in Table~\ref{tab:andvulns} in our analysis.
-We have chosen these vulnerabilities to fit the attack vectors introduced in \S\ref{sec:threatmodel}.
-These vulnerabilities affect all Android devices regardless of manufacturer, and as a result our selected vulnerabilities will dominate any security analysis of Android.
-Hence, with our chosen set of vulnerabilities, our analysis represents a lower-bound on the vulnerability of devices in the \da\ data.
+%\avo\ currently contains \avoNumVulnerabilities\ vulnerabilities, of which \avoNumVulnAllAndroid\ affect all Android devices and \avoNumVulnSpecific\ are specific to particular devices or device manufacturers.
 
@@
 We use historical data collected by the \da\ project~\cite{Wagner2013}.
-\da\ collects data\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/collected.html}} from study participants who install the Android app from Google Play.
-Most study participants allow researchers a\-round the world to access a subset of their device data, including the data presented here.
+\da\ collects data\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/collected.html}} from study participants who install the Android app from the Play Store.
+Most study participants allow external researchers to access the subset of the device data needed for this analysis.
 
 We extracted the build string and API version for each device each day.
-The build string is a user-readable version string.
-The API version is a positive integer that increases when new features are added to the API.
+The build string is a user-readable version string and the API version is a positive integer that increases when new features are added to the API.
 Consequently security (bug) fixes do not always result in a change in the API version.
@@
 On a large proportion of devices `opaque\_marker' is a well defined build number\footnote{\url{https://source.android.com/source/build-numbers.html}} however different device manufacturers use different schema.
-Google provides API version distribution information but not the OS and build version information we need.
+%Google provides API version distribution information but not the OS and build version information we need.
+
+\da\ has collected data from \daNumDevices\ devices with a total of \daDeviceDays\ device days.
+The majority of devices only contribute data for a short period of time, however \daMonthsDevices\ devices have contributed data for more than \daMonths~months.
 We verify that the \da\ data is representative in~\S\ref{sec:representative}.
-
-\da\ has collected data from \daNumDevices\ devices with a total of \daDeviceDays device days.
-The majority of devices only contribute data for a short period of time, however \daMonthsDevices\ devices have contributed data for more than \daMonths~months.
 
@@
 This normalisation is a manual task that involves removing invalid values (such as `manufacturer' or `airplane mode is on'), collating across company name changes (e.g.\ `lge' to `LG'), normalising punctuation, removing extra strings sometimes added such as (`(2g)' or `communications') and mapping some incorrectly placed model names back to their manufacturer.
-This normalisation is not perfect and so these are overestimates on the \da\ data but they are likely still underestimates as there will be some device manufacturers and network operators that are not included in the \da\ data.
-The representativeness of \da\ is discussed in \S\ref{sec:representative}.
+This normalisation is not perfect so these are likely overestimates on the \da\ data.
+We believe they nevertheless are likely to underestimate the total number of device manufacturers and network operators worldwide.
+%The representativeness of \da\ is discussed in \S\ref{sec:representative}.
 
@@
 To compute the latency between upstream releases and the release of the first version of Android containing that release we scraped the release pages, to obtain the version numbers and release dates.
-This allows us to compute the latency between an upstream project being released and it being included in Android, this is shown in Table~\ref{tab:update_ecosystem}.
+This allows us to compute the latency between an upstream project being released and it being included in Android; this is shown in Table~\ref{tab:update_ecosystem}.
 The versions included in Android were about half a year old when the first version of Android containing it was released.

### 2015-07-28 security_scoring.tex
Rewordings to make it fit on 12 pages again

@@
 By combining data on critical vulnerabilities in Android and the versions of Android running on devices we can determine which vulnerabilities each device was vulnerable to each day.
-Then we consider a device is \emph{insecure} if it is running a vulnerable version of Android and the device has not received an update which might fix it;
+We consider a device is \emph{insecure} if it is running a vulnerable version of Android and the device has not received an update which might fix it;
 it is \emph{maybe secure} if it is running a vulnerable version but received an update which could have fixed the vulnerability if it contained a backported fix;
@@
 This demonstrates the importance of a longitudinal study: this type of analysis requires years of data.
-Once \avovuln{zergRush} was discovered in October 2011 then most devices are recorded as \emph{insecure} (red) as they were exposed to that vulnerability.
+Once \avovuln{zergRush} was discovered in October 2011 then most devices are recorded as \emph{insecure} (red) as they were vulnerable.
 The remaining devices were already running a version of Android which fixed the \avovuln{zergRush} vulnerability and are therefore marked as \emph{secure} (green).
-From October 2011 until the discovery of \avovuln[APK_duplicate_file]{APK duplicate file} in February 2013 the graph shows progressive improvement as devices are upgraded.
+From October 2011 until the discovery of \avovuln[APK_duplicate_file]{APK duplicate file} in February 2013 the graph shows progressive improvement as devices are upgraded or replaced.
 This means more and more devices are marked as \emph{secure} because they are now running a secure version of Android, or marked as \emph{maybe secure} because they received an OS update that did not update to a known-good version of Android but which may still have included a backport of a fix, as the update was made available after the vulnerability was disclosed.
-From February 2013 onwards regular discovery of critical vulnerabilities ensures that most devices are exposed to known critical vulnerabilities.
+From February 2013 onwards regular discovery of critical vulnerabilities ensures that most devices are vulnerable.
 This means that \daMeanInsecurityPerc\ of devices were classified as \emph{insecure} to at least one critical vulnerability between \daStartDateMonth\ and \daEndDateMonth\ (ignoring devices classed as \emph{maybe secure}).
@@
 
-We want to provide a score out of 10 as this is easy for phone buyers to understand because many ratings are given as a score out of 10.
+We want to provide a score out of 10 as many other ratings are given as a score out of 10.
 Since $f$ is the most important metric we weight it more highly.
@@
 Figure~\ref{fig:security_scores} shows this for manufacturers, device models, network operators and for Nexus and non-Nexus devices.
-These show how the scores for different entities are different and change over time, while there is correlated behaviour for different entities (for instance, when new vulnerabilities affecting all Android are discovered), these lines still have crossings due to the different behaviour of the various entities.
+These show how the scores for different entities are different and change over time, while there is correlated behaviour for different entities (for instance, when new vulnerabilities affecting all Android are discovered), these lines still cross due to the different behaviour of the entities.
 It also shows that we do not have sufficient data for all the entities all of the time, resulting in gaps in the data.

### 2015-07-28 spsm04-thomas.tex
Rewordings to make it fit on 12 pages again

@@
 The FUM score also supports better regulation, and indeed there is ongoing legal action to force network operators to ship updates for security vulnerabilities~\cite{Soghoian2013}.\todolater{Check on the status of this legal action}
-We will continue to provide updated versions of our FUM scores on our companion website~\cite{androidvulnerabilities.org}.
+We will continue to provide updated versions of our FUM scores on our website~\cite{androidvulnerabilities.org}.
 
@@
 For example, the addJavascriptInterface vulnerability (CVE-2012-6636) allows an attacker to inject JavaScript into HTTP traffic destined for the device and execute arbitrary code with all the privileges of the app.
-The fix for the addJavascriptInterface vulnerability breaks backwards compatibility and requires a two-sided fix.
+The fix for this vulnerability breaks backwards compatibility and requires a two-sided fix.
 While the fix was released in December 2012, by \daGPAPISeventeenLaterDateMonth, \daGPAPISeventeenEarlierProportion\ of handsets connecting to the Play Store were still vulnerable to this attack~\cite{Thomas2015a}.
@@
 Fortunately most (\daOSVersionPercValidLines) entries in these data have a build string of the form `x.y.z opaque\_marker' and so it is possible to extract the Android version number `x.y.z'.
-On a large proportion of devices `opaque\_marker' is a well defined build number\footnote{\url{https://source.android.com/source/build-numbers.html}} however different device manufacturers use different schema.
+On a large proportion of devices `opaque\_marker' is a well defined build number\footnote{\url{https://source.android.com/source/build-numbers.html}} however this is not universal.
 %Google provides API version distribution information but not the OS and build version information we need.
@@
 
-We extracted data on the external projects used in Android and have included this and the scripts which generated it in AndroidVulnerabilities.org (\avo)~\cite{androidvulnerabilities.org}.
+We extracted data on the external projects used in Android and have included this and the scripts which generated it in \avo.
 These scripts analysed the Android Open Source Project's source tree to examine the source code of each of the external projects to find the project version associated with each Android version tag on the repository.
@@
 \section{Comparison with other data}\label{sec:representative}
-We compare the Device Analyzer data with three other sources of data and find similar distributions which bound the Device Analyzer data, indicating that it is likely to be typical of the true distribution.
+We compare the Device Analyzer data with three other sources of data and find similar distributions which bound the Device Analyzer data, indicating that it is likely to be typical.
 We have obtained comparable data on 5\,290 devices from a multinational FTSE 100 company's mobile device management database, which includes company and employee owned devices, and from 5\,170\,000 matching User-Agent headers on all HTTP traffic for 30\% of Rwanda for a week.

### 2015-07-28 spsm04-thomas.tex
We built the list of Android vulnerabilities
Eject the introduction section title onto the next column and make it fit on 12 pages again

@@
 \keywords{Android; updates; vulnerabilities; metrics; ecosystems}
-
+\vfill\eject
 \section{Introduction}
@@
 
-We obtained our list of critical Android vulnerabilities from our AndroidVulnerabilities.org~(\avo) website~\cite{androidvulnerabilities.org}.
+We built a list of critical Android vulnerabilities for our AndroidVulnerabilities.org~(\avo) website~\cite{androidvulnerabilities.org}.
 The site contains \avoNumVulnerabilities\ critical vulnerabilities such as root vulnerabilities that do not require USB debugging to exploit.
@@
 There is a complex Android ecosystem that creates and distributes updates which fix vulnerabilities.
-In this section we describe how the Android ecosystem functions and how Android versions are produced, as seen by the \da\ project and by analysing the Android source code and upstream projects.
+In this section we describe how the Android ecosystem functions and how Android versions are produced, using \da\ data and by analysing the Android source code and upstream projects.
 We quantify the number of updates shipped by various entities in the ecosystem and the number of entities.

### 2015-07-28 spsm04-thomas.tex
Tweak to abstract; updated conclusion.

--- 2015-07-28 spsm04-thomas.tex
@@
 \begin{abstract}
-The security of Android depends on the timely delivery of security updates for critical vulnerabilities across a complex ecosystem.
-In this paper we map the major players in the Android ecosystem who must collaborate to provide updates, and determine that inaction by some manufacturers and network operators means many handsets are vulnerable to critical vulnerabilities.
+The security of Android depends on the timely delivery of updates to fix critical vulnerabilities.
+In this paper we map the complex network of players in the Android ecosystem who must collaborate to provide updates, and determine that inaction by some manufacturers and network operators means many handsets are vulnerable to critical vulnerabilities.
 We define the FUM security metric to rank the performance of device manufacturers and network operators, based on their provision of updates and exposure to critical vulnerabilities.
@@
 Unfortunately something has gone wrong with the provision of security updates in the Android market. 
-Many smartphones are sold on 12--24 month contracts, and yet our data shows few Android devices receive many security updates, with an overall average of just \daUpdatesPerYearNominal\ updates per year, leaving devices unpatched for many months or years.
+Many smartphones are sold on 12--24 month contracts, and yet our data shows few Android devices receive many security updates, with an overall average of just \daUpdatesPerYearNominal\ updates per year, leaving devices unpatched for long periods of time.
 
@@
 \section{Conclusion}
-We quantified how the Android ecosystem results in different device models receiving different levels of security due to whether or not they get security updates.
-We found that on average \daMeanInsecurityPercNominal\ of Android devices are exposed to at least one of \daNumVulnsUsed\ known critical vulnerabilities.
-On average \daMeanOutstandingVulnerabilitiesNominal\ outstanding vulnerabilities are not fixed on any device and on average only \daUpdatednessPercNominal\ of devices run the current version of Android.
-We have defined the FUM score, which evaluates the security performance of companies and organisations in shipping security updates.
-We have used this to compare different device models, device manufacturers and network operators and found that there are differences between them, which the discerning purchaser might use to influence their decision about which device model to buy from which device manufacturer and network operator.
-While Android has a score of \daSecurityScoreNominal\ out of 10, Nexus devices get \daSecScoreBestsummaryScoreNominal\ and \daSecScoreBestmanufacturer\ \daSecScoreBestmanufacturerScoreNominal.
-We hope that this analysis will encourage device manufacturers and network operators to improve the support they provide for devices after sale.
+The security of Android depends on the timely delivery of updates to fix critical vulnerabilities.
+Unfortunately few devices receive prompt updates, with an overall average of \daUpdatesPerYearNominal\ updates per year, leaving devices unpatched for long periods.
+We showed that the bottleneck for the delivery of updates in the Android ecosystem rests with the manufacturers and network operators, who fail to provide updates to fix critical vulnerabilities.
+This arises in part because the market for Android security today is like the market for lemons: there is information asymmetry between the manufacturer, who knows whether the device is currently secure and will receive updates, and the consumer, who does not.
 
-The FUM score is applicable to other software ecosystems, as long as information on vulnerabilities and the versions they affected, as well as data on which versions were running on real devices is available.
+Consequently there is little incentive for manufacturers to provide updates.
+To address this issue we developed the FUM security metric to quantify and rank the performance of device manufacturers and network operators, based on their provision of updates and exposure to critical vulnerabilities.
+The metric enables purchasers and regulators to determine which device manufacturers and network operators provide updates and which do not.
+
+Using a corpus of \daNumOSDevices\ devices we demonstrated that there is significant variability in the timely delivery of security updates across different device manufacturers and network operators.
+We find that on average \daMeanInsecurityPercNominal\ of Android devices are exposed to at least one of \daNumVulnsUsed\ known critical vulnerabilities and, across the ecosystem as a whole, assign a FUM security score of \daSecurityScoreNominal\ out of 10.
+In our data, Nexus devices do considerably better than average with a score of \daSecScoreBestsummaryScoreNominal; \daSecScoreBestmanufacturer\ is the best manufacturer with a score of \daSecScoreBestmanufacturerScoreNominal.
 

### 2015-07-28 spsm04-thomas.tex
Abstract tweak.
Removed network operator from conclusions.

@@
 We find that on average \daMeanInsecurityPercNominal\ of Android devices are exposed to at least one of \daNumVulnsUsed\ known critical vulnerabilities and, across the ecosystem as a whole, assign a FUM security score of \daSecurityScoreNominal\ out of 10.
-In our data, Nexus devices do considerably better than average with a score of \daSecScoreBestsummaryScoreNominal; \daSecScoreBestmanufacturer\ is the best manufacturer with a score of \daSecScoreBestmanufacturerScoreNominal.
+In our data, Nexus devices do considerably better than average with a score of \daSecScoreBestsummaryScoreNominal; and \daSecScoreBestmanufacturer\ is the best manufacturer with a score of \daSecScoreBestmanufacturerScoreNominal.
 %On average \daMeanOutstandingVulnerabilitiesNominal\ outstanding vulnerabilities are not fixed on any device and on average only \daUpdatednessPercNominal\ of devices run the current version of Android.
@@
 Unfortunately few devices receive prompt updates, with an overall average of \daUpdatesPerYearNominal\ updates per year, leaving devices unpatched for long periods.
-We showed that the bottleneck for the delivery of updates in the Android ecosystem rests with the manufacturers and network operators, who fail to provide updates to fix critical vulnerabilities.
+We showed that the bottleneck for the delivery of updates in the Android ecosystem rests with the manufacturers, who fail to provide updates to fix critical vulnerabilities.
 This arises in part because the market for Android security today is like the market for lemons: there is information asymmetry between the manufacturer, who knows whether the device is currently secure and will receive updates, and the consumer, who does not.

### 2015-07-28 security_scoring.tex
Minor fixes

@@
 From February 2013 onwards regular discovery of critical vulnerabilities ensures that most devices are vulnerable.
-This means that \daMeanInsecurityPerc\ of devices were classified as \emph{insecure} to at least one critical vulnerability between \daStartDateMonth\ and \daEndDateMonth\ (ignoring devices classed as \emph{maybe secure}).
+Ignoring devices classed as \emph{maybe secure}, we find that on average \daMeanInsecurityPerc\ of devices were classed as \emph{insecure} and \daMeanSecurityPercNominal\ classified as \emph{secure} between \daStartDateMonth\ and \daEndDateMonth.
 %This is $1 - f$, the proportion of devices free from critical vulnerabilities, however this is only one component of a more robust score, the FUM score.
@@
 This does not capture systematic errors.
-For example, we do not include manufacturer specific vulnerabilities, we expect that performance in fixing manufacturer specific vulnerabilities is strongly correlated with performance fixing vulnerabilities affecting all of Android.
+For example, we do not include manufacturer specific vulnerabilities, however we expect that performance in fixing manufacturer specific vulnerabilities is strongly correlated with performance fixing vulnerabilities affecting all of Android.
 
@@
 \subsection{Results: Security scores}\label{sec:security_scoring:results}
-The scores across the whole of Android are that \daMeanInsecurityPerc\ of devices are exposed to known root exploits.
-There are on average \daMeanOutstandingVulnerabilities\ outstanding vulnerabilities not fixed on any device.
-On average \daUpdatednessPerc\ of devices run the most recent version of Android.
+On average, between \daStartDateMonth\ and \daEndDateMonth\, we found \daMeanOutstandingVulnerabilities\ outstanding vulnerabilities not fixed on any device and \daUpdatednessPerc\ of devices to run the most recent version of Android.
 This gives a security score of \daSecurityScore\ out of 10.
@@
 However there are a wide variety of scores depending on the source of the device.
-There is anecdotal evidence that Google's Nexus devices are better at getting updates than other Android devices because Google makes the original updates and ships them to its devices\footnote{\url{http://www.howtogeek.com/139391/htg-explains-why-android-geeks-buy-nexus-devices/}}.
+There is anecdotal evidence that Google's Nexus devices are better at getting updates than other Android devices because Google makes the original updates and ships them to its devices.\footnote{\url{http://www.howtogeek.com/139391/htg-explains-why-android-geeks-buy-nexus-devices/}}
 Table~\ref{tab:sec_summary} shows that this is the case with Nexus devices getting much better scores than non-Nexus devices.
@@
 Figure~\ref{fig:full_version_comp}c shows the full version distribution for \emph{\daSecScoreWorstmodel}, which we only observe running one version.
-Figure~\ref{fig:full_version_comp}b shows the full version distribution for \emph{HTC Desire HD A9191}, which used to be our worst model and for which we have more historical data; it shows how it received one update at the beginning of 2012, which was deployed fairly rapidly to most devices, but received no further updates.
+Figure~\ref{fig:full_version_comp}b shows the full version distribution for \emph{HTC Desire HD A9191}, which used to be our worst model and for which we have more historical data; it shows it received one update at the beginning of 2012, which was deployed fairly rapidly to most devices, but received no further updates.
 Figure~\ref{fig:full_version_comp}a shows the same information for \emph{\daSecScoreBestmodel} which received \daSecScoreBestmodelNumFullVersions\ different versions, some of which were only deployed to small numbers of devices, but the distribution for all devices regularly and rapidly transitions from one version to another before ending up on `4.3 JWR66Y'.
@@
 We also analysed the \daNumSigOperators\ network operators with a significant presence in our data.
-Table~\ref{tab:sec_operator} shows the results \emph{\daSecScoreBestoperator} (\daSecScoreBestoperatorScore\ out of 10) scoring highest and \emph{\daSecScoreWorstoperator} (\daSecScoreWorstoperatorScore\ out of 10) scoring lowest.
+Table~\ref{tab:sec_operator} shows the results with \emph{\daSecScoreBestoperator} (\daSecScoreBestoperatorScore\ out of 10) scoring highest and \emph{\daSecScoreWorstoperator} (\daSecScoreWorstoperatorScore\ out of 10) scoring lowest.
 However, the score of a network operator is affected by the manufacturers of the devices which are in use on its network.
@@
 To evaluate whether the ranking of different manufacturers is sensitive to the form of the scoring metric we computed the normalised Spearman's Rank correlation coefficient between the lists ordered using different forms of the scoring metric, this is shown in Table~\ref{tab:spearman_ranks}.
-The `equal' metric weights $f$, $u$ and $m$ equally rather than favouring $f$ and makes little difference.
+In the table, the `equal' metric weights $f$, $u$ and $m$ equally rather than favouring $f$ and makes little difference.
 Similarly weighting $u$ or $m$ more highly rather than $f$ makes little difference.

### 2015-07-28 spsm04-thomas.tex
Drop DOIs for lack of space :-(
Minor fixes

@@
 \newcommand{\todolater}[1]{}%{\todo{#1}}%
-\RequirePackage[date=terse, isbn=true, doi=true, url=false, urldate=iso8601, maxbibnames=9, backref=false, backend=bibtex, style=ieee]{biblatex}
+\RequirePackage[date=terse, isbn=true, doi=false, url=false, urldate=iso8601, maxbibnames=9, backref=false, backend=bibtex, style=ieee]{biblatex}
 \addbibresource{securityupdates.bib}
@@
 It is a complex system with many parties involved in a long multi-stage pipeline~\cite{HTC2013}.
-We map and quantify the major players in this space who must collaborate to provide updates~(\S\ref{sec:android_ecosystem}) and determine that inaction~(\S\ref{sec:bottleneck}) by some of the manufacturers (we know of \daNumManufacturers) and network operators (\daNumOperators) means many handsets are vulnerable to critical vulnerabilities.
+We map and quantify the major players in this space who must collaborate to provide updates~(\S\ref{sec:android_ecosystem}) and determine that inaction~(\S\ref{sec:bottleneck}) by some of the manufacturers and network operators means many handsets are vulnerable to critical vulnerabilities.
 Understanding this ecosystem is all the more important because device manufacturers have introduced additional vulnerabilities in the past~\cite{Grace2012}.
@@
 Security for the Android ecosystem can be deployed at three levels: in an online marketplace, at app installation time on the device, and during app execution. 
-Indeed, Google provides its users with security in all these places: through analysis of apps by the Play Store, using the Verify Apps feature on the smartphone at installation time, and by an app sandbox on the smartphone during execution.
-The best place to prevent attacks is by sandboxing the app during execution, since all three attack vectors can be prevented at this level whereas not all users install apps exclusively via the Play Store or enable Verify Apps.
+Google provides its users with security in all these places: through analysis of apps by the Play Store, using the Verify Apps feature on the smartphone at installation time, and by an app sandbox on the smartphone during execution.
+The best place to prevent attacks is by sandboxing the app during execution, since all three attack vectors can be prevented at this level, whereas not all users install apps exclusively via the Play Store or enable Verify Apps.
 In addition, dynamic code loading and injection attacks cannot be discovered at installation time and can be difficult for a marketplace to detect.
 Unfortunately, as we shall see, the security sandbox for Android has known critical vulnerabilities on most devices.
-This does not mean these devices are attacked, but that they are susceptible to it.
+This does not mean these devices are attacked, but that they are vulnerable.
 The likelihood of a successful attack then depends on what apps the user installs and where from, as well as the computer networks the device is connected to and the actions the user takes whilst connected.
@@
 We used \da\ to record data on the OS versions of Android installed on devices.
-Nappa et al.\ performed a similar analysis using WINE to record data on installed versions of Windows client programs and used cleaned NVD data rather than \avo\ for vulnerability data, they did not assign client applications scores~\cite{Nappa2015}.
+Nappa et al.\ performed a similar analysis using WINE to record data on installed versions of Windows client programs and used cleaned NVD data rather than \avo\ for vulnerability data.
+They did not assign client applications scores~\cite{Nappa2015}.
 
@@
 It is possible that many updates arrive at handsets, but are simply not installed.
-Anecdotal evidence at least suggests that it is the lack of updates rather the lack of installation, which is the major problem at present and \S\ref{sec:security_scoring:results} supports this.
-Further work is required to tease these numbers apart.
-Partly this is the result of the fact that an operating system update is being installed and so a reboot is required, but Chrome installs the new version side by side with the old one and switches the next time it is restarted.
+Anecdotal evidence suggests that it is the lack of updates rather the lack of installation, which is the major problem at present.
+Our analysis in \S\ref{sec:bottleneck} supports this view.
+Nevertheless it is problematic that an operating system update requires a reboot.
+Chrome installs the new version side by side with the old one and switches the next time it is restarted.
 The same technique would be more difficult on phones with limited storage space (as many cheap Android phones have barely enough space to install just the update) but is a plausible improvement for more high-end devices.
@@
 Data from \da\ and \avo\ used in this paper is already available.
-The Rwanda and FTSE data cannot be made available due to the terms they were obtained under.
+The Rwanda and FTSE data cannot be made available.
 

### 2015-07-29 spsm04-thomas.tex
Use the ACM bibliography style

@@
 \documentclass{sig-alternate-2013}
+\setlength{\paperheight}{11in}
+\setlength{\paperwidth}{8.5in}
 
@@
 \newcommand{\todolater}[1]{}%{\todo{#1}}%
-\RequirePackage[date=terse, isbn=true, doi=false, url=false, urldate=iso8601, maxbibnames=9, backref=false, backend=bibtex, style=ieee]{biblatex}
+\RequirePackage[date=terse, isbn=true, doi=false, url=false, urldate=iso8601, maxbibnames=9, backref=false, backend=bibtex,firstinits=true]{biblatex}
+
+\DeclareNameAlias{default}{last-first}
+
+\AtBeginBibliography{%
+  \renewcommand*{\mkbibnamelast}[1]{\textsc{#1}}%
+  %% commas between authors
+  \renewcommand{\multinamedelim}{\addcomma\space}
+  \renewcommand{\finalnamedelim}{\addcomma\addspace\textsc{and}\space}
+}
+
+\DefineBibliographyStrings{english}{%
+ andothers = {\addcomma\addspace\textsc{et\addabbrvspace al}\adddot},
+ and = {\textsc{and}}
+}
+
+\renewcommand*{\labelnamepunct}{\space\space}
+
+\DeclareFieldFormat
+  [article,inbook,incollection,inproceedings,patent,thesis,unpublished]
+  {title}{#1}
+
+\renewbibmacro{in:}{%
+  \ifentrytype{article}{%
+  }{%
+    \printtext{\bibstring{in}\intitlepunct}%
+  }%
+}
+
+\renewbibmacro*{volume+number+eid}{%
+  \printfield{volume}%
+  \setunit*{\addcomma\space}%
+  \printfield{number}%
+  \setunit{\addcomma\space}%
+  \printfield{eid}}
+
+\DeclareFieldFormat{pages}{#1}
+
+\renewbibmacro*{publisher+location+date}{%
+  \printlist{publisher}%
+  \setunit*{\addcomma\space}%
+  \printlist{location}%
+  \setunit*{\addcomma\space}%
+  \usebibmacro{date}%
+  \newunit}
+
 \addbibresource{securityupdates.bib}
@@
 }
-\sloppy
+\raggedright
 \printbibliography

### 2015-09-16 cuttingfloor.tex
Remove some old stuff that is no longer relevant

@@
 %\todo{Get data from Richard so that we can do this comparison}
-
-\subsection{Experiment 4: Predicting future version distributions}\label{sec:exp:predicting_distributions}
-\subsubsection{Method}
-Examining the way that the OS version distribution changes over time in Figure~\ref{fig:norm_os} we see some regularity to the behaviour.
-To show this behaviour more clearly we normalised the data to the number of days since the version was released and smoothed it by forcing it to be monotonically decreasing.\footnote{While sometimes an older version becomes more popular after a later version has been released causing the values to actually increase, mostly this was just noise or changes in the distribution of devices (e.g.\ the sudden arrival of 1000 devices from Bangladesh due to a study being conducted there) and this assumption makes the graph clearer.}
-Then we can plot Figure~\ref{fig:frvh_os_versions} which shows the proportion of of devices in \da\ not upgraded to each Android version or a later Android version (i.e.
-for 2.1.0 it includes 2.1.1, 2.2.0 and indeed all devices as they all run later versions of Android).
-The mean and standard deviation of the same data are shown in Figure~\ref{fig:frms_os_versions}.
-
-\begin{figure}
- \includegraphics[width=\columnwidth]{figures/frvh_os_versions}
- \caption{Proportion of devices not updated to particular versions of Android or any later version}
- \label{fig:frvh_os_versions}
-\end{figure}
-\begin{figure}
- \includegraphics[width=\columnwidth]{figures/frms_os_versions}
- \caption{Proportion of devices not updated to particular versions of Android or any later version}
- \label{fig:frms_os_versions}
-\end{figure}
-
-The data appears to have an exponential decay as it tends to 1 but with a slow start as few devices are upgraded initially.
-We approximate this with a $e^{-x}$ for the main behaviour and a which is offset to start later when uptake takes off.
-The equation is then:
-\begin{equation}
-y =
- \begin{cases}
-  1.0 & \text{if }x<\mathrm{undeployed}\\
-  e^{-\mathrm{decay}(x-\mathrm{undeployed})} & \text{otherwise}
- \end{cases}
-\end{equation}
-
-\subsubsection{Results}
-Fitting that curve to the data using \texttt{scipy.optimize}, we get a RMSE of \daOSCurveFitRMSE\ and parameters $\mathrm{undeployed} = \daOSCurveFitParamFirst ~\si{days}$, $\mathrm{decay} = \daOSCurveFitParamSecond ~\si{days^{-1}}$ for OS versions and RMSE of \daAPICurveFitRMSE\, $\mathrm{undeployed} = \daAPICurveFitParamFirst ~\si{days}$, $\mathrm{decay} = \daAPICurveFitParamSecond ~\si{days^{-1}}$ for API versions.
-From this the number of days from release of a new version of Android until 50\% of devices are running that version or higher is \daOSCurveHalfDeployed\ and full deployment to \daFullDeployedAt\ of devices takes \daOSCurveFullDeployed\ days.
-For API versions this is \daAPICurveHalfDeployed\ and \daAPICurveFullDeployed\ respectively.
-
-Hence if a security vulnerability is fixed through the release of a particular Android version it will be \daOSCurveFullDeployed\ days after until the fix is fully deployed.
-
-The equation we chose for the model has a root mean squared error (RMSE) of \daOSCurveFitRMSE\ for OS versions, which is a comparably good fit to a standard polynomial fit (3 degree polynomial fit gave a RMSE of \daOSCurvePolyRMSE) or a spline fit (RMSE of \daOSCurveSplineRMSE) but gives a meaningful model of behaviour rather than a generic curve.
-
-\todo{Our handling of errors here is not quite good enough as the uncertainty from our y value interacts badly with the logarithm function when computing the inverse and we compute the log of ~0 which means our error ends up much bigger than our value - but that is an upper bound, not a +- bound.}
-
-We hope that a consequence of this paper will be a change in behaviour of the Android ecosystem which invalidates this model by speeding up the deployment of new OS versions.

### 2015-10-02 avostats.tex
./update.sh

@@
-\newcommand{\avoNumSubmitters}{7}
-\newcommand{\avoTotalExternalLines}{25 Million}
+\newcommand{\avoNumSubmitters}{8}
+\newcommand{\avoTotalExternalLines}{25 million}
 \newcommand{\avoNumExternalProjects}{176}
-\newcommand{\avoNumBigExternalLinesOfCode}{25 Million}
+\newcommand{\avoNumBigExternalLinesOfCode}{25 million}
 \newcommand{\avoBigExternalLinesOfCodePerc}{99.7\%}
-\newcommand{\avoNumVulnerabilities}{32}
-\newcommand{\avoNumVulnAllAndroid}{15}
+\newcommand{\avoNumVulnerabilities}{33}
+\newcommand{\avoNumVulnAllAndroid}{16}
 \newcommand{\avoNumVulnSpecific}{17}
@@
 \newcommand{\avoLastDataDate}{2015-03-12}
-\newcommand{\avoVulnsPerYearNominal}{6.86}
-\newcommand{\avoVulnsPerYear}{$6.86 \pm 1.21$}
-\newcommand{\avoVulnsPerYearAllAndroidNominal}{3.21}
-\newcommand{\avoVulnsPerYearAllAndroid}{$3.21 \pm 0.83$}
-\newcommand{\avoVulnsPerYearTwosfNominal}{6.9}
-\newcommand{\avoVulnsPerYearTwosf}{$6.9 \pm 1.2$}
-\newcommand{\avoVulnsPerYearAllAndroidTwosfNominal}{3.2}
-\newcommand{\avoVulnsPerYearAllAndroidTwosf}{$3.2 \pm 0.0$}
-\newcommand{\avoTabAndVulns}{\begin{table} \centering \small \begin{tabular}{l|l|c|c} Vulnerability & How known & Date & Categories\\ \hline KillingInTheNameOf & Fixed & 2010-07-13 & sys, kern\\ exploid udev & Discovered & 2010-07-15 & kernel\\ levitator & Discovered & 2011-03-10 & kernel\\ Gingerbreak & Fixed & 2011-04-18 & system\\ zergRush & Discovered & 2011-10-06 & system\\ APK duplicate file & Discovered & 2013-02-18 & signature\\ APK unchecked name & Discovered & 2013-06-30 & signature\\ APK unsigned shorts & Fixed & 2013-07-03 & signature\\ vold asec & Fixed & 2014-01-27 & system\\ Fake ID & Fixed & 2014-04-17 & signature\\ TowelRoot & Discovered & 2014-05-03 & kernel\\\end{tabular} \caption{Critical vulnerabilities in Android} \label{tab:andvulns} \end{table}}
+\newcommand{\avoVulnsPerYearNominal}{7.07}
+\newcommand{\avoVulnsPerYear}{$7.07 \pm 1.23$}
+\newcommand{\avoVulnsPerYearAllAndroidNominal}{3.43}
+\newcommand{\avoVulnsPerYearAllAndroid}{$3.43 \pm 0.85$}
+\newcommand{\avoVulnsPerYearTwosfNominal}{7.1}
+\newcommand{\avoVulnsPerYearTwosf}{$7.1 \pm 1.2$}
+\newcommand{\avoVulnsPerYearAllAndroidTwosfNominal}{3.4}
+\newcommand{\avoVulnsPerYearAllAndroidTwosf}{$3.4 \pm 0.0$}
+\newcommand{\avoTabAndVulns}{\begin{table} \centering \small \begin{tabular}{l|l|c|c} Vulnerability & How known & Date & Categories\\ \hline KillingInTheNameOf & Fixed on & 2010-07-13 & system, kernel\\ exploid udev & Discovered on & 2010-07-15 & kernel\\ levitator & Discovered on & 2011-03-10 & kernel\\ Gingerbreak & Fixed on & 2011-04-18 & system\\ zergRush & Discovered on & 2011-10-06 & system\\ APK duplicate file & Discovered on & 2013-02-18 & signature\\ APK unchecked name & Discovered on & 2013-06-30 & signature\\ APK unsigned shorts & Fixed on & 2013-07-03 & signature\\ vold asec & Fixed on & 2014-01-27 & system\\ Fake ID & Fixed on & 2014-04-17 & signature\\ TowelRoot & Discovered on & 2014-05-03 & kernel\\ ObjectInputStream_deserializable & Discovered on & 2014-06-22 & system\\\end{tabular} \caption{Critical vulnerabilities in Android} \label{tab:andvulns} \end{table}}
 \newcommand{\avoNumBigExternalProjects}{40}
 \newcommand{\avoNumAnalysedExternalProjects}{28}
-\newcommand{\avoNumAnalysedExternalLinesOfCode}{6 Million}
+\newcommand{\avoNumAnalysedExternalLinesOfCode}{6 million}
 \newcommand{\avoAnalysedExternalLinesOfCodePerc}{24.9\%}

### 2015-10-02 dastats.tex
./update.sh

@@
 \newcommand{\daOSYearsOfData}{4}
-\newcommand{\daOSMonthsOfData}{49}
+\newcommand{\daOSMonthsOfData}{51}
 \newcommand{\daStartDate}{2011-07-01}
-\newcommand{\daEndDate}{2015-07-19}
+\newcommand{\daEndDate}{2015-09-22}
 \newcommand{\daFullDeployedAt}{95\%}
-\newcommand{\daNumManufacturers}{301}
+\newcommand{\daNumManufacturers}{302}
 \newcommand{\daNumSigManufacturers}{10}
-\newcommand{\daNumModels}{2\,740}
-\newcommand{\daNumOperators}{1\,460}
+\newcommand{\daNumModels}{2\,770}
+\newcommand{\daNumOperators}{1\,470}
 \newcommand{\daNumSigOperators}{14}
-\newcommand{\daNumOSDevices}{20\,400}
+\newcommand{\daNumOSDevices}{20\,500}
 \newcommand{\daOSVersionPercValidLines}{99.9\%}
-\newcommand{\daOSTotalDaysData}{1\,330\,000}
-\newcommand{\daMeanInsecurityPercNominal}{87.7\%}
-\newcommand{\daMeanInsecurityPerc}{$87.7 \pm 0.0\%$}
-\newcommand{\daMeanInsecurityPercTwosfNominal}{88\%}
-\newcommand{\daMeanInsecurityPercTwosf}{$88 \pm 0\%$}
-\newcommand{\daMeanSecurityPercNominal}{12.3\%}
-\newcommand{\daMeanSecurityPerc}{$12.3 \pm 0.0\%$}
-\newcommand{\daMeanSecurityPercTwosfNominal}{12\%}
-\newcommand{\daMeanSecurityPercTwosf}{$12 \pm 0\%$}
-\newcommand{\daVulnFreeNominal}{0.123}
-\newcommand{\daVulnFree}{$0.123 \pm 0.0$}
-\newcommand{\daMeanOutstandingVulnerabilities}{$0.53 \pm 0.0$}
-\newcommand{\daUpdatednessNominal}{0.0523}
-\newcommand{\daUpdatedness}{$0.0523 \pm 0.0$}
-\newcommand{\daUpdatednessPercNominal}{5.23\%}
-\newcommand{\daUpdatednessPerc}{$5.23 \pm 0.0\%$}
-\newcommand{\daUpdatednessPercTwosfNominal}{5.2\%}
-\newcommand{\daUpdatednessPercTwosf}{$5.2 \pm 0.0\%$}
-\newcommand{\daSecurityScoreNominal}{2.87}
-\newcommand{\daSecurityScore}{$2.87 \pm 0.0$}
-\newcommand{\daCurrentMeanInsecurityPercNominal}{79.5\%}
-\newcommand{\daCurrentMeanInsecurityPerc}{$79.5 \pm 0.0\%$}
-\newcommand{\daCurrentMeanInsecurityPercTwosfNominal}{79\%}
-\newcommand{\daCurrentMeanInsecurityPercTwosf}{$79 \pm 0\%$}
-\newcommand{\daCurrentMeanSecurityPercNominal}{20.5\%}
-\newcommand{\daCurrentMeanSecurityPerc}{$20.5 \pm 0.2\%$}
-\newcommand{\daCurrentMeanSecurityPercTwosfNominal}{21\%}
-\newcommand{\daCurrentMeanSecurityPercTwosf}{$21 \pm 0\%$}
-\newcommand{\daCurrentVulnFreeNominal}{0.205}
-\newcommand{\daCurrentVulnFree}{$0.205 \pm 0.002$}
-\newcommand{\daCurrentMeanOutstandingVulnerabilities}{$0.119 \pm 0.0$}
-\newcommand{\daCurrentUpdatednessNominal}{0.0561}
-\newcommand{\daCurrentUpdatedness}{$0.0561 \pm 0.0016$}
-\newcommand{\daCurrentUpdatednessPercNominal}{5.61\%}
-\newcommand{\daCurrentUpdatednessPerc}{$5.61 \pm 0.16\%$}
-\newcommand{\daCurrentUpdatednessPercTwosfNominal}{5.6\%}
-\newcommand{\daCurrentUpdatednessPercTwosf}{$5.6 \pm 0.0\%$}
-\newcommand{\daCurrentSecurityScoreNominal}{3.81}
-\newcommand{\daCurrentSecurityScore}{$3.81 \pm 0.0$}
-\newcommand{\daDailyParticipationNominal}{896}
-\newcommand{\daDailyParticipation}{$896 \pm 249$}
-\newcommand{\daNumFullVersions}{1\,270}
-\newcommand{\daNumSigFullVersions}{95}
+\newcommand{\daOSTotalDaysData}{1\,350\,000}
+\newcommand{\daMeanInsecurityPercNominal}{87.3\%}
+\newcommand{\daMeanInsecurityPerc}{$87.3 \pm 0.0\%$}
+\newcommand{\daMeanInsecurityPercTwosfNominal}{87\%}
+\newcommand{\daMeanInsecurityPercTwosf}{$87 \pm 0\%$}
+\newcommand{\daMeanSecurityPercNominal}{12.7\%}
+\newcommand{\daMeanSecurityPerc}{$12.7 \pm 0.0\%$}
+\newcommand{\daMeanSecurityPercTwosfNominal}{13\%}
+\newcommand{\daMeanSecurityPercTwosf}{$13 \pm 0\%$}
+\newcommand{\daVulnFreeNominal}{0.127}
+\newcommand{\daVulnFree}{$0.127 \pm 0.0$}
+\newcommand{\daMeanOutstandingVulnerabilities}{$0.509 \pm 0.0$}
+\newcommand{\daUpdatednessNominal}{0.0553}
+\newcommand{\daUpdatedness}{$0.0553 \pm 0.0$}
+\newcommand{\daUpdatednessPercNominal}{5.53\%}
+\newcommand{\daUpdatednessPerc}{$5.53 \pm 0.0\%$}
+\newcommand{\daUpdatednessPercTwosfNominal}{5.5\%}
+\newcommand{\daUpdatednessPercTwosf}{$5.5 \pm 0.0\%$}
+\newcommand{\daSecurityScoreNominal}{2.93}
+\newcommand{\daSecurityScore}{$2.93 \pm 0.0$}
+\newcommand{\daCurrentMeanInsecurityPercNominal}{68.8\%}
+\newcommand{\daCurrentMeanInsecurityPerc}{$68.8 \pm 0.0\%$}
+\newcommand{\daCurrentMeanInsecurityPercTwosfNominal}{69\%}
+\newcommand{\daCurrentMeanInsecurityPercTwosf}{$69 \pm 0\%$}
+\newcommand{\daCurrentMeanSecurityPercNominal}{31.2\%}
+\newcommand{\daCurrentMeanSecurityPerc}{$31.2 \pm 0.2\%$}
+\newcommand{\daCurrentMeanSecurityPercTwosfNominal}{31\%}
+\newcommand{\daCurrentMeanSecurityPercTwosf}{$31 \pm 0\%$}
+\newcommand{\daCurrentVulnFreeNominal}{0.312}
+\newcommand{\daCurrentVulnFree}{$0.312 \pm 0.002$}
+\newcommand{\daCurrentMeanOutstandingVulnerabilities}{$0.0645 \pm 0.0$}
+\newcommand{\daCurrentUpdatednessNominal}{0.0934}
+\newcommand{\daCurrentUpdatedness}{$0.0934 \pm 0.0019$}
+\newcommand{\daCurrentUpdatednessPercNominal}{9.34\%}
+\newcommand{\daCurrentUpdatednessPerc}{$9.34 \pm 0.19\%$}
+\newcommand{\daCurrentUpdatednessPercTwosfNominal}{9.3\%}
+\newcommand{\daCurrentUpdatednessPercTwosf}{$9.3 \pm 0.0\%$}
+\newcommand{\daCurrentSecurityScoreNominal}{4.43}
+\newcommand{\daCurrentSecurityScore}{$4.43 \pm 0.0$}
+\newcommand{\daDailyParticipationNominal}{872}
+\newcommand{\daDailyParticipation}{$872 \pm 271$}
+\newcommand{\daNumFullVersions}{1\,300}
+\newcommand{\daNumSigFullVersions}{101}
 \newcommand{\daNumOSVersions}{50}
@@
 \newcommand{\daNumAPIVersions}{16}
-\newcommand{\daNumFullVersionUpdates}{4\,590}
-\newcommand{\daNumOSVersionUpdates}{3\,900}
-\newcommand{\daUpdatesPerYearNominal}{1.26}
-\newcommand{\daUpdatesPerYear}{$1.26 \pm 0.01$}
+\newcommand{\daNumFullVersionUpdates}{4\,670}
+\newcommand{\daNumOSVersionUpdates}{3\,940}
+\newcommand{\daUpdatesPerYearNominal}{1.27}
+\newcommand{\daUpdatesPerYear}{$1.27 \pm 0.01$}
 \newcommand{\daUpdatesPerYearTwosfNominal}{1.3}
 \newcommand{\daUpdatesPerYearTwosf}{$1.3 \pm 0.0$}
-\newcommand{\daNumFullOnlyVersionUpdates}{677}
-\newcommand{\daNumSecurityUpdates}{1\,250}
-\newcommand{\daNumPossibleSecurityUpdates}{2\,300}
-\newcommand{\daOSCurveFitParamFirst}{84.4}
-\newcommand{\daOSCurveFitParamSecond}{\num{0.00285}}
-\newcommand{\daOSCurveFitRMSE}{0.115}
-\newcommand{\daOSCurvePolyRMSE}{0.115}
-\newcommand{\daOSCurveSplineRMSE}{0.115}
-\newcommand{\daOSCurveHalfDeployedYearsNominal}{0.897}
-\newcommand{\daOSCurveHalfDeployedYears}{$0.897 \pm 0.252$}
-\newcommand{\daOSCurveHalfDeployedNominal}{328}
-\newcommand{\daOSCurveHalfDeployed}{$328 \pm 91$}
-\newcommand{\daOSCurveFullDeployedYearsNominal}{3.11}
-\newcommand{\daOSCurveFullDeployedYears}{$3.11 \pm 71$}
-\newcommand{\daOSCurveFullDeployedNominal}{1\,140}
-\newcommand{\daOSCurveFullDeployed}{$1\,140 \pm 260\,000$}
-\newcommand{\daAPICurveFitParamFirst}{99.4}
-\newcommand{\daAPICurveFitParamSecond}{\num{0.00285}}
-\newcommand{\daAPICurveFitRMSE}{0.112}
-\newcommand{\daAPICurvePolyRMSE}{0.112}
-\newcommand{\daAPICurveSplineRMSE}{0.112}
-\newcommand{\daAPICurveHalfDeployedYearsNominal}{0.939}
-\newcommand{\daAPICurveHalfDeployedYears}{$0.939 \pm 0.244$}
-\newcommand{\daAPICurveHalfDeployedNominal}{343}
-\newcommand{\daAPICurveHalfDeployed}{$343 \pm 89$}
-\newcommand{\daAPICurveFullDeployedYearsNominal}{3.15}
-\newcommand{\daAPICurveFullDeployedYears}{$3.15 \pm 71$}
-\newcommand{\daAPICurveFullDeployedNominal}{1\,150}
-\newcommand{\daAPICurveFullDeployed}{$1\,150 \pm 260\,000$}
-\newcommand{\daVulnAPKDuplicateFileOctoberPerc}{91.6\%}
+\newcommand{\daNumFullOnlyVersionUpdates}{724}
+\newcommand{\daNumSecurityUpdates}{1\,270}
+\newcommand{\daNumPossibleSecurityUpdates}{2\,310}
+\newcommand{\daOSCurveFitParamFirst}{83.6}
+\newcommand{\daOSCurveFitParamSecond}{\num{0.0028}}
+\newcommand{\daOSCurveFitRMSE}{0.114}
+\newcommand{\daOSCurvePolyRMSE}{0.114}
+\newcommand{\daOSCurveSplineRMSE}{0.114}
+\newcommand{\daOSCurveHalfDeployedYearsNominal}{0.908}
+\newcommand{\daOSCurveHalfDeployedYears}{$0.908 \pm 0.254$}
+\newcommand{\daOSCurveHalfDeployedNominal}{332}
+\newcommand{\daOSCurveHalfDeployed}{$332 \pm 92$}
+\newcommand{\daOSCurveFullDeployedYearsNominal}{3.16}
+\newcommand{\daOSCurveFullDeployedYears}{$3.16 \pm 72$}
+\newcommand{\daOSCurveFullDeployedNominal}{1\,150}
+\newcommand{\daOSCurveFullDeployed}{$1\,150 \pm 265\,000$}
+\newcommand{\daAPICurveFitParamFirst}{97.8}
+\newcommand{\daAPICurveFitParamSecond}{\num{0.0028}}
+\newcommand{\daAPICurveFitRMSE}{0.11}
+\newcommand{\daAPICurvePolyRMSE}{0.111}
+\newcommand{\daAPICurveSplineRMSE}{0.111}
+\newcommand{\daAPICurveHalfDeployedYearsNominal}{0.946}
+\newcommand{\daAPICurveHalfDeployedYears}{$0.946 \pm 0.244$}
+\newcommand{\daAPICurveHalfDeployedNominal}{345}
+\newcommand{\daAPICurveHalfDeployed}{$345 \pm 89$}
+\newcommand{\daAPICurveFullDeployedYearsNominal}{3.2}
+\newcommand{\daAPICurveFullDeployedYears}{$3.2 \pm 7$}
+\newcommand{\daAPICurveFullDeployedNominal}{1\,170}
+\newcommand{\daAPICurveFullDeployed}{$1\,170 \pm 265\,000$}
+\newcommand{\daVulnAPKDuplicateFileOctoberPerc}{91.7\%}
 \newcommand{\daVulnZergRushMonthsDefFixDeployed}{27.3}
-\newcommand{\daTabSecScoresmanufacturer}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline LG & $0.22 \pm 0.00$ & $0.33 \pm 0.00$ & $0.62 \pm 0.01$ & $3.97 \pm 0.02$ \\  Motorola & $0.18 \pm 0.00$ & $0.12 \pm 0.00$ & $0.71 \pm 0.02$ & $3.07 \pm 0.02$ \\  Samsung & $0.13 \pm 0.00$ & $0.04 \pm 0.00$ & $0.61 \pm 0.00$ & $2.75 \pm 0.00$ \\  Sony & $0.14 \pm 0.00$ & $0.19 \pm 0.00$ & $1.09 \pm 0.02$ & $2.63 \pm 0.02$ \\  HTC & $0.14 \pm 0.00$ & $0.10 \pm 0.00$ & $0.87 \pm 0.01$ & $2.63 \pm 0.02$ \\  asus & $0.20 \pm 0.00$ & $0.51 \pm 0.01$ & $6.01 \pm 0.07$ & $2.35 \pm 0.02$ \\  other & $0.06 \pm 0.00$ & $0.05 \pm 0.00$ & $1.04 \pm 0.01$ & $1.97 \pm 0.02$ \\  alps & $0.03 \pm 0.00$ & $0.19 \pm 0.01$ & $3.99 \pm 0.08$ & $0.80 \pm 0.02$ \\  Symphony & $0.00 \pm 0.00$ & $0.08 \pm 0.00$ & $5.00 \pm 0.05$ & $0.30 \pm 0.01$ \\  walton & $0.00 \pm 0.00$ & $0.09 \pm 0.00$ & $6.00 \pm 0.08$ & $0.27 \pm 0.01$ \\ \end{tabular} \caption{Security scores for manufacturers} \label{tab:sec_manufacturer} \end{table*}}
+\newcommand{\daTabSecScoresmanufacturer}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline LG & $0.24 \pm 0.00$ & $0.34 \pm 0.00$ & $0.60 \pm 0.01$ & $4.11 \pm 0.02$ \\  Motorola & $0.19 \pm 0.00$ & $0.12 \pm 0.00$ & $0.71 \pm 0.02$ & $3.11 \pm 0.02$ \\  Samsung & $0.13 \pm 0.00$ & $0.04 \pm 0.00$ & $0.58 \pm 0.00$ & $2.80 \pm 0.00$ \\  Sony & $0.14 \pm 0.00$ & $0.19 \pm 0.00$ & $1.09 \pm 0.02$ & $2.66 \pm 0.02$ \\  HTC & $0.14 \pm 0.00$ & $0.10 \pm 0.00$ & $0.86 \pm 0.01$ & $2.65 \pm 0.02$ \\  asus & $0.21 \pm 0.00$ & $0.51 \pm 0.01$ & $6.01 \pm 0.07$ & $2.38 \pm 0.02$ \\  other & $0.06 \pm 0.00$ & $0.05 \pm 0.00$ & $1.00 \pm 0.01$ & $2.02 \pm 0.02$ \\  alps & $0.03 \pm 0.00$ & $0.19 \pm 0.01$ & $3.99 \pm 0.08$ & $0.80 \pm 0.02$ \\  Symphony & $0.00 \pm 0.00$ & $0.09 \pm 0.00$ & $5.00 \pm 0.05$ & $0.32 \pm 0.01$ \\  walton & $0.00 \pm 0.00$ & $0.09 \pm 0.00$ & $6.00 \pm 0.08$ & $0.27 \pm 0.01$ \\ \end{tabular} \caption{Security scores for manufacturers} \label{tab:sec_manufacturer} \end{table*}}
 \newcommand{\daSecScoreBestmanufacturer}{LG}
-\newcommand{\daSecScoreBestmanufacturerScoreNominal}{3.97}
-\newcommand{\daSecScoreBestmanufacturerScore}{$3.97 \pm 0.0$}
-\newcommand{\daSecScoreBestmanufacturerNumFullVersions}{135}
+\newcommand{\daSecScoreBestmanufacturerScoreNominal}{4.11}
+\newcommand{\daSecScoreBestmanufacturerScore}{$4.11 \pm 0.0$}
+\newcommand{\daSecScoreBestmanufacturerNumFullVersions}{142}
 \newcommand{\daSecScoreWorstmanufacturer}{walton}
-\newcommand{\daSecScoreWorstmanufacturerScoreNominal}{0.272}
-\newcommand{\daSecScoreWorstmanufacturerScore}{$0.272 \pm 0.007$}
+\newcommand{\daSecScoreWorstmanufacturerScoreNominal}{0.273}
+\newcommand{\daSecScoreWorstmanufacturerScore}{$0.273 \pm 0.007$}
 \newcommand{\daSecScoreWorstmanufacturerNumFullVersions}{21}
-\newcommand{\daTabSecScoresmodel}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline Galaxy Nexus & $0.50 \pm 0.00$ & $0.54 \pm 0.01$ & $1.53 \pm 0.04$ & $4.71 \pm 0.04$ \\  Nexus 4 & $0.30 \pm 0.00$ & $0.82 \pm 0.01$ & $6.06 \pm 0.09$ & $3.69 \pm 0.04$ \\  Nexus 7 & $0.26 \pm 0.00$ & $0.74 \pm 0.01$ & $5.92 \pm 0.09$ & $3.25 \pm 0.04$ \\  other & $0.10 \pm 0.00$ & $0.14 \pm 0.00$ & $0.53 \pm 0.00$ & $3.03 \pm 0.00$ \\  Desire HD & $0.08 \pm 0.00$ & $0.05 \pm 0.00$ & $0.38 \pm 0.02$ & $2.91 \pm 0.04$ \\  HTC Sensation & $0.35 \pm 0.00$ & $0.01 \pm 0.01$ & $1.57 \pm 0.05$ & $2.44 \pm 0.05$ \\  GT-I9100 & $0.22 \pm 0.00$ & $0.02 \pm 0.00$ & $1.23 \pm 0.02$ & $2.27 \pm 0.02$ \\  HTC Desire S & $0.02 \pm 0.00$ & $0.02 \pm 0.00$ & $1.00 \pm 0.06$ & $1.74 \pm 0.07$ \\  GT-N7000 & $0.25 \pm 0.00$ & $0.00 \pm 0.00$ & $2.52 \pm 0.05$ & $1.43 \pm 0.02$ \\  GT-P1000 & $0.01 \pm 0.00$ & $0.00 \pm 0.01$ & $1.79 \pm 0.06$ & $0.90 \pm 0.05$ \\  GT-I9300 & $0.13 \pm 0.00$ & $0.01 \pm 0.00$ & $6.23 \pm 0.04$ & $0.58 \pm 0.01$ \\  GT-I9505 & $0.03 \pm 0.00$ & $0.13 \pm 0.00$ & $6.82 \pm 0.07$ & $0.52 \pm 0.01$ \\  HTC Desire HD & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & $3.03 \pm 0.05$ & $0.28 \pm 0.03$ \\  GT-N7100 & $0.06 \pm 0.00$ & $0.00 \pm 0.01$ & $6.93 \pm 0.08$ & $0.24 \pm 0.02$ \\  Symphony W68 & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & $11.00 \pm 0.12$ & $0.00 \pm 0.03$ \\ \end{tabular} \caption{Security scores for models} \label{tab:sec_model} \end{table*}}
+\newcommand{\daTabSecScoresmodel}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline Galaxy Nexus & $0.50 \pm 0.00$ & $0.54 \pm 0.01$ & $1.53 \pm 0.04$ & $4.70 \pm 0.04$ \\  Nexus 4 & $0.33 \pm 0.00$ & $0.82 \pm 0.01$ & $6.06 \pm 0.09$ & $3.78 \pm 0.04$ \\  Nexus 7 & $0.27 \pm 0.00$ & $0.73 \pm 0.01$ & $5.92 \pm 0.09$ & $3.29 \pm 0.04$ \\  other & $0.10 \pm 0.00$ & $0.14 \pm 0.00$ & $0.51 \pm 0.00$ & $3.08 \pm 0.00$ \\  Desire HD & $0.08 \pm 0.00$ & $0.05 \pm 0.00$ & $0.38 \pm 0.02$ & $2.91 \pm 0.04$ \\  HTC Sensation & $0.35 \pm 0.00$ & $0.01 \pm 0.01$ & $1.57 \pm 0.05$ & $2.44 \pm 0.05$ \\  GT-I9100 & $0.22 \pm 0.00$ & $0.02 \pm 0.00$ & $1.23 \pm 0.02$ & $2.27 \pm 0.02$ \\  HTC Desire S & $0.02 \pm 0.00$ & $0.02 \pm 0.00$ & $1.00 \pm 0.06$ & $1.74 \pm 0.07$ \\  GT-N7000 & $0.25 \pm 0.00$ & $0.00 \pm 0.00$ & $2.52 \pm 0.05$ & $1.43 \pm 0.02$ \\  GT-P1000 & $0.01 \pm 0.00$ & $0.00 \pm 0.01$ & $1.79 \pm 0.06$ & $0.90 \pm 0.05$ \\  GT-I9505 & $0.06 \pm 0.00$ & $0.13 \pm 0.00$ & $6.82 \pm 0.07$ & $0.62 \pm 0.01$ \\  GT-I9300 & $0.13 \pm 0.00$ & $0.01 \pm 0.00$ & $6.23 \pm 0.04$ & $0.57 \pm 0.01$ \\  HTC Desire HD & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & $3.03 \pm 0.05$ & $0.28 \pm 0.03$ \\  GT-N7100 & $0.06 \pm 0.00$ & $0.00 \pm 0.01$ & $6.93 \pm 0.08$ & $0.24 \pm 0.02$ \\  Symphony W68 & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & $11.00 \pm 0.12$ & $0.00 \pm 0.03$ \\ \end{tabular} \caption{Security scores for models} \label{tab:sec_model} \end{table*}}
 \newcommand{\daSecScoreBestmodel}{Galaxy Nexus}
-\newcommand{\daSecScoreBestmodelScoreNominal}{4.71}
-\newcommand{\daSecScoreBestmodelScore}{$4.71 \pm 0.0$}
+\newcommand{\daSecScoreBestmodelScoreNominal}{4.7}
+\newcommand{\daSecScoreBestmodelScore}{$4.7 \pm 0.0$}
 \newcommand{\daSecScoreBestmodelNumFullVersions}{49}
@@
 \newcommand{\daSecScoreWorstmodelScoreNominal}{0.0001}
-\newcommand{\daSecScoreWorstmodelScore}{$0.0001 \pm 0.0273$}
+\newcommand{\daSecScoreWorstmodelScore}{$0.0001 \pm 0.0272$}
 \newcommand{\daSecScoreWorstmodelNumFullVersions}{1}
-\newcommand{\daTabSecScoressummary}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline nexus & $0.39 \pm 0.00$ & $0.48 \pm 0.00$ & $0.56 \pm 0.01$ & $5.17 \pm 0.02$ \\  notnexus & $0.10 \pm 0.00$ & $0.02 \pm 0.00$ & $0.53 \pm 0.00$ & $2.70 \pm 0.00$ \\ \end{tabular} \caption{Security scores for nexus} \label{tab:sec_summary} \end{table*}}
-\newcommand{\daTabSecScoresoperator}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline O2 uk & $0.27 \pm 0.00$ & $0.12 \pm 0.00$ & $0.37 \pm 0.02$ & $3.87 \pm 0.03$ \\  T-Mobile & $0.21 \pm 0.00$ & $0.18 \pm 0.00$ & $0.40 \pm 0.01$ & $3.81 \pm 0.02$ \\  Orange & $0.22 \pm 0.00$ & $0.10 \pm 0.00$ & $0.36 \pm 0.02$ & $3.65 \pm 0.04$ \\  Sprint & $0.18 \pm 0.00$ & $0.11 \pm 0.00$ & $0.43 \pm 0.02$ & $3.42 \pm 0.03$ \\  3 & $0.20 \pm 0.00$ & $0.09 \pm 0.00$ & $0.47 \pm 0.02$ & $3.39 \pm 0.03$ \\  Vodafone uk & $0.14 \pm 0.00$ & $0.13 \pm 0.00$ & $0.52 \pm 0.03$ & $3.17 \pm 0.04$ \\  AT\&T & $0.14 \pm 0.00$ & $0.08 \pm 0.00$ & $0.43 \pm 0.02$ & $3.13 \pm 0.02$ \\  unknown & $0.11 \pm 0.00$ & $0.20 \pm 0.00$ & $0.84 \pm 0.01$ & $2.88 \pm 0.02$ \\  Verizon & $0.19 \pm 0.00$ & $0.09 \pm 0.00$ & $0.82 \pm 0.02$ & $2.84 \pm 0.02$ \\  n Telenor & $0.04 \pm 0.00$ & $0.12 \pm 0.00$ & $1.21 \pm 0.02$ & $1.89 \pm 0.02$ \\  Airtel & $0.05 \pm 0.00$ & $0.03 \pm 0.00$ & $1.47 \pm 0.03$ & $1.41 \pm 0.03$ \\  Grameenphone & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & $1.88 \pm 0.02$ & $0.94 \pm 0.01$ \\  Robi & $0.00 \pm 0.00$ & $0.08 \pm 0.00$ & $2.07 \pm 0.04$ & $0.91 \pm 0.03$ \\  banglalink & $0.00 \pm 0.00$ & $0.03 \pm 0.00$ & $2.56 \pm 0.04$ & $0.54 \pm 0.02$ \\ \end{tabular} \caption{Security scores for operators} \label{tab:sec_operator} \end{table*}}
+\newcommand{\daTabSecScoressummary}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline nexus & $0.40 \pm 0.00$ & $0.49 \pm 0.00$ & $0.55 \pm 0.01$ & $5.28 \pm 0.02$ \\  non-nexus & $0.11 \pm 0.00$ & $0.03 \pm 0.00$ & $0.51 \pm 0.00$ & $2.76 \pm 0.00$ \\ \end{tabular} \caption{Security scores for nexus} \label{tab:sec_summary} \end{table*}}
+\newcommand{\daTabSecScoresoperator}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline O2 uk & $0.28 \pm 0.00$ & $0.12 \pm 0.00$ & $0.37 \pm 0.02$ & $3.91 \pm 0.03$ \\  T-Mobile & $0.22 \pm 0.00$ & $0.19 \pm 0.00$ & $0.39 \pm 0.01$ & $3.87 \pm 0.02$ \\  Orange & $0.22 \pm 0.00$ & $0.10 \pm 0.00$ & $0.36 \pm 0.02$ & $3.64 \pm 0.04$ \\  3 & $0.22 \pm 0.00$ & $0.09 \pm 0.00$ & $0.47 \pm 0.02$ & $3.45 \pm 0.03$ \\  Sprint & $0.18 \pm 0.00$ & $0.11 \pm 0.00$ & $0.43 \pm 0.02$ & $3.41 \pm 0.03$ \\  Vodafone uk & $0.14 \pm 0.00$ & $0.13 \pm 0.00$ & $0.52 \pm 0.03$ & $3.19 \pm 0.04$ \\  AT\&T & $0.15 \pm 0.00$ & $0.08 \pm 0.00$ & $0.43 \pm 0.02$ & $3.17 \pm 0.02$ \\  unknown & $0.12 \pm 0.00$ & $0.20 \pm 0.00$ & $0.80 \pm 0.01$ & $2.95 \pm 0.02$ \\  Verizon & $0.19 \pm 0.00$ & $0.09 \pm 0.00$ & $0.82 \pm 0.02$ & $2.86 \pm 0.02$ \\  n Telenor & $0.05 \pm 0.00$ & $0.12 \pm 0.00$ & $1.14 \pm 0.02$ & $2.01 \pm 0.02$ \\  Airtel & $0.05 \pm 0.00$ & $0.03 \pm 0.00$ & $1.44 \pm 0.03$ & $1.45 \pm 0.03$ \\  Grameenphone & $0.01 \pm 0.00$ & $0.04 \pm 0.00$ & $1.66 \pm 0.02$ & $1.10 \pm 0.01$ \\  Robi & $0.00 \pm 0.00$ & $0.08 \pm 0.00$ & $2.08 \pm 0.04$ & $0.91 \pm 0.03$ \\  banglalink & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & $2.55 \pm 0.04$ & $0.55 \pm 0.02$ \\ \end{tabular} \caption{Security scores for operators} \label{tab:sec_operator} \end{table*}}
 \newcommand{\daSecScoreBestoperator}{O2 uk}
-\newcommand{\daSecScoreBestoperatorScoreNominal}{3.87}
-\newcommand{\daSecScoreBestoperatorScore}{$3.87 \pm 0.0$}
-\newcommand{\daSecScoreBestoperatorNumFullVersions}{100}
+\newcommand{\daSecScoreBestoperatorScoreNominal}{3.91}
+\newcommand{\daSecScoreBestoperatorScore}{$3.91 \pm 0.0$}
+\newcommand{\daSecScoreBestoperatorNumFullVersions}{103}
 \newcommand{\daSecScoreWorstoperator}{banglalink}
-\newcommand{\daSecScoreWorstoperatorScoreNominal}{0.536}
-\newcommand{\daSecScoreWorstoperatorScore}{$0.536 \pm 0.018$}
-\newcommand{\daSecScoreWorstoperatorNumFullVersions}{73}
-\newcommand{\daTabDLDistances}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & nexus \\ \hline equal & 0.0 & 0.0667 & 0.0 & 0.0 \\weight $m$ & 0.2 & 0.4 & 0.0714 & 0.0 \\$f$ & 0.4 & 0.667 & 0.5 & 0.0 \\weight $u$ & 0.4 & 0.0667 & 0.143 & 0.0 \\$m$ & 0.6 & 0.733 & 0.357 & 0.5 \\$u$ & 0.9 & 0.6 & 0.786 & 0.0 \\\end{tabular} \caption{Normalised Damerau-Levenshtein distances for different metrics} \label{tab:dl_distances} \end{table}}
-\newcommand{\daTabChangeInScores}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & nexus \\ \hline $m$ & $-1.67 \pm 2.25$ & $-0.66 \pm 2.54$ & $-3.3 \pm 1.3$ & $-3.4 \pm 1.8$ \\weight $m$ & $-0.264 \pm 0.277$ & $-0.0991 \pm 0.321$ & $-0.461 \pm 0.167$ & $-0.488 \pm 0.217$ \\equal & $-0.108 \pm 0.062$ & $-0.0367 \pm 0.112$ & $-0.146 \pm 0.045$ & $-0.165 \pm 0.033$ \\weight $u$ & $-0.0608 \pm 0.122$ & $-0.0112 \pm 0.232$ & $0.0242 \pm 0.0808$ & $-0.00596 \pm 0.12$ \\$u$ & $0.367 \pm 1.61$ & $0.219 \pm 2.12$ & $1.55 \pm 0.94$ & $1.42 \pm 1.48$ \\$f$ & $0.975 \pm 0.564$ & $0.331 \pm 1.01$ & $1.31 \pm 0.41$ & $1.48 \pm 0.3$ \\\end{tabular} \caption{Mean change in scores for different metrics} \label{tab:change_in_scores} \end{table*}}
-\newcommand{\daGPAPICurveFitParamFirst}{80.8}
-\newcommand{\daGPAPICurveFitParamSecond}{\num{0.00258}}
-\newcommand{\daGPAPICurveFitRMSE}{0.179}
-\newcommand{\daGPAPICurvePolyRMSE}{0.178}
-\newcommand{\daGPAPICurveSplineRMSE}{0.178}
-\newcommand{\daGPAPICurveHalfDeployedYearsNominal}{0.957}
-\newcommand{\daGPAPICurveHalfDeployedYears}{$0.957 \pm 0.469$}
-\newcommand{\daGPAPICurveHalfDeployedNominal}{349}
-\newcommand{\daGPAPICurveHalfDeployed}{$349 \pm 171$}
-\newcommand{\daGPAPICurveFullDeployedYearsNominal}{3.4}
-\newcommand{\daGPAPICurveFullDeployedYears}{$3.4 \pm 7$}
-\newcommand{\daGPAPICurveFullDeployedNominal}{1\,240}
-\newcommand{\daGPAPICurveFullDeployed}{$1\,240 \pm 287\,000$}
-\newcommand{\daGPAPIPerAPIMeanNominal}{0.00452}
-\newcommand{\daGPAPIPerAPIMean}{$0.00452 \pm 0.00341$}
+\newcommand{\daSecScoreWorstoperatorScoreNominal}{0.547}
+\newcommand{\daSecScoreWorstoperatorScore}{$0.547 \pm 0.017$}
+\newcommand{\daSecScoreWorstoperatorNumFullVersions}{75}
+\newcommand{\daTabDLDistances}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & nexus \\ \hline equal & 0.0 & 0.0 & 0.0714 & 0.0 \\weight $m$ & 0.2 & 0.333 & 0.214 & 0.0 \\weight $u$ & 0.2 & 0.0 & 0.143 & 0.0 \\$f$ & 0.3 & 0.733 & 0.357 & 0.0 \\$m$ & 0.6 & 0.733 & 0.357 & 0.5 \\$u$ & 0.9 & 0.667 & 0.786 & 0.0 \\\end{tabular} \caption{Normalised Damerau-Levenshtein distances for different metrics} \label{tab:dl_distances} \end{table}}
+\newcommand{\daTabChangeInScores}{\begin{table*}\small \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & nexus \\ \hline $m$ & $-1.67 \pm 2.26$ & $-0.65 \pm 2.56$ & $-3.34 \pm 1.27$ & $-3.4 \pm 1.9$ \\weight $m$ & $-0.263 \pm 0.278$ & $-0.0958 \pm 0.325$ & $-0.464 \pm 0.159$ & $-0.488 \pm 0.225$ \\equal & $-0.106 \pm 0.061$ & $-0.0343 \pm 0.111$ & $-0.145 \pm 0.043$ & $-0.164 \pm 0.035$ \\weight $u$ & $-0.0569 \pm 0.121$ & $-0.00706 \pm 0.225$ & $0.029 \pm 0.082$ & $-0.00391 \pm 0.12$ \\$u$ & $0.389 \pm 1.62$ & $0.238 \pm 2.09$ & $1.6 \pm 0.9$ & $1.44 \pm 1.51$ \\$f$ & $0.958 \pm 0.556$ & $0.309 \pm 1.0$ & $1.31 \pm 0.39$ & $1.48 \pm 0.32$ \\\end{tabular} \caption{Mean change in scores for different metrics} \label{tab:change_in_scores} \end{table*}}
+\newcommand{\daGPAPICurveFitParamFirst}{82.1}
+\newcommand{\daGPAPICurveFitParamSecond}{\num{0.00257}}
+\newcommand{\daGPAPICurveFitRMSE}{0.176}
+\newcommand{\daGPAPICurvePolyRMSE}{0.176}
+\newcommand{\daGPAPICurveSplineRMSE}{0.176}
+\newcommand{\daGPAPICurveHalfDeployedYearsNominal}{0.964}
+\newcommand{\daGPAPICurveHalfDeployedYears}{$0.964 \pm 0.462$}
+\newcommand{\daGPAPICurveHalfDeployedNominal}{352}
+\newcommand{\daGPAPICurveHalfDeployed}{$352 \pm 169$}
+\newcommand{\daGPAPICurveFullDeployedYearsNominal}{3.42}
+\newcommand{\daGPAPICurveFullDeployedYears}{$3.42 \pm 79$}
+\newcommand{\daGPAPICurveFullDeployedNominal}{1\,250}
+\newcommand{\daGPAPICurveFullDeployed}{$1\,250 \pm 289\,000$}
+\newcommand{\daGPAPIPerAPIMeanNominal}{0.00472}
+\newcommand{\daGPAPIPerAPIMean}{$0.00472 \pm 0.00327$}
 \newcommand{\daGPAPIPerAPIStdevNominal}{0.0209}
-\newcommand{\daGPAPIPerAPIStdev}{$0.0209 \pm 0.0125$}
-\newcommand{\daGPAPIPerAPIAbsMeanNominal}{0.0145}
-\newcommand{\daGPAPIPerAPIAbsMean}{$0.0145 \pm 0.0067$}
-\newcommand{\daGPAPIPerAPIAbsStdevNominal}{0.0152}
-\newcommand{\daGPAPIPerAPIAbsStdev}{$0.0152 \pm 0.0117$}
-\newcommand{\daGPAPIPerAPIParametersTable}{\begin{table} \centering \begin{tabular}{c|S|S} API Version & ${t_0}$ & $\mathrm{decay}$ \\ \hline2	&	0.0	&	0.0237\\3	&	0.0	&	0.0235\\4	&	0.0	&	0.00689\\5	&	36.3	&	0.00429\\6	&	0.0	&	0.0043\\7	&	0.0	&	0.00499\\8	&	47.7	&	0.00415\\9	&	155	&	0.00365\\10	&	90.3	&	0.00364\\11	&	444	&	0.00235\\12	&	356	&	0.00229\\13	&	304	&	0.00236\\14	&	219	&	0.00239\\15	&	161	&	0.0024\\16	&	142	&	0.00228\\17	&	290	&	0.00197\\18	&	121	&	0.00143\\19	&	120	&	0.00155\\\end{tabular} \caption{Parameters for different API versions} \label{tab:per_api_parameters} \end{table}}
-\newcommand{\daGPAPISeventeenLaterDate}{2015-06-01}
-\newcommand{\daGPAPISeventeenLaterProportion}{74.3\%}
-\newcommand{\daGPAPISeventeenEarlierProportion}{25.7\%}
+\newcommand{\daGPAPIPerAPIStdev}{$0.0209 \pm 0.0121$}
+\newcommand{\daGPAPIPerAPIAbsMeanNominal}{0.0143}
+\newcommand{\daGPAPIPerAPIAbsMean}{$0.0143 \pm 0.0066$}
+\newcommand{\daGPAPIPerAPIAbsStdevNominal}{0.0155}
+\newcommand{\daGPAPIPerAPIAbsStdev}{$0.0155 \pm 0.0112$}
+\newcommand{\daGPAPIPerAPIParametersTable}{\begin{table} \centering \begin{tabular}{c|S|S} API Version & ${t_0}$ & $\mathrm{decay}$ \\ \hline2	&	0.0	&	0.0237\\3	&	0.0	&	0.0235\\4	&	0.0	&	0.00689\\5	&	36.3	&	0.00429\\6	&	0.0	&	0.0043\\7	&	0.0	&	0.00499\\8	&	47.7	&	0.00415\\9	&	155	&	0.00365\\10	&	90.3	&	0.00364\\11	&	445	&	0.00236\\12	&	356	&	0.00229\\13	&	305	&	0.00237\\14	&	219	&	0.0024\\15	&	161	&	0.00241\\16	&	143	&	0.00229\\17	&	292	&	0.00199\\18	&	136	&	0.00152\\19	&	123	&	0.00158\\\end{tabular} \caption{Parameters for different API versions} \label{tab:per_api_parameters} \end{table}}
+\newcommand{\daGPAPISeventeenLaterDate}{2015-09-07}
+\newcommand{\daGPAPISeventeenLaterProportion}{79.9\%}
+\newcommand{\daGPAPISeventeenEarlierProportion}{20.1\%}
 \newcommand{\daAPISeventeenReleaseDate}{2012-10-29}
-\newcommand{\daGPAPISeventeenFullDeployment}{2017-10-16}
-\newcommand{\daGPAPISeventeenCurveFitParamFirst}{290}
-\newcommand{\daGPAPISeventeenCurveFitParamSecond}{\num{0.00197}}
-\newcommand{\daGPAPISeventeenCurveFitRMSE}{0.0273}
-\newcommand{\daGPAPISeventeenCurvePolyRMSE}{0.00867}
-\newcommand{\daGPAPISeventeenCurveSplineRMSE}{0.00867}
-\newcommand{\daGPAPISeventeenCurveHalfDeployedYearsNominal}{1.76}
-\newcommand{\daGPAPISeventeenCurveHalfDeployedYears}{$1.76 \pm 0.07$}
-\newcommand{\daGPAPISeventeenCurveHalfDeployedNominal}{642}
-\newcommand{\daGPAPISeventeenCurveHalfDeployed}{$642 \pm 28$}
-\newcommand{\daGPAPISeventeenCurveFullDeployedYearsNominal}{4.97}
-\newcommand{\daGPAPISeventeenCurveFullDeployedYears}{$4.97 \pm 1.1$}
-\newcommand{\daGPAPISeventeenCurveFullDeployedNominal}{1\,810}
-\newcommand{\daGPAPISeventeenCurveFullDeployed}{$1\,810 \pm 401$}
-\newcommand{\daNumUpdatesUpgrades}{3\,660}
-\newcommand{\daUpdatesPerMonthPerVersionNominal}{2.67}
-\newcommand{\daUpdatesPerMonthPerVersion}{$2.67 \pm 0.04$}
+\newcommand{\daGPAPISeventeenFullDeployment}{2017-09-28}
+\newcommand{\daGPAPISeventeenCurveFitParamFirst}{292}
+\newcommand{\daGPAPISeventeenCurveFitParamSecond}{\num{0.00199}}
+\newcommand{\daGPAPISeventeenCurveFitRMSE}{0.027}
+\newcommand{\daGPAPISeventeenCurvePolyRMSE}{0.0103}
+\newcommand{\daGPAPISeventeenCurveSplineRMSE}{0.0103}
+\newcommand{\daGPAPISeventeenCurveHalfDeployedYearsNominal}{1.75}
+\newcommand{\daGPAPISeventeenCurveHalfDeployedYears}{$1.75 \pm 0.07$}
+\newcommand{\daGPAPISeventeenCurveHalfDeployedNominal}{639}
+\newcommand{\daGPAPISeventeenCurveHalfDeployed}{$639 \pm 27$}
+\newcommand{\daGPAPISeventeenCurveFullDeployedYearsNominal}{4.92}
+\newcommand{\daGPAPISeventeenCurveFullDeployedYears}{$4.92 \pm 1.07$}
+\newcommand{\daGPAPISeventeenCurveFullDeployedNominal}{1\,800}
+\newcommand{\daGPAPISeventeenCurveFullDeployed}{$1\,800 \pm 390$}
+\newcommand{\daNumUpdatesUpgrades}{3\,690}
+\newcommand{\daUpdatesPerMonthPerVersionNominal}{2.59}
+\newcommand{\daUpdatesPerMonthPerVersion}{$2.59 \pm 0.04$}
 \newcommand{\daNumUpdatesSkippedBig}{3}
-\newcommand{\daNumUpdatesBigUpgrades}{833}
+\newcommand{\daNumUpdatesBigUpgrades}{843}
 \newcommand{\daNumUpdatesDowngrades}{146}
-\newcommand{\daNumUpdateFullOnly}{674}
-\newcommand{\daPercBigUpgradesNominal}{18.6\%}
-\newcommand{\daPercBigUpgrades}{$18.6 \pm 0.7\%$}
-\newcommand{\daPercUpdatesDowngradesNominal}{3.26\%}
-\newcommand{\daPercUpdatesDowngrades}{$3.26 \pm 0.27\%$}
-\newcommand{\daNumDataPoints}{184 Billion}
-\newcommand{\daMonthsDevices}{2\,110}
+\newcommand{\daNumUpdateFullOnly}{721}
+\newcommand{\daPercBigUpgradesNominal}{18.5\%}
+\newcommand{\daPercBigUpgrades}{$18.5 \pm 0.6\%$}
+\newcommand{\daPercUpdatesDowngradesNominal}{3.2\%}
+\newcommand{\daPercUpdatesDowngrades}{$3.2 \pm 0.2\%$}
+\newcommand{\daNumDataPoints}{103 billion}
+\newcommand{\daMonthsDevices}{2\,140}
 \newcommand{\daMonths}{6}
@@
 \newcommand{\daAdbEnabledPerc}{$19.2 \pm 0.0\%$}
-\newcommand{\daNumStartedApps}{128\,000}
-\newcommand{\daNumInstalledApps}{182\,000}
+\newcommand{\daNumStartedApps}{129\,000}
+\newcommand{\daNumInstalledApps}{184\,000}
 \newcommand{\daAPISeventeenReleaseDateMonth}{October 2012}
-\newcommand{\daGPAPISeventeenFullDeploymentMonth}{October 2017}
-\newcommand{\daMeanOutstandingVulnerabilitiesNominal}{0.53}
-\newcommand{\daCurrentMeanOutstandingVulnerabilitiesNominal}{0.119}
+\newcommand{\daGPAPISeventeenFullDeploymentMonth}{September 2017}
+\newcommand{\daMeanOutstandingVulnerabilitiesNominal}{0.509}
+\newcommand{\daCurrentMeanOutstandingVulnerabilitiesNominal}{0.0645}
 \newcommand{\daModelHalfDeploymentYears}{0.0822}
@@
 \newcommand{\daMeanOutstandingUncertaintyEquation}{$0.99^n$}
-\newcommand{\daGPAPISeventeenLaterDateMonth}{June 2015}
-\newcommand{\daTabSpearmanRanks}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & nexus \\ $\pm\sigma$ & 0.211 & 0.169 & 0.175 & 0.632 \\ \hline $u$ & 0.297 & 0.804 & 0.618 & 1.0 \\$m$ & 0.794 & 0.593 & 0.969 & -1.0 \\$f$ & 0.83 & 0.775 & 0.934 & 1.0 \\weight $u$ & 0.939 & 0.996 & 0.991 & 1.0 \\weight $m$ & 0.976 & 0.964 & 0.996 & 1.0 \\equal & 1.0 & 0.996 & 1.0 & 1.0 \\\end{tabular} \caption{Spearman Rank correlation coefficients for different metrics.
+\newcommand{\daGPAPISeventeenLaterDateMonth}{September 2015}
+\newcommand{\daTabSpearmanRanks}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & nexus \\ $\pm\sigma$ & 0.211 & 0.169 & 0.175 & 0.632 \\ \hline $u$ & 0.309 & 0.8 & 0.609 & 1.0 \\$m$ & 0.794 & 0.589 & 0.965 & -1.0 \\$f$ & 0.842 & 0.754 & 0.943 & 1.0 \\weight $m$ & 0.976 & 0.968 & 0.987 & 1.0 \\weight $u$ & 0.976 & 1.0 & 0.991 & 1.0 \\equal & 1.0 & 1.0 & 0.996 & 1.0 \\\end{tabular} \caption{Spearman Rank correlation coefficients for different metrics.
 The uncertainty is constant for each column but does not take into account the uncertainty in the score which produced the ranking.} \label{tab:spearman_ranks} \end{table}}
 \newcommand{\daStartDateMonth}{July 2011}
-\newcommand{\daEndDateMonth}{July 2015}
+\newcommand{\daEndDateMonth}{September 2015}
 \newcommand{\daSecScoreBestsummary}{nexus}
-\newcommand{\daSecScoreBestsummaryScoreNominal}{5.17}
-\newcommand{\daSecScoreBestsummaryScore}{$5.17 \pm 0.0$}
-\newcommand{\daSecScoreBestsummaryNumFullVersions}{92}
-\newcommand{\daSecScoreWorstsummary}{notnexus}
-\newcommand{\daSecScoreWorstsummaryScoreNominal}{2.7}
-\newcommand{\daSecScoreWorstsummaryScore}{$2.7 \pm 0.0$}
-\newcommand{\daSecScoreWorstsummaryNumFullVersions}{1\,250}
+\newcommand{\daSecScoreBestsummaryScoreNominal}{5.28}
+\newcommand{\daSecScoreBestsummaryScore}{$5.28 \pm 0.0$}
+\newcommand{\daSecScoreBestsummaryNumFullVersions}{99}
+\newcommand{\daSecScoreWorstsummary}{non-nexus}
+\newcommand{\daSecScoreWorstsummaryScoreNominal}{2.76}
+\newcommand{\daSecScoreWorstsummaryScore}{$2.76 \pm 0.0$}
+\newcommand{\daSecScoreWorstsummaryNumFullVersions}{1\,280}

### 2015-10-02 update.sh
Add googleplayapi_single to figures

@@ -14,6 +14,8 @@ cp "$fromdir"device_data/device-data-all-os.pdf device-data-all-os.pdf
 cp "$fromdir"from_to_updates.pdf from_to_updates.pdf
 cp "$fromdir"googleplayapi.pdf googleplayapi.pdf
 cp "$fromdir"googleplayapi-notitle.pdf googleplayapi-notitle.pdf
+cp "$fromdir"googleplayapi_single.pdf googleplayapi_single.pdf
+cp "$fromdir"googleplayapi_single.svg googleplayapi_single.svg
 cp "$fromdir"norm_api_gpcomp.pdf norm_api_gpcomp.pdf
 cp "$fromdir"norm_api_heat.pdf norm_api_heat.pdf
 cp "$fromdir"norm_os_heat.pdf norm_os_heat.pdf

### 2015-10-02 update.sh
Add SVG version of security_score_summary

@@ -46,6 +46,7 @@ cp "$fromdir"security_score_manufacturer.pdf security_score_manufacturer.pdf
 cp "$fromdir"security_score_model.pdf security_score_model.pdf
 cp "$fromdir"security_score_operator.pdf security_score_operator.pdf
 cp "$fromdir"security_score_summary.pdf security_score_summary.pdf
+cp "$fromdir"security_score_summary.svg security_score_summary.svg
 cp ~drt24/git/da/graphing/pdf/dists.pdf dists.pdf
 popd
 update_bib 2> /dev/null

### 2016-04-11 avostats.tex
./update.sh

@@
-\newcommand{\avoNumSubmitters}{8}
+\newcommand{\avoNumSubmitters}{11}
 \newcommand{\avoTotalExternalLines}{25 million}
@@
 \newcommand{\avoBigExternalLinesOfCodePerc}{99.7\%}
-\newcommand{\avoNumVulnerabilities}{33}
-\newcommand{\avoNumVulnAllAndroid}{16}
-\newcommand{\avoNumVulnSpecific}{17}
+\newcommand{\avoNumVulnerabilities}{42}
+\newcommand{\avoNumVulnAllAndroid}{20}
+\newcommand{\avoNumVulnSpecific}{22}
 \newcommand{\avoStartDate}{2013-08-28}
-\newcommand{\avoEndDate}{2015-03-24}
+\newcommand{\avoEndDate}{2016-03-21}
 \newcommand{\avoFirstDataDateMonth}{July 2010}
 \newcommand{\avoFirstDataDate}{2010-07-13}
-\newcommand{\avoLastDataDateMonth}{March 2015}
-\newcommand{\avoLastDataDate}{2015-03-12}
-\newcommand{\avoVulnsPerYearNominal}{7.07}
-\newcommand{\avoVulnsPerYear}{$7.07 \pm 1.23$}
-\newcommand{\avoVulnsPerYearAllAndroidNominal}{3.43}
-\newcommand{\avoVulnsPerYearAllAndroid}{$3.43 \pm 0.85$}
-\newcommand{\avoVulnsPerYearTwosfNominal}{7.1}
-\newcommand{\avoVulnsPerYearTwosf}{$7.1 \pm 1.2$}
-\newcommand{\avoVulnsPerYearAllAndroidTwosfNominal}{3.4}
-\newcommand{\avoVulnsPerYearAllAndroidTwosf}{$3.4 \pm 0.0$}
-\newcommand{\avoTabAndVulns}{\begin{table} \centering \small \begin{tabular}{l|l|c|c} Vulnerability & How known & Date & Categories\\ \hline KillingInTheNameOf & Fixed on & 2010-07-13 & system, kernel\\ exploid udev & Discovered on & 2010-07-15 & kernel\\ levitator & Discovered on & 2011-03-10 & kernel\\ Gingerbreak & Fixed on & 2011-04-18 & system\\ zergRush & Discovered on & 2011-10-06 & system\\ APK duplicate file & Discovered on & 2013-02-18 & signature\\ APK unchecked name & Discovered on & 2013-06-30 & signature\\ APK unsigned shorts & Fixed on & 2013-07-03 & signature\\ vold asec & Fixed on & 2014-01-27 & system\\ Fake ID & Fixed on & 2014-04-17 & signature\\ TowelRoot & Discovered on & 2014-05-03 & kernel\\ ObjectInputStream_deserializable & Discovered on & 2014-06-22 & system\\\end{tabular} \caption{Critical vulnerabilities in Android} \label{tab:andvulns} \end{table}}
+\newcommand{\avoLastDataDateMonth}{October 2015}
+\newcommand{\avoLastDataDate}{2015-10-22}
+\newcommand{\avoVulnsPerYearNominal}{7.96}
+\newcommand{\avoVulnsPerYear}{$7.96 \pm 1.23$}
+\newcommand{\avoVulnsPerYearAllAndroidNominal}{3.79}
+\newcommand{\avoVulnsPerYearAllAndroid}{$3.79 \pm 0.84$}
+\newcommand{\avoVulnsPerYearTwosfNominal}{8.0}
+\newcommand{\avoVulnsPerYearTwosf}{$8.0 \pm 1.2$}
+\newcommand{\avoVulnsPerYearAllAndroidTwosfNominal}{3.8}
+\newcommand{\avoVulnsPerYearAllAndroidTwosf}{$3.8 \pm 0.0$}
+\newcommand{\avoTabAndVulns}{\begin{table} \centering \begin{tabular}{l|l|c|c|p{19.5em}} Vulnerability & How known & Date & Categories & CVEs\\ \hline \href{http://androidvulnerabilities.org/vulnerabilities/KillingInTheNameOf_psneuter_ashmem}{KillingInTheNameOf} & Fixed on & 2010-07-13 & system, kernel & CVE-2011-1149~\cite{CVE-2011-1149}\\ \href{http://androidvulnerabilities.org/vulnerabilities/exploid_udev}{exploid udev} & Discovered on & 2010-07-15 & kernel & CVE-2009-1185~\cite{CVE-2009-1185}\\ \href{http://androidvulnerabilities.org/vulnerabilities/levitator}{levitator} & Discovered on & 2011-03-10 & kernel & CVE-2011-1350~\cite{CVE-2011-1350}, CVE-2011-1352~\cite{CVE-2011-1352}\\ \href{http://androidvulnerabilities.org/vulnerabilities/Gingerbreak}{Gingerbreak} & Fixed on & 2011-04-18 & system & CVE-2011-1823~\cite{CVE-2011-1823}\\ \href{http://androidvulnerabilities.org/vulnerabilities/zergRush}{zergRush} & Discovered on & 2011-10-06 & system & CVE-2011-3874~\cite{CVE-2011-3874}\\ \href{http://androidvulnerabilities.org/vulnerabilities/APK_duplicate_file}{APK duplicate file} & Discovered on & 2013-02-18 & signature & ANDROID-8219321, CVE-2013-4787~\cite{CVE-2013-4787}\\ \href{http://androidvulnerabilities.org/vulnerabilities/APK_unchecked_name}{APK unchecked name} & Discovered on & 2013-06-30 & signature & ANDROID-9950697\\ \href{http://androidvulnerabilities.org/vulnerabilities/APK_unsigned_shorts}{APK unsigned shorts} & Fixed on & 2013-07-03 & signature & ANDROID-9695860\\ \href{http://androidvulnerabilities.org/vulnerabilities/Fake_ID}{Fake ID} & Fixed on & 2014-04-17 & signature & \\ \href{http://androidvulnerabilities.org/vulnerabilities/TowelRoot}{TowelRoot} & Discovered on & 2014-05-03 & kernel & CVE-2014-3153~\cite{CVE-2014-3153}\\ \href{http://androidvulnerabilities.org/vulnerabilities/ObjectInputStream_deserializable}{ObjectInputStream} & Discovered on & 2014-06-22 & system & CVE-2014-7911~\cite{CVE-2014-7911}\\ \href{http://androidvulnerabilities.org/vulnerabilities/Stagefright}{Stagefright} & Fixed on & 2015-04-08 & system, network & CVE-2015-1538~\cite{CVE-2015-1538}, CVE-2015-1539~\cite{CVE-2015-1539}, CVE-2015-3824~\cite{CVE-2015-3824}, CVE-2015-3826~\cite{CVE-2015-3826}, CVE-2015-3827~\cite{CVE-2015-3827}, CVE-2015-3828~\cite{CVE-2015-3828}, CVE-2015-3829~\cite{CVE-2015-3829}\\ \href{http://androidvulnerabilities.org/vulnerabilities/One_class_to_rule_them_all}{One class to rule} & Discovered on & 2015-05-22 & system & CVE-2015-3837~\cite{CVE-2015-3837}, CVE-2015-3825~\cite{CVE-2015-3825}, ANDROID-21437603, ANDROID-21583849\\ \href{http://androidvulnerabilities.org/vulnerabilities/Stagefright2}{Stagefright2} & Discovered on & 2015-08-15 & system, network & CVE-2015-6602~\cite{CVE-2015-6602}, CVE-2015-3876~\cite{CVE-2015-3876}\\\end{tabular} \caption{Critical vulnerabilities in Android} \label{tab:andvulns} \end{table}}
 \newcommand{\avoNumBigExternalProjects}{40}

### 2016-04-11 dastats.tex
./update.sh

@@
-\newcommand{\daNumSigModels}{18}
+\newcommand{\daNumSigModels}{20}
 \newcommand{\daSigNumDevices}{100}
@@
 \newcommand{\daSigVersionDays}{10}
-\newcommand{\daNumVulnsUsed}{11}
+\newcommand{\daNumVulnsUsed}{14}
 \newcommand{\daNumDeviceDataDevices}{50}
 \newcommand{\daOSYearsOfData}{4}
-\newcommand{\daOSMonthsOfData}{51}
+\newcommand{\daOSMonthsOfData}{57}
 \newcommand{\daStartDate}{2011-07-01}
-\newcommand{\daEndDate}{2015-09-22}
+\newcommand{\daEndDate}{2016-03-13}
 \newcommand{\daFullDeployedAt}{95\%}
-\newcommand{\daNumManufacturers}{302}
-\newcommand{\daNumSigManufacturers}{10}
-\newcommand{\daNumModels}{2\,770}
-\newcommand{\daNumOperators}{1\,470}
-\newcommand{\daNumSigOperators}{14}
-\newcommand{\daNumOSDevices}{20\,500}
+\newcommand{\daNumManufacturers}{402}
+\newcommand{\daNumSigManufacturers}{11}
+\newcommand{\daNumModels}{3\,720}
+\newcommand{\daNumOperators}{1\,650}
+\newcommand{\daNumSigOperators}{18}
+\newcommand{\daNumOSDevices}{24\,600}
 \newcommand{\daOSVersionPercValidLines}{99.9\%}
-\newcommand{\daOSTotalDaysData}{1\,350\,000}
-\newcommand{\daMeanInsecurityPercNominal}{87.3\%}
-\newcommand{\daMeanInsecurityPerc}{$87.3 \pm 0.0\%$}
-\newcommand{\daMeanInsecurityPercTwosfNominal}{87\%}
-\newcommand{\daMeanInsecurityPercTwosf}{$87 \pm 0\%$}
-\newcommand{\daMeanSecurityPercNominal}{12.7\%}
-\newcommand{\daMeanSecurityPerc}{$12.7 \pm 0.0\%$}
-\newcommand{\daMeanSecurityPercTwosfNominal}{13\%}
-\newcommand{\daMeanSecurityPercTwosf}{$13 \pm 0\%$}
-\newcommand{\daVulnFreeNominal}{0.127}
-\newcommand{\daVulnFree}{$0.127 \pm 0.0$}
-\newcommand{\daMeanOutstandingVulnerabilities}{$0.509 \pm 0.0$}
-\newcommand{\daUpdatednessNominal}{0.0553}
-\newcommand{\daUpdatedness}{$0.0553 \pm 0.0$}
-\newcommand{\daUpdatednessPercNominal}{5.53\%}
-\newcommand{\daUpdatednessPerc}{$5.53 \pm 0.0\%$}
-\newcommand{\daUpdatednessPercTwosfNominal}{5.5\%}
-\newcommand{\daUpdatednessPercTwosf}{$5.5 \pm 0.0\%$}
-\newcommand{\daSecurityScoreNominal}{2.93}
-\newcommand{\daSecurityScore}{$2.93 \pm 0.0$}
-\newcommand{\daCurrentMeanInsecurityPercNominal}{68.8\%}
-\newcommand{\daCurrentMeanInsecurityPerc}{$68.8 \pm 0.0\%$}
-\newcommand{\daCurrentMeanInsecurityPercTwosfNominal}{69\%}
-\newcommand{\daCurrentMeanInsecurityPercTwosf}{$69 \pm 0\%$}
-\newcommand{\daCurrentMeanSecurityPercNominal}{31.2\%}
-\newcommand{\daCurrentMeanSecurityPerc}{$31.2 \pm 0.2\%$}
-\newcommand{\daCurrentMeanSecurityPercTwosfNominal}{31\%}
-\newcommand{\daCurrentMeanSecurityPercTwosf}{$31 \pm 0\%$}
-\newcommand{\daCurrentVulnFreeNominal}{0.312}
-\newcommand{\daCurrentVulnFree}{$0.312 \pm 0.002$}
-\newcommand{\daCurrentMeanOutstandingVulnerabilities}{$0.0645 \pm 0.0$}
-\newcommand{\daCurrentUpdatednessNominal}{0.0934}
-\newcommand{\daCurrentUpdatedness}{$0.0934 \pm 0.0019$}
-\newcommand{\daCurrentUpdatednessPercNominal}{9.34\%}
-\newcommand{\daCurrentUpdatednessPerc}{$9.34 \pm 0.19\%$}
-\newcommand{\daCurrentUpdatednessPercTwosfNominal}{9.3\%}
-\newcommand{\daCurrentUpdatednessPercTwosf}{$9.3 \pm 0.0\%$}
-\newcommand{\daCurrentSecurityScoreNominal}{4.43}
-\newcommand{\daCurrentSecurityScore}{$4.43 \pm 0.0$}
-\newcommand{\daDailyParticipationNominal}{872}
-\newcommand{\daDailyParticipation}{$872 \pm 271$}
-\newcommand{\daNumFullVersions}{1\,300}
-\newcommand{\daNumSigFullVersions}{101}
-\newcommand{\daNumOSVersions}{50}
-\newcommand{\daNumSigOSVersions}{28}
-\newcommand{\daNumAPIVersions}{16}
-\newcommand{\daNumFullVersionUpdates}{4\,670}
-\newcommand{\daNumOSVersionUpdates}{3\,940}
-\newcommand{\daUpdatesPerYearNominal}{1.27}
-\newcommand{\daUpdatesPerYear}{$1.27 \pm 0.01$}
-\newcommand{\daUpdatesPerYearTwosfNominal}{1.3}
-\newcommand{\daUpdatesPerYearTwosf}{$1.3 \pm 0.0$}
-\newcommand{\daNumFullOnlyVersionUpdates}{724}
-\newcommand{\daNumSecurityUpdates}{1\,270}
-\newcommand{\daNumPossibleSecurityUpdates}{2\,310}
-\newcommand{\daOSCurveFitParamFirst}{83.6}
-\newcommand{\daOSCurveFitParamSecond}{\num{0.0028}}
-\newcommand{\daOSCurveFitRMSE}{0.114}
-\newcommand{\daOSCurvePolyRMSE}{0.114}
-\newcommand{\daOSCurveSplineRMSE}{0.114}
+\newcommand{\daOSTotalDaysData}{1\,520\,000}
+\newcommand{\daMeanInsecurityPercNominal}{83.1\%}
+\newcommand{\daMeanInsecurityPerc}{$83.1 \pm 0.0\%$}
+\newcommand{\daMeanInsecurityPercTwosfNominal}{83\%}
+\newcommand{\daMeanInsecurityPercTwosf}{$83 \pm 0\%$}
+\newcommand{\daMeanSecurityPercNominal}{16.9\%}
+\newcommand{\daMeanSecurityPerc}{$16.9 \pm 0.0\%$}
+\newcommand{\daMeanSecurityPercTwosfNominal}{17\%}
+\newcommand{\daMeanSecurityPercTwosf}{$17 \pm 0\%$}
+\newcommand{\daVulnFreeNominal}{0.169}
+\newcommand{\daVulnFree}{$0.169 \pm 0.0$}
+\newcommand{\daMeanOutstandingVulnerabilities}{$0.468 \pm 0.0$}
+\newcommand{\daUpdatednessNominal}{0.0567}
+\newcommand{\daUpdatedness}{$0.0567 \pm 0.0$}
+\newcommand{\daUpdatednessPercNominal}{5.67\%}
+\newcommand{\daUpdatednessPerc}{$5.67 \pm 0.0\%$}
+\newcommand{\daUpdatednessPercTwosfNominal}{5.7\%}
+\newcommand{\daUpdatednessPercTwosf}{$5.7 \pm 0.0\%$}
+\newcommand{\daSecurityScoreNominal}{3.16}
+\newcommand{\daSecurityScore}{$3.16 \pm 0.0$}
+\newcommand{\daCurrentMeanInsecurityPercNominal}{63.7\%}
+\newcommand{\daCurrentMeanInsecurityPerc}{$63.7 \pm 0.0\%$}
+\newcommand{\daCurrentMeanInsecurityPercTwosfNominal}{64\%}
+\newcommand{\daCurrentMeanInsecurityPercTwosf}{$64 \pm 0\%$}
+\newcommand{\daCurrentMeanSecurityPercNominal}{36.3\%}
+\newcommand{\daCurrentMeanSecurityPerc}{$36.3 \pm 0.0\%$}
+\newcommand{\daCurrentMeanSecurityPercTwosfNominal}{36\%}
+\newcommand{\daCurrentMeanSecurityPercTwosf}{$36 \pm 0\%$}
+\newcommand{\daCurrentVulnFreeNominal}{0.363}
+\newcommand{\daCurrentVulnFree}{$0.363 \pm 0.0$}
+\newcommand{\daCurrentMeanOutstandingVulnerabilities}{$0.0257 \pm 0.0$}
+\newcommand{\daCurrentUpdatednessNominal}{0.0963}
+\newcommand{\daCurrentUpdatedness}{$0.0963 \pm 0.0$}
+\newcommand{\daCurrentUpdatednessPercNominal}{9.63\%}
+\newcommand{\daCurrentUpdatednessPerc}{$9.63 \pm 0.0\%$}
+\newcommand{\daCurrentUpdatednessPercTwosfNominal}{9.6\%}
+\newcommand{\daCurrentUpdatednessPercTwosf}{$9.6 \pm 0.0\%$}
+\newcommand{\daCurrentSecurityScoreNominal}{4.7}
+\newcommand{\daCurrentSecurityScore}{$4.7 \pm 0.0$}
+\newcommand{\daDailyParticipationNominal}{885}
+\newcommand{\daDailyParticipation}{$885 \pm 340$}
+\newcommand{\daNumFullVersions}{1\,660}
+\newcommand{\daNumSigFullVersions}{117}
+\newcommand{\daNumOSVersions}{55}
+\newcommand{\daNumSigOSVersions}{30}
+\newcommand{\daNumAPIVersions}{17}
+\newcommand{\daNumFullVersionUpdates}{5\,970}
+\newcommand{\daNumOSVersionUpdates}{4\,310}
+\newcommand{\daUpdatesPerYearNominal}{1.43}
+\newcommand{\daUpdatesPerYear}{$1.43 \pm 0.01$}
+\newcommand{\daUpdatesPerYearTwosfNominal}{1.4}
+\newcommand{\daUpdatesPerYearTwosf}{$1.4 \pm 0.0$}
+\newcommand{\daNumFullOnlyVersionUpdates}{1\,650}
+\newcommand{\daNumSecurityUpdates}{1\,460}
+\newcommand{\daNumPossibleSecurityUpdates}{2\,340}
+\newcommand{\daOSCurveFitParamFirst}{81.0}
+\newcommand{\daOSCurveFitParamSecond}{\num{0.00277}}
+\newcommand{\daOSCurveFitRMSE}{0.109}
+\newcommand{\daOSCurvePolyRMSE}{0.109}
+\newcommand{\daOSCurveSplineRMSE}{0.109}
 \newcommand{\daOSCurveHalfDeployedYearsNominal}{0.908}
-\newcommand{\daOSCurveHalfDeployedYears}{$0.908 \pm 0.254$}
+\newcommand{\daOSCurveHalfDeployedYears}{$0.908 \pm 0.243$}
 \newcommand{\daOSCurveHalfDeployedNominal}{332}
-\newcommand{\daOSCurveHalfDeployed}{$332 \pm 92$}
-\newcommand{\daOSCurveFullDeployedYearsNominal}{3.16}
-\newcommand{\daOSCurveFullDeployedYears}{$3.16 \pm 72$}
-\newcommand{\daOSCurveFullDeployedNominal}{1\,150}
-\newcommand{\daOSCurveFullDeployed}{$1\,150 \pm 265\,000$}
-\newcommand{\daAPICurveFitParamFirst}{97.8}
+\newcommand{\daOSCurveHalfDeployed}{$332 \pm 88$}
+\newcommand{\daOSCurveFullDeployedYearsNominal}{3.19}
+\newcommand{\daOSCurveFullDeployedYears}{$3.19 \pm 73$}
+\newcommand{\daOSCurveFullDeployedNominal}{1\,160}
+\newcommand{\daOSCurveFullDeployed}{$1\,160 \pm 268\,000$}
+\newcommand{\daAPICurveFitParamFirst}{96.9}
 \newcommand{\daAPICurveFitParamSecond}{\num{0.0028}}
-\newcommand{\daAPICurveFitRMSE}{0.11}
-\newcommand{\daAPICurvePolyRMSE}{0.111}
-\newcommand{\daAPICurveSplineRMSE}{0.111}
-\newcommand{\daAPICurveHalfDeployedYearsNominal}{0.946}
-\newcommand{\daAPICurveHalfDeployedYears}{$0.946 \pm 0.244$}
-\newcommand{\daAPICurveHalfDeployedNominal}{345}
-\newcommand{\daAPICurveHalfDeployed}{$345 \pm 89$}
-\newcommand{\daAPICurveFullDeployedYearsNominal}{3.2}
-\newcommand{\daAPICurveFullDeployedYears}{$3.2 \pm 7$}
+\newcommand{\daAPICurveFitRMSE}{0.106}
+\newcommand{\daAPICurvePolyRMSE}{0.107}
+\newcommand{\daAPICurveSplineRMSE}{0.107}
+\newcommand{\daAPICurveHalfDeployedYearsNominal}{0.943}
+\newcommand{\daAPICurveHalfDeployedYears}{$0.943 \pm 0.232$}
+\newcommand{\daAPICurveHalfDeployedNominal}{344}
+\newcommand{\daAPICurveHalfDeployed}{$344 \pm 84$}
+\newcommand{\daAPICurveFullDeployedYearsNominal}{3.19}
+\newcommand{\daAPICurveFullDeployedYears}{$3.19 \pm 72$}
 \newcommand{\daAPICurveFullDeployedNominal}{1\,170}
-\newcommand{\daAPICurveFullDeployed}{$1\,170 \pm 265\,000$}
-\newcommand{\daVulnAPKDuplicateFileOctoberPerc}{91.7\%}
-\newcommand{\daVulnZergRushMonthsDefFixDeployed}{27.3}
-\newcommand{\daTabSecScoresmanufacturer}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline LG & $0.24 \pm 0.00$ & $0.34 \pm 0.00$ & $0.60 \pm 0.01$ & $4.11 \pm 0.02$ \\  Motorola & $0.19 \pm 0.00$ & $0.12 \pm 0.00$ & $0.71 \pm 0.02$ & $3.11 \pm 0.02$ \\  Samsung & $0.13 \pm 0.00$ & $0.04 \pm 0.00$ & $0.58 \pm 0.00$ & $2.80 \pm 0.00$ \\  Sony & $0.14 \pm 0.00$ & $0.19 \pm 0.00$ & $1.09 \pm 0.02$ & $2.66 \pm 0.02$ \\  HTC & $0.14 \pm 0.00$ & $0.10 \pm 0.00$ & $0.86 \pm 0.01$ & $2.65 \pm 0.02$ \\  asus & $0.21 \pm 0.00$ & $0.51 \pm 0.01$ & $6.01 \pm 0.07$ & $2.38 \pm 0.02$ \\  other & $0.06 \pm 0.00$ & $0.05 \pm 0.00$ & $1.00 \pm 0.01$ & $2.02 \pm 0.02$ \\  alps & $0.03 \pm 0.00$ & $0.19 \pm 0.01$ & $3.99 \pm 0.08$ & $0.80 \pm 0.02$ \\  Symphony & $0.00 \pm 0.00$ & $0.09 \pm 0.00$ & $5.00 \pm 0.05$ & $0.32 \pm 0.01$ \\  walton & $0.00 \pm 0.00$ & $0.09 \pm 0.00$ & $6.00 \pm 0.08$ & $0.27 \pm 0.01$ \\ \end{tabular} \caption{Security scores for manufacturers} \label{tab:sec_manufacturer} \end{table*}}
+\newcommand{\daAPICurveFullDeployed}{$1\,170 \pm 264\,000$}
+\newcommand{\daVulnAPKDuplicateFileOctoberPerc}{92.2\%}
+\newcommand{\daVulnZergRushMonthsDefFixDeployed}{27.6}
+\newcommand{\daTabSecScoresmanufacturer}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline LG & $0.44 \pm 0.00$ & $0.33 \pm 0.00$ & $0.55 \pm 0.01$ & $4.93 \pm 0.02$ \\  oneplus & $0.88 \pm 0.00$ & $0.31 \pm 0.01$ & $6.95 \pm 0.12$ & $4.44 \pm 0.03$ \\  Motorola & $0.32 \pm 0.00$ & $0.14 \pm 0.00$ & $0.65 \pm 0.01$ & $3.74 \pm 0.02$ \\  Sony & $0.29 \pm 0.00$ & $0.18 \pm 0.00$ & $1.06 \pm 0.02$ & $3.25 \pm 0.02$ \\  Samsung & $0.15 \pm 0.00$ & $0.06 \pm 0.00$ & $0.55 \pm 0.00$ & $2.95 \pm 0.00$ \\  HTC & $0.13 \pm 0.00$ & $0.09 \pm 0.00$ & $0.81 \pm 0.01$ & $2.66 \pm 0.02$ \\  Asus & $0.31 \pm 0.00$ & $0.46 \pm 0.01$ & $5.60 \pm 0.06$ & $2.64 \pm 0.02$ \\  \textit{other} & $0.05 \pm 0.00$ & $0.05 \pm 0.00$ & $0.97 \pm 0.01$ & $2.03 \pm 0.02$ \\  alps & $0.02 \pm 0.00$ & $0.20 \pm 0.01$ & $5.00 \pm 0.10$ & $0.73 \pm 0.02$ \\  Symphony & $0.00 \pm 0.00$ & $0.10 \pm 0.00$ & $5.45 \pm 0.06$ & $0.33 \pm 0.01$ \\  walton & $0.00 \pm 0.00$ & $0.09 \pm 0.00$ & $6.08 \pm 0.08$ & $0.28 \pm 0.01$ \\ \end{tabular} \caption{Security scores for manufacturers} \label{tab:sec_manufacturer} \end{table*}}
 \newcommand{\daSecScoreBestmanufacturer}{LG}
-\newcommand{\daSecScoreBestmanufacturerScoreNominal}{4.11}
-\newcommand{\daSecScoreBestmanufacturerScore}{$4.11 \pm 0.0$}
-\newcommand{\daSecScoreBestmanufacturerNumFullVersions}{142}
+\newcommand{\daSecScoreBestmanufacturerScoreNominal}{4.93}
+\newcommand{\daSecScoreBestmanufacturerScore}{$4.93 \pm 0.0$}
+\newcommand{\daSecScoreBestmanufacturerNumFullVersions}{242}
 \newcommand{\daSecScoreWorstmanufacturer}{walton}
-\newcommand{\daSecScoreWorstmanufacturerScoreNominal}{0.273}
-\newcommand{\daSecScoreWorstmanufacturerScore}{$0.273 \pm 0.007$}
-\newcommand{\daSecScoreWorstmanufacturerNumFullVersions}{21}
-\newcommand{\daTabSecScoresmodel}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline Galaxy Nexus & $0.50 \pm 0.00$ & $0.54 \pm 0.01$ & $1.53 \pm 0.04$ & $4.70 \pm 0.04$ \\  Nexus 4 & $0.33 \pm 0.00$ & $0.82 \pm 0.01$ & $6.06 \pm 0.09$ & $3.78 \pm 0.04$ \\  Nexus 7 & $0.27 \pm 0.00$ & $0.73 \pm 0.01$ & $5.92 \pm 0.09$ & $3.29 \pm 0.04$ \\  other & $0.10 \pm 0.00$ & $0.14 \pm 0.00$ & $0.51 \pm 0.00$ & $3.08 \pm 0.00$ \\  Desire HD & $0.08 \pm 0.00$ & $0.05 \pm 0.00$ & $0.38 \pm 0.02$ & $2.91 \pm 0.04$ \\  HTC Sensation & $0.35 \pm 0.00$ & $0.01 \pm 0.01$ & $1.57 \pm 0.05$ & $2.44 \pm 0.05$ \\  GT-I9100 & $0.22 \pm 0.00$ & $0.02 \pm 0.00$ & $1.23 \pm 0.02$ & $2.27 \pm 0.02$ \\  HTC Desire S & $0.02 \pm 0.00$ & $0.02 \pm 0.00$ & $1.00 \pm 0.06$ & $1.74 \pm 0.07$ \\  GT-N7000 & $0.25 \pm 0.00$ & $0.00 \pm 0.00$ & $2.52 \pm 0.05$ & $1.43 \pm 0.02$ \\  GT-P1000 & $0.01 \pm 0.00$ & $0.00 \pm 0.01$ & $1.79 \pm 0.06$ & $0.90 \pm 0.05$ \\  GT-I9505 & $0.06 \pm 0.00$ & $0.13 \pm 0.00$ & $6.82 \pm 0.07$ & $0.62 \pm 0.01$ \\  GT-I9300 & $0.13 \pm 0.00$ & $0.01 \pm 0.00$ & $6.23 \pm 0.04$ & $0.57 \pm 0.01$ \\  HTC Desire HD & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & $3.03 \pm 0.05$ & $0.28 \pm 0.03$ \\  GT-N7100 & $0.06 \pm 0.00$ & $0.00 \pm 0.01$ & $6.93 \pm 0.08$ & $0.24 \pm 0.02$ \\  Symphony W68 & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & $11.00 \pm 0.12$ & $0.00 \pm 0.03$ \\ \end{tabular} \caption{Security scores for models} \label{tab:sec_model} \end{table*}}
-\newcommand{\daSecScoreBestmodel}{Galaxy Nexus}
-\newcommand{\daSecScoreBestmodelScoreNominal}{4.7}
-\newcommand{\daSecScoreBestmodelScore}{$4.7 \pm 0.0$}
-\newcommand{\daSecScoreBestmodelNumFullVersions}{49}
+\newcommand{\daSecScoreWorstmanufacturerScoreNominal}{0.284}
+\newcommand{\daSecScoreWorstmanufacturerScore}{$0.284 \pm 0.008$}
+\newcommand{\daSecScoreWorstmanufacturerNumFullVersions}{22}
+\newcommand{\daTabSecScoresmodel}{\begin{table*} \centering \begin{tabular}{l|l|c|c|c|c} Made by & Model name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline LG & Nexus 5 & $0.78 \pm 0.00$ & $0.65 \pm 0.01$ & $6.02 \pm 0.09$ & $5.10 \pm 0.03$ \\  Samsung & Galaxy Nexus & $0.51 \pm 0.00$ & $0.53 \pm 0.01$ & $1.50 \pm 0.04$ & $4.75 \pm 0.05$ \\  LG & Nexus 4 & $0.60 \pm 0.00$ & $0.57 \pm 0.01$ & $5.52 \pm 0.08$ & $4.16 \pm 0.03$ \\  Asus & Nexus 7 & $0.46 \pm 0.00$ & $0.67 \pm 0.01$ & $5.69 \pm 0.08$ & $3.87 \pm 0.03$ \\  \textit{unknown} & \textit{other} & $0.14 \pm 0.00$ & $0.13 \pm 0.00$ & $0.54 \pm 0.00$ & $3.15 \pm 0.00$ \\  HTC & HTC Desire & $0.14 \pm 0.00$ & $0.07 \pm 0.01$ & $0.51 \pm 0.03$ & $3.03 \pm 0.05$ \\  HTC & Desire HD & $0.09 \pm 0.00$ & $0.05 \pm 0.00$ & $0.43 \pm 0.03$ & $2.86 \pm 0.04$ \\  Samsung & GT-I9000 & $0.03 \pm 0.00$ & $0.03 \pm 0.00$ & $0.44 \pm 0.04$ & $2.55 \pm 0.05$ \\  HTC & HTC Sensation & $0.34 \pm 0.00$ & $0.01 \pm 0.01$ & $1.49 \pm 0.06$ & $2.49 \pm 0.06$ \\  Motorola & DROIDX & $0.02 \pm 0.00$ & $0.04 \pm 0.01$ & $0.55 \pm 0.04$ & $2.39 \pm 0.06$ \\  Samsung & GT-I9100 & $0.21 \pm 0.00$ & $0.01 \pm 0.00$ & $1.26 \pm 0.02$ & $2.23 \pm 0.02$ \\  HTC & HTC Desire S & $0.02 \pm 0.00$ & $0.02 \pm 0.00$ & $1.00 \pm 0.08$ & $1.75 \pm 0.09$ \\  HTC & HTC One & $0.10 \pm 0.00$ & $0.38 \pm 0.01$ & $5.75 \pm 0.12$ & $1.56 \pm 0.04$ \\  Samsung & GT-N7000 & $0.25 \pm 0.00$ & $0.00 \pm 0.00$ & $2.46 \pm 0.05$ & $1.46 \pm 0.03$ \\  Samsung & GT-P1000 & $0.01 \pm 0.00$ & $0.00 \pm 0.01$ & $1.87 \pm 0.06$ & $0.84 \pm 0.05$ \\  Samsung & GT-I9505 & $0.10 \pm 0.00$ & $0.13 \pm 0.00$ & $6.79 \pm 0.07$ & $0.79 \pm 0.02$ \\  Samsung & SM-N9005 & $0.07 \pm 0.00$ & $0.15 \pm 0.01$ & $7.17 \pm 0.18$ & $0.72 \pm 0.03$ \\  Samsung & GT-I9300 & $0.14 \pm 0.00$ & $0.02 \pm 0.00$ & $6.26 \pm 0.05$ & $0.63 \pm 0.01$ \\  HTC & HTC Desire HD & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & $3.03 \pm 0.05$ & $0.28 \pm 0.03$ \\  Samsung & GT-N7100 & $0.05 \pm 0.00$ & $0.00 \pm 0.01$ & $6.64 \pm 0.09$ & $0.21 \pm 0.02$ \\  Symphony & Symphony W68 & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & $11.00 \pm 0.13$ & $0.00 \pm 0.03$ \\ \end{tabular} \caption{Security scores for models} \label{tab:sec_model} \end{table*}}
+\newcommand{\daSecScoreBestmodel}{Nexus 5}
+\newcommand{\daSecScoreBestmodelScoreNominal}{5.1}
+\newcommand{\daSecScoreBestmodelScore}{$5.1 \pm 0.0$}
+\newcommand{\daSecScoreBestmodelNumFullVersions}{47}
 \newcommand{\daSecScoreWorstmodel}{Symphony W68}
 \newcommand{\daSecScoreWorstmodelScoreNominal}{0.0001}
-\newcommand{\daSecScoreWorstmodelScore}{$0.0001 \pm 0.0272$}
+\newcommand{\daSecScoreWorstmodelScore}{$0.0001 \pm 0.0283$}
 \newcommand{\daSecScoreWorstmodelNumFullVersions}{1}
-\newcommand{\daTabSecScoressummary}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline nexus & $0.40 \pm 0.00$ & $0.49 \pm 0.00$ & $0.55 \pm 0.01$ & $5.28 \pm 0.02$ \\  non-nexus & $0.11 \pm 0.00$ & $0.03 \pm 0.00$ & $0.51 \pm 0.00$ & $2.76 \pm 0.00$ \\ \end{tabular} \caption{Security scores for nexus} \label{tab:sec_summary} \end{table*}}
-\newcommand{\daTabSecScoresoperator}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline O2 uk & $0.28 \pm 0.00$ & $0.12 \pm 0.00$ & $0.37 \pm 0.02$ & $3.91 \pm 0.03$ \\  T-Mobile & $0.22 \pm 0.00$ & $0.19 \pm 0.00$ & $0.39 \pm 0.01$ & $3.87 \pm 0.02$ \\  Orange & $0.22 \pm 0.00$ & $0.10 \pm 0.00$ & $0.36 \pm 0.02$ & $3.64 \pm 0.04$ \\  3 & $0.22 \pm 0.00$ & $0.09 \pm 0.00$ & $0.47 \pm 0.02$ & $3.45 \pm 0.03$ \\  Sprint & $0.18 \pm 0.00$ & $0.11 \pm 0.00$ & $0.43 \pm 0.02$ & $3.41 \pm 0.03$ \\  Vodafone uk & $0.14 \pm 0.00$ & $0.13 \pm 0.00$ & $0.52 \pm 0.03$ & $3.19 \pm 0.04$ \\  AT\&T & $0.15 \pm 0.00$ & $0.08 \pm 0.00$ & $0.43 \pm 0.02$ & $3.17 \pm 0.02$ \\  unknown & $0.12 \pm 0.00$ & $0.20 \pm 0.00$ & $0.80 \pm 0.01$ & $2.95 \pm 0.02$ \\  Verizon & $0.19 \pm 0.00$ & $0.09 \pm 0.00$ & $0.82 \pm 0.02$ & $2.86 \pm 0.02$ \\  n Telenor & $0.05 \pm 0.00$ & $0.12 \pm 0.00$ & $1.14 \pm 0.02$ & $2.01 \pm 0.02$ \\  Airtel & $0.05 \pm 0.00$ & $0.03 \pm 0.00$ & $1.44 \pm 0.03$ & $1.45 \pm 0.03$ \\  Grameenphone & $0.01 \pm 0.00$ & $0.04 \pm 0.00$ & $1.66 \pm 0.02$ & $1.10 \pm 0.01$ \\  Robi & $0.00 \pm 0.00$ & $0.08 \pm 0.00$ & $2.08 \pm 0.04$ & $0.91 \pm 0.03$ \\  banglalink & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & $2.55 \pm 0.04$ & $0.55 \pm 0.02$ \\ \end{tabular} \caption{Security scores for operators} \label{tab:sec_operator} \end{table*}}
-\newcommand{\daSecScoreBestoperator}{O2 uk}
-\newcommand{\daSecScoreBestoperatorScoreNominal}{3.91}
-\newcommand{\daSecScoreBestoperatorScore}{$3.91 \pm 0.0$}
-\newcommand{\daSecScoreBestoperatorNumFullVersions}{103}
+\newcommand{\daTabSecScoressummary}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline Nexus devices & $0.62 \pm 0.00$ & $0.50 \pm 0.00$ & $0.50 \pm 0.01$ & $6.26 \pm 0.02$ \\  non-Nexus & $0.13 \pm 0.00$ & $0.02 \pm 0.00$ & $0.54 \pm 0.00$ & $2.81 \pm 0.00$ \\ \end{tabular} \caption{Security scores for Nexus} \label{tab:sec_summary} \end{table*}}
+\newcommand{\daTabSecScoresoperator}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline T-Mobile & $0.31 \pm 0.00$ & $0.17 \pm 0.00$ & $0.35 \pm 0.01$ & $4.24 \pm 0.02$ \\  O2 uk & $0.31 \pm 0.00$ & $0.12 \pm 0.00$ & $0.30 \pm 0.01$ & $4.15 \pm 0.02$ \\  Sprint & $0.23 \pm 0.00$ & $0.12 \pm 0.00$ & $0.30 \pm 0.01$ & $3.85 \pm 0.02$ \\  Orange & $0.23 \pm 0.00$ & $0.11 \pm 0.00$ & $0.34 \pm 0.02$ & $3.75 \pm 0.04$ \\  3 & $0.24 \pm 0.00$ & $0.11 \pm 0.00$ & $0.36 \pm 0.02$ & $3.72 \pm 0.03$ \\  AT\&T & $0.20 \pm 0.00$ & $0.09 \pm 0.00$ & $0.37 \pm 0.01$ & $3.53 \pm 0.02$ \\  Verizon & $0.25 \pm 0.00$ & $0.08 \pm 0.00$ & $0.58 \pm 0.01$ & $3.38 \pm 0.02$ \\  \textit{unknown} & $0.18 \pm 0.00$ & $0.19 \pm 0.00$ & $0.73 \pm 0.01$ & $3.26 \pm 0.02$ \\  Vodafone uk & $0.14 \pm 0.00$ & $0.14 \pm 0.00$ & $0.54 \pm 0.03$ & $3.18 \pm 0.04$ \\  n Telenor & $0.09 \pm 0.00$ & $0.13 \pm 0.00$ & $1.17 \pm 0.02$ & $2.19 \pm 0.02$ \\  Airtel & $0.05 \pm 0.00$ & $0.04 \pm 0.00$ & $2.48 \pm 0.04$ & $0.78 \pm 0.02$ \\  Grameenphone & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & $2.72 \pm 0.03$ & $0.51 \pm 0.01$ \\  Robi & $0.00 \pm 0.00$ & $0.08 \pm 0.00$ & $3.13 \pm 0.05$ & $0.50 \pm 0.02$ \\  banglalink & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & $3.12 \pm 0.05$ & $0.37 \pm 0.01$ \\ \end{tabular} \caption{Security scores for operators} \label{tab:sec_operator} \end{table*}}
+\newcommand{\daSecScoreBestoperator}{T-Mobile}
+\newcommand{\daSecScoreBestoperatorScoreNominal}{4.24}
+\newcommand{\daSecScoreBestoperatorScore}{$4.24 \pm 0.0$}
+\newcommand{\daSecScoreBestoperatorNumFullVersions}{201}
 \newcommand{\daSecScoreWorstoperator}{banglalink}
-\newcommand{\daSecScoreWorstoperatorScoreNominal}{0.547}
-\newcommand{\daSecScoreWorstoperatorScore}{$0.547 \pm 0.017$}
-\newcommand{\daSecScoreWorstoperatorNumFullVersions}{75}
-\newcommand{\daTabDLDistances}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & nexus \\ \hline equal & 0.0 & 0.0 & 0.0714 & 0.0 \\weight $m$ & 0.2 & 0.333 & 0.214 & 0.0 \\weight $u$ & 0.2 & 0.0 & 0.143 & 0.0 \\$f$ & 0.3 & 0.733 & 0.357 & 0.0 \\$m$ & 0.6 & 0.733 & 0.357 & 0.5 \\$u$ & 0.9 & 0.667 & 0.786 & 0.0 \\\end{tabular} \caption{Normalised Damerau-Levenshtein distances for different metrics} \label{tab:dl_distances} \end{table}}
-\newcommand{\daTabChangeInScores}{\begin{table*}\small \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & nexus \\ \hline $m$ & $-1.67 \pm 2.26$ & $-0.65 \pm 2.56$ & $-3.34 \pm 1.27$ & $-3.4 \pm 1.9$ \\weight $m$ & $-0.263 \pm 0.278$ & $-0.0958 \pm 0.325$ & $-0.464 \pm 0.159$ & $-0.488 \pm 0.225$ \\equal & $-0.106 \pm 0.061$ & $-0.0343 \pm 0.111$ & $-0.145 \pm 0.043$ & $-0.164 \pm 0.035$ \\weight $u$ & $-0.0569 \pm 0.121$ & $-0.00706 \pm 0.225$ & $0.029 \pm 0.082$ & $-0.00391 \pm 0.12$ \\$u$ & $0.389 \pm 1.62$ & $0.238 \pm 2.09$ & $1.6 \pm 0.9$ & $1.44 \pm 1.51$ \\$f$ & $0.958 \pm 0.556$ & $0.309 \pm 1.0$ & $1.31 \pm 0.39$ & $1.48 \pm 0.32$ \\\end{tabular} \caption{Mean change in scores for different metrics} \label{tab:change_in_scores} \end{table*}}
-\newcommand{\daGPAPICurveFitParamFirst}{82.1}
-\newcommand{\daGPAPICurveFitParamSecond}{\num{0.00257}}
-\newcommand{\daGPAPICurveFitRMSE}{0.176}
-\newcommand{\daGPAPICurvePolyRMSE}{0.176}
-\newcommand{\daGPAPICurveSplineRMSE}{0.176}
-\newcommand{\daGPAPICurveHalfDeployedYearsNominal}{0.964}
-\newcommand{\daGPAPICurveHalfDeployedYears}{$0.964 \pm 0.462$}
-\newcommand{\daGPAPICurveHalfDeployedNominal}{352}
-\newcommand{\daGPAPICurveHalfDeployed}{$352 \pm 169$}
-\newcommand{\daGPAPICurveFullDeployedYearsNominal}{3.42}
-\newcommand{\daGPAPICurveFullDeployedYears}{$3.42 \pm 79$}
-\newcommand{\daGPAPICurveFullDeployedNominal}{1\,250}
-\newcommand{\daGPAPICurveFullDeployed}{$1\,250 \pm 289\,000$}
-\newcommand{\daGPAPIPerAPIMeanNominal}{0.00472}
-\newcommand{\daGPAPIPerAPIMean}{$0.00472 \pm 0.00327$}
-\newcommand{\daGPAPIPerAPIStdevNominal}{0.0209}
-\newcommand{\daGPAPIPerAPIStdev}{$0.0209 \pm 0.0121$}
-\newcommand{\daGPAPIPerAPIAbsMeanNominal}{0.0143}
-\newcommand{\daGPAPIPerAPIAbsMean}{$0.0143 \pm 0.0066$}
-\newcommand{\daGPAPIPerAPIAbsStdevNominal}{0.0155}
-\newcommand{\daGPAPIPerAPIAbsStdev}{$0.0155 \pm 0.0112$}
-\newcommand{\daGPAPIPerAPIParametersTable}{\begin{table} \centering \begin{tabular}{c|S|S} API Version & ${t_0}$ & $\mathrm{decay}$ \\ \hline2	&	0.0	&	0.0237\\3	&	0.0	&	0.0235\\4	&	0.0	&	0.00689\\5	&	36.3	&	0.00429\\6	&	0.0	&	0.0043\\7	&	0.0	&	0.00499\\8	&	47.7	&	0.00415\\9	&	155	&	0.00365\\10	&	90.3	&	0.00364\\11	&	445	&	0.00236\\12	&	356	&	0.00229\\13	&	305	&	0.00237\\14	&	219	&	0.0024\\15	&	161	&	0.00241\\16	&	143	&	0.00229\\17	&	292	&	0.00199\\18	&	136	&	0.00152\\19	&	123	&	0.00158\\\end{tabular} \caption{Parameters for different API versions} \label{tab:per_api_parameters} \end{table}}
-\newcommand{\daGPAPISeventeenLaterDate}{2015-09-07}
-\newcommand{\daGPAPISeventeenLaterProportion}{79.9\%}
-\newcommand{\daGPAPISeventeenEarlierProportion}{20.1\%}
+\newcommand{\daSecScoreWorstoperatorScoreNominal}{0.366}
+\newcommand{\daSecScoreWorstoperatorScore}{$0.366 \pm 0.013$}
+\newcommand{\daSecScoreWorstoperatorNumFullVersions}{76}
+\newcommand{\daTabDLDistances}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & Nexus \\ \hline equal & 0.0 & 0.0476 & 0.0714 & 0.0 \\weight $u$ & 0.0909 & 0.0952 & 0.143 & 0.0 \\weight $m$ & 0.273 & 0.286 & 0.0714 & 0.0 \\$f$ & 0.273 & 0.81 & 0.357 & 0.0 \\$m$ & 0.636 & 0.857 & 0.429 & 0.0 \\$u$ & 0.818 & 0.714 & 0.786 & 0.0 \\\end{tabular} \caption{Normalised Damerau-Levenshtein distances for different metrics} \label{tab:dl_distances} \end{table}}
+\newcommand{\daTabChangeInScores}{\begin{table*}\small \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & Nexus \\ \hline $m$ & $-0.968 \pm 2.81$ & $-0.75 \pm 3.09$ & $-3.06 \pm 1.75$ & $-2.94 \pm 2.33$ \\weight $m$ & $-0.116 \pm 0.426$ & $-0.095 \pm 0.446$ & $-0.413 \pm 0.228$ & $-0.369 \pm 0.346$ \\equal & $-0.0211 \pm 0.18$ & $-0.0223 \pm 0.163$ & $-0.119 \pm 0.064$ & $-0.0839 \pm 0.126$ \\weight $u$ & $0.0524 \pm 0.208$ & $0.0282 \pm 0.15$ & $0.057 \pm 0.101$ & $0.118 \pm 0.019$ \\$f$ & $0.19 \pm 1.62$ & $0.201 \pm 1.36$ & $1.07 \pm 0.50$ & $0.755 \pm 1.04$ \\$u$ & $0.715 \pm 1.54$ & $0.482 \pm 1.65$ & $1.64 \pm 1.21$ & $1.93 \pm 0.96$ \\\end{tabular} \caption{Mean change in scores for different metrics} \label{tab:change_in_scores} \end{table*}}
+\newcommand{\daGPAPICurveFitParamFirst}{83.6}
+\newcommand{\daGPAPICurveFitParamSecond}{\num{0.00249}}
+\newcommand{\daGPAPICurveFitRMSE}{0.167}
+\newcommand{\daGPAPICurvePolyRMSE}{0.167}
+\newcommand{\daGPAPICurveSplineRMSE}{0.167}
+\newcommand{\daGPAPICurveHalfDeployedYearsNominal}{0.991}
+\newcommand{\daGPAPICurveHalfDeployedYears}{$0.991 \pm 0.448$}
+\newcommand{\daGPAPICurveHalfDeployedNominal}{362}
+\newcommand{\daGPAPICurveHalfDeployed}{$362 \pm 164$}
+\newcommand{\daGPAPICurveFullDeployedYearsNominal}{3.52}
+\newcommand{\daGPAPICurveFullDeployedYears}{$3.52 \pm 81$}
+\newcommand{\daGPAPICurveFullDeployedNominal}{1\,290}
+\newcommand{\daGPAPICurveFullDeployed}{$1\,290 \pm 297\,000$}
+\newcommand{\daGPAPIPerAPIMeanNominal}{0.00465}
+\newcommand{\daGPAPIPerAPIMean}{$0.00465 \pm 0.00294$}
+\newcommand{\daGPAPIPerAPIStdevNominal}{0.0197}
+\newcommand{\daGPAPIPerAPIStdev}{$0.0197 \pm 0.0118$}
+\newcommand{\daGPAPIPerAPIAbsMeanNominal}{0.0135}
+\newcommand{\daGPAPIPerAPIAbsMean}{$0.0135 \pm 0.0061$}
+\newcommand{\daGPAPIPerAPIAbsStdevNominal}{0.0146}
+\newcommand{\daGPAPIPerAPIAbsStdev}{$0.0146 \pm 0.011$}
+\newcommand{\daGPAPIPerAPIParametersTable}{\begin{table} \centering \begin{tabular}{c|S|S} API Version & ${t_0}$ & $\mathrm{decay}$ \\ \hline2	&	0.0	&	0.0237\\3	&	0.0	&	0.0235\\4	&	0.0	&	0.00689\\5	&	36.3	&	0.00429\\6	&	0.0	&	0.0043\\7	&	0.0	&	0.00499\\8	&	0.0	&	0.00346\\9	&	155	&	0.00365\\10	&	90.3	&	0.00364\\11	&	446	&	0.00237\\12	&	369	&	0.00237\\13	&	306	&	0.00238\\14	&	220	&	0.00241\\15	&	162	&	0.00242\\16	&	153	&	0.00236\\17	&	296	&	0.00205\\18	&	146	&	0.00162\\19	&	137	&	0.0017\\21	&	103	&	0.00121\\22	&	127	&	0.00102\\\end{tabular} \caption{Parameters for different API versions} \label{tab:per_api_parameters} \end{table}}
+\newcommand{\daGPAPISeventeenLaterDate}{2016-04-04}
+\newcommand{\daGPAPISeventeenLaterProportion}{87.3\%}
+\newcommand{\daGPAPISeventeenEarlierProportion}{12.7\%}
 \newcommand{\daAPISeventeenReleaseDate}{2012-10-29}
-\newcommand{\daGPAPISeventeenFullDeployment}{2017-09-28}
-\newcommand{\daGPAPISeventeenCurveFitParamFirst}{292}
-\newcommand{\daGPAPISeventeenCurveFitParamSecond}{\num{0.00199}}
-\newcommand{\daGPAPISeventeenCurveFitRMSE}{0.027}
-\newcommand{\daGPAPISeventeenCurvePolyRMSE}{0.0103}
-\newcommand{\daGPAPISeventeenCurveSplineRMSE}{0.0103}
-\newcommand{\daGPAPISeventeenCurveHalfDeployedYearsNominal}{1.75}
-\newcommand{\daGPAPISeventeenCurveHalfDeployedYears}{$1.75 \pm 0.07$}
-\newcommand{\daGPAPISeventeenCurveHalfDeployedNominal}{639}
-\newcommand{\daGPAPISeventeenCurveHalfDeployed}{$639 \pm 27$}
-\newcommand{\daGPAPISeventeenCurveFullDeployedYearsNominal}{4.92}
-\newcommand{\daGPAPISeventeenCurveFullDeployedYears}{$4.92 \pm 1.07$}
-\newcommand{\daGPAPISeventeenCurveFullDeployedNominal}{1\,800}
-\newcommand{\daGPAPISeventeenCurveFullDeployed}{$1\,800 \pm 390$}
-\newcommand{\daNumUpdatesUpgrades}{3\,690}
-\newcommand{\daUpdatesPerMonthPerVersionNominal}{2.59}
-\newcommand{\daUpdatesPerMonthPerVersion}{$2.59 \pm 0.04$}
+\newcommand{\daGPAPISeventeenFullDeployment}{2017-08-23}
+\newcommand{\daGPAPISeventeenCurveFitParamFirst}{296}
+\newcommand{\daGPAPISeventeenCurveFitParamSecond}{\num{0.00205}}
+\newcommand{\daGPAPISeventeenCurveFitRMSE}{0.0258}
+\newcommand{\daGPAPISeventeenCurvePolyRMSE}{0.0182}
+\newcommand{\daGPAPISeventeenCurveSplineRMSE}{0.0182}
+\newcommand{\daGPAPISeventeenCurveHalfDeployedYearsNominal}{1.74}
+\newcommand{\daGPAPISeventeenCurveHalfDeployedYears}{$1.74 \pm 0.07$}
+\newcommand{\daGPAPISeventeenCurveHalfDeployedNominal}{634}
+\newcommand{\daGPAPISeventeenCurveHalfDeployed}{$634 \pm 25$}
+\newcommand{\daGPAPISeventeenCurveFullDeployedYearsNominal}{4.82}
+\newcommand{\daGPAPISeventeenCurveFullDeployedYears}{$4.82 \pm 0.97$}
+\newcommand{\daGPAPISeventeenCurveFullDeployedNominal}{1\,760}
+\newcommand{\daGPAPISeventeenCurveFullDeployed}{$1\,760 \pm 355$}
+\newcommand{\daNumUpdatesUpgrades}{4\,080}
+\newcommand{\daUpdatesPerMonthPerVersionNominal}{2.38}
+\newcommand{\daUpdatesPerMonthPerVersion}{$2.38 \pm 0.03$}
 \newcommand{\daNumUpdatesSkippedBig}{3}
-\newcommand{\daNumUpdatesBigUpgrades}{843}
+\newcommand{\daNumUpdatesBigUpgrades}{890}
 \newcommand{\daNumUpdatesDowngrades}{146}
-\newcommand{\daNumUpdateFullOnly}{721}
-\newcommand{\daPercBigUpgradesNominal}{18.5\%}
-\newcommand{\daPercBigUpgrades}{$18.5 \pm 0.6\%$}
-\newcommand{\daPercUpdatesDowngradesNominal}{3.2\%}
-\newcommand{\daPercUpdatesDowngrades}{$3.2 \pm 0.2\%$}
-\newcommand{\daNumDataPoints}{103 billion}
-\newcommand{\daMonthsDevices}{2\,140}
+\newcommand{\daNumUpdateFullOnly}{1\,640}
+\newcommand{\daPercBigUpgradesNominal}{15.2\%}
+\newcommand{\daPercBigUpgrades}{$15.2 \pm 0.5\%$}
+\newcommand{\daPercUpdatesDowngradesNominal}{2.49\%}
+\newcommand{\daPercUpdatesDowngrades}{$2.49 \pm 0.20\%$}
+\newcommand{\daNumDataPoints}{58 billion}
+\newcommand{\daMonthsDevices}{2\,110}
 \newcommand{\daMonths}{6}
-\newcommand{\daAdbEnabledPercNominal}{19.2\%}
-\newcommand{\daAdbEnabledPerc}{$19.2 \pm 0.0\%$}
-\newcommand{\daNumStartedApps}{129\,000}
-\newcommand{\daNumInstalledApps}{184\,000}
+\newcommand{\daAdbEnabledPercNominal}{19.4\%}
+\newcommand{\daAdbEnabledPerc}{$19.4 \pm 0.0\%$}
+\newcommand{\daNumStartedApps}{134\,000}
+\newcommand{\daNumInstalledApps}{234\,000}
 \newcommand{\daAPISeventeenReleaseDateMonth}{October 2012}
-\newcommand{\daGPAPISeventeenFullDeploymentMonth}{September 2017}
-\newcommand{\daMeanOutstandingVulnerabilitiesNominal}{0.509}
-\newcommand{\daCurrentMeanOutstandingVulnerabilitiesNominal}{0.0645}
-\newcommand{\daModelHalfDeploymentYears}{0.0822}
-\newcommand{\daModelHalfDeployment}{30}
-\newcommand{\daModelFullDeploymentYears}{0.888}
-\newcommand{\daModelFullDeployment}{324}
+\newcommand{\daGPAPISeventeenFullDeploymentMonth}{August 2017}
+\newcommand{\daMeanOutstandingVulnerabilitiesNominal}{0.468}
+\newcommand{\daCurrentMeanOutstandingVulnerabilitiesNominal}{0.0257}
+\newcommand{\daModelHalfDeploymentYears}{0.0384}
+\newcommand{\daModelHalfDeployment}{14}
+\newcommand{\daModelFullDeploymentYears}{0.69}
+\newcommand{\daModelFullDeployment}{252}
 \newcommand{\daMeanOutstandingProbability}{1.0\%}
 \newcommand{\daMeanOutstandingUncertaintyEquation}{$0.99^n$}
-\newcommand{\daGPAPISeventeenLaterDateMonth}{September 2015}
-\newcommand{\daTabSpearmanRanks}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & nexus \\ $\pm\sigma$ & 0.211 & 0.169 & 0.175 & 0.632 \\ \hline $u$ & 0.309 & 0.8 & 0.609 & 1.0 \\$m$ & 0.794 & 0.589 & 0.965 & -1.0 \\$f$ & 0.842 & 0.754 & 0.943 & 1.0 \\weight $m$ & 0.976 & 0.968 & 0.987 & 1.0 \\weight $u$ & 0.976 & 1.0 & 0.991 & 1.0 \\equal & 1.0 & 1.0 & 0.996 & 1.0 \\\end{tabular} \caption{Spearman Rank correlation coefficients for different metrics.
+\newcommand{\daGPAPISeventeenLaterDateMonth}{April 2016}
+\newcommand{\daTabSpearmanRanks}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & Nexus \\ $\pm\sigma$ & 0.2 & 0.141 & 0.175 & 0.632 \\ \hline $u$ & 0.436 & 0.705 & 0.552 & 1.0 \\$m$ & 0.455 & 0.517 & 0.952 & 1.0 \\$f$ & 0.936 & 0.694 & 0.938 & 1.0 \\weight $m$ & 0.973 & 0.978 & 0.996 & 1.0 \\weight $u$ & 0.991 & 0.997 & 0.991 & 1.0 \\equal & 1.0 & 0.999 & 0.996 & 1.0 \\\end{tabular} \caption{Spearman Rank correlation coefficients for different metrics.
 The uncertainty is constant for each column but does not take into account the uncertainty in the score which produced the ranking.} \label{tab:spearman_ranks} \end{table}}
 \newcommand{\daStartDateMonth}{July 2011}
-\newcommand{\daEndDateMonth}{September 2015}
-\newcommand{\daSecScoreBestsummary}{nexus}
-\newcommand{\daSecScoreBestsummaryScoreNominal}{5.28}
-\newcommand{\daSecScoreBestsummaryScore}{$5.28 \pm 0.0$}
-\newcommand{\daSecScoreBestsummaryNumFullVersions}{99}
-\newcommand{\daSecScoreWorstsummary}{non-nexus}
-\newcommand{\daSecScoreWorstsummaryScoreNominal}{2.76}
-\newcommand{\daSecScoreWorstsummaryScore}{$2.76 \pm 0.0$}
-\newcommand{\daSecScoreWorstsummaryNumFullVersions}{1\,280}
+\newcommand{\daEndDateMonth}{March 2016}
+\newcommand{\daSecScoreBestsummary}{Nexus devices}
+\newcommand{\daSecScoreBestsummaryScoreNominal}{6.26}
+\newcommand{\daSecScoreBestsummaryScore}{$6.26 \pm 0.0$}
+\newcommand{\daSecScoreBestsummaryNumFullVersions}{150}
+\newcommand{\daSecScoreWorstsummary}{non-Nexus devices}
+\newcommand{\daSecScoreWorstsummaryScoreNominal}{2.81}
+\newcommand{\daSecScoreWorstsummaryScore}{$2.81 \pm 0.0$}
+\newcommand{\daSecScoreWorstsummaryNumFullVersions}{1\,630}
+\newcommand{\daMeanDefInsecurityPercNominal}{70.9\%}
+\newcommand{\daMeanDefInsecurityPerc}{$70.9 \pm 0.0\%$}
+\newcommand{\daMeanDefInsecurityPercTwosfNominal}{71\%}
+\newcommand{\daMeanDefInsecurityPercTwosf}{$71 \pm 0\%$}
+\newcommand{\daMeanMaybeSecurePercNominal}{14.6\%}
+\newcommand{\daMeanMaybeSecurePerc}{$14.6 \pm 0.0\%$}
+\newcommand{\daMeanMaybeSecurePercTwosfNominal}{15\%}
+\newcommand{\daMeanMaybeSecurePercTwosf}{$15 \pm 0\%$}
+\newcommand{\daOptSecurityScoreNominal}{3.64}
+\newcommand{\daOptSecurityScore}{$3.64 \pm 0.0$}
+\newcommand{\daCurrentMeanDefInsecurityPercNominal}{58.1\%}
+\newcommand{\daCurrentMeanDefInsecurityPerc}{$58.1 \pm 0.0\%$}
+\newcommand{\daCurrentMeanDefInsecurityPercTwosfNominal}{58\%}
+\newcommand{\daCurrentMeanDefInsecurityPercTwosf}{$58 \pm 0\%$}
+\newcommand{\daCurrentMeanMaybeSecurePercNominal}{9.49\%}
+\newcommand{\daCurrentMeanMaybeSecurePerc}{$9.49 \pm 0.0\%$}
+\newcommand{\daCurrentMeanMaybeSecurePercTwosfNominal}{9.5\%}
+\newcommand{\daCurrentMeanMaybeSecurePercTwosf}{$9.5 \pm 0.0\%$}
+\newcommand{\daCurrentOptSecurityScoreNominal}{4.93}
+\newcommand{\daCurrentOptSecurityScore}{$4.93 \pm 0.0$}

### 2016-04-11 update.sh

@@ -13,6 +13,12 @@ cp "$fromdir"device_data/device-data-manufacturer-LG-os.pdf device_data-LG-os.pd
 cp "$fromdir"device_data/device-data-manufacturer-LG-os.svg device_data-LG-os.svg
 cp "$fromdir"device_data/device-data-manufacturer-HTC-os.pdf device_data-HTC-os.pdf
 cp "$fromdir"device_data/device-data-manufacturer-HTC-os.svg device_data-HTC-os.svg
+cp "$fromdir"device_data/device-data-manufacturer-HTC-security.pdf device_data-HTC-security.pdf
+cp "$fromdir"device_data/device-data-manufacturer-HTC-security.svg device_data-HTC-security.svg
+cp "$fromdir"device_data/device-data-manufacturer-Samsung-os.pdf device_data-Samsung-os.pdf
+cp "$fromdir"device_data/device-data-manufacturer-Samsung-os.svg device_data-Samsung-os.svg
+cp "$fromdir"device_data/device-data-manufacturer-Samsung-security.pdf device_data-Samsung-security.pdf
+cp "$fromdir"device_data/device-data-manufacturer-Samsung-security.svg device_data-Samsung-security.svg
 cp "$fromdir"device_data/device-data-all-security.pdf device-data-all-security.pdf
 cp "$fromdir"device_data/device-data-all-os.pdf device-data-all-os.pdf
 cp "$fromdir"from_to_updates.pdf from_to_updates.pdf

### 2016-04-11 update.sh
update norm_versionsecurity plot

@@ -30,6 +30,7 @@ cp "$fromdir"norm_api_gpcomp.pdf norm_api_gpcomp.pdf
 cp "$fromdir"norm_api_heat.pdf norm_api_heat.pdf
 cp "$fromdir"norm_os_heat.pdf norm_os_heat.pdf
 cp "$fromdir"norm_versionsecurity.pdf proportioninsecure.pdf
+cp "$fromdir"norm_versionsecurity.svg norm_versionsecurity.svg
 cp "$fromdir"norm_versionsecurity-notitle.pdf proportioninsecure-notitle.pdf
 cp "$fromdir"updates_between.pdf updates_between_versions.pdf
 cp "$fromdir"nvulnerabilities.pdf vulnerabilities.pdf

### 2016-06-07 dastats.tex
Updated figures etc

@@
 \newcommand{\daStartDate}{2011-07-01}
-\newcommand{\daEndDate}{2016-03-13}
+\newcommand{\daEndDate}{2016-03-16}
 \newcommand{\daFullDeployedAt}{95\%}
@@
 \newcommand{\daOSTotalDaysData}{1\,520\,000}
-\newcommand{\daMeanInsecurityPercNominal}{83.1\%}
-\newcommand{\daMeanInsecurityPerc}{$83.1 \pm 0.0\%$}
-\newcommand{\daMeanInsecurityPercTwosfNominal}{83\%}
-\newcommand{\daMeanInsecurityPercTwosf}{$83 \pm 0\%$}
-\newcommand{\daMeanSecurityPercNominal}{16.9\%}
-\newcommand{\daMeanSecurityPerc}{$16.9 \pm 0.0\%$}
-\newcommand{\daMeanSecurityPercTwosfNominal}{17\%}
-\newcommand{\daMeanSecurityPercTwosf}{$17 \pm 0\%$}
-\newcommand{\daVulnFreeNominal}{0.169}
-\newcommand{\daVulnFree}{$0.169 \pm 0.0$}
-\newcommand{\daMeanOutstandingVulnerabilities}{$0.468 \pm 0.0$}
+\newcommand{\daMeanInsecurityPercNominal}{87.6\%}
+\newcommand{\daMeanInsecurityPerc}{$87.6 \pm 0.0\%$}
+\newcommand{\daMeanInsecurityPercTwosfNominal}{88\%}
+\newcommand{\daMeanInsecurityPercTwosf}{$88 \pm 0\%$}
+\newcommand{\daMeanSecurityPercNominal}{12.4\%}
+\newcommand{\daMeanSecurityPerc}{$12.4 \pm 0.0\%$}
+\newcommand{\daMeanSecurityPercTwosfNominal}{12\%}
+\newcommand{\daMeanSecurityPercTwosf}{$12 \pm 0\%$}
+\newcommand{\daVulnFreeNominal}{0.124}
+\newcommand{\daVulnFree}{$0.124 \pm 0.0$}
+\newcommand{\daMeanOutstandingVulnerabilities}{$0.661 \pm 0.0$}
 \newcommand{\daUpdatednessNominal}{0.0567}
@@
 \newcommand{\daUpdatednessPercTwosf}{$5.7 \pm 0.0\%$}
-\newcommand{\daSecurityScoreNominal}{3.16}
-\newcommand{\daSecurityScore}{$3.16 \pm 0.0$}
-\newcommand{\daCurrentMeanInsecurityPercNominal}{63.7\%}
-\newcommand{\daCurrentMeanInsecurityPerc}{$63.7 \pm 0.0\%$}
-\newcommand{\daCurrentMeanInsecurityPercTwosfNominal}{64\%}
-\newcommand{\daCurrentMeanInsecurityPercTwosf}{$64 \pm 0\%$}
-\newcommand{\daCurrentMeanSecurityPercNominal}{36.3\%}
-\newcommand{\daCurrentMeanSecurityPerc}{$36.3 \pm 0.0\%$}
-\newcommand{\daCurrentMeanSecurityPercTwosfNominal}{36\%}
-\newcommand{\daCurrentMeanSecurityPercTwosf}{$36 \pm 0\%$}
-\newcommand{\daCurrentVulnFreeNominal}{0.363}
-\newcommand{\daCurrentVulnFree}{$0.363 \pm 0.0$}
-\newcommand{\daCurrentMeanOutstandingVulnerabilities}{$0.0257 \pm 0.0$}
-\newcommand{\daCurrentUpdatednessNominal}{0.0963}
-\newcommand{\daCurrentUpdatedness}{$0.0963 \pm 0.0$}
-\newcommand{\daCurrentUpdatednessPercNominal}{9.63\%}
-\newcommand{\daCurrentUpdatednessPerc}{$9.63 \pm 0.0\%$}
-\newcommand{\daCurrentUpdatednessPercTwosfNominal}{9.6\%}
-\newcommand{\daCurrentUpdatednessPercTwosf}{$9.6 \pm 0.0\%$}
-\newcommand{\daCurrentSecurityScoreNominal}{4.7}
-\newcommand{\daCurrentSecurityScore}{$4.7 \pm 0.0$}
-\newcommand{\daDailyParticipationNominal}{885}
-\newcommand{\daDailyParticipation}{$885 \pm 340$}
+\newcommand{\daSecurityScoreNominal}{2.71}
+\newcommand{\daSecurityScore}{$2.71 \pm 0.0$}
+\newcommand{\daCurrentMeanInsecurityPercNominal}{85.2\%}
+\newcommand{\daCurrentMeanInsecurityPerc}{$85.2 \pm 0.0\%$}
+\newcommand{\daCurrentMeanInsecurityPercTwosfNominal}{85\%}
+\newcommand{\daCurrentMeanInsecurityPercTwosf}{$85 \pm 0\%$}
+\newcommand{\daCurrentMeanSecurityPercNominal}{14.8\%}
+\newcommand{\daCurrentMeanSecurityPerc}{$14.8 \pm 0.1\%$}
+\newcommand{\daCurrentMeanSecurityPercTwosfNominal}{15\%}
+\newcommand{\daCurrentMeanSecurityPercTwosf}{$15 \pm 0\%$}
+\newcommand{\daCurrentVulnFreeNominal}{0.148}
+\newcommand{\daCurrentVulnFree}{$0.148 \pm 0.001$}
+\newcommand{\daCurrentMeanOutstandingVulnerabilities}{$0.397 \pm 0.0$}
+\newcommand{\daCurrentUpdatednessNominal}{0.0951}
+\newcommand{\daCurrentUpdatedness}{$0.0951 \pm 0.0$}
+\newcommand{\daCurrentUpdatednessPercNominal}{9.51\%}
+\newcommand{\daCurrentUpdatednessPerc}{$9.51 \pm 0.0\%$}
+\newcommand{\daCurrentUpdatednessPercTwosfNominal}{9.5\%}
+\newcommand{\daCurrentUpdatednessPercTwosf}{$9.5 \pm 0.0\%$}
+\newcommand{\daCurrentSecurityScoreNominal}{3.29}
+\newcommand{\daCurrentSecurityScore}{$3.29 \pm 0.0$}
+\newcommand{\daDailyParticipationNominal}{884}
+\newcommand{\daDailyParticipation}{$884 \pm 341$}
 \newcommand{\daNumFullVersions}{1\,660}
@@
 \newcommand{\daNumSecurityUpdates}{1\,460}
-\newcommand{\daNumPossibleSecurityUpdates}{2\,340}
-\newcommand{\daOSCurveFitParamFirst}{81.0}
+\newcommand{\daNumPossibleSecurityUpdates}{2\,860}
+\newcommand{\daOSCurveFitParamFirst}{80.9}
 \newcommand{\daOSCurveFitParamSecond}{\num{0.00277}}
@@
 \newcommand{\daOSCurveHalfDeployedYearsNominal}{0.908}
-\newcommand{\daOSCurveHalfDeployedYears}{$0.908 \pm 0.243$}
+\newcommand{\daOSCurveHalfDeployedYears}{$0.908 \pm 0.242$}
 \newcommand{\daOSCurveHalfDeployedNominal}{332}
@@
 \newcommand{\daVulnZergRushMonthsDefFixDeployed}{27.6}
-\newcommand{\daTabSecScoresmanufacturer}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline LG & $0.44 \pm 0.00$ & $0.33 \pm 0.00$ & $0.55 \pm 0.01$ & $4.93 \pm 0.02$ \\  oneplus & $0.88 \pm 0.00$ & $0.31 \pm 0.01$ & $6.95 \pm 0.12$ & $4.44 \pm 0.03$ \\  Motorola & $0.32 \pm 0.00$ & $0.14 \pm 0.00$ & $0.65 \pm 0.01$ & $3.74 \pm 0.02$ \\  Sony & $0.29 \pm 0.00$ & $0.18 \pm 0.00$ & $1.06 \pm 0.02$ & $3.25 \pm 0.02$ \\  Samsung & $0.15 \pm 0.00$ & $0.06 \pm 0.00$ & $0.55 \pm 0.00$ & $2.95 \pm 0.00$ \\  HTC & $0.13 \pm 0.00$ & $0.09 \pm 0.00$ & $0.81 \pm 0.01$ & $2.66 \pm 0.02$ \\  Asus & $0.31 \pm 0.00$ & $0.46 \pm 0.01$ & $5.60 \pm 0.06$ & $2.64 \pm 0.02$ \\  \textit{other} & $0.05 \pm 0.00$ & $0.05 \pm 0.00$ & $0.97 \pm 0.01$ & $2.03 \pm 0.02$ \\  alps & $0.02 \pm 0.00$ & $0.20 \pm 0.01$ & $5.00 \pm 0.10$ & $0.73 \pm 0.02$ \\  Symphony & $0.00 \pm 0.00$ & $0.10 \pm 0.00$ & $5.45 \pm 0.06$ & $0.33 \pm 0.01$ \\  walton & $0.00 \pm 0.00$ & $0.09 \pm 0.00$ & $6.08 \pm 0.08$ & $0.28 \pm 0.01$ \\ \end{tabular} \caption{Security scores for manufacturers} \label{tab:sec_manufacturer} \end{table*}}
+\newcommand{\daTabSecScoresmanufacturer}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline LG & $0.34 \pm 0.00$ & $0.33 \pm 0.00$ & $0.74 \pm 0.01$ & $4.28 \pm 0.02$ \\  Motorola & $0.26 \pm 0.00$ & $0.14 \pm 0.00$ & $0.65 \pm 0.01$ & $3.50 \pm 0.02$ \\  HTC & $0.13 \pm 0.00$ & $0.09 \pm 0.00$ & $0.87 \pm 0.01$ & $2.59 \pm 0.02$ \\  Sony & $0.13 \pm 0.00$ & $0.18 \pm 0.00$ & $1.09 \pm 0.02$ & $2.57 \pm 0.02$ \\  Asus & $0.23 \pm 0.00$ & $0.46 \pm 0.01$ & $5.61 \pm 0.06$ & $2.29 \pm 0.02$ \\  Samsung & $0.11 \pm 0.00$ & $0.06 \pm 0.00$ & $0.99 \pm 0.00$ & $2.24 \pm 0.00$ \\  \textit{other} & $0.04 \pm 0.00$ & $0.05 \pm 0.00$ & $1.14 \pm 0.01$ & $1.79 \pm 0.02$ \\  oneplus & $0.02 \pm 0.00$ & $0.31 \pm 0.01$ & $7.85 \pm 0.13$ & $1.00 \pm 0.02$ \\  alps & $0.02 \pm 0.00$ & $0.20 \pm 0.01$ & $5.00 \pm 0.10$ & $0.73 \pm 0.02$ \\  Symphony & $0.00 \pm 0.00$ & $0.10 \pm 0.00$ & $5.74 \pm 0.06$ & $0.32 \pm 0.01$ \\  walton & $0.00 \pm 0.00$ & $0.09 \pm 0.00$ & $6.08 \pm 0.08$ & $0.28 \pm 0.01$ \\ \end{tabular} \caption{Security scores for manufacturers} \label{tab:sec_manufacturer} \end{table*}}
 \newcommand{\daSecScoreBestmanufacturer}{LG}
-\newcommand{\daSecScoreBestmanufacturerScoreNominal}{4.93}
-\newcommand{\daSecScoreBestmanufacturerScore}{$4.93 \pm 0.0$}
+\newcommand{\daSecScoreBestmanufacturerScoreNominal}{4.28}
+\newcommand{\daSecScoreBestmanufacturerScore}{$4.28 \pm 0.0$}
 \newcommand{\daSecScoreBestmanufacturerNumFullVersions}{242}
@@
 \newcommand{\daSecScoreWorstmanufacturerNumFullVersions}{22}
-\newcommand{\daTabSecScoresmodel}{\begin{table*} \centering \begin{tabular}{l|l|c|c|c|c} Made by & Model name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline LG & Nexus 5 & $0.78 \pm 0.00$ & $0.65 \pm 0.01$ & $6.02 \pm 0.09$ & $5.10 \pm 0.03$ \\  Samsung & Galaxy Nexus & $0.51 \pm 0.00$ & $0.53 \pm 0.01$ & $1.50 \pm 0.04$ & $4.75 \pm 0.05$ \\  LG & Nexus 4 & $0.60 \pm 0.00$ & $0.57 \pm 0.01$ & $5.52 \pm 0.08$ & $4.16 \pm 0.03$ \\  Asus & Nexus 7 & $0.46 \pm 0.00$ & $0.67 \pm 0.01$ & $5.69 \pm 0.08$ & $3.87 \pm 0.03$ \\  \textit{unknown} & \textit{other} & $0.14 \pm 0.00$ & $0.13 \pm 0.00$ & $0.54 \pm 0.00$ & $3.15 \pm 0.00$ \\  HTC & HTC Desire & $0.14 \pm 0.00$ & $0.07 \pm 0.01$ & $0.51 \pm 0.03$ & $3.03 \pm 0.05$ \\  HTC & Desire HD & $0.09 \pm 0.00$ & $0.05 \pm 0.00$ & $0.43 \pm 0.03$ & $2.86 \pm 0.04$ \\  Samsung & GT-I9000 & $0.03 \pm 0.00$ & $0.03 \pm 0.00$ & $0.44 \pm 0.04$ & $2.55 \pm 0.05$ \\  HTC & HTC Sensation & $0.34 \pm 0.00$ & $0.01 \pm 0.01$ & $1.49 \pm 0.06$ & $2.49 \pm 0.06$ \\  Motorola & DROIDX & $0.02 \pm 0.00$ & $0.04 \pm 0.01$ & $0.55 \pm 0.04$ & $2.39 \pm 0.06$ \\  Samsung & GT-I9100 & $0.21 \pm 0.00$ & $0.01 \pm 0.00$ & $1.26 \pm 0.02$ & $2.23 \pm 0.02$ \\  HTC & HTC Desire S & $0.02 \pm 0.00$ & $0.02 \pm 0.00$ & $1.00 \pm 0.08$ & $1.75 \pm 0.09$ \\  HTC & HTC One & $0.10 \pm 0.00$ & $0.38 \pm 0.01$ & $5.75 \pm 0.12$ & $1.56 \pm 0.04$ \\  Samsung & GT-N7000 & $0.25 \pm 0.00$ & $0.00 \pm 0.00$ & $2.46 \pm 0.05$ & $1.46 \pm 0.03$ \\  Samsung & GT-P1000 & $0.01 \pm 0.00$ & $0.00 \pm 0.01$ & $1.87 \pm 0.06$ & $0.84 \pm 0.05$ \\  Samsung & GT-I9505 & $0.10 \pm 0.00$ & $0.13 \pm 0.00$ & $6.79 \pm 0.07$ & $0.79 \pm 0.02$ \\  Samsung & SM-N9005 & $0.07 \pm 0.00$ & $0.15 \pm 0.01$ & $7.17 \pm 0.18$ & $0.72 \pm 0.03$ \\  Samsung & GT-I9300 & $0.14 \pm 0.00$ & $0.02 \pm 0.00$ & $6.26 \pm 0.05$ & $0.63 \pm 0.01$ \\  HTC & HTC Desire HD & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & $3.03 \pm 0.05$ & $0.28 \pm 0.03$ \\  Samsung & GT-N7100 & $0.05 \pm 0.00$ & $0.00 \pm 0.01$ & $6.64 \pm 0.09$ & $0.21 \pm 0.02$ \\  Symphony & Symphony W68 & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & $11.00 \pm 0.13$ & $0.00 \pm 0.03$ \\ \end{tabular} \caption{Security scores for models} \label{tab:sec_model} \end{table*}}
+\newcommand{\daTabSecScoresmodel}{\begin{table*} \centering \begin{tabular}{l|l|c|c|c|c} Made by & Model name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline LG & Nexus 5 & $0.71 \pm 0.00$ & $0.65 \pm 0.01$ & $6.09 \pm 0.09$ & $4.80 \pm 0.03$ \\  Samsung & Galaxy Nexus & $0.51 \pm 0.00$ & $0.53 \pm 0.01$ & $1.50 \pm 0.04$ & $4.74 \pm 0.05$ \\  Asus & Nexus 7 & $0.34 \pm 0.00$ & $0.67 \pm 0.01$ & $5.69 \pm 0.08$ & $3.39 \pm 0.03$ \\  LG & Nexus 4 & $0.36 \pm 0.00$ & $0.57 \pm 0.01$ & $5.52 \pm 0.08$ & $3.21 \pm 0.03$ \\  HTC & HTC Desire & $0.14 \pm 0.00$ & $0.07 \pm 0.01$ & $0.51 \pm 0.03$ & $3.03 \pm 0.05$ \\  HTC & Desire HD & $0.09 \pm 0.00$ & $0.05 \pm 0.00$ & $0.43 \pm 0.03$ & $2.86 \pm 0.04$ \\  \textit{unknown} & \textit{other} & $0.08 \pm 0.00$ & $0.13 \pm 0.00$ & $0.74 \pm 0.00$ & $2.65 \pm 0.00$ \\  Samsung & GT-I9000 & $0.03 \pm 0.00$ & $0.03 \pm 0.00$ & $0.44 \pm 0.04$ & $2.55 \pm 0.05$ \\  HTC & HTC Sensation & $0.34 \pm 0.00$ & $0.01 \pm 0.01$ & $1.49 \pm 0.06$ & $2.49 \pm 0.06$ \\  Motorola & DROIDX & $0.02 \pm 0.00$ & $0.04 \pm 0.01$ & $0.55 \pm 0.04$ & $2.39 \pm 0.06$ \\  Samsung & GT-I9100 & $0.21 \pm 0.00$ & $0.01 \pm 0.00$ & $1.26 \pm 0.02$ & $2.23 \pm 0.02$ \\  HTC & HTC Desire S & $0.02 \pm 0.00$ & $0.02 \pm 0.00$ & $1.00 \pm 0.08$ & $1.75 \pm 0.09$ \\  HTC & HTC One & $0.10 \pm 0.00$ & $0.38 \pm 0.01$ & $6.08 \pm 0.12$ & $1.56 \pm 0.04$ \\  Samsung & GT-N7000 & $0.25 \pm 0.00$ & $0.00 \pm 0.00$ & $2.46 \pm 0.05$ & $1.46 \pm 0.03$ \\  Samsung & GT-P1000 & $0.01 \pm 0.00$ & $0.00 \pm 0.01$ & $1.87 \pm 0.06$ & $0.84 \pm 0.05$ \\  Samsung & GT-I9505 & $0.07 \pm 0.00$ & $0.13 \pm 0.00$ & $7.03 \pm 0.07$ & $0.70 \pm 0.01$ \\  Samsung & SM-N9005 & $0.05 \pm 0.00$ & $0.15 \pm 0.01$ & $8.65 \pm 0.20$ & $0.65 \pm 0.03$ \\  Samsung & GT-I9300 & $0.14 \pm 0.00$ & $0.02 \pm 0.00$ & $6.46 \pm 0.05$ & $0.62 \pm 0.01$ \\  HTC & HTC Desire HD & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & $3.03 \pm 0.05$ & $0.28 \pm 0.03$ \\  Samsung & GT-N7100 & $0.05 \pm 0.00$ & $0.00 \pm 0.01$ & $6.64 \pm 0.09$ & $0.21 \pm 0.02$ \\  Symphony & Symphony W68 & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & $11.00 \pm 0.13$ & $0.00 \pm 0.03$ \\ \end{tabular} \caption{Security scores for models} \label{tab:sec_model} \end{table*}}
 \newcommand{\daSecScoreBestmodel}{Nexus 5}
-\newcommand{\daSecScoreBestmodelScoreNominal}{5.1}
-\newcommand{\daSecScoreBestmodelScore}{$5.1 \pm 0.0$}
+\newcommand{\daSecScoreBestmodelScoreNominal}{4.8}
+\newcommand{\daSecScoreBestmodelScore}{$4.8 \pm 0.0$}
 \newcommand{\daSecScoreBestmodelNumFullVersions}{47}
@@
 \newcommand{\daSecScoreWorstmodelNumFullVersions}{1}
-\newcommand{\daTabSecScoressummary}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline Nexus devices & $0.62 \pm 0.00$ & $0.50 \pm 0.00$ & $0.50 \pm 0.01$ & $6.26 \pm 0.02$ \\  non-Nexus & $0.13 \pm 0.00$ & $0.02 \pm 0.00$ & $0.54 \pm 0.00$ & $2.81 \pm 0.00$ \\ \end{tabular} \caption{Security scores for Nexus} \label{tab:sec_summary} \end{table*}}
-\newcommand{\daTabSecScoresoperator}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline T-Mobile & $0.31 \pm 0.00$ & $0.17 \pm 0.00$ & $0.35 \pm 0.01$ & $4.24 \pm 0.02$ \\  O2 uk & $0.31 \pm 0.00$ & $0.12 \pm 0.00$ & $0.30 \pm 0.01$ & $4.15 \pm 0.02$ \\  Sprint & $0.23 \pm 0.00$ & $0.12 \pm 0.00$ & $0.30 \pm 0.01$ & $3.85 \pm 0.02$ \\  Orange & $0.23 \pm 0.00$ & $0.11 \pm 0.00$ & $0.34 \pm 0.02$ & $3.75 \pm 0.04$ \\  3 & $0.24 \pm 0.00$ & $0.11 \pm 0.00$ & $0.36 \pm 0.02$ & $3.72 \pm 0.03$ \\  AT\&T & $0.20 \pm 0.00$ & $0.09 \pm 0.00$ & $0.37 \pm 0.01$ & $3.53 \pm 0.02$ \\  Verizon & $0.25 \pm 0.00$ & $0.08 \pm 0.00$ & $0.58 \pm 0.01$ & $3.38 \pm 0.02$ \\  \textit{unknown} & $0.18 \pm 0.00$ & $0.19 \pm 0.00$ & $0.73 \pm 0.01$ & $3.26 \pm 0.02$ \\  Vodafone uk & $0.14 \pm 0.00$ & $0.14 \pm 0.00$ & $0.54 \pm 0.03$ & $3.18 \pm 0.04$ \\  n Telenor & $0.09 \pm 0.00$ & $0.13 \pm 0.00$ & $1.17 \pm 0.02$ & $2.19 \pm 0.02$ \\  Airtel & $0.05 \pm 0.00$ & $0.04 \pm 0.00$ & $2.48 \pm 0.04$ & $0.78 \pm 0.02$ \\  Grameenphone & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & $2.72 \pm 0.03$ & $0.51 \pm 0.01$ \\  Robi & $0.00 \pm 0.00$ & $0.08 \pm 0.00$ & $3.13 \pm 0.05$ & $0.50 \pm 0.02$ \\  banglalink & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & $3.12 \pm 0.05$ & $0.37 \pm 0.01$ \\ \end{tabular} \caption{Security scores for operators} \label{tab:sec_operator} \end{table*}}
-\newcommand{\daSecScoreBestoperator}{T-Mobile}
-\newcommand{\daSecScoreBestoperatorScoreNominal}{4.24}
-\newcommand{\daSecScoreBestoperatorScore}{$4.24 \pm 0.0$}
-\newcommand{\daSecScoreBestoperatorNumFullVersions}{201}
+\newcommand{\daTabSecScoressummary}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline Nexus devices & $0.53 \pm 0.00$ & $0.50 \pm 0.00$ & $0.69 \pm 0.01$ & $5.63 \pm 0.02$ \\  non-Nexus & $0.09 \pm 0.00$ & $0.02 \pm 0.00$ & $0.74 \pm 0.00$ & $2.35 \pm 0.00$ \\ \end{tabular} \caption{Security scores for Nexus} \label{tab:sec_summary} \end{table*}}
+\newcommand{\daTabSecScoresoperator}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline O2 uk & $0.27 \pm 0.00$ & $0.12 \pm 0.00$ & $0.30 \pm 0.01$ & $3.97 \pm 0.02$ \\  Sprint & $0.19 \pm 0.00$ & $0.12 \pm 0.00$ & $0.30 \pm 0.01$ & $3.67 \pm 0.02$ \\  T-Mobile & $0.21 \pm 0.00$ & $0.17 \pm 0.00$ & $0.56 \pm 0.02$ & $3.56 \pm 0.03$ \\  3 & $0.16 \pm 0.00$ & $0.11 \pm 0.00$ & $0.36 \pm 0.02$ & $3.44 \pm 0.03$ \\  Orange & $0.20 \pm 0.00$ & $0.11 \pm 0.00$ & $0.53 \pm 0.03$ & $3.35 \pm 0.04$ \\  AT\&T & $0.12 \pm 0.00$ & $0.09 \pm 0.00$ & $0.37 \pm 0.01$ & $3.20 \pm 0.02$ \\  Vodafone uk & $0.11 \pm 0.00$ & $0.14 \pm 0.00$ & $0.54 \pm 0.03$ & $3.08 \pm 0.04$ \\  Verizon & $0.16 \pm 0.00$ & $0.08 \pm 0.00$ & $0.58 \pm 0.01$ & $3.04 \pm 0.02$ \\  \textit{unknown} & $0.12 \pm 0.00$ & $0.19 \pm 0.00$ & $1.01 \pm 0.02$ & $2.66 \pm 0.02$ \\  n Telenor & $0.06 \pm 0.00$ & $0.13 \pm 0.00$ & $1.31 \pm 0.02$ & $1.93 \pm 0.02$ \\  Airtel & $0.05 \pm 0.00$ & $0.04 \pm 0.00$ & $2.48 \pm 0.04$ & $0.76 \pm 0.02$ \\  Robi & $0.00 \pm 0.00$ & $0.09 \pm 0.00$ & $3.14 \pm 0.05$ & $0.51 \pm 0.02$ \\  Grameenphone & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & $3.08 \pm 0.03$ & $0.40 \pm 0.01$ \\  banglalink & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & $3.12 \pm 0.05$ & $0.37 \pm 0.01$ \\ \end{tabular} \caption{Security scores for operators} \label{tab:sec_operator} \end{table*}}
+\newcommand{\daSecScoreBestoperator}{O2 uk}
+\newcommand{\daSecScoreBestoperatorScoreNominal}{3.97}
+\newcommand{\daSecScoreBestoperatorScore}{$3.97 \pm 0.0$}
+\newcommand{\daSecScoreBestoperatorNumFullVersions}{124}
 \newcommand{\daSecScoreWorstoperator}{banglalink}
@@
 \newcommand{\daSecScoreWorstoperatorNumFullVersions}{76}
-\newcommand{\daTabDLDistances}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & Nexus \\ \hline equal & 0.0 & 0.0476 & 0.0714 & 0.0 \\weight $u$ & 0.0909 & 0.0952 & 0.143 & 0.0 \\weight $m$ & 0.273 & 0.286 & 0.0714 & 0.0 \\$f$ & 0.273 & 0.81 & 0.357 & 0.0 \\$m$ & 0.636 & 0.857 & 0.429 & 0.0 \\$u$ & 0.818 & 0.714 & 0.786 & 0.0 \\\end{tabular} \caption{Normalised Damerau-Levenshtein distances for different metrics} \label{tab:dl_distances} \end{table}}
-\newcommand{\daTabChangeInScores}{\begin{table*}\small \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & Nexus \\ \hline $m$ & $-0.968 \pm 2.81$ & $-0.75 \pm 3.09$ & $-3.06 \pm 1.75$ & $-2.94 \pm 2.33$ \\weight $m$ & $-0.116 \pm 0.426$ & $-0.095 \pm 0.446$ & $-0.413 \pm 0.228$ & $-0.369 \pm 0.346$ \\equal & $-0.0211 \pm 0.18$ & $-0.0223 \pm 0.163$ & $-0.119 \pm 0.064$ & $-0.0839 \pm 0.126$ \\weight $u$ & $0.0524 \pm 0.208$ & $0.0282 \pm 0.15$ & $0.057 \pm 0.101$ & $0.118 \pm 0.019$ \\$f$ & $0.19 \pm 1.62$ & $0.201 \pm 1.36$ & $1.07 \pm 0.50$ & $0.755 \pm 1.04$ \\$u$ & $0.715 \pm 1.54$ & $0.482 \pm 1.65$ & $1.64 \pm 1.21$ & $1.93 \pm 0.96$ \\\end{tabular} \caption{Mean change in scores for different metrics} \label{tab:change_in_scores} \end{table*}}
+\newcommand{\daTabDLDistances}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & Nexus \\ \hline equal & 0.0909 & 0.143 & 0.0 & 0.0 \\weight $u$ & 0.0909 & 0.143 & 0.0 & 0.0 \\weight $m$ & 0.182 & 0.381 & 0.143 & 0.0 \\$f$ & 0.273 & 0.762 & 0.571 & 0.0 \\$m$ & 0.636 & 0.857 & 0.429 & 0.0 \\$u$ & 0.909 & 0.762 & 0.857 & 0.0 \\\end{tabular} \caption{Normalised Damerau-Levenshtein distances for different metrics} \label{tab:dl_distances} \end{table}}
+\newcommand{\daTabChangeInScores}{\begin{table*}\small \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & Nexus \\ \hline $m$ & $-1.21 \pm 2.08$ & $-0.819 \pm 2.93$ & $-3.03 \pm 1.84$ & $-2.58 \pm 2.2$ \\weight $m$ & $-0.201 \pm 0.249$ & $-0.116 \pm 0.412$ & $-0.426 \pm 0.25$ & $-0.347 \pm 0.314$ \\equal & $-0.0894 \pm 0.0513$ & $-0.0377 \pm 0.147$ & $-0.137 \pm 0.077$ & $-0.0991 \pm 0.105$ \\weight $u$ & $-0.0671 \pm 0.128$ & $0.00267 \pm 0.17$ & $0.0162 \pm 0.0816$ & $0.0494 \pm 0.0373$ \\$u$ & $0.134 \pm 1.58$ & $0.366 \pm 1.78$ & $1.39 \pm 1.11$ & $1.39 \pm 1.08$ \\$f$ & $0.804 \pm 0.462$ & $0.34 \pm 1.21$ & $1.23 \pm 0.61$ & $0.892 \pm 0.84$ \\\end{tabular} \caption{Mean change in scores for different metrics} \label{tab:change_in_scores} \end{table*}}
 \newcommand{\daGPAPICurveFitParamFirst}{83.6}
@@
 \newcommand{\daNumUpdatesSkippedBig}{3}
-\newcommand{\daNumUpdatesBigUpgrades}{890}
+\newcommand{\daNumUpdatesBigUpgrades}{891}
 \newcommand{\daNumUpdatesDowngrades}{146}
@@
 \newcommand{\daPercUpdatesDowngrades}{$2.49 \pm 0.20\%$}
-\newcommand{\daNumDataPoints}{58 billion}
+\newcommand{\daNumDataPoints}{44 billion}
 \newcommand{\daMonthsDevices}{2\,110}
@@
 \newcommand{\daGPAPISeventeenFullDeploymentMonth}{August 2017}
-\newcommand{\daMeanOutstandingVulnerabilitiesNominal}{0.468}
-\newcommand{\daCurrentMeanOutstandingVulnerabilitiesNominal}{0.0257}
+\newcommand{\daMeanOutstandingVulnerabilitiesNominal}{0.661}
+\newcommand{\daCurrentMeanOutstandingVulnerabilitiesNominal}{0.397}
 \newcommand{\daModelHalfDeploymentYears}{0.0384}
@@
 \newcommand{\daGPAPISeventeenLaterDateMonth}{April 2016}
-\newcommand{\daTabSpearmanRanks}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & Nexus \\ $\pm\sigma$ & 0.2 & 0.141 & 0.175 & 0.632 \\ \hline $u$ & 0.436 & 0.705 & 0.552 & 1.0 \\$m$ & 0.455 & 0.517 & 0.952 & 1.0 \\$f$ & 0.936 & 0.694 & 0.938 & 1.0 \\weight $m$ & 0.973 & 0.978 & 0.996 & 1.0 \\weight $u$ & 0.991 & 0.997 & 0.991 & 1.0 \\equal & 1.0 & 0.999 & 0.996 & 1.0 \\\end{tabular} \caption{Spearman Rank correlation coefficients for different metrics.
+\newcommand{\daTabSpearmanRanks}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & Nexus \\ $\pm\sigma$ & 0.2 & 0.141 & 0.175 & 0.632 \\ \hline $u$ & 0.3 & 0.712 & 0.552 & 1.0 \\$m$ & 0.855 & 0.517 & 0.934 & 1.0 \\$f$ & 0.964 & 0.694 & 0.943 & 1.0 \\weight $m$ & 0.973 & 0.948 & 0.991 & 1.0 \\equal & 0.991 & 0.996 & 1.0 & 1.0 \\weight $u$ & 0.991 & 0.996 & 1.0 & 1.0 \\\end{tabular} \caption{Spearman Rank correlation coefficients for different metrics.
 The uncertainty is constant for each column but does not take into account the uncertainty in the score which produced the ranking.} \label{tab:spearman_ranks} \end{table}}
@@
 \newcommand{\daSecScoreBestsummary}{Nexus devices}
-\newcommand{\daSecScoreBestsummaryScoreNominal}{6.26}
-\newcommand{\daSecScoreBestsummaryScore}{$6.26 \pm 0.0$}
+\newcommand{\daSecScoreBestsummaryScoreNominal}{5.63}
+\newcommand{\daSecScoreBestsummaryScore}{$5.63 \pm 0.0$}
 \newcommand{\daSecScoreBestsummaryNumFullVersions}{150}
 \newcommand{\daSecScoreWorstsummary}{non-Nexus devices}
-\newcommand{\daSecScoreWorstsummaryScoreNominal}{2.81}
-\newcommand{\daSecScoreWorstsummaryScore}{$2.81 \pm 0.0$}
+\newcommand{\daSecScoreWorstsummaryScoreNominal}{2.35}
+\newcommand{\daSecScoreWorstsummaryScore}{$2.35 \pm 0.0$}
 \newcommand{\daSecScoreWorstsummaryNumFullVersions}{1\,630}
-\newcommand{\daMeanDefInsecurityPercNominal}{70.9\%}
-\newcommand{\daMeanDefInsecurityPerc}{$70.9 \pm 0.0\%$}
-\newcommand{\daMeanDefInsecurityPercTwosfNominal}{71\%}
-\newcommand{\daMeanDefInsecurityPercTwosf}{$71 \pm 0\%$}
-\newcommand{\daMeanMaybeSecurePercNominal}{14.6\%}
-\newcommand{\daMeanMaybeSecurePerc}{$14.6 \pm 0.0\%$}
-\newcommand{\daMeanMaybeSecurePercTwosfNominal}{15\%}
-\newcommand{\daMeanMaybeSecurePercTwosf}{$15 \pm 0\%$}
-\newcommand{\daOptSecurityScoreNominal}{3.64}
-\newcommand{\daOptSecurityScore}{$3.64 \pm 0.0$}
-\newcommand{\daCurrentMeanDefInsecurityPercNominal}{58.1\%}
-\newcommand{\daCurrentMeanDefInsecurityPerc}{$58.1 \pm 0.0\%$}
-\newcommand{\daCurrentMeanDefInsecurityPercTwosfNominal}{58\%}
-\newcommand{\daCurrentMeanDefInsecurityPercTwosf}{$58 \pm 0\%$}
-\newcommand{\daCurrentMeanMaybeSecurePercNominal}{9.49\%}
-\newcommand{\daCurrentMeanMaybeSecurePerc}{$9.49 \pm 0.0\%$}
-\newcommand{\daCurrentMeanMaybeSecurePercTwosfNominal}{9.5\%}
-\newcommand{\daCurrentMeanMaybeSecurePercTwosf}{$9.5 \pm 0.0\%$}
-\newcommand{\daCurrentOptSecurityScoreNominal}{4.93}
-\newcommand{\daCurrentOptSecurityScore}{$4.93 \pm 0.0$}
+\newcommand{\daMeanDefInsecurityPercNominal}{73.2\%}
+\newcommand{\daMeanDefInsecurityPerc}{$73.2 \pm 0.0\%$}
+\newcommand{\daMeanDefInsecurityPercTwosfNominal}{73\%}
+\newcommand{\daMeanDefInsecurityPercTwosf}{$73 \pm 0\%$}
+\newcommand{\daMeanMaybeSecurePercNominal}{16.4\%}
+\newcommand{\daMeanMaybeSecurePerc}{$16.4 \pm 0.0\%$}
+\newcommand{\daMeanMaybeSecurePercTwosfNominal}{16\%}
+\newcommand{\daMeanMaybeSecurePercTwosf}{$16 \pm 0\%$}
+\newcommand{\daOptSecurityScoreNominal}{3.29}
+\newcommand{\daOptSecurityScore}{$3.29 \pm 0.0$}
+\newcommand{\daCurrentMeanDefInsecurityPercNominal}{69.4\%}
+\newcommand{\daCurrentMeanDefInsecurityPerc}{$69.4 \pm 0.0\%$}
+\newcommand{\daCurrentMeanDefInsecurityPercTwosfNominal}{69\%}
+\newcommand{\daCurrentMeanDefInsecurityPercTwosf}{$69 \pm 0\%$}
+\newcommand{\daCurrentMeanMaybeSecurePercNominal}{19.1\%}
+\newcommand{\daCurrentMeanMaybeSecurePerc}{$19.1 \pm 0.1\%$}
+\newcommand{\daCurrentMeanMaybeSecurePercTwosfNominal}{19\%}
+\newcommand{\daCurrentMeanMaybeSecurePercTwosf}{$19 \pm 0\%$}
+\newcommand{\daCurrentOptSecurityScoreNominal}{3.92}
+\newcommand{\daCurrentOptSecurityScore}{$3.92 \pm 0.0$}
